### 2017-05-04 maxNt.tex
update luca's bib

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

\usepackage[final,nonatbib]{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[utf8]{inputenc} % allow utf-8 input
%\usepackage{hyperref}       % hyperlinks
%\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
%\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{textcomp}
\usepackage{amsmath}

\usepackage{mathtools}
\setlength{\multlinegap}{0pt}
% \newlength{\Normalbaselineskip}
% \setlength{\Normalbaselineskip}{\baselineskip}
% \appto\MultlinedHook{\setlength\baselineskip{\Normalbaselineskip}}

\usepackage{amssymb}
\usepackage{amsxtra}
\usepackage{amsthm}
\usepackage{bm}

\usepackage[british]{babel}
\selectlanguage{british}
% \newcommand*{\langfrench}{\foreignlanguage{french}}
% \newcommand*{\langgerman}{\foreignlanguage{german}}
% \newcommand*{\langitalian}{\foreignlanguage{italian}}
% \newcommand*{\langswedish}{\foreignlanguage{swedish}}
% \newcommand*{\langlatin}{\foreignlanguage{latin}}
% \newcommand*{\langnohyph}{\foreignlanguage{nohyphenation}}
% \newcommand*{\latin}[1]{\langlatin{#1}}
% \newcommand*{\french}[1]{\emph{\langfrench{#1}}}
% \newcommand*{\german}[1]{\emph{\langgerman{#1}}}

\usepackage[autostyle=false,autopunct=false,english=british]{csquotes}
\setquotestyle{american}
\iffalse
\usepackage[shortlabels,inline]{enumitem}
% to replace paralist environments with enumitem ones:
\SetEnumitemKey{para}{itemindent=\parindent,leftmargin=0pt,listparindent=\parindent,parsep=0pt,itemsep=\topsep}
\setlist[enumerate,2]{label=\alph*.}
\setlist[enumerate]{leftmargin=\parindent}
\setlist[itemize]{leftmargin=\parindent}
\setlist[description]{leftmargin=\parindent}
\fi

\usepackage[backend=biber,mcite,subentry,citestyle=numeric-comp,bibstyle=numericbringhurst,autopunct=false,sorting=none,sortcites=false,natbib=false,maxnames=8,minnames=8,giveninits=true,block=space,hyperref=true,defernumbers=false,useprefix=true,language=british]{biblatex}
\renewcommand*{\finalnamedelim}{, }
\setcounter{biburlnumpenalty}{1}
\setcounter{biburlucpenalty}{0}
\setcounter{biburllcpenalty}{1}

\usepackage{graphicx}

\usepackage{hyperref}
\providecommand{\href}[2]{#2}
\providecommand{\eprint}[2]{\texttt{\href{#1}{#2}}}
\providecommand*{\urlalt}{\href}

\usepackage[depth=4]{bookmark}
\usepackage{mathdots}

%%%%%%%%%%%%%% macros
\newcommand*{\widebar}[1]{{\mkern 1.5mu\skew{2}\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}}
\newcommand*{\av}{\overline} %pop average
\newcommand*{\sav}{\widehat} %subpop average
\newcommand*{\yxx}{x}
\newcommand*{\yx}{\bm{\yxx}}%subpop state
\newcommand*{\yxs}{\sav{\yx}}%subpop av state
\newcommand*{\yxxs}{\sav{\yx\yx}}
\newcommand*{\yX}{\bm{X}}%pop state
\newcommand*{\yXf}{\av{\yX}}%pop av state
\newcommand*{\yr}{\bm{r}}%subpop value
\newcommand*{\yrs}{\sav{\yr}}%subpop av value
\newcommand*{\yR}{\bm{R}}%pop value
\newcommand*{\yRf}{\av{\yR}}%pop av value
\newcommand*{\yN}{N}
\newcommand*{\yM}{M}
\newcommand*{\yee}{L}

\newcommand*{\mee}{maximum entropy}
\newcommand*{\me}{maximum-entropy}
\title{Hidden network-size assumptions\\in the maximum-entropy method}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{ V. Rostami\\
  Forschungszentrum Jülich INM-6\\
  \texttt{v.rostami@fz-juelich.de}
  \And
  P.G.L. Porta Mana\\who knows
  \And E. Torre\\Chocolateland
  %
% David S.~Hippocampus\thanks{Use footnote for providing further information
%   about author (webpage, alternative
%   address)---\emph{not} for acknowledging funding agencies.} \\
% Department of Computer Science\\
% Cranberry-Lemon University\\
% Pittsburgh, PA 15213 \\
% \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch
  (3~picas) on both the left- and right-hand margins. Use 10~point
  type, with a vertical spacing (leading) of 11~points.  The word
  \textbf{Abstract} must be centered, bold, and in point size 12. Two
  line spaces precede the abstract. The abstract must be limited to
  one paragraph.
\end{abstract}

\section{Implicit assumptions in the maximum-entropy method}

The \me\ method is been used in neuroscience blahblah***

We would like to warn about an implicit assumption of the \me\ method,
which can informally be put this way:
\begin{quote}
  \emph{\me\ method applied to $\yN$ units assumes that no other
  units exist.}
\end{quote}
Less informally, the method implicitly assumes that the $\yN$ units are not a
subnetwork sample of a network with $\yM$ units, $\yM>\yN$.

This may come as a surprise. The method simply requires a number of
exhaustive and mutually exclusive events, and if these are composite events
the final distribution may have a multiplicity factor. When we consider the
$2^{\yN}$ states of $\yN$ units we are not excluding that these might be
marginals of $2^{\yM}$ states of $\yM$ units. Each one has the same
multiplicity $2^{\yM-\yN}$, but this constant multiplicity factor disappears
by normalization. So the method applies just as in the case of $\yN$ units
only, right?

Wrong. In our reasoning we have made subtle assumptions of independence
between the full network and the subnetwork. The problem is that the
counting of multiplicities is not based on simple enumeration, but already
involves probability considerations. Consider three cases with a full
population of two units, $\yM=2$, of which we consider one unit, $\yN=1$.
\begin{itemize}
\item First case: all four states are \emph{possible}. The two states of
  the first unit have multiplicity $2$ each. The usual \me\ distribution
  obtains.
\item Second case: only the states with at most one active unit are
  possible. The state $\yxx_1=1$ of the first unit has multiplicity
  $2$, and $\yxx_1=1$ has multiplicity $1$. The \me\ distribution has
  multiplicity factors.
\item Third case: states with at most one active unit are, say, $10^{9}$
  times more probable than the state with no active units. But all four
  states are \emph{possible}. By enumeration this case is like the first:
  multiplicities $(1,1)$. But by common sense it is more similar to the
  second: multiplicities $(2,1)$ for most practical purposes.
\end{itemize}

This simple example shows that the multiplicity inspection that must
precede a \me\ application already involve probability considerations at
the level of the full network. The usual reasoning by enumeration
implicitly assumes a uniform distribution, or a ***

We:
\begin{enumerate}
\item have a network with $\yN$ units, $2^{\yN}$ possible states
\item expect averages of $\yxs$ active neurons and $\yxxs$ active pairs
\item use maximum-entropy to choose a probability distribution for the
  states of the $\yN$ units conforming to our expectations.
\end{enumerate}

We:
\begin{enumerate}
\item have a network with $\yM$ units, $2^{\yM}$ possible states
\item expect that any subpopulation of $\yN$ units has $\yxs$ active
  neurons and $\yxxs$ active pairs
\item use maximum-entropy to choose a probability distribution for the
  states of the $\yM$ units conforming to our expectations
\item marginalize to find the probability distribution for the states of
  $\yN$ units.
\end{enumerate}

NIPS requires electronic submissions.  The electronic submission site
is
\begin{center}
  \url{https://cmt.research.microsoft.com/NIPS2017/}
\end{center}

Please read carefully the instructions below and follow them
faithfully.

\subsection{Style}

Papers to be submitted to NIPS 2017 must be prepared according to the
instructions presented here. Papers may only be up to eight pages
long, including figures. This does not include acknowledgments and
cited references which are allowed on subsequent pages.
Papers that exceed these limits will not be reviewed, or in any
other way considered for presentation at the conference.

The margins in 2017 are the same as since 2007, which allow for
$\sim$$15\%$ more words in the paper compared to earlier years.

Authors are required to use the NIPS \LaTeX{} style files obtainable
at the NIPS website as indicated below. Please make sure you use the
current files and not previous versions. Tweaking the style files may
be grounds for rejection.

\subsection{Retrieval of style files}

The style files for NIPS and other conference information are
available on the World Wide Web at
\begin{center}
  \url{http://www.nips.cc/}
\end{center}
The file \verb+nips_2017.pdf+ contains these instructions and
illustrates the various formatting requirements your NIPS paper must
satisfy.

The only supported style file for NIPS 2017 is \verb+nips_2017.sty+,
rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{}
  2.09, Microsoft Word, and RTF are no longer supported!}

The new \LaTeX{} style file contains two optional arguments:
\verb+final+, which creates a camera-ready copy, and \verb+nonatbib+,
which will not load the \verb+natbib+ package for you in case of
package clash.

At submission time, please omit the \verb+final+ option. This will
anonymize your submission and add line numbers to aid review.  Please
do \emph{not} refer to these line numbers in your paper as they will
be removed during generation of camera-ready copies.

The file \verb+nips_2017.tex+ may be used as a ``shell'' for writing
your paper. All you have to do is replace the author, title, abstract,
and text of the paper with your own.

The formatting instructions contained in these style files are
summarized in Sections \ref{gen_inst}, \ref{headings}, and
\ref{others} below.

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas)
wide and 9~inches (54~picas) long. The left margin is 1.5~inch
(9~picas).  Use 10~point type with a vertical spacing (leading) of
11~points.  Times New Roman is the preferred typeface throughout, and
will be selected for you by default.  Paragraphs are separated by
\nicefrac{1}{2}~line space (5.5 points), with no indentation.

The paper title should be 17~point, initial caps/lower case, bold,
centered between two horizontal rules. The top rule should be 4~points
thick and the bottom rule should be 1~point thick. Allow
\nicefrac{1}{4}~inch space above and below the title to rules. All
pages should start at 1~inch (6~picas) from the top of the page.

For the final version, authors' names are set in boldface, and each
name is centered above the corresponding address. The lead author's
name is to be listed first (left-most), and the co-authors' names (if
different address) are set to follow. If there is only one co-author,
list both author and co-author side by side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

All headings should be lower case (except for first word and proper
nouns), flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the
heading in bold, flush left, and inline with the text, with the
heading followed by 1\,em of space.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.
Citations may be author/year or numeric, as long as you maintain
internal consistency.  As to the format of the references themselves,
any style is acceptable as long as it is used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may
add the following before loading the \verb+nips_2017+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add
the optional argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{nips_2017}
\end{verbatim}

As submission is double blind, refer to your own published work in the
third person. That is, use ``In the previous work of Jones et
al.\ [4],'' not ``In our previous work [4].'' If you cite your other
papers that are not widely available (e.g., a journal paper under
review), use anonymous author names in the citation, e.g., an author
of the form ``A.\ Anonymous.''

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote,
indicate footnotes with a number\footnote{Sample of the first
  footnote.} in the text. Place the footnotes at the bottom of the
page on which they appear.  Precede the footnote with a horizontal
rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

All artwork must be neat, clean, and legible. Lines should be dark
enough for purposes of reproduction. The figure number and caption
always appear after the figure. Place one line space before the figure
caption and one line space after the figure. The figure caption should
be lower case (except for first word and proper nouns); figures are
numbered consecutively.

You may use color figures.  However, it is best for the figure
captions and the paper body to be legible if the paper is printed in
either black/white or in color.
\begin{figure}[h]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table
number and title always appear before the table.  See
Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical
  rules.} We strongly suggest the use of the \verb+booktabs+ package,
which allows for typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}[t]
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule{1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Final instructions}

Do not change any aspects of the formatting parameters in the style
files.  In particular, do not modify the width or length of the
rectangle the text should fit into, and do not change font sizes
(except perhaps in the \textbf{References} section; see below). Please
note that pages should be numbered.

\section{Preparing PDF files}

Please prepare submission files with paper size ``US Letter,'' and
not, for example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file
must only contain Type 1 or Embedded TrueType fonts. Here are a few
instructions to achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader,
  select the menu Files$>$Document Properties$>$Fonts and select Show
  All Fonts. You can also use the program \verb+pdffonts+ which comes
  with \verb+xpdf+ and is available out-of-the-box on most Linux
  machines.

\item The IEEE has recommendations for generating PDF files whose
  fonts are also acceptable for NIPS. Please see
  \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item \verb+xfig+ "patterned" shapes are implemented with bitmap
  fonts.  Use "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You
  should use the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or
\verb+\mathbb{C}+ for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You
can also use the following workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the
\verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we
will ask you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always
specify the figure width as a multiple of the line width as in the
example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly
hyphenate a line. Please give LaTeX hyphenation hints using the
\verb+\-+ command when necessary.

\subsubsection*{Acknowledgments}

Use unnumbered third level headings for the acknowledgments. All
acknowledgments go at the end of the paper. Do not include
acknowledgments in the anonymized submission, only in the final paper.

\section*{References}

References follow the acknowledgments. Use unnumbered first-level
heading for the references. Any choice of citation style is acceptable
as long as you are consistent. It is permissible to reduce the font
size to \verb+small+ (9 point) when listing the references. {\bf
  Remember that you can go over 8 pages as long as the subsequent ones contain
  \emph{only} cited references.}
\medskip

\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
  Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
  Exploring Realistic Neural Models with the GEneral NEural SImulation
  System.}  New York: TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
learning and recall at excitatory recurrent synapses and cholinergic
modulation in rat hippocampal region CA3. {\it Journal of
  Neuroscience} {\bf 15}(7):5249-5262.

\end{document}

### 2017-05-04 nips_2017.tex
update luca's bib

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

\usepackage[nonatbib]{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{mathtools}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{lmodern}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsxtra}
\usepackage[babel,theoremfont]{newpxtext}
\usepackage[bigdelims,nosymbolsc]{newpxmath}
\usepackage{bm}
\usepackage{mathdots}
\usepackage[usenames]{xcolor}
\usepackage[defblank,neveradjust]{paralist}
\usepackage[shortlabels]{enumitem}
\usepackage[british]{babel}
\selectlanguage{british}
\newcommand*{\langfrench}{\foreignlanguage{french}}
\newcommand*{\langgerman}{\foreignlanguage{german}}
\newcommand*{\langitalian}{\foreignlanguage{italian}}
\newcommand*{\langswedish}{\foreignlanguage{swedish}}
\newcommand*{\langlatin}{\foreignlanguage{latin}}
\newcommand*{\langnohyph}{\foreignlanguage{nohyphenation}}
\newcommand*{\latin}[1]{\langlatin{#1}}
\newcommand*{\french}[1]{\emph{\langfrench{#1}}}
\newcommand*{\german}[1]{\emph{\langgerman{#1}}}
\usepackage[autostyle=false,autopunct=false,english=british]{csquotes}
\setquotestyle{american}

\usepackage[%
backend=biber,%
mcite,%
subentry,%
citestyle=numeric-comp,%
bibstyle=numericbringhurst,%
%bibstyle=chem-biochem,%
autopunct=false,%
sorting=none,%
sortcites=false,%
%style=verbose,%
natbib=false,
maxnames=8,minnames=8,% maxnames=2,minnames=1,
giveninits=true,%
block=space,%
hyperref=true,%
defernumbers=false,%
%refsegment=chapter,%
useprefix=true%
,language=british%
]{biblatex}
\DeclareDelimFormat{multicitedelim}{\addsemicolon\space}
\DeclareDelimFormat{postnotedelim}{\space}
\addbibresource{portamanabib.bib}
\renewcommand{\bibfont}{\footnotesize}
\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
}
\renewcommand*{\finalnamedelim}{, }
\setcounter{biburlnumpenalty}{1}
\setcounter{biburlucpenalty}{0}
\setcounter{biburllcpenalty}{1}
\newcommand*{\citep}{\parencites}
\newcommand*{\citey}{\parencites*}
\renewcommand{\cite}{\citep}
\providecommand{\href}[2]{#2}
\providecommand{\eprint}[2]{\texttt{\href{#1}{#2}}}
%\renewcommand{\eprint}[2]{\texttt{\href{#1#2}{#2}}}

%\DeclareBibliographyCategory{extras}

\def\arxivp{}
\def\mparcp{}
\def\philscip{}
\def\biorxivp{}

\providecommand*{\urlalt}{\href}

\newcommand*{\arxivsi}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.\\}
\newcommand*{\mparcsi}{\texttt{mp\_arc} eprints available at \url{http://www.ma.utexas.edu/mp_arc/}.\\}
\newcommand*{\philscisi}{\texttt{philsci} eprints available at \url{http://philsci-archive.pitt.edu/}.\\}
\newcommand*{\biorxivsi}{\texttt{bioRxiv} eprints available at \url{http://biorxiv.org/}.\\}

\newcommand*{\arxiveprint}[1]{\global\def\arxivp{\arxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
\texttt{\urlalt{http://arxiv.org/abs/#1}{arXiv:\bd #1}}%
%\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
%\renewcommand{\arxivnote}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.}
}

\newcommand*{\mparceprint}[1]{\global\def\mparcp{\mparcsi}%\citeauthor{0mparccite}\addtocategory{ifarchcit}{0mparccite}%eprint
\texttt{\urlalt{http://www.ma.utexas.edu/mp_arc-bin/mpa?yn=#1}{mp\_arc:\bd #1}}%
%\texttt{\href{http://www.ma.utexas.edu/mp_arc-bin/mpa?yn=#1}{\protect\url{mp_arc:#1}}}%
%\providecommand{\mparcnote}{\texttt{mp_arc} eprints available at \url{http://www.ma.utexas.edu/mp_arc/}.}
}

\newcommand*{\philscieprint}[1]{\global\def\philscip{\philscisi}%\citeauthor{0philscicite}\addtocategory{ifarchcit}{0philscicite}%eprint
\texttt{\urlalt{http://philsci-archive.pitt.edu/archive/#1}{PhilSci:\bd #1}}%
%\texttt{\href{http://philsci-archive.pitt.edu/archive/#1}{\protect\url{PhilSci:#1}}}%
%\providecommand{\mparcnote}{\texttt{philsci} eprints available at \url{http://philsci-archive.pitt.edu/}.}
}

\newcommand*{\biorxiveprint}[1]{\global\def\biorxivp{\biorxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
\texttt{\urlalt{http://biorxiv.org/content/early/#1}{bioRxiv:\bd #1}}%
%\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
%\renewcommand{\arxivnote}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.}
}

\newtheoremstyle{innote}%
{\parsep}%
{\parsep}%
{\footnotesize}%
{}%
{}%
{}%
{0pt}%
{}%
\theoremstyle{innote}
\newtheorem*{innote}{}

\title{Inferences from a network to a subnetwork, and vice versa, under an assumption of symmetry}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  David S.~Hippocampus\thanks{Use footnote for providing further
    information about author (webpage, alternative
    address)---\emph{not} for acknowledging funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

%@@@@@@@@@@@@@@@@@@@@@@@ new commands @@@@@@@@@@@@@@@@@@@@@

\providecommand{\defaultlists}{}
%\newcommand*{\av}{\widebar} %pop average
\newcommand*{\av}{\overline} %pop average
\newcommand*{\sav}{\widehat} %subpop average

\newcommand*{\yx}{\bm{x}}%subpop state
\newcommand*{\yxs}{\sav{\yx}}%subpop av state
\newcommand*{\yX}{\bm{X}}%pop state
\newcommand*{\yXf}{\av{\yX}}%pop av state
\newcommand*{\yr}{\bm{r}}%subpop value
\newcommand*{\yrs}{\sav{\yr}}%subpop av value
\newcommand*{\yR}{\bm{R}}%pop value
\newcommand*{\yRf}{\av{\yR}}%pop av value
%conditional assumptions
\newcommand*{\yH}{\varEta}
\newcommand*{\yD}{\varDelta}
\newcommand*{\yHa}{\varEta'}
\newcommand*{\yHb}{\varEta''}
%Lagr multipliers
\newcommand*{\yL}{\varLambda}
\newcommand*{\yl}{\lambda}
\newcommand*{\yk}{\kappa}
%entropies
\newcommand*{\ysh}{H}
\newcommand*{\ybu}{H_{\text{B}}}

\newcommand*{\prop}[1]{`#1'}
\DeclarePairedDelimiter\set{\{}{\}}
\newcommand*{\amp}{\&}

%Latin abbrev.
\newcommand*{\vs}{{vs.}}
\newcommand*{\etc}{{etc.}}
\newcommand*{\ie}{{i.e.}}
\newcommand*{\ca}{{c.}}
\newcommand*{\Ie}{{I.e.}}
\newcommand*{\Eg}{{E.g.}}
\newcommand*{\eg}{{e.g.}}
\newcommand*{\viz}{{viz}}
\newcommand*{\cf}{{cf.}}
\newcommand*{\Cf}{{Cf.}}
\newcommand*{\vd}{{v.}}
\newcommand*{\Vd}{{V.}}
\newcommand*{\etal}{{et al.}}
\newcommand*{\etsim}{{et sim.}}
\newcommand*{\ibid}{{ibid.}}
\newcommand*{\sic}{{sic}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand*{\defd}{\coloneqq}
\newcommand*{\defs}{\eqqcolon}
\newcommand*{\Land}{\bigwedge}
\newcommand*{\Lor}{\bigvee}
\newcommand*{\lland}{\mathbin{\ \land\ }}
\newcommand*{\llor}{\mathbin{\ \lor\ }}
\newcommand*{\lonlyif}{\mathbin{\Rightarrow}}%implies
\newcommand*{\limplies}{\mathbin{\Rightarrow}}%implies
\newcommand*{\mimplies}{\Rightarrow}%implies
\newcommand*{\liff}{\mathbin{\Leftrightarrow}}%if and only if
%\newcommand*{\lequal}{=}%identical propositions
%\newcommand*{\ldef}{\defd}%equivalent by def.
%\newcommand*{\ldefin}{\stackrel{_\text{def}}{\Longleftrightarrow}}
\newcommand*{\cond}%{|}%
{\mathpunct{|}}%conditional sign (in probabilities)
\newcommand*{\lcond}%{|}%
{\mathpunct{|\ }}%conditional sign (in probabilities)
\newcommand*{\bigcond}{\mathpunct{\big|}}%conditional sign (in probabilities)
\newcommand*{\lbigcond}{\mathpunct{\big|\ }}%conditional sign (in probabilities)
\newcommand*{\suchthat}{\mid}%{\mathpunct{|}}%such that (eg in sets)
\newcommand*{\bigst}{\mathpunct{\big|}}%such that (eg in sets)
\newcommand*{\with}{\colon}%with (list of indices)
\newcommand*{\mul}{\times}%multiplication
\newcommand*{\inn}{\cdot}%inner product
\newcommand*{\dotv}{\mathord{\,\cdot\,}}%variable place
%\newcommand*{\cross}{\times}%vector product
%\newcommand*{\cart}{\times}%cartesian product
\newcommand*{\comp}{\circ}%composition of functions
%\DeclareMathOperator{\conv}{conv}%convex hull
\DeclareMathOperator*{\argmax}{arg\,max}%convex hull
%\newcommand*{\convl}{*}%convolution
\newcommand*{\con}{\mathbin{:}}%scal prod of tensors
%\newcommand*{\tpro}{\otimes}%tens prod
\newcommand*{\equi}{\sim}%equivalent to
\newcommand*{\corr}{\mathrel{\hat{=}}}%corresponds to
\providecommand{\varparallel}{\ensuremath{\mathbin{/\mkern-7mu/}}}%parallel (tentative symbol)
%\newcommand*{\paral}{\varparallel}%parallel (tentative symbol)
%\newcommand*{\asym}{\simeq}%asymptotically equal to
%\newcommand*{\prop}{\simeq}%asymptotically equal to
\renewcommand{\le}{\leqslant}%less or equal
\renewcommand{\ge}{\geqslant}%greater or equal

\newcommand*{\vect}[1]{\bm{#1}}%vectors
%probability
\DeclareMathOperator{\pr}{P}%probability
\newcommand*{\pf}{\mathrm{p}}%probability
%\DeclareMathOperator{\G}{\Gamma}%probability density
\newcommand*{\p}{\mathrm{P}}%probability
\newcommand*{\tf}{\mathrm{T}}%probability
\renewcommand*{\|}{\cond}
\newcommand*{\+}{\lor}
\newcommand*{\un}{\;\mathrm}
%\newcommand*{\comma}{\mathord{.}}% space
%\newcommand*{\uni}[1]{\textnormal{#1}}% unity
%\newcommand*{\tim}{\cdot}% multipl
%\newcommand*{\per}{/}% per
%Sections and theorem-like environments
\newcommand*{\sect}{\S}% Sect.~
\newcommand*{\sects}{\S\S}% Sect.~
\newcommand*{\chap}{ch.}%
\newcommand*{\chaps}{chs}%
\newcommand*{\fn}{fn}%
\newcommand*{\eqn}{eq.}%
\newcommand*{\eqns}{eqs}%
\newcommand*{\fig}{fig.}%
\newcommand*{\figs}{figs}%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*{\ordm}[2]{(#1,#2)}%matrix order
%\DeclareMathOperator{\rank}{rank}%rank
%\DeclareMathOperator{\range}{range}%range
\newcommand*{\id}{\mathte{I}}%id matrix
\newcommand*{\nbd}{\nobreakdash}%
\newcommand*{\bd}{\hspace{0pt}}%
\def\hy{-\penalty0\hskip0pt\relax}

\newcommand*{\labelbis}[1]{\tag*{(\ref{#1})$_\text{r}$}}
\newcommand*{\mathbox}[2][.8]{\parbox[t]{#1\columnwidth}{#2}}
\newcommand*{\zerob}[1]{\makebox[0pt][l]{#1}}

\newcommand*{\rtext}[1]{\mathrel{\text{#1}}}

\let\varAlpha A
\newcommand*{\Beta}{\mathrm{B}}
\let\varBeta B
\newcommand*{\Epsilon}{\mathrm{E}}
\let\varEpsilon E
\newcommand*{\Zeta}{\mathrm{Z}}
\let\varZeta Z
\newcommand*{\Eta}{\mathrm{H}}
\let\varEta H
\newcommand*{\Iota}{\mathrm{I}}
\let\varIota I
\newcommand*{\Kappa}{\mathrm{K}}
\let\varKappa K
\newcommand*{\Mu}{\mathrm{M}}
\let\varMu M
\newcommand*{\Nu}{\mathrm{N}}
\let\varNu N
\newcommand*{\Omicron}{\mathrm{O}}
\let\varOmicron O
\newcommand*{\Rho}{\mathrm{P}}
\let\varRho P
\newcommand*{\Tau}{\mathrm{T}}
\let\varTau T
\newcommand*{\Chi}{\mathrm{X}}
\let\varChi X
\newcommand*{\omicron}{o}

\newcommand*{\di}{\mathrm{d}}%differential
\newcommand*{\RR}{\mathbb{R}}

\newcommand*{\tprod}{\mathop{\textstyle\prod}\nolimits}
\newcommand*{\tsum}{\mathop{\textstyle\sum}\nolimits}
\newcommand*{\tint}{\begingroup\textstyle\int\endgroup\nolimits}
\newcommand*{\tland}{\mathop{\textstyle\bigwedge}\nolimits}
\newcommand*{\tlor}{\mathop{\textstyle\bigvee}\nolimits}
\newcommand*{\sprod}{\mathop{\textstyle\prod}}
\newcommand*{\ssum}{\mathop{\textstyle\sum}}
\newcommand*{\sint}{\begingroup\textstyle\int\endgroup}
\newcommand*{\sland}{\mathop{\textstyle\bigwedge}}
\newcommand*{\slor}{\mathop{\textstyle\bigvee}}
%\newcommand*{\lsum}{\mathop{\textstyle\sum}\limits}
%\newcommand*{\lprod}{\mathop{\textstyle\prod}\limits}
\newcommand*{\T}{^\intercal}%transpose
%\newcommand*{\iT}{^{-\mathord{\intercal}}}
%\newcommand*{\herm}{^\dagger}%hermitian conj
%\newcommand*{\ccon}{{}^{\mathord{*}}}%complex conj
%\DeclareMathOperator{\ve}{T}
\newcommand*{\E}{\mathrm{E}}
\DeclarePairedDelimiter\expp{(}{)}
\newcommand*{\expe}{\E\expp}%round
\newcommand*{\expeb}{\E\clcl}%square
%\renewcommand{\expe}[1]{\E(#1)}
%\newcommand*{\aver}{\overline}
%\newcommand*{\cont}{\models} %\vdash or \models
\newcommand*{\QEM}%{\textnormal{$\Box$}}
{\ding{167}}
\newcommand*{\qem}{\leavevmode\unskip\penalty9999 \hbox{}\nobreak\hfill
  \quad\hbox{\QEM}}
\DeclareMathOperator{\sgn}{sgn}%signum
\DeclareMathOperator{\ent}{ent}%integer less or equal to
\DeclareMathOperator{\Ord}{O}%same order as
\DeclareMathOperator{\ord}{o}%lower order than
\DeclarePairedDelimiter\clcl{[}{]}
\DeclarePairedDelimiter\clop{[}{[}
\DeclarePairedDelimiter\opcl{]}{]}
\DeclarePairedDelimiter\opop{]}{[}

\newcommand*{\pu}{\piup}%constant pi
\newcommand*{\delt}{\deltaup}%Kronecker, Dirac

%@@@@@@@@@@@@@@@@@@@@@@@ new commands end @@@@@@@@@@@@@@@@@@

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}\labelsep 0pt%
\noindent
This note summarizes some mathematical relations between the probability
distributions for the states of a network of binary units and a subnetwork
thereof, under an assumption of symmetry. These relations are standard
results of probability theory, but seem to be rarely used in neuroscience.
Some of their consequences for inferences between network and subnetwork,
especially in connection with the maximum-entropy principle, are briefly
discussed. The meanings and applicability of the assumption of symmetry
are also discussed.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% A pawn on a chessboard
% A false move by god will now destroy me
% But wait, on the horizon
% a new dawn seems to be rising
% never to recall this passerby, born to die.

\section{Uncertainties about networks and subnetworks}
\label{sec:intro}

If we are uncertain about the state of a network of neurons,
what is our uncertainty about the state of a subnetwork? And
vice versa, if we are uncertain about the state of a subnetwork, what is
our uncertainty about the state of the whole network?

If our uncertainties are expressed by two probability distributions for
the states of network and subnetwork, and these distributions satisfy a
particular symmetry property, then they are related by precise and relatively
simple mathematical formulae. These formulae are of essential importance
when we want to make inferences about the whole network given data about
the subnetwork and vice versa.

While well-known in survey sampling and in the pedagogic problem of
\enquote{drawing from an urn without replacement}, such formulae are
somewhat hard to find explicitly written in the neuroscientific literature.
We have therefore decided to summarize them in this Note, as a reference,
for the case of neurons modelled as \enquote{units} with binary
$\set{0,1} \equiv \set{\text{inactive}, \text{active}}$ states. The
formulae also apply to any network or population of any kind of units
modelled in the same way: \eg, the values of the physical
connectivities of a neuronal network.

The formulae and the assumptions underlying them are given in the next two
sections, followed by examples of their use in both inferential directions:
network to subnetwork and reverse. We also discuss the meanings of the
symmetry property, and conclude with a brief summary. Proofs of the
formulae are sketched in an appendix.

Our notation follows ISO \citep{iso1993}, ANSI/IEEE \citep{ieee1993}, and
NIST \citep{nist1995} standards but for the use of the comma \enquote{,} to
denote logical conjunction (\enquote{and}, \enquote{$\land$}), for the sake
of typographic poise. %clarity
%and not to scare away readers unfamiliar with formal logic.
Probability-calculus notation follows Jaynes's book \citep{jaynes1994_r2003}.

\section{Formulae connecting network and subnetwork probabilities}
\label{sec:setup_assumptions}

\subsection{Setup}
\label{sec:setup}

Consider a network of $N$ binary neurons with states $(X_1,\dotsc,X_N)$
having fixed but unknown binary values $(R_1,\dotsc,R_N)$, with
$R_i \text{ in } \set{0,1}$, vectorially written $\yX=\yR$. For example,
$\yX$ can represent the state of a network at a particular time. We call
the neurons \enquote{units} to lend some generality to our discussion. We
shall make statements about the whole network of $N$ units and about a
subnetwork of $n$ units; the word \enquote{network} will always refer to
the \emph{whole} network. The subnetwork states and their values are
denoted by lowercase letters: $(x_1, \dotsc, x_n)\equiv\yx$ and
$(r_1, \dotsc, r_n)\equiv\yr$; but note that $x_i \equiv X_{j_i}$ and
$r_i \equiv R_{j_i}$ for some distinct $j_1,\dotsc,j_n$. We shall also make
statements about the network-averaged state, or \emph{network average}:
\begin{align}
  \label{eq:network-av}
   \yXf &\defd (X_1 + \dotsb + X_N)/N,
\\
\intertext{and the subnetwork-averaged state, or \emph{subnetwork average}:}
\label{eq:subnetwork-av}
\yxs &\defd (x_1 + \dotsc + x_n)/n.
\end{align}
The quantities $N\yXf$ and $n\yxs$ represent the total number of active
units in the network and the subnetwork. Quantities like $\yRf$ and $\yrs$
are defined analogously. The averaging operators $\sav{\dotv}$ and
$\av{\dotv}$ are also extended to averages of $\tbinom{n}{m}$ or
$\tbinom{N}{m}$ products of $m$ states; \eg,
\begin{align}
\av{\yX \yX} &\defd
\tbinom{N}{2}^{-1} (X_1 X_2 + X_1 X_3  + \dotsb +  X_{N-1} X_N),
\\[2\jot]
\sav{\yx\yx\yx} &\defd
\tbinom{n}{3}^{-1} (x_1 x_2 x_3 + x_1 x_2 x_4 + \dotsb + x_{n-2} x_{n-1} x_n),
\end{align}
and so on.

\subsection{Assumptions}
\label{sec:assumptions}

Our uncertainty about the network state is represented by the joint
probability distribution of the individual states, from which we can
derive all other probabilities of interest. We denote it by
\begin{equation}
  \label{eq:joint_plaus}
  \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \cond \yH) \quad\text{or}\quad
\p(\yX =\yR \cond \yH).
\end{equation}
Such probability is conditional on our state of knowledge, \ie\ the
evidence and assumptions backing our probability assignments, denoted by
the proposition $\yH$.

In the present discussion, $\yH$ is a state of knowledge that leads to two
specific properties in our probability assignments:
\defaultlists

\medskip
\begin{asparaenum}[$\yH$1.]
\item \emph{Permutation symmetry}, expressed as the invariance of the
  joint distribution~\eqref{eq:joint_plaus} under arbitrary permutations of
  the units's labels:
\begin{multline}
  \label{eq:homog_math}
  \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \cond \yH) ={}
\\
\shoveright{\p(X_1=R_{\pi(1)}, X_2=R_{\pi(2)}, \dotsc, X_N=R_{\pi(N)} \cond \yH)}
\\
\text{for any permutation $\pi$}.
\end{multline}
This property can reflect two very different states of knowledge: physical
homogeneity of the network, or symmetry in our ignorance about the network.
This property is called \emph{finite exchangeability} in the Bayesian
literature and its basis, consequences, and alternatives to it are discussed
in \sect~\ref{sec:stoch_symm}.

\medskip

\item The network average $\yXf$ has a particular distribution $Q$:
\begin{equation}
  \label{eq:pop_average_Q}
  \p(\yXf = A \cond \yH)
% \equiv \pr(x_1+ \dotsb + x_N = N A \cond \varEta)
=  Q(A),
\qquad
A \in \set*{0,\tfrac{1}{N},\tfrac{2}{N},\dotsc,1}.
\end{equation}
For the moment we are not concerned about the specific form of $Q$ and
about how it was assigned: it could, \eg, arise from maximum-entropy
arguments
\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,aczeletal1975,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013}
used with data on the network.
\end{asparaenum}

\subsection{Formulae}
%\section{\ding{41} {\itshape The mathematical relations} \ding{41}}
%\section*{3.\hspace{\stretch{1}}The mathematical relations\hspace{\stretch{1}}\mbox{}}
%\addcontentsline{toc}{section}{The mathematical relations}
\label{sec:main_formulae}

The state of knowledge $\yH$ has the following six (not independent) main
consequences for our probability assignments:

\begin{asparaenum}[\bfseries\itshape I.]
  \medskip\item\label{item:tot_plaus}Probability for the network state:
\begin{equation}
  \label{eq:joint_plaus_N_homog}
  \p(\yX = \yR \cond H) = \binom{N}{N\yRf}^{-1} Q(\yRf).
\end{equation}

\medskip
\item\label{item:marginal}Probability for the state $\yx$ of any subnetwork
  of $n$ units:
\begin{equation}
  \label{eq:marginal}
    \p(\yx = \yr \cond \yH) =
\sum_{N A =0}^{N}
\binom{N-n}{N A-n\yrs}\binom{N}{N A}^{-1}\,
Q(A).
\end{equation}
Note that the only summands contributing to this sum are those for which
$n\yrs \le NA \le N$; the others are zero because by definition
$\tbinom{M}{y}=0$ if $y<0$. This remark applies to all the sums of this
kind in the rest of this Note.

\medskip
\item\label{item:conditional}Probability for the subnetwork
  state conditional on a network state:
\begin{equation}
  \label{eq:conditional_total_state}
  \p(\yx = \yr \cond \yX = \yR, \yH)
=
\binom{N-n}{N \yRf-n \yrs}.
\end{equation}

\medskip
\item\label{item:marginal_average}Probability for the
  subnetwork average $\yxs$:
\begin{multline}
  \label{eq:subpop_average}
  \p(\yxs = a  \cond \yH) = \binom{n}{n a}
 \sum_{N A=0}^{N}
\binom{N-n}{N A-n a}\binom{N}{N A}^{-1}\,
Q(A),
\\
a \in \set*{0,\tfrac{1}{n},\tfrac{2}{n},\dotsc,1}.
\end{multline}

\medskip
\item\label{item:conditional_average}Probability for the subnetwork
  average conditional on the network average:
\begin{equation}
  \label{eq:conditional_prob}
  \p(\yxs = a \cond \yXf = A, \yH)
=
\binom{n}{n a}\binom{N-n}{N A-n a}\binom{N}{N A}^{-1}.
\end{equation}

\medskip
\item\label{item:moments}The product of the states of any $m$ distinct
  units from a given subnetwork,
  \begin{equation*}
    x_{i_1} x_{i_2} \dotsm x_{i_m},
    \qquad 1\le i_1 < i_2 < \dotsb < i_m \le n
  \end{equation*}
  has an expectation equal to that of the subnetwork average of such
  products, is independent of the subnetwork size $n$:
\begin{subequations}
\label{eq:expe_products}
\begin{align}
  \expe{x_{i_1} \dotsm x_{i_m} \cond \yH}
&=
\expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \cond \yH}
=
\expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \cond \yH},
\\
\intertext{and has an explicit expression in terms of $Q$:}
  \begin{split}
\expe{x_{i_1} \dotsm x_{i_m} \cond \yH}
&=
\binom{N}{m}^{-1}
%\frac{(N-m)!}{N!}
\sum_{N A=0}^N %m!\,
\binom{N A}{m}\, Q(A)
\\
&\equiv
%\frac{(N-m)!}{N!}
\sum_{N A=0}^N %m!\,
\binom{N-m}{N A-m}
\binom{N}{N A}^{-1} Q(A).
\end{split}
\end{align}
\end{subequations}
\end{asparaenum}

\medskip

A useful relation connects the expectation of a
product~\eqref{eq:expe_products} and the $m$th \emph{factorial moment}
\citep{potts1953} of the probability distributions for the averages. The
$m$th factorial moment of the subnetwork average $\yxs$ is defined by
\begin{equation}
  \label{eq:fact_mom_defin}
\vphantom{\underbrace{\yxs}_{m}}\expeb*{
\smash{\underbrace{n\yxs\,(n\yxs-1)\dotsm \bigl(n\yxs-(m-1)\bigr)}_{\text{$m$ factors}}}
\cond \yH}
\equiv \expeb*{\frac{(n\yxs)!}{(n\yxs-m)!} \cond \yH},
\end{equation}
an analogous definition holding for $\yXf$. We have that
\begin{equation}
  \label{eq:eq_factmom_prod}
\expe{x_{i_1} \dotsm x_{i_m} \cond \yH}
=
  \tfrac{(n-m)!}{n!}\expeb*{\frac{(n\yxs)!}{(n\yxs-m)!} \cond \yH}
=
  \tfrac{(N-m)!}{N!}\expeb*{\frac{(N\yXf)!}{(N\yXf-m)!} \cond \yH}.
\end{equation}

As a consequence of the above relation, the first three moments of the
probability distributions $\p(\yxs=a\cond\yH)$ and $\p(\yXf=A\cond\yH)$,
are related by
\begin{subequations}\label{eq:moments}
  \begin{align}
    \expe{\yxs \cond \yH} &= \expe{\yXf \cond \yH},
    \\[2\jot]
    \begin{split}
    \expe{\yxs^2 \cond \yH} &= \expe{\yXf \cond \yH}
                                  \frac{N-n}{(N-1)n}
 +{}
\\&\quad\quad \expe{\yXf^2 \cond \yH}\frac{N\,(n-1)}{(N-1)n},
\end{split}
    \\[2\jot]
    \begin{split}
    \expe{\yxs^3 \cond \yH} &=
                              \expe{\yXf \cond \yH}\frac{(N-n)(N-2n)}{(N-1)(N-2)n^2}+{}
                                    \\
                                    &\quad\quad\expe{\yXf^2 \cond \yH}
                                    \frac{3N\,(N-n)(n-1)}{(N-1)(N-2)n^2} +{}
                                    \\
                                    &\quad\quad\quad\expe{\yXf^3 \cond \yH}
                                    \frac{N^2\,(n-1)(n-2)}{(N-1)(N-2)n^2}.
                                  \end{split}
                                    % \\[2\jot]
                                    % \expe{\yxs^4 \cond \varEta} &=
                                    % \!\begin{aligned}[t] &\expe{\yxf
                                    %   \cond \varEta}
                                    %   \frac{(N-n)(N^2+N-6Nn+6n^2)}{(N-1)(N-2)(N-3)n^3}+{}
                                    %   \\%[1.5\jot]
                                    %   &\quad\expe{\yxf^2 \cond \varEta}
                                    %   \frac{(7N-11n+1)(N-n)(n-1)N}{(N-1)(N-2)(N-3)n^3}+{}
                                    %   \\
                                    %   &\quad\expe{\yxf^3 \cond \varEta}
                                    %   \frac{6(n-1)(n-2)(N-n)N^2}{(N-1)(N-2)(N-3)n^3}
                                    %   +{}
                                    %   \\
                                    %   &\quad\expe{\yxf^4 \cond \varEta}
                                    %   \frac{(n-1)(n-2)(n-3)N^3}{(N-1)(N-2)(N-3)n^3}.
                                    % \end{aligned}
  \end{align}
\end{subequations}
Relations for higher moments can be obtained recursively from
\eqn~\eqref{eq:eq_factmom_prod}. In general, this means that the two sets of
first $m$ moments are related by a homogeneous linear transformation,
\begin{equation}
  \label{eq:lin_transf_moments}
  \expe{\yxs^m \cond \yH} = \sum_{l=1}^m M_{ml}(n,N) \expe{\yXf^l \cond \yH},
\end{equation}
with a universal, lower-triangular transformation matrix $M_{ml}(n,N)$ that
depends only on $n$, $N$, and the condition of symmetry~\eqref{eq:homog_math}.

As intuition suggests, we have
%$\expe{\yxs^m \cond \yH} \xrightarrow{n\to N} \expe{\yXf^m \cond \yH}$.
\begin{equation}
  \label{eq:limit_moments}
  \expe{\yxs^m \cond \yH} \xrightarrow{n\to N}
  \expe{\yXf^m \cond \yH},
\qquad
  \expe{\yxs^m \cond \yH} \xrightarrow{n\to 1}
  \expe{\yXf \cond \yH},
\end{equation}
the latter because ${x_i}^m=x_i$, since states are $\set{0,1}$-valued.

\bigskip

The core of the six mathematical relations above are
\eqns~\eqref{eq:marginal} and~\eqref{eq:subpop_average}. The latter
expresses the probability for the subnetwork average as a mixture of
hypergeometric distributions
\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
with parameters $N,N\yXf,n$, weighted by the probabilities
$\p(\yXf = A \cond\yH)$ \citep[\cf][\sect~4, esp.\ \eqn~(22)]{kendall1967}.
The connection between this mixture representation and the condition of
symmetry~\eqref{eq:homog_math} is well-known in the Bayesian literature
\citep{kendall1967,definetti1969b,heathetal1976,diaconis1977,diaconisetal1980,jaynes1986c}.

Proofs of the above formulae are sketched in appendix
\ref{sec:derivations}.

\subsection{Asymptotic approximations and generalizations}
\label{sec:limits}

Recall the definition of the Shannon and Burg
\citep{burg1975,ericksonetal1988b} entropies for a binary probability
distribution $(p,1-p)$ with $p$ in $\clcl{0,1}$
\begin{equation}
  \label{eq:entropies}
\ysh(p) \defd -p\ln p -(1-p)\ln(1-p),
\qquad
\ybu(p) \defd \ln p +\ln(1-p).
\end{equation}
The binomial coefficient has an asymptotic form that involves the two
entropies above
\citep{robbins1955,weissman1983,mermin1983,dence1983,bowers1983,vedder1983,pathria1985}:
\begin{equation}
  \label{eq:binomial_asymp}
  \ln\binom{M}{Mp} = M\,\ysh(p) -\tfrac{1}{2}\ln(2\pu M)-\tfrac{1}{2}\ybu(p)
  + \Ord\Bigl(\tfrac{1}{\sqrt{M}}\Bigr),\qquad
\text{$M$ large},
\end{equation}
%file: limits_hypergeometric.nb
and this expression can be used to obtain asymptotic forms of the
mathematical formulae
\eqref{eq:joint_plaus_N_homog}--\eqref{eq:eq_factmom_prod} for large $N$,
depending on the magnitudes of $n$ and $n/N$. For example, if $N$ and $n$
are large and their ratio $k\defd n/N$ finite, the probability
distributions for the averages can be approximated (provided some
regularity conditions) by continuum densities,
\begin{equation}
  \label{eq:cont_approx}
  \begin{gathered}
  n\,\p(\yxs = a \cond\yH)\approx f(a),
\qquad
  N\,\p(\yXf = A \cond\yH)=N\,Q(A)\approx F(A),
\\
f,F\colon \clcl{0,1}\to\clop{0,+\infty},
\end{gathered}
\end{equation}
and sums by integrals, and \eqn~\eqref{eq:marginal} takes on the
approximate form
\begin{equation}
  \label{eq:marginal_continuum}
  f(a)
=
\int\limits_0^1
\exp\biggl\{
N\,\biggl[
k\,\ysh(a) - \ysh(A) + (1-k)\,\ysh\biggl( \frac{A-ka}{1-k} \biggr)
 \biggr]
\biggr\}\,
F(A)\,\di A,
\end{equation}
which shows an interplay of three entropies that depends on the ratio $n/N$. In
some cases this allows us to approximate the integral by the value at a mode
determined by a generalized entropy principle. (In fact, the standard
maximum-entropy procedure can be derived from the probability calculus via
a similar approximation \citep{portamana2009}.)

The relations above can also be generalized to $K$-valued states $X_i$ in
$\set{0,\dotsc,K-1}$, leading to the appearance of the generalized
hypergeometric distribution
\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
or to real-valued states $X_i$ in $\RR$.

We do not pursue any of these approximations or generalizations in this Note.

\section{Examples of inferential use of the formulae}

\subsection{From network to subnetwork}
\label{sec:pop2sub_examples}

Let us illustrate with an example how the probability distribution for the
subnetwork average $\yxs$, determined by \eqn~\eqref{eq:subpop_average},
changes with the subnetwork size $n$. Choose a network-average distribution
$\p(\yXf=A\cond \yH)$ belonging to the exponential family
\cites[\sect~4.5.3]{bernardoetal1994}[see also][]{fortinietal2000}:
\begin{equation}
  \label{eq:full_pop_av_example_ME}
  \p(\yXf = A \cond \yH) = Q(A) \propto
\binom{N}{N A}\exp[
\yl_2\, N A\,(N A-1)/2 + \yl_1\, N A].
\end{equation}
This is the form obtained from the principle of maximum relative entropy
\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,aczeletal1975,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013}
with first and second moments as constraints and the reference distribution
$Q_0$ defined by $Q_0(A) = 2^{-N}\binom{N}{N A}$, corresponding to a
uniform probability distribution for the network state $\yX$.
%file: distr_scaling_from_full_population.nb
\begin{figure}[!p]
\centering
\includegraphics[width=0.95\columnwidth]{scaled_subpop_probs.pdf}%
\caption{Probability distributions $\p(\yxs = a \cond \yH)$ for different
  subnetwork sizes $n$, obtained from a network probability
  distribution $\p(\yXf = A \cond \yH)$ having the maximum-etropy
  form~\eqref{eq:full_pop_av_example_ME}.}
\label{scaling_distr}
\end{figure}
\begin{figure}[!p]
\centering
\includegraphics[width=0.95\columnwidth]{scaling_subpop_moments.pdf}%
\caption{Moments of the probability distributions $\p(\yxs =a \cond
  \yH)$ as functions of the subnetwork size $n$.}
\label{scaling_moments}
\end{figure}

The probability distribution of \eqn~\eqref{eq:full_pop_av_example_ME} is
plotted in \fig~\ref{scaling_distr}, together with the resulting
subnetwork-average distributions $\p(\yxs = a \cond\yH)$, for the case in
which $N = 1000$ units, $\lambda_1=-2.55$, $\lambda_2=0.005$, and
$n=10, 50, 100, 250$. The distributions become broader as $n$ decreases,
and the minimum of the original distribution disappears; at the same time
the finite-difference
\[\frac{\p(\yxs=a+1/n\cond\yH)-\p(\yxs=a\cond\yH)}{1/n}\]
presents a sharp jump at this minimum when $n\approx 100$.

To the eye familiar with maximum-entropy distributions, the
subnetwork-average distributions of \fig~\ref{scaling_distr} do not look
like maximum-entropy ones with second-moment constraints. In fact,
they are not and \emph{cannot} be:
\begin{equation}
  \label{eq:maxent_form_sub}
  \p(\yxs =a  \cond \yH)
\ne\kappa\binom{n}{n a}\exp[\yk_2\, n a\,(n a-1)/2 + \yk_1\, n a]
%\quad\text{(false)}
\end{equation}
for any $\kappa, \yk_1, \yk_2$, unless $n=2$. This impossibility holds more
generally for any number of constraints $m$ and subnetwork size $n$ such
that $m<n$. The reason is simple: suppose we have assigned a
maximum-entropy distribution with $m$ moment constraints as the
distribution for the network average. If we want the same kind of
distribution for a subnetwork of size $n$, we are free to play with $m+1$
parameters (normalization included), but we must also satisfy the $n+1$
equations corresponding to the marginalization~\eqref{eq:subpop_average}.
This is generally impossible unless $m \ge n$. (Impossibilities of a
similar kind appear in statistical mechanics, see \eg\
ref.~\citep{maesetal1999}.)
%file: test_ME_same_under_scaling.nb

This fact can be significant for recent works
\citep[\eg,][]{schneidmanetal2006,shlensetal2006,tkaciketal2006,marreetal2009,tkaciketal2009,ganmoretal2011,shimazakietal2012,tkaciketal2013,shimazakietal2015}
in which a maximum-entropy probability distribution with second- or
third-moment constraints is assigned to relatively small subnetworks
($n < 200$) of neurons. If we assume that such subnetwork is part of a
larger network, and assume the condition of symmetry~\eqref{eq:homog_math},
then the larger network \emph{cannot} be assigned a maximum-entropy
distribution with the same number of constraints. Vice versa, if we assign
such a maximum-entropy distribution to the larger network, then none of its
subnetwork of enough large size $n$ can be assigned a similar
maximum-entropy distribution. See ref.~\citep{rostamietal2016} for a
broader discussion of this fact and of its consequences.

\medskip

The dependence of the first four moments $\expe{\yxs^m \cond \yH}$ as a
function of size $n$ is shown in \fig~\ref{scaling_moments}. The moments
become practically constant when $n \approx 100$ or larger. The
expectations of $m$-tuple products of states
$\expe{x_{i_1}\dotsm x_{i_m} \cond \yH}$, proportional to the factorial
moments, are not shown as they do not depend on $n$.

\subsection{From subnetwork to network}
\label{sec:from_sub_to_full}

We have seen that, given the condition of symmetry~\eqref{eq:homog_math},
the probability $\p(\yXf = A \cond\yH)$ for the network average determines
that of each subnetwork average, $\p(\yxs \cond\yH)$, by the
marginalization \eqn~\eqref{eq:subpop_average}. The reverse is trivially
not true, since \eqn~\eqref{eq:subpop_average}, as a linear mapping from
$\RR^{N+1}$ to $\RR^{n+1}$, with $N$ larger than $n$, is onto but not into.
Assigning a probability distribution $\p(\yxs = a \cond\yH)$ to a
subnetwork average $\yxs$ does not determine a network distribution
$\p(\yXf = A\cond\yH)$: it only restricts the set of possible ones; this
set can in principle be determined via linear-programming methods
\citep{hailperin1965,hailperin1984,hailperin1996,hailperin2006,hailperin2011}.
\begin{innote}
  Analogous situations appear in the truth-valued logical calculus: if the
  composite proposition $\varAlpha\limplies \varBeta$ is assigned the
  truth-value \enquote{true}, then assigning $\varAlpha$ the value
  \enquote{true} also determines the value of $\varBeta$, whereas assigning
  $\varBeta$ the value \enquote{true} leaves the value of $\varAlpha$
  undetermined.
\end{innote}
The same linear-programming methods show that any inference from subnetwork
properties to network ones must necessarily start from some assumptions
$\varIota$ that assign a probability distribution
$\p(\yX = \yR \cond \varIota)$ for the network states. The approaches to
this task and reformulations of it have become uncountable: they include
exchangeable models, parametric and non-parametric models, hierarchical
models, general linear models, models via sufficiency, maximum-entropy
models, and whatnot
\citep[\eg:][]{jeffreys1931_r1973,jeffreys1939_r2003,jaynes1994_r2003,bernardoetal1994,gelmanetal1995_r2014,ghoshetal1997,kallenberg2005,gregory2005,sivia1996_r2006,ferreiraetal2007,dawid2013,damienetal2013}.
We now show two examples, based on a maximum-entropy approach, that to our
knowledge have not yet been explored in the neuroscientific literature. For
a concrete application see \citep{rostamietal2016b}.

\paragraph{First example: moment constraints for the network.}
\label{sec:maxent_moments}
Consider a state of knowledge $\yHa$ leading to the following properties:
\begin{enumerate}[$\yHa$1.]
\item the expectations of the single and pair averages $\yxs$ and
  $\sav{\yx\yx}$ of a particular subnetwork have given values
  \begin{equation}
    \label{eq:constr_ex1}
    \expe{\yxs \cond \yHa} = c_1, \qquad \expe{\sav{x_i x_j}\cond\yHa} = c_2;
  \end{equation}
\item the network probability distribution $\pr(\yX = \yR \cond\yHa)$
  has maximum relative entropy with respect to the uniform one, given the
  constraints above.
\end{enumerate}

\medskip Then the probability distribution for the network conditional on $\yHa$
is completely determined: it satisfies the symmetry
property~\eqref{eq:homog_math} and is defined by
\begin{multline}
  \label{eq:pop_distr_maxent1}
  \pf(\yX =\yR \cond\yHa) =
\varKappa\exp[\yL_2 N\yRf\,(N\yRf-1)/2 + \yL_1 N\yRf]
\\
\!\begin{aligned}[b]
&\text{with $\varKappa, \yL_m$, such that the distribution is normalized and}
\\
&\varKappa\sum_{N A=0}^N %m!\,
\binom{N-m}{N A-m}
\exp[\yL_2 N A\,(N A-1)/2 + \yL_1 N A]
= c_m, \quad m=1,2.
\end{aligned}
\end{multline}
We omit the full proof of this statement: it is a standard application of
the maximum-entropy procedure
\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013},
combined with the equality~\eqref{eq:expe_products} of subnetwork and
network expectations, \eg
\begin{equation}
  \label{eq:exploit_eq_expects}
  c_2 = \expe{\sav{\yx\yx}\cond\yHa} =
\binom{N}{2}^{-1}
\sum_{N A=0}^N %m!\,
\binom{N A}{2} \, \p(\yXf = A\cond\yHa),
\end{equation}
and with relations~\eqref{eq:joint_plaus_N_homog},
\eqref{eq:conditional_total_state}. This example is easily generalized to
any number $m$ of constraints such that  $m\le n$.

Note again that, as remarked in \sect~\ref{sec:pop2sub_examples}, the
subnetwork from which the averages in the
expectations~\eqref{eq:constr_ex1} are calculated has a probability
distribution $\p(\yxs =a \cond\yH)$ determined by the
marginalization~\eqref{eq:subpop_average} and does \emph{not} have a
maximum-entropy form with the same number of constraints.
%file: test_ME_same_under_scaling.nb

\paragraph{Second example: subnetwork-distribution constraint.}
\label{sec:maxent_fullsubpop}
Consider another state of knowledge $\yHb$ leading to the following
properties:
\begin{asparaenum}[$\yHb$1.]
\item the average $\yxs$ of a particular
  subnetwork has a probability distribution $q$:
  \begin{equation}
    \label{eq:constr_subpop_distr}
    \pr(\yxs =a \cond\yHb) = q(a);
  \end{equation}
\item the probability distribution for the network, $\p(\yX = \yR
  \cond\yHb)$, has maximum relative entropy with respect to the uniform
  one, given the constraint above.
\end{asparaenum}

\medskip Then the probability distribution for the network given $\yHb$ is
completely determined and satisfies the symmetry
property~\eqref{eq:homog_math}:
\begin{multline}
  \label{eq:pop_distr_maxent2}
  \p(\yX= \yR \cond\yHb) =
\exp\Biggl[
\sum_{n a=0}^n \yL_{ a}
%\binom{N}{N\yXf}^{-1}
\binom{n}{n a}
\binom{N-n}{N\yRf-n a}\Biggr]
\\
\!\begin{aligned}[b]
&\text{with $\yL_{ a}$ such that}
\\
&\sum_{N A=0}^N
%\binom{N}{N A}^{-1}
\binom{n}{n a}
\binom{N-n}{N A-n a}\,
\exp\Biggl[
\sum_{n a=0}^n \yL_{ a}
%\binom{N}{N A}^{-1}
\binom{n}{n a}
\binom{N-n}{N A-n a}\Biggr]
=q( a)
\end{aligned}
\end{multline}
(the normalization constraint being unnecessary since $q$ is normalized).
This result is just another application of the maximum-entropy procedure
with $n+1$ (linear) constraints given by \eqn~\eqref{eq:marginal}, where
the left-hand side is now given and equal to $q(a)$.

This example is equivalent to the generalization of the previous one with
$n$ moment constraints, since knowledge of $\pr(\yxs=a \cond \yHb)$ is
equivalent to knowledge its first $n$ moments.

\bigskip

The above examples do not mention how the values of the expectation
constraints or of the subnetwork-average probability distribution can have
been assigned. They cannot be assigned by a measurement, of course, since
probabilities and expectations are not physical quantities and cannot be
physically measured -- they represent guesses of an observer and depend on
the observer's state of knowledge and assumptions. Rather, such values
usually come from measurements made on \enquote{copies} -- in a very
general sense -- of the states of the network; \eg, when we have a time
sequence of them and measure the frequencies of their occurrence. Such
situations can again be fully analysed with the probability calculus, and
one can show \citep{portamana2009} that the maximum-entropy formulae in the
examples above are just limit forms of such an analysis, employing
measured physical data $\varDelta$ like, \eg, frequencies, and
repeated applications of Bayes's theorem,
\begin{equation}
  \label{eq:bayes_thm}
  \p(\yX =\yR\cond \varDelta, \varIota) \propto
\p(\varDelta \cond \yX=\yR, \varIota)\, \p(\yX =\yR \cond \varIota),
\end{equation}
which updates the initial probability assignments on such data. But the
discussion of this is again outside the scope of this Note.

\bigskip

\section{On the  symmetry property}
\label{sec:stoch_symm}

The symmetry property~\eqref{eq:homog_math} is called \emph{finite
  exchangeability} in the Bayesian literature and, as was mentioned in
\sect~\ref{sec:main_formulae}, its relation to the hypergeometric
distribution in expression~\eqref{eq:subpop_average} for the subnetwork
average is
well-known
\citep{kendall1967,definetti1969b,heathetal1976,diaconis1977,diaconisetal1980,jaynes1986c}.

This property can reflect two very different states of knowledge:
either \begin{inparaenum}[(a)]\item knowledge that the network is somehow
  \emph{physically homogeneous}, or
\item \emph{complete ignorance} about the network's homogeneity or inhomogeneity.
\end{inparaenum}
In the second case we are saying that the indices or labels \enquote{$i$}
of the units are \emph{uninformative} -- because, for example, we have no
idea of how the units were labelled, hence we cannot presuppose any
relation among the units, nor can we presuppose any structural or
topological properties of the network they constitute. In the first case we
are saying that the labels are \emph{irrelevant}, even though they might be
informative. The reason could be that, even if there is a connection
between labels and, say, spatial locations of the units, each unit is
nevertheless physically, homogeneously, identically connected to all the
others; so we assume that spatial location does not play a relevant role.

An important consequence of the symmetry property is that no amount of new
evidence about an \emph{a}symmetry in the labelling of some units can lead
to asymmetric predictions about the remaining ones. For example, if we have
new data $\yD$ saying that the first $n$ units are in state $1$ and the
last $n$ in state $0$,
\begin{equation}
 \yD \defd (X_1 = X_2 = \dotsb = X_n =1,\quad
X_{N-(n-1)}= \dotsb = X_{N-1} = X_N = 0),
\end{equation}
with $n$ large, our updated probabily distribution will still assign (as
can be shown using \eqns~\eqref{eq:homog_math} and
\eqref{eq:joint_plaus_N_homog}) the same probability for their respective
neighbouring units with labels $n+1$ and $N-n$ to be in state $1$ or $0$:
\begin{equation}
  \label{eq:same_prob_symm}
  \begin{split}
  \p(X_{n+1} =1 \cond \yD,\yH) &= \p(X_{N-n} =1 \cond \yD,\yH),
  \\
  \p(X_{n+1} =0 \cond \yD,\yH) &= \p(X_{N-n} =0 \cond \yD,\yH).
\end{split}
\end{equation}
This may seem unreasonable -- we would now say that the unit $n+1$ is more
likely to be in state $1$, as the first $n$ are, than in $0$; and that the
unit $N-n$ is more likely in state $0$, as the last $n$ are, than in $1$;
\ie
\begin{equation}
  \label{eq:wished_result}
  \begin{split}
  \p(X_{n+1} = 1 \cond \yD,\yH) &>  \p(X_{n+1} = 0 \cond \yD,\yH),
\\
  \p(X_{N-n} = 1 \cond \yD,\yH) &<  \p(X_{N-n} = 0 \cond \yD,\yH),
\end{split}
\end{equation}
which is equivalent to $\p(X_{n+1} = 1 \cond \yD,\yH) > \tfrac{1}{2}$,
$\p(X_{N-n} = 1 \cond \yD,\yH) < \tfrac{1}{2}$; but the
equalities~\eqref{eq:same_prob_symm} make this impossible.

The two very different motivations behind the symmetry property -- lack of
information and lack of relevance -- can of course be handled differently
by the probability calculus, in such a way that the appearance of asymmetry
in new data leads to asymmetry in updated predictions. But this requires a
more complex set of assumptions than those embodied in $\yH$; it requires,
in particular, some sort of probability distribution for the degree of
physical symmetry or homogeneity of the network, appropriately quantified.

The moral is that we should use the symmetric assumption $\yH$ only if we
can safely exclude the presence of inhomogeneity or are uninterested in
detecting its presence. Otherwise, we must resort to more appropriate (and
complex) assumptions. If we repeatedly observe new values that happen to
have a very low probability according to the updated distribution
$\p(\yX = \yR \cond \yD_1,\yD_2,\dotsc, \yH)$, this could be an indication
that the symmetry property is unreasonable. Any strong departure of higher
powers of measured averages from their expected behaviour given in
\eqn~\eqref{eq:moments} and illustrated in \fig~\ref{scaling_moments} can
also be an indication that the symmetry property may have to be abandoned;
hence the usefulness of \eqn~\eqref{eq:moments}.

\section{Summary and remarks}
\label{sec:summary}

The main point of this Note was to explicitly collect and illustrate the
mathematical formulae \ref{item:tot_plaus}--\ref{item:moments},
\sect~\ref{sec:main_formulae}, between the probabilities assigned to the
state of a network of neurons (or similar entities), and those assigned to
the state of a subnetwork thereof. The formulae hold in the simple case of
binary states and under an assumption of symmetry. Such relations can be
found in several classical texts on probability and statistics; but we
deemed it useful to restate them in a neuroscientific context, given their
fundamental importance in relating a whole to its parts.

We have indeed seen that these formulae lead to straightforward but in some
cases unexpected consequences: \eg, if a network is assigned a
maximum-entropy probability distribution, then its subnetworks
\emph{cannot} be assigned a maximum-entropy distribution of the same form,
and vice versa. The formulae also readily suggest new ways of making
inferences or of formulating starting assumptions. Discussion of these points
is left to forthcoming works \citep{rostamietal2016,rostamietal2016b}.

We also hope to have given the readers a feeling of the agility and
simplicity with which the probability calculus (\enquote{Bayesian theory})
leads us from assumptions to consequences (albeit sometimes with non-simple
mathematics), and from consequences to the assessment of how suitable our
initial assumptions are, as shown with the assumption of symmetry.

\defbibnote{postnote}{\small\par\medskip\noindent{\footnotesize% Note:
\arxivp \mparcp \philscip \biorxivp}%
}

\newcommand{\citein}[2][]{\textnormal{\textcite[#1]{#2}}%\addtocategory{extras}{#2}
}
\newcommand{\citebi}[2][]{ref.\ \citep[#1]{#2}%\addtocategory{extras}{#2}
}
\newcommand{\subtitleproc}[1]{}

\printbibliography[postnote=postnote]

\end{document}

### 2017-05-04 maxNt.tex
update luca's bib

@@
 This simple example shows that the multiplicity inspection that must
-precede a \me\ application already involve probability considerations at
+precede a \me\ application already involves probability considerations at
 the level of the full network. The usual reasoning by enumeration
-implicitly assumes a uniform distribution, or a ***
+implicitly assumes a uniform distribution or at least a \emph{factorizable}
+distribution.
 
-We:
+***If the distribution is factorizable, however, it means that examination of
+the subnetwork \emph{cannot give us any insights about the network it is a
+  part of}. This is obviously contrary to the reason why we made neuronal observations.
+
+Consider the following ways of proceeding. We:
 \begin{enumerate}

### 2017-05-04 maxNt.tex
update intro

@@
 
+\usepackage[usenames]{xcolor}
+\definecolor{myblue}{RGB}{51,34,136}
+\definecolor{mygreen}{RGB}{17,119,51}
+\definecolor{myred}{RGB}{136,34,85}
+\definecolor{myyellow}{RGB}{153,153,51}
+\definecolor{mylightyellow}{RGB}{221,204,119}
+\newcommand*{\mynote}[1]{ {\color[RGB]{68,170,153}\maltese\ #1}}
+
 \usepackage{hyperref}
@@
 \providecommand*{\urlalt}{\href}
-
+\hypersetup{
+colorlinks=true,%
+bookmarksnumbered,%
+pdfborder={0 0 0.25},%
+citebordercolor={0.2 0.1333 0.5333},%bluish
+citecolor=myblue,%
+linkbordercolor={0.0667 0.4667 0.2},%greenish
+linkcolor=myred,%
+urlbordercolor={0.5333 0.1333 0.3333},%reddish
+urlcolor=mygreen,%
+breaklinks=true,%
+pdftitle = {Hidden network-size assumptions in the maximum-entropy method},%***
+pdfauthor = {Rostami, Porta Mana, Torre}%***
+}
 \usepackage[depth=4]{bookmark}
@@
 \begin{abstract}
-  The abstract paragraph should be indented \nicefrac{1}{2}~inch
-  (3~picas) on both the left- and right-hand margins. Use 10~point
-  type, with a vertical spacing (leading) of 11~points.  The word
-  \textbf{Abstract} must be centered, bold, and in point size 12. Two
-  line spaces precede the abstract. The abstract must be limited to
-  one paragraph.
+***%  The abstract must be limited to one paragraph.
 \end{abstract}
@@
 
-We would like to warn about an implicit assumption of the \me\ method,
-which can informally be put this way:
+In this note we would like to analyse and warn about a subtle assumption
+behind the \me\ method when it is applied to a network. It can informally
+be put this way:
 \begin{quote}
-  \emph{\me\ method applied to $\yN$ units assumes that no other
-  units exist.}
+  \emph{the \me\ method assumes that the network it is applied to is
+    completely isolated from any larger network.}
 \end{quote}
-Less informally, the method implicitly assumes that the $\yN$ units are not a
-subnetwork sample of a network with $\yM$ units, $\yM>\yN$.
 
+\mynote{Version 1}
+The \me\ method does not construct a probability distribution out of
+nothing, but starting from a uniform distribution. A uniform distribution
+is an innocuous assumption for a set of non-composite events, like the
+outcomes of a die roll, and also for some sets of composite events, like
+the outcomes of the roll of two dice. In the latter case multiplicities
+appear.
+
+\mynote{Version 2}
 This may come as a surprise. The method simply requires a number of
@@
 only, right?
+
+Right, but
+
+Right, and that is where the problem lies. This way of counting of
+multiplicities assumes an underlying
 
@@
 
-NIPS requires electronic submissions.  The electronic submission site
-is
-\begin{center}
-  \url{https://cmt.research.microsoft.com/NIPS2017/}
-\end{center}
+\subsection{***}
 
-Please read carefully the instructions below and follow them
-faithfully.
+\subsubsection{***}
 
-\subsection{Style}
+\paragraph{***}
 
-Papers to be submitted to NIPS 2017 must be prepared according to the
-instructions presented here. Papers may only be up to eight pages
-long, including figures. This does not include acknowledgments and
-cited references which are allowed on subsequent pages.
-Papers that exceed these limits will not be reviewed, or in any
-other way considered for presentation at the conference.
-
-The margins in 2017 are the same as since 2007, which allow for
-$\sim$$15\%$ more words in the paper compared to earlier years.
-
-Authors are required to use the NIPS \LaTeX{} style files obtainable
-at the NIPS website as indicated below. Please make sure you use the
-current files and not previous versions. Tweaking the style files may
-be grounds for rejection.
-
-\subsection{Retrieval of style files}
-
-The style files for NIPS and other conference information are
-available on the World Wide Web at
-\begin{center}
-  \url{http://www.nips.cc/}
-\end{center}
-The file \verb+nips_2017.pdf+ contains these instructions and
-illustrates the various formatting requirements your NIPS paper must
-satisfy.
-
-The only supported style file for NIPS 2017 is \verb+nips_2017.sty+,
-rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{}
-  2.09, Microsoft Word, and RTF are no longer supported!}
-
-The new \LaTeX{} style file contains two optional arguments:
-\verb+final+, which creates a camera-ready copy, and \verb+nonatbib+,
-which will not load the \verb+natbib+ package for you in case of
-package clash.
-
-At submission time, please omit the \verb+final+ option. This will
-anonymize your submission and add line numbers to aid review.  Please
-do \emph{not} refer to these line numbers in your paper as they will
-be removed during generation of camera-ready copies.
-
-The file \verb+nips_2017.tex+ may be used as a ``shell'' for writing
-your paper. All you have to do is replace the author, title, abstract,
-and text of the paper with your own.
-
-The formatting instructions contained in these style files are
-summarized in Sections \ref{gen_inst}, \ref{headings}, and
-\ref{others} below.
-
-\section{General formatting instructions}
-\label{gen_inst}
-
-The text must be confined within a rectangle 5.5~inches (33~picas)
-wide and 9~inches (54~picas) long. The left margin is 1.5~inch
-(9~picas).  Use 10~point type with a vertical spacing (leading) of
-11~points.  Times New Roman is the preferred typeface throughout, and
-will be selected for you by default.  Paragraphs are separated by
-\nicefrac{1}{2}~line space (5.5 points), with no indentation.
-
-The paper title should be 17~point, initial caps/lower case, bold,
-centered between two horizontal rules. The top rule should be 4~points
-thick and the bottom rule should be 1~point thick. Allow
-\nicefrac{1}{4}~inch space above and below the title to rules. All
-pages should start at 1~inch (6~picas) from the top of the page.
-
-For the final version, authors' names are set in boldface, and each
-name is centered above the corresponding address. The lead author's
-name is to be listed first (left-most), and the co-authors' names (if
-different address) are set to follow. If there is only one co-author,
-list both author and co-author side by side.
-
-Please pay special attention to the instructions in Section \ref{others}
-regarding figures, tables, acknowledgments, and references.
-
-\section{Headings: first level}
-\label{headings}
-
-All headings should be lower case (except for first word and proper
-nouns), flush left, and bold.
-
-First-level headings should be in 12-point type.
-
-\subsection{Headings: second level}
-
-Second-level headings should be in 10-point type.
-
-\subsubsection{Headings: third level}
-
-Third-level headings should be in 10-point type.
-
-\paragraph{Paragraphs}
-
-There is also a \verb+\paragraph+ command available, which sets the
-heading in bold, flush left, and inline with the text, with the
-heading followed by 1\,em of space.
-
-\section{Citations, figures, tables, references}
-\label{others}
-
-These instructions apply to everyone.
-
-\subsection{Citations within the text}
-
-The \verb+natbib+ package will be loaded for you by default.
-Citations may be author/year or numeric, as long as you maintain
-internal consistency.  As to the format of the references themselves,
-any style is acceptable as long as it is used consistently.
-
-The documentation for \verb+natbib+ may be found at
-\begin{center}
-  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
-\end{center}
-Of note is the command \verb+\citet+, which produces citations
-appropriate for use in inline text.  For example,
-\begin{verbatim}
-   \citet{hasselmo} investigated\dots
-\end{verbatim}
-produces
-\begin{quote}
-  Hasselmo, et al.\ (1995) investigated\dots
-\end{quote}
-
-If you wish to load the \verb+natbib+ package with options, you may
-add the following before loading the \verb+nips_2017+ package:
-\begin{verbatim}
-   \PassOptionsToPackage{options}{natbib}
-\end{verbatim}
-
-If \verb+natbib+ clashes with another package you load, you can add
-the optional argument \verb+nonatbib+ when loading the style file:
-\begin{verbatim}
-   \usepackage[nonatbib]{nips_2017}
-\end{verbatim}
-
-As submission is double blind, refer to your own published work in the
-third person. That is, use ``In the previous work of Jones et
-al.\ [4],'' not ``In our previous work [4].'' If you cite your other
-papers that are not widely available (e.g., a journal paper under
-review), use anonymous author names in the citation, e.g., an author
-of the form ``A.\ Anonymous.''
-
-\subsection{Footnotes}
-
-Footnotes should be used sparingly.  If you do require a footnote,
-indicate footnotes with a number\footnote{Sample of the first
-  footnote.} in the text. Place the footnotes at the bottom of the
-page on which they appear.  Precede the footnote with a horizontal
-rule of 2~inches (12~picas).
-
-Note that footnotes are properly typeset \emph{after} punctuation
-marks.\footnote{As in this example.}
-
-\subsection{Figures}
-
-All artwork must be neat, clean, and legible. Lines should be dark
-enough for purposes of reproduction. The figure number and caption
-always appear after the figure. Place one line space before the figure
-caption and one line space after the figure. The figure caption should
-be lower case (except for first word and proper nouns); figures are
-numbered consecutively.
-
-You may use color figures.  However, it is best for the figure
-captions and the paper body to be legible if the paper is printed in
-either black/white or in color.
-\begin{figure}[h]
-  \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  \caption{Sample figure caption.}
-\end{figure}
-
-\subsection{Tables}
-
-All tables must be centered, neat, clean and legible.  The table
-number and title always appear before the table.  See
-Table~\ref{sample-table}.
-
-Place one line space before the table title, one line space after the
-table title, and one line space after the table. The table title must
-be lower case (except for first word and proper nouns); tables are
-numbered consecutively.
-
-Note that publication-quality tables \emph{do not contain vertical
-  rules.} We strongly suggest the use of the \verb+booktabs+ package,
-which allows for typesetting high-quality, professional tables:
-\begin{center}
-  \url{https://www.ctan.org/pkg/booktabs}
-\end{center}
-This package was used to typeset Table~\ref{sample-table}.
-
-\begin{table}[t]
-  \caption{Sample table title}
-  \label{sample-table}
-  \centering
-  \begin{tabular}{lll}
-    \toprule
-    \multicolumn{2}{c}{Part}                   \\
-    \cmidrule{1-2}
-    Name     & Description     & Size ($\mu$m) \\
-    \midrule
-    Dendrite & Input terminal  & $\sim$100     \\
-    Axon     & Output terminal & $\sim$10      \\
-    Soma     & Cell body       & up to $10^6$  \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\section{Final instructions}
-
-Do not change any aspects of the formatting parameters in the style
-files.  In particular, do not modify the width or length of the
-rectangle the text should fit into, and do not change font sizes
-(except perhaps in the \textbf{References} section; see below). Please
-note that pages should be numbered.
-
-\section{Preparing PDF files}
-
-Please prepare submission files with paper size ``US Letter,'' and
-not, for example, ``A4.''
-
-Fonts were the main cause of problems in the past years. Your PDF file
-must only contain Type 1 or Embedded TrueType fonts. Here are a few
-instructions to achieve this.
-
-\begin{itemize}
-
-\item You should directly generate PDF files using \verb+pdflatex+.
-
-\item You can check which fonts a PDF files uses.  In Acrobat Reader,
-  select the menu Files$>$Document Properties$>$Fonts and select Show
-  All Fonts. You can also use the program \verb+pdffonts+ which comes
-  with \verb+xpdf+ and is available out-of-the-box on most Linux
-  machines.
-
-\item The IEEE has recommendations for generating PDF files whose
-  fonts are also acceptable for NIPS. Please see
-  \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}
-
-\item \verb+xfig+ "patterned" shapes are implemented with bitmap
-  fonts.  Use "solid" shapes instead.
-
-\item The \verb+\bbold+ package almost always uses bitmap fonts.  You
-  should use the equivalent AMS Fonts:
-\begin{verbatim}
-   \usepackage{amsfonts}
-\end{verbatim}
-followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or
-\verb+\mathbb{C}+ for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You
-can also use the following workaround for reals, natural and complex:
-\begin{verbatim}
-   \newcommand{\RR}{I\!\!R} %real numbers
-   \newcommand{\Nat}{I\!\!N} %natural numbers
-   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
-\end{verbatim}
-Note that \verb+amsfonts+ is automatically loaded by the
-\verb+amssymb+ package.
-
-\end{itemize}
-
-If your file contains type 3 fonts or non embedded TrueType fonts, we
-will ask you to fix it.
-
-\subsection{Margins in \LaTeX{}}
-
-Most of the margin problems come from figures positioned by hand using
-\verb+\special+ or other commands. We suggest using the command
-\verb+\includegraphics+ from the \verb+graphicx+ package. Always
-specify the figure width as a multiple of the line width as in the
-example below:
-\begin{verbatim}
-   \usepackage[pdftex]{graphicx} ...
-   \includegraphics[width=0.8\linewidth]{myfile.pdf}
-\end{verbatim}
-See Section 4.4 in the graphics bundle documentation
-(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})
-
-A number of width problems arise when \LaTeX{} cannot properly
-hyphenate a line. Please give LaTeX hyphenation hints using the
-\verb+\-+ command when necessary.
+% do not change font sizes (except perhaps in the \textbf{References}
+% section; see below).
 
@@
 
-Use unnumbered third level headings for the acknowledgments. All
-acknowledgments go at the end of the paper. Do not include
-acknowledgments in the anonymized submission, only in the final paper.
+% Do not include acknowledgments in the anonymized submission, only in the
+% final paper.
 
@@
 
-References follow the acknowledgments. Use unnumbered first-level
-heading for the references. Any choice of citation style is acceptable
-as long as you are consistent. It is permissible to reduce the font
-size to \verb+small+ (9 point) when listing the references. {\bf
-  Remember that you can go over 8 pages as long as the subsequent ones contain
-  \emph{only} cited references.}
-\medskip
-
-\small
-
-[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
-for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
-T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
-  Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.
-
-[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
-  Exploring Realistic Neural Models with the GEneral NEural SImulation
-  System.}  New York: TELOS/Springer--Verlag.
-
-[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
-learning and recall at excitatory recurrent synapses and cholinergic
-modulation in rat hippocampal region CA3. {\it Journal of
-  Neuroscience} {\bf 15}(7):5249-5262.
+% References follow the acknowledgments. Use unnumbered first-level
+% heading for the references. Any choice of citation style is acceptable
+% as long as you are consistent. It is permissible to reduce the font
+% size to \verb+small+ (9 point) when listing the references. {\bf
+%   Remember that you can go over 8 pages as long as the subsequent ones contain
+%   \emph{only} cited references.}
+% \medskip
 

### 2017-05-04 maxNt.tex
update intro

@@
 
+When applied to a subnetwork, the \me\ method assumes that the uniform over
+the larger network is uniform, and therefore factorizable. The new
+distribution of the subnetwork will not be uniform, but that of the full
+network will still be factorizable into the one for the subnetwork and the
+rest.
+
+A uniform distribution, however, is not the right one when we suppose that
+learning about an event may tell us something about a related event. For
+example, consider 1\,000 tosses of a particular coin and assume a uniform
+distribution over the possible $2^{1\,000}$ outcomes. If we learn that the
+first 999 tosses yielded all \enquote{heads}, the probability calculus
+tells us that the probability for the 1\,000th toss is still 50\%/50\%. It
+is a consequence of our choice of a uniform distribution: we have
+implicitly declared all tosses to be completely independent, completely
+\emph{irrelevant} to one another. This fact is well-known in sampling
+theory. A more telling example in fact is that of a presidential election
+with two candidates: each citizen will vote for one or the other. We do
+survey sampling on a large number of citizens to guess the election's
+outcome. If we assumed a uniform distribution over the possible
+combinations of choices of all citizens, our sampling would be completely
+irrelevant for the choices of the rest of the population.
+
+The latter example has many similarities with that of a neuronal binary
+network. When we record the neuronal activity of a sample of neurons from a
+brain area, we assume that our measurements can tell us something -- no
+matter how vague or imprecise -- about the whole brain area. This means
+that we are not assuming a uniform distribution over all possible states of
+the area.
+
 \mynote{Version 2}

### 2017-05-12 maxNt.tex
add to intro

@@
+%% Author: PGL  Porta Mana
+%% Created: 2015-11-04T09:20:08+0100
+%% Last-Updated: 2017-05-12T13:15:15+0200
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Report-no: ***
 \documentclass{article}
@@
 \usepackage[final,nonatbib]{nips_2017}
+\newcommand*{\firstdraft}{4 November 2015}
+\newcommand*{\pdftitle}{Hidden network-size assumptions\\in the maximum-entropy method}
+\newcommand*{\headtitle}{\pdftitle}
+\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, V. Rostami, E. Torre}
 
-% to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage[final]{nips_2017}
+%\usepackage{pifont}
+%\usepackage{fontawesome}
+\usepackage[T1]{fontenc} 
+\input{glyphtounicode} \pdfgentounicode=1
+\usepackage[utf8]{inputenx}
+\usepackage{newunicodechar}
+% \newunicodechar{Ĕ}{\u{E}}
+% \newunicodechar{ĕ}{\u{e}}
+% \newunicodechar{Ĭ}{\u{I}}
+% \newunicodechar{ĭ}{\u{\i}}
+% \newunicodechar{Ŏ}{\u{O}}
+% \newunicodechar{ŏ}{\u{o}}
+% \newunicodechar{Ŭ}{\u{U}}
+% \newunicodechar{ŭ}{\u{u}}
+% \newunicodechar{Ā}{\=A}
+% \newunicodechar{ā}{\=a}
+% \newunicodechar{Ē}{\=E}
+% \newunicodechar{ē}{\=e}
+% \newunicodechar{Ī}{\=I}
+% \newunicodechar{ī}{\={\i}}
+% \newunicodechar{Ō}{\=O}
+% \newunicodechar{ō}{\=o}
+% \newunicodechar{Ū}{\=U}
+% \newunicodechar{ū}{\=u}
+% \newunicodechar{Ȳ}{\=Y}
+% \newunicodechar{ȳ}{\=y}
 
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage[utf8]{inputenc} % allow utf-8 input
-%\usepackage{hyperref}       % hyperlinks
-%\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-%\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
+%\newcommand*{\bmmax}{0} % reduce number of bold fonts, before bm
+\newcommand*{\hmmax}{0} % reduce number of heavy fonts, before bm
 \usepackage{textcomp}
 \usepackage{amsmath}
-
 \usepackage{mathtools}
 \setlength{\multlinegap}{0pt}
-% \newlength{\Normalbaselineskip}
-% \setlength{\Normalbaselineskip}{\baselineskip}
-% \appto\MultlinedHook{\setlength\baselineskip{\Normalbaselineskip}}
 
@@
 \usepackage{amsxtra}
-\usepackage{amsthm}
-\usepackage{bm}
 
-\usepackage[british]{babel}
-\selectlanguage{british}
-% \newcommand*{\langfrench}{\foreignlanguage{french}}
-% \newcommand*{\langgerman}{\foreignlanguage{german}}
-% \newcommand*{\langitalian}{\foreignlanguage{italian}}
-% \newcommand*{\langswedish}{\foreignlanguage{swedish}}
-% \newcommand*{\langlatin}{\foreignlanguage{latin}}
-% \newcommand*{\langnohyph}{\foreignlanguage{nohyphenation}}
-% \newcommand*{\latin}[1]{\langlatin{#1}}
-% \newcommand*{\french}[1]{\emph{\langfrench{#1}}}
-% \newcommand*{\german}[1]{\emph{\langgerman{#1}}}
+\usepackage[british]{babel}\selectlanguage{british}
+\newcommand*{\langfrench}{\foreignlanguage{french}}
+\newcommand*{\langgerman}{\foreignlanguage{german}}
+\newcommand*{\langitalian}{\foreignlanguage{italian}}
+\newcommand*{\langswedish}{\foreignlanguage{swedish}}
+\newcommand*{\langlatin}{\foreignlanguage{latin}}
+\newcommand*{\langnohyph}{\foreignlanguage{nohyphenation}}
 
@@
 \setquotestyle{american}
+
+\usepackage{amsthm}
+\newcommand*{\QED}{\textsc{q.e.d.}}
+\renewcommand*{\qedsymbol}{\QED}
+\theoremstyle{remark}
+\newtheorem{note}{Note}
+\newtheorem*{remark}{Note}
+\newtheoremstyle{innote}{\parsep}{\parsep}{\footnotesize}{}{}{}{0pt}{}
+\theoremstyle{innote}
+\newtheorem*{innote}{}
+
+\usepackage{bm}
+
 \iffalse
 \usepackage[shortlabels,inline]{enumitem}
-% to replace paralist environments with enumitem ones:
 \SetEnumitemKey{para}{itemindent=\parindent,leftmargin=0pt,listparindent=\parindent,parsep=0pt,itemsep=\topsep}
+% \begin{asparaenum} = \begin{enumerate}[para]
+% \begin{inparaenum} = \begin{enumerate*}
 \setlist[enumerate,2]{label=\alph*.}
@@
 
-\usepackage[backend=biber,mcite,subentry,citestyle=numeric-comp,bibstyle=numericbringhurst,autopunct=false,sorting=none,sortcites=false,natbib=false,maxnames=8,minnames=8,giveninits=true,block=space,hyperref=true,defernumbers=false,useprefix=true,language=british]{biblatex}
-\renewcommand*{\finalnamedelim}{, }
-\setcounter{biburlnumpenalty}{1}
-\setcounter{biburlucpenalty}{0}
-\setcounter{biburllcpenalty}{1}
+%% With euler font cursive for Greek letters - the [1] means 100% scaling
+\DeclareFontFamily{U}{egreek}{\skewchar\font'177}%
+\DeclareFontShape{U}{egreek}{m}{n}{<-6>s*[0.95]eurm5 <6-8>s*[0.95]eurm7 <8->s*[0.95]eurm10}{}%
+\DeclareFontShape{U}{egreek}{m}{it}{<->s*[0.95]eurmo10}{}%
+\DeclareFontShape{U}{egreek}{b}{n}{<-6>s*[0.95]eurb5 <6-8>s*[0.95]eurb7 <8->s*[0.95]eurb10}{}%
+\DeclareFontShape{U}{egreek}{b}{it}{<->s*[0.95]eurbo10}{}%
+\DeclareSymbolFont{egreeki}{U}{egreek}{m}{it}%
+\SetSymbolFont{egreeki}{bold}{U}{egreek}{b}{it}% from the amsfonts package
+\DeclareSymbolFont{egreekr}{U}{egreek}{m}{n}%
+\SetSymbolFont{egreekr}{bold}{U}{egreek}{b}{n}% from the amsfonts package
+% Take also \sum, \prod, \coprod symbols from Euler fonts
+\DeclareFontFamily{U}{egreekx}{\skewchar\font'177}
+\DeclareFontShape{U}{egreekx}{m}{n}{%
+       <-7.5>s*[0.9]euex7%
+    <7.5-8.5>s*[0.9]euex8%
+    <8.5-9.5>s*[0.9]euex9%
+    <9.5->s*[0.9]euex10%
+}{}
+\DeclareSymbolFont{egreekx}{U}{egreekx}{m}{n}
+\DeclareMathSymbol{\sumop}{\mathop}{egreekx}{"50}
+\DeclareMathSymbol{\prodop}{\mathop}{egreekx}{"51}
+\DeclareMathSymbol{\coprodop}{\mathop}{egreekx}{"60}
+\makeatletter
+\def\sum{\DOTSI\sumop\slimits@}
+\def\prod{\DOTSI\prodop\slimits@}
+\def\coprod{\DOTSI\coprodop\slimits@}
+\makeatother
+% Greek letters not usually given in LaTeX. Comment the unneeded ones
+\DeclareMathSymbol{\varpartial}{\mathalpha}{egreeki}{"40}
+\DeclareMathSymbol{\partialup}{\mathalpha}{egreekr}{"40}
+\DeclareMathSymbol{\alpha}{\mathalpha}{egreeki}{"0B}
+\DeclareMathSymbol{\beta}{\mathalpha}{egreeki}{"0C}
+\DeclareMathSymbol{\gamma}{\mathalpha}{egreeki}{"0D}
+\DeclareMathSymbol{\delta}{\mathalpha}{egreeki}{"0E}
+\DeclareMathSymbol{\epsilon}{\mathalpha}{egreeki}{"0F}
+\DeclareMathSymbol{\zeta}{\mathalpha}{egreeki}{"10}
+\DeclareMathSymbol{\eta}{\mathalpha}{egreeki}{"11}
+\DeclareMathSymbol{\theta}{\mathalpha}{egreeki}{"12}
+\DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
+\DeclareMathSymbol{\kappa}{\mathalpha}{egreeki}{"14}
+\DeclareMathSymbol{\lambda}{\mathalpha}{egreeki}{"15}
+\DeclareMathSymbol{\mu}{\mathalpha}{egreeki}{"16}
+\DeclareMathSymbol{\nu}{\mathalpha}{egreeki}{"17}
+\DeclareMathSymbol{\xi}{\mathalpha}{egreeki}{"18}
+\DeclareMathSymbol{\omicron}{\mathalpha}{egreeki}{"6F}
+\DeclareMathSymbol{\pi}{\mathalpha}{egreeki}{"19}
+\DeclareMathSymbol{\rho}{\mathalpha}{egreeki}{"1A}
+\DeclareMathSymbol{\sigma}{\mathalpha}{egreeki}{"1B}
+\DeclareMathSymbol{\tau}{\mathalpha}{egreeki}{"1C}
+\DeclareMathSymbol{\upsilon}{\mathalpha}{egreeki}{"1D}
+\DeclareMathSymbol{\phi}{\mathalpha}{egreeki}{"1E}
+\DeclareMathSymbol{\chi}{\mathalpha}{egreeki}{"1F}
+\DeclareMathSymbol{\psi}{\mathalpha}{egreeki}{"20}
+\DeclareMathSymbol{\omega}{\mathalpha}{egreeki}{"21}
+\DeclareMathSymbol{\varepsilon}{\mathalpha}{egreeki}{"22}
+\DeclareMathSymbol{\vartheta}{\mathalpha}{egreeki}{"23}
+\DeclareMathSymbol{\varpi}{\mathalpha}{egreeki}{"24}
+\let\varrho\rho 
+\let\varsigma\sigma
+\let\varkappa\kappa
+\DeclareMathSymbol{\varphi}{\mathalpha}{egreeki}{"27}
+%
+\DeclareMathSymbol{\varAlpha}{\mathalpha}{egreeki}{"41}
+\DeclareMathSymbol{\varBeta}{\mathalpha}{egreeki}{"42}
+\DeclareMathSymbol{\varGamma}{\mathalpha}{egreeki}{"00}
+\DeclareMathSymbol{\varDelta}{\mathalpha}{egreeki}{"01}
+\DeclareMathSymbol{\varEpsilon}{\mathalpha}{egreeki}{"45}
+\DeclareMathSymbol{\varZeta}{\mathalpha}{egreeki}{"5A}
+\DeclareMathSymbol{\varEta}{\mathalpha}{egreeki}{"48}
+\DeclareMathSymbol{\varTheta}{\mathalpha}{egreeki}{"02}
+\DeclareMathSymbol{\varIota}{\mathalpha}{egreeki}{"49}
+\DeclareMathSymbol{\varKappa}{\mathalpha}{egreeki}{"4B}
+\DeclareMathSymbol{\varLambda}{\mathalpha}{egreeki}{"03}
+\DeclareMathSymbol{\varMu}{\mathalpha}{egreeki}{"4D}
+\DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
+\DeclareMathSymbol{\varXi}{\mathalpha}{egreeki}{"04}
+\DeclareMathSymbol{\varOmicron}{\mathalpha}{egreeki}{"4F}
+\DeclareMathSymbol{\varPi}{\mathalpha}{egreeki}{"05}
+\DeclareMathSymbol{\varRho}{\mathalpha}{egreeki}{"50}
+\DeclareMathSymbol{\varSigma}{\mathalpha}{egreeki}{"06}
+\DeclareMathSymbol{\varTau}{\mathalpha}{egreeki}{"54}
+\DeclareMathSymbol{\varUpsilon}{\mathalpha}{egreeki}{"07}
+\DeclareMathSymbol{\varPhi}{\mathalpha}{egreeki}{"08}
+\DeclareMathSymbol{\varChi}{\mathalpha}{egreeki}{"58}
+\DeclareMathSymbol{\varPsi}{\mathalpha}{egreeki}{"09}
+\DeclareMathSymbol{\varOmega}{\mathalpha}{egreeki}{"0A} 
+%
+\DeclareMathSymbol{\Alpha}{\mathalpha}{egreekr}{"41}
+\DeclareMathSymbol{\Beta}{\mathalpha}{egreekr}{"42}
+\DeclareMathSymbol{\Gamma}{\mathalpha}{egreekr}{"00}
+\DeclareMathSymbol{\Delta}{\mathalpha}{egreekr}{"01}
+\DeclareMathSymbol{\Epsilon}{\mathalpha}{egreekr}{"45}
+\DeclareMathSymbol{\Zeta}{\mathalpha}{egreekr}{"5A}
+\DeclareMathSymbol{\Eta}{\mathalpha}{egreekr}{"48}
+\DeclareMathSymbol{\Theta}{\mathalpha}{egreekr}{"02}
+\DeclareMathSymbol{\Iota}{\mathalpha}{egreekr}{"49}
+\DeclareMathSymbol{\Kappa}{\mathalpha}{egreekr}{"4B}
+\DeclareMathSymbol{\Lambda}{\mathalpha}{egreekr}{"03}
+\DeclareMathSymbol{\Mu}{\mathalpha}{egreekr}{"4D}
+\DeclareMathSymbol{\Nu}{\mathalpha}{egreekr}{"4E}
+\DeclareMathSymbol{\Xi}{\mathalpha}{egreekr}{"04}
+\DeclareMathSymbol{\Omicron}{\mathalpha}{egreekr}{"4F}
+\DeclareMathSymbol{\Pi}{\mathalpha}{egreekr}{"05}
+\DeclareMathSymbol{\Rho}{\mathalpha}{egreekr}{"50}
+\DeclareMathSymbol{\Sigma}{\mathalpha}{egreekr}{"06}
+\DeclareMathSymbol{\Tau}{\mathalpha}{egreekr}{"54}
+\DeclareMathSymbol{\Upsilon}{\mathalpha}{egreekr}{"07}
+\DeclareMathSymbol{\Phi}{\mathalpha}{egreekr}{"08}
+\DeclareMathSymbol{\Chi}{\mathalpha}{egreekr}{"58}
+\DeclareMathSymbol{\Psi}{\mathalpha}{egreekr}{"09}
+\DeclareMathSymbol{\Omega}{\mathalpha}{egreekr}{"0A}
+%
+\DeclareMathSymbol{\alphaup}{\mathalpha}{egreekr}{"0B}
+\DeclareMathSymbol{\betaup}{\mathalpha}{egreekr}{"0C}
+\DeclareMathSymbol{\gammaup}{\mathalpha}{egreekr}{"0D}
+\DeclareMathSymbol{\deltaup}{\mathalpha}{egreekr}{"0E}
+\DeclareMathSymbol{\epsilonup}{\mathalpha}{egreekr}{"0F}
+\DeclareMathSymbol{\zetaup}{\mathalpha}{egreekr}{"10}
+\DeclareMathSymbol{\etaup}{\mathalpha}{egreekr}{"11}
+\DeclareMathSymbol{\thetaup}{\mathalpha}{egreekr}{"12}
+\DeclareMathSymbol{\iotaup}{\mathalpha}{egreekr}{"13}
+\DeclareMathSymbol{\kappaup}{\mathalpha}{egreekr}{"14}
+\DeclareMathSymbol{\lambdaup}{\mathalpha}{egreekr}{"15}
+\DeclareMathSymbol{\muup}{\mathalpha}{egreekr}{"16}
+\DeclareMathSymbol{\nuup}{\mathalpha}{egreekr}{"17}
+\DeclareMathSymbol{\xiup}{\mathalpha}{egreekr}{"18}
+\DeclareMathSymbol{\omicronup}{\mathalpha}{egreekr}{"6F}
+ \DeclareMathSymbol{\piup}{\mathalpha}{egreekr}{"19}
+\DeclareMathSymbol{\rhoup}{\mathalpha}{egreekr}{"1A}
+\DeclareMathSymbol{\sigmaup}{\mathalpha}{egreekr}{"1B}
+\DeclareMathSymbol{\tauup}{\mathalpha}{egreekr}{"1C}
+\DeclareMathSymbol{\upsilonup}{\mathalpha}{egreekr}{"1D}
+\DeclareMathSymbol{\phiup}{\mathalpha}{egreekr}{"1E}
+\DeclareMathSymbol{\chiup}{\mathalpha}{egreekr}{"1F}
+\DeclareMathSymbol{\psiup}{\mathalpha}{egreekr}{"20}
+\DeclareMathSymbol{\omegaup}{\mathalpha}{egreekr}{"21}
+\DeclareMathSymbol{\varepsilonup}{\mathalpha}{egreekr}{"22}
+\DeclareMathSymbol{\varthetaup}{\mathalpha}{egreekr}{"23}
+\DeclareMathSymbol{\varpiup}{\mathalpha}{egreekr}{"24}
+\let\varrhoup\rhoup 
+\let\varsigmaup\sigmaup
+\let\varkappaup\kappaup
+\DeclareMathSymbol{\varphiup}{\mathalpha}{egreekr}{"27}
 
-\usepackage{graphicx}
+%\newcommand*{\mathte}[1]{\textbf{\textit{\textsf{#1}}}}
+% Upright sans-serif math alphabet
+% \DeclareMathAlphabet{\mathsu}  {T1}{\sfdefault}{m}{n}
+% \SetMathAlphabet{\mathsu}{bold}{T1}{\sfdefault}{b}{n}
+
+\usepackage{mathdots}
 
@@
 \definecolor{mylightyellow}{RGB}{221,204,119}
-\newcommand*{\mynote}[1]{ {\color[RGB]{68,170,153}\maltese\ #1}}
+
+\usepackage{microtype}
+
+\usepackage[backend=biber,mcite,subentry,citestyle=numeric-comp,bibstyle=numericbringhurst,autopunct=false,sorting=none,sortcites=false,natbib=false,maxnames=8,minnames=8,giveninits=true,block=space,hyperref=true,defernumbers=false,useprefix=true,language=british]{biblatex}
+\renewcommand*{\finalnamedelim}{, }
+\setcounter{biburlnumpenalty}{1}
+\setcounter{biburlucpenalty}{0}
+\setcounter{biburllcpenalty}{1}
+\DeclareDelimFormat{multicitedelim}{\addsemicolon\space}
+\DeclareDelimFormat{postnotedelim}{\space}
+\addbibresource{portamanabib.bib}
+\renewcommand{\bibfont}{\footnotesize}
+\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
+}
+\newcommand*{\citep}{\parencites}
+\newcommand*{\citey}{\parencites*}
+\renewcommand*{\cite}{\citep}
+\providecommand{\href}[2]{#2}
+\providecommand{\eprint}[2]{\texttt{\href{#1}{#2}}}
+\newcommand*{\amp}{\&}
+
+\def\arxivp{}
+\def\mparcp{}
+\def\philscip{}
+\def\biorxivp{}
+\newcommand*{\arxivsi}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.\\}
+\newcommand*{\mparcsi}{\texttt{mp\_arc} eprints available at \url{http://www.ma.utexas.edu/mp_arc/}.\\}
+\newcommand*{\philscisi}{\texttt{philsci} eprints available at \url{http://philsci-archive.pitt.edu/}.\\}
+\newcommand*{\biorxivsi}{\texttt{bioRxiv} eprints available at \url{http://biorxiv.org/}.\\}
+\newcommand*{\arxiveprint}[1]{\global\def\arxivp{\arxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
+\texttt{\urlalt{http://arxiv.org/abs/#1}{arXiv:\hspace{0pt}#1}}%
+%\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
+%\renewcommand{\arxivnote}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.}
+}
+\newcommand*{\mparceprint}[1]{\global\def\mparcp{\mparcsi}%\citeauthor{0mparccite}\addtocategory{ifarchcit}{0mparccite}%eprint
+\texttt{\urlalt{http://www.ma.utexas.edu/mp_arc-bin/mpa?yn=#1}{mp\_arc:\hspace{0pt}#1}}%
+%\texttt{\href{http://www.ma.utexas.edu/mp_arc-bin/mpa?yn=#1}{\protect\url{mp_arc:#1}}}%
+%\providecommand{\mparcnote}{\texttt{mp_arc} eprints available at \url{http://www.ma.utexas.edu/mp_arc/}.}
+}
+\newcommand*{\philscieprint}[1]{\global\def\philscip{\philscisi}%\citeauthor{0philscicite}\addtocategory{ifarchcit}{0philscicite}%eprint
+\texttt{\urlalt{http://philsci-archive.pitt.edu/archive/#1}{PhilSci:\hspace{0pt}#1}}%
+%\texttt{\href{http://philsci-archive.pitt.edu/archive/#1}{\protect\url{PhilSci:#1}}}%
+%\providecommand{\mparcnote}{\texttt{philsci} eprints available at \url{http://philsci-archive.pitt.edu/}.}
+}
+\newcommand*{\biorxiveprint}[1]{\global\def\biorxivp{\biorxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
+\texttt{\urlalt{http://biorxiv.org/content/early/#1}{bioRxiv:\hspace{0pt}#1}}%
+%\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
+%\renewcommand{\arxivnote}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.}
+}
+
+\usepackage{graphicx}
 
 \usepackage{hyperref}
-\providecommand{\href}[2]{#2}
-\providecommand{\eprint}[2]{\texttt{\href{#1}{#2}}}
+\usepackage[depth=4]{bookmark}
+\hypersetup{colorlinks=true,bookmarksnumbered,pdfborder={0 0 0.25},citebordercolor={0.2 0.1333 0.5333},%bluish
+citecolor=myblue,linkbordercolor={0.0667 0.4667 0.2},%greenish
+linkcolor=myred,urlbordercolor={0.5333 0.1333 0.3333},%reddish
+urlcolor=mygreen,breaklinks=true,pdftitle={\pdftitle},pdfauthor={\pdfauthor}}
+
+% \usepackage[vertfit=local]{breakurl}% only for arXiv
 \providecommand*{\urlalt}{\href}
-\hypersetup{
-colorlinks=true,%
-bookmarksnumbered,%
-pdfborder={0 0 0.25},%
-citebordercolor={0.2 0.1333 0.5333},%bluish
-citecolor=myblue,%
-linkbordercolor={0.0667 0.4667 0.2},%greenish
-linkcolor=myred,%
-urlbordercolor={0.5333 0.1333 0.3333},%reddish
-urlcolor=mygreen,%
-breaklinks=true,%
-pdftitle = {Hidden network-size assumptions in the maximum-entropy method},%***
-pdfauthor = {Rostami, Porta Mana, Torre}%***
+
+\usepackage{datetime2}
+\DTMnewdatestyle{mydate}%
+{% definitions
+\renewcommand*{\DTMdisplaydate}[4]{%
+\number##3\ \DTMenglishmonthname{##2} ##1}%
+\renewcommand*{\DTMDisplaydate}{\DTMdisplaydate}%
 }
-\usepackage[depth=4]{bookmark}
-\usepackage{mathdots}
+\DTMsetdatestyle{mydate}
 
-%%%%%%%%%%%%%% macros
-\newcommand*{\widebar}[1]{{\mkern 1.5mu\skew{2}\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}}
+% handling orphan/widow lines:
+\clubpenalty=10000
+\widowpenalty=10000
+\raggedbottom
+
+\selectlanguage{british}\frenchspacing
+
+% to compile a camera-ready version, add the [final] option, e.g.:
+% \usepackage[final]{nips_2017}
+
+%\usepackage[utf8]{inputenc} % allow utf-8 input
+%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+%\usepackage{hyperref}       % hyperlinks
+%\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+%\usepackage{amsfonts}       % blackboard math symbols
+%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+%\usepackage{microtype}      % microtypography
+
+% @@@@@@@@@@ new macros @@@@@@@@@@
+% Common ones - uncomment as needed
+%\providecommand{\nequiv}{\not\equiv}
+%\providecommand{\coloneqq}{\mathrel{\mathop:}=}
+%\providecommand{\eqqcolon}{=\mathrel{\mathop:}}
+%\providecommand{\varprod}{\prod}
+%\newcommand*{\de}{\partialup}%partial diff
+%\newcommand*{\pu}{\piup}%constant pi
+%\newcommand*{\delt}{\deltaup}%Kronecker, Dirac
+%\newcommand*{\eps}{\varepsilonup}%Levi-Civita, Heaviside
+%\newcommand*{\riem}{\zetaup}%Riemann zeta
+%\providecommand{\degree}{\textdegree}% degree
+%\newcommand*{\celsius}{\textcelsius}% degree Celsius
+%\newcommand*{\micro}{\textmu}% degree Celsius
+%\newcommand*{\I}{\mathrm{i}}%imaginary unit
+%\newcommand*{\e}{\mathrm{e}}%Neper
+%\newcommand*{\di}{\mathrm{d}}%differential
+%\newcommand*{\Di}{\mathrm{D}}%capital differential
+%\newcommand*{\planckc}{\hslash}
+%\newcommand*{\avogn}{N_{\textrm{A}}}
+%\newcommand*{\NN}{\bm{\mathrm{N}}}
+%\newcommand*{\ZZ}{\bm{\mathrm{Z}}}
+%\newcommand*{\QQ}{\bm{\mathrm{Q}}}
+%\newcommand*{\RR}{\bm{\mathrm{R}}}
+%\newcommand*{\CC}{\bm{\mathrm{C}}}
+%\newcommand*{\nabl}{\bm{\nabla}}%nabla
+%\DeclareMathOperator{\lb}{lb}%base 2 log
+%\DeclareMathOperator{\tr}{tr}%trace
+%\DeclareMathOperator{\card}{card}%cardinality
+%\DeclareMathOperator{\im}{Im}%im part
+%\DeclareMathOperator{\re}{Re}%re part
+%\DeclareMathOperator{\sgn}{sgn}%signum
+%\DeclareMathOperator{\ent}{ent}%integer less or equal to
+%\DeclareMathOperator{\Ord}{O}%same order as
+%\DeclareMathOperator{\ord}{o}%lower order than
+%\newcommand*{\incr}{\triangle}%finite increment
+%\newcommand*{\defd}{\coloneqq}
+%\newcommand*{\defs}{\eqqcolon}
+%\newcommand*{\Land}{\bigwedge}
+%\newcommand*{\Lor}{\bigvee}
+%\newcommand*{\lland}{\mathbin{\ \land\ }}
+%\newcommand*{\llor}{\mathbin{\ \lor\ }}
+%\newcommand*{\lonlyif}{\mathbin{\Rightarrow}}%implies
+%\newcommand*{\limplies}{\mathbin{\Rightarrow}}%implies
+%\newcommand*{\mimplies}{\Rightarrow}%implies
+%\newcommand*{\liff}{\mathbin{\Leftrightarrow}}%if and only if
+%\newcommand*{\cond}{\mathpunct{|}}%conditional sign (in probabilities)
+%\newcommand*{\lcond}{\mathpunct{|\ }}%conditional sign (in probabilities)
+%\newcommand*{\bigcond}{\mathpunct{\big|}}%conditional sign (in probabilities)
+%\newcommand*{\lbigcond}{\mathpunct{\big|\ }}%conditional sign (in probabilities)
+%\newcommand*{\suchthat}{\mid}%{\mathpunct{|}}%such that (eg in sets)
+%\newcommand*{\bigst}{\mathpunct{\big|}}%such that (eg in sets)
+%\newcommand*{\with}{\colon}%with (list of indices)
+%\newcommand*{\mul}{\times}%multiplication
+%\newcommand*{\inn}{\cdot}%inner product
+%\newcommand*{\dotv}{\mathord{\,\cdot\,}}%variable place
+%\newcommand*{\comp}{\circ}%composition of functions
+%\newcommand*{\con}{\mathbin{:}}%scal prod of tensors
+%\newcommand*{\equi}{\sim}%equivalent to 
+%\newcommand*{\corr}{\mathrel{\hat{=}}}%corresponds to
+%\providecommand{\varparallel}{\ensuremath{\mathbin{/\mkern-7mu/}}}%parallel (tentative symbol)
+%\renewcommand{\le}{\leqslant}%less or equal
+%\renewcommand{\ge}{\geqslant}%greater or equal
+%\DeclarePairedDelimiter\clcl{[}{]}
+%\DeclarePairedDelimiter\clop{[}{[}
+%\DeclarePairedDelimiter\opcl{]}{]}
+%\DeclarePairedDelimiter\opop{]}{[}
+%\DeclarePairedDelimiter\abs{\lvert}{\rvert}
+%\DeclarePairedDelimiter\norm{\lVert}{\rVert}
+%\DeclarePairedDelimiter\set{\{}{\}}
+%\DeclareMathOperator{\pr}{P}%probability
+%\newcommand*{\pf}{\mathrm{p}}%probability
+%\newcommand*{\p}{\mathrm{P}}%probability
+%\newcommand*{\tf}{\mathrm{T}}%probability
+%\renewcommand*{\|}{\cond}
+%\newcommand*{\+}{\lor}
+%\renewcommand{\*}{\land}
+%\newcommand*{\sect}{\S}% Sect.~
+%\newcommand*{\sects}{\S\S}% Sect.~
+%\newcommand*{\chap}{ch.}%
+%\newcommand*{\chaps}{chs}%
+%\newcommand*{\fn}{fn}%
+%\newcommand*{\eqn}{eq.}%
+%\newcommand*{\eqns}{eqs}%
+%\newcommand*{\fig}{fig.}%
+%\newcommand*{\figs}{figs}%
+%\newcommand*{\vs}{{vs.}}
+%\newcommand*{\etc}{{etc.}}
+%\newcommand*{\ie}{{i.e.}}
+%\newcommand*{\ca}{{c.}}
+%\newcommand*{\Ie}{{I.e.}}
+%\newcommand*{\Eg}{{E.g.}}
+%\newcommand*{\eg}{{e.g.}}
+%\newcommand*{\viz}{{viz}}
+%\newcommand*{\cf}{{cf.}}
+%\newcommand*{\Cf}{{Cf.}}
+%\newcommand*{\vd}{{v.}}
+%\newcommand*{\Vd}{{V.}}
+%\newcommand*{\etal}{{et al.}}
+%\newcommand*{\etsim}{{et sim.}}
+%\newcommand*{\ibid}{{ibid.}}
+%\newcommand*{\sic}{{sic}}
+%\newcommand*{\id}{\mathte{I}}%id matrix
+%\newcommand*{\nbd}{\nobreakdash}%
+%\newcommand*{\bd}{\hspace{0pt}}%
+%\def\hy{-\penalty0\hskip0pt\relax}
+%\newcommand*{\labelbis}[1]{\tag*{(\ref{#1})$_\text{r}$}}
+%\newcommand*{\mathbox}[2][.8]{\parbox[t]{#1\columnwidth}{#2}}
+%\newcommand*{\zerob}[1]{\makebox[0pt][l]{#1}}
+%\newcommand*{\tprod}{\mathop{\textstyle\prod}\nolimits}
+%\newcommand*{\tsum}{\mathop{\textstyle\sum}\nolimits}
+%\newcommand*{\tint}{\begingroup\textstyle\int\endgroup\nolimits}
+%\newcommand*{\tland}{\mathop{\textstyle\bigwedge}\nolimits}
+%\newcommand*{\tlor}{\mathop{\textstyle\bigvee}\nolimits}
+%\newcommand*{\sprod}{\mathop{\textstyle\prod}}
+%\newcommand*{\ssum}{\mathop{\textstyle\sum}}
+%\newcommand*{\sint}{\begingroup\textstyle\int\endgroup}
+%\newcommand*{\sland}{\mathop{\textstyle\bigwedge}}
+%\newcommand*{\slor}{\mathop{\textstyle\bigvee}}
+%\newcommand*{\T}{^\intercal}%transpose
+%\newcommand*{\E}{\mathrm{E}}
+%\DeclarePairedDelimiter\expp{(}{)}
+%\newcommand*{\expe}{\E\expp}%round
+%\newcommand*{\expeb}{\E\clcl}%square
+%%\newcommand*{\QEM}%{\textnormal{$\Box$}}%{\ding{167}}
+%\newcommand*{\qem}{\leavevmode\unskip\penalty9999 \hbox{}\nobreak\hfill
+%\quad\hbox{\QEM}}
+\definecolor{notecolour}{RGB}{68,170,153}
+\newcommand*{\puzzle}{{\fontencoding{U}\fontfamily{fontawesometwo}\selectfont\symbol{225}}}
+\newcommand*{\mynote}[1]{ {\color{notecolour}\puzzle\ #1}}
+\newcommand*{\widebar}[1]{{\mkern1.5mu\skew{2}\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}}
+
 \newcommand*{\av}{\overline} %pop average
@@
 \newcommand*{\me}{maximum-entropy}
-\title{Hidden network-size assumptions\\in the maximum-entropy method}
+
+%@@@@@@@@@@ new macros end @@@@@@@@@@
+
+\selectlanguage{british}\frenchspacing
+
+\title{\pdftitle}
 
@@
   \And E. Torre\\Chocolateland  
-  %
-% David S.~Hippocampus\thanks{Use footnote for providing further information
-%   about author (webpage, alternative
+  % David S.~Hippocampus\thanks{Use footnote for providing further
+  %   information about author (webpage, alternative
   %   address)---\emph{not} for acknowledging funding agencies.} \\
@@
 
-The \me\ method is been used in neuroscience blahblah***
+The \me\ method is used in neuroscience blahblah***
+
+Let's clarify at once that our discussion pertains to binary networks at
+one specific instant, or short window, in time. It is possible to
+generalize it to multi-state neuron models and network dynamics; but for
+simplicity neurons are here assumed to be in a fixed, active \enquote{$1$}
+or inactive \enquote{$0$} state; and evolution, change, time correlations,
+and similar concepts do not concern us.
+
+Maximum-entropy models are used for a variety of reasons, sometimes not
+completely clear. For our purposes we can simply say that the \me\ method
+is assumed to generate a \enquote{maximally noncommital} \cite{jaynes1963}
+probability distribution. We intend the quoted expression only as an
+umbrella term, also because it  means very little without further
+clarifications.
+
+The neurons recorded in these kinds of experiments are not usually
+specifically chosen, and from their observation we expect to learn
+something about all other neurons in the same brain area. That is, they are
+assumed to be a \enquote{representative random sample} of the neurons in
+that area. Our discussion hinges on this often just tacitly understood
+assumption.
+
+\mynote{V1 }In this note we would like to show that there is a contradiction in
+applying \me\ to a \enquote{representative random sample} of neurons to
+generate a \enquote{maximally noncommital} distribution.\\
+\mynote{V2 }In this note we would like to show that there is a clash between the
+\enquote{representative random sample} spirit and the \enquote{maximally
+  noncommital} spirit of \me\ models applied to such a sample of neurons.
+
+The contradiction is this. Assume that our sample of neurons is
+representative of some population. If we assign a \me\ distribution to the
+sample, then we cannot assign a \me\ distribution to the full population.
+Vice versa, If we assign a \me\ distribution to the full population, then
+we cannot assign a \me\ distribution to the sample. This impossibility
+holds at least for \me\ distribution with moment constraints, and even if
+the orders of the moments constrained in the sample and in the full
+population are different.
 
@@
 
-% Do not include acknowledgments in the anonymized submission, only in the
-% final paper.
+Use unnumbered third level headings for the acknowledgments. All
+acknowledgments go at the end of the paper. Do not include
+acknowledgments in the anonymized submission, only in the final paper.
 
@@
 
+% \small
+
+% [1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
+% for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
+% T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
+%   Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.
+
+% [2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
+%   Exploring Realistic Neural Models with the GEneral NEural SImulation
+%   System.}  New York: TELOS/Springer--Verlag.
+
+% [3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
+% learning and recall at excitatory recurrent synapses and cholinergic
+% modulation in rat hippocampal region CA3. {\it Journal of
+%   Neuroscience} {\bf 15}(7):5249-5262.
+
 \end{document}

### 2017-05-12 maxNt.tex
add to intro

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-12T13:15:15+0200
+%% Last-Updated: 2017-05-12T13:32:47+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Maximum-entropy models are used for a variety of reasons, sometimes not
-completely clear. For our purposes we can simply say that the \me\ method
-is assumed to generate a \enquote{maximally noncommital} \cite{jaynes1963}
-probability distribution. We intend the quoted expression only as an
-umbrella term, also because it  means very little without further
+Maximum-entropy models are used for a variety of reasons; sometimes not
+completely clear ones. For our purposes we can simply say that the \me\
+method is assumed to generate a \enquote{maximally noncommital}
+\cite{jaynes1963} probability distribution. We intend the quoted expression
+only as an umbrella term, also because it means very little without further
 clarifications.
 
-The neurons recorded in these kinds of experiments are not usually
+The neurons recorded in these kinds of experiments are usually not
 specifically chosen, and from their observation we expect to learn
@@
 population are different.
+
+In other words, if our sample is \enquote{random} and
+\enquote{representative} of some population, then we must choose which of
+them has a \enquote{maximally noncommittal} distribution: the sample or the
+population. We can't choose both. Which choice is most meaningful? It goes
+without saying that if either is \enquote{maximally noncommittal}, the
+other must be somehow \enquote{committal}.
+
+***
 

### 2017-05-13 maxNt.tex
add to intro
Complete intro

--- 2017-05-12 maxNt.tex
@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-12T13:15:15+0200
+%% Last-Updated: 2017-05-13T17:24:25+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\firstdraft}{4 November 2015}
-\newcommand*{\pdftitle}{Hidden network-size assumptions\\in the maximum-entropy method}
+\newcommand*{\pdftitle}{Hidden  assumptions on population size\\in the maximum-entropy method}
 \newcommand*{\headtitle}{\pdftitle}
@@
 
+\usepackage{newtxmath}
 %\usepackage{pifont}
@@
 
-%\newcommand*{\bmmax}{0} % reduce number of bold fonts, before bm
+\newcommand*{\bmmax}{3} % reduce number of bold fonts, before bm
 \newcommand*{\hmmax}{0} % reduce number of heavy fonts, before bm
@@
 
+\let\openbox\relax
 \usepackage{amsthm}
@@
 %\newcommand*{\QQ}{\bm{\mathrm{Q}}}
-%\newcommand*{\RR}{\bm{\mathrm{R}}}
+\newcommand*{\RR}{\bm{\mathrm{R}}}
 %\newcommand*{\CC}{\bm{\mathrm{C}}}
@@
 %\newcommand*{\incr}{\triangle}%finite increment
-%\newcommand*{\defd}{\coloneqq}
+\newcommand*{\defd}{\coloneqq}
 %\newcommand*{\defs}{\eqqcolon}
@@
 %\newcommand*{\lonlyif}{\mathbin{\Rightarrow}}%implies
-%\newcommand*{\limplies}{\mathbin{\Rightarrow}}%implies
+\newcommand*{\limplies}{\mathbin{\Rightarrow}}%implies
 %\newcommand*{\mimplies}{\Rightarrow}%implies
 %\newcommand*{\liff}{\mathbin{\Leftrightarrow}}%if and only if
-%\newcommand*{\cond}{\mathpunct{|}}%conditional sign (in probabilities)
+\renewcommand*{\|}{\mathpunct{|}}%conditional sign (in probabilities)
 %\newcommand*{\lcond}{\mathpunct{|\ }}%conditional sign (in probabilities)
@@
 %\newcommand*{\inn}{\cdot}%inner product
-%\newcommand*{\dotv}{\mathord{\,\cdot\,}}%variable place
+\newcommand*{\dotv}{\mathord{\,\cdot\,}}%variable place
 %\newcommand*{\comp}{\circ}%composition of functions
@@
 %\providecommand{\varparallel}{\ensuremath{\mathbin{/\mkern-7mu/}}}%parallel (tentative symbol)
-%\renewcommand{\le}{\leqslant}%less or equal
-%\renewcommand{\ge}{\geqslant}%greater or equal
-%\DeclarePairedDelimiter\clcl{[}{]}
+\renewcommand{\le}{\leqslant}%less or equal
+\renewcommand{\ge}{\geqslant}%greater or equal
+\DeclarePairedDelimiter\clcl{[}{]}
 %\DeclarePairedDelimiter\clop{[}{[}
@@
 %\DeclarePairedDelimiter\norm{\lVert}{\rVert}
-%\DeclarePairedDelimiter\set{\{}{\}}
+\DeclarePairedDelimiter\set{\{}{\}}
 %\DeclareMathOperator{\pr}{P}%probability
-%\newcommand*{\pf}{\mathrm{p}}%probability
-%\newcommand*{\p}{\mathrm{P}}%probability
+\newcommand*{\pf}{\mathrm{p}}%probability
+\newcommand*{\p}{\mathrm{P}}%probability
 %\newcommand*{\tf}{\mathrm{T}}%probability
-%\renewcommand*{\|}{\cond}
+%\renewcommand*{\|}{\|}
 %\newcommand*{\+}{\lor}
 %\renewcommand{\*}{\land}
-%\newcommand*{\sect}{\S}% Sect.~
-%\newcommand*{\sects}{\S\S}% Sect.~
-%\newcommand*{\chap}{ch.}%
-%\newcommand*{\chaps}{chs}%
+\newcommand*{\sect}{\S}% Sect.~
+\newcommand*{\sects}{\S\S}% Sect.~
+\newcommand*{\chap}{ch.}%
+\newcommand*{\chaps}{chs}%
 %\newcommand*{\fn}{fn}%
-%\newcommand*{\eqn}{eq.}%
-%\newcommand*{\eqns}{eqs}%
-%\newcommand*{\fig}{fig.}%
-%\newcommand*{\figs}{figs}%
+\newcommand*{\eqn}{eq.}%
+\newcommand*{\eqns}{eqs}%
+\newcommand*{\fig}{fig.}%
+\newcommand*{\figs}{figs}%
 %\newcommand*{\vs}{{vs.}}
 %\newcommand*{\etc}{{etc.}}
-%\newcommand*{\ie}{{i.e.}}
+\newcommand*{\ie}{{i.e.}}
 %\newcommand*{\ca}{{c.}}
@@
 %\newcommand*{\Eg}{{E.g.}}
-%\newcommand*{\eg}{{e.g.}}
+\newcommand*{\eg}{{e.g.}}
 %\newcommand*{\viz}{{viz}}
-%\newcommand*{\cf}{{cf.}}
+\newcommand*{\cf}{{cf.}}
 %\newcommand*{\Cf}{{Cf.}}
@@
 %\newcommand*{\T}{^\intercal}%transpose
-%\newcommand*{\E}{\mathrm{E}}
-%\DeclarePairedDelimiter\expp{(}{)}
-%\newcommand*{\expe}{\E\expp}%round
-%\newcommand*{\expeb}{\E\clcl}%square
+\newcommand*{\E}{\mathrm{E}}
+\DeclarePairedDelimiter\expp{(}{)}
+\newcommand*{\expe}{\E\expp}%round
+\newcommand*{\expeb}{\E\clcl}%square
 %%\newcommand*{\QEM}%{\textnormal{$\Box$}}%{\ding{167}}
@@
 %\quad\hbox{\QEM}}
+
+\newtheoremstyle{simple}%
+{0\parsep}%
+{0\parsep}%
+{\footnotesize}%
+{}%
+{}%
+{}%
+{0pt}%
+{}%
+\theoremstyle{simple}
+\newtheorem*{simplenote}{}
+
 \definecolor{notecolour}{RGB}{68,170,153}
@@
 
-\newcommand*{\av}{\overline} %pop average
+\newcommand*{\av}{\widebar} %pop average
 \newcommand*{\sav}{\widehat} %subpop average
-\newcommand*{\yxx}{x}
+
+\newcommand*{\yxx}{x}%subpop state
 \newcommand*{\yx}{\bm{\yxx}}%subpop state
 \newcommand*{\yxs}{\sav{\yx}}%subpop av state
-\newcommand*{\yxxs}{\sav{\yx\yx}}
+\newcommand*{\yxxs}{\sav{\yx\yx}}%subpop av state
 \newcommand*{\yX}{\bm{X}}%pop state
@@
 \newcommand*{\yRf}{\av{\yR}}%pop av value
-\newcommand*{\yN}{N}
-\newcommand*{\yM}{M}
-\newcommand*{\yee}{L}
+%conditional assumptions
+\newcommand*{\yH}{\varEta}
+\newcommand*{\yD}{\varDelta}
+\newcommand*{\yHa}{\varEta'}
+\newcommand*{\yHb}{\varEta''}
+%Lagr multipliers
+\newcommand*{\yL}{\varLambda}
+\newcommand*{\yl}{\lambda}
+\newcommand*{\yk}{\kappa}
+%entropies
+\newcommand*{\ysh}{H}
+\newcommand*{\ybu}{H_{\text{B}}}
+
+\newcommand*{\prop}[1]{`#1'}
 
@@
 \author{ V. Rostami\\
-  Forschungszentrum Jülich INM-6\\
+  Forschungszentrum Jülich INM-6,\\Hitlerland\\
   \texttt{v.rostami@fz-juelich.de}
@@
 
-Let's clarify at once that our discussion pertains to binary networks at
-one specific instant, or short window, in time. It is possible to
-generalize it to multi-state neuron models and network dynamics; but for
+Let's clarify at once that our discussion pertains to a binary population at
+one specific instant or short window in time. It is possible to
+similarly discuss multi-state neuron models and population dynamics; but for
 simplicity neurons are here assumed to be in a fixed, active \enquote{$1$}
@@
 
-Maximum-entropy models are used for a variety of reasons, sometimes not
-completely clear. For our purposes we can simply say that the \me\ method
-is assumed to generate a \enquote{maximally noncommital} \cite{jaynes1963}
-probability distribution. We intend the quoted expression only as an
-umbrella term, also because it  means very little without further
-clarifications.
+Maximum-entropy models are used for a variety of (sometimes not completely
+clear) reasons. For our purposes let's say that the \me\ method is assumed
+to generate a \enquote{maximally noncommital} \cite{jaynes1963} probability
+distribution. We intend the quoted expression only as an umbrella term,
+also because it means very little without further clarifications.
 
-The neurons recorded in these kinds of experiments are not usually
-specifically chosen, and from their observation we expect to learn
+The neurons recorded in these kinds of experiments are usually picked out
+in a way beyond our control, and from their observation we expect to learn
 something about all other neurons in the same brain area. That is, they are
@@
 
-\mynote{V1 }In this note we would like to show that there is a contradiction in
+In this note we would like to show that there is a contradiction in
 applying \me\ to a \enquote{representative random sample} of neurons to
-generate a \enquote{maximally noncommital} distribution.\\
-\mynote{V2 }In this note we would like to show that there is a clash between the
-\enquote{representative random sample} spirit and the \enquote{maximally
-  noncommital} spirit of \me\ models applied to such a sample of neurons.
+generate a \enquote{maximally noncommital} distribution for that sample.
+% \\
+% \mynote{V2 }In this note we would like to show that there is a clash between the
+% \enquote{representative random sample} spirit and the \enquote{maximally
+%   noncommital} spirit of \me\ models applied to such a sample of neurons.
 
@@
 
+In other words, if our sample is a \enquote{random representative} of some
+population, then we must choose which has a \enquote{maximally
+  noncommittal} distribution: the sample or the population. We can't choose
+both. It goes without saying that if one is \enquote{maximally
+  noncommittal}, the other must be somehow \enquote{committal}. Which
+choice is most meaningful?
+
+In the rest of the paper we will mathematically show the contradiction
+above and generate a \me\ distribution for the full population rather than
+the sample. We will also present some probability relations relevant to
+random sampling. These relations are well-known in survey sampling and in
+the pedagogic problem of drawing from an urn without replacement, yet they
+are somewhat hard to find explicitly written in the neuroscientific
+literature.
+
+Our notation follows ISO and ANSI standards
+\citep{iso1993,ieee1993,nist1995} but for the use of the comma \enquote{,}
+to denote logical conjunction. Probability notation follows Jaynes
+\citep{jaynes1994_r2003}.
+
+\section{Probability of random representative samples}
+\label{sec:prob_samples}
+
+\subsection{Setup}
+\label{sec:setup}
+
+Consider a population of $N$ binary neurons with states $(X_1,\dotsc,X_N)$
+having fixed but unknown binary values $(R_1,\dotsc,R_N)$, with
+$R_i \text{ in } \set{0,1}$, vectorially written $\yX=\yR$. For example,
+$\yX$ can represent the state of a population at a particular time. We call
+the neurons \enquote{units} to lend some generality to our discussion. We
+shall make statements about the whole population of $N$ units and about a
+subpopulation of $n$ units; the word \enquote{population} will always refer to
+the \emph{whole} population. The subpopulation states and their values are
+denoted by lowercase letters: $(x_1, \dotsc, x_n)\equiv\yx$ and
+$(r_1, \dotsc, r_n)\equiv\yr$; but note that $x_i \equiv X_{j_i}$ and
+$r_i \equiv R_{j_i}$ for some distinct $j_1,\dotsc,j_n$. We shall also make
+statements about the population-averaged state, or \emph{population average}:
+\begin{align}
+  \label{eq:population-av}
+   \yXf &\defd (X_1 + \dotsb + X_N)/N,
+\\
+\intertext{and the subpopulation-averaged state, or \emph{subpopulation average}:}
+\label{eq:subpopulation-av}
+\yxs &\defd (x_1 + \dotsc + x_n)/n.
+\end{align}
+The quantities $N\yXf$ and $n\yxs$ represent the total number of active
+units in the population and the subpopulation. Quantities like $\yRf$ and $\yrs$
+are defined analogously. The averaging operators $\sav{\dotv}$ and
+$\av{\dotv}$ are also extended to averages of $\tbinom{n}{m}$ or
+$\tbinom{N}{m}$ products of $m$ states; \eg,
+\begin{align}
+\av{\yX \yX} &\defd
+\tbinom{N}{2}^{-1} (X_1 X_2 + X_1 X_3  + \dotsb +  X_{N-1} X_N),
+\\[2\jot]
+\sav{\yx\yx\yx} &\defd
+\tbinom{n}{3}^{-1} (x_1 x_2 x_3 + x_1 x_2 x_4 + \dotsb + x_{n-2} x_{n-1} x_n),
+\end{align}
+and so on.
+
+\subsection{Assumptions}
+\label{sec:assumptions}
+
+Our uncertainty about the population state is represented by the joint
+probability distribution of the individual states, from which we can
+derive all other probabilities of interest. We denote it by
+\begin{equation}
+  \label{eq:joint_plaus}
+  \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \| \yH) \quad\text{or}\quad
+\p(\yX =\yR \| \yH).
+\end{equation}
+Such probability is conditional on our state of knowledge, \ie\ the
+evidence and assumptions backing our probability assignments, denoted by
+the proposition $\yH$.
+
+In the present discussion, $\yH$ is a state of knowledge that leads to two
+specific properties in our probability assignments:
+
+\medskip
+\begin{enumerate}%[wide,label=$\yH$\arabic*.]
+\item \emph{Permutation symmetry}, expressed as the invariance of the
+  joint distribution~\eqref{eq:joint_plaus} under arbitrary permutations of
+  the units's labels:
+\begin{multline}
+  \label{eq:homog_math}
+  \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \| \yH) ={}
+\\ 
+\shoveright{\p(X_1=R_{\pi(1)}, X_2=R_{\pi(2)}, \dotsc, X_N=R_{\pi(N)} \| \yH)}
+\\
+\text{for any permutation $\pi$}.
+\end{multline}
+This property can reflect two very different states of knowledge: physical
+homogeneity of the population, or symmetry in our ignorance about the population.
+This property is called \emph{finite exchangeability} in the Bayesian
+literature and its basis, consequences, and alternatives to it are discussed
+in \sect~\ref{sec:stoch_symm}.
+
+\medskip
+
+\item The population average $\yXf$ has a particular distribution $Q$:
+\begin{equation}
+  \label{eq:pop_average_Q}
+  \p(\yXf = A \| \yH)
+% \equiv \pr(x_1+ \dotsb + x_N = N A \| \varEta) 
+=  Q(A),
+\qquad
+A \in \set*{0,\tfrac{1}{N},\tfrac{2}{N},\dotsc,1}.
+\end{equation}
+For the moment we are not concerned about the specific form of $Q$ and
+about how it was assigned: it could, \eg, arise from maximum-entropy
+arguments
+\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,aczeletal1975,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013}
+used with data on the population.
+\end{enumerate}
+
+\subsection{Formulae}
+%\section{\ding{41} {\itshape The mathematical relations} \ding{41}}
+%\section*{3.\hspace{\stretch{1}}The mathematical relations\hspace{\stretch{1}}\mbox{}}
+%\addcontentsline{toc}{section}{The mathematical relations}
+\label{sec:main_formulae}
+
+The state of knowledge $\yH$ has the following six (not independent) main
+consequences for our probability assignments:
+
+\begin{enumerate}%[wide,label=\textbf{\emph{\Roman*}}.]
+  \medskip\item\label{item:tot_plaus}Probability for the population state:
+\begin{equation}
+  \label{eq:joint_plaus_N_homog}
+  \p(\yX = \yR \| H) = \binom{N}{N\yRf}^{-1} Q(\yRf).
+\end{equation}
+
+\medskip
+\item\label{item:marginal}Probability for the state $\yx$ of any subpopulation
+  of $n$ units:
+\begin{equation}
+  \label{eq:marginal}
+    \p(\yx = \yr \| \yH) =
+\sum_{N A =0}^{N} 
+\binom{N-n}{N A-n\yrs}\binom{N}{N A}^{-1}\,
+Q(A).
+\end{equation}
+Note that the only summands contributing to this sum are those for which
+$n\yrs \le NA \le N$; the others are zero because by definition
+$\tbinom{M}{y}=0$ if $y<0$. This remark applies to all the sums of this
+kind in the rest of this Note.
+
+\medskip
+\item\label{item:conditional}Probability for the subpopulation
+  state conditional on a population state:
+\begin{equation}
+  \label{eq:conditional_total_state}
+  \p(\yx = \yr \| \yX = \yR, \yH)
+=
+\binom{N-n}{N \yRf-n \yrs}.
+\end{equation}
+
+\medskip
+\item\label{item:marginal_average}Probability for the
+  subpopulation average $\yxs$:
+\begin{multline}
+  \label{eq:subpop_average}
+  \p(\yxs = a  \| \yH) = \binom{n}{n a}
+ \sum_{N A=0}^{N}
+\binom{N-n}{N A-n a}\binom{N}{N A}^{-1}\,
+Q(A),
+\\
+a \in \set*{0,\tfrac{1}{n},\tfrac{2}{n},\dotsc,1}.
+\end{multline}
+
+\medskip
+\item\label{item:conditional_average}Probability for the subpopulation
+  average conditional on the population average:
+\begin{equation}
+  \label{eq:conditional_prob}
+  \p(\yxs = a \| \yXf = A, \yH)
+=
+\binom{n}{n a}\binom{N-n}{N A-n a}\binom{N}{N A}^{-1}.
+\end{equation}
+
+\medskip
+\item\label{item:moments}The product of the states of any $m$ distinct
+  units from a given subpopulation,
+  \begin{equation*}
+    x_{i_1} x_{i_2} \dotsm x_{i_m},
+    \qquad 1\le i_1 < i_2 < \dotsb < i_m \le n
+  \end{equation*}
+  has an expectation equal to that of the subpopulation average of such
+  products, is independent of the subpopulation size $n$:
+\begin{subequations}
+\label{eq:expe_products}
+\begin{align}
+  \expe{x_{i_1} \dotsm x_{i_m} \| \yH}
+&=
+\expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
+=
+\expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH},
+\\
+\intertext{and has an explicit expression in terms of $Q$:}
+  \begin{split}
+\expe{x_{i_1} \dotsm x_{i_m} \| \yH}
+&=
+\binom{N}{m}^{-1}
+%\frac{(N-m)!}{N!}
+\sum_{N A=0}^N %m!\, 
+\binom{N A}{m}\, Q(A)
+\\
+&\equiv
+%\frac{(N-m)!}{N!}
+\sum_{N A=0}^N %m!\, 
+\binom{N-m}{N A-m}
+\binom{N}{N A}^{-1} Q(A).
+\end{split}
+\end{align}
+\end{subequations}
+\end{enumerate}
+
+\medskip
+
+A useful relation connects the expectation of a
+product~\eqref{eq:expe_products} and the $m$th \emph{factorial moment}
+\citep{potts1953} of the probability distributions for the averages. The
+$m$th factorial moment of the subpopulation average $\yxs$ is defined by
+\begin{equation}
+  \label{eq:fact_mom_defin}
+\vphantom{\underbrace{\yxs}_{m}}\expeb*{
+\smash{\underbrace{n\yxs\,(n\yxs-1)\dotsm \bigl(n\yxs-(m-1)\bigr)}_{\text{$m$ factors}}}
+\| \yH}
+\equiv \expeb*{\frac{(n\yxs)!}{(n\yxs-m)!} \| \yH},
+\end{equation}
+an analogous definition holding for $\yXf$. We have that
+\begin{equation}
+  \label{eq:eq_factmom_prod}
+\expe{x_{i_1} \dotsm x_{i_m} \| \yH}
+=
+  \tfrac{(n-m)!}{n!}\expeb*{\frac{(n\yxs)!}{(n\yxs-m)!} \| \yH}
+=
+  \tfrac{(N-m)!}{N!}\expeb*{\frac{(N\yXf)!}{(N\yXf-m)!} \| \yH}.
+\end{equation}
+
+As a consequence of the above relation, the first three moments of the
+probability distributions $\p(\yxs=a\|\yH)$ and $\p(\yXf=A\|\yH)$,
+are related by
+\begin{subequations}\label{eq:moments}
+  \begin{align}
+    \expe{\yxs \| \yH} &= \expe{\yXf \| \yH},
+    \\[2\jot]
+    \begin{split}
+    \expe{\yxs^2 \| \yH} &= \expe{\yXf \| \yH}
+                                  \frac{N-n}{(N-1)n}
+ +{}
+\\&\quad\quad \expe{\yXf^2 \| \yH}\frac{N\,(n-1)}{(N-1)n},
+\end{split}
+    \\[2\jot]
+    \begin{split}
+    \expe{\yxs^3 \| \yH} &=
+                              \expe{\yXf \| \yH}\frac{(N-n)(N-2n)}{(N-1)(N-2)n^2}+{}
+                                    \\
+                                    &\quad\quad\expe{\yXf^2 \| \yH}
+                                    \frac{3N\,(N-n)(n-1)}{(N-1)(N-2)n^2} +{}
+                                    \\
+                                    &\quad\quad\quad\expe{\yXf^3 \| \yH}
+                                    \frac{N^2\,(n-1)(n-2)}{(N-1)(N-2)n^2}.
+                                  \end{split}
+                                    % \\[2\jot]
+                                    % \expe{\yxs^4 \| \varEta} &=
+                                    % \!\begin{aligned}[t] &\expe{\yxf
+                                    %   \| \varEta}
+                                    %   \frac{(N-n)(N^2+N-6Nn+6n^2)}{(N-1)(N-2)(N-3)n^3}+{}
+                                    %   \\%[1.5\jot]
+                                    %   &\quad\expe{\yxf^2 \| \varEta}
+                                    %   \frac{(7N-11n+1)(N-n)(n-1)N}{(N-1)(N-2)(N-3)n^3}+{}
+                                    %   \\
+                                    %   &\quad\expe{\yxf^3 \| \varEta}
+                                    %   \frac{6(n-1)(n-2)(N-n)N^2}{(N-1)(N-2)(N-3)n^3}
+                                    %   +{}
+                                    %   \\
+                                    %   &\quad\expe{\yxf^4 \| \varEta}
+                                    %   \frac{(n-1)(n-2)(n-3)N^3}{(N-1)(N-2)(N-3)n^3}.
+                                    % \end{aligned}
+  \end{align}
+\end{subequations}
+Relations for higher moments can be obtained recursively from
+\eqn~\eqref{eq:eq_factmom_prod}. In general, this means that the two sets of
+first $m$ moments are related by a homogeneous linear transformation,
+\begin{equation}
+  \label{eq:lin_transf_moments}
+  \expe{\yxs^m \| \yH} = \sum_{l=1}^m M_{ml}(n,N) \expe{\yXf^l \| \yH},
+\end{equation}
+with a universal, lower-triangular transformation matrix $M_{ml}(n,N)$ that
+depends only on $n$, $N$, and the condition of symmetry~\eqref{eq:homog_math}.
+
+As intuition suggests, we have 
+%$\expe{\yxs^m \| \yH} \xrightarrow{n\to N} \expe{\yXf^m \| \yH}$.
+\begin{equation}
+  \label{eq:limit_moments}
+  \expe{\yxs^m \| \yH} \xrightarrow{n\to N} 
+  \expe{\yXf^m \| \yH},
+\qquad
+  \expe{\yxs^m \| \yH} \xrightarrow{n\to 1} 
+  \expe{\yXf \| \yH},
+\end{equation}
+the latter because ${x_i}^m=x_i$, since states are $\set{0,1}$-valued.
+
+\bigskip
+
+The core of the six mathematical relations above are
+\eqns~\eqref{eq:marginal} and~\eqref{eq:subpop_average}. The latter
+expresses the probability for the subpopulation average as a mixture of
+hypergeometric distributions
+\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
+with parameters $N,N\yXf,n$, weighted by the probabilities
+$\p(\yXf = A \|\yH)$ \citep[\cf][\sect~4, esp.\ \eqn~(22)]{kendall1967}.
+The connection between this mixture representation and the condition of
+symmetry~\eqref{eq:homog_math} is well-known in the Bayesian literature
+\citep{kendall1967,definetti1969b,heathetal1976,diaconis1977,diaconisetal1980,jaynes1986c}.
+
+\section{Examples of inferential use of the formulae}
+
+\subsection{From network to subnetwork}
+\label{sec:pop2sub_examples}
+
+Let us illustrate with an example how the probability distribution for the
+subnetwork average $\yxs$, determined by \eqn~\eqref{eq:subpop_average},
+changes with the subnetwork size $n$. Choose a network-average distribution
+$\p(\yXf=A\| \yH)$ belonging to the exponential family
+\cites[\sect~4.5.3]{bernardoetal1994}[see also][]{fortinietal2000}:
+\begin{equation}
+  \label{eq:full_pop_av_example_ME}
+  \p(\yXf = A \| \yH) = Q(A) \propto
+\binom{N}{N A}\exp[
+\yl_2\, N A\,(N A-1)/2 + \yl_1\, N A].
+\end{equation}
+This is the form obtained from the principle of maximum relative entropy
+\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,aczeletal1975,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013}
+with first and second moments as constraints and the reference distribution
+$Q_0$ defined by $Q_0(A) = 2^{-N}\binom{N}{N A}$, corresponding to a
+uniform probability distribution for the network state $\yX$.
+%file: distr_scaling_from_full_population.nb
+\begin{figure}[!b]
+\centering
+\includegraphics[width=0.95\columnwidth]{scaled_subpop_probs.pdf}%
+\caption{Probability distributions $\p(\yxs = a \| \yH)$ for different
+  subnetwork sizes $n$, obtained from a network probability
+  distribution $\p(\yXf = A \| \yH)$ having the maximum-etropy
+  form~\eqref{eq:full_pop_av_example_ME}.}
+\label{scaling_distr}
+\end{figure}
+\begin{figure}[!b]
+\centering
+\includegraphics[width=0.95\columnwidth]{scaling_subpop_moments.pdf}%
+\caption{Moments of the probability distributions $\p(\yxs =a \|
+  \yH)$ as functions of the subnetwork size $n$.}
+\label{scaling_moments}
+\end{figure}
+
+The probability distribution of \eqn~\eqref{eq:full_pop_av_example_ME} is
+plotted in \fig~\ref{scaling_distr}, together with the resulting
+subnetwork-average distributions $\p(\yxs = a \|\yH)$, for the case in
+which $N = 1000$ units, $\lambda_1=-2.55$, $\lambda_2=0.005$, and
+$n=10, 50, 100, 250$. The distributions become broader as $n$ decreases,
+and the minimum of the original distribution disappears; at the same time
+the finite-difference 
+\[\frac{\p(\yxs=a+1/n\|\yH)-\p(\yxs=a\|\yH)}{1/n}\]
+presents a sharp jump at this minimum when $n\approx 100$.
+
+To the eye familiar with maximum-entropy distributions, the
+subnetwork-average distributions of \fig~\ref{scaling_distr} do not look
+like maximum-entropy ones with second-moment constraints. In fact,
+they are not and \emph{cannot} be:
+\begin{equation}
+  \label{eq:maxent_form_sub}
+  \p(\yxs =a  \| \yH)
+\ne\kappa\binom{n}{n a}\exp[\yk_2\, n a\,(n a-1)/2 + \yk_1\, n a]
+%\quad\text{(false)}
+\end{equation}
+for any $\kappa, \yk_1, \yk_2$, unless $n=2$. This impossibility holds more
+generally for any number of constraints $m$ and subnetwork size $n$ such
+that $m<n$. The reason is simple: suppose we have assigned a
+maximum-entropy distribution with $m$ moment constraints as the
+distribution for the network average. If we want the same kind of
+distribution for a subnetwork of size $n$, we are free to play with $m+1$
+parameters (normalization included), but we must also satisfy the $n+1$
+equations corresponding to the marginalization~\eqref{eq:subpop_average}.
+This is generally impossible unless $m \ge n$. (Impossibilities of a
+similar kind appear in statistical mechanics, see \eg\
+ref.~\citep{maesetal1999}.)
+%file: test_ME_same_under_scaling.nb
+
+This fact can be significant for recent works
+\citep[\eg,][]{schneidmanetal2006,shlensetal2006,tkaciketal2006,marreetal2009,tkaciketal2009,ganmoretal2011,shimazakietal2012,tkaciketal2013,shimazakietal2015}
+in which a maximum-entropy probability distribution with second- or
+third-moment constraints is assigned to relatively small subnetworks
+($n < 200$) of neurons. If we assume that such subnetwork is part of a
+larger network, and assume the condition of symmetry~\eqref{eq:homog_math},
+then the larger network \emph{cannot} be assigned a maximum-entropy
+distribution with the same number of constraints. Vice versa, if we assign
+such a maximum-entropy distribution to the larger network, then none of its
+subnetwork of enough large size $n$ can be assigned a similar
+maximum-entropy distribution. See ref.~\citep{rostamietal2016} for a
+broader discussion of this fact and of its consequences.
+
+\medskip
+
+The dependence of the first four moments $\expe{\yxs^m \| \yH}$ as a
+function of size $n$ is shown in \fig~\ref{scaling_moments}. The moments
+become practically constant when $n \approx 100$ or larger. The
+expectations of $m$-tuple products of states
+$\expe{x_{i_1}\dotsm x_{i_m} \| \yH}$, proportional to the factorial
+moments, are not shown as they do not depend on $n$.
+
+\subsection{From subnetwork to network}
+\label{sec:from_sub_to_full}
+
+We have seen that, given the condition of symmetry~\eqref{eq:homog_math},
+the probability $\p(\yXf = A \|\yH)$ for the network average determines
+that of each subnetwork average, $\p(\yxs \|\yH)$, by the
+marginalization \eqn~\eqref{eq:subpop_average}. The reverse is trivially
+not true, since \eqn~\eqref{eq:subpop_average}, as a linear mapping from
+$\RR^{N+1}$ to $\RR^{n+1}$, with $N$ larger than $n$, is onto but not into.
+Assigning a probability distribution $\p(\yxs = a \|\yH)$ to a
+subnetwork average $\yxs$ does not determine a network distribution
+$\p(\yXf = A\|\yH)$: it only restricts the set of possible ones; this
+set can in principle be determined via linear-programming methods
+\citep{hailperin1965,hailperin1984,hailperin1996,hailperin2006,hailperin2011}.
+\begin{innote}
+  Analogous situations appear in the truth-valued logical calculus: if the
+  composite proposition $\varAlpha\limplies \varBeta$ is assigned the
+  truth-value \enquote{true}, then assigning $\varAlpha$ the value
+  \enquote{true} also determines the value of $\varBeta$, whereas assigning
+  $\varBeta$ the value \enquote{true} leaves the value of $\varAlpha$
+  undetermined.
+\end{innote}
+The same linear-programming methods show that any inference from subnetwork
+properties to network ones must necessarily start from some assumptions
+$\varIota$ that assign a probability distribution
+$\p(\yX = \yR \| \varIota)$ for the network states. The approaches to
+this task and reformulations of it have become uncountable: they include
+exchangeable models, parametric and non-parametric models, hierarchical
+models, general linear models, models via sufficiency, maximum-entropy
+models, and whatnot
+\citep[\eg:][]{jeffreys1931_r1973,jeffreys1939_r2003,jaynes1994_r2003,bernardoetal1994,gelmanetal1995_r2014,ghoshetal1997,kallenberg2005,gregory2005,sivia1996_r2006,ferreiraetal2007,dawid2013,damienetal2013}.
+We now show two examples, based on a maximum-entropy approach, that to our
+knowledge have not yet been explored in the neuroscientific literature. For
+a concrete application see \citep{rostamietal2016b}.
+
+\paragraph{First example: moment constraints for the network.}
+\label{sec:maxent_moments}
+Consider a state of knowledge $\yHa$ leading to the following properties:
+\begin{enumerate}%[$\yHa$1.]
+\item the expectations of the single and pair averages $\yxs$ and
+  $\sav{\yx\yx}$ of a particular subnetwork have given values
+  \begin{equation}
+    \label{eq:constr_ex1}
+    \expe{\yxs \| \yHa} = c_1, \qquad \expe{\sav{x_i x_j}\|\yHa} = c_2;
+  \end{equation}
+\item the network probability distribution $\p(\yX = \yR \|\yHa)$
+  has maximum relative entropy with respect to the uniform one, given the
+  constraints above.
+\end{enumerate}
+
+\medskip Then the probability distribution for the network conditional on $\yHa$
+is completely determined: it satisfies the symmetry
+property~\eqref{eq:homog_math} and is defined by
+\begin{multline}
+  \label{eq:pop_distr_maxent1}
+  \pf(\yX =\yR \|\yHa) =
+\varKappa\exp[\yL_2 N\yRf\,(N\yRf-1)/2 + \yL_1 N\yRf]
+\\
+\!\begin{aligned}[b]
+&\text{with $\varKappa, \yL_m$, such that the distribution is normalized and}
+\\
+&\varKappa\sum_{N A=0}^N %m!\, 
+\binom{N-m}{N A-m}
+\exp[\yL_2 N A\,(N A-1)/2 + \yL_1 N A]
+= c_m, \quad m=1,2.
+\end{aligned}
+\end{multline}
+We omit the full proof of this statement: it is a standard application of
+the maximum-entropy procedure
+\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013},
+combined with the equality~\eqref{eq:expe_products} of subnetwork and
+network expectations, \eg
+\begin{equation}
+  \label{eq:exploit_eq_expects}
+  c_2 = \expe{\sav{\yx\yx}\|\yHa} = 
+\binom{N}{2}^{-1}
+\sum_{N A=0}^N %m!\, 
+\binom{N A}{2} \, \p(\yXf = A\|\yHa),
+\end{equation}
+and with relations~\eqref{eq:joint_plaus_N_homog},
+\eqref{eq:conditional_total_state}. This example is easily generalized to
+any number $m$ of constraints such that  $m\le n$.
+
+Note again that, as remarked in \sect~\ref{sec:pop2sub_examples}, the
+subnetwork from which the averages in the
+expectations~\eqref{eq:constr_ex1} are calculated has a probability
+distribution $\p(\yxs =a \|\yH)$ determined by the
+marginalization~\eqref{eq:subpop_average} and does \emph{not} have a
+maximum-entropy form with the same number of constraints.
+%file: test_ME_same_under_scaling.nb
+
+\paragraph{Second example: subnetwork-distribution constraint.}
+\label{sec:maxent_fullsubpop}
+Consider another state of knowledge $\yHb$ leading to the following
+properties:
+\begin{enumerate}%[wide,label=$\yHb$\arabic*.]
+\item the average $\yxs$ of a particular
+  subnetwork has a probability distribution $q$:
+  \begin{equation}
+    \label{eq:constr_subpop_distr}
+    \p(\yxs =a \|\yHb) = q(a);
+  \end{equation}
+\item the probability distribution for the network, $\p(\yX = \yR
+  \|\yHb)$, has maximum relative entropy with respect to the uniform
+  one, given the constraint above.
+\end{enumerate}
+
+\medskip Then the probability distribution for the network given $\yHb$ is
+completely determined and satisfies the symmetry
+property~\eqref{eq:homog_math}:
+\begin{multline}
+  \label{eq:pop_distr_maxent2}
+  \p(\yX= \yR \|\yHb) =
+\exp\Biggl[
+\sum_{n a=0}^n \yL_{ a}
+%\binom{N}{N\yXf}^{-1}
+\binom{n}{n a}
+\binom{N-n}{N\yRf-n a}\Biggr]
+\\
+\!\begin{aligned}[b]
+&\text{with $\yL_{ a}$ such that}
+\\
+&\sum_{N A=0}^N 
+%\binom{N}{N A}^{-1}
+\binom{n}{n a}
+\binom{N-n}{N A-n a}\,
+\exp\Biggl[
+\sum_{n a=0}^n \yL_{ a}
+%\binom{N}{N A}^{-1}
+\binom{n}{n a}
+\binom{N-n}{N A-n a}\Biggr]
+=q( a)
+\end{aligned}
+\end{multline}
+(the normalization constraint being unnecessary since $q$ is normalized).
+This result is just another application of the maximum-entropy procedure
+with $n+1$ (linear) constraints given by \eqn~\eqref{eq:marginal}, where
+the left-hand side is now given and equal to $q(a)$.
+
+This example is equivalent to the generalization of the previous one with
+$n$ moment constraints, since knowledge of $\p(\yxs=a \| \yHb)$ is
+equivalent to knowledge its first $n$ moments.
+
+***
+
 In this note we would like to analyse and warn about a subtle assumption
-behind the \me\ method when it is applied to a network. It can informally
+behind the \me\ method when it is applied to a population. It can informally
 be put this way:
 \begin{quote}
-  \emph{the \me\ method assumes that the network it is applied to is
-    completely isolated from any larger network.}
+  \emph{the \me\ method assumes that the population it is applied to is
+    completely isolated from any larger population.}
 \end{quote}
@@
 
-When applied to a subnetwork, the \me\ method assumes that the uniform over
-the larger network is uniform, and therefore factorizable. The new
-distribution of the subnetwork will not be uniform, but that of the full
-network will still be factorizable into the one for the subnetwork and the
+When applied to a subpopulation, the \me\ method assumes that the uniform over
+the larger population is uniform, and therefore factorizable. The new
+distribution of the subpopulation will not be uniform, but that of the full
+population will still be factorizable into the one for the subpopulation and the
 rest.
@@
 The latter example has many similarities with that of a neuronal binary
-network. When we record the neuronal activity of a sample of neurons from a
+population. When we record the neuronal activity of a sample of neurons from a
 brain area, we assume that our measurements can tell us something -- no
@@
 the final distribution may have a multiplicity factor. When we consider the
-$2^{\yN}$ states of $\yN$ units we are not excluding that these might be
-marginals of $2^{\yM}$ states of $\yM$ units. Each one has the same
-multiplicity $2^{\yM-\yN}$, but this constant multiplicity factor disappears
-by normalization. So the method applies just as in the case of $\yN$ units
+$2^{N}$ states of $N$ units we are not excluding that these might be
+marginals of $2^{M}$ states of $M$ units. Each one has the same
+multiplicity $2^{M-N}$, but this constant multiplicity factor disappears
+by normalization. So the method applies just as in the case of $N$ units
 only, right?
@@
 Wrong. In our reasoning we have made subtle assumptions of independence
-between the full network and the subnetwork. The problem is that the
+between the full population and the subpopulation. The problem is that the
 counting of multiplicities is not based on simple enumeration, but already
 involves probability considerations. Consider three cases with a full
-population of two units, $\yM=2$, of which we consider one unit, $\yN=1$.
+population of two units, $M=2$, of which we consider one unit, $N=1$.
 \begin{itemize}
@@
 precede a \me\ application already involves probability considerations at
-the level of the full network. The usual reasoning by enumeration
+the level of the full population. The usual reasoning by enumeration
 implicitly assumes a uniform distribution or at least a \emph{factorizable}
@@
 ***If the distribution is factorizable, however, it means that examination of
-the subnetwork \emph{cannot give us any insights about the network it is a
+the subpopulation \emph{cannot give us any insights about the population it is a
   part of}. This is obviously contrary to the reason why we made neuronal observations.
@@
 \begin{enumerate}
-\item have a network with $\yN$ units, $2^{\yN}$ possible states
-\item expect averages of $\yxs$ active neurons and $\yxxs$ active pairs
+\item have a population with $N$ units, $2^{N}$ possible states
+\item expect averages of $\yx$ active neurons and $\yxxs$ active pairs
 \item use maximum-entropy to choose a probability distribution for the
-  states of the $\yN$ units conforming to our expectations.
+  states of the $N$ units conforming to our expectations.
 \end{enumerate}
@@
 \begin{enumerate}
-\item have a network with $\yM$ units, $2^{\yM}$ possible states
-\item expect that any subpopulation of $\yN$ units has $\yxs$ active
+\item have a population with $M$ units, $2^{M}$ possible states
+\item expect that any subpopulation of $N$ units has $\yxs$ active
   neurons and $\yxxs$ active pairs
 \item use maximum-entropy to choose a probability distribution for the
-  states of the $\yM$ units conforming to our expectations
+  states of the $M$ units conforming to our expectations
 \item marginalize to find the probability distribution for the states of
-  $\yN$ units.
+  $N$ units.
 \end{enumerate}

### 2017-05-14 maxNt.tex
Complete intro

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-13T17:24:25+0200
+%% Last-Updated: 2017-05-14T12:06:05+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\newcommand*{\pu}{\piup}%constant pi
-%\newcommand*{\delt}{\deltaup}%Kronecker, Dirac
+\newcommand*{\delt}{\deltaup}%Kronecker, Dirac
 %\newcommand*{\eps}{\varepsilonup}%Levi-Civita, Heaviside
@@
 \newcommand*{\defd}{\coloneqq}
-%\newcommand*{\defs}{\eqqcolon}
+\newcommand*{\defs}{\eqqcolon}
 %\newcommand*{\Land}{\bigwedge}
@@
 %\def\hy{-\penalty0\hskip0pt\relax}
-%\newcommand*{\labelbis}[1]{\tag*{(\ref{#1})$_\text{r}$}}
+\newcommand*{\labelbis}[1]{\tag*{(\ref{#1})$_\text{r}$}}
 %\newcommand*{\mathbox}[2][.8]{\parbox[t]{#1\columnwidth}{#2}}
@@
 \newtheoremstyle{simple}%
-{0\parsep}%
-{0\parsep}%
+{}%.5\baselineskip plusminus.2\baselineskip % \parsep
+{}%
 {\footnotesize}%
@@
 \newcommand*{\av}{\widebar} %pop average
-\newcommand*{\sav}{\widehat} %subpop average
+\newcommand*{\sav}{\widebar} %subpop average
 
@@
 \newcommand*{\yxs}{\sav{\yx}}%subpop av state
+\newcommand*{\yX}{\bm{X}}%pop state
+\newcommand*{\yy}{\bm{y}}%pop state
+\newcommand*{\yXf}{\av{\yX}}%pop av state
 \newcommand*{\yxxs}{\sav{\yx\yx}}%subpop av state
-\newcommand*{\yX}{\bm{X}}%pop state
-\newcommand*{\yXf}{\av{\yX}}%pop av state
+\newcommand*{\yXXf}{\av{\yX\yX}}%subpop av state
 \newcommand*{\yr}{\bm{r}}%subpop value
+\newcommand*{\ys}{\bm{s}}%subpop value
 \newcommand*{\yrs}{\sav{\yr}}%subpop av value
@@
 
+The present note has three nested purposes.
+
+The first and more general is to remind our readers that the \me\ method
+rests on subtle assumptions that may be contradictory with other
+assumptions we want to make on the data under study.
+
+As a concrete example we show -- and this is the second purpose -- that a
+sample of a neuronal population cannot be assigned a \enquote{maximally
+  noncommittal} \me\ distribution and at the same time be considered a
+\enquote{random representative} of the population.
+
+The proof of the example above rests on some probability relations from
+classical random sampling. These relations are somewhat hard to find
+explicitly stated in the neuroscientific literature; the third, minor
+purpose of this note is to state them explicitly.
+
 Let's clarify at once that our discussion pertains to a binary population at
@@
 clear) reasons. For our purposes let's say that the \me\ method is assumed
-to generate a \enquote{maximally noncommital} \cite{jaynes1963} probability
+to generate a \enquote{maximally noncommittal} \cite{jaynes1963} probability
 distribution. We intend the quoted expression only as an umbrella term,
@@
 The neurons recorded in these kinds of experiments are usually picked out
-in a way beyond our control, and from their observation we expect to learn
-something about all other neurons in the same brain area. That is, they are
-assumed to be a \enquote{representative random sample} of the neurons in
-that area. Our discussion hinges on this often just tacitly understood
-assumption.
+according to an unknown process, and from their observation we expect to
+learn something about all other neurons in the same brain area. That is,
+they are assumed to be a \enquote{representative random sample} of the
+neurons in that area. Our discussion hinges on this often just tacitly
+understood assumption.
 
@@
 
-Consider a population of $N$ binary neurons with states $(X_1,\dotsc,X_N)$
-having fixed but unknown binary values $(R_1,\dotsc,R_N)$, with
-$R_i \text{ in } \set{0,1}$, vectorially written $\yX=\yR$. For example,
-$\yX$ can represent the state of a population at a particular time. We call
-the neurons \enquote{units} to lend some generality to our discussion. We
-shall make statements about the whole population of $N$ units and about a
-subpopulation of $n$ units; the word \enquote{population} will always refer to
-the \emph{whole} population. The subpopulation states and their values are
-denoted by lowercase letters: $(x_1, \dotsc, x_n)\equiv\yx$ and
+We have a population of $N$ binary neurons. Let's assume that the
+identities of the neurons can be distinguished, by their spike shapes for
+example; but other details, like their locations, are unknown. The neurons
+have a joint state $(X_1,\dotsc,X_N) \defs \yX$ having fixed but unknown
+binary values $(R_1,\dotsc,R_N)\defs \yR \in \set{0,1}^N$. A particular
+sample of $n$ neurons from this population has joint state
+$(x_1, \dotsc, x_n) \defs \yx$ having fixed binary values
+$(r_1, \dotsc, r_n)\defs\yr$. We will consider various averages for the
+population and the sample. For this purpose we introduce a general
+averaging operator $\sav{\dotv}$ defined by
+\begin{align}
+  \av{\yX}  &\defd \tfrac{1}{N} (X_1 + X_2 + \dotsb + X_N),
+  &
+  \sav{\yx}  &\defd \tfrac{1}{n} (x_1 + x_2 + \dotsb + x_n),
+    \\[2\jot]
+  \av{\yX \yX} &\defd
+\tbinom{N}{2}^{-1} (X_1 X_2 + X_1 X_3  + \dotsb +  X_{N-1} X_N),
+  &
+    \sav{\yx \yx} &\defd
+\tbinom{n}{2}^{-1} (x_1 x_2 + x_1 x_3  + \dotsb +  x_{n-1} x_n),
+\\[2\jot]
+\av{\yX\yX\yX} &\defd
+                  \tbinom{N}{3}^{-1}
+                  (X_1 X_2 X_3 +% X_1 X_2 X_4
+                   \dotsb + X_{N-2} X_{N-1}X_N),
+  &
+    \sav{\yx\yx\yx} &\defd
+                      \tbinom{n}{3}^{-1} (x_1 x_2 x_3 + %x_1 x_2 x_4
+                       \dotsb + x_{n-2} x_{n-1} x_n),
+\end{align}
+and so on; analogously for the quantities $\yR$ and $\yr$. These formulae
+say that $\yXf$ is the fraction of active neurons, $\yXXf$ the fraction of
+simultaneously active pairs out of all $\binom{N}{2}$ pairs, $\av{\yX\yX\yX}$
+the fraction of simultaneously active triplets, and so on.
+
+\begin{simplenote}
+  A sample can be \enquote{random} in two different ways. Consider the
+  population of neurons named $\set{1,2,3}$ for example. The sample
+  $\set{1,3}$ is random because neurons $1$ and $3$ were chosen according
+  to an unknown process. The sample $\set{\textrm{X}_1, \textrm{X}_2}$ is
+  random because we don't know the identity of neurons $\textrm{X}_1$ and
+  $\textrm{X}_2$. Given the probability for the population, the
+  probabilities for the states of these two samples are different: the
+  former is obtained by marginalization, the latter by symmetrization and
+  marginalization.
+\end{simplenote}
+
+Our uncertainty about the state of the population is completely expressed by the
+probability distribution
+\begin{equation}
+  \label{eq:joint_plaus}
+  \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \| \yH) \quad\text{or}\quad
+\p(\yX =\yR \| \yH), \quad \yR \in \set{0,1}^N,
+\end{equation}
+where $\yH$ denotes our state of knowledge (the evidence and assumptions
+backing our probability assignments). Our uncertainty about the state of
+the sample is likewise expressed by
+\begin{equation}
+  \label{eq:sample_plaus}
+  \p(x_1=r_1, x_2=r_2, \dotsc, x_n=r_n \| \yH) \quad\text{or}\quad
+\p(\yx =\yr \| \yH), \quad \yr \in \set{0,1}^n.
+\end{equation}
+
+\subsection{Initial assumptions}
+\label{sec:initial_assumptions}
+
+Before any experimental observations are made we need to make an initial
+probability assignment, no matter what kinds of predictions we are
+interested in. This initial assignment will be modified by the experimental
+observations.
+
+We yet know very little about the physical details of the individual
+neurons, like their locations. Our state of knowledge is therefore
+symmetric, or \enquote{exchangeable}, under their permutations. This
+symmetry must be reflected in our initial probability, and the
+\emph{representation theorem for finite exchangeability} states that it
+must obey
+\begin{equation}
+  \label{eq:joint_plaus_N_homog}
+  \p(\yX = \yR \| \yH) = \binom{N}{N\yRf}^{-1} \p(\yXf=\yRf \| \yH),
+\end{equation}
+the latter being the probability for the population average $\yX$. Proof
+and generalizations to non-binary and continuum cases are given by
+de~Finetti \cite{definetti1959b}, Ericson \cite{ericson1976}, Diaconis
+\cite{diaconis1977}, Heath \amp\ Sudderth \cite{heathetal1976}. This
+theorem is intuitive: owing to symmetry, all states with $N\yRf$ active
+neurons must have equal probabilities.
+
+By marginalization we obtain the probability for the state of the sample:
+\begin{gather}
+  \label{eq:marginal}
+  \p(\yx = \yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs \| \yH),
+  \\
+  \shortintertext{with}
+  \label{eq:subpop_average}
+  \p(\yxs=\yrs \| \yH) = \binom{n}{n\yrs}
+  \sum_{N \yRf =0}^{N} \binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}\,
+  \p(\yXf=\yRf \| \yH).
+\end{gather}
+The combinatorial proof is the same as for the theory of drawing without
+replacement, leading to the hypergeometric distribution, and can be found
+in many texts
+\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
+The symmetry of our initial ignorance should intuitively also apply to the
+sample. This is indeed the case: the probability for the state of the
+sample~\eqref{eq:subpop_average} automatically satisfies the representation
+theorem~\eqref{eq:joint_plaus_N_homog} as well.
+
+Formula~\eqref{eq:subpop_average}, connecting our initial probabilities for
+the population and sample averages, has a very important consequence: \emph{our
+expectations for all averages of the sample and the full population are the
+same}:
+\begin{multline}
+  \label{eq:expe_products}
+\expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
+=
+\expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
+\\[2\jot]
+{}=\sum_{N \yRf=0}^N %m!\, 
+\binom{N-m}{N \yRf-m}
+\binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
+\equiv
+\binom{N}{m}^{-1}
+%\frac{(N-m)!}{N!}
+\sum_{N \yRf=0}^N %m!\, 
+\binom{N \yRf}{m}\, \p(\yXf=\yRf \| \yH).
+\end{multline}
+
+\iffalse
+\begin{align}
+  \expe{x_{i_1} \dotsm x_{i_m} \| \yH}
+&=
+\expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
+=
+\expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH},
+\\
+\intertext{and has an explicit expression in terms of $Q$:}
+  \begin{split}
+\expe{x_{i_1} \dotsm x_{i_m} \| \yH}
+&=
+\binom{N}{m}^{-1}
+%\frac{(N-m)!}{N!}
+\sum_{N A=0}^N %m!\, 
+\binom{N A}{m}\, Q(A)
+\\
+&\equiv
+%\frac{(N-m)!}{N!}
+\sum_{N A=0}^N %m!\, 
+\binom{N-m}{N A-m}
+\binom{N}{N A}^{-1} Q(A).
+\end{split}
+\end{align}
+\end{subequations}
+\fi
+***
+
+What do we mean by saying that we can learn something about the population
+by observing the sample? At the very least we mean that our uncertainty
+about the rest of the population can change upon observing the sample. That
+is, the probability for the state $\yy$ of the rest of the population is
+conditionally dependent on the sample's values:
+\begin{equation}
+  \label{eq:conditionally_dependent_sample}
+  \p(\yy =\ys \| \yx=\yr,\yH) \ne \p(\yy =\ys \|\yH)
+  \qquad\text{for some $\ys,\yr$},
+\end{equation}
+This also means that the probability for the population cannot be
+factorized into that of the sample times that of the complement.
+
+What do we mean when we say that the sample is representative of the
+population? It means that we expect some collective properties of the
+sample, like the fraction of active neurons, or the fraction of
+simultaneously active pairs, to be roughly equal to those of the full
+population.
+
+***
+
+For example, $\yX$ can
+represent the state of a population at a particular time. We call the
+neurons \enquote{units} to lend some generality to our discussion. We shall
+make statements about the whole population of $N$ units and about a
+subpopulation of $n$ units; the word \enquote{population} will always refer
+to the \emph{whole} population. The subpopulation states and their values
+are denoted by lowercase letters: $(x_1, \dotsc, x_n)\equiv\yx$ and
 $(r_1, \dotsc, r_n)\equiv\yr$; but note that $x_i \equiv X_{j_i}$ and
 $r_i \equiv R_{j_i}$ for some distinct $j_1,\dotsc,j_n$. We shall also make
-statements about the population-averaged state, or \emph{population average}:
+statements about the population-averaged state, or \emph{population
+  average}:
 \begin{align}
@@
 
+  \section{Sketched proofs}
+  \label{sec:derivations}
+
+  Variants of the following derivations and combinatorial considerations
+  can be found \eg\ in
+  \cites[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003};
+  see also \citep{whitworth1897}.
+
+  To derive the joint probability distribution
+  \eqref{eq:joint_plaus_N_homog} from that for the network
+  average~\eqref{eq:pop_average_Q}, consider that if the network total
+  is $N\yXf$, then $N\yXf$ out of $N$ units are active, and there are
+  $\tbinom{N}{N\yXf}$ possible states for which this can be true; therefore
+  \begin{equation}
+    \labelbis{eq:joint_plaus_N_homog}
+    \p(\yX = \yR \| H) = \binom{N}{N\yRf}^{-1}Q(\yRf).
+  \end{equation}
+An analogous reasoning for $n$ and $\yxs$ leads to an analogous equality,
+\begin{equation}
+  \label{eq:aver_subpop}
+  \p(\yx=\yr \| H) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs\|\yH),
+\end{equation}
+for the subnetwork.
+
+\bigskip
+
+Let us next consider the probability $\p(\yxs =a \| \yXf = A,\yH)$ for
+the subnetwork average $\yxs$ conditional on the network average $\yXf$.
+There are $\tbinom{N}{N\yXf}$ possible network states if the network
+average is $\yXf$, \ie\ if $N\yXf$ units are in state $1$; the conditional
+probability of each is therefore $1/\tbinom{N}{N\yXf}$, owing to the
+symmetry assumption~\eqref{eq:homog_math}. Now consider the subnetwork of the
+first $n$ units. The conditional probability of having $n\yxs$ specific
+ones in state $1$ is the sum of the probabilities of all states for
+which $N\yXf-n\yxs$ of the remaining $N-n$ units are in state $1$; there are
+$\tbinom{N-n}{N\yXf-n\yxs}$ such states, all equally probable. Finally,
+there are $\tbinom{n}{n\yxs}$ possible ways, all equally probable, in which
+$n\yxs$ of the first $n$ units can be in state $1$. In formulae,
+\begin{equation}
+\label{eq:conditional_prob_explained}
+\begin{split}
+      \p(\yxs =a \| \yXf=A, \yH)
+      &=
+      \sum_{\yr}^{\yrs=a} \sum_{\yR}^{\yRf=A} 
+      \p(\yx =\yr \|  \yX = \yR, \yH) \,
+      \p(\yX =\yR  \| \yXf=A,\yH),
+      \\
+      &=
+      \sum_{\yr}^{\yrs=a} \sum_{\yR}^{\yRf=A} 
+      \p(\yx=\yr \|  \yX =\yR, \yH) \,
+      \binom{N}{N A}^{-1},
+      \\
+      &=\sum_{\yr}^{\yrs=a} 
+      \binom{N-n}{N A-n a}\binom{N}{N A}^{-1},
+      \\
+      &=
+      \binom{n}{n a}\binom{N-n}{N A-n a}\binom{N}{N A}^{-1},
+\end{split}
+\end{equation}
+  which is the conditional probability~\eqref{eq:conditional_prob}. Note
+  that this is just a derivation of the hypergeometric distribution
+  \cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
+  which describes the probability of, say, drawing a proportion of $\yxs$
+  blue balls in $n$ drawings without replacement from an urn with $N$
+  balls, a fraction $\yXf$ of which are blue.
+
+  \bigskip The probability of a subnetwork average $\yxs$ is then, by
+  marginalization,
+  \begin{equation}
+    \label{eq:subpop_average_explained}
+    \begin{split}
+      \p(\yxs =a \| \yH)
+      &= \sum_{N A=0}^N\p(\yxs = a \| \yXf = A, \yH)\,
+      \p(\yXf = A \| \yH),
+      \\
+      &=\sum_{N A=0}^N
+      \binom{n}{n a}\binom{N-n}{N A-n a}\binom{N}{N A}^{-1}
+      Q(A).
+    \end{split}
+  \end{equation}
+  which proves the subnetwork-average formula~\eqref{eq:subpop_average}.
+  This formula, combined with \eqns~\eqref{eq:aver_subpop} and
+  \eqref{eq:joint_plaus_N_homog}, leads to the conditional
+  probability~\eqref{eq:conditional_total_state}.
+
+  \bigskip The independence of the expectation of products of states from
+  the subnetwork size is trivial by marginalization:
+  \begin{multline}
+    \label{eq:margin_products}
+    \sum_{\yx} x_1\dotsm x_m \p(\yx =\yr \| \yH)
+      =\sum_{\yr,\yR}  r_1\dotsm r_m \p(\yx=\yr \| \yX=\yR,\yH)
+      \, \p(\yX=\yR \| \yH),
+      \\
+      \begin{aligned}
+      &=\sum_{\yr,\yR} r_1\dotsm r_m\,\delt(R_1-r_1)\dotsm\delt(R_m-R_m)
+      \p(\yX = \yR \| \yH),
+      \\
+      &=\sum_{\yR} R_1\dotsm R_m \p(\yX =\yR \| \yH).
+    \end{aligned}
+  \end{multline}
+  All such $m$-fold products have the same expectation by symmetry,
+  therefore their subnetwork average will do, too, being an average of
+  equal terms.
+
+  Now consider the sum of all distinct products of states of two units in
+  the subnetwork:
+  \begin{equation*}
+    x_1 x_2 + x_1 x_3 + \dotsb + x_{n-1} x_n.
+  \end{equation*}
+  The terms in this sum are either $0$ or $1$. The non-vanishing ones are
+  those with index pairs chosen from the $n\yxs$ units of the subnetwork
+  which are in state $1$, and there are $\tbinom{n\yxs}{2}$ such choices,
+  so the sum above is equal to $\tbinom{n\yxs}{2}$. The sum has
+  $\tbinom{n}{2}$ terms, so their average is
+  $\tbinom{n\yxs}{2}/\tbinom{n}{2}$. Generalizing the argument to products
+  of $m$ units, we have that
+  \begin{equation}
+    \sav{x_{i_1} \dotsm x_{i_m}} = \binom{n\yxs}{m} \binom{n}{m}^{-1}.
+  \end{equation}
+  Then, using \eqn~\eqref{eq:subpop_average},
+  \begin{multline}
+    \label{eq:expect_product_m}
+    \expe*{\sav{x_{i_1} \dotsm x_{i_m}} \| \yH}
+    \\
+    \begin{aligned}[b]
+      &=\tfrac{(n-m)!}{n!}
+      \expe*{m!\,\tbinom{n\yxs}{m}  \| \yH}
+      \\
+      &\equiv\frac{(n-m)!}{n!}
+      \sum_{n a=0}^n m!\, \binom{n a}{m} \p(\yxs =a \|\yH),
+      \\
+      &=\frac{(n-m)!}{n!} \sum_{N A=0}^N
+      Q( A)
+      \Biggl[ 
+      \sum_{n a=0}^n 
+      m!\,\binom{n a}{m}\, \binom{n}{n a}\binom{N-n}{N A-n a}\binom{N}{N A}^{-1} 
+      \Biggr].
+    \end{aligned}
+  \end{multline}
+  The expression in brackets is the $m$th factorial moment of the
+  hypergeometric function, and is given by \citep{potts1953}
+%file: factorial_moments_etc.nb
+  \begin{equation}
+    \sum_{n a=0}^n m!\,\binom{n a}{m} \binom{n}{n a}\binom{N-n}{N\yXf-n a}
+    \binom{N}{N\yXf}^{-1} 
+    =
+    m!\,\binom{n}{m} \binom{n a}{m}\binom{N}{m},
+  \end{equation}
+  which combined with the previous equation yields the second line of
+  \eqn~\eqref{eq:expe_products}; its last equality comes from the identity
+  \begin{equation}
+    \label{eq:binom_identity}
+    \binom{N}{M}\binom{M}{m}=\binom{N}{m}\binom{N-m}{M-m},
+  \end{equation}
+  easily derived by writing the binomial coefficients in terms of
+  factorials. Finally, \eqns~\eqref{eq:moments}, relating the moments of
+  the distributions for subnetwork and network averages, is obtained
+  from the definition of moments,
+  \begin{equation}
+    \label{eq:moments_def}
+    \expe*{\yxs^m \| \yH} \defd
+    \sum_{n a=0}^{n} a^m \p(\yxs = a  \| H),
+    \quad
+    \expe{\yXf^m \| \yH} \defd
+    \sum_{N A=0}^{N} A^m \p(\yXf = A  \| H),
+  \end{equation}
+  replaced in the equalities for the factorial
+  moments~\eqref{eq:eq_factmom_prod}, by recursively solving in terms of
+  the moments of the network distribution.
+
 \subsection{***}

### 2017-05-14 maxNt.tex
add discrepancy section

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-14T12:06:05+0200
+%% Last-Updated: 2017-05-15T00:15:08+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %conditional assumptions
-\newcommand*{\yH}{\varEta}
+\newcommand*{\yH}{\varIota}
 \newcommand*{\yD}{\varDelta}
@@
 \end{equation}
-where $\yH$ denotes our state of knowledge (the evidence and assumptions
+where $\yH$ denotes our initial state of knowledge (the evidence and assumptions
 backing our probability assignments). Our uncertainty about the state of
@@
   \label{eq:subpop_average}
-  \p(\yxs=\yrs \| \yH) = \binom{n}{n\yrs}
-  \sum_{N \yRf =0}^{N} \binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}\,
-  \p(\yXf=\yRf \| \yH).
+  \p(\yxs=\yrs \| \yH) = \sum_{N \yRf =0}^{N}
+    \p(\yxs = \yrs \|\yXf=\yRf, \yH)\,
+  \p(\yXf=\yRf \| \yH),
+  \\[2\jot]
+  \p(\yxs = \yrs \|\yXf=\yRf, \yH)=
+  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}.
 \end{gather}
-The combinatorial proof is the same as for the theory of drawing without
-replacement, leading to the hypergeometric distribution, and can be found
-in many texts
+The latter conditional probability is a hypergeometric distribution,
+typical of \enquote{drawing without replacement} problems. The
+combinatorial proof of the formulae above is in fact the same as for this
+class of problems
 \cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-The symmetry of our initial ignorance should intuitively also apply to the
+
+The symmetry present in the probabilities above is not a physical property
+of the neuronal population. It only expresses the symmetry of our initial
+ignorance about the population, and does not imply any sort of physical
+similarity between the neurons. Subsequent observations may in fact break
+this symmetry. The initial symmetry should intuitively also apply to the
 sample. This is indeed the case: the probability for the state of the
@@
 
-Formula~\eqref{eq:subpop_average}, connecting our initial probabilities for
-the population and sample averages, has a very important consequence: \emph{our
-expectations for all averages of the sample and the full population are the
-same}:
-\begin{multline}
+Our initial knowledge (or ignorance) $\yH$ has two very important
+properties derivable from the formulae above. First, \emph{our initial
+  expectations for all averages of the sample and the full population are
+  the same}:
+\begin{equation}
   \label{eq:expe_products}
@@
 \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
-\\[2\jot]
-{}=\sum_{N \yRf=0}^N %m!\, 
-\binom{N-m}{N \yRf-m}
-\binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
+%\\[2\jot]
+%{}
+=
+\binom{n}{m}^{-1}
+%\frac{(n-m)!}{n!}
+\sum_{n \yrs=0}^n %m!\, 
+\binom{n \yrs}{m}\, \p(\yxs=\yrs \| \yH)
+% \sum_{N \yRf=0}^N %m!\, 
+% \binom{N-m}{N \yRf-m}
+% \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
 \equiv
@@
 \binom{N \yRf}{m}\, \p(\yXf=\yRf \| \yH).
-\end{multline}
-
+\end{equation}
+Second, upon observation of a sample average, say $\yxxs=a$, the updated
+expectations for such average in the population \emph{and in any new
+  sample} will usually be shifted towards the observed value, as follows
+from Bayes's theorem and the formulae above.
+
+These two properties express the fact that the sample is
+\emph{representative} of the population. Our initial knowledge $\yH$ thus
+assumes that we can make predictions about the population by observing a
+sample. Contrast this with an assumption that assigns an independent
+probability to each neuron: in this case, the observation of one sample
+does not change our initial probability about another; we can't learn from
+observation.
+
+\subsection{Specifying the initial probability}
+\label{sec:specific_initial_probability}
+
+Formulae \eqref{eq:joint_plaus_N_homog}--\eqref{eq:expe_products} are
+constraints on our initial probability assignment, but do not determine it
+numerically yet. The probability $\p(\yXf=\yRf \| \yH)$ for the population
+average needs to be numerically specified, and by \eqref{eq:subpop_average}
+it will determine that of the sample average, $\p(\yxs=\yrs \| \yH)$. If we
+numerically specify the latter, the former is not completely specified,
+because \eqn~\eqref{eq:subpop_average} linearly constrains $N+1$ unknowns
+by only $n+1$ equations in this case.
+
+We may want to specify the probability by enforcing several of its sample
+expectations to have specific values, for example $\expe{\yxs}=a$,
+$\expe{\yxxs}=b$. But this is still and underdetermined linear problem
+because several distributions can have the same desired expectations, as
+clear from \eqns~\eqref{eq:expe_products}.
+
+It is at this point that the \me\ method enters the scene. It selects one
+distribution, purported to be \enquote{maximally noncommittal}, among those
+that have the desired expectations. But we are in a quandary:
+formulae~\eqref{eq:expe_products} allow us to apply the method to find the
+probability of the population $\p(\yXf=\yRf \| \yH)$, or of the sample
+$\p(\yxs=\yrs \| \yH)$; the two applications, however, are
+\emph{inequivalent}. They lead to numerically different
+$\p(\yxs=\yrs \| \yH)$; see \fig~\ref{fig:diff_maxent_pop_sample}. Which
+should we choose?
+
+\begin{figure}[!b]
+\centering
+\includegraphics[width=0.95\columnwidth]{different_maxent_pop_sample.pdf}%
+\caption{Example of discrepancy in the probability $\p(\yxs = \yrs \| \yH)$
+  constructed by \me\ at the population level followed by sample
+  marginalization (blue triangles), and at the sample level (red circles).
+  The constraints are the same in both applications:
+  $\expe{\yxs \|\yH}=0.45$, $\expe{\yxxs \|\yH}=0.35$, and $N=5\,000$,
+  $n=200$.}
+\label{fig:diff_maxent_pop_sample}
+\end{figure}%maxent_pop_or_sample.nb
 \iffalse

### 2017-05-14 maxNt.tex
add discrepancy section

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T00:15:08+0200
+%% Last-Updated: 2017-05-15T01:20:54+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \emph{inequivalent}. They lead to numerically different
-$\p(\yxs=\yrs \| \yH)$; see \fig~\ref{fig:diff_maxent_pop_sample}. Which
-should we choose?
-
-\begin{figure}[!b]
+$\p(\yxs=\yrs \| \yH)$; see \fig~\ref{fig:diff_maxent_pop_sample} for an
+example. Suppose our constraints are $\expe{\yxs \|\yH}$ and
+$\expe{\yxxs \|\yH}$. Application of \me\ at the population level leads to
+the distribution
+\begin{equation}
+  \label{eq:full_pop_av_example_ME}
+  \p(\yXf = \yRf \| \yH)  \propto
+\binom{N}{N \yRf}\exp\biggl[
+\yL_2\, \frac{\yRf\,(N \yRf-1)}{N-1} + \yL_1\, \yRf\biggr]
+\end{equation}
+with such parameters $\yL_1$, $\yL_2$ as to satisfy the constraints.
+Marginalization by \eqn~\eqref{eq:marginal} leads to 
+\begin{equation}
+  \label{eq:full_pop_av_example_ME_marg}
+  \p(\yxs = \yrs \| \yH)  \propto
+  \sum_{N\yRf=0}^N
+  \binom{n}{n\yrs} \binom{N-n}{N\yRf-n\yrs} \exp\biggl[
+\yL_2\, \frac{\yRf\,(N \yRf-1)}{N-1} + \yL_1\, \yRf\biggr].
+\end{equation}
+Application of \me\ at the sample level leads to the distribution
+\begin{equation}
+  \label{eq:sample_av_example_ME}
+  \p(\yxs = \yrs \| \yH)  \propto
+  \binom{n}{n\yrs}
+  \exp\biggl[\yl_2\, \frac{\yrs\,(n \yrs-1)}{n-1} + \yl_1\, \yrs\biggr]
+\end{equation}
+with such parameters $\yl_1$, $\yl_2$ as to satisfy the constraints.
+Expressions \eqref{eq:full_pop_av_example_ME_marg} and
+\eqref{eq:sample_av_example_ME} cannot be equal when $n>6$ because, even if
+we could choose their four parameters and two normalization constants
+freely (and we can't), for their equality we should still satisfy $n+1$ equations,
+\ie\ more than six.
+
+Which should we choose?
+
+\begin{figure}[!t]
 \centering
-\includegraphics[width=0.95\columnwidth]{different_maxent_pop_sample.pdf}%
-\caption{Example of discrepancy in the probability $\p(\yxs = \yrs \| \yH)$
-  constructed by \me\ at the population level followed by sample
-  marginalization (blue triangles), and at the sample level (red circles).
-  The constraints are the same in both applications:
-  $\expe{\yxs \|\yH}=0.45$, $\expe{\yxxs \|\yH}=0.35$, and $N=5\,000$,
-  $n=200$.}
+\includegraphics[width=0.95\columnwidth]{different_maxent_pop_sample_100.pdf}%
+\caption{Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
+  \me\ at the population level followed by sample marginalization (blue
+  triangles), and at the sample level (red circles). The constraints are
+  the same in either case: $\expe{\yxs \|\yH}=0.45$,
+  $\expe{\yxxs \|\yH}=0.35$, and $N=5\,000$, $n=100$.}
 \label{fig:diff_maxent_pop_sample}

### 2017-05-15 maxNt.tex
add discrepancy section
start discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T01:20:54+0200
+%% Last-Updated: 2017-05-15T11:15:45+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\yD}{\varDelta}
-\newcommand*{\yHa}{\varEta'}
-\newcommand*{\yHb}{\varEta''}
+\newcommand*{\yHa}{\varIota'}
+\newcommand*{\yHb}{\varIota''}
 %Lagr multipliers
@@
 \newcommand*{\yk}{\kappa}
+\newcommand*{\yK}{\varKappa}
 %entropies
@@
 \author{ V. Rostami\\
-  Forschungszentrum Jülich INM-6,\\Hitlerland\\
+  Forschungszentrum Jülich INM-6\\Hitlerland\\
   \texttt{v.rostami@fz-juelich.de}
   \And
-  P.G.L. Porta Mana\\who knows
-  \And E. Torre\\Chocolateland  
+  P.G.L. Porta Mana\\Mussoliniland
+  \And E. Torre\\Moneylaunderingland  
   % David S.~Hippocampus\thanks{Use footnote for providing further
@@
 
+\mynote{Version 1}
 The present note has three nested purposes.
@@
 
-In this note we would like to show that there is a contradiction in
-applying \me\ to a \enquote{representative random sample} of neurons to
-generate a \enquote{maximally noncommital} distribution for that sample.
+\mynote{Version 2 }
+In this note we would like to show that there is a
+contradiction in applying \me\ to a \enquote{representative random sample}
+of neurons to generate a \enquote{maximally noncommital} distribution for
+that sample.
 % \\
@@
 
+\mynote{Add formula $\av{\underbrace{\yX\dotsm\yX}_{\text{$m$ factors}}} = \binom{N\yXf}{m}/\binom{N}{m}$.}
+
 \begin{simplenote}
-  A sample can be \enquote{random} in two different ways. Consider the
-  population of neurons named $\set{1,2,3}$ for example. The sample
-  $\set{1,3}$ is random because neurons $1$ and $3$ were chosen according
-  to an unknown process. The sample $\set{\textrm{X}_1, \textrm{X}_2}$ is
-  random because we don't know the identity of neurons $\textrm{X}_1$ and
-  $\textrm{X}_2$. Given the probability for the population, the
-  probabilities for the states of these two samples are different: the
-  former is obtained by marginalization, the latter by symmetrization and
-  marginalization.
+  \mynote{Remove this} A sample can be \enquote{random} in two different
+  ways. Consider the population of neurons named $\set{1,2,3}$ for example.
+  The sample $\set{1,3}$ is random because neurons $1$ and $3$ were chosen
+  according to an unknown process. The sample
+  $\set{\textrm{X}_1, \textrm{X}_2}$ is random because we don't know the
+  identity of neurons $\textrm{X}_1$ and $\textrm{X}_2$. Given the
+  probability for the population, the probabilities for the states of these
+  two samples are different: the former is obtained by marginalization, the
+  latter by symmetrization and marginalization.
 \end{simplenote}
@@
 
+\mynote{Add paragraph on moments}
+
 These two properties express the fact that the sample is
@@
 probability of the population $\p(\yXf=\yRf \| \yH)$, or of the sample
-$\p(\yxs=\yrs \| \yH)$; the two applications, however, are
-\emph{inequivalent}. They lead to numerically different
-$\p(\yxs=\yrs \| \yH)$; see \fig~\ref{fig:diff_maxent_pop_sample} for an
-example. Suppose our constraints are $\expe{\yxs \|\yH}$ and
-$\expe{\yxxs \|\yH}$. Application of \me\ at the population level leads to
-the distribution
+$\p(\yxs=\yrs \| \yH)$; \emph{the two applications, however, are
+  inequivalent}. They lead to numerically different $\p(\yxs=\yrs \| \yH)$.
+Here is an example.
+
+Suppose our desired constraints are $\expe{\yxs } = c_1$ and
+$\expe{\yxxs} = c_2$, the expectations being given by the fourth term of
+\eqn~\eqref{eq:expe_products}. Applying \me\ to the distribution of the
+average of the full population we find
 \begin{equation}
   \label{eq:full_pop_av_example_ME}
-  \p(\yXf = \yRf \| \yH)  \propto
+  \p(\yXf = \yRf \| \yHa)  = \yL\,
 \binom{N}{N \yRf}\exp\biggl[
@@
 \end{equation}
-with such parameters $\yL_1$, $\yL_2$ as to satisfy the constraints.
-Marginalization by \eqn~\eqref{eq:marginal} leads to 
+with such parameters $\yL$, $\yL_1$, $\yL_2$ as to satisfy normalization
+and constraints. We call this state of knowledge $\yHa$. Marginalization by
+\eqn~\eqref{eq:marginal} then gives
 \begin{equation}
   \label{eq:full_pop_av_example_ME_marg}
-  \p(\yxs = \yrs \| \yH)  \propto
+  \p(\yxs = \yrs \| \yHa)  = \yL
   \sum_{N\yRf=0}^N
@@
 \end{equation}
-Application of \me\ at the sample level leads to the distribution
+Applying \me\ at the sample level, using the third term of
+\eqn~\eqref{eq:expe_products}, we find
 \begin{equation}
   \label{eq:sample_av_example_ME}
-  \p(\yxs = \yrs \| \yH)  \propto
+  \p(\yxs = \yrs \| \yHb)  = \yl\,
   \binom{n}{n\yrs}
@@
 \end{equation}
-with such parameters $\yl_1$, $\yl_2$ as to satisfy the constraints.
-Expressions \eqref{eq:full_pop_av_example_ME_marg} and
-\eqref{eq:sample_av_example_ME} cannot be equal when $n>6$ because, even if
-we could choose their four parameters and two normalization constants
-freely (and we can't), for their equality we should still satisfy $n+1$ equations,
-\ie\ more than six.
-
-Which should we choose?
-
+with such parameters $\yl$, $\yl_1$, $\yl_2$ as to satisfy the constraints.
+We call this state of knowledge $\yHb$. For example, with $N=5\,000$,
+$n=100$, $c_1=0.45$, $c_2=0.35$, we find $\yL_1 = -13\,321.9$,
+$\yL_2 = 13\,321.5$, $\yl_1=-269.3$, $\yl_2 = 268.9$, and appropriate
+normalization constants. The resulting distributions
+$\p(\yxs = \yrs \| \yHa)$, $\p(\yxs = \yrs \| \yHb)$ are shown in
+\fig~\ref{fig:diff_maxent_pop_sample}; they are clearly different.
 \begin{figure}[!t]
@@
 \end{figure}%maxent_pop_or_sample.nb
+
+The inequivalence of the two \me\ applications is generally valid for small
+number of constraints, even if the two states of knowledge $\yHa$, $\yHb$
+use different numbers $m'$, $m''$ \emph{and values} of constraints. The
+equality of the third and fourth term of \eqn~\eqref{eq:expe_products} in
+this case implies $n+1$ equations in $m'+m''+1$ unknowns (one normalization
+is taken care of), and in general can only be satisfied if $n\le m'+m''$.
+
+\medskip
+
+Which application of the \me\ method should we choose, then?
+
+\section{Discussion}
+\label{sec:discussion}
+
+{\centering\mynote{*** From here down old text to be discarded ***}\par}
+
 \iffalse
@@
 \fi
-***
 

### 2017-05-15 maxNt.tex
start discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T11:15:45+0200
+%% Last-Updated: 2017-05-15T11:42:02+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \renewcommand{\bibfont}{\footnotesize}
-\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
-}
+%\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
+%}
 \newcommand*{\citep}{\parencites}
@@
 
-\section{Probability of random representative samples}
-\label{sec:prob_samples}
-
-\subsection{Setup}
+\section{Setup}
 \label{sec:setup}
@@
 
-\subsection{Initial assumptions}
-\label{sec:initial_assumptions}
+\section{Initial assumptions: the probability of random representative samples}
+\label{sec:prob_samples}
 
@@
 
-\subsection{Specifying the initial probability}
+\section{Enter maximum-entropy: quandary}
 \label{sec:specific_initial_probability}
@@
 
+\mynote{We don't give an answer. Different purposes. Bright side: no real
+  contradiction because we are expressing two states of knowledge, not
+  physical situations. Also, in some case numerically very similar results.
+  Same phenomenon in statistical mechanics \cite{maesetal1999}. Somehow
+  similar quandary for equilibrium state at the end of non-equilibrium
+  process: Gibbs state vs initial Gibbs state + Liouville
+  evolution \cite[\sect~4]{jaynes1985d_r1993}; but in this case both
+  distributions are immensely sharp, and this doesn't affect predictions}
+
+\bigskip
 {\centering\mynote{*** From here down old text to be discarded ***}\par}
+\section{***}
 
@@
 
-\subsection{***}
-
-\subsubsection{***}
-
-\paragraph{***}
-
 % do not change font sizes (except perhaps in the \textbf{References}
@@
 
-\section*{References}
+%\section*{References}
+\renewcommand*{\bibname}{References}
+\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
+}
 
@@
 %   Neuroscience} {\bf 15}(7):5249-5262.
+\defbibnote{postnote}{\small\par\medskip\noindent{\footnotesize% Note:
+\arxivp \mparcp \philscip \biorxivp}%
+}
+
+\newcommand*{\citein}[2][]{\textnormal{\textcite[#1]{#2}}%\addtocategory{extras}{#2}
+}
+\newcommand*{\citebi}[2][]{ref.\ \citep[#1]{#2}%\addtocategory{extras}{#2}
+}
+\newcommand*{\subtitleproc}[1]{}
+
+\printbibliography[postnote=postnote]
 

### 2017-05-15 maxNt.tex
start discussion
add luca's ackn

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T11:42:02+0200
+%% Last-Updated: 2017-05-15T11:55:06+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\puzzle}{{\fontencoding{U}\fontfamily{fontawesometwo}\selectfont\symbol{225}}}
-\newcommand*{\mynote}[1]{ {\color{notecolour}\puzzle\ #1}}
+\newcommand{\mynote}[1]{ {\color{notecolour}\puzzle\ #1}}
 \newcommand*{\widebar}[1]{{\mkern1.5mu\skew{2}\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}}
@@
   physical situations. Also, in some case numerically very similar results.
+
+  The distribution obtained from the population-level application has lower
+  entropy; this usually means that in includes additional assumptions than
+  the other. Intuitively, this additional assumption is that the $n$
+  neurons are part of a larger population, of which they are
+  representative.
+
   Same phenomenon in statistical mechanics \cite{maesetal1999}. Somehow
@@
 
-Use unnumbered third level headings for the acknowledgments. All
-acknowledgments go at the end of the paper. Do not include
-acknowledgments in the anonymized submission, only in the final paper.
+PGLPM thanks Mari \amp\ Miri for continuous encouragement and affection,
+Buster Keaton for filling life with awe and inspiration, and the developers
+and maintainers of \LaTeX, Emacs, AUC\TeX, MiK\TeX, arXiv, biorXiv,
+PhilSci, Hal archives, Python, Inkscape, Sci-Hub for making a free and
+unfiltered scientific exchange possible.
+
+% Use unnumbered third level headings for the acknowledgments. All
+% acknowledgments go at the end of the paper. Do not include
+% acknowledgments in the anonymized submission, only in the final paper.
 

### 2017-05-15 maxNt.tex
add luca's ackn
more discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T11:55:06+0200
+%% Last-Updated: 2017-05-15T14:30:55+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+We prefer not to answer the question that closed the preceding section,
+because the optimal answer can only be given case by case, depending on the
+inferences we are trying to make, on the assumptions needed and negligible,
+and on our computational power. We would just like to stress that the
+answer is not unique, and some answers may imply assumptions we were not
+aware of.
+
+The quandary we found in the present note has at least two bright sides.
+The first is that the two different \me\ applications may give numerically
+similar results in some cases of interest, especially with numerically
+small constraints. One could thus be used as an approximation of the other
+depending on the circumstances -- \eg\ if probability tails are
+unimportant. See \fig~\ref{fig:diff_maxent_pop_sample_realdata} for an
+example.
+\begin{figure}[!t]
+\centering
+\includegraphics[width=0.95\columnwidth]{different_maxent_pop_sample_200_realdata_both.pdf}
+% 
+\caption{Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
+  \me\ at the population level followed by sample marginalization (blue
+  triangles), and at the sample level (red circles). The constraints are
+  the same in either case: $\expe{\yxs \|\yH}=0.045$,
+  $\expe{\yxxs \|\yH}=0.0025$, and $N=5\,000$, $n=200$.}
+\label{fig:diff_maxent_pop_sample_realdata}
+\end{figure}%maxent_pop_or_sample.nb
+
 \mynote{We don't give an answer. Different purposes. Bright side: no real

### 2017-05-15 maxNt.tex
more discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T14:30:55+0200
+%% Last-Updated: 2017-05-15T15:53:25+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\usepackage{wrapfig}
 \usepackage{datetime2}
@@
 \end{equation}
+%\setlength{\intextsep}{0.5ex}%
+%\begin{wrapfigure}{r}{0.5\linewidth}
+\begin{figure}[!t]
+\centering
+\includegraphics[width=0.95\linewidth]{different_maxent_pop_sample_100.pdf}%
+\caption{Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
+  \me\ at the population level followed by sample marginalization (blue
+  triangles), and at the sample level (red circles). The constraints are
+  the same in either case: $\expe{\yxs \|\yH}=0.45$,
+  $\expe{\yxxs \|\yH}=0.35$, and $N=5\,000$, $n=100$.}
+\label{fig:diff_maxent_pop_sample}
+\end{figure}%maxent_pop_or_sample.nb
+%\end{wrapfigure}
 with such parameters $\yl$, $\yl_1$, $\yl_2$ as to satisfy the constraints.
@@
 \fig~\ref{fig:diff_maxent_pop_sample}; they are clearly different.
-\begin{figure}[!t]
-\centering
-\includegraphics[width=0.95\columnwidth]{different_maxent_pop_sample_100.pdf}%
-\caption{Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
-  \me\ at the population level followed by sample marginalization (blue
-  triangles), and at the sample level (red circles). The constraints are
-  the same in either case: $\expe{\yxs \|\yH}=0.45$,
-  $\expe{\yxxs \|\yH}=0.35$, and $N=5\,000$, $n=100$.}
-\label{fig:diff_maxent_pop_sample}
-\end{figure}%maxent_pop_or_sample.nb
 
@@
 
+The probability distribution for the sample average obtained by applying
+\me\ at the population level has by construction a lower entropy than the
+one obtained by application at the sample level. It says that some sample
+states $\yx$ should be considered less likely than in the other
+distribution. The lower entropy signals the presence of an additional
+physical assumption behind the former distribution. Such assumption,
+intuitively, is that the sample is part of a larger population, of which it
+is representative. And this assumption may affect our predictions.
+
 The quandary we found in the present note has at least two bright sides.
 The first is that the two different \me\ applications may give numerically
-similar results in some cases of interest, especially with numerically
-small constraints. One could thus be used as an approximation of the other
-depending on the circumstances -- \eg\ if probability tails are
-unimportant. See \fig~\ref{fig:diff_maxent_pop_sample_realdata} for an
+similar results in some cases. One could thus be used as an approximation
+of the other, depending on the circumstances -- \eg\ if probability tails
+are unimportant. See \fig~\ref{fig:diff_maxent_pop_sample_realdata} for an
 example.
@@
 \end{figure}%maxent_pop_or_sample.nb
-
-\mynote{We don't give an answer. Different purposes. Bright side: no real
-  contradiction because we are expressing two states of knowledge, not
-  physical situations. Also, in some case numerically very similar results.
-
-  The distribution obtained from the population-level application has lower
-  entropy; this usually means that in includes additional assumptions than
-  the other. Intuitively, this additional assumption is that the $n$
-  neurons are part of a larger population, of which they are
-  representative.
-
-  Same phenomenon in statistical mechanics \cite{maesetal1999}. Somehow
-  similar quandary for equilibrium state at the end of non-equilibrium
-  process: Gibbs state vs initial Gibbs state + Liouville
-  evolution \cite[\sect~4]{jaynes1985d_r1993}; but in this case both
-  distributions are immensely sharp, and this doesn't affect predictions}
-
+The second is that the possibility of using two different distributions is
+not a physical contradiction. Similar situations arise in statistical
+mechanics. It is known that if a system is described by a \me\ Gibbs state,
+its subsystems need not be \cite{maesetal1999}. A quandary somehow similar
+to ours also appears in the statistical description of an equilibrium state
+at the end of a non-equilibrium process: we can describe our knowledge
+about it either by a Gibbs distribution, or by the Liouville-evolved Gibbs
+distribution associated with equilibrium state at the beginning of the
+process. The two descriptions differ -- even though the final physical
+state is exactly the same \cite[\sect~4]{jaynes1985d_r1993}. The appearance
+of this difference is natural: in one case we can make sharper predictions
+about the state thanks to our knowledge of its preceding dynamics.
+\mynote{but in this case both distributions are immensely sharp, and this
+  doesn't affect predictions}
+
+\mynote{Importance of the formulae relating sample and population.}
+
+% do not change font sizes (except perhaps in the \textbf{References}
+% section; see below).
+
+\subsubsection*{Acknowledgments}
+
+PGLPM thanks Mari \amp\ Miri for continuous encouragement and affection,
+Buster Keaton for filling life with awe and inspiration, and the developers
+and maintainers of \LaTeX, Emacs, AUC\TeX, MiK\TeX, arXiv, biorXiv,
+PhilSci, Hal archives, Python, Inkscape, Sci-Hub for making a free and
+unfiltered scientific exchange possible.
+
+% Use unnumbered third level headings for the acknowledgments. All
+% acknowledgments go at the end of the paper. Do not include
+% acknowledgments in the anonymized submission, only in the final paper.
+
+%\section*{References}
+\renewcommand*{\bibname}{References}
+\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
+}
+
+% References follow the acknowledgments. Use unnumbered first-level
+% heading for the references. Any choice of citation style is acceptable
+% as long as you are consistent. It is permissible to reduce the font
+% size to \verb+small+ (9 point) when listing the references. {\bf
+%   Remember that you can go over 8 pages as long as the subsequent ones contain
+%   \emph{only} cited references.}
+% \medskip
+
+% \small
+
+% [1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
+% for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
+% T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
+%   Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.
+
+% [2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
+%   Exploring Realistic Neural Models with the GEneral NEural SImulation
+%   System.}  New York: TELOS/Springer--Verlag.
+
+% [3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
+% learning and recall at excitatory recurrent synapses and cholinergic
+% modulation in rat hippocampal region CA3. {\it Journal of
+%   Neuroscience} {\bf 15}(7):5249-5262.
+\defbibnote{postnote}{\small\par\medskip\noindent{\footnotesize% Note:
+\arxivp \mparcp \philscip \biorxivp}%
+}
+
+\newcommand*{\citein}[2][]{\textnormal{\textcite[#1]{#2}}%\addtocategory{extras}{#2}
+}
+\newcommand*{\citebi}[2][]{ref.\ \citep[#1]{#2}%\addtocategory{extras}{#2}
+}
+\newcommand*{\subtitleproc}[1]{}
+
+\printbibliography[postnote=postnote]
+
+\newrefsection
 \bigskip
@@
 
-% do not change font sizes (except perhaps in the \textbf{References}
-% section; see below).
-
-\subsubsection*{Acknowledgments}
-
-PGLPM thanks Mari \amp\ Miri for continuous encouragement and affection,
-Buster Keaton for filling life with awe and inspiration, and the developers
-and maintainers of \LaTeX, Emacs, AUC\TeX, MiK\TeX, arXiv, biorXiv,
-PhilSci, Hal archives, Python, Inkscape, Sci-Hub for making a free and
-unfiltered scientific exchange possible.
-
-% Use unnumbered third level headings for the acknowledgments. All
-% acknowledgments go at the end of the paper. Do not include
-% acknowledgments in the anonymized submission, only in the final paper.
-
-%\section*{References}
-\renewcommand*{\bibname}{References}
-\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
-}
-
-% References follow the acknowledgments. Use unnumbered first-level
-% heading for the references. Any choice of citation style is acceptable
-% as long as you are consistent. It is permissible to reduce the font
-% size to \verb+small+ (9 point) when listing the references. {\bf
-%   Remember that you can go over 8 pages as long as the subsequent ones contain
-%   \emph{only} cited references.}
-% \medskip
-
-% \small
-
-% [1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
-% for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
-% T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
-%   Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.
-
-% [2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
-%   Exploring Realistic Neural Models with the GEneral NEural SImulation
-%   System.}  New York: TELOS/Springer--Verlag.
-
-% [3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
-% learning and recall at excitatory recurrent synapses and cholinergic
-% modulation in rat hippocampal region CA3. {\it Journal of
-%   Neuroscience} {\bf 15}(7):5249-5262.
-\defbibnote{postnote}{\small\par\medskip\noindent{\footnotesize% Note:
-\arxivp \mparcp \philscip \biorxivp}%
-}
-
-\newcommand*{\citein}[2][]{\textnormal{\textcite[#1]{#2}}%\addtocategory{extras}{#2}
-}
-\newcommand*{\citebi}[2][]{ref.\ \citep[#1]{#2}%\addtocategory{extras}{#2}
-}
-\newcommand*{\subtitleproc}[1]{}
-
-\printbibliography[postnote=postnote]
-
 \end{document}

### 2017-05-15 maxNt.tex
more discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T15:53:25+0200
+%% Last-Updated: 2017-05-15T17:05:34+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\yHb}{\varIota''}
+\newcommand*{\yHc}{\varKappa}
 %Lagr multipliers
@@
 
-We have a population of $N$ binary neurons. Let's assume that the
-identities of the neurons can be distinguished, by their spike shapes for
-example; but other details, like their locations, are unknown. The neurons
-have a joint state $(X_1,\dotsc,X_N) \defs \yX$ having fixed but unknown
-binary values $(R_1,\dotsc,R_N)\defs \yR \in \set{0,1}^N$. A particular
-sample of $n$ neurons from this population has joint state
-$(x_1, \dotsc, x_n) \defs \yx$ having fixed binary values
-$(r_1, \dotsc, r_n)\defs\yr$. We will consider various averages for the
-population and the sample. For this purpose we introduce a general
-averaging operator $\sav{\dotv}$ defined by
+We have a population of $N$ binary neurons. We assume that they can be
+distinguished, by their spike shapes for example; but other details, like
+their locations, are unknown. The neurons have a joint state
+$(X_1,\dotsc,X_N) \defs \yX$ having fixed but unknown binary values
+$(R_1,\dotsc,R_N)\defs \yR \in \set{0,1}^N$. A particular sample of $n$
+neurons from this population has joint state $(x_1, \dotsc, x_n) \defs \yx$
+having fixed binary values $(r_1, \dotsc, r_n)\defs\yr$. We will consider
+various averages of the population and the sample. For this purpose we
+introduce a general averaging operator $\sav{\dotv}$ defined by
 \begin{align}
@@
 \end{align}
-and so on; analogously for the quantities $\yR$ and $\yr$. These formulae
-say that $\yXf$ is the fraction of active neurons, $\yXXf$ the fraction of
-simultaneously active pairs out of all $\binom{N}{2}$ pairs, $\av{\yX\yX\yX}$
-the fraction of simultaneously active triplets, and so on.
-
-\mynote{Add formula $\av{\underbrace{\yX\dotsm\yX}_{\text{$m$ factors}}} = \binom{N\yXf}{m}/\binom{N}{m}$.}
-
+and so on; analogously for the quantities like $\yRf$ and $\sav{\yr\yr}$.
+These formulae say that $\yXf$ is the fraction of active neurons, $\yXXf$
+the fraction of simultaneously active pairs out of all $\binom{N}{2}$
+pairs, $\av{\yX\yX\yX}$ the fraction of simultaneously active triplets, and
+so on. Products of states like $X_i \dotsm X_j$ also have values in
+$\set{0,1}$; from this it can combinatorially be proved that
+\begin{equation}
+  \label{eq:products_intermsof_average}
+  \av{\underbrace{\yX\dotsm\yX}_{\text{$m$ factors}}}
+  = \binom{N}{m}^{-1}\binom{N\yXf}{m}
+\end{equation}
+and similarly for $\yx$, $\yR$, $\yr$.
+
+\iffalse
 \begin{simplenote}
@@
 \end{simplenote}
-
-Our uncertainty about the state of the population is completely expressed by the
-probability distribution
+\fi
+
+Our uncertainty about the actual state of the population is completely
+expressed by the joint probability distribution
 \begin{equation}
   \label{eq:joint_plaus}
-  \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \| \yH) \quad\text{or}\quad
-\p(\yX =\yR \| \yH), \quad \yR \in \set{0,1}^N,
-\end{equation}
-where $\yH$ denotes our initial state of knowledge (the evidence and assumptions
-backing our probability assignments). Our uncertainty about the state of
-the sample is likewise expressed by
+  \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \| \yHc) \quad\text{or}\quad
+\p(\yX =\yR \| \yHc), \quad \yR \in \set{0,1}^N,
+\end{equation}
+where $\yHc$ denotes our state of knowledge, \ie\ the evidence and
+assumptions backing this particular probability assignment. Our uncertainty
+about the state of the sample is likewise expressed by
 \begin{equation}
   \label{eq:sample_plaus}
-  \p(x_1=r_1, x_2=r_2, \dotsc, x_n=r_n \| \yH) \quad\text{or}\quad
-\p(\yx =\yr \| \yH), \quad \yr \in \set{0,1}^n.
+  \p(x_1=r_1, x_2=r_2, \dotsc, x_n=r_n \| \yHc) \quad\text{or}\quad
+\p(\yx =\yr \| \yHc), \quad \yr \in \set{0,1}^n.
 \end{equation}
@@
 
-Before any experimental observations are made we need to make an initial
-probability assignment, no matter what kinds of predictions we are
+We need to make an initial probability assignment before any experimental
+observations are made, no matter what kinds of predictions we are
 interested in. This initial assignment will be modified by the experimental
@@
 We yet know very little about the physical details of the individual
-neurons, like their locations. Our state of knowledge is therefore
-symmetric, or \enquote{exchangeable}, under their permutations. This
-symmetry must be reflected in our initial probability, and the
-\emph{representation theorem for finite exchangeability} states that it
+neurons, their locations for example. Our initial state of knowledge $\yH$
+is therefore symmetric, or \enquote{exchangeable}, under their
+permutations. This symmetry must be reflected in our initial probability:
+the \emph{representation theorem for finite exchangeability} states that it
 must obey
@@
 this symmetry. The initial symmetry should intuitively also apply to the
-sample. This is indeed the case: the probability for the state of the
+sample, and this is indeed the case: the probability for the state of the
 sample~\eqref{eq:subpop_average} automatically satisfies the representation
@@
 
-Our initial knowledge (or ignorance) $\yH$ has two very important
-properties derivable from the formulae above. First, \emph{our initial
-  expectations for all averages of the sample and the full population are
-  the same}:
+Our initial knowledge (ignorance) $\yH$ has two important properties
+derivable from the formulae above.
+
+First, our initial expectations for the averages of activity products
+\emph{are the same for the sample and for the full population}:
 \begin{equation}
@@
 % \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
-\equiv
+=
 \binom{N}{m}^{-1}
@@
 \end{equation}
+Raw moments like $\expe{\yxs^m}$, which can be written in terms of the
+product expectations above via \eqn~\eqref{eq:expe_products}, similarly
+satisfy $\expe{\yxs^m} \to \expe{\yXf^m}$ as $n \to N$.
+
 Second, upon observation of a sample average, say $\yxxs=a$, the updated
@@
 
-\mynote{Add paragraph on moments}
-
 These two properties express the fact that the sample is
 \emph{representative} of the population. Our initial knowledge $\yH$ thus
-assumes that we can make predictions about the population by observing a
-sample. Contrast this with an assumption that assigns an independent
-probability to each neuron: in this case, the observation of one sample
-does not change our initial probability about another; we can't learn from
-observation.
+assumes that by observing a sample we can make predictions about the full
+population and about new samples. Contrast this with an assumption that
+assigns an independent probability to each neuron: in this case, the
+observation of one sample does not change our initial probability about
+another; we can't learn from observation.
 
@@
 
+\mynote{Discussion of case when individual $X_i\dotsm X_j$ are constrained}
+
 % do not change font sizes (except perhaps in the \textbf{References}
@@
 \begin{align}
-  \label{eq:population-av}
+  \label{noeq:population-av}
    \yXf &\defd (X_1 + \dotsb + X_N)/N,
@@
 \intertext{and the subpopulation-averaged state, or \emph{subpopulation average}:}
-\label{eq:subpopulation-av}
+\label{noeq:subpopulation-av}
 \yxs &\defd (x_1 + \dotsc + x_n)/n.
@@
 \subsection{Assumptions}
-\label{sec:assumptions}
+\label{nosec:assumptions}
 
@@
 \begin{equation}
-  \label{eq:joint_plaus}
+  \label{noeq:joint_plaus}
   \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \| \yH) \quad\text{or}\quad
@@
 \item \emph{Permutation symmetry}, expressed as the invariance of the
-  joint distribution~\eqref{eq:joint_plaus} under arbitrary permutations of
+  joint distribution~\eqref{noeq:joint_plaus} under arbitrary permutations of
   the units's labels:
 \begin{multline}
-  \label{eq:homog_math}
+  \label{noeq:homog_math}
   \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \| \yH) ={}
@@
 literature and its basis, consequences, and alternatives to it are discussed
-in \sect~\ref{sec:stoch_symm}.
+in \sect~***%\ref{nosec:stoch_symm}.
 
@@
 \begin{equation}
-  \label{eq:pop_average_Q}
+  \label{noeq:pop_average_Q}
   \p(\yXf = A \| \yH)
@@
 %\addcontentsline{toc}{section}{The mathematical relations}
-\label{sec:main_formulae}
+\label{nosec:main_formulae}
 
@@
 \begin{enumerate}%[wide,label=\textbf{\emph{\Roman*}}.]
-  \medskip\item\label{item:tot_plaus}Probability for the population state:
-\begin{equation}
-  \label{eq:joint_plaus_N_homog}
+  \medskip\item\label{noitem:tot_plaus}Probability for the population state:
+\begin{equation}
+  \label{noeq:joint_plaus_N_homog}
   \p(\yX = \yR \| H) = \binom{N}{N\yRf}^{-1} Q(\yRf).
@@
 \medskip
-\item\label{item:marginal}Probability for the state $\yx$ of any subpopulation
+\item\label{noitem:marginal}Probability for the state $\yx$ of any subpopulation
   of $n$ units:
 \begin{equation}
-  \label{eq:marginal}
+  \label{noeq:marginal}
     \p(\yx = \yr \| \yH) =
@@
 \medskip
-\item\label{item:conditional}Probability for the subpopulation
+\item\label{noitem:conditional}Probability for the subpopulation
   state conditional on a population state:
 \begin{equation}
-  \label{eq:conditional_total_state}
+  \label{noeq:conditional_total_state}
   \p(\yx = \yr \| \yX = \yR, \yH)
@@
 \medskip
-\item\label{item:marginal_average}Probability for the
+\item\label{noitem:marginal_average}Probability for the
   subpopulation average $\yxs$:
 \begin{multline}
-  \label{eq:subpop_average}
+  \label{noeq:subpop_average}
   \p(\yxs = a  \| \yH) = \binom{n}{n a}
@@
 \medskip
-\item\label{item:conditional_average}Probability for the subpopulation
+\item\label{noitem:conditional_average}Probability for the subpopulation
   average conditional on the population average:
 \begin{equation}
-  \label{eq:conditional_prob}
+  \label{noeq:conditional_prob}
   \p(\yxs = a \| \yXf = A, \yH)
@@
 \medskip
-\item\label{item:moments}The product of the states of any $m$ distinct
+\item\label{noitem:moments}The product of the states of any $m$ distinct
   units from a given subpopulation,
@@
 \begin{subequations}
-\label{eq:expe_products}
+\label{noeq:expe_products}
 \begin{align}
@@
 A useful relation connects the expectation of a
-product~\eqref{eq:expe_products} and the $m$th \emph{factorial moment}
+product~\eqref{noeq:expe_products} and the $m$th \emph{factorial moment}
 \citep{potts1953} of the probability distributions for the averages. The
@@
 \begin{equation}
-  \label{eq:fact_mom_defin}
+  \label{noeq:fact_mom_defin}
 \vphantom{\underbrace{\yxs}_{m}}\expeb*{
@@
 \begin{equation}
-  \label{eq:eq_factmom_prod}
+  \label{noeq:eq_factmom_prod}
 \expe{x_{i_1} \dotsm x_{i_m} \| \yH}
@@
 are related by
-\begin{subequations}\label{eq:moments}
+\begin{subequations}\label{noeq:moments}
   \begin{align}
@@
 Relations for higher moments can be obtained recursively from
-\eqn~\eqref{eq:eq_factmom_prod}. In general, this means that the two sets of
+\eqn~\eqref{noeq:eq_factmom_prod}. In general, this means that the two sets of
 first $m$ moments are related by a homogeneous linear transformation,
 \begin{equation}
-  \label{eq:lin_transf_moments}
+  \label{noeq:lin_transf_moments}
   \expe{\yxs^m \| \yH} = \sum_{l=1}^m M_{ml}(n,N) \expe{\yXf^l \| \yH},
@@
 with a universal, lower-triangular transformation matrix $M_{ml}(n,N)$ that
-depends only on $n$, $N$, and the condition of symmetry~\eqref{eq:homog_math}.
+depends only on $n$, $N$, and the condition of symmetry~\eqref{noeq:homog_math}.
 
@@
 \begin{equation}
-  \label{eq:limit_moments}
+  \label{noeq:limit_moments}
   \expe{\yxs^m \| \yH} \xrightarrow{n\to N} 
@@
 The core of the six mathematical relations above are
-\eqns~\eqref{eq:marginal} and~\eqref{eq:subpop_average}. The latter
+\eqns~\eqref{noeq:marginal} and~\eqref{noeq:subpop_average}. The latter
 expresses the probability for the subpopulation average as a mixture of
@@
 The connection between this mixture representation and the condition of
-symmetry~\eqref{eq:homog_math} is well-known in the Bayesian literature
+symmetry~\eqref{noeq:homog_math} is well-known in the Bayesian literature
 \citep{kendall1967,definetti1969b,heathetal1976,diaconis1977,diaconisetal1980,jaynes1986c}.
@@
 \subsection{From network to subnetwork}
-\label{sec:pop2sub_examples}
+\label{nosec:pop2sub_examples}
 
 Let us illustrate with an example how the probability distribution for the
-subnetwork average $\yxs$, determined by \eqn~\eqref{eq:subpop_average},
+subnetwork average $\yxs$, determined by \eqn~\eqref{noeq:subpop_average},
 changes with the subnetwork size $n$. Choose a network-average distribution
@@
 \begin{equation}
-  \label{eq:full_pop_av_example_ME}
+  \label{noeq:full_pop_av_example_ME}
   \p(\yXf = A \| \yH) = Q(A) \propto
@@
   distribution $\p(\yXf = A \| \yH)$ having the maximum-etropy
-  form~\eqref{eq:full_pop_av_example_ME}.}
-\label{scaling_distr}
+  form~\eqref{noeq:full_pop_av_example_ME}.}
+\label{noscaling_distr}
 \end{figure}
@@
   \yH)$ as functions of the subnetwork size $n$.}
-\label{scaling_moments}
+\label{noscaling_moments}
 \end{figure}
 
-The probability distribution of \eqn~\eqref{eq:full_pop_av_example_ME} is
-plotted in \fig~\ref{scaling_distr}, together with the resulting
+The probability distribution of \eqn~\eqref{noeq:full_pop_av_example_ME} is
+plotted in \fig~\ref{noscaling_distr}, together with the resulting
 subnetwork-average distributions $\p(\yxs = a \|\yH)$, for the case in
@@
 To the eye familiar with maximum-entropy distributions, the
-subnetwork-average distributions of \fig~\ref{scaling_distr} do not look
+subnetwork-average distributions of \fig~\ref{noscaling_distr} do not look
 like maximum-entropy ones with second-moment constraints. In fact,
@@
 \begin{equation}
-  \label{eq:maxent_form_sub}
+  \label{noeq:maxent_form_sub}
   \p(\yxs =a  \| \yH)
@@
 parameters (normalization included), but we must also satisfy the $n+1$
-equations corresponding to the marginalization~\eqref{eq:subpop_average}.
+equations corresponding to the marginalization~\eqref{noeq:subpop_average}.
 This is generally impossible unless $m \ge n$. (Impossibilities of a
@@
 ($n < 200$) of neurons. If we assume that such subnetwork is part of a
-larger network, and assume the condition of symmetry~\eqref{eq:homog_math},
+larger network, and assume the condition of symmetry~\eqref{noeq:homog_math},
 then the larger network \emph{cannot} be assigned a maximum-entropy
@@
 The dependence of the first four moments $\expe{\yxs^m \| \yH}$ as a
-function of size $n$ is shown in \fig~\ref{scaling_moments}. The moments
+function of size $n$ is shown in \fig~\ref{noscaling_moments}. The moments
 become practically constant when $n \approx 100$ or larger. The
@@
 \subsection{From subnetwork to network}
-\label{sec:from_sub_to_full}
-
-We have seen that, given the condition of symmetry~\eqref{eq:homog_math},
+\label{nosec:from_sub_to_full}
+
+We have seen that, given the condition of symmetry~\eqref{noeq:homog_math},
 the probability $\p(\yXf = A \|\yH)$ for the network average determines
 that of each subnetwork average, $\p(\yxs \|\yH)$, by the
-marginalization \eqn~\eqref{eq:subpop_average}. The reverse is trivially
-not true, since \eqn~\eqref{eq:subpop_average}, as a linear mapping from
+marginalization \eqn~\eqref{noeq:subpop_average}. The reverse is trivially
+not true, since \eqn~\eqref{noeq:subpop_average}, as a linear mapping from
 $\RR^{N+1}$ to $\RR^{n+1}$, with $N$ larger than $n$, is onto but not into.
@@
 \paragraph{First example: moment constraints for the network.}
-\label{sec:maxent_moments}
+\label{nosec:maxent_moments}
 Consider a state of knowledge $\yHa$ leading to the following properties:
@@
   \begin{equation}
-    \label{eq:constr_ex1}
+    \label{noeq:constr_ex1}
     \expe{\yxs \| \yHa} = c_1, \qquad \expe{\sav{x_i x_j}\|\yHa} = c_2;
@@
 is completely determined: it satisfies the symmetry
-property~\eqref{eq:homog_math} and is defined by
+property~\eqref{noeq:homog_math} and is defined by
 \begin{multline}
-  \label{eq:pop_distr_maxent1}
+  \label{noeq:pop_distr_maxent1}
   \pf(\yX =\yR \|\yHa) =
@@
 \citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013},
-combined with the equality~\eqref{eq:expe_products} of subnetwork and
+combined with the equality~\eqref{noeq:expe_products} of subnetwork and
 network expectations, \eg
 \begin{equation}
-  \label{eq:exploit_eq_expects}
+  \label{noeq:exploit_eq_expects}
   c_2 = \expe{\sav{\yx\yx}\|\yHa} = 
@@
 \end{equation}
-and with relations~\eqref{eq:joint_plaus_N_homog},
-\eqref{eq:conditional_total_state}. This example is easily generalized to
+and with relations~\eqref{noeq:joint_plaus_N_homog},
+\eqref{noeq:conditional_total_state}. This example is easily generalized to
 any number $m$ of constraints such that  $m\le n$.
 
-Note again that, as remarked in \sect~\ref{sec:pop2sub_examples}, the
+Note again that, as remarked in \sect~\ref{nosec:pop2sub_examples}, the
 subnetwork from which the averages in the
-expectations~\eqref{eq:constr_ex1} are calculated has a probability
+expectations~\eqref{noeq:constr_ex1} are calculated has a probability
 distribution $\p(\yxs =a \|\yH)$ determined by the
-marginalization~\eqref{eq:subpop_average} and does \emph{not} have a
+marginalization~\eqref{noeq:subpop_average} and does \emph{not} have a
 maximum-entropy form with the same number of constraints.
@@
 \paragraph{Second example: subnetwork-distribution constraint.}
-\label{sec:maxent_fullsubpop}
+\label{nosec:maxent_fullsubpop}
 Consider another state of knowledge $\yHb$ leading to the following
@@
   \begin{equation}
-    \label{eq:constr_subpop_distr}
+    \label{noeq:constr_subpop_distr}
     \p(\yxs =a \|\yHb) = q(a);
@@
 completely determined and satisfies the symmetry
-property~\eqref{eq:homog_math}:
+property~\eqref{noeq:homog_math}:
 \begin{multline}
-  \label{eq:pop_distr_maxent2}
+  \label{noeq:pop_distr_maxent2}
   \p(\yX= \yR \|\yHb) =
@@
 This result is just another application of the maximum-entropy procedure
-with $n+1$ (linear) constraints given by \eqn~\eqref{eq:marginal}, where
+with $n+1$ (linear) constraints given by \eqn~\eqref{noeq:marginal}, where
 the left-hand side is now given and equal to $q(a)$.
@@
   \section{Sketched proofs}
-  \label{sec:derivations}
+  \label{nosec:derivations}
 
@@
   To derive the joint probability distribution
-  \eqref{eq:joint_plaus_N_homog} from that for the network
-  average~\eqref{eq:pop_average_Q}, consider that if the network total
+  \eqref{noeq:joint_plaus_N_homog} from that for the network
+  average~\eqref{noeq:pop_average_Q}, consider that if the network total
   is $N\yXf$, then $N\yXf$ out of $N$ units are active, and there are
@@
 \begin{equation}
-  \label{eq:aver_subpop}
+  \label{noeq:aver_subpop}
   \p(\yx=\yr \| H) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs\|\yH),
@@
 probability of each is therefore $1/\tbinom{N}{N\yXf}$, owing to the
-symmetry assumption~\eqref{eq:homog_math}. Now consider the subnetwork of the
+symmetry assumption~\eqref{noeq:homog_math}. Now consider the subnetwork of the
 first $n$ units. The conditional probability of having $n\yxs$ specific
@@
 \begin{equation}
-\label{eq:conditional_prob_explained}
+\label{noeq:conditional_prob_explained}
 \begin{split}
@@
 \end{equation}
-  which is the conditional probability~\eqref{eq:conditional_prob}. Note
+  which is the conditional probability~\eqref{noeq:conditional_prob}. Note
   that this is just a derivation of the hypergeometric distribution
@@
   \begin{equation}
-    \label{eq:subpop_average_explained}
+    \label{noeq:subpop_average_explained}
     \begin{split}
@@
   \end{equation}
-  which proves the subnetwork-average formula~\eqref{eq:subpop_average}.
-  This formula, combined with \eqns~\eqref{eq:aver_subpop} and
-  \eqref{eq:joint_plaus_N_homog}, leads to the conditional
-  probability~\eqref{eq:conditional_total_state}.
+  which proves the subnetwork-average formula~\eqref{noeq:subpop_average}.
+  This formula, combined with \eqns~\eqref{noeq:aver_subpop} and
+  \eqref{noeq:joint_plaus_N_homog}, leads to the conditional
+  probability~\eqref{noeq:conditional_total_state}.
 
@@
   \begin{multline}
-    \label{eq:margin_products}
+    \label{noeq:margin_products}
     \sum_{\yx} x_1\dotsm x_m \p(\yx =\yr \| \yH)
@@
   \end{equation}
-  Then, using \eqn~\eqref{eq:subpop_average},
+  Then, using \eqn~\eqref{noeq:subpop_average},
   \begin{multline}
-    \label{eq:expect_product_m}
+    \label{noeq:expect_product_m}
     \expe*{\sav{x_{i_1} \dotsm x_{i_m}} \| \yH}
@@
   which combined with the previous equation yields the second line of
-  \eqn~\eqref{eq:expe_products}; its last equality comes from the identity
-  \begin{equation}
-    \label{eq:binom_identity}
+  \eqn~\eqref{noeq:expe_products}; its last equality comes from the identity
+  \begin{equation}
+    \label{noeq:binom_identity}
     \binom{N}{M}\binom{M}{m}=\binom{N}{m}\binom{N-m}{M-m},
@@
   easily derived by writing the binomial coefficients in terms of
-  factorials. Finally, \eqns~\eqref{eq:moments}, relating the moments of
+  factorials. Finally, \eqns~\eqref{noeq:moments}, relating the moments of
   the distributions for subnetwork and network averages, is obtained
@@
   \begin{equation}
-    \label{eq:moments_def}
+    \label{noeq:moments_def}
     \expe*{\yxs^m \| \yH} \defd
@@
   replaced in the equalities for the factorial
-  moments~\eqref{eq:eq_factmom_prod}, by recursively solving in terms of
+  moments~\eqref{noeq:eq_factmom_prod}, by recursively solving in terms of
   the moments of the network distribution.

### 2017-05-15 maxNt.tex
more discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T17:05:34+0200
+%% Last-Updated: 2017-05-15T17:29:54+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\author{ V. Rostami\\
-  Forschungszentrum Jülich INM-6\\Hitlerland\\
-  \texttt{v.rostami@fz-juelich.de}
+\author{
+  V. Rostami\\Hitlerland\\\texttt{v.rostami@fz-juelich.de}
   \And
-  P.G.L. Porta Mana\\Mussoliniland
-  \And E. Torre\\Moneylaunderingland  
+  P.G.L. Porta Mana\\Mussoliniland\\\texttt{pgl@portamana.org}
+  \And
+E. Torre\\Moneylaunderingland  
   % David S.~Hippocampus\thanks{Use footnote for providing further
@@
 observation of one sample does not change our initial probability about
-another; we can't learn from observation.
+another -- we can't learn from observation.
 
@@
 constraints on our initial probability assignment, but do not determine it
-numerically yet. The probability $\p(\yXf=\yRf \| \yH)$ for the population
+numerically. The probability $\p(\yXf=\yRf \| \yH)$ for the population
 average needs to be numerically specified, and by \eqref{eq:subpop_average}
@@
 We may want to specify the probability by enforcing several of its sample
-expectations to have specific values, for example $\expe{\yxs}=a$,
-$\expe{\yxxs}=b$. But this is still and underdetermined linear problem
-because several distributions can have the same desired expectations, as
+expectations to have specific values, for example $\expe{\yxs}=c_1$,
+$\expe{\yxxs}=c_2$. This is still an underdetermined linear problem:
+ several distributions can have the same desired expectations, as
 clear from \eqns~\eqref{eq:expe_products}.
 
-It is at this point that the \me\ method enters the scene. It selects one
-distribution, purported to be \enquote{maximally noncommittal}, among those
-that have the desired expectations. But we are in a quandary:
+The \me\ method is brought into play to solve this indeterminacy. It selects
+one distribution, purported to be \enquote{maximally noncommittal}, among
+those that have the desired expectations. But we are in a quandary:
 formulae~\eqref{eq:expe_products} allow us to apply the method to find the
 probability of the population $\p(\yXf=\yRf \| \yH)$, or of the sample
-$\p(\yxs=\yrs \| \yH)$; \emph{the two applications, however, are
+$\p(\yxs=\yrs \| \yH)$. \emph{The two applications, however, are
   inequivalent}. They lead to numerically different $\p(\yxs=\yrs \| \yH)$.
@@
 \end{equation}
-Applying \me\ at the sample level, using the third term of
-\eqn~\eqref{eq:expe_products}, we find
+Applying \me\ to the distribution of the average of the sample, using the
+third term of \eqn~\eqref{eq:expe_products}, we find
 \begin{equation}
@@
 with such parameters $\yl$, $\yl_1$, $\yl_2$ as to satisfy the constraints.
-We call this state of knowledge $\yHb$. For example, with $N=5\,000$,
-$n=100$, $c_1=0.45$, $c_2=0.35$, we find $\yL_1 = -13\,321.9$,
+We call this state of knowledge $\yHb$.
+
+The resulting distributions $\p(\yxs = \yrs \| \yHa)$,
+$\p(\yxs = \yrs \| \yHb)$ are different.
+Figure~\ref{fig:diff_maxent_pop_sample} shows an example, for values
+$N=5\,000$, $n=100$, $c_1=0.45$, $c_2=0.35$, yielding $\yL_1 = -13\,321.9$,
 $\yL_2 = 13\,321.5$, $\yl_1=-269.3$, $\yl_2 = 268.9$, and appropriate
-normalization constants. The resulting distributions
-$\p(\yxs = \yrs \| \yHa)$, $\p(\yxs = \yrs \| \yHb)$ are shown in
-\fig~\ref{fig:diff_maxent_pop_sample}; they are clearly different.
+normalization constants.
 
@@
 use different numbers $m'$, $m''$ \emph{and values} of constraints. The
-equality of the third and fourth term of \eqn~\eqref{eq:expe_products} in
-this case implies $n+1$ equations in $m'+m''+1$ unknowns (one normalization
-is taken care of), and in general can only be satisfied if $n\le m'+m''$.
+equality $\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$ in this case
+requires solving $n$ equations in $m'+m''$ unknowns (normalization
+is taken care of), so in general it can only be satisfied if $n\le m'+m''$.
 

### 2017-05-15 maxNt.tex
more discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T17:29:54+0200
+%% Last-Updated: 2017-05-15T17:53:53+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{equation}
+with such parameters $\yl$, $\yl_1$, $\yl_2$ as to satisfy the constraints.
+We call this state of knowledge $\yHb$.
 %\setlength{\intextsep}{0.5ex}%
@@
 \centering
-\includegraphics[width=0.95\linewidth]{different_maxent_pop_sample_100.pdf}%
-\caption{Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
+\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_100.pdf}%
+\caption{\mynote{shorten }Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
   \me\ at the population level followed by sample marginalization (blue
@@
 \end{figure}%maxent_pop_or_sample.nb
-%\end{wrapfigure}
-with such parameters $\yl$, $\yl_1$, $\yl_2$ as to satisfy the constraints.
-We call this state of knowledge $\yHb$.
+% \end{wrapfigure}
+\begin{figure}[!t]
+\centering
+\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_both.pdf}
+% 
+\caption{\mynote{shorten }Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
+  \me\ at the population level followed by sample marginalization (blue
+  triangles), and at the sample level (red circles). The constraints are
+  the same in either case: $\expe{\yxs \|\yH}=0.045$,
+  $\expe{\yxxs \|\yH}=0.0025$, and $N=5\,000$, $n=200$.}
+\label{fig:diff_maxent_pop_sample_realdata}
+\end{figure}%maxent_pop_or_sample.nb
 
@@
 $\p(\yxs = \yrs \| \yHb)$ are different.
-Figure~\ref{fig:diff_maxent_pop_sample} shows an example, for values
+
+Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for values
 $N=5\,000$, $n=100$, $c_1=0.45$, $c_2=0.35$, yielding $\yL_1 = -13\,321.9$,
 $\yL_2 = 13\,321.5$, $\yl_1=-269.3$, $\yl_2 = 268.9$, and appropriate
-normalization constants.
+normalization constants. Another example is shown in
+\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for neurobiologically more
+realistic constraints $c_1=0.045$, $c_2=0.0025$, and $N=5\,000$, $n=200$,
+yielding $\yL_1 = -16\,834.0$, $\yL_2 = 16\,825.7$, $\yl_1=-673.5$,
+$\yl_2 = 665.1$. The discrepancy, although not as large as in the previous
+example, is still relevant for the tails of the distributions.
 
@@
 
-Which application of the \me\ method should we choose, then?
+Where should we apply the \me\ method then, on the sample or on the
+population?
 
@@
 
+\iffalse
 The quandary we found in the present note has at least two bright sides.
@@
 \end{figure}%maxent_pop_or_sample.nb
-The second is that the possibility of using two different distributions is
+The second is that
+\fi
+The possibility of using two different distributions is
 not a physical contradiction. Similar situations arise in statistical

### 2017-05-15 maxNt.tex
more discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T17:53:53+0200
+%% Last-Updated: 2017-05-15T18:21:39+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 is representative. And this assumption may affect our predictions.
+\mynote{It is a very natural assumption in the neuroscientific context and
+  therefore we find it preferable. In fact, also physical neuronal network
+  model add an external input to the neurons, mimicking their embedding in a
+  larger network}
 
@@
 
-\mynote{Discussion of case when individual $X_i\dotsm X_j$ are constrained}
+\mynote{Discussion of case when individual $X_i\dotsm X_j$ are constrained.
+Likely breakdown of \me\ in this case \cite{portamana2009}. Hints at full
+Bayesian treatment of which \me\ is a limit? \cite[***]{mackay1995_r2003}}
 

### 2017-05-15 maxNt.tex
fix green todos

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T18:21:39+0200
+%% Last-Updated: 2017-05-15T22:57:55+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_100.pdf}%
-\caption{\mynote{shorten }Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
-  \me\ at the population level followed by sample marginalization (blue
-  triangles), and at the sample level (red circles). The constraints are
-  the same in either case: $\expe{\yxs \|\yH}=0.45$,
-  $\expe{\yxxs \|\yH}=0.35$, and $N=5\,000$, $n=100$.}
+\caption{Plots of $\p(\yxs = \yrs)$ constructed by \me\ at the population
+  level followed by sample marginalization (blue triangles),
+  \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
+  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5\,000$,
+  $n=100$, $\expe{\yxs}=0.45$, $\expe{\yxxs}=0.35$, $\yL_1 = -13\,321.9$,
+  $\yL_2 = 13\,321.5$, $\yl_1=-269.3$, $\yl_2 = 268.9$.}
 \label{fig:diff_maxent_pop_sample}
@@
 % 
-\caption{\mynote{shorten }Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
-  \me\ at the population level followed by sample marginalization (blue
-  triangles), and at the sample level (red circles). The constraints are
-  the same in either case: $\expe{\yxs \|\yH}=0.045$,
-  $\expe{\yxxs \|\yH}=0.0025$, and $N=5\,000$, $n=200$.}
+\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at
+  the population level followed by sample marginalization (blue triangles),
+  \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
+  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5\,000$,
+  $n=200$, $\expe{\yxs}=0.045$, $\expe{\yxxs}=0.0025$,
+  $\yL_1 = -16\,834.0$, $\yL_2 = 16\,825.7$, $\yl_1=-673.5$,
+  $\yl_2 = 665.1$.}
 \label{fig:diff_maxent_pop_sample_realdata}
@@
 $\p(\yxs = \yrs \| \yHb)$ are different.
-
-Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for values
-$N=5\,000$, $n=100$, $c_1=0.45$, $c_2=0.35$, yielding $\yL_1 = -13\,321.9$,
-$\yL_2 = 13\,321.5$, $\yl_1=-269.3$, $\yl_2 = 268.9$, and appropriate
-normalization constants. Another example is shown in
-\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for neurobiologically more
-realistic constraints $c_1=0.045$, $c_2=0.0025$, and $N=5\,000$, $n=200$,
-yielding $\yL_1 = -16\,834.0$, $\yL_2 = 16\,825.7$, $\yl_1=-673.5$,
-$\yl_2 = 665.1$. The discrepancy, although not as large as in the previous
-example, is still relevant for the tails of the distributions.
+Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for $N=5\,000$,
+$n=100$, $c_1=0.45$, $c_2=0.35$. The two distributions have slightly
+different modes and the one obtained from \me\ at the population level is
+more peaked. Another example is shown in
+\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for $N=5\,000$, $n=200$,
+and neurobiologically more realistic constraints $c_1=0.045$, $c_2=0.0025$
+\cite{rostamietal2016}. The discrepancy is maybe not as large as in the
+previous example, but the tails of the distributions are still very
+different.
 

### 2017-05-16 maxNt.tex
adding a short introduction on the usage of MaxEnt in neuroscence

--- 2017-05-15 maxNt.tex
@@
 
-The \me\ method is used in neuroscience blahblah***
-
+Recent advances in electrophysiological recordings allow to
+ study the collective behavior of large population of neurons 
+in relation to stimuli and bahavioral tasks [\cite]. Huge amount of parallel
+ recordings from hundreds of neurons, available today, is accompanied by
+ the statistical challenges to shed light on the mechanism behind neuronal coding. 
+The principle of \me\ was proposed as a way to overcome these challenges by constructing a full description of the 
+neuronal firing patterns based on only first and second order statistics of the observed data [\cite].
+
+ used in neuroscience to study the patterns of neuronal
+ firing in relation to stimuli or behavioral tasks. These studies strive for understanding the 
+mechanism behind neuronal coding. 
+ by answering
+ questions such as, which level of interaction (pairwise or higher-order) between neurons
+ is suffiecient to explain the observed experimental data [\cite]? Or how collective behavior
+ of neurons can be related to stimuli or behavioral tasks [\cite]? These questions 
+**********
 \mynote{Version 1}

### 2017-05-16 maxNt.tex
working on a short introduction regarding the neuroscience background

@@
 
- used in neuroscience to study the patterns of neuronal
- firing in relation to stimuli or behavioral tasks. These studies strive for understanding the 
-mechanism behind neuronal coding. 
- by answering
+ These studies strive for answering 
  questions such as, which level of interaction (pairwise or higher-order) between neurons

### 2017-05-16 maxNt.tex
worked on Intro

@@
 
-\section{Implicit assumptions in the maximum-entropy method}
+%\section{Implicit assumptions in the maximum-entropy method}
+\section{Introduction}
 
@@
 
- These studies strive for answering 
+ Using the pairwise \me\ models many studies strived for answering 
  questions such as, which level of interaction (pairwise or higher-order) between neurons
  is suffiecient to explain the observed experimental data [\cite]? Or how collective behavior
- of neurons can be related to stimuli or behavioral tasks [\cite]? These questions 
-**********
+ of neurons can be related to stimuli or behavioral tasks [\cite]?
+
+Here we show that there is a
+contradiction in applying \me\ to the observed data \enquote{representative random sample}
+of neurons to generate a \enquote{maximally noncommital} distribution for
+that sample.
+
+The contradiction is this. Assume that our sample of neurons is
+representative of some population. If we assign a \me\ distribution to the
+sample, then we cannot assign a \me\ distribution to the full population.
+Vice versa, If we assign a \me\ distribution to the full population, then
+we cannot assign a \me\ distribution to the sample. This impossibility
+holds at least for \me\ distribution with moment constraints, and even if
+the orders of the moments constrained in the sample and in the full
+population are different.
+
+In other words, if our sample is a \enquote{random representative} of some
+population, then we must choose which has a \enquote{maximally
+  noncommittal} distribution: the sample or the population. We can't choose
+both. It goes without saying that if one is \enquote{maximally
+  noncommittal}, the other must be somehow \enquote{committal}. Which
+choice is most meaningful?
+
+In the rest of the paper we will mathematically show the contradiction
+above and generate a \me\ distribution for the full population rather than
+the sample. We will also present some probability relations relevant to
+random sampling. These relations are well-known in survey sampling and in
+the pedagogic problem of drawing from an urn without replacement, yet they
+are somewhat hard to find explicitly written in the neuroscientific
+literature.
+
  \mynote{Version 1}

### 2017-05-16 maxNt.tex
worked on Intro

@@
 Recent advances in electrophysiological recordings allow to
- study the collective behavior of large population of neurons 
+ study the activity of large population of neurons 
 in relation to stimuli and bahavioral tasks [\cite]. Huge amount of parallel
  recordings from hundreds of neurons, available today, is accompanied by
- the statistical challenges to shed light on the mechanism behind neuronal coding. 
+ the statistical challenges to be employed for shedding light on the mechanism behind neuronal coding. 
 The principle of \me\ was proposed as a way to overcome these challenges by constructing a full description of the 
-neuronal firing patterns based on only first and second order statistics of the observed data [\cite].
+neuronal firing patterns. The \me\ method exploits only first and second
+order statistics of the observed data and construct a least structured or
+a \enquote{maximally noncommital} probability distribution of firing patterns [\cite].
 

### 2017-05-16 maxNt.tex
worked on Intro

@@
 
-Recent advances in electrophysiological recordings allow to
- study the activity of large population of neurons 
-in relation to stimuli and bahavioral tasks [\cite]. Huge amount of parallel
- recordings from hundreds of neurons, available today, is accompanied by
- the statistical challenges to be employed for shedding light on the mechanism behind neuronal coding. 
-The principle of \me\ was proposed as a way to overcome these challenges by constructing a full description of the 
-neuronal firing patterns. The \me\ method exploits only first and second
-order statistics of the observed data and construct a least structured or
-a \enquote{maximally noncommital} probability distribution of firing patterns [\cite].
+Recent advances in electrophysiological techniques allow to
+record from large population of neurons [\cite]. The recorded neurons
+are usually picked out according to an unknown process, and from their
+observation we expect to learn something about all other neurons in the
+same brain area. That is, they are assumed to be a
+\enquote{representative random sample} of the neurons in that area.
+
+Huge amount of parallel recordings from hundreds of neurons,
+available today, is accompanied by the statistical challenges
+to be employed for shedding light on the mechanism behind neuronal coding. 
+The principle of \me\ was proposed as a way to overcome these challenges by
+constructing a full description of the 
+neuronal firing patterns. The \me\ method employs only first and second
+order statistics of the representative random sample to construct 
+a \enquote{maximally noncommittal} \cite{jaynes1963} 
+probability distribution of the firing patterns [\cite].
 
@@
  is suffiecient to explain the observed experimental data [\cite]? Or how collective behavior
- of neurons can be related to stimuli or behavioral tasks [\cite]?
+ of neurons can be related to stimuli or behavioral tasks [\cite]? add more **
 
 Here we show that there is a
-contradiction in applying \me\ to the observed data \enquote{representative random sample}
-of neurons to generate a \enquote{maximally noncommital} distribution for
+contradiction in applying \me\ to the representative random sample
+of neurons to generate a maximally noncommital distribution for
 that sample.

### 2017-05-16 maxNt.tex
worked on Intro

@@
 record from large population of neurons [\cite]. The recorded neurons
-are usually picked out according to an unknown process, and from their
-observation we expect to learn something about all other neurons in the
-same brain area. That is, they are assumed to be a
+are usually picked out from a specific area according to an unknown process, and from their
+observation in relation to stimuli or behavioral task, we strive for
+understanding the mechanism behind neuronal coding. That is, they are assumed to be a
 \enquote{representative random sample} of the neurons in that area.
@@
 
-Using the pairwise \me\ models many studies strived for answering 
+Using the pairwise \me\ models many studies investigated the
  questions such as, which level of interaction (pairwise or higher-order) between neurons

### 2017-05-16 maxNt.tex
worked on Intro

@@
 record from large population of neurons [\cite]. The recorded neurons
-are usually picked out from a specific area according to an unknown process, and from their
+are usually picked out from a specific brain area according to an unknown process, and from their
 observation in relation to stimuli or behavioral task, we strive for

### 2017-05-16 maxNt.tex
worked on Intro

@@
 
-Huge amount of parallel recordings from hundreds of neurons,
-available today, is accompanied by the statistical challenges
+Huge amounts of parallel recordings from hundreds of neurons,
+available today, are accompanied by the statistical challenges
 to be employed for shedding light on the mechanism behind neuronal coding. 

### 2017-05-16 maxNt.tex
fix green todos

--- 2017-05-16 maxNt.tex
@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-15T22:57:55+0200
+%% Last-Updated: 2017-05-16T01:37:28+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}
+  \mynote{Better title?}
   ***%  The abstract must be limited to one paragraph.
@@
 
-%\section{Implicit assumptions in the maximum-entropy method}
-\section{Introduction}
-
-Recent advances in electrophysiological techniques allow to
-record from large population of neurons [\cite]. The recorded neurons
-are usually picked out from a specific area according to an unknown process, and from their
-observation in relation to stimuli or behavioral task, we strive for
-understanding the mechanism behind neuronal coding. That is, they are assumed to be a
-\enquote{representative random sample} of the neurons in that area.
-
-Huge amount of parallel recordings from hundreds of neurons,
-available today, is accompanied by the statistical challenges
-to be employed for shedding light on the mechanism behind neuronal coding. 
-The principle of \me\ was proposed as a way to overcome these challenges by
-constructing a full description of the 
-neuronal firing patterns. The \me\ method employs only first and second
-order statistics of the representative random sample to construct 
-a \enquote{maximally noncommittal} \cite{jaynes1963} 
-probability distribution of the firing patterns [\cite].
-
-Using the pairwise \me\ models many studies investigated the
- questions such as, which level of interaction (pairwise or higher-order) between neurons
- is suffiecient to explain the observed experimental data [\cite]? Or how collective behavior
- of neurons can be related to stimuli or behavioral tasks [\cite]? add more **
-
-Here we show that there is a
-contradiction in applying \me\ to the representative random sample
-of neurons to generate a maximally noncommital distribution for
-that sample.
-
-The contradiction is this. Assume that our sample of neurons is
-representative of some population. If we assign a \me\ distribution to the
-sample, then we cannot assign a \me\ distribution to the full population.
-Vice versa, If we assign a \me\ distribution to the full population, then
-we cannot assign a \me\ distribution to the sample. This impossibility
-holds at least for \me\ distribution with moment constraints, and even if
-the orders of the moments constrained in the sample and in the full
-population are different.
-
-In other words, if our sample is a \enquote{random representative} of some
-population, then we must choose which has a \enquote{maximally
-  noncommittal} distribution: the sample or the population. We can't choose
-both. It goes without saying that if one is \enquote{maximally
-  noncommittal}, the other must be somehow \enquote{committal}. Which
-choice is most meaningful?
-
-In the rest of the paper we will mathematically show the contradiction
-above and generate a \me\ distribution for the full population rather than
-the sample. We will also present some probability relations relevant to
-random sampling. These relations are well-known in survey sampling and in
-the pedagogic problem of drawing from an urn without replacement, yet they
-are somewhat hard to find explicitly written in the neuroscientific
-literature.
+\section{Implicit assumptions in the maximum-entropy method}
+
+The \me\ method is used in neuroscience blahblah***
 
@@
 \binom{N}{N \yRf}\exp\biggl[
-\yL_2\, \frac{\yRf\,(N \yRf-1)}{N-1} + \yL_1\, \yRf\biggr]
+\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr]
 \end{equation}
@@
   \binom{n}{n\yrs} \binom{N-n}{N\yRf-n\yrs} \exp\biggl[
-\yL_2\, \frac{\yRf\,(N \yRf-1)}{N-1} + \yL_1\, \yRf\biggr].
+\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr].
 \end{equation}
@@
   \binom{n}{n\yrs}
-  \exp\biggl[\yl_2\, \frac{\yrs\,(n \yrs-1)}{n-1} + \yl_1\, \yrs\biggr]
+  \exp\biggl[\yl_2 \yrs \,\frac{n \yrs-1}{n-1} + \yl_1\, \yrs\biggr]
 \end{equation}
@@
 
-PGLPM thanks Mari \amp\ Miri for continuous encouragement and affection,
+To be added after review.
+\iffalse PGLPM thanks Mari \amp\ Miri for continuous encouragement and affection,
 Buster Keaton for filling life with awe and inspiration, and the developers
@@
 PhilSci, Hal archives, Python, Inkscape, Sci-Hub for making a free and
-unfiltered scientific exchange possible.
+unfiltered scientific exchange possible.\fi
 

### 2017-05-16 maxNt.tex
comments on intro

--- 2017-05-16 maxNt.tex
@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-16T01:37:28+0200
+%% Last-Updated: 2017-05-16T16:44:08+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Implicit assumptions in the maximum-entropy method}
-
-The \me\ method is used in neuroscience blahblah***
+%\section{Implicit assumptions in the maximum-entropy method}
+\section{Introduction}
+
+Recent advances in electrophysiological techniques allow to record from
+large\mynote{\enquote{from large} or just \enquote{large}? But is the
+  largeness relevant in our paper?} population of neurons [\cite]. The recorded
+neurons are usually picked out from a specific brain area according to an
+unknown process, and from their observation in relation to stimuli or
+behavioral task, we strive for understanding the mechanism behind neuronal
+coding\mynote{LM still don't know what people mean by this, can we find an
+  unambiguous term or rephrase?}. That is, they are assumed to be a
+\enquote{representative random sample} of the neurons in that
+area.\mynote{This assumption says that from the sample we can learn about
+  the population. The relation of population activity with stimuli or
+  behaviour is a separate problem, so \enquote{this is} doesn't really fit...}
+
+Huge amounts of parallel recordings from hundreds of neurons, available
+today, are accompanied by the statistical challenges to be employed for
+shedding light on the mechanism behind neuronal coding. The principle of
+\me\ was proposed as a way to overcome these challenges by constructing a
+full description of the neuronal firing patterns. The \me\ method employs
+only first and second order statistics of the representative random sample
+to construct a \enquote{maximally noncommittal} \cite{jaynes1963}
+probability distribution of the firing patterns [\cite]. \mynote{LM:
+  \enquote{noncommittal} doesn't mean anything to me to be honest, even if
+  Jaynes used that term. I would like to openly say this, backed up by
+  references, as done some paragraphs below.\\It seems they have used \me\
+  for other purposes too, sometimes unclear ones. And with all kinds of
+  constraints. So we shouldn't be too specific here in relation to purpose
+  and constraints. I'd rather give references and the readers can form
+  their own ideas about its use.}
+
+Using the pairwise \me\ models many studies investigated the questions such
+as, which level of interaction\mynote{ambiguous word} (pairwise or higher-order)
+between neurons is suffiecient to explain the observed experimental data
+[\cite]? Or how collective behavior of neurons can be related to stimuli or
+behavioral tasks [\cite]? add more **
+
+Here we show that there is a quandary in applying \me\ to a representative
+random sample of neurons to generate a maximally noncommital distribution
+for that sample.
+
+The quandary is this. Assume that our sample of neurons is
+representative of some population. If we assign a \me\ distribution to the
+sample, then we cannot assign a \me\ distribution to the full population.
+Vice versa, If we assign a \me\ distribution to the full population, then
+we cannot assign a \me\ distribution to the sample. This impossibility
+holds at least for \me\ distribution with moment constraints, and even if
+the orders of the moments constrained in the sample and in the full
+population are different.
+
+In other words, if our sample is a \enquote{random representative} of some
+population, then we must choose which has a \enquote{maximally
+  noncommittal} distribution: the sample or the population. We can't choose
+both. It goes without saying that if one is \enquote{maximally
+  noncommittal}, the other must be somehow \enquote{committal}. Which
+choice is most meaningful?
+
+In the rest of the paper we will mathematically show the contradiction
+above and generate a \me\ distribution for the full population rather than
+the sample. We will also present some probability relations relevant to
+random sampling. These relations are well-known in survey sampling and in
+the pedagogic problem of drawing from an urn without replacement, yet they
+are somewhat hard to find explicitly written in the neuroscientific
+literature.
 

### 2017-05-16 maxNt.tex
comments on intro

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-16T16:44:08+0200
+%% Last-Updated: 2017-05-17T00:18:49+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \centering
-\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_100.pdf}%
+\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
 \caption{Plots of $\p(\yxs = \yrs)$ constructed by \me\ at the population
@@
   \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5\,000$,
-  $n=100$, $\expe{\yxs}=0.45$, $\expe{\yxxs}=0.35$, $\yL_1 = -13\,321.9$,
-  $\yL_2 = 13\,321.5$, $\yl_1=-269.3$, $\yl_2 = 268.9$.}
+  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5000$,
+  $n=200$, $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$.}
 \label{fig:diff_maxent_pop_sample}
@@
 \centering
-\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_both.pdf}
+\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}
 % 
@@
   \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5\,000$,
-  $n=200$, $\expe{\yxs}=0.045$, $\expe{\yxxs}=0.0025$,
-  $\yL_1 = -16\,834.0$, $\yL_2 = 16\,825.7$, $\yl_1=-673.5$,
-  $\yl_2 = 665.1$.}
+  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5000$,
+  $n=200$, $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$,
+  $\expe{\sav{\yx\yx\yx}}=1.48\times10^{-4}$,
+  $\expe{\sav{\yx\yx\yx\yx}}=8.81\times 10^{-6}$.}
 \label{fig:diff_maxent_pop_sample_realdata}
@@
 $\p(\yxs = \yrs \| \yHb)$ are different.
-Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for $N=5\,000$,
+Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for $N=5000$,
 $n=100$, $c_1=0.45$, $c_2=0.35$. The two distributions have slightly
@@
 more peaked. Another example is shown in
-\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for $N=5\,000$, $n=200$,
+\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for $N=5000$, $n=200$,
 and neurobiologically more realistic constraints $c_1=0.045$, $c_2=0.0025$
@@
   the same in either case: $\expe{\yxs \|\yH}=0.045$,
-  $\expe{\yxxs \|\yH}=0.0025$, and $N=5\,000$, $n=200$.}
+  $\expe{\yxxs \|\yH}=0.0025$, and $N=5000$, $n=200$.}
 \label{fig:diff_maxent_pop_sample_realdata}
@@
 learning about an event may tell us something about a related event. For
-example, consider 1\,000 tosses of a particular coin and assume a uniform
-distribution over the possible $2^{1\,000}$ outcomes. If we learn that the
+example, consider 1000 tosses of a particular coin and assume a uniform
+distribution over the possible $2^{1000}$ outcomes. If we learn that the
 first 999 tosses yielded all \enquote{heads}, the probability calculus
-tells us that the probability for the 1\,000th toss is still 50\%/50\%. It
+tells us that the probability for the 1000th toss is still 50\%/50\%. It
 is a consequence of our choice of a uniform distribution: we have

### 2017-05-17 maxNt.tex
new material on projection

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T00:18:49+0200
+%% Last-Updated: 2017-05-17T10:31:16+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \av{\yX}  &\defd \tfrac{1}{N} (X_1 + X_2 + \dotsb + X_N),
-  &
-  \sav{\yx}  &\defd \tfrac{1}{n} (x_1 + x_2 + \dotsb + x_n),
+%  &
+%  \sav{\yx}  &\defd \tfrac{1}{n} (x_1 + x_2 + \dotsb + x_n),
     \\[2\jot]
@@
 \tbinom{N}{2}^{-1} (X_1 X_2 + X_1 X_3  + \dotsb +  X_{N-1} X_N),
-  &
-    \sav{\yx \yx} &\defd
-\tbinom{n}{2}^{-1} (x_1 x_2 + x_1 x_3  + \dotsb +  x_{n-1} x_n),
+%  &
+%    \sav{\yx \yx} &\defd\tbinom{n}{2}^{-1} (x_1 x_2 + x_1 x_3  + \dotsb +  x_{n-1} x_n),
 \\[2\jot]
@@
                    \dotsb + X_{N-2} X_{N-1}X_N),
-  &
-    \sav{\yx\yx\yx} &\defd
-                      \tbinom{n}{3}^{-1} (x_1 x_2 x_3 + %x_1 x_2 x_4
-                       \dotsb + x_{n-2} x_{n-1} x_n),
+%  &
+%    \sav{\yx\yx\yx} &\defd
+%                      \tbinom{n}{3}^{-1} (x_1 x_2 x_3 + %x_1 x_2 x_4
+%                       \dotsb + x_{n-2} x_{n-1} x_n),
 \end{align}
-and so on; analogously for the quantities like $\yRf$ and $\sav{\yr\yr}$.
+and so on
 These formulae say that $\yXf$ is the fraction of active neurons, $\yXXf$
@@
   \av{\underbrace{\yX\dotsm\yX}_{\text{$m$ factors}}}
-  = \binom{N}{m}^{-1}\binom{N\yXf}{m}
-\end{equation}
-and similarly for $\yx$, $\yR$, $\yr$.
+  = \binom{N}{m}^{-1}\binom{N\yXf}{m}.
+\end{equation}
+Analogous formulae  hold for quantities like $\yx$, $\yRf$, $\yrs$.
 
@@
 \end{gather}
+
 The latter conditional probability is a hypergeometric distribution,
@@
 \cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
+This conditional probability acts a
 
@@
 
+\begin{figure}[!t]
+\centering
+%\includegraphics[width=0.99\linewidth]{pop_sample_projection.pdf}\\
+\includegraphics[width=0.99\linewidth]{pop_sample_projection2.pdf}%
+\caption{hyperg}
+\label{fig:hypergeom_proj}
+\end{figure}%hypergeometric_identity.nb
+% \end{wrapfigure}
 Our initial knowledge (ignorance) $\yH$ has two important properties

### 2017-05-17 maxNt.tex
new material on projection

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T10:31:16+0200
+%% Last-Updated: 2017-05-17T12:40:23+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\newcommand*{\ypp}{\varPi}
+
 \newcommand*{\prop}[1]{`#1'}
@@
   \p(\yxs = \yrs \|\yXf=\yRf, \yH)=
-  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}.
+  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}
+  \defs \ypp(\yrs\|\yRf).
 \end{gather}
@@
 \cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-This conditional probability acts a
-
-The symmetry present in the probabilities above is not a physical property
-of the neuronal population. It only expresses the symmetry of our initial
-ignorance about the population, and does not imply any sort of physical
-similarity between the neurons. Subsequent observations may in fact break
-this symmetry. The initial symmetry should intuitively also apply to the
-sample, and this is indeed the case: the probability for the state of the
-sample~\eqref{eq:subpop_average} automatically satisfies the representation
-theorem~\eqref{eq:joint_plaus_N_homog} as well.
+
+The conditional probability $\ypp(\yrs,\yRf)$ connects the spaces of the
+sample variable $\yXf$ and of the population variable $\yxs$. It is a
+projector $\ypp$ of probabilities $p$ for $\yXf$ onto marginal
+probabilities for $\yxs$:
+\begin{equation}
+  \label{eq:projection_P}
+  \ypp\colon p(\yXf=\dotv) \to p_*(\yxs=\dotv)
+  \defd \sum_{N\yRf=0}^N \ypp(\dotv\|\yRf) p(\yRf).
+\end{equation}
+It is also a pull-back $\ypp^*$ of functions $f$ of $\yxs$ to functions of
+$\yXf$:
+\begin{equation}
+  \label{eq:pullback_P}
+  \ypp^*\colon f(\yxs=\dotv) \to f^*(\yXf=\dotv) \defd \sum_{N\yrs=0}^n f(\yrs)\ypp(\yrs\|\dotv),
+\end{equation}
+with a caveat: if $f$ is a physical quantity and doesn't depend on our
+knowledge, $f_*$ is \emph{not} a physical quantity but a conditional
+expectation, $f^*(\yRf) \equiv \expe{f(\yxs)\|\yXf=\yRf,\yH}$, and depends
+on our knowledge.
 
@@
 \includegraphics[width=0.99\linewidth]{pop_sample_projection2.pdf}%
-\caption{hyperg}
+\caption{Log-plot of the hypergeometric distribution
+  $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
+  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1} \defs
+  \ypp(\yrs\|\yRf)$ for $N=5000$, $n=200$. (Band artifacts may appear in the
+  colourbar depending on your \textsc{pdf} viewer.)}
 \label{fig:hypergeom_proj}
 \end{figure}%hypergeometric_identity.nb
-% \end{wrapfigure}
-Our initial knowledge (ignorance) $\yH$ has two important properties
-derivable from the formulae above.
-
-First, our initial expectations for the averages of activity products
-\emph{are the same for the sample and for the full population}:
+A look at a plot of the hypergeometric distribution $\ypp(\yrs\|\yXf)$, for
+example \fig~\ref{fig:hypergeom_proj}, reveals that the maps $\ypp$ and
+$\ypp^*$ are \enquote{fuzzy identities} between the $\yXf$ space
+$\set{0,\dotsc,N}$ and $\yxs$-space $\set{0,\dotsc,n}$. When $n=N$ they are
+exact identities. We thus have very roughly that
+\begin{equation}
+  \label{eq:roughly_equal_nN}
+  \p(\yxs=a) \sim \p(\yXf=a),\qquad
+\expeb{f(\yxs)} \sim \expeb{f(\yXf)}.
+\end{equation}
+These formulae express the fact that \emph{the sample is representative of
+  the population}, and vice versa. Our initial knowledge $\yH$, which fully
+determines $\ypp$, thus assumes that by observing a sample we can make
+predictions about the full population, and thence about new samples.
+Upon observation of a sample average, say $\yxxs=a$, the updated
+expectations for such average in the population and in any new sample will
+usually be shifted towards the observed value, as follows from Bayes's
+theorem and the formulae above.
+
+Note that the approximate formulae~\eqref{eq:roughly_equal_nN} say more
+than the limits $\p(\yxs=a) \to \p(\yXf=a)$,
+$\expeb{f(\yxs)} \to \expeb{f(\yXf)}$ as $n\to N$. These limits are
+trivially valid because as $n\to N$ the sample becomes the full population.
+For conditional probabilities $\p(\yxs = \yrs \|\yXf=\yRf)$ very different
+from the fuzzy identity $\ypp(\yrs\|\yRf)$ (see
+\fig~\ref{fig:hypergeom_proj}) these limits would hold, but the
+probabilities and expectations from a small sample could wildly differ from those
+from the population.
+
+\bigskip***
+
+with the property that
+\begin{equation}
+  \label{eq:expectation_pullback}
+  \expeb{f(\yxs) \|\yH} = \expeb{f^*(\yXf)\|\yH},
+  \qquad\text{that is,}\qquad
+  \sum_{n\yrs=0}^n f(\yrs)\,p_*(\yxs=\yrs) = \sum_{N\yRf=0}^N f^*(\yRf)\,p(\yXf=\yRf).
+\end{equation}
+
+A special case of the pull-back above is the function
+$f(\yxs) \defd \sav{\yx\dots\yx}=\binom{\yxs}{m}/\binom{n}{m}$, for which
+\begin{equation}
+  \label{eq:pullback_m_expectations}
+  \ypp^*\colon \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}
+  \mapsto \bigl( \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}\bigr)^*
+  \equiv
+  \av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}};
+\end{equation}
+thus, the averages of activity products \emph{are the same for the sample
+  and for the full population}:
 \begin{equation}
@@
 \end{equation}
-Raw moments like $\expe{\yxs^m}$, which can be written in terms of the
-product expectations above via \eqn~\eqref{eq:expe_products}, similarly
-satisfy $\expe{\yxs^m} \to \expe{\yXf^m}$ as $n \to N$.
-
-Second, upon observation of a sample average, say $\yxxs=a$, the updated
-expectations for such average in the population \emph{and in any new
-  sample} will usually be shifted towards the observed value, as follows
-from Bayes's theorem and the formulae above.
-
-These two properties express the fact that the sample is
-\emph{representative} of the population. Our initial knowledge $\yH$ thus
-assumes that by observing a sample we can make predictions about the full
-population and about new samples. Contrast this with an assumption that
-assigns an independent probability to each neuron: in this case, the
-observation of one sample does not change our initial probability about
-another -- we can't learn from observation.
+
+This is true, for example, for the raw moments $\expe{\yxs^m}$ and
+$\expe{\yXf^m}$, which can be written in terms of the product expectations
+via \eqn~\eqref{eq:expe_products}.
+
+The symmetry present in the probabilities above is not a physical property
+of the neuronal population. It only expresses the symmetry of our initial
+ignorance about the population, and does not imply any sort of physical
+similarity between the neurons. Subsequent observations may in fact break
+this symmetry. The initial symmetry should intuitively also apply to the
+sample, and this is indeed the case: the probability for the state of the
+sample~\eqref{eq:subpop_average} automatically satisfies the representation
+theorem~\eqref{eq:joint_plaus_N_homog} as well.
 

### 2017-05-17 maxNt.tex
new material on projection

--- 2017-05-17 maxNt.tex
@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T10:31:16+0200
+%% Last-Updated: 2017-05-17T14:42:01+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \\[2\jot]
+  \label{eq:conditional_hypergeometric}
   \p(\yxs = \yrs \|\yXf=\yRf, \yH)=
-  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}.
+  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}
+  \defs \ypp(\yrs\|\yRf).
 \end{gather}
 
-The latter conditional probability is a hypergeometric distribution,
-typical of \enquote{drawing without replacement} problems. The
-combinatorial proof of the formulae above is in fact the same as for this
-class of problems
+The latter conditional probability is a hypergeometric distribution
+$\ypp(\yrs\|\yRf)$, typical of \enquote{drawing without replacement}
+problems. The combinatorial proof of the formulae above is in fact the same
+as for this class of problems
 \cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-This conditional probability acts a
+
+The conditional probability $\ypp(\yrs,\yRf)$ connects the spaces of the
+sample average $\yXf\in\set{0,\dotsc,N}$ and of the population average
+$\yxs\in\set{0,\dotsc,n}$. It is a projector $\ypp$ of any probability $p$
+for $\yXf$ onto a marginal probability $p_*$ for $\yxs$:
+\begin{equation}
+  \label{eq:projection_P}
+  % \ypp\colon p(\yXf=\dotv) \to
+  p_*(\yxs=\yrs)
+  = \sum_{N\yRf=0}^N \ypp(\yrs\|\yRf) p(\yXf=\yRf).
+\end{equation}
+Therefore it also a pulls back expectations of functions $f$ of the sample
+average $\yxs$ to expectations of functions $f^*$ the population average
+$\yXf$:
+\begin{equation}
+  \label{eq:pullback_P}
+%  \ypp^*\colon f(\yxs=\dotv) \to f^*(\yXf=\dotv) \defd \sum_{N\yrs=0}^n
+%  f(\yrs)\ypp(\yrs\|\dotv),
+  \begin{gathered}
+    f^*(\yXf) \defd \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yXf),
+    \\
+   \expeb{f(\yxs)}= \expeb{f^*(\yXf)}
+    =
+    \sum_{n\yrs=0}^n f(\yrs)\,\p(\yxs=\yrs \| \yH)
+    =
+    \sum_{N\yRf=0}^N f^*(\yRf)\,\p(\yXf=\yRf \| \yH).
+  \end{gathered}
+\end{equation}
+
+\begin{figure}[!t]
+\centering
+%\includegraphics[width=0.75\linewidth]{pop_sample_projection.pdf}\\
+\includegraphics[width=0.75\linewidth]{pop_sample_projection2.pdf}%
+\caption{Log-plot of the hypergeometric distribution
+  $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
+  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1} \defs
+  \ypp(\yrs\|\yRf)$ for $N=5000$, $n=200$. (Band artifacts may appear in the
+  colourbar depending on your \textsc{pdf} viewer.)}
+\label{fig:hypergeom_proj}
+\end{figure}%hypergeometric_identity.nb
+A look at a plot of the hypergeometric distribution $\ypp(\yrs\|\yRf)$, for
+example \fig~\ref{fig:hypergeom_proj}, reveals that it is a sort of
+\enquote{fuzzy identity matrix} between the $\yXf$-space $\set{0,\dotsc,N}$
+and $\yxs$-space $\set{0,\dotsc,n}$. When $n=N$ it is the identity matrix.
+We thus have that, very roughly,
+\begin{equation}
+  \label{eq:roughly_equal_nN}
+  \p(\yxs=a) \sim \p(\yXf=a),\qquad
+\expeb{f(\yxs)} \sim \expeb{f(\yXf)}.
+\end{equation}
+These formulae express the fact that \emph{our uncertainty about the sample
+  is representative of our uncertainty about the population and about other
+  samples}, and vice versa. Upon observation of a sample average, say
+$\yxxs=a$, the updated expectations for such average in the population and
+in any new sample will usually be shifted towards the observed value, as
+follows from Bayes's theorem and the formulae above.
+
+Note that the approximate formulae~\eqref{eq:roughly_equal_nN} say more
+than the limits $\p(\yxs=a) \to \p(\yXf=a)$,
+$\expeb{f(\yxs)} \to \expeb{f(\yXf)}$ as $n\to N$. These limits are
+trivially valid because the sample becomes the full population as $n\to N$.
+In particular, these limits hold even in cases where the conditional
+probability $\p(\yxs = \yrs \|\yXf=\yRf)$ is not a fuzzy identity and our
+uncertainties about sample and about population can differ wildly.
 
@@
 of the neuronal population. It only expresses the symmetry of our initial
-ignorance about the population, and does not imply any sort of physical
+uncertainty about the population, and does not imply any sort of physical
 similarity between the neurons. Subsequent observations may in fact break
@@
 
-\begin{figure}[!t]
-\centering
-%\includegraphics[width=0.99\linewidth]{pop_sample_projection.pdf}\\
-\includegraphics[width=0.99\linewidth]{pop_sample_projection2.pdf}%
-\caption{hyperg}
-\label{fig:hypergeom_proj}
-\end{figure}%hypergeometric_identity.nb
-% \end{wrapfigure}
-Our initial knowledge (ignorance) $\yH$ has two important properties
-derivable from the formulae above.
-
-First, our initial expectations for the averages of activity products
-\emph{are the same for the sample and for the full population}:
+\bigskip
+
+% The projection~\eqref{eq:projection_P} and pull-back~\eqref{eq:pullback_P}
+% have the property
+% \begin{equation}
+%   \label{eq:expectation_pullback}
+%   \expeb{f(\yxs) \|\yH} = \expeb{f^*(\yXf)\|\yH}
+%   \qquad\text{or}\qquad
+%   \sum_{n\yrs=0}^n f(\yrs)\,p_*(\yxs=\yrs) = \sum_{N\yRf=0}^N f^*(\yRf)\,p(\yXf=\yRf).
+% \end{equation}
+For functions representing averaged products,
+$f(\yxs) \defd \sav{\yx\dots\yx}=\binom{\yxs}{m}/\binom{n}{m}$,
+formulae~\eqref{eq:pullback_P} have the useful form
 \begin{equation}
   \label{eq:expe_products}
-\expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
-=
-\expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
-%\\[2\jot]
-%{}
-=
-\binom{n}{m}^{-1}
-%\frac{(n-m)!}{n!}
+%  \label{eq:pullback_m_expectations}
+%  \ypp^*\colon \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}
+%  \mapsto
+  \begin{gathered}
+    \bigl( \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}\bigr)^* =
+    \av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}},
+    \\[2\jot]
+    % \expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH} =
+    % \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
+    \expe{\yx \dotsm \yx \| \yH} =
+    \expe{\yX \dotsm \yX \| \yH}
+    % \\[2\jot]
+    % {}
+    = \binom{n}{m}^{-1}
+    % \frac{(n-m)!}{n!}
     \sum_{n \yrs=0}^n %m!\,
@@
     % \sum_{N \yRf=0}^N %m!\,
-% \binom{N-m}{N \yRf-m}
-% \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
-=
-\binom{N}{m}^{-1}
-%\frac{(N-m)!}{N!}
+    % \binom{N-m}{N \yRf-m} \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
+    = \binom{N}{m}^{-1}
+    % \frac{(N-m)!}{N!}
     \sum_{N \yRf=0}^N %m!\,
     \binom{N \yRf}{m}\, \p(\yXf=\yRf \| \yH).
-\end{equation}
-Raw moments like $\expe{\yxs^m}$, which can be written in terms of the
-product expectations above via \eqn~\eqref{eq:expe_products}, similarly
-satisfy $\expe{\yxs^m} \to \expe{\yXf^m}$ as $n \to N$.
-
-Second, upon observation of a sample average, say $\yxxs=a$, the updated
-expectations for such average in the population \emph{and in any new
-  sample} will usually be shifted towards the observed value, as follows
-from Bayes's theorem and the formulae above.
-
-These two properties express the fact that the sample is
-\emph{representative} of the population. Our initial knowledge $\yH$ thus
-assumes that by observing a sample we can make predictions about the full
-population and about new samples. Contrast this with an assumption that
-assigns an independent probability to each neuron: in this case, the
-observation of one sample does not change our initial probability about
-another -- we can't learn from observation.
+  \end{gathered}
+\end{equation}
+The proof uses the expression for the $m$th factorial moment of the
+hypergeometric distribution \cite{potts1953}. Thus, the averages of
+activity products \emph{are the same for the sample and for the full
+  population}. Similar relations can be found for the raw moments
+$\expe{\yxs^m}$ and $\expe{\yXf^m}$, which can be written in terms of the
+product expectations via \eqn~\eqref{eq:products_intermsof_average}.
 
@@
 
-Formulae \eqref{eq:joint_plaus_N_homog}--\eqref{eq:expe_products} are
+Formulae
+\eqref{eq:joint_plaus_N_homog}--~\eqref{eq:conditional_hypergeometric} are
 constraints on our initial probability assignment, but do not determine it
@@
 
-We may want to specify the probability by enforcing several of its sample
-expectations to have specific values, for example $\expe{\yxs}=c_1$,
-$\expe{\yxxs}=c_2$. This is still an underdetermined linear problem:
- several distributions can have the same desired expectations, as
-clear from \eqns~\eqref{eq:expe_products}.
+We may want to specify the probability by enforcing the sample expectations
+of several functions to have specific values, for example
+$\expe{\yxs}=c_1$, $\expe{\yxxs}=c_2$. This is still an underdetermined
+linear problem: several distributions can have the same desired
+expectations, as clear from \eqns~\eqref{eq:expe_products}.
 
@@
 those that have the desired expectations. But we are in a quandary:
-formulae~\eqref{eq:expe_products} allow us to apply the method to find the
+formulae~\eqref{eq:pullback_P} allow us to apply the method to find the
 probability of the population $\p(\yXf=\yRf \| \yH)$, or of the sample

### 2017-05-17 maxNt.tex
new material on projection

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T14:42:01+0200
+%% Last-Updated: 2017-05-17T15:27:36+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\newcommand*{\slor}{\mathop{\textstyle\bigvee}}
-%\newcommand*{\T}{^\intercal}%transpose
+\newcommand*{\T}{^\intercal}%transpose
 \newcommand*{\E}{\mathrm{E}}
@@
 sample average $\yXf\in\set{0,\dotsc,N}$ and of the population average
-$\yxs\in\set{0,\dotsc,n}$. It is a projector $\ypp$ of any probability $p$
-for $\yXf$ onto a marginal probability $p_*$ for $\yxs$:
+$\yxs\in\set{0,\dotsc,n}$. It is a coarsening projector of any probability
+$p$ for $\yXf$ onto a marginal probability $p_*$ for $\yxs$:
 \begin{equation}
@@
 and $\yxs$-space $\set{0,\dotsc,n}$. When $n=N$ it is the identity matrix.
-We thus have that, very roughly,
+We thus have that
 \begin{equation}
   \label{eq:roughly_equal_nN}
-  \p(\yxs=a) \sim \p(\yXf=a),\qquad
-\expeb{f(\yxs)} \sim \expeb{f(\yXf)}.
-\end{equation}
-These formulae express the fact that \emph{our uncertainty about the sample
-  is representative of our uncertainty about the population and about other
-  samples}, and vice versa. Upon observation of a sample average, say
-$\yxxs=a$, the updated expectations for such average in the population and
-in any new sample will usually be shifted towards the observed value, as
-follows from Bayes's theorem and the formulae above.
-
-Note that the approximate formulae~\eqref{eq:roughly_equal_nN} say more
-than the limits $\p(\yxs=a) \to \p(\yXf=a)$,
-$\expeb{f(\yxs)} \to \expeb{f(\yXf)}$ as $n\to N$. These limits are
-trivially valid because the sample becomes the full population as $n\to N$.
-In particular, these limits hold even in cases where the conditional
-probability $\p(\yxs = \yrs \|\yXf=\yRf)$ is not a fuzzy identity and our
-uncertainties about sample and about population can differ wildly.
+  \p(\yxs=a) \approx \p(\yXf=a),\qquad
+\expeb{f(\yxs)} \approx \expeb{f(\yXf)}.
+\end{equation}
+These are only very approximate equalities; they may miss important
+features of the two probability distributions. If the distribution for the
+population average $\yXf$ is bimodal, for example, the bimodality can be
+lost in the distribution for the sample average $\yxs$, owing to the
+coarsening effect of $\ypp(\yrs\|\yRf)$.
+
+The approximate equalities above express the fact that \emph{our
+  uncertainty about the sample is representative of our uncertainty about
+  the population and about other samples}, and vice versa. Upon observation
+of a sample average, say $\yxxs=a$, the updated expectations for such
+average in the population and in any new sample will usually be shifted
+towards the observed value, as follows from Bayes's theorem and the
+formulae above.
+
+Note that formulae~\eqref{eq:roughly_equal_nN} say more than the limits
+$\p(\yxs=a) \to \p(\yXf=a)$, $\expeb{f(\yxs)} \to \expeb{f(\yXf)}$ as
+$n\to N$. These limits are trivially valid because the sample becomes the
+full population as $n\to N$. In particular, these limits hold even in cases
+where the conditional probability $\p(\yxs = \yrs \|\yXf=\yRf)$ is not a
+fuzzy identity and our uncertainties about sample and about population can
+differ wildly.
 
@@
     % \frac{(n-m)!}{n!}
-    \sum_{n \yrs=0}^n %m!\,
+    \sum_{\mathclap{n \yrs=0}}^n %m!\,
     \binom{n \yrs}{m}\, \p(\yxs=\yrs \| \yH)
@@
     % \frac{(N-m)!}{N!}
-    \sum_{N \yRf=0}^N %m!\,
+    \sum_{\mathclap{N \yRf=0}}^N %m!\,
     \binom{N \yRf}{m}\, \p(\yXf=\yRf \| \yH).
@@
 Formulae
-\eqref{eq:joint_plaus_N_homog}--~\eqref{eq:conditional_hypergeometric} are
+\eqref{eq:joint_plaus_N_homog}--\eqref{eq:conditional_hypergeometric} are
 constraints on our initial probability assignment, but do not determine it
@@
 
-The \me\ method is brought into play to solve this indeterminacy. It selects
-one distribution, purported to be \enquote{maximally noncommittal}, among
-those that have the desired expectations. But we are in a quandary:
+The \me\ method is brought into play to solve this indeterminacy. It
+selects one distribution, purported to be \enquote{maximally noncommittal},
+among those that have the desired expectations. But we are in a quandary:
 formulae~\eqref{eq:pullback_P} allow us to apply the method to find the
@@
   inequivalent}. They lead to numerically different $\p(\yxs=\yrs \| \yH)$.
+
+Suppose we want to constrain the sample expectations of a vector function
+$f$ to the vector values $c$, that is, $\expeb{f(\yxs)}=c$. Application of
+\me\ at the population level, denoted by $\yHa$, gives
+\begin{equation}
+  \label{eq:app_maxent_pop}
+  \p(\yXf = \yRf \| \yHa)  = \yK\,
+  \binom{N}{N \yRf}\exp\Bigl[\yL\T
+  \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\Bigr],
+\end{equation}
+and then by marginalization with \eqn~\eqref{eq:marginal}
+\begin{equation}
+  \label{eq:app_maxent_pop_marg}
+  \p(\yxs = \yrs \| \yHa)  = \yK\, 
+  \sum_{N\yRf=0}^N \ypp(\yrs \|\yRf)
+\,
+  \binom{N}{N \yRf}\exp\Bigl[\yL\T
+  \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\Bigr],
+\end{equation}
+where $\yK$ is a normalization constant and $\yL\T$ are Lagrange
+multipliers such that
+\begin{equation}
+  \label{eq:app_maxent_pop_constraints}
+c = \yK\sum_{n\yrs=0}^n \sum_{N\yRf=0}^N f(\yrs) \ypp(\yrs \|\yRf)
+\,
+  \binom{N}{N \yRf}\exp\Bigl[\yL\T
+  \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\Bigr].
+\end{equation}
+Application of \me\ at the sample level, denoted by $\yHb$, gives
+\begin{equation}
+  \label{eq:app_maxent_sample}
+  \p(\yxs=\yrs \|\yHb)
+  =\yk\,
+  \binom{n}{n\yrs}
+  \exp[\yl\T f(\yrs)]
+\end{equation}
+where $\yk$ is a normalization constant and $\yl\T$ are Lagrange
+multipliers such that
+\begin{equation}
+  \label{eq:app_maxent_sample_constraints}
+c = \yk\sum_{n\yrs=0}^n  f(\yrs) \binom{n}{n\yrs}
+  \exp[\yl\T f(\yrs)].
+\end{equation}
+
+The probabilities for the sample average obtained from application at the
+population level~\eqref{eq:app_maxent_pop_marg} and at the sample level
+\eqref{eq:app_maxent_sample} will be approximately equal, by our previous
+observation~\eqref{eq:roughly_equal_nN} and also by the fact that they must
+satisfy the same expectations for $f$. But the latter may miss important
+features present in the former, like additional modes or particular tail
+behaviour.
+
+We show two examples of this discrepancy in
+\figs~\ref{fig:diff_maxent_pop_sample}
+and~\ref{fig:diff_maxent_pop_sample_realdata}. In the former the
+constraints are $\expe{\yxs} = c_1$ and
+$\expe{\yxxs} = c_2$, with
+
 Here is an example.

### 2017-05-17 maxNt.tex
new material on projection

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T15:27:36+0200
+%% Last-Updated: 2017-05-17T15:41:05+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We show two examples of this discrepancy in
-\figs~\ref{fig:diff_maxent_pop_sample}
-and~\ref{fig:diff_maxent_pop_sample_realdata}. In the former the
-constraints are $\expe{\yxs} = c_1$ and
-$\expe{\yxxs} = c_2$, with
-
-Here is an example.
-
-Suppose our desired constraints are $\expe{\yxs } = c_1$ and
-$\expe{\yxxs} = c_2$, the expectations being given by the fourth term of
-\eqn~\eqref{eq:expe_products}. Applying \me\ to the distribution of the
-average of the full population we find
-\begin{equation}
-  \label{eq:full_pop_av_example_ME}
-  \p(\yXf = \yRf \| \yHa)  = \yL\,
-\binom{N}{N \yRf}\exp\biggl[
-\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr]
-\end{equation}
-with such parameters $\yL$, $\yL_1$, $\yL_2$ as to satisfy normalization
-and constraints. We call this state of knowledge $\yHa$. Marginalization by
-\eqn~\eqref{eq:marginal} then gives
-\begin{equation}
-  \label{eq:full_pop_av_example_ME_marg}
-  \p(\yxs = \yrs \| \yHa)  = \yL
-  \sum_{N\yRf=0}^N
-  \binom{n}{n\yrs} \binom{N-n}{N\yRf-n\yrs} \exp\biggl[
-\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr].
-\end{equation}
-Applying \me\ to the distribution of the average of the sample, using the
-third term of \eqn~\eqref{eq:expe_products}, we find
-\begin{equation}
-  \label{eq:sample_av_example_ME}
-  \p(\yxs = \yrs \| \yHb)  = \yl\,
-  \binom{n}{n\yrs}
-  \exp\biggl[\yl_2 \yrs \,\frac{n \yrs-1}{n-1} + \yl_1\, \yrs\biggr]
-\end{equation}
-with such parameters $\yl$, $\yl_1$, $\yl_2$ as to satisfy the constraints.
-We call this state of knowledge $\yHb$.
 %\setlength{\intextsep}{0.5ex}%
@@
 \centering
-\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
+\includegraphics[width=0.75\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
 \caption{Plots of $\p(\yxs = \yrs)$ constructed by \me\ at the population
@@
 \centering
-\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}
+\includegraphics[width=0.75\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}
 % 
@@
 \end{figure}%maxent_pop_or_sample.nb
-
-The resulting distributions $\p(\yxs = \yrs \| \yHa)$,
-$\p(\yxs = \yrs \| \yHb)$ are different.
-Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for $N=5000$,
-$n=100$, $c_1=0.45$, $c_2=0.35$. The two distributions have slightly
-different modes and the one obtained from \me\ at the population level is
-more peaked. Another example is shown in
-\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for $N=5000$, $n=200$,
-and neurobiologically more realistic constraints $c_1=0.045$, $c_2=0.0025$
-\cite{rostamietal2016}. The discrepancy is maybe not as large as in the
-previous example, but the tails of the distributions are still very
-different.
+We show two examples of this discrepancy in
+\figs~\ref{fig:diff_maxent_pop_sample}
+and~\ref{fig:diff_maxent_pop_sample_realdata}, for $N=5000$, $n=200$. In
+the first example the constraints are $\expe{\yxs} = c_1$ and
+$\expe{\yxxs} = c_2$, with $c_1=0.0478$ and $c_2=0.00257$. The distribution
+obtained at the sample level is broader than the one obtained at the
+population level; the tails of the two distributions are very different.
+The second example includes two additional constraints
+$\expe{\sav{\yx\yx\yx}}= c_3$, $\expe{\sav{\yx\yx\yx\yx}}=c_4$ with
+$c_3=0.000148$, $c_4=8.81\times 10^{-6}$. The distribution obtained at the
+population level has two modes, replaced by only one in the distribution
+obtained at the sample level; the tails are very different also in this
+case. The constraints used in these examples have neurobiologically
+realistic values \cite{rostamietal2016}.
 
@@
 
+  *******
+
+Here is an example.
+
+Suppose our desired constraints are $\expe{\yxs } = c_1$ and
+$\expe{\yxxs} = c_2$, the expectations being given by the fourth term of
+\eqn~\eqref{eq:expe_products}. Applying \me\ to the distribution of the
+average of the full population we find
+\begin{equation}
+  \label{eq:full_pop_av_example_ME}
+  \p(\yXf = \yRf \| \yHa)  = \yL\,
+\binom{N}{N \yRf}\exp\biggl[
+\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr]
+\end{equation}
+with such parameters $\yL$, $\yL_1$, $\yL_2$ as to satisfy normalization
+and constraints. We call this state of knowledge $\yHa$. Marginalization by
+\eqn~\eqref{eq:marginal} then gives
+\begin{equation}
+  \label{eq:full_pop_av_example_ME_marg}
+  \p(\yxs = \yrs \| \yHa)  = \yL
+  \sum_{N\yRf=0}^N
+  \binom{n}{n\yrs} \binom{N-n}{N\yRf-n\yrs} \exp\biggl[
+\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr].
+\end{equation}
+Applying \me\ to the distribution of the average of the sample, using the
+third term of \eqn~\eqref{eq:expe_products}, we find
+\begin{equation}
+  \label{eq:sample_av_example_ME}
+  \p(\yxs = \yrs \| \yHb)  = \yl\,
+  \binom{n}{n\yrs}
+  \exp\biggl[\yl_2 \yrs \,\frac{n \yrs-1}{n-1} + \yl_1\, \yrs\biggr]
+\end{equation}
+with such parameters $\yl$, $\yl_1$, $\yl_2$ as to satisfy the constraints.
+We call this state of knowledge $\yHb$.
+
+The resulting distributions $\p(\yxs = \yrs \| \yHa)$,
+$\p(\yxs = \yrs \| \yHb)$ are different.
+Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for $N=5000$,
+$n=100$, $c_1=0.45$, $c_2=0.35$. The two distributions have slightly
+different modes and the one obtained from \me\ at the population level is
+more peaked. Another example is shown in
+\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for $N=5000$, $n=200$,
+and neurobiologically more realistic constraints $c_1=0.045$, $c_2=0.0025$
+\cite{rostamietal2016}. The discrepancy is maybe not as large as in the
+previous example, but the tails of the distributions are still very
+different.
+
 \end{document}

### 2017-05-17 maxNt.tex
new material on projection
intro

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T15:41:05+0200
+%% Last-Updated: 2017-05-17T21:23:29+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\section{Implicit assumptions in the maximum-entropy method}
-\section{Introduction}
-
-Recent advances in electrophysiological techniques allow to record from
-large\mynote{\enquote{from large} or just \enquote{large}? But is the
-  largeness relevant in our paper?} population of neurons [\cite]. The recorded
-neurons are usually picked out from a specific brain area according to an
-unknown process, and from their observation in relation to stimuli or
-behavioral task, we strive for understanding the mechanism behind neuronal
-coding\mynote{LM still don't know what people mean by this, can we find an
-  unambiguous term or rephrase?}. That is, they are assumed to be a
-\enquote{representative random sample} of the neurons in that
-area.\mynote{This assumption says that from the sample we can learn about
-  the population. The relation of population activity with stimuli or
-  behaviour is a separate problem, so \enquote{this is} doesn't really fit...}
+\section{Introduction: \me\ and sampling in neuroscience}
+
+% LM: I'd like to avoid common beginnings like "Today blah blah is
+% important" or "Blah blah is interesting/fundamental". They feel more like
+% a mechanical ritual than a real tentative of connecting with the reader.
+This note is mainly addressed to neuroscientists interested in \me\
+methods, but its discussion of \enquote{random sampling} formulae may be
+useful to all neuroscientists involved in understanding the physical and
+dynamical characteristics of brain areas via neuronal recording.
+
+Recent electrophysiological techniques in fact allow experimenters to
+record samples of even a couple hundreds neurons from specific brain areas.
+We don't elaborate on the various important purposes of such recordings
+here, but stress that these sample sizes can be considered \enquote{large},
+because their statistical analysis requires considerable computational
+power.
+
+These computational costs are one, probably not the earliest, of the many
+reasons why \me\ methods have been introduced in neuroscience. It would be
+useful to somehow compress the statistical wealth of large neuronal
+recordings into few quantities, like sample moments for example. This
+compression would also entail interesting physico-biological properties of
+neuronal activity. Maximum-entropy methods make this kind of compression:
+they associate a unique probability distribution with few experimental
+quantities. This distribution is usually said to be \enquote{maximally
+  noncommittal} \cite{jaynes1963}.
+
+We intend the quoted expression only as
+an umbrella term, also because it means very little without further
+clarifications.
+
+The neurons recorded in these kinds of experiments are usually picked out
+according to an unknown process, and from their observation we expect to
+learn something about all other neurons in the same brain area. That is,
+they are assumed to be a \enquote{representative random sample} of the
+neurons in that area. Our discussion hinges on this often just tacitly
+understood assumption.
+
+One of the uses of \me\ methods
+
+number. It is large from large
+population of neurons [\cite]. The recorded neurons are usually picked out
+from a specific brain area according to an unknown process, and from their
+observation in relation to stimuli or behavioral task, we strive for
+understanding the mechanism behind neuronal coding\mynote{LM still don't
+  know what people mean by this, can we find an unambiguous term or
+  rephrase?}. That is, they are assumed to be a \enquote{representative
+  random sample} of the neurons in that area.\mynote{This assumption says
+  that from the sample we can learn about the population. The relation of
+  population activity with stimuli or behaviour is a separate problem, so
+  \enquote{this is} doesn't really fit...}
 
@@
 to denote logical conjunction. Probability notation follows Jaynes
-\citep{jaynes1994_r2003}.
+\citep{jaynes1994_r2003}. We intend \enquote{probability} as the degree of
+belief that \enquote{would be agreed by all rational men if there were any
+  rational men} \cite{good1966}.
 
@@
     % \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
-    \expe{\yx \dotsm \yx \| \yH} =
-    \expe{\yX \dotsm \yX \| \yH}
+    \expe{\sav{\yx \dotsm \yx} \| \yH} =
+    \expe{\av{\yX \dotsm \yX} \| \yH}
     % \\[2\jot]

### 2017-05-17 maxNt.tex
intro

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T21:23:29+0200
+%% Last-Updated: 2017-05-17T22:40:04+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %Lagr multipliers
-\newcommand*{\yL}{\varLambda}
-\newcommand*{\yl}{\lambda}
+\newcommand*{\yf}{\bm{f}}
+\newcommand*{\yc}{\bm{c}}
+\newcommand*{\yL}{\bm{\varLambda}}
+\newcommand*{\yl}{\bm{\lambda}}
 \newcommand*{\yk}{\kappa}
@@
 
-Recent electrophysiological techniques in fact allow experimenters to
-record samples of even a couple hundreds neurons from specific brain areas.
+Recent electrophysiological techniques \cite{***} in fact allow
+experimenters to record samples of even a couple hundreds neurons from
+specific brain areas. These samples are usually picked out according to an
+unknown process, but from their observation we expect to learn something
+about all other neurons in the same brain area. That is, we assumed
+that they are a \enquote{representative random sample}.
+
 We don't elaborate on the various important purposes of such recordings
-here, but stress that these sample sizes can be considered \enquote{large},
-because their statistical analysis requires considerable computational
-power.
+here \cite{?***}, but stress that these sample sizes can be considered
+\enquote{large}, because their statistical analysis requires considerable
+computational power.
 
@@
 compression would also entail interesting physico-biological properties of
-neuronal activity. Maximum-entropy methods make this kind of compression:
-they associate a unique probability distribution with few experimental
-quantities. This distribution is usually said to be \enquote{maximally
-  noncommittal} \cite{jaynes1963}.
-
-We intend the quoted expression only as
-an umbrella term, also because it means very little without further
-clarifications.
-
-The neurons recorded in these kinds of experiments are usually picked out
-according to an unknown process, and from their observation we expect to
-learn something about all other neurons in the same brain area. That is,
-they are assumed to be a \enquote{representative random sample} of the
-neurons in that area. Our discussion hinges on this often just tacitly
-understood assumption.
-
-One of the uses of \me\ methods
-
-number. It is large from large
-population of neurons [\cite]. The recorded neurons are usually picked out
-from a specific brain area according to an unknown process, and from their
-observation in relation to stimuli or behavioral task, we strive for
-understanding the mechanism behind neuronal coding\mynote{LM still don't
-  know what people mean by this, can we find an unambiguous term or
-  rephrase?}. That is, they are assumed to be a \enquote{representative
-  random sample} of the neurons in that area.\mynote{This assumption says
-  that from the sample we can learn about the population. The relation of
-  population activity with stimuli or behaviour is a separate problem, so
-  \enquote{this is} doesn't really fit...}
-
-Huge amounts of parallel recordings from hundreds of neurons, available
-today, are accompanied by the statistical challenges to be employed for
-shedding light on the mechanism behind neuronal coding. The principle of
-\me\ was proposed as a way to overcome these challenges by constructing a
-full description of the neuronal firing patterns. The \me\ method employs
-only first and second order statistics of the representative random sample
-to construct a \enquote{maximally noncommittal} \cite{jaynes1963}
-probability distribution of the firing patterns [\cite]. \mynote{LM:
-  \enquote{noncommittal} doesn't mean anything to me to be honest, even if
-  Jaynes used that term. I would like to openly say this, backed up by
-  references, as done some paragraphs below.\\It seems they have used \me\
-  for other purposes too, sometimes unclear ones. And with all kinds of
-  constraints. So we shouldn't be too specific here in relation to purpose
-  and constraints. I'd rather give references and the readers can form
-  their own ideas about its use.}
-
-Using the pairwise \me\ models many studies investigated the questions such
-as, which level of interaction\mynote{ambiguous word} (pairwise or higher-order)
-between neurons is suffiecient to explain the observed experimental data
-[\cite]? Or how collective behavior of neurons can be related to stimuli or
-behavioral tasks [\cite]? add more **
-
-Here we show that there is a quandary in applying \me\ to a representative
-random sample of neurons to generate a maximally noncommital distribution
-for that sample.
-
-The quandary is this. Assume that our sample of neurons is
-representative of some population. If we assign a \me\ distribution to the
-sample, then we cannot assign a \me\ distribution to the full population.
-Vice versa, If we assign a \me\ distribution to the full population, then
-we cannot assign a \me\ distribution to the sample. This impossibility
-holds at least for \me\ distribution with moment constraints, and even if
-the orders of the moments constrained in the sample and in the full
-population are different.
-
-In other words, if our sample is a \enquote{random representative} of some
-population, then we must choose which has a \enquote{maximally
-  noncommittal} distribution: the sample or the population. We can't choose
-both. It goes without saying that if one is \enquote{maximally
-  noncommittal}, the other must be somehow \enquote{committal}. Which
-choice is most meaningful?
-
-In the rest of the paper we will mathematically show the contradiction
-above and generate a \me\ distribution for the full population rather than
-the sample. We will also present some probability relations relevant to
-random sampling. These relations are well-known in survey sampling and in
-the pedagogic problem of drawing from an urn without replacement, yet they
-are somewhat hard to find explicitly written in the neuroscientific
-literature.
-
- \mynote{Version 1}
-The present note has three nested purposes.
-
-The first and more general is to remind our readers that the \me\ method
-rests on subtle assumptions that may be contradictory with other
-assumptions we want to make on the data under study.
-
-As a concrete example we show -- and this is the second purpose -- that a
-sample of a neuronal population cannot be assigned a \enquote{maximally
-  noncommittal} \me\ distribution and at the same time be considered a
-\enquote{random representative} of the population.
-
-The proof of the example above rests on some probability relations from
-classical random sampling. These relations are somewhat hard to find
-explicitly stated in the neuroscientific literature; the third, minor
-purpose of this note is to state them explicitly.
-
-Let's clarify at once that our discussion pertains to a binary population at
-one specific instant or short window in time. It is possible to
-similarly discuss multi-state neuron models and population dynamics; but for
-simplicity neurons are here assumed to be in a fixed, active \enquote{$1$}
-or inactive \enquote{$0$} state; and evolution, change, time correlations,
-and similar concepts do not concern us.
-
-Maximum-entropy models are used for a variety of (sometimes not completely
-clear) reasons. For our purposes let's say that the \me\ method is assumed
-to generate a \enquote{maximally noncommittal} \cite{jaynes1963} probability
-distribution. We intend the quoted expression only as an umbrella term,
-also because it means very little without further clarifications.
-
-The neurons recorded in these kinds of experiments are usually picked out
-according to an unknown process, and from their observation we expect to
-learn something about all other neurons in the same brain area. That is,
-they are assumed to be a \enquote{representative random sample} of the
-neurons in that area. Our discussion hinges on this often just tacitly
-understood assumption.
-
-\mynote{Version 2 }
-In this note we would like to show that there is a
-contradiction in applying \me\ to a \enquote{representative random sample}
-of neurons to generate a \enquote{maximally noncommital} distribution for
-that sample.
-% \\
-% \mynote{V2 }In this note we would like to show that there is a clash between the
-% \enquote{representative random sample} spirit and the \enquote{maximally
-%   noncommital} spirit of \me\ models applied to such a sample of neurons.
-
-The contradiction is this. Assume that our sample of neurons is
-representative of some population. If we assign a \me\ distribution to the
-sample, then we cannot assign a \me\ distribution to the full population.
-Vice versa, If we assign a \me\ distribution to the full population, then
-we cannot assign a \me\ distribution to the sample. This impossibility
-holds at least for \me\ distribution with moment constraints, and even if
-the orders of the moments constrained in the sample and in the full
-population are different.
-
-In other words, if our sample is a \enquote{random representative} of some
-population, then we must choose which has a \enquote{maximally
-  noncommittal} distribution: the sample or the population. We can't choose
-both. It goes without saying that if one is \enquote{maximally
-  noncommittal}, the other must be somehow \enquote{committal}. Which
-choice is most meaningful?
-
-In the rest of the paper we will mathematically show the contradiction
-above and generate a \me\ distribution for the full population rather than
-the sample. We will also present some probability relations relevant to
-random sampling. These relations are well-known in survey sampling and in
-the pedagogic problem of drawing from an urn without replacement, yet they
-are somewhat hard to find explicitly written in the neuroscientific
-literature.
+neuronal activity. The standard \me\ method
+\cite{jaynes1954_r1974,sivia1996_r2006,meadetal1984} achieve this kind of
+compression: they associate a unique probability distribution with few
+experimental quantities \cite{***}. But this is only one of their uses.
+They are also used for various information-theoretic purposes or to
+generate reference probability distributions \cite{***}. In all these uses
+the \me\ distribution is chosen as the \enquote{maximally noncommittal} one
+\cite{jaynes1963}. This adjective means little without further technical
+characterizations. Different works give different characterizations, and in
+this paper we will use the quoted expression as an umbrella term for all of
+them. Our results will not depend on the specific characterization of
+\enquote{noncommittal}.
+
+\bigskip
+
+In this note we want to show that we have to face a dilemma if we want to
+apply the \me\ method to a representative random sample to find a maximally
+noncommittal distribution.
+
+The dilemma is this. We can apply the \me\ method to the sample, using a
+specified set of experimental constraints, and generate a probability
+distribution for its state. But our sample is representative of a larger
+population. We can apply the \me\ method to the larger population, using
+the same constraints, and generate a probability distribution for its
+larger state, and then find the distribution for the sample by
+marginalization. Either application seems to have some commendable
+features. However, \emph{the distributions obtained by these two
+  applications differ}. It goes without saying that if one is
+\enquote{maximally noncommittal}, the other must be somehow
+\enquote{committal}. Which choice is most meaningful?
+
+In the rest of the paper we mathematically formulate this dilemma. To this
+purpose we also present some probability relations relevant to
+\enquote{random sampling}. These relations are well-known in survey
+sampling and in the pedagogic problem of drawing from an urn without
+replacement, yet they are somewhat hard to find explicitly written in the
+neuroscientific literature, so they may be of interest on their own.
 
@@
 Suppose we want to constrain the sample expectations of a vector function
-$f$ to the vector values $c$, that is, $\expeb{f(\yxs)}=c$. Application of
-\me\ at the population level, denoted by $\yHa$, gives
+$\yf = (f_1,\dotsc,f_m)$ to the vector values $\yc=(c_1,\dotsc,c_m)$, that
+is, $\expeb{\yf(\yxs)}=\yc$. Application of \me\ at the population level,
+denoted by $\yHa$, gives
 \begin{equation}
@@
   \binom{N}{N \yRf}\exp\Bigl[\yL\T
-  \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\Bigr],
+  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
 \end{equation}
@@
   \binom{N}{N \yRf}\exp\Bigl[\yL\T
-  \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\Bigr],
-\end{equation}
-where $\yK$ is a normalization constant and $\yL\T$ are Lagrange
-multipliers such that
+  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
+\end{equation}
+where $\yK$ is a normalization constant and
+$\yL\T=(\varLambda_1,\dotsc,\varLambda_m)\T$ are Lagrange multipliers such
+that
 \begin{equation}
   \label{eq:app_maxent_pop_constraints}
-c = \yK\sum_{n\yrs=0}^n \sum_{N\yRf=0}^N f(\yrs) \ypp(\yrs \|\yRf)
+\yc = \yK\sum_{n\yrs=0}^n \sum_{N\yRf=0}^N \yf(\yrs) \ypp(\yrs \|\yRf)
 \,
   \binom{N}{N \yRf}\exp\Bigl[\yL\T
-  \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\Bigr].
+  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
 \end{equation}
@@
   \binom{n}{n\yrs}
-  \exp[\yl\T f(\yrs)]
+  \exp[\yl\T \yf(\yrs)]
 \end{equation}
@@
   \label{eq:app_maxent_sample_constraints}
-c = \yk\sum_{n\yrs=0}^n  f(\yrs) \binom{n}{n\yrs}
-  \exp[\yl\T f(\yrs)].
+\yc = \yk\sum_{n\yrs=0}^n  \yf(\yrs) \binom{n}{n\yrs}
+  \exp[\yl\T \yf(\yrs)].
 \end{equation}
@@
 observation~\eqref{eq:roughly_equal_nN} and also by the fact that they must
-satisfy the same expectations for $f$. But the latter may miss important
+satisfy the same expectations for $\yf$. But the latter may miss important
 features present in the former, like additional modes or particular tail
@@
 
+**********************************************************
+
+The neurons recorded in these kinds of experiments are usually picked out
+according to an unknown process, and from their observation we expect to
+learn something about all other neurons in the same brain area. That is,
+they are assumed to be a \enquote{representative random sample} of the
+neurons in that area. Our discussion hinges on this often just tacitly
+understood assumption.
+
+One of the uses of \me\ methods
+
+number. It is large from large
+population of neurons [\cite]. The recorded neurons are usually picked out
+from a specific brain area according to an unknown process, and from their
+observation in relation to stimuli or behavioral task, we strive for
+understanding the mechanism behind neuronal coding\mynote{LM still don't
+  know what people mean by this, can we find an unambiguous term or
+  rephrase?}. That is, they are assumed to be a \enquote{representative
+  random sample} of the neurons in that area.\mynote{This assumption says
+  that from the sample we can learn about the population. The relation of
+  population activity with stimuli or behaviour is a separate problem, so
+  \enquote{this is} doesn't really fit...}
+
+Huge amounts of parallel recordings from hundreds of neurons, available
+today, are accompanied by the statistical challenges to be employed for
+shedding light on the mechanism behind neuronal coding. The principle of
+\me\ was proposed as a way to overcome these challenges by constructing a
+full description of the neuronal firing patterns. The \me\ method employs
+only first and second order statistics of the representative random sample
+to construct a \enquote{maximally noncommittal} \cite{jaynes1963}
+probability distribution of the firing patterns [\cite]. \mynote{LM:
+  \enquote{noncommittal} doesn't mean anything to me to be honest, even if
+  Jaynes used that term. I would like to openly say this, backed up by
+  references, as done some paragraphs below.\\It seems they have used \me\
+  for other purposes too, sometimes unclear ones. And with all kinds of
+  constraints. So we shouldn't be too specific here in relation to purpose
+  and constraints. I'd rather give references and the readers can form
+  their own ideas about its use.}
+
+Using the pairwise \me\ models many studies investigated the questions such
+as, which level of interaction\mynote{ambiguous word} (pairwise or higher-order)
+between neurons is suffiecient to explain the observed experimental data
+[\cite]? Or how collective behavior of neurons can be related to stimuli or
+behavioral tasks [\cite]? add more **
+
+Here we show that there is a quandary in applying \me\ to a representative
+random sample of neurons to generate a maximally noncommital distribution
+for that sample.
+
+The quandary is this. Assume that our sample of neurons is
+representative of some population. If we assign a \me\ distribution to the
+sample, then we cannot assign a \me\ distribution to the full population.
+Vice versa, If we assign a \me\ distribution to the full population, then
+we cannot assign a \me\ distribution to the sample. This impossibility
+holds at least for \me\ distribution with moment constraints, and even if
+the orders of the moments constrained in the sample and in the full
+population are different.
+
+In other words, if our sample is a \enquote{random representative} of some
+population, then we must choose which has a \enquote{maximally
+  noncommittal} distribution: the sample or the population. We can't choose
+both. It goes without saying that if one is \enquote{maximally
+  noncommittal}, the other must be somehow \enquote{committal}. Which
+choice is most meaningful?
+
+In the rest of the paper we will mathematically show the contradiction
+above and generate a \me\ distribution for the full population rather than
+the sample. We will also present some probability relations relevant to
+random sampling. These relations are well-known in survey sampling and in
+the pedagogic problem of drawing from an urn without replacement, yet they
+are somewhat hard to find explicitly written in the neuroscientific
+literature.
+
+ \mynote{Version 1}
+The present note has three nested purposes.
+
+The first and more general is to remind our readers that the \me\ method
+rests on subtle assumptions that may be contradictory with other
+assumptions we want to make on the data under study.
+
+As a concrete example we show -- and this is the second purpose -- that a
+sample of a neuronal population cannot be assigned a \enquote{maximally
+  noncommittal} \me\ distribution and at the same time be considered a
+\enquote{random representative} of the population.
+
+The proof of the example above rests on some probability relations from
+classical random sampling. These relations are somewhat hard to find
+explicitly stated in the neuroscientific literature; the third, minor
+purpose of this note is to state them explicitly.
+
+Let's clarify at once that our discussion pertains to a binary population at
+one specific instant or short window in time. It is possible to
+similarly discuss multi-state neuron models and population dynamics; but for
+simplicity neurons are here assumed to be in a fixed, active \enquote{$1$}
+or inactive \enquote{$0$} state; and evolution, change, time correlations,
+and similar concepts do not concern us.
+
+Maximum-entropy models are used for a variety of (sometimes not completely
+clear) reasons. For our purposes let's say that the \me\ method is assumed
+to generate a \enquote{maximally noncommittal} \cite{jaynes1963} probability
+distribution. We intend the quoted expression only as an umbrella term,
+also because it means very little without further clarifications.
+
+The neurons recorded in these kinds of experiments are usually picked out
+according to an unknown process, and from their observation we expect to
+learn something about all other neurons in the same brain area. That is,
+they are assumed to be a \enquote{representative random sample} of the
+neurons in that area. Our discussion hinges on this often just tacitly
+understood assumption.
+
+\mynote{Version 2 }
+In this note we would like to show that there is a
+contradiction in applying \me\ to a \enquote{representative random sample}
+of neurons to generate a \enquote{maximally noncommital} distribution for
+that sample.
+% \\
+% \mynote{V2 }In this note we would like to show that there is a clash between the
+% \enquote{representative random sample} spirit and the \enquote{maximally
+%   noncommital} spirit of \me\ models applied to such a sample of neurons.
+
+The contradiction is this. Assume that our sample of neurons is
+representative of some population. If we assign a \me\ distribution to the
+sample, then we cannot assign a \me\ distribution to the full population.
+Vice versa, If we assign a \me\ distribution to the full population, then
+we cannot assign a \me\ distribution to the sample. This impossibility
+holds at least for \me\ distribution with moment constraints, and even if
+the orders of the moments constrained in the sample and in the full
+population are different.
+
+In other words, if our sample is a \enquote{random representative} of some
+population, then we must choose which has a \enquote{maximally
+  noncommittal} distribution: the sample or the population. We can't choose
+both. It goes without saying that if one is \enquote{maximally
+  noncommittal}, the other must be somehow \enquote{committal}. Which
+choice is most meaningful?
+
+In the rest of the paper we will mathematically show the contradiction
+above and generate a \me\ distribution for the full population rather than
+the sample. We will also present some probability relations relevant to
+random sampling. These relations are well-known in survey sampling and in
+the pedagogic problem of drawing from an urn without replacement, yet they
+are somewhat hard to find explicitly written in the neuroscientific
+literature.
+
 \end{document}

### 2017-05-17 maxNt.tex
intro

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T22:40:04+0200
+%% Last-Updated: 2017-05-17T22:42:16+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \cite{jaynes1954_r1974,sivia1996_r2006,meadetal1984} achieve this kind of
-compression: they associate a unique probability distribution with few
-experimental quantities \cite{***}. But this is only one of their uses.
-They are also used for various information-theoretic purposes or to
-generate reference probability distributions \cite{***}. In all these uses
-the \me\ distribution is chosen as the \enquote{maximally noncommittal} one
+compression: it associates a unique probability distribution with few
+experimental quantities \cite{***}. But this is only one of its uses. It is
+also used for various information-theoretic purposes or to generate
+reference probability distributions \cite{***}. In all these uses the \me\
+distribution is chosen as the \enquote{maximally noncommittal} one
 \cite{jaynes1963}. This adjective means little without further technical
-characterizations. Different works give different characterizations, and in
+characterizations. Different works give different characterizations, but in
 this paper we will use the quoted expression as an umbrella term for all of

### 2017-05-17 maxNt.tex
intro

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T22:42:16+0200
+%% Last-Updated: 2017-05-17T22:47:16+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 neuronal activity. The standard \me\ method
-\cite{jaynes1954_r1974,sivia1996_r2006,meadetal1984} achieve this kind of
+\cite{jaynes1954_r1974,sivia1996_r2006,meadetal1984} accomplishes this kind of
 compression: it associates a unique probability distribution with few

### 2017-05-17 maxNt.tex
intro

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T22:47:16+0200
+%% Last-Updated: 2017-05-17T22:51:02+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\firstdraft}{4 November 2015}
-\newcommand*{\pdftitle}{Hidden  assumptions on population size\\in the maximum-entropy method}
+\newcommand*{\pdftitle}{Maximum-entropy distributions\\ and representative
+  random samples: a dilemma}
 \newcommand*{\headtitle}{\pdftitle}
@@
 
+Our discussion pertains to a neurons modelled as binary units at one
+specific instant or short window in time. It is possible to similarly
+discuss multi-state neuron models and population dynamics; but for
+simplicity neurons are here assumed to be in a fixed, active \enquote{$1$}
+or inactive \enquote{$0$} state; and evolution, change, time correlations,
+and similar concepts do not concern us.
+
 Our notation follows ISO and ANSI standards

### 2017-05-17 maxNt.tex
intro

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T22:51:02+0200
+%% Last-Updated: 2017-05-17T23:04:18+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 This note is mainly addressed to neuroscientists interested in \me\
-methods, but its discussion of \enquote{random sampling} formulae may be
-useful to all neuroscientists involved in understanding the physical and
-dynamical characteristics of brain areas via neuronal recording.
+methods, but we would be pleased if its discussion of the probability of
+\enquote{random sampling} were useful to neuroscientists that use other
+statistical methods to study the physical and dynamical characteristics of
+brain areas via neuronal recording.
 

### 2017-05-17 maxNt.tex
discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-17T23:04:18+0200
+%% Last-Updated: 2017-05-18T01:18:11+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\firstdraft}{4 November 2015}
-\newcommand*{\pdftitle}{Maximum-entropy distributions\\ and representative
-  random samples: a dilemma}
+\newcommand*{\pdftitle}{Representative random samples\\ and maximum-entropy
+  distributions:\\ a dilemma}
 \newcommand*{\headtitle}{\pdftitle}
@@
 \author{
-  V. Rostami\\Hitlerland\\\texttt{v.rostami@fz-juelich.de}
+  P.G.L. Porta Mana\thanks{Equally contributing first authors}\\Independent researcher\\\texttt{pgl@portamana.org}
   \And
-  P.G.L. Porta Mana\\Mussoliniland\\\texttt{pgl@portamana.org}
+  V. Rostami$^*$\\Forschungszentrum J{\"u}lich INM-6\\Germany\\\texttt{v.rostami@fz-juelich.de}
   \And
@@
 neuronal activity. The standard \me\ method
-\cite{jaynes1954_r1974,sivia1996_r2006,meadetal1984} accomplishes this kind of
+\cite{jaynes1957,sivia1996_r2006,meadetal1984} accomplishes this kind of
 compression: it associates a unique probability distribution with few
@@
 
-Our discussion pertains to a neurons modelled as binary units at one
-specific instant or short window in time. It is possible to similarly
+In the final discussion we present some arguments in favour of one choice
+in our dilemma, but we do not mean that to be our final answer. This note,
+and its dilemma,  also have the more general purpose of reminding our
+readers the importance of asking \emph{what is the question we're trying to
+  answer with probability theory?} and \emph{which assumptions do we want
+  or need to make?}
+
+Our mathematical analysis pertains to a neurons modelled as binary units at
+one specific instant or short window in time. It is possible to similarly
 discuss multi-state neuron models and population dynamics; but for
@@
 or inactive \enquote{$0$} state; and evolution, change, time correlations,
-and similar concepts do not concern us.
-
-Our notation follows ISO and ANSI standards
+and similar concepts do not concern us. We consider \me\ models that have
+constraints based on some kinds of sample or population averages; they are
+often called \enquote{homogeneous}. The final discussion touches upon
+\enquote{inhomogeneous} models as well.
+
+The notation in this note follows ISO and ANSI standards
 \citep{iso1993,ieee1993,nist1995} but for the use of the comma \enquote{,}
@@
 \citep{jaynes1994_r2003}. We intend \enquote{probability} as the degree of
-belief that \enquote{would be agreed by all rational men if there were any
+belief which \enquote{would be agreed by all rational men if there were any
   rational men} \cite{good1966}.
@@
 introduce a general averaging operator $\sav{\dotv}$ defined by
-\begin{align}
-  \av{\yX}  &\defd \tfrac{1}{N} (X_1 + X_2 + \dotsb + X_N),
-%  &
-%  \sav{\yx}  &\defd \tfrac{1}{n} (x_1 + x_2 + \dotsb + x_n),
+\begin{equation}
+  \begin{gathered}
+    \av{\yX} \defd \tfrac{1}{N} (X_1 + X_2 + \dotsb + X_N),
+    % & \sav{\yx} &\defd \tfrac{1}{n} (x_1 + x_2 + \dotsb + x_n),
+   \qquad% \\[2\jot]
+    \av{\yX \yX} \defd \tbinom{N}{2}^{-1} (X_1 X_2 + X_1 X_3 + \dotsb +
+    X_{N-1} X_N),
+    % & \sav{\yx \yx} &\defd\tbinom{n}{2}^{-1} (x_1 x_2 + x_1 x_3 + \dotsb
+    % + x_{n-1} x_n),
     \\[2\jot]
-  \av{\yX \yX} &\defd
-\tbinom{N}{2}^{-1} (X_1 X_2 + X_1 X_3  + \dotsb +  X_{N-1} X_N),
-%  &
-%    \sav{\yx \yx} &\defd\tbinom{n}{2}^{-1} (x_1 x_2 + x_1 x_3  + \dotsb +  x_{n-1} x_n),
-\\[2\jot]
-\av{\yX\yX\yX} &\defd
-                  \tbinom{N}{3}^{-1}
-                  (X_1 X_2 X_3 +% X_1 X_2 X_4
+    \av{\yX\yX\yX} \defd \tbinom{N}{3}^{-1} (X_1 X_2 X_3 +% X_1 X_2 X_4
     \dotsb + X_{N-2} X_{N-1}X_N),
-%  &
-%    \sav{\yx\yx\yx} &\defd
-%                      \tbinom{n}{3}^{-1} (x_1 x_2 x_3 + %x_1 x_2 x_4
+    % & \sav{\yx\yx\yx} &\defd \tbinom{n}{3}^{-1} (x_1 x_2 x_3
+    % + %x_1 x_2 x_4
     % \dotsb + x_{n-2} x_{n-1} x_n),
-\end{align}
-and so on
+  \end{gathered}
+\end{equation}
+and so on.
 These formulae say that $\yXf$ is the fraction of active neurons, $\yXXf$
@@
 so on. Products of states like $X_i \dotsm X_j$ also have values in
-$\set{0,1}$; from this it can combinatorially be proved that
+$\set{0,1}$; from this we can combinatorially  prove that
 \begin{equation}
@@
 \end{equation}
-Analogous formulae  hold for quantities like $\yx$, $\yRf$, $\yrs$.
+Analogous formulae  hold for quantities like $\yx$, $\yR$, $\yr$.
 
@@
 We need to make an initial probability assignment before any experimental
-observations are made, no matter what kinds of predictions we are
-interested in. This initial assignment will be modified by the experimental
+observations are made. %, no matter what kinds of predictions we are interested in.
+This initial assignment will be modified by our experimental
 observations.
@@
 We yet know very little about the physical details of the individual
-neurons, their locations for example. Our initial state of knowledge $\yH$
+neurons; their locations for example. Our initial state of knowledge $\yH$
 is therefore symmetric, or \enquote{exchangeable}, under their
@@
 \end{equation}
-the latter being the probability for the population average $\yX$. Proof
-and generalizations to non-binary and continuum cases are given by
-de~Finetti \cite{definetti1959b}, Ericson \cite{ericson1976}, Diaconis
-\cite{diaconis1977}, Heath \amp\ Sudderth \cite{heathetal1976}. This
-theorem is intuitive: owing to symmetry, all states with $N\yRf$ active
-neurons must have equal probabilities.
+the latter being the probability for the population average $\yX$. Proof of
+this theorem and generalizations to non-binary and continuum cases are
+given by de~Finetti \cite{definetti1959b}, Ericson \cite{ericson1976},
+Diaconis \cite{diaconis1977}, Heath \amp\ Sudderth \cite{heathetal1976}.
+This theorem is intuitive: owing to symmetry, all states with $N\yRf$
+active neurons must have equal probabilities.
 
@@
 
-The conditional probability $\ypp(\yrs,\yRf)$ connects the spaces of the
+The conditional probability $\ypp(\yrs,\yRf)$ relates the spaces of the
 sample average $\yXf\in\set{0,\dotsc,N}$ and of the population average
-$\yxs\in\set{0,\dotsc,n}$. It is a coarsening projector of any probability
-$p$ for $\yXf$ onto a marginal probability $p_*$ for $\yxs$:
+$\yxs\in\set{0,\dotsc,n}$ in a special way. It is a coarsening projector of
+any probability $p$ for $\yXf$ onto a marginal probability $p_*$ for
+$\yxs$:
 \begin{equation}
@@
 \end{equation}
-Therefore it also a pulls back expectations of functions $f$ of the sample
-average $\yxs$ to expectations of functions $f^*$ the population average
+Conversely it also a pulls back expectations of functions $f$ of the sample
+average $\yxs$ to expectations of functions $f^*$ of the population average
 $\yXf$:
@@
 These are only very approximate equalities; they may miss important
-features of the two probability distributions. If the distribution for the
-population average $\yXf$ is bimodal, for example, the bimodality can be
-lost in the distribution for the sample average $\yxs$, owing to the
-coarsening effect of $\ypp(\yrs\|\yRf)$.
-
-The approximate equalities above express the fact that \emph{our
+features of the two probability distributions. In the next section we will
+in fact emphasize their differences. If the distribution for the population
+average $\yXf$ is bimodal, for example, the bimodality can be lost in the
+distribution for the sample average $\yxs$, owing to the coarsening effect
+of $\ypp(\yrs\|\yRf)$.
+
+Yet the approximate equalities above express the fact that \emph{our
   uncertainty about the sample is representative of our uncertainty about
@@
 
-We prefer not to answer the question that closed the preceding section,
-because the optimal answer can only be given case by case, depending on the
-inferences we are trying to make, on the assumptions needed and negligible,
-and on our computational power. We would just like to stress that the
-answer is not unique, and some answers may imply assumptions we were not
-aware of.
+The question that closed the preceding section cannot receive a categorical
+answer. An answer can only be given case by case, depending on the
+inferences we are trying to make, on the assumptions needed and those
+negligible, and on our computational power. We want to stress, in fact, the
+extreme importance and priority of the following two questions:
+\begin{itemize}
+\item Of which quantities do we need to quantify the uncertainty? and why?
+\item What are the assumptions we need or want to make to answer the
+  question above?
+\end{itemize}
+
+The first purpose of this note was to remind our readers that probability
+models can have hidden assumptions that we don't want to make or that are
+in contradiction with assumptions we want to make. The dilemma generated by
+the \me\ method and the assumption of random representative sampling is an
+example. The usual application of \me\ at the sample level generates a
+different distribution than the application at the population level. Then
+the former application must implicitly assume that no larger population
+exist; or in other words that the neurons of the sample are isolated from
+their environment.
 
@@
 \me\ at the population level has by construction a lower entropy than the
-one obtained by application at the sample level. It says that some sample
-states $\yx$ should be considered less likely than in the other
-distribution. The lower entropy signals the presence of an additional
-physical assumption behind the former distribution. Such assumption,
-intuitively, is that the sample is part of a larger population, of which it
-is representative. And this assumption may affect our predictions.
-\mynote{It is a very natural assumption in the neuroscientific context and
-  therefore we find it preferable. In fact, also physical neuronal network
-  model add an external input to the neurons, mimicking their embedding in a
-  larger network}
+one obtained by application at the sample level. The former distribution
+says that some sample states $\yx$ should be considered less likely than in
+the latter distribution. A lower entropy signals the presence of an
+additional assumption. Such assumption, intuitively, is that the sample is
+part of a larger population, and our ignorance about the former is
+representative of our ignorance about the latter. This assumption may
+affect our predictions.
+
+In a neuroscientific context we find this assumption very natural, and
+therefore the \me\ method at the population level preferable. Also
+\emph{physical} neuronal-network models add indeed an external input to the
+neurons, mimicking their embedding in a larger network.
 
@@
 about the state thanks to our knowledge of its preceding dynamics.
-\mynote{but in this case both distributions are immensely sharp, and this
-  doesn't affect predictions}
+%\mynote{but in this case both distributions are immensely sharp, and this
+%  doesn't affect predictions}
 

### 2017-05-17 maxNt.tex
discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-18T01:18:11+0200
+%% Last-Updated: 2017-05-18T01:43:11+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}
-  \mynote{Better title?}
-  ***%  The abstract must be limited to one paragraph.
+  This note has three nested purposes. The first purpose is to show that
+  the \me\ method can be applied to a representative random sample of a
+  population, to generate its probability distribution, along two different
+  routes. Both routes appear legitimate, but they give inequivalent
+  results. Which route should be chosen? Some arguments are presented in
+  favour of one. The second more general purpose, motivated by the above
+  dilemma, is to remind readers that models like \me\ may contain hidden
+  assumptions; in this case the hidden an unnatural assumption that the
+  sample modelled is isolated from the rest of the population. The third
+  purpose is to promote some old but possibly forgotten probability
+  formulae that may be useful in neuroscientific sampling contexts.
+  % The abstract must be limited to one paragraph.
 \end{abstract}

### 2017-05-18 maxNt.tex
discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-18T01:43:11+0200
+%% Last-Updated: 2017-05-18T11:35:39+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 to denote logical conjunction. Probability notation follows Jaynes
-\citep{jaynes1994_r2003}. We intend \enquote{probability} as the degree of
-belief which \enquote{would be agreed by all rational men if there were any
-  rational men} \cite{good1966}.
+\citep{jaynes1994_r2003}. By \enquote{probability} we mean plausibility or
+the degree of belief which \enquote{would be agreed by all rational men if
+  there were any rational men} \cite{good1966}.
 
@@
 We need to make an initial probability assignment before any experimental
-observations are made. %, no matter what kinds of predictions we are interested in.
-This initial assignment will be modified by our experimental
-observations.
-
-We yet know very little about the physical details of the individual
-neurons; their locations for example. Our initial state of knowledge $\yH$
-is therefore symmetric, or \enquote{exchangeable}, under their
-permutations. This symmetry must be reflected in our initial probability:
-the \emph{representation theorem for finite exchangeability} states that it
-must obey
+observations are
+made. %, no matter what kinds of predictions we are interested in.
+This initial assignment will be modified by our experimental observations.
+We would also like our probability assignment to reflect that the sample is
+somehow \enquote{representative} of the population.
+
+Let's assume that we initially know very little about the physical details
+of the individual neurons; their locations for example. Our initial state
+of knowledge or ignorance $\yH$ is therefore symmetric, or
+\enquote{exchangeable}, under their permutations. This symmetry must be
+reflected in our initial probability: the \emph{representation theorem for
+  finite exchangeability} states that it must obey
 \begin{equation}
@@
 Diaconis \cite{diaconis1977}, Heath \amp\ Sudderth \cite{heathetal1976}.
-This theorem is intuitive: owing to symmetry, all states with $N\yRf$
-active neurons must have equal probabilities.
+This theorem is intuitive: owing to symmetry, we must assign equal
+probabilities to all states with $N\yRf$ active neurons.
 
@@
 \end{gather}
-
-The latter conditional probability is a hypergeometric distribution
-$\ypp(\yrs\|\yRf)$, typical of \enquote{drawing without replacement}
-problems. The combinatorial proof of the formulae above is in fact the same
-as for this class of problems
+Our initial symmetric ignorance should intuitively also apply to the
+sample; indeed, the probability for the state of the
+sample~\eqref{eq:subpop_average} automatically satisfies the representation
+theorem~\eqref{eq:joint_plaus_N_homog} as well. The conditional probability
+in the last formula is a hypergeometric distribution $\ypp(\yrs\|\yRf)$,
+typical of \enquote{drawing without replacement} problems. The
+combinatorial proof of the formulae above is in fact the same as for this
+class of problems
 \cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
@@
   $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
-  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1} \defs
-  \ypp(\yrs\|\yRf)$ for $N=5000$, $n=200$. (Band artifacts may appear in the
+\ypp(\yrs\|\yRf) \defd  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}$ for $N=5000$, $n=200$. (Band artifacts may appear in the
   colourbar depending on your \textsc{pdf} viewer.)}
@@
 \end{figure}%hypergeometric_identity.nb
-A look at a plot of the hypergeometric distribution $\ypp(\yrs\|\yRf)$, for
-example \fig~\ref{fig:hypergeom_proj}, reveals that it is a sort of
-\enquote{fuzzy identity matrix} between the $\yXf$-space $\set{0,\dotsc,N}$
-and $\yxs$-space $\set{0,\dotsc,n}$. When $n=N$ it is the identity matrix.
-We thus have that
+A look at a plot of the hypergeometric distribution $\ypp(\yrs\|\yRf)$, see
+\fig~\ref{fig:hypergeom_proj}, reveals that it is a sort of \enquote{fuzzy
+  identity matrix} between the $\yXf$-space $\set{0,\dotsc,N}$ and
+$\yxs$-space $\set{0,\dotsc,n}$. When $n=N$ it is the identity matrix. We
+thus have that
 \begin{equation}
@@
 \end{equation}
-These are only very approximate equalities; they may miss important
+These are only very approximate equalities: they may miss important
 features of the two probability distributions. In the next section we will
@@
 
-Yet the approximate equalities above express the fact that \emph{our
+Yet, the approximate equalities above express the fact that \emph{our
   uncertainty about the sample is representative of our uncertainty about
-  the population and about other samples}, and vice versa. Upon observation
-of a sample average, say $\yxxs=a$, the updated expectations for such
-average in the population and in any new sample will usually be shifted
-towards the observed value, as follows from Bayes's theorem and the
-formulae above.
+  the population and about other samples}, and vice versa. This is how the
+idea of representativeness is translated into our probabilities. The
+symmetry present in our initial probabilities is not a physical property of
+the neuronal population or sample. It only expresses the symmetry of our
+initial uncertainty about them, and does not imply any sort of physical
+similarity between the neurons. Subsequent observations may in fact break
+this symmetry. Upon observation of a sample average, say $\yxxs=a$, the
+updated expectations for such average in the population and in any new
+sample will usually be shifted towards the observed value, as follows from
+Bayes's theorem and the formulae above.
 
@@
 differ wildly.
-
-The symmetry present in the probabilities above is not a physical property
-of the neuronal population. It only expresses the symmetry of our initial
-uncertainty about the population, and does not imply any sort of physical
-similarity between the neurons. Subsequent observations may in fact break
-this symmetry. The initial symmetry should intuitively also apply to the
-sample, and this is indeed the case: the probability for the state of the
-sample~\eqref{eq:subpop_average} automatically satisfies the representation
-theorem~\eqref{eq:joint_plaus_N_homog} as well.
 
@@
 
-\section{Enter maximum-entropy: quandary}
+\section{Enter maximum-entropy: dilemma}
 \label{sec:specific_initial_probability}
@@
 selects one distribution, purported to be \enquote{maximally noncommittal},
-among those that have the desired expectations. But we are in a quandary:
+among those that have the desired expectations. But here's a dilemma:
 formulae~\eqref{eq:pullback_P} allow us to apply the method to find the
@@
 population level~\eqref{eq:app_maxent_pop_marg} and at the sample level
-\eqref{eq:app_maxent_sample} will be approximately equal, by our previous
-observation~\eqref{eq:roughly_equal_nN} and also by the fact that they must
-satisfy the same expectations for $\yf$. But the latter may miss important
-features present in the former, like additional modes or particular tail
-behaviour.
+\eqref{eq:app_maxent_sample} should be approximately equal, by our previous
+observation about representativity~\eqref{eq:roughly_equal_nN} and also by
+the fact that they must satisfy the same expectations for $\yf$.
+
+Yet they cannot be exactly equal, because their equality would require the
+Lagrange multipliers $\yL$ and $\yl$ to satisfy equations
+\eqref{eq:app_maxent_pop_constraints},
+\eqref{eq:app_maxent_sample_constraints}, and
+$\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$ -- that is, $2m+n$
+equations in $m$ unknowns. This could be possible, if at all, only for very
+special choices of constraints functions $\yf$ and values $\yc$.
+
+The sample distribution obtained from \me\ at the sample level will
+therefore likely miss important features present in the one obtained at the
+population level, like additional modes or particular tail behaviour.
 
@@
 \centering
-\includegraphics[width=0.75\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
-\caption{Plots of $\p(\yxs = \yrs)$ constructed by \me\ at the population
+\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
+\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at the population
   level followed by sample marginalization (blue triangles),
   \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5000$,
-  $n=200$, $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$.}
+  circles), \eqn~\eqref{eq:sample_av_example_ME}, with $N=5000$,
+  $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$.}
 \label{fig:diff_maxent_pop_sample}
@@
 \centering
-\includegraphics[width=0.75\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}
+\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}
 % 
@@
   \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5000$,
-  $n=200$, $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$,
+  circles), \eqn~\eqref{eq:sample_av_example_ME}, with $N=5000$,
+  $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$,
   $\expe{\sav{\yx\yx\yx}}=1.48\times10^{-4}$,
@@
 
-The inequivalence of the two \me\ applications is generally valid for small
-number of constraints, even if the two states of knowledge $\yHa$, $\yHb$
-use different numbers $m'$, $m''$ \emph{and values} of constraints. The
-equality $\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$ in this case
-requires solving $n$ equations in $m'+m''$ unknowns (normalization
-is taken care of), so in general it can only be satisfied if $n\le m'+m''$.
-
 \medskip
 
-Where should we apply the \me\ method then, on the sample or on the
-population?
+How should we apply the \me\ method then? on the sample or on the
+population? Which one is maximally noncommittal?
 
@@
 The question that closed the preceding section cannot receive a categorical
-answer. An answer can only be given case by case, depending on the
-inferences we are trying to make, on the assumptions needed and those
-negligible, and on our computational power. We want to stress, in fact, the
+answer. An optimal answer can only be given case by case, depending on the
+computational power available, on which inferences we are trying to make,
+on which assumptions we need or want to make and those we wish to avoid.
+
+\iffalse We want to stress, in fact, the
 extreme importance and priority of the following two questions:
@@
   question above?
-\end{itemize}
-
-The first purpose of this note was to remind our readers that probability
-models can have hidden assumptions that we don't want to make or that are
-in contradiction with assumptions we want to make. The dilemma generated by
-the \me\ method and the assumption of random representative sampling is an
-example. The usual application of \me\ at the sample level generates a
-different distribution than the application at the population level. Then
-the former application must implicitly assume that no larger population
-exist; or in other words that the neurons of the sample are isolated from
-their environment.
-
-The probability distribution for the sample average obtained by applying
-\me\ at the population level has by construction a lower entropy than the
-one obtained by application at the sample level. The former distribution
-says that some sample states $\yx$ should be considered less likely than in
-the latter distribution. A lower entropy signals the presence of an
-additional assumption. Such assumption, intuitively, is that the sample is
-part of a larger population, and our ignorance about the former is
-representative of our ignorance about the latter. This assumption may
-affect our predictions.
+\end{itemize}\fi
+
+The first purpose of this note was indeed to remind ourselves that
+probability models require a continuous scrutiny, because they can rest on
+hidden assumptions that we don't want to make or that contradict others we
+are making. The dilemma generated by the \me\ method and the assumption of
+random representative sampling is an example. The usual application of \me\
+at the sample level yields a different distribution from the application at
+the population level. The former application must therefore implicitly
+assume that no larger population exist, or that the neurons of the sample
+are isolated from their environment.
+
+The \me\ application at the population level has in fact by construction a
+lower entropy than the one obtained by application at the sample level:
+both satisfy the same constraints, but only the latter is obtained by
+direct maximization. The lower entropy of the distribution obtained from
+application at the population level is clearly visible from its higher
+sharpness in \fig~\ref{fig:diff_maxent_pop_sample}. This distribution
+therefore says that some sample states $\yx$ should be considered less
+likely than in the distribution obtained at the sample level. A lower
+entropy signals the presence of additional assumptions or constraints. In
+this case the assumption is that the sample is part of a larger population,
+and our ignorance about the sample should be representative of our
+ignorance about a larger population. This assumption affects our
+predictions.
 
 In a neuroscientific context we find this assumption very natural, and
-therefore the \me\ method at the population level preferable. Also
-\emph{physical} neuronal-network models add indeed an external input to the
-neurons, mimicking their embedding in a larger network.
+therefore the \me\ method at the population level preferable. After all,
+also \emph{physical} neuronal-network models usually include some sort of
+external input to the neurons, mimicking their embedding in a larger
+network.
 

### 2017-05-18 maxNt.tex
discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-18T11:35:39+0200
+%% Last-Updated: 2017-05-18T12:03:23+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \usepackage[utf8]{inputenx}
-\usepackage{newunicodechar}
+%\usepackage{newunicodechar}
 % \newunicodechar{Ĕ}{\u{E}}
@@
 % Greek letters not usually given in LaTeX. Comment the unneeded ones
-\DeclareMathSymbol{\varpartial}{\mathalpha}{egreeki}{"40}
-\DeclareMathSymbol{\partialup}{\mathalpha}{egreekr}{"40}
-\DeclareMathSymbol{\alpha}{\mathalpha}{egreeki}{"0B}
-\DeclareMathSymbol{\beta}{\mathalpha}{egreeki}{"0C}
-\DeclareMathSymbol{\gamma}{\mathalpha}{egreeki}{"0D}
-\DeclareMathSymbol{\delta}{\mathalpha}{egreeki}{"0E}
-\DeclareMathSymbol{\epsilon}{\mathalpha}{egreeki}{"0F}
-\DeclareMathSymbol{\zeta}{\mathalpha}{egreeki}{"10}
-\DeclareMathSymbol{\eta}{\mathalpha}{egreeki}{"11}
-\DeclareMathSymbol{\theta}{\mathalpha}{egreeki}{"12}
-\DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
-\DeclareMathSymbol{\kappa}{\mathalpha}{egreeki}{"14}
-\DeclareMathSymbol{\lambda}{\mathalpha}{egreeki}{"15}
-\DeclareMathSymbol{\mu}{\mathalpha}{egreeki}{"16}
-\DeclareMathSymbol{\nu}{\mathalpha}{egreeki}{"17}
-\DeclareMathSymbol{\xi}{\mathalpha}{egreeki}{"18}
-\DeclareMathSymbol{\omicron}{\mathalpha}{egreeki}{"6F}
-\DeclareMathSymbol{\pi}{\mathalpha}{egreeki}{"19}
-\DeclareMathSymbol{\rho}{\mathalpha}{egreeki}{"1A}
-\DeclareMathSymbol{\sigma}{\mathalpha}{egreeki}{"1B}
-\DeclareMathSymbol{\tau}{\mathalpha}{egreeki}{"1C}
-\DeclareMathSymbol{\upsilon}{\mathalpha}{egreeki}{"1D}
-\DeclareMathSymbol{\phi}{\mathalpha}{egreeki}{"1E}
-\DeclareMathSymbol{\chi}{\mathalpha}{egreeki}{"1F}
-\DeclareMathSymbol{\psi}{\mathalpha}{egreeki}{"20}
-\DeclareMathSymbol{\omega}{\mathalpha}{egreeki}{"21}
-\DeclareMathSymbol{\varepsilon}{\mathalpha}{egreeki}{"22}
-\DeclareMathSymbol{\vartheta}{\mathalpha}{egreeki}{"23}
-\DeclareMathSymbol{\varpi}{\mathalpha}{egreeki}{"24}
-\let\varrho\rho 
-\let\varsigma\sigma
-\let\varkappa\kappa
-\DeclareMathSymbol{\varphi}{\mathalpha}{egreeki}{"27}
-%
-\DeclareMathSymbol{\varAlpha}{\mathalpha}{egreeki}{"41}
-\DeclareMathSymbol{\varBeta}{\mathalpha}{egreeki}{"42}
-\DeclareMathSymbol{\varGamma}{\mathalpha}{egreeki}{"00}
-\DeclareMathSymbol{\varDelta}{\mathalpha}{egreeki}{"01}
-\DeclareMathSymbol{\varEpsilon}{\mathalpha}{egreeki}{"45}
-\DeclareMathSymbol{\varZeta}{\mathalpha}{egreeki}{"5A}
-\DeclareMathSymbol{\varEta}{\mathalpha}{egreeki}{"48}
-\DeclareMathSymbol{\varTheta}{\mathalpha}{egreeki}{"02}
+% \DeclareMathSymbol{\varpartial}{\mathalpha}{egreeki}{"40}
+% \DeclareMathSymbol{\partialup}{\mathalpha}{egreekr}{"40}
+% \DeclareMathSymbol{\alpha}{\mathalpha}{egreeki}{"0B}
+% \DeclareMathSymbol{\beta}{\mathalpha}{egreeki}{"0C}
+% \DeclareMathSymbol{\gamma}{\mathalpha}{egreeki}{"0D}
+% \DeclareMathSymbol{\delta}{\mathalpha}{egreeki}{"0E}
+% \DeclareMathSymbol{\epsilon}{\mathalpha}{egreeki}{"0F}
+% \DeclareMathSymbol{\zeta}{\mathalpha}{egreeki}{"10}
+% \DeclareMathSymbol{\eta}{\mathalpha}{egreeki}{"11}
+% \DeclareMathSymbol{\theta}{\mathalpha}{egreeki}{"12}
+% \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
+% \DeclareMathSymbol{\kappa}{\mathalpha}{egreeki}{"14}
+% \DeclareMathSymbol{\lambda}{\mathalpha}{egreeki}{"15}
+% \DeclareMathSymbol{\mu}{\mathalpha}{egreeki}{"16}
+% \DeclareMathSymbol{\nu}{\mathalpha}{egreeki}{"17}
+% \DeclareMathSymbol{\xi}{\mathalpha}{egreeki}{"18}
+% \DeclareMathSymbol{\omicron}{\mathalpha}{egreeki}{"6F}
+% \DeclareMathSymbol{\pi}{\mathalpha}{egreeki}{"19}
+% \DeclareMathSymbol{\rho}{\mathalpha}{egreeki}{"1A}
+% \DeclareMathSymbol{\sigma}{\mathalpha}{egreeki}{"1B}
+% \DeclareMathSymbol{\tau}{\mathalpha}{egreeki}{"1C}
+% \DeclareMathSymbol{\upsilon}{\mathalpha}{egreeki}{"1D}
+% \DeclareMathSymbol{\phi}{\mathalpha}{egreeki}{"1E}
+% \DeclareMathSymbol{\chi}{\mathalpha}{egreeki}{"1F}
+% \DeclareMathSymbol{\psi}{\mathalpha}{egreeki}{"20}
+% \DeclareMathSymbol{\omega}{\mathalpha}{egreeki}{"21}
+% \DeclareMathSymbol{\varepsilon}{\mathalpha}{egreeki}{"22}
+% \DeclareMathSymbol{\vartheta}{\mathalpha}{egreeki}{"23}
+% \DeclareMathSymbol{\varpi}{\mathalpha}{egreeki}{"24}
+% \let\varrho\rho 
+% \let\varsigma\sigma
+% \let\varkappa\kappa
+% \DeclareMathSymbol{\varphi}{\mathalpha}{egreeki}{"27}
+% %
+% \DeclareMathSymbol{\varAlpha}{\mathalpha}{egreeki}{"41}
+% \DeclareMathSymbol{\varBeta}{\mathalpha}{egreeki}{"42}
+% \DeclareMathSymbol{\varGamma}{\mathalpha}{egreeki}{"00}
+% \DeclareMathSymbol{\varDelta}{\mathalpha}{egreeki}{"01}
+% \DeclareMathSymbol{\varEpsilon}{\mathalpha}{egreeki}{"45}
+% \DeclareMathSymbol{\varZeta}{\mathalpha}{egreeki}{"5A}
+% \DeclareMathSymbol{\varEta}{\mathalpha}{egreeki}{"48}
+% \DeclareMathSymbol{\varTheta}{\mathalpha}{egreeki}{"02}
  \DeclareMathSymbol{\varIota}{\mathalpha}{egreeki}{"49}
  \DeclareMathSymbol{\varKappa}{\mathalpha}{egreeki}{"4B}
-\DeclareMathSymbol{\varLambda}{\mathalpha}{egreeki}{"03}
-\DeclareMathSymbol{\varMu}{\mathalpha}{egreeki}{"4D}
-\DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
-\DeclareMathSymbol{\varXi}{\mathalpha}{egreeki}{"04}
-\DeclareMathSymbol{\varOmicron}{\mathalpha}{egreeki}{"4F}
-\DeclareMathSymbol{\varPi}{\mathalpha}{egreeki}{"05}
-\DeclareMathSymbol{\varRho}{\mathalpha}{egreeki}{"50}
-\DeclareMathSymbol{\varSigma}{\mathalpha}{egreeki}{"06}
-\DeclareMathSymbol{\varTau}{\mathalpha}{egreeki}{"54}
-\DeclareMathSymbol{\varUpsilon}{\mathalpha}{egreeki}{"07}
-\DeclareMathSymbol{\varPhi}{\mathalpha}{egreeki}{"08}
-\DeclareMathSymbol{\varChi}{\mathalpha}{egreeki}{"58}
-\DeclareMathSymbol{\varPsi}{\mathalpha}{egreeki}{"09}
-\DeclareMathSymbol{\varOmega}{\mathalpha}{egreeki}{"0A} 
-%
-\DeclareMathSymbol{\Alpha}{\mathalpha}{egreekr}{"41}
-\DeclareMathSymbol{\Beta}{\mathalpha}{egreekr}{"42}
-\DeclareMathSymbol{\Gamma}{\mathalpha}{egreekr}{"00}
-\DeclareMathSymbol{\Delta}{\mathalpha}{egreekr}{"01}
-\DeclareMathSymbol{\Epsilon}{\mathalpha}{egreekr}{"45}
-\DeclareMathSymbol{\Zeta}{\mathalpha}{egreekr}{"5A}
-\DeclareMathSymbol{\Eta}{\mathalpha}{egreekr}{"48}
-\DeclareMathSymbol{\Theta}{\mathalpha}{egreekr}{"02}
-\DeclareMathSymbol{\Iota}{\mathalpha}{egreekr}{"49}
-\DeclareMathSymbol{\Kappa}{\mathalpha}{egreekr}{"4B}
-\DeclareMathSymbol{\Lambda}{\mathalpha}{egreekr}{"03}
-\DeclareMathSymbol{\Mu}{\mathalpha}{egreekr}{"4D}
-\DeclareMathSymbol{\Nu}{\mathalpha}{egreekr}{"4E}
-\DeclareMathSymbol{\Xi}{\mathalpha}{egreekr}{"04}
-\DeclareMathSymbol{\Omicron}{\mathalpha}{egreekr}{"4F}
-\DeclareMathSymbol{\Pi}{\mathalpha}{egreekr}{"05}
-\DeclareMathSymbol{\Rho}{\mathalpha}{egreekr}{"50}
-\DeclareMathSymbol{\Sigma}{\mathalpha}{egreekr}{"06}
-\DeclareMathSymbol{\Tau}{\mathalpha}{egreekr}{"54}
-\DeclareMathSymbol{\Upsilon}{\mathalpha}{egreekr}{"07}
-\DeclareMathSymbol{\Phi}{\mathalpha}{egreekr}{"08}
-\DeclareMathSymbol{\Chi}{\mathalpha}{egreekr}{"58}
-\DeclareMathSymbol{\Psi}{\mathalpha}{egreekr}{"09}
-\DeclareMathSymbol{\Omega}{\mathalpha}{egreekr}{"0A}
-%
-\DeclareMathSymbol{\alphaup}{\mathalpha}{egreekr}{"0B}
-\DeclareMathSymbol{\betaup}{\mathalpha}{egreekr}{"0C}
-\DeclareMathSymbol{\gammaup}{\mathalpha}{egreekr}{"0D}
-\DeclareMathSymbol{\deltaup}{\mathalpha}{egreekr}{"0E}
-\DeclareMathSymbol{\epsilonup}{\mathalpha}{egreekr}{"0F}
-\DeclareMathSymbol{\zetaup}{\mathalpha}{egreekr}{"10}
-\DeclareMathSymbol{\etaup}{\mathalpha}{egreekr}{"11}
-\DeclareMathSymbol{\thetaup}{\mathalpha}{egreekr}{"12}
-\DeclareMathSymbol{\iotaup}{\mathalpha}{egreekr}{"13}
-\DeclareMathSymbol{\kappaup}{\mathalpha}{egreekr}{"14}
-\DeclareMathSymbol{\lambdaup}{\mathalpha}{egreekr}{"15}
-\DeclareMathSymbol{\muup}{\mathalpha}{egreekr}{"16}
-\DeclareMathSymbol{\nuup}{\mathalpha}{egreekr}{"17}
-\DeclareMathSymbol{\xiup}{\mathalpha}{egreekr}{"18}
-\DeclareMathSymbol{\omicronup}{\mathalpha}{egreekr}{"6F}
- \DeclareMathSymbol{\piup}{\mathalpha}{egreekr}{"19}
-\DeclareMathSymbol{\rhoup}{\mathalpha}{egreekr}{"1A}
-\DeclareMathSymbol{\sigmaup}{\mathalpha}{egreekr}{"1B}
-\DeclareMathSymbol{\tauup}{\mathalpha}{egreekr}{"1C}
-\DeclareMathSymbol{\upsilonup}{\mathalpha}{egreekr}{"1D}
-\DeclareMathSymbol{\phiup}{\mathalpha}{egreekr}{"1E}
-\DeclareMathSymbol{\chiup}{\mathalpha}{egreekr}{"1F}
-\DeclareMathSymbol{\psiup}{\mathalpha}{egreekr}{"20}
-\DeclareMathSymbol{\omegaup}{\mathalpha}{egreekr}{"21}
-\DeclareMathSymbol{\varepsilonup}{\mathalpha}{egreekr}{"22}
-\DeclareMathSymbol{\varthetaup}{\mathalpha}{egreekr}{"23}
-\DeclareMathSymbol{\varpiup}{\mathalpha}{egreekr}{"24}
-\let\varrhoup\rhoup 
-\let\varsigmaup\sigmaup
-\let\varkappaup\kappaup
-\DeclareMathSymbol{\varphiup}{\mathalpha}{egreekr}{"27}
+% \DeclareMathSymbol{\varLambda}{\mathalpha}{egreeki}{"03}
+% \DeclareMathSymbol{\varMu}{\mathalpha}{egreeki}{"4D}
+% \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
+% \DeclareMathSymbol{\varXi}{\mathalpha}{egreeki}{"04}
+% \DeclareMathSymbol{\varOmicron}{\mathalpha}{egreeki}{"4F}
+% \DeclareMathSymbol{\varPi}{\mathalpha}{egreeki}{"05}
+% \DeclareMathSymbol{\varRho}{\mathalpha}{egreeki}{"50}
+% \DeclareMathSymbol{\varSigma}{\mathalpha}{egreeki}{"06}
+% \DeclareMathSymbol{\varTau}{\mathalpha}{egreeki}{"54}
+% \DeclareMathSymbol{\varUpsilon}{\mathalpha}{egreeki}{"07}
+% \DeclareMathSymbol{\varPhi}{\mathalpha}{egreeki}{"08}
+% \DeclareMathSymbol{\varChi}{\mathalpha}{egreeki}{"58}
+% \DeclareMathSymbol{\varPsi}{\mathalpha}{egreeki}{"09}
+% \DeclareMathSymbol{\varOmega}{\mathalpha}{egreeki}{"0A} 
+% %
+% \DeclareMathSymbol{\Alpha}{\mathalpha}{egreekr}{"41}
+% \DeclareMathSymbol{\Beta}{\mathalpha}{egreekr}{"42}
+% \DeclareMathSymbol{\Gamma}{\mathalpha}{egreekr}{"00}
+% \DeclareMathSymbol{\Delta}{\mathalpha}{egreekr}{"01}
+% \DeclareMathSymbol{\Epsilon}{\mathalpha}{egreekr}{"45}
+% \DeclareMathSymbol{\Zeta}{\mathalpha}{egreekr}{"5A}
+% \DeclareMathSymbol{\Eta}{\mathalpha}{egreekr}{"48}
+% \DeclareMathSymbol{\Theta}{\mathalpha}{egreekr}{"02}
+% \DeclareMathSymbol{\Iota}{\mathalpha}{egreekr}{"49}
+% \DeclareMathSymbol{\Kappa}{\mathalpha}{egreekr}{"4B}
+% \DeclareMathSymbol{\Lambda}{\mathalpha}{egreekr}{"03}
+% \DeclareMathSymbol{\Mu}{\mathalpha}{egreekr}{"4D}
+% \DeclareMathSymbol{\Nu}{\mathalpha}{egreekr}{"4E}
+% \DeclareMathSymbol{\Xi}{\mathalpha}{egreekr}{"04}
+% \DeclareMathSymbol{\Omicron}{\mathalpha}{egreekr}{"4F}
+% \DeclareMathSymbol{\Pi}{\mathalpha}{egreekr}{"05}
+% \DeclareMathSymbol{\Rho}{\mathalpha}{egreekr}{"50}
+% \DeclareMathSymbol{\Sigma}{\mathalpha}{egreekr}{"06}
+% \DeclareMathSymbol{\Tau}{\mathalpha}{egreekr}{"54}
+% \DeclareMathSymbol{\Upsilon}{\mathalpha}{egreekr}{"07}
+% \DeclareMathSymbol{\Phi}{\mathalpha}{egreekr}{"08}
+% \DeclareMathSymbol{\Chi}{\mathalpha}{egreekr}{"58}
+% \DeclareMathSymbol{\Psi}{\mathalpha}{egreekr}{"09}
+% \DeclareMathSymbol{\Omega}{\mathalpha}{egreekr}{"0A}
+% %
+% \DeclareMathSymbol{\alphaup}{\mathalpha}{egreekr}{"0B}
+% \DeclareMathSymbol{\betaup}{\mathalpha}{egreekr}{"0C}
+% \DeclareMathSymbol{\gammaup}{\mathalpha}{egreekr}{"0D}
+% \DeclareMathSymbol{\deltaup}{\mathalpha}{egreekr}{"0E}
+% \DeclareMathSymbol{\epsilonup}{\mathalpha}{egreekr}{"0F}
+% \DeclareMathSymbol{\zetaup}{\mathalpha}{egreekr}{"10}
+% \DeclareMathSymbol{\etaup}{\mathalpha}{egreekr}{"11}
+% \DeclareMathSymbol{\thetaup}{\mathalpha}{egreekr}{"12}
+% \DeclareMathSymbol{\iotaup}{\mathalpha}{egreekr}{"13}
+% \DeclareMathSymbol{\kappaup}{\mathalpha}{egreekr}{"14}
+% \DeclareMathSymbol{\lambdaup}{\mathalpha}{egreekr}{"15}
+% \DeclareMathSymbol{\muup}{\mathalpha}{egreekr}{"16}
+% \DeclareMathSymbol{\nuup}{\mathalpha}{egreekr}{"17}
+% \DeclareMathSymbol{\xiup}{\mathalpha}{egreekr}{"18}
+% \DeclareMathSymbol{\omicronup}{\mathalpha}{egreekr}{"6F}
+%  \DeclareMathSymbol{\piup}{\mathalpha}{egreekr}{"19}
+% \DeclareMathSymbol{\rhoup}{\mathalpha}{egreekr}{"1A}
+% \DeclareMathSymbol{\sigmaup}{\mathalpha}{egreekr}{"1B}
+% \DeclareMathSymbol{\tauup}{\mathalpha}{egreekr}{"1C}
+% \DeclareMathSymbol{\upsilonup}{\mathalpha}{egreekr}{"1D}
+% \DeclareMathSymbol{\phiup}{\mathalpha}{egreekr}{"1E}
+% \DeclareMathSymbol{\chiup}{\mathalpha}{egreekr}{"1F}
+% \DeclareMathSymbol{\psiup}{\mathalpha}{egreekr}{"20}
+% \DeclareMathSymbol{\omegaup}{\mathalpha}{egreekr}{"21}
+% \DeclareMathSymbol{\varepsilonup}{\mathalpha}{egreekr}{"22}
+% \DeclareMathSymbol{\varthetaup}{\mathalpha}{egreekr}{"23}
+% \DeclareMathSymbol{\varpiup}{\mathalpha}{egreekr}{"24}
+% \let\varrhoup\rhoup 
+% \let\varsigmaup\sigmaup
+% \let\varkappaup\kappaup
+% \DeclareMathSymbol{\varphiup}{\mathalpha}{egreekr}{"27}
 
@@
 \centering
-%\includegraphics[width=0.75\linewidth]{pop_sample_projection.pdf}\\
-\includegraphics[width=0.75\linewidth]{pop_sample_projection2.pdf}%
+\includegraphics[width=0.99\linewidth]{pop_sample_projection3.pdf}%
 \caption{Log-plot of the hypergeometric distribution
@@
   level followed by sample marginalization (blue triangles),
-  \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:sample_av_example_ME}, with $N=5000$,
+  \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
+  circles), \eqn~\eqref{eq:app_maxent_sample}, with $N=5000$,
   $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$.}
@@
   the population level followed by sample marginalization (blue triangles),
-  \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:sample_av_example_ME}, with $N=5000$,
+  \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
+  circles), \eqn~\eqref{eq:app_maxent_sample}, with $N=5000$,
   $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$,
@@
 
+\iffalse
 \newrefsection
@@
 
-\iffalse
 \begin{align}
@@
 \end{subequations}
-\fi
 
@@
 literature.
-
+\fi
 \end{document}

### 2017-05-18 maxNt.tex
discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-18T12:03:23+0200
+%% Last-Updated: 2017-05-18T13:36:45+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \DeclareMathSymbol{\varMu}{\mathalpha}{egreeki}{"4D}
-% \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
+ \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
 % \DeclareMathSymbol{\varXi}{\mathalpha}{egreeki}{"04}
@@
 
+\newcommand*{\ypp}{\varPi}
+
 \newcommand*{\yxx}{x}%subpop state
@@
 \newcommand*{\yHc}{\varKappa}
+\newcommand*{\yHd}{\Hat{\varIota}}
 %Lagr multipliers
@@
 
-\newcommand*{\ypp}{\varPi}
-
 \newcommand*{\prop}[1]{`#1'}
@@
   \And
-E. Torre\\Moneylaunderingland  
+E. Torre\\ETH Z{\"u}rich\\Switzerland\\
+\texttt{torre@ibk.baug.ethz.ch}
   % David S.~Hippocampus\thanks{Use footnote for providing further
@@
 
-\section{Initial assumptions: the probability of random representative samples}
+\section{Initial assumptions: the probability of representative samples}
 \label{sec:prob_samples}
@@
 For functions representing averaged products,
-$f(\yxs) \defd \sav{\yx\dots\yx}=\binom{\yxs}{m}/\binom{n}{m}$,
+$f(\yxs) \defd \sav{\yx\dots\yx}=\binom{n\yxs}{m}/\binom{n}{m}$,
 formulae~\eqref{eq:pullback_P} have the useful form
@@
 
-%\setlength{\intextsep}{0.5ex}%
-%\begin{wrapfigure}{r}{0.5\linewidth}
 \begin{figure}[!t]
@@
 \figs~\ref{fig:diff_maxent_pop_sample}
-and~\ref{fig:diff_maxent_pop_sample_realdata}, for $N=5000$, $n=200$. In
-the first example the constraints are $\expe{\yxs} = c_1$ and
-$\expe{\yxxs} = c_2$, with $c_1=0.0478$ and $c_2=0.00257$. The distribution
-obtained at the sample level is broader than the one obtained at the
-population level; the tails of the two distributions are very different.
-The second example includes two additional constraints
-$\expe{\sav{\yx\yx\yx}}= c_3$, $\expe{\sav{\yx\yx\yx\yx}}=c_4$ with
-$c_3=0.000148$, $c_4=8.81\times 10^{-6}$. The distribution obtained at the
-population level has two modes, replaced by only one in the distribution
-obtained at the sample level; the tails are very different also in this
-case. The constraints used in these examples have neurobiologically
-realistic values \cite{rostamietal2016}.
-
-\medskip
+and~\ref{fig:diff_maxent_pop_sample_realdata}, for $N=5000$, $n=200$, and
+constraint functions of the form
+$\yf(\yxs)=(\yxs,\yxxs,\dotsc)\equiv\bigl(
+\yxs, \raisebox{0pt}[0pt][1ex]{}\binom{n\yxs}{2}/\binom{n}{2},\dotsc \bigr)$. In the first example the
+constraints are $\expe{\yxs} = c_1$ and $\expe{\yxxs} = c_2$, with
+$c_1=0.0478$ and $c_2=0.00257$. The distribution obtained at the sample
+level is broader than the one obtained at the population level; the tails
+of the two distributions are very different. The second example includes
+two additional constraints $\expe{\sav{\yx\yx\yx}}= c_3$,
+$\expe{\sav{\yx\yx\yx\yx}}=c_4$ with $c_3=0.000148$,
+$c_4=8.81\times 10^{-6}$. The distribution obtained at the population level
+has two modes, replaced by only one in the distribution obtained at the
+sample level; the tails are very different also in this case. The
+constraints used in these examples have neurobiologically realistic values
+\cite{rostamietal2016}.
+
+\bigskip
 
 How should we apply the \me\ method then? on the sample or on the
-population? Which one is maximally noncommittal?
+population? Which application is maximally noncommittal?
 
@@
 
-The first purpose of this note was indeed to remind ourselves that
-probability models require a continuous scrutiny, because they can rest on
+The first purpose of this note is indeed to remind ourselves that
+probability models require careful scrutiny, because they can rest on
 hidden assumptions that we don't want to make or that contradict others we
-are making. The dilemma generated by the \me\ method and the assumption of
-random representative sampling is an example. The usual application of \me\
-at the sample level yields a different distribution from the application at
-the population level. The former application must therefore implicitly
-assume that no larger population exist, or that the neurons of the sample
-are isolated from their environment.
-
-The \me\ application at the population level has in fact by construction a
-lower entropy than the one obtained by application at the sample level:
-both satisfy the same constraints, but only the latter is obtained by
-direct maximization. The lower entropy of the distribution obtained from
-application at the population level is clearly visible from its higher
-sharpness in \fig~\ref{fig:diff_maxent_pop_sample}. This distribution
-therefore says that some sample states $\yx$ should be considered less
-likely than in the distribution obtained at the sample level. A lower
-entropy signals the presence of additional assumptions or constraints. In
-this case the assumption is that the sample is part of a larger population,
-and our ignorance about the sample should be representative of our
-ignorance about a larger population. This assumption affects our
-predictions.
+are making, and we may not be aware of this. The dilemma generated by the
+\me\ method and the assumption of representative sampling is an example.
+
+The tricky point is this. Maximum-entropy applied at the population level
+and applied at the sample level give different results; they are different
+statistical models. The former model clearly assumes, by construction, the
+existence of a larger population from which the sample is taken. What does
+the latter model assume in this respect? is it \enquote{unassuming}, as
+often claimed in the literature? or is it actually assuming that \emph{no}
+larger population exists? In the latter case it would not be the correct
+model to use in our problem.
+
+A perfunctory intuitive reasoning seems insufficient for clarifying this
+point. Suppose we do not know whether the sample is really part of a larger
+population; we do not know if $N=n$ or how large is $N$ otherwise. Call
+this state of ignorance $\yHd$. From the
+point of view of probability theory, th
+ ignorance about the
+value of $N$ is expressed by assigning a probability distribution
+$\p(N \| \yH)$ that vanishes if $N<n$, since we know that $N\ge n$. Let's
+call thi
+Probability assignments for the sample state that do not assume a specific
+value of $N$ are then obtained by marginalization
+
+The former application must therefore implicitly assume that no
+larger population exist, or that the neurons of the sample are isolated
+from their environment.
+
+The \me\ application at the population level gives in fact, by
+construction, a distribution with lower Shannon relative entropy than the
+one given by the application at the sample level: both satisfy the same
+constraints, but only the latter is obtained by direct maximization of the
+entropy. The lower entropy of the distribution obtained from application at
+the population level is clearly visible from its higher sharpness in
+\fig~\ref{fig:diff_maxent_pop_sample}. Such distribution therefore
+excludes, probabilistically speaking, more sample states $\yx$ than the
+distribution obtained at the sample level does. A lower entropy signals the
+presence of additional assumptions or constraints. In this case the
+assumption is that the sample is part of a larger population, and our
+ignorance about the sample should be representative of our ignorance about
+a larger population. This assumption affects our predictions.
 
@@
 \end{figure}%maxent_pop_or_sample.nb
-The second is that
-\fi
-The possibility of using two different distributions is
-not a physical contradiction. Similar situations arise in statistical
+The second is that \fi The possibility of using two different distributions
+is not a physical contradiction. Similar situations arise in statistical
 mechanics. It is known that if a system is described by a \me\ Gibbs state,
-its subsystems need not be \cite{maesetal1999}. A quandary somehow similar
-to ours also appears in the statistical description of an equilibrium state
-at the end of a non-equilibrium process: we can describe our knowledge
-about it either by a Gibbs distribution, or by the Liouville-evolved Gibbs
-distribution associated with equilibrium state at the beginning of the
-process. The two descriptions differ -- even though the final physical
-state is exactly the same \cite[\sect~4]{jaynes1985d_r1993}. The appearance
-of this difference is natural: in one case we can make sharper predictions
-about the state thanks to our knowledge of its preceding dynamics.
-%\mynote{but in this case both distributions are immensely sharp, and this
-%  doesn't affect predictions}
+its subsystems need not be \cite{maesetal1999}. A dilemma quite similar to
+ours also appears in the statistical description of the final state of a
+non-equilibrium process starting and ending in two equilibrium states: we
+can describe our knowledge about the final state either by a Gibbs
+distribution, or by the distribution obtained from the Liouville evolution
+of the Gibbs distribution assigned to the initial state. The two
+descriptions differ -- even though the final \emph{physical} state is
+obviously exactly the same \cite[\sect~4]{jaynes1985d_r1993}. The
+difference in the two descriptions appears because in one case we can make
+sharper predictions about the state thanks to our knowledge of its
+preceding dynamics. In this example, though, both distributions are usually
+immensely sharp and practically lead to the same predictions. In the cases
+considered in this note the difference in predictions may be relevant, and
+the ones made by \me\ at the population level are more informed.
 

### 2017-05-18 maxNt.tex
discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-18T13:36:45+0200
+%% Last-Updated: 2017-05-18T14:02:33+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \DeclareMathSymbol{\varMu}{\mathalpha}{egreeki}{"4D}
- \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
+% \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
 % \DeclareMathSymbol{\varXi}{\mathalpha}{egreeki}{"04}
@@
 % \DeclareMathSymbol{\varTau}{\mathalpha}{egreeki}{"54}
-% \DeclareMathSymbol{\varUpsilon}{\mathalpha}{egreeki}{"07}
+ \DeclareMathSymbol{\varUpsilon}{\mathalpha}{egreeki}{"07}
 % \DeclareMathSymbol{\varPhi}{\mathalpha}{egreeki}{"08}
@@
 \newcommand*{\yHc}{\varKappa}
-\newcommand*{\yHd}{\Hat{\varIota}}
+\newcommand*{\yHd}{\varUpsilon}
 %Lagr multipliers
@@
 A perfunctory intuitive reasoning seems insufficient for clarifying this
-point. Suppose we do not know whether the sample is really part of a larger
-population; we do not know if $N=n$ or how large is $N$ otherwise. Call
-this state of ignorance $\yHd$. From the
-point of view of probability theory, th
- ignorance about the
-value of $N$ is expressed by assigning a probability distribution
-$\p(N \| \yH)$ that vanishes if $N<n$, since we know that $N\ge n$. Let's
-call thi
-Probability assignments for the sample state that do not assume a specific
-value of $N$ are then obtained by marginalization
-
+point. Let's express it via the probability calculus. Suppose we do not
+know whether the sample is really part of a larger population; we do not
+know if $N=n$ or how large is $N$ otherwise. Call this state of ignorance
+$\yHd$. From the point of view of the probability calculus, this ignorance
+about $N$ is expressed by assigning a probability distribution
+$\p(N \| \yHd)$ that vanishes if $N<n$, since we know that $N\ge n$.
+Maintaining our assumption of symmetric ignorance, probability assignments
+that do not assume a specific value of $N$ are then obtained by
+multiplication of all $N$-dependent probabilities by $\p(N \| \yHd)$ and
+subsequent marginalization over $N$ -- technically speaking, $N$ is a
+\emph{nuisance parameter} that we eliminate. The probability obtained from
+\me\ at the population level, \eqn~\eqref{eq:app_maxent_pop_marg} then
+generalizes to
+\begin{equation}
+  \label{eq:app_maxent_pop_marg_unknown_N}
+  \p(\yxs = \yrs \| \yHd)  = \sum_N \biggl\{\yK_N\, 
+  \sum_{N\yRf=0}^N \ypp_N(\yrs \|\yRf)
+\,
+  \binom{N}{N \yRf}\exp\Bigl[{\yL_N}\T
+  \sum_{n\yrs=0}^n \yf(\yrs)\ypp_N(\yrs\|\yRf)\Bigr]\biggr\}
+  \,\p(N \| \yHd).
+\end{equation}
+This is a formidable expression. But the answer to our question, whether
+the usual \me\ at the sample level~\eqref{eq:app_maxent_sample} does not
+assume anything about a larger population, translates into the precise
+mathematical question: are the
+distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
+\eqref{eq:app_maxent_sample} equal, for some choice of $\p(N \| \yHd)$? We
+leave this mathematical problem for future work.
+
+\iffalse
 The former application must therefore implicitly assume that no
@@
 a larger population. This assumption affects our predictions.
+\fi
 

### 2017-05-18 maxNt.tex
added some references

--- 2017-05-18 maxNt.tex
@@
 
-Recent electrophysiological techniques \cite{***} in fact allow
+Recent electrophysiological techniques \cite{berenyietal2014} in fact allow
 experimenters to record samples of even a couple hundreds neurons from
@@
 compression: it associates a unique probability distribution with few
-experimental quantities \cite{***}. But this is only one of its uses. It is
+experimental quantities \cite{***}.
+But this is only one of its uses. It is
 also used for various information-theoretic purposes or to generate
-reference probability distributions \cite{***}. In all these uses the \me\
+reference probability distributions \cite{schneidmanetal2006,shlensetal2006,roudietal2009b,tkaciketal2014b}.
+In all these uses the \me\
 distribution is chosen as the \enquote{maximally noncommittal} one

### 2017-05-18 maxNt.tex
discussion

--- 2017-05-18 maxNt.tex
@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-18T14:02:33+0200
+%% Last-Updated: 2017-05-18T14:18:19+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 $\yf = (f_1,\dotsc,f_m)$ to the vector values $\yc=(c_1,\dotsc,c_m)$, that
-is, $\expeb{\yf(\yxs)}=\yc$. Application of \me\ at the population level,
-denoted by $\yHa$, gives
+is, $\expeb{\yf(\yxs)}=\yc$. Application of \me\
+\cite{sivia1996_r2006,meadetal1984} at the population level, denoted by
+$\yHa$, gives
 \begin{equation}
@@
 $\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$ -- that is, $2m+n$
-equations in $m$ unknowns. This could be possible, if at all, only for very
-special choices of constraints functions $\yf$ and values $\yc$.
+equations (normalization is taken care of) in $2m$ unknowns. A solution can
+exist, if at all, only for very special choices of constraints functions
+$\yf$ and values $\yc$.
 
@@
 A perfunctory intuitive reasoning seems insufficient for clarifying this
-point. Let's express it via the probability calculus. Suppose we do not
-know whether the sample is really part of a larger population; we do not
-know if $N=n$ or how large is $N$ otherwise. Call this state of ignorance
-$\yHd$. From the point of view of the probability calculus, this ignorance
-about $N$ is expressed by assigning a probability distribution
-$\p(N \| \yHd)$ that vanishes if $N<n$, since we know that $N\ge n$.
-Maintaining our assumption of symmetric ignorance, probability assignments
-that do not assume a specific value of $N$ are then obtained by
-multiplication of all $N$-dependent probabilities by $\p(N \| \yHd)$ and
-subsequent marginalization over $N$ -- technically speaking, $N$ is a
-\emph{nuisance parameter} that we eliminate. The probability obtained from
-\me\ at the population level, \eqn~\eqref{eq:app_maxent_pop_marg} then
-generalizes to
+point. Let's express it in the language of the probability calculus.
+Suppose we do not know whether the sample is really part of a larger
+population: we do not know whether $N=n$ or how large is $N$ otherwise.
+Call this state of ignorance $\yHd$. From the point of view of the
+probability calculus, this ignorance about $N$ is expressed by assigning a
+probability distribution $\p(N \| \yHd)$ that vanishes if $N<n$, since we
+know that $N\ge n$. Maintaining our assumption of symmetric ignorance,
+probability assignments that do not assume a specific value of $N$ are then
+obtained by multiplication of all $N$-dependent probabilities by
+$\p(N \| \yHd)$ and subsequent marginalization over $N$ -- technically
+speaking, $N$ is a \emph{nuisance parameter} that we eliminate
+\cite{jaynes1994_r2003,lindley1965b_r2008,bernardoetal1994_r2000}. The
+probability obtained from \me\ at the population level,
+\eqn~\eqref{eq:app_maxent_pop_marg} then generalizes to
 \begin{equation}

### 2017-05-18 maxNt.tex
discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-18T14:18:19+0200
+%% Last-Updated: 2017-05-18T16:09:42+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\usepackage[final,nonatbib]{nips_2017}
+\usepackage[%final,
+nonatbib]{nips_2017}
+
 \newcommand*{\firstdraft}{4 November 2015}
-\newcommand*{\pdftitle}{Representative random samples\\ and maximum-entropy
-  distributions:\\ a dilemma}
-\newcommand*{\headtitle}{\pdftitle}
-\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, V. Rostami, E. Torre}
+
+\newcommand*{\pdftitle}{Representative samples\\ and maximum-entropy distributions\\ in neuroscience: a dilemma}
+
+\newcommand*{\headtitle}{\pdftitle} \newcommand*{\pdfauthor}{P.G.L. Porta Mana, V. Rostami, E. Torre}
 
@@
 methods, but we would be pleased if its discussion of the probability of
-\enquote{random sampling} were useful to neuroscientists that use other
+\enquote{sampling} were useful to neuroscientists that use other
 statistical methods to study the physical and dynamical characteristics of
@@
 unknown process, but from their observation we expect to learn something
-about all other neurons in the same brain area. That is, we assumed
+about all other neurons in the same brain area. That is, we assume
 that they are a \enquote{representative random sample}.
@@
 We don't elaborate on the various important purposes of such recordings
-here \cite{?***}, but stress that these sample sizes can be considered
+here, but stress that these sample sizes can be considered
 \enquote{large}, because their statistical analysis requires considerable
@@
 compression: it associates a unique probability distribution with few
-experimental quantities \cite{***}.
-But this is only one of its uses. It is
-also used for various information-theoretic purposes or to generate
-reference probability distributions \cite{schneidmanetal2006,shlensetal2006,roudietal2009b,tkaciketal2014b}.
-In all these uses the \me\
-distribution is chosen as the \enquote{maximally noncommittal} one
-\cite{jaynes1963}. This adjective means little without further technical
-characterizations. Different works give different characterizations, but in
-this paper we will use the quoted expression as an umbrella term for all of
-them. Our results will not depend on the specific characterization of
-\enquote{noncommittal}.
+experimental quantities. But this is only one of its uses. It is also used
+for various information-theoretic purposes or to generate reference
+probability distributions
+\cite{bohteetal2000,schneidmanetal2006,shlensetal2006,roudietal2009b,mackeetal2011,tkaciketal2014b,shimazakietal2015}.
+In all these uses the \me\ distribution is chosen as the \enquote{maximally
+  noncommittal} one \cite{jaynes1963}. This adjective means little without
+further technical characterizations. Different works give different
+characterizations, but in this paper we will use the quoted expression as
+an umbrella term for all of them. Our results will not depend on the
+specific characterization of \enquote{noncommittal}.
 
@@
 In this note we want to show that we have to face a dilemma if we want to
-apply the \me\ method to a representative random sample to find a maximally
+apply the \me\ method to a representative  sample to find a maximally
 noncommittal distribution.
@@
 purpose we also present some probability relations relevant to
-\enquote{random sampling}. These relations are well-known in survey
+sampling. These relations are well-known in survey
 sampling and in the pedagogic problem of drawing from an urn without
@@
   $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
-\ypp(\yrs\|\yRf) \defd  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}$ for $N=5000$, $n=200$. (Band artifacts may appear in the
+\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}$ for $N=5000$, $n=200$. (Band artifacts may appear in the
   colourbar depending on your \textsc{pdf} viewer.)}
@@
   \p(\yXf = \yRf \| \yHa)  = \yK\,
-  \binom{N}{N \yRf}\exp\Bigl[\yL\T
+  \binom{N}{N \yRf}\,\exp\Bigl[\yL\T
   \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
@@
 \,
-  \binom{N}{N \yRf}\exp\Bigl[\yL\T
+  \binom{N}{N \yRf}\,\exp\Bigl[\yL\T
   \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
@@
 \,
-  \binom{N}{N \yRf}\exp\Bigl[\yL\T
+  \binom{N}{N \yRf}\,\exp\Bigl[\yL\T
   \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
@@
 population: we do not know whether $N=n$ or how large is $N$ otherwise.
-Call this state of ignorance $\yHd$. From the point of view of the
-probability calculus, this ignorance about $N$ is expressed by assigning a
-probability distribution $\p(N \| \yHd)$ that vanishes if $N<n$, since we
-know that $N\ge n$. Maintaining our assumption of symmetric ignorance,
-probability assignments that do not assume a specific value of $N$ are then
-obtained by multiplication of all $N$-dependent probabilities by
-$\p(N \| \yHd)$ and subsequent marginalization over $N$ -- technically
-speaking, $N$ is a \emph{nuisance parameter} that we eliminate
+Call this state of ignorance $\yHd$. In the probability calculus, this
+ignorance about $N$ is expressed by assigning a probability distribution
+$\p(N \| \yHd)$ that vanishes if $N<n$, since we know that $N\ge n$; see
+Good \cite{good1965,good1967b} and Rissanen \cite{rissanen1983} for
+examples of such distributions over the integers. Maintaining our
+assumption of symmetric ignorance, probability assignments that do not
+assume a specific value of $N$ are then obtained by multiplication of all
+$N$-dependent probabilities by $\p(N \| \yHd)$ and subsequent
+marginalization over $N$. Technically speaking, $N$ becomes a
+\emph{nuisance parameter}
 \cite{jaynes1994_r2003,lindley1965b_r2008,bernardoetal1994_r2000}. The
 probability obtained from \me\ at the population level,
-\eqn~\eqref{eq:app_maxent_pop_marg} then generalizes to
+\eqn~\eqref{eq:app_maxent_pop_marg}, then generalizes to
 \begin{equation}
@@
 \,
-  \binom{N}{N \yRf}\exp\Bigl[{\yL_N}\T
+  \binom{N}{N \yRf}\,\exp\Bigl[{\yL_N}\T
   \sum_{n\yrs=0}^n \yf(\yrs)\ypp_N(\yrs\|\yRf)\Bigr]\biggr\}
-  \,\p(N \| \yHd).
-\end{equation}
-This is a formidable expression. But the answer to our question, whether
-the usual \me\ at the sample level~\eqref{eq:app_maxent_sample} does not
-assume anything about a larger population, translates into the precise
-mathematical question: are the
-distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
-\eqref{eq:app_maxent_sample} equal, for some choice of $\p(N \| \yHd)$? We
-leave this mathematical problem for future work.
+  \,\p(N \| \yHd),
+\end{equation}
+where $N$-dependencies have been made explicit. This is a formidable
+expression. But the answer to our question, whether the usual \me\ at the
+sample level~\eqref{eq:app_maxent_sample} does not assume anything about a
+larger population, translates now into the precise mathematical question:
+are the distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
+\eqref{eq:app_maxent_sample} equal for some choice of $\p(N \| \yHd)$, with
+$\p(N \| \yHd)\ne 0$ for $N>n$? We leave this mathematical problem for
+future work. Note, however, that this equality is satisfied if
+$\p(N=1 \| \yHd)=1$, which means that the usual \me\ model can also be
+interpreted as assuming that \emph{no} larger population exists.
 
@@
 
-In a neuroscientific context we find this assumption very natural, and
-therefore the \me\ method at the population level preferable. After all,
-also \emph{physical} neuronal-network models usually include some sort of
-external input to the neurons, mimicking their embedding in a larger
-network.
+Rough estimates of $N$ may be available in neuroscientific contexts, so we
+find the \me\ model constructed at the population level very natural and
+preferable. After all, \emph{physical} neuronal-network models usually
+include some sort of external input to the neurons as well, mimicking their
+embedding in a larger network. The distribution of this \me\ model, when
+used as a reference distribution for surprise analysis, may reveal features
+in a dataset that were unnoticed by the standard \me\ model.
 
@@
 \end{figure}%maxent_pop_or_sample.nb
-The second is that \fi The possibility of using two different distributions
-is not a physical contradiction. Similar situations arise in statistical
-mechanics. It is known that if a system is described by a \me\ Gibbs state,
-its subsystems need not be \cite{maesetal1999}. A dilemma quite similar to
-ours also appears in the statistical description of the final state of a
+The second is that \fi
+
+The possibility of using two different distributions is not a physical
+contradiction. Similar situations arise in statistical mechanics. It is
+known that if a system is described by a \me\ Gibbs state, its subsystems
+need not be \cite{maesetal1999}. A dilemma quite similar to ours also
+appears in the statistical description of the final state of a
 non-equilibrium process starting and ending in two equilibrium states: we
@@
 immensely sharp and practically lead to the same predictions. In the cases
-considered in this note the difference in predictions may be relevant, and
-the ones made by \me\ at the population level are more informed.
-
-\mynote{Importance of the formulae relating sample and population.}
-
-\mynote{Discussion of case when individual $X_i\dotsm X_j$ are constrained.
-Likely breakdown of \me\ in this case \cite{portamana2009}. Hints at full
-Bayesian treatment of which \me\ is a limit? \cite[***]{mackay1995_r2003}}
+considered in this note the difference in predictions may be relevant
+instead.
+
+Our analysis touched only constraints of the sample average. The
+corresponding models are usually called \enquote{homogeneous} in the
+literature. Purely \enquote{inhomogeneous} models have also been used
+\cite{schneidmanetal2006,shlensetal2006,roudietal2009b}, in which
+expectations for individual neurons or groups of neurons are constrained,
+for example $\expe{x_2}$ or $\expe{x_1x_8x_9}$. A short computation shows
+that the \me\ method with this kind of constraints gives the same result
+whether applied at the sample or at the population level: the states of any
+unconstrained neurons marginalize out. This is understandable: expressing
+different uncertainties about neurons $2$ and $5$ we are breaking the
+symmetry of our uncertainty, which thus cannot be representative of other
+neurons in the sample or in the population.
+
+Inhomogeneous models, however, require enormous computational power for
+large sample sizes; homogeneous models therefore retain their importance.
+Our analysis and dilemma also persist for hybrid homogeneous-inhomogeneous
+models \cite{tkaciketal2014b,shimazakietal2015}.
+
+%\mynote{Importance of the formulae relating sample and population.}
+
+% \mynote{Discussion of case when individual $X_i\dotsm X_j$ are constrained.
+% Likely breakdown of \me\ in this case \cite{portamana2009}. Hints at full
+% Bayesian treatment of which \me\ is a limit? \cite[***]{mackay1995_r2003}}
 

### 2017-05-19 maxNt.tex
discussion

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-18T16:09:42+0200
+%% Last-Updated: 2017-05-19T02:51:01+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\usepackage[%final,
+\usepackage[%final,%
 nonatbib]{nips_2017}
@@
 \newcommand*{\yD}{\varDelta}
-\newcommand*{\yHa}{\varIota'}
-\newcommand*{\yHb}{\varIota''}
+\newcommand*{\yHa}{\varIota_\textrm{p}}
+\newcommand*{\yHb}{\varIota_\textrm{s}}
 \newcommand*{\yHc}{\varKappa}
 \newcommand*{\yHd}{\varUpsilon}
+\newcommand*{\yHp}{\varIota'}
+\newcommand*{\yHi}{\varIota''}
 %Lagr multipliers
@@
 Recent electrophysiological techniques \cite{berenyietal2014} in fact allow
-experimenters to record samples of even a couple hundreds neurons from
-specific brain areas. These samples are usually picked out according to an
-unknown process, but from their observation we expect to learn something
-about all other neurons in the same brain area. That is, we assume
-that they are a \enquote{representative random sample}.
-
-We don't elaborate on the various important purposes of such recordings
-here, but stress that these sample sizes can be considered
-\enquote{large}, because their statistical analysis requires considerable
-computational power.
+experimenters to record samples comprising even a couple hundreds neurons
+from specific brain areas. From the observation of these samples we expect
+to learn something about all other neurons in the same brain area. That is,
+we assume that they are a \enquote{representative sample}. We don't
+elaborate on the various important purposes of such recordings here, but
+stress that these sample sizes can be considered \enquote{large}, because
+their statistical analysis requires considerable computational power.
 
@@
 
-In this note we want to show that we have to face a dilemma if we want to
-apply the \me\ method to a representative  sample to find a maximally
-noncommittal distribution.
+In this note we show that we must face a dilemma if we want to apply the
+\me\ method to a representative sample to find a maximally noncommittal
+distribution.
 
@@
 
+In the final discussion we present some arguments in favour of one choice
+in our dilemma.
+
 In the rest of the paper we mathematically formulate this dilemma. To this
-purpose we also present some probability relations relevant to
-sampling. These relations are well-known in survey
-sampling and in the pedagogic problem of drawing from an urn without
-replacement, yet they are somewhat hard to find explicitly written in the
-neuroscientific literature, so they may be of interest on their own.
-
-In the final discussion we present some arguments in favour of one choice
+purpose we also present some probability relations relevant to the sampling
+problem. The relations we presentare well-known in survey sampling and in
+the pedagogic problem of drawing from an urn without replacement, yet they
+are somewhat hard to find explicitly written in the neuroscientific
+literature, so they may be of interest on their own. We briefly discuss
+also two very different assumptions that lead to similar initial
+probability assignments for sampling.
+
+\iffalse In the final discussion we present some arguments in favour of one choice
 in our dilemma, but we do not mean that to be our final answer. This note,
@@
   answer with probability theory?} and \emph{which assumptions do we want
-  or need to make?}
+  or need to make?} \fi
 
@@
 We would also like our probability assignment to reflect that the sample is
-somehow \enquote{representative} of the population.
-
-Let's assume that we initially know very little about the physical details
-of the individual neurons; their locations for example. Our initial state
-of knowledge or ignorance $\yH$ is therefore symmetric, or
-\enquote{exchangeable}, under their permutations. This symmetry must be
-reflected in our initial probability: the \emph{representation theorem for
-  finite exchangeability} states that it must obey
+somehow \enquote{representative} of the population. We consider here two
+states of knowledge that express this representativeness in different ways
+but lead to identical \emph{initial} probability assignments.
+
+In the first, denoted $\yHp$, we know that the neurons in the population
+are physico-biologically similar, for example in morphology and kind of
+input they receive. Knowledge of this similarity leads us to assign a
+probability distribution for the population state $\yX$ that is symmetric,
+also called \emph{exchangeable}, under permutations of neuron identities.
+
+In the second, denoted $\yHi$, we are completely ignorant about the
+physical details of the individual neurons. Our ignorance is therefore
+symmetric under permutations of neuron identities. This also leads to an
+exchangeable probability distribution for $\yX$.
+
+Let us use $\yH$ to denote either of these two states of knowledge in those
+probabilities that are identical in both.
+
+The \emph{representation theorem for finite exchangeability} states that
+the symmetric distribution of $\yH$ must obey
 \begin{equation}
   \label{eq:joint_plaus_N_homog}
-  \p(\yX = \yR \| \yH) = \binom{N}{N\yRf}^{-1} \p(\yXf=\yRf \| \yH),
+  \p(\yX = \yR \|  \yH) \equiv
+  \p(\yX = \yR \| \yXf = \yRf, \yH)\,
+  \p(\yXf=\yRf \| \yH) =
+  \binom{N}{N\yRf}^{-1} \p(\yXf=\yRf \| \yH),
 \end{equation}
@@
 this theorem and generalizations to non-binary and continuum cases are
-given by de~Finetti \cite{definetti1959b}, Ericson \cite{ericson1976},
-Diaconis \cite{diaconis1977}, Heath \amp\ Sudderth \cite{heathetal1976}.
-This theorem is intuitive: owing to symmetry, we must assign equal
-probabilities to all states with $N\yRf$ active neurons.
+given by de~Finetti \cite{definetti1959b}, Kendall \cite{kendall1967},
+Ericson \cite{ericson1976}, Diaconis \amp\ Freedman
+\cite{diaconis1977,diaconisetal1980}, Heath \amp\ Sudderth
+\cite{heathetal1976}. This theorem is intuitive: owing to symmetry, we must
+assign equal probabilities to all states with $N\yRf$ active neurons.
 
@@
 
+How is it possible that the very different states of knowledge $\yHp$ and
+$\yHi$ lead to the same formulae above? Their difference appears as soon as
+we make an experimental observation, say $X_2 = R_2\in\set{0,1}$ and update our
+probabilities \eqref{eq:joint_plaus_N_homog},
+\begin{equation}
+  \label{eq:updated_joint}
+    \p(\yX = \yR \| X_2 = R_2, \yH) \equiv
+  \p(\yX = \yR \| \yXf = \yRf, X_2 = R_2, \yH)\,
+  \p(\yXf=\yRf \| X_2 = R_2,\yH)
+\end{equation}
+Both the conditional probability and the one for the average on the right
+side will update in very different ways for $\yHp$ and $\yHi$. The
+discussion of this update process in the two cases is outside the purposes
+of this note, and we hope to address it in full in a future work.
+
 The conditional probability $\ypp(\yrs,\yRf)$ relates the spaces of the
@@
 \centering
-\includegraphics[width=0.99\linewidth]{pop_sample_projection3.pdf}%
+\includegraphics[width=0.9\linewidth]{pop_sample_projection3.pdf}%
 \caption{Log-plot of the hypergeometric distribution
@@
 
-Yet, the approximate equalities above express the fact that \emph{our
+The approximate equalities above express the fact that \emph{our
   uncertainty about the sample is representative of our uncertainty about
-  the population and about other samples}, and vice versa. This is how the
-idea of representativeness is translated into our probabilities. The
-symmetry present in our initial probabilities is not a physical property of
-the neuronal population or sample. It only expresses the symmetry of our
-initial uncertainty about them, and does not imply any sort of physical
-similarity between the neurons. Subsequent observations may in fact break
-this symmetry. Upon observation of a sample average, say $\yxxs=a$, the
-updated expectations for such average in the population and in any new
-sample will usually be shifted towards the observed value, as follows from
-Bayes's theorem and the formulae above.
+  the population and about other samples}, and vice versa. This fact comes
+about for very different reasons in the states of knowledge $\yHp$ and
+$\yHi$. In $\yHp$, because our sample is \emph{physically} representative
+of the population and of other samples; we could write this as
+$\yxs\approx\yXf$. In $\yHi$, because we are ignorant about sample and
+population in similar symmetric ways; but this does not imply that
+$\yxs\approx\yXf$. New observations may in fact break this symmetry by
+\eqn~\eqref{eq:updated_joint}.
 
@@
 The proof uses the expression for the $m$th factorial moment of the
-hypergeometric distribution \cite{potts1953}. Thus, the averages of
-activity products \emph{are the same for the sample and for the full
-  population}. Similar relations can be found for the raw moments
-$\expe{\yxs^m}$ and $\expe{\yXf^m}$, which can be written in terms of the
-product expectations via \eqn~\eqref{eq:products_intermsof_average}.
+hypergeometric distribution \cite{potts1953}. Thus, in the states of
+knowledge $\yHp$ and $\yHi$ the averages of activity products \emph{are
+  initially the same for the sample and for the full population}. Similar
+relations can be found for the raw moments $\expe{\yxs^m}$ and
+$\expe{\yXf^m}$, which can be written in terms of the product expectations
+via \eqn~\eqref{eq:products_intermsof_average}.
 
@@
 \eqref{eq:app_maxent_sample_constraints}, and
-$\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$ -- that is, $2m+n$
+$\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$; that is, $2m+n$
 equations (normalization is taken care of) in $2m$ unknowns. A solution can
@@
 \centering
-\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
+\includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
 \caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at the population
@@
 \centering
-\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}
+\includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}
 % 
@@
 constraint functions of the form
-$\yf(\yxs)=(\yxs,\yxxs,\dotsc)\equiv\bigl(
-\yxs, \raisebox{0pt}[0pt][1ex]{}\binom{n\yxs}{2}/\binom{n}{2},\dotsc \bigr)$. In the first example the
-constraints are $\expe{\yxs} = c_1$ and $\expe{\yxxs} = c_2$, with
-$c_1=0.0478$ and $c_2=0.00257$. The distribution obtained at the sample
-level is broader than the one obtained at the population level; the tails
-of the two distributions are very different. The second example includes
-two additional constraints $\expe{\sav{\yx\yx\yx}}= c_3$,
-$\expe{\sav{\yx\yx\yx\yx}}=c_4$ with $c_3=0.000148$,
-$c_4=8.81\times 10^{-6}$. The distribution obtained at the population level
-has two modes, replaced by only one in the distribution obtained at the
-sample level; the tails are very different also in this case. The
-constraints used in these examples have neurobiologically realistic values
-\cite{rostamietal2016}.
+$\yf(\yxs)=(\yxs,\yxxs,\dotsc)\equiv\bigl( \yxs,
+\raisebox{0pt}[0pt][1ex]{}\binom{n\yxs}{2}/\binom{n}{2},\dotsc \bigr)$. In
+the first example the constraint functions are $\expe{\yxs}$ and
+$\expe{\yxxs}$, The distribution obtained at the sample level is broader
+than the one obtained at the population level; the tails of the two
+distributions are very different. The second example uses two additional
+constraint function $\expe{\sav{\yx\yx\yx}}$, $\expe{\sav{\yx\yx\yx\yx}}$.
+The distribution obtained at the population level has two modes, replaced
+by only one in the distribution obtained at the sample level; the tails are
+very different also in this case. The constraint values used in these
+examples, reported in the figure captions, have neurobiologically realistic
+values \cite{rostamietal2016}.
 
@@
 
-The first purpose of this note is indeed to remind ourselves that
+\iffalse The first purpose of this note is indeed to remind ourselves that
 probability models require careful scrutiny, because they can rest on
@@
 \me\ method and the assumption of representative sampling is an example.
+\fi
 

### 2017-07-27 maxNt.tex
update plots and scripts

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-19T02:51:01+0200
+%% Last-Updated: 2017-05-19T15:42:30+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\newcommand*{\pdftitle}{Representative samples\\ and maximum-entropy distributions\\ in neuroscience: a dilemma}
-
-\newcommand*{\headtitle}{\pdftitle} \newcommand*{\pdfauthor}{P.G.L. Porta Mana, V. Rostami, E. Torre}
-
+\newcommand*{\pdftitle}{Representative samples\\ and maximum-entropy distributions:\\ a dilemma}
+
+\newcommand*{\headtitle}{\pdftitle} \newcommand*{\pdfauthor}{%P.G.L. Porta Mana, V. Rostami, E. Torre
+Anonymous Author(s)}
 \usepackage{newtxmath}
@@
 % \DeclareMathSymbol{\epsilon}{\mathalpha}{egreeki}{"0F}
-% \DeclareMathSymbol{\zeta}{\mathalpha}{egreeki}{"10}
+ \DeclareMathSymbol{\zeta}{\mathalpha}{egreeki}{"10}
 % \DeclareMathSymbol{\eta}{\mathalpha}{egreeki}{"11}
@@
 % \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
-% \DeclareMathSymbol{\kappa}{\mathalpha}{egreeki}{"14}
-% \DeclareMathSymbol{\lambda}{\mathalpha}{egreeki}{"15}
+ \DeclareMathSymbol{\kappa}{\mathalpha}{egreeki}{"14}
+ \DeclareMathSymbol{\lambda}{\mathalpha}{egreeki}{"15}
 % \DeclareMathSymbol{\mu}{\mathalpha}{egreeki}{"16}
@@
 % \DeclareMathSymbol{\varEpsilon}{\mathalpha}{egreeki}{"45}
-% \DeclareMathSymbol{\varZeta}{\mathalpha}{egreeki}{"5A}
+ \DeclareMathSymbol{\varZeta}{\mathalpha}{egreeki}{"5A}
 % \DeclareMathSymbol{\varEta}{\mathalpha}{egreeki}{"48}
@@
  \DeclareMathSymbol{\varKappa}{\mathalpha}{egreeki}{"4B}
-% \DeclareMathSymbol{\varLambda}{\mathalpha}{egreeki}{"03}
+ \DeclareMathSymbol{\varLambda}{\mathalpha}{egreeki}{"03}
 % \DeclareMathSymbol{\varMu}{\mathalpha}{egreeki}{"4D}
@@
 \newcommand*{\yl}{\bm{\lambda}}
-\newcommand*{\yk}{\kappa}
-\newcommand*{\yK}{\varKappa}
+\newcommand*{\yk}{\zeta}
+\newcommand*{\yK}{\varZeta}
 %entropies
@@
 \begin{abstract}
-  This note has three nested purposes. The first purpose is to show that
-  the \me\ method can be applied to a representative random sample of a
-  population, to generate its probability distribution, along two different
-  routes. Both routes appear legitimate, but they give inequivalent
-  results. Which route should be chosen? Some arguments are presented in
-  favour of one. The second more general purpose, motivated by the above
-  dilemma, is to remind readers that models like \me\ may contain hidden
-  assumptions; in this case the hidden an unnatural assumption that the
-  sample modelled is isolated from the rest of the population. The third
-  purpose is to promote some old but possibly forgotten probability
-  formulae that may be useful in neuroscientific sampling contexts.
+  This note shows that the \me\ method can be applied to a representative
+  sample from a neuronal population along two different routes: (1) apply
+  to the sample; or (2) apply to the population and marginalize to the
+  sample. These two routes give inequivalent results. Which route should be
+  chosen? Some arguments are presented in favour of the second. The note
+  also touches upon probability formulae of representative sampling and
+  discusses their possible meanings, a discussion that may be useful for
+   sampling problems in neuroscience.
   % The abstract must be limited to one paragraph.
@@
 statistical methods to study the physical and dynamical characteristics of
-brain areas via neuronal recording.
+brain areas via neuronal recordings.
 
@@
 to learn something about all other neurons in the same brain area. That is,
-we assume that they are a \enquote{representative sample}. We don't
+we assume that they are a \enquote{representative sample}. We do not
 elaborate on the various important purposes of such recordings here, but
-stress that these sample sizes can be considered \enquote{large}, because
+stress that these sample sizes can be considered \enquote{large} because
 their statistical analysis requires considerable computational power.
@@
 recordings into few quantities, like sample moments for example. This
-compression would also entail interesting physico-biological properties of
-neuronal activity. The standard \me\ method
+compression might also entail interesting biological or functional
+properties of neuronal activity. The standard \me\ method
 \cite{jaynes1957,sivia1996_r2006,meadetal1984} accomplishes this kind of
@@
 specified set of experimental constraints, and generate a probability
-distribution for its state. But our sample is representative of a larger
-population. We can apply the \me\ method to the larger population, using
-the same constraints, and generate a probability distribution for its
-larger state, and then find the distribution for the sample by
-marginalization. Either application seems to have some commendable
-features. However, \emph{the distributions obtained by these two
-  applications differ}. It goes without saying that if one is
-\enquote{maximally noncommittal}, the other must be somehow
-\enquote{committal}. Which choice is most meaningful?
-
-In the final discussion we present some arguments in favour of one choice
-in our dilemma.
+distribution for the state of the sample. But our sample is representative
+of a larger population, in which it is biologically and functionally
+embedded. We can apply the \me\ method to the larger population, using the
+same constraints, and generate a probability distribution for its larger
+state, and then find the distribution for the sample by marginalization.
+Either application seems to have some commendable features. However,
+\emph{the distributions obtained from these two applications differ}. It
+goes without saying that if one is \enquote{noncommittal}, the other must
+be \enquote{committal}. Which choice is most meaningful?
+
+In the final discussion we present some arguments in favour of one choice.
 
 In the rest of the paper we mathematically formulate this dilemma. To this
-purpose we also present some probability relations relevant to the sampling
-problem. The relations we presentare well-known in survey sampling and in
-the pedagogic problem of drawing from an urn without replacement, yet they
-are somewhat hard to find explicitly written in the neuroscientific
-literature, so they may be of interest on their own. We briefly discuss
-also two very different assumptions that lead to similar initial
+purpose we also present some probability relations relevant to sampling.
+The relations we present are well-known in survey sampling and in the
+pedagogic problem of drawing from an urn without replacement, yet they are
+somewhat hard to find explicitly written in the neuroscientific literature,
+so they may be of interest on their own. We briefly discuss also two very
+different meaning of \enquote{representative} that lead to similar initial
 probability assignments for sampling.
@@
 
-Our mathematical analysis pertains to a neurons modelled as binary units at
+Our mathematical analysis pertains to neurons modelled as binary units at
 one specific instant or short window in time. It is possible to similarly
@@
 or inactive \enquote{$0$} state; and evolution, change, time correlations,
-and similar concepts do not concern us. We consider \me\ models that have
-constraints based on some kinds of sample or population averages; they are
+and similar concepts do not concern us. We consider \me\ models that
+constrain various kinds of sample or population averages; these models are
 often called \enquote{homogeneous}. The final discussion touches upon
@@
 to denote logical conjunction. Probability notation follows Jaynes
-\citep{jaynes1994_r2003}. By \enquote{probability} we mean plausibility or
-the degree of belief which \enquote{would be agreed by all rational men if
-  there were any rational men} \cite{good1966}.
-
-\section{Setup}
+\citep{jaynes1994_r2003}. By \enquote{probability} we mean a degree of
+belief which \enquote{would be agreed by all rational men if there were any
+  rational men} \cite{good1966}.
+
+\section{Setup: population, sample, probabilities}
 \label{sec:setup}
@@
 neurons from this population has joint state $(x_1, \dotsc, x_n) \defs \yx$
-having fixed binary values $(r_1, \dotsc, r_n)\defs\yr$. We will consider
-various averages of the population and the sample. For this purpose we
-introduce a general averaging operator $\sav{\dotv}$ defined by
+having fixed binary values $(r_1, \dotsc, r_n)\defs\yr \in \set{0,1}^n$. We
+will consider various averages of the population and the sample. For this
+purpose we introduce a general averaging operator $\sav{\dotv}$ defined by
 \begin{equation}
@@
   according to an unknown process. The sample
-  $\set{\textrm{X}_1, \textrm{X}_2}$ is random because we don't know the
+  $\set{\textrm{X}_1, \textrm{X}_2}$ is random because we do not know the
   identity of neurons $\textrm{X}_1$ and $\textrm{X}_2$. Given the
   probability for the population, the probabilities for the states of these
-  two samples are different: the former is obtained by marginalization, the
+  two samples are different: the former is obtained from marginalization, the
   latter by symmetrization and marginalization.
@@
 This initial assignment will be modified by our experimental observations.
-We would also like our probability assignment to reflect that the sample is
-somehow \enquote{representative} of the population. We consider here two
-states of knowledge that express this representativeness in different ways
-but lead to identical \emph{initial} probability assignments.
-
-In the first, denoted $\yHp$, we know that the neurons in the population
-are physico-biologically similar, for example in morphology and kind of
-input they receive. Knowledge of this similarity leads us to assign a
-probability distribution for the population state $\yX$ that is symmetric,
-also called \emph{exchangeable}, under permutations of neuron identities.
-
-In the second, denoted $\yHi$, we are completely ignorant about the
-physical details of the individual neurons. Our ignorance is therefore
-symmetric under permutations of neuron identities. This also leads to an
-exchangeable probability distribution for $\yX$.
-
-Let us use $\yH$ to denote either of these two states of knowledge in those
-probabilities that are identical in both.
+Our probability assignment should reflect that the sample is somehow
+\enquote{representative} of the population. We consider here two states of
+knowledge that express this representativeness in different ways but lead
+to identical \emph{initial} probability assignments.
+
+In the first state of knowledge, denoted $\yHp$, we know that the neurons
+in the population are biologically or functionally similar, for
+example in morphology and kind of input or output they receive or give.
+Knowledge of this similarity leads us to assign a probability distribution
+for the population state $\yX$ that is symmetric under permutations of
+neuron identities, or \emph{exchangeable} as it is usually called.
+
+In the second state of knowledge or ignorance, denoted $\yHi$, we are
+completely ignorant about the physical details of the individual neurons.
+Our ignorance is therefore symmetric under permutations of neuron
+identities. This also leads to an exchangeable probability distribution for
+$\yX$.
+
+Let us use $\yH$ to denote either of these two states of knowledge, in those
+probabilities that are identical for $\yHp$ and $\yHi$.
 
@@
 \end{equation}
-the latter being the probability for the population average $\yX$. Proof of
-this theorem and generalizations to non-binary and continuum cases are
-given by de~Finetti \cite{definetti1959b}, Kendall \cite{kendall1967},
-Ericson \cite{ericson1976}, Diaconis \amp\ Freedman
+the latter being the probability for the population average $\yX$. A sum is
+only apparently missing in the central term: its summands $\yXf=A$ are all
+zero except for $A=\yRf$. Proof of this theorem and generalizations to
+non-binary and continuum cases are given by de~Finetti
+\cite{definetti1959b}, Kendall \cite{kendall1967}, Ericson
+\cite{ericson1976}, Diaconis \amp\ Freedman
 \cite{diaconis1977,diaconisetal1980}, Heath \amp\ Sudderth
@@
 \end{gather}
-Our initial symmetric ignorance should intuitively also apply to the
+The conditional probability in the last formula is a hypergeometric
+distribution $\ypp(\yrs\|\yRf)$, typical of \enquote{drawing without
+  replacement} problems. The combinatorial proof of the formulae above is
+in fact the same as for this class of problems
+\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
+Our initial symmetric knowledge should intuitively also apply to the
 sample; indeed, the probability for the state of the
 sample~\eqref{eq:subpop_average} automatically satisfies the representation
-theorem~\eqref{eq:joint_plaus_N_homog} as well. The conditional probability
-in the last formula is a hypergeometric distribution $\ypp(\yrs\|\yRf)$,
-typical of \enquote{drawing without replacement} problems. The
-combinatorial proof of the formulae above is in fact the same as for this
-class of problems
-\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
+theorem~\eqref{eq:joint_plaus_N_homog} as well.
 
@@
 we make an experimental observation, say $X_2 = R_2\in\set{0,1}$ and update our
-probabilities \eqref{eq:joint_plaus_N_homog},
+initial probabilities \eqref{eq:joint_plaus_N_homog}:
 \begin{equation}
@@
   \p(\yX = \yR \| \yXf = \yRf, X_2 = R_2, \yH)\,
-  \p(\yXf=\yRf \| X_2 = R_2,\yH)
-\end{equation}
-Both the conditional probability and the one for the average on the right
-side will update in very different ways for $\yHp$ and $\yHi$. The
-discussion of this update process in the two cases is outside the purposes
-of this note, and we hope to address it in full in a future work.
+  \p(\yXf=\yRf \| X_2 = R_2,\yH).
+\end{equation}
+The conditional probability and the probability for the average on the
+right side will update in very different ways for $\yHp$ and $\yHi$. The
+discussion of this update process in the two cases is unnecessary for the
+purposes of this note and outside their scope; we hope to address it in
+full in a future work.
 
@@
 
-The approximate equalities above express the fact that \emph{our
+Yet, the approximate equalities above express the fact that \emph{our
   uncertainty about the sample is representative of our uncertainty about
@@
 $\yxs\approx\yXf$. In $\yHi$, because we are ignorant about sample and
-population in similar symmetric ways; but this does not imply that
-$\yxs\approx\yXf$. New observations may in fact break this symmetry by
+population in similar symmetric ways; but this does \emph{not} imply that
+$\yxs\approx\yXf$. New observations may in fact break this symmetry via
 \eqn~\eqref{eq:updated_joint}.
@@
 Note that formulae~\eqref{eq:roughly_equal_nN} say more than the limits
-$\p(\yxs=a) \to \p(\yXf=a)$, $\expeb{f(\yxs)} \to \expeb{f(\yXf)}$ as
-$n\to N$. These limits are trivially valid because the sample becomes the
-full population as $n\to N$. In particular, these limits hold even in cases
-where the conditional probability $\p(\yxs = \yrs \|\yXf=\yRf)$ is not a
-fuzzy identity and our uncertainties about sample and about population can
-differ wildly.
+$\p(\yxs=a) \to \p(\yXf=a)$ and $\expeb{f(\yxs)} \to \expeb{f(\yXf)}$, as
+$n\to N$, do. These limits are trivially valid because the sample becomes
+the full population as $n\to N$. In particular, these limits hold even in
+cases where the conditional probability $\p(\yxs = \yrs \|\yXf=\yRf)$ is
+not a fuzzy identity and our uncertainties about sample and about
+population can differ wildly.
 
@@
 For functions representing averaged products,
-$f(\yxs) \defd \sav{\yx\dots\yx}=\binom{n\yxs}{m}/\binom{n}{m}$,
+$f(\yxs) = \sav{\yx\dots\yx}\equiv\binom{n\yxs}{m}/\binom{n}{m}$,
 formulae~\eqref{eq:pullback_P} have the useful form
@@
 
-Formulae
+The probability formulae
 \eqref{eq:joint_plaus_N_homog}--\eqref{eq:conditional_hypergeometric} are
@@
 numerically. The probability $\p(\yXf=\yRf \| \yH)$ for the population
-average needs to be numerically specified, and by \eqref{eq:subpop_average}
-it will determine that of the sample average, $\p(\yxs=\yrs \| \yH)$. If we
-numerically specify the latter, the former is not completely specified,
-because \eqn~\eqref{eq:subpop_average} linearly constrains $N+1$ unknowns
-by only $n+1$ equations in this case.
+average needs to be numerically specified, and by marginalization
+\eqref{eq:subpop_average} it will determine that of the sample average,
+$\p(\yxs=\yrs \| \yH)$. If we numerically specify the latter, the former is
+not completely specified, because \eqn~\eqref{eq:subpop_average} linearly
+constrains $N+1$ unknowns by only $n+1$ equations.
 
@@
 $\expe{\yxs}=c_1$, $\expe{\yxxs}=c_2$. This is still an underdetermined
-linear problem: several distributions can have the same desired
-expectations, as clear from \eqns~\eqref{eq:expe_products}.
+problem: several distributions can have the same desired expectations, as
+clear from \eqns~\eqref{eq:expe_products}.
 
@@
 selects one distribution, purported to be \enquote{maximally noncommittal},
-among those that have the desired expectations. But here's a dilemma:
-formulae~\eqref{eq:pullback_P} allow us to apply the method to find the
-probability of the population $\p(\yXf=\yRf \| \yH)$, or of the sample
-$\p(\yxs=\yrs \| \yH)$. \emph{The two applications, however, are
-  inequivalent}. They lead to numerically different $\p(\yxs=\yrs \| \yH)$.
+among those that have the desired expectations. But here's a dilemma: the
+expectation formulae~\eqref{eq:pullback_P} allow us to apply the method to
+find the probability of the population $\p(\yXf=\yRf \| \yH)$, or of the
+sample $\p(\yxs=\yrs \| \yH)$. \emph{The two applications, however, are
+  inequivalent}. They lead to numerically different distributions for the
+sample average $\p(\yxs=\yrs \| \yH)$.
 
@@
 \end{equation}
-and then by marginalization with \eqn~\eqref{eq:marginal}
+and then by marginalization~\eqref{eq:marginal}
 \begin{equation}
@@
 Yet they cannot be exactly equal, because their equality would require the
-Lagrange multipliers $\yL$ and $\yl$ to satisfy equations
+Lagrange multipliers $\yL$ and $\yl$ to satisfy the constraint equations
 \eqref{eq:app_maxent_pop_constraints},
-\eqref{eq:app_maxent_sample_constraints}, and
+\eqref{eq:app_maxent_sample_constraints}, and also
 $\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$; that is, $2m+n$
-equations (normalization is taken care of) in $2m$ unknowns. A solution can
+equations (one normalization is taken care of) in $2m$ unknowns. A solution can
 exist, if at all, only for very special choices of constraints functions
@@
 population level, like additional modes or particular tail behaviour.
-
 \begin{figure}[!t]
@@
 $\yf(\yxs)=(\yxs,\yxxs,\dotsc)\equiv\bigl( \yxs,
-\raisebox{0pt}[0pt][1ex]{}\binom{n\yxs}{2}/\binom{n}{2},\dotsc \bigr)$. In
-the first example the constraint functions are $\expe{\yxs}$ and
-$\expe{\yxxs}$, The distribution obtained at the sample level is broader
-than the one obtained at the population level; the tails of the two
-distributions are very different. The second example uses two additional
-constraint function $\expe{\sav{\yx\yx\yx}}$, $\expe{\sav{\yx\yx\yx\yx}}$.
-The distribution obtained at the population level has two modes, replaced
-by only one in the distribution obtained at the sample level; the tails are
-very different also in this case. The constraint values used in these
+\raisebox{0pt}[0pt][1ex]{}\binom{n\yxs}{2}/\binom{n}{2},\dotsc \bigr)$,
+equivalent to moments constraints. The constraint values used in these
 examples, reported in the figure captions, have neurobiologically realistic
@@
 
+In the first example the constraint functions are $\expe{\yxs}$ and
+$\expe{\yxxs}$. The distribution obtained at the sample level is broader
+than the one obtained at the population level; the tails of the two
+distributions are very different.
+
+The second
+example uses two additional constraint functions $\expe{\sav{\yx\yx\yx}}$,
+$\expe{\sav{\yx\yx\yx\yx}}$. The distribution obtained at the population
+level has two modes, replaced by only one in the distribution obtained at
+the sample level; the tails are very different also in this case.
+
 \bigskip
@@
 How should we apply the \me\ method then? on the sample or on the
-population? Which application is maximally noncommittal?
+population? Which application is \enquote{maximally noncommittal}?
 
@@
 computational power available, on which inferences we are trying to make,
-on which assumptions we need or want to make and those we wish to avoid.
+on which assumptions we need or want to make, and those we wish to avoid.
 
@@
 probability models require careful scrutiny, because they can rest on
-hidden assumptions that we don't want to make or that contradict others we
+hidden assumptions that we do not want to make or that contradict others we
 are making, and we may not be aware of this. The dilemma generated by the
@@
 
-The tricky point is this. Maximum-entropy applied at the population level
-and applied at the sample level give different results; they are different
-statistical models. The former model clearly assumes, by construction, the
-existence of a larger population from which the sample is taken. What does
-the latter model assume in this respect? is it \enquote{unassuming}, as
-often claimed in the literature? or is it actually assuming that \emph{no}
-larger population exists? In the latter case it would not be the correct
-model to use in our problem.
+The tricky point is this. The \me\ application at the population level and
+the application at the sample level give different results; they are two
+different statistical models. The former model clearly assumes, by
+construction, the existence of a larger population from which the sample is
+taken. What does the latter model assume in this respect? is it
+\enquote{unassuming}, as often claimed in the literature? or does it
+actually assume that \emph{no} larger population exists? In the latter case
+it would not be correct to use this model in our problem.
 
@@
 Suppose we do not know whether the sample is really part of a larger
-population: we do not know whether $N=n$ or how large is $N$ otherwise.
-Call this state of ignorance $\yHd$. In the probability calculus, this
+population: we do not know whether $N=n$ or how large $N$ is otherwise.
+Call this state of ignorance $\yHd$. In the probability calculus this
 ignorance about $N$ is expressed by assigning a probability distribution
@@
 assumption of symmetric ignorance, probability assignments that do not
-assume a specific value of $N$ are then obtained by multiplication of all
+assume a specific value of $N$ are then obtained via multiplication of all
 $N$-dependent probabilities by $\p(N \| \yHd)$ and subsequent
@@
 where $N$-dependencies have been made explicit. This is a formidable
-expression. But the answer to our question, whether the usual \me\ at the
-sample level~\eqref{eq:app_maxent_sample} does not assume anything about a
-larger population, translates now into the precise mathematical question:
-are the distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
+expression. But our question, \enquote{is the usual \me\ at the sample
+  level~\eqref{eq:app_maxent_sample} unassuming with regard to the
+  existence of a larger population?}, translates now into the precise
+mathematical question: \enquote{are the
+distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
 \eqref{eq:app_maxent_sample} equal for some choice of $\p(N \| \yHd)$, with
-$\p(N \| \yHd)\ne 0$ for $N>n$? We leave this mathematical problem for
+$\p(N \| \yHd)\ne 0$ for $N>n$?}. We leave this mathematical problem for
 future work. Note, however, that this equality is satisfied if
@@
 one given by the application at the sample level: both satisfy the same
-constraints, but only the latter is obtained by direct maximization of the
+constraints, but only the latter is obtained via direct maximization of the
 entropy. The lower entropy of the distribution obtained from application at
@@
 
-Rough estimates of $N$ may be available in neuroscientific contexts, so we
-find the \me\ model constructed at the population level very natural and
-preferable. After all, \emph{physical} neuronal-network models usually
-include some sort of external input to the neurons as well, mimicking their
-embedding in a larger network. The distribution of this \me\ model, when
-used as a reference distribution for surprise analysis, may reveal features
-in a dataset that were unnoticed by the standard \me\ model.
+We find the \me\ model constructed at the population level very natural and
+preferable. After all, physical models of neuronal networks usually include
+some sort of external input to the neurons as well, mimicking their
+embedding in a larger network. The sample distribution given by the \me\
+model at the population level, when used as a reference distribution for
+surprise analysis, may reveal features in a dataset that were unnoticed by
+the standard \me\ model. The question remains of how to specify $N$,
+though. We have tacitly intended $N$ as the size of the largest
+biologically or functionally homogeneous population from which our sample
+was recorded. It could be the amount of neurons in a functional brain area,
+for example the primary visual cortex, for which $N\sim 10^8$
+\cite{leubaetal1994}. For large $N$ -- unfortunately we are not yet able to
+translate this \enquote{large} into a numeric order of magnitude -- the
+final distribution becomes independent of $N$, and continuous
+approximations become available.
 
@@
 descriptions differ -- even though the final \emph{physical} state is
-obviously exactly the same \cite[\sect~4]{jaynes1985d_r1993}. The
-difference in the two descriptions appears because in one case we can make
-sharper predictions about the state thanks to our knowledge of its
-preceding dynamics. In this example, though, both distributions are usually
-immensely sharp and practically lead to the same predictions. In the cases
-considered in this note the difference in predictions may be relevant
-instead.
-
-Our analysis touched only constraints of the sample average. The
-corresponding models are usually called \enquote{homogeneous} in the
-literature. Purely \enquote{inhomogeneous} models have also been used
+obviously exactly the same \cite[\sect~4]{jaynes1985d_r1993}. The two
+descriptions differs because in one case we can make sharper predictions
+about the state thanks to our knowledge of its preceding dynamics. In this
+example, though, both distributions are usually immensely sharp and
+practically lead to the same predictions. In the neuroscientific
+applications considered in this note the difference in predictions may be
+relevant instead.
+
+Our analysis touched only constraints of the sample average,
+$\expeb{f(\yxs)}$. The corresponding models are usually called
+\enquote{homogeneous} in the literature. Purely \enquote{inhomogeneous}
+models have also been used
 \cite{schneidmanetal2006,shlensetal2006,roudietal2009b}, in which
@@
 unconstrained neurons marginalize out. This is understandable: expressing
-different uncertainties about neurons $2$ and $5$ we are breaking the
+different uncertainties about, say, neurons $2$ and $5$ we are breaking the
 symmetry of our uncertainty, which thus cannot be representative of other
-neurons in the sample or in the population.
-
-Inhomogeneous models, however, require enormous computational power for
-large sample sizes; homogeneous models therefore retain their importance.
-Our analysis and dilemma also persist for hybrid homogeneous-inhomogeneous
-models \cite{tkaciketal2014b,shimazakietal2015}.
+neurons in the sample or in the population. Inhomogeneous models, however,
+require enormous computational power for large sample sizes; homogeneous
+models therefore retain their importance. Our analysis and dilemma also
+persist for hybrid homogeneous-inhomogeneous models
+\cite{tkaciketal2014b,shimazakietal2015}.
 

### 2017-07-30 rebuttal.txt
update text, scripts, bib

The author(s) thank the three reviewers for their feedback. Here are our replies point-by-point.

Reviewer 1 ("Assigned_Reviewer_1")

1. We agree that the introduction part did not give enough neuroscientific introductory context and literature. The maximum-entropy method at the "sample level" has actually been repeatedly used and analysed in computational neuroscience since the work of Bohte et al (2000) to the present. A more complete bibliography is given at the end of this response and can be easily extended. The method is also used by non-experts, and our point would be to clarify some assumptions implicit the method also for non-experts.

2. The reviewer is again right. Our discussion starts from the *constatation* that most of the neuroscientific literature implicitly or explicitly generalizes the statistical properties of a sample of recorded neurons, analysed via maximum-entropy, to whole areas the neurons are recorded from, eg retina or motor cortex. This generalization is a complex argument lying outside our competence; we cannot justify or deprecate it in this manuscript. Our purpose is to warn neuroscientists who make this generalization: the typical sample-level method commonly used in the literature is inconsistent with this generalization. We are proposing a more consistent approach.

The case of absolute ignorance (I'') is not used in the rest of the manuscript and was confusing; we would omit it in a revised version.

We invite the reviewer to read also our responses 1 & 2 to the second reviewer.

3. We again agree: our principal equations are:

(15) to be used when a larger population is assumed and its size roughly known;

(17) *the equation used in the literature*, to be used when no larger population is assumed;

(19) to be used when the the size of the larger population is unknown.

Equations (15) and (19) are novel to our knowledge. Unfortunately we failed to emphasize these three equations and give clear guidance to their use. We would amend this in a revised version.

4. The captions should indeed give a clearer explanation of the figures. We would amend this in a revised version.

Reviewer 2 ("Assigned_Reviewer_2")

1. We agree with the reviewer, we did not give enough context and literature. Our starting point is indeed that

- the *sample-level* maximum-entropy method is increasingly being used in the neuroscientific literature (references below);

- this literature often generalizes results obtained by this method from a recorded sample to a whole area;

- worries have already been expressed (eg Ganmor et al 2011, Roudi et al PlosCB 2009) about the sample-size dependence of maximum-entropy results.

Our manuscript would like to explain from simple probability considerations why these worries are motivated, why the results depend on the sample size, and what alternative formulae can be used: eqns (15), (19).

2. The reviewer is right, we should have mentioned and referenced an important fact: the sample approach is routinely used in the literature. We agree that it is not an option if we assume sampling from a larger population. Our manuscript wanted to explain why it is not an option, and propose more valid alternatives (eqns 15, 19).

The phrase "left for future work" was unfortunate. Eqn (19), for unknown population size, can actually be used and numerically computed as is. Left for future work is the analysis of its mathematical properties. In a revised version of the paper we would add to each figure also the distribution obtained by eqn (19).

We invite the reviewer to read also our responses 1 & 2 to the first reviewer.

3. It is true, we give much space to the old formulae of sampling without replacement. This was on purpose, see lines 42-45: "The relations we present are well-known in survey sampling and in the pedagogic problem of drawing from an urn without replacement, yet they are somewhat hard to find explicitly written in the neuroscientific literature, so they may be of interest on their own". We should have stressed more that our manuscript wanted to "advertize" the importance of these formulae, not much used in neuroscience but very central to it.

4. Equation (12) and its discussion (lines 113-129) are irrelevant to the rest of the paper, to the discussion of maximum-entropy and to the main equations. We would eliminate this whole passage in a revised version.

5. We would have liked to present a Bayesian analysis of the problem, but the page limit prevented this; the manuscript could become confusing. In a revised version we would add a measure of discrepancy, eg Kullback-Leibler divergence, between the "sample" and "population" approach.

Reviewer 3 ("Assigned_Reviewer_5")

We cordially thank the reviewer for the appreciative words. We would modify the manuscript to address the concerns of the other reviewers but leaving the theoretical points. We want to add that our figures do use actual recorded neural activity data, but because of the double-blind review procedure we could not give more details and made comparison with recorded frequency distributions in the reviewed version.ÿ

Sample references of maximum-entropy use in neuroscientific literature:
Martignon & al, Biol Cybern (1995)
Bohte & al, Neural Comp (2000)
Schneidman & al,  Nature (2006)
Shlens & al,  J Neurosci (2006)
Tkačik & al, arXiv:q-bio/0611072 (2006)
Macke & al, Neural Comp (2009)
Roudi & al, PLoS CB (2009)
Roudi & al, Front Comp Neurosci (2009)
Roudi & al, Phys Rev E (2009)
Tkačik & al, arXiv:09125409 (2009)
Macke & al, Tech Rep (2009)
Gerwinn & al, Front Comput Neurosci (2010)
Barreiro & al, arXiv:10112797 (2010)
Macke & al, NIPS (2011)
Macke & al, Phys Rev Lett (2011)
Ganmor & al, Proc Natl Acad Sci (2011)
Granot-Atedgi & al, PLoS CB (2013)
Macke & al, Entropy (2013)
Tkačik & al, Proc Natl Acad Sci (2014)
Mora & al, Phys Rev Lett (2015)
Shimazaki & al,  Sci Rep (2015)

### 2017-07-30 rebuttal2.txt
update text, scripts, bib

The author(s) thank the three reviewers for their feedback, which is extremely useful for a revised version of the manuscript, even in case of non-acceptance. Here are our comments point-by-point.

Reviewer 1 ("Assigned_Reviewer_1")

1. The "sample level" maximum-entropy method has been repeatedly used and discussed in computational neuroscience since the work of Bohte et al (2000). A more complete bibliography is given at the end of this response and could be easily extended. Unfortunately our introduction did not clarify and reference this context enough; it would surely be amended in a revised version. The method is also used by non-experts; our purpose is to clarify some implicit assumptions that experts and non-experts seem to be unaware of.

2. Our discussion starts from an observation: in the neuroscientific literature some authors generalize conclusions drawn via maximum-entropy on a sample of recorded neurons, to whole areas the sample is recorded from, eg retina or motor cortex. Unfortunately this observation was not made clear in our introduction. Our purpose is to give a warning: the typical sample-level approach (Eq. 17) is incompatible with the beyond-sample generalizations made by some authors. We propose a simple alternative consistent with this assumption (Eqs 15, 19).

The case of absolute ignorance (I'') is not used in the rest of the manuscript; we apologize for the confusion it creates. We would omit it in a revised version.

We invite the reviewer to read also our responses 1 & 2 to the second reviewer.

3. In the manuscript the principal equations are:

(15) to be used when a larger population is assumed and its size roughly known;

(17) to be used when no larger population is assumed: most literature uses this equation but also assumes a larger population;

(19) to be used when the the size of the larger population is uncertain.

Equations (15) and (19) are novel, to our knowledge. Unfortunately we did not emphasize the role of these three equations and give clear guidance to their use. This should be amended in a revised version.

4. The captions should indeed give a clearer explanation of the figures. Owing to the page limit we preferred not to repeat explanations given in the main text.

Reviewer 2 ("Assigned_Reviewer_2")

1. The starting points of the manuscripts are three:

- the *sample-level* maximum-entropy method is increasingly being used in the neuroscientific literature (references below);

- this literature often generalizes results obtained with this method from a recorded sample to a whole area;

- worries have been expressed (eg Ganmor et al 2011, Roudi et al PlosCB 2009) about an excessive sample-size dependence of maximum-entropy results.

Unfortunately these three starting points are not clarified enough in our introduction; this should be amended in a revised version. We motivate the sample-size worries by simple probabilistic considerations, and provide alternative formulae to be used: Eq. (15) when the size of the full population is known, Eq. (19) when it is uncertain.

2. The sample approach is routinely used in the literature and we agree with the reviewer: "it is not even an option" if the recorded neurons are assumed to be a sample of a larger population. This is the message of the manuscript, unfortunately not clear enough. Its analysis is meant to explain why the sample approach is not an option and to propose two corrected formulae (Eqs 15, 19).

Our phrase "left for future work" was unfortunate: Eq. (19) for uncertain population size is actually an end result, ready to be used computationally if not analytically. What we left for future work was the analysis of its mathematical difference from (15) and (17). In a revised version of the paper we would add an example of distribution (19) to Figs 2 & 3.

We invite the reviewer to read also our responses 1, 2, 3 to the first reviewer.

3. Large space was dedicated to deriving the formulae of sampling without replacement. As mentioned on lines 42-45 this was on purpose because these old formulae are central in many neuroscientific problems and yet we have rarely seen them used in the neuroscience literature. We wanted to stress their importance.

But from those formulae the manuscript also presents novel results: Eqs (15) & (19), which can immediately be used instead of Eq. (17), currently used in the literature.

4. Equation (12) and its discussion (lines 113-129) are an excursion irrelevant to the rest of the paper, to the discussion of maximum-entropy and to the main equations we derive. We would eliminate this whole passage in a revised version.

5. We would have liked to present a Bayesian analysis of the problem, but the NIPS page limit prevented this. A measure of discrepancy, eg Kullback-Leibler divergence, between Eqs (15), (17), (19) could be added in a revised version.

Reviewer 3 ("Assigned_Reviewer_5")

We cordially thank the reviewer for the appreciative words. We want to add that our figures *do* use actual recorded neural activity data, but because of the double-blind review procedure we omitted this information, which we plan to add in the final version if the manuscript were accepted.

Sample references of maximum-entropy use in neuroscientific literature:
Martignon & al, Biol Cybern (1995)
Bohte & al, Neural Comp (2000)
Schneidman & al,  Nature (2006)
Shlens & al,  J Neurosci (2006)
Tkačik & al, arXiv:q-bio/0611072 (2006)
Macke & al, Neural Comp (2009)
Roudi & al, PLoS CB (2009)
Roudi & al, Front Comp Neurosci (2009)
Roudi & al, Phys Rev E (2009)
Tkačik & al, arXiv:09125409 (2009)
Macke & al, Tech Rep (2009)
Gerwinn & al, Front Comput Neurosci (2010)
Barreiro & al, arXiv:10112797 (2010)
Macke & al, NIPS (2011)
Macke & al, Phys Rev Lett (2011)
Ganmor & al, Proc Natl Acad Sci (2011)
Granot-Atedgi & al, PLoS CB (2013)
Macke & al, Entropy (2013)
Tkačik & al, Proc Natl Acad Sci (2014)
Mora & al, Phys Rev Lett (2015)
Shimazaki & al,  Sci Rep (2015)

### 2017-07-30 rebuttal3.txt
update text, scripts, bib

The author(s) thank the reviewers for their feedback, which is extremely useful for a revised version of the manuscript even in case of non-acceptance. Here are our comments point-by-point.

* Reviewer 1 (Assigned_Reviewer_1)

1. The sample-level maximum-entropy method has been repeatedly used and discussed in computational neuroscience at least since Bohte & al (2000). Unfortunately our Introduction did not clarify nor reference this context enough; this should be amended in a revised version. An unabridged bibliography is given at the end of this response and can be easily extended. The method is also used by non-experts; the manuscript aims to clarify implicit assumptions that users seem to be unaware of.

2. Our discussion starts from an observation: in the neuroscientific literature, several authors generalize conclusions drawn via maximum-entropy on recorded neurons, to whole areas the neurons are recorded from, eg retina or motor cortex. Unfortunately this observation was not clear in our Introduction. Whether such generalization makes biological sense is a difficult question to be analysed case-by-case. Our purpose is to give a statistician's warning: the sample-level equation (17) and the beyond-sample generalization, both typical of the literature, are incompatible. We propose two simple alternative equations (15, 19) compatible with the generalization.

The case of absolute ignorance (I'') is not used in the rest of the manuscript; we apologize for the confusion it creates. We would omit it in a revised version.

We invite the reviewer to read also our responses 1, 2 to the second reviewer.

3. The manuscript's principal equations are:

(15) to be used when a larger population is assumed and its size roughly known;

(17) to be used when no larger population is assumed. Most literature uses this equation but assumes a larger population;

(19) to be used when the the size of the larger population is uncertain.

Equations (15) & (19) are novel to our knowledge. We apologize for not emphasizing the role of the three equations above nor giving clear guidance to their use. This should be amended in a revised version.

4. The captions should be amended and give a clearer explanation of the figures. Owing to the page limit we chose not to repeat explanations given in the main text.

* Reviewer 2 (Assigned_Reviewer_2)

1. The manuscript's starting points are three:

- the *sample-level* maximum-entropy method is increasingly being used in the neuroscientific literature (refs below);

- this literature often generalizes results obtained via this method from a recorded sample to a whole area;

- worries have been expressed (eg Ganmor & al 2011, Roudi & al PlosCB 2009) about excessive sample-size dependence of maximum-entropy results.

We apologize for not clarifying enough these three starting points in our Introduction, which should be amended in a revised version. We aim to motivate the sample-size worries by simple probabilistic considerations, and to provide alternative correct formulae: Eq. (15) when the size of the full population is known, Eq. (19) when it is uncertain.

2. The sample approach is routinely used in the literature and we agree with the reviewer: "it is not even an option" if the recorded neurons are assumed to be a sample of a larger population. This is the message of the manuscript, unfortunately not clear enough. Its analysis is meant to explain why the sample approach is not an option and to propose two corrected formulae (Eqs 15, 19).

Our phrase "left for future work" was unfortunate: Eq. (19) for uncertain population size is actually an end result, ready to be used. What we left for future work was the analysis of its mathematical difference from (15) & (17). In a revised version we would add an example of distribution (19) to Figs 2 & 3.

We invite the reviewer to read also our responses 1-3 to the first reviewer.

3. Large space was given to deriving the formulae of sampling without replacement. As mentioned on lines 42-45 this was on purpose because these old formulae are central in many neuroscientific problems, yet we rarely see them used in the neuroscience literature. We wanted to stress their importance. But, starting from those formulae, the manuscript also presents novel results: Eqs (15) & (19), which can immediately be used instead of Eq. (17), currently used in the literature.

4. Eq. (12) and its discussion (lines 113-129) are an excursion irrelevant to the rest of the paper, to the discussion of maximum-entropy and to the main equations we derive. We would eliminate this whole passage in a revised version.

5. We would have liked to present a full Bayesian analysis of the problem, but the NIPS page limit prevents this. A measure of discrepancy, eg Kullback-Leibler divergence, between Eqs (15), (17), (19) would be added in a revised version.

* Reviewer 3 (Assigned_Reviewer_5)

We cordially thank the reviewer for the appreciative words. We want to add that our figures *do* use actual recorded neural activity data, but because of the double-blind review procedure we had to omit the related discussion at first. It would be added in the final version if the manuscript were accepted.

Sample references of maximum-entropy use in neuroscientific literature:
Martignon & al, Biol Cybern 1995
Bohte & al, Neural Comp 2000
Amari & al, Neural Comp 2003
Schneidman & al, Nature 2006
Shlens & al, J Neurosci 2006
Tkačik & al, arXiv:q-bio/0611072 2006
Mcdonald & al, NIPS 2009
Macke & al, Neural Comp 2009
Roudi & al, PLoS CB 2009
Roudi & al, Front Comp Neurosci 2009
Roudi & al, Phys Rev E 2009
Tkačik & al, arXiv:0912.5409 2009
Macke & al, Tech Rep 2009
Gerwinn & al, NIPS 2009
Barreiro & al, BMC Neurosci 2010
Macke & al, NIPS 2011
Macke & al, Phys Rev Lett 2011
Ganmor & al, Proc Natl Acad Sci 2011
Granot-Atedgi & al, PLoS CB 2013
Macke & al, Entropy 2013
Tkačik & al, Proc Natl Acad Sci 2014
Mora & al, Phys Rev Lett 2015
Shimazaki & al, Sci Rep 2015

### 2017-12-04 maxNt.tex
results

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-05-19T15:42:30+0200
+%% Last-Updated: 2017-12-04T11:34:56+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\usepackage[%final,%
+\usepackage[final,%
 nonatbib]{nips_2017}
@@
 
-\newcommand*{\headtitle}{\pdftitle} \newcommand*{\pdfauthor}{%P.G.L. Porta Mana, V. Rostami, E. Torre
-Anonymous Author(s)}
+\newcommand*{\headtitle}{\pdftitle} \newcommand*{\pdfauthor}{P.G.L. Porta Mana, V. Rostami, E. Torre}
 \usepackage{newtxmath}
@@
 \author{
-  P.G.L. Porta Mana\thanks{Equally contributing first authors}\\Independent researcher\\\texttt{pgl@portamana.org}
+  P.G.L. Porta Mana\\\texttt{mana@kth.org}
   \And
-  V. Rostami$^*$\\Forschungszentrum J{\"u}lich INM-6\\Germany\\\texttt{v.rostami@fz-juelich.de}
+  V. Rostami$^*$\\\texttt{v.rostami@fz-juelich.de}
   \And
-E. Torre\\ETH Z{\"u}rich\\Switzerland\\
-\texttt{torre@ibk.baug.ethz.ch}
+E. Torre\\\texttt{torre@ibk.baug.ethz.ch}
   % David S.~Hippocampus\thanks{Use footnote for providing further
@@
 examples, reported in the figure captions, have neurobiologically realistic
-values \cite{rostamietal2016}.
+values \cite{rostamietal2016_r2017}.
 
@@
 subnetwork of enough large size $n$ can be assigned a similar
-maximum-entropy distribution. See ref.~\citep{rostamietal2016} for a
+maximum-entropy distribution. See ref.~\citep{rostamietal2016_r2017} for a
 broader discussion of this fact and of its consequences.

### 2018-05-12 maxNt_v2.tex
V2

--- 2017-12-04 maxNt.tex
@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-12-04T11:34:56+0100
+%% Last-Updated: 2018-05-12T15:29:50+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \usepackage[autostyle=false,autopunct=false,english=british]{csquotes}
-\setquotestyle{american}
+\setquotestyle{british}
 
@@
 % \DeclareMathSymbol{\eta}{\mathalpha}{egreeki}{"11}
-% \DeclareMathSymbol{\theta}{\mathalpha}{egreeki}{"12}
+ \DeclareMathSymbol{\theta}{\mathalpha}{egreeki}{"12}
 % \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
@@
 
+\pdfmapline{+dummy-space <dummy-space.pfb}\pdfinterwordspaceon%
+
 % @@@@@@@@@@ new macros @@@@@@@@@@
@@
 %\newcommand*{\e}{\mathrm{e}}%Neper
-%\newcommand*{\di}{\mathrm{d}}%differential
+\newcommand*{\di}{\mathrm{d}}%differential
 %\newcommand*{\Di}{\mathrm{D}}%capital differential
@@
 \newcommand*{\me}{maximum-entropy}
-
+\newcommand*{\ytheta}{\bm{\theta}}
 %@@@@@@@@@@ new macros end @@@@@@@@@@
@@
 \author{
-  P.G.L. Porta Mana\\\texttt{mana@kth.org}
+  P.G.L. Porta Mana\\\texttt{piero.mana@ntnu.org}
   \And
-  V. Rostami$^*$\\\texttt{v.rostami@fz-juelich.de}
+  V. Rostami\\\texttt{v.rostami@fz-juelich.de}
   \And
@@
 % a mechanical ritual than a real tentative of connecting with the reader.
+
+Imagine that we have recorded the firing activity of a hundred neurons from
+a particular brain area. This scenario is a concrete possibility thanks to
+recent electrophysiological techniques \cite{berenyietal2014}. We bin and
+binarize the recorded activity into sequences of $0$s (inactive) and $1$s
+(firing) \cite{caianiello1961,caianiello1986}. We notice that groups of two
+or more $1$s can appear in any time bin. We may ask several questions
+looking at these groups. A recurring question is of this kind: can the
+presence of groups of three or more simultaneous firings be
+\enquote{explained} in terms of groups of just two or three? This kind of
+question is very vague until we define what we mean with \enquote{explain}
+or similar words.
+
+A possible, precise definition is as follows. We want to predict, \ie\
+assign a probability to, the population activity in a time bin, given the
+population activities in the other bins:
+\begin{equation}
+  \label{eq:general_prediction}
+  \p(\text{activity at one time bin} \|
+  \text{activities at the remaining bins, and other info}).
+\end{equation}
+We can say that the simultaneous firing of at most two neurons
+\enquote{explains} the activity if only the values of the first and second
+empirical moments of the activities are relevant for the probability above:
+\begin{multline}
+  \label{eq:informal_sufficiency}
+  \p(\text{activity at one time bin} \|
+  \text{activities at the other bins, and other info}) ={}\\
+  \p(\text{activity at some time bin} \|
+  \text{1st \amp\ 2nd empirical moments, and other info});
+\end{multline}
+in statistical terms, the first and second moments are \emph{sufficient
+  statistics}. Other definitions are possible, of course; the definition
+above has the advantage of being mathematically very precise. In fact it
+determines, from just the three basic laws of the probability calculus, the
+mathematical form of the probability for the activity at any number of bins:
+\begin{multline}
+  \label{eq:sufficiency_maxent_family}
+  \p(\text{activity at bins $1,2,\dotsc$} \|
+  \text{other info}) ={}\\
+  \shoveright{\int
+    \pf(\text{activity at bin $1$} \|\ytheta)\,
+    \pf(\text{activity at bin $2$} \|\ytheta)\dotsm
+ \pf(\ytheta\| \text{other info})\,
+  \di\ytheta,}
+  \\[2\jot]
+  \text{with \enquote{$\pf(\text{activity at time bin $i$} \|\ytheta)$}
+    having maximum-entropy form}.
+\end{multline}
+The proof that the assumption of sufficient statistics leads to the
+probability expression above is the Koopman-Pitman theorem
+\cite{koopman1936,pitman1936,darmois1935} and can be found in
+\citetext[\sect~4.5]{bernardoetal1994_r2000} and references therein. It can
+be extended to higher moments and other statistics.
+
 This note is mainly addressed to neuroscientists interested in \me\

### 2018-05-12 maxNt_v2.tex
V2

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-12T15:29:50+0200
+%% Last-Updated: 2018-05-12T16:21:17+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \text{with \enquote{$\pf(\text{activity at time bin $i$} \|\ytheta)$}
-    having maximum-entropy form}.
+    having maximum-entropy form};
 \end{multline}
-The proof that the assumption of sufficient statistics leads to the
-probability expression above is the Koopman-Pitman theorem
+the $\ytheta$ being the Lagrange multipliers of the maximum-entropy
+distribution. The proof that the assumption of sufficient statistics leads
+to the probability expression above is the Koopman-Pitman theorem
 \cite{koopman1936,pitman1936,darmois1935} and can be found in
-\citetext[\sect~4.5]{bernardoetal1994_r2000} and references therein. It can
+\textcite[\sect~4.5]{bernardoetal1994_r2000} and references therein. It can
 be extended to higher moments and other statistics.
+
+When many data are available (and unfortunately also when they aren't), the
+conditional density for the Lagrange multipliers,
+$\pf(\ytheta\| \text{data, other info})$ is strongly peaked at the unique
+values determined by the inverse problem \cite{meadetal1984}, and the
+probability for the population activity is well-approximated by a
+maximum-entropy one with those Lagrange multipliers. This kind of
+distribution has been used at least since the 1990s
+\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,tkaciketal2006,mackeetal2009,mackeetal2009b,roudietal2009,roudietal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,barreiroetal2010b,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,moraetal2015,shimazakietal2015}
 

### 2018-05-13 maxNt_v2.tex
V2

--- 2018-05-12 maxNt_v2.tex
@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-12T15:29:50+0200
+%% Last-Updated: 2018-05-13T13:23:37+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \usepackage[final,%
-nonatbib]{nips_2017}
+nonatbib]{nips_2018}
 
@@
 
-\pdfmapline{+dummy-space <dummy-space.pfb}\pdfinterwordspaceon%
+%\pdfmapline{+dummy-space <dummy-space.pfb}\pdfinterwordspaceon%
 
@@
   \label{eq:general_prediction}
-  \p(\text{activity at one time bin} \|
-  \text{activities at the remaining bins, and other info}).
+  \p(\text{\small activity at one time bin} \|
+  \text{\small activities at the remaining bins},\text{\small other info}).
 \end{equation}
@@
   \label{eq:informal_sufficiency}
-  \p(\text{activity at one time bin} \|
-  \text{activities at the other bins, and other info}) ={}\\
-  \p(\text{activity at some time bin} \|
-  \text{1st \amp\ 2nd empirical moments, and other info});
+  \p(\text{\small activity at one time bin} \|
+  \text{\small activities at the other bins},\text{\small other info}) ={}\\
+  \p(\text{\small activity at some time bin} \|
+  \text{\small 1st \amp\ 2nd empirical moments},\text{\small other info});
 \end{multline}
@@
 mathematical form of the probability for the activity at any number of bins:
-\begin{multline}
+\begin{equation}
   \label{eq:sufficiency_maxent_family}
-  \p(\text{activity at bins $1,2,\dotsc$} \|
-  \text{other info}) ={}\\
-  \shoveright{\int
-    \pf(\text{activity at bin $1$} \|\ytheta)\,
-    \pf(\text{activity at bin $2$} \|\ytheta)\dotsm
- \pf(\ytheta\| \text{other info})\,
-  \di\ytheta,}
-  \\[2\jot]
-  \text{with \enquote{$\pf(\text{activity at time bin $i$} \|\ytheta)$}
-    having maximum-entropy form}.
-\end{multline}
-The proof that the assumption of sufficient statistics leads to the
-probability expression above is the Koopman-Pitman theorem
+  \begin{gathered}
+  \p(\text{\small activity at bins $1,2,\dotsc$} \|
+  \text{\small other info}) =
+  \int\Bigl[\prod_i 
+    \pf(\text{\small activity at bin } i \|\ytheta) \Bigr]\,
+ \pf(\ytheta\| \text{\small other info})\,
+  \di\ytheta,
+  \\[\jot]
+  \text{with \enquote{$\pf(\text{\small activity at time bin } i \|\ytheta)$}
+    having maximum-entropy form with multipliers $\ytheta$};
+\end{gathered}
+\end{equation}
+The proof that the assumption of sufficient statistics leads
+to the probability expression above is the Koopman-Pitman theorem
 \cite{koopman1936,pitman1936,darmois1935} and can be found in
-\citetext[\sect~4.5]{bernardoetal1994_r2000} and references therein. It can
+\textcite[\sect~4.5]{bernardoetal1994_r2000} and references therein. It can
 be extended to higher moments and other statistics.
+
+When many data are available (and unfortunately also when they aren't), the
+conditional density for the Lagrange multipliers,
+$\pf(\ytheta\| \text{data, other info})$ is strongly peaked at the unique
+values determined by the inverse problem \cite{meadetal1984}, and the
+probability for the population activity is well-approximated by a
+maximum-entropy distribution \cite{jaynes1957,sivia1996_r2006,meadetal1984}
+with those Lagrange multipliers. This kind of distributions has been used
+in neuroscience at least since the 1990s to study correlations from diverse
+points of view, for diverse purposes, and with data recorded from diverse
+brain areas
+\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015}.
+Part of this literature focused on the range of applicability of maximum-entropy
+distributions
+\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013}.
+
+This kind of studies would have little significance if we did not expect
+their results to generalize beyond the particular recorded sample. In other
+words, we implicitly or explicitly expect that the conclusions drawn with
+maximum-entropy methods about the recorded neurons should somehow
+generalize or extrapolate at least to whole areas those neurons are recorded from,
+\eg\ retina or motor cortex. We are thus assuming that the recorded sample
+is a \emph{representative sample} of that area.
+
+Whether such extrapolation makes biological sense is a difficult question,
+to be analysed case by case. The purpose
+
+The point that we want to stress is that
+
+From the observation of these samples we expect
+to learn something about all other neurons in the same brain area. That is,
+we assume that they are a \enquote{representative sample}.
 

### 2018-05-13 maxNt_v2.tex
V2

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-13T13:23:37+0200
+%% Last-Updated: 2018-05-13T22:11:43+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\def\arxivp{}
-\def\mparcp{}
-\def\philscip{}
-\def\biorxivp{}
-\newcommand*{\arxivsi}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.\\}
-\newcommand*{\mparcsi}{\texttt{mp\_arc} eprints available at \url{http://www.ma.utexas.edu/mp_arc/}.\\}
-\newcommand*{\philscisi}{\texttt{philsci} eprints available at \url{http://philsci-archive.pitt.edu/}.\\}
-\newcommand*{\biorxivsi}{\texttt{bioRxiv} eprints available at \url{http://biorxiv.org/}.\\}
-\newcommand*{\arxiveprint}[1]{\global\def\arxivp{\arxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
-\texttt{\urlalt{http://arxiv.org/abs/#1}{arXiv:\hspace{0pt}#1}}%
+\newcommand*{\arxiveprint}[1]{%\global\def\arxivp{\arxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
+\texttt{\urlalt{https://arxiv.org/abs/#1}{arXiv:\hspace{0pt}#1}}%
 %\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
@@
 }
-\newcommand*{\mparceprint}[1]{\global\def\mparcp{\mparcsi}%\citeauthor{0mparccite}\addtocategory{ifarchcit}{0mparccite}%eprint
+\newcommand*{\mparceprint}[1]{%\global\def\mparcp{\mparcsi}%\citeauthor{0mparccite}\addtocategory{ifarchcit}{0mparccite}%eprint
 \texttt{\urlalt{http://www.ma.utexas.edu/mp_arc-bin/mpa?yn=#1}{mp\_arc:\hspace{0pt}#1}}%
@@
 }
-\newcommand*{\philscieprint}[1]{\global\def\philscip{\philscisi}%\citeauthor{0philscicite}\addtocategory{ifarchcit}{0philscicite}%eprint
+\newcommand*{\philscieprint}[1]{%\global\def\philscip{\philscisi}%\citeauthor{0philscicite}\addtocategory{ifarchcit}{0philscicite}%eprint
 \texttt{\urlalt{http://philsci-archive.pitt.edu/archive/#1}{PhilSci:\hspace{0pt}#1}}%
@@
 }
-\newcommand*{\biorxiveprint}[1]{\global\def\biorxivp{\biorxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
+\newcommand*{\biorxiveprint}[1]{%\global\def\biorxivp{\biorxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
 \texttt{\urlalt{http://biorxiv.org/content/early/#1}{bioRxiv:\hspace{0pt}#1}}%
@@
 %\renewcommand{\arxivnote}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.}
+}
+\newcommand*{\osfeprint}[1]{%
+\texttt{\urlalt{https://doi.org/10.17605/osf.io/#1}{doi:10.17605/osf.io/#1}}%
 }
@@
 \author{
-  P.G.L. Porta Mana\\\texttt{piero.mana@ntnu.org}
+  P.G.L. Porta Mana\\\texttt{piero.mana@ntnu.no}
   \And
@@
 a particular brain area. This scenario is a concrete possibility thanks to
-recent electrophysiological techniques \cite{berenyietal2014}. We bin and
-binarize the recorded activity into sequences of $0$s (inactive) and $1$s
-(firing) \cite{caianiello1961,caianiello1986}. We notice that groups of two
-or more $1$s can appear in any time bin. We may ask several questions
-looking at these groups. A recurring question is of this kind: can the
-presence of groups of three or more simultaneous firings be
-\enquote{explained} in terms of groups of just two or three? This kind of
+recent electrophysiological techniques \cite{berenyietal2014}. In 1985,
+Gerstein, Perkel, \amp\ Dayhoff wrote:
+\begin{quote}
+  the very profusion of data produced in such experiments creates serious
+  problems of analysis and interpretation; these problems are conceptual as
+  well as practical. The principal conceptual problems are (1) defining
+  cooperativity or functional grouping among neurons and (2) formulating
+  quantitative criteria for recognizing and characterizing such
+  cooperativity. \cite{gersteinetal1985}
+\end{quote}
+ More than thirty years later these problems are
+still under investigation.
+
+Some investigations in the literature proceed as follows. We bin in time
+and binarize in intensity \cite{caianiello1961,caianiello1986} the recorded
+activity, transforming it into a sequence of $0$s (inactive) and $1$s
+(firing) for each neuron. Groups of two or more $1$s can appear in any time
+bin. Then we ask a question of this kind: can the presence of groups of
+three or more simultaneous firings be \enquote{explained} in terms of
+groups of just two, or three, or\ldots? This kind of question is very vague
+until we define what we mean with \enquote{explain} or similar words.
+Usually such \enquote{explanation} has a probabilistic meaning: we wonder
+whether the sole knowledge about two, or three, or more simultaneous
+firings can yield as good predictions about the whole activity as a more
+complete knowledge would do. To make this question concrete we must choose
+a probabilistic model.
+
+Maximum-entropy models
+\cite{jaynes1957,jaynes1963,sivia1996_r2006,meadetal1984} have been used in
+neuroscience with this kind of purpose and for diverse brain areas at least
+since the 1990s
+\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015}.
+Part of this literature focused on the range of applicability of \me\
+distributions
+\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.
+We assume that our readers are at least vaguely familiar with such models.
+In brief, they build a probability distribution for neuronal activity using
+only data about simultaneous firings of two neurons, or of three, and so
+on. Maximum-entropy probability distributions can be introduced and justified in
+three main ways:
+\begin{enumerate}
+\item from the maximum-entropy method
+  \cite{jaynes1957,jaynes1963,sivia1996_r2006};
+\item as asymptotic distributions of Bayesian models called
+  \enquote{exchangeable} with entropic priors
+  \cites{jaynes1986d_r1996,skilling1989b,skilling1990,rodriguez1989,rodriguez2002,portamana2017}[see
+  also][\chap~VI]{schroedinger1946_r1948}%[pp.~52--55]
+{jaynes1979b};
+\item as asymptotic distributions of Bayesian models based on sufficient
+  statistics
+  \cite{jaynes1982,%[\sect~I]
+jaynes1982b,portamana2017}.
+\end{enumerate}
+
+*** usually \enquote{explain} has a probabilistic meaning. We want to see
+whether information about moments give reasonable predictions about the
+rest. Necessity of making this precise. Three ways:
+
+-- \me\ method,
+
+-- Bayesian with entropic prior,
+
+-- Bayesian sufficiency.
+
+Applied to sample, all three are asymptotically equivalent. However, sample
+is representative of population. Need to take full population into
+consideration. This leads to different formulae, and also to
+non-equivalence of third way. Here we focus on first two.
+
+This kind of
 question is very vague until we define what we mean with \enquote{explain}
@@
   \text{with \enquote{$\pf(\text{\small activity at time bin } i \|\ytheta)$}
-    having maximum-entropy form with multipliers $\ytheta$};
+    having \me\ form with multipliers $\ytheta$};
 \end{gathered}
@@
 probability for the population activity is well-approximated by a
-maximum-entropy distribution \cite{jaynes1957,sivia1996_r2006,meadetal1984}
-with those Lagrange multipliers. This kind of distributions has been used
-in neuroscience at least since the 1990s to study correlations from diverse
+\me\ distribution
+\cite{jaynes1957,jaynes1963,sivia1996_r2006,meadetal1984} with those
+Lagrange multipliers. This kind of distributions has been used in
+neuroscience at least since the 1990s to study correlations from diverse
 points of view, for diverse purposes, and with data recorded from diverse
@@
 \cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015}.
-Part of this literature focused on the range of applicability of maximum-entropy
-distributions
+Part of this literature focused on the range of applicability of
+\me\ distributions
 \cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013}.
@@
 words, we implicitly or explicitly expect that the conclusions drawn with
-maximum-entropy methods about the recorded neurons should somehow
-generalize or extrapolate at least to whole areas those neurons are recorded from,
-\eg\ retina or motor cortex. We are thus assuming that the recorded sample
-is a \emph{representative sample} of that area.
-
-Whether such extrapolation makes biological sense is a difficult question,
-to be analysed case by case. The purpose
-
-The point that we want to stress is that
-
-From the observation of these samples we expect
-to learn something about all other neurons in the same brain area. That is,
-we assume that they are a \enquote{representative sample}.
-
-This note is mainly addressed to neuroscientists interested in \me\
-methods, but we would be pleased if its discussion of the probability of
-\enquote{sampling} were useful to neuroscientists that use other
-statistical methods to study the physical and dynamical characteristics of
-brain areas via neuronal recordings.
-
-Recent electrophysiological techniques \cite{berenyietal2014} in fact allow
-experimenters to record samples comprising even a couple hundreds neurons
-from specific brain areas. From the observation of these samples we expect
-to learn something about all other neurons in the same brain area. That is,
-we assume that they are a \enquote{representative sample}. We do not
-elaborate on the various important purposes of such recordings here, but
-stress that these sample sizes can be considered \enquote{large} because
-their statistical analysis requires considerable computational power.
-
-These computational costs are one, probably not the earliest, of the many
-reasons why \me\ methods have been introduced in neuroscience. It would be
-useful to somehow compress the statistical wealth of large neuronal
-recordings into few quantities, like sample moments for example. This
-compression might also entail interesting biological or functional
-properties of neuronal activity. The standard \me\ method
-\cite{jaynes1957,sivia1996_r2006,meadetal1984} accomplishes this kind of
-compression: it associates a unique probability distribution with few
-experimental quantities. But this is only one of its uses. It is also used
-for various information-theoretic purposes or to generate reference
-probability distributions
-\cite{bohteetal2000,schneidmanetal2006,shlensetal2006,roudietal2009b,mackeetal2011,tkaciketal2014b,shimazakietal2015}.
-In all these uses the \me\ distribution is chosen as the \enquote{maximally
-  noncommittal} one \cite{jaynes1963}. This adjective means little without
-further technical characterizations. Different works give different
-characterizations, but in this paper we will use the quoted expression as
-an umbrella term for all of them. Our results will not depend on the
-specific characterization of \enquote{noncommittal}.
+\me\ methods about the recorded neurons should somehow generalize or
+extrapolate at least to whole areas those neurons are recorded from, \eg\
+retina or motor cortex. We are thus assuming that the recorded sample is a
+\emph{representative sample} of that area. Whether such extrapolation makes
+biological sense is a difficult question, to be analysed case by case.
 
@@
 
-In this note we show that we must face a dilemma if we want to apply the
-\me\ method to a representative sample to find a maximally noncommittal
-distribution.
-
-The dilemma is this. We can apply the \me\ method to the sample, using a
-specified set of experimental constraints, and generate a probability
-distribution for the state of the sample. But our sample is representative
-of a larger population, in which it is biologically and functionally
-embedded. We can apply the \me\ method to the larger population, using the
-same constraints, and generate a probability distribution for its larger
-state, and then find the distribution for the sample by marginalization.
-Either application seems to have some commendable features. However,
-\emph{the distributions obtained from these two applications differ}. It
-goes without saying that if one is \enquote{noncommittal}, the other must
-be \enquote{committal}. Which choice is most meaningful?
+But a problem remains even if we assume that the sample is representative
+of its brain area. The problem lies in the way the \me\ methods
+is typically used in the literature. In this note we want to discuss this
+problem and offer a solution to it.
+
+The problem is this. We can apply the \me\ method to the sample, choosing
+some sufficient statistics, and generate a probability distribution for the
+state of the sample. But we consider our sample to be representative of a
+larger population, in which it is biologically and functionally embedded.
+We can apply the \me\ method to the larger population, using the same kind
+of sufficient statistics, and generate a probability distribution for its
+larger state, and then find the distribution for the sample by
+marginalization. Either application seems to have some commendable
+features. However, \emph{the distributions obtained from these two
+  applications differ}. It goes without saying that if one is
+\enquote{noncommittal}, the other must be \enquote{committal}. Which choice
+is most meaningful?
 
@@
   \\
-  \shortintertext{with}
+  \shortintertext{where}
   \label{eq:subpop_average}
@@
   \\[2\jot]
+  \shortintertext{with the conditional probability}
   \label{eq:conditional_hypergeometric}
@@
 
-The conditional probability $\ypp(\yrs,\yRf)$ relates the spaces of the
-sample average $\yXf\in\set{0,\dotsc,N}$ and of the population average
-$\yxs\in\set{0,\dotsc,n}$ in a special way. It is a coarsening projector of
-any probability $p$ for $\yXf$ onto a marginal probability $p_*$ for
-$\yxs$:
-\begin{equation}
-  \label{eq:projection_P}
-  % \ypp\colon p(\yXf=\dotv) \to
-  p_*(\yxs=\yrs)
-  = \sum_{N\yRf=0}^N \ypp(\yrs\|\yRf) p(\yXf=\yRf).
-\end{equation}
-Conversely it also a pulls back expectations of functions $f$ of the sample
-average $\yxs$ to expectations of functions $f^*$ of the population average
-$\yXf$:
+The conditional probability
+$\p(\yxs = \yrs \|\yXf=\yRf, \yH)\equiv\ypp(\yrs\|\yRf)$ relates the spaces
+of the sample average $\yXf\in\set{0,\dotsc,N}$ and of the population
+average $\yxs\in\set{0,\dotsc,n}$. It is a coarsening projector of any
+probability $p$ for $\yXf$ onto a marginal probability, as
+\eqn~\eqref{eq:subpop_average} shows. It also allows us to construct a
+function of the population average $f^*(\yXf)$, given a function of the
+sample average $f(\yxs)$:
 \begin{equation}
   \label{eq:pullback_P}
-%  \ypp^*\colon f(\yxs=\dotv) \to f^*(\yXf=\dotv) \defd \sum_{N\yrs=0}^n
-%  f(\yrs)\ypp(\yrs\|\dotv),
-  \begin{gathered}
   f^*(\yXf) \defd \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yXf),
-    \\
-   \expeb{f(\yxs)}= \expeb{f^*(\yXf)}
+\end{equation}  
+with the property of having the same expectation value as the latter:
+\begin{equation}
+  \label{eq:same_expect}
+  \expeb{f^*(\yXf)} \equiv
+  \sum_{N\yRf=0}^N f^*(\yRf)\,\p(\yXf=\yRf \| \yH)
   =
   \sum_{n\yrs=0}^n f(\yrs)\,\p(\yxs=\yrs \| \yH)
-    =
-    \sum_{N\yRf=0}^N f^*(\yRf)\,\p(\yXf=\yRf \| \yH).
-  \end{gathered}
+  \equiv
+  \expeb{f(\yxs)}. 
 \end{equation}
@@
 \end{figure}%hypergeometric_identity.nb
-A look at a plot of the hypergeometric distribution $\ypp(\yrs\|\yRf)$, see
+A look at a plot of the hypergeometric distribution $\ypp(\yrs\|\yRf)$,
 \fig~\ref{fig:hypergeom_proj}, reveals that it is a sort of \enquote{fuzzy
@@
 $f(\yxs) = \sav{\yx\dots\yx}\equiv\binom{n\yxs}{m}/\binom{n}{m}$,
-formulae~\eqref{eq:pullback_P} have the useful form
+ the expectations at the population and at the sample
+levels are exactly the same, as can be seen from \eqn~\eqref{eq:same_expect}:
 \begin{equation}
@@
 %  \mapsto
-  \begin{gathered}
-    \bigl( \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}\bigr)^* =
-    \av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}},
-    \\[2\jot]
+    % \bigl( \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}\bigr)^* =
+    % \av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}},
+    % \\[2\jot]
     % \expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH} =
@@
     \expe{\av{\yX \dotsm \yX} \| \yH}
-    % \\[2\jot]
-    % {}
-    = \binom{n}{m}^{-1}
-    % \frac{(n-m)!}{n!}
-    \sum_{\mathclap{n \yrs=0}}^n %m!\,
-    \binom{n \yrs}{m}\, \p(\yxs=\yrs \| \yH)
-    % \sum_{N \yRf=0}^N %m!\,
-    % \binom{N-m}{N \yRf-m} \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
-    = \binom{N}{m}^{-1}
-    % \frac{(N-m)!}{N!}
-    \sum_{\mathclap{N \yRf=0}}^N %m!\,
-    \binom{N \yRf}{m}\, \p(\yXf=\yRf \| \yH).
-  \end{gathered}
+    % % \\[2\jot]
+    % % {}
+    % = \binom{n}{m}^{-1}
+    % % \frac{(n-m)!}{n!}
+    % \sum_{\mathclap{n \yrs=0}}^n %m!\,
+    % \binom{n \yrs}{m}\, \p(\yxs=\yrs \| \yH)
+    % % \sum_{N \yRf=0}^N %m!\,
+    % % \binom{N-m}{N \yRf-m} \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
+    % = \binom{N}{m}^{-1}
+    % % \frac{(N-m)!}{N!}
+    % \sum_{\mathclap{N \yRf=0}}^N %m!\,
+    % \binom{N \yRf}{m}\, \p(\yXf=\yRf \| \yH).
 \end{equation}
@@
 %   Neuroscience} {\bf 15}(7):5249-5262.
-\defbibnote{postnote}{\small\par\medskip\noindent{\footnotesize% Note:
-\arxivp \mparcp \philscip \biorxivp}%
-}
+% \defbibnote{postnote}{\small\par\medskip\noindent{\footnotesize% Note:
+% \arxivp \mparcp \philscip \biorxivp}%
+% }
 
@@
 
-\printbibliography[postnote=postnote]
+\printbibliography%[postnote=postnote]
 

### 2018-05-13 maxNt_v2.tex
V2

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-13T22:11:43+0200
+%% Last-Updated: 2018-05-13T23:58:06+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 only data about simultaneous firings of two neurons, or of three, and so
-on. Maximum-entropy probability distributions can be introduced and justified in
+on. Let us take the case of the first two moments just as an example. The
+corresponding maximum-entropy probability distributions can be derived in
 three main ways:
 \begin{enumerate}
-\item from the maximum-entropy method
-  \cite{jaynes1957,jaynes1963,sivia1996_r2006};
-\item as asymptotic distributions of Bayesian models called
+\item From the maximum-entropy method
+  \cite{jaynes1957,jaynes1963,sivia1996_r2006}. The distribution is chosen
+  as the broadest one, in the Shannon-entropy sense, among those having
+  particular first and second moments.
+\item As asymptotic distributions of Bayesian models called
   \enquote{exchangeable} with entropic priors
@@
   also][\chap~VI]{schroedinger1946_r1948}%[pp.~52--55]
-{jaynes1979b};
-\item as asymptotic distributions of Bayesian models based on sufficient
+  {jaynes1979b}. The distribution is the limit, when the number of time
+  bins is very large, of
+  \begin{equation}
+    \label{eq:predictive_exchangeability}
+    \p(\text{\small activity at some time bin} \|
+    \text{\small 1st \amp\ 2nd empirical moments from observed bins},\text{\small other info});
+  \end{equation}
+  obtained from Bayes's theorem \emph{given} the observed empirical moments
+  and a particular initial joint distribution.
+\item As asymptotic distributions of Bayesian models based on sufficient
   statistics
   \cite{jaynes1982,%[\sect~I]
-jaynes1982b,portamana2017}.
+jaynes1982b,portamana2017}. The distribution is the limit, when the number of time
+bins is very large, of
+\begin{equation}
+  \label{eq:general_prediction}
+  \p(\text{\small activity at one time bin} \|
+  \text{\small activities at the remaining bins},\text{\small other info})
+\end{equation}
+under the assumption that only the values of the first and second empirical
+moments of the activities are \emph{relevant} for the probability above, \ie\
+\begin{multline}
+  \label{eq:informal_sufficiency}
+  \p(\text{\small activity at one time bin} \|
+  \text{\small activities at the other bins},\text{\small other info}) ={}\\
+  \p(\text{\small activity at some time bin} \|
+  \text{\small 1st \amp\ 2nd empirical moments},\text{\small other info});
+\end{multline}
+in other words, the assumption that the first and second moments are
+\emph{sufficient statistics}. This limit generally does not depend on the initial
+joint distribution.
 \end{enumerate}
 
-*** usually \enquote{explain} has a probabilistic meaning. We want to see
+***
+
+usually \enquote{explain} has a probabilistic meaning. We want to see
 whether information about moments give reasonable predictions about the

### 2018-05-14 maxNt_v2.tex
V2

--- 2018-05-13 maxNt_v2.tex
@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-13T22:11:43+0200
+%% Last-Updated: 2018-05-14T14:56:53+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Some investigations in the literature proceed as follows. We bin in time
-and binarize in intensity \cite{caianiello1961,caianiello1986} the recorded
-activity, transforming it into a sequence of $0$s (inactive) and $1$s
-(firing) for each neuron. Groups of two or more $1$s can appear in any time
-bin. Then we ask a question of this kind: can the presence of groups of
-three or more simultaneous firings be \enquote{explained} in terms of
-groups of just two, or three, or\ldots? This kind of question is very vague
-until we define what we mean with \enquote{explain} or similar words.
-Usually such \enquote{explanation} has a probabilistic meaning: we wonder
-whether the sole knowledge about two, or three, or more simultaneous
-firings can yield as good predictions about the whole activity as a more
-complete knowledge would do. To make this question concrete we must choose
-a probabilistic model.
-
-Maximum-entropy models
+Some investigations in the neuroscientific literature proceed as follows.
+We bin in time and binarize in intensity
+\cite{caianiello1961,caianiello1986} the recorded activity, transforming it
+into a sequence of $0$s (inactive) and $1$s (firing) for each neuron.
+Groups of two or more $1$s can appear in any time bin. Then we ask a
+question of this kind: can the presence of groups of three or more
+simultaneous firings be \enquote{explained} in terms of groups of just two,
+or three, or\ldots? This question is very vague until we define what we
+mean with \enquote{explain} or similar words. Usually such
+\enquote{explanation} has a probabilistic meaning: we wonder whether the
+sole knowledge about two, or three, or more simultaneous firings can yield
+as good predictions about the whole activity as a more complete knowledge
+would do. To make this question concrete we must first choose a
+probabilistic model from some well-defined assumptions, and then assess
+whether its final probabilities are \enquote{good} according to some
+well-defined measure. In the following we consider only the first,
+model-building, step.
+
+Maximum-entropy probability models
 \cite{jaynes1957,jaynes1963,sivia1996_r2006,meadetal1984} have been used in
-neuroscience with this kind of purpose and for diverse brain areas at least
-since the 1990s
+neuroscience with this kind of purpose, for recordings from brain areas as
+diverse as retina and motor cortex, at least since the 1990s
 \cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015}.
-Part of this literature focused on the range of applicability of \me\
+Part of this literature focuses on the range of applicability of \me\
 distributions
 \cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.
-We assume that our readers are at least vaguely familiar with such models.
-In brief, they build a probability distribution for neuronal activity using
+We assume that our readers are at least vaguely familiar with this kind of model.
+In brief, it builds a probability distribution for neuronal activity using
 only data about simultaneous firings of two neurons, or of three, and so
-on. Maximum-entropy probability distributions can be introduced and justified in
-three main ways:
+on. Let us take the case of the first two moments as an example. The
+maximum-entropy distribution can be derived in three main ways:
 \begin{enumerate}
-\item from the maximum-entropy method
-  \cite{jaynes1957,jaynes1963,sivia1996_r2006};
-\item as asymptotic distributions of Bayesian models called
-  \enquote{exchangeable} with entropic priors
+\item From the maximum-entropy method
+  \cite{jaynes1957,jaynes1963,sivia1996_r2006}. The distribution is chosen
+  as the one with the highest Shannon entropy among those having particular
+  first and second moments. This distribution is deemed to be
+  \enquote{maximally noncommittal} \cite{jaynes1963}, but this adjective
+  means little without further technical characterization. Two
+  characterizations are given by the next two derivations.
+
+\item As the asymptotic distribution of a Bayesian model called
+  \enquote{exchangeable}, with a particular prior
   \cites{jaynes1986d_r1996,skilling1989b,skilling1990,rodriguez1989,rodriguez2002,portamana2017}[see
-  also][\chap~VI]{schroedinger1946_r1948}%[pp.~52--55]
-{jaynes1979b};
-\item as asymptotic distributions of Bayesian models based on sufficient
-  statistics
-  \cite{jaynes1982,%[\sect~I]
-jaynes1982b,portamana2017}.
-\end{enumerate}
-
-*** usually \enquote{explain} has a probabilistic meaning. We want to see
-whether information about moments give reasonable predictions about the
-rest. Necessity of making this precise. Three ways:
-
--- \me\ method,
-
--- Bayesian with entropic prior,
-
--- Bayesian sufficiency.
-
-Applied to sample, all three are asymptotically equivalent. However, sample
-is representative of population. Need to take full population into
-consideration. This leads to different formulae, and also to
-non-equivalence of third way. Here we focus on first two.
-
-This kind of
-question is very vague until we define what we mean with \enquote{explain}
-or similar words.
-
-A possible, precise definition is as follows. We want to predict, \ie\
-assign a probability to, the population activity in a time bin, given the
-population activities in the other bins:
+  also:][\chap~VI]{schroedinger1946_r1948}%[pp.~52--55]
+  {jaynes1979b}. This distribution is the limit, when the number of time
+  bins is very large, of
+  \begin{equation}
+    \label{eq:predictive_exchangeability}
+    \p(\text{\small activity in one time bin} \|
+    \text{\small 1st \amp\ 2nd empirical moments from other bins}%,\text{\small other info}
+    ),
+  \end{equation}
+  calculated with Bayes's theorem from a particular initial joint
+  distribution in which different time bins are almost independent. This
+  derivation produces a probability just by taking into account those two
+  moments, but does not exclude that other information might be relevant.
+  It can therefore be compared with a probability that uses additional
+  information, or with empirical frequencies.
+\item As the asymptotic distribution of a Bayesian model based on
+  sufficient statistics \cite{jaynes1982,%[\sect~I]
+    jaynes1982b,portamana2017}. This distribution is the limit, when the
+  number of time bins is very large, of
 \begin{equation}
   \label{eq:general_prediction}
-  \p(\text{\small activity at one time bin} \|
-  \text{\small activities at the remaining bins},\text{\small other info}).
-\end{equation}
-We can say that the simultaneous firing of at most two neurons
-\enquote{explains} the activity if only the values of the first and second
-empirical moments of the activities are relevant for the probability above:
+  \p(\text{\small activity in one time bin} \|
+  \text{\small activities in other bins}%,\text{\small other info}
+  )
+\end{equation}
+under the assumption that \emph{only} the values of the first and second
+empirical moments of the activities are relevant for the probability above,
+\ie\
 \begin{multline}
   \label{eq:informal_sufficiency}
-  \p(\text{\small activity at one time bin} \|
-  \text{\small activities at the other bins},\text{\small other info}) ={}\\
-  \p(\text{\small activity at some time bin} \|
-  \text{\small 1st \amp\ 2nd empirical moments},\text{\small other info});
+  \p(\text{\small activity in one time bin} \|
+  \text{\small activities in other bins}%,\text{\small other info}
+  ) \equiv{}\\
+  \p(\text{\small activity in one time bin} \|
+  \text{\small 1st \amp\ 2nd empirical moments from other bins}%,\text{\small other info}
+  );
 \end{multline}
-in statistical terms, the first and second moments are \emph{sufficient
-  statistics}. Other definitions are possible, of course; the definition
-above has the advantage of being mathematically very precise. In fact it
-determines, from just the three basic laws of the probability calculus, the
-mathematical form of the probability for the activity at any number of bins:
-\begin{equation}
-  \label{eq:sufficiency_maxent_family}
-  \begin{gathered}
-  \p(\text{\small activity at bins $1,2,\dotsc$} \|
-  \text{\small other info}) =
-  \int\Bigl[\prod_i 
-    \pf(\text{\small activity at bin } i \|\ytheta) \Bigr]\,
- \pf(\ytheta\| \text{\small other info})\,
-  \di\ytheta,
-  \\[\jot]
-  \text{with \enquote{$\pf(\text{\small activity at time bin } i \|\ytheta)$}
-    having \me\ form with multipliers $\ytheta$};
-\end{gathered}
-\end{equation}
-The proof that the assumption of sufficient statistics leads
-to the probability expression above is the Koopman-Pitman theorem
-\cite{koopman1936,pitman1936,darmois1935} and can be found in
-\textcite[\sect~4.5]{bernardoetal1994_r2000} and references therein. It can
-be extended to higher moments and other statistics.
-
-When many data are available (and unfortunately also when they aren't), the
-conditional density for the Lagrange multipliers,
-$\pf(\ytheta\| \text{data, other info})$ is strongly peaked at the unique
-values determined by the inverse problem \cite{meadetal1984}, and the
-probability for the population activity is well-approximated by a
-\me\ distribution
-\cite{jaynes1957,jaynes1963,sivia1996_r2006,meadetal1984} with those
-Lagrange multipliers. This kind of distributions has been used in
-neuroscience at least since the 1990s to study correlations from diverse
-points of view, for diverse purposes, and with data recorded from diverse
-brain areas
-\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015}.
-Part of this literature focused on the range of applicability of
-\me\ distributions
-\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013}.
-
-This kind of studies would have little significance if we did not expect
-their results to generalize beyond the particular recorded sample. In other
-words, we implicitly or explicitly expect that the conclusions drawn with
-\me\ methods about the recorded neurons should somehow generalize or
-extrapolate at least to whole areas those neurons are recorded from, \eg\
-retina or motor cortex. We are thus assuming that the recorded sample is a
-\emph{representative sample} of that area. Whether such extrapolation makes
-biological sense is a difficult question, to be analysed case by case.
+in other words, the assumption that the first and second moments are
+sufficient statistics. This limit generally does not depend on the initial
+joint distribution. This derivation produces a probability by explicitly
+assuming that only the first and second moments are informationally
+relevant: any other information would lead to the same conditional
+probability. It can therefore only be compared with empirical frequencies.
+\end{enumerate}
+We refer our readers to the cited references for a deeper discussion of
+these derivations. It is remarkable that all three lead to the same
+distribution. We shall now see, however, that if we apply these derivations
+more carefully they lead to different distributions.
 
@@
 
-But a problem remains even if we assume that the sample is representative
-of its brain area. The problem lies in the way the \me\ methods
-is typically used in the literature. In this note we want to discuss this
-problem and offer a solution to it.
+This kind of investigation of recorded samples of neuronal activity would
+have little significance if we did not expect its results to generalize
+beyond the particular recorded sample. In other words, we implicitly or
+explicitly expect that the conclusions drawn with \me\ methods about the
+recorded neurons should somehow generalize or extrapolate at least to whole
+areas those neurons are recorded from, \eg\ retina or motor cortex. We are
+thus assuming that the recorded sample is a \emph{representative sample} of
+that area. Whether such extrapolation makes biological sense is a difficult
+question, to be analysed case by case. For the moment let's assume that it
+does make biological sense, and that our sample can be assumed to be
+representative of its brain area.
+
+Then a problem appears in the way the \me\ method is typically applied, in
+the literature, to recorded or artificial samples. In this note we want to
+discuss this problem and offer a solution to it.
 
@@
 \end{document}
+
+----------- cut text ---------------------
+
+***
+
+usually \enquote{explain} has a probabilistic meaning. We want to see
+whether information about moments give reasonable predictions about the
+rest. Necessity of making this precise. Three ways:
+
+-- \me\ method,
+
+-- Bayesian with entropic prior,
+
+-- Bayesian sufficiency.
+
+Applied to sample, all three are asymptotically equivalent. However, sample
+is representative of population. Need to take full population into
+consideration. This leads to different formulae, and also to
+non-equivalence of third way. Here we focus on first two.
+
+This kind of
+question is very vague until we define what we mean with \enquote{explain}
+or similar words.
+
+A possible, precise definition is as follows. We want to predict, \ie\
+assign a probability to, the population activity in a time bin, given the
+population activities in the other bins:
+\begin{equation}
+  \label{eq:general_prediction}
+  \p(\text{\small activity at one time bin} \|
+  \text{\small activities at the remaining bins},\text{\small other info}).
+\end{equation}
+We can say that the simultaneous firing of at most two neurons
+\enquote{explains} the activity if only the values of the first and second
+empirical moments of the activities are relevant for the probability above:
+\begin{multline}
+  \label{eq:informal_sufficiency}
+  \p(\text{\small activity at one time bin} \|
+  \text{\small activities at the other bins},\text{\small other info}) ={}\\
+  \p(\text{\small activity at some time bin} \|
+  \text{\small 1st \amp\ 2nd empirical moments},\text{\small other info});
+\end{multline}
+in statistical terms, the first and second moments are \emph{sufficient
+  statistics}. Other definitions are possible, of course; the definition
+above has the advantage of being mathematically very precise. In fact it
+determines, from just the three basic laws of the probability calculus, the
+mathematical form of the probability for the activity at any number of bins:
+\begin{equation}
+  \label{eq:sufficiency_maxent_family}
+  \begin{gathered}
+  \p(\text{\small activity at bins $1,2,\dotsc$} \|
+  \text{\small other info}) =
+  \int\Bigl[\prod_i 
+    \pf(\text{\small activity at bin } i \|\ytheta) \Bigr]\,
+ \pf(\ytheta\| \text{\small other info})\,
+  \di\ytheta,
+  \\[\jot]
+  \text{with \enquote{$\pf(\text{\small activity at time bin } i \|\ytheta)$}
+    having \me\ form with multipliers $\ytheta$};
+\end{gathered}
+\end{equation}
+The proof that the assumption of sufficient statistics leads
+to the probability expression above is the Koopman-Pitman theorem
+\cite{koopman1936,pitman1936,darmois1935} and can be found in
+\textcite[\sect~4.5]{bernardoetal1994_r2000} and references therein. It can
+be extended to higher moments and other statistics.
+
+When many data are available (and unfortunately also when they aren't), the
+conditional density for the Lagrange multipliers,
+$\pf(\ytheta\| \text{data, other info})$ is strongly peaked at the unique
+values determined by the inverse problem \cite{meadetal1984}, and the
+probability for the population activity is well-approximated by a
+\me\ distribution
+\cite{jaynes1957,jaynes1963,sivia1996_r2006,meadetal1984} with those
+Lagrange multipliers. This kind of distributions has been used in
+neuroscience at least since the 1990s to study correlations from diverse
+points of view, for diverse purposes, and with data recorded from diverse
+brain areas
+\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015}.
+Part of this literature focused on the range of applicability of
+\me\ distributions
+\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013}.

### 2018-05-14 maxNt_v2.tex
V2

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-14T14:56:53+0200
+%% Last-Updated: 2018-05-14T15:43:37+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{enumerate}
-\item From the maximum-entropy method
+\item\label{item:maxent}From the maximum-entropy method
   \cite{jaynes1957,jaynes1963,sivia1996_r2006}. The distribution is chosen
@@
 
-\item As the asymptotic distribution of a Bayesian model called
+\item\label{item:entropic_prior}As the asymptotic distribution of a Bayesian model called
   \enquote{exchangeable}, with a particular prior
@@
   information, or with empirical frequencies.
-\item As the asymptotic distribution of a Bayesian model based on
+
+\item\label{item:sufficiency_model}As the asymptotic distribution of a Bayesian model based on
   sufficient statistics \cite{jaynes1982,%[\sect~I]
@@
 have little significance if we did not expect its results to generalize
-beyond the particular recorded sample. In other words, we implicitly or
-explicitly expect that the conclusions drawn with \me\ methods about the
-recorded neurons should somehow generalize or extrapolate at least to whole
-areas those neurons are recorded from, \eg\ retina or motor cortex. We are
-thus assuming that the recorded sample is a \emph{representative sample} of
-that area. Whether such extrapolation makes biological sense is a difficult
-question, to be analysed case by case. For the moment let's assume that it
-does make biological sense, and that our sample can be assumed to be
-representative of its brain area.
-
-Then a problem appears in the way the \me\ method is typically applied, in
-the literature, to recorded or artificial samples. In this note we want to
-discuss this problem and offer a solution to it.
-
-The problem is this. We can apply the \me\ method to the sample, choosing
-some sufficient statistics, and generate a probability distribution for the
-state of the sample. But we consider our sample to be representative of a
-larger population, in which it is biologically and functionally embedded.
-We can apply the \me\ method to the larger population, using the same kind
-of sufficient statistics, and generate a probability distribution for its
-larger state, and then find the distribution for the sample by
-marginalization. Either application seems to have some commendable
-features. However, \emph{the distributions obtained from these two
-  applications differ}. It goes without saying that if one is
-\enquote{noncommittal}, the other must be \enquote{committal}. Which choice
-is most meaningful?
+beyond the particular recorded sample. In other words, we expect that the
+conclusions drawn with \me\ methods about the recorded neurons should
+somehow generalize or extrapolate at least to larger areas from which those
+neurons are recorded. We are thus assuming that the recorded sample is a
+\emph{representative sample} of some larger area. Whether such
+extrapolation makes biological sense is a difficult question, to be
+analysed case by case. For the moment let's assume that it does make
+biological sense, and that our sample can be assumed to be representative
+of a larger area.
+
+Then a double dilemma appears in the way the \me\ method is typically
+applied in the neuroscientific literature to recorded or artificial
+samples. In this note we want to discuss one part of this dilemma and offer
+a solution to it.
+
+We can apply the \me\ method to the sample, choosing some set of moments,
+and generate a probability distribution for the state of the sample. But if
+we consider our sample to be representative of a larger population, it
+makes sense to apply the \me\ method to the larger population, using one of
+the three derivations above, and generate a probability distribution for
+the larger population. The distribution for the sample can then be found by
+marginalization. Here's the double dilemma:
+\begin{itemize}
+\item \emph{the distribution obtained from the application at the sample
+    level differs from those obtained from the application at the
+    population level};
+\item \emph{at the population level, the
+    derivations~\ref{item:entropic_prior} and~\ref{item:sufficiency_model}
+    lead to different distributions} (the derivations \ref{item:maxent}
+  and~\ref{item:entropic_prior} always lead to the same distribution).
+\end{itemize}
+
+Either application seems to have some commendable features. However,
+\emph{the distributions obtained from these two applications differ}. It
+goes without saying that if one is \enquote{noncommittal}, the other must
+be \enquote{committal}. Which choice is most meaningful?
 

### 2018-05-14 maxNt_v2.tex
V2

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-14T15:43:37+0200
+%% Last-Updated: 2018-05-14T16:10:02+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\hrule
+\mynote{delete this part}
+
 Either application seems to have some commendable features. However,
@@
 In the final discussion we present some arguments in favour of one choice.
-
-In the rest of the paper we mathematically formulate this dilemma. To this
-purpose we also present some probability relations relevant to sampling.
-The relations we present are well-known in survey sampling and in the
-pedagogic problem of drawing from an urn without replacement, yet they are
-somewhat hard to find explicitly written in the neuroscientific literature,
-so they may be of interest on their own. We briefly discuss also two very
-different meaning of \enquote{representative} that lead to similar initial
-probability assignments for sampling.
+\hrule
+
+In the rest of the paper we mathematically formulate and analyse the first
+part of this dilemma, focusing on the derivation~\ref{item:entropic_prior}.
+To this purpose we also present some probability relations relevant to
+sampling. The relations we present are well-known in survey sampling and in
+the pedagogic problem of drawing from an urn without replacement, yet they
+are somewhat hard to find explicitly written in the neuroscientific
+literature, so they may be of interest on their own. We briefly discuss
+also two very different meaning of \enquote{representative} that lead to
+similar initial probability assignments for sampling.
 
@@
 
+\hrule
+
+\mynote{DISCUSSION ABOUT THE SECOND STATE OF KNOWLEDGE IS TO BE DISCARDED}
+
 In the second state of knowledge or ignorance, denoted $\yHi$, we are
@@
 probabilities that are identical for $\yHp$ and $\yHi$.
+
+\hrule
 
@@
 
+\hrule
+
+\mynote{DISCUSSION ABOUT THE SECOND STATE OF KNOWLEDGE IS TO BE DISCARDED}
+
 How is it possible that the very different states of knowledge $\yHp$ and
@@
 full in a future work.
+
+\hrule
 
@@
 
-The second
-example uses two additional constraint functions $\expe{\sav{\yx\yx\yx}}$,
-$\expe{\sav{\yx\yx\yx\yx}}$. The distribution obtained at the population
-level has two modes, replaced by only one in the distribution obtained at
-the sample level; the tails are very different also in this case.
+The second example uses two additional constraint functions
+$\expe{\sav{\yx\yx\yx}}$, $\expe{\sav{\yx\yx\yx\yx}}$. The distribution
+obtained at the population level has two modes, replaced by only one in the
+distribution obtained at the sample level; the tails are very different
+also in this case.
+
+\begin{figure}[!t]
+\centering
+\includegraphics[width=0.9\linewidth]{newplot.pdf}
+% 
+\caption{\mynote{to be written} distributions obtained using
+  formulae~\eqref{eq:app_maxent_pop_marg_unknown_N} and the other two}
+\label{fig:newplot}
+\end{figure}%maxent_pop_or_sample.nb
 
@@
 \label{sec:discussion}
+
+\mynote{needs to be changed}
 

### 2018-05-14 maxNt_v3.tex
V2

--- 2017-12-04 maxNt.tex
@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2017-12-04T11:34:56+0100
+%% Last-Updated: 2018-05-15T01:35:24+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \author{
-  P.G.L. Porta Mana\\\texttt{mana@kth.org}
+  P.G.L. Porta Mana\\\texttt{piero.mana@ntnu.org}
   \And
-  V. Rostami$^*$\\\texttt{v.rostami@fz-juelich.de}
+  V. Rostami\\\texttt{v.rostami@fz-juelich.de}
+  \AND
+  E. Torre\\\texttt{torre@ibk.baug.ethz.ch}
     \And
-E. Torre\\\texttt{torre@ibk.baug.ethz.ch}
+Y. Roudi\\\texttt{yasser.roudi@ntnu.no}
   % David S.~Hippocampus\thanks{Use footnote for providing further
@@
 \section{Introduction: \me\ and sampling in neuroscience}
+
+Suppose that we have recorded the firing activity of a hundred neurons,
+sampled from a particular brain area. This scenario is a concrete
+possibility thanks to recent electrophysiological techniques
+\cite{berenyietal2014}. In 1985, Gerstein, Perkel, \amp\ Dayhoff wrote (our
+emphasis):
+\begin{quote}
+  the very profusion of data produced in such experiments creates serious
+  problems of analysis and interpretation; these problems are conceptual as
+  well as practical. The principal conceptual problems are (1)
+  \emph{defining cooperativity or functional grouping} among neurons and
+  (2) \emph{formulating quantitative criteria} for recognizing and
+  characterizing such cooperativity. \cite{gersteinetal1985}
+\end{quote}
+The neuroscientific literature has since offered several mathematical
+definitions of \enquote{cooperativity} or \enquote{functional grouping} and
+criteria to quantify it.
+
+One such quantitative criterion relies on the \me\ method
+\cite{jaynes1957,jaynes1963,sivia1996_r2006,meadetal1984}. This criterion
+has been used in neuroscience at least since the 1990s, applied to data
+recorded from brain areas as diverse as retina and motor cortex
+\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015},
+and it has been subjected to mathematical and conceptual scrutiny
+\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.
+
+\enquote{Cooperativity} can quantified and characterized with \me\ method
+in several ways. The simplest way proceeds along the following steps.
+Consider the recorded activity of a sample of $n$ neurons.
+\begin{enumerate}
+\item The activity of each neuron, ideally a continuous signal, is
+  transformed into into a sequence of \enquote{$0$}s (inactive) and
+  \enquote{$1$}s (firing) by binning the time and binarizing the signal
+  intensity \cite{caianiello1961,caianiello1986}.
+
+\item The total activity $n\yxs$ of the sample in each time bin is
+  therefore an integer between $0$ and $n$. The total pairwise activity in
+  each time bin, \ie\ the total number of simultaneously active pairs, is
+  combinatorially given by $\tbinom{n\yxs}{2}\equiv n\yxs\,(n\yxs-1)/2$,
+  and so on for higher-order simultaneous activities.
+
+\item We construct a collection of \me\ distribution for the total
+  activity, using as constraints: the time-averaged total activity; the
+  time-averaged total activity \emph{and} total pairwise activity; and so
+  on up to the time-averaged total activities of orders $1$ to $k$. Call
+  the resulting distributions $p_1(n\yxs), p_2(n\yxs),\dotsc,p_k(n\yxs)$.
+  We also have the empirical frequency distribution of the total activity,
+  $f(n\yxs)$, counted from the time bins.
+
+\item We can now compare the distributions above with one another and with
+  the frequency distribution, using some probability-space distance like
+  the relative entropy or discrimination information
+  \citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}. If we find,
+  for example, that such distance is very high between $p_1$ and $f$, but
+  very low between $p_k$ and $f$ for $k \ge 2$, then we can say that there
+  is a \enquote{pairwise cooperativity}, and that any higher-order
+  cooperativity is just a reflection or consequence of the pairwise one.
+\end{enumerate}
+The protocol above needs to be made precise by specifying various
+parameters, for example the width of the time bins.
+
+We hurry to say that the description just given is not \emph{the} way to
+quantify and characterize cooperativity and functional grouping, but just
+\emph{one} way. It can surely be criticized from many points of view. Yet,
+it is quantitative and more precise than an undefined notion of
+\enquote{cooperativity}. Two people who apply it to the same data will
+obtain the same numbers. Other protocols can be based on the \me\ method,
+for example protocols that take into account the activities or joint
+activities of individual neurons.
+
+The purpose of the present work is to show that there is a problem in this
+general way of applying the \me\ method to the data recorded from our
+sampled neurons.
+
+Suppose that the neurons were recorded with electrodes covering an area of
+some square millimetres. This recording is a \emph{sample} of the activity
+of the population covered by the electrode; or we can consider it as a
+sample of an even more extended brain area.
+
+The characterization of the cooperativity of the recorded sample would have
+little meaning if we did not expect its results to generalize to a larger
+area, at least that covered by the electrodes. In other words, we expect
+that the conclusions drawn with \me\ methods about the sampled neurons
+should somehow generalize or extrapolate at least to larger areas from
+which those neurons were sampled. We are thus assuming that the recorded
+sample is a \emph{representative sample} of some larger area.
+
+The problem is this. We can apply the \me\ method to the sample, as
+described in the above protocol, to generate probability distributions for
+the state of the sample. But, if our sample is representative of a larger
+population, we can apply the \me\ method to the larger population, using
+the constraints obtained from the sample, and generate a probability
+distribution for the larger population; the distribution for the sample is
+then obtained by marginalization. Either application seems to have some
+commendable features. However, \emph{the distributions obtained from these
+  two applications differ}. Which choice is most meaningful?
+
+***
 

### 2018-05-15 maxNt_v3.tex
V2

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-15T01:35:24+0200
+%% Last-Updated: 2018-05-15T11:12:59+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 One such quantitative criterion relies on the \me\ method
-\cite{jaynes1957,jaynes1963,sivia1996_r2006,meadetal1984}. This criterion
-has been used in neuroscience at least since the 1990s, applied to data
-recorded from brain areas as diverse as retina and motor cortex
+\cite{jaynes1957,jaynes1963,hobsonetal1973,sivia1996_r2006,meadetal1984}.
+This criterion has been used in neuroscience at least since the 1990s,
+applied to data recorded from brain areas as diverse as retina and motor
+cortex
 \cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015},
@@
 for example protocols that take into account the activities or joint
-activities of individual neurons.
-
-The purpose of the present work is to show that there is a problem in this
-general way of applying the \me\ method to the data recorded from our
-sampled neurons.
+activities of individual neurons, or even protocols that take into account
+time dependence.
+
+The purpose of the present work is to show that there is a problem in the
+way the \me\ method as sketched above, or more complex variations of it, is
+applied to the activity of the recorded neurons. The problem stems from the
+fact that the recorded neurons are a sample from a larger population.
 
 Suppose that the neurons were recorded with electrodes covering an area of
-some square millimetres. This recording is a \emph{sample} of the activity
-of the population covered by the electrode; or we can consider it as a
-sample of an even more extended brain area.
-
-The characterization of the cooperativity of the recorded sample would have
-little meaning if we did not expect its results to generalize to a larger
-area, at least that covered by the electrodes. In other words, we expect
-that the conclusions drawn with \me\ methods about the sampled neurons
-should somehow generalize or extrapolate at least to larger areas from
-which those neurons were sampled. We are thus assuming that the recorded
-sample is a \emph{representative sample} of some larger area.
-
-The problem is this. We can apply the \me\ method to the sample, as
+some square millimetres. This recording is a sample of the activity of the
+population covered by the electrode; we can consider it as a sample of an
+even more extended brain area. The characterization of the cooperativity of
+the recorded sample would have little meaning if we did not expect its
+results to generalize to a larger area, at least that covered by the
+electrodes. In other words, we expect that the conclusions drawn with the
+\me\ methods about the sampled neurons should somehow generalize or
+extrapolate at least to larger areas from which those neurons were sampled.
+We are thus assuming that the recorded sample is a \emph{representative
+  sample} of some larger area. Probability theory tells us how to make
+inferences from a sample to the larger population from which it is sampled
+\parentext{see references below}.
+
+Here is the problem. We can apply the \me\ method to the sample, as
 described in the above protocol, to generate probability distributions for
@@
 
-***
-
-% LM: I'd like to avoid common beginnings like "Today blah blah is
-% important" or "Blah blah is interesting/fundamental". They feel more like
-% a mechanical ritual than a real tentative of connecting with the reader.
-This note is mainly addressed to neuroscientists interested in \me\
-methods, but we would be pleased if its discussion of the probability of
-\enquote{sampling} were useful to neuroscientists that use other
-statistical methods to study the physical and dynamical characteristics of
-brain areas via neuronal recordings.
-
-Recent electrophysiological techniques \cite{berenyietal2014} in fact allow
-experimenters to record samples comprising even a couple hundreds neurons
-from specific brain areas. From the observation of these samples we expect
-to learn something about all other neurons in the same brain area. That is,
-we assume that they are a \enquote{representative sample}. We do not
-elaborate on the various important purposes of such recordings here, but
-stress that these sample sizes can be considered \enquote{large} because
-their statistical analysis requires considerable computational power.
-
-These computational costs are one, probably not the earliest, of the many
-reasons why \me\ methods have been introduced in neuroscience. It would be
-useful to somehow compress the statistical wealth of large neuronal
-recordings into few quantities, like sample moments for example. This
-compression might also entail interesting biological or functional
-properties of neuronal activity. The standard \me\ method
-\cite{jaynes1957,sivia1996_r2006,meadetal1984} accomplishes this kind of
-compression: it associates a unique probability distribution with few
-experimental quantities. But this is only one of its uses. It is also used
-for various information-theoretic purposes or to generate reference
-probability distributions
-\cite{bohteetal2000,schneidmanetal2006,shlensetal2006,roudietal2009b,mackeetal2011,tkaciketal2014b,shimazakietal2015}.
-In all these uses the \me\ distribution is chosen as the \enquote{maximally
-  noncommittal} one \cite{jaynes1963}. This adjective means little without
-further technical characterizations. Different works give different
-characterizations, but in this paper we will use the quoted expression as
-an umbrella term for all of them. Our results will not depend on the
-specific characterization of \enquote{noncommittal}.
-
-\bigskip
-
-In this note we show that we must face a dilemma if we want to apply the
-\me\ method to a representative sample to find a maximally noncommittal
-distribution.
-
-The dilemma is this. We can apply the \me\ method to the sample, using a
-specified set of experimental constraints, and generate a probability
-distribution for the state of the sample. But our sample is representative
-of a larger population, in which it is biologically and functionally
-embedded. We can apply the \me\ method to the larger population, using the
-same constraints, and generate a probability distribution for its larger
-state, and then find the distribution for the sample by marginalization.
-Either application seems to have some commendable features. However,
-\emph{the distributions obtained from these two applications differ}. It
-goes without saying that if one is \enquote{noncommittal}, the other must
-be \enquote{committal}. Which choice is most meaningful?
-
-In the final discussion we present some arguments in favour of one choice.
-
-In the rest of the paper we mathematically formulate this dilemma. To this
-purpose we also present some probability relations relevant to sampling.
+In this paper we develop the second way of applying the \me, at the level
+of the larger population, and show that its results differ from the
+application at the sample level. We also consider the case where the size
+of the larger population is unknown.
+
+To apply the \me\ method to the larger, unsampled population, it is
+necessary to use probability relations relevant to sampling
+\cites{ghoshetal1997}[parts~I,VI]{freedmanetal1978_r2007}[\chap~8]{gelmanetal1995_r2014}[\chap~3]{jaynes1994_r2003}.
 The relations we present are well-known in survey sampling and in the
@@
 somewhat hard to find explicitly written in the neuroscientific literature,
-so they may be of interest on their own. We briefly discuss also two very
-different meaning of \enquote{representative} that lead to similar initial
-probability assignments for sampling.
+so they may be of interest on their own.
 
@@
   answer with probability theory?} and \emph{which assumptions do we want
-  or need to make?} \fi
+  or need to make?}
 
@@
 \enquote{inhomogeneous} models as well.
-
-The notation in this note follows ISO and ANSI standards
-\citep{iso1993,ieee1993,nist1995} but for the use of the comma \enquote{,}
-to denote logical conjunction. Probability notation follows Jaynes
-\citep{jaynes1994_r2003}. By \enquote{probability} we mean a degree of
-belief which \enquote{would be agreed by all rational men if there were any
-  rational men} \cite{good1966}.
+\fi
+
+The notation and terminology in this note follows ISO and ANSI standards
+\citep{iso1993,ieee1993,nist1995,iso2006,iso2006b} but for the use of the
+comma \enquote{,} to denote logical conjunction. Probability notation
+follows Jaynes \citep{jaynes1994_r2003}. By \enquote{probability} we mean a
+degree of belief which \enquote{would be agreed by all rational men if
+  there were any rational men} \cite{good1966}.
 

### 2018-05-15 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-15T11:12:59+0200
+%% Last-Updated: 2018-05-15T12:04:44+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\newcommand*{\Vd}{{V.}}
-%\newcommand*{\etal}{{et al.}}
+\newcommand*{\etal}{{et al.}}
 %\newcommand*{\etsim}{{et sim.}}
@@
 \item The activity of each neuron, ideally a continuous signal, is
-  transformed into into a sequence of \enquote{$0$}s (inactive) and
+  transformed into a sequence of \enquote{$0$}s (inactive) and
   \enquote{$1$}s (firing) by binning the time and binarizing the signal
@@
 
-We hurry to say that the description just given is not \emph{the} way to
-quantify and characterize cooperativity and functional grouping, but just
-\emph{one} way. It can surely be criticized from many points of view. Yet,
-it is quantitative and more precise than an undefined notion of
+We hurry to say that the description just given is just \emph{one} way to
+quantify and characterize cooperativity and functional grouping, not
+\emph{the} only way. It can surely be criticized from many points of view.
+Yet, it is quantitative and more precise than an undefined notion of
 \enquote{cooperativity}. Two people who apply it to the same data will
@@
 necessary to use probability relations relevant to sampling
-\cites{ghoshetal1997}[parts~I,VI]{freedmanetal1978_r2007}[\chap~8]{gelmanetal1995_r2014}[\chap~3]{jaynes1994_r2003}.
+\cites{ghoshetal1997}[parts~I, VI]{freedmanetal1978_r2007}[\chap~8]{gelmanetal1995_r2014}[\chap~3]{jaynes1994_r2003}.
 The relations we present are well-known in survey sampling and in the
@@
 
-\section{Setup: population, sample, probabilities}
+\section{Notation: population, sample, probabilities}
 \label{sec:setup}
@@
 $(X_1,\dotsc,X_N) \defs \yX$ having fixed but unknown binary values
-$(R_1,\dotsc,R_N)\defs \yR \in \set{0,1}^N$. A particular sample of $n$
-neurons from this population has joint state $(x_1, \dotsc, x_n) \defs \yx$
-having fixed binary values $(r_1, \dotsc, r_n)\defs\yr \in \set{0,1}^n$. We
-will consider various averages of the population and the sample. For this
-purpose we introduce a general averaging operator $\sav{\dotv}$ defined by
+$(R_1,\dotsc,R_N)\defs \yR \in \set{0,1}^N$; \enquote{$X$} denote the
+variable, and \enquote{$R$} its value. A particular sample of $n$ neurons
+from this population has joint state $(x_1, \dotsc, x_n) \defs \yx$ having
+fixed binary values $(r_1, \dotsc, r_n)\defs\yr \in \set{0,1}^n$. We will
+consider various averages of the population and the sample. For this
+purpose we introduce a population-averaging operator $\sav{\dotv}$ defined
+by
 \begin{equation}
@@
 
-\section{Initial assumptions: the probability of representative samples}
+\section{Probability: relations between population and  sample}
 \label{sec:prob_samples}
 
-We need to make an initial probability assignment before any experimental
-observations are
+The theory of statistical sampling is covered in many excellent texts, for
+example Ghosh \amp\ Meeden \citey{ghoshetal1997} or Freedman, Pisani, \amp\
+Purves \citey[parts~I, VI]{freedmanetal1978_r2007}; summaries can be found
+in Gelman \etal\ \citey[\chap~8]{gelmanetal1995_r2014} and Jaynes
+\citey[\chap~3]{jaynes1994_r2003}.
+
+We need to make an initial probability assignment for the state of the
+population before any experimental observations are
 made. %, no matter what kinds of predictions we are interested in.
-This initial assignment will be modified by our experimental observations.
-Our probability assignment should reflect that the sample is somehow
-\enquote{representative} of the population. We consider here two states of
-knowledge that express this representativeness in different ways but lead
-to identical \emph{initial} probability assignments.
-
-In the first state of knowledge, denoted $\yHp$, we know that the neurons
-in the population are biologically or functionally similar, for
-example in morphology and kind of input or output they receive or give.
-Knowledge of this similarity leads us to assign a probability distribution
-for the population state $\yX$ that is symmetric under permutations of
-neuron identities, or \emph{exchangeable} as it is usually called.
-
-In the second state of knowledge or ignorance, denoted $\yHi$, we are
+This initial assignment will be modified by our experimental observations,
+which can involve only a sample of the population. Our initial probability
+assignment should reflect that samples are somehow representative of the
+whole population. We consider here a state of knowledge that express
+this representativeness.
+
+In this state of knowledge, denoted $\yH$, we know that the neurons in the
+population are biologically or functionally similar, for example in
+morphology or the kind of input or output they receive or give. But we are
 completely ignorant about the physical details of the individual neurons.
 Our ignorance is therefore symmetric under permutations of neuron
-identities. This also leads to an exchangeable probability distribution for
-$\yX$.
-
-Let us use $\yH$ to denote either of these two states of knowledge, in those
-probabilities that are identical for $\yHp$ and $\yHi$.
+identities. This ignorance leads us to assign a probability distribution
+for the population state $\yX$ that is symmetric under permutations of
+neuron identities; such a distribution is usually called \emph{finitely
+  exchangeable} \cites{ericson1969}[\chap~1]{ghoshetal1997}. We stress that
+this probability assignment is just an expression of the symmetry of our
+\emph{ignorance} about the state of the population, not an expression of
+some biologic or physical identity of the neurons.
 
 The \emph{representation theorem for finite exchangeability} states that
-the symmetric distribution of $\yH$ must obey
+the symmetric distribution in the state of knowledge $\yH$ must obey
 \begin{equation}

### 2018-05-15 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-15T12:04:44+0200
+%% Last-Updated: 2018-05-15T13:47:43+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 $(X_1,\dotsc,X_N) \defs \yX$ having fixed but unknown binary values
-$(R_1,\dotsc,R_N)\defs \yR \in \set{0,1}^N$; \enquote{$X$} denote the
-variable, and \enquote{$R$} its value. A particular sample of $n$ neurons
+$(R_1,\dotsc,R_N)\defs \yR \in \set{0,1}^N$; the \enquote{$X$} denote the
+variables, and \enquote{$R$} their values. A particular sample of $n$ neurons
 from this population has joint state $(x_1, \dotsc, x_n) \defs \yx$ having
@@
     \dotsb + X_{N-2} X_{N-1}X_N),
+    \qquad\text{and so on.}
     % & \sav{\yx\yx\yx} &\defd \tbinom{n}{3}^{-1} (x_1 x_2 x_3
@@
 \end{equation}
-and so on.
 These formulae say that $\yXf$ is the fraction of active neurons, $\yXXf$
@@
 Our ignorance is therefore symmetric under permutations of neuron
-identities. This ignorance leads us to assign a probability distribution
-for the population state $\yX$ that is symmetric under permutations of
-neuron identities; such a distribution is usually called \emph{finitely
-  exchangeable} \cites{ericson1969}[\chap~1]{ghoshetal1997}. We stress that
-this probability assignment is just an expression of the symmetry of our
+identities. This ignorance is represented by a probability distribution
+that is symmetric under permutations of neuron identities; such a
+distribution is usually called \emph{finitely exchangeable}
+\cites{ericson1969}[\chap~1]{ghoshetal1997}. We stress that this
+probability assignment is just an expression of the symmetry of our
 \emph{ignorance} about the state of the population, not an expression of
@@
 \cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-Our initial symmetric knowledge should intuitively also apply to the
-sample; indeed, the probability for the state of the
-sample~\eqref{eq:subpop_average} automatically satisfies the representation
+Our initial symmetric ignorance should intuitively also apply to the
+sample; the probability for the state of the
+sample~\eqref{eq:subpop_average} indeed satisfies the representation
 theorem~\eqref{eq:joint_plaus_N_homog} as well.
 
-How is it possible that the very different states of knowledge $\yHp$ and
-$\yHi$ lead to the same formulae above? Their difference appears as soon as
-we make an experimental observation, say $X_2 = R_2\in\set{0,1}$ and update our
-initial probabilities \eqref{eq:joint_plaus_N_homog}:
-\begin{equation}
-  \label{eq:updated_joint}
-    \p(\yX = \yR \| X_2 = R_2, \yH) \equiv
-  \p(\yX = \yR \| \yXf = \yRf, X_2 = R_2, \yH)\,
-  \p(\yXf=\yRf \| X_2 = R_2,\yH).
-\end{equation}
-The conditional probability and the probability for the average on the
-right side will update in very different ways for $\yHp$ and $\yHi$. The
-discussion of this update process in the two cases is unnecessary for the
-purposes of this note and outside their scope; we hope to address it in
-full in a future work.
-
-The conditional probability $\ypp(\yrs,\yRf)$ relates the spaces of the
-sample average $\yXf\in\set{0,\dotsc,N}$ and of the population average
-$\yxs\in\set{0,\dotsc,n}$ in a special way. It is a coarsening projector of
-any probability $p$ for $\yXf$ onto a marginal probability $p_*$ for
-$\yxs$:
-\begin{equation}
-  \label{eq:projection_P}
-  % \ypp\colon p(\yXf=\dotv) \to
-  p_*(\yxs=\yrs)
-  = \sum_{N\yRf=0}^N \ypp(\yrs\|\yRf) p(\yXf=\yRf).
-\end{equation}
-Conversely it also a pulls back expectations of functions $f$ of the sample
-average $\yxs$ to expectations of functions $f^*$ of the population average
-$\yXf$:
+The conditional probability
+$\p(\yxs = \yrs \|\yXf=\yRf, \yH) \equiv \ypp(\yrs,\yRf)$ relates the
+spaces of the sample average $\yXf\in\set{0,\dotsc,N}$ and of the
+population average $\yxs\in\set{0,\dotsc,n}$. It is a coarsening projector
+of any probability for $\yXf$ onto a marginal probability for $\yxs$, as
+\eqn~\eqref{eq:subpop_average} shows. For each function $f$ of the sample
+average $\yxs$, the  hypergeometric distribution 
+$\ypp(\yrs,\yRf)$ also allows us to construct a function $f^*$ of the
+population average $\yXf$
 \begin{equation}
   \label{eq:pullback_P}
+    f^*(\yXf) \defd \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yXf),
 %  \ypp^*\colon f(\yxs=\dotv) \to f^*(\yXf=\dotv) \defd \sum_{N\yrs=0}^n
 %  f(\yrs)\ypp(\yrs\|\dotv),
-  \begin{gathered}
-    f^*(\yXf) \defd \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yXf),
-    \\
-   \expeb{f(\yxs)}= \expeb{f^*(\yXf)}
+\end{equation}
+that has the same expectation as $f$:
+\begin{multline}
+  \expeb{f^*(\yXf)}
+  \equiv
+  \sum_{N\yRf=0}^N f^*(\yRf)\,\p(\yXf=\yRf \| \yH)
   =
+  \sum_{N\yRf=0}^N \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
+  ={}\\
   \sum_{n\yrs=0}^n f(\yrs)\,\p(\yxs=\yrs \| \yH)
-    =
-    \sum_{N\yRf=0}^N f^*(\yRf)\,\p(\yXf=\yRf \| \yH).
-  \end{gathered}
-\end{equation}
+  \equiv  \expeb{f(\yxs)}
+\end{multline}
 
@@
 
-Yet, the approximate equalities above express the fact that \emph{our
-  uncertainty about the sample is representative of our uncertainty about
-  the population and about other samples}, and vice versa. This fact comes
-about for very different reasons in the states of knowledge $\yHp$ and
-$\yHi$. In $\yHp$, because our sample is \emph{physically} representative
-of the population and of other samples; we could write this as
-$\yxs\approx\yXf$. In $\yHi$, because we are ignorant about sample and
-population in similar symmetric ways; but this does \emph{not} imply that
-$\yxs\approx\yXf$. New observations may in fact break this symmetry via
-\eqn~\eqref{eq:updated_joint}.
-
-Note that formulae~\eqref{eq:roughly_equal_nN} say more than the limits
+The approximate equalities above, on the other hand, express the fact that
+\emph{our uncertainty about the sample is representative of our uncertainty
+  about the population and about other samples}, and vice versa. These
+approximate equalities are not generally valid for other probability
+distributions for the population. They should be contrasted with the limits
 $\p(\yxs=a) \to \p(\yXf=a)$ and $\expeb{f(\yxs)} \to \expeb{f(\yXf)}$, as
-$n\to N$, do. These limits are trivially valid because the sample becomes
-the full population as $n\to N$. In particular, these limits hold even in
-cases where the conditional probability $\p(\yxs = \yrs \|\yXf=\yRf)$ is
-not a fuzzy identity and our uncertainties about sample and about
-population can differ wildly.
+$n\to N$, do: these limits are trivially, universally valid because the
+sample becomes the full population as $n\to N$. In particular, these limits
+hold even in cases where the conditional probability
+$\p(\yxs = \yrs \|\yXf=\yRf)$ is not a fuzzy identity, our uncertainties
+about sample and about population can differ wildly, and the approximate
+equalities~\eqref{eq:roughly_equal_nN} do not hold.
 
@@
 % \end{equation}
-For functions representing averaged products,
-$f(\yxs) = \sav{\yx\dots\yx}\equiv\binom{n\yxs}{m}/\binom{n}{m}$,
-formulae~\eqref{eq:pullback_P} have the useful form
+For averaged products, $\sav{\underbrace{\yx\dots\yx}_{\text{$m$
+      factors}}}\equiv\binom{n\yxs}{m}/\binom{n}{m}$,
+formulae~\eqref{eq:pullback_P} yield the useful equality
 \begin{equation}
@@
 %  \mapsto
-  \begin{gathered}
-    \bigl( \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}\bigr)^* =
-    \av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}},
-    \\[2\jot]
+    % \bigl( \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}\bigr)^* =
+    % \av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}},
+    % \\[2\jot]
     % \expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH} =
     % \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
-    \expe{\sav{\yx \dotsm \yx} \| \yH} =
-    \expe{\av{\yX \dotsm \yX} \| \yH}
+  \expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
+\equiv
     % \\[2\jot]
     % {}
-    = \binom{n}{m}^{-1}
+    \binom{n}{m}^{-1}
     % \frac{(n-m)!}{n!}
@@
     \sum_{\mathclap{N \yRf=0}}^N %m!\,
-    \binom{N \yRf}{m}\, \p(\yXf=\yRf \| \yH).
-  \end{gathered}
-\end{equation}
-The proof uses the expression for the $m$th factorial moment of the
-hypergeometric distribution \cite{potts1953}. Thus, in the states of
-knowledge $\yHp$ and $\yHi$ the averages of activity products \emph{are
-  initially the same for the sample and for the full population}. Similar
-relations can be found for the raw moments $\expe{\yxs^m}$ and
-$\expe{\yXf^m}$, which can be written in terms of the product expectations
-via \eqn~\eqref{eq:products_intermsof_average}.
+    \binom{N \yRf}{m}\, \p(\yXf=\yRf \| \yH)
+\equiv    \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}.
+\end{equation}
+The proof of the middle equality uses the expression for the $m$th
+factorial moment of the hypergeometric distribution \cite{potts1953}. Thus,
+in the state of knowledge $\yH$ the expectations of the averages of
+activity products \emph{are initially the same for the sample and for the
+  full population}. Similar relations can be found for the raw moments
+$\expe{\yxs^m}$ and $\expe{\yXf^m}$, which can be written in terms of the
+product expectations via \eqn~\eqref{eq:products_intermsof_average}.
 

### 2018-05-15 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-15T13:47:43+0200
+%% Last-Updated: 2018-05-15T22:50:14+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \DeclareMathSymbol{\mu}{\mathalpha}{egreeki}{"16}
-% \DeclareMathSymbol{\nu}{\mathalpha}{egreeki}{"17}
-% \DeclareMathSymbol{\xi}{\mathalpha}{egreeki}{"18}
+ \DeclareMathSymbol{\nu}{\mathalpha}{egreeki}{"17}
+ \DeclareMathSymbol{\xi}{\mathalpha}{egreeki}{"18}
 % \DeclareMathSymbol{\omicron}{\mathalpha}{egreeki}{"6F}
@@
 % \DeclareMathSymbol{\rho}{\mathalpha}{egreeki}{"1A}
-% \DeclareMathSymbol{\sigma}{\mathalpha}{egreeki}{"1B}
+ \DeclareMathSymbol{\sigma}{\mathalpha}{egreeki}{"1B}
 % \DeclareMathSymbol{\tau}{\mathalpha}{egreeki}{"1C}
@@
 % \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
-% \DeclareMathSymbol{\varXi}{\mathalpha}{egreeki}{"04}
+ \DeclareMathSymbol{\varXi}{\mathalpha}{egreeki}{"04}
 % \DeclareMathSymbol{\varOmicron}{\mathalpha}{egreeki}{"4F}
@@
 % \DeclareMathSymbol{\varRho}{\mathalpha}{egreeki}{"50}
-% \DeclareMathSymbol{\varSigma}{\mathalpha}{egreeki}{"06}
+ \DeclareMathSymbol{\varSigma}{\mathalpha}{egreeki}{"06}
 % \DeclareMathSymbol{\varTau}{\mathalpha}{egreeki}{"54}
@@
 
-\newcommand*{\yxx}{x}%subpop state
-\newcommand*{\yx}{\bm{\yxx}}%subpop state
+\newcommand*{\yvX}{\varSigma}
+\newcommand*{\yvR}{\sigma}
+\newcommand*{\yvx}{S}
+\newcommand*{\yvr}{s}
+\newcommand*{\yvN}{\nu}
+
+\newcommand*{\yx}{\bm{\yvx}}%subpop state
 \newcommand*{\yxs}{\sav{\yx}}%subpop av state
-\newcommand*{\yX}{\bm{X}}%pop state
-\newcommand*{\yy}{\bm{y}}%pop state
+\newcommand*{\yX}{\bm{\yvX}}%pop state
 \newcommand*{\yXf}{\av{\yX}}%pop av state
@@
 \newcommand*{\yXXf}{\av{\yX\yX}}%subpop av state
-\newcommand*{\yr}{\bm{r}}%subpop value
-\newcommand*{\ys}{\bm{s}}%subpop value
+\newcommand*{\yr}{\bm{\yvr}}%subpop value
 \newcommand*{\yrs}{\sav{\yr}}%subpop av value
-\newcommand*{\yR}{\bm{R}}%pop value
+\newcommand*{\yR}{\bm{\yvR}}%pop value
 \newcommand*{\yRf}{\av{\yR}}%pop av value
@@
 \begin{abstract}
-  This note shows that the \me\ method can be applied to a representative
-  sample from a neuronal population along two different routes: (1) apply
-  to the sample; or (2) apply to the population and marginalize to the
-  sample. These two routes give inequivalent results. Which route should be
-  chosen? Some arguments are presented in favour of the second. The note
-  also touches upon probability formulae of representative sampling and
-  discusses their possible meanings, a discussion that may be useful for
-   sampling problems in neuroscience.
+  \mynote{to be rewritten} This note shows that the \me\ method can be
+  applied to a representative sample from a neuronal population along two
+  different routes: (1) apply to the sample; or (2) apply to the population
+  and marginalize to the sample. These two routes give inequivalent
+  results. Which route should be chosen? Some arguments are presented in
+  favour of the second. The note also touches upon probability formulae of
+  representative sampling and discusses their possible meanings, a
+  discussion that may be useful for sampling problems in neuroscience.
   % The abstract must be limited to one paragraph.
@@
 Suppose that we have recorded the firing activity of a hundred neurons,
-sampled from a particular brain area. This scenario is a concrete
-possibility thanks to recent electrophysiological techniques
-\cite{berenyietal2014}. In 1985, Gerstein, Perkel, \amp\ Dayhoff wrote (our
-emphasis):
+sampled from a particular brain area. What are we to do with such data?
+Gerstein, Perkel, \amp\ Dayhoff \cite{gersteinetal1985} posed this question
+very tersely (our emphasis):
 \begin{quote}
-  the very profusion of data produced in such experiments creates serious
-  problems of analysis and interpretation; these problems are conceptual as
-  well as practical. The principal conceptual problems are (1)
-  \emph{defining cooperativity or functional grouping} among neurons and
-  (2) \emph{formulating quantitative criteria} for recognizing and
-  characterizing such cooperativity. \cite{gersteinetal1985}
+  The principal conceptual problems are (1) \emph{defining cooperativity or
+    functional grouping} among neurons and (2) \emph{formulating
+    quantitative criteria} for recognizing and characterizing such
+  cooperativity.
 \end{quote}
-The neuroscientific literature has since offered several mathematical
-definitions of \enquote{cooperativity} or \enquote{functional grouping} and
-criteria to quantify it.
-
-One such quantitative criterion relies on the \me\ method
+These questions have a long history, of course; see \eg\ the 1966 review by
+Moore \etal\ \cite{mooreetal1966}. % This scenario is a concrete possibility
+% thanks to recent electrophysiological techniques \cite{berenyietal2014}.
+The neuroscientific literature has offered several mathematical definitions
+of \enquote{cooperativity} or \enquote{functional grouping} and criteria to
+quantify it.
+
+One such quantitative criterion relies on the \me\ or relative-\me\ method
 \cite{jaynes1957,jaynes1963,hobsonetal1973,sivia1996_r2006,meadetal1984}.
@@
 \enquote{Cooperativity} can quantified and characterized with \me\ method
-in several ways. The simplest way proceeds along the following steps.
-Consider the recorded activity of a sample of $n$ neurons.
+in several ways. The simplest way roughly proceeds along the following
+steps. Consider the recorded activity of a sample of $n$ neurons.
 \begin{enumerate}
-\item The activity of each neuron, ideally a continuous signal, is
-  transformed into a sequence of \enquote{$0$}s (inactive) and
-  \enquote{$1$}s (firing) by binning the time and binarizing the signal
-  intensity \cite{caianiello1961,caianiello1986}.
-
-\item The total activity $n\yxs$ of the sample in each time bin is
-  therefore an integer between $0$ and $n$. The total pairwise activity in
-  each time bin, \ie\ the total number of simultaneously active pairs, is
-  combinatorially given by $\tbinom{n\yxs}{2}\equiv n\yxs\,(n\yxs-1)/2$,
-  and so on for higher-order simultaneous activities.
-
-\item We construct a collection of \me\ distribution for the total
-  activity, using as constraints: the time-averaged total activity; the
-  time-averaged total activity \emph{and} total pairwise activity; and so
-  on up to the time-averaged total activities of orders $1$ to $k$. Call
-  the resulting distributions $p_1(n\yxs), p_2(n\yxs),\dotsc,p_k(n\yxs)$.
+\item The activity of each neuron, a continuous signal, is binned in time
+  and binarized in intensity, and thus transformed into a sequence of
+  digits \enquote{$0$}s (inactive) and \enquote{$1$}s
+  \cite[\cf][]{caianiello1961,caianiello1986}.
+
+  Let the variable $\yvx_i(t)\in\set{0,1}$ denote the activity of the $i$th
+  sampled neuron at time bin $t$. The average activity at that bin is
+  therefore $\yxs(t) \defd \sum_i \yvx_i(t)/n$. If we count the number of
+  distinct pairs of active neurons at that bin we combinatorially find
+  $\tbinom{n\yxs(t)}{2}\equiv n\yxs(t)\,[n\yxs(t)-1]/2$. There can be at
+  most $\tbinom{n}{2}$ simultaneously active pairs, so the average pair
+  activity is
+  $\av{\yx \yx}(t) \defd \tbinom{n}{2}^{-1}\tbinom{n\yxs(t)}{2}$. With some
+  combinatorics we see that the average activity of $m$-tuples of neurons
+  is
+  \begin{equation}
+    \label{eq:products_intermsof_average}
+    \av{\underbrace{\yx\dotsm\yx}_{\text{$m$ terms}}}
+    = \binom{n}{m}^{-1}\binom{n\yxs(t)}{m}.
+  \end{equation}
+  For brevity let us agree to simply call \enquote{activity} the average
+  $\yxs$, \enquote{pair-activity} the average $\av{\yx \yx}$, and so on.
+
+\item Construct a sequence of relative-\me\ distributions for the activity
+  $\yxs$, using this sequence of constraints:
+  \begin{itemize}
+  \item the time average of the activities $\yxs(t)$;
+  \item the time averages of the activities $\yxs(t)$ and pair-activities
+    $\av{\yx \yx}(t)$;
+  \item \ldots and so on, including the time averages of $\yxs(t)$ up to
+    $\av{\yx\dotsm\yx}(t)$ with $k$ terms.
+  \end{itemize}
+  Call the resulting distributions $p_1(\yxs), p_2(\yxs),\dotsc,p_k(\yxs)$.
+  The time-bin dependence is now absent because these distributions can be
+  interpreted as referring to any one of the time bins $t$, or to a new
+  time bin containing new data.
+
   We also have the empirical frequency distribution of the total activity,
-  $f(n\yxs)$, counted from the time bins.
-
-\item We can now compare the distributions above with one another and with
-  the frequency distribution, using some probability-space distance like
-  the relative entropy or discrimination information
+  $f(\yxs)$, counted from the time bins.
+
+\item Now compare the distributions above with one another and with the
+  frequency distribution, using some probability-space distance like the
+  relative entropy or discrimination information
   \citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}. If we find,
-  for example, that such distance is very high between $p_1$ and $f$, but
-  very low between $p_k$ and $f$ for $k \ge 2$, then we can say that there
-  is a \enquote{pairwise cooperativity}, and that any higher-order
-  cooperativity is just a reflection or consequence of the pairwise one.
+  say, that such distance is very high between $p_1$ and $f$, very low
+  between $p_2$ and $f$, and is more or less the same between all $p_m$ and
+  $f$ for $m \ge 2$, then we can say that there is a \enquote{pairwise
+    cooperativity}, and that any higher-order cooperativity is just a
+  reflection or consequence of the pairwise one. The reason is that the
+  information from higher-order simultaneous activities did not lead to
+  appreciable changes in the distribution obtained from pair activities.
 \end{enumerate}
 The protocol above needs to be made precise by specifying various
-parameters, for example the width of the time bins.
+parameters, such as the width of the time bins or the probability
+distance used.
 
@@
 quantify and characterize cooperativity and functional grouping, not
-\emph{the} only way. It can surely be criticized from many points of view.
-Yet, it is quantitative and more precise than an undefined notion of
-\enquote{cooperativity}. Two people who apply it to the same data will
-obtain the same numbers. Other protocols can be based on the \me\ method,
-for example protocols that take into account the activities or joint
-activities of individual neurons, or even protocols that take into account
-time dependence.
-
-The purpose of the present work is to show that there is a problem in the
-way the \me\ method as sketched above, or more complex variations of it, is
-applied to the activity of the recorded neurons. The problem stems from the
-fact that the recorded neurons are a sample from a larger population.
+\emph{the only} way. It can surely be criticized from many points of view.
+Yet, it is quantitative and bears a more precise meaning than an undefined,
+vague notion of \enquote{cooperativity}. Two persons who apply this
+procedure to the same data will obtain the same numbers. Different
+protocols can be based on the \me\ method, for instance protocols that take
+into account the activities or pair activities of specific neurons rather
+than population averages, or even protocols that take into account time
+dependence.
+
+The purpose of the present work is not to assess the merits of \me\ methods
+with respect to other methods. Its main purpose is to show that there is a
+problem in the way the \me\ method itself, as sketched above, is applied to
+the activity of the recorded neurons. We believe that this problem is at
+the root of some quirks about this method that were pointed out in the
+literature \cite{roudietal2009b}. This problems extends also to more
+complex versions of the method, possibly except version that use
+\enquote{hidden} neurons. The problem stems from the fact that the recorded
+neurons are a \emph{sample} from a larger, unrecorded population.
 
 Suppose that the neurons were recorded with electrodes covering an area of
-some square millimetres. This recording is a sample of the activity of the
-population covered by the electrode; we can consider it as a sample of an
-even more extended brain area. The characterization of the cooperativity of
-the recorded sample would have little meaning if we did not expect its
-results to generalize to a larger area, at least that covered by the
-electrodes. In other words, we expect that the conclusions drawn with the
-\me\ methods about the sampled neurons should somehow generalize or
-extrapolate at least to larger areas from which those neurons were sampled.
-We are thus assuming that the recorded sample is a \emph{representative
-  sample} of some larger area. Probability theory tells us how to make
-inferences from a sample to the larger population from which it is sampled
-\parentext{see references below}.
+some square millimetres \cite[\cf][]{berenyietal2014}. This recording is a
+sample of the activity of the neuronal population under the recording
+device; such population can amount to tens of thousands of neurons
+\cite{abeles1991}. We could even consider the recorded neurons as a sample
+of a brain area more extended than the recording device.
+
+The characterization of the cooperativity of the recorded sample would have
+little meaning if we did not expect its results to generalize to a larger,
+unrecorded population -- at the very least that under the recording device.
+In other words, we expect that the conclusions drawn with the \me\ methods
+about the sampled neurons should somehow extrapolate to unrecorded neurons
+in some larger area, from which the recorded neurons were sampled. In
+statistical terms we are assuming that the recorded neurons are a
+\emph{representative sample} of some larger neuronal population.
+Probability theory tells us how to make inferences from a sample to the
+larger population from which it is sampled \parentext{see references
+  below}.
 
@@
 described in the above protocol, to generate probability distributions for
-the state of the sample. But, if our sample is representative of a larger
-population, we can apply the \me\ method to the larger population, using
-the constraints obtained from the sample, and generate a probability
-distribution for the larger population; the distribution for the sample is
-then obtained by marginalization. Either application seems to have some
-commendable features. However, \emph{the distributions obtained from these
-  two applications differ}. Which choice is most meaningful?
+the activity of the sample. But, if our sample is representative of a
+larger population, we can apply the \me\ method to the larger (unrecorded)
+population, using the constraints obtained from the recorded sample, and
+generate a probability distribution for the larger population. The
+distribution for the sample is then obtained by marginalization. Either
+application seems to have some commendable features. However, \emph{the
+  distributions obtained from these two applications differ}. Which choice
+is most meaningful?
 
@@
 
-We have a population of $N$ binary neurons. We assume that they can be
+We have a population of $\yvN$ binary neurons. We assume that they can be
 distinguished, by their spike shapes for example; but other details, like
 their locations, are unknown. The neurons have a joint state
-$(X_1,\dotsc,X_N) \defs \yX$ having fixed but unknown binary values
-$(R_1,\dotsc,R_N)\defs \yR \in \set{0,1}^N$; the \enquote{$X$} denote the
-variables, and \enquote{$R$} their values. A particular sample of $n$ neurons
-from this population has joint state $(x_1, \dotsc, x_n) \defs \yx$ having
-fixed binary values $(r_1, \dotsc, r_n)\defs\yr \in \set{0,1}^n$. We will
+$(\yvX_1,\dotsc,\yvX_\yvN) \defs \yX$ having fixed but unknown binary values
+$(\yvR_1,\dotsc,\yvR_\yvN)\defs \yR \in \set{0,1}^\yvN$; the \enquote{$\yvX$} denote the
+variables, and \enquote{$\yvR$} their values. A particular sample of $n$ neurons
+from this population has joint state $(\yvx_1, \dotsc, \yvx_n) \defs \yx$ having
+fixed binary values $(\yvr_1, \dotsc, \yvr_n)\defs\yr \in \set{0,1}^n$. We will
 consider various averages of the population and the sample. For this
@@
   \begin{gathered}
-    \av{\yX} \defd \tfrac{1}{N} (X_1 + X_2 + \dotsb + X_N),
+    \av{\yX} \defd \tfrac{1}{\yvN} (\yvX_1 + \yvX_2 + \dotsb + \yvX_\yvN),
     % & \sav{\yx} &\defd \tfrac{1}{n} (x_1 + x_2 + \dotsb + x_n),
    \qquad% \\[2\jot]
-    \av{\yX \yX} \defd \tbinom{N}{2}^{-1} (X_1 X_2 + X_1 X_3 + \dotsb +
-    X_{N-1} X_N),
+    \av{\yX \yX} \defd \tbinom{\yvN}{2}^{-1} (\yvX_1 \yvX_2 + \yvX_1 \yvX_3 + \dotsb +
+    \yvX_{\yvN-1} \yvX_\yvN),
     % & \sav{\yx \yx} &\defd\tbinom{n}{2}^{-1} (x_1 x_2 + x_1 x_3 + \dotsb
@@
     \\[2\jot]
-    \av{\yX\yX\yX} \defd \tbinom{N}{3}^{-1} (X_1 X_2 X_3 +% X_1 X_2 X_4
-    \dotsb + X_{N-2} X_{N-1}X_N),
+    \av{\yX\yX\yX} \defd \tbinom{\yvN}{3}^{-1} (\yvX_1 \yvX_2 \yvX_3 +% \yvX_1 \yvX_2 \yvX_4
+    \dotsb + \yvX_{\yvN-2} \yvX_{\yvN-1}\yvX_\yvN),
     \qquad\text{and so on.}
@@
 These formulae say that $\yXf$ is the fraction of active neurons, $\yXXf$
-the fraction of simultaneously active pairs out of all $\binom{N}{2}$
+the fraction of simultaneously active pairs out of all $\binom{\yvN}{2}$
 pairs, $\av{\yX\yX\yX}$ the fraction of simultaneously active triplets, and
-so on. Products of states like $X_i \dotsm X_j$ also have values in
+so on. Products of states like $\yvX_i \dotsm \yvX_j$ also have values in
 $\set{0,1}$; from this we can combinatorially  prove that
@@
   \av{\underbrace{\yX\dotsm\yX}_{\text{$m$ factors}}}
-  = \binom{N}{m}^{-1}\binom{N\yXf}{m}.
+  = \binom{\yvN}{m}^{-1}\binom{\yvN\yXf}{m}.
 \end{equation}
@@
   \label{eq:joint_plaus}
-  \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \| \yHc) \quad\text{or}\quad
-\p(\yX =\yR \| \yHc), \quad \yR \in \set{0,1}^N,
+  \p(\yvX_1=\yvR_1, \yvX_2=\yvR_2, \dotsc, \yvX_\yvN=\yvR_\yvN \| \yHc) \quad\text{or}\quad
+\p(\yX =\yR \| \yHc), \quad \yR \in \set{0,1}^\yvN,
 \end{equation}
@@
   \label{eq:sample_plaus}
-  \p(x_1=r_1, x_2=r_2, \dotsc, x_n=r_n \| \yHc) \quad\text{or}\quad
+  \p(\yvx_1=\yvr_1, \yvx_2=\yvr_2, \dotsc, \yvx_n=\yvr_n \| \yHc) \quad\text{or}\quad
 \p(\yx =\yr \| \yHc), \quad \yr \in \set{0,1}^n.
@@
   \p(\yXf=\yRf \| \yH) =
-  \binom{N}{N\yRf}^{-1} \p(\yXf=\yRf \| \yH),
+  \binom{\yvN}{\yvN\yRf}^{-1} \p(\yXf=\yRf \| \yH),
 \end{equation}
@@
 \cite{heathetal1976}. This theorem is intuitive: owing to symmetry, we must
-assign equal probabilities to all states with $N\yRf$ active neurons.
+assign equal probabilities to all states with $\yvN\yRf$ active neurons.
 
@@
   \label{eq:subpop_average}
-  \p(\yxs=\yrs \| \yH) = \sum_{N \yRf =0}^{N}
+  \p(\yxs=\yrs \| \yH) = \sum_{\yvN \yRf =0}^{\yvN}
     \p(\yxs = \yrs \|\yXf=\yRf, \yH)\,
@@
   \p(\yxs = \yrs \|\yXf=\yRf, \yH)=
-  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}
+  \binom{n}{n\yrs}\binom{\yvN-n}{\yvN \yRf-n\yrs}\binom{\yvN}{\yvN \yRf}^{-1}
   \defs \ypp(\yrs\|\yRf).
@@
 $\p(\yxs = \yrs \|\yXf=\yRf, \yH) \equiv \ypp(\yrs,\yRf)$ relates the
-spaces of the sample average $\yXf\in\set{0,\dotsc,N}$ and of the
+spaces of the sample average $\yXf\in\set{0,\dotsc,\yvN}$ and of the
 population average $\yxs\in\set{0,\dotsc,n}$. It is a coarsening projector
@@
 that has the same expectation as $f$:
-\begin{multline}
+\begin{equation}
   \expeb{f^*(\yXf)}
   \equiv
-  \sum_{N\yRf=0}^N f^*(\yRf)\,\p(\yXf=\yRf \| \yH)
+  \sum_{\yvN\yRf=0}^\yvN f^*(\yRf)\,\p(\yXf=\yRf \| \yH)
+%  =
+%  \sum_{N\yRf=0}^N \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
   =
-  \sum_{N\yRf=0}^N \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
-  ={}\\
   \sum_{n\yrs=0}^n f(\yrs)\,\p(\yxs=\yrs \| \yH)
-  \equiv  \expeb{f(\yxs)}
-\end{multline}
+  \equiv  \expeb{f(\yxs)}.
+\end{equation}
+The middle equality uses the combination of formulae~\eqref{eq:pullback_P}
+and~\eqref{eq:subpop_average}.
 
@@
   $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
-\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}$ for $N=5000$, $n=200$. (Band artifacts may appear in the
+\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yvN-n}{\yvN \yRf-n\yrs}\binom{\yvN}{\yvN \yRf}^{-1}$ for $\yvN=5000$, $n=200$. (Band artifacts may appear in the
   colourbar depending on your \textsc{pdf} viewer.)}
@@
 \fig~\ref{fig:hypergeom_proj}, reveals that it is a sort of \enquote{fuzzy
-  identity matrix} between the $\yXf$-space $\set{0,\dotsc,N}$ and
-$\yxs$-space $\set{0,\dotsc,n}$. When $n=N$ it is the identity matrix. We
+  identity matrix} between the $\yXf$-space $\set{0,\dotsc,\yvN}$ and
+$\yxs$-space $\set{0,\dotsc,n}$. When $n=\yvN$ it is the identity matrix. We
 thus have that
@@
 $\p(\yxs=a) \to \p(\yXf=a)$ and $\expeb{f(\yxs)} \to \expeb{f(\yXf)}$, as
-$n\to N$, do: these limits are trivially, universally valid because the
-sample becomes the full population as $n\to N$. In particular, these limits
+$n\to \yvN$, do: these limits are trivially, universally valid because the
+sample becomes the full population as $n\to \yvN$. In particular, these limits
 hold even in cases where the conditional probability
@@
 %   \qquad\text{or}\qquad
-%   \sum_{n\yrs=0}^n f(\yrs)\,p_*(\yxs=\yrs) = \sum_{N\yRf=0}^N f^*(\yRf)\,p(\yXf=\yRf).
+%   \sum_{n\yrs=0}^n f(\yrs)\,p_*(\yxs=\yrs) = \sum_{\yvN\yRf=0}^\yvN f^*(\yRf)\,p(\yXf=\yRf).
 % \end{equation}
@@
     % \binom{N-m}{N \yRf-m} \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
-    = \binom{N}{m}^{-1}
+    = \binom{\yvN}{m}^{-1}
     % \frac{(N-m)!}{N!}
-    \sum_{\mathclap{N \yRf=0}}^N %m!\,
-    \binom{N \yRf}{m}\, \p(\yXf=\yRf \| \yH)
+    \sum_{\mathclap{\yvN \yRf=0}}^\yvN %m!\,
+    \binom{\yvN \yRf}{m}\, \p(\yXf=\yRf \| \yH)
 \equiv    \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}.
@@
 
-\section{Enter maximum-entropy: dilemma}
+\section{Maximum-entropy at the population level with constraints at the
+  sample level}
 \label{sec:specific_initial_probability}
@@
 not completely specified, because \eqn~\eqref{eq:subpop_average} linearly
-constrains $N+1$ unknowns by only $n+1$ equations.
+constrains $\yvN+1$ unknowns by only $n+1$ equations.
 
@@
   \p(\yXf = \yRf \| \yHa)  = \yK\,
-  \binom{N}{N \yRf}\,\exp\Bigl[\yL\T
+  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[\yL\T
   \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
@@
   \p(\yxs = \yrs \| \yHa)  = \yK\, 
-  \sum_{N\yRf=0}^N \ypp(\yrs \|\yRf)
+  \sum_{\yvN\yRf=0}^\yvN \ypp(\yrs \|\yRf)
 \,
-  \binom{N}{N \yRf}\,\exp\Bigl[\yL\T
+  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[\yL\T
   \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
@@
   \label{eq:app_maxent_pop_constraints}
-\yc = \yK\sum_{n\yrs=0}^n \sum_{N\yRf=0}^N \yf(\yrs) \ypp(\yrs \|\yRf)
+\yc = \yK\sum_{n\yrs=0}^n \sum_{\yvN\yRf=0}^\yvN \yf(\yrs) \ypp(\yrs \|\yRf)
 \,
-  \binom{N}{N \yRf}\,\exp\Bigl[\yL\T
+  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[\yL\T
   \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
@@
   \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:app_maxent_sample}, with $N=5000$,
+  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yvN=5000$,
   $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$.}
@@
   \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:app_maxent_sample}, with $N=5000$,
+  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yvN=5000$,
   $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$,
@@
 \figs~\ref{fig:diff_maxent_pop_sample}
-and~\ref{fig:diff_maxent_pop_sample_realdata}, for $N=5000$, $n=200$, and
+and~\ref{fig:diff_maxent_pop_sample_realdata}, for $\yvN=5000$, $n=200$, and
 constraint functions of the form
@@
 Suppose we do not know whether the sample is really part of a larger
-population: we do not know whether $N=n$ or how large $N$ is otherwise.
+population: we do not know whether $\yvN=n$ or how large $\yvN$ is otherwise.
 Call this state of ignorance $\yHd$. In the probability calculus this
-ignorance about $N$ is expressed by assigning a probability distribution
-$\p(N \| \yHd)$ that vanishes if $N<n$, since we know that $N\ge n$; see
+ignorance about $\yvN$ is expressed by assigning a probability distribution
+$\p(\yvN \| \yHd)$ that vanishes if $\yvN<n$, since we know that $\yvN\ge n$; see
 Good \cite{good1965,good1967b} and Rissanen \cite{rissanen1983} for
@@
 assumption of symmetric ignorance, probability assignments that do not
-assume a specific value of $N$ are then obtained via multiplication of all
-$N$-dependent probabilities by $\p(N \| \yHd)$ and subsequent
-marginalization over $N$. Technically speaking, $N$ becomes a
+assume a specific value of $\yvN$ are then obtained via multiplication of all
+$\yvN$-dependent probabilities by $\p(\yvN \| \yHd)$ and subsequent
+marginalization over $\yvN$. Technically speaking, $\yvN$ becomes a
 \emph{nuisance parameter}
@@
   \label{eq:app_maxent_pop_marg_unknown_N}
-  \p(\yxs = \yrs \| \yHd)  = \sum_N \biggl\{\yK_N\, 
-  \sum_{N\yRf=0}^N \ypp_N(\yrs \|\yRf)
+  \p(\yxs = \yrs \| \yHd)  = \sum_\yvN \biggl\{\yK_\yvN\, 
+  \sum_{\yvN\yRf=0}^\yvN \ypp_\yvN(\yrs \|\yRf)
 \,
-  \binom{N}{N \yRf}\,\exp\Bigl[{\yL_N}\T
-  \sum_{n\yrs=0}^n \yf(\yrs)\ypp_N(\yrs\|\yRf)\Bigr]\biggr\}
-  \,\p(N \| \yHd),
-\end{equation}
-where $N$-dependencies have been made explicit. This is a formidable
+  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[{\yL_\yvN}\T
+  \sum_{n\yrs=0}^n \yf(\yrs)\ypp_\yvN(\yrs\|\yRf)\Bigr]\biggr\}
+  \,\p(\yvN \| \yHd),
+\end{equation}
+where $\yvN$-dependencies have been made explicit. This is a formidable
 expression. But our question, \enquote{is the usual \me\ at the sample
@@
 distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
-\eqref{eq:app_maxent_sample} equal for some choice of $\p(N \| \yHd)$, with
-$\p(N \| \yHd)\ne 0$ for $N>n$?}. We leave this mathematical problem for
+\eqref{eq:app_maxent_sample} equal for some choice of $\p(\yvN \| \yHd)$, with
+$\p(\yvN \| \yHd)\ne 0$ for $\yvN>n$?}. We leave this mathematical problem for
 future work. Note, however, that this equality is satisfied if
-$\p(N=1 \| \yHd)=1$, which means that the usual \me\ model can also be
+$\p(\yvN=1 \| \yHd)=1$, which means that the usual \me\ model can also be
 interpreted as assuming that \emph{no} larger population exists.
@@
 surprise analysis, may reveal features in a dataset that were unnoticed by
-the standard \me\ model. The question remains of how to specify $N$,
-though. We have tacitly intended $N$ as the size of the largest
+the standard \me\ model. The question remains of how to specify $\yvN$,
+though. We have tacitly intended $\yvN$ as the size of the largest
 biologically or functionally homogeneous population from which our sample
 was recorded. It could be the amount of neurons in a functional brain area,
-for example the primary visual cortex, for which $N\sim 10^8$
-\cite{leubaetal1994}. For large $N$ -- unfortunately we are not yet able to
+for example the primary visual cortex, for which $\yvN\sim 10^8$
+\cite{leubaetal1994}. For large $\yvN$ -- unfortunately we are not yet able to
 translate this \enquote{large} into a numeric order of magnitude -- the
-final distribution becomes independent of $N$, and continuous
+final distribution becomes independent of $\yvN$, and continuous
 approximations become available.
@@
   the same in either case: $\expe{\yxs \|\yH}=0.045$,
-  $\expe{\yxxs \|\yH}=0.0025$, and $N=5000$, $n=200$.}
+  $\expe{\yxxs \|\yH}=0.0025$, and $\yvN=5000$, $n=200$.}
 \label{fig:diff_maxent_pop_sample_realdata}
@@
 expectations for individual neurons or groups of neurons are constrained,
-for example $\expe{x_2}$ or $\expe{x_1x_8x_9}$. A short computation shows
+for example $\expe{\yvx_2}$ or $\expe{\yvx_1x_8x_9}$. A short computation shows
 that the \me\ method with this kind of constraints gives the same result
@@
 
-% \mynote{Discussion of case when individual $X_i\dotsm X_j$ are constrained.
+% \mynote{Discussion of case when individual $\yvX_i\dotsm \yvX_j$ are constrained.
 % Likely breakdown of \me\ in this case \cite{portamana2009}. Hints at full
@@
 \begin{align}
-  \expe{x_{i_1} \dotsm x_{i_m} \| \yH}
+  \expe{\yvx_{i_1} \dotsm \yvx_{i_m} \| \yH}
 &=
@@
   \begin{split}
-\expe{x_{i_1} \dotsm x_{i_m} \| \yH}
+\expe{\yvx_{i_1} \dotsm \yvx_{i_m} \| \yH}
 &=
-\binom{N}{m}^{-1}
+\binom{\yvN}{m}^{-1}
 %\frac{(N-m)!}{N!}
-\sum_{N A=0}^N %m!\, 
-\binom{N A}{m}\, Q(A)
+\sum_{\yvN A=0}^\yvN %m!\, 
+\binom{\yvN A}{m}\, Q(A)
 \\
@@
 %\frac{(N-m)!}{N!}
-\sum_{N A=0}^N %m!\, 
-\binom{N-m}{N A-m}
-\binom{N}{N A}^{-1} Q(A).
+\sum_{\yvN A=0}^\yvN %m!\, 
+\binom{\yvN-m}{\yvN A-m}
+\binom{\yvN}{\yvN A}^{-1} Q(A).
 \end{split}
@@
 neurons \enquote{units} to lend some generality to our discussion. We shall
-make statements about the whole population of $N$ units and about a
+make statements about the whole population of $\yvN$ units and about a
 subpopulation of $n$ units; the word \enquote{population} will always refer
 to the \emph{whole} population. The subpopulation states and their values
-are denoted by lowercase letters: $(x_1, \dotsc, x_n)\equiv\yx$ and
-$(r_1, \dotsc, r_n)\equiv\yr$; but note that $x_i \equiv X_{j_i}$ and
-$r_i \equiv R_{j_i}$ for some distinct $j_1,\dotsc,j_n$. We shall also make
+are denoted by lowercase letters: $(\yvx_1, \dotsc, \yvx_n)\equiv\yx$ and
+$(\yvr_1, \dotsc, \yvr_n)\equiv\yr$; but note that $\yvx_i \equiv \yvX_{j_i}$ and
+$\yvr_i \equiv \yvR_{j_i}$ for some distinct $j_1,\dotsc,j_n$. We shall also make
 statements about the population-averaged state, or \emph{population
@@
   \label{noeq:population-av}
-   \yXf &\defd (X_1 + \dotsb + X_N)/N,
+   \yXf &\defd (\yvX_1 + \dotsb + \yvX_\yvN)/\yvN,
 \\
@@
 \label{noeq:subpopulation-av}
-\yxs &\defd (x_1 + \dotsc + x_n)/n.
+\yxs &\defd (\yvx_1 + \dotsc + \yvx_n)/n.
 \end{align}
-The quantities $N\yXf$ and $n\yxs$ represent the total number of active
+The quantities $\yvN\yXf$ and $n\yxs$ represent the total number of active
 units in the population and the subpopulation. Quantities like $\yRf$ and $\yrs$
@@
 $\av{\dotv}$ are also extended to averages of $\tbinom{n}{m}$ or
-$\tbinom{N}{m}$ products of $m$ states; \eg,
+$\tbinom{\yvN}{m}$ products of $m$ states; \eg,
 \begin{align}
 \av{\yX \yX} &\defd
-\tbinom{N}{2}^{-1} (X_1 X_2 + X_1 X_3  + \dotsb +  X_{N-1} X_N),
+\tbinom{\yvN}{2}^{-1} (\yvX_1 \yvX_2 + \yvX_1 \yvX_3  + \dotsb +  \yvX_{\yvN-1} \yvX_\yvN),
 \\[2\jot]
 \sav{\yx\yx\yx} &\defd
-\tbinom{n}{3}^{-1} (x_1 x_2 x_3 + x_1 x_2 x_4 + \dotsb + x_{n-2} x_{n-1} x_n),
+\tbinom{n}{3}^{-1} (\yvx_1 \yvx_2 \yvx_3 + \yvx_1 \yvx_2 \yvx_4 + \dotsb + \yvx_{n-2} \yvx_{n-1} \yvx_n),
 \end{align}
@@
   \label{noeq:joint_plaus}
-  \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \| \yH) \quad\text{or}\quad
+  \p(\yvX_1=\yvR_1, \yvX_2=\yvR_2, \dotsc, \yvX_\yvN=\yvR_\yvN \| \yH) \quad\text{or}\quad
 \p(\yX =\yR \| \yH).
@@
   \label{noeq:homog_math}
-  \p(X_1=R_1, X_2=R_2, \dotsc, X_N=R_N \| \yH) ={}
+  \p(\yvX_1=\yvR_1, \yvX_2=\yvR_2, \dotsc, \yvX_\yvN=\yvR_\yvN \| \yH) ={}
 \\ 
-\shoveright{\p(X_1=R_{\pi(1)}, X_2=R_{\pi(2)}, \dotsc, X_N=R_{\pi(N)} \| \yH)}
+\shoveright{\p(\yvX_1=\yvR_{\pi(1)}, \yvX_2=\yvR_{\pi(2)}, \dotsc, \yvX_\yvN=\yvR_{\pi(\yvN)} \| \yH)}
 \\
@@
 \qquad
-A \in \set*{0,\tfrac{1}{N},\tfrac{2}{N},\dotsc,1}.
+A \in \set*{0,\tfrac{1}{\yvN},\tfrac{2}{\yvN},\dotsc,1}.
 \end{equation}
@@
   \label{noeq:joint_plaus_N_homog}
-  \p(\yX = \yR \| H) = \binom{N}{N\yRf}^{-1} Q(\yRf).
+  \p(\yX = \yR \| H) = \binom{\yvN}{\yvN\yRf}^{-1} Q(\yRf).
 \end{equation}
@@
     \p(\yx = \yr \| \yH) =
-\sum_{N A =0}^{N} 
-\binom{N-n}{N A-n\yrs}\binom{N}{N A}^{-1}\,
+\sum_{\yvN A =0}^{\yvN} 
+\binom{\yvN-n}{\yvN A-n\yrs}\binom{\yvN}{\yvN A}^{-1}\,
 Q(A).
@@
 Note that the only summands contributing to this sum are those for which
-$n\yrs \le NA \le N$; the others are zero because by definition
+$n\yrs \le NA \le \yvN$; the others are zero because by definition
 $\tbinom{M}{y}=0$ if $y<0$. This remark applies to all the sums of this
@@
 =
-\binom{N-n}{N \yRf-n \yrs}.
+\binom{\yvN-n}{\yvN \yRf-n \yrs}.
 \end{equation}
@@
   \p(\yxs = a  \| \yH) = \binom{n}{n a}
- \sum_{N A=0}^{N}
-\binom{N-n}{N A-n a}\binom{N}{N A}^{-1}\,
+ \sum_{\yvN A=0}^{\yvN}
+\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1}\,
 Q(A),
@@
 =
-\binom{n}{n a}\binom{N-n}{N A-n a}\binom{N}{N A}^{-1}.
+\binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1}.
 \end{equation}
@@
   \begin{equation*}
-    x_{i_1} x_{i_2} \dotsm x_{i_m},
+    \yvx_{i_1} \yvx_{i_2} \dotsm \yvx_{i_m},
     \qquad 1\le i_1 < i_2 < \dotsb < i_m \le n
@@
   \label{noeq:eq_factmom_prod}
-\expe{x_{i_1} \dotsm x_{i_m} \| \yH}
+\expe{\yvx_{i_1} \dotsm \yvx_{i_m} \| \yH}
 =
@@
 =
-  \tfrac{(N-m)!}{N!}\expeb*{\frac{(N\yXf)!}{(N\yXf-m)!} \| \yH}.
+  \tfrac{(\yvN-m)!}{\yvN!}\expeb*{\frac{(\yvN\yXf)!}{(\yvN\yXf-m)!} \| \yH}.
 \end{equation}
@@
     \expe{\yxs^2 \| \yH} &= \expe{\yXf \| \yH}
-                                  \frac{N-n}{(N-1)n}
+                                  \frac{\yvN-n}{(\yvN-1)n}
  +{}
-\\&\quad\quad \expe{\yXf^2 \| \yH}\frac{N\,(n-1)}{(N-1)n},
+\\&\quad\quad \expe{\yXf^2 \| \yH}\frac{\yvN\,(n-1)}{(\yvN-1)n},
 \end{split}
@@
     \expe{\yxs^3 \| \yH} &=
-                              \expe{\yXf \| \yH}\frac{(N-n)(N-2n)}{(N-1)(N-2)n^2}+{}
+                              \expe{\yXf \| \yH}\frac{(\yvN-n)(\yvN-2n)}{(\yvN-1)(\yvN-2)n^2}+{}
                                     \\
                                     &\quad\quad\expe{\yXf^2 \| \yH}
-                                    \frac{3N\,(N-n)(n-1)}{(N-1)(N-2)n^2} +{}
+                                    \frac{3N\,(\yvN-n)(n-1)}{(\yvN-1)(\yvN-2)n^2} +{}
                                     \\
                                     &\quad\quad\quad\expe{\yXf^3 \| \yH}
-                                    \frac{N^2\,(n-1)(n-2)}{(N-1)(N-2)n^2}.
+                                    \frac{\yvN^2\,(n-1)(n-2)}{(\yvN-1)(\yvN-2)n^2}.
                                   \end{split}
@@
   \label{noeq:lin_transf_moments}
-  \expe{\yxs^m \| \yH} = \sum_{l=1}^m M_{ml}(n,N) \expe{\yXf^l \| \yH},
-\end{equation}
-with a universal, lower-triangular transformation matrix $M_{ml}(n,N)$ that
-depends only on $n$, $N$, and the condition of symmetry~\eqref{noeq:homog_math}.
+  \expe{\yxs^m \| \yH} = \sum_{l=1}^m M_{ml}(n,\yvN) \expe{\yXf^l \| \yH},
+\end{equation}
+with a universal, lower-triangular transformation matrix $M_{ml}(n,\yvN)$ that
+depends only on $n$, $\yvN$, and the condition of symmetry~\eqref{noeq:homog_math}.
 
 As intuition suggests, we have 
-%$\expe{\yxs^m \| \yH} \xrightarrow{n\to N} \expe{\yXf^m \| \yH}$.
+%$\expe{\yxs^m \| \yH} \xrightarrow{n\to \yvN} \expe{\yXf^m \| \yH}$.
 \begin{equation}
   \label{noeq:limit_moments}
-  \expe{\yxs^m \| \yH} \xrightarrow{n\to N} 
+  \expe{\yxs^m \| \yH} \xrightarrow{n\to \yvN} 
   \expe{\yXf^m \| \yH},
@@
 \end{equation}
-the latter because ${x_i}^m=x_i$, since states are $\set{0,1}$-valued.
+the latter because ${\yvx_i}^m=\yvx_i$, since states are $\set{0,1}$-valued.
 
@@
 \cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
-with parameters $N,N\yXf,n$, weighted by the probabilities
+with parameters $\yvN,\yvN\yXf,n$, weighted by the probabilities
 $\p(\yXf = A \|\yH)$ \citep[\cf][\sect~4, esp.\ \eqn~(22)]{kendall1967}.
@@
   \p(\yXf = A \| \yH) = Q(A) \propto
-\binom{N}{N A}\exp[
-\yl_2\, N A\,(N A-1)/2 + \yl_1\, N A].
+\binom{\yvN}{\yvN A}\exp[
+\yl_2\, \yvN A\,(\yvN A-1)/2 + \yl_1\, \yvN A].
 \end{equation}
@@
 with first and second moments as constraints and the reference distribution
-$Q_0$ defined by $Q_0(A) = 2^{-N}\binom{N}{N A}$, corresponding to a
+$Q_0$ defined by $Q_0(A) = 2^{-\yvN}\binom{\yvN}{\yvN A}$, corresponding to a
 uniform probability distribution for the network state $\yX$.
@@
 subnetwork-average distributions $\p(\yxs = a \|\yH)$, for the case in
-which $N = 1000$ units, $\lambda_1=-2.55$, $\lambda_2=0.005$, and
+which $\yvN = 1000$ units, $\lambda_1=-2.55$, $\lambda_2=0.005$, and
 $n=10, 50, 100, 250$. The distributions become broader as $n$ decreases,
@@
 expectations of $m$-tuple products of states
-$\expe{x_{i_1}\dotsm x_{i_m} \| \yH}$, proportional to the factorial
+$\expe{\yvx_{i_1}\dotsm \yvx_{i_m} \| \yH}$, proportional to the factorial
 moments, are not shown as they do not depend on $n$.
@@
 not true, since \eqn~\eqref{noeq:subpop_average}, as a linear mapping from
-$\RR^{N+1}$ to $\RR^{n+1}$, with $N$ larger than $n$, is onto but not into.
+$\RR^{\yvN+1}$ to $\RR^{n+1}$, with $\yvN$ larger than $n$, is onto but not into.
 Assigning a probability distribution $\p(\yxs = a \|\yH)$ to a
@@
     \label{noeq:constr_ex1}
-    \expe{\yxs \| \yHa} = c_1, \qquad \expe{\sav{x_i x_j}\|\yHa} = c_2;
+    \expe{\yxs \| \yHa} = c_1, \qquad \expe{\sav{\yvx_i \yvx_j}\|\yHa} = c_2;
   \end{equation}
@@
   \pf(\yX =\yR \|\yHa) =
-\varKappa\exp[\yL_2 N\yRf\,(N\yRf-1)/2 + \yL_1 N\yRf]
+\varKappa\exp[\yL_2 \yvN\yRf\,(\yvN\yRf-1)/2 + \yL_1 \yvN\yRf]
 \\
@@
 \\
-&\varKappa\sum_{N A=0}^N %m!\, 
-\binom{N-m}{N A-m}
-\exp[\yL_2 N A\,(N A-1)/2 + \yL_1 N A]
+&\varKappa\sum_{\yvN A=0}^\yvN %m!\, 
+\binom{\yvN-m}{\yvN A-m}
+\exp[\yL_2 \yvN A\,(\yvN A-1)/2 + \yL_1 \yvN A]
 = c_m, \quad m=1,2.
@@
   c_2 = \expe{\sav{\yx\yx}\|\yHa} = 
-\binom{N}{2}^{-1}
-\sum_{N A=0}^N %m!\, 
-\binom{N A}{2} \, \p(\yXf = A\|\yHa),
+\binom{\yvN}{2}^{-1}
+\sum_{\yvN A=0}^\yvN %m!\, 
+\binom{\yvN A}{2} \, \p(\yXf = A\|\yHa),
 \end{equation}
@@
 \binom{n}{n a}
-\binom{N-n}{N\yRf-n a}\Biggr]
+\binom{\yvN-n}{\yvN\yRf-n a}\Biggr]
 \\
@@
 \\
-&\sum_{N A=0}^N 
+&\sum_{\yvN A=0}^\yvN 
 %\binom{N}{N A}^{-1}
 \binom{n}{n a}
-\binom{N-n}{N A-n a}\,
+\binom{\yvN-n}{\yvN A-n a}\,
 \exp\Biggl[
@@
 \binom{n}{n a}
-\binom{N-n}{N A-n a}\Biggr]
+\binom{\yvN-n}{\yvN A-n a}\Biggr]
 =q( a)
@@
 the final distribution may have a multiplicity factor. When we consider the
-$2^{N}$ states of $N$ units we are not excluding that these might be
+$2^{\yvN}$ states of $\yvN$ units we are not excluding that these might be
 marginals of $2^{M}$ states of $M$ units. Each one has the same
-multiplicity $2^{M-N}$, but this constant multiplicity factor disappears
-by normalization. So the method applies just as in the case of $N$ units
+multiplicity $2^{M-\yvN}$, but this constant multiplicity factor disappears
+by normalization. So the method applies just as in the case of $\yvN$ units
 only, right?
@@
 involves probability considerations. Consider three cases with a full
-population of two units, $M=2$, of which we consider one unit, $N=1$.
+population of two units, $M=2$, of which we consider one unit, $\yvN=1$.
 \begin{itemize}
@@
 \begin{enumerate}
-\item have a population with $N$ units, $2^{N}$ possible states
+\item have a population with $\yvN$ units, $2^{\yvN}$ possible states
 \item expect averages of $\yx$ active neurons and $\yxxs$ active pairs
 \item use maximum-entropy to choose a probability distribution for the
-  states of the $N$ units conforming to our expectations.
+  states of the $\yvN$ units conforming to our expectations.
 \end{enumerate}
@@
 \item have a population with $M$ units, $2^{M}$ possible states
-\item expect that any subpopulation of $N$ units has $\yxs$ active
+\item expect that any subpopulation of $\yvN$ units has $\yxs$ active
   neurons and $\yxxs$ active pairs
@@
 \item marginalize to find the probability distribution for the states of
-  $N$ units.
+  $\yvN$ units.
 \end{enumerate}
@@
   average~\eqref{noeq:pop_average_Q}, consider that if the network total
-  is $N\yXf$, then $N\yXf$ out of $N$ units are active, and there are
-  $\tbinom{N}{N\yXf}$ possible states for which this can be true; therefore
-  \begin{equation}
-    \labelbis{eq:joint_plaus_N_homog}
-    \p(\yX = \yR \| H) = \binom{N}{N\yRf}^{-1}Q(\yRf).
+  is $\yvN\yXf$, then $\yvN\yXf$ out of $\yvN$ units are active, and there are
+  $\tbinom{\yvN}{\yvN\yXf}$ possible states for which this can be true; therefore
+  \begin{equation}
+    \labelbis{eq:joint_plaus_\yvN_homog}
+    \p(\yX = \yR \| H) = \binom{\yvN}{\yvN\yRf}^{-1}Q(\yRf).
   \end{equation}
@@
 the subnetwork average $\yxs$ conditional on the network average $\yXf$.
-There are $\tbinom{N}{N\yXf}$ possible network states if the network
-average is $\yXf$, \ie\ if $N\yXf$ units are in state $1$; the conditional
-probability of each is therefore $1/\tbinom{N}{N\yXf}$, owing to the
+There are $\tbinom{\yvN}{\yvN\yXf}$ possible network states if the network
+average is $\yXf$, \ie\ if $\yvN\yXf$ units are in state $1$; the conditional
+probability of each is therefore $1/\tbinom{\yvN}{\yvN\yXf}$, owing to the
 symmetry assumption~\eqref{noeq:homog_math}. Now consider the subnetwork of the
@@
 ones in state $1$ is the sum of the probabilities of all states for
-which $N\yXf-n\yxs$ of the remaining $N-n$ units are in state $1$; there are
-$\tbinom{N-n}{N\yXf-n\yxs}$ such states, all equally probable. Finally,
+which $\yvN\yXf-n\yxs$ of the remaining $\yvN-n$ units are in state $1$; there are
+$\tbinom{\yvN-n}{\yvN\yXf-n\yxs}$ such states, all equally probable. Finally,
 there are $\tbinom{n}{n\yxs}$ possible ways, all equally probable, in which
@@
       \p(\yx=\yr \|  \yX =\yR, \yH) \,
-      \binom{N}{N A}^{-1},
+      \binom{\yvN}{\yvN A}^{-1},
       \\
       &=\sum_{\yr}^{\yrs=a} 
-      \binom{N-n}{N A-n a}\binom{N}{N A}^{-1},
+      \binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1},
       \\
       &=
-      \binom{n}{n a}\binom{N-n}{N A-n a}\binom{N}{N A}^{-1},
+      \binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1},
 \end{split}
@@
   which describes the probability of, say, drawing a proportion of $\yxs$
-  blue balls in $n$ drawings without replacement from an urn with $N$
+  blue balls in $n$ drawings without replacement from an urn with $\yvN$
   balls, a fraction $\yXf$ of which are blue.
@@
       \p(\yxs =a \| \yH)
-      &= \sum_{N A=0}^N\p(\yxs = a \| \yXf = A, \yH)\,
+      &= \sum_{\yvN A=0}^\yvN\p(\yxs = a \| \yXf = A, \yH)\,
       \p(\yXf = A \| \yH),
       \\
-      &=\sum_{N A=0}^N
-      \binom{n}{n a}\binom{N-n}{N A-n a}\binom{N}{N A}^{-1}
+      &=\sum_{\yvN A=0}^\yvN
+      \binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1}
       Q(A).
@@
     \label{noeq:margin_products}
-    \sum_{\yx} x_1\dotsm x_m \p(\yx =\yr \| \yH)
-      =\sum_{\yr,\yR}  r_1\dotsm r_m \p(\yx=\yr \| \yX=\yR,\yH)
+    \sum_{\yx} \yvx_1\dotsm \yvx_m \p(\yx =\yr \| \yH)
+      =\sum_{\yr,\yR}  \yvr_1\dotsm \yvr_m \p(\yx=\yr \| \yX=\yR,\yH)
       \, \p(\yX=\yR \| \yH),
@@
       \begin{aligned}
-      &=\sum_{\yr,\yR} r_1\dotsm r_m\,\delt(R_1-r_1)\dotsm\delt(R_m-R_m)
+      &=\sum_{\yr,\yR} \yvr_1\dotsm \yvr_m\,\delt(\yvR_1-\yvr_1)\dotsm\delt(\yvR_m-\yvR_m)
       \p(\yX = \yR \| \yH),
       \\
-      &=\sum_{\yR} R_1\dotsm R_m \p(\yX =\yR \| \yH).
+      &=\sum_{\yR} \yvR_1\dotsm \yvR_m \p(\yX =\yR \| \yH).
     \end{aligned}
@@
   \begin{equation*}
-    x_1 x_2 + x_1 x_3 + \dotsb + x_{n-1} x_n.
+    \yvx_1 \yvx_2 + \yvx_1 \yvx_3 + \dotsb + \yvx_{n-1} \yvx_n.
   \end{equation*}
@@
   \begin{equation}
-    \sav{x_{i_1} \dotsm x_{i_m}} = \binom{n\yxs}{m} \binom{n}{m}^{-1}.
+    \sav{\yvx_{i_1} \dotsm \yvx_{i_m}} = \binom{n\yxs}{m} \binom{n}{m}^{-1}.
   \end{equation}
@@
     \label{noeq:expect_product_m}
-    \expe*{\sav{x_{i_1} \dotsm x_{i_m}} \| \yH}
+    \expe*{\sav{\yvx_{i_1} \dotsm \yvx_{i_m}} \| \yH}
     \\
@@
       \\
-      &=\frac{(n-m)!}{n!} \sum_{N A=0}^N
+      &=\frac{(n-m)!}{n!} \sum_{\yvN A=0}^\yvN
       Q( A)
@@
       \sum_{n a=0}^n 
-      m!\,\binom{n a}{m}\, \binom{n}{n a}\binom{N-n}{N A-n a}\binom{N}{N A}^{-1} 
+      m!\,\binom{n a}{m}\, \binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1} 
       \Biggr].
@@
   \begin{equation}
-    \sum_{n a=0}^n m!\,\binom{n a}{m} \binom{n}{n a}\binom{N-n}{N\yXf-n a}
-    \binom{N}{N\yXf}^{-1} 
+    \sum_{n a=0}^n m!\,\binom{n a}{m} \binom{n}{n a}\binom{\yvN-n}{\yvN\yXf-n a}
+    \binom{\yvN}{\yvN\yXf}^{-1} 
     =
-    m!\,\binom{n}{m} \binom{n a}{m}\binom{N}{m},
+    m!\,\binom{n}{m} \binom{n a}{m}\binom{\yvN}{m},
   \end{equation}
@@
     \label{noeq:binom_identity}
-    \binom{N}{M}\binom{M}{m}=\binom{N}{m}\binom{N-m}{M-m},
+    \binom{\yvN}{M}\binom{M}{m}=\binom{\yvN}{m}\binom{\yvN-m}{M-m},
   \end{equation}
@@
     \expe{\yXf^m \| \yH} \defd
-    \sum_{N A=0}^{N} A^m \p(\yXf = A  \| H),
+    \sum_{\yvN A=0}^{\yvN} A^m \p(\yXf = A  \| H),
   \end{equation}
@@
   \p(\yXf = \yRf \| \yHa)  = \yL\,
-\binom{N}{N \yRf}\exp\biggl[
-\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr]
+\binom{\yvN}{\yvN \yRf}\exp\biggl[
+\yL_2 \yRf \,\frac{\yvN \yRf-1}{\yvN-1} + \yL_1\, \yRf\biggr]
 \end{equation}
@@
   \p(\yxs = \yrs \| \yHa)  = \yL
-  \sum_{N\yRf=0}^N
-  \binom{n}{n\yrs} \binom{N-n}{N\yRf-n\yrs} \exp\biggl[
-\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr].
+  \sum_{\yvN\yRf=0}^\yvN
+  \binom{n}{n\yrs} \binom{\yvN-n}{\yvN\yRf-n\yrs} \exp\biggl[
+\yL_2 \yRf \,\frac{\yvN \yRf-1}{\yvN-1} + \yL_1\, \yRf\biggr].
 \end{equation}
@@
 $\p(\yxs = \yrs \| \yHb)$ are different.
-Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for $N=5000$,
+Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for $\yvN=5000$,
 $n=100$, $c_1=0.45$, $c_2=0.35$. The two distributions have slightly
@@
 more peaked. Another example is shown in
-\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for $N=5000$, $n=200$,
+\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for $\yvN=5000$, $n=200$,
 and neurobiologically more realistic constraints $c_1=0.045$, $c_2=0.0025$

### 2018-05-15 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-15T22:50:14+0200
+%% Last-Updated: 2018-05-15T22:52:43+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 sample of the activity of the neuronal population under the recording
-device; such population can amount to tens of thousands of neurons
+device; which can amount to tens of thousands of neurons
 \cite{abeles1991}. We could even consider the recorded neurons as a sample

### 2018-05-16 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-15T22:52:43+0200
+%% Last-Updated: 2018-05-16T11:37:42+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \DeclareMathSymbol{\theta}{\mathalpha}{egreeki}{"12}
-% \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
+ \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
  \DeclareMathSymbol{\kappa}{\mathalpha}{egreeki}{"14}
@@
 %\newcommand*{\vs}{{vs.}}
-%\newcommand*{\etc}{{etc.}}
+\newcommand*{\etc}{{etc.}}
 \newcommand*{\ie}{{i.e.}}
@@
 
+\newcommand*{\tav}{\widehat} %time average
 \newcommand*{\av}{\widebar} %pop average
@@
 
-\newcommand*{\yvX}{\varSigma}
-\newcommand*{\yvR}{\sigma}
-\newcommand*{\yvx}{S}
-\newcommand*{\yvr}{s}
-\newcommand*{\yvN}{\nu}
-
-\newcommand*{\yx}{\bm{\yvx}}%subpop state
+\newcommand*{\yXv}{\varSigma}
+\newcommand*{\yRv}{\sigma}
+\newcommand*{\yxv}{S}
+\newcommand*{\yrv}{s}
+\newcommand*{\yNv}{\nu}
+
+\newcommand*{\yx}{\bm{\yxv}}%subpop state
 \newcommand*{\yxs}{\sav{\yx}}%subpop av state
-\newcommand*{\yX}{\bm{\yvX}}%pop state
+\newcommand*{\yX}{\bm{\yXv}}%pop state
 \newcommand*{\yXf}{\av{\yX}}%pop av state
@@
 \newcommand*{\yXXf}{\av{\yX\yX}}%subpop av state
-\newcommand*{\yr}{\bm{\yvr}}%subpop value
+\newcommand*{\yr}{\bm{\yrv}}%subpop value
 \newcommand*{\yrs}{\sav{\yr}}%subpop av value
-\newcommand*{\yR}{\bm{\yvR}}%pop value
+\newcommand*{\yR}{\bm{\yRv}}%pop value
 \newcommand*{\yRf}{\av{\yR}}%pop av value
@@
 \begin{enumerate}
-\item The activity of each neuron, a continuous signal, is binned in time
-  and binarized in intensity, and thus transformed into a sequence of
-  digits \enquote{$0$}s (inactive) and \enquote{$1$}s
+\item The activity of each neuron, a continuous signal, is divided into $T$
+  time bins and binarized in intensity, and thus transformed into a
+  sequence of digits \enquote{$0$}s (inactive) and \enquote{$1$}s
   \cite[\cf][]{caianiello1961,caianiello1986}.
 
-  Let the variable $\yvx_i(t)\in\set{0,1}$ denote the activity of the $i$th
-  sampled neuron at time bin $t$. The average activity at that bin is
-  therefore $\yxs(t) \defd \sum_i \yvx_i(t)/n$. If we count the number of
-  distinct pairs of active neurons at that bin we combinatorially find
-  $\tbinom{n\yxs(t)}{2}\equiv n\yxs(t)\,[n\yxs(t)-1]/2$. There can be at
-  most $\tbinom{n}{2}$ simultaneously active pairs, so the average pair
-  activity is
-  $\av{\yx \yx}(t) \defd \tbinom{n}{2}^{-1}\tbinom{n\yxs(t)}{2}$. With some
-  combinatorics we see that the average activity of $m$-tuples of neurons
-  is
+  Let the variable $\yrv_i(t)\in\set{0,1}$ denote the activity of the $i$th
+  sampled neuron at time bin $t$. Collectively denote the $n$ activities
+  with $\yr(t) \defd \bigl(\yrv_1(t),\dotsc,\yrv_n(t)\bigr)$. The
+  population-averaged activity at that bin is
+  $\yrs(t) \defd \sum_i \yrv_i(t)/n$. If we count the number of distinct
+  pairs of active neurons at that bin we combinatorially find
+  $\tbinom{n\yrs(t)}{2}\equiv n\yrs(t)\,[n\yrs(t)-1]/2$. There can be at
+  most $\tbinom{n}{2}$ simultaneously active pairs, so the
+  population-averaged pair activity is
+  $\av{\yr \yr}(t) \defd \tbinom{n}{2}^{-1}\tbinom{n\yrs(t)}{2}$. With some
+  combinatorics we see that the population-averaged activity of $m$-tuples
+  of neurons is
   \begin{equation}
     \label{eq:products_intermsof_average}
-    \av{\underbrace{\yx\dotsm\yx}_{\text{$m$ terms}}}
-    = \binom{n}{m}^{-1}\binom{n\yxs(t)}{m}.
-  \end{equation}
+    \av{\underbrace{\yr\dotsm\yr}_{\text{$m$ terms}}}(t)
+    = \binom{n}{m}^{-1}\binom{n\yrs(t)}{m}.
+  \end{equation}
+
   For brevity let us agree to simply call \enquote{activity} the average
-  $\yxs$, \enquote{pair-activity} the average $\av{\yx \yx}$, and so on.
+  $\yrs$, \enquote{pair-activity} the average $\av{\yr \yr}$, and so on.
 
 \item Construct a sequence of relative-\me\ distributions for the activity
-  $\yxs$, using this sequence of constraints:
+  $\yrs$, using this sequence of constraints:
   \begin{itemize}
-  \item the time average of the activities $\yxs(t)$;
-  \item the time averages of the activities $\yxs(t)$ and pair-activities
-    $\av{\yx \yx}(t)$;
-  \item \ldots and so on, including the time averages of $\yxs(t)$ up to
-    $\av{\yx\dotsm\yx}(t)$ with $k$ terms.
+  \item the time average of the activity: $\tav{\yrs} \defd \sum_t\yrs(t)/T$;
+  \item the time averages of the activity and of the pair-activity
+    $\tav{\av{\yr \yr}} \defd \sum_t\av{\yr \yr}(t)/T$;
+  \item \ldots
+  \item the time averages of the activity, of the pair-activity, \etc\ up
+    to the $k$-activity.
   \end{itemize}
-  Call the resulting distributions $p_1(\yxs), p_2(\yxs),\dotsc,p_k(\yxs)$.
+  Call the resulting distributions $p_1(\yrs), p_2(\yrs),\dotsc,p_k(\yrs)$.
   The time-bin dependence is now absent because these distributions can be
   interpreted as referring to any one of the time bins $t$, or to a new
-  time bin containing new data.
+  time bin (in the future or in the past) containing new data.
 
   We also have the empirical frequency distribution of the total activity,
-  $f(\yxs)$, counted from the time bins.
+  $f(\yrs)$, counted from the time bins.
 
@@
 complex versions of the method, possibly except version that use
-\enquote{hidden} neurons. The problem stems from the fact that the recorded
-neurons are a \emph{sample} from a larger, unrecorded population.
+\enquote{hidden} neurons. The problem is that the recorded neurons are a
+\emph{sample} from a larger, unrecorded population, but the \me\ method as
+applied above is treating them as isolated from the rest of the brain.
+Hence, the results it provides cannot rightfully be extrapolated. We will
+give a mathematical proof of this. Let us first analyse this issue in more
+detail.
 
@@
 
-Here is the problem. We can apply the \me\ method to the sample, as
-described in the above protocol, to generate probability distributions for
-the activity of the sample. But, if our sample is representative of a
-larger population, we can apply the \me\ method to the larger (unrecorded)
-population, using the constraints obtained from the recorded sample, and
-generate a probability distribution for the larger population. The
-distribution for the sample is then obtained by marginalization. Either
-application seems to have some commendable features. However, \emph{the
-  distributions obtained from these two applications differ}. Which choice
-is most meaningful?
-
-In this paper we develop the second way of applying the \me, at the level
-of the larger population, and show that its results differ from the
+We can apply the \me\ method to the sample, as described in the above
+protocol, to generate probability distributions for the activity of the
+sample. But, given that our sample is representative of a larger
+population, we can also apply the \me\ method to the larger (unrecorded)
+population. The constraints are the same: the time averages of the sampled
+data, since they constitute representative data about the larger population
+as well. The method thus yields a probability distribution for the larger
+population, and the distribution for the sample is obtained by
+marginalization. The problem is that \emph{the distributions obtained from
+  these two applications differ}. Which choice is most meaningful?
+
+In this paper we develop the second way of applying the \me\ method, at the
+level of the larger population, and show that its results differ from the
 application at the sample level. We also consider the case where the size
@@
 necessary to use probability relations relevant to sampling
-\cites{ghoshetal1997}[parts~I, VI]{freedmanetal1978_r2007}[\chap~8]{gelmanetal1995_r2014}[\chap~3]{jaynes1994_r2003}.
+\cites{ghoshetal1997}[parts~I,
+VI]{freedmanetal1978_r2007}[\chap~8]{gelmanetal1995_r2014}[\chap~3]{jaynes1994_r2003}.
 The relations we present are well-known in survey sampling and in the
 pedagogic problem of drawing from an urn without replacement, yet they are
-somewhat hard to find explicitly written in the neuroscientific literature,
-so they may be of interest on their own.
+somewhat hard to find explicitly written in the neuroscientific literature.
+We present and discuss them in the next section. A minor purpose of this
+paper is to make these relations more widely known, because they can be
+useful independently of \me\ methods.
 
@@
 
-The notation and terminology in this note follows ISO and ANSI standards
+The notation and terminology in this note follow ISO and ANSI standards
 \citep{iso1993,ieee1993,nist1995,iso2006,iso2006b} but for the use of the
@@
 
-\section{Notation: population, sample, probabilities}
-\label{sec:setup}
-
-We have a population of $\yvN$ binary neurons. We assume that they can be
+\section{Probability relations between population and  sample}
+\label{sec:prob_samples}
+
+We have already introduced the notation for the sample neurons. We
+introduce an analogous notation for the $\yNv$ neurons constituting the
+larger population, but using the corresponding Greek letters:
+$\yRv_{\iota}(t)$ is the activity of the $\iota$th neuron at time bin $t$,
+$\yRf(t) \defd \sum_{\iota} \yRv_{\iota}(t)/\yNv$ is the activity at that
+bin averaged over the larger population, \etc.
+
+The probability relations between sample and larger population are valid at
+every time bin. As we mentioned above, the \me\ distribution refers to any
+time bin or to a new bin. For these reasons we will now omit the time-bin
+argument \enquote{(t)} from our expressions.
+
+Probabilities refer to statements about the quantities we observe. We use
+the standard notation:
+\begin{equation}
+  \label{eq:notation_statements}
+  \begin{aligned}
+    &\text{\enquote{$\yxv_i = \yrv_i$} 
+      stands for \enquote{the activity of the $i$th sample neuron is $\yrv_i$}},
+    \\
+    &\text{\enquote{$\yxs = \yrs$} 
+      stands for \enquote{the (population-averaged) activity of the sample neurons is $\yrv_i$}},
+  \end{aligned}
+\end{equation}
+and similarly for other quantities and for the activities of the neurons of
+the larger population.
+
+\iffalse We have a population of $\yNv$ binary neurons. We assume that they can be
 distinguished, by their spike shapes for example; but other details, like
 their locations, are unknown. The neurons have a joint state
-$(\yvX_1,\dotsc,\yvX_\yvN) \defs \yX$ having fixed but unknown binary values
-$(\yvR_1,\dotsc,\yvR_\yvN)\defs \yR \in \set{0,1}^\yvN$; the \enquote{$\yvX$} denote the
-variables, and \enquote{$\yvR$} their values. A particular sample of $n$ neurons
-from this population has joint state $(\yvx_1, \dotsc, \yvx_n) \defs \yx$ having
-fixed binary values $(\yvr_1, \dotsc, \yvr_n)\defs\yr \in \set{0,1}^n$. We will
+$(\yXv_1,\dotsc,\yXv_\yNv) \defs \yX$ having fixed but unknown binary values
+$(\yRv_1,\dotsc,\yRv_\yNv)\defs \yR \in \set{0,1}^\yNv$; the \enquote{$\yXv$} denote the
+variables, and \enquote{$\yRv$} their values. A particular sample of $n$ neurons
+from this population has joint state $(\yxv_1, \dotsc, \yxv_n) \defs \yx$ having
+fixed binary values $(\yrv_1, \dotsc, \yrv_n)\defs\yr \in \set{0,1}^n$. We will
 consider various averages of the population and the sample. For this
@@
   \begin{gathered}
-    \av{\yX} \defd \tfrac{1}{\yvN} (\yvX_1 + \yvX_2 + \dotsb + \yvX_\yvN),
+    \av{\yX} \defd \tfrac{1}{\yNv} (\yXv_1 + \yXv_2 + \dotsb + \yXv_\yNv),
     % & \sav{\yx} &\defd \tfrac{1}{n} (x_1 + x_2 + \dotsb + x_n),
    \qquad% \\[2\jot]
-    \av{\yX \yX} \defd \tbinom{\yvN}{2}^{-1} (\yvX_1 \yvX_2 + \yvX_1 \yvX_3 + \dotsb +
-    \yvX_{\yvN-1} \yvX_\yvN),
+    \av{\yX \yX} \defd \tbinom{\yNv}{2}^{-1} (\yXv_1 \yXv_2 + \yXv_1 \yXv_3 + \dotsb +
+    \yXv_{\yNv-1} \yXv_\yNv),
     % & \sav{\yx \yx} &\defd\tbinom{n}{2}^{-1} (x_1 x_2 + x_1 x_3 + \dotsb
@@
     \\[2\jot]
-    \av{\yX\yX\yX} \defd \tbinom{\yvN}{3}^{-1} (\yvX_1 \yvX_2 \yvX_3 +% \yvX_1 \yvX_2 \yvX_4
-    \dotsb + \yvX_{\yvN-2} \yvX_{\yvN-1}\yvX_\yvN),
+    \av{\yX\yX\yX} \defd \tbinom{\yNv}{3}^{-1} (\yXv_1 \yXv_2 \yXv_3 +% \yXv_1 \yXv_2 \yXv_4
+    \dotsb + \yXv_{\yNv-2} \yXv_{\yNv-1}\yXv_\yNv),
     \qquad\text{and so on.}
@@
 These formulae say that $\yXf$ is the fraction of active neurons, $\yXXf$
-the fraction of simultaneously active pairs out of all $\binom{\yvN}{2}$
+the fraction of simultaneously active pairs out of all $\binom{\yNv}{2}$
 pairs, $\av{\yX\yX\yX}$ the fraction of simultaneously active triplets, and
-so on. Products of states like $\yvX_i \dotsm \yvX_j$ also have values in
+so on. Products of states like $\yXv_i \dotsm \yXv_j$ also have values in
 $\set{0,1}$; from this we can combinatorially  prove that
@@
   \av{\underbrace{\yX\dotsm\yX}_{\text{$m$ factors}}}
-  = \binom{\yvN}{m}^{-1}\binom{\yvN\yXf}{m}.
+  = \binom{\yNv}{m}^{-1}\binom{\yNv\yXf}{m}.
 \end{equation}
@@
 
-\iffalse
 \begin{simplenote}
@@
   probability for the population, the probabilities for the states of these
-  two samples are different: the former is obtained from marginalization, the
-  latter by symmetrization and marginalization.
+  two samples are different: the former is obtained from marginalization,
+  the latter by symmetrization and marginalization.
 \end{simplenote}
@@
 
-Our uncertainty about the actual state of the population is completely
-expressed by the joint probability distribution
+If $\yHc$ denotes our state of knowledge, \ie\ the evidence and assumptions
+backing our probability assignments, our uncertainty about the full
+activity of the larger population is expressed by the joint probability
+distribution
 \begin{equation}
   \label{eq:joint_plaus}
-  \p(\yvX_1=\yvR_1, \yvX_2=\yvR_2, \dotsc, \yvX_\yvN=\yvR_\yvN \| \yHc) \quad\text{or}\quad
-\p(\yX =\yR \| \yHc), \quad \yR \in \set{0,1}^\yvN,
-\end{equation}
-where $\yHc$ denotes our state of knowledge, \ie\ the evidence and
-assumptions backing this particular probability assignment. Our uncertainty
-about the state of the sample is likewise expressed by
+  \p(\yXv_1=\yRv_1, \yXv_2=\yRv_2, \dotsc, \yXv_\yNv=\yRv_\yNv \| \yHc)
+  \quad\text{or}\quad
+\p(\yX =\yR \| \yHc), \quad \yR \in \set{0,1}^\yNv.
+\end{equation}
+Our uncertainty about the state of the sample is likewise expressed by
 \begin{equation}
   \label{eq:sample_plaus}
-  \p(\yvx_1=\yvr_1, \yvx_2=\yvr_2, \dotsc, \yvx_n=\yvr_n \| \yHc) \quad\text{or}\quad
+  \p(\yxv_1=\yrv_1, \yxv_2=\yrv_2, \dotsc, \yxv_n=\yrv_n \| \yHc) \quad\text{or}\quad
 \p(\yx =\yr \| \yHc), \quad \yr \in \set{0,1}^n.
 \end{equation}
-
-\section{Probability: relations between population and  sample}
-\label{sec:prob_samples}
 
@@
   \p(\yXf=\yRf \| \yH) =
-  \binom{\yvN}{\yvN\yRf}^{-1} \p(\yXf=\yRf \| \yH),
+  \binom{\yNv}{\yNv\yRf}^{-1} \p(\yXf=\yRf \| \yH),
 \end{equation}
@@
 \cite{heathetal1976}. This theorem is intuitive: owing to symmetry, we must
-assign equal probabilities to all states with $\yvN\yRf$ active neurons.
+assign equal probabilities to all states with $\yNv\yRf$ active neurons.
 
@@
   \label{eq:subpop_average}
-  \p(\yxs=\yrs \| \yH) = \sum_{\yvN \yRf =0}^{\yvN}
+  \p(\yxs=\yrs \| \yH) = \sum_{\yNv \yRf =0}^{\yNv}
     \p(\yxs = \yrs \|\yXf=\yRf, \yH)\,
@@
   \p(\yxs = \yrs \|\yXf=\yRf, \yH)=
-  \binom{n}{n\yrs}\binom{\yvN-n}{\yvN \yRf-n\yrs}\binom{\yvN}{\yvN \yRf}^{-1}
+  \binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}
   \defs \ypp(\yrs\|\yRf).
@@
 $\p(\yxs = \yrs \|\yXf=\yRf, \yH) \equiv \ypp(\yrs,\yRf)$ relates the
-spaces of the sample average $\yXf\in\set{0,\dotsc,\yvN}$ and of the
+spaces of the sample average $\yXf\in\set{0,\dotsc,\yNv}$ and of the
 population average $\yxs\in\set{0,\dotsc,n}$. It is a coarsening projector
@@
   \equiv
-  \sum_{\yvN\yRf=0}^\yvN f^*(\yRf)\,\p(\yXf=\yRf \| \yH)
+  \sum_{\yNv\yRf=0}^\yNv f^*(\yRf)\,\p(\yXf=\yRf \| \yH)
 %  =
@@
   $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
-\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yvN-n}{\yvN \yRf-n\yrs}\binom{\yvN}{\yvN \yRf}^{-1}$ for $\yvN=5000$, $n=200$. (Band artifacts may appear in the
+\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
   colourbar depending on your \textsc{pdf} viewer.)}
@@
 \fig~\ref{fig:hypergeom_proj}, reveals that it is a sort of \enquote{fuzzy
-  identity matrix} between the $\yXf$-space $\set{0,\dotsc,\yvN}$ and
-$\yxs$-space $\set{0,\dotsc,n}$. When $n=\yvN$ it is the identity matrix. We
+  identity matrix} between the $\yXf$-space $\set{0,\dotsc,\yNv}$ and
+$\yxs$-space $\set{0,\dotsc,n}$. When $n=\yNv$ it is the identity matrix. We
 thus have that
@@
 $\p(\yxs=a) \to \p(\yXf=a)$ and $\expeb{f(\yxs)} \to \expeb{f(\yXf)}$, as
-$n\to \yvN$, do: these limits are trivially, universally valid because the
-sample becomes the full population as $n\to \yvN$. In particular, these limits
+$n\to \yNv$, do: these limits are trivially, universally valid because the
+sample becomes the full population as $n\to \yNv$. In particular, these limits
 hold even in cases where the conditional probability
@@
 %   \qquad\text{or}\qquad
-%   \sum_{n\yrs=0}^n f(\yrs)\,p_*(\yxs=\yrs) = \sum_{\yvN\yRf=0}^\yvN f^*(\yRf)\,p(\yXf=\yRf).
+%   \sum_{n\yrs=0}^n f(\yrs)\,p_*(\yxs=\yrs) = \sum_{\yNv\yRf=0}^\yNv f^*(\yRf)\,p(\yXf=\yRf).
 % \end{equation}
@@
     % \binom{N-m}{N \yRf-m} \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
-    = \binom{\yvN}{m}^{-1}
+    = \binom{\yNv}{m}^{-1}
     % \frac{(N-m)!}{N!}
-    \sum_{\mathclap{\yvN \yRf=0}}^\yvN %m!\,
-    \binom{\yvN \yRf}{m}\, \p(\yXf=\yRf \| \yH)
+    \sum_{\mathclap{\yNv \yRf=0}}^\yNv %m!\,
+    \binom{\yNv \yRf}{m}\, \p(\yXf=\yRf \| \yH)
 \equiv    \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}.
@@
 not completely specified, because \eqn~\eqref{eq:subpop_average} linearly
-constrains $\yvN+1$ unknowns by only $n+1$ equations.
+constrains $\yNv+1$ unknowns by only $n+1$ equations.
 
@@
   \p(\yXf = \yRf \| \yHa)  = \yK\,
-  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[\yL\T
+  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
   \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
@@
   \p(\yxs = \yrs \| \yHa)  = \yK\, 
-  \sum_{\yvN\yRf=0}^\yvN \ypp(\yrs \|\yRf)
+  \sum_{\yNv\yRf=0}^\yNv \ypp(\yrs \|\yRf)
 \,
-  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[\yL\T
+  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
   \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
@@
   \label{eq:app_maxent_pop_constraints}
-\yc = \yK\sum_{n\yrs=0}^n \sum_{\yvN\yRf=0}^\yvN \yf(\yrs) \ypp(\yrs \|\yRf)
+\yc = \yK\sum_{n\yrs=0}^n \sum_{\yNv\yRf=0}^\yNv \yf(\yrs) \ypp(\yrs \|\yRf)
 \,
-  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[\yL\T
+  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
   \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
@@
   \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yvN=5000$,
+  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yNv=5000$,
   $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$.}
@@
   \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yvN=5000$,
+  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yNv=5000$,
   $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$,
@@
 \figs~\ref{fig:diff_maxent_pop_sample}
-and~\ref{fig:diff_maxent_pop_sample_realdata}, for $\yvN=5000$, $n=200$, and
+and~\ref{fig:diff_maxent_pop_sample_realdata}, for $\yNv=5000$, $n=200$, and
 constraint functions of the form
@@
 Suppose we do not know whether the sample is really part of a larger
-population: we do not know whether $\yvN=n$ or how large $\yvN$ is otherwise.
+population: we do not know whether $\yNv=n$ or how large $\yNv$ is otherwise.
 Call this state of ignorance $\yHd$. In the probability calculus this
-ignorance about $\yvN$ is expressed by assigning a probability distribution
-$\p(\yvN \| \yHd)$ that vanishes if $\yvN<n$, since we know that $\yvN\ge n$; see
+ignorance about $\yNv$ is expressed by assigning a probability distribution
+$\p(\yNv \| \yHd)$ that vanishes if $\yNv<n$, since we know that $\yNv\ge n$; see
 Good \cite{good1965,good1967b} and Rissanen \cite{rissanen1983} for
@@
 assumption of symmetric ignorance, probability assignments that do not
-assume a specific value of $\yvN$ are then obtained via multiplication of all
-$\yvN$-dependent probabilities by $\p(\yvN \| \yHd)$ and subsequent
-marginalization over $\yvN$. Technically speaking, $\yvN$ becomes a
+assume a specific value of $\yNv$ are then obtained via multiplication of all
+$\yNv$-dependent probabilities by $\p(\yNv \| \yHd)$ and subsequent
+marginalization over $\yNv$. Technically speaking, $\yNv$ becomes a
 \emph{nuisance parameter}
@@
   \label{eq:app_maxent_pop_marg_unknown_N}
-  \p(\yxs = \yrs \| \yHd)  = \sum_\yvN \biggl\{\yK_\yvN\, 
-  \sum_{\yvN\yRf=0}^\yvN \ypp_\yvN(\yrs \|\yRf)
+  \p(\yxs = \yrs \| \yHd)  = \sum_\yNv \biggl\{\yK_\yNv\, 
+  \sum_{\yNv\yRf=0}^\yNv \ypp_\yNv(\yrs \|\yRf)
 \,
-  \binom{\yvN}{\yvN \yRf}\,\exp\Bigl[{\yL_\yvN}\T
-  \sum_{n\yrs=0}^n \yf(\yrs)\ypp_\yvN(\yrs\|\yRf)\Bigr]\biggr\}
-  \,\p(\yvN \| \yHd),
-\end{equation}
-where $\yvN$-dependencies have been made explicit. This is a formidable
+  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[{\yL_\yNv}\T
+  \sum_{n\yrs=0}^n \yf(\yrs)\ypp_\yNv(\yrs\|\yRf)\Bigr]\biggr\}
+  \,\p(\yNv \| \yHd),
+\end{equation}
+where $\yNv$-dependencies have been made explicit. This is a formidable
 expression. But our question, \enquote{is the usual \me\ at the sample
@@
 distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
-\eqref{eq:app_maxent_sample} equal for some choice of $\p(\yvN \| \yHd)$, with
-$\p(\yvN \| \yHd)\ne 0$ for $\yvN>n$?}. We leave this mathematical problem for
+\eqref{eq:app_maxent_sample} equal for some choice of $\p(\yNv \| \yHd)$, with
+$\p(\yNv \| \yHd)\ne 0$ for $\yNv>n$?}. We leave this mathematical problem for
 future work. Note, however, that this equality is satisfied if
-$\p(\yvN=1 \| \yHd)=1$, which means that the usual \me\ model can also be
+$\p(\yNv=1 \| \yHd)=1$, which means that the usual \me\ model can also be
 interpreted as assuming that \emph{no} larger population exists.
@@
 surprise analysis, may reveal features in a dataset that were unnoticed by
-the standard \me\ model. The question remains of how to specify $\yvN$,
-though. We have tacitly intended $\yvN$ as the size of the largest
+the standard \me\ model. The question remains of how to specify $\yNv$,
+though. We have tacitly intended $\yNv$ as the size of the largest
 biologically or functionally homogeneous population from which our sample
 was recorded. It could be the amount of neurons in a functional brain area,
-for example the primary visual cortex, for which $\yvN\sim 10^8$
-\cite{leubaetal1994}. For large $\yvN$ -- unfortunately we are not yet able to
+for example the primary visual cortex, for which $\yNv\sim 10^8$
+\cite{leubaetal1994}. For large $\yNv$ -- unfortunately we are not yet able to
 translate this \enquote{large} into a numeric order of magnitude -- the
-final distribution becomes independent of $\yvN$, and continuous
+final distribution becomes independent of $\yNv$, and continuous
 approximations become available.
@@
   the same in either case: $\expe{\yxs \|\yH}=0.045$,
-  $\expe{\yxxs \|\yH}=0.0025$, and $\yvN=5000$, $n=200$.}
+  $\expe{\yxxs \|\yH}=0.0025$, and $\yNv=5000$, $n=200$.}
 \label{fig:diff_maxent_pop_sample_realdata}
@@
 expectations for individual neurons or groups of neurons are constrained,
-for example $\expe{\yvx_2}$ or $\expe{\yvx_1x_8x_9}$. A short computation shows
+for example $\expe{\yxv_2}$ or $\expe{\yxv_1x_8x_9}$. A short computation shows
 that the \me\ method with this kind of constraints gives the same result
@@
 
-% \mynote{Discussion of case when individual $\yvX_i\dotsm \yvX_j$ are constrained.
+% \mynote{Discussion of case when individual $\yXv_i\dotsm \yXv_j$ are constrained.
 % Likely breakdown of \me\ in this case \cite{portamana2009}. Hints at full
@@
 \begin{align}
-  \expe{\yvx_{i_1} \dotsm \yvx_{i_m} \| \yH}
+  \expe{\yxv_{i_1} \dotsm \yxv_{i_m} \| \yH}
 &=
@@
   \begin{split}
-\expe{\yvx_{i_1} \dotsm \yvx_{i_m} \| \yH}
+\expe{\yxv_{i_1} \dotsm \yxv_{i_m} \| \yH}
 &=
-\binom{\yvN}{m}^{-1}
+\binom{\yNv}{m}^{-1}
 %\frac{(N-m)!}{N!}
-\sum_{\yvN A=0}^\yvN %m!\, 
-\binom{\yvN A}{m}\, Q(A)
+\sum_{\yNv A=0}^\yNv %m!\, 
+\binom{\yNv A}{m}\, Q(A)
 \\
@@
 %\frac{(N-m)!}{N!}
-\sum_{\yvN A=0}^\yvN %m!\, 
-\binom{\yvN-m}{\yvN A-m}
-\binom{\yvN}{\yvN A}^{-1} Q(A).
+\sum_{\yNv A=0}^\yNv %m!\, 
+\binom{\yNv-m}{\yNv A-m}
+\binom{\yNv}{\yNv A}^{-1} Q(A).
 \end{split}
@@
 neurons \enquote{units} to lend some generality to our discussion. We shall
-make statements about the whole population of $\yvN$ units and about a
+make statements about the whole population of $\yNv$ units and about a
 subpopulation of $n$ units; the word \enquote{population} will always refer
 to the \emph{whole} population. The subpopulation states and their values
-are denoted by lowercase letters: $(\yvx_1, \dotsc, \yvx_n)\equiv\yx$ and
-$(\yvr_1, \dotsc, \yvr_n)\equiv\yr$; but note that $\yvx_i \equiv \yvX_{j_i}$ and
-$\yvr_i \equiv \yvR_{j_i}$ for some distinct $j_1,\dotsc,j_n$. We shall also make
+are denoted by lowercase letters: $(\yxv_1, \dotsc, \yxv_n)\equiv\yx$ and
+$(\yrv_1, \dotsc, \yrv_n)\equiv\yr$; but note that $\yxv_i \equiv \yXv_{j_i}$ and
+$\yrv_i \equiv \yRv_{j_i}$ for some distinct $j_1,\dotsc,j_n$. We shall also make
 statements about the population-averaged state, or \emph{population
@@
   \label{noeq:population-av}
-   \yXf &\defd (\yvX_1 + \dotsb + \yvX_\yvN)/\yvN,
+   \yXf &\defd (\yXv_1 + \dotsb + \yXv_\yNv)/\yNv,
 \\
@@
 \label{noeq:subpopulation-av}
-\yxs &\defd (\yvx_1 + \dotsc + \yvx_n)/n.
+\yxs &\defd (\yxv_1 + \dotsc + \yxv_n)/n.
 \end{align}
-The quantities $\yvN\yXf$ and $n\yxs$ represent the total number of active
+The quantities $\yNv\yXf$ and $n\yxs$ represent the total number of active
 units in the population and the subpopulation. Quantities like $\yRf$ and $\yrs$
@@
 $\av{\dotv}$ are also extended to averages of $\tbinom{n}{m}$ or
-$\tbinom{\yvN}{m}$ products of $m$ states; \eg,
+$\tbinom{\yNv}{m}$ products of $m$ states; \eg,
 \begin{align}
 \av{\yX \yX} &\defd
-\tbinom{\yvN}{2}^{-1} (\yvX_1 \yvX_2 + \yvX_1 \yvX_3  + \dotsb +  \yvX_{\yvN-1} \yvX_\yvN),
+\tbinom{\yNv}{2}^{-1} (\yXv_1 \yXv_2 + \yXv_1 \yXv_3  + \dotsb +  \yXv_{\yNv-1} \yXv_\yNv),
 \\[2\jot]
 \sav{\yx\yx\yx} &\defd
-\tbinom{n}{3}^{-1} (\yvx_1 \yvx_2 \yvx_3 + \yvx_1 \yvx_2 \yvx_4 + \dotsb + \yvx_{n-2} \yvx_{n-1} \yvx_n),
+\tbinom{n}{3}^{-1} (\yxv_1 \yxv_2 \yxv_3 + \yxv_1 \yxv_2 \yxv_4 + \dotsb + \yxv_{n-2} \yxv_{n-1} \yxv_n),
 \end{align}
@@
   \label{noeq:joint_plaus}
-  \p(\yvX_1=\yvR_1, \yvX_2=\yvR_2, \dotsc, \yvX_\yvN=\yvR_\yvN \| \yH) \quad\text{or}\quad
+  \p(\yXv_1=\yRv_1, \yXv_2=\yRv_2, \dotsc, \yXv_\yNv=\yRv_\yNv \| \yH) \quad\text{or}\quad
 \p(\yX =\yR \| \yH).
@@
   \label{noeq:homog_math}
-  \p(\yvX_1=\yvR_1, \yvX_2=\yvR_2, \dotsc, \yvX_\yvN=\yvR_\yvN \| \yH) ={}
+  \p(\yXv_1=\yRv_1, \yXv_2=\yRv_2, \dotsc, \yXv_\yNv=\yRv_\yNv \| \yH) ={}
 \\ 
-\shoveright{\p(\yvX_1=\yvR_{\pi(1)}, \yvX_2=\yvR_{\pi(2)}, \dotsc, \yvX_\yvN=\yvR_{\pi(\yvN)} \| \yH)}
+\shoveright{\p(\yXv_1=\yRv_{\pi(1)}, \yXv_2=\yRv_{\pi(2)}, \dotsc, \yXv_\yNv=\yRv_{\pi(\yNv)} \| \yH)}
 \\
@@
 \qquad
-A \in \set*{0,\tfrac{1}{\yvN},\tfrac{2}{\yvN},\dotsc,1}.
+A \in \set*{0,\tfrac{1}{\yNv},\tfrac{2}{\yNv},\dotsc,1}.
 \end{equation}
@@
   \label{noeq:joint_plaus_N_homog}
-  \p(\yX = \yR \| H) = \binom{\yvN}{\yvN\yRf}^{-1} Q(\yRf).
+  \p(\yX = \yR \| H) = \binom{\yNv}{\yNv\yRf}^{-1} Q(\yRf).
 \end{equation}
@@
     \p(\yx = \yr \| \yH) =
-\sum_{\yvN A =0}^{\yvN} 
-\binom{\yvN-n}{\yvN A-n\yrs}\binom{\yvN}{\yvN A}^{-1}\,
+\sum_{\yNv A =0}^{\yNv} 
+\binom{\yNv-n}{\yNv A-n\yrs}\binom{\yNv}{\yNv A}^{-1}\,
 Q(A).
@@
 Note that the only summands contributing to this sum are those for which
-$n\yrs \le NA \le \yvN$; the others are zero because by definition
+$n\yrs \le NA \le \yNv$; the others are zero because by definition
 $\tbinom{M}{y}=0$ if $y<0$. This remark applies to all the sums of this
@@
 =
-\binom{\yvN-n}{\yvN \yRf-n \yrs}.
+\binom{\yNv-n}{\yNv \yRf-n \yrs}.
 \end{equation}
@@
   \p(\yxs = a  \| \yH) = \binom{n}{n a}
- \sum_{\yvN A=0}^{\yvN}
-\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1}\,
+ \sum_{\yNv A=0}^{\yNv}
+\binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1}\,
 Q(A),
@@
 =
-\binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1}.
+\binom{n}{n a}\binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1}.
 \end{equation}
@@
   \begin{equation*}
-    \yvx_{i_1} \yvx_{i_2} \dotsm \yvx_{i_m},
+    \yxv_{i_1} \yxv_{i_2} \dotsm \yxv_{i_m},
     \qquad 1\le i_1 < i_2 < \dotsb < i_m \le n
@@
   \label{noeq:eq_factmom_prod}
-\expe{\yvx_{i_1} \dotsm \yvx_{i_m} \| \yH}
+\expe{\yxv_{i_1} \dotsm \yxv_{i_m} \| \yH}
 =
@@
 =
-  \tfrac{(\yvN-m)!}{\yvN!}\expeb*{\frac{(\yvN\yXf)!}{(\yvN\yXf-m)!} \| \yH}.
+  \tfrac{(\yNv-m)!}{\yNv!}\expeb*{\frac{(\yNv\yXf)!}{(\yNv\yXf-m)!} \| \yH}.
 \end{equation}
@@
     \expe{\yxs^2 \| \yH} &= \expe{\yXf \| \yH}
-                                  \frac{\yvN-n}{(\yvN-1)n}
+                                  \frac{\yNv-n}{(\yNv-1)n}
  +{}
-\\&\quad\quad \expe{\yXf^2 \| \yH}\frac{\yvN\,(n-1)}{(\yvN-1)n},
+\\&\quad\quad \expe{\yXf^2 \| \yH}\frac{\yNv\,(n-1)}{(\yNv-1)n},
 \end{split}
@@
     \expe{\yxs^3 \| \yH} &=
-                              \expe{\yXf \| \yH}\frac{(\yvN-n)(\yvN-2n)}{(\yvN-1)(\yvN-2)n^2}+{}
+                              \expe{\yXf \| \yH}\frac{(\yNv-n)(\yNv-2n)}{(\yNv-1)(\yNv-2)n^2}+{}
                                     \\
                                     &\quad\quad\expe{\yXf^2 \| \yH}
-                                    \frac{3N\,(\yvN-n)(n-1)}{(\yvN-1)(\yvN-2)n^2} +{}
+                                    \frac{3N\,(\yNv-n)(n-1)}{(\yNv-1)(\yNv-2)n^2} +{}
                                     \\
                                     &\quad\quad\quad\expe{\yXf^3 \| \yH}
-                                    \frac{\yvN^2\,(n-1)(n-2)}{(\yvN-1)(\yvN-2)n^2}.
+                                    \frac{\yNv^2\,(n-1)(n-2)}{(\yNv-1)(\yNv-2)n^2}.
                                   \end{split}
@@
   \label{noeq:lin_transf_moments}
-  \expe{\yxs^m \| \yH} = \sum_{l=1}^m M_{ml}(n,\yvN) \expe{\yXf^l \| \yH},
-\end{equation}
-with a universal, lower-triangular transformation matrix $M_{ml}(n,\yvN)$ that
-depends only on $n$, $\yvN$, and the condition of symmetry~\eqref{noeq:homog_math}.
+  \expe{\yxs^m \| \yH} = \sum_{l=1}^m M_{ml}(n,\yNv) \expe{\yXf^l \| \yH},
+\end{equation}
+with a universal, lower-triangular transformation matrix $M_{ml}(n,\yNv)$ that
+depends only on $n$, $\yNv$, and the condition of symmetry~\eqref{noeq:homog_math}.
 
 As intuition suggests, we have 
-%$\expe{\yxs^m \| \yH} \xrightarrow{n\to \yvN} \expe{\yXf^m \| \yH}$.
+%$\expe{\yxs^m \| \yH} \xrightarrow{n\to \yNv} \expe{\yXf^m \| \yH}$.
 \begin{equation}
   \label{noeq:limit_moments}
-  \expe{\yxs^m \| \yH} \xrightarrow{n\to \yvN} 
+  \expe{\yxs^m \| \yH} \xrightarrow{n\to \yNv} 
   \expe{\yXf^m \| \yH},
@@
 \end{equation}
-the latter because ${\yvx_i}^m=\yvx_i$, since states are $\set{0,1}$-valued.
+the latter because ${\yxv_i}^m=\yxv_i$, since states are $\set{0,1}$-valued.
 
@@
 \cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
-with parameters $\yvN,\yvN\yXf,n$, weighted by the probabilities
+with parameters $\yNv,\yNv\yXf,n$, weighted by the probabilities
 $\p(\yXf = A \|\yH)$ \citep[\cf][\sect~4, esp.\ \eqn~(22)]{kendall1967}.
@@
   \p(\yXf = A \| \yH) = Q(A) \propto
-\binom{\yvN}{\yvN A}\exp[
-\yl_2\, \yvN A\,(\yvN A-1)/2 + \yl_1\, \yvN A].
+\binom{\yNv}{\yNv A}\exp[
+\yl_2\, \yNv A\,(\yNv A-1)/2 + \yl_1\, \yNv A].
 \end{equation}
@@
 with first and second moments as constraints and the reference distribution
-$Q_0$ defined by $Q_0(A) = 2^{-\yvN}\binom{\yvN}{\yvN A}$, corresponding to a
+$Q_0$ defined by $Q_0(A) = 2^{-\yNv}\binom{\yNv}{\yNv A}$, corresponding to a
 uniform probability distribution for the network state $\yX$.
@@
 subnetwork-average distributions $\p(\yxs = a \|\yH)$, for the case in
-which $\yvN = 1000$ units, $\lambda_1=-2.55$, $\lambda_2=0.005$, and
+which $\yNv = 1000$ units, $\lambda_1=-2.55$, $\lambda_2=0.005$, and
 $n=10, 50, 100, 250$. The distributions become broader as $n$ decreases,
@@
 expectations of $m$-tuple products of states
-$\expe{\yvx_{i_1}\dotsm \yvx_{i_m} \| \yH}$, proportional to the factorial
+$\expe{\yxv_{i_1}\dotsm \yxv_{i_m} \| \yH}$, proportional to the factorial
 moments, are not shown as they do not depend on $n$.
@@
 not true, since \eqn~\eqref{noeq:subpop_average}, as a linear mapping from
-$\RR^{\yvN+1}$ to $\RR^{n+1}$, with $\yvN$ larger than $n$, is onto but not into.
+$\RR^{\yNv+1}$ to $\RR^{n+1}$, with $\yNv$ larger than $n$, is onto but not into.
 Assigning a probability distribution $\p(\yxs = a \|\yH)$ to a
@@
     \label{noeq:constr_ex1}
-    \expe{\yxs \| \yHa} = c_1, \qquad \expe{\sav{\yvx_i \yvx_j}\|\yHa} = c_2;
+    \expe{\yxs \| \yHa} = c_1, \qquad \expe{\sav{\yxv_i \yxv_j}\|\yHa} = c_2;
   \end{equation}
@@
   \pf(\yX =\yR \|\yHa) =
-\varKappa\exp[\yL_2 \yvN\yRf\,(\yvN\yRf-1)/2 + \yL_1 \yvN\yRf]
+\varKappa\exp[\yL_2 \yNv\yRf\,(\yNv\yRf-1)/2 + \yL_1 \yNv\yRf]
 \\
@@
 \\
-&\varKappa\sum_{\yvN A=0}^\yvN %m!\, 
-\binom{\yvN-m}{\yvN A-m}
-\exp[\yL_2 \yvN A\,(\yvN A-1)/2 + \yL_1 \yvN A]
+&\varKappa\sum_{\yNv A=0}^\yNv %m!\, 
+\binom{\yNv-m}{\yNv A-m}
+\exp[\yL_2 \yNv A\,(\yNv A-1)/2 + \yL_1 \yNv A]
 = c_m, \quad m=1,2.
@@
   c_2 = \expe{\sav{\yx\yx}\|\yHa} = 
-\binom{\yvN}{2}^{-1}
-\sum_{\yvN A=0}^\yvN %m!\, 
-\binom{\yvN A}{2} \, \p(\yXf = A\|\yHa),
+\binom{\yNv}{2}^{-1}
+\sum_{\yNv A=0}^\yNv %m!\, 
+\binom{\yNv A}{2} \, \p(\yXf = A\|\yHa),
 \end{equation}
@@
 \binom{n}{n a}
-\binom{\yvN-n}{\yvN\yRf-n a}\Biggr]
+\binom{\yNv-n}{\yNv\yRf-n a}\Biggr]
 \\
@@
 \\
-&\sum_{\yvN A=0}^\yvN 
+&\sum_{\yNv A=0}^\yNv 
 %\binom{N}{N A}^{-1}
 \binom{n}{n a}
-\binom{\yvN-n}{\yvN A-n a}\,
+\binom{\yNv-n}{\yNv A-n a}\,
 \exp\Biggl[
@@
 \binom{n}{n a}
-\binom{\yvN-n}{\yvN A-n a}\Biggr]
+\binom{\yNv-n}{\yNv A-n a}\Biggr]
 =q( a)
@@
 the final distribution may have a multiplicity factor. When we consider the
-$2^{\yvN}$ states of $\yvN$ units we are not excluding that these might be
+$2^{\yNv}$ states of $\yNv$ units we are not excluding that these might be
 marginals of $2^{M}$ states of $M$ units. Each one has the same
-multiplicity $2^{M-\yvN}$, but this constant multiplicity factor disappears
-by normalization. So the method applies just as in the case of $\yvN$ units
+multiplicity $2^{M-\yNv}$, but this constant multiplicity factor disappears
+by normalization. So the method applies just as in the case of $\yNv$ units
 only, right?
@@
 involves probability considerations. Consider three cases with a full
-population of two units, $M=2$, of which we consider one unit, $\yvN=1$.
+population of two units, $M=2$, of which we consider one unit, $\yNv=1$.
 \begin{itemize}
@@
 \begin{enumerate}
-\item have a population with $\yvN$ units, $2^{\yvN}$ possible states
+\item have a population with $\yNv$ units, $2^{\yNv}$ possible states
 \item expect averages of $\yx$ active neurons and $\yxxs$ active pairs
 \item use maximum-entropy to choose a probability distribution for the
-  states of the $\yvN$ units conforming to our expectations.
+  states of the $\yNv$ units conforming to our expectations.
 \end{enumerate}
@@
 \item have a population with $M$ units, $2^{M}$ possible states
-\item expect that any subpopulation of $\yvN$ units has $\yxs$ active
+\item expect that any subpopulation of $\yNv$ units has $\yxs$ active
   neurons and $\yxxs$ active pairs
@@
 \item marginalize to find the probability distribution for the states of
-  $\yvN$ units.
+  $\yNv$ units.
 \end{enumerate}
@@
   average~\eqref{noeq:pop_average_Q}, consider that if the network total
-  is $\yvN\yXf$, then $\yvN\yXf$ out of $\yvN$ units are active, and there are
-  $\tbinom{\yvN}{\yvN\yXf}$ possible states for which this can be true; therefore
-  \begin{equation}
-    \labelbis{eq:joint_plaus_\yvN_homog}
-    \p(\yX = \yR \| H) = \binom{\yvN}{\yvN\yRf}^{-1}Q(\yRf).
+  is $\yNv\yXf$, then $\yNv\yXf$ out of $\yNv$ units are active, and there are
+  $\tbinom{\yNv}{\yNv\yXf}$ possible states for which this can be true; therefore
+  \begin{equation}
+    \labelbis{eq:joint_plaus_\yNv_homog}
+    \p(\yX = \yR \| H) = \binom{\yNv}{\yNv\yRf}^{-1}Q(\yRf).
   \end{equation}
@@
 the subnetwork average $\yxs$ conditional on the network average $\yXf$.
-There are $\tbinom{\yvN}{\yvN\yXf}$ possible network states if the network
-average is $\yXf$, \ie\ if $\yvN\yXf$ units are in state $1$; the conditional
-probability of each is therefore $1/\tbinom{\yvN}{\yvN\yXf}$, owing to the
+There are $\tbinom{\yNv}{\yNv\yXf}$ possible network states if the network
+average is $\yXf$, \ie\ if $\yNv\yXf$ units are in state $1$; the conditional
+probability of each is therefore $1/\tbinom{\yNv}{\yNv\yXf}$, owing to the
 symmetry assumption~\eqref{noeq:homog_math}. Now consider the subnetwork of the
@@
 ones in state $1$ is the sum of the probabilities of all states for
-which $\yvN\yXf-n\yxs$ of the remaining $\yvN-n$ units are in state $1$; there are
-$\tbinom{\yvN-n}{\yvN\yXf-n\yxs}$ such states, all equally probable. Finally,
+which $\yNv\yXf-n\yxs$ of the remaining $\yNv-n$ units are in state $1$; there are
+$\tbinom{\yNv-n}{\yNv\yXf-n\yxs}$ such states, all equally probable. Finally,
 there are $\tbinom{n}{n\yxs}$ possible ways, all equally probable, in which
@@
       \p(\yx=\yr \|  \yX =\yR, \yH) \,
-      \binom{\yvN}{\yvN A}^{-1},
+      \binom{\yNv}{\yNv A}^{-1},
       \\
       &=\sum_{\yr}^{\yrs=a} 
-      \binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1},
+      \binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1},
       \\
       &=
-      \binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1},
+      \binom{n}{n a}\binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1},
 \end{split}
@@
   which describes the probability of, say, drawing a proportion of $\yxs$
-  blue balls in $n$ drawings without replacement from an urn with $\yvN$
+  blue balls in $n$ drawings without replacement from an urn with $\yNv$
   balls, a fraction $\yXf$ of which are blue.
@@
       \p(\yxs =a \| \yH)
-      &= \sum_{\yvN A=0}^\yvN\p(\yxs = a \| \yXf = A, \yH)\,
+      &= \sum_{\yNv A=0}^\yNv\p(\yxs = a \| \yXf = A, \yH)\,
       \p(\yXf = A \| \yH),
       \\
-      &=\sum_{\yvN A=0}^\yvN
-      \binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1}
+      &=\sum_{\yNv A=0}^\yNv
+      \binom{n}{n a}\binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1}
       Q(A).
@@
     \label{noeq:margin_products}
-    \sum_{\yx} \yvx_1\dotsm \yvx_m \p(\yx =\yr \| \yH)
-      =\sum_{\yr,\yR}  \yvr_1\dotsm \yvr_m \p(\yx=\yr \| \yX=\yR,\yH)
+    \sum_{\yx} \yxv_1\dotsm \yxv_m \p(\yx =\yr \| \yH)
+      =\sum_{\yr,\yR}  \yrv_1\dotsm \yrv_m \p(\yx=\yr \| \yX=\yR,\yH)
       \, \p(\yX=\yR \| \yH),
@@
       \begin{aligned}
-      &=\sum_{\yr,\yR} \yvr_1\dotsm \yvr_m\,\delt(\yvR_1-\yvr_1)\dotsm\delt(\yvR_m-\yvR_m)
+      &=\sum_{\yr,\yR} \yrv_1\dotsm \yrv_m\,\delt(\yRv_1-\yrv_1)\dotsm\delt(\yRv_m-\yRv_m)
       \p(\yX = \yR \| \yH),
       \\
-      &=\sum_{\yR} \yvR_1\dotsm \yvR_m \p(\yX =\yR \| \yH).
+      &=\sum_{\yR} \yRv_1\dotsm \yRv_m \p(\yX =\yR \| \yH).
     \end{aligned}
@@
   \begin{equation*}
-    \yvx_1 \yvx_2 + \yvx_1 \yvx_3 + \dotsb + \yvx_{n-1} \yvx_n.
+    \yxv_1 \yxv_2 + \yxv_1 \yxv_3 + \dotsb + \yxv_{n-1} \yxv_n.
   \end{equation*}
@@
   \begin{equation}
-    \sav{\yvx_{i_1} \dotsm \yvx_{i_m}} = \binom{n\yxs}{m} \binom{n}{m}^{-1}.
+    \sav{\yxv_{i_1} \dotsm \yxv_{i_m}} = \binom{n\yxs}{m} \binom{n}{m}^{-1}.
   \end{equation}
@@
     \label{noeq:expect_product_m}
-    \expe*{\sav{\yvx_{i_1} \dotsm \yvx_{i_m}} \| \yH}
+    \expe*{\sav{\yxv_{i_1} \dotsm \yxv_{i_m}} \| \yH}
     \\
@@
       \\
-      &=\frac{(n-m)!}{n!} \sum_{\yvN A=0}^\yvN
+      &=\frac{(n-m)!}{n!} \sum_{\yNv A=0}^\yNv
       Q( A)
@@
       \sum_{n a=0}^n 
-      m!\,\binom{n a}{m}\, \binom{n}{n a}\binom{\yvN-n}{\yvN A-n a}\binom{\yvN}{\yvN A}^{-1} 
+      m!\,\binom{n a}{m}\, \binom{n}{n a}\binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1} 
       \Biggr].
@@
   \begin{equation}
-    \sum_{n a=0}^n m!\,\binom{n a}{m} \binom{n}{n a}\binom{\yvN-n}{\yvN\yXf-n a}
-    \binom{\yvN}{\yvN\yXf}^{-1} 
+    \sum_{n a=0}^n m!\,\binom{n a}{m} \binom{n}{n a}\binom{\yNv-n}{\yNv\yXf-n a}
+    \binom{\yNv}{\yNv\yXf}^{-1} 
     =
-    m!\,\binom{n}{m} \binom{n a}{m}\binom{\yvN}{m},
+    m!\,\binom{n}{m} \binom{n a}{m}\binom{\yNv}{m},
   \end{equation}
@@
     \label{noeq:binom_identity}
-    \binom{\yvN}{M}\binom{M}{m}=\binom{\yvN}{m}\binom{\yvN-m}{M-m},
+    \binom{\yNv}{M}\binom{M}{m}=\binom{\yNv}{m}\binom{\yNv-m}{M-m},
   \end{equation}
@@
     \expe{\yXf^m \| \yH} \defd
-    \sum_{\yvN A=0}^{\yvN} A^m \p(\yXf = A  \| H),
+    \sum_{\yNv A=0}^{\yNv} A^m \p(\yXf = A  \| H),
   \end{equation}
@@
   \p(\yXf = \yRf \| \yHa)  = \yL\,
-\binom{\yvN}{\yvN \yRf}\exp\biggl[
-\yL_2 \yRf \,\frac{\yvN \yRf-1}{\yvN-1} + \yL_1\, \yRf\biggr]
+\binom{\yNv}{\yNv \yRf}\exp\biggl[
+\yL_2 \yRf \,\frac{\yNv \yRf-1}{\yNv-1} + \yL_1\, \yRf\biggr]
 \end{equation}
@@
   \p(\yxs = \yrs \| \yHa)  = \yL
-  \sum_{\yvN\yRf=0}^\yvN
-  \binom{n}{n\yrs} \binom{\yvN-n}{\yvN\yRf-n\yrs} \exp\biggl[
-\yL_2 \yRf \,\frac{\yvN \yRf-1}{\yvN-1} + \yL_1\, \yRf\biggr].
+  \sum_{\yNv\yRf=0}^\yNv
+  \binom{n}{n\yrs} \binom{\yNv-n}{\yNv\yRf-n\yrs} \exp\biggl[
+\yL_2 \yRf \,\frac{\yNv \yRf-1}{\yNv-1} + \yL_1\, \yRf\biggr].
 \end{equation}
@@
 $\p(\yxs = \yrs \| \yHb)$ are different.
-Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for $\yvN=5000$,
+Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for $\yNv=5000$,
 $n=100$, $c_1=0.45$, $c_2=0.35$. The two distributions have slightly
@@
 more peaked. Another example is shown in
-\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for $\yvN=5000$, $n=200$,
+\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for $\yNv=5000$, $n=200$,
 and neurobiologically more realistic constraints $c_1=0.045$, $c_2=0.0025$

### 2018-05-17 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-16T11:37:42+0200
+%% Last-Updated: 2018-05-17T18:47:41+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \DeclareMathSymbol{\varMu}{\mathalpha}{egreeki}{"4D}
-% \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
+ \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
  \DeclareMathSymbol{\varXi}{\mathalpha}{egreeki}{"04}
@@
 \newcommand*{\ie}{{i.e.}}
-%\newcommand*{\ca}{{c.}}
+\newcommand*{\ca}{{c.}}
 %\newcommand*{\Ie}{{I.e.}}
@@
 \newcommand*{\yNv}{\nu}
+\newcommand*{\yNN}{\varNu}
 
@@
 \newcommand*{\yXf}{\av{\yX}}%pop av state
-\newcommand*{\yxxs}{\sav{\yx\yx}}%subpop av state
-\newcommand*{\yXXf}{\av{\yX\yX}}%subpop av state
+\newcommand*{\yxxs}{\sav{\yr\yr}}%subpop av state
+\newcommand*{\yXXf}{\av{\yR\yR}}%subpop av state
 \newcommand*{\yr}{\bm{\yrv}}%subpop value
@@
 \newcommand*{\yHc}{\varKappa}
-\newcommand*{\yHd}{\varUpsilon}
+\newcommand*{\yHd}{\varIota_\textrm{u}}
 \newcommand*{\yHp}{\varIota'}
@@
 %Lagr multipliers
-\newcommand*{\yf}{\bm{f}}
+\newcommand*{\yf}{\bm{g}}
 \newcommand*{\yc}{\bm{c}}
-\newcommand*{\yL}{\bm{\varLambda}}
-\newcommand*{\yl}{\bm{\lambda}}
-\newcommand*{\yk}{\zeta}
-\newcommand*{\yK}{\varZeta}
+\newcommand*{\yL}{\bm{\lambda}}
+\newcommand*{\yl}{\bm{l}}
+\newcommand*{\yk}{z}
+\newcommand*{\yK}{\zeta}
 %entropies
@@
 \section{Introduction: \me\ and sampling in neuroscience}
+\label{sec:intro}
 
@@
   \begin{aligned}
+    &\text{\enquote{$\yXv_{\iota} = \yRv_{\iota}$} 
+      stands for \enquote{the activity of the $\iota$th neuron is $\yRv_{\iota}$}},
+    \\
+    &\text{\enquote{$\yXf = \yRf$} 
+      stands for \enquote{the (population-averaged) activity of the neurons is $\yRf$}},
+    \\
     &\text{\enquote{$\yxv_i = \yrv_i$} 
       stands for \enquote{the activity of the $i$th sample neuron is $\yrv_i$}},
-    \\
-    &\text{\enquote{$\yxs = \yrs$} 
-      stands for \enquote{the (population-averaged) activity of the sample neurons is $\yrv_i$}},
   \end{aligned}
 \end{equation}
-and similarly for other quantities and for the activities of the neurons of
-the larger population.
+and similarly for other quantities.
 
@@
 
+\bigskip
+
 The theory of statistical sampling is covered in many excellent texts, for
@@
 
-We need to make an initial probability assignment for the state of the
-population before any experimental observations are
-made. %, no matter what kinds of predictions we are interested in.
-This initial assignment will be modified by our experimental observations,
-which can involve only a sample of the population. Our initial probability
-assignment should reflect that samples are somehow representative of the
-whole population. We consider here a state of knowledge that express
-this representativeness.
+We need to make an initial probability assignment for the state of the full
+population before any experimental observations are made. This initial
+assignment will be modified by our experimental observations, and these can
+involve just a sample of the population. Our state of knowledge and initial
+probability assignment should reflect that samples are somehow
+representative of the whole population.
 
@@
 \emph{ignorance} about the state of the population, not an expression of
-some biologic or physical identity of the neurons.
-
-The \emph{representation theorem for finite exchangeability} states that
-the symmetric distribution in the state of knowledge $\yH$ must obey
+some biologic or physical symmetry or identity of the neurons.
+
+The \emph{representation theorem for finite exchangeability} states that,
+in the state of knowledge $\yH$, the symmetric distribution for the full
+activity is completely determined by the distribution for its
+population-average:
 \begin{equation}
@@
   \p(\yX = \yR \|  \yH) \equiv
-  \p(\yX = \yR \| \yXf = \yRf, \yH)\,
+  \sum_{\yRf}\p(\yX = \yR \| \yXf = \yRf, \yH)\,
   \p(\yXf=\yRf \| \yH) =
-  \binom{\yNv}{\yNv\yRf}^{-1} \p(\yXf=\yRf \| \yH),
-\end{equation}
-the latter being the probability for the population average $\yX$. A sum is
-only apparently missing in the central term: its summands $\yXf=A$ are all
-zero except for $A=\yRf$. Proof of this theorem and generalizations to
+  \binom{\yNv}{\yNv\yRf}^{-1} \p(\yXf=\yRf \| \yH).
+\end{equation}
+The equivalence on the left is just an application of the law of total
+probability; the equality on the right is the statement of the theorem.
+This result is intuitive: owing to symmetry, we must assign equal
+probabilities to all $\tbinom{\yNv}{\yNv\yRf}$ activity vectors with
+$\yNv\yRf$ active neurons; the probability of each activity vector is
+therefore given by that of the average activity divided by the number of
+possible vector values. Proof of this theorem and generalizations to
 non-binary and continuum cases are given by de~Finetti
@@
 \cite{diaconis1977,diaconisetal1980}, Heath \amp\ Sudderth
-\cite{heathetal1976}. This theorem is intuitive: owing to symmetry, we must
-assign equal probabilities to all states with $\yNv\yRf$ active neurons.
-
-By marginalization we obtain the probability for the state of the sample:
-\begin{gather}
-  \label{eq:marginal}
-  \p(\yx = \yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs \| \yH),
-  \\
-  \shortintertext{with}
-  \label{eq:subpop_average}
-  \p(\yxs=\yrs \| \yH) = \sum_{\yNv \yRf =0}^{\yNv}
-    \p(\yxs = \yrs \|\yXf=\yRf, \yH)\,
-  \p(\yXf=\yRf \| \yH),
-  \\[2\jot]
+\cite{heathetal1976}.
+
+Our uncertainties about the full population and the sample are connected
+via the conditional probability
+\begin{equation}
   \label{eq:conditional_hypergeometric}
@@
   \binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}
-  \defs \ypp(\yrs\|\yRf).
+  \defs \ypp(\yrs\|\yRf),
+\end{equation}
+which is a hypergeometric distribution, typical of \enquote{drawing without
+  replacement} problems. The combinatorial proof of this expression is in
+fact the same as for this class of problems
+\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
+
+Using the conditional probability above we obtain the probability for the
+activity of the sample:
+\begin{gather}
+  \label{eq:subpop_average}
+  \p(\yxs=\yrs \| \yH) = \sum_{\yRf}
+  \p(\yxs = \yrs \|\yXf=\yRf, \yH)\,
+  \p(\yXf=\yRf \| \yH)
+  = 
+  \sum_{\yRf}
+  \ypp(\yrs\|\yRf)\,
+  \p(\yXf=\yRf \| \yH).
 \end{gather}
-The conditional probability in the last formula is a hypergeometric
-distribution $\ypp(\yrs\|\yRf)$, typical of \enquote{drawing without
-  replacement} problems. The combinatorial proof of the formulae above is
-in fact the same as for this class of problems
-\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-Our initial symmetric ignorance should intuitively also apply to the
-sample; the probability for the state of the
-sample~\eqref{eq:subpop_average} indeed satisfies the representation
-theorem~\eqref{eq:joint_plaus_N_homog} as well.
+It should be proved that the probability distribution for the full activity
+of the sample is also symmetric and completely determined by the
+distribution of its population-averaged activity:
+\begin{equation}
+  \label{eq:marginal}
+  \p(\yx = \yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs \| \yH).
+\end{equation}
+This is intuitively clear: our initial symmetric ignorance should also
+apply to the sample. The distribution for the
+sample~\eqref{eq:subpop_average} indeed satisfies the same representation
+theorem~\eqref{eq:joint_plaus_N_homog} as the distribution for the full
+population.
 
 The conditional probability
-$\p(\yxs = \yrs \|\yXf=\yRf, \yH) \equiv \ypp(\yrs,\yRf)$ relates the
-spaces of the sample average $\yXf\in\set{0,\dotsc,\yNv}$ and of the
-population average $\yxs\in\set{0,\dotsc,n}$. It is a coarsening projector
-of any probability for $\yXf$ onto a marginal probability for $\yxs$, as
-\eqn~\eqref{eq:subpop_average} shows. For each function $f$ of the sample
-average $\yxs$, the  hypergeometric distribution 
-$\ypp(\yrs,\yRf)$ also allows us to construct a function $f^*$ of the
-population average $\yXf$
-\begin{equation}
+$\p(\yxs = \yrs \|\yXf=\yRf, \yH) \equiv \ypp(\yrs \|\yRf)$, besides
+relating the distributions for the population and sample activities via
+marginalization, also allows us to express the expectation value of any
+function of the sample activity, $g(\yrs)$, in terms of the distribution
+for the full population, as follows:
+\begin{multline}
   \label{eq:pullback_P}
-    f^*(\yXf) \defd \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yXf),
-%  \ypp^*\colon f(\yxs=\dotv) \to f^*(\yXf=\dotv) \defd \sum_{N\yrs=0}^n
-%  f(\yrs)\ypp(\yrs\|\dotv),
-\end{equation}
-that has the same expectation as $f$:
-\begin{equation}
-  \expeb{f^*(\yXf)}
+  \expe{g\|I}
   \equiv
-  \sum_{\yNv\yRf=0}^\yNv f^*(\yRf)\,\p(\yXf=\yRf \| \yH)
+  \sum_{\yrs} g(\yrs)\,\p(\yxs=\yrs \| \yH)
+  =
+  \sum_{\yrs} g(\yrs) \sum_{\yRf} \ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
+  ={}\\
+  \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
+  \p(\yXf=\yRf \| \yH),
 %  =
 %  \sum_{N\yRf=0}^N \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
-  =
-  \sum_{n\yrs=0}^n f(\yrs)\,\p(\yxs=\yrs \| \yH)
-  \equiv  \expeb{f(\yxs)}.
-\end{equation}
-The middle equality uses the combination of formulae~\eqref{eq:pullback_P}
-and~\eqref{eq:subpop_average}.
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=0.9\linewidth]{pop_sample_projection3.pdf}%
-\caption{Log-plot of the hypergeometric distribution
-  $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
-\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
-  colourbar depending on your \textsc{pdf} viewer.)}
-\label{fig:hypergeom_proj}
-\end{figure}%hypergeometric_identity.nb
-A look at a plot of the hypergeometric distribution $\ypp(\yrs\|\yRf)$, see
-\fig~\ref{fig:hypergeom_proj}, reveals that it is a sort of \enquote{fuzzy
-  identity matrix} between the $\yXf$-space $\set{0,\dotsc,\yNv}$ and
-$\yxs$-space $\set{0,\dotsc,n}$. When $n=\yNv$ it is the identity matrix. We
-thus have that
-\begin{equation}
-  \label{eq:roughly_equal_nN}
-  \p(\yxs=a) \approx \p(\yXf=a),\qquad
-\expeb{f(\yxs)} \approx \expeb{f(\yXf)}.
-\end{equation}
-These are only very approximate equalities: they may miss important
-features of the two probability distributions. In the next section we will
-in fact emphasize their differences. If the distribution for the population
-average $\yXf$ is bimodal, for example, the bimodality can be lost in the
-distribution for the sample average $\yxs$, owing to the coarsening effect
-of $\ypp(\yrs\|\yRf)$.
-
-The approximate equalities above, on the other hand, express the fact that
-\emph{our uncertainty about the sample is representative of our uncertainty
-  about the population and about other samples}, and vice versa. These
-approximate equalities are not generally valid for other probability
-distributions for the population. They should be contrasted with the limits
-$\p(\yxs=a) \to \p(\yXf=a)$ and $\expeb{f(\yxs)} \to \expeb{f(\yXf)}$, as
-$n\to \yNv$, do: these limits are trivially, universally valid because the
-sample becomes the full population as $n\to \yNv$. In particular, these limits
-hold even in cases where the conditional probability
-$\p(\yxs = \yrs \|\yXf=\yRf)$ is not a fuzzy identity, our uncertainties
-about sample and about population can differ wildly, and the approximate
-equalities~\eqref{eq:roughly_equal_nN} do not hold.
+%  \equiv  \expeb{g^*}.
+\end{multline}
+where the second step uses \eqn~\eqref{eq:subpop_average}. The last
+expression shows that the expectation of the function $g$, whose argument
+is the sample activity $\yrs$, is equal to the expectation of the function
+$g^*$, whose argument is the full-population activity $\yRf$, defined as
+$g^*(\yRf) \defd \sum_{\yrs} g(\yrs)\,\ypp(\yrs\|\yRf)$.
 
@@
 
-% The projection~\eqref{eq:projection_P} and pull-back~\eqref{eq:pullback_P}
-% have the property
-% \begin{equation}
-%   \label{eq:expectation_pullback}
-%   \expeb{f(\yxs) \|\yH} = \expeb{f^*(\yXf)\|\yH}
-%   \qquad\text{or}\qquad
-%   \sum_{n\yrs=0}^n f(\yrs)\,p_*(\yxs=\yrs) = \sum_{\yNv\yRf=0}^\yNv f^*(\yRf)\,p(\yXf=\yRf).
-% \end{equation}
-For averaged products, $\sav{\underbrace{\yx\dots\yx}_{\text{$m$
-      factors}}}\equiv\binom{n\yxs}{m}/\binom{n}{m}$,
-formulae~\eqref{eq:pullback_P} yield the useful equality
+The final expression in \eqn~\eqref{eq:pullback_P} is important for our
+\me\ application: the requirement that the function $g$, defined for the
+sample, have a  value $c$ obtained from observed data,
+\emph{translates into a linear constraint for the distribution of the full
+  population}:
+\begin{equation}
+  \label{eq:constraint_extended}
+  c = \expe{g \| \yH} \equiv \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
+  \p(\yXf=\yRf \| \yH).
+\end{equation}
+
+In particular, when the function $g$ is the $m$-activity of the sample,
+$g(\yrs) = \sav{\yr\dotso\yr} \equiv \binom{n\yrs}{m}/\binom{n}{m}$, we
+find
 \begin{equation}
@@
     % \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
-  \expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
+  \expe{\sav{\underbrace{\yr \dotsm \yr}_{\text{$m$ factors}}} \| \yH}
 \equiv
@@
     % {}
+    % \frac{(n-m)!}{n!}
+    \sum_{\yrs} %m!\,
     \binom{n}{m}^{-1}
-    % \frac{(n-m)!}{n!}
-    \sum_{\mathclap{n \yrs=0}}^n %m!\,
     \binom{n \yrs}{m}\, \p(\yxs=\yrs \| \yH)
@@
     % \frac{(N-m)!}{N!}
-    \sum_{\mathclap{\yNv \yRf=0}}^\yNv %m!\,
+    \sum_{\yRf} %m!\,
     \binom{\yNv \yRf}{m}\, \p(\yXf=\yRf \| \yH)
-\equiv    \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}.
-\end{equation}
-The proof of the middle equality uses the expression for the $m$th
-factorial moment of the hypergeometric distribution \cite{potts1953}. Thus,
-in the state of knowledge $\yH$ the expectations of the averages of
-activity products \emph{are initially the same for the sample and for the
-  full population}. Similar relations can be found for the raw moments
-$\expe{\yxs^m}$ and $\expe{\yXf^m}$, which can be written in terms of the
-product expectations via \eqn~\eqref{eq:products_intermsof_average}.
-
-\section{Maximum-entropy at the population level with constraints at the
-  sample level}
+\equiv    \expe{\av{\underbrace{\yR \dotsm \yR}_{\text{$m$ factors}}} \| \yH},
+\end{equation}
+that is, \emph{the expected values of the $m$-activities of the sample and
+  of the full population are equal}. The proof of the middle equality uses
+the expression for the $m$th factorial moment of the hypergeometric
+distribution and can be found in \textcite{potts1953}. Similar relations
+can be found for the raw moments $\expe{\yrs^m}$ and $\expe{\yRf^m}$, which
+can be written in terms of the product expectations using
+\eqn~\eqref{eq:products_intermsof_average}.
+
+Thus, in a \me\ application, when we require the expectation of the
+$m$-activity of a sample to have a particular value, we are also
+requiring  the expectation of the $m$-activity of the full population to
+have the same value.
+
+\bigskip
+
+\begin{figure}[!b]
+\centering
+\includegraphics[width=0.9\linewidth]{pop_sample_projection3.pdf}%
+\caption{Log-density plot of the hypergeometric distribution
+  $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
+\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
+  colourbar depending on your \textsc{pdf} viewer.)}
+\label{fig:hypergeom_proj}
+\end{figure}%hypergeometric_identity.nb
+These expectation equalities between sample and full population should not
+be surprising: we intuitively \emph{expect} that the proportion of coloured
+balls sampled from an urn should be roughly equal to the proportion of
+coloured ball contained in the urn. The formulae in the present section
+formalize and mathematically express our intuition. The hypergeometric
+distribution $\ypp(\yrs\|\yRf)$ plays an important role in this
+formalization. A look at its plot, \fig~\ref{fig:hypergeom_proj}, reveals
+that it is a sort of \enquote{fuzzy identity transformation}, or fuzzy
+Kronecker delta, between the $\yRf$-space $\set{0,\dotsc,\yNv}$ and
+$\yrs$-space $\set{0,\dotsc,n}$. From \eqn~\eqref{eq:marginal} we thus have
+that
+\begin{equation}
+  \label{eq:roughly_equal_nN}
+  \p(\yxs=a \|\yH) \approx \p(\yXf=a \|\yH),\qquad
+\expeb{g(\yrs) \|\yH} \approx \expeb{g(\yRf) \|\yH},
+\end{equation}
+where $g$ is any smooth function defined on $\clcl{0,1}$. These approximate
+equalities express the intuitive fact that \emph{our uncertainty about the
+  sample is representative of our uncertainty about the population and
+  about other samples}, and vice versa. When $n=\yNv$, $\ypp(\yrs\|\yRf)$
+becomes the identity matrix and the approximate equalities above become
+exact -- of course, since we have sampled the full population.
+
+But the approximate equalities above may miss important features of the two
+probability distributions. In the next section we will in fact emphasize
+their differences. If the distribution for the population average $\yRf$ is
+bimodal, for example, the bimodality can be lost in the distribution for
+the sample average $\yrs$, owing to the coarsening effect of
+$\ypp(\yrs\|\yRf)$.
+
+% They should be contrasted with the limits
+% $\p(\yxs=a) \to \p(\yXf=a)$ and $\expeb{g(\yxs)} \to \expeb{g(\yXf)}$, as
+% $n\to \yNv$, do: these limits are trivially, universally valid because the
+% sample becomes the full population as $n\to \yNv$. In particular, these limits
+% hold even in cases where the conditional probability
+% $\p(\yxs = \yrs \|\yXf=\yRf)$ is not a fuzzy identity, our uncertainties
+% about sample and about population can differ wildly, and the approximate
+% equalities~\eqref{eq:roughly_equal_nN} do not hold.
+
+\section{Maximum-entropy: sample level vs full-population level}
 \label{sec:specific_initial_probability}
 
-The probability formulae
-\eqref{eq:joint_plaus_N_homog}--\eqref{eq:conditional_hypergeometric} are
-constraints on our initial probability assignment, but do not determine it
-numerically. The probability $\p(\yXf=\yRf \| \yH)$ for the population
-average needs to be numerically specified, and by marginalization
-\eqref{eq:subpop_average} it will determine that of the sample average,
-$\p(\yxs=\yrs \| \yH)$. If we numerically specify the latter, the former is
-not completely specified, because \eqn~\eqref{eq:subpop_average} linearly
-constrains $\yNv+1$ unknowns by only $n+1$ equations.
-
-We may want to specify the probability by enforcing the sample expectations
-of several functions to have specific values, for example
-$\expe{\yxs}=c_1$, $\expe{\yxxs}=c_2$. This is still an underdetermined
-problem: several distributions can have the same desired expectations, as
-clear from \eqns~\eqref{eq:expe_products}.
-
-The \me\ method is brought into play to solve this indeterminacy. It
-selects one distribution, purported to be \enquote{maximally noncommittal},
-among those that have the desired expectations. But here's a dilemma: the
-expectation formulae~\eqref{eq:pullback_P} allow us to apply the method to
-find the probability of the population $\p(\yXf=\yRf \| \yH)$, or of the
-sample $\p(\yxs=\yrs \| \yH)$. \emph{The two applications, however, are
-  inequivalent}. They lead to numerically different distributions for the
-sample average $\p(\yxs=\yrs \| \yH)$.
-
-Suppose we want to constrain the sample expectations of a vector function
-$\yf = (f_1,\dotsc,f_m)$ to the vector values $\yc=(c_1,\dotsc,c_m)$, that
-is, $\expeb{\yf(\yxs)}=\yc$. Application of \me\
-\cite{sivia1996_r2006,meadetal1984} at the population level, denoted by
-$\yHa$, gives
+In the previous section we have seen that observations about a sample can
+be used as constraints on the distribution for the activity of the full
+population. Let us use such constraints with the \me\ method. Suppose that
+we want to constrain $m$ functions of the sample activity, vectorially
+written $\yf \defd (g_1,\dotsc,g_m)$, to $m$ values
+$\yc \defd (c_1,\dotsc,c_m)$. These functions are typically $k$-activities
+$\sav{\yr\dotso \yr}$, and the values are typically the time averages of
+the observed sample, as discussed in \sect~\ref{sec:intro}:
+$\yc = \sum_t \yf[\yrs(t)]/T$.
+
+Let us apply the relative-\me\ method \cite{sivia1996_r2006,meadetal1984}
+directly to sampled neurons; denote this approach by $\yHb$. Then we apply
+the method to the full population of neurons, most of which are unsampled;
+denote this approach by $\yHa$.
+
+Applied directly to the sampled neurons, the method yields the distribution
+\begin{equation}
+  \label{eq:app_maxent_sample}
+  \p(\yxs=\yrs \|\yHb)
+  =\frac{1}{\yk(\yl)}\,
+  \binom{n}{n\yrs}
+  \exp[\yl\T \yf(\yrs)]
+\end{equation}
+where $\yk(\yl)$ is a normalization constant. The binomial in front of the
+exponential appears because we must account for the multiplicity by which
+the population-average activity $\yrs$ can be realized: $\yrs=0$ can be
+realized in only one way (all neurons inactive), $\yrs=1/n$ can be realized
+in $n$ ways (one active neuron out of $n$), \etc\ This term is analogous to
+the \enquote{density of states} in front of a Boltzmann factor in
+statistical mechanics \cite[\chap~16]{callen1960_r1985}. The $m$ Lagrange
+multipliers $\yl\defd (l_1,\dotsc,l_m)$ must satisfy the $m$ constraint
+equations
+\begin{equation}
+  \label{eq:app_maxent_sample_constraints}
+  c_k = \expe{g_k \|\yHb} \equiv
+  \frac{1}{\yk(\yl)}\sum_{\yrs}  g_k(\yrs) \binom{n}{n\yrs}
+  \exp[\yl\T \yf(\yrs)], \qquad k=1,\dotsc,m.
+\end{equation}
+
+Applied to the full population, using the constraint
+expression~\eqref{eq:constraint_extended} derived in the previous section,
+the method yields the distribution for the full-population activity
 \begin{equation}
   \label{eq:app_maxent_pop}
-  \p(\yXf = \yRf \| \yHa)  = \yK\,
+  \p(\yXf = \yRf \| \yHa)  = \frac{1}{\yK(\yL)}\,
   \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
-  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
-\end{equation}
-and then by marginalization~\eqref{eq:marginal}
-\begin{equation}
-  \label{eq:app_maxent_pop_marg}
-  \p(\yxs = \yrs \| \yHa)  = \yK\, 
-  \sum_{\yNv\yRf=0}^\yNv \ypp(\yrs \|\yRf)
+  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
+\end{equation}
+The $m$ Lagrange multipliers $\yL\defd (\lambda_1,\dotsc,\lambda_m)$ must
+satisfy the $m$ constraint equations
+\begin{equation}
+  \label{eq:app_maxent_pop_constraints}
+  c_k = \expe{g_k \|\yHa}\equiv
+  \frac{1}{\yK(\yL)}\sum_{\yrs} \sum_{\yRf} g_k(\yrs) \ypp(\yrs \|\yRf)
 \,
   \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
-  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
-\end{equation}
-where $\yK$ is a normalization constant and
-$\yL\T=(\varLambda_1,\dotsc,\varLambda_m)\T$ are Lagrange multipliers such
-that
-\begin{equation}
-  \label{eq:app_maxent_pop_constraints}
-\yc = \yK\sum_{n\yrs=0}^n \sum_{\yNv\yRf=0}^\yNv \yf(\yrs) \ypp(\yrs \|\yRf)
+  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
+  \qquad k=1,\dotsc,m.
+\end{equation}
+
+We obtain the distribution for the sample activity by marginalization, using
+\eqn~\eqref{eq:marginal}:
+\begin{equation}
+  \label{eq:app_maxent_pop_marg}
+  \p(\yxs = \yrs \| \yHa)  = \frac{1}{\yK(\yL)}\, 
+  \sum_{\yRf} \ypp(\yrs \|\yRf)
 \,
   \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
-  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
-\end{equation}
-Application of \me\ at the sample level, denoted by $\yHb$, gives
-\begin{equation}
-  \label{eq:app_maxent_sample}
-  \p(\yxs=\yrs \|\yHb)
-  =\yk\,
-  \binom{n}{n\yrs}
-  \exp[\yl\T \yf(\yrs)]
-\end{equation}
-where $\yk$ is a normalization constant and $\yl\T$ are Lagrange
-multipliers such that
-\begin{equation}
-  \label{eq:app_maxent_sample_constraints}
-\yc = \yk\sum_{n\yrs=0}^n  \yf(\yrs) \binom{n}{n\yrs}
-  \exp[\yl\T \yf(\yrs)].
-\end{equation}
-
-The probabilities for the sample average obtained from application at the
-population level~\eqref{eq:app_maxent_pop_marg} and at the sample level
-\eqref{eq:app_maxent_sample} should be approximately equal, by our previous
-observation about representativity~\eqref{eq:roughly_equal_nN} and also by
-the fact that they must satisfy the same expectations for $\yf$.
-
-Yet they cannot be exactly equal, because their equality would require the
-Lagrange multipliers $\yL$ and $\yl$ to satisfy the constraint equations
-\eqref{eq:app_maxent_pop_constraints},
-\eqref{eq:app_maxent_sample_constraints}, and also
-$\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$; that is, $2m+n$
-equations (one normalization is taken care of) in $2m$ unknowns. A solution can
-exist, if at all, only for very special choices of constraints functions
-$\yf$ and values $\yc$.
-
-The sample distribution obtained from \me\ at the sample level will
-therefore likely miss important features present in the one obtained at the
-population level, like additional modes or particular tail behaviour.
+  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
+\end{equation}
+
+The distributions for the sample activity,
+\eqns~\eqref{eq:app_maxent_pop_marg} and \eqref{eq:app_maxent_sample},
+obtained with the two approaches $\yHb$ and $\yHa$, are different. From the
+discussion in the previous section we expect them to be vaguely similar;
+yet they cannot be exactly equal, because their equality would require the
+$2m$ quantities $\yL$ and $\yl$ to satisfy the constraint equations
+\eqref{eq:app_maxent_pop_constraints} and
+\eqref{eq:app_maxent_sample_constraints}, and in addition also the $n$
+equations $\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$,
+$\yrs=1/n,\dotsc,1$ (one equation is taken care of by the normalization of
+the distributions). We would have a set of $2m+n$ equations in $2m$
+unknowns.
+
+Hence, \emph{the applications of \me\ at the sample level and at the
+  full-population level are inequivalent}. They lead to numerically
+different distributions for the sample activity $\yr$.
+
+%%% FIGURES %%%
+\iffalse
 \begin{figure}[!t]
@@
 \includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
-\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at the population
-  level followed by sample marginalization (blue triangles),
-  \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yNv=5000$,
-  $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$.}
-\label{fig:diff_maxent_pop_sample}
-\end{figure}%maxent_pop_or_sample.nb
-% \end{wrapfigure}
-\begin{figure}[!t]
-\centering
-\includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}
-% 
 \caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at
@@
   circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yNv=5000$,
-  $n=200$, constraints $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$,
-  $\expe{\sav{\yx\yx\yx}}=1.48\times10^{-4}$,
-  $\expe{\sav{\yx\yx\yx\yx}}=8.81\times 10^{-6}$.}
+  $n=200$, and constraints as in \eqn~\eqref{eq:constraints}.}
+\label{fig:diff_maxent_pop_sample}
+\end{figure}%maxent_pop_or_sample.nb
+%
+\begin{figure}[!t]
+\centering
+\includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}% 
+\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at
+  the population level followed by sample marginalization (blue triangles),
+  \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
+  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yNv=5000$,
+  $n=200$, and constraints as in \eqn~\eqref{eq:constraints}.}
 \label{fig:diff_maxent_pop_sample_realdata}
 \end{figure}%maxent_pop_or_sample.nb
-We show two examples of this discrepancy in
-\figs~\ref{fig:diff_maxent_pop_sample}
-and~\ref{fig:diff_maxent_pop_sample_realdata}, for $\yNv=5000$, $n=200$, and
-constraint functions of the form
-$\yf(\yxs)=(\yxs,\yxxs,\dotsc)\equiv\bigl( \yxs,
-\raisebox{0pt}[0pt][1ex]{}\binom{n\yxs}{2}/\binom{n}{2},\dotsc \bigr)$,
-equivalent to moments constraints. The constraint values used in these
-examples, reported in the figure captions, have neurobiologically realistic
-values \cite{rostamietal2016_r2017}.
-
-In the first example the constraint functions are $\expe{\yxs}$ and
-$\expe{\yxxs}$. The distribution obtained at the sample level is broader
-than the one obtained at the population level; the tails of the two
-distributions are very different.
-
-The second
-example uses two additional constraint functions $\expe{\sav{\yx\yx\yx}}$,
-$\expe{\sav{\yx\yx\yx\yx}}$. The distribution obtained at the population
-level has two modes, replaced by only one in the distribution obtained at
-the sample level; the tails are very different also in this case.
+\fi%
+\begin{figure}[!t]
+\centering
+\includegraphics[width=0.9\linewidth]{comparison3.pdf}%
+\caption{Linear and log-plots of $\p(\yxs = \yrs)$ for a sample of $n=200$
+  and constraints as in \eqn~\eqref{eq:constraints}, constructed by: (red
+  circles) \me\ at the sample level, \eqn~\eqref{eq:app_maxent_sample};
+  (blue triangles) \me\ at the population level,
+  \eqn~\eqref{eq:app_maxent_pop_marg} with $\yNv=10\,000$, followed by
+  sample marginalization; (yellow squares) \me\ at the population level
+  with unknown population size,
+  \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}, according to the
+  distribution~\eqref{eq:pdf_popsize}.}
+\label{fig:all_three}
+\end{figure}%maxent_pop_or_sample.nb\fi
+The distribution obtained at the sample level will show different
+features from the one obtained at the population level, like displaced or
+additional modes or particular tail behaviour. We show an example of this
+discrepancy in \fig~\ref{fig:all_three}, for
+$\yNv=10\,000$, $n=200$, and the two constraints
+\begin{equation}
+  \label{eq:constraints}
+  \expe{\yrs} = 0.0478,\qquad
+  \expe{\yxxs} = 0.00257,
+  % ,\quad
+  % \expe{\sav{\yr\yr\yr}} = 1.48\times 10^{-4},\quad
+  % \expe{\sav{\yr\yr\yr\yr}} = 8.81 \times 10^{-6}.
+\end{equation}
+which come from the actual recording of circa 200 neurons from macaque
+motor cortex \cite{rostamietal2016_r2017}. The distribution obtained at the
+population level (blue triangles) has a higher and displaced mode and a quite
+different behaviour for activities around $0.5$ than the distribution
+obtained at the sample level (red squares).
 
@@
 
-How should we apply the \me\ method then? on the sample or on the
-population? Which application is \enquote{maximally noncommittal}?
+In our discussion we have so far assumed the size $\yNv$ of the larger
+population to be known. This is rarely the case, however. We usually are
+uncertain about $\yNv$ and can only guess its order of magnitude. In such a
+state of knowledge $\yHd$ our ignorance about the possible value of $\yNv$
+is expressed by a probability distribution $\p(\yNN=\yNv \|\yHd)=h(\yNv)$,
+and the marginal distribution for the sample
+activity~\eqref{eq:app_maxent_pop_marg} is modified, by the law of total
+probability, to
+\begin{multline}
+  \label{eq:app_maxent_pop_marg_unknown_N}
+  \p(\yxs = \yrs \| \yHd)  =
+  \sum_{\yNv} \p(\yxs = \yrs \| \yNN=\yNv,\yHd) \,
+  \p(\yNN=\yNv \|\yHd)
+  ={}\\[-\jot]
+  \sum_\yNv \biggl\{\frac{1}{\yK(\yL_\yNv)}\, 
+  \sum_{\yRf} \ypp_\yNv(\yrs \|\yRf)
+\,
+  \tbinom{\yNv}{\yNv \yRf}\,\exp\Bigl[{\yL_\yNv}\T
+  \sum_{\yrs} \yf(\yrs)\ypp_\yNv(\yrs\|\yRf)\Bigr]\biggr\}
+  \,h(\yNv),
+\end{multline}
+where the Lagrange multipliers $\yL_{\yNv}$ and the summation range for
+$\yRf$ depend on $\yNv$.
+
+As a proof of concept, \fig~\ref{fig:all_three} also shows such a distribution
+(yellow circles) for the same constraints as above, and a probability
+distribution for $\yNv$ inspired by Jeffreys
+\cite[\sect~4.8]{jeffreys1939_r2003}:
+\begin{equation}
+  \label{eq:pdf_popsize}
+  h(\yNv) \propto 1/\yNv, \qquad
+  \yNv \in\set{1\,000,\; 2\,000,\; \dotsc,\;10\,000}.
+\end{equation}
 
@@
 
-The question that closed the preceding section cannot receive a categorical
-answer. An optimal answer can only be given case by case, depending on the
-computational power available, on which inferences we are trying to make,
-on which assumptions we need or want to make, and those we wish to avoid.
-
-\iffalse We want to stress, in fact, the
-extreme importance and priority of the following two questions:
+The purpose of the present note was to simply to point out and show, in a
+simple set-up, that the \me\ method can be applied to recorded neuronal
+data in a way that accounts for the larger population from which the data
+are sampled,
+\eqns~\eqref{eq:app_maxent_pop}--\eqref{eq:app_maxent_pop_marg}. This
+application leads to results that differ from the standard application
+focused on the sample only,
+\eqns~\eqref{eq:app_maxent_sample}--\eqref{eq:app_maxent_sample_constraints}.
+We gave a numerical example of this difference. We have also shown how to
+extend the new application when the size of the larger population is
+unknown, \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}.
+
+The latter formula, in particular, shows that the standard way of applying
+\me\ (without \enquote{hidden neurons}) implicitly assumes that \emph{no}
+larger population exists beyond the recorded sample of neurons. One could
+in fact object to the application at the population level, and say that the
+traditional way of applying \me, \eqn~\eqref{eq:app_maxent_sample}, yields
+different results because it does not make assumptions about the size
+$\yNv$ of a possibly existing larger population. Such a state of
+uncertainty, however, is correctly formalized according to the laws of
+probability by introducing a probability distribution for $\yNv$, and is
+expressed by \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}. This expression
+cannot generally be equal to~\eqref{eq:app_maxent_sample} unless the
+distribution for $\yNv$ gives unit probability to $\yNv=n$; that is,
+unless the sample \emph{is} the full population, and no larger population
+exists or the recorded neurons are completely isolated from it.
+
+The latter notion is unreasonable. It amounts to say that the neurons we
+distinguished and tracked with our recording device are a very special,
+isolated set among all those that could have been recorded. It is for this
+reason that we find the \me\ application at the population level
+preferable. Physical models of neuronal networks usually include some sort
+of external input, mimicking an embedding in a larger network. The \me\
+distribution obtained at the population level may reveal features in a
+data set that were unnoticed by the standard \me\ model.
+
+The difference between the applications at the sample and at the population
+levels appears also in statistical mechanics. If a system is statistically
+described by a \me\ Gibbs state, its subsystems cannot be described by a
+Gibbs state \cite{maesetal1999}. A somewhat similar situation also appears
+in the statistical description of the final state of a non-equilibrium
+process starting and ending in two equilibrium states: we can describe our
+knowledge about the final state either by (1) a Gibbs distribution,
+calculated from the final equilibrium macrovariables, or (2) by the
+distribution obtained from the Liouville evolution of the Gibbs
+distribution assigned to the initial state. The two distributions differ
+(even though the final \emph{physical} state is obviously exactly the same
+\cite[\sect~4]{jaynes1985d_r1993}), and the second allows us to make
+sharper predictions about the final physical state thanks to our knowledge
+of its preceding dynamics. In this example, though, both distributions are
+usually immensely sharp and practically lead to the same predictions. In
+the neuroscientific applications considered in this note the difference in
+predictions of the sample vs full-population applications may be very
+relevant instead.
+
+The idea of the new application leads in fact to more questions. For
+instance:
 \begin{itemize}
-\item Of which quantities do we need to quantify the uncertainty? and why?
-\item What are the assumptions we need or want to make to answer the
-  question above?
-\end{itemize}\fi
-
-\iffalse The first purpose of this note is indeed to remind ourselves that
-probability models require careful scrutiny, because they can rest on
-hidden assumptions that we do not want to make or that contradict others we
-are making, and we may not be aware of this. The dilemma generated by the
-\me\ method and the assumption of representative sampling is an example.
-\fi
-
-The tricky point is this. The \me\ application at the population level and
-the application at the sample level give different results; they are two
-different statistical models. The former model clearly assumes, by
-construction, the existence of a larger population from which the sample is
-taken. What does the latter model assume in this respect? is it
-\enquote{unassuming}, as often claimed in the literature? or does it
-actually assume that \emph{no} larger population exists? In the latter case
-it would not be correct to use this model in our problem.
-
-A perfunctory intuitive reasoning seems insufficient for clarifying this
-point. Let's express it in the language of the probability calculus.
-Suppose we do not know whether the sample is really part of a larger
-population: we do not know whether $\yNv=n$ or how large $\yNv$ is otherwise.
-Call this state of ignorance $\yHd$. In the probability calculus this
-ignorance about $\yNv$ is expressed by assigning a probability distribution
-$\p(\yNv \| \yHd)$ that vanishes if $\yNv<n$, since we know that $\yNv\ge n$; see
-Good \cite{good1965,good1967b} and Rissanen \cite{rissanen1983} for
-examples of such distributions over the integers. Maintaining our
-assumption of symmetric ignorance, probability assignments that do not
-assume a specific value of $\yNv$ are then obtained via multiplication of all
-$\yNv$-dependent probabilities by $\p(\yNv \| \yHd)$ and subsequent
-marginalization over $\yNv$. Technically speaking, $\yNv$ becomes a
-\emph{nuisance parameter}
-\cite{jaynes1994_r2003,lindley1965b_r2008,bernardoetal1994_r2000}. The
-probability obtained from \me\ at the population level,
-\eqn~\eqref{eq:app_maxent_pop_marg}, then generalizes to
-\begin{equation}
-  \label{eq:app_maxent_pop_marg_unknown_N}
-  \p(\yxs = \yrs \| \yHd)  = \sum_\yNv \biggl\{\yK_\yNv\, 
-  \sum_{\yNv\yRf=0}^\yNv \ypp_\yNv(\yrs \|\yRf)
-\,
-  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[{\yL_\yNv}\T
-  \sum_{n\yrs=0}^n \yf(\yrs)\ypp_\yNv(\yrs\|\yRf)\Bigr]\biggr\}
-  \,\p(\yNv \| \yHd),
-\end{equation}
-where $\yNv$-dependencies have been made explicit. This is a formidable
-expression. But our question, \enquote{is the usual \me\ at the sample
-  level~\eqref{eq:app_maxent_sample} unassuming with regard to the
-  existence of a larger population?}, translates now into the precise
-mathematical question: \enquote{are the
-distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
-\eqref{eq:app_maxent_sample} equal for some choice of $\p(\yNv \| \yHd)$, with
-$\p(\yNv \| \yHd)\ne 0$ for $\yNv>n$?}. We leave this mathematical problem for
-future work. Note, however, that this equality is satisfied if
-$\p(\yNv=1 \| \yHd)=1$, which means that the usual \me\ model can also be
-interpreted as assuming that \emph{no} larger population exists.
+\item Do the standard and new applications lead to different or contrasting
+  conclusions about \enquote{cooperativity}, when applied to real data
+  sets?
+\item How to extend the new application when we want to study the 
+  activities of individual neurons, of individual pairs, triplets, and so
+  on?
+\item What is the mathematical relation between the new application and
+  \me\ models with hidden neurons?
+\end{itemize}
+Owing to space limitations we cannot investigate these questions here, but
+we hope to in future work.
+
+Finally, we would like to point out the usefulness and importance of the
+probability formulae that relate our states of knowledge about a population
+and its samples, presented in \sect~\ref{sec:prob_samples}. This kind of
+formulae is essential in neuroscience, where we try to understand
+properties of extended regions from partial observations of them. The
+formulae presented here reflected a very basic, symmetric state of
+ignorance. It would be useful to explore formulae more appropriate to our
+real knowledge of the brain and of its networking properties.
 
@@
 \fi
-
-We find the \me\ model constructed at the population level very natural and
-preferable. After all, physical models of neuronal networks usually include
-some sort of external input to the neurons as well, mimicking their
-embedding in a larger network. The sample distribution given by the \me\
-model at the population level, when used as a reference distribution for
-surprise analysis, may reveal features in a dataset that were unnoticed by
-the standard \me\ model. The question remains of how to specify $\yNv$,
-though. We have tacitly intended $\yNv$ as the size of the largest
-biologically or functionally homogeneous population from which our sample
-was recorded. It could be the amount of neurons in a functional brain area,
-for example the primary visual cortex, for which $\yNv\sim 10^8$
-\cite{leubaetal1994}. For large $\yNv$ -- unfortunately we are not yet able to
-translate this \enquote{large} into a numeric order of magnitude -- the
-final distribution becomes independent of $\yNv$, and continuous
-approximations become available.
 
@@
 
-The possibility of using two different distributions is not a physical
-contradiction. Similar situations arise in statistical mechanics. It is
-known that if a system is described by a \me\ Gibbs state, its subsystems
-need not be \cite{maesetal1999}. A dilemma quite similar to ours also
-appears in the statistical description of the final state of a
-non-equilibrium process starting and ending in two equilibrium states: we
-can describe our knowledge about the final state either by a Gibbs
-distribution, or by the distribution obtained from the Liouville evolution
-of the Gibbs distribution assigned to the initial state. The two
-descriptions differ -- even though the final \emph{physical} state is
-obviously exactly the same \cite[\sect~4]{jaynes1985d_r1993}. The two
-descriptions differs because in one case we can make sharper predictions
-about the state thanks to our knowledge of its preceding dynamics. In this
-example, though, both distributions are usually immensely sharp and
-practically lead to the same predictions. In the neuroscientific
-applications considered in this note the difference in predictions may be
-relevant instead.
-
-Our analysis touched only constraints of the sample average,
-$\expeb{f(\yxs)}$. The corresponding models are usually called
+\iffalse Our analysis touched only constraints of the sample average,
+$\expeb{g(\yxs)}$. The corresponding models are usually called
 \enquote{homogeneous} in the literature. Purely \enquote{inhomogeneous}
@@
 \cite{tkaciketal2014b,shimazakietal2015}.
+\fi
 

### 2018-05-18 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-17T18:47:41+0200
+%% Last-Updated: 2018-05-18T10:23:24+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\newcommand*{\pdftitle}{Representative samples\\ and maximum-entropy distributions:\\ a dilemma}
+\newcommand*{\pdftitle}{Maximum-entropy and neuronal samples:\\an alternative approach}
 
@@
 \usepackage[usenames]{xcolor}
-\definecolor{myblue}{RGB}{51,34,136}
+\definecolor{mybluishpurple}{RGB}{51,34,136}
+\definecolor{myblue}{RGB}{136,204,238}
+\definecolor{mybluishgreen}{RGB}{68,170,153}
 \definecolor{mygreen}{RGB}{17,119,51}
-\definecolor{myred}{RGB}{136,34,85}
-\definecolor{myyellow}{RGB}{153,153,51}
-\definecolor{mylightyellow}{RGB}{221,204,119}
+\definecolor{mygreenishyellow}{RGB}{153,153,51}
+\definecolor{myyellow}{RGB}{221,204,119}
+\definecolor{myred}{RGB}{204,102,119}
+\definecolor{mypurplishred}{RGB}{136,34,85}
+\definecolor{myreddishpurple}{RGB}{170,68,153}
+\definecolor{mygrey}{RGB}{221,221,221}
 
@@
 
-\usepackage{hyperref}
+\PassOptionsToPackage{hyphens}{url}\usepackage[hypertexnames=false]{hyperref}
 \usepackage[depth=4]{bookmark}
 \hypersetup{colorlinks=true,bookmarksnumbered,pdfborder={0 0 0.25},citebordercolor={0.2 0.1333 0.5333},%bluish
-citecolor=myblue,linkbordercolor={0.0667 0.4667 0.2},%greenish
-linkcolor=myred,urlbordercolor={0.5333 0.1333 0.3333},%reddish
+citecolor=mybluishpurple,linkbordercolor={0.0667 0.4667 0.2},%greenish
+linkcolor=mypurplishred,urlbordercolor={0.5333 0.1333 0.3333},%reddish
 urlcolor=mygreen,breaklinks=true,pdftitle={\pdftitle},pdfauthor={\pdfauthor}}
-
 % \usepackage[vertfit=local]{breakurl}% only for arXiv
@@
 
-\usepackage{wrapfig}
+%\usepackage{wrapfig}
 \usepackage{datetime2}
@@
 %\newcommand*{\vs}{{vs.}}
-\newcommand*{\etc}{{etc.}}
-\newcommand*{\ie}{{i.e.}}
-\newcommand*{\ca}{{c.}}
+%\newcommand*{\etc}{{etc.}}
+%\newcommand*{\ie}{{i.e.}}
+%\newcommand*{\ca}{{c.}}
 %\newcommand*{\Ie}{{I.e.}}
 %\newcommand*{\Eg}{{E.g.}}
-\newcommand*{\eg}{{e.g.}}
+%\newcommand*{\eg}{{e.g.}}
 %\newcommand*{\viz}{{viz}}
@@
 \begin{abstract}
-  \mynote{to be rewritten} This note shows that the \me\ method can be
-  applied to a representative sample from a neuronal population along two
-  different routes: (1) apply to the sample; or (2) apply to the population
-  and marginalize to the sample. These two routes give inequivalent
-  results. Which route should be chosen? Some arguments are presented in
-  favour of the second. The note also touches upon probability formulae of
-  representative sampling and discusses their possible meanings, a
-  discussion that may be useful for sampling problems in neuroscience.
+  This note shows that the \me\ method can be applied to a sample of
+  neuronal activity along two different routes: (1) apply to the sample; or
+  (2) apply to a larger, unsampled neuronal population from which the
+  sample is drawn, and then marginalize to the sample. The second route can
+  be generalized to the case where the size of the larger population is
+  unknown. These two routes give inequivalent results. Which route should
+  be chosen? Some arguments are presented in favour of the second. This
+  note also presents and discusses probability formulae that relate our
+  states of knowledge about a population and its samples, and that may be
+  useful for sampling problems in neuroscience.
   % The abstract must be limited to one paragraph.
@@
 %\section{Implicit assumptions in the maximum-entropy method}
-\section{Introduction: \me\ and sampling in neuroscience}
+\section{Introduction: \me\ and recordings of neuronal activity}
 \label{sec:intro}
@@
 sampled from a particular brain area. What are we to do with such data?
-Gerstein, Perkel, \amp\ Dayhoff \cite{gersteinetal1985} posed this question
-very tersely (our emphasis):
+Gerstein, Perkel, Dayhoff \cite{gersteinetal1985} posed this question very
+tersely (our emphasis):
 \begin{quote}
@@
 \end{quote}
-These questions have a long history, of course; see \eg\ the 1966 review by
-Moore \etal\ \cite{mooreetal1966}. % This scenario is a concrete possibility
+These questions have a long history, of course; see for instance the 1966
+review by Moore \etal\
+\cite{mooreetal1966}. % This scenario is a concrete possibility
 % thanks to recent electrophysiological techniques \cite{berenyietal2014}.
@@
   \item \ldots
-  \item the time averages of the activity, of the pair-activity, \etc\ up
+  \item the time averages of the activity, of the pair-activity, and so on, up
     to the $k$-activity.
@@
 
-\iffalse In the final discussion we present some arguments in favour of one choice
-in our dilemma, but we do not mean that to be our final answer. This note,
-and its dilemma,  also have the more general purpose of reminding our
-readers the importance of asking \emph{what is the question we're trying to
-  answer with probability theory?} and \emph{which assumptions do we want
-  or need to make?}
-
-Our mathematical analysis pertains to neurons modelled as binary units at
-one specific instant or short window in time. It is possible to similarly
-discuss multi-state neuron models and population dynamics; but for
-simplicity neurons are here assumed to be in a fixed, active \enquote{$1$}
-or inactive \enquote{$0$} state; and evolution, change, time correlations,
-and similar concepts do not concern us. We consider \me\ models that
-constrain various kinds of sample or population averages; these models are
-often called \enquote{homogeneous}. The final discussion touches upon
-\enquote{inhomogeneous} models as well.
-\fi
-
 The notation and terminology in this note follow ISO and ANSI standards
@@
 $\yRf(t) \defd \sum_{\iota} \yRv_{\iota}(t)/\yNv$ is the activity at that
-bin averaged over the larger population, \etc.
+bin averaged over the larger population, and so on.
 
@@
 
-\iffalse We have a population of $\yNv$ binary neurons. We assume that they can be
-distinguished, by their spike shapes for example; but other details, like
-their locations, are unknown. The neurons have a joint state
-$(\yXv_1,\dotsc,\yXv_\yNv) \defs \yX$ having fixed but unknown binary values
-$(\yRv_1,\dotsc,\yRv_\yNv)\defs \yR \in \set{0,1}^\yNv$; the \enquote{$\yXv$} denote the
-variables, and \enquote{$\yRv$} their values. A particular sample of $n$ neurons
-from this population has joint state $(\yxv_1, \dotsc, \yxv_n) \defs \yx$ having
-fixed binary values $(\yrv_1, \dotsc, \yrv_n)\defs\yr \in \set{0,1}^n$. We will
-consider various averages of the population and the sample. For this
-purpose we introduce a population-averaging operator $\sav{\dotv}$ defined
-by
-\begin{equation}
-  \begin{gathered}
-    \av{\yX} \defd \tfrac{1}{\yNv} (\yXv_1 + \yXv_2 + \dotsb + \yXv_\yNv),
-    % & \sav{\yx} &\defd \tfrac{1}{n} (x_1 + x_2 + \dotsb + x_n),
-   \qquad% \\[2\jot]
-    \av{\yX \yX} \defd \tbinom{\yNv}{2}^{-1} (\yXv_1 \yXv_2 + \yXv_1 \yXv_3 + \dotsb +
-    \yXv_{\yNv-1} \yXv_\yNv),
-    % & \sav{\yx \yx} &\defd\tbinom{n}{2}^{-1} (x_1 x_2 + x_1 x_3 + \dotsb
-    % + x_{n-1} x_n),
-    \\[2\jot]
-    \av{\yX\yX\yX} \defd \tbinom{\yNv}{3}^{-1} (\yXv_1 \yXv_2 \yXv_3 +% \yXv_1 \yXv_2 \yXv_4
-    \dotsb + \yXv_{\yNv-2} \yXv_{\yNv-1}\yXv_\yNv),
-    \qquad\text{and so on.}
-    % & \sav{\yx\yx\yx} &\defd \tbinom{n}{3}^{-1} (x_1 x_2 x_3
-    % + %x_1 x_2 x_4
-    % \dotsb + x_{n-2} x_{n-1} x_n),
-  \end{gathered}
-\end{equation}
-These formulae say that $\yXf$ is the fraction of active neurons, $\yXXf$
-the fraction of simultaneously active pairs out of all $\binom{\yNv}{2}$
-pairs, $\av{\yX\yX\yX}$ the fraction of simultaneously active triplets, and
-so on. Products of states like $\yXv_i \dotsm \yXv_j$ also have values in
-$\set{0,1}$; from this we can combinatorially  prove that
-\begin{equation}
-  \label{eq:products_intermsof_average}
-  \av{\underbrace{\yX\dotsm\yX}_{\text{$m$ factors}}}
-  = \binom{\yNv}{m}^{-1}\binom{\yNv\yXf}{m}.
-\end{equation}
-Analogous formulae  hold for quantities like $\yx$, $\yR$, $\yr$.
-
-\begin{simplenote}
-  \mynote{Remove this} A sample can be \enquote{random} in two different
-  ways. Consider the population of neurons named $\set{1,2,3}$ for example.
-  The sample $\set{1,3}$ is random because neurons $1$ and $3$ were chosen
-  according to an unknown process. The sample
-  $\set{\textrm{X}_1, \textrm{X}_2}$ is random because we do not know the
-  identity of neurons $\textrm{X}_1$ and $\textrm{X}_2$. Given the
-  probability for the population, the probabilities for the states of these
-  two samples are different: the former is obtained from marginalization,
-  the latter by symmetrization and marginalization.
-\end{simplenote}
-\fi
-
-If $\yHc$ denotes our state of knowledge, \ie\ the evidence and assumptions
-backing our probability assignments, our uncertainty about the full
+If $\yHc$ denotes our state of knowledge -- the evidence and assumptions
+backing our probability assignments -- our uncertainty about the full
 activity of the larger population is expressed by the joint probability
@@
 where the second step uses \eqn~\eqref{eq:subpop_average}. The last
-expression shows that the expectation of the function $g$, whose argument
-is the sample activity $\yrs$, is equal to the expectation of the function
-$g^*$, whose argument is the full-population activity $\yRf$, defined as
+expression shows that the expectation of the function $g(\yrs)$ is equal to
+the expectation of the function
 $g^*(\yRf) \defd \sum_{\yrs} g(\yrs)\,\ypp(\yrs\|\yRf)$.
@@
 realized in only one way (all neurons inactive), $\yrs=1/n$ can be realized
-in $n$ ways (one active neuron out of $n$), \etc\ This term is analogous to
-the \enquote{density of states} in front of a Boltzmann factor in
+in $n$ ways (one active neuron out of $n$), and so on. This term is analogous to
+the \enquote{density of states} in front of the Boltzmann factor in
 statistical mechanics \cite[\chap~16]{callen1960_r1985}. The $m$ Lagrange
@@
 \caption{Linear and log-plots of $\p(\yxs = \yrs)$ for a sample of $n=200$
-  and constraints as in \eqn~\eqref{eq:constraints}, constructed by: (red
-  circles) \me\ at the sample level, \eqn~\eqref{eq:app_maxent_sample};
-  (blue triangles) \me\ at the population level,
-  \eqn~\eqref{eq:app_maxent_pop_marg} with $\yNv=10\,000$, followed by
-  sample marginalization; (yellow squares) \me\ at the population level
-  with unknown population size,
+  and constraints as in \eqn~\eqref{eq:constraints}, constructed by:
+  \textcolor{myred}{\textbf{red squares:}} \me\ at the sample level,
+  \eqn~\eqref{eq:app_maxent_sample};
+  \textcolor{mybluishpurple}{\textbf{blue triangles:}} \me\ at the
+  population level, \eqn~\eqref{eq:app_maxent_pop_marg} with
+  $\yNv=10\,000$, followed by sample marginalization;
+  \textcolor{myyellow}{\textbf{yellow circles:}} \me\ at the population
+  level with unknown population size,
   \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}, according to the
-  distribution~\eqref{eq:pdf_popsize}.}
+  distribution~\eqref{eq:pdf_popsize} for the population.}
 \label{fig:all_three}
@@
 
-\iffalse
-The former application must therefore implicitly assume that no
-larger population exist, or that the neurons of the sample are isolated
-from their environment.
-
-The \me\ application at the population level gives in fact, by
-construction, a distribution with lower Shannon relative entropy than the
-one given by the application at the sample level: both satisfy the same
-constraints, but only the latter is obtained via direct maximization of the
-entropy. The lower entropy of the distribution obtained from application at
-the population level is clearly visible from its higher sharpness in
-\fig~\ref{fig:diff_maxent_pop_sample}. Such distribution therefore
-excludes, probabilistically speaking, more sample states $\yx$ than the
-distribution obtained at the sample level does. A lower entropy signals the
-presence of additional assumptions or constraints. In this case the
-assumption is that the sample is part of a larger population, and our
-ignorance about the sample should be representative of our ignorance about
-a larger population. This assumption affects our predictions.
-\fi
-
-\iffalse
-The quandary we found in the present note has at least two bright sides.
-The first is that the two different \me\ applications may give numerically
-similar results in some cases. One could thus be used as an approximation
-of the other, depending on the circumstances -- \eg\ if probability tails
-are unimportant. See \fig~\ref{fig:diff_maxent_pop_sample_realdata} for an
-example.
-\begin{figure}[!t]
-\centering
-\includegraphics[width=0.95\columnwidth]{different_maxent_pop_sample_200_realdata_both.pdf}
-% 
-\caption{Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
-  \me\ at the population level followed by sample marginalization (blue
-  triangles), and at the sample level (red circles). The constraints are
-  the same in either case: $\expe{\yxs \|\yH}=0.045$,
-  $\expe{\yxxs \|\yH}=0.0025$, and $\yNv=5000$, $n=200$.}
-\label{fig:diff_maxent_pop_sample_realdata}
-\end{figure}%maxent_pop_or_sample.nb
-The second is that \fi
-
-\iffalse Our analysis touched only constraints of the sample average,
-$\expeb{g(\yxs)}$. The corresponding models are usually called
-\enquote{homogeneous} in the literature. Purely \enquote{inhomogeneous}
-models have also been used
-\cite{schneidmanetal2006,shlensetal2006,roudietal2009b}, in which
-expectations for individual neurons or groups of neurons are constrained,
-for example $\expe{\yxv_2}$ or $\expe{\yxv_1x_8x_9}$. A short computation shows
-that the \me\ method with this kind of constraints gives the same result
-whether applied at the sample or at the population level: the states of any
-unconstrained neurons marginalize out. This is understandable: expressing
-different uncertainties about, say, neurons $2$ and $5$ we are breaking the
-symmetry of our uncertainty, which thus cannot be representative of other
-neurons in the sample or in the population. Inhomogeneous models, however,
-require enormous computational power for large sample sizes; homogeneous
-models therefore retain their importance. Our analysis and dilemma also
-persist for hybrid homogeneous-inhomogeneous models
-\cite{tkaciketal2014b,shimazakietal2015}.
-\fi
-
 %\mynote{Importance of the formulae relating sample and population.}
@@
 
-\iffalse
-\newrefsection
-\bigskip
-{\centering\mynote{*** From here down old text to be discarded ***}\par}
-\section{***}
-
-\begin{align}
-  \expe{\yxv_{i_1} \dotsm \yxv_{i_m} \| \yH}
-&=
-\expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
-=
-\expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH},
-\\
-\intertext{and has an explicit expression in terms of $Q$:}
-  \begin{split}
-\expe{\yxv_{i_1} \dotsm \yxv_{i_m} \| \yH}
-&=
-\binom{\yNv}{m}^{-1}
-%\frac{(N-m)!}{N!}
-\sum_{\yNv A=0}^\yNv %m!\, 
-\binom{\yNv A}{m}\, Q(A)
-\\
-&\equiv
-%\frac{(N-m)!}{N!}
-\sum_{\yNv A=0}^\yNv %m!\, 
-\binom{\yNv-m}{\yNv A-m}
-\binom{\yNv}{\yNv A}^{-1} Q(A).
-\end{split}
-\end{align}
-\end{subequations}
-
-What do we mean by saying that we can learn something about the population
-by observing the sample? At the very least we mean that our uncertainty
-about the rest of the population can change upon observing the sample. That
-is, the probability for the state $\yy$ of the rest of the population is
-conditionally dependent on the sample's values:
-\begin{equation}
-  \label{eq:conditionally_dependent_sample}
-  \p(\yy =\ys \| \yx=\yr,\yH) \ne \p(\yy =\ys \|\yH)
-  \qquad\text{for some $\ys,\yr$},
-\end{equation}
-This also means that the probability for the population cannot be
-factorized into that of the sample times that of the complement.
-
-What do we mean when we say that the sample is representative of the
-population? It means that we expect some collective properties of the
-sample, like the fraction of active neurons, or the fraction of
-simultaneously active pairs, to be roughly equal to those of the full
-population.
-
-***
-
-For example, $\yX$ can
-represent the state of a population at a particular time. We call the
-neurons \enquote{units} to lend some generality to our discussion. We shall
-make statements about the whole population of $\yNv$ units and about a
-subpopulation of $n$ units; the word \enquote{population} will always refer
-to the \emph{whole} population. The subpopulation states and their values
-are denoted by lowercase letters: $(\yxv_1, \dotsc, \yxv_n)\equiv\yx$ and
-$(\yrv_1, \dotsc, \yrv_n)\equiv\yr$; but note that $\yxv_i \equiv \yXv_{j_i}$ and
-$\yrv_i \equiv \yRv_{j_i}$ for some distinct $j_1,\dotsc,j_n$. We shall also make
-statements about the population-averaged state, or \emph{population
-  average}:
-\begin{align}
-  \label{noeq:population-av}
-   \yXf &\defd (\yXv_1 + \dotsb + \yXv_\yNv)/\yNv,
-\\
-\intertext{and the subpopulation-averaged state, or \emph{subpopulation average}:}
-\label{noeq:subpopulation-av}
-\yxs &\defd (\yxv_1 + \dotsc + \yxv_n)/n.
-\end{align}
-The quantities $\yNv\yXf$ and $n\yxs$ represent the total number of active
-units in the population and the subpopulation. Quantities like $\yRf$ and $\yrs$
-are defined analogously. The averaging operators $\sav{\dotv}$ and
-$\av{\dotv}$ are also extended to averages of $\tbinom{n}{m}$ or
-$\tbinom{\yNv}{m}$ products of $m$ states; \eg,
-\begin{align}
-\av{\yX \yX} &\defd
-\tbinom{\yNv}{2}^{-1} (\yXv_1 \yXv_2 + \yXv_1 \yXv_3  + \dotsb +  \yXv_{\yNv-1} \yXv_\yNv),
-\\[2\jot]
-\sav{\yx\yx\yx} &\defd
-\tbinom{n}{3}^{-1} (\yxv_1 \yxv_2 \yxv_3 + \yxv_1 \yxv_2 \yxv_4 + \dotsb + \yxv_{n-2} \yxv_{n-1} \yxv_n),
-\end{align}
-and so on.
-
-\subsection{Assumptions}
-\label{nosec:assumptions}
-
-Our uncertainty about the population state is represented by the joint
-probability distribution of the individual states, from which we can
-derive all other probabilities of interest. We denote it by
-\begin{equation}
-  \label{noeq:joint_plaus}
-  \p(\yXv_1=\yRv_1, \yXv_2=\yRv_2, \dotsc, \yXv_\yNv=\yRv_\yNv \| \yH) \quad\text{or}\quad
-\p(\yX =\yR \| \yH).
-\end{equation}
-Such probability is conditional on our state of knowledge, \ie\ the
-evidence and assumptions backing our probability assignments, denoted by
-the proposition $\yH$.
-
-In the present discussion, $\yH$ is a state of knowledge that leads to two
-specific properties in our probability assignments:
-
-\medskip
-\begin{enumerate}%[wide,label=$\yH$\arabic*.]
-\item \emph{Permutation symmetry}, expressed as the invariance of the
-  joint distribution~\eqref{noeq:joint_plaus} under arbitrary permutations of
-  the units's labels:
-\begin{multline}
-  \label{noeq:homog_math}
-  \p(\yXv_1=\yRv_1, \yXv_2=\yRv_2, \dotsc, \yXv_\yNv=\yRv_\yNv \| \yH) ={}
-\\ 
-\shoveright{\p(\yXv_1=\yRv_{\pi(1)}, \yXv_2=\yRv_{\pi(2)}, \dotsc, \yXv_\yNv=\yRv_{\pi(\yNv)} \| \yH)}
-\\
-\text{for any permutation $\pi$}.
-\end{multline}
-This property can reflect two very different states of knowledge: physical
-homogeneity of the population, or symmetry in our ignorance about the population.
-This property is called \emph{finite exchangeability} in the Bayesian
-literature and its basis, consequences, and alternatives to it are discussed
-in \sect~***%\ref{nosec:stoch_symm}.
-
-\medskip
-
-\item The population average $\yXf$ has a particular distribution $Q$:
-\begin{equation}
-  \label{noeq:pop_average_Q}
-  \p(\yXf = A \| \yH)
-% \equiv \pr(x_1+ \dotsb + x_N = N A \| \varEta) 
-=  Q(A),
-\qquad
-A \in \set*{0,\tfrac{1}{\yNv},\tfrac{2}{\yNv},\dotsc,1}.
-\end{equation}
-For the moment we are not concerned about the specific form of $Q$ and
-about how it was assigned: it could, \eg, arise from maximum-entropy
-arguments
-\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,aczeletal1975,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013}
-used with data on the population.
-\end{enumerate}
-
-\subsection{Formulae}
-%\section{\ding{41} {\itshape The mathematical relations} \ding{41}}
-%\section*{3.\hspace{\stretch{1}}The mathematical relations\hspace{\stretch{1}}\mbox{}}
-%\addcontentsline{toc}{section}{The mathematical relations}
-\label{nosec:main_formulae}
-
-The state of knowledge $\yH$ has the following six (not independent) main
-consequences for our probability assignments:
-
-\begin{enumerate}%[wide,label=\textbf{\emph{\Roman*}}.]
-  \medskip\item\label{noitem:tot_plaus}Probability for the population state:
-\begin{equation}
-  \label{noeq:joint_plaus_N_homog}
-  \p(\yX = \yR \| H) = \binom{\yNv}{\yNv\yRf}^{-1} Q(\yRf).
-\end{equation}
-
-\medskip
-\item\label{noitem:marginal}Probability for the state $\yx$ of any subpopulation
-  of $n$ units:
-\begin{equation}
-  \label{noeq:marginal}
-    \p(\yx = \yr \| \yH) =
-\sum_{\yNv A =0}^{\yNv} 
-\binom{\yNv-n}{\yNv A-n\yrs}\binom{\yNv}{\yNv A}^{-1}\,
-Q(A).
-\end{equation}
-Note that the only summands contributing to this sum are those for which
-$n\yrs \le NA \le \yNv$; the others are zero because by definition
-$\tbinom{M}{y}=0$ if $y<0$. This remark applies to all the sums of this
-kind in the rest of this Note.
-
-\medskip
-\item\label{noitem:conditional}Probability for the subpopulation
-  state conditional on a population state:
-\begin{equation}
-  \label{noeq:conditional_total_state}
-  \p(\yx = \yr \| \yX = \yR, \yH)
-=
-\binom{\yNv-n}{\yNv \yRf-n \yrs}.
-\end{equation}
-
-\medskip
-\item\label{noitem:marginal_average}Probability for the
-  subpopulation average $\yxs$:
-\begin{multline}
-  \label{noeq:subpop_average}
-  \p(\yxs = a  \| \yH) = \binom{n}{n a}
- \sum_{\yNv A=0}^{\yNv}
-\binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1}\,
-Q(A),
-\\
-a \in \set*{0,\tfrac{1}{n},\tfrac{2}{n},\dotsc,1}.
-\end{multline}
-
-\medskip
-\item\label{noitem:conditional_average}Probability for the subpopulation
-  average conditional on the population average:
-\begin{equation}
-  \label{noeq:conditional_prob}
-  \p(\yxs = a \| \yXf = A, \yH)
-=
-\binom{n}{n a}\binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1}.
-\end{equation}
-
-\medskip
-\item\label{noitem:moments}The product of the states of any $m$ distinct
-  units from a given subpopulation,
-  \begin{equation*}
-    \yxv_{i_1} \yxv_{i_2} \dotsm \yxv_{i_m},
-    \qquad 1\le i_1 < i_2 < \dotsb < i_m \le n
-  \end{equation*}
-  has an expectation equal to that of the subpopulation average of such
-  products, is independent of the subpopulation size $n$:
-\begin{subequations}
-\label{noeq:expe_products}
-\begin{align}
-  \expe{\yxv_{i_1} \dotsm \yxv_{i_m} \| \yH}
-&=
-\expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
-=
-\expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH},
-\\
-\intertext{and has an explicit expression in terms of $Q$:}
-  \begin{split}
-\expe{\yxv_{i_1} \dotsm \yxv_{i_m} \| \yH}
-&=
-\binom{\yNv}{m}^{-1}
-%\frac{(N-m)!}{N!}
-\sum_{\yNv A=0}^\yNv %m!\, 
-\binom{\yNv A}{m}\, Q(A)
-\\
-&\equiv
-%\frac{(N-m)!}{N!}
-\sum_{\yNv A=0}^\yNv %m!\, 
-\binom{\yNv-m}{\yNv A-m}
-\binom{\yNv}{\yNv A}^{-1} Q(A).
-\end{split}
-\end{align}
-\end{subequations}
-\end{enumerate}
-
-\medskip
-
-A useful relation connects the expectation of a
-product~\eqref{noeq:expe_products} and the $m$th \emph{factorial moment}
-\citep{potts1953} of the probability distributions for the averages. The
-$m$th factorial moment of the subpopulation average $\yxs$ is defined by
-\begin{equation}
-  \label{noeq:fact_mom_defin}
-\vphantom{\underbrace{\yxs}_{m}}\expeb*{
-\smash{\underbrace{n\yxs\,(n\yxs-1)\dotsm \bigl(n\yxs-(m-1)\bigr)}_{\text{$m$ factors}}}
-\| \yH}
-\equiv \expeb*{\frac{(n\yxs)!}{(n\yxs-m)!} \| \yH},
-\end{equation}
-an analogous definition holding for $\yXf$. We have that
-\begin{equation}
-  \label{noeq:eq_factmom_prod}
-\expe{\yxv_{i_1} \dotsm \yxv_{i_m} \| \yH}
-=
-  \tfrac{(n-m)!}{n!}\expeb*{\frac{(n\yxs)!}{(n\yxs-m)!} \| \yH}
-=
-  \tfrac{(\yNv-m)!}{\yNv!}\expeb*{\frac{(\yNv\yXf)!}{(\yNv\yXf-m)!} \| \yH}.
-\end{equation}
-
-As a consequence of the above relation, the first three moments of the
-probability distributions $\p(\yxs=a\|\yH)$ and $\p(\yXf=A\|\yH)$,
-are related by
-\begin{subequations}\label{noeq:moments}
-  \begin{align}
-    \expe{\yxs \| \yH} &= \expe{\yXf \| \yH},
-    \\[2\jot]
-    \begin{split}
-    \expe{\yxs^2 \| \yH} &= \expe{\yXf \| \yH}
-                                  \frac{\yNv-n}{(\yNv-1)n}
- +{}
-\\&\quad\quad \expe{\yXf^2 \| \yH}\frac{\yNv\,(n-1)}{(\yNv-1)n},
-\end{split}
-    \\[2\jot]
-    \begin{split}
-    \expe{\yxs^3 \| \yH} &=
-                              \expe{\yXf \| \yH}\frac{(\yNv-n)(\yNv-2n)}{(\yNv-1)(\yNv-2)n^2}+{}
-                                    \\
-                                    &\quad\quad\expe{\yXf^2 \| \yH}
-                                    \frac{3N\,(\yNv-n)(n-1)}{(\yNv-1)(\yNv-2)n^2} +{}
-                                    \\
-                                    &\quad\quad\quad\expe{\yXf^3 \| \yH}
-                                    \frac{\yNv^2\,(n-1)(n-2)}{(\yNv-1)(\yNv-2)n^2}.
-                                  \end{split}
-                                    % \\[2\jot]
-                                    % \expe{\yxs^4 \| \varEta} &=
-                                    % \!\begin{aligned}[t] &\expe{\yxf
-                                    %   \| \varEta}
-                                    %   \frac{(N-n)(N^2+N-6Nn+6n^2)}{(N-1)(N-2)(N-3)n^3}+{}
-                                    %   \\%[1.5\jot]
-                                    %   &\quad\expe{\yxf^2 \| \varEta}
-                                    %   \frac{(7N-11n+1)(N-n)(n-1)N}{(N-1)(N-2)(N-3)n^3}+{}
-                                    %   \\
-                                    %   &\quad\expe{\yxf^3 \| \varEta}
-                                    %   \frac{6(n-1)(n-2)(N-n)N^2}{(N-1)(N-2)(N-3)n^3}
-                                    %   +{}
-                                    %   \\
-                                    %   &\quad\expe{\yxf^4 \| \varEta}
-                                    %   \frac{(n-1)(n-2)(n-3)N^3}{(N-1)(N-2)(N-3)n^3}.
-                                    % \end{aligned}
-  \end{align}
-\end{subequations}
-Relations for higher moments can be obtained recursively from
-\eqn~\eqref{noeq:eq_factmom_prod}. In general, this means that the two sets of
-first $m$ moments are related by a homogeneous linear transformation,
-\begin{equation}
-  \label{noeq:lin_transf_moments}
-  \expe{\yxs^m \| \yH} = \sum_{l=1}^m M_{ml}(n,\yNv) \expe{\yXf^l \| \yH},
-\end{equation}
-with a universal, lower-triangular transformation matrix $M_{ml}(n,\yNv)$ that
-depends only on $n$, $\yNv$, and the condition of symmetry~\eqref{noeq:homog_math}.
-
-As intuition suggests, we have 
-%$\expe{\yxs^m \| \yH} \xrightarrow{n\to \yNv} \expe{\yXf^m \| \yH}$.
-\begin{equation}
-  \label{noeq:limit_moments}
-  \expe{\yxs^m \| \yH} \xrightarrow{n\to \yNv} 
-  \expe{\yXf^m \| \yH},
-\qquad
-  \expe{\yxs^m \| \yH} \xrightarrow{n\to 1} 
-  \expe{\yXf \| \yH},
-\end{equation}
-the latter because ${\yxv_i}^m=\yxv_i$, since states are $\set{0,1}$-valued.
-
-\bigskip
-
-The core of the six mathematical relations above are
-\eqns~\eqref{noeq:marginal} and~\eqref{noeq:subpop_average}. The latter
-expresses the probability for the subpopulation average as a mixture of
-hypergeometric distributions
-\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
-with parameters $\yNv,\yNv\yXf,n$, weighted by the probabilities
-$\p(\yXf = A \|\yH)$ \citep[\cf][\sect~4, esp.\ \eqn~(22)]{kendall1967}.
-The connection between this mixture representation and the condition of
-symmetry~\eqref{noeq:homog_math} is well-known in the Bayesian literature
-\citep{kendall1967,definetti1969b,heathetal1976,diaconis1977,diaconisetal1980,jaynes1986c}.
-
-\section{Examples of inferential use of the formulae}
-
-\subsection{From network to subnetwork}
-\label{nosec:pop2sub_examples}
-
-Let us illustrate with an example how the probability distribution for the
-subnetwork average $\yxs$, determined by \eqn~\eqref{noeq:subpop_average},
-changes with the subnetwork size $n$. Choose a network-average distribution
-$\p(\yXf=A\| \yH)$ belonging to the exponential family
-\cites[\sect~4.5.3]{bernardoetal1994}[see also][]{fortinietal2000}:
-\begin{equation}
-  \label{noeq:full_pop_av_example_ME}
-  \p(\yXf = A \| \yH) = Q(A) \propto
-\binom{\yNv}{\yNv A}\exp[
-\yl_2\, \yNv A\,(\yNv A-1)/2 + \yl_1\, \yNv A].
-\end{equation}
-This is the form obtained from the principle of maximum relative entropy
-\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,aczeletal1975,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013}
-with first and second moments as constraints and the reference distribution
-$Q_0$ defined by $Q_0(A) = 2^{-\yNv}\binom{\yNv}{\yNv A}$, corresponding to a
-uniform probability distribution for the network state $\yX$.
-%file: distr_scaling_from_full_population.nb
-\begin{figure}[!b]
-\centering
-\includegraphics[width=0.95\columnwidth]{scaled_subpop_probs.pdf}%
-\caption{Probability distributions $\p(\yxs = a \| \yH)$ for different
-  subnetwork sizes $n$, obtained from a network probability
-  distribution $\p(\yXf = A \| \yH)$ having the maximum-etropy
-  form~\eqref{noeq:full_pop_av_example_ME}.}
-\label{noscaling_distr}
-\end{figure}
-\begin{figure}[!b]
-\centering
-\includegraphics[width=0.95\columnwidth]{scaling_subpop_moments.pdf}%
-\caption{Moments of the probability distributions $\p(\yxs =a \|
-  \yH)$ as functions of the subnetwork size $n$.}
-\label{noscaling_moments}
-\end{figure}
-
-The probability distribution of \eqn~\eqref{noeq:full_pop_av_example_ME} is
-plotted in \fig~\ref{noscaling_distr}, together with the resulting
-subnetwork-average distributions $\p(\yxs = a \|\yH)$, for the case in
-which $\yNv = 1000$ units, $\lambda_1=-2.55$, $\lambda_2=0.005$, and
-$n=10, 50, 100, 250$. The distributions become broader as $n$ decreases,
-and the minimum of the original distribution disappears; at the same time
-the finite-difference 
-\[\frac{\p(\yxs=a+1/n\|\yH)-\p(\yxs=a\|\yH)}{1/n}\]
-presents a sharp jump at this minimum when $n\approx 100$.
-
-To the eye familiar with maximum-entropy distributions, the
-subnetwork-average distributions of \fig~\ref{noscaling_distr} do not look
-like maximum-entropy ones with second-moment constraints. In fact,
-they are not and \emph{cannot} be:
-\begin{equation}
-  \label{noeq:maxent_form_sub}
-  \p(\yxs =a  \| \yH)
-\ne\kappa\binom{n}{n a}\exp[\yk_2\, n a\,(n a-1)/2 + \yk_1\, n a]
-%\quad\text{(false)}
-\end{equation}
-for any $\kappa, \yk_1, \yk_2$, unless $n=2$. This impossibility holds more
-generally for any number of constraints $m$ and subnetwork size $n$ such
-that $m<n$. The reason is simple: suppose we have assigned a
-maximum-entropy distribution with $m$ moment constraints as the
-distribution for the network average. If we want the same kind of
-distribution for a subnetwork of size $n$, we are free to play with $m+1$
-parameters (normalization included), but we must also satisfy the $n+1$
-equations corresponding to the marginalization~\eqref{noeq:subpop_average}.
-This is generally impossible unless $m \ge n$. (Impossibilities of a
-similar kind appear in statistical mechanics, see \eg\
-ref.~\citep{maesetal1999}.)
-%file: test_ME_same_under_scaling.nb
-
-This fact can be significant for recent works
-\citep[\eg,][]{schneidmanetal2006,shlensetal2006,tkaciketal2006,marreetal2009,tkaciketal2009,ganmoretal2011,shimazakietal2012,tkaciketal2013,shimazakietal2015}
-in which a maximum-entropy probability distribution with second- or
-third-moment constraints is assigned to relatively small subnetworks
-($n < 200$) of neurons. If we assume that such subnetwork is part of a
-larger network, and assume the condition of symmetry~\eqref{noeq:homog_math},
-then the larger network \emph{cannot} be assigned a maximum-entropy
-distribution with the same number of constraints. Vice versa, if we assign
-such a maximum-entropy distribution to the larger network, then none of its
-subnetwork of enough large size $n$ can be assigned a similar
-maximum-entropy distribution. See ref.~\citep{rostamietal2016_r2017} for a
-broader discussion of this fact and of its consequences.
-
-\medskip
-
-The dependence of the first four moments $\expe{\yxs^m \| \yH}$ as a
-function of size $n$ is shown in \fig~\ref{noscaling_moments}. The moments
-become practically constant when $n \approx 100$ or larger. The
-expectations of $m$-tuple products of states
-$\expe{\yxv_{i_1}\dotsm \yxv_{i_m} \| \yH}$, proportional to the factorial
-moments, are not shown as they do not depend on $n$.
-
-\subsection{From subnetwork to network}
-\label{nosec:from_sub_to_full}
-
-We have seen that, given the condition of symmetry~\eqref{noeq:homog_math},
-the probability $\p(\yXf = A \|\yH)$ for the network average determines
-that of each subnetwork average, $\p(\yxs \|\yH)$, by the
-marginalization \eqn~\eqref{noeq:subpop_average}. The reverse is trivially
-not true, since \eqn~\eqref{noeq:subpop_average}, as a linear mapping from
-$\RR^{\yNv+1}$ to $\RR^{n+1}$, with $\yNv$ larger than $n$, is onto but not into.
-Assigning a probability distribution $\p(\yxs = a \|\yH)$ to a
-subnetwork average $\yxs$ does not determine a network distribution
-$\p(\yXf = A\|\yH)$: it only restricts the set of possible ones; this
-set can in principle be determined via linear-programming methods
-\citep{hailperin1965,hailperin1984,hailperin1996,hailperin2006,hailperin2011}.
-\begin{innote}
-  Analogous situations appear in the truth-valued logical calculus: if the
-  composite proposition $\varAlpha\limplies \varBeta$ is assigned the
-  truth-value \enquote{true}, then assigning $\varAlpha$ the value
-  \enquote{true} also determines the value of $\varBeta$, whereas assigning
-  $\varBeta$ the value \enquote{true} leaves the value of $\varAlpha$
-  undetermined.
-\end{innote}
-The same linear-programming methods show that any inference from subnetwork
-properties to network ones must necessarily start from some assumptions
-$\varIota$ that assign a probability distribution
-$\p(\yX = \yR \| \varIota)$ for the network states. The approaches to
-this task and reformulations of it have become uncountable: they include
-exchangeable models, parametric and non-parametric models, hierarchical
-models, general linear models, models via sufficiency, maximum-entropy
-models, and whatnot
-\citep[\eg:][]{jeffreys1931_r1973,jeffreys1939_r2003,jaynes1994_r2003,bernardoetal1994,gelmanetal1995_r2014,ghoshetal1997,kallenberg2005,gregory2005,sivia1996_r2006,ferreiraetal2007,dawid2013,damienetal2013}.
-We now show two examples, based on a maximum-entropy approach, that to our
-knowledge have not yet been explored in the neuroscientific literature. For
-a concrete application see \citep{rostamietal2016b}.
-
-\paragraph{First example: moment constraints for the network.}
-\label{nosec:maxent_moments}
-Consider a state of knowledge $\yHa$ leading to the following properties:
-\begin{enumerate}%[$\yHa$1.]
-\item the expectations of the single and pair averages $\yxs$ and
-  $\sav{\yx\yx}$ of a particular subnetwork have given values
-  \begin{equation}
-    \label{noeq:constr_ex1}
-    \expe{\yxs \| \yHa} = c_1, \qquad \expe{\sav{\yxv_i \yxv_j}\|\yHa} = c_2;
-  \end{equation}
-\item the network probability distribution $\p(\yX = \yR \|\yHa)$
-  has maximum relative entropy with respect to the uniform one, given the
-  constraints above.
-\end{enumerate}
-
-\medskip Then the probability distribution for the network conditional on $\yHa$
-is completely determined: it satisfies the symmetry
-property~\eqref{noeq:homog_math} and is defined by
-\begin{multline}
-  \label{noeq:pop_distr_maxent1}
-  \pf(\yX =\yR \|\yHa) =
-\varKappa\exp[\yL_2 \yNv\yRf\,(\yNv\yRf-1)/2 + \yL_1 \yNv\yRf]
-\\
-\!\begin{aligned}[b]
-&\text{with $\varKappa, \yL_m$, such that the distribution is normalized and}
-\\
-&\varKappa\sum_{\yNv A=0}^\yNv %m!\, 
-\binom{\yNv-m}{\yNv A-m}
-\exp[\yL_2 \yNv A\,(\yNv A-1)/2 + \yL_1 \yNv A]
-= c_m, \quad m=1,2.
-\end{aligned}
-\end{multline}
-We omit the full proof of this statement: it is a standard application of
-the maximum-entropy procedure
-\citep[\eg:][]{jaynes1957,jaynes1963,good1963,jaynes1967,jaynes1979b,vancampenhoutetal1981,sivia1990,fangetal1997,bretthorst2013},
-combined with the equality~\eqref{noeq:expe_products} of subnetwork and
-network expectations, \eg
-\begin{equation}
-  \label{noeq:exploit_eq_expects}
-  c_2 = \expe{\sav{\yx\yx}\|\yHa} = 
-\binom{\yNv}{2}^{-1}
-\sum_{\yNv A=0}^\yNv %m!\, 
-\binom{\yNv A}{2} \, \p(\yXf = A\|\yHa),
-\end{equation}
-and with relations~\eqref{noeq:joint_plaus_N_homog},
-\eqref{noeq:conditional_total_state}. This example is easily generalized to
-any number $m$ of constraints such that  $m\le n$.
-
-Note again that, as remarked in \sect~\ref{nosec:pop2sub_examples}, the
-subnetwork from which the averages in the
-expectations~\eqref{noeq:constr_ex1} are calculated has a probability
-distribution $\p(\yxs =a \|\yH)$ determined by the
-marginalization~\eqref{noeq:subpop_average} and does \emph{not} have a
-maximum-entropy form with the same number of constraints.
-%file: test_ME_same_under_scaling.nb
-
-\paragraph{Second example: subnetwork-distribution constraint.}
-\label{nosec:maxent_fullsubpop}
-Consider another state of knowledge $\yHb$ leading to the following
-properties:
-\begin{enumerate}%[wide,label=$\yHb$\arabic*.]
-\item the average $\yxs$ of a particular
-  subnetwork has a probability distribution $q$:
-  \begin{equation}
-    \label{noeq:constr_subpop_distr}
-    \p(\yxs =a \|\yHb) = q(a);
-  \end{equation}
-\item the probability distribution for the network, $\p(\yX = \yR
-  \|\yHb)$, has maximum relative entropy with respect to the uniform
-  one, given the constraint above.
-\end{enumerate}
-
-\medskip Then the probability distribution for the network given $\yHb$ is
-completely determined and satisfies the symmetry
-property~\eqref{noeq:homog_math}:
-\begin{multline}
-  \label{noeq:pop_distr_maxent2}
-  \p(\yX= \yR \|\yHb) =
-\exp\Biggl[
-\sum_{n a=0}^n \yL_{ a}
-%\binom{N}{N\yXf}^{-1}
-\binom{n}{n a}
-\binom{\yNv-n}{\yNv\yRf-n a}\Biggr]
-\\
-\!\begin{aligned}[b]
-&\text{with $\yL_{ a}$ such that}
-\\
-&\sum_{\yNv A=0}^\yNv 
-%\binom{N}{N A}^{-1}
-\binom{n}{n a}
-\binom{\yNv-n}{\yNv A-n a}\,
-\exp\Biggl[
-\sum_{n a=0}^n \yL_{ a}
-%\binom{N}{N A}^{-1}
-\binom{n}{n a}
-\binom{\yNv-n}{\yNv A-n a}\Biggr]
-=q( a)
-\end{aligned}
-\end{multline}
-(the normalization constraint being unnecessary since $q$ is normalized).
-This result is just another application of the maximum-entropy procedure
-with $n+1$ (linear) constraints given by \eqn~\eqref{noeq:marginal}, where
-the left-hand side is now given and equal to $q(a)$.
-
-This example is equivalent to the generalization of the previous one with
-$n$ moment constraints, since knowledge of $\p(\yxs=a \| \yHb)$ is
-equivalent to knowledge its first $n$ moments.
-
-***
-
-In this note we would like to analyse and warn about a subtle assumption
-behind the \me\ method when it is applied to a population. It can informally
-be put this way:
-\begin{quote}
-  \emph{the \me\ method assumes that the population it is applied to is
-    completely isolated from any larger population.}
-\end{quote}
-
-\mynote{Version 1}
-The \me\ method does not construct a probability distribution out of
-nothing, but starting from a uniform distribution. A uniform distribution
-is an innocuous assumption for a set of non-composite events, like the
-outcomes of a die roll, and also for some sets of composite events, like
-the outcomes of the roll of two dice. In the latter case multiplicities
-appear.
-
-When applied to a subpopulation, the \me\ method assumes that the uniform over
-the larger population is uniform, and therefore factorizable. The new
-distribution of the subpopulation will not be uniform, but that of the full
-population will still be factorizable into the one for the subpopulation and the
-rest.
-
-A uniform distribution, however, is not the right one when we suppose that
-learning about an event may tell us something about a related event. For
-example, consider 1000 tosses of a particular coin and assume a uniform
-distribution over the possible $2^{1000}$ outcomes. If we learn that the
-first 999 tosses yielded all \enquote{heads}, the probability calculus
-tells us that the probability for the 1000th toss is still 50\%/50\%. It
-is a consequence of our choice of a uniform distribution: we have
-implicitly declared all tosses to be completely independent, completely
-\emph{irrelevant} to one another. This fact is well-known in sampling
-theory. A more telling example in fact is that of a presidential election
-with two candidates: each citizen will vote for one or the other. We do
-survey sampling on a large number of citizens to guess the election's
-outcome. If we assumed a uniform distribution over the possible
-combinations of choices of all citizens, our sampling would be completely
-irrelevant for the choices of the rest of the population.
-
-The latter example has many similarities with that of a neuronal binary
-population. When we record the neuronal activity of a sample of neurons from a
-brain area, we assume that our measurements can tell us something -- no
-matter how vague or imprecise -- about the whole brain area. This means
-that we are not assuming a uniform distribution over all possible states of
-the area.
-
-\mynote{Version 2}
-This may come as a surprise. The method simply requires a number of
-exhaustive and mutually exclusive events, and if these are composite events
-the final distribution may have a multiplicity factor. When we consider the
-$2^{\yNv}$ states of $\yNv$ units we are not excluding that these might be
-marginals of $2^{M}$ states of $M$ units. Each one has the same
-multiplicity $2^{M-\yNv}$, but this constant multiplicity factor disappears
-by normalization. So the method applies just as in the case of $\yNv$ units
-only, right?
-
-Right, but
-
-Right, and that is where the problem lies. This way of counting of
-multiplicities assumes an underlying
-
-Wrong. In our reasoning we have made subtle assumptions of independence
-between the full population and the subpopulation. The problem is that the
-counting of multiplicities is not based on simple enumeration, but already
-involves probability considerations. Consider three cases with a full
-population of two units, $M=2$, of which we consider one unit, $\yNv=1$.
-\begin{itemize}
-\item First case: all four states are \emph{possible}. The two states of
-  the first unit have multiplicity $2$ each. The usual \me\ distribution
-  obtains.
-\item Second case: only the states with at most one active unit are
-  possible. The state $\yxx_1=1$ of the first unit has multiplicity
-  $2$, and $\yxx_1=1$ has multiplicity $1$. The \me\ distribution has
-  multiplicity factors.
-\item Third case: states with at most one active unit are, say, $10^{9}$
-  times more probable than the state with no active units. But all four
-  states are \emph{possible}. By enumeration this case is like the first:
-  multiplicities $(1,1)$. But by common sense it is more similar to the
-  second: multiplicities $(2,1)$ for most practical purposes.
-\end{itemize}
-
-This simple example shows that the multiplicity inspection that must
-precede a \me\ application already involves probability considerations at
-the level of the full population. The usual reasoning by enumeration
-implicitly assumes a uniform distribution or at least a \emph{factorizable}
-distribution.
-
-***If the distribution is factorizable, however, it means that examination of
-the subpopulation \emph{cannot give us any insights about the population it is a
-  part of}. This is obviously contrary to the reason why we made neuronal observations.
-
-Consider the following ways of proceeding. We:
-\begin{enumerate}
-\item have a population with $\yNv$ units, $2^{\yNv}$ possible states
-\item expect averages of $\yx$ active neurons and $\yxxs$ active pairs
-\item use maximum-entropy to choose a probability distribution for the
-  states of the $\yNv$ units conforming to our expectations.
-\end{enumerate}
-
-We:
-\begin{enumerate}
-\item have a population with $M$ units, $2^{M}$ possible states
-\item expect that any subpopulation of $\yNv$ units has $\yxs$ active
-  neurons and $\yxxs$ active pairs
-\item use maximum-entropy to choose a probability distribution for the
-  states of the $M$ units conforming to our expectations
-\item marginalize to find the probability distribution for the states of
-  $\yNv$ units.
-\end{enumerate}
-
-  \section{Sketched proofs}
-  \label{nosec:derivations}
-
-  Variants of the following derivations and combinatorial considerations
-  can be found \eg\ in
-  \cites[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003};
-  see also \citep{whitworth1897}.
-
-  To derive the joint probability distribution
-  \eqref{noeq:joint_plaus_N_homog} from that for the network
-  average~\eqref{noeq:pop_average_Q}, consider that if the network total
-  is $\yNv\yXf$, then $\yNv\yXf$ out of $\yNv$ units are active, and there are
-  $\tbinom{\yNv}{\yNv\yXf}$ possible states for which this can be true; therefore
-  \begin{equation}
-    \labelbis{eq:joint_plaus_\yNv_homog}
-    \p(\yX = \yR \| H) = \binom{\yNv}{\yNv\yRf}^{-1}Q(\yRf).
-  \end{equation}
-An analogous reasoning for $n$ and $\yxs$ leads to an analogous equality,
-\begin{equation}
-  \label{noeq:aver_subpop}
-  \p(\yx=\yr \| H) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs\|\yH),
-\end{equation}
-for the subnetwork.
-
-\bigskip
-
-Let us next consider the probability $\p(\yxs =a \| \yXf = A,\yH)$ for
-the subnetwork average $\yxs$ conditional on the network average $\yXf$.
-There are $\tbinom{\yNv}{\yNv\yXf}$ possible network states if the network
-average is $\yXf$, \ie\ if $\yNv\yXf$ units are in state $1$; the conditional
-probability of each is therefore $1/\tbinom{\yNv}{\yNv\yXf}$, owing to the
-symmetry assumption~\eqref{noeq:homog_math}. Now consider the subnetwork of the
-first $n$ units. The conditional probability of having $n\yxs$ specific
-ones in state $1$ is the sum of the probabilities of all states for
-which $\yNv\yXf-n\yxs$ of the remaining $\yNv-n$ units are in state $1$; there are
-$\tbinom{\yNv-n}{\yNv\yXf-n\yxs}$ such states, all equally probable. Finally,
-there are $\tbinom{n}{n\yxs}$ possible ways, all equally probable, in which
-$n\yxs$ of the first $n$ units can be in state $1$. In formulae,
-\begin{equation}
-\label{noeq:conditional_prob_explained}
-\begin{split}
-      \p(\yxs =a \| \yXf=A, \yH)
-      &=
-      \sum_{\yr}^{\yrs=a} \sum_{\yR}^{\yRf=A} 
-      \p(\yx =\yr \|  \yX = \yR, \yH) \,
-      \p(\yX =\yR  \| \yXf=A,\yH),
-      \\
-      &=
-      \sum_{\yr}^{\yrs=a} \sum_{\yR}^{\yRf=A} 
-      \p(\yx=\yr \|  \yX =\yR, \yH) \,
-      \binom{\yNv}{\yNv A}^{-1},
-      \\
-      &=\sum_{\yr}^{\yrs=a} 
-      \binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1},
-      \\
-      &=
-      \binom{n}{n a}\binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1},
-\end{split}
-\end{equation}
-  which is the conditional probability~\eqref{noeq:conditional_prob}. Note
-  that this is just a derivation of the hypergeometric distribution
-  \cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
-  which describes the probability of, say, drawing a proportion of $\yxs$
-  blue balls in $n$ drawings without replacement from an urn with $\yNv$
-  balls, a fraction $\yXf$ of which are blue.
-
-  \bigskip The probability of a subnetwork average $\yxs$ is then, by
-  marginalization,
-  \begin{equation}
-    \label{noeq:subpop_average_explained}
-    \begin{split}
-      \p(\yxs =a \| \yH)
-      &= \sum_{\yNv A=0}^\yNv\p(\yxs = a \| \yXf = A, \yH)\,
-      \p(\yXf = A \| \yH),
-      \\
-      &=\sum_{\yNv A=0}^\yNv
-      \binom{n}{n a}\binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1}
-      Q(A).
-    \end{split}
-  \end{equation}
-  which proves the subnetwork-average formula~\eqref{noeq:subpop_average}.
-  This formula, combined with \eqns~\eqref{noeq:aver_subpop} and
-  \eqref{noeq:joint_plaus_N_homog}, leads to the conditional
-  probability~\eqref{noeq:conditional_total_state}.
-
-  \bigskip The independence of the expectation of products of states from
-  the subnetwork size is trivial by marginalization:
-  \begin{multline}
-    \label{noeq:margin_products}
-    \sum_{\yx} \yxv_1\dotsm \yxv_m \p(\yx =\yr \| \yH)
-      =\sum_{\yr,\yR}  \yrv_1\dotsm \yrv_m \p(\yx=\yr \| \yX=\yR,\yH)
-      \, \p(\yX=\yR \| \yH),
-      \\
-      \begin{aligned}
-      &=\sum_{\yr,\yR} \yrv_1\dotsm \yrv_m\,\delt(\yRv_1-\yrv_1)\dotsm\delt(\yRv_m-\yRv_m)
-      \p(\yX = \yR \| \yH),
-      \\
-      &=\sum_{\yR} \yRv_1\dotsm \yRv_m \p(\yX =\yR \| \yH).
-    \end{aligned}
-  \end{multline}
-  All such $m$-fold products have the same expectation by symmetry,
-  therefore their subnetwork average will do, too, being an average of
-  equal terms.
-
-  Now consider the sum of all distinct products of states of two units in
-  the subnetwork:
-  \begin{equation*}
-    \yxv_1 \yxv_2 + \yxv_1 \yxv_3 + \dotsb + \yxv_{n-1} \yxv_n.
-  \end{equation*}
-  The terms in this sum are either $0$ or $1$. The non-vanishing ones are
-  those with index pairs chosen from the $n\yxs$ units of the subnetwork
-  which are in state $1$, and there are $\tbinom{n\yxs}{2}$ such choices,
-  so the sum above is equal to $\tbinom{n\yxs}{2}$. The sum has
-  $\tbinom{n}{2}$ terms, so their average is
-  $\tbinom{n\yxs}{2}/\tbinom{n}{2}$. Generalizing the argument to products
-  of $m$ units, we have that
-  \begin{equation}
-    \sav{\yxv_{i_1} \dotsm \yxv_{i_m}} = \binom{n\yxs}{m} \binom{n}{m}^{-1}.
-  \end{equation}
-  Then, using \eqn~\eqref{noeq:subpop_average},
-  \begin{multline}
-    \label{noeq:expect_product_m}
-    \expe*{\sav{\yxv_{i_1} \dotsm \yxv_{i_m}} \| \yH}
-    \\
-    \begin{aligned}[b]
-      &=\tfrac{(n-m)!}{n!}
-      \expe*{m!\,\tbinom{n\yxs}{m}  \| \yH}
-      \\
-      &\equiv\frac{(n-m)!}{n!}
-      \sum_{n a=0}^n m!\, \binom{n a}{m} \p(\yxs =a \|\yH),
-      \\
-      &=\frac{(n-m)!}{n!} \sum_{\yNv A=0}^\yNv
-      Q( A)
-      \Biggl[ 
-      \sum_{n a=0}^n 
-      m!\,\binom{n a}{m}\, \binom{n}{n a}\binom{\yNv-n}{\yNv A-n a}\binom{\yNv}{\yNv A}^{-1} 
-      \Biggr].
-    \end{aligned}
-  \end{multline}
-  The expression in brackets is the $m$th factorial moment of the
-  hypergeometric function, and is given by \citep{potts1953}
-%file: factorial_moments_etc.nb
-  \begin{equation}
-    \sum_{n a=0}^n m!\,\binom{n a}{m} \binom{n}{n a}\binom{\yNv-n}{\yNv\yXf-n a}
-    \binom{\yNv}{\yNv\yXf}^{-1} 
-    =
-    m!\,\binom{n}{m} \binom{n a}{m}\binom{\yNv}{m},
-  \end{equation}
-  which combined with the previous equation yields the second line of
-  \eqn~\eqref{noeq:expe_products}; its last equality comes from the identity
-  \begin{equation}
-    \label{noeq:binom_identity}
-    \binom{\yNv}{M}\binom{M}{m}=\binom{\yNv}{m}\binom{\yNv-m}{M-m},
-  \end{equation}
-  easily derived by writing the binomial coefficients in terms of
-  factorials. Finally, \eqns~\eqref{noeq:moments}, relating the moments of
-  the distributions for subnetwork and network averages, is obtained
-  from the definition of moments,
-  \begin{equation}
-    \label{noeq:moments_def}
-    \expe*{\yxs^m \| \yH} \defd
-    \sum_{n a=0}^{n} a^m \p(\yxs = a  \| H),
-    \quad
-    \expe{\yXf^m \| \yH} \defd
-    \sum_{\yNv A=0}^{\yNv} A^m \p(\yXf = A  \| H),
-  \end{equation}
-  replaced in the equalities for the factorial
-  moments~\eqref{noeq:eq_factmom_prod}, by recursively solving in terms of
-  the moments of the network distribution.
-
-  *******
-
-Here is an example.
-
-Suppose our desired constraints are $\expe{\yxs } = c_1$ and
-$\expe{\yxxs} = c_2$, the expectations being given by the fourth term of
-\eqn~\eqref{eq:expe_products}. Applying \me\ to the distribution of the
-average of the full population we find
-\begin{equation}
-  \label{eq:full_pop_av_example_ME}
-  \p(\yXf = \yRf \| \yHa)  = \yL\,
-\binom{\yNv}{\yNv \yRf}\exp\biggl[
-\yL_2 \yRf \,\frac{\yNv \yRf-1}{\yNv-1} + \yL_1\, \yRf\biggr]
-\end{equation}
-with such parameters $\yL$, $\yL_1$, $\yL_2$ as to satisfy normalization
-and constraints. We call this state of knowledge $\yHa$. Marginalization by
-\eqn~\eqref{eq:marginal} then gives
-\begin{equation}
-  \label{eq:full_pop_av_example_ME_marg}
-  \p(\yxs = \yrs \| \yHa)  = \yL
-  \sum_{\yNv\yRf=0}^\yNv
-  \binom{n}{n\yrs} \binom{\yNv-n}{\yNv\yRf-n\yrs} \exp\biggl[
-\yL_2 \yRf \,\frac{\yNv \yRf-1}{\yNv-1} + \yL_1\, \yRf\biggr].
-\end{equation}
-Applying \me\ to the distribution of the average of the sample, using the
-third term of \eqn~\eqref{eq:expe_products}, we find
-\begin{equation}
-  \label{eq:sample_av_example_ME}
-  \p(\yxs = \yrs \| \yHb)  = \yl\,
-  \binom{n}{n\yrs}
-  \exp\biggl[\yl_2 \yrs \,\frac{n \yrs-1}{n-1} + \yl_1\, \yrs\biggr]
-\end{equation}
-with such parameters $\yl$, $\yl_1$, $\yl_2$ as to satisfy the constraints.
-We call this state of knowledge $\yHb$.
-
-The resulting distributions $\p(\yxs = \yrs \| \yHa)$,
-$\p(\yxs = \yrs \| \yHb)$ are different.
-Figure~\ref{fig:diff_maxent_pop_sample} shows one example, for $\yNv=5000$,
-$n=100$, $c_1=0.45$, $c_2=0.35$. The two distributions have slightly
-different modes and the one obtained from \me\ at the population level is
-more peaked. Another example is shown in
-\fig~\ref{fig:diff_maxent_pop_sample_realdata}, for $\yNv=5000$, $n=200$,
-and neurobiologically more realistic constraints $c_1=0.045$, $c_2=0.0025$
-\cite{rostamietal2016}. The discrepancy is maybe not as large as in the
-previous example, but the tails of the distributions are still very
-different.
-
-**********************************************************
-
-The neurons recorded in these kinds of experiments are usually picked out
-according to an unknown process, and from their observation we expect to
-learn something about all other neurons in the same brain area. That is,
-they are assumed to be a \enquote{representative random sample} of the
-neurons in that area. Our discussion hinges on this often just tacitly
-understood assumption.
-
-One of the uses of \me\ methods
-
-number. It is large from large
-population of neurons [\cite]. The recorded neurons are usually picked out
-from a specific brain area according to an unknown process, and from their
-observation in relation to stimuli or behavioral task, we strive for
-understanding the mechanism behind neuronal coding\mynote{LM still don't
-  know what people mean by this, can we find an unambiguous term or
-  rephrase?}. That is, they are assumed to be a \enquote{representative
-  random sample} of the neurons in that area.\mynote{This assumption says
-  that from the sample we can learn about the population. The relation of
-  population activity with stimuli or behaviour is a separate problem, so
-  \enquote{this is} doesn't really fit...}
-
-Huge amounts of parallel recordings from hundreds of neurons, available
-today, are accompanied by the statistical challenges to be employed for
-shedding light on the mechanism behind neuronal coding. The principle of
-\me\ was proposed as a way to overcome these challenges by constructing a
-full description of the neuronal firing patterns. The \me\ method employs
-only first and second order statistics of the representative random sample
-to construct a \enquote{maximally noncommittal} \cite{jaynes1963}
-probability distribution of the firing patterns [\cite]. \mynote{LM:
-  \enquote{noncommittal} doesn't mean anything to me to be honest, even if
-  Jaynes used that term. I would like to openly say this, backed up by
-  references, as done some paragraphs below.\\It seems they have used \me\
-  for other purposes too, sometimes unclear ones. And with all kinds of
-  constraints. So we shouldn't be too specific here in relation to purpose
-  and constraints. I'd rather give references and the readers can form
-  their own ideas about its use.}
-
-Using the pairwise \me\ models many studies investigated the questions such
-as, which level of interaction\mynote{ambiguous word} (pairwise or higher-order)
-between neurons is suffiecient to explain the observed experimental data
-[\cite]? Or how collective behavior of neurons can be related to stimuli or
-behavioral tasks [\cite]? add more **
-
-Here we show that there is a quandary in applying \me\ to a representative
-random sample of neurons to generate a maximally noncommital distribution
-for that sample.
-
-The quandary is this. Assume that our sample of neurons is
-representative of some population. If we assign a \me\ distribution to the
-sample, then we cannot assign a \me\ distribution to the full population.
-Vice versa, If we assign a \me\ distribution to the full population, then
-we cannot assign a \me\ distribution to the sample. This impossibility
-holds at least for \me\ distribution with moment constraints, and even if
-the orders of the moments constrained in the sample and in the full
-population are different.
-
-In other words, if our sample is a \enquote{random representative} of some
-population, then we must choose which has a \enquote{maximally
-  noncommittal} distribution: the sample or the population. We can't choose
-both. It goes without saying that if one is \enquote{maximally
-  noncommittal}, the other must be somehow \enquote{committal}. Which
-choice is most meaningful?
-
-In the rest of the paper we will mathematically show the contradiction
-above and generate a \me\ distribution for the full population rather than
-the sample. We will also present some probability relations relevant to
-random sampling. These relations are well-known in survey sampling and in
-the pedagogic problem of drawing from an urn without replacement, yet they
-are somewhat hard to find explicitly written in the neuroscientific
-literature.
-
- \mynote{Version 1}
-The present note has three nested purposes.
-
-The first and more general is to remind our readers that the \me\ method
-rests on subtle assumptions that may be contradictory with other
-assumptions we want to make on the data under study.
-
-As a concrete example we show -- and this is the second purpose -- that a
-sample of a neuronal population cannot be assigned a \enquote{maximally
-  noncommittal} \me\ distribution and at the same time be considered a
-\enquote{random representative} of the population.
-
-The proof of the example above rests on some probability relations from
-classical random sampling. These relations are somewhat hard to find
-explicitly stated in the neuroscientific literature; the third, minor
-purpose of this note is to state them explicitly.
-
-Let's clarify at once that our discussion pertains to a binary population at
-one specific instant or short window in time. It is possible to
-similarly discuss multi-state neuron models and population dynamics; but for
-simplicity neurons are here assumed to be in a fixed, active \enquote{$1$}
-or inactive \enquote{$0$} state; and evolution, change, time correlations,
-and similar concepts do not concern us.
-
-Maximum-entropy models are used for a variety of (sometimes not completely
-clear) reasons. For our purposes let's say that the \me\ method is assumed
-to generate a \enquote{maximally noncommittal} \cite{jaynes1963} probability
-distribution. We intend the quoted expression only as an umbrella term,
-also because it means very little without further clarifications.
-
-The neurons recorded in these kinds of experiments are usually picked out
-according to an unknown process, and from their observation we expect to
-learn something about all other neurons in the same brain area. That is,
-they are assumed to be a \enquote{representative random sample} of the
-neurons in that area. Our discussion hinges on this often just tacitly
-understood assumption.
-
-\mynote{Version 2 }
-In this note we would like to show that there is a
-contradiction in applying \me\ to a \enquote{representative random sample}
-of neurons to generate a \enquote{maximally noncommital} distribution for
-that sample.
-% \\
-% \mynote{V2 }In this note we would like to show that there is a clash between the
-% \enquote{representative random sample} spirit and the \enquote{maximally
-%   noncommital} spirit of \me\ models applied to such a sample of neurons.
-
-The contradiction is this. Assume that our sample of neurons is
-representative of some population. If we assign a \me\ distribution to the
-sample, then we cannot assign a \me\ distribution to the full population.
-Vice versa, If we assign a \me\ distribution to the full population, then
-we cannot assign a \me\ distribution to the sample. This impossibility
-holds at least for \me\ distribution with moment constraints, and even if
-the orders of the moments constrained in the sample and in the full
-population are different.
-
-In other words, if our sample is a \enquote{random representative} of some
-population, then we must choose which has a \enquote{maximally
-  noncommittal} distribution: the sample or the population. We can't choose
-both. It goes without saying that if one is \enquote{maximally
-  noncommittal}, the other must be somehow \enquote{committal}. Which
-choice is most meaningful?
-
-In the rest of the paper we will mathematically show the contradiction
-above and generate a \me\ distribution for the full population rather than
-the sample. We will also present some probability relations relevant to
-random sampling. These relations are well-known in survey sampling and in
-the pedagogic problem of drawing from an urn without replacement, yet they
-are somewhat hard to find explicitly written in the neuroscientific
-literature.
-\fi
 \end{document}

### 2018-05-18 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-18T10:23:24+0200
+%% Last-Updated: 2018-05-18T10:25:17+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\figs}{figs}%
-%\newcommand*{\vs}{{vs.}}
+\newcommand*{\vs}{{vs}}
 %\newcommand*{\etc}{{etc.}}
@@
 
-\section{Maximum-entropy: sample level vs full-population level}
+\section{Maximum-entropy: sample level \vs\ full-population level}
 \label{sec:specific_initial_probability}
@@
 the neuroscientific applications considered in this note the difference in
-predictions of the sample vs full-population applications may be very
+predictions of the sample \vs\ full-population applications may be very
 relevant instead.

### 2018-05-18 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-18T10:25:17+0200
+%% Last-Updated: 2018-05-18T10:28:29+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   sets?
-\item How to extend the new application when we want to study the 
-  activities of individual neurons, of individual pairs, triplets, and so
-  on?
+\item How to extend the new application to the \enquote{inhomogeneous} case
+  \cite{schneidmanetal2006,shlensetal2006,roudietal2009b}, in which
+  expectations for individual neurons or groups of neurons are constrained?
 \item What is the mathematical relation between the new application and

### 2018-05-18 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-18T10:28:29+0200
+%% Last-Updated: 2018-05-18T11:33:51+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 complex versions of the method, possibly except version that use
-\enquote{hidden} neurons. The problem is that the recorded neurons are a
-\emph{sample} from a larger, unrecorded population, but the \me\ method as
-applied above is treating them as isolated from the rest of the brain.
-Hence, the results it provides cannot rightfully be extrapolated. We will
-give a mathematical proof of this. Let us first analyse this issue in more
-detail.
+\enquote{hidden} neurons \cite{smolensky1986,hintonetal2006}. The problem
+is that the recorded neurons are a \emph{sample} from a larger, unrecorded
+population, but the \me\ method as applied above is treating them as
+isolated from the rest of the brain. Hence, the results it provides cannot
+rightfully be extrapolated. We will give a mathematical proof of this. Let
+us first analyse this issue in more detail.
 
@@
 The latter formula, in particular, shows that the standard way of applying
-\me\ (without \enquote{hidden neurons}) implicitly assumes that \emph{no}
-larger population exists beyond the recorded sample of neurons. One could
-in fact object to the application at the population level, and say that the
-traditional way of applying \me, \eqn~\eqref{eq:app_maxent_sample}, yields
-different results because it does not make assumptions about the size
-$\yNv$ of a possibly existing larger population. Such a state of
-uncertainty, however, is correctly formalized according to the laws of
-probability by introducing a probability distribution for $\yNv$, and is
-expressed by \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}. This expression
-cannot generally be equal to~\eqref{eq:app_maxent_sample} unless the
-distribution for $\yNv$ gives unit probability to $\yNv=n$; that is,
-unless the sample \emph{is} the full population, and no larger population
-exists or the recorded neurons are completely isolated from it.
-
-The latter notion is unreasonable. It amounts to say that the neurons we
-distinguished and tracked with our recording device are a very special,
-isolated set among all those that could have been recorded. It is for this
-reason that we find the \me\ application at the population level
-preferable. Physical models of neuronal networks usually include some sort
-of external input, mimicking an embedding in a larger network. The \me\
-distribution obtained at the population level may reveal features in a
+\me\ 
+implicitly assumes that \emph{no} larger population exists beyond the
+recorded sample of neurons. One could in fact object to the application at
+the population level, and say that the traditional way of applying \me,
+\eqn~\eqref{eq:app_maxent_sample}, yields different results because it does
+not make assumptions about the size $\yNv$ of a possibly existing larger
+population. Such a state of uncertainty, however, is correctly formalized
+according to the laws of probability by introducing a probability
+distribution for $\yNv$, and is expressed by
+\eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}. This expression cannot
+generally be equal to~\eqref{eq:app_maxent_sample} unless the distribution
+for $\yNv$ gives unit probability to $\yNv=n$; that is, unless the sample
+\emph{is} the full population, and no larger population exists or the
+recorded neurons are completely isolated from it.
+
+This hidden assumption of isolation is unreasonable. It amounts to say that
+the neurons we distinguished and tracked with our recording device are a
+very special, isolated set among all those that could have been recorded.
+It is for this reason that we find the \me\ application at the population
+level preferable. Physical models of neuronal networks usually include some
+sort of external input, mimicking an embedding in a larger network. The
+\me\ distribution obtained at the population level may reveal features in a
 data set that were unnoticed by the standard \me\ model.
 
-The difference between the applications at the sample and at the population
-levels appears also in statistical mechanics. If a system is statistically
-described by a \me\ Gibbs state, its subsystems cannot be described by a
-Gibbs state \cite{maesetal1999}. A somewhat similar situation also appears
-in the statistical description of the final state of a non-equilibrium
-process starting and ending in two equilibrium states: we can describe our
-knowledge about the final state either by (1) a Gibbs distribution,
-calculated from the final equilibrium macrovariables, or (2) by the
-distribution obtained from the Liouville evolution of the Gibbs
+The difference in the resulting distributions between the applications at
+the sample and at the population levels appears in the use of Boltzmann
+machines with hidden units \cite{lerouxetal2008}, although by a different
+conceptual route. It also appears in statistical mechanics: if a system is
+statistically described by a \me\ Gibbs state, its subsystems cannot be
+described by a Gibbs state \cite{maesetal1999}. A somewhat similar
+situation also appears in the statistical description of the final state of
+a non-equilibrium process starting and ending in two equilibrium states: we
+can describe our knowledge about the final state either by (1) a Gibbs
+distribution, calculated from the final equilibrium macrovariables, or (2)
+by the distribution obtained from the Liouville evolution of the Gibbs
 distribution assigned to the initial state. The two distributions differ
@@
 usually immensely sharp and practically lead to the same predictions. In
-the neuroscientific applications considered in this note the difference in
-predictions of the sample \vs\ full-population applications may be very
-relevant instead.
+neuroscientific applications, the difference in predictions of the sample
+\vs\ full-population applications can instead be very relevant.
 
@@
 \item What is the mathematical relation between the new application and
-  \me\ models with hidden neurons?
+  \me\ models with hidden neurons \cite{smolensky1986,hintonetal2006}?
 \end{itemize}
@@
 \newcommand*{\subtitleproc}[1]{}
+\newcommand*{\chapb}{ch.}
 

### 2018-05-18 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-18T11:33:51+0200
+%% Last-Updated: 2018-05-18T11:50:42+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\newcommand*{\pdftitle}{Maximum-entropy and neuronal samples:\\an alternative approach}
-
-\newcommand*{\headtitle}{\pdftitle} \newcommand*{\pdfauthor}{P.G.L. Porta Mana, V. Rostami, E. Torre}
+\newcommand*{\pdftitle}{Maximum-entropy and samples of neuronal activity:\\an alternative approach}
+
+\newcommand*{\headtitle}{\pdftitle} \newcommand*{\pdfauthor}{P.G.L.  Porta Mana, V. Rostami, E. Torre, Y. Roudi}
 \usepackage{newtxmath}
@@
 and it has been subjected to mathematical and conceptual scrutiny
-\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.
+\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.
 

### 2018-05-18 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-18T11:50:42+0200
+%% Last-Updated: 2018-05-18T15:43:51+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\usepackage[final,%
+\usepackage[%final,%
 nonatbib]{nips_2017}
@@
 
-\newcommand*{\pdftitle}{Maximum-entropy and samples of neuronal activity:\\an alternative approach}
+\newcommand*{\pdftitle}{Maximum-entropy and representative samples\\ of neuronal
+  activity: a dilemma}
 
@@
 \author{
-  P.G.L. Porta Mana\\\texttt{piero.mana@ntnu.org}
+  P.G.L. Porta Mana\\\texttt{piero.mana@ntnu.no}
   \And
-  V. Rostami\\\texttt{v.rostami@fz-juelich.de}
+  V. Rostami\\\texttt{vrostami@uni-koeln.de}
   \AND
@@
 \begin{abstract}
-  This note shows that the \me\ method can be applied to a sample of
-  neuronal activity along two different routes: (1) apply to the sample; or
-  (2) apply to a larger, unsampled neuronal population from which the
-  sample is drawn, and then marginalize to the sample. The second route can
-  be generalized to the case where the size of the larger population is
-  unknown. These two routes give inequivalent results. Which route should
-  be chosen? Some arguments are presented in favour of the second. This
-  note also presents and discusses probability formulae that relate our
+  The present work shows that the \me\ method can be applied to a sample of
+  neuronal recordings along two different routes: (1) apply to the sample;
+  or (2) apply to a larger, unsampled neuronal population from which the
+  sample is drawn, and then marginalize to the sample. These two routes
+  give inequivalent results. The second route can be further generalized to
+  the case where the size of the larger population is unknown. Which route
+  should be chosen? Some arguments are presented in favour of the second.
+  This work also presents and discusses probability formulae that relate
   states of knowledge about a population and its samples, and that may be
@@
 
-\enquote{Cooperativity} can quantified and characterized with \me\ method
-in several ways. The simplest way roughly proceeds along the following
-steps. Consider the recorded activity of a sample of $n$ neurons.
+\enquote{Cooperativity} can be quantified and characterized with \me\
+methods in several ways. The simplest way roughly proceeds along the
+following steps. Consider the recorded activity of a sample of $n$ neurons.
 \begin{enumerate}
@@
 complex versions of the method, possibly except version that use
-\enquote{hidden} neurons \cite{smolensky1986,hintonetal2006}. The problem
-is that the recorded neurons are a \emph{sample} from a larger, unrecorded
+\enquote{hidden} neurons
+\cite{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}. The problem is
+that the recorded neurons are a \emph{sample} from a larger, unrecorded
 population, but the \me\ method as applied above is treating them as
@@
 
-In this paper we develop the second way of applying the \me\ method, at the
+In this work we develop the second way of applying the \me\ method, at the
 level of the larger population, and show that its results differ from the
@@
 
-The notation and terminology in this note follow ISO and ANSI standards
+The notation and terminology in the present work follow ISO and ANSI standards
 \citep{iso1993,ieee1993,nist1995,iso2006,iso2006b} but for the use of the
@@
 
-The purpose of the present note was to simply to point out and show, in a
-simple set-up, that the \me\ method can be applied to recorded neuronal
-data in a way that accounts for the larger population from which the data
-are sampled,
-\eqns~\eqref{eq:app_maxent_pop}--\eqref{eq:app_maxent_pop_marg}. This
-application leads to results that differ from the standard application
-focused on the sample only,
+The purpose of the present work was to point out and show, in a simple
+set-up, that the \me\ method can be applied to recorded neuronal data in a
+way that accounts for the larger population from which the data are
+sampled, \eqns~\eqref{eq:app_maxent_pop}--\eqref{eq:app_maxent_pop_marg}.
+This application leads to results that differ from the standard application
+which only considers the sample in isolation,
 \eqns~\eqref{eq:app_maxent_sample}--\eqref{eq:app_maxent_sample_constraints}.
@@
 for $\yNv$ gives unit probability to $\yNv=n$; that is, unless the sample
-\emph{is} the full population, and no larger population exists or the
-recorded neurons are completely isolated from it.
-
-This hidden assumption of isolation is unreasonable. It amounts to say that
-the neurons we distinguished and tracked with our recording device are a
-very special, isolated set among all those that could have been recorded.
-It is for this reason that we find the \me\ application at the population
-level preferable. Physical models of neuronal networks usually include some
-sort of external input, mimicking an embedding in a larger network. The
-\me\ distribution obtained at the population level may reveal features in a
-data set that were unnoticed by the standard \me\ model.
+\emph{is} the full population, and no larger population exists.
+
+The standard \me\ approach therefore assumes that the recorded neurons
+constitute a special subnetwork, isolated from the larger network of
+neurons in which it is embedded, and which was also present under the
+recording device. This assumption is unrealistic. The \me\ approach at the
+population level does not make such assumption and is therefore preferable.
+It may reveal features in a data set that were unnoticed by the standard
+\me\ approach.
+
+% This hidden assumption of isolation is unreasonable. It amounts to say that
+% the neurons we distinguished and tracked with our recording device are a
+% very special, isolated set among all those that could have been recorded.
+% It is for this reason that we find the \me\ application at the population
+% level preferable. Physical models of neuronal networks usually include some
+% sort of external input, mimicking an embedding in a larger network. The
+% \me\ distribution obtained at the population level may reveal features in a
+% data set that were unnoticed by the standard \me\ model.
 
@@
 of its preceding dynamics. In this example, though, both distributions are
-usually immensely sharp and practically lead to the same predictions. In
+usually extremely sharp and practically lead to the same predictions. In
 neuroscientific applications, the difference in predictions of the sample
@@
 \item What is the mathematical relation between the new application and
-  \me\ models with hidden neurons \cite{smolensky1986,hintonetal2006}?
+  \me\ models with hidden neurons
+  \cite{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}?
 \end{itemize}
-Owing to space limitations we cannot investigate these questions here, but
-we hope to in future work.
+Owing to space limitations we must leave a thorough investigation of these
+questions to future work.
 
@@
 formulae is essential in neuroscience, where we try to understand
-properties of extended regions from partial observations of them. The
-formulae presented here reflected a very basic, symmetric state of
-ignorance. It would be useful to explore formulae more appropriate to our
-real knowledge of the brain and of its networking properties.
+properties of extended brain regions from partial observations. The
+formulae presented here reflect a simple, symmetric state of ignorance.
+More work is needed \cite[\cf][]{levinaetal2017} to extend these formulae
+to account for finer knowledge of the cerebral cortex and its network
+properties.
 

### 2018-05-23 maxNt_v3.tex
V3

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-18T15:43:51+0200
+%% Last-Updated: 2018-05-23T09:12:59+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\usepackage[%final,%
-nonatbib]{nips_2017}
+\usepackage[preprint,%
+nonatbib]{nips_2018}
 
 \newcommand*{\firstdraft}{4 November 2015}
+\newcommand*{\published}{22 May 2018}
 
@@
 
+\vspace{-2em}
+\begin{center}
+  \footnotesize\published
+\end{center}
+\medskip
+
 \begin{abstract}
@@
 
-\subsubsection*{Acknowledgments}
-
-To be added after review.
-\iffalse PGLPM thanks Mari \amp\ Miri for continuous encouragement and affection,
-Buster Keaton for filling life with awe and inspiration, and the developers
-and maintainers of \LaTeX, Emacs, AUC\TeX, MiK\TeX, arXiv, biorXiv,
-PhilSci, Hal archives, Python, Inkscape, Sci-Hub for making a free and
-unfiltered scientific exchange possible.\fi
+\subsubsection*{Acknowledgements}
+
+PGLPM thanks Mari \amp\ Miri for continuous
+encouragement and affection; Buster Keaton for filling life with awe and
+inspiration; the developers and maintainers of \LaTeX, Emacs, AUC\TeX, Open
+Science Framework, Python, Inkscape, Sci-Hub for making a free and
+unfiltered scientific exchange possible.
 

### 2018-07-26 maxNt_v3.tex
V3
reply

@@
+\pdfoutput=1
 %% Author: PGL  Porta Mana
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-05-23T09:12:59+0200
+%% Last-Updated: 2018-07-19T11:05:41+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \DeclareDelimFormat{postnotedelim}{\space}
-\addbibresource{portamanabib.bib}
+\addbibresource{portamanabib.bib}%comment for arxiv
 \renewcommand{\bibfont}{\footnotesize}
@@
   time bins and binarized in intensity, and thus transformed into a
-  sequence of digits \enquote{$0$}s (inactive) and \enquote{$1$}s
+  sequence of digits \enquote{$0$}s (inactive) and \enquote{$1$}s (active)
   \cite[\cf][]{caianiello1961,caianiello1986}.
@@
 literature \cite{roudietal2009b}. This problems extends also to more
-complex versions of the method, possibly except version that use
+complex versions of the method, possibly except versions that use
 \enquote{hidden} neurons
@@
 isolated from the rest of the brain. Hence, the results it provides cannot
-rightfully be extrapolated. We will give a mathematical proof of this. Let
+be rightfully extrapolated. We will give a mathematical proof of this. Let
 us first analyse this issue in more detail.
@@
 sample of the activity of the neuronal population under the recording
-device; which can amount to tens of thousands of neurons
+device, which can amount to tens of thousands of neurons
 \cite{abeles1991}. We could even consider the recorded neurons as a sample
@@
 little meaning if we did not expect its results to generalize to a larger,
-unrecorded population -- at the very least that under the recording device.
-In other words, we expect that the conclusions drawn with the \me\ methods
-about the sampled neurons should somehow extrapolate to unrecorded neurons
-in some larger area, from which the recorded neurons were sampled. In
-statistical terms we are assuming that the recorded neurons are a
-\emph{representative sample} of some larger neuronal population.
-Probability theory tells us how to make inferences from a sample to the
-larger population from which it is sampled \parentext{see references
-  below}.
+unrecorded population -- at the very least the population under the
+recording device. In other words, we expect that the conclusions drawn with
+the \me\ methods about the sampled neurons should somehow extrapolate to
+unrecorded neurons in some larger area, from which the recorded neurons
+were sampled. In statistical terms we are assuming that the recorded
+neurons are a \emph{representative sample} of some larger neuronal
+population. Probability theory tells us how to make inferences from a
+sample to the larger population from which it is sampled \parentext{see
+  references below}.
 

### 2018-07-26 nips_2018_author_response.tex
reply

--- 2018-07-26 maxNt_v3.tex
@@
-\pdfoutput=1
-%% Author: PGL  Porta Mana
-%% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-07-19T11:05:41+0200
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%Report-no: ***
 \documentclass{article}
 
-% if you need to pass options to natbib, use, e.g.:
-% \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading nips_2017
-%
-% to avoid loading the natbib package, add option nonatbib:
-% \usepackage[nonatbib]{nips_2017}
-
-\usepackage[preprint,%
-nonatbib]{nips_2018}
+\usepackage{nips_2018_author_response}
 
@@
 \newcommand*{\headtitle}{\pdftitle} \newcommand*{\pdfauthor}{P.G.L.  Porta Mana, V. Rostami, E. Torre, Y. Roudi}
-\usepackage{newtxmath}
-%\usepackage{pifont}
-%\usepackage{fontawesome}
+
 \usepackage[T1]{fontenc} 
@@
 \usepackage[utf8]{inputenx}
-%\usepackage{newunicodechar}
-% \newunicodechar{Ĕ}{\u{E}}
-% \newunicodechar{ĕ}{\u{e}}
-% \newunicodechar{Ĭ}{\u{I}}
-% \newunicodechar{ĭ}{\u{\i}}
-% \newunicodechar{Ŏ}{\u{O}}
-% \newunicodechar{ŏ}{\u{o}}
-% \newunicodechar{Ŭ}{\u{U}}
-% \newunicodechar{ŭ}{\u{u}}
-% \newunicodechar{Ā}{\=A}
-% \newunicodechar{ā}{\=a}
-% \newunicodechar{Ē}{\=E}
-% \newunicodechar{ē}{\=e}
-% \newunicodechar{Ī}{\=I}
-% \newunicodechar{ī}{\={\i}}
-% \newunicodechar{Ō}{\=O}
-% \newunicodechar{ō}{\=o}
-% \newunicodechar{Ū}{\=U}
-% \newunicodechar{ū}{\=u}
-% \newunicodechar{Ȳ}{\=Y}
-% \newunicodechar{ȳ}{\=y}
-
 \newcommand*{\bmmax}{3} % reduce number of bold fonts, before bm
@@
 \setquotestyle{british}
-
-\let\openbox\relax
-\usepackage{amsthm}
-\newcommand*{\QED}{\textsc{q.e.d.}}
-\renewcommand*{\qedsymbol}{\QED}
-\theoremstyle{remark}
-\newtheorem{note}{Note}
-\newtheorem*{remark}{Note}
-\newtheoremstyle{innote}{\parsep}{\parsep}{\footnotesize}{}{}{}{0pt}{}
-\theoremstyle{innote}
-\newtheorem*{innote}{}
-
 \usepackage{bm}
 
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
 \iffalse
@@
 \fi
-
 %% With euler font cursive for Greek letters - the [1] means 100% scaling
@@
 \makeatother
-% Greek letters not usually given in LaTeX. Comment the unneeded ones
-% \DeclareMathSymbol{\varpartial}{\mathalpha}{egreeki}{"40}
-% \DeclareMathSymbol{\partialup}{\mathalpha}{egreekr}{"40}
-% \DeclareMathSymbol{\alpha}{\mathalpha}{egreeki}{"0B}
-% \DeclareMathSymbol{\beta}{\mathalpha}{egreeki}{"0C}
-% \DeclareMathSymbol{\gamma}{\mathalpha}{egreeki}{"0D}
-% \DeclareMathSymbol{\delta}{\mathalpha}{egreeki}{"0E}
-% \DeclareMathSymbol{\epsilon}{\mathalpha}{egreeki}{"0F}
- \DeclareMathSymbol{\zeta}{\mathalpha}{egreeki}{"10}
-% \DeclareMathSymbol{\eta}{\mathalpha}{egreeki}{"11}
-% \DeclareMathSymbol{\theta}{\mathalpha}{egreeki}{"12}
- \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
- \DeclareMathSymbol{\kappa}{\mathalpha}{egreeki}{"14}
- \DeclareMathSymbol{\lambda}{\mathalpha}{egreeki}{"15}
-% \DeclareMathSymbol{\mu}{\mathalpha}{egreeki}{"16}
- \DeclareMathSymbol{\nu}{\mathalpha}{egreeki}{"17}
- \DeclareMathSymbol{\xi}{\mathalpha}{egreeki}{"18}
-% \DeclareMathSymbol{\omicron}{\mathalpha}{egreeki}{"6F}
-% \DeclareMathSymbol{\pi}{\mathalpha}{egreeki}{"19}
-% \DeclareMathSymbol{\rho}{\mathalpha}{egreeki}{"1A}
- \DeclareMathSymbol{\sigma}{\mathalpha}{egreeki}{"1B}
-% \DeclareMathSymbol{\tau}{\mathalpha}{egreeki}{"1C}
-% \DeclareMathSymbol{\upsilon}{\mathalpha}{egreeki}{"1D}
-% \DeclareMathSymbol{\phi}{\mathalpha}{egreeki}{"1E}
-% \DeclareMathSymbol{\chi}{\mathalpha}{egreeki}{"1F}
-% \DeclareMathSymbol{\psi}{\mathalpha}{egreeki}{"20}
-% \DeclareMathSymbol{\omega}{\mathalpha}{egreeki}{"21}
-% \DeclareMathSymbol{\varepsilon}{\mathalpha}{egreeki}{"22}
-% \DeclareMathSymbol{\vartheta}{\mathalpha}{egreeki}{"23}
-% \DeclareMathSymbol{\varpi}{\mathalpha}{egreeki}{"24}
-% \let\varrho\rho 
-% \let\varsigma\sigma
-% \let\varkappa\kappa
-% \DeclareMathSymbol{\varphi}{\mathalpha}{egreeki}{"27}
-% %
-% \DeclareMathSymbol{\varAlpha}{\mathalpha}{egreeki}{"41}
-% \DeclareMathSymbol{\varBeta}{\mathalpha}{egreeki}{"42}
-% \DeclareMathSymbol{\varGamma}{\mathalpha}{egreeki}{"00}
-% \DeclareMathSymbol{\varDelta}{\mathalpha}{egreeki}{"01}
-% \DeclareMathSymbol{\varEpsilon}{\mathalpha}{egreeki}{"45}
- \DeclareMathSymbol{\varZeta}{\mathalpha}{egreeki}{"5A}
-% \DeclareMathSymbol{\varEta}{\mathalpha}{egreeki}{"48}
-% \DeclareMathSymbol{\varTheta}{\mathalpha}{egreeki}{"02}
- \DeclareMathSymbol{\varIota}{\mathalpha}{egreeki}{"49}
- \DeclareMathSymbol{\varKappa}{\mathalpha}{egreeki}{"4B}
- \DeclareMathSymbol{\varLambda}{\mathalpha}{egreeki}{"03}
-% \DeclareMathSymbol{\varMu}{\mathalpha}{egreeki}{"4D}
- \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
- \DeclareMathSymbol{\varXi}{\mathalpha}{egreeki}{"04}
-% \DeclareMathSymbol{\varOmicron}{\mathalpha}{egreeki}{"4F}
-% \DeclareMathSymbol{\varPi}{\mathalpha}{egreeki}{"05}
-% \DeclareMathSymbol{\varRho}{\mathalpha}{egreeki}{"50}
- \DeclareMathSymbol{\varSigma}{\mathalpha}{egreeki}{"06}
-% \DeclareMathSymbol{\varTau}{\mathalpha}{egreeki}{"54}
- \DeclareMathSymbol{\varUpsilon}{\mathalpha}{egreeki}{"07}
-% \DeclareMathSymbol{\varPhi}{\mathalpha}{egreeki}{"08}
-% \DeclareMathSymbol{\varChi}{\mathalpha}{egreeki}{"58}
-% \DeclareMathSymbol{\varPsi}{\mathalpha}{egreeki}{"09}
-% \DeclareMathSymbol{\varOmega}{\mathalpha}{egreeki}{"0A} 
-% %
-% \DeclareMathSymbol{\Alpha}{\mathalpha}{egreekr}{"41}
-% \DeclareMathSymbol{\Beta}{\mathalpha}{egreekr}{"42}
-% \DeclareMathSymbol{\Gamma}{\mathalpha}{egreekr}{"00}
-% \DeclareMathSymbol{\Delta}{\mathalpha}{egreekr}{"01}
-% \DeclareMathSymbol{\Epsilon}{\mathalpha}{egreekr}{"45}
-% \DeclareMathSymbol{\Zeta}{\mathalpha}{egreekr}{"5A}
-% \DeclareMathSymbol{\Eta}{\mathalpha}{egreekr}{"48}
-% \DeclareMathSymbol{\Theta}{\mathalpha}{egreekr}{"02}
-% \DeclareMathSymbol{\Iota}{\mathalpha}{egreekr}{"49}
-% \DeclareMathSymbol{\Kappa}{\mathalpha}{egreekr}{"4B}
-% \DeclareMathSymbol{\Lambda}{\mathalpha}{egreekr}{"03}
-% \DeclareMathSymbol{\Mu}{\mathalpha}{egreekr}{"4D}
-% \DeclareMathSymbol{\Nu}{\mathalpha}{egreekr}{"4E}
-% \DeclareMathSymbol{\Xi}{\mathalpha}{egreekr}{"04}
-% \DeclareMathSymbol{\Omicron}{\mathalpha}{egreekr}{"4F}
-% \DeclareMathSymbol{\Pi}{\mathalpha}{egreekr}{"05}
-% \DeclareMathSymbol{\Rho}{\mathalpha}{egreekr}{"50}
-% \DeclareMathSymbol{\Sigma}{\mathalpha}{egreekr}{"06}
-% \DeclareMathSymbol{\Tau}{\mathalpha}{egreekr}{"54}
-% \DeclareMathSymbol{\Upsilon}{\mathalpha}{egreekr}{"07}
-% \DeclareMathSymbol{\Phi}{\mathalpha}{egreekr}{"08}
-% \DeclareMathSymbol{\Chi}{\mathalpha}{egreekr}{"58}
-% \DeclareMathSymbol{\Psi}{\mathalpha}{egreekr}{"09}
-% \DeclareMathSymbol{\Omega}{\mathalpha}{egreekr}{"0A}
-% %
-% \DeclareMathSymbol{\alphaup}{\mathalpha}{egreekr}{"0B}
-% \DeclareMathSymbol{\betaup}{\mathalpha}{egreekr}{"0C}
-% \DeclareMathSymbol{\gammaup}{\mathalpha}{egreekr}{"0D}
-% \DeclareMathSymbol{\deltaup}{\mathalpha}{egreekr}{"0E}
-% \DeclareMathSymbol{\epsilonup}{\mathalpha}{egreekr}{"0F}
-% \DeclareMathSymbol{\zetaup}{\mathalpha}{egreekr}{"10}
-% \DeclareMathSymbol{\etaup}{\mathalpha}{egreekr}{"11}
-% \DeclareMathSymbol{\thetaup}{\mathalpha}{egreekr}{"12}
-% \DeclareMathSymbol{\iotaup}{\mathalpha}{egreekr}{"13}
-% \DeclareMathSymbol{\kappaup}{\mathalpha}{egreekr}{"14}
-% \DeclareMathSymbol{\lambdaup}{\mathalpha}{egreekr}{"15}
-% \DeclareMathSymbol{\muup}{\mathalpha}{egreekr}{"16}
-% \DeclareMathSymbol{\nuup}{\mathalpha}{egreekr}{"17}
-% \DeclareMathSymbol{\xiup}{\mathalpha}{egreekr}{"18}
-% \DeclareMathSymbol{\omicronup}{\mathalpha}{egreekr}{"6F}
-%  \DeclareMathSymbol{\piup}{\mathalpha}{egreekr}{"19}
-% \DeclareMathSymbol{\rhoup}{\mathalpha}{egreekr}{"1A}
-% \DeclareMathSymbol{\sigmaup}{\mathalpha}{egreekr}{"1B}
-% \DeclareMathSymbol{\tauup}{\mathalpha}{egreekr}{"1C}
-% \DeclareMathSymbol{\upsilonup}{\mathalpha}{egreekr}{"1D}
-% \DeclareMathSymbol{\phiup}{\mathalpha}{egreekr}{"1E}
-% \DeclareMathSymbol{\chiup}{\mathalpha}{egreekr}{"1F}
-% \DeclareMathSymbol{\psiup}{\mathalpha}{egreekr}{"20}
-% \DeclareMathSymbol{\omegaup}{\mathalpha}{egreekr}{"21}
-% \DeclareMathSymbol{\varepsilonup}{\mathalpha}{egreekr}{"22}
-% \DeclareMathSymbol{\varthetaup}{\mathalpha}{egreekr}{"23}
-% \DeclareMathSymbol{\varpiup}{\mathalpha}{egreekr}{"24}
-% \let\varrhoup\rhoup 
-% \let\varsigmaup\sigmaup
-% \let\varkappaup\kappaup
-% \DeclareMathSymbol{\varphiup}{\mathalpha}{egreekr}{"27}
-
-%\newcommand*{\mathte}[1]{\textbf{\textit{\textsf{#1}}}}
-% Upright sans-serif math alphabet
-% \DeclareMathAlphabet{\mathsu}  {T1}{\sfdefault}{m}{n}
-% \SetMathAlphabet{\mathsu}{bold}{T1}{\sfdefault}{b}{n}
 
 \usepackage{mathdots}
-
-\usepackage[usenames]{xcolor}
-\definecolor{mybluishpurple}{RGB}{51,34,136}
-\definecolor{myblue}{RGB}{136,204,238}
-\definecolor{mybluishgreen}{RGB}{68,170,153}
-\definecolor{mygreen}{RGB}{17,119,51}
-\definecolor{mygreenishyellow}{RGB}{153,153,51}
-\definecolor{myyellow}{RGB}{221,204,119}
-\definecolor{myred}{RGB}{204,102,119}
-\definecolor{mypurplishred}{RGB}{136,34,85}
-\definecolor{myreddishpurple}{RGB}{170,68,153}
-\definecolor{mygrey}{RGB}{221,221,221}
-
 \usepackage{microtype}
@@
 }
-
 \usepackage{graphicx}
@@
 
-%\usepackage{wrapfig}
-\usepackage{datetime2}
-\DTMnewdatestyle{mydate}%
-{% definitions
-\renewcommand*{\DTMdisplaydate}[4]{%
-\number##3\ \DTMenglishmonthname{##2} ##1}%
-\renewcommand*{\DTMDisplaydate}{\DTMdisplaydate}%
-}
-\DTMsetdatestyle{mydate}
-
-% handling orphan/widow lines:
-\clubpenalty=10000
-\widowpenalty=10000
-\raggedbottom
-
 \selectlanguage{british}\frenchspacing
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage[final]{nips_2017}
-
-%\usepackage[utf8]{inputenc} % allow utf-8 input
-%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-%\usepackage{hyperref}       % hyperlinks
-%\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-%\usepackage{amsfonts}       % blackboard math symbols
-%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-%\usepackage{microtype}      % microtypography
-
-% @@@@@@@@@@ new macros @@@@@@@@@@
-% Common ones - uncomment as needed
-%\providecommand{\nequiv}{\not\equiv}
-%\providecommand{\coloneqq}{\mathrel{\mathop:}=}
-%\providecommand{\eqqcolon}{=\mathrel{\mathop:}}
-%\providecommand{\varprod}{\prod}
-%\newcommand*{\de}{\partialup}%partial diff
-%\newcommand*{\pu}{\piup}%constant pi
-\newcommand*{\delt}{\deltaup}%Kronecker, Dirac
-%\newcommand*{\eps}{\varepsilonup}%Levi-Civita, Heaviside
-%\newcommand*{\riem}{\zetaup}%Riemann zeta
-%\providecommand{\degree}{\textdegree}% degree
-%\newcommand*{\celsius}{\textcelsius}% degree Celsius
-%\newcommand*{\micro}{\textmu}% degree Celsius
-%\newcommand*{\I}{\mathrm{i}}%imaginary unit
-%\newcommand*{\e}{\mathrm{e}}%Neper
-%\newcommand*{\di}{\mathrm{d}}%differential
-%\newcommand*{\Di}{\mathrm{D}}%capital differential
-%\newcommand*{\planckc}{\hslash}
-%\newcommand*{\avogn}{N_{\textrm{A}}}
-%\newcommand*{\NN}{\bm{\mathrm{N}}}
-%\newcommand*{\ZZ}{\bm{\mathrm{Z}}}
-%\newcommand*{\QQ}{\bm{\mathrm{Q}}}
-\newcommand*{\RR}{\bm{\mathrm{R}}}
-%\newcommand*{\CC}{\bm{\mathrm{C}}}
-%\newcommand*{\nabl}{\bm{\nabla}}%nabla
-%\DeclareMathOperator{\lb}{lb}%base 2 log
-%\DeclareMathOperator{\tr}{tr}%trace
-%\DeclareMathOperator{\card}{card}%cardinality
-%\DeclareMathOperator{\im}{Im}%im part
-%\DeclareMathOperator{\re}{Re}%re part
-%\DeclareMathOperator{\sgn}{sgn}%signum
-%\DeclareMathOperator{\ent}{ent}%integer less or equal to
-%\DeclareMathOperator{\Ord}{O}%same order as
-%\DeclareMathOperator{\ord}{o}%lower order than
-%\newcommand*{\incr}{\triangle}%finite increment
-\newcommand*{\defd}{\coloneqq}
-\newcommand*{\defs}{\eqqcolon}
-%\newcommand*{\Land}{\bigwedge}
-%\newcommand*{\Lor}{\bigvee}
-%\newcommand*{\lland}{\mathbin{\ \land\ }}
-%\newcommand*{\llor}{\mathbin{\ \lor\ }}
-%\newcommand*{\lonlyif}{\mathbin{\Rightarrow}}%implies
-\newcommand*{\limplies}{\mathbin{\Rightarrow}}%implies
-%\newcommand*{\mimplies}{\Rightarrow}%implies
-%\newcommand*{\liff}{\mathbin{\Leftrightarrow}}%if and only if
-\renewcommand*{\|}{\mathpunct{|}}%conditional sign (in probabilities)
-%\newcommand*{\lcond}{\mathpunct{|\ }}%conditional sign (in probabilities)
-%\newcommand*{\bigcond}{\mathpunct{\big|}}%conditional sign (in probabilities)
-%\newcommand*{\lbigcond}{\mathpunct{\big|\ }}%conditional sign (in probabilities)
-%\newcommand*{\suchthat}{\mid}%{\mathpunct{|}}%such that (eg in sets)
-%\newcommand*{\bigst}{\mathpunct{\big|}}%such that (eg in sets)
-%\newcommand*{\with}{\colon}%with (list of indices)
-%\newcommand*{\mul}{\times}%multiplication
-%\newcommand*{\inn}{\cdot}%inner product
-\newcommand*{\dotv}{\mathord{\,\cdot\,}}%variable place
-%\newcommand*{\comp}{\circ}%composition of functions
-%\newcommand*{\con}{\mathbin{:}}%scal prod of tensors
-%\newcommand*{\equi}{\sim}%equivalent to 
-%\newcommand*{\corr}{\mathrel{\hat{=}}}%corresponds to
-%\providecommand{\varparallel}{\ensuremath{\mathbin{/\mkern-7mu/}}}%parallel (tentative symbol)
-\renewcommand{\le}{\leqslant}%less or equal
-\renewcommand{\ge}{\geqslant}%greater or equal
-\DeclarePairedDelimiter\clcl{[}{]}
-%\DeclarePairedDelimiter\clop{[}{[}
-%\DeclarePairedDelimiter\opcl{]}{]}
-%\DeclarePairedDelimiter\opop{]}{[}
-%\DeclarePairedDelimiter\abs{\lvert}{\rvert}
-%\DeclarePairedDelimiter\norm{\lVert}{\rVert}
-\DeclarePairedDelimiter\set{\{}{\}}
-%\DeclareMathOperator{\pr}{P}%probability
-\newcommand*{\pf}{\mathrm{p}}%probability
-\newcommand*{\p}{\mathrm{P}}%probability
-%\newcommand*{\tf}{\mathrm{T}}%probability
-%\renewcommand*{\|}{\|}
-%\newcommand*{\+}{\lor}
-%\renewcommand{\*}{\land}
-\newcommand*{\sect}{\S}% Sect.~
-\newcommand*{\sects}{\S\S}% Sect.~
-\newcommand*{\chap}{ch.}%
-\newcommand*{\chaps}{chs}%
-%\newcommand*{\fn}{fn}%
-\newcommand*{\eqn}{eq.}%
-\newcommand*{\eqns}{eqs}%
-\newcommand*{\fig}{fig.}%
-\newcommand*{\figs}{figs}%
-\newcommand*{\vs}{{vs}}
-%\newcommand*{\etc}{{etc.}}
-%\newcommand*{\ie}{{i.e.}}
-%\newcommand*{\ca}{{c.}}
-%\newcommand*{\Ie}{{I.e.}}
-%\newcommand*{\Eg}{{E.g.}}
-%\newcommand*{\eg}{{e.g.}}
-%\newcommand*{\viz}{{viz}}
-\newcommand*{\cf}{{cf.}}
-%\newcommand*{\Cf}{{Cf.}}
-%\newcommand*{\vd}{{v.}}
-%\newcommand*{\Vd}{{V.}}
-\newcommand*{\etal}{{et al.}}
-%\newcommand*{\etsim}{{et sim.}}
-%\newcommand*{\ibid}{{ibid.}}
-%\newcommand*{\sic}{{sic}}
-%\newcommand*{\id}{\mathte{I}}%id matrix
-%\newcommand*{\nbd}{\nobreakdash}%
-%\newcommand*{\bd}{\hspace{0pt}}%
-%\def\hy{-\penalty0\hskip0pt\relax}
-\newcommand*{\labelbis}[1]{\tag*{(\ref{#1})$_\text{r}$}}
-%\newcommand*{\mathbox}[2][.8]{\parbox[t]{#1\columnwidth}{#2}}
-%\newcommand*{\zerob}[1]{\makebox[0pt][l]{#1}}
-%\newcommand*{\tprod}{\mathop{\textstyle\prod}\nolimits}
-%\newcommand*{\tsum}{\mathop{\textstyle\sum}\nolimits}
-%\newcommand*{\tint}{\begingroup\textstyle\int\endgroup\nolimits}
-%\newcommand*{\tland}{\mathop{\textstyle\bigwedge}\nolimits}
-%\newcommand*{\tlor}{\mathop{\textstyle\bigvee}\nolimits}
-%\newcommand*{\sprod}{\mathop{\textstyle\prod}}
-%\newcommand*{\ssum}{\mathop{\textstyle\sum}}
-%\newcommand*{\sint}{\begingroup\textstyle\int\endgroup}
-%\newcommand*{\sland}{\mathop{\textstyle\bigwedge}}
-%\newcommand*{\slor}{\mathop{\textstyle\bigvee}}
-\newcommand*{\T}{^\intercal}%transpose
-\newcommand*{\E}{\mathrm{E}}
-\DeclarePairedDelimiter\expp{(}{)}
-\newcommand*{\expe}{\E\expp}%round
-\newcommand*{\expeb}{\E\clcl}%square
-%%\newcommand*{\QEM}%{\textnormal{$\Box$}}%{\ding{167}}
-%\newcommand*{\qem}{\leavevmode\unskip\penalty9999 \hbox{}\nobreak\hfill
-%\quad\hbox{\QEM}}
-
-\newtheoremstyle{simple}%
-{}%.5\baselineskip plusminus.2\baselineskip % \parsep
-{}%
-{\footnotesize}%
-{}%
-{}%
-{}%
-{0pt}%
-{}%
-\theoremstyle{simple}
-\newtheorem*{simplenote}{}
-
-\definecolor{notecolour}{RGB}{68,170,153}
-\newcommand*{\puzzle}{{\fontencoding{U}\fontfamily{fontawesometwo}\selectfont\symbol{225}}}
-\newcommand{\mynote}[1]{ {\color{notecolour}\puzzle\ #1}}
-\newcommand*{\widebar}[1]{{\mkern1.5mu\skew{2}\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}}
-
-\newcommand*{\tav}{\widehat} %time average
-\newcommand*{\av}{\widebar} %pop average
-\newcommand*{\sav}{\widebar} %subpop average
-
-\newcommand*{\ypp}{\varPi}
-
-\newcommand*{\yXv}{\varSigma}
-\newcommand*{\yRv}{\sigma}
-\newcommand*{\yxv}{S}
-\newcommand*{\yrv}{s}
-\newcommand*{\yNv}{\nu}
-\newcommand*{\yNN}{\varNu}
-
-\newcommand*{\yx}{\bm{\yxv}}%subpop state
-\newcommand*{\yxs}{\sav{\yx}}%subpop av state
-\newcommand*{\yX}{\bm{\yXv}}%pop state
-\newcommand*{\yXf}{\av{\yX}}%pop av state
-\newcommand*{\yxxs}{\sav{\yr\yr}}%subpop av state
-\newcommand*{\yXXf}{\av{\yR\yR}}%subpop av state
-\newcommand*{\yr}{\bm{\yrv}}%subpop value
-\newcommand*{\yrs}{\sav{\yr}}%subpop av value
-\newcommand*{\yR}{\bm{\yRv}}%pop value
-\newcommand*{\yRf}{\av{\yR}}%pop av value
-%conditional assumptions
-\newcommand*{\yH}{\varIota}
-\newcommand*{\yD}{\varDelta}
-\newcommand*{\yHa}{\varIota_\textrm{p}}
-\newcommand*{\yHb}{\varIota_\textrm{s}}
-\newcommand*{\yHc}{\varKappa}
-\newcommand*{\yHd}{\varIota_\textrm{u}}
-\newcommand*{\yHp}{\varIota'}
-\newcommand*{\yHi}{\varIota''}
-%Lagr multipliers
-\newcommand*{\yf}{\bm{g}}
-\newcommand*{\yc}{\bm{c}}
-\newcommand*{\yL}{\bm{\lambda}}
-\newcommand*{\yl}{\bm{l}}
-\newcommand*{\yk}{z}
-\newcommand*{\yK}{\zeta}
-%entropies
-\newcommand*{\ysh}{H}
-\newcommand*{\ybu}{H_{\text{B}}}
-
-\newcommand*{\prop}[1]{`#1'}
-
-\newcommand*{\mee}{maximum entropy}
-\newcommand*{\me}{maximum-entropy}
-
-%@@@@@@@@@@ new macros end @@@@@@@@@@
-
-\selectlanguage{british}\frenchspacing
-
-\title{\pdftitle}
-
-% The \author macro works with any number of authors. There are two
-% commands used to separate the names and addresses of multiple
-% authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to
-% break the lines. Using \AND forces a line break at that point. So,
-% if LaTeX puts 3 of 4 authors names on the first line, and the last
-% on the second line, try using \AND instead of \And before the third
-% author name.
-
-\author{
-  P.G.L. Porta Mana\\\texttt{piero.mana@ntnu.no}
-  \And
-  V. Rostami\\\texttt{vrostami@uni-koeln.de}
-  \AND
-  E. Torre\\\texttt{torre@ibk.baug.ethz.ch}
-    \And
-Y. Roudi\\\texttt{yasser.roudi@ntnu.no}
-  % David S.~Hippocampus\thanks{Use footnote for providing further
-  %   information about author (webpage, alternative
-  %   address)---\emph{not} for acknowledging funding agencies.} \\
-  % Department of Computer Science\\
-  % Cranberry-Lemon University\\
-  % Pittsburgh, PA 15213 \\
-  % \texttt{hippo@cs.cranberry-lemon.edu} \\
-  %% examples of more authors
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \AND
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-}
+% \usepackage{booktabs}       % professional-quality tables
+% \usepackage{amsfonts}       % blackboard math symbols
+% \usepackage{nicefrac}       % compact symbols for 1/2, etc.
+%\usepackage{lipsum}
 
 \begin{document}
-% \nipsfinalcopy is no longer used
-
-\maketitle
-
-\vspace{-2em}
-\begin{center}
-  \footnotesize\published
-\end{center}
-\medskip
-
-\begin{abstract}
-  The present work shows that the \me\ method can be applied to a sample of
-  neuronal recordings along two different routes: (1) apply to the sample;
-  or (2) apply to a larger, unsampled neuronal population from which the
-  sample is drawn, and then marginalize to the sample. These two routes
-  give inequivalent results. The second route can be further generalized to
-  the case where the size of the larger population is unknown. Which route
-  should be chosen? Some arguments are presented in favour of the second.
-  This work also presents and discusses probability formulae that relate
-  states of knowledge about a population and its samples, and that may be
-  useful for sampling problems in neuroscience.
-  % The abstract must be limited to one paragraph.
-\end{abstract}
-
-%\section{Implicit assumptions in the maximum-entropy method}
-\section{Introduction: \me\ and recordings of neuronal activity}
-\label{sec:intro}
-
-Suppose that we have recorded the firing activity of a hundred neurons,
-sampled from a particular brain area. What are we to do with such data?
-Gerstein, Perkel, Dayhoff \cite{gersteinetal1985} posed this question very
-tersely (our emphasis):
-\begin{quote}
-  The principal conceptual problems are (1) \emph{defining cooperativity or
-    functional grouping} among neurons and (2) \emph{formulating
-    quantitative criteria} for recognizing and characterizing such
-  cooperativity.
-\end{quote}
-These questions have a long history, of course; see for instance the 1966
-review by Moore \etal\
-\cite{mooreetal1966}. % This scenario is a concrete possibility
-% thanks to recent electrophysiological techniques \cite{berenyietal2014}.
-The neuroscientific literature has offered several mathematical definitions
-of \enquote{cooperativity} or \enquote{functional grouping} and criteria to
-quantify it.
-
-One such quantitative criterion relies on the \me\ or relative-\me\ method
-\cite{jaynes1957,jaynes1963,hobsonetal1973,sivia1996_r2006,meadetal1984}.
-This criterion has been used in neuroscience at least since the 1990s,
-applied to data recorded from brain areas as diverse as retina and motor
-cortex
-\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015},
-and it has been subjected to mathematical and conceptual scrutiny
-\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.
-
-\enquote{Cooperativity} can be quantified and characterized with \me\
-methods in several ways. The simplest way roughly proceeds along the
-following steps. Consider the recorded activity of a sample of $n$ neurons.
+We thank the reviewers for their analysis and suggestions. We would like to
+reply to specific points:
 \begin{enumerate}
-\item The activity of each neuron, a continuous signal, is divided into $T$
-  time bins and binarized in intensity, and thus transformed into a
-  sequence of digits \enquote{$0$}s (inactive) and \enquote{$1$}s (active)
-  \cite[\cf][]{caianiello1961,caianiello1986}.
-
-  Let the variable $\yrv_i(t)\in\set{0,1}$ denote the activity of the $i$th
-  sampled neuron at time bin $t$. Collectively denote the $n$ activities
-  with $\yr(t) \defd \bigl(\yrv_1(t),\dotsc,\yrv_n(t)\bigr)$. The
-  population-averaged activity at that bin is
-  $\yrs(t) \defd \sum_i \yrv_i(t)/n$. If we count the number of distinct
-  pairs of active neurons at that bin we combinatorially find
-  $\tbinom{n\yrs(t)}{2}\equiv n\yrs(t)\,[n\yrs(t)-1]/2$. There can be at
-  most $\tbinom{n}{2}$ simultaneously active pairs, so the
-  population-averaged pair activity is
-  $\av{\yr \yr}(t) \defd \tbinom{n}{2}^{-1}\tbinom{n\yrs(t)}{2}$. With some
-  combinatorics we see that the population-averaged activity of $m$-tuples
-  of neurons is
-  \begin{equation}
-    \label{eq:products_intermsof_average}
-    \av{\underbrace{\yr\dotsm\yr}_{\text{$m$ terms}}}(t)
-    = \binom{n}{m}^{-1}\binom{n\yrs(t)}{m}.
-  \end{equation}
-
-  For brevity let us agree to simply call \enquote{activity} the average
-  $\yrs$, \enquote{pair-activity} the average $\av{\yr \yr}$, and so on.
-
-\item Construct a sequence of relative-\me\ distributions for the activity
-  $\yrs$, using this sequence of constraints:
-  \begin{itemize}
-  \item the time average of the activity: $\tav{\yrs} \defd \sum_t\yrs(t)/T$;
-  \item the time averages of the activity and of the pair-activity
-    $\tav{\av{\yr \yr}} \defd \sum_t\av{\yr \yr}(t)/T$;
-  \item \ldots
-  \item the time averages of the activity, of the pair-activity, and so on, up
-    to the $k$-activity.
-  \end{itemize}
-  Call the resulting distributions $p_1(\yrs), p_2(\yrs),\dotsc,p_k(\yrs)$.
-  The time-bin dependence is now absent because these distributions can be
-  interpreted as referring to any one of the time bins $t$, or to a new
-  time bin (in the future or in the past) containing new data.
-
-  We also have the empirical frequency distribution of the total activity,
-  $f(\yrs)$, counted from the time bins.
-
-\item Now compare the distributions above with one another and with the
-  frequency distribution, using some probability-space distance like the
-  relative entropy or discrimination information
-  \citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}. If we find,
-  say, that such distance is very high between $p_1$ and $f$, very low
-  between $p_2$ and $f$, and is more or less the same between all $p_m$ and
-  $f$ for $m \ge 2$, then we can say that there is a \enquote{pairwise
-    cooperativity}, and that any higher-order cooperativity is just a
-  reflection or consequence of the pairwise one. The reason is that the
-  information from higher-order simultaneous activities did not lead to
-  appreciable changes in the distribution obtained from pair activities.
+\item The 
 \end{enumerate}
-The protocol above needs to be made precise by specifying various
-parameters, such as the width of the time bins or the probability
-distance used.
-
-We hurry to say that the description just given is just \emph{one} way to
-quantify and characterize cooperativity and functional grouping, not
-\emph{the only} way. It can surely be criticized from many points of view.
-Yet, it is quantitative and bears a more precise meaning than an undefined,
-vague notion of \enquote{cooperativity}. Two persons who apply this
-procedure to the same data will obtain the same numbers. Different
-protocols can be based on the \me\ method, for instance protocols that take
-into account the activities or pair activities of specific neurons rather
-than population averages, or even protocols that take into account time
-dependence.
-
-The purpose of the present work is not to assess the merits of \me\ methods
-with respect to other methods. Its main purpose is to show that there is a
-problem in the way the \me\ method itself, as sketched above, is applied to
-the activity of the recorded neurons. We believe that this problem is at
-the root of some quirks about this method that were pointed out in the
-literature \cite{roudietal2009b}. This problems extends also to more
-complex versions of the method, possibly except versions that use
-\enquote{hidden} neurons
-\cite{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}. The problem is
-that the recorded neurons are a \emph{sample} from a larger, unrecorded
-population, but the \me\ method as applied above is treating them as
-isolated from the rest of the brain. Hence, the results it provides cannot
-be rightfully extrapolated. We will give a mathematical proof of this. Let
-us first analyse this issue in more detail.
-
-Suppose that the neurons were recorded with electrodes covering an area of
-some square millimetres \cite[\cf][]{berenyietal2014}. This recording is a
-sample of the activity of the neuronal population under the recording
-device, which can amount to tens of thousands of neurons
-\cite{abeles1991}. We could even consider the recorded neurons as a sample
-of a brain area more extended than the recording device.
-
-The characterization of the cooperativity of the recorded sample would have
-little meaning if we did not expect its results to generalize to a larger,
-unrecorded population -- at the very least the population under the
-recording device. In other words, we expect that the conclusions drawn with
-the \me\ methods about the sampled neurons should somehow extrapolate to
-unrecorded neurons in some larger area, from which the recorded neurons
-were sampled. In statistical terms we are assuming that the recorded
-neurons are a \emph{representative sample} of some larger neuronal
-population. Probability theory tells us how to make inferences from a
-sample to the larger population from which it is sampled \parentext{see
-  references below}.
-
-We can apply the \me\ method to the sample, as described in the above
-protocol, to generate probability distributions for the activity of the
-sample. But, given that our sample is representative of a larger
-population, we can also apply the \me\ method to the larger (unrecorded)
-population. The constraints are the same: the time averages of the sampled
-data, since they constitute representative data about the larger population
-as well. The method thus yields a probability distribution for the larger
-population, and the distribution for the sample is obtained by
-marginalization. The problem is that \emph{the distributions obtained from
-  these two applications differ}. Which choice is most meaningful?
-
-In this work we develop the second way of applying the \me\ method, at the
-level of the larger population, and show that its results differ from the
-application at the sample level. We also consider the case where the size
-of the larger population is unknown.
-
-To apply the \me\ method to the larger, unsampled population, it is
-necessary to use probability relations relevant to sampling
-\cites{ghoshetal1997}[parts~I,
-VI]{freedmanetal1978_r2007}[\chap~8]{gelmanetal1995_r2014}[\chap~3]{jaynes1994_r2003}.
-The relations we present are well-known in survey sampling and in the
-pedagogic problem of drawing from an urn without replacement, yet they are
-somewhat hard to find explicitly written in the neuroscientific literature.
-We present and discuss them in the next section. A minor purpose of this
-paper is to make these relations more widely known, because they can be
-useful independently of \me\ methods.
-
-The notation and terminology in the present work follow ISO and ANSI standards
-\citep{iso1993,ieee1993,nist1995,iso2006,iso2006b} but for the use of the
-comma \enquote{,} to denote logical conjunction. Probability notation
-follows Jaynes \citep{jaynes1994_r2003}. By \enquote{probability} we mean a
-degree of belief which \enquote{would be agreed by all rational men if
-  there were any rational men} \cite{good1966}.
-
-\section{Probability relations between population and  sample}
-\label{sec:prob_samples}
-
-We have already introduced the notation for the sample neurons. We
-introduce an analogous notation for the $\yNv$ neurons constituting the
-larger population, but using the corresponding Greek letters:
-$\yRv_{\iota}(t)$ is the activity of the $\iota$th neuron at time bin $t$,
-$\yRf(t) \defd \sum_{\iota} \yRv_{\iota}(t)/\yNv$ is the activity at that
-bin averaged over the larger population, and so on.
-
-The probability relations between sample and larger population are valid at
-every time bin. As we mentioned above, the \me\ distribution refers to any
-time bin or to a new bin. For these reasons we will now omit the time-bin
-argument \enquote{(t)} from our expressions.
-
-Probabilities refer to statements about the quantities we observe. We use
-the standard notation:
-\begin{equation}
-  \label{eq:notation_statements}
-  \begin{aligned}
-    &\text{\enquote{$\yXv_{\iota} = \yRv_{\iota}$} 
-      stands for \enquote{the activity of the $\iota$th neuron is $\yRv_{\iota}$}},
-    \\
-    &\text{\enquote{$\yXf = \yRf$} 
-      stands for \enquote{the (population-averaged) activity of the neurons is $\yRf$}},
-    \\
-    &\text{\enquote{$\yxv_i = \yrv_i$} 
-      stands for \enquote{the activity of the $i$th sample neuron is $\yrv_i$}},
-  \end{aligned}
-\end{equation}
-and similarly for other quantities.
-
-If $\yHc$ denotes our state of knowledge -- the evidence and assumptions
-backing our probability assignments -- our uncertainty about the full
-activity of the larger population is expressed by the joint probability
-distribution
-\begin{equation}
-  \label{eq:joint_plaus}
-  \p(\yXv_1=\yRv_1, \yXv_2=\yRv_2, \dotsc, \yXv_\yNv=\yRv_\yNv \| \yHc)
-  \quad\text{or}\quad
-\p(\yX =\yR \| \yHc), \quad \yR \in \set{0,1}^\yNv.
-\end{equation}
-Our uncertainty about the state of the sample is likewise expressed by
-\begin{equation}
-  \label{eq:sample_plaus}
-  \p(\yxv_1=\yrv_1, \yxv_2=\yrv_2, \dotsc, \yxv_n=\yrv_n \| \yHc) \quad\text{or}\quad
-\p(\yx =\yr \| \yHc), \quad \yr \in \set{0,1}^n.
-\end{equation}
-
-\bigskip
-
-The theory of statistical sampling is covered in many excellent texts, for
-example Ghosh \amp\ Meeden \citey{ghoshetal1997} or Freedman, Pisani, \amp\
-Purves \citey[parts~I, VI]{freedmanetal1978_r2007}; summaries can be found
-in Gelman \etal\ \citey[\chap~8]{gelmanetal1995_r2014} and Jaynes
-\citey[\chap~3]{jaynes1994_r2003}.
-
-We need to make an initial probability assignment for the state of the full
-population before any experimental observations are made. This initial
-assignment will be modified by our experimental observations, and these can
-involve just a sample of the population. Our state of knowledge and initial
-probability assignment should reflect that samples are somehow
-representative of the whole population.
-
-In this state of knowledge, denoted $\yH$, we know that the neurons in the
-population are biologically or functionally similar, for example in
-morphology or the kind of input or output they receive or give. But we are
-completely ignorant about the physical details of the individual neurons.
-Our ignorance is therefore symmetric under permutations of neuron
-identities. This ignorance is represented by a probability distribution
-that is symmetric under permutations of neuron identities; such a
-distribution is usually called \emph{finitely exchangeable}
-\cites{ericson1969}[\chap~1]{ghoshetal1997}. We stress that this
-probability assignment is just an expression of the symmetry of our
-\emph{ignorance} about the state of the population, not an expression of
-some biologic or physical symmetry or identity of the neurons.
-
-The \emph{representation theorem for finite exchangeability} states that,
-in the state of knowledge $\yH$, the symmetric distribution for the full
-activity is completely determined by the distribution for its
-population-average:
-\begin{equation}
-  \label{eq:joint_plaus_N_homog}
-  \p(\yX = \yR \|  \yH) \equiv
-  \sum_{\yRf}\p(\yX = \yR \| \yXf = \yRf, \yH)\,
-  \p(\yXf=\yRf \| \yH) =
-  \binom{\yNv}{\yNv\yRf}^{-1} \p(\yXf=\yRf \| \yH).
-\end{equation}
-The equivalence on the left is just an application of the law of total
-probability; the equality on the right is the statement of the theorem.
-This result is intuitive: owing to symmetry, we must assign equal
-probabilities to all $\tbinom{\yNv}{\yNv\yRf}$ activity vectors with
-$\yNv\yRf$ active neurons; the probability of each activity vector is
-therefore given by that of the average activity divided by the number of
-possible vector values. Proof of this theorem and generalizations to
-non-binary and continuum cases are given by de~Finetti
-\cite{definetti1959b}, Kendall \cite{kendall1967}, Ericson
-\cite{ericson1976}, Diaconis \amp\ Freedman
-\cite{diaconis1977,diaconisetal1980}, Heath \amp\ Sudderth
-\cite{heathetal1976}.
-
-Our uncertainties about the full population and the sample are connected
-via the conditional probability
-\begin{equation}
-  \label{eq:conditional_hypergeometric}
-  \p(\yxs = \yrs \|\yXf=\yRf, \yH)=
-  \binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}
-  \defs \ypp(\yrs\|\yRf),
-\end{equation}
-which is a hypergeometric distribution, typical of \enquote{drawing without
-  replacement} problems. The combinatorial proof of this expression is in
-fact the same as for this class of problems
-\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-
-Using the conditional probability above we obtain the probability for the
-activity of the sample:
-\begin{gather}
-  \label{eq:subpop_average}
-  \p(\yxs=\yrs \| \yH) = \sum_{\yRf}
-  \p(\yxs = \yrs \|\yXf=\yRf, \yH)\,
-  \p(\yXf=\yRf \| \yH)
-  = 
-  \sum_{\yRf}
-  \ypp(\yrs\|\yRf)\,
-  \p(\yXf=\yRf \| \yH).
-\end{gather}
-It should be proved that the probability distribution for the full activity
-of the sample is also symmetric and completely determined by the
-distribution of its population-averaged activity:
-\begin{equation}
-  \label{eq:marginal}
-  \p(\yx = \yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs \| \yH).
-\end{equation}
-This is intuitively clear: our initial symmetric ignorance should also
-apply to the sample. The distribution for the
-sample~\eqref{eq:subpop_average} indeed satisfies the same representation
-theorem~\eqref{eq:joint_plaus_N_homog} as the distribution for the full
-population.
-
-The conditional probability
-$\p(\yxs = \yrs \|\yXf=\yRf, \yH) \equiv \ypp(\yrs \|\yRf)$, besides
-relating the distributions for the population and sample activities via
-marginalization, also allows us to express the expectation value of any
-function of the sample activity, $g(\yrs)$, in terms of the distribution
-for the full population, as follows:
-\begin{multline}
-  \label{eq:pullback_P}
-  \expe{g\|I}
-  \equiv
-  \sum_{\yrs} g(\yrs)\,\p(\yxs=\yrs \| \yH)
-  =
-  \sum_{\yrs} g(\yrs) \sum_{\yRf} \ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
-  ={}\\
-  \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
-  \p(\yXf=\yRf \| \yH),
-%  =
-%  \sum_{N\yRf=0}^N \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
-%  \equiv  \expeb{g^*}.
-\end{multline}
-where the second step uses \eqn~\eqref{eq:subpop_average}. The last
-expression shows that the expectation of the function $g(\yrs)$ is equal to
-the expectation of the function
-$g^*(\yRf) \defd \sum_{\yrs} g(\yrs)\,\ypp(\yrs\|\yRf)$.
-
-\bigskip
-
-The final expression in \eqn~\eqref{eq:pullback_P} is important for our
-\me\ application: the requirement that the function $g$, defined for the
-sample, have a  value $c$ obtained from observed data,
-\emph{translates into a linear constraint for the distribution of the full
-  population}:
-\begin{equation}
-  \label{eq:constraint_extended}
-  c = \expe{g \| \yH} \equiv \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
-  \p(\yXf=\yRf \| \yH).
-\end{equation}
-
-In particular, when the function $g$ is the $m$-activity of the sample,
-$g(\yrs) = \sav{\yr\dotso\yr} \equiv \binom{n\yrs}{m}/\binom{n}{m}$, we
-find
-\begin{equation}
-  \label{eq:expe_products}
-%  \label{eq:pullback_m_expectations}
-%  \ypp^*\colon \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}
-%  \mapsto
-    % \bigl( \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}\bigr)^* =
-    % \av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}},
-    % \\[2\jot]
-    % \expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH} =
-    % \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
-  \expe{\sav{\underbrace{\yr \dotsm \yr}_{\text{$m$ factors}}} \| \yH}
-\equiv
-    % \\[2\jot]
-    % {}
-    % \frac{(n-m)!}{n!}
-    \sum_{\yrs} %m!\,
-    \binom{n}{m}^{-1}
-    \binom{n \yrs}{m}\, \p(\yxs=\yrs \| \yH)
-    % \sum_{N \yRf=0}^N %m!\,
-    % \binom{N-m}{N \yRf-m} \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
-    = \binom{\yNv}{m}^{-1}
-    % \frac{(N-m)!}{N!}
-    \sum_{\yRf} %m!\,
-    \binom{\yNv \yRf}{m}\, \p(\yXf=\yRf \| \yH)
-\equiv    \expe{\av{\underbrace{\yR \dotsm \yR}_{\text{$m$ factors}}} \| \yH},
-\end{equation}
-that is, \emph{the expected values of the $m$-activities of the sample and
-  of the full population are equal}. The proof of the middle equality uses
-the expression for the $m$th factorial moment of the hypergeometric
-distribution and can be found in \textcite{potts1953}. Similar relations
-can be found for the raw moments $\expe{\yrs^m}$ and $\expe{\yRf^m}$, which
-can be written in terms of the product expectations using
-\eqn~\eqref{eq:products_intermsof_average}.
-
-Thus, in a \me\ application, when we require the expectation of the
-$m$-activity of a sample to have a particular value, we are also
-requiring  the expectation of the $m$-activity of the full population to
-have the same value.
-
-\bigskip
-
-\begin{figure}[!b]
-\centering
-\includegraphics[width=0.9\linewidth]{pop_sample_projection3.pdf}%
-\caption{Log-density plot of the hypergeometric distribution
-  $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
-\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
-  colourbar depending on your \textsc{pdf} viewer.)}
-\label{fig:hypergeom_proj}
-\end{figure}%hypergeometric_identity.nb
-These expectation equalities between sample and full population should not
-be surprising: we intuitively \emph{expect} that the proportion of coloured
-balls sampled from an urn should be roughly equal to the proportion of
-coloured ball contained in the urn. The formulae in the present section
-formalize and mathematically express our intuition. The hypergeometric
-distribution $\ypp(\yrs\|\yRf)$ plays an important role in this
-formalization. A look at its plot, \fig~\ref{fig:hypergeom_proj}, reveals
-that it is a sort of \enquote{fuzzy identity transformation}, or fuzzy
-Kronecker delta, between the $\yRf$-space $\set{0,\dotsc,\yNv}$ and
-$\yrs$-space $\set{0,\dotsc,n}$. From \eqn~\eqref{eq:marginal} we thus have
-that
-\begin{equation}
-  \label{eq:roughly_equal_nN}
-  \p(\yxs=a \|\yH) \approx \p(\yXf=a \|\yH),\qquad
-\expeb{g(\yrs) \|\yH} \approx \expeb{g(\yRf) \|\yH},
-\end{equation}
-where $g$ is any smooth function defined on $\clcl{0,1}$. These approximate
-equalities express the intuitive fact that \emph{our uncertainty about the
-  sample is representative of our uncertainty about the population and
-  about other samples}, and vice versa. When $n=\yNv$, $\ypp(\yrs\|\yRf)$
-becomes the identity matrix and the approximate equalities above become
-exact -- of course, since we have sampled the full population.
-
-But the approximate equalities above may miss important features of the two
-probability distributions. In the next section we will in fact emphasize
-their differences. If the distribution for the population average $\yRf$ is
-bimodal, for example, the bimodality can be lost in the distribution for
-the sample average $\yrs$, owing to the coarsening effect of
-$\ypp(\yrs\|\yRf)$.
-
-% They should be contrasted with the limits
-% $\p(\yxs=a) \to \p(\yXf=a)$ and $\expeb{g(\yxs)} \to \expeb{g(\yXf)}$, as
-% $n\to \yNv$, do: these limits are trivially, universally valid because the
-% sample becomes the full population as $n\to \yNv$. In particular, these limits
-% hold even in cases where the conditional probability
-% $\p(\yxs = \yrs \|\yXf=\yRf)$ is not a fuzzy identity, our uncertainties
-% about sample and about population can differ wildly, and the approximate
-% equalities~\eqref{eq:roughly_equal_nN} do not hold.
-
-\section{Maximum-entropy: sample level \vs\ full-population level}
-\label{sec:specific_initial_probability}
-
-In the previous section we have seen that observations about a sample can
-be used as constraints on the distribution for the activity of the full
-population. Let us use such constraints with the \me\ method. Suppose that
-we want to constrain $m$ functions of the sample activity, vectorially
-written $\yf \defd (g_1,\dotsc,g_m)$, to $m$ values
-$\yc \defd (c_1,\dotsc,c_m)$. These functions are typically $k$-activities
-$\sav{\yr\dotso \yr}$, and the values are typically the time averages of
-the observed sample, as discussed in \sect~\ref{sec:intro}:
-$\yc = \sum_t \yf[\yrs(t)]/T$.
-
-Let us apply the relative-\me\ method \cite{sivia1996_r2006,meadetal1984}
-directly to sampled neurons; denote this approach by $\yHb$. Then we apply
-the method to the full population of neurons, most of which are unsampled;
-denote this approach by $\yHa$.
-
-Applied directly to the sampled neurons, the method yields the distribution
-\begin{equation}
-  \label{eq:app_maxent_sample}
-  \p(\yxs=\yrs \|\yHb)
-  =\frac{1}{\yk(\yl)}\,
-  \binom{n}{n\yrs}
-  \exp[\yl\T \yf(\yrs)]
-\end{equation}
-where $\yk(\yl)$ is a normalization constant. The binomial in front of the
-exponential appears because we must account for the multiplicity by which
-the population-average activity $\yrs$ can be realized: $\yrs=0$ can be
-realized in only one way (all neurons inactive), $\yrs=1/n$ can be realized
-in $n$ ways (one active neuron out of $n$), and so on. This term is analogous to
-the \enquote{density of states} in front of the Boltzmann factor in
-statistical mechanics \cite[\chap~16]{callen1960_r1985}. The $m$ Lagrange
-multipliers $\yl\defd (l_1,\dotsc,l_m)$ must satisfy the $m$ constraint
-equations
-\begin{equation}
-  \label{eq:app_maxent_sample_constraints}
-  c_k = \expe{g_k \|\yHb} \equiv
-  \frac{1}{\yk(\yl)}\sum_{\yrs}  g_k(\yrs) \binom{n}{n\yrs}
-  \exp[\yl\T \yf(\yrs)], \qquad k=1,\dotsc,m.
-\end{equation}
-
-Applied to the full population, using the constraint
-expression~\eqref{eq:constraint_extended} derived in the previous section,
-the method yields the distribution for the full-population activity
-\begin{equation}
-  \label{eq:app_maxent_pop}
-  \p(\yXf = \yRf \| \yHa)  = \frac{1}{\yK(\yL)}\,
-  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
-  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
-\end{equation}
-The $m$ Lagrange multipliers $\yL\defd (\lambda_1,\dotsc,\lambda_m)$ must
-satisfy the $m$ constraint equations
-\begin{equation}
-  \label{eq:app_maxent_pop_constraints}
-  c_k = \expe{g_k \|\yHa}\equiv
-  \frac{1}{\yK(\yL)}\sum_{\yrs} \sum_{\yRf} g_k(\yrs) \ypp(\yrs \|\yRf)
-\,
-  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
-  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
-  \qquad k=1,\dotsc,m.
-\end{equation}
-
-We obtain the distribution for the sample activity by marginalization, using
-\eqn~\eqref{eq:marginal}:
-\begin{equation}
-  \label{eq:app_maxent_pop_marg}
-  \p(\yxs = \yrs \| \yHa)  = \frac{1}{\yK(\yL)}\, 
-  \sum_{\yRf} \ypp(\yrs \|\yRf)
-\,
-  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
-  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
-\end{equation}
-
-The distributions for the sample activity,
-\eqns~\eqref{eq:app_maxent_pop_marg} and \eqref{eq:app_maxent_sample},
-obtained with the two approaches $\yHb$ and $\yHa$, are different. From the
-discussion in the previous section we expect them to be vaguely similar;
-yet they cannot be exactly equal, because their equality would require the
-$2m$ quantities $\yL$ and $\yl$ to satisfy the constraint equations
-\eqref{eq:app_maxent_pop_constraints} and
-\eqref{eq:app_maxent_sample_constraints}, and in addition also the $n$
-equations $\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$,
-$\yrs=1/n,\dotsc,1$ (one equation is taken care of by the normalization of
-the distributions). We would have a set of $2m+n$ equations in $2m$
-unknowns.
-
-Hence, \emph{the applications of \me\ at the sample level and at the
-  full-population level are inequivalent}. They lead to numerically
-different distributions for the sample activity $\yr$.
-
-%%% FIGURES %%%
-\iffalse
-\begin{figure}[!t]
-\centering
-\includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
-\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at
-  the population level followed by sample marginalization (blue triangles),
-  \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yNv=5000$,
-  $n=200$, and constraints as in \eqn~\eqref{eq:constraints}.}
-\label{fig:diff_maxent_pop_sample}
-\end{figure}%maxent_pop_or_sample.nb
-%
-\begin{figure}[!t]
-\centering
-\includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}% 
-\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at
-  the population level followed by sample marginalization (blue triangles),
-  \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yNv=5000$,
-  $n=200$, and constraints as in \eqn~\eqref{eq:constraints}.}
-\label{fig:diff_maxent_pop_sample_realdata}
-\end{figure}%maxent_pop_or_sample.nb
-\fi%
-\begin{figure}[!t]
-\centering
-\includegraphics[width=0.9\linewidth]{comparison3.pdf}%
-\caption{Linear and log-plots of $\p(\yxs = \yrs)$ for a sample of $n=200$
-  and constraints as in \eqn~\eqref{eq:constraints}, constructed by:
-  \textcolor{myred}{\textbf{red squares:}} \me\ at the sample level,
-  \eqn~\eqref{eq:app_maxent_sample};
-  \textcolor{mybluishpurple}{\textbf{blue triangles:}} \me\ at the
-  population level, \eqn~\eqref{eq:app_maxent_pop_marg} with
-  $\yNv=10\,000$, followed by sample marginalization;
-  \textcolor{myyellow}{\textbf{yellow circles:}} \me\ at the population
-  level with unknown population size,
-  \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}, according to the
-  distribution~\eqref{eq:pdf_popsize} for the population.}
-\label{fig:all_three}
-\end{figure}%maxent_pop_or_sample.nb\fi
-The distribution obtained at the sample level will show different
-features from the one obtained at the population level, like displaced or
-additional modes or particular tail behaviour. We show an example of this
-discrepancy in \fig~\ref{fig:all_three}, for
-$\yNv=10\,000$, $n=200$, and the two constraints
-\begin{equation}
-  \label{eq:constraints}
-  \expe{\yrs} = 0.0478,\qquad
-  \expe{\yxxs} = 0.00257,
-  % ,\quad
-  % \expe{\sav{\yr\yr\yr}} = 1.48\times 10^{-4},\quad
-  % \expe{\sav{\yr\yr\yr\yr}} = 8.81 \times 10^{-6}.
-\end{equation}
-which come from the actual recording of circa 200 neurons from macaque
-motor cortex \cite{rostamietal2016_r2017}. The distribution obtained at the
-population level (blue triangles) has a higher and displaced mode and a quite
-different behaviour for activities around $0.5$ than the distribution
-obtained at the sample level (red squares).
-
-\bigskip
-
-In our discussion we have so far assumed the size $\yNv$ of the larger
-population to be known. This is rarely the case, however. We usually are
-uncertain about $\yNv$ and can only guess its order of magnitude. In such a
-state of knowledge $\yHd$ our ignorance about the possible value of $\yNv$
-is expressed by a probability distribution $\p(\yNN=\yNv \|\yHd)=h(\yNv)$,
-and the marginal distribution for the sample
-activity~\eqref{eq:app_maxent_pop_marg} is modified, by the law of total
-probability, to
-\begin{multline}
-  \label{eq:app_maxent_pop_marg_unknown_N}
-  \p(\yxs = \yrs \| \yHd)  =
-  \sum_{\yNv} \p(\yxs = \yrs \| \yNN=\yNv,\yHd) \,
-  \p(\yNN=\yNv \|\yHd)
-  ={}\\[-\jot]
-  \sum_\yNv \biggl\{\frac{1}{\yK(\yL_\yNv)}\, 
-  \sum_{\yRf} \ypp_\yNv(\yrs \|\yRf)
-\,
-  \tbinom{\yNv}{\yNv \yRf}\,\exp\Bigl[{\yL_\yNv}\T
-  \sum_{\yrs} \yf(\yrs)\ypp_\yNv(\yrs\|\yRf)\Bigr]\biggr\}
-  \,h(\yNv),
-\end{multline}
-where the Lagrange multipliers $\yL_{\yNv}$ and the summation range for
-$\yRf$ depend on $\yNv$.
-
-As a proof of concept, \fig~\ref{fig:all_three} also shows such a distribution
-(yellow circles) for the same constraints as above, and a probability
-distribution for $\yNv$ inspired by Jeffreys
-\cite[\sect~4.8]{jeffreys1939_r2003}:
-\begin{equation}
-  \label{eq:pdf_popsize}
-  h(\yNv) \propto 1/\yNv, \qquad
-  \yNv \in\set{1\,000,\; 2\,000,\; \dotsc,\;10\,000}.
-\end{equation}
-
-\section{Discussion}
-\label{sec:discussion}
-
-The purpose of the present work was to point out and show, in a simple
-set-up, that the \me\ method can be applied to recorded neuronal data in a
-way that accounts for the larger population from which the data are
-sampled, \eqns~\eqref{eq:app_maxent_pop}--\eqref{eq:app_maxent_pop_marg}.
-This application leads to results that differ from the standard application
-which only considers the sample in isolation,
-\eqns~\eqref{eq:app_maxent_sample}--\eqref{eq:app_maxent_sample_constraints}.
-We gave a numerical example of this difference. We have also shown how to
-extend the new application when the size of the larger population is
-unknown, \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}.
-
-The latter formula, in particular, shows that the standard way of applying
-\me\ 
-implicitly assumes that \emph{no} larger population exists beyond the
-recorded sample of neurons. One could in fact object to the application at
-the population level, and say that the traditional way of applying \me,
-\eqn~\eqref{eq:app_maxent_sample}, yields different results because it does
-not make assumptions about the size $\yNv$ of a possibly existing larger
-population. Such a state of uncertainty, however, is correctly formalized
-according to the laws of probability by introducing a probability
-distribution for $\yNv$, and is expressed by
-\eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}. This expression cannot
-generally be equal to~\eqref{eq:app_maxent_sample} unless the distribution
-for $\yNv$ gives unit probability to $\yNv=n$; that is, unless the sample
-\emph{is} the full population, and no larger population exists.
-
-The standard \me\ approach therefore assumes that the recorded neurons
-constitute a special subnetwork, isolated from the larger network of
-neurons in which it is embedded, and which was also present under the
-recording device. This assumption is unrealistic. The \me\ approach at the
-population level does not make such assumption and is therefore preferable.
-It may reveal features in a data set that were unnoticed by the standard
-\me\ approach.
-
-% This hidden assumption of isolation is unreasonable. It amounts to say that
-% the neurons we distinguished and tracked with our recording device are a
-% very special, isolated set among all those that could have been recorded.
-% It is for this reason that we find the \me\ application at the population
-% level preferable. Physical models of neuronal networks usually include some
-% sort of external input, mimicking an embedding in a larger network. The
-% \me\ distribution obtained at the population level may reveal features in a
-% data set that were unnoticed by the standard \me\ model.
-
-The difference in the resulting distributions between the applications at
-the sample and at the population levels appears in the use of Boltzmann
-machines with hidden units \cite{lerouxetal2008}, although by a different
-conceptual route. It also appears in statistical mechanics: if a system is
-statistically described by a \me\ Gibbs state, its subsystems cannot be
-described by a Gibbs state \cite{maesetal1999}. A somewhat similar
-situation also appears in the statistical description of the final state of
-a non-equilibrium process starting and ending in two equilibrium states: we
-can describe our knowledge about the final state either by (1) a Gibbs
-distribution, calculated from the final equilibrium macrovariables, or (2)
-by the distribution obtained from the Liouville evolution of the Gibbs
-distribution assigned to the initial state. The two distributions differ
-(even though the final \emph{physical} state is obviously exactly the same
-\cite[\sect~4]{jaynes1985d_r1993}), and the second allows us to make
-sharper predictions about the final physical state thanks to our knowledge
-of its preceding dynamics. In this example, though, both distributions are
-usually extremely sharp and practically lead to the same predictions. In
-neuroscientific applications, the difference in predictions of the sample
-\vs\ full-population applications can instead be very relevant.
-
-The idea of the new application leads in fact to more questions. For
-instance:
-\begin{itemize}
-\item Do the standard and new applications lead to different or contrasting
-  conclusions about \enquote{cooperativity}, when applied to real data
-  sets?
-\item How to extend the new application to the \enquote{inhomogeneous} case
-  \cite{schneidmanetal2006,shlensetal2006,roudietal2009b}, in which
-  expectations for individual neurons or groups of neurons are constrained?
-\item What is the mathematical relation between the new application and
-  \me\ models with hidden neurons
-  \cite{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}?
-\end{itemize}
-Owing to space limitations we must leave a thorough investigation of these
-questions to future work.
-
-Finally, we would like to point out the usefulness and importance of the
-probability formulae that relate our states of knowledge about a population
-and its samples, presented in \sect~\ref{sec:prob_samples}. This kind of
-formulae is essential in neuroscience, where we try to understand
-properties of extended brain regions from partial observations. The
-formulae presented here reflect a simple, symmetric state of ignorance.
-More work is needed \cite[\cf][]{levinaetal2017} to extend these formulae
-to account for finer knowledge of the cerebral cortex and its network
-properties.
-
-%\mynote{Importance of the formulae relating sample and population.}
-
-% \mynote{Discussion of case when individual $\yXv_i\dotsm \yXv_j$ are constrained.
-% Likely breakdown of \me\ in this case \cite{portamana2009}. Hints at full
-% Bayesian treatment of which \me\ is a limit? \cite[***]{mackay1995_r2003}}
-
-% do not change font sizes (except perhaps in the \textbf{References}
-% section; see below).
-
-\subsubsection*{Acknowledgements}
-
-PGLPM thanks Mari \amp\ Miri for continuous
-encouragement and affection; Buster Keaton for filling life with awe and
-inspiration; the developers and maintainers of \LaTeX, Emacs, AUC\TeX, Open
-Science Framework, Python, Inkscape, Sci-Hub for making a free and
-unfiltered scientific exchange possible.
-
-% Use unnumbered third level headings for the acknowledgments. All
-% acknowledgments go at the end of the paper. Do not include
-% acknowledgments in the anonymized submission, only in the final paper.
-
-%\section*{References}
-\renewcommand*{\bibname}{References}
-\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
-}
-
-% References follow the acknowledgments. Use unnumbered first-level
-% heading for the references. Any choice of citation style is acceptable
-% as long as you are consistent. It is permissible to reduce the font
-% size to \verb+small+ (9 point) when listing the references. {\bf
-%   Remember that you can go over 8 pages as long as the subsequent ones contain
-%   \emph{only} cited references.}
-% \medskip
-
-% \small
-
-% [1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
-% for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
-% T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
-%   Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.
-
-% [2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
-%   Exploring Realistic Neural Models with the GEneral NEural SImulation
-%   System.}  New York: TELOS/Springer--Verlag.
-
-% [3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
-% learning and recall at excitatory recurrent synapses and cholinergic
-% modulation in rat hippocampal region CA3. {\it Journal of
-%   Neuroscience} {\bf 15}(7):5249-5262.
-% \defbibnote{postnote}{\small\par\medskip\noindent{\footnotesize% Note:
-% \arxivp \mparcp \philscip \biorxivp}%
-% }
-
-\newcommand*{\citein}[2][]{\textnormal{\textcite[#1]{#2}}%\addtocategory{extras}{#2}
-}
-\newcommand*{\citebi}[2][]{ref.\ \citep[#1]{#2}%\addtocategory{extras}{#2}
-}
-\newcommand*{\subtitleproc}[1]{}
-\newcommand*{\chapb}{ch.}
-
-\printbibliography%[postnote=postnote]
-
 \end{document}

### 2018-07-26 nips_2018_author_response.tex
update text and scripts

@@
 
-\newcommand*{\pdftitle}{Maximum-entropy and representative samples\\ of neuronal
-  activity: a dilemma}
+\newcommand*{\pdftitle}{Response to reviewers}
 
-\newcommand*{\headtitle}{\pdftitle} \newcommand*{\pdfauthor}{P.G.L.  Porta Mana, V. Rostami, E. Torre, Y. Roudi}
+\newcommand*{\headtitle}{\pdftitle} %\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, V. Rostami, E. Torre, Y. Roudi}
 
@@
 
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
 \iffalse
@@
 
+\iffalse
 \usepackage[backend=biber,mcite,subentry,citestyle=numeric-comp,bibstyle=numericbringhurst,autopunct=false,sorting=none,sortcites=false,natbib=false,maxnames=8,minnames=8,giveninits=true,block=space,hyperref=true,defernumbers=false,useprefix=true,language=british]{biblatex}
@@
 }
+\fi
+
 \usepackage{graphicx}
@@
 \PassOptionsToPackage{hyphens}{url}\usepackage[hypertexnames=false]{hyperref}
-\usepackage[depth=4]{bookmark}
+%\usepackage[depth=4]{bookmark}
 \hypersetup{colorlinks=true,bookmarksnumbered,pdfborder={0 0 0.25},citebordercolor={0.2 0.1333 0.5333},%bluish
@@
 linkcolor=mypurplishred,urlbordercolor={0.5333 0.1333 0.3333},%reddish
-urlcolor=mygreen,breaklinks=true,pdftitle={\pdftitle},pdfauthor={\pdfauthor}}
+urlcolor=mygreen,breaklinks=true,pdftitle={\pdftitle}}
 % \usepackage[vertfit=local]{breakurl}% only for arXiv
@@
 \begin{enumerate}
-\item The 
+\item We agree with Reviewer 3 that the paper emphasizes some philosophical
+  aspects; intentionally so. Sadly we can't make all kinds of readers
+  happy; a choice of audience is necessary. As readers ourselves we
+  appreciate when a paper begins by asking: \enquote{What is the question?
+    is it possible to translate it into a mathematical problem? which
+    principles can we use to make such translation?}. There are papers that
+  develop very refined mathematical techniques but leave us unsatisfied,
+  with the lingering question \enquote{why are we doing this?}. This is the
+  reason why we try to emphasize these kinds of questions. But we are sure
+  that part of the NIPS audience will appreciate this emphasis.
+
+\item We also agree with Reviewer 3 that the main formulae of the paper,
+  (17) and (19), probably have little experimental use today. But the point
+  they make is important:
+  \begin{itemize}
+  \item the Reviewer says \enquote{the differences between the sample level
+      and population level distribution seem minor} -- but we could not
+    have known this, if we hadn't faced the whole problem and derived a
+    formula showing that the difference is minor;
+  \item the sample-level distribution is actually \emph{less} entropic. By
+    lifting the constraints to the population level, the maximum-entropy
+    principle allows us to explore even \emph{more entropic} sample
+    distributions than those allowed by a direct application at the sample
+    level. Thus this approach agrees with the \enquote{maximum entropy}
+    spirit. -- These points should be mentioned, of course, in an amended
+    version of the paper.
+  \item 
+  \end{itemize}
+
 \end{enumerate}

### 2018-07-26 nips_2018_author_response.tex
update text and scripts

@@
 
-\iffalse
 \usepackage[shortlabels,inline]{enumitem}
@@
 \setlist[enumerate,2]{label=\alph*.}
-\setlist[enumerate]{leftmargin=\parindent}
-\setlist[itemize]{leftmargin=\parindent}
+\setlist[enumerate]{leftmargin=1.5em}
+\setlist[itemize]{leftmargin=1.5em}
 \setlist[description]{leftmargin=\parindent}
-\fi
+
 %% With euler font cursive for Greek letters - the [1] means 100% scaling
@@
 \begin{document}
-We thank the reviewers for their analysis and suggestions. We would like to
-reply to specific points:
+We thank the reviewers for their analysis, appreciative words, and suggestions.
+
+\textbf{Reviewer 1}
 \begin{enumerate}
-\item We agree with Reviewer 3 that the paper emphasizes some philosophical
-  aspects; intentionally so. Sadly we can't make all kinds of readers
-  happy; a choice of audience is necessary. As readers ourselves we
-  appreciate when a paper begins by asking: \enquote{What is the question?
-    is it possible to translate it into a mathematical problem? which
-    principles can we use to make such translation?}. There are papers that
-  develop very refined mathematical techniques but leave us unsatisfied,
-  with the lingering question \enquote{why are we doing this?}. This is the
-  reason why we try to emphasize these kinds of questions. But we are sure
-  that part of the NIPS audience will appreciate this emphasis.
+\item It would be interesting to use the frequency distribution of the
+  sample activity, measured in the recording, as a constraint, equivalent
+  to $n$ moment constraints. In fact, using it at the population level is
+  less questionable than using at the sample level (activities with zero
+  frequencies lead to zeroes in the maximum-entropy distribution, and these
+  are unreasonable because the sample size is much smaller than the
+  state-space size. This indicates that maximum-entropy is used beyond its
+  range of validity). Its use, however, addresses a different question than
+  the one discussed in our introduction, namely to define and measure
+  \enquote{cooperativity} by comparing maximum-entropy distributions
+  constructed from different numbers of moments.
+\item Unfortunately the reviewer forgot the bibliographic reference to Tkacik et al
+  2012. We'd like to read that paper.
+\item It's true, we forgot to discuss how the population-level application
+  solves the sample-size-dependence of maximum-entropy results pointed out in
+  previous literature. This should be addressed in an amended version of
+  the paper.
+\end{enumerate}
 
+\medskip
+
+\textbf{Reviewer 2}
+\begin{enumerate}
+\item We agree that the comparison with other approaches is important. But
+  we believe that it's also important to carefully examine (1) the question
+  we're asking, and (2) which first principles, if any, can be used to
+  translate it into a mathematical problem. A lot of literature develops
+  very refined -- though often ad hoc -- mathematical techniques; and yet
+  it leaves us unsatisfied, with the lingering question \enquote{why were
+    we doing this?}. This is the reason for our long introductory
+  discussion about the problem and the first principles that can be used to
+  address it.
+\item Regarding the effect of correlations on the result: the original
+  question is how \enquote{cooperativity} can be related to the measured
+  correlations, or defined in terms of them. The method is devised to give
+  one answer to this question; see the \enquote{protocol} 1--3 in the
+  Introduction. The crucial point is that there is no underlying
+  distribution; the point of the method is to construct it for us.
+\item We agree that there are many approaches in the literature. In the
+  literature we have explored we've seen very little use of the basic
+  sampling formulae presented in our paper. Honestly this surprised us,
+  because they're surely essential even in a very basic analysis of the data.
+\end{enumerate}
+
+\medskip
+
+\textbf{Reviewer 3}
+\begin{enumerate}
 \item We also agree with Reviewer 3 that the main formulae of the paper,
@@
   \end{itemize}
+\item We agree with Reviewer 3 that the paper emphasizes some philosophical
+  aspects; intentionally so. Sadly we can't make all kinds of readers
+  happy; a choice of audience is necessary. As readers ourselves we
+  appreciate when a paper begins by asking: \enquote{What is the question?
+    is it possible to translate it into a mathematical problem? which
+    principles can we use to make such translation?}. There are papers that
+  develop very refined mathematical techniques but leave us unsatisfied,
+  with the lingering question \enquote{why are we doing this?}. This is the
+  reason why we try to emphasize these kinds of questions. But we are sure
+  that part of the NIPS audience will appreciate this emphasis.
 

### 2018-07-27 nips_2018_author_response.tex
update text and scripts

@@
 \usepackage[shortlabels,inline]{enumitem}
-\SetEnumitemKey{para}{itemindent=\parindent,leftmargin=0pt,listparindent=\parindent,parsep=0pt,itemsep=\topsep}
+\SetEnumitemKey{para}{itemindent=\parindent,leftmargin=0pt,listparindent=\parindent,parsep=0pt,itemsep=\topsep,topsep=0pt}
 % \begin{asparaenum} = \begin{enumerate}[para]
@@
 \setlist[enumerate,2]{label=\alph*.}
-\setlist[enumerate]{leftmargin=1.5em}
+\setlist[enumerate]{leftmargin=1.5em,topsep=-1ex}
 \setlist[itemize]{leftmargin=1.5em}
@@
 
-\medskip
+\bigskip
 
@@
   we believe that it's also important to carefully examine (1) the question
-  we're asking, and (2) which first principles, if any, can be used to
-  translate it into a mathematical problem. A lot of literature develops
-  very refined -- though often ad hoc -- mathematical techniques; and yet
-  it leaves us unsatisfied, with the lingering question \enquote{why were
-    we doing this?}. This is the reason for our long introductory
-  discussion about the problem and the first principles that can be used to
-  address it.
-\item Regarding the effect of correlations on the result: the original
-  question is how \enquote{cooperativity} can be related to the measured
-  correlations, or defined in terms of them. The method is devised to give
-  one answer to this question; see the \enquote{protocol} 1--3 in the
-  Introduction. The crucial point is that there is no underlying
-  distribution; the point of the method is to construct it for us.
+  we're asking, and (2) which first principles can be used to translate it
+  into a mathematical problem. Some literature develops very refined --
+  though often ad hoc -- mathematical techniques, but then leaves us
+  unsatisfied, with the lingering question \enquote{why were we doing
+    this?}. This is the reason for our long introductory discussion about
+  the problem and the first principles that can be used to address it.
+\item The firing rate likely affects the results and the quantification of
+  \enquote{cooperativity}. Is this an issue? is our idea of
+  \enquote{cooperativity} invariant with respect to the firing rate, or
+  not? Requirements of this kind constrain the translation of cooperativity
+  into a mathematical quantity. This again shows the importance of defining
+  the question we're asking first.
+\item Regarding correlations: the method quantifies cooperativity using the
+  measured correlations; see the \enquote{protocol} 1--3 in the
+  Introduction. Thus, correlations determine the result.
 \item We agree that there are many approaches in the literature. In the
@@
 \begin{enumerate}
-\item We also agree with Reviewer 3 that the main formulae of the paper,
-  (17) and (19), probably have little experimental use today. But the point
-  they make is important:
+\item We agree with the Reviewer that the main formulae of the paper, (17)
+  and (19), probably have little \emph{experimental} use today. But the
+  formulae and the analysis behind them are important exactly for the
+  points raised by the Reviewer:
   \begin{itemize}
-  \item the Reviewer says \enquote{the differences between the sample level
-      and population level distribution seem minor} -- but we could not
-    have known this, if we hadn't faced the whole problem and derived a
+  \item \begin{minipage}[t]{0.8\linewidth} the Reviewer says \enquote{the
+        differences between the sample level and population level
+        distribution seem minor} -- this can be true, but we could not have
+      known this, if we hadn't faced the whole problem and derived a
       formula showing that the difference is minor;
-  \item the sample-level distribution is actually \emph{less} entropic. By
+    \end{minipage}
+  \item
+    \begin{minipage}[t]{0.8\linewidth}
+      application of the method with higher-order constrains can show more
+      interesting differences between sample- and population-level
+      distributions. See e.g. the plot on the right. This should be added
+      in an amended version of the paper. Unfortunately there are space
+      constraints.
+    \end{minipage}\hspace{\fill}
+    \smash{\begin{minipage}[c]{0.19\linewidth}\vspace{0pt}
+      \includegraphics[width=\linewidth]{zoom_4mom.png}
+    \end{minipage}}
+
+%     % \setlength{\intextsep}{0.5ex}% with wrapfigure
+%     \begin{figure}[h!]%{r}{0.4\linewidth} % with wrapfigure
+%   \centering
+% %\caption{***}\label{fig:comparison_a5}
+% \end{figure}
+  \item The population-level distribution can be more or less entropic of
+    the sample-level one. is actually \emph{less} entropic. By
     lifting the constraints to the population level, the maximum-entropy

### 2018-07-27 nips_2018_author_response.tex
update text and scripts

@@
 \begin{document}
-We thank the reviewers for their analysis, appreciative words, and suggestions.
+We thank the reviewers for their analysis, appreciative words, and
+suggestions. We reply to major points only. The manuscript can be easily
+amended to address the points we don't discuss here.
 
@@
   constructed from different numbers of moments.
-\item Unfortunately the reviewer forgot the bibliographic reference to Tkacik et al
-  2012. We'd like to read that paper.
 \item It's true, we forgot to discuss how the population-level application
@@
   \begin{itemize}
-  \item \begin{minipage}[t]{0.8\linewidth} the Reviewer says \enquote{the
+  \item \begin{minipage}[t]{0.74\linewidth}The Reviewer says \enquote{the
         differences between the sample level and population level
-        distribution seem minor} -- this can be true, but we could not have
+        distribution seem minor}: this could be true, but we could not have
       known this, if we hadn't faced the whole problem and derived a
-      formula showing that the difference is minor;
-    \end{minipage}
-  \item
-    \begin{minipage}[t]{0.8\linewidth}
-      application of the method with higher-order constrains can show more
-      interesting differences between sample- and population-level
-      distributions. See e.g. the plot on the right. This should be added
-      in an amended version of the paper. Unfortunately there are space
-      constraints.
+      formula showing that the difference is minor.
     \end{minipage}\hspace{\fill}
-    \smash{\begin{minipage}[c]{0.19\linewidth}\vspace{0pt}
+    \smash{\begin{minipage}[t]{0.24\linewidth}\vspace{0pt}
       \includegraphics[width=\linewidth]{zoom_4mom.png}
     \end{minipage}}
-
+  \item
+    \begin{minipage}[t]{0.74\linewidth}
+      Application of the method with higher-order constrains can show more
+      interesting differences between sample- and population-level
+      distributions. See e.g. the plot on the right: the bimodal
+      distribution is the marginalized, population-level one.
+    \end{minipage}
 %     % \setlength{\intextsep}{0.5ex}% with wrapfigure
@@
 % \end{figure}
-  \item The population-level distribution can be more or less entropic of
-    the sample-level one. is actually \emph{less} entropic. By
-    lifting the constraints to the population level, the maximum-entropy
-    principle allows us to explore even \emph{more entropic} sample
-    distributions than those allowed by a direct application at the sample
-    level. Thus this approach agrees with the \enquote{maximum entropy}
-    spirit. -- These points should be mentioned, of course, in an amended
-    version of the paper.
-  \item 
+\item Calculations show that the population-level distribution can be less
+  or more entropic than the sample-level one. But a more entropic
+  distribution is not necessarily more correct (we could just use a uniform
+  distribution then). The question is how the constraints are appropriately
+  applied in the method. 
 \end{itemize}
-\item We agree with Reviewer 3 that the paper emphasizes some philosophical
-  aspects; intentionally so. Sadly we can't make all kinds of readers
-  happy; a choice of audience is necessary. As readers ourselves we
-  appreciate when a paper begins by asking: \enquote{What is the question?
-    is it possible to translate it into a mathematical problem? which
-    principles can we use to make such translation?}. There are papers that
-  develop very refined mathematical techniques but leave us unsatisfied,
-  with the lingering question \enquote{why are we doing this?}. This is the
-  reason why we try to emphasize these kinds of questions. But we are sure
-  that part of the NIPS audience will appreciate this emphasis.
-
+As this Reviewer and Reviewer 2 remark, the points and findings above
+deserve more discussion. Unfortunately there are space constraints, and we
+preferred to give more room to a careful analysis of the principles behind
+the method, to see whether the method is correctly applied. Our main wish
+is that our readers will give more thoughts to these matters. We could
+hurry up and produce a lot of mathematical results; but, whether positive or
+negative, such would be meaningless if the method isn't meaningfully
+applied. And it isn't always the case that a meaningful application of a
+method can be judged by its results alone.
+% \item We agree with Reviewer 3 that the paper emphasizes some philosophical
+%   aspects; intentionally so. Sadly we can't make all kinds of readers
+%   happy; a choice of audience is necessary. As readers ourselves we
+%   appreciate when a paper begins by asking: \enquote{What is the question?
+%     is it possible to translate it into a mathematical problem? which
+%     principles can we use to make such translation?}. There are papers that
+%   develop very refined mathematical techniques but leave us unsatisfied,
+%   with the lingering question \enquote{why are we doing this?}. This is the
+%   reason why we try to emphasize these kinds of questions. But we are sure
+%   that part of the NIPS audience will appreciate this emphasis.
 \end{enumerate}

### 2018-07-27 nips_2018_author_response.tex
update text and scripts

@@
     this?}. This is the reason for our long introductory discussion about
-  the problem and the first principles that can be used to address it.
+  the problem and about the first principles that can be used to address
+  it. See also our reply to Reviewer 3 below.
 \item The firing rate likely affects the results and the quantification of
-  \enquote{cooperativity}. Is this an issue? is our idea of
-  \enquote{cooperativity} invariant with respect to the firing rate, or
+  \enquote{cooperativity}. The question is: Is this an issue? is our idea
+  of \enquote{cooperativity} invariant with respect to the firing rate, or
   not? Requirements of this kind constrain the translation of cooperativity
@@
   measured correlations; see the \enquote{protocol} 1--3 in the
-  Introduction. Thus, correlations determine the result.
+  Introduction. Thus, correlations determine the result rather than
+  affecting it.
 \item We agree that there are many approaches in the literature. In the
@@
       interesting differences between sample- and population-level
-      distributions. See e.g. the plot on the right: the bimodal
+      distributions. See e.g. the plot on the right, obtained with
+      fourth-moment constraints: the bimodal
       distribution is the marginalized, population-level one.
@@
 is that our readers will give more thoughts to these matters. We could
-hurry up and produce a lot of mathematical results; but, whether positive or
-negative, such would be meaningless if the method isn't meaningfully
+hurry up and produce a lot of mathematical results; but, whether positive
+or negative, any results are meaningless if the method isn't meaningfully
 applied. And it isn't always the case that a meaningful application of a

### 2018-07-31 nips_2018_author_response.tex
update text and scripts

--- 2018-07-27 nips_2018_author_response.tex
@@
 % \begin{inparaenum} = \begin{enumerate*}
-\setlist[enumerate,2]{label=\alph*.}
-\setlist[enumerate]{leftmargin=1.5em,topsep=-1ex}
+\setlist[enumerate,2]{label=\alph*}
+\setlist[enumerate]{leftmargin=1.5em,topsep=-1ex,itemsep=0.5ex}
 \setlist[itemize]{leftmargin=1.5em}
@@
 We thank the reviewers for their analysis, appreciative words, and
-suggestions. We reply to major points only. The manuscript can be easily
-amended to address the points we don't discuss here.
+suggestions. Our reply below does not address the points that can be easily
+addressed by amending our manuscript.
+
+\bigskip
 
 \textbf{Reviewer 1}
-\begin{enumerate}
-\item It would be interesting to use the frequency distribution of the
-  sample activity, measured in the recording, as a constraint, equivalent
-  to $n$ moment constraints. In fact, using it at the population level is
-  less questionable than using at the sample level (activities with zero
-  frequencies lead to zeroes in the maximum-entropy distribution, and these
-  are unreasonable because the sample size is much smaller than the
-  state-space size. This indicates that maximum-entropy is used beyond its
-  range of validity). Its use, however, addresses a different question than
-  the one discussed in our introduction, namely to define and measure
-  \enquote{cooperativity} by comparing maximum-entropy distributions
-  constructed from different numbers of moments.
-\item It's true, we forgot to discuss how the population-level application
-  solves the sample-size-dependence of maximum-entropy results pointed out in
-  previous literature. This should be addressed in an amended version of
-  the paper.
+\begin{enumerate}[wide]
+\item \emph{\enquote{$P(s)$}:} To be honest we are not sure what the
+  Reviewer means with \enquote{$P(s)$-type constraints} (the Reviewer's
+  comment is truncated). If it means the empirical frequency distribution
+  of the sample activity, equivalent to $n$ moment constraints, then it is
+  one of the distributions calculated at step 2. of the protocol (lines
+  27--56), to be compared with the other distributions. We may add that
+  using the empirical frequencies as constraints at the sample level
+  is questionable: activities with zero frequencies lead to zeroes in the
+  maximum-entropy distribution, and these are unreasonable because the
+  sample size is much smaller than the state-space size. This indicates
+  that maximum-entropy is used beyond its range of validity. The approach
+  at the population level + marginalization cures this problem somewhat.
+% Its use, however, addresses a different question than
+%   the one discussed in our introduction, namely to define and measure
+%   \enquote{cooperativity} by comparing maximum-entropy distributions
+%   constructed from different numbers of moments.
+\item \emph{Puzzling finds in the literature:} We apologize for having
+  forgotten to discuss this; this oversight can be easily fixed.
 \end{enumerate}
@@
 \textbf{Reviewer 2}
-\begin{enumerate}
-\item We agree that the comparison with other approaches is important. But
-  we believe that it's also important to carefully examine (1) the question
-  we're asking, and (2) which first principles can be used to translate it
-  into a mathematical problem. Some literature develops very refined --
-  though often ad hoc -- mathematical techniques, but then leaves us
-  unsatisfied, with the lingering question \enquote{why were we doing
-    this?}. This is the reason for our long introductory discussion about
-  the problem and the first principles that can be used to address it.
-\item The firing rate likely affects the results and the quantification of
-  \enquote{cooperativity}. Is this an issue? is our idea of
-  \enquote{cooperativity} invariant with respect to the firing rate, or
-  not? Requirements of this kind constrain the translation of cooperativity
-  into a mathematical quantity. This again shows the importance of defining
-  the question we're asking first.
-\item Regarding correlations: the method quantifies cooperativity using the
-  measured correlations; see the \enquote{protocol} 1--3 in the
-  Introduction. Thus, correlations determine the result.
-\item We agree that there are many approaches in the literature. In the
-  literature we have explored we've seen very little use of the basic
-  sampling formulae presented in our paper. Honestly this surprised us,
-  because they're surely essential even in a very basic analysis of the data.
+\begin{enumerate}[wide]
+\item \emph{Comparison with other frameworks and long intro:} The
+  comparison with other approaches and the exploration of more results are
+  indeed important. There are two reasons why we gave less space to them:
+  \begin{enumerate}[label=\alph*.,itemsep=0ex]
+  \item We think that most literature hurries in deriving results and does
+    not give enough space to the examination of the principles behind a
+    method. This worries us. The misapplication of a method often cannot be
+    detected by its results alone. This is especially true for
+    maximum-entropy. We risk finding a plethora of results that may even
+    look interesting but are meaningless, because the method was not
+    meaningfully applied. In turn, this often happens because the problem
+    we wanted to solve with that method is too vaguely defined. For this
+    reason we wanted to give large space to defining the problem and to
+    discussing how the method and its principles should be applied to it.
+    Our wish is that our readers will give more thoughts to these matters.
+  \item We focus on the comparison between the \enquote{naive}
+    maximum-entropy method, and the maximum-entropy method that accounts
+    for the sampling context. The eight-page limit was not enough to make
+    comparisons with more methods and discuss results. Also, a long
+    discussion about further results would steal focus from the examination
+    of the problem and the principles, which are our main concern and message.
+  \end{enumerate}
+
+%   As this Reviewer and Reviewer 2 remark, the points and findings above
+% deserve more discussion. Unfortunately there are space constraints, and we
+% preferred to give more room to a careful analysis of the principles behind
+% the method, to see whether the method is correctly applied. Our main wish
+% is that our readers will give more thoughts to these matters. We could
+% hurry up and produce a lot of mathematical results; but, whether positive
+% or negative, any results are meaningless if the method isn't meaningfully
+% applied. And it isn't always the case that a meaningful application of a
+% method can be judged by its results alone.
+  % But we believe that it's also important to carefully
+  % examine (1) the question we're asking, and (2) which first principles can
+  % be used to translate it into a mathematical problem. Some literature
+  % develops very refined -- though often ad hoc -- mathematical techniques,
+  % but then leaves us unsatisfied, with the lingering question \enquote{why
+  %   were we doing this?}. This is the reason for our long introductory
+  % discussion about the problem and about the first principles that can be
+  % used to address it. See also our reply to Reviewer 3 below.
+\item \emph{The effect of firing rates and correlations:} The empirical
+  correlations determine (rather than affect) the result; see the protocol
+  1--3, lines 27--56. Firing rates surely affect the results, and this is
+  an interesting question: most literature seem to seek a measure of
+  \enquote{cooperativity} that is invariant under firing-rate changes. We
+  wonder whether maximum-entropy can provide this. Again, this is a
+  question that should be addressed in defining the problem and the
+  principles behind the method to solve it.
+%   likely affects the results and the
+%   quantification of \enquote{cooperativity}. The question is: Is this an
+%   issue? is our idea of \enquote{cooperativity} invariant with respect to
+%   the firing rate, or not? Requirements of this kind constrain the
+%   translation of cooperativity into a mathematical quantity. This again
+%   shows the importance of defining the question we're asking first.
+% \item Regarding correlations: the method quantifies cooperativity using the
+%   measured correlations; see the \enquote{protocol} 1--3 in the
+%   Introduction. Thus, correlations determine the result rather than
+%   affecting it.
+\item \emph{Bayesian approaches and sampling:} The literature indeed offers
+  fully Bayesian approaches, which we also prefer. But maximum-entropy
+  methods still abound in the literature, and maybe they can be reasonably
+  motivated. Our goal is to point out the subtlety in apply these methods.
+  In the literature we have explored we've seen very little use of the
+  basic sampling formulae presented in our paper. Honestly this surprised
+  us, because they're surely essential even in a very basic analysis of the
+  data.
 \end{enumerate}
@@
 \textbf{Reviewer 3}
-\begin{enumerate}
-\item We agree with the Reviewer that the main formulae of the paper, (17)
-  and (19), probably have little \emph{experimental} use today. But the
-  formulae and the analysis behind them are important exactly for the
-  points raised by the Reviewer:
-  \begin{itemize}
-  \item \begin{minipage}[t]{0.74\linewidth}The Reviewer says \enquote{the
-        differences between the sample level and population level
-        distribution seem minor}: this could be true, but we could not have
-      known this, if we hadn't faced the whole problem and derived a
-      formula showing that the difference is minor.
+\begin{enumerate}[wide]
+\item \emph{Whether it matters in practice:} We agree with the Reviewer
+  that the consequences of the differences between the two applications
+  should be discussed further in the manuscript. In particular:
+  \begin{enumerate}[label=\alph*.,itemsep=0ex]
+  \item Minor differences may lead to considerably different
+    interpretations of the underlying activity. For instance, the thinner
+    right tail of the population-level distribution indicates that
+    higher-order correlations observed in the data have larger statistical
+    significance (this is the type of question Schneidman et al. were
+    addressing in their research, which popularized maximum-entropy in
+    neuroscience). From this point of view, a more entropic distribution is
+    not necessarily more correct (we could just use a uniform distribution).
+  \item
+    \begin{minipage}[t]{0.74\linewidth}
+      Application of the method with higher-order constrains can show more
+      interesting differences between sample- and population-level
+      distributions. See e.g. the plot on the right, obtained with
+      fourth-moment constraints: the bimodal
+      distribution is the marginalized, population-level one.
     \end{minipage}\hspace{\fill}
@@
     \end{minipage}}
-  \item
-    \begin{minipage}[t]{0.74\linewidth}
-      Application of the method with higher-order constrains can show more
-      interesting differences between sample- and population-level
-      distributions. See e.g. the plot on the right: the bimodal
-      distribution is the marginalized, population-level one.
+  \item \begin{minipage}[t]{0.74\linewidth}Even if the differences were
+      negligible, we believe that the results of the paper would matter:
+      the point is that we could not have known about the presence or
+      absence of differences, if we had not faced the whole problem and
+      derived a formula showing that the difference were negligible.
     \end{minipage}
@@
 % \end{figure}
-\item Calculations show that the population-level distribution can be less
-  or more entropic than the sample-level one. But a more entropic
-  distribution is not necessarily more correct (we could just use a uniform
-  distribution then). The question is how the constraints are appropriately
-  applied in the method. 
-\end{itemize}
-As this Reviewer and Reviewer 2 remark, the points and findings above
-deserve more discussion. Unfortunately there are space constraints, and we
-preferred to give more room to a careful analysis of the principles behind
-the method, to see whether the method is correctly applied. Our main wish
-is that our readers will give more thoughts to these matters. We could
-hurry up and produce a lot of mathematical results; but, whether positive or
-negative, such would be meaningless if the method isn't meaningfully
-applied. And it isn't always the case that a meaningful application of a
-method can be judged by its results alone.
+  \end{enumerate}
+\item\emph{Unnecessarily principled:} For our motivation in emphasizing
+  principles, we kindly ask the Reviewer to read our reply 1. to Reviewer~2
+  above. Probably the writing could be improved on our part to make the
+  paper flow more smoothly, without sacrificing the more philosophical
+  discussion?
+  % We agree with the Reviewer that the main formulae of the paper, (17)
+  % and (19), probably have little \emph{experimental} use today. But the
+  % formulae and the analysis behind them are important exactly for the
+  % points raised by the Reviewer:
+% \item Calculations show that the population-level distribution can be less
+%   or more entropic than the sample-level one. But a more entropic
+%   distribution is not necessarily more correct (we could just use a uniform
+%   distribution then). The question is how the constraints are appropriately
+%   applied in the method. 
+% \item As this Reviewer and Reviewer 2 remark, the points and findings above
+% deserve more discussion. Unfortunately there are space constraints, and we
+% preferred to give more room to a careful analysis of the principles behind
+% the method, to see whether the method is correctly applied. Our main wish
+% is that our readers will give more thoughts to these matters. We could
+% hurry up and produce a lot of mathematical results; but, whether positive
+% or negative, any results are meaningless if the method isn't meaningfully
+% applied. And it isn't always the case that a meaningful application of a
+% method can be judged by its results alone.
 % \item We agree with Reviewer 3 that the paper emphasizes some philosophical

### 2018-09-05 nips_2018_author_response2.tex
update lit

--- 2018-07-31 nips_2018_author_response.tex
@@
 We thank the reviewers for their analysis, appreciative words, and
-suggestions. Our reply below does not address the points that can be easily
-addressed by amending our manuscript.
+suggestions. Owing to space constraints our reply below does not address
+the points that can be easily addressed by amending our manuscript.
 
@@
 \begin{enumerate}[wide]
-\item \emph{\enquote{$P(s)$}:} To be honest we are not sure what the
-  Reviewer means with \enquote{$P(s)$-type constraints} (the Reviewer's
-  comment is truncated). If it means the empirical frequency distribution
-  of the sample activity, equivalent to $n$ moment constraints, then it is
-  one of the distributions calculated at step 2. of the protocol (lines
-  27--56), to be compared with the other distributions. We may add that
-  using the empirical frequencies as constraints at the sample level
-  is questionable: activities with zero frequencies lead to zeroes in the
-  maximum-entropy distribution, and these are unreasonable because the
-  sample size is much smaller than the state-space size. This indicates
-  that maximum-entropy is used beyond its range of validity. The approach
-  at the population level + marginalization cures this problem somewhat.
+\item \emph{\enquote{$P(s)$}:} Unfortunately the Reviewer's comment is
+  truncated; we aren't fully sure if we are interpreting \enquote{$P(s)$}
+  correctly. We assume it to be the empirical frequency distribution of the
+  sample activity, equivalent to $n$ moment constraints. This constraint is
+  one of those used in step 2. of the protocol (lines 27--56), to be
+  compared with the other constraints. We may add that using the
+  empirical frequencies as constraints at the sample level is questionable:
+  activities with zero frequencies lead to zeroes in the maximum-entropy
+  distribution, and these are unreasonable because the sample size is much
+  smaller than the state-space size. This indicates that maximum-entropy is
+  used beyond its range of validity. The approach at the population level +
+  marginalization cures this problem.
 % Its use, however, addresses a different question than
@@
 \begin{enumerate}[wide]
-\item \emph{Comparison with other frameworks and long intro:} The
-  comparison with other approaches and the exploration of more results are
-  indeed important. There are two reasons why we gave less space to them:
-  \begin{enumerate}[label=\alph*.,itemsep=0ex]
-  \item We think that most literature hurries in deriving results and does
-    not give enough space to the examination of the principles behind a
-    method. This worries us. The misapplication of a method often cannot be
-    detected by its results alone. This is especially true for
-    maximum-entropy. We risk finding a plethora of results that may even
-    look interesting but are meaningless, because the method was not
-    meaningfully applied. In turn, this often happens because the problem
-    we wanted to solve with that method is too vaguely defined. For this
-    reason we wanted to give large space to defining the problem and to
-    discussing how the method and its principles should be applied to it.
-    Our wish is that our readers will give more thoughts to these matters.
-  \item We focus on the comparison between the \enquote{naive}
-    maximum-entropy method, and the maximum-entropy method that accounts
-    for the sampling context. The eight-page limit was not enough to make
-    comparisons with more methods and discuss results. Also, a long
-    discussion about further results would steal focus from the examination
-    of the problem and the principles, which are our main concern and message.
-  \end{enumerate}
+\item \emph{Comparison with other frameworks and long intro:} We
+  respectfully acknowledge the Reviewer's personal stand on the
+  \enquote{results vs introduction} point. Our opinion is that at times it
+  is important to focus on the problem and the principles of the method
+  purported to solve it, leaving more detailed results till later. The
+  reason is this: if the method is misapplied, the results are actually
+  irrelevant or even misleading -- and yet they may \emph{appear}
+  interesting, risking to steal our focus from the principles. What's worse
+  is that the results themselves may not reveal, or may even conceal, that
+  the method was misapplied. We think that this has happened in many uses
+  of the maximum-entropy method in neuroscience.
+
+  The modicum of results we presented is meant to bring to light this
+  possible misapplication. We called this a \enquote{dilemma} because we do
+  not wish to advertise the newly derived equations as the final solution
+  to the problem. At this stage we consider a careful description of the
+  problem critical, and a comparison with more advanced methods still
+  premature.
+
+  Also, we would like to reach a broader audience than the one specifically
+  interested in \enquote{cooperativity}. Our primary goal is to raise
+  awareness about this possible pitfall of the maximum-entropy method, and
+  to show a possible solution. We fear that focusing on further results
+  specific to cooperativity would conceal the wider bearing of our message.
 
@@
   % used to address it. See also our reply to Reviewer 3 below.
-\item \emph{The effect of firing rates and correlations:} The empirical
+\item \emph{Effect of firing rates and correlations:} The empirical
   correlations determine (rather than affect) the result; see the protocol
-  1--3, lines 27--56. Firing rates surely affect the results, and this is
-  an interesting question: most literature seem to seek a measure of
-  \enquote{cooperativity} that is invariant under firing-rate changes. We
-  wonder whether maximum-entropy can provide this. Again, this is a
-  question that should be addressed in defining the problem and the
-  principles behind the method to solve it.
+  1--3, lines 27--56. Firing rates surely affect the results, but they do
+  so through the intermediary of the correlations. The question about the
+  effect of firing rates actually brings us back to an analysis of the
+  initial problem: do we want a measure of \enquote{cooperativity} that is
+  invariant under firing-rate changes? do the principles behind the
+  maximum-entropy method provide this? from this point of view, are we
+  misapplying the method?
+
+  % We wonder whether maximum-entropy can provide this. Again, this is a
+  % question that should be addressed in defining the problem and the
+  % principles behind the method to solve it.
 %   likely affects the results and the
@@
   fully Bayesian approaches, which we also prefer. But maximum-entropy
-  methods still abound in the literature, and maybe they can be reasonably
-  motivated. Our goal is to point out the subtlety in apply these methods.
-  In the literature we have explored we've seen very little use of the
-  basic sampling formulae presented in our paper. Honestly this surprised
-  us, because they're surely essential even in a very basic analysis of the
-  data.
+  methods still abound in the neuroscientific literature, and maybe they
+  can be reasonably motivated. Our goal is to point out the subtlety in
+  applying these methods. The literature we have explored makes very little
+  use of the basic sampling formulae presented in our paper. This is
+  surprising: such formulae are surely essential even in a very basic
+  analysis of the data.
 \end{enumerate}
@@
 \item \emph{Whether it matters in practice:} We agree with the Reviewer
-  that the consequences of the differences between the two applications
-  should be discussed further in the manuscript. In particular:
-  \begin{enumerate}[label=\alph*.,itemsep=0ex]
-  \item Minor differences may lead to considerably different
-    interpretations of the underlying activity. For instance, the thinner
-    right tail of the population-level distribution indicates that
-    higher-order correlations observed in the data have larger statistical
-    significance (this is the type of question Schneidman et al. were
-    addressing in their research, which popularized maximum-entropy in
-    neuroscience). From this point of view, a more entropic distribution is
-    not necessarily more correct (we could just use a uniform distribution).
-  \item
-    \begin{minipage}[t]{0.74\linewidth}
-      Application of the method with higher-order constrains can show more
-      interesting differences between sample- and population-level
-      distributions. See e.g. the plot on the right, obtained with
-      fourth-moment constraints: the bimodal
-      distribution is the marginalized, population-level one.
-    \end{minipage}\hspace{\fill}
+  that pointing out the circumstances where the difference matters would be
+  useful. For instance, the thinner
+  right tail of the population-level distribution observed by the Reviewer\\
+  \begin{minipage}[t]{0.74\linewidth}indicates that higher-order
+    correlations observed in the data are important (this is the question
+    Schneidman et al. were addressing in their research, which popularized
+    maximum-entropy in neuroscience). Note that a more entropic
+    distribution is not necessarily more correct (after all, we could just
+    use a uniform distribution otherwise). Application of the method with
+    higher-order constraints can show more interesting differences between
+    sample- and population-level distributions, e.g. a bimodality not
+    observed in the empirical frequencies, as in the plot on the right
+    (bimodal distribution = population-level). This also leads to different
+    conclusion about the relative importance of the correlations.
+  \end{minipage}
+  \hspace{\fill}
     \smash{\begin{minipage}[t]{0.24\linewidth}\vspace{0pt}
@@
     \end{minipage}}
-  \item \begin{minipage}[t]{0.74\linewidth}Even if the differences were
-      negligible, we believe that the results of the paper would matter:
-      the point is that we could not have known about the presence or
-      absence of differences, if we had not faced the whole problem and
-      derived a formula showing that the difference were negligible.
-    \end{minipage}
 %     % \setlength{\intextsep}{0.5ex}% with wrapfigure
@@
 % \end{figure}
-  \end{enumerate}
-\item\emph{Unnecessarily principled:} For our motivation in emphasizing
+
+  But we believe that \emph{even if the differences were negligible, the
+    results of the paper would still matter}: the point is that we could
+  not have known about the presence or absence of differences, if we had
+  not faced the whole problem and derived a formula showing that the
+  difference were negligible.
+\item\emph{Unnecessarily principled:} Regarding our motivation for emphasizing
   principles, we kindly ask the Reviewer to read our reply 1. to Reviewer~2
-  above. Probably the writing could be improved on our part to make the
+  above. Possibly the writing could be improved on our part to make the
   paper flow more smoothly, without sacrificing the more philosophical

### 2018-12-17 maxNt_v4.tex
update paper

--- 2018-07-26 maxNt_v3.tex
@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-07-19T11:05:41+0200
+%% Last-Updated: 2018-09-13T14:10:47+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Suppose that we have recorded the firing activity of a hundred neurons,
+Suppose we have recorded the firing activity of a hundred neurons,
 sampled from a particular brain area. What are we to do with such data?
@@
 
-We hurry to say that the description just given is just \emph{one} way to
-quantify and characterize cooperativity and functional grouping, not
-\emph{the only} way. It can surely be criticized from many points of view.
-Yet, it is quantitative and bears a more precise meaning than an undefined,
-vague notion of \enquote{cooperativity}. Two persons who apply this
-procedure to the same data will obtain the same numbers. Different
-protocols can be based on the \me\ method, for instance protocols that take
-into account the activities or pair activities of specific neurons rather
-than population averages, or even protocols that take into account time
-dependence.
-
-The purpose of the present work is not to assess the merits of \me\ methods
-with respect to other methods. Its main purpose is to show that there is a
-problem in the way the \me\ method itself, as sketched above, is applied to
-the activity of the recorded neurons. We believe that this problem is at
-the root of some quirks about this method that were pointed out in the
-literature \cite{roudietal2009b}. This problems extends also to more
-complex versions of the method, possibly except versions that use
-\enquote{hidden} neurons
-\cite{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}. The problem is
-that the recorded neurons are a \emph{sample} from a larger, unrecorded
+% We hurry to say that the description just given is just \emph{one} way to
+% quantify and characterize cooperativity and functional grouping, not
+% \emph{the only} way. It can surely be criticized from many points of view.
+% Yet, it is quantitative and bears a more precise meaning than an undefined,
+% vague notion of \enquote{cooperativity}. Two persons who apply this
+% procedure to the same data will obtain the same numbers. Different
+% protocols can be based on the \me\ method, for instance protocols that take
+% into account the activities or pair activities of specific neurons rather
+% than population averages, or even protocols that take into account time
+% dependence.
+
+The purpose of the present work is to show that there is a problem in the
+way the \me\ method itself, as sketched above, is applied to the activity
+of the recorded neurons. We believe that this problem is at the root of
+some quirks about this method that were pointed out in the literature
+\cite{roudietal2009b}. This problems extends also to more complex versions
+of the method, possibly except versions that use \enquote{hidden} neurons
+\cite{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}. The problem
+is that the recorded neurons are a \emph{sample} from a larger, unrecorded
 population, but the \me\ method as applied above is treating them as
@@
 
-We need to make an initial probability assignment for the state of the full
+%% from Yasser:
+** we want to describe average activities so our initial knowledge
+disregards details
+
+** we assume sample is representative of full population, so our initial
+knowledge is symmetric
+
+\mynote{modidy this}We need to make an initial probability assignment for the state of the full
 population before any experimental observations are made. This initial
@@
 
-In this state of knowledge, denoted $\yH$, we know that the neurons in the
-population are biologically or functionally similar, for example in
-morphology or the kind of input or output they receive or give. But we are
-completely ignorant about the physical details of the individual neurons.
+% In this state of knowledge, denoted $\yH$, we know that the neurons in the
+% population are biologically or functionally similar, for example in
+% morphology or the kind of input or output they receive or give. But we are
+% completely ignorant about the physical details of the individual neurons.
+We as
 Our ignorance is therefore symmetric under permutations of neuron
@@
 \end{document}
+
+%%% Local Variables: 
+%%% mode: LaTeX
+%%% TeX-PDF-mode: t
+%%% TeX-master: t
+%%% End:

### 2019-01-24 maxNt_v3.tex
that removed
update

@@
 %% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-07-19T11:05:41+0200
+%% Last-Updated: 2018-12-18T13:21:16+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \DeclareMathSymbol{\theta}{\mathalpha}{egreeki}{"12}
- \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
+% \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
  \DeclareMathSymbol{\kappa}{\mathalpha}{egreeki}{"14}
@@
 \newcommand*{\pf}{\mathrm{p}}%probability
-\newcommand*{\p}{\mathrm{P}}%probability
+\newcommand*{\p}{\mathrm{p}}%probability
 %\newcommand*{\tf}{\mathrm{T}}%probability
@@
 
-\newcommand*{\yXv}{\varSigma}
-\newcommand*{\yRv}{\sigma}
-\newcommand*{\yxv}{S}
+%\newcommand*{\yXv}{\varSigma}
+\newcommand*{\yRv}{S}
+%\newcommand*{\yxv}{s}
 \newcommand*{\yrv}{s}
-\newcommand*{\yNv}{\nu}
+\newcommand*{\yNv}{N}
 \newcommand*{\yNN}{\varNu}
 
-\newcommand*{\yx}{\bm{\yxv}}%subpop state
-\newcommand*{\yxs}{\sav{\yx}}%subpop av state
-\newcommand*{\yX}{\bm{\yXv}}%pop state
-\newcommand*{\yXf}{\av{\yX}}%pop av state
-\newcommand*{\yxxs}{\sav{\yr\yr}}%subpop av state
-\newcommand*{\yXXf}{\av{\yR\yR}}%subpop av state
+%\newcommand*{\yx}{\bm{\yxv}}%subpop state
+%\newcommand*{\yxs}{\sav{\yx}}%subpop av state
+%\newcommand*{\yX}{\bm{\yXv}}%pop state
+%\newcommand*{\yXf}{\av{\yX}}%pop av state
+\newcommand*{\yrrs}{\sav{\yr\yr}}%subpop av state
+%\newcommand*{\yXXf}{\av{\yR\yR}}%subpop av state
 \newcommand*{\yr}{\bm{\yrv}}%subpop value
@@
 introduce an analogous notation for the $\yNv$ neurons constituting the
-larger population, but using the corresponding Greek letters:
-$\yRv_{\iota}(t)$ is the activity of the $\iota$th neuron at time bin $t$,
-$\yRf(t) \defd \sum_{\iota} \yRv_{\iota}(t)/\yNv$ is the activity at that
+larger population, but using the corresponding capital letters:
+$\yRv_{i}(t)$ is the activity of the $i$th neuron at time bin $t$,
+$\yRf(t) \defd \sum_{i} \yRv_{i}(t)/\yNv$ is the activity at that
 bin averaged over the larger population, and so on.
@@
 time bin or to a new bin. For these reasons we will now omit the time-bin
-argument \enquote{(t)} from our expressions.
-
-Probabilities refer to statements about the quantities we observe. We use
-the standard notation:
-\begin{equation}
-  \label{eq:notation_statements}
-  \begin{aligned}
-    &\text{\enquote{$\yXv_{\iota} = \yRv_{\iota}$} 
-      stands for \enquote{the activity of the $\iota$th neuron is $\yRv_{\iota}$}},
-    \\
-    &\text{\enquote{$\yXf = \yRf$} 
-      stands for \enquote{the (population-averaged) activity of the neurons is $\yRf$}},
-    \\
-    &\text{\enquote{$\yxv_i = \yrv_i$} 
-      stands for \enquote{the activity of the $i$th sample neuron is $\yrv_i$}},
-  \end{aligned}
-\end{equation}
-and similarly for other quantities.
+argument \enquote{$(t)$} from our expressions.
+
+% Probabilities refer to statements about the quantities we observe. We use
+% the standard notation:
+% \begin{equation}
+%   \label{eq:notation_statements}
+%   \begin{aligned}
+%     &\text{\enquote{$\yXv_{\iota} = \yRv_{\iota}$} 
+%       stands for \enquote{the activity of the $\iota$th neuron is $\yRv_{\iota}$}},
+%     \\
+%     &\text{\enquote{$\yXf = \yRf$} 
+%       stands for \enquote{the (population-averaged) activity of the neurons is $\yRf$}},
+%     \\
+%     &\text{\enquote{$\yxv_i = \yrv_i$} 
+%       stands for \enquote{the activity of the $i$th sample neuron is $\yrv_i$}},
+%   \end{aligned}
+% \end{equation}
+% and similarly for other quantities.
 
@@
   \label{eq:joint_plaus}
-  \p(\yXv_1=\yRv_1, \yXv_2=\yRv_2, \dotsc, \yXv_\yNv=\yRv_\yNv \| \yHc)
+  \p(\yRv_1, \yRv_2, \dotsc, \yRv_\yNv \| \yHc)
   \quad\text{or}\quad
-\p(\yX =\yR \| \yHc), \quad \yR \in \set{0,1}^\yNv.
+\p(\yR \| \yHc), \quad \yR \in \set{0,1}^\yNv.
 \end{equation}
@@
   \label{eq:sample_plaus}
-  \p(\yxv_1=\yrv_1, \yxv_2=\yrv_2, \dotsc, \yxv_n=\yrv_n \| \yHc) \quad\text{or}\quad
-\p(\yx =\yr \| \yHc), \quad \yr \in \set{0,1}^n.
+  \p(\yrv_1, \yrv_2, \dotsc, \yrv_n \| \yHc) \quad\text{or}\quad
+\p(\yr \| \yHc), \quad \yr \in \set{0,1}^n.
 \end{equation}
@@
   \label{eq:joint_plaus_N_homog}
-  \p(\yX = \yR \|  \yH) \equiv
-  \sum_{\yRf}\p(\yX = \yR \| \yXf = \yRf, \yH)\,
-  \p(\yXf=\yRf \| \yH) =
-  \binom{\yNv}{\yNv\yRf}^{-1} \p(\yXf=\yRf \| \yH).
+  \p(\yR \|  \yH) \equiv
+  \sum_{\yRf}\p(\yR \| \yRf, \yH)\,
+  \p(\yRf \| \yH) =
+  \binom{\yNv}{\yNv\yRf}^{-1} \p(\yRf \| \yH).
 \end{equation}
@@
   \label{eq:conditional_hypergeometric}
-  \p(\yxs = \yrs \|\yXf=\yRf, \yH)=
+  \p(\yrs \|\yRf, \yH)=
   \binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}
@@
   \label{eq:subpop_average}
-  \p(\yxs=\yrs \| \yH) = \sum_{\yRf}
-  \p(\yxs = \yrs \|\yXf=\yRf, \yH)\,
-  \p(\yXf=\yRf \| \yH)
+  \p(\yrs \| \yH) = \sum_{\yRf}
+  \p(\yrs \|\yRf, \yH)\,
+  \p(\yRf \| \yH)
   = 
@@
   \ypp(\yrs\|\yRf)\,
-  \p(\yXf=\yRf \| \yH).
+  \p(\yRf \| \yH).
 \end{gather}
@@
   \label{eq:marginal}
-  \p(\yx = \yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs \| \yH).
+  \p(\yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yrs \| \yH).
 \end{equation}
@@
 The conditional probability
-$\p(\yxs = \yrs \|\yXf=\yRf, \yH) \equiv \ypp(\yrs \|\yRf)$, besides
+$\p(\yrs \|\yRf, \yH) \equiv \ypp(\yrs \|\yRf)$, besides
 relating the distributions for the population and sample activities via
@@
 for the full population, as follows:
-\begin{multline}
+\begin{equation}
   \label{eq:pullback_P}
@@
   \equiv
-  \sum_{\yrs} g(\yrs)\,\p(\yxs=\yrs \| \yH)
+  \sum_{\yrs} g(\yrs)\,\p(\yrs \| \yH)
   =
-  \sum_{\yrs} g(\yrs) \sum_{\yRf} \ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
-  ={}\\
+  \sum_{\yrs} g(\yrs) \sum_{\yRf} \ypp(\yrs\|\yRf)\,\p(\yRf \| \yH)
+  = %{}\\
   \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
-  \p(\yXf=\yRf \| \yH),
+  \p(\yRf \| \yH),
 %  =
@@
 %  \equiv  \expeb{g^*}.
-\end{multline}
+\end{equation}
 where the second step uses \eqn~\eqref{eq:subpop_average}. The last
@@
   c = \expe{g \| \yH} \equiv \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
-  \p(\yXf=\yRf \| \yH).
+  \p(\yRf \| \yH).
 \end{equation}
@@
     \binom{n}{m}^{-1}
-    \binom{n \yrs}{m}\, \p(\yxs=\yrs \| \yH)
+    \binom{n \yrs}{m}\, \p(\yrs \| \yH)
     % \sum_{N \yRf=0}^N %m!\,
@@
     \sum_{\yRf} %m!\,
-    \binom{\yNv \yRf}{m}\, \p(\yXf=\yRf \| \yH)
+    \binom{\yNv \yRf}{m}\, \p(\yRf \| \yH)
 \equiv    \expe{\av{\underbrace{\yR \dotsm \yR}_{\text{$m$ factors}}} \| \yH},
@@
 \caption{Log-density plot of the hypergeometric distribution
-  $%\p(\yxs = \yrs \|\yXf=\yRf, \yH)=
+  $%\p(\yrs \|\yRf, \yH)=
 \ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
@@
   \label{eq:roughly_equal_nN}
-  \p(\yxs=a \|\yH) \approx \p(\yXf=a \|\yH),\qquad
+  \p(\yrs=a \|\yH) \approx \p(\yRf=a \|\yH),\qquad
 \expeb{g(\yrs) \|\yH} \approx \expeb{g(\yRf) \|\yH},
@@
   \label{eq:app_maxent_sample}
-  \p(\yxs=\yrs \|\yHb)
+  \p(\yrs \|\yHb)
   =\frac{1}{\yk(\yl)}\,
@@
   \label{eq:app_maxent_pop}
-  \p(\yXf = \yRf \| \yHa)  = \frac{1}{\yK(\yL)}\,
+  \p(\yRf \| \yHa)  = \frac{1}{\yK(\yL)}\,
   \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
@@
   \label{eq:app_maxent_pop_marg}
-  \p(\yxs = \yrs \| \yHa)  = \frac{1}{\yK(\yL)}\, 
+  \p(\yrs \| \yHa)  = \frac{1}{\yK(\yL)}\, 
   \sum_{\yRf} \ypp(\yrs \|\yRf)
@@
 \eqref{eq:app_maxent_sample_constraints}, and in addition also the $n$
-equations $\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$,
+equations $\p(\yrs \| \yHa) = \p(\yrs \| \yHb)$,
 $\yrs=1/n,\dotsc,1$ (one equation is taken care of by the normalization of
@@
 \includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
-\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at
+\caption{Linear and log-plots of $\p(\yrs)$ constructed by \me\ at
   the population level followed by sample marginalization (blue triangles),
@@
 \includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}% 
-\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at
+\caption{Linear and log-plots of $\p(\yrs)$ constructed by \me\ at
   the population level followed by sample marginalization (blue triangles),
@@
 \includegraphics[width=0.9\linewidth]{comparison3.pdf}%
-\caption{Linear and log-plots of $\p(\yxs = \yrs)$ for a sample of $n=200$
+\caption{Linear and log-plots of $\p(\yrs)$ for a sample of $n=200$
   and constraints as in \eqn~\eqref{eq:constraints}, constructed by:
@@
   \expe{\yrs} = 0.0478,\qquad
-  \expe{\yxxs} = 0.00257,
+  \expe{\yrrs} = 0.00257,
   % ,\quad
@@
   \label{eq:app_maxent_pop_marg_unknown_N}
-  \p(\yxs = \yrs \| \yHd)  =
-  \sum_{\yNv} \p(\yxs = \yrs \| \yNN=\yNv,\yHd) \,
-  \p(\yNN=\yNv \|\yHd)
+  \p(\yrs \| \yHd)  =
+  \sum_{\yNv} \p(\yrs \| \yNv,\yHd) \,
+  \p(\yNv \|\yHd)
   ={}\\[-\jot]

### 2019-01-24 scripts/ME_4momsolution.m

@@ -1,0 +1,91 @@
+(* Homog maxent with 4 moments - corresponds to constraining variances *)
+
+Clear[MEhom4moments,MEhom4thmoment,MEhomvmoments,MEhom4momentsprob]; 
+
+MEhom4moments[m1_, m2_, m3_, m4_, nn_, wp_:MachinePrecision] := 
+ Block[{totalprob,
+	zzzm, js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
+	nn2=Binomial[nn,2], nn3=Binomial[nn,3], nn4=Binomial[nn,4], 
+	bvalues},
+       freeenm = Log@Sum[Binomial[nn, m]*
+			 Exp[js1*(m/nn-m1) + js2*(Binomial[m,2]/nn2-m2) +
+			     js3*(Binomial[m,3]/nn3-m3) +
+			     js4*(Binomial[m,4]/nn4-m4)],
+       {m, 0, nn}];
+  soluzm = 
+   NMinimize[freeenm, {js1, js2, js3, js4}, MaxIterations -> 10000, 
+    PrecisionGoal -> 4, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+  prob = Table[
+	  Binomial[nn, m]*Exp[js1*m/nn + js2*Binomial[m,2]/nn2 +
+		      js3*Binomial[m,3]/nn3 +
+		      js4*Binomial[m,4]/nn4]/. soluzm[[2]],
+	  {m, 0, nn}];
+  totalprob = Total[prob];prob=prob/totalprob;
+    bvalues={(ran.prob/nn)/m1-1,
+	     ((Binomial[ran,2]/nn2).prob)/m2-1,
+	     ((Binomial[ran,3]/nn3).prob)/m3-1,
+	     ((Binomial[ran,4]/nn4).prob)/m4-1};
+ Print["rel. discrepancies %: ", bvalues*100//N];
+ {js1 /. soluzm[[2]], js2 /. soluzm[[2]],
+  js3 /. soluzm[[2]], js4 /. soluzm[[2]],prob,totalprob}]
+
+MEhom4thmoment[m1_, m2_, m4_, nn_, wp_:MachinePrecision] := 
+ Block[{totcorr = meancorr,
+   totmm = meanmm,totalprob,
[…]

### 2019-01-24 scripts/ME_4momsolution2.m

@@ -1,0 +1,93 @@
+(* Homog maxent with 4 moments - corresponds to constraining variances *)
+
+Clear[MEhom4moments2,MEhom4thmoment,MEhomvmoments,MEhom4momentsprob]; 
+
+MEhom4moments2[m1_, m2_, m3_, m4_, nn_, wp_:MachinePrecision] := 
+ Block[{totalprob,
+	zzzm, js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
+	nn2=Binomial[nn,2], nn3=Binomial[nn,3], nn4=Binomial[nn,4], 
+	bvalues},
+       zzzm = Sum[Binomial[nn, m]*
+		  Exp[js1*m + js2*Binomial[m,2] +
+		      js3*Binomial[m,3] +
+		      js4*Binomial[m,4]],
+       {m, 0, nn}];
+  freeenm = Log[zzzm] - js1*m1*nn - js2*m2*nn2 -js3*m3*nn3 -js4*m4*nn4;
+  soluzm = 
+   NMinimize[freeenm, {js1, js2, js3, js4}, MaxIterations -> 10000, 
+    PrecisionGoal -> 6, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+  prob = Table[
+	  Binomial[nn, m]*
+	  		  Exp[js1*m + js2*Binomial[m,2] +
+		      js3*Binomial[m,3] +
+		      js4*Binomial[m,4]]/. soluzm[[2]],
+	  {m, 0, nn}];
+  totalprob = Total[prob];prob=prob/totalprob;
+    bvalues={(ran.prob/nn)/m1-1,
+	     ((Binomial[ran,2]/nn2).prob)/m2-1,
+	     ((Binomial[ran,3]/nn3).prob)/m3-1,
+	     ((Binomial[ran,4]/nn4).prob)/m4-1};
+ Print["rel. discrepancies %: ", bvalues*100//N];
+ {js1/nn /. soluzm[[2]], js2/nn2 /. soluzm[[2]],
+  js3/nn3 /. soluzm[[2]], js4/nn4 /. soluzm[[2]],prob,totalprob}]
+
+MEhom4thmoment[m1_, m2_, m4_, nn_, wp_:MachinePrecision] := 
+ Block[{totcorr = meancorr,
[…]

### 2019-01-24 scripts/ME_4momsolution3.m

@@ -1,0 +1,93 @@
+(* Homog maxent with 4 moments - corresponds to constraining variances *)
+
+Clear[MEhom4moments3,MEhom4thmoment,MEhomvmoments,MEhom4momentsprob]; 
+
+MEhom4moments3[m1_, m2_, m3_, m4_, nn_, wp_:MachinePrecision] := 
+ Block[{totalprob,
+	zzzm, js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
+	nn1=Sqrt[nn], nn2=Sqrt@Binomial[nn,2], nn3=Sqrt@Binomial[nn,3], nn4=Sqrt@Binomial[nn,4], 
+	bvalues},
+       zzzm = Sum[Binomial[nn, m]*
+		  Exp[js1*m/nn1 + js2*Binomial[m,2]/nn2 +
+		      js3*Binomial[m,3]/nn2 +
+		      js4*Binomial[m,4]/nn4],
+       {m, 0, nn}];
+  freeenm = Log[zzzm] - js1*m1*nn1 - js2*m2*nn2 -js3*m3*nn3 -js4*m4*nn4;
+  soluzm = 
+   NMinimize[freeenm, {js1, js2, js3, js4}, MaxIterations -> 10000, 
+    PrecisionGoal -> 6, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+  prob = Table[
+	  Binomial[nn, m]*
+	  		  Exp[js1*m/nn1 + js2*Binomial[m,2]/nn2 +
+		      js3*Binomial[m,3]/nn3 +
+		      js4*Binomial[m,4]/nn4]/. soluzm[[2]],
+	  {m, 0, nn}];
+  totalprob = Total[prob];prob=prob/totalprob;
+    bvalues={(ran.prob/nn)/m1-1,
+	     ((Binomial[ran,2]/nn2^2).prob)/m2-1,
+	     ((Binomial[ran,3]/nn3^2).prob)/m3-1,
+	     ((Binomial[ran,4]/nn4^2).prob)/m4-1};
+ Print["rel. discrepancies %: ", bvalues*100//N];
+ {js1/nn1 /. soluzm[[2]], js2/nn2 /. soluzm[[2]],
+  js3/nn3 /. soluzm[[2]], js4/nn4 /. soluzm[[2]],prob,totalprob}]
+
[…]

### 2019-01-24 scripts/ME_homsolution.m

@@ -1,0 +1,89 @@
+(* Homog maxent *)
+
+Clear[MEhomsolution,MEhomsolutiontruncated,MEhomprob]; 
+
+MEhomsolution::usage = "Input: pop-avgd activity [0,1], pop-avgd couple activity, pop size, machine precision. Output: Lagr multipl, Lagr multipl, distribution, Z"
+(* it uses normalized values in the optimization; usually gives more precise results than MEhomsolution2, although it reports difficulties in finding the optimum *)
+MEhomsolution[meanmm_, meancorr_, nn_, wp_:MachinePrecision] := 
+ Block[{totcorr = meancorr,
+   totmm = meanmm,totalprob,
+   zzzm, js2, js1, freeenm, varm, soluzm, prob,ran=Range[0,nn],bvalues},
+  zzzm = Sum[Binomial[nn, m]*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn], {m, 0, nn}]/2^nn;
+  freeenm = Log[zzzm] - js2*totcorr - js1*totmm;
+  soluzm = 
+   NMinimize[freeenm, {js1, js2}, MaxIterations -> 10000, 
+    PrecisionGoal -> 6,Method->Automatic, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+  prob = Table[
+    Binomial[nn, m]*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn]/2^nn /. soluzm[[2]], {m, 
+     0, nn}];
+  totalprob = Total[prob];prob=prob/totalprob;
+  bvalues={(ran.prob/nn)/meanmm-1,((ran^2-ran).prob/(nn^2-nn))/meancorr-1};
+ Print["rel. discrepancies %: ", bvalues*100//N];
+  {js1 /. soluzm[[2]], js2/. soluzm[[2]], prob,totalprob}]
+
+MEhomsolutiont[meanmm_, meancorr_, nn_, wp_:MachinePrecision] := 
+ Block[{totcorr = meancorr,
+   totmm = meanmm,totalprob,
+   js2, js1, freeenm, varm, soluzm, prob,ran=Range[0,nn],bvalues},
[…]

### 2019-01-24 scripts/ME_homsolution2.m

@@ -1,0 +1,49 @@
+(* Homog maxent *)
+
+Clear[MEhomsolution2,MEhomsolutiontruncated,MEhomprob]; 
+
+MEhomsolution2::usage = "Input: pop-avgd activity [0,1], pop-avgd couple activity, pop size, machine precision. Output: Lagr multipl, Lagr multipl, distribution, Z"
+(* it uses absolute values in the optimization; usually gives much less precise results than MEhomsolutio2, but it doesn't complain about difficulties in finding the optimum *)
+MEhomsolution2[meanmm_, meancorr_, nn_, wp_:MachinePrecision] := 
+  Block[{totcorr = meancorr*(nn^2-nn)/2,(* it uses absolute values in the optimization *)
+   totmm = meanmm*nn,totalprob,
+   zzzm, js2, js1, freeenm, varm, soluzm, prob,ran=Range[0,nn],bvalues},
+  zzzm = Sum[Binomial[nn, m]*Exp[js2*(m^2-m)/2 + js1*m], {m, 0, nn}]/2^nn;
+  freeenm = Log[zzzm] - js2*totcorr - js1*totmm;
+  soluzm = 
+   NMinimize[freeenm, {js1, js2}, MaxIterations -> 10000, 
+    PrecisionGoal -> 6,Method->Automatic, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+  prob = Table[
+    Binomial[nn, m]*Exp[js2*(m^2-m)/2 + js1*m]/2^nn /. soluzm[[2]], {m, 
+     0, nn}];
+  totalprob = Total[prob];prob=prob/totalprob;
+  bvalues={(ran.prob/nn)/meanmm-1,((ran^2-ran).prob/(nn^2-nn))/meancorr-1};
+ Print["rel. discrepancies %: ", bvalues*100//N];
+  {js1*nn /. soluzm[[2]], js2*(nn^2-nn)/2 /. soluzm[[2]], prob,totalprob}]
+
+ MEhomsolutiontruncated[meanmm_, meancorr_, nn_, trunc_, wp_:MachinePrecision] := 
+ Block[{totcorr = meancorr,
+   totmm = meanmm,totalprob,
[…]

### 2019-01-24 scripts/me_solution.R

@@ -1,0 +1,94 @@
+## Calculation of expected value, std, thetas, spread of conditional frequencies
+## of each symptom given each SNP - extra data set
+
+#### PREAMBLE
+## libraries and colour-blind palette from http://www.sron.nl/~pault/
+##runfunction <- function(aa=1e3L){
+library('ggplot2')
+library('RColorBrewer')
+library('cowplot')
+library('png')
+library('plot3D')
+library('doParallel')
+library('LaplacesDemon')
+library('dplyr')
+library('GA')
+#
+mypurpleblue <- '#4477AA'
+myblue <- '#66CCEE'
+mygreen <- '#228833'
+myyellow <- '#CCBB44'
+myred <- '#EE6677'
+myredpurple <- '#AA3377'
+mygrey <- '#BBBBBB'
+mycolours <- c(myblue, myred, mygreen, myyellow, myredpurple, mypurpleblue, mygrey, 'black')
+palette(mycolours)
+barpalette <- colorRampPalette(c(mypurpleblue,'white',myredpurple),space='Lab')
+barpalettepos <- colorRampPalette(c('white','black'),space='Lab')
+dev.off()
+mmtoin <- 0.0393701
+#### END PREAMBLE
+
+mehomsolution <- function(popsize,popmean,popcouple){
+    popsize2 <- (popsize^2-popsize)/2
+    rpopmean <- popmean*popsize
+    rpopcouple <- popcouple*popsize2
+
+    zterm <- function(js){
+        m <- 0:popsize
+        log(sum(exp(lchoose(popsize,m) + js[1]*m + js[2]*(m^2-m)/2))) - sum(js*c(rpopmean,rpopcouple))
+    }
+    
+    zterm <- function(js){
+        m <- 0:popsize
+        log(sum(choose(popsize,m)*exp( js[1]*m + js[2]*(m^2-m)/2))) - sum(js*c(rpopmean,rpopcouple))
+    }
+    
+    result <- nlm(f=zterm,p=c(0,0),iterlim=1e6)
+
+    result$estimate/c(popsize,popsize2)
+
+
+    # this seems to lead to solutions closer to the maximum
+    mehomsolution2 <- function(popsize,popmean,popcouple){
+    popsize2 <- (popsize^2-popsize)/2
+    rpopmean <- popmean*popsize
+    rpopcouple <- popcouple*popsize2
+
+    zterm <- function(js,cadd){
[…]

### 2019-01-24 scripts/stensola2mom_1000.m

@@ -1,0 +1,1028 @@
+solution = {-4492.3242675362289306417`12., 4482.4132898545298672193`12., 
+     {7.91049480882382492940182`10.245920169588343*^-6, 
+      0.00008855478022495072293256842`10.212842729433458, 
+      0.00049963532236660560823268416`10.18204799473724, 
+      0.00189437225335087367806430619`10.153237536519883, 
+      0.00543001046596453906883719244`10.126167884545389, 
+      0.01255126353402920705452869399`10.100637809304118, 
+      0.02436991410004108910458430064`10.076479082515338, 
+      0.04088220714041069442770645545`10.0535496334263, 
+      0.06048989901344833504487199744`10.031728392045089, 
+      0.08019338723911787409716781042`10.010911343680236, 
+      0.0964486203856138722022566098`9.99100846861537, 
+      0.10629665761519102946062402021`9.97194133882831, 
+      0.10824628638591237160821731032`9.953641209422251, 
+      0.10256571731826156785665190501`9.936047487382966, 
+      0.09096288670528169348707970491`9.919106491539287, 
+      0.07589620447317053861615569252`9.90277043969945, 
+      0.05984167552199169136054199095`9.886996614781005, 
+      0.04476253086483744240544677639`9.871746673268413, 
+      0.03187554578584053757605284233`9.856986067806504, 
+      0.02167578187149012157329206734`9.842683562045144, 
+      0.01411465642404770967850727911`9.828810820594862, 
+      0.00882329686349376817112523009`9.815342060557198, 
+      0.00530691307202063999265358193`9.802253753856865, 
+      0.0030775232034981838089741028`9.789524371739512, 
+      0.00172397226008121479270307176`9.777134164464652, 
+      0.00093450788430625434518160674`9.765064970531691, 
+      0.00049096936792351994826438634`9.75330005081194, 
+      0.00025037272778846636271894682`9.741823943783958, 
+      0.0001241013578150558368484851`9.730622338729932, 
+      0.00005986552225615228004782495`9.719681964283682, 
+      0.00002813869797707134225200621`9.70899049015261, 
+      0.00001290151042887337239506663`9.698536440187993, 
+      5.77614580985609244796574`9.68830911526665*^-6, 
+      2.52767148123571328409346`9.678298524684296*^-6, 
[…]

### 2019-01-24 scripts/stensola2mom_2000.m

@@ -1,0 +1,2029 @@
+solution = {-8984.458335089092200426`12., 8974.5617346377262884321`12., 
+     {6.254288766969503158307937147`9.944888418162927*^-11, 
+      1.40041768302555586537`9.928035278746478*^-9, 
+      1.574127194066642161159`9.911795866655991*^-8, 
+      1.1843041861433611615281`9.896126514217817*^-7, 
+      6.7093481423872238495371`9.880988055396363*^-7, 
+      3.05294953965692856960397`9.866345227537005*^-6, 
+      0.00001162277466172515831400089`9.852166169305178, 
+      0.00003807897505175728747525119`9.838421996837198, 
+      0.00010959750222130764606758603`9.82508644392277, 
+      0.00028151098857478048824779587`9.812135554950752, 
+      0.00065337762642330043262614681`9.799547421597557, 
+      0.00138411467510685312208798115`9.787301955987006, 
+      0.00269850240672702727640368293`9.775380694422683, 
+      0.00487576553393763342216950065`9.763766626877793, 
+      0.00821314322405807904886564485`9.752444048289629, 
+      0.01296414588223358774336704108`9.741398428395946, 
+      0.01926108474425174904092816907`9.730616297406243, 
+      0.0270408020461045611467811233`9.720085145251064, 
+      0.03599694418108770883512835727`9.70979333251912, 
+      0.04557863520474713725995722896`9.69973001149192, 
+      0.05504415454645637519078988157`9.689885055932672, 
+      0.0635626926139278233153036728`9.680248998489898, 
+      0.0703429149498906334963942346`9.670812974745893, 
+      0.07475901145721372524040370242`9.661568673081046, 
+      0.07644577564356084658778017427`9.652508289643364, 
+      0.0753433130093457957725536779`9.643624487811849, 
+      0.07168570016730366610697886513`9.634910361626023, 
+      0.06594164213169877062408644517`9.626359402724878, 
+      0.05872491957779981910189655543`9.61796547039877, 
+      0.05069612188990133590580330799`9.609722764408916, 
+      0.04247498938167560733758591464`9.601625800273256, 
+      0.03457647841101547515907396116`9.593669386754812, 
+      0.02737595132150500872562935649`9.585848605321157, 
+      0.02110196381113306424275790549`9.578158791371477, 
[…]

### 2019-01-24 scripts/stensola2mom_3000.m

@@ -1,0 +1,3029 @@
+solution = {-13476.5957602289250291442`12., 13466.7017593618299366635`12., 
+     {4.9450536240888997986249`9.768806083397697*^-16, 
+      1.660944811545583822196237`9.757498153465962*^-14, 
+      2.7968212956657890262585908`9.74646994527079*^-13, 
+      3.14802548934408574316757522`9.735707785060455*^-12, 
+      2.664576807970195091446769444`9.725198978002046*^-11, 
+      1.8091034557436583447973008979`9.714931716912574*^-10, 
+      1.02629553092884057782`9.704895001384962*^-9, 
+      5.00368262252023408745`9.695078565919847*^-9, 
+      2.140277813653743357685`9.685472815885978*^-8, 
+      8.159301796362427996728`9.676068770307596*^-8, 
+      2.8069409805726585582392`9.666858010623727*^-7, 
+      8.8018863878737450115031`9.657832634686548*^-7, 
+      2.53678992921510293722873`9.64898521536897*^-6, 
+      6.76684028877159936089221`9.640308763238044*^-6, 
+      0.00001680573058825192194836474`9.631796692824215, 
+      0.00003905892879783672318056651`9.62344279207855, 
+      0.00008533145904396469370207295`9.615241194663081, 
+      0.00017592342126078209916544123`9.607186354764682, 
+      0.00034345441798536218784315929`9.599273024161533, 
+      0.00063692448595711845745944356`9.591496231304763, 
+      0.001125083689380344453468716`9.583851262206348, 
+      0.00189778410232215226662048889`9.576333642949367, 
+      0.00306379523284637776881764341`9.568939123658016, 
+      0.00474375014088645035319450637`9.561663663783685, 
+      0.00705756220898055074905670595`9.554503418579413, 
+      0.01010678558345099902797547508`9.547454726649466, 
+      0.01395379044506264657612005414`9.540514098473047, 
+      0.01860094557527406827377274857`9.533678205812079, 
+      0.02397384143751447765251383255`9.526943871922619, 
+      0.02991260717365706832075161663`9.520308062497756, 
+      0.03617442193274817568587863377`9.513767877277443, 
+      0.04244850443509610357743520725`9.507320542267115, 
+      0.04838255002865275073976560432`9.5009634025129, 
+      0.05361730188864450941723893569`9.494693915386284, 
[…]

### 2019-01-24 scripts/stensola4mom_1000.m

@@ -1,0 +1,1032 @@
+solution = {-5419.6389678540528791535`12., 28178.0964201717551818647`12., 
+     985508.741383959412433281`12., -2.5031294027926990031736749`12.*^7, 
+     {0.00280393134755988956370734367`10.093572468336756, 
+      0.01241789771715728419457546011`10.065315721404486, 
+      0.02906457364212882915843706918`10.038517883966996, 
+      0.04822020363073988904342681755`10.012998855779081, 
+      0.06413700392239619376221089162`9.988604975431608, 
+      0.07329627556044720922565956407`9.965203881342811, 
+      0.07527753486281266184204331432`9.942680587819266, 
+      0.07171647904685520693115311169`9.920934449167154, 
+      0.06488740062541725490500518443`9.899876783063382, 
+      0.05677138684713285015428326466`9.879428990240589, 
+      0.04871462174857456945778579718`9.859521052535706, 
+      0.04145626415533736004320326483`9.840090322644082, 
+      0.03529710800597576121454664751`9.821080541034187, 
+      0.03027493419707517139662285351`9.80244103133644, 
+      0.02629589142721978087358994873`9.78412603704891, 
+      0.02321668656595953385216496635`9.766094170892318, 
+      0.02088900901212392318438074894`9.748307954473846, 
+      0.01917966920547630432921858814`9.730733430689549, 
+      0.01797684923738874374063320792`9.713339834931295, 
+      0.01718911026368950474545492826`9.696099313966407, 
+      0.01674088769623622631006397007`9.678986683540204, 
+      0.01656634335330100390665545123`9.661979217468655, 
+      0.01660244653512568622765940231`9.645056462353214, 
+      0.01678180956053743699968752207`9.628200073145274, 
+      0.01702596767595115099337711833`9.611393665675077, 
+      0.017240370879043155805426629`9.5946226829852, 
+      0.01731318485996050033744350326`9.577874272905566, 
+      0.01712070675810454727181747077`9.561137174801425, 
+      0.01654208775768085276078179138`9.544401613837437, 
+      0.01548417552144179601194198749`9.527659201444319, 
+      0.01391297939428918190391657242`9.510902840961464, 
+      0.01188215680156781899004872504`9.494126637667689, 
[…]

### 2019-01-24 scripts/stensola4mom_2000.m

@@ -1,0 +1,2031 @@
+solution = {-11073.3425032128154787224`12., 68134.1436286657463508225`12., 
+     1.954354673527362413545413`12.*^6, -5.39774526459431462506552932`12.*^7, 
+     {0.00003229783582380820038945255`9.77720794285961, 
+      0.00025448233769449348741647171`9.76304542440075, 
+      0.00103680531275094557036235516`9.749247098647201, 
+      0.00291654234051088066685441054`9.735789665565916, 
+      0.00638154058408607994113801663`9.722651645807964, 
+      0.01160017579080585081452896874`9.709813187617028, 
+      0.01827030394605966589781165974`9.697255899140325, 
+      0.02567437868567287679027748586`9.684962702256223, 
+      0.032895992984126955949814578`9.672917704709029, 
+      0.03907870627965901514447978546`9.66110608788637, 
+      0.04361913461595346384864161776`9.64951400801606, 
+      0.04624594702842306571718299589`9.638128508919182, 
+      0.04699527544067604694591451485`9.626937444750551, 
+      0.04612430095475955658060910499`9.615929411400286, 
+      0.04400756674504106330620422653`9.605093685430559, 
+      0.04104710481547589896517810671`9.594420169588044, 
+      0.03761046852603559861204519736`9.583899344071408, 
+      0.03399769211549870353109635557`9.573522222849357, 
+      0.03043114902918669135889699434`9.563280314422528, 
+      0.02706020560405539755552154022`9.553165586504871, 
+      0.02397343191220924864178861954`9.54317043417003, 
+      0.02121316058091820729101569106`9.53328765106748, 
+      0.01878926643377339864965283693`9.523510403363828, 
+      0.01669067183234651286348808618`9.513832206107843, 
+      0.01489415544216130848902947755`9.504246901754971, 
+      0.01337063885734328431181964839`9.494748640618973, 
+      0.01208938783931758099575635866`9.485331863045912, 
+      0.01102062503441797051695116451`9.475991283129629, 
+      0.01013700712684124309805747361`9.466721873808424, 
+      0.00941433429398817799383986775`9.457518853200922, 
+      0.0088317686911109207206562393`9.448377672054574, 
+      0.00837175809618270708610054744`9.43929400219434, 
+      0.0080197964408379683308204932`9.43026372587088, 
[…]

### 2019-01-24 scripts/stensola4mom_3000.m

@@ -1,0 +1,3031 @@
+solution = {-17655.4357444881600165945`12., 222141.7746610999886213283`12., 
+     -2.3667602556706461314249378`12.*^6, -762518.2460782486535458305`12., 
+     {4.29400665744575109053649`9.555501447167142*^-6, 
+      0.0000358177137796819052864324`9.546412977029526, 
+      0.00015689324699414859583182498`9.537436879013416, 
+      0.00048094113678364604103310602`9.528569342552734, 
+      0.00116007112400643244076879965`9.51980674590417, 
+      0.00234737328995038574872103926`9.51114564417596, 
+      0.00414843539046771186658470796`9.502582758285302, 
+      0.00658258541371086380306141263`9.494114964759326, 
+      0.00956851042436266817298214046`9.485739286303366, 
+      0.01293713071254016985498595236`9.477452883068333, 
+      0.01646429159905120112051233345`9.469253044555892, 
+      0.01991108423697868204552985318`9.461137182106325, 
+      0.023060466785172671561971072`9.453102821919499, 
+      0.02574301326383966826999597063`9.445147598564127, 
+      0.02784943685154799559211658869`9.43726924893497, 
+      0.02933126493468785469990289399`9.429465606621296, 
+      0.03019305333384663919529666658`9.4217345966536, 
+      0.03047999261187247995447558621`9.414074230598377, 
+      0.03026421750582350327218544446`9.406482601973757, 
+      0.02963214104426714584894517024`9.398957881961113, 
+      0.02867410966704016103329773414`9.391498315390017, 
+      0.02747684168386266052161014039`9.384102216975943, 
+      0.0261185457591448732896076964`9.376767967791865, 
+      0.02466630266472541788648145879`9.36949401195654, 
+      0.02317517349881681497525099074`9.362278853523728, 
+      0.02168850359502915846661797995`9.355121053557914, 
+      0.02023896563675340823553183039`9.348019227383315, 
+      0.01884998621655019515935490802`9.34097204199397, 
+      0.01753730109004084865949818332`9.333978213613825, 
+      0.01631047195079823558902416145`9.327036505396467, 
+      0.01517426649691671450142742602`9.320145725255097, 
+      0.01412985377522468949207181749`9.313304723813973, 
+      0.01317580075927555910436813058`9.306512392473339, 
+      0.01230887726101849950765788352`9.299767661580354, 
[…]

### 2019-01-24 scripts/testsave.m

@@ -1,0 +1,2 @@
+xj1 = -44921.5380444505662705784`12.
+xj2 = 44911.6307105902829471578`12.

### 2019-04-08 maxNt_v5.tex
update

--- 2019-01-24 maxNt_v3.tex
@@
 %% Author: PGL  Porta Mana
-%% Created: 2015-11-04T09:20:08+0100
-%% Last-Updated: 2018-12-18T13:21:16+0100
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%Report-no: ***
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-% \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading nips_2017
-%
-% to avoid loading the natbib package, add option nonatbib:
-% \usepackage[nonatbib]{nips_2017}
-
-\usepackage[preprint,%
-nonatbib]{nips_2018}
-
+%% Created: 2015-05-01T20:53:34+0200
+%% Last-Updated: 2019-04-08T11:32:41+0200
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newif\ifarxiv
+\arxivfalse
+\ifarxiv\pdfmapfile{+classico.map}\fi
+\newif\ifafour
+\afourfalse% true = A4, false = A5
+\newif\iftypodisclaim % typographical disclaim on the side
+\typodisclaimtrue
+\newcommand*{\memfontfamily}{zplx}
+\newcommand*{\memfontpack}{newpxtext}
+\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
+onecolumn,oneside,article,%french,italian,german,swedish,latin,
+british%
+]{memoir}
+\newcommand*{\updated}{\today}
 \newcommand*{\firstdraft}{4 November 2015}
-\newcommand*{\published}{22 May 2018}
-
-\newcommand*{\pdftitle}{Maximum-entropy and representative samples\\ of neuronal
-  activity: a dilemma}
-
-\newcommand*{\headtitle}{\pdftitle} \newcommand*{\pdfauthor}{P.G.L.  Porta Mana, V. Rostami, E. Torre, Y. Roudi}
-\usepackage{newtxmath}
+\newcommand*{\firstpublished}{***}
+\newcommand*{\propertitle}{Maximum-entropy distribution for a 
+  neuronal population\\ from subpopulation constraints%\\{\large ***}%
+}
+\newcommand*{\pdftitle}{\propertitle}
+\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
+\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
+\newcommand*{\headauthor}{Porta Mana \etal}
+\newcommand*{\reporthead}{}% Report number
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Calls to packages (uncomment as needed)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %\usepackage{pifont}
+
 %\usepackage{fontawesome}
+
 \usepackage[T1]{fontenc} 
 \input{glyphtounicode} \pdfgentounicode=1
+
 \usepackage[utf8]{inputenx}
+
 %\usepackage{newunicodechar}
@@
 
-\newcommand*{\bmmax}{3} % reduce number of bold fonts, before bm
-\newcommand*{\hmmax}{0} % reduce number of heavy fonts, before bm
+\newcommand*{\bmmax}{0} % reduce number of bold fonts, before font packages
+\newcommand*{\hmmax}{0} % reduce number of heavy fonts, before font packages
+
 \usepackage{textcomp}
+
+%\usepackage[normalem]{ulem}% package for underlining
+% \makeatletter
+% \def\ssout{\bgroup \ULdepth=-.35ex%\UL@setULdepth
+%  \markoverwith{\lower\ULdepth\hbox
+%    {\kern-.03em\vbox{\hrule width.2em\kern1.2\p@\hrule}\kern-.03em}}%
+%  \ULon}
+% \makeatother
+
 \usepackage{amsmath}
+
 \usepackage{mathtools}
+\addtolength{\jot}{\jot} % increase spacing in multiline formulae
 \setlength{\multlinegap}{0pt}
 
+%\usepackage{empheq}% automatically calls amsmath and mathtools
+\newcommand*{\widefbox}[1]{\fbox{\hspace{1em}#1\hspace{1em}}}
+
+%\usepackage{fancybox}
+
+%\usepackage{framed}
+
+% \usepackage[misc]{ifsym} % for dice
+% \newcommand*{\diceone}{{\scriptsize\Cube{1}}}
+
 \usepackage{amssymb}
+
 \usepackage{amsxtra}
 
-\usepackage[british]{babel}\selectlanguage{british}
+\usepackage[main=british,french,italian,german,swedish,latin,esperanto]{babel}\selectlanguage{british}
 \newcommand*{\langfrench}{\foreignlanguage{french}}
@@
 
-\let\openbox\relax
 \usepackage{amsthm}
@@
 
-\usepackage{bm}
-
-\iffalse
 \usepackage[shortlabels,inline]{enumitem}
@@
 \setlist[enumerate,2]{label=\alph*.}
-\setlist[enumerate]{leftmargin=\parindent}
-\setlist[itemize]{leftmargin=\parindent}
-\setlist[description]{leftmargin=\parindent}
-\fi
-
+\setlist[enumerate]{label=\arabic*.,leftmargin=1.5\parindent}
+\setlist[itemize]{leftmargin=1.5\parindent}
+\setlist[description]{leftmargin=1.5\parindent}
+% old alternative:
+% \setlist[enumerate,2]{label=\alph*.}
+% \setlist[enumerate]{leftmargin=\parindent}
+% \setlist[itemize]{leftmargin=\parindent}
+% \setlist[description]{leftmargin=\parindent}
+
+\usepackage[babel,theoremfont,largesc]{newpxtext}
+
+\usepackage[bigdelims,nosymbolsc%,smallerops % probably arXiv doesn't have it
+]{newpxmath}
+\useosf\linespread{1.083}
+%% smaller operators for old version of newpxmath
+\makeatletter
+\def\re@DeclareMathSymbol#1#2#3#4{%
+    \let#1=\undefined
+    \DeclareMathSymbol{#1}{#2}{#3}{#4}}
+%\re@DeclareMathSymbol{\bigsqcupop}{\mathop}{largesymbols}{"46}
+%\re@DeclareMathSymbol{\bigodotop}{\mathop}{largesymbols}{"4A}
+\re@DeclareMathSymbol{\bigoplusop}{\mathop}{largesymbols}{"4C}
+\re@DeclareMathSymbol{\bigotimesop}{\mathop}{largesymbols}{"4E}
+\re@DeclareMathSymbol{\sumop}{\mathop}{largesymbols}{"50}
+\re@DeclareMathSymbol{\prodop}{\mathop}{largesymbols}{"51}
+\re@DeclareMathSymbol{\bigcupop}{\mathop}{largesymbols}{"53}
+\re@DeclareMathSymbol{\bigcapop}{\mathop}{largesymbols}{"54}
+%\re@DeclareMathSymbol{\biguplusop}{\mathop}{largesymbols}{"55}
+\re@DeclareMathSymbol{\bigwedgeop}{\mathop}{largesymbols}{"56}
+\re@DeclareMathSymbol{\bigveeop}{\mathop}{largesymbols}{"57}
+%\re@DeclareMathSymbol{\bigcupdotop}{\mathop}{largesymbols}{"DF}
+%\re@DeclareMathSymbol{\bigcapplusop}{\mathop}{largesymbolsPXA}{"00}
+%\re@DeclareMathSymbol{\bigsqcupplusop}{\mathop}{largesymbolsPXA}{"02}
+%\re@DeclareMathSymbol{\bigsqcapplusop}{\mathop}{largesymbolsPXA}{"04}
+%\re@DeclareMathSymbol{\bigsqcapop}{\mathop}{largesymbolsPXA}{"06}
+\re@DeclareMathSymbol{\bigtimesop}{\mathop}{largesymbolsPXA}{"10}
+%\re@DeclareMathSymbol{\coprodop}{\mathop}{largesymbols}{"60}
+%\re@DeclareMathSymbol{\varprod}{\mathop}{largesymbolsPXA}{16}
+\makeatother
+%%
 %% With euler font cursive for Greek letters - the [1] means 100% scaling
 \DeclareFontFamily{U}{egreek}{\skewchar\font'177}%
-\DeclareFontShape{U}{egreek}{m}{n}{<-6>s*[0.95]eurm5 <6-8>s*[0.95]eurm7 <8->s*[0.95]eurm10}{}%
-\DeclareFontShape{U}{egreek}{m}{it}{<->s*[0.95]eurmo10}{}%
-\DeclareFontShape{U}{egreek}{b}{n}{<-6>s*[0.95]eurb5 <6-8>s*[0.95]eurb7 <8->s*[0.95]eurb10}{}%
-\DeclareFontShape{U}{egreek}{b}{it}{<->s*[0.95]eurbo10}{}%
+\DeclareFontShape{U}{egreek}{m}{n}{<-6>s*[1]eurm5 <6-8>s*[1]eurm7 <8->s*[1]eurm10}{}%
+\DeclareFontShape{U}{egreek}{m}{it}{<->s*[1]eurmo10}{}%
+\DeclareFontShape{U}{egreek}{b}{n}{<-6>s*[1]eurb5 <6-8>s*[1]eurb7 <8->s*[1]eurb10}{}%
+\DeclareFontShape{U}{egreek}{b}{it}{<->s*[1]eurbo10}{}%
 \DeclareSymbolFont{egreeki}{U}{egreek}{m}{it}%
@@
 \makeatother
-% Greek letters not usually given in LaTeX. Comment the unneeded ones
-% \DeclareMathSymbol{\varpartial}{\mathalpha}{egreeki}{"40}
-% \DeclareMathSymbol{\partialup}{\mathalpha}{egreekr}{"40}
-% \DeclareMathSymbol{\alpha}{\mathalpha}{egreeki}{"0B}
-% \DeclareMathSymbol{\beta}{\mathalpha}{egreeki}{"0C}
-% \DeclareMathSymbol{\gamma}{\mathalpha}{egreeki}{"0D}
-% \DeclareMathSymbol{\delta}{\mathalpha}{egreeki}{"0E}
-% \DeclareMathSymbol{\epsilon}{\mathalpha}{egreeki}{"0F}
- \DeclareMathSymbol{\zeta}{\mathalpha}{egreeki}{"10}
-% \DeclareMathSymbol{\eta}{\mathalpha}{egreeki}{"11}
-% \DeclareMathSymbol{\theta}{\mathalpha}{egreeki}{"12}
-% \DeclareMathSymbol{\iota}{\mathalpha}{egreeki}{"13}
- \DeclareMathSymbol{\kappa}{\mathalpha}{egreeki}{"14}
- \DeclareMathSymbol{\lambda}{\mathalpha}{egreeki}{"15}
-% \DeclareMathSymbol{\mu}{\mathalpha}{egreeki}{"16}
- \DeclareMathSymbol{\nu}{\mathalpha}{egreeki}{"17}
- \DeclareMathSymbol{\xi}{\mathalpha}{egreeki}{"18}
-% \DeclareMathSymbol{\omicron}{\mathalpha}{egreeki}{"6F}
-% \DeclareMathSymbol{\pi}{\mathalpha}{egreeki}{"19}
-% \DeclareMathSymbol{\rho}{\mathalpha}{egreeki}{"1A}
- \DeclareMathSymbol{\sigma}{\mathalpha}{egreeki}{"1B}
-% \DeclareMathSymbol{\tau}{\mathalpha}{egreeki}{"1C}
-% \DeclareMathSymbol{\upsilon}{\mathalpha}{egreeki}{"1D}
-% \DeclareMathSymbol{\phi}{\mathalpha}{egreeki}{"1E}
-% \DeclareMathSymbol{\chi}{\mathalpha}{egreeki}{"1F}
-% \DeclareMathSymbol{\psi}{\mathalpha}{egreeki}{"20}
-% \DeclareMathSymbol{\omega}{\mathalpha}{egreeki}{"21}
-% \DeclareMathSymbol{\varepsilon}{\mathalpha}{egreeki}{"22}
-% \DeclareMathSymbol{\vartheta}{\mathalpha}{egreeki}{"23}
-% \DeclareMathSymbol{\varpi}{\mathalpha}{egreeki}{"24}
-% \let\varrho\rho 
-% \let\varsigma\sigma
-% \let\varkappa\kappa
-% \DeclareMathSymbol{\varphi}{\mathalpha}{egreeki}{"27}
-% %
-% \DeclareMathSymbol{\varAlpha}{\mathalpha}{egreeki}{"41}
-% \DeclareMathSymbol{\varBeta}{\mathalpha}{egreeki}{"42}
-% \DeclareMathSymbol{\varGamma}{\mathalpha}{egreeki}{"00}
-% \DeclareMathSymbol{\varDelta}{\mathalpha}{egreeki}{"01}
-% \DeclareMathSymbol{\varEpsilon}{\mathalpha}{egreeki}{"45}
- \DeclareMathSymbol{\varZeta}{\mathalpha}{egreeki}{"5A}
-% \DeclareMathSymbol{\varEta}{\mathalpha}{egreeki}{"48}
-% \DeclareMathSymbol{\varTheta}{\mathalpha}{egreeki}{"02}
- \DeclareMathSymbol{\varIota}{\mathalpha}{egreeki}{"49}
- \DeclareMathSymbol{\varKappa}{\mathalpha}{egreeki}{"4B}
- \DeclareMathSymbol{\varLambda}{\mathalpha}{egreeki}{"03}
-% \DeclareMathSymbol{\varMu}{\mathalpha}{egreeki}{"4D}
- \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}
- \DeclareMathSymbol{\varXi}{\mathalpha}{egreeki}{"04}
-% \DeclareMathSymbol{\varOmicron}{\mathalpha}{egreeki}{"4F}
-% \DeclareMathSymbol{\varPi}{\mathalpha}{egreeki}{"05}
-% \DeclareMathSymbol{\varRho}{\mathalpha}{egreeki}{"50}
- \DeclareMathSymbol{\varSigma}{\mathalpha}{egreeki}{"06}
-% \DeclareMathSymbol{\varTau}{\mathalpha}{egreeki}{"54}
- \DeclareMathSymbol{\varUpsilon}{\mathalpha}{egreeki}{"07}
-% \DeclareMathSymbol{\varPhi}{\mathalpha}{egreeki}{"08}
-% \DeclareMathSymbol{\varChi}{\mathalpha}{egreeki}{"58}
-% \DeclareMathSymbol{\varPsi}{\mathalpha}{egreeki}{"09}
-% \DeclareMathSymbol{\varOmega}{\mathalpha}{egreeki}{"0A} 
-% %
-% \DeclareMathSymbol{\Alpha}{\mathalpha}{egreekr}{"41}
-% \DeclareMathSymbol{\Beta}{\mathalpha}{egreekr}{"42}
-% \DeclareMathSymbol{\Gamma}{\mathalpha}{egreekr}{"00}
-% \DeclareMathSymbol{\Delta}{\mathalpha}{egreekr}{"01}
-% \DeclareMathSymbol{\Epsilon}{\mathalpha}{egreekr}{"45}
-% \DeclareMathSymbol{\Zeta}{\mathalpha}{egreekr}{"5A}
-% \DeclareMathSymbol{\Eta}{\mathalpha}{egreekr}{"48}
-% \DeclareMathSymbol{\Theta}{\mathalpha}{egreekr}{"02}
-% \DeclareMathSymbol{\Iota}{\mathalpha}{egreekr}{"49}
-% \DeclareMathSymbol{\Kappa}{\mathalpha}{egreekr}{"4B}
-% \DeclareMathSymbol{\Lambda}{\mathalpha}{egreekr}{"03}
-% \DeclareMathSymbol{\Mu}{\mathalpha}{egreekr}{"4D}
-% \DeclareMathSymbol{\Nu}{\mathalpha}{egreekr}{"4E}
-% \DeclareMathSymbol{\Xi}{\mathalpha}{egreekr}{"04}
-% \DeclareMathSymbol{\Omicron}{\mathalpha}{egreekr}{"4F}
-% \DeclareMathSymbol{\Pi}{\mathalpha}{egreekr}{"05}
-% \DeclareMathSymbol{\Rho}{\mathalpha}{egreekr}{"50}
-% \DeclareMathSymbol{\Sigma}{\mathalpha}{egreekr}{"06}
-% \DeclareMathSymbol{\Tau}{\mathalpha}{egreekr}{"54}
-% \DeclareMathSymbol{\Upsilon}{\mathalpha}{egreekr}{"07}
-% \DeclareMathSymbol{\Phi}{\mathalpha}{egreekr}{"08}
-% \DeclareMathSymbol{\Chi}{\mathalpha}{egreekr}{"58}
-% \DeclareMathSymbol{\Psi}{\mathalpha}{egreekr}{"09}
-% \DeclareMathSymbol{\Omega}{\mathalpha}{egreekr}{"0A}
-% %
-% \DeclareMathSymbol{\alphaup}{\mathalpha}{egreekr}{"0B}
-% \DeclareMathSymbol{\betaup}{\mathalpha}{egreekr}{"0C}
-% \DeclareMathSymbol{\gammaup}{\mathalpha}{egreekr}{"0D}
-% \DeclareMathSymbol{\deltaup}{\mathalpha}{egreekr}{"0E}
-% \DeclareMathSymbol{\epsilonup}{\mathalpha}{egreekr}{"0F}
-% \DeclareMathSymbol{\zetaup}{\mathalpha}{egreekr}{"10}
-% \DeclareMathSymbol{\etaup}{\mathalpha}{egreekr}{"11}
-% \DeclareMathSymbol{\thetaup}{\mathalpha}{egreekr}{"12}
-% \DeclareMathSymbol{\iotaup}{\mathalpha}{egreekr}{"13}
-% \DeclareMathSymbol{\kappaup}{\mathalpha}{egreekr}{"14}
-% \DeclareMathSymbol{\lambdaup}{\mathalpha}{egreekr}{"15}
-% \DeclareMathSymbol{\muup}{\mathalpha}{egreekr}{"16}
-% \DeclareMathSymbol{\nuup}{\mathalpha}{egreekr}{"17}
-% \DeclareMathSymbol{\xiup}{\mathalpha}{egreekr}{"18}
-% \DeclareMathSymbol{\omicronup}{\mathalpha}{egreekr}{"6F}
-%  \DeclareMathSymbol{\piup}{\mathalpha}{egreekr}{"19}
-% \DeclareMathSymbol{\rhoup}{\mathalpha}{egreekr}{"1A}
-% \DeclareMathSymbol{\sigmaup}{\mathalpha}{egreekr}{"1B}
-% \DeclareMathSymbol{\tauup}{\mathalpha}{egreekr}{"1C}
-% \DeclareMathSymbol{\upsilonup}{\mathalpha}{egreekr}{"1D}
-% \DeclareMathSymbol{\phiup}{\mathalpha}{egreekr}{"1E}
-% \DeclareMathSymbol{\chiup}{\mathalpha}{egreekr}{"1F}
-% \DeclareMathSymbol{\psiup}{\mathalpha}{egreekr}{"20}
-% \DeclareMathSymbol{\omegaup}{\mathalpha}{egreekr}{"21}
-% \DeclareMathSymbol{\varepsilonup}{\mathalpha}{egreekr}{"22}
-% \DeclareMathSymbol{\varthetaup}{\mathalpha}{egreekr}{"23}
-% \DeclareMathSymbol{\varpiup}{\mathalpha}{egreekr}{"24}
-% \let\varrhoup\rhoup 
-% \let\varsigmaup\sigmaup
-% \let\varkappaup\kappaup
-% \DeclareMathSymbol{\varphiup}{\mathalpha}{egreekr}{"27}
-
+\input{definegreek.tex}% Greek letters not usually given in LaTeX.
+
+%\usepackage%[scaled=0.9]%
+%{classico}%  Optima as sans-serif font
+\renewcommand\sfdefault{uop}
+\DeclareMathAlphabet{\mathsf}  {T1}{\sfdefault}{m}{sl}
+\SetMathAlphabet{\mathsf}{bold}{T1}{\sfdefault}{b}{sl}
 %\newcommand*{\mathte}[1]{\textbf{\textit{\textsf{#1}}}}
@@
 
+% DejaVu Mono as typewriter text
+\usepackage[scaled=0.84]{DejaVuSansMono}
+
 \usepackage{mathdots}
@@
 \usepackage[usenames]{xcolor}
-\definecolor{mybluishpurple}{RGB}{51,34,136}
-\definecolor{myblue}{RGB}{136,204,238}
-\definecolor{mybluishgreen}{RGB}{68,170,153}
-\definecolor{mygreen}{RGB}{17,119,51}
-\definecolor{mygreenishyellow}{RGB}{153,153,51}
-\definecolor{myyellow}{RGB}{221,204,119}
-\definecolor{myred}{RGB}{204,102,119}
-\definecolor{mypurplishred}{RGB}{136,34,85}
-\definecolor{myreddishpurple}{RGB}{170,68,153}
-\definecolor{mygrey}{RGB}{221,221,221}
+% Tol (2012) colour-blind-, print-, screen-friendly colours, alternative scheme; Munsell terminology
+\definecolor{mypurpleblue}{RGB}{68,119,170}
+\definecolor{myblue}{RGB}{102,204,238}
+\definecolor{mygreen}{RGB}{34,136,51}
+\definecolor{myyellow}{RGB}{204,187,68}
+\definecolor{myred}{RGB}{238,102,119}
+\definecolor{myredpurple}{RGB}{170,51,119}
+\definecolor{mygrey}{RGB}{187,187,187}
+% Tol (2012) colour-blind-, print-, screen-friendly colours; Munsell terminology
+% \definecolor{lbpurple}{RGB}{51,34,136}
+% \definecolor{lblue}{RGB}{136,204,238}
+% \definecolor{lbgreen}{RGB}{68,170,153}
+% \definecolor{lgreen}{RGB}{17,119,51}
+% \definecolor{lgyellow}{RGB}{153,153,51}
+% \definecolor{lyellow}{RGB}{221,204,119}
+% \definecolor{lred}{RGB}{204,102,119}
+% \definecolor{lpred}{RGB}{136,34,85}
+% \definecolor{lrpurple}{RGB}{170,68,153}
+\definecolor{lgrey}{RGB}{221,221,221}
+%\newcommand*\mycolourbox[1]{%
+%\colorbox{mygrey}{\hspace{1em}#1\hspace{1em}}}
+\colorlet{shadecolor}{lgrey}
+
+\usepackage{bm}
 
@@
 
-\usepackage[backend=biber,mcite,subentry,citestyle=numeric-comp,bibstyle=numericbringhurst,autopunct=false,sorting=none,sortcites=false,natbib=false,maxnames=8,minnames=8,giveninits=true,block=space,hyperref=true,defernumbers=false,useprefix=true,language=british]{biblatex}
-\renewcommand*{\finalnamedelim}{, }
+\usepackage[backend=biber,mcite,%subentry,
+citestyle=authoryear-comp,bibstyle=pglpm-authoryear,autopunct=false,sorting=ny,sortcites=false,natbib=false,maxcitenames=1,maxbibnames=8,minbibnames=8,giveninits=true,uniquename=false,uniquelist=false,maxalphanames=1,block=space,hyperref=true,defernumbers=false,useprefix=true,sortupper=false,language=british,parentracker=false]{biblatex}
+\DeclareSortingScheme{ny}{\sort{\field{sortname}\field{author}\field{editor}}\sort{\field{year}}}
+\iffalse\makeatletter%%% replace parenthesis with brackets
+\newrobustcmd*{\parentexttrack}[1]{%
+  \begingroup
+  \blx@blxinit
+  \blx@setsfcodes
+  \blx@bibopenparen#1\blx@bibcloseparen
+  \endgroup}
+\AtEveryCite{%
+  \let\parentext=\parentexttrack%
+  \let\bibopenparen=\bibopenbracket%
+  \let\bibcloseparen=\bibclosebracket}
+\makeatother\fi
+\DefineBibliographyExtras{british}{\def\finalandcomma{\addcomma}}
+\renewcommand*{\finalnamedelim}{\addcomma\space}
 \setcounter{biburlnumpenalty}{1}
@@
 \DeclareDelimFormat{multicitedelim}{\addsemicolon\space}
+\DeclareDelimFormat{compcitedelim}{\addsemicolon\space}
 \DeclareDelimFormat{postnotedelim}{\space}
-\addbibresource{portamanabib.bib}%comment for arxiv
+\ifarxiv\else\addbibresource{portamanabib.bib}\fi
 \renewcommand{\bibfont}{\footnotesize}
-%\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
-%}
+%\appto{\citesetup}{\footnotesize}% smaller font for citations
+\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
+}
 \newcommand*{\citep}{\parencites}
 \newcommand*{\citey}{\parencites*}
-\renewcommand*{\cite}{\citep}
+%\renewcommand*{\cite}{\parencite}
+\renewcommand*{\cites}{\parencites}
 \providecommand{\href}[2]{#2}
@@
 \newcommand*{\amp}{\&}
-
+% \newcommand*{\citein}[2][]{\textnormal{\textcite[#1]{#2}}%\addtocategory{extras}{#2}
+% }
+\newcommand*{\citein}[2][]{\textnormal{\textcite[#1]{#2}}%\addtocategory{extras}{#2}
+}
+\newcommand*{\citebi}[2][]{\textcite[#1]{#2}%\addtocategory{extras}{#2}
+}
+\newcommand*{\subtitleproc}[1]{}
+\newcommand*{\chapb}{ch.}
+%
+% \def\arxivp{}
+% \def\mparcp{}
+% \def\philscip{}
+% \def\biorxivp{}
+% \newcommand*{\arxivsi}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.\\}
+% \newcommand*{\mparcsi}{\texttt{mp\_arc} eprints available at \url{http://www.ma.utexas.edu/mp_arc/}.\\}
+% \newcommand*{\philscisi}{\texttt{philsci} eprints available at \url{http://philsci-archive.pitt.edu/}.\\}
+% \newcommand*{\biorxivsi}{\texttt{bioRxiv} eprints available at \url{http://biorxiv.org/}.\\}
 \newcommand*{\arxiveprint}[1]{%\global\def\arxivp{\arxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
@@
 \newcommand*{\biorxiveprint}[1]{%\global\def\biorxivp{\biorxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
-\texttt{\urlalt{http://biorxiv.org/content/early/#1}{bioRxiv:\hspace{0pt}#1}}%
+\texttt{\urlalt{https://doi.org/10.1101/#1}{bioRxiv doi:\hspace{0pt}10.1101/#1}}%
 %\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
@@
 \newcommand*{\osfeprint}[1]{%
-\texttt{\urlalt{https://doi.org/10.17605/osf.io/#1}{doi:10.17605/osf.io/#1}}%
+\texttt{\urlalt{https://doi.org/10.17605/osf.io/#1}{Open Science Framework doi:10.17605/osf.io/#1}}%
 }
@@
 
+%\usepackage{wrapfig}
+
+%\usepackage{tikz-cd}
+
 \PassOptionsToPackage{hyphens}{url}\usepackage[hypertexnames=false]{hyperref}
+
 \usepackage[depth=4]{bookmark}
-\hypersetup{colorlinks=true,bookmarksnumbered,pdfborder={0 0 0.25},citebordercolor={0.2 0.1333 0.5333},%bluish
-citecolor=mybluishpurple,linkbordercolor={0.0667 0.4667 0.2},%greenish
-linkcolor=mypurplishred,urlbordercolor={0.5333 0.1333 0.3333},%reddish
-urlcolor=mygreen,breaklinks=true,pdftitle={\pdftitle},pdfauthor={\pdfauthor}}
+\hypersetup{colorlinks=true,bookmarksnumbered,pdfborder={0 0 0.25},citebordercolor={0.2667 0.4667 0.6667},citecolor=mypurpleblue,linkbordercolor={0.6667 0.2 0.4667},linkcolor=myredpurple,urlbordercolor={0.1333 0.5333 0.2},urlcolor=mygreen,breaklinks=true,pdftitle={\pdftitle},pdfauthor={\pdfauthor}}
 % \usepackage[vertfit=local]{breakurl}% only for arXiv
@@
 
-%\usepackage{wrapfig}
-\usepackage{datetime2}
+\usepackage[british]{datetime2}
 \DTMnewdatestyle{mydate}%
@@
 
-% handling orphan/widow lines:
-\clubpenalty=10000
-\widowpenalty=10000
-\raggedbottom
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Layout. I do not know on which kind of paper the reader will print the
+%%% paper on (A4? letter? one-sided? double-sided?). So I choose A5, which
+%%% provides a good layout for reading on screen and save paper if printed
+%%% two pages per sheet. Average length line is 66 characters and page
+%%% numbers are centred.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\ifafour\setstocksize{297mm}{210mm}%{*}% A4
+\else\setstocksize{210mm}{5.5in}%{*}% 210x139.7
+\fi
+\settrimmedsize{\stockheight}{\stockwidth}{*}
+\setlxvchars[\normalfont] %313.3632pt for a 66-characters line
+\setxlvchars[\normalfont]
+\setlength{\trimtop}{0pt}
+\setlength{\trimedge}{\stockwidth}
+\addtolength{\trimedge}{-\paperwidth}
+% The length of the normalsize alphabet is 133.05988pt - 10 pt = 26.1408pc
+% The length of the normalsize alphabet is 159.6719pt - 12pt = 30.3586pc
+% Bringhurst gives 32pc as boundary optimal with 69 ch per line
+% The length of the normalsize alphabet is 191.60612pt - 14pt = 35.8634pc
+\ifafour\settypeblocksize{*}{32pc}{1.618} % A4
+%\setulmargins{*}{*}{1.667}%gives 5/3 margins % 2 or 1.667
+\else\settypeblocksize{*}{26pc}{1.618}% nearer to a 66-line newpx and preserves GR
+\fi
+\setulmargins{*}{*}{1}%gives equal margins
+\setlrmargins{*}{*}{*}
+\setheadfoot{\onelineskip}{2.5\onelineskip}
+\setheaderspaces{*}{2\onelineskip}{*}
+\setmarginnotes{2ex}{10mm}{0pt}
+\checkandfixthelayout[nearest]
+\fixpdflayout
+%%% End layout
+%% this fixes missing white spaces
+\pdfmapline{+dummy-space <dummy-space.pfb}\pdfinterwordspaceon%
+
+%%% Sectioning
+\newcommand*{\asudedication}[1]{%
+{\par\centering\textit{#1}\par}}
+\newenvironment{acknowledgements}{\section*{Thanks}\addcontentsline{toc}{section}{Thanks}}{\par}
+\makeatletter\renewcommand{\appendix}{\par
+  \bigskip{\centering
+   \interlinepenalty \@M
+   \normalfont
+   \printchaptertitle{\sffamily\appendixpagename}\par}
+  \setcounter{section}{0}%
+  \gdef\@chapapp{\appendixname}%
+  \gdef\thesection{\@Alph\c@section}%
+  \anappendixtrue}\makeatother
+\counterwithout{section}{chapter}
+\setsecnumformat{\upshape\csname the#1\endcsname\quad}
+\setsecheadstyle{\large\bfseries\sffamily%
+\centering}
+\setsubsecheadstyle{\bfseries\sffamily%
+\raggedright}
+%\setbeforesecskip{-1.5ex plus 1ex minus .2ex}% plus 1ex minus .2ex}
+%\setaftersecskip{1.3ex plus .2ex }% plus 1ex minus .2ex}
+%\setsubsubsecheadstyle{\bfseries\sffamily\slshape\raggedright}
+%\setbeforesubsecskip{1.25ex plus 1ex minus .2ex }% plus 1ex minus .2ex}
+%\setaftersubsecskip{-1em}%{-0.5ex plus .2ex}% plus 1ex minus .2ex}
+\setsubsecindent{0pt}%0ex plus 1ex minus .2ex}
+\setparaheadstyle{\bfseries\sffamily%
+\raggedright}
+\setcounter{secnumdepth}{2}
+\setlength{\headwidth}{\textwidth}
+\newcommand{\addchap}[1]{\chapter*[#1]{#1}\addcontentsline{toc}{chapter}{#1}}
+\newcommand{\addsec}[1]{\section*{#1}\addcontentsline{toc}{section}{#1}}
+\newcommand{\addsubsec}[1]{\subsection*{#1}\addcontentsline{toc}{subsection}{#1}}
+\newcommand{\addpara}[1]{\paragraph*{#1.}\addcontentsline{toc}{subsubsection}{#1}}
+\newcommand{\addparap}[1]{\paragraph*{#1}\addcontentsline{toc}{subsubsection}{#1}}
+
+%%% Headers, footers, pagestyle
+\copypagestyle{manaart}{plain}
+\makeheadrule{manaart}{\headwidth}{0.5\normalrulethickness}
+\makeoddhead{manaart}{%
+{\footnotesize%\sffamily%
+\scshape\headauthor}}{}{{\footnotesize\sffamily%
+\headtitle}}
+\makeoddfoot{manaart}{}{\thepage}{}
+\newcommand*\autanet{\includegraphics[height=\heightof{M}]{autanet.pdf}}
+\definecolor{mygray}{gray}{0.333}
+\iftypodisclaim%
+\ifafour\newcommand\addprintnote{\begin{picture}(0,0)%
+\put(245,149){\makebox(0,0){\rotatebox{90}{\tiny\color{mygray}\textsf{This
+            document is designed for screen reading and
+            two-up printing on A4 or Letter paper}}}}%
+\end{picture}}% A4
+\else\newcommand\addprintnote{\begin{picture}(0,0)%
+\put(176,112){\makebox(0,0){\rotatebox{90}{\tiny\color{mygray}\textsf{This
+            document is designed for screen reading and
+            two-up printing on A4 or Letter paper}}}}%
+\end{picture}}\fi%afourtrue
+\makeoddfoot{plain}{}{\makebox[0pt]{\thepage}\addprintnote}{}
+\else
+\makeoddfoot{plain}{}{\makebox[0pt]{\thepage}}{}
+\fi%typodisclaimtrue
+\makeoddhead{plain}{}{}{\footnotesize\reporthead}
+% \copypagestyle{manainitial}{plain}
+% \makeheadrule{manainitial}{\headwidth}{0.5\normalrulethickness}
+% \makeoddhead{manainitial}{%
+% \footnotesize\sffamily%
+% \scshape\headauthor}{}{\footnotesize\sffamily%
+% \headtitle}
+% \makeoddfoot{manaart}{}{\thepage}{}
+
+\pagestyle{manaart}
+
+\setlength{\droptitle}{-3.9\onelineskip}
+\pretitle{\begin{center}\Large\sffamily%
+\bfseries}
+\posttitle{\bigskip\end{center}}
+
+\makeatletter\newcommand*{\atf}{\includegraphics[%trim=1pt 1pt 0pt 0pt,
+totalheight=\heightof{@}]{atblack.png}}\makeatother
+\providecommand{\affiliation}[1]{\textsl{\textsf{\footnotesize #1}}}
+\providecommand{\epost}[1]{\texttt{\footnotesize\textless#1\textgreater}}
+\providecommand{\email}[2]{\href{mailto:#1ZZ@#2 ((remove ZZ))}{#1\protect\atf#2}}
+
+\preauthor{\vspace{-0.5\baselineskip}\begin{center}
+\normalsize\sffamily%
+\lineskip  0.5em}
+\postauthor{\par\end{center}}
+\predate{\DTMsetdatestyle{mydate}\begin{center}\footnotesize}
+\postdate{\end{center}\vspace{-\medskipamount}}
+
+\setfloatadjustment{figure}{\footnotesize}
+\captiondelim{\quad}
+\captionnamefont{\footnotesize\sffamily%
+}
+\captiontitlefont{\footnotesize}
+\firmlists*
+\midsloppy
+% handling orphan/widow lines, memman.pdf
+% \clubpenalty=10000
+% \widowpenalty=10000
+% \raggedbottom
+% Downes, memman.pdf
+\clubpenalty=9996
+\widowpenalty=9999
+\brokenpenalty=4991
+\predisplaypenalty=10000
+\postdisplaypenalty=1549
+\displaywidowpenalty=1602
 \selectlanguage{british}\frenchspacing
 
-% to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage[final]{nips_2017}
-
-%\usepackage[utf8]{inputenc} % allow utf-8 input
-%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-%\usepackage{hyperref}       % hyperlinks
-%\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-%\usepackage{amsfonts}       % blackboard math symbols
-%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-%\usepackage{microtype}      % microtypography
-
-% @@@@@@@@@@ new macros @@@@@@@@@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Paper's details
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{\propertitle}
+\author{%
+\hspace*{\stretch{1}}%
+\parbox{0.5\linewidth}%\makebox[0pt][c]%
+{\protect\centering P.G.L.  Porta Mana\\%
+\footnotesize\epost{\email{piero.mana}{ntnu.no}}}%
+\hspace*{\stretch{1}}%
+%% uncomment if additional authors present
+\parbox{0.5\linewidth}%\makebox[0pt][c]%
+{\protect\centering Y. Roudi\\%
+\footnotesize\epost{\email{yasser.roudi}{ntnu.no}}}%
+\hspace*{\stretch{1}}\\%
+\parbox{0.5\linewidth}%\makebox[0pt][c]%
+{\protect\centering V. Rostami\\%
+\footnotesize\epost{\email{vrostami}{uni-koeln.de}}}%
+\hspace*{\stretch{1}}%
+\parbox{0.5\linewidth}%\makebox[0pt][c]%
+{\protect\centering E. Torre\\%
+\footnotesize\epost{\email{torre}{ibk.baug.ethz.ch}}}%
+\hspace*{\stretch{1}}%
+%\quad\href{https://orcid.org/0000-0002-6070-0784}{\protect\includegraphics[scale=0.16]{orcid_32x32.png}\textsc{orcid}:0000-0002-6070-0784}%
+}
+
+\date{Draft of \today\ (first drafted \firstdraft)}
+%\date{\firstpublished; updated \updated}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Macros @@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % Common ones - uncomment as needed
@@
 %\providecommand{\varprod}{\prod}
-%\newcommand*{\de}{\partialup}%partial diff
-%\newcommand*{\pu}{\piup}%constant pi
+\newcommand*{\de}{\partialup}%partial diff
+\newcommand*{\pu}{\piup}%constant pi
 \newcommand*{\delt}{\deltaup}%Kronecker, Dirac
@@
 %\newcommand*{\micro}{\textmu}% degree Celsius
-%\newcommand*{\I}{\mathrm{i}}%imaginary unit
-%\newcommand*{\e}{\mathrm{e}}%Neper
-%\newcommand*{\di}{\mathrm{d}}%differential
+\newcommand*{\I}{\mathrm{i}}%imaginary unit
+\newcommand*{\e}{\mathrm{e}}%Neper
+\newcommand*{\di}{\mathrm{d}}%differential
 %\newcommand*{\Di}{\mathrm{D}}%capital differential
@@
 %\newcommand*{\Lor}{\bigvee}
-%\newcommand*{\lland}{\mathbin{\ \land\ }}
-%\newcommand*{\llor}{\mathbin{\ \lor\ }}
-%\newcommand*{\lonlyif}{\mathbin{\Rightarrow}}%implies
-\newcommand*{\limplies}{\mathbin{\Rightarrow}}%implies
-%\newcommand*{\mimplies}{\Rightarrow}%implies
-%\newcommand*{\liff}{\mathbin{\Leftrightarrow}}%if and only if
-\renewcommand*{\|}{\mathpunct{|}}%conditional sign (in probabilities)
-%\newcommand*{\lcond}{\mathpunct{|\ }}%conditional sign (in probabilities)
-%\newcommand*{\bigcond}{\mathpunct{\big|}}%conditional sign (in probabilities)
-%\newcommand*{\lbigcond}{\mathpunct{\big|\ }}%conditional sign (in probabilities)
+%\newcommand*{\lland}{\DOTSB\;\land\;}
+%\newcommand*{\llor}{\DOTSB\;\lor\;}
+%\newcommand*{\limplies}{\mathbin{\Rightarrow}}%implies
 %\newcommand*{\suchthat}{\mid}%{\mathpunct{|}}%such that (eg in sets)
-%\newcommand*{\bigst}{\mathpunct{\big|}}%such that (eg in sets)
 %\newcommand*{\with}{\colon}%with (list of indices)
@@
 %\newcommand*{\inn}{\cdot}%inner product
-\newcommand*{\dotv}{\mathord{\,\cdot\,}}%variable place
+%\newcommand*{\dotv}{\mathord{\,\cdot\,}}%variable place
 %\newcommand*{\comp}{\circ}%composition of functions
@@
 %\newcommand*{\equi}{\sim}%equivalent to 
+\renewcommand*{\asymp}{\simeq}%equivalent to 
 %\newcommand*{\corr}{\mathrel{\hat{=}}}%corresponds to
 %\providecommand{\varparallel}{\ensuremath{\mathbin{/\mkern-7mu/}}}%parallel (tentative symbol)
-\renewcommand{\le}{\leqslant}%less or equal
-\renewcommand{\ge}{\geqslant}%greater or equal
+\renewcommand*{\le}{\leqslant}%less or equal
+\renewcommand*{\ge}{\geqslant}%greater or equal
 \DeclarePairedDelimiter\clcl{[}{]}
@@
 %\DeclarePairedDelimiter\opop{]}{[}
-%\DeclarePairedDelimiter\abs{\lvert}{\rvert}
+\DeclarePairedDelimiter\abs{\lvert}{\rvert}
 %\DeclarePairedDelimiter\norm{\lVert}{\rVert}
@@
 \newcommand*{\pf}{\mathrm{p}}%probability
-\newcommand*{\p}{\mathrm{p}}%probability
-%\newcommand*{\tf}{\mathrm{T}}%probability
-%\renewcommand*{\|}{\|}
+\newcommand*{\p}{\mathrm{P}}%probability
+\newcommand*{\E}{\mathrm{E}}
+\renewcommand*{\|}{\nonscript\,\vert\nonscript\;\mathopen{}}
+%\DeclarePairedDelimiterX{\cond}[2]{(}{)}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
+%\DeclarePairedDelimiterX{\condt}[2]{[}{]}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
+%\DeclarePairedDelimiterX{\conds}[2]{\{}{\}}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
 %\newcommand*{\+}{\lor}
@@
 \newcommand*{\chaps}{chs}%
+\newcommand*{\bref}{ref.}%
+\newcommand*{\brefs}{refs}%
 %\newcommand*{\fn}{fn}%
@@
 %\newcommand*{\ca}{{c.}}
-%\newcommand*{\Ie}{{I.e.}}
-%\newcommand*{\Eg}{{E.g.}}
 %\newcommand*{\eg}{{e.g.}}
+\newcommand*{\foll}{{ff.}}
 %\newcommand*{\viz}{{viz}}
@@
 %\newcommand*{\vd}{{v.}}
-%\newcommand*{\Vd}{{V.}}
 \newcommand*{\etal}{{et al.}}
@@
 %\def\hy{-\penalty0\hskip0pt\relax}
-\newcommand*{\labelbis}[1]{\tag*{(\ref{#1})$_\text{r}$}}
+%\newcommand*{\labelbis}[1]{\tag*{(\ref{#1})$_\text{r}$}}
 %\newcommand*{\mathbox}[2][.8]{\parbox[t]{#1\columnwidth}{#2}}
 %\newcommand*{\zerob}[1]{\makebox[0pt][l]{#1}}
-%\newcommand*{\tprod}{\mathop{\textstyle\prod}\nolimits}
-%\newcommand*{\tsum}{\mathop{\textstyle\sum}\nolimits}
+\newcommand*{\tprod}{\mathop{\textstyle\prod}\nolimits}
+\newcommand*{\tsum}{\mathop{\textstyle\sum}\nolimits}
 %\newcommand*{\tint}{\begingroup\textstyle\int\endgroup\nolimits}
@@
 \newcommand*{\T}{^\intercal}%transpose
-\newcommand*{\E}{\mathrm{E}}
-\DeclarePairedDelimiter\expp{(}{)}
-\newcommand*{\expe}{\E\expp}%round
-\newcommand*{\expeb}{\E\clcl}%square
 %%\newcommand*{\QEM}%{\textnormal{$\Box$}}%{\ding{167}}
@@
 
-\newtheoremstyle{simple}%
-{}%.5\baselineskip plusminus.2\baselineskip % \parsep
-{}%
-{\footnotesize}%
-{}%
-{}%
-{}%
-{0pt}%
-{}%
-\theoremstyle{simple}
-\newtheorem*{simplenote}{}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Custom macros for this file @@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \definecolor{notecolour}{RGB}{68,170,153}
 \newcommand*{\puzzle}{{\fontencoding{U}\fontfamily{fontawesometwo}\selectfont\symbol{225}}}
+%\newcommand*{\puzzle}{\maltese}
 \newcommand{\mynote}[1]{ {\color{notecolour}\puzzle\ #1}}
 \newcommand*{\widebar}[1]{{\mkern1.5mu\skew{2}\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}}
+
+% \newcommand{\explanation}[4][t]{%\setlength{\tabcolsep}{-1ex}
+% %\smash{
+% \begin{tabular}[#1]{c}#2\\[0.5\jot]\rule{1pt}{#3}\\#4\end{tabular}}%}
+% \newcommand*{\ptext}[1]{\text{\small #1}}
+%\DeclareMathOperator*{\argsup}{arg\,sup}
+\newcommand*{\dob}{degree of belief}
+\newcommand*{\dobs}{degrees of belief}
 
@@
 \newcommand*{\me}{maximum-entropy}
-
-%@@@@@@@@@@ new macros end @@@@@@@@@@
-
+%%% Custom macros end @@@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Beginning of document
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\firmlists
+\begin{document}
+\captiondelim{\quad}\captionnamefont{\footnotesize}\captiontitlefont{\footnotesize}
 \selectlanguage{british}\frenchspacing
-
-\title{\pdftitle}
-
-% The \author macro works with any number of authors. There are two
-% commands used to separate the names and addresses of multiple
-% authors: \And and \AND.
+\maketitle
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Abstract
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\abstractrunin
+\abslabeldelim{}
+\renewcommand*{\abstractname}{}
+\setlength{\absleftindent}{0pt}
+\setlength{\absrightindent}{0pt}
+\setlength{\abstitleskip}{-\absparindent}
+\begin{abstract}\labelsep 0pt%
+  \noindent This work shows how to build a maximum-entropy probabilistic
+  model for the total activity of a population of neurons, given only some
+  activity statistics -- for example, empirical moments -- of a
+  \emph{subpopulation} thereof. This kind of model is useful because
+  neuronal recordings are always limited to a very small sample of a
+  population of neurons.
 %
-% Using \And between authors leaves it to LaTeX to determine where to
-% break the lines. Using \AND forces a line break at that point. So,
-% if LaTeX puts 3 of 4 authors names on the first line, and the last
-% on the second line, try using \AND instead of \And before the third
-% author name.
-
-\author{
-  P.G.L. Porta Mana\\\texttt{piero.mana@ntnu.no}
-  \And
-  V. Rostami\\\texttt{vrostami@uni-koeln.de}
-  \AND
-  E. Torre\\\texttt{torre@ibk.baug.ethz.ch}
-    \And
-Y. Roudi\\\texttt{yasser.roudi@ntnu.no}
-  % David S.~Hippocampus\thanks{Use footnote for providing further
-  %   information about author (webpage, alternative
-  %   address)---\emph{not} for acknowledging funding agencies.} \\
-  % Department of Computer Science\\
-  % Cranberry-Lemon University\\
-  % Pittsburgh, PA 15213 \\
-  % \texttt{hippo@cs.cranberry-lemon.edu} \\
-  %% examples of more authors
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \AND
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-}
-
-\begin{document}
-% \nipsfinalcopy is no longer used
-
-\maketitle
-
-\vspace{-2em}
-\begin{center}
-  \footnotesize\published
-\end{center}
-\medskip
-
-\begin{abstract}
-  The present work shows that the \me\ method can be applied to a sample of
-  neuronal recordings along two different routes: (1) apply to the sample;
-  or (2) apply to a larger, unsampled neuronal population from which the
-  sample is drawn, and then marginalize to the sample. These two routes
-  give inequivalent results. The second route can be further generalized to
-  the case where the size of the larger population is unknown. Which route
-  should be chosen? Some arguments are presented in favour of the second.
-  This work also presents and discusses probability formulae that relate
-  states of knowledge about a population and its samples, and that may be
-  useful for sampling problems in neuroscience.
-  % The abstract must be limited to one paragraph.
+  The model is applied to two sets of neuronal data available in the
+  literature. In some cases it predicts the larger population to have
+  interesting features -- for example, two modes in the probability for the
+  total activity at low regimes -- that are not visible in the sample or in
+  a maximum-entropy model built for the sample alone.
+%
+  For the two datasets, the maximum-entropy probability model applied only
+  to the subpopulation is compared with the marginal probability
+  distribution obtained from the maximum-entropy model applied to the full
+  population. On a linear probability scale no large differences are
+  visible, but on a logarithmic scale the two distributions show very
+  different behaviours, especially in the tails.
+  \\\noindent\emph{\footnotesize Note: Dear Reader \amp\ Peer, this
+    manuscript is being peer-reviewed by you. Thank you.}
+% \par%\\[\jot]
+% \noindent
+% {\footnotesize PACS: ***}\qquad%
+% {\footnotesize MSC: ***}%
+%\qquad{\footnotesize Keywords: ***}
 \end{abstract}
-
-%\section{Implicit assumptions in the maximum-entropy method}
-\section{Introduction: \me\ and recordings of neuronal activity}
+\selectlanguage{british}\frenchspacing
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Epigraph
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \asudedication{\small ***}
+% \vspace{\bigskipamount}
+% \setlength{\epigraphwidth}{.7\columnwidth}
+% %\epigraphposition{flushright}
+% \epigraphtextposition{flushright}
+% %\epigraphsourceposition{flushright}
+% \epigraphfontsize{\footnotesize}
+% \setlength{\epigraphrule}{0pt}
+% %\setlength{\beforeepigraphskip}{0pt}
+% %\setlength{\afterepigraphskip}{0pt}
+% \epigraph{\emph{text}}{source}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% BEGINNING OF MAIN TEXT
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Maximum-entropy and recordings of neuronal activity}
 \label{sec:intro}
@@
 sampled from a particular brain area. What are we to do with such data?
-Gerstein, Perkel, Dayhoff \cite{gersteinetal1985} posed this question very
+Gerstein \etal\ \citey{gersteinetal1985} posed this question very
 tersely (our emphasis):
-\begin{quote}
+\begin{quote}\small
   The principal conceptual problems are (1) \emph{defining cooperativity or
@@
 review by Moore \etal\
-\cite{mooreetal1966}. % This scenario is a concrete possibility
-% thanks to recent electrophysiological techniques \cite{berenyietal2014}.
+\citey{mooreetal1966}. % This scenario is a concrete possibility
+% thanks to recent electrophysiological techniques \citep{berenyietal2014}.
 The neuroscientific literature has offered several mathematical definitions
@@
 One such quantitative criterion relies on the \me\ or relative-\me\ method
-\cite{jaynes1957,jaynes1963,hobsonetal1973,sivia1996_r2006,meadetal1984}.
+\citep{jaynes1957,jaynes1963,hobsonetal1973,sivia1996_r2006,meadetal1984}.
 This criterion has been used in neuroscience at least since the 1990s,
@@
 cortex
-\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015},
+\citep{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015},
 and it has been subjected to mathematical and conceptual scrutiny
-\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.
+\citep{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.
 
@@
   sequence of digits \enquote{$0$}s (inactive) and \enquote{$1$}s (active)
-  \cite[\cf][]{caianiello1961,caianiello1986}.
+  \citep[\cf][]{caianiello1961,caianiello1986}.
 
@@
 dependence.
+\textcolor{white}{If you find this you can claim a postcard from me.}
 
@@
 the root of some quirks about this method that were pointed out in the
-literature \cite{roudietal2009b}. This problems extends also to more
+literature \citep{roudietal2009b}. This problems extends also to more
 complex versions of the method, possibly except versions that use
 \enquote{hidden} neurons
-\cite{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}. The problem is
+\citep{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}. The problem is
 that the recorded neurons are a \emph{sample} from a larger, unrecorded
@@
 Suppose that the neurons were recorded with electrodes covering an area of
-some square millimetres \cite[\cf][]{berenyietal2014}. This recording is a
+some square millimetres \citep[\cf][]{berenyietal2014}. This recording is a
 sample of the activity of the neuronal population under the recording
 device, which can amount to tens of thousands of neurons
-\cite{abeles1991}. We could even consider the recorded neurons as a sample
+\citep{abeles1991}. We could even consider the recorded neurons as a sample
 of a brain area more extended than the recording device.
@@
 degree of belief which \enquote{would be agreed by all rational men if
-  there were any rational men} \cite{good1966}.
+  there were any rational men} \citep{good1966}.
 
@@
 distribution is usually called \emph{finitely exchangeable}
-\cites{ericson1969}[\chap~1]{ghoshetal1997}. We stress that this
+\citep{ericson1969}[\chap~1]{ghoshetal1997}. We stress that this
 probability assignment is just an expression of the symmetry of our
@@
 non-binary and continuum cases are given by de~Finetti
-\cite{definetti1959b}, Kendall \cite{kendall1967}, Ericson
-\cite{ericson1976}, Diaconis \amp\ Freedman
-\cite{diaconis1977,diaconisetal1980}, Heath \amp\ Sudderth
-\cite{heathetal1976}.
+\citey{definetti1959b}, Kendall \citey{kendall1967}, Ericson
+\citey{ericson1976}, Diaconis \amp\ Freedman
+\citey{diaconis1977,diaconisetal1980}, Heath \amp\ Sudderth
+\citey{heathetal1976}.
 
@@
 for the full population, as follows:
-\begin{equation}
+\begin{multline}
   \label{eq:pullback_P}
-  \expe{g\|I}
+  \E(g\|I)
   \equiv
@@
   \sum_{\yrs} g(\yrs) \sum_{\yRf} \ypp(\yrs\|\yRf)\,\p(\yRf \| \yH)
-  = %{}\\
+  ={}\\
   \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
@@
 %  \equiv  \expeb{g^*}.
-\end{equation}
+\end{multline}
 where the second step uses \eqn~\eqref{eq:subpop_average}. The last
@@
   \label{eq:constraint_extended}
-  c = \expe{g \| \yH} \equiv \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
+  c = \E(g \| \yH) \equiv \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
   \p(\yRf \| \yH).
@@
 find
-\begin{equation}
+\begin{multline}
   \label{eq:expe_products}
@@
     % \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
-  \expe{\sav{\underbrace{\yr \dotsm \yr}_{\text{$m$ factors}}} \| \yH}
+  \E(\sav{\underbrace{\yr \dotsm \yr}_{\text{$m$ factors}}} \| \yH)
 \equiv
@@
     % \binom{N-m}{N \yRf-m} \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
-    = \binom{\yNv}{m}^{-1}
+    ={}\\
+    \binom{\yNv}{m}^{-1}
     % \frac{(N-m)!}{N!}
@@
     \binom{\yNv \yRf}{m}\, \p(\yRf \| \yH)
-\equiv    \expe{\av{\underbrace{\yR \dotsm \yR}_{\text{$m$ factors}}} \| \yH},
-\end{equation}
+\equiv    \E(\av{\underbrace{\yR \dotsm \yR}_{\text{$m$ factors}}} \| \yH),
+\end{multline}
 that is, \emph{the expected values of the $m$-activities of the sample and
@@
 distribution and can be found in \textcite{potts1953}. Similar relations
-can be found for the raw moments $\expe{\yrs^m}$ and $\expe{\yRf^m}$, which
+can be found for the raw moments $\E(\yrs^m)$ and $\E(\yRf^m)$, which
 can be written in terms of the product expectations using
@@
 \centering
-\includegraphics[width=0.9\linewidth]{pop_sample_projection3.pdf}%
+\includegraphics[width=\linewidth]{pop_sample_projection3.pdf}%
 \caption{Log-density plot of the hypergeometric distribution
@@
   \p(\yrs=a \|\yH) \approx \p(\yRf=a \|\yH),\qquad
-\expeb{g(\yrs) \|\yH} \approx \expeb{g(\yRf) \|\yH},
+\E[g(\yrs) \|\yH] \approx \E[g(\yRf) \|\yH],
 \end{equation}
@@
 
-Let us apply the relative-\me\ method \cite{sivia1996_r2006,meadetal1984}
+Let us apply the relative-\me\ method \citep{sivia1996_r2006,meadetal1984}
 directly to sampled neurons; denote this approach by $\yHb$. Then we apply
@@
 the \enquote{density of states} in front of the Boltzmann factor in
-statistical mechanics \cite[\chap~16]{callen1960_r1985}. The $m$ Lagrange
+statistical mechanics \citep[\chap~16]{callen1960_r1985}. The $m$ Lagrange
 multipliers $\yl\defd (l_1,\dotsc,l_m)$ must satisfy the $m$ constraint
@@
   \label{eq:app_maxent_sample_constraints}
-  c_k = \expe{g_k \|\yHb} \equiv
+  c_k = \E(g_k \|\yHb) \equiv
   \frac{1}{\yk(\yl)}\sum_{\yrs}  g_k(\yrs) \binom{n}{n\yrs}
@@
 satisfy the $m$ constraint equations
-\begin{equation}
+\begin{multline}
   \label{eq:app_maxent_pop_constraints}
-  c_k = \expe{g_k \|\yHa}\equiv
-  \frac{1}{\yK(\yL)}\sum_{\yrs} \sum_{\yRf} g_k(\yrs) \ypp(\yrs \|\yRf)
+  c_k = \E(g_k \|\yHa)\equiv{}\\
+\shoveright{  \frac{1}{\yK(\yL)}\sum_{\yrs} \sum_{\yRf} g_k(\yrs) \ypp(\yrs \|\yRf)
 \,
   \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
-  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
-  \qquad k=1,\dotsc,m.
-\end{equation}
+  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],}
+  \\ k=1,\dotsc,m.
+\end{multline}
 
@@
 \centering
-\includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
+\includegraphics[width=\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
 \caption{Linear and log-plots of $\p(\yrs)$ constructed by \me\ at
@@
 \centering
-\includegraphics[width=0.9\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}% 
+\includegraphics[width=\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}% 
 \caption{Linear and log-plots of $\p(\yrs)$ constructed by \me\ at
@@
 \centering
-\includegraphics[width=0.9\linewidth]{comparison3.pdf}%
+\includegraphics[width=\linewidth]{comparison3.pdf}%
 \caption{Linear and log-plots of $\p(\yrs)$ for a sample of $n=200$
@@
   \eqn~\eqref{eq:app_maxent_sample};
-  \textcolor{mybluishpurple}{\textbf{blue triangles:}} \me\ at the
+  \textcolor{mypurpleblue}{\textbf{blue triangles:}} \me\ at the
   population level, \eqn~\eqref{eq:app_maxent_pop_marg} with
@@
   \label{eq:constraints}
-  \expe{\yrs} = 0.0478,\qquad
-  \expe{\yrrs} = 0.00257,
+  \E(\yrs) = 0.0478,\qquad
+  \E(\yrrs) = 0.00257,
   % ,\quad
@@
 which come from the actual recording of circa 200 neurons from macaque
-motor cortex \cite{rostamietal2016_r2017}. The distribution obtained at the
+motor cortex \citep{rostamietal2016_r2017}. The distribution obtained at the
 population level (blue triangles) has a higher and displaced mode and a quite
@@
 distribution for $\yNv$ inspired by Jeffreys
-\cite[\sect~4.8]{jeffreys1939_r2003}:
+\citep[\sect~4.8]{jeffreys1939_r1983}:
 \begin{equation}
@@
 the sample and at the population levels appears in the use of Boltzmann
-machines with hidden units \cite{lerouxetal2008}, although by a different
+machines with hidden units \citep{lerouxetal2008}, although by a different
 conceptual route. It also appears in statistical mechanics: if a system is
 statistically described by a \me\ Gibbs state, its subsystems cannot be
-described by a Gibbs state \cite{maesetal1999}. A somewhat similar
+described by a Gibbs state \citep{maesetal1999}. A somewhat similar
 situation also appears in the statistical description of the final state of
@@
 (even though the final \emph{physical} state is obviously exactly the same
-\cite[\sect~4]{jaynes1985d_r1993}), and the second allows us to make
+\citep[\sect~4]{jaynes1985d_r1993}), and the second allows us to make
 sharper predictions about the final physical state thanks to our knowledge
@@
 \item How to extend the new application to the \enquote{inhomogeneous} case
-  \cite{schneidmanetal2006,shlensetal2006,roudietal2009b}, in which
+  \citep{schneidmanetal2006,shlensetal2006,roudietal2009b}, in which
   expectations for individual neurons or groups of neurons are constrained?
@@
   \me\ models with hidden neurons
-  \cite{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}?
+  \citep{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}?
 \end{itemize}
@@
 formulae presented here reflect a simple, symmetric state of ignorance.
-More work is needed \cite[\cf][]{levinaetal2017} to extend these formulae
+More work is needed \citep[\cf][]{levinaetal2017} to extend these formulae
 to account for finer knowledge of the cerebral cortex and its network
@@
 
-%\mynote{Importance of the formulae relating sample and population.}
-
-% \mynote{Discussion of case when individual $\yXv_i\dotsm \yXv_j$ are constrained.
-% Likely breakdown of \me\ in this case \cite{portamana2009}. Hints at full
-% Bayesian treatment of which \me\ is a limit? \cite[***]{mackay1995_r2003}}
-
-% do not change font sizes (except perhaps in the \textbf{References}
-% section; see below).
-
-\subsubsection*{Acknowledgements}
-
-PGLPM thanks Mari \amp\ Miri for continuous
-encouragement and affection; Buster Keaton for filling life with awe and
-inspiration; the developers and maintainers of \LaTeX, Emacs, AUC\TeX, Open
-Science Framework, Python, Inkscape, Sci-Hub for making a free and
+%%\setlength{\intextsep}{0.5ex}% with wrapfigure
+%\begin{figure}[p!]%{r}{0.4\linewidth} % with wrapfigure
+%  \centering\includegraphics[trim={12ex 0 18ex 0},clip,width=\linewidth]{maxent_saddle.png}\\
+%\caption{caption}\label{fig:comparison_a5}
+%\end{figure}% exp_family_maxent.nb
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Acknowledgements
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\iffalse
+\begin{acknowledgements}
+  PGLPM thanks Mari, Miri, \amp\ Emma for continuous encouragement and
+  affection; Buster Keaton and Saitama for filling life with awe and
+  inspiration; the developers and maintainers of \LaTeX, Emacs, AUC\TeX,
+  Open Science Framework, Python, Inkscape, Sci-Hub for making a free and
   unfiltered scientific exchange possible.
-
-% Use unnumbered third level headings for the acknowledgments. All
-% acknowledgments go at the end of the paper. Do not include
-% acknowledgments in the anonymized submission, only in the final paper.
-
-%\section*{References}
-\renewcommand*{\bibname}{References}
-\defbibheading{bibliography}[\bibname]{\section*{#1}\addcontentsline{toc}{section}{#1}%\markboth{#1}{#1}
-}
-
-% References follow the acknowledgments. Use unnumbered first-level
-% heading for the references. Any choice of citation style is acceptable
-% as long as you are consistent. It is permissible to reduce the font
-% size to \verb+small+ (9 point) when listing the references. {\bf
-%   Remember that you can go over 8 pages as long as the subsequent ones contain
-%   \emph{only} cited references.}
-% \medskip
-
-% \small
-
-% [1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
-% for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
-% T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
-%   Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.
-
-% [2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
-%   Exploring Realistic Neural Models with the GEneral NEural SImulation
-%   System.}  New York: TELOS/Springer--Verlag.
-
-% [3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
-% learning and recall at excitatory recurrent synapses and cholinergic
-% modulation in rat hippocampal region CA3. {\it Journal of
-%   Neuroscience} {\bf 15}(7):5249-5262.
-% \defbibnote{postnote}{\small\par\medskip\noindent{\footnotesize% Note:
-% \arxivp \mparcp \philscip \biorxivp}%
-% }
-
-\newcommand*{\citein}[2][]{\textnormal{\textcite[#1]{#2}}%\addtocategory{extras}{#2}
-}
-\newcommand*{\citebi}[2][]{ref.\ \citep[#1]{#2}%\addtocategory{extras}{#2}
-}
-\newcommand*{\subtitleproc}[1]{}
-\newcommand*{\chapb}{ch.}
-
-\printbibliography%[postnote=postnote]
+%\rotatebox{15}{P}\rotatebox{5}{I}\rotatebox{-10}{P}\rotatebox{10}{\reflectbox{P}}\rotatebox{-5}{O}.
+\sourceatright{\autanet}
+\end{acknowledgements}
+\fi
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Appendices
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\clearpage
+% %\renewcommand*{\appendixpagename}{Appendix}
+% %\renewcommand*{\appendixname}{Appendix}
+% %\appendixpage
+% \appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Bibliography
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\defbibnote{prenote}{{\footnotesize (\enquote{de $X$} is listed under D,
+    \enquote{van $X$} under V, and so on, regardless of national
+    conventions.)\par}}
+% \defbibnote{postnote}{\par\medskip\noindent{\footnotesize% Note:
+%     \arxivp \mparcp \philscip \biorxivp}}
+
+\printbibliography[prenote=prenote%,postnote=postnote
+]
 
 \end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Cut text (won't be compiled)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2019-04-08 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-08T11:32:41+0200
+%% Last-Updated: 2019-04-08T13:52:04+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\firstpublished}{***}
-\newcommand*{\propertitle}{Maximum-entropy distribution for a 
+\newcommand*{\propertitle}{Maximum-entropy distributions for a 
   neuronal population\\ from subpopulation constraints%\\{\large ***}%
@@
 
-\newcommand*{\ypp}{\varPi}
+\newcommand*{\ypp}{G}
 
@@
 \newcommand*{\yr}{\bm{\yrv}}%subpop value
-\newcommand*{\yrs}{\sav{\yr}}%subpop av value
+\newcommand*{\yrs}{\yrv}%subpop av value
 \newcommand*{\yR}{\bm{\yRv}}%pop value
-\newcommand*{\yRf}{\av{\yR}}%pop av value
+\newcommand*{\yRf}{\yRv}%pop av value
 %conditional assumptions
@@
 %Lagr multipliers
-\newcommand*{\yf}{\bm{g}}
-\newcommand*{\yc}{\bm{c}}
+\newcommand*{\yg}{\bm{c}}
+\newcommand*{\yc}{\widehat{\bm{c}}}
 \newcommand*{\yL}{\bm{\lambda}}
@@
   \label{eq:joint_plaus}
-  \p(\yRv_1, \yRv_2, \dotsc, \yRv_\yNv \| \yHc)
+  \pf(\yRv_1, \yRv_2, \dotsc, \yRv_\yNv \| \yHc)
   \quad\text{or}\quad
-\p(\yR \| \yHc), \quad \yR \in \set{0,1}^\yNv.
+\pf(\yR \| \yHc), \quad \yR \in \set{0,1}^\yNv.
 \end{equation}
@@
   \label{eq:sample_plaus}
-  \p(\yrv_1, \yrv_2, \dotsc, \yrv_n \| \yHc) \quad\text{or}\quad
-\p(\yr \| \yHc), \quad \yr \in \set{0,1}^n.
+  \pf(\yrv_1, \yrv_2, \dotsc, \yrv_n \| \yHc) \quad\text{or}\quad
+\pf(\yr \| \yHc), \quad \yr \in \set{0,1}^n.
 \end{equation}
@@
   \label{eq:joint_plaus_N_homog}
-  \p(\yR \|  \yH) \equiv
-  \sum_{\yRf}\p(\yR \| \yRf, \yH)\,
-  \p(\yRf \| \yH) =
-  \binom{\yNv}{\yNv\yRf}^{-1} \p(\yRf \| \yH).
+  \pf(\yR \|  \yH) \equiv
+  \sum_{\yRf}\pf(\yR \| \yRf, \yH)\,
+  \pf(\yRf \| \yH) =
+  \binom{\yNv}{\yNv\yRf}^{-1} \pf(\yRf \| \yH).
 \end{equation}
@@
   \label{eq:conditional_hypergeometric}
-  \p(\yrs \|\yRf, \yH)=
+  \pf(\yrs \|\yRf, \yH)=
   \binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}
-  \defs \ypp(\yrs\|\yRf),
+  \defs \ypp_{\yrs\yRf},
 \end{equation}
@@
   \label{eq:subpop_average}
-  \p(\yrs \| \yH) = \sum_{\yRf}
-  \p(\yrs \|\yRf, \yH)\,
-  \p(\yRf \| \yH)
+  \pf(\yrs \| \yH) = \sum_{\yRf}
+  \pf(\yrs \|\yRf, \yH)\,
+  \pf(\yRf \| \yH)
   = 
   \sum_{\yRf}
-  \ypp(\yrs\|\yRf)\,
-  \p(\yRf \| \yH).
+  \ypp_{\yrs\yRf}\,
+  \pf(\yRf \| \yH).
 \end{gather}
@@
   \label{eq:marginal}
-  \p(\yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yrs \| \yH).
+  \pf(\yr \| \yH) = \binom{n}{n\yrs}^{-1} \pf(\yrs \| \yH).
 \end{equation}
@@
 The conditional probability
-$\p(\yrs \|\yRf, \yH) \equiv \ypp(\yrs \|\yRf)$, besides
+$\pf(\yrs \|\yRf, \yH) \equiv \ypp_{\yrs \yRf}$, besides
 relating the distributions for the population and sample activities via
 marginalization, also allows us to express the expectation value of any
-function of the sample activity, $g(\yrs)$, in terms of the distribution
+function of the sample activity, $\yg_{\yrs}$, in terms of the distribution
 for the full population, as follows:
-\begin{multline}
+\begin{equation}
   \label{eq:pullback_P}
-  \E(g\|I)
+  \E(\yg\|I)
   \equiv
-  \sum_{\yrs} g(\yrs)\,\p(\yrs \| \yH)
+  \sum_{\yrs} \yg_{\yrs}\,\pf(\yrs \| \yH)
   =
-  \sum_{\yrs} g(\yrs) \sum_{\yRf} \ypp(\yrs\|\yRf)\,\p(\yRf \| \yH)
-  ={}\\
-  \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
-  \p(\yRf \| \yH),
+  \sum_{\yrs} \yg_{\yrs} \sum_{\yRf} \ypp_{\yrs\yRf}\,\pf(\yRf \| \yH)
+  =%{}\\
+  \sum_{\yRf} \bigl( \tsum_{\yrs} \yg_{\yrs}  \ypp_{\yrs\yRf} \bigr)\,
+  \pf(\yRf \| \yH),
 %  =
-%  \sum_{N\yRf=0}^N \sum_{n\yrs=0}^n f(\yrs)\ypp(\yrs\|\yRf)\,\p(\yXf=\yRf \| \yH)
-%  \equiv  \expeb{g^*}.
-\end{multline}
+%  \sum_{N\yRf=0}^N \sum_{n\yrs=0}^n f(\yrs)\ypp_{\yrs\yRf}\,\p(\yXf=\yRf \| \yH)
+%  \equiv  \expeb{\yg^*}.
+\end{equation}
 where the second step uses \eqn~\eqref{eq:subpop_average}. The last
-expression shows that the expectation of the function $g(\yrs)$ is equal to
+expression shows that the expectation of the function $\yg_{\yrs}$ is equal to
 the expectation of the function
-$g^*(\yRf) \defd \sum_{\yrs} g(\yrs)\,\ypp(\yrs\|\yRf)$.
+$\yg^*(\yRf) \defd \sum_{\yrs} \yg_{\yrs}\,\ypp_{\yrs\yRf}$.
 
@@
 The final expression in \eqn~\eqref{eq:pullback_P} is important for our
-\me\ application: the requirement that the function $g$, defined for the
-sample, have a  value $c$ obtained from observed data,
+\me\ application: the requirement that the function $\yg$, defined for the
+sample, have a  value $\yc$ obtained from observed data,
 \emph{translates into a linear constraint for the distribution of the full
@@
   \label{eq:constraint_extended}
-  c = \E(g \| \yH) \equiv \sum_{\yRf} \biggl[ \sum_{\yrs} g(\yrs)  \ypp(\yrs\|\yRf) \biggr]\,
-  \p(\yRf \| \yH).
+  \yc = \E(\yg \| \yH) \equiv \sum_{\yRf} \bigl( \tsum_{\yrs} \yg_{\yrs}  \ypp_{\yrs\yRf} \bigr)\;
+  \pf(\yRf \| \yH).
 \end{equation}
 
-In particular, when the function $g$ is the $m$-activity of the sample,
-$g(\yrs) = \sav{\yr\dotso\yr} \equiv \binom{n\yrs}{m}/\binom{n}{m}$, we
+In particular, when the function $\yg$ is the $m$-activity of the sample,
+$\yg_{\yrs} = \sav{\yr\dotso\yr} \equiv \binom{n\yrs}{m}/\binom{n}{m}$, we
 find
@@
     \binom{n}{m}^{-1}
-    \binom{n \yrs}{m}\, \p(\yrs \| \yH)
+    \binom{n \yrs}{m}\, \pf(\yrs \| \yH)
     % \sum_{N \yRf=0}^N %m!\,
@@
     \sum_{\yRf} %m!\,
-    \binom{\yNv \yRf}{m}\, \p(\yRf \| \yH)
+    \binom{\yNv \yRf}{m}\, \pf(\yRf \| \yH)
 \equiv    \E(\av{\underbrace{\yR \dotsm \yR}_{\text{$m$ factors}}} \| \yH),
@@
 \caption{Log-density plot of the hypergeometric distribution
-  $%\p(\yrs \|\yRf, \yH)=
-\ypp(\yrs\|\yRf) \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
+  $%\pf(\yrs \|\yRf, \yH)=
+\ypp_{\yrs\yRf} \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
   colourbar depending on your \textsc{pdf} viewer.)}
@@
 formalize and mathematically express our intuition. The hypergeometric
-distribution $\ypp(\yrs\|\yRf)$ plays an important role in this
+distribution $\ypp_{\yrs\yRf}$ plays an important role in this
 formalization. A look at its plot, \fig~\ref{fig:hypergeom_proj}, reveals
@@
   \label{eq:roughly_equal_nN}
-  \p(\yrs=a \|\yH) \approx \p(\yRf=a \|\yH),\qquad
-\E[g(\yrs) \|\yH] \approx \E[g(\yRf) \|\yH],
+  \pf(\yrs=a \|\yH) \approx \pf(\yRf=a \|\yH),\qquad
+\E[\yg_{\yrs} \|\yH] \approx \E[\yg_{\yRf} \|\yH],
 \end{equation}
-where $g$ is any smooth function defined on $\clcl{0,1}$. These approximate
+where $\yg$ is any smooth function defined on $\clcl{0,1}$. These approximate
 equalities express the intuitive fact that \emph{our uncertainty about the
   sample is representative of our uncertainty about the population and
-  about other samples}, and vice versa. When $n=\yNv$, $\ypp(\yrs\|\yRf)$
+  about other samples}, and vice versa. When $n=\yNv$, $\ypp_{\yrs\yRf}$
 becomes the identity matrix and the approximate equalities above become
@@
 the sample average $\yrs$, owing to the coarsening effect of
-$\ypp(\yrs\|\yRf)$.
+$\ypp_{\yrs\yRf}$.
 
 % They should be contrasted with the limits
-% $\p(\yxs=a) \to \p(\yXf=a)$ and $\expeb{g(\yxs)} \to \expeb{g(\yXf)}$, as
+% $\pf(\yxs=a) \to \pf(\yXf=a)$ and $\expeb{\yg_{\yxs}} \to \expeb{\yg_{\yXf}}$, as
 % $n\to \yNv$, do: these limits are trivially, universally valid because the
@@
 % hold even in cases where the conditional probability
-% $\p(\yxs = \yrs \|\yXf=\yRf)$ is not a fuzzy identity, our uncertainties
+% $\pf(\yxs = \yrs \|\yXf=\yRf)$ is not a fuzzy identity, our uncertainties
 % about sample and about population can differ wildly, and the approximate
@@
 we want to constrain $m$ functions of the sample activity, vectorially
-written $\yf \defd (g_1,\dotsc,g_m)$, to $m$ values
-$\yc \defd (c_1,\dotsc,c_m)$. These functions are typically $k$-activities
+written $\yg \defd (c_1,\dotsc,c_m)$, to $m$ values
+$\yc \defd (\widehat{c}_1,\dotsc,\widehat{c}_m)$. These functions are typically $k$-activities
 $\sav{\yr\dotso \yr}$, and the values are typically the time averages of
 the observed sample, as discussed in \sect~\ref{sec:intro}:
-$\yc = \sum_t \yf[\yrs(t)]/T$.
+$\yc = \sum_t \yg[\yrs(t)]/T$.
 
@@
   \label{eq:app_maxent_sample}
-  \p(\yrs \|\yHb)
+  \pf(\yrs \|\yHb)
   =\frac{1}{\yk(\yl)}\,
   \binom{n}{n\yrs}
-  \exp[\yl\T \yf(\yrs)]
+  \exp[\yl\T \yg_{\yrs}]
 \end{equation}
@@
   \label{eq:app_maxent_sample_constraints}
-  c_k = \E(g_k \|\yHb) \equiv
-  \frac{1}{\yk(\yl)}\sum_{\yrs}  g_k(\yrs) \binom{n}{n\yrs}
-  \exp[\yl\T \yf(\yrs)], \qquad k=1,\dotsc,m.
+  \yc = \E(\yg \|\yHb) \equiv
+  \frac{1}{\yk(\yl)}\sum_{\yrs}  \yg_{\yrs} \binom{n}{n\yrs}
+  \exp[\yl\T \yg_{\yrs}].%, \qquad k=1,\dotsc,m.
 \end{equation}
@@
   \label{eq:app_maxent_pop}
-  \p(\yRf \| \yHa)  = \frac{1}{\yK(\yL)}\,
-  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
-  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
+  \pf(\yRf \| \yHa)  = \frac{1}{\yK(\yL)}\,
+  \binom{\yNv}{\yNv \yRf}\,\exp\bigl(\yL\T
+  \tsum_{\yrs} \yg_{\yrs}\ypp_{\yrs\yRf}\bigr).
 \end{equation}
@@
 satisfy the $m$ constraint equations
-\begin{multline}
+\begin{equation}
   \label{eq:app_maxent_pop_constraints}
-  c_k = \E(g_k \|\yHa)\equiv{}\\
-\shoveright{  \frac{1}{\yK(\yL)}\sum_{\yrs} \sum_{\yRf} g_k(\yrs) \ypp(\yrs \|\yRf)
+  \yc = \E(\yg \|\yHa)\equiv %{}\\
+ \frac{1}{\yK(\yL)}\sum_{\yrs} \sum_{\yRf} \yg_{\yrs} \ypp_{\yrs \yRf}
 \,
-  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
-  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],}
-  \\ k=1,\dotsc,m.
-\end{multline}
+  \binom{\yNv}{\yNv \yRf}\,\exp\bigl(\yL\T
+  \tsum_{\yrs} \yg_{\yrs}\ypp_{\yrs\yRf}\bigr).
+%  \\ k=1,\dotsc,m.
+\end{equation}
 
@@
   \label{eq:app_maxent_pop_marg}
-  \p(\yrs \| \yHa)  = \frac{1}{\yK(\yL)}\, 
-  \sum_{\yRf} \ypp(\yrs \|\yRf)
+  \pf(\yrs \| \yHa)  = \frac{1}{\yK(\yL)}\, 
+  \sum_{\yRf} \ypp_{\yrs \yRf}
 \,
-  \binom{\yNv}{\yNv \yRf}\,\exp\Bigl[\yL\T
-  \sum_{\yrs} \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
+  \binom{\yNv}{\yNv \yRf}\,\exp\bigl(\yL\T
+  \tsum_{\yrs} \yg_{\yrs}\ypp_{\yrs\yRf}\bigr).
 \end{equation}
@@
 \eqref{eq:app_maxent_sample_constraints}, and in addition also the $n$
-equations $\p(\yrs \| \yHa) = \p(\yrs \| \yHb)$,
+equations $\pf(\yrs \| \yHa) = \pf(\yrs \| \yHb)$,
 $\yrs=1/n,\dotsc,1$ (one equation is taken care of by the normalization of
@@
 \includegraphics[width=\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
-\caption{Linear and log-plots of $\p(\yrs)$ constructed by \me\ at
+\caption{Linear and log-plots of $\pf(\yrs)$ constructed by \me\ at
   the population level followed by sample marginalization (blue triangles),
@@
 \includegraphics[width=\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}% 
-\caption{Linear and log-plots of $\p(\yrs)$ constructed by \me\ at
+\caption{Linear and log-plots of $\pf(\yrs)$ constructed by \me\ at
   the population level followed by sample marginalization (blue triangles),
@@
 \includegraphics[width=\linewidth]{comparison3.pdf}%
-\caption{Linear and log-plots of $\p(\yrs)$ for a sample of $n=200$
+\caption{Linear and log-plots of $\pf(\yrs)$ for a sample of $n=200$
   and constraints as in \eqn~\eqref{eq:constraints}, constructed by:
@@
 state of knowledge $\yHd$ our ignorance about the possible value of $\yNv$
-is expressed by a probability distribution $\p(\yNN=\yNv \|\yHd)=h(\yNv)$,
+is expressed by a probability distribution $\pf(\yNN=\yNv \|\yHd)=h(\yNv)$,
 and the marginal distribution for the sample
@@
   \label{eq:app_maxent_pop_marg_unknown_N}
-  \p(\yrs \| \yHd)  =
-  \sum_{\yNv} \p(\yrs \| \yNv,\yHd) \,
-  \p(\yNv \|\yHd)
+  \pf(\yrs \| \yHd)  =
+  \sum_{\yNv} \pf(\yrs \| \yNv,\yHd) \,
+  \pf(\yNv \|\yHd)
   ={}\\[-\jot]
   \sum_\yNv \biggl\{\frac{1}{\yK(\yL_\yNv)}\, 
-  \sum_{\yRf} \ypp_\yNv(\yrs \|\yRf)
+  \sum_{\yRf} \ypp^{(\yNv)}_{\yrs\yRf}
 \,
   \tbinom{\yNv}{\yNv \yRf}\,\exp\Bigl[{\yL_\yNv}\T
-  \sum_{\yrs} \yf(\yrs)\ypp_\yNv(\yrs\|\yRf)\Bigr]\biggr\}
-  \,h(\yNv),
+  \tsum_{\yrs} \yg_{\yrs}\ypp^{(\yNv)}_{\yrs\yRf}\Bigr]\biggr\}
+  \;h(\yNv),
 \end{multline}
@@
 \end{equation}
+
+\section{Derivation from the probability calculus}
+\label{sec:derivation_prob_calculus}
 

### 2019-04-08 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-08T13:52:04+0200
+%% Last-Updated: 2019-04-08T16:12:02+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+The maximum-entropy method is used to construct a complete plausibility
+distribution given only a quantitative \emph{feeling} of some of its
+properties, such as expectations. This way of constructing probabilities
+may be needed at the very beginning of an inference
+problem\footnote{\enquote{Now the axioms of probability enable us to infer
+    any probability-conclusion \emph{only} from probability-premisses. In
+    other words, the calculus of probability does not enable us to infer
+    any probability-value unless we have some probabilities or probability
+    relations \emph{given}. Such data cannot be supplied by the
+    mathematician. E.g. the rules of arithmetic and the axioms of the
+    probability-calculus are utterly impotent to determine, on the supposed
+    knowledge that the throw of a coin must yield either head or tail and
+    cannot yield both, the probability that it will yield head or that it
+    will yield tail. We must assume that the two co-exclusive and
+    co-exhaustive possibilities are \emph{equally probable}, before we can
+    estimate the probability of either as being a half of certitude}
+  \citep[\emph{Appendix on eduction}, \sect~5, p.~182]{johnson1924}.}
+\mynote{refs to Jaynes and others}.
+
+In the present case, however, the probability distribution can be derived
+through the probability calculus from the precise empirical statistics and
+observations made; the initial probabilities and assumptions needed in such
+an inference are of a different kind. It can be shown that the
+maximum-entropy distribution is an \emph{approximation} of the exact result
+obtained this way.
+
 \section{Discussion}

### 2019-04-09 maxNt_v5.tex
update

--- 2019-04-08 maxNt_v5.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-08T13:52:04+0200
+%% Last-Updated: 2019-04-09T17:16:14+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\renewcommand*{\cite}{\parencite}
-\renewcommand*{\cites}{\parencites}
+%\newcommand*{\citeps}{\parencites}
 \providecommand{\href}[2]{#2}
@@
 
+\newcommand*{\ynu}{\bm{\nu}}
+
 \newcommand*{\prop}[1]{`#1'}
@@
 necessary to use probability relations relevant to sampling
-\cites{ghoshetal1997}[parts~I,
+\citep{ghoshetal1997}[parts~I,
 VI]{freedmanetal1978_r2007}[\chap~8]{gelmanetal1995_r2014}[\chap~3]{jaynes1994_r2003}.
@@
 fact the same as for this class of problems
-\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
+\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
 
@@
 
+There are three inequivalent main routes that lead to a probability
+distribution of the maximum-entropy form~\eqref{eq:app_maxent_sample}
+or~\eqref{eq:app_maxent_pop}. The distribution carries a different
+interpretation under each route \citep[pp.~52--55,
+72--77]{jaynes1979b}[pp.~25--28]{jaynes1982}[\sect~I]{jaynes1982b}{jaynes1986d_r1996}[\sect~11.1]{jaynes1994_r2003}.
+
+\begin{enumerate}[wide,label=(\alph*)]
+\item \label{item:entropy_route}One route is the choice of the distribution having the highest
+  Shannon entropy, given only a quantitative assessment some of its
+  properties, such as expectations. The numerical choice of the value of
+  such properties is a (subjective) assumption.
+
+  \medskip
+
+  In the two other routes the maximum-entropy distribution is obtained as
+  an \emph{approximation} of a distribution obtained via the probability
+  calculus, using data coming from a set of $T$ measurements -- as in our
+  present case.\mynote{refs here} Also in this case some (subjective)
+  assumptions are necessary: they concern our beliefs about the long-run
+  relative frequencies of the measurement outcomes:\footnote{Such
+    assumptions are always necessary at the beginning of an inference:
+    \enquote{Now the axioms of probability enable us to infer any
+      probability-conclusion \emph{only} from probability-premisses. In
+      other words, the calculus of probability does not enable us to infer
+      any probability-value unless we have some probabilities or
+      probability relations \emph{given}. Such data cannot be supplied by
+      the mathematician. E.g. the rules of arithmetic and the axioms of the
+      probability-calculus are utterly impotent to determine, on the
+      supposed knowledge that the throw of a coin must yield either head or
+      tail and cannot yield both, the probability that it will yield head
+      or that it will yield tail. We must assume that the two co-exclusive
+      and co-exhaustive possibilities are \emph{equally probable}, before
+      we can estimate the probability of either as being a half of
+      certitude} \citep[\emph{Appendix on eduction}, \sect~5,
+    p.~182]{johnson1924}.}
+
+\item \label{item:entr_prior_route}In one case we consider all possible
+  sets of measurement outcomes to be \emph{roughly} equally likely; this
+  leads to a probability for the frequencies $\ynu$ proportional to a
+  multinomial coefficient $\binom{L}{L\ynu}$, with $L$ large but smaller
+  than $T$. (We cannot assume the sets of measurement outcomes to be
+  exactly equally likely, because this is equivalent to their independence:
+  \cf\
+  \cites[\sect~6.7]{jaynes1994_r2003}[\sect~B]{portamana2009}[\sect~2]{portamana2017}.)
+\item \label{item:suff_route}In the other case we assume that the
+  measurements have a \emph{sufficient statistics}: the same as appears in
+  the exponential of the maximum-entropy distribution.
+\end{enumerate}
+
+It's important to keep in mind that the approximate equivalence of these
+three routes only holds under very specific assumptions -- which \emph{have
+  physical and biological meanings and consequences}. In particular,
+route~\ref{item:suff_route} implies that we can discard other empirical
+statistics of the data, if they are known; whereas
+route~\ref{item:entr_prior_route} requires us to specify all known
+empirical statistics, because using only a subset of them may lead to
+different results. Route~\ref{item:entropy_route} is also supposed to be
+used with all known data. Moreover, the approximate equivalence of
+route~\ref{item:entropy_route} with routes~\ref{item:entr_prior_route} and
+\ref{item:suff_route} \emph{only holds if $T$ is much larger than the
+  possible values of the activity $\yRf$}.
+
 \section{Discussion}

### 2019-04-10 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-09T17:16:14+0200
+%% Last-Updated: 2019-04-11T00:12:54+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\hrule
+Yasser's intro
+\hrule
+
+Experimental technologies to record the activity of many neurons at the
+same time in different species and brain areas are rapidly advancing. These
+experimental advancements are paralleled by advances in theoretical and
+computational methods for analyzing the data accumulated using the
+recording technologies. Such theoretical methods usually take the form of
+probabilistic models that try to describe the multi-neuronal activity of
+the recorded neurons. With such probabilistic models one aims to address
+numerous issues: What correlations are important in describing the
+multi-neuronal pattern? How does the pattern of activity covary with
+external stimuli or experimental conditions? What dimensionality does the
+neural data live in and how is this related to the underlying network
+interactions? The probabilistic models can also be used to make predictions
+about the structure of the neural code, by studying the properties of the
+fitted model, or by generating synthetic data from it.
+
+In general, despite the rapid advances in recording technology, the best
+experimental measurements of neuronal activity still only provide data from
+a small subset of neurons that comprise a neuronal network. A wealth of
+studies on building probabilistic models of neural data focuses on
+describing such subsets, ignoring the fact that the observed neurons is a
+small group in a much bigger set of hidden neurons. Some other studies do
+include hidden variables which, amongst other things, aim to model the
+global features of the unrecorded neurons, but we still lack a through
+understanding of how to include the role of hidden neurons in probabilistic
+models, how our inferences about the recorded activity would be affected by
+them, and what we can we say about the rest of the network by studying the
+heavily subsampled recordings. In this paper, we aim to address these
+questions in the case of a simple maximum entropy model, namely the
+homogeneous maximum entropy model.
+
+The maximum entropy approach has been used in a variety of setting for
+building statistical models of complex systems and datasets, ranging from
+neuronal activity in the retina, in the cortex, protein sequences, gene
+regulatory networks and natural images. The general idea is, for a dataset,
+to write down the distribution that maximizes the entropy of state
+variables, given some low order statistics. Now given the fact that the
+recorded neurons are a fraction of the neurons in the network, several
+quantitative questions arise that we will address in this paper: given the
+data from the sampled neurons, can we build a maximum entropy model over
+the whole network? Once we build such a network level maximum entropy
+model, can we see features in the neural activity which cannot be directly
+seen from a model build from the sampled neurons? Since we can marginalize
+down the network level maximum entropy model to the sampled network, how
+does this margianzlied maximum entropy model match the sample level model?
+
+All these questions can be answered in the case of the homogeneous maximum
+entropy model. First we show how to go from the sample level maximum
+entropy level to the network level, by assuming different sizes of the
+network and also by assuming an uninformative prior over the size of the
+network. This is done by inferring the statistics of correlation functions
+at the network level from those of the sample level by using simple
+counting arguments. We then find that, when applied to experimental
+recordings from the Medial Entorhinal Cortex of rats and the monkey visual
+cortex, this network level maximum entropy model may exhibit features that
+the sample level model does not predict. Specifically, we observed modes in
+the distribution of the activity in the network level model that do not
+show up in the sample level. We study how the assumed size of the full
+network affects the appearance of these modes and find that there is a
+minimum size of the full network for which such modes can be observed. We
+then compared the distribution found by marginalizing the full network
+maximum entropy model down to the sample level, and the distribution fit
+directly to the sample level. For the two datasets that we tested, we found
+that the two distributions match each other to a large degree but that
+there are also differences between them. We quantify how these differences
+also depend on the assumed size of the network and find that …(WE SHOULD
+TEST SHI) This predicts that for a large enough population (DO WE PREDICT
+THAT IF THE FULL NETWORK GETS BIGGER THE DIFFERENCE ALSO GET BIGGER)?…
+
+The rest of the paper is organized as follows. We first describe how to go
+from the sample level maximum entropy model to the full network maximum
+entropy model. In section 2, we apply this to the two experimental datasets
+and study the effect of the assumed size of the network as well as the
+moments that we use for building the maximum entropy model. In section 3 we
+compare the distributions found from marginalizing the maximum entropy
+model down to the sample level and the original sample level model.
+
+\hrule
+\bigskip
+
 Suppose we have recorded the firing activity of a hundred neurons,
@@
   \cites[\sect~6.7]{jaynes1994_r2003}[\sect~B]{portamana2009}[\sect~2]{portamana2017}.)
+  The exact expression is\mynote{equation here}
+  %   \begin{multline}
+  %   \label{eq:prob_conditional_average_exch}
+  %   \p\bigl[\yE{N+1}{k} \bigcond \tsum\yo\yf{N}\in\yA, \yX\bigr] ={}\\
+  %   \frac{\int\yqq_k\sum_{\yf{N}}  \delt(\tsum \yo\yf{N}\in \yA)\,
+  %     \binom{N}{N\yf{N}}\,\bigl( \tprod \yq^{N\yf{N}} \bigr)  \,
+  %     \pf(\yq \|\yX)  \,\di\yq}{ \int\sum_{\yf{N}} \delt(\tsum \yo\yf{N}\in \yA)\,
+  %     \binom{N}{N\yf{N}}\,\bigl( \tprod \yq^{N\yf{N}} \bigr) \,
+  %     \pf(\yq \|\yX) \,\di\yq },
+  % \end{multline}
+
 \item \label{item:suff_route}In the other case we assume that the
   measurements have a \emph{sufficient statistics}: the same as appears in
-  the exponential of the maximum-entropy distribution.
+  the exponential of the maximum-entropy distribution. The exact expression is\mynote{equation here}
+  % \begin{multline}
+  %   \label{eq:prob_conditional_average_suff}
+  %   \p\bigl[\yE{N+1}{k} \bigcond \tsum\yo\yf{N}\in\yA, \yS\bigr] ={}
+  %       \\[3\jot]
+  %    \frac{\int \pf(k \| \yl,\yr,\yS) \sum_{\yf{N}} \delt(\tsum \yo\yf{N}\in \yA)\,
+  %     \binom{N}{N\yf{N}}\,\bigl[ \tprod  \pf(\yk \| \yl,\yr,\yS)^{N\yf{N}} \bigr]
+  %      \,
+  %     \pf(\yl \|\yS) \,\di\yl}{\int\sum_{\yf{N}} \delt(\tsum \yo\yf{N}\in \yA)\,
+  %     \binom{N}{N\yf{N}}\,\bigl[ \tprod \pf(\yk \| \yl,\yr,\yS)^{N\yf{N}} \bigr]
+  %      \,
+  %     \pf(\yl \|\yS) \,\di\yl}.
+  % \end{multline}
 \end{enumerate}
@@
 \ref{item:suff_route} \emph{only holds if $T$ is much larger than the
-  possible values of the activity $\yRf$}.
+  possible values of the activity $\yRf$}. Finally, we also obtain very
+different expressions depending on whether we're asking about \emph{the
+  activity in one of the \textbf{recorded} time bins} or about \emph{the
+  activity in a \textbf{new} time bin}. Works that use maximum-entropy
+distributions are often very vague about the latter point.
 

### 2019-04-18 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-11T00:12:54+0200
+%% Last-Updated: 2019-04-18T16:17:13+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%\usepackage{wrapfig}
+\usepackage{wrapfig}
 
@@
 
+A necessary step of every plausible inference is the assignment of some
+probabilistic premisses -- as many as necessary to arrive at our desired
+probabilistic conclusions given the evidence. The necessity of this step
+was tersely stated by W.~E.~Johnson %(J.~M.~Keynes's supervisor)
+a hundred years ago:
+\begin{quote}\footnotesize
+  Now the axioms of probability enable us to infer any
+  probability-conclusion \emph{only} from probability-premisses. In other
+  words, the calculus of probability does not enable us to infer any
+  probability-value unless we have some probabilities or probability
+  relations \emph{given}. Such data cannot be supplied by the
+  mathematician. E.g. the rules of arithmetic and the axioms of the
+  probability-calculus are utterly impotent to determine, on the supposed
+  knowledge that the throw of a coin must yield either head or tail and
+  cannot yield both, the probability that it will yield head or that it
+  will yield tail. We must assume that the two co-exclusive and
+  co-exhaustive possibilities are \emph{equally probable}, before we can
+  estimate the probability of either as being a half of certitude.
+  \sourceatright{\citep[Appendix on eduction, \sect~5,
+    p.~182]{johnson1924}}
+\end{quote}
+
+These premisses may represent either actual knowledge or just some
+hypotheses whose consequences we want to explore. They can be
+\setlength{\intextsep}{0.5ex}% with wrapfigure
+\begin{wrapfigure}{l}{0.33\linewidth}% with wrapfigure
+\centering\includegraphics[width=\linewidth]{statements_scheme1.pdf}%\\
+%\caption{caption}\label{fig:comparison_a5}
+\end{wrapfigure}% statements_scheme.svg statements_scheme2.svg
+assigned at different depths of the logical analysis of our inference
+problem, and assignments at different depths can turn out to be
+equivalent.
+
+We can for example assign probabilistic premisses at a very deep
+level, as schematically illustrated on the left. They will determine the
+probabilities of statements at all higher levels, including the statements
+we're ultimately interested in.
+
+Or we can assign probabilistic premisses at
+a level of intermediate
+\begin{wrapfigure}{r}{0.33\linewidth}% with wrapfigure
+\centering\includegraphics[width=\linewidth]{statements_scheme2.pdf}%\\
+%\caption{caption}\label{fig:comparison_a5}
+\end{wrapfigure}% statements_scheme.svg statements_scheme2.svg
+depth, as schematically illustrated on the right. In this case they will
+determine the probabilities of statements at all higher level, and partly
+constrain the probabilities of statements at deeper levels.
+
+\clearpage
+
 \hrule
@@
 \hrule
-
+{\footnotesize
 Experimental technologies to record the activity of many neurons at the
@@
 model down to the sample level and the original sample level model.
+
+}
+\mynote{Luca: add this: from sampling theory we know that important
+  features of the full network may not be visible in a sample because
+  smoothed out. But using sampling theory in the inverse direction we can
+  infer such full-network features from the sample.}
 

### 2019-04-18 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-18T16:17:13+0200
+%% Last-Updated: 2019-04-18T21:40:56+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-These premisses may represent either actual knowledge or just some
-hypotheses whose consequences we want to explore. They can be
+These premisses may represent either actual knowledge or just
 \setlength{\intextsep}{0.5ex}% with wrapfigure
@@
 \end{wrapfigure}% statements_scheme.svg statements_scheme2.svg
-assigned at different depths of the logical analysis of our inference
-problem, and assignments at different depths can turn out to be
-equivalent.
+some hypotheses whose consequences we want to explore. They can be assigned
+at different depths of the logical analysis of our inference problem.
 
 We can for example assign probabilistic premisses at a very deep
-level, as schematically illustrated on the left. They will determine the
+level, as schematically illustrated on the left. They determine the
 probabilities of statements at all higher levels, including the statements
@@
 
-Or we can assign probabilistic premisses at
-a level of intermediate
 \begin{wrapfigure}{r}{0.33\linewidth}% with wrapfigure
-\centering\includegraphics[width=\linewidth]{statements_scheme2.pdf}%\\
+\centering\includegraphics[width=\linewidth]{statements_scheme2b.pdf}%\\
 %\caption{caption}\label{fig:comparison_a5}
 \end{wrapfigure}% statements_scheme.svg statements_scheme2.svg
-depth, as schematically illustrated on the right. In this case they will
-determine the probabilities of statements at all higher level, and partly
-constrain the probabilities of statements at deeper levels.
+Or we can assign probabilistic premisses at a level of intermediate depth,
+as schematically illustrated on the right. In this case they determine the
+probabilities of statements at all higher levels, and partly constrain
+the probabilities of statements at deeper levels.
+
+The choice of the logical level at which to assign the premisses depends on
+many factors: on the difficulty of assessing our state of knowledge at
+particular levels, on the cost of the ensuing calculations, and especially
+on the general purpose of our inference. But two assignments of premisses at
+different levels can still turn out to be completely equivalent for the
+probabilities of the statements of interest.
+
+The maximum-entropy principle was proposed first by Jaynes
+\parentext{\cite*{jaynes1957,jaynes1963,jaynes1994_r2003};
+  \cite{sivia1996_r2006,hobsonetal1973}} as a way of assigning probability
+premisses at a level of intermediate depth -- this point is made
+especially clear in \textcite[\sect~11.1]{jaynes1994_r2003}.
 

### 2019-04-19 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-18T21:40:56+0200
+%% Last-Updated: 2019-04-19T15:23:02+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \postauthor{\par\end{center}}
-\predate{\DTMsetdatestyle{mydate}\begin{center}\footnotesize}
+\predate{\DTMsetdatestyle{mydate}\vspace{0.5\baselineskip}%
+\begin{center}\footnotesize}
 \postdate{\end{center}\vspace{-\medskipamount}}
@@
 \displaywidowpenalty=1602
+\raggedbottom
 \selectlanguage{british}\frenchspacing
@@
 \firmlists
+\raggedbottom
 \begin{document}
@@
 was tersely stated by W.~E.~Johnson %(J.~M.~Keynes's supervisor)
-a hundred years ago:
+a hundred years ago:\vspace{\topsep}
 \begin{quote}\footnotesize
@@
 
-The maximum-entropy principle was proposed first by Jaynes
+Here's an example. Suppose that $n$ neurons have been recorded in a
+sequence of time bins. We're given the time averages of the numbers of
+active neurons and neuron pairs in the sequence. We don't know the exact
+sequence of activities or the number $T$ of time bins, although we know
+that the latter is large. We want to forecast how many neurons will be
+active in a \emph{new} time bin.
+
+In a deep approach to this inference problem we assign the probability for
+the possible value of $T$, and the probabilities for several hypotheses
+about the sequence of activities in $T+1$ or more time bins. One hypothesis
+could be, for example, that the $n$ neurons are a sample of a larger
+population of $N$ neurons and for this population there are some sufficient
+statistics, out of biological reasons. We could of course deepen this
+approach further, with no end, down to assigning the probabilities for the
+constitution and motion of the molecules of the $N$ neurons and their
+environment, the probabilities for the position and response of the
+recording electrodes, and so on. From all these probabilities and the data
+we could finally calculate the probability for the activity in the new time
+bin.
+
+At the other end, the shallowest approach to this inference problem is to
+directly assign the probabilities for the number of active neurons in the
+new time bin, somehow using the data we were given. (It must be remarked
+that such assignment partly constrains, in a backward fashion, all assignments we
+would do at a deeper level.)
+
+The maximum-entropy principle was proposed by Jaynes
 \parentext{\cite*{jaynes1957,jaynes1963,jaynes1994_r2003};
   \cite{sivia1996_r2006,hobsonetal1973}} as a way of assigning probability
-premisses at a level of intermediate depth -- this point is made
-especially clear in \textcite[\sect~11.1]{jaynes1994_r2003}.
+premisses at a more direct level, as in the approach just described. This
+point is made especially clear in \textcite{jaynes1994_r2003},
+\sect~11.1:\vspace{0.5\topsep}
+\begin{quote}\footnotesize
+  If we knew the numbers $N$, $L$, $W$, then this could be solved by direct
+  application of Bayes' theorem; without that information, we could still
+  introduce the unknowns $N$, $L$, $W$ as nuisance parameters and use
+  Bayes' theorem, eliminating them at the end. \textelp{} However, the
+  Bayesian solution would not really address our problem; it only transfers
+  it to the problem of assigning priors to $N$, $L$, $W$, leaving us back
+  in essentially the same situation; how do we assign informative
+  probabilities?
+\end{quote}
+Its idea is to translate \emph{all} data we're given as constraints for our
+probability assignment, resolving the leftover freedom by choosing the
+probability distribution having highest Shannon entropy, as a measure of
+its uncertainty or broadness. It can be shown that the use of this
+principle is approximately equivalent, under specific conditions, to a
+probability assignment at a deeper level based on an assumption of
+exchangeability together with a multinomial probability distribution for
+long-run frequencies, or with an assumption about sufficient statistics
+\citep{jaynes1986d_r1996,rodriguez1989,rodriguez2002,skilling1989b,skilling1990}[\sect~4.5]{bernardoetal1994_r2000}{diaconisetal1981,portamana2009,portamana2017}.
+
+Owing to its directness, the maximum-entropy principle is straightforward
+to use -- and to misuse. In our previous example, for instance, it gives
+silly results unless the number of time bins is much larger than the
+possible activity values (it has thus been variously misused in the
+neuroscientific literature \mynote{refs, especially to Tkacik}).
+
+In the present work we show how to use the maximum-entropy principle to...
 

### 2019-04-19 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-19T15:23:02+0200
+%% Last-Updated: 2019-04-19T19:09:13+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{quote}
-Its idea is to translate \emph{all} data we're given as constraints for our
-probability assignment, resolving the leftover freedom by choosing the
+The idea behind this principle is to translate \emph{all} the data we're
+given into constraints -- such as expectations -- for our probability
+assignment, resolving the leftover freedom by choosing the constrained
 probability distribution having highest Shannon entropy, as a measure of
@@
 long-run frequencies, or with an assumption about sufficient statistics
-\citep{jaynes1986d_r1996,rodriguez1989,rodriguez2002,skilling1989b,skilling1990}[\sect~4.5]{bernardoetal1994_r2000}{diaconisetal1981,portamana2009,portamana2017}.
+\citep{jaynes1986d_r1996,rodriguez1989,rodriguez2002,skilling1989b,skilling1990}[\sect~4.5]{bernardoetal1994_r2000}{diaconisetal1981,portamana2017,portamana2009}.
+
+Sometimes this principle is applied with only \emph{part} of the available
+data used as constraint; such an altered approach can be a heuristic way of
+testing for sufficient statistics.
 
 Owing to its directness, the maximum-entropy principle is straightforward
-to use -- and to misuse. In our previous example, for instance, it gives
-silly results unless the number of time bins is much larger than the
+to use -- and to misuse. In our previous example, for instance, it would
+give silly results if the number of time bins were not much larger than the
 possible activity values (it has thus been variously misused in the
@@
 
-In the present work we show how to use the maximum-entropy principle to...
+\bigskip
+
+In the present work we show how to use the maximum-entropy principle to
+assign a probability distribution for the activity of a population of
+neurons at a new time bin, given specific statistics about the activity of
+a \emph{subpopulation}, that is, a sample, of the full population. This
+kind of probability assignment is extremely important in neuroscience: we
+try to understand the workings of brain areas that may include tens of
+thousands of neurons, by analysing recording of only a few tens or hundreds
+of them.
 

### 2019-04-19 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-19T19:09:13+0200
+%% Last-Updated: 2019-04-19T19:24:02+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The maximum-entropy principle was proposed by Jaynes
+The maximum-entropy method was proposed by Jaynes
 \parentext{\cite*{jaynes1957,jaynes1963,jaynes1994_r2003};
@@
 \end{quote}
-The idea behind this principle is to translate \emph{all} the data we're
+The idea behind this method is to translate \emph{all} the data we're
 given into constraints -- such as expectations -- for our probability
@@
 its uncertainty or broadness. It can be shown that the use of this
-principle is approximately equivalent, under specific conditions, to a
+method is approximately equivalent, under specific conditions, to a
 probability assignment at a deeper level based on an assumption of
@@
 
-Sometimes this principle is applied with only \emph{part} of the available
+Sometimes this method is applied with only \emph{part} of the available
 data used as constraint; such an altered approach can be a heuristic way of
@@
 
-Owing to its directness, the maximum-entropy principle is straightforward
+Owing to its directness, the maximum-entropy method is straightforward
 to use -- and to misuse. In our previous example, for instance, it would
@@
 
-In the present work we show how to use the maximum-entropy principle to
+In the present work we show how to use the maximum-entropy method to
 assign a probability distribution for the activity of a population of
@@
 try to understand the workings of brain areas that may include tens of
-thousands of neurons, by analysing recording of only a few tens or hundreds
+thousands of neurons, but we can record only few tens or hundreds
 of them.
+
+Our approach also corrects a slight misuse of the maximum-entropy method as
+it has appeared in the literature. When this method is applied to a set of
+neurons, it implicitly assumes that the set is completely isolated from
+other neurons. We'll reveal this implicit assumption in several ways. The
+probability distribution thus constructed is therefore affected by this
+hidden assumption. We'll study in which circumstances the application of
+the method at the sample level gives results similar to the application at
+the full-population level.
 

### 2019-04-20 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-19T19:24:02+0200
+%% Last-Updated: 2019-04-20T15:37:10+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\DeclarePairedDelimiterX{\cond}[2]{(}{)}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
-%\DeclarePairedDelimiterX{\condt}[2]{[}{]}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
+\DeclarePairedDelimiterX{\condt}[2]{[}{]}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
 %\DeclarePairedDelimiterX{\conds}[2]{\{}{\}}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
@@
 
-\section{Maximum-entropy and recordings of neuronal activity}
-\label{sec:intro}
+\section{Why polls?}
+\label{sec:polls}
+
+When political elections are approaching, mass media are constantly
+bubbling with results and discussions of pre-election polls. Why do we
+conduct pre-election polls and other kinds of survey sampling? The
+simplest, intuitive answer is \enquote{because they give us an idea of what
+  the final election outcomes will be}. Can this answer be made more
+quantitative?
+
+Yes. In the simplest cases, the probability calculus give us simple and
+precise quantitative relations between our expectations from a population
+-- of voters, for instance -- and from a sample thereof. Suppose there are
+$N$ voters, $S$ of which prefer one political candidate and $N-S$ of which
+prefer another. We don't know $S$ or $N-S$. Now take a sample of $n$
+voters, assuming that every set of $n$ voters is equally likely to be
+sampled. In this sample, $s$ voters prefer the first candidate, and $n-s$
+the second. The probability calculus tells us that our expectations of
+$s/n$ and $S/N$ are equal:
+\begin{equation}
+  \label{eq:equal_expectations_poll}
+  \E(s/n \| \yH) = \E(S/N \| \yH),
+\end{equation}
+where $\yH$ represents our state of knowledge in this example.
+
+A more precise answer to our original question is then: \enquote{because we
+  \emph{expect} to see the same proportion of political preferences in a
+  sample as in the full population}.
+
+The equality above extends to other quantities beyond the means $S/N$ and
+$s/n$: it holds for all normalized \emph{factorial} moments \citep{potts1953,broca2005}:
+\begin{multline}
+  \label{eq:equal_factorialmoments_polls}
+  \E\condt[\Big]{\tfrac{s\;(s-1)}{n\;(n-1)}}{\yH} = 
+  \E\condt[\Big]{\tfrac{S\;(S-1)}{N\;(N-1)}}{\yH},\\
+  \E\condt[\Big]{\tfrac{s\;(s-1)\;(s-2)}{n\;(n-1)\;(n-2)}}{\yH} = 
+  \E\condt[\Big]{\tfrac{S\;(S-1)\;(S-2)}{N\;(N-1)\;(N-2)}}{\yH},\quad\dotsc
+  \\
+  \E\condt[\Big]{\tbinom{s}{r}\;\tbinom{n}{r}^{-1}}{\yH} =
+  \E\condt[\Big]{\tbinom{S}{r}\;\tbinom{N}{r}^{-1}}{\yH},
+\end{multline}
+with $r<N$.
+
+\section{Maximum-entropy method}
+\label{sec:maxent_remarks}
 
 A necessary step of every plausible inference is the assignment of some
-probabilistic premisses -- as many as necessary to arrive at our desired
+probabilistic premisses; as many as necessary to arrive at our desired
 probabilistic conclusions given the evidence. The necessity of this step
 was tersely stated by W.~E.~Johnson %(J.~M.~Keynes's supervisor)
-a hundred years ago:\vspace{\topsep}
+a century ago:\vspace{\topsep}
 \begin{quote}\footnotesize
@@
 
-We can for example assign probabilistic premisses at a very deep
+We can for example assign probabilistic premisses at a very deep analytical
 level, as schematically illustrated on the left. They determine the
-probabilities of statements at all higher levels, including the statements
-we're ultimately interested in.
+probabilities of statements at more superficial levels, including the
+statements which we're ultimately interested in or which we can directly
+observe.
 
@@
 \end{wrapfigure}% statements_scheme.svg statements_scheme2.svg
-Or we can assign probabilistic premisses at a level of intermediate depth,
-as schematically illustrated on the right. In this case they determine the
-probabilities of statements at all higher levels, and partly constrain
-the probabilities of statements at deeper levels.
-
-The choice of the logical level at which to assign the premisses depends on
-many factors: on the difficulty of assessing our state of knowledge at
-particular levels, on the cost of the ensuing calculations, and especially
-on the general purpose of our inference. But two assignments of premisses at
+Or we can assign probabilistic premisses at a level of superficial or
+intermediate depth, as schematically illustrated on the right. In this case
+they determine the probabilities of statements at more superficial levels, and
+partly constrain the probabilities of statements at deeper levels.
+
+The choice of the analytical depth at which to assign the premisses depends
+on many factors: on the difficulty of assessing our state of uncertainty at
+that depth, on the cost of the ensuing calculations, and especially on the
+general purpose of our inference. Assignments of probabilistic premisses at
 different levels can still turn out to be completely equivalent for the
@@
 sequence of time bins. We're given the time averages of the numbers of
-active neurons and neuron pairs in the sequence. We don't know the exact
+active neurons and of neuron pairs in the sequence. We don't know the exact
 sequence of activities or the number $T$ of time bins, although we know
-that the latter is large. We want to forecast how many neurons will be
-active in a \emph{new} time bin.
+that the latter is large. We want to forecast how many of the $n$ neurons
+will be active in a \emph{new} time bin.
 
@@
 about the sequence of activities in $T+1$ or more time bins. One hypothesis
-could be, for example, that the $n$ neurons are a sample of a larger
-population of $N$ neurons and for this population there are some sufficient
-statistics, out of biological reasons. We could of course deepen this
-approach further, with no end, down to assigning the probabilities for the
-constitution and motion of the molecules of the $N$ neurons and their
-environment, the probabilities for the position and response of the
-recording electrodes, and so on. From all these probabilities and the data
-we could finally calculate the probability for the activity in the new time
-bin.
+could be, for instance, that the $n$ neurons are a sample of a larger
+population of $N$ neurons, and a second hypothesis could be that for this
+population there are some sufficient statistics, out of biological reasons.
+We could of course deepen this approach further, with no end, down to
+assigning the probabilities for the constitution and motion of the
+molecules of the neurons and their environment, the probabilities for the
+position and response of the recording electrodes, and so on. From all
+these probabilities and the data we could finally calculate the probability
+for the activity in the new time bin.
 
@@
 new time bin, somehow using the data we were given. (It must be remarked
-that such assignment partly constrains, in a backward fashion, all assignments we
-would do at a deeper level.)
+that such assignment partly constrains, in a backward fashion, all
+assignments we could do at a deeper level.)
 
@@
 \begin{quote}\footnotesize
-  If we knew the numbers $N$, $L$, $W$, then this could be solved by direct
-  application of Bayes' theorem; without that information, we could still
-  introduce the unknowns $N$, $L$, $W$ as nuisance parameters and use
-  Bayes' theorem, eliminating them at the end. \textelp{} However, the
-  Bayesian solution would not really address our problem; it only transfers
-  it to the problem of assigning priors to $N$, $L$, $W$, leaving us back
-  in essentially the same situation; how do we assign informative
-  probabilities?
+  If we knew the numbers $N$, $L$, $W$, \textins{some quantities relevant
+    to our inference} then this could be solved by direct application of
+  Bayes' theorem; without that information, we could still introduce the
+  unknowns $N$, $L$, $W$ as nuisance parameters and use Bayes' theorem,
+  eliminating them at the end. \textelp{} However, the Bayesian solution
+  would not really address our problem; it only transfers it to the problem
+  of assigning priors to $N$, $L$, $W$, leaving us back in essentially the
+  same situation; how do we assign informative probabilities?
 \end{quote}
@@
 
-In the present work we show how to use the maximum-entropy method to
-assign a probability distribution for the activity of a population of
-neurons at a new time bin, given specific statistics about the activity of
-a \emph{subpopulation}, that is, a sample, of the full population. This
-kind of probability assignment is extremely important in neuroscience: we
-try to understand the workings of brain areas that may include tens of
-thousands of neurons, but we can record only few tens or hundreds
-of them.
-
-Our approach also corrects a slight misuse of the maximum-entropy method as
-it has appeared in the literature. When this method is applied to a set of
-neurons, it implicitly assumes that the set is completely isolated from
-other neurons. We'll reveal this implicit assumption in several ways. The
-probability distribution thus constructed is therefore affected by this
+In the present work we show how to use the maximum-entropy method to assign
+a probability distribution for the total activity at a new time bin of a
+population of neurons, \emph{most of which can be unrecorded}, given
+specific statistics about the activity of a recorded \emph{sample} of the
+full population. This kind of probability assignment is extremely important
+in neuroscience: we try to understand the workings of brain areas that may
+include tens of thousands of neurons, but we can record only few tens or
+hundreds of them.
+
+Our approach also corrects a slight misuse of the maximum-entropy method
+as sometimes appears in the literature. When this method is applied to a
+set of neurons, it implicitly assumes that the set is completely isolated
+from other neurons. We'll reveal this implicit assumption in several ways.
+The probability distribution thus constructed is therefore affected by this
 hidden assumption. We'll study in which circumstances the application of
-the method at the sample level gives results similar to the application at
-the full-population level.
+the method at the sample level and at the full-population level give
+similar or different probabilities.
+
+The maximum-entropy method can be used this way thanks to a very simple ***
 

### 2019-04-20 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-20T15:37:10+0200
+%% Last-Updated: 2019-04-20T15:41:30+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The equality above extends to other quantities beyond the means $S/N$ and
-$s/n$: it holds for all normalized \emph{factorial} moments \citep{potts1953,broca2005}:
+$s/n$: it holds for all normalized \emph{factorial} moments
+\citep{potts1953,broca2005}, no matter what probability distribution we
+have assigned to $S$:
 \begin{multline}
@@
 \end{multline}
-with $r<N$.
+with $r<N$. Note that from the first $r$ factorial moments we can calculate
+the first $r$ raw power moments and vice versa; but the simple equalities
+above don't hold for the power moments.
 

### 2019-04-20 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-20T15:41:30+0200
+%% Last-Updated: 2019-04-20T17:16:05+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\usepackage{empheq}% automatically calls amsmath and mathtools
-\newcommand*{\widefbox}[1]{\fbox{\hspace{1em}#1\hspace{1em}}}
+%\newcommand*{\widefbox}[1]{\fbox{\hspace{1em}#1\hspace{1em}}}
 
@@
 
-A more precise answer to our original question is then: \enquote{because we
+A more precise answer to our original question then is: \enquote{because we
   \emph{expect} to see the same proportion of political preferences in a
@@
 $s/n$: it holds for all normalized \emph{factorial} moments
-\citep{potts1953,broca2005}, no matter what probability distribution we
-have assigned to $S$:
-\begin{multline}
+\citep{potts1953,broca2005}, no matter what probability distribution
+$\pf(S\|\yH)$ we have assigned to $S$:
+\begin{equation}
   \label{eq:equal_factorialmoments_polls}
-  \E\condt[\Big]{\tfrac{s\;(s-1)}{n\;(n-1)}}{\yH} = 
+  \begin{split}
+  \E\condt[\Big]{\tfrac{s\;(s-1)}{n\;(n-1)}}{\yH} &= 
   \E\condt[\Big]{\tfrac{S\;(S-1)}{N\;(N-1)}}{\yH},\\
-  \E\condt[\Big]{\tfrac{s\;(s-1)\;(s-2)}{n\;(n-1)\;(n-2)}}{\yH} = 
+  \E\condt[\Big]{\tfrac{s\;(s-1)\;(s-2)}{n\;(n-1)\;(n-2)}}{\yH} &= 
   \E\condt[\Big]{\tfrac{S\;(S-1)\;(S-2)}{N\;(N-1)\;(N-2)}}{\yH},\quad\dotsc
   \\
-  \E\condt[\Big]{\tbinom{s}{r}\;\tbinom{n}{r}^{-1}}{\yH} =
+  \text{and in general}\quad
+  \E\condt[\Big]{\tbinom{s}{r}\;\tbinom{n}{r}^{-1}}{\yH} &=
   \E\condt[\Big]{\tbinom{S}{r}\;\tbinom{N}{r}^{-1}}{\yH},
-\end{multline}
-with $r<N$. Note that from the first $r$ factorial moments we can calculate
-the first $r$ raw power moments and vice versa; but the simple equalities
-above don't hold for the power moments.
+  \quad r<N.
+\end{split}
+\end{equation}
+Note that from the first $r$ factorial moments we can calculate the first
+$r$ raw power moments and vice versa; but the simple equalities above don't
+hold for the power moments \citep[\sect~A]{portamanaetal2015}.
 

### 2019-04-21 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-20T17:16:05+0200
+%% Last-Updated: 2019-04-20T23:54:37+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Yes. In the simplest cases, the probability calculus give us simple and
-precise quantitative relations between our expectations from a population
--- of voters, for instance -- and from a sample thereof. Suppose there are
+precise quantitative relations between our expectations about a population
+-- of voters, for instance -- and about a sample thereof. Suppose there are
 $N$ voters, $S$ of which prefer one political candidate and $N-S$ of which
 prefer another. We don't know $S$ or $N-S$. Now take a sample of $n$
-voters, assuming that every set of $n$ voters is equally likely to be
-sampled. In this sample, $s$ voters prefer the first candidate, and $n-s$
-the second. The probability calculus tells us that our expectations of
-$s/n$ and $S/N$ are equal:
+voters, in such a way that we can assume that every set of $n$ voters is
+equally likely to be sampled. In this sample, $s$ voters prefer the first
+candidate, and $n-s$ the second. The probability calculus tells us that our
+expectations of $s/n$ and $S/N$ are equal:
 \begin{equation}
@@
 $s/n$: it holds for all normalized \emph{factorial} moments
-\citep{potts1953,broca2005}, no matter what probability distribution
-$\pf(S\|\yH)$ we have assigned to $S$:
+\citep{potts1953,broca2005}:
 \begin{equation}
@@
 hold for the power moments \citep[\sect~A]{portamanaetal2015}.
+
+The relations above are powerful because they hold no matter what the
+probability distribution $\pf(S\|\yH)$ be.
 

### 2019-04-21 scripts/ME_4momsolution.m

@@ -2,7 +2,7 @@
 
 Clear[MEhom4moments,MEhom4thmoment,MEhomvmoments,MEhom4momentsprob]; 
 
-MEhom4moments[m1_, m2_, m3_, m4_, nn_, wp_:MachinePrecision,pg_:3] := 
+MEhom4moments[m1_, m2_, m3_, m4_, nn_, wp_:MachinePrecision,pg_:6,mi_:10000] := 
  Block[{totalprob,
 	zzzm, js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
 	nn2=Binomial[nn,2], nn3=Binomial[nn,3], nn4=Binomial[nn,4], 
@@ -13,8 +13,8 @@ MEhom4moments[m1_, m2_, m3_, m4_, nn_, wp_:MachinePrecision,pg_:3] :=
 			     js4*(Binomial[m,4]/nn4-m4)],
        {m, 0, nn}];
   soluzm = 
-   NMinimize[freeenm, {js1, js2, js3, js4}, MaxIterations -> 10000, 
-    PrecisionGoal -> 4, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+   NMinimize[freeenm, {js1, js2, js3, js4}, MaxIterations -> mi, 
+    PrecisionGoal -> pg, AccuracyGoal -> Infinity,WorkingPrecision->wp];
   prob = Table[
 	  Binomial[nn, m]*Exp[js1*m/nn + js2*Binomial[m,2]/nn2 +
 		      js3*Binomial[m,3]/nn3 +

### 2019-04-21 scripts/ME_homsolution_prior2.m

@@ -1,0 +1,96 @@
+(* Homog maxent *)
+
+Clear[MEhomsolution,MEhomsolutiontruncated,MEhomprob]; 
+
+MEhomsolution::usage = "Input: pop-avgd activity [0,1], pop-avgd couple activity, pop size, machine precision. Output: Lagr multipl, Lagr multipl, distribution, Z"
+(* it uses normalized values in the optimization; usually gives more precise results than MEhomsolution2, although it reports difficulties in finding the optimum *)
+MEhomsolution[meanmm_, meancorr_, nn_, wp_:MachinePrecision] := 
+ Block[{totcorr = meancorr,
+   totmm = meanmm,totalprob,
+   zzzm, js2, js1, freeenm, varm, soluzm, prob,ran=Range[0,nn],bvalues},
+       zzzm = Sum[(1-m+nn)/((nn+1)*(nn+2))*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn], {m, 0, nn}]/2^nn;
+  freeenm = Log[zzzm] - js2*totcorr - js1*totmm;
+  soluzm = 
+   NMinimize[freeenm, {js1, js2}, MaxIterations -> 10000, 
+    PrecisionGoal -> 6,Method->Automatic, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+  prob = Table[
+    (1-m+nn)/((nn+1)*(nn+2))*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn]/2^nn /. soluzm[[2]], {m, 
+     0, nn}];
+  totalprob = Total[prob];prob=prob/totalprob;
+  bvalues={(ran.prob/nn)/meanmm-1,((ran^2-ran).prob/(nn^2-nn))/meancorr-1};
+ Print["rel. discrepancies %: ", bvalues*100//N];
+  {js1 /. soluzm[[2]], js2/. soluzm[[2]], prob,totalprob}]
+
+MEhomsolutiont[meanmm_, meancorr_, nn_, wp_:MachinePrecision] := 
+ Block[{totcorr = meancorr,
+   totmm = meanmm,totalprob,
+   js2, js1, freeenm, varm, soluzm, prob,ran=Range[0,nn],bvalues},
[…]

### 2019-04-23 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-20T23:54:37+0200
+%% Last-Updated: 2019-04-23T09:55:00+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\usepackage{wrapfig}
+%\usepackage{wrapfig}
 
@@
 
-\section{Why polls?}
+\section{Polling neurons}
 \label{sec:polls}
@@
 probability distribution $\pf(S\|\yH)$ be.
-
-\section{Maximum-entropy method}
-\label{sec:maxent_remarks}
-
-A necessary step of every plausible inference is the assignment of some
-probabilistic premisses; as many as necessary to arrive at our desired
-probabilistic conclusions given the evidence. The necessity of this step
-was tersely stated by W.~E.~Johnson %(J.~M.~Keynes's supervisor)
-a century ago:\vspace{\topsep}
-\begin{quote}\footnotesize
-  Now the axioms of probability enable us to infer any
-  probability-conclusion \emph{only} from probability-premisses. In other
-  words, the calculus of probability does not enable us to infer any
-  probability-value unless we have some probabilities or probability
-  relations \emph{given}. Such data cannot be supplied by the
-  mathematician. E.g. the rules of arithmetic and the axioms of the
-  probability-calculus are utterly impotent to determine, on the supposed
-  knowledge that the throw of a coin must yield either head or tail and
-  cannot yield both, the probability that it will yield head or that it
-  will yield tail. We must assume that the two co-exclusive and
-  co-exhaustive possibilities are \emph{equally probable}, before we can
-  estimate the probability of either as being a half of certitude.
-  \sourceatright{\citep[Appendix on eduction, \sect~5,
-    p.~182]{johnson1924}}
-\end{quote}
-
-These premisses may represent either actual knowledge or just
-\setlength{\intextsep}{0.5ex}% with wrapfigure
-\begin{wrapfigure}{l}{0.33\linewidth}% with wrapfigure
-\centering\includegraphics[width=\linewidth]{statements_scheme1.pdf}%\\
-%\caption{caption}\label{fig:comparison_a5}
-\end{wrapfigure}% statements_scheme.svg statements_scheme2.svg
-some hypotheses whose consequences we want to explore. They can be assigned
-at different depths of the logical analysis of our inference problem.
-
-We can for example assign probabilistic premisses at a very deep analytical
-level, as schematically illustrated on the left. They determine the
-probabilities of statements at more superficial levels, including the
-statements which we're ultimately interested in or which we can directly
-observe.
-
-\begin{wrapfigure}{r}{0.33\linewidth}% with wrapfigure
-\centering\includegraphics[width=\linewidth]{statements_scheme2b.pdf}%\\
-%\caption{caption}\label{fig:comparison_a5}
-\end{wrapfigure}% statements_scheme.svg statements_scheme2.svg
-Or we can assign probabilistic premisses at a level of superficial or
-intermediate depth, as schematically illustrated on the right. In this case
-they determine the probabilities of statements at more superficial levels, and
-partly constrain the probabilities of statements at deeper levels.
-
-The choice of the analytical depth at which to assign the premisses depends
-on many factors: on the difficulty of assessing our state of uncertainty at
-that depth, on the cost of the ensuing calculations, and especially on the
-general purpose of our inference. Assignments of probabilistic premisses at
-different levels can still turn out to be completely equivalent for the
-probabilities of the statements of interest.
-
-Here's an example. Suppose that $n$ neurons have been recorded in a
-sequence of time bins. We're given the time averages of the numbers of
-active neurons and of neuron pairs in the sequence. We don't know the exact
-sequence of activities or the number $T$ of time bins, although we know
-that the latter is large. We want to forecast how many of the $n$ neurons
-will be active in a \emph{new} time bin.
-
-In a deep approach to this inference problem we assign the probability for
-the possible value of $T$, and the probabilities for several hypotheses
-about the sequence of activities in $T+1$ or more time bins. One hypothesis
-could be, for instance, that the $n$ neurons are a sample of a larger
-population of $N$ neurons, and a second hypothesis could be that for this
-population there are some sufficient statistics, out of biological reasons.
-We could of course deepen this approach further, with no end, down to
-assigning the probabilities for the constitution and motion of the
-molecules of the neurons and their environment, the probabilities for the
-position and response of the recording electrodes, and so on. From all
-these probabilities and the data we could finally calculate the probability
-for the activity in the new time bin.
-
-At the other end, the shallowest approach to this inference problem is to
-directly assign the probabilities for the number of active neurons in the
-new time bin, somehow using the data we were given. (It must be remarked
-that such assignment partly constrains, in a backward fashion, all
-assignments we could do at a deeper level.)
-
-The maximum-entropy method was proposed by Jaynes
-\parentext{\cite*{jaynes1957,jaynes1963,jaynes1994_r2003};
-  \cite{sivia1996_r2006,hobsonetal1973}} as a way of assigning probability
-premisses at a more direct level, as in the approach just described. This
-point is made especially clear in \textcite{jaynes1994_r2003},
-\sect~11.1:\vspace{0.5\topsep}
-\begin{quote}\footnotesize
-  If we knew the numbers $N$, $L$, $W$, \textins{some quantities relevant
-    to our inference} then this could be solved by direct application of
-  Bayes' theorem; without that information, we could still introduce the
-  unknowns $N$, $L$, $W$ as nuisance parameters and use Bayes' theorem,
-  eliminating them at the end. \textelp{} However, the Bayesian solution
-  would not really address our problem; it only transfers it to the problem
-  of assigning priors to $N$, $L$, $W$, leaving us back in essentially the
-  same situation; how do we assign informative probabilities?
-\end{quote}
-The idea behind this method is to translate \emph{all} the data we're
-given into constraints -- such as expectations -- for our probability
-assignment, resolving the leftover freedom by choosing the constrained
-probability distribution having highest Shannon entropy, as a measure of
-its uncertainty or broadness. It can be shown that the use of this
-method is approximately equivalent, under specific conditions, to a
-probability assignment at a deeper level based on an assumption of
-exchangeability together with a multinomial probability distribution for
-long-run frequencies, or with an assumption about sufficient statistics
-\citep{jaynes1986d_r1996,rodriguez1989,rodriguez2002,skilling1989b,skilling1990}[\sect~4.5]{bernardoetal1994_r2000}{diaconisetal1981,portamana2017,portamana2009}.
-
-Sometimes this method is applied with only \emph{part} of the available
-data used as constraint; such an altered approach can be a heuristic way of
-testing for sufficient statistics.
-
-Owing to its directness, the maximum-entropy method is straightforward
-to use -- and to misuse. In our previous example, for instance, it would
-give silly results if the number of time bins were not much larger than the
-possible activity values (it has thus been variously misused in the
-neuroscientific literature \mynote{refs, especially to Tkacik}).
 

### 2019-04-23 scripts/ME_1fmoms.m

@@ -1,0 +1,46 @@
+(* Homog maxent with 4 moments - corresponds to constraining variances *)
+
+Clear[ME1fmoms,ME1fmoms0]; 
+
+ME1fmoms[eemoms_, nn_, wp_:MachinePrecision,pg_:4] := 
+  Block[{ll=Length[eemoms],js=Table[Unique["js"],Length[eemoms]], freeenm, soluzm, prob, bvalues,fmoms,emoms},
+	emoms=1-eemoms;
+	fmoms[s_]=Table[1-Binomial[s,r]/Binomial[nn,r], {r,ll}];
+	freeenm = Log@Sum[(1-ss+nn)/((nn+1)*(nn+2))*
+			 Exp[js.(fmoms[ss]-emoms)],{ss, 0, nn}];
+
+  soluzm = 
+   NMinimize[freeenm, js, MaxIterations -> 10000, 
+    PrecisionGoal -> pg, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+
+  prob = Table[
+    (1-ss+nn)/((nn+1)*(nn+2))*
+    Exp[(js /. soluzm[[2]]).fmoms[ss]] ,{ss, 0, nn}];
+  totalprob = Total[prob];prob=prob/totalprob;
+  
+  bvalues=(Total[Table[fmoms[ss],{ss,0,nn}]*prob]/emoms-1);
+  Print["rel. discrepancies %: ", bvalues*100//N];
+  
+ {js /. soluzm[[2]],prob,totalprob}]
+
+
+ME1fmoms0[eemoms_, nn_, wp_:MachinePrecision,pg_:4] := 
+  Block[{ll=Length[eemoms],js=Table[Unique["js"],Length[eemoms]], freeenm, soluzm, prob, bvalues,fmoms,emoms},
+	emoms=1-eemoms;
+	fmoms[s_]=Table[1-Binomial[s,r]/Binomial[nn,r], {r,ll}];
+	freeenm = Log@Sum[Binomial[nn,ss]*
+			 Exp[js.(fmoms[ss]-emoms)],{ss, 0, nn}]/2^nn;
+
+  soluzm = 
+   NMinimize[freeenm, js, MaxIterations -> 10000, 
+    PrecisionGoal -> pg, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+
+  prob = Table[
+    Binomial[nn,ss]*
[…]

### 2019-04-23 scripts/ME_4momsolution.m

@@ -2,9 +2,10 @@
 
 Clear[MEhom4moments,MEhom4thmoment,MEhomvmoments,MEhom4momentsprob]; 
 
-MEhom4moments[m1_, m2_, m3_, m4_, nn_, wp_:MachinePrecision,pg_:6,mi_:10000] := 
- Block[{totalprob,
-	zzzm, js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
+MEhom4moments[mr1_, mr2_, mr3_, mr4_, nn_, wp_:MachinePrecision,pg_:6,mi_:10000] := 
+  Block[{totalprob,
+	 m1=Rationalize[mr1,0],m2=Rationalize[mr2,0],m3=Rationalize[mr3,0],m4=Rationalize[mr4,0],
+	js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
 	nn2=Binomial[nn,2], nn3=Binomial[nn,3], nn4=Binomial[nn,4], 
 	bvalues},
        freeenm = Log@Sum[Binomial[nn, m]*

### 2019-04-23 scripts/ME_4momsolution_prior2.m

@@ -2,9 +2,10 @@
 
 Clear[MEhom4moments,MEhom4thmoment,MEhomvmoments,MEhom4momentsprob]; 
 
-MEhom4moments[m1_, m2_, m3_, m4_, nn_, wp_:MachinePrecision,pg_:3] := 
- Block[{totalprob,
-	zzzm, js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
+MEhom4moments[mr1_, mr2_, mr3_, mr4_, nn_, wp_:MachinePrecision,pg_:4] := 
+  Block[{totalprob,
+	 	 m1=Rationalize[mr1,0],m2=Rationalize[mr2,0],m3=Rationalize[mr3,0],m4=Rationalize[mr4,0],
+	js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
 	nn2=Binomial[nn,2], nn3=Binomial[nn,3], nn4=Binomial[nn,4], 
 	bvalues},
        freeenm = Log@Sum[(1-m+nn)/((nn+1)*(nn+2))*
@@ -14,7 +15,7 @@ MEhom4moments[m1_, m2_, m3_, m4_, nn_, wp_:MachinePrecision,pg_:3] :=
        {m, 0, nn}];
   soluzm = 
    NMinimize[freeenm, {js1, js2, js3, js4}, MaxIterations -> 10000, 
-    PrecisionGoal -> 4, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+    PrecisionGoal -> pg, AccuracyGoal -> Infinity,WorkingPrecision->wp];
   prob = Table[
 	  (1-m+nn)/((nn+1)*(nn+2))*Exp[js1*m/nn + js2*Binomial[m,2]/nn2 +
 		      js3*Binomial[m,3]/nn3 +

### 2019-04-23 scripts/ME_freqs.m

@@ -1,0 +1,25 @@
+(* Homog maxent with 4 moments - corresponds to constraining variances *)
+
+Clear[MEfreqs]; 
+
+MEfreqs[freqs_, n_,nn_, wp_:MachinePrecision,pg_:4] := 
+  Block[{r=Length[freqs]-1,js=Table[Unique["js"],Length[freqs]], freeenm, varm, soluzm, prob, bvalues},
+	
+       freeenm = Log@Sum[(1-ss+nn)/((nn+1)*(nn+2))*
+			 Exp[js.(Table[Binomial[n,s]*Binomial[nn-nn,ss-s]/Binomial[nn,ss],{s,0,r}]-freqs)]
+       ,{ss, 0, nn}];
+
+  soluzm = 
+   NMinimize[freeenm, js, MaxIterations -> 10000, 
+    PrecisionGoal -> pg, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+
+  prob = Table[
+    (1-ss+nn)/((nn+1)*(nn+2))*
+    Exp[js.(Table[Binomial[n,s]*Binomial[nn-nn,ss-s]/Binomial[nn,ss],{s,r}]-freqs)]
+	 /. soluzm[[2]],{ss, 0, nn}];
+  totalprob = Total[prob];prob=prob/totalprob;
+  bvalues=(Table[
+    Table[Binomial[n,s]*Binomial[nn-nn,ss-s]/Binomial[nn,ss],{ss,0,nn}].prob
+	   ,{s,0,r}]/freqs-1);
+ Print["rel. discrepancies %: ", bvalues*100//N];
+ {js /. soluzm[[2]],prob,totalprob}]

### 2019-04-23 scripts/ME_homsolution.m

@@ -4,15 +4,15 @@ Clear[MEhomsolution,MEhomsolutiontruncated,MEhomprob];
 
 MEhomsolution::usage = "Input: pop-avgd activity [0,1], pop-avgd couple activity, pop size, machine precision. Output: Lagr multipl, Lagr multipl, distribution, Z"
 (* it uses normalized values in the optimization; usually gives more precise results than MEhomsolution2, although it reports difficulties in finding the optimum *)
-MEhomsolution[meanmm_, meancorr_, nn_, wp_:MachinePrecision] := 
- Block[{totcorr = meancorr,
-   totmm = meanmm,totalprob,
+MEhomsolution[meanmm_, meancorr_, nn_, wp_:MachinePrecision, pg_:6] := 
+  Block[{totcorr = Rationalize[meancorr,0],
+	 totmm = Rationalize[meanmm,0],totalprob,
    zzzm, js2, js1, freeenm, varm, soluzm, prob,ran=Range[0,nn],bvalues},
   zzzm = Sum[Binomial[nn, m]*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn], {m, 0, nn}]/2^nn;
   freeenm = Log[zzzm] - js2*totcorr - js1*totmm;
   soluzm = 
    NMinimize[freeenm, {js1, js2}, MaxIterations -> 10000, 
-    PrecisionGoal -> 6,Method->Automatic, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+    PrecisionGoal -> pg,Method->Automatic, AccuracyGoal -> Infinity,WorkingPrecision->wp];
   prob = Table[
     Binomial[nn, m]*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn]/2^nn /. soluzm[[2]], {m, 
      0, nn}];

### 2019-04-23 scripts/ME_homsolution_prior2.m

@@ -4,17 +4,17 @@ Clear[MEhomsolution,MEhomsolutiontruncated,MEhomprob];
 
 MEhomsolution::usage = "Input: pop-avgd activity [0,1], pop-avgd couple activity, pop size, machine precision. Output: Lagr multipl, Lagr multipl, distribution, Z"
 (* it uses normalized values in the optimization; usually gives more precise results than MEhomsolution2, although it reports difficulties in finding the optimum *)
-MEhomsolution[meanmm_, meancorr_, nn_, wp_:MachinePrecision] := 
- Block[{totcorr = meancorr,
-   totmm = meanmm,totalprob,
+MEhomsolution[meanmm_, meancorr_, nn_, wp_:MachinePrecision,pg_:6] := 
+  Block[{totcorr = Rationalize[meancorr,0],
+	 totmm = Rationalize[meanmm,0],totalprob,
    zzzm, js2, js1, freeenm, varm, soluzm, prob,ran=Range[0,nn],bvalues},
-       zzzm = Sum[(1-m+nn)/((nn+1)*(nn+2))*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn], {m, 0, nn}]/2^nn;
+       zzzm = Sum[(1-m+nn)/((nn+1)*(nn+2))*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn], {m, 0, nn}];
   freeenm = Log[zzzm] - js2*totcorr - js1*totmm;
   soluzm = 
    NMinimize[freeenm, {js1, js2}, MaxIterations -> 10000, 
-    PrecisionGoal -> 6,Method->Automatic, AccuracyGoal -> Infinity,WorkingPrecision->wp];
+    PrecisionGoal -> pg,Method->Automatic, AccuracyGoal -> Infinity,WorkingPrecision->wp];
   prob = Table[
-    (1-m+nn)/((nn+1)*(nn+2))*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn]/2^nn /. soluzm[[2]], {m, 
+    (1-m+nn)/((nn+1)*(nn+2))*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn] /. soluzm[[2]], {m, 
      0, nn}];
   totalprob = Total[prob];prob=prob/totalprob;
[…]

### 2019-04-23 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-23T09:55:00+0200
+%% Last-Updated: 2019-04-23T09:59:41+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The maximum-entropy method can be used this way thanks to a very simple ***
-
-\clearpage
-
-\hrule
-Yasser's intro
-\hrule
-{\footnotesize
+\mynote{integrate above with Yasser's below}
+
 Experimental technologies to record the activity of many neurons at the
@@
 
-}
 \mynote{Luca: add this: from sampling theory we know that important

### 2019-04-23 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-23T09:59:41+0200
+%% Last-Updated: 2019-04-23T10:42:47+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 expectations of $s/n$ and $S/N$ are equal:
+\begin{subequations}
+  \label{eq:all_equal_expectations_poll}
 \begin{equation}
-  \label{eq:equal_expectations_poll}
   \E(s/n \| \yH) = \E(S/N \| \yH),
@@
 \end{equation}
+\end{subequations}
 Note that from the first $r$ factorial moments we can calculate the first
-$r$ raw power moments and vice versa; but the simple equalities above don't
+$r$ power moments and vice versa; but the simple equalities above don't
 hold for the power moments \citep[\sect~A]{portamanaetal2015}.
@@
 probability distribution $\pf(S\|\yH)$ be.
+
+\bigskip
+
+Whenever we record the activity of $n$ neurons we're effectively polling
+them on the activity of all $N$ neurons under the recording device -- those
+that could equally likely have been sampled instead. Suppose their
+activities are time-binned and binarized. At every time bin $t$, $s(t)$ out
+of the $n$ neurons \enquote{vote} for activity, and $n-s(t)$ vote for
+silence. The probability calculus tells us that the expectation
+formulae~\eqref{eq:all_equal_expectations_poll} hold for the total activity
+$S(t)$ of the full population of $N$ neurons that could have been recorded.
 

### 2019-04-23 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-23T10:42:47+0200
+%% Last-Updated: 2019-04-23T11:28:39+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \begin{split}
-  \E\condt[\Big]{\tfrac{s\;(s-1)}{n\;(n-1)}}{\yH} &= 
-  \E\condt[\Big]{\tfrac{S\;(S-1)}{N\;(N-1)}}{\yH},\\
-  \E\condt[\Big]{\tfrac{s\;(s-1)\;(s-2)}{n\;(n-1)\;(n-2)}}{\yH} &= 
-  \E\condt[\Big]{\tfrac{S\;(S-1)\;(S-2)}{N\;(N-1)\;(N-2)}}{\yH},\quad\dotsc
+  \E\condt[\big]{\tfrac{s\;(s-1)}{n\;(n-1)}}{\yH} &= 
+  \E\condt[\big]{\tfrac{S\;(S-1)}{N\;(N-1)}}{\yH},\\
+  \E\condt[\big]{\tfrac{s\;(s-1)\;(s-2)}{n\;(n-1)\;(n-2)}}{\yH} &= 
+  \E\condt[\big]{\tfrac{S\;(S-1)\;(S-2)}{N\;(N-1)\;(N-2)}}{\yH},
   \\
   \text{and in general}\quad
-  \E\condt[\Big]{\tbinom{s}{r}\;\tbinom{n}{r}^{-1}}{\yH} &=
-  \E\condt[\Big]{\tbinom{S}{r}\;\tbinom{N}{r}^{-1}}{\yH},
+  \E\condt[\big]{\tbinom{s}{r}\;\tbinom{n}{r}^{-1}}{\yH} &=
+  \E\condt[\big]{\tbinom{S}{r}\;\tbinom{N}{r}^{-1}}{\yH},
   \quad r<N.
@@
 \end{equation}
-\end{subequations}
 Note that from the first $r$ factorial moments we can calculate the first
 $r$ power moments and vice versa; but the simple equalities above don't
-hold for the power moments \citep[\sect~A]{portamanaetal2015}.
-
+hold for the power moments \citep[\sect~A]{portamanaetal2015}. The formulae
+above follow from a precise relation between our \dobs\ about the polled
+voters and about all voters:
+\begin{equation}
+  \label{eq:relation_ps_pS}
+  \pf(s \| \yH) = \sum_{S} \binom{n}{s}\;\binom{N-n}{S-s}\;\binom{N}{S}^{-1}\;
+  \pf(S \| \yH).
+\end{equation}
+\end{subequations}
 The relations above are powerful because they hold no matter what the
-probability distribution $\pf(S\|\yH)$ be.
+probability distribution $\pf(S\|\yH)$ may be.
 
@@
 
+This probabilistic fact is of outstanding importance in neuroscience, where
+we try to understand the workings of brain areas that may include tens of
+thousands of neurons, but we can record only few tens or hundreds of them.
 Whenever we record the activity of $n$ neurons we're effectively polling
 them on the activity of all $N$ neurons under the recording device -- those
-that could equally likely have been sampled instead. Suppose their
+that could equally likely have been sampled instead. Suppose the recorded
 activities are time-binned and binarized. At every time bin $t$, $s(t)$ out
@@
 $S(t)$ of the full population of $N$ neurons that could have been recorded.
+
+The simple expectation equalities are
+
+Expectations are often used with
 

### 2019-04-23 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-23T11:28:39+0200
+%% Last-Updated: 2019-04-23T13:06:43+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 expectations of $s/n$ and $S/N$ are equal:
-\begin{subequations}
-  \label{eq:all_equal_expectations_poll}
-\begin{equation}
+\begin{equation}\label{eq:poll_equal_expectations}
   \E(s/n \| \yH) = \E(S/N \| \yH),
 \end{equation}
-where $\yH$ represents our state of knowledge in this example.
-
-A more precise answer to our original question then is: \enquote{because we
-  \emph{expect} to see the same proportion of political preferences in a
-  sample as in the full population}.
-
-The equality above extends to other quantities beyond the means $S/N$ and
-$s/n$: it holds for all normalized \emph{factorial} moments
-\citep{potts1953,broca2005}:
-\begin{equation}
-  \label{eq:equal_factorialmoments_polls}
-  \begin{split}
-  \E\condt[\big]{\tfrac{s\;(s-1)}{n\;(n-1)}}{\yH} &= 
-  \E\condt[\big]{\tfrac{S\;(S-1)}{N\;(N-1)}}{\yH},\\
-  \E\condt[\big]{\tfrac{s\;(s-1)\;(s-2)}{n\;(n-1)\;(n-2)}}{\yH} &= 
-  \E\condt[\big]{\tfrac{S\;(S-1)\;(S-2)}{N\;(N-1)\;(N-2)}}{\yH},
-  \\
-  \text{and in general}\quad
-  \E\condt[\big]{\tbinom{s}{r}\;\tbinom{n}{r}^{-1}}{\yH} &=
-  \E\condt[\big]{\tbinom{S}{r}\;\tbinom{N}{r}^{-1}}{\yH},
-  \quad r<N.
-\end{split}
-\end{equation}
-Note that from the first $r$ factorial moments we can calculate the first
-$r$ power moments and vice versa; but the simple equalities above don't
-hold for the power moments \citep[\sect~A]{portamanaetal2015}. The formulae
-above follow from a precise relation between our \dobs\ about the polled
-voters and about all voters:
-\begin{equation}
-  \label{eq:relation_ps_pS}
-  \pf(s \| \yH) = \sum_{S} \binom{n}{s}\;\binom{N-n}{S-s}\;\binom{N}{S}^{-1}\;
-  \pf(S \| \yH).
-\end{equation}
-\end{subequations}
-The relations above are powerful because they hold no matter what the
-probability distribution $\pf(S\|\yH)$ may be.
+where $\yH$ represents our state of knowledge in this example. A more
+precise answer to our original question then is: \enquote{because we
+  \emph{expect} -- in the full probabilistic sense -- to see the same
+  proportion of political preferences in a sample as in the full
+  population}.
+
+The exact equality above extends to other quantities beyond the means $S/N$
+and $s/n$: it holds for all normalized \emph{factorial} moments
+\citep{potts1953,broca2005}[\sect~A]{portamanaetal2015}. These expectation
+equalities follow from a precise relation between our \dobs\ about the
+polled voters and about all voters. The relations above are powerful
+because they hold no matter what the probability distribution $\pf(S\|\yH)$
+may be.
 
@@
 of the $n$ neurons \enquote{vote} for activity, and $n-s(t)$ vote for
-silence. The probability calculus tells us that the expectation
-formulae~\eqref{eq:all_equal_expectations_poll} hold for the total activity
+silence. The probability calculus tells us that the expectation equalities
+such as~\eqref{eq:poll_equal_expectations} hold for the total activity
 $S(t)$ of the full population of $N$ neurons that could have been recorded.
@@
 \bigskip
+
+\section{Mathematical development}
+\label{sec:maths_development}
+
+\begin{subequations}
+  \label{eq:all_equal_expectations_poll}
+\begin{equation}
+  \label{eq:equal_factorialmoments_polls}
+  \begin{split}
+  \E\condt[\big]{\tfrac{s\;(s-1)}{n\;(n-1)}}{\yH} &= 
+  \E\condt[\big]{\tfrac{S\;(S-1)}{N\;(N-1)}}{\yH},\\
+  \E\condt[\big]{\tfrac{s\;(s-1)\;(s-2)}{n\;(n-1)\;(n-2)}}{\yH} &= 
+  \E\condt[\big]{\tfrac{S\;(S-1)\;(S-2)}{N\;(N-1)\;(N-2)}}{\yH},
+  \\
+  \text{and in general}\quad
+  \E\condt[\big]{\tbinom{s}{r}\;\tbinom{n}{r}^{-1}}{\yH} &=
+  \E\condt[\big]{\tbinom{S}{r}\;\tbinom{N}{r}^{-1}}{\yH},
+  \quad r<N.
+\end{split}
+\end{equation}
+
+Note that from the first $r$ factorial moments we can calculate the first
+$r$ power moments and vice versa; but the simple equalities above don't
+hold for the power moments \citep[\sect~A]{portamanaetal2015}.
+
+\begin{equation}
+  \label{eq:relation_ps_pS}
+  \pf(s \| \yH) = \sum_{S} \binom{n}{s}\;\binom{N-n}{S-s}\;\binom{N}{S}^{-1}\;
+  \pf(S \| \yH).
+\end{equation}
+\end{subequations}
 

### 2019-04-23 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-23T13:06:43+0200
+%% Last-Updated: 2019-04-23T13:50:31+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-This probabilistic fact is of outstanding importance in neuroscience, where
-we try to understand the workings of brain areas that may include tens of
-thousands of neurons, but we can record only few tens or hundreds of them.
-Whenever we record the activity of $n$ neurons we're effectively polling
-them on the activity of all $N$ neurons under the recording device -- those
-that could equally likely have been sampled instead. Suppose the recorded
-activities are time-binned and binarized. At every time bin $t$, $s(t)$ out
-of the $n$ neurons \enquote{vote} for activity, and $n-s(t)$ vote for
-silence. The probability calculus tells us that the expectation equalities
-such as~\eqref{eq:poll_equal_expectations} hold for the total activity
-$S(t)$ of the full population of $N$ neurons that could have been recorded.
-
-The simple expectation equalities are
-
-Expectations are often used with
-
-\bigskip
-
-In the present work we show how to use the maximum-entropy method to assign
-a probability distribution for the total activity at a new time bin of a
-population of neurons, \emph{most of which can be unrecorded}, given
-specific statistics about the activity of a recorded \emph{sample} of the
-full population. This kind of probability assignment is extremely important
-in neuroscience: we try to understand the workings of brain areas that may
-include tens of thousands of neurons, but we can record only few tens or
-hundreds of them.
-
-Our approach also corrects a slight misuse of the maximum-entropy method
-as sometimes appears in the literature. When this method is applied to a
-set of neurons, it implicitly assumes that the set is completely isolated
-from other neurons. We'll reveal this implicit assumption in several ways.
-The probability distribution thus constructed is therefore affected by this
-hidden assumption. We'll study in which circumstances the application of
-the method at the sample level and at the full-population level give
-similar or different probabilities.
+These probabilistic properties are of outstanding importance in
+neuroscience, where we try to understand the workings of brain areas that
+may include tens of thousands of neurons, but we can record only few tens
+or hundreds of them. Whenever we record the activity of $n$ neurons we're
+effectively polling them on the activity of all $N$ neurons under the
+recording device -- those that could equally likely have been sampled
+instead. Suppose the recorded activities are time-binned and binarized. At
+every time bin $t$, $s(t)$ out of the $n$ neurons \enquote{vote} for
+activity, and $n-s(t)$ vote for silence. The probability calculus tells us
+that expectation equalities such as~\eqref{eq:poll_equal_expectations} hold
+for the total activity $S(t)$ of the full population of $N$ neurons that
+could have been recorded.
+
+Expectation properties are especially important for the maximum-entropy
+method, which is used for various purposes in the neuroscientific
+literature. This method is typically applied constraining the equality of
+expectations and empirical averages. Expectation equalities such
+as~\eqref{eq:poll_equal_expectations} then have a far-reaching consequence:
+\emph{from empirical statistics of the activity of a \textbf{sample} of
+  neurons we can construct maximum-entropy distributions for the total
+  activity of the \textbf{full population} of neurons that could have been
+  sampled by the recording device}.
+
+% In the present work we show how to use the maximum-entropy method to assign
+% a probability distribution for the total activity at a new time bin of a
+% population of neurons, \emph{most of which can be unrecorded}, given
+% specific statistics about the activity of a recorded \emph{sample} of the
+% full population. This kind of probability assignment is extremely important
+% in neuroscience: we try to understand the workings of brain areas that may
+% include tens of thousands of neurons, but we can record only few tens or
+% hundreds of them.
+
+% Our approach also corrects a slight misuse of the maximum-entropy method
+% as sometimes appears in the literature. When this method is applied to a
+% set of neurons, it implicitly assumes that the set is completely isolated
+% from other neurons. We'll reveal this implicit assumption in several ways.
+% The probability distribution thus constructed is therefore affected by this
+% hidden assumption. We'll study in which circumstances the application of
+% the method at the sample level and at the full-population level give
+% similar or different probabilities.
 

### 2019-05-12 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-23T13:50:31+0200
+%% Last-Updated: 2019-05-12T18:09:32+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Polling neurons}
-\label{sec:polls}
-
-When political elections are approaching, mass media are constantly
-bubbling with results and discussions of pre-election polls. Why do we
-conduct pre-election polls and other kinds of survey sampling? The
-simplest, intuitive answer is \enquote{because they give us an idea of what
-  the final election outcomes will be}. Can this answer be made more
-quantitative?
-
-Yes. In the simplest cases, the probability calculus give us simple and
-precise quantitative relations between our expectations about a population
--- of voters, for instance -- and about a sample thereof. Suppose there are
-$N$ voters, $S$ of which prefer one political candidate and $N-S$ of which
-prefer another. We don't know $S$ or $N-S$. Now take a sample of $n$
-voters, in such a way that we can assume that every set of $n$ voters is
-equally likely to be sampled. In this sample, $s$ voters prefer the first
-candidate, and $n-s$ the second. The probability calculus tells us that our
-expectations of $s/n$ and $S/N$ are equal:
-\begin{equation}\label{eq:poll_equal_expectations}
-  \E(s/n \| \yH) = \E(S/N \| \yH),
-\end{equation}
-where $\yH$ represents our state of knowledge in this example. A more
-precise answer to our original question then is: \enquote{because we
-  \emph{expect} -- in the full probabilistic sense -- to see the same
-  proportion of political preferences in a sample as in the full
-  population}.
-
-The exact equality above extends to other quantities beyond the means $S/N$
-and $s/n$: it holds for all normalized \emph{factorial} moments
-\citep{potts1953,broca2005}[\sect~A]{portamanaetal2015}. These expectation
-equalities follow from a precise relation between our \dobs\ about the
-polled voters and about all voters. The relations above are powerful
-because they hold no matter what the probability distribution $\pf(S\|\yH)$
-may be.
-
-\bigskip
-
-These probabilistic properties are of outstanding importance in
-neuroscience, where we try to understand the workings of brain areas that
-may include tens of thousands of neurons, but we can record only few tens
-or hundreds of them. Whenever we record the activity of $n$ neurons we're
-effectively polling them on the activity of all $N$ neurons under the
-recording device -- those that could equally likely have been sampled
-instead. Suppose the recorded activities are time-binned and binarized. At
-every time bin $t$, $s(t)$ out of the $n$ neurons \enquote{vote} for
-activity, and $n-s(t)$ vote for silence. The probability calculus tells us
-that expectation equalities such as~\eqref{eq:poll_equal_expectations} hold
-for the total activity $S(t)$ of the full population of $N$ neurons that
-could have been recorded.
-
-Expectation properties are especially important for the maximum-entropy
-method, which is used for various purposes in the neuroscientific
-literature. This method is typically applied constraining the equality of
-expectations and empirical averages. Expectation equalities such
-as~\eqref{eq:poll_equal_expectations} then have a far-reaching consequence:
-\emph{from empirical statistics of the activity of a \textbf{sample} of
-  neurons we can construct maximum-entropy distributions for the total
-  activity of the \textbf{full population} of neurons that could have been
-  sampled by the recording device}.
-
-% In the present work we show how to use the maximum-entropy method to assign
-% a probability distribution for the total activity at a new time bin of a
-% population of neurons, \emph{most of which can be unrecorded}, given
-% specific statistics about the activity of a recorded \emph{sample} of the
-% full population. This kind of probability assignment is extremely important
-% in neuroscience: we try to understand the workings of brain areas that may
-% include tens of thousands of neurons, but we can record only few tens or
-% hundreds of them.
-
-% Our approach also corrects a slight misuse of the maximum-entropy method
-% as sometimes appears in the literature. When this method is applied to a
-% set of neurons, it implicitly assumes that the set is completely isolated
-% from other neurons. We'll reveal this implicit assumption in several ways.
-% The probability distribution thus constructed is therefore affected by this
-% hidden assumption. We'll study in which circumstances the application of
-% the method at the sample level and at the full-population level give
-% similar or different probabilities.
-
-\mynote{integrate above with Yasser's below}
+\mynote{comment about the possibility of drawing conclusions about a brain
+  area using different sets of neurons (eg because of recording across many
+  sessions)}
+
+\section{Introduction}
+\label{sec:intro}
+
+What correlations are important for the description of the multi-neuronal
+activity in a specific brain area? How does such activity change when
+external stimuli or experimental conditions change? Does such activity
+range over all its theoretically possible values, or only over a
+submanifold thereof?
+
+Answering this kind of questions always involves an element of uncertainty.
+Our answers therefore involve experimental data, such as neuronal
+recordings from specific brain areas, but also probabilistic assumptions --
+probabilistic \enquote{models} -- about biological conditions and
+mechanisms that cannot be experimentally ascertained.
+
+Despite fast advances in recording technologies, our best experimental
+measurements of neuronal activity can still only sample a very small number
+of neurons compared to the number that constitutes a brain area. Many
+probabilistic models focus on such samples only,
 
@@
 
+When political elections are approaching, mass media are constantly
+bubbling with results and discussions of pre-election polls. Why do we
+conduct pre-election polls and other kinds of survey sampling? The
+simplest, intuitive answer is \enquote{because they give us an idea of what
+  the final election outcomes will be}. Can this answer be made more
+quantitative?
+
+Yes. In the simplest cases, the probability calculus give us simple and
+precise quantitative relations between our expectations about a population
+-- of voters, for instance -- and about a sample thereof. Suppose there are
+$N$ voters, $S$ of which prefer one political candidate and $N-S$ of which
+prefer another. We don't know $S$ or $N-S$. Now take a sample of $n$
+voters, in such a way that we can assume that every set of $n$ voters is
+equally likely to be sampled. In this sample, $s$ voters prefer the first
+candidate, and $n-s$ the second. The probability calculus tells us that our
+expectations of $s/n$ and $S/N$ are equal:
+\begin{equation}\label{eq:poll_equal_expectations}
+  \E(s/n \| \yH) = \E(S/N \| \yH),
+\end{equation}
+where $\yH$ represents our state of knowledge in this example. A more
+precise answer to our original question then is: \enquote{because we
+  \emph{expect} -- in the full probabilistic sense -- to see the same
+  proportion of political preferences in a sample as in the full
+  population}.
+
+The exact equality above extends to other quantities beyond the means $S/N$
+and $s/n$: it holds for all normalized \emph{factorial} moments
+\citep{potts1953,broca2005}[\sect~A]{portamanaetal2015}. These expectation
+equalities follow from a precise relation between our \dobs\ about the
+polled voters and about all voters. The relations above are powerful
+because they hold no matter what the probability distribution $\pf(S\|\yH)$
+may be.
+
+\bigskip
+
+These probabilistic properties are of outstanding importance in
+neuroscience, where we try to understand the workings of brain areas that
+may include tens of thousands of neurons, but we can record only few tens
+or hundreds of them. Whenever we record the activity of $n$ neurons we're
+effectively polling them on the activity of all $N$ neurons under the
+recording device -- those that could equally likely have been sampled
+instead. Suppose the recorded activities are time-binned and binarized. At
+every time bin $t$, $s(t)$ out of the $n$ neurons \enquote{vote} for
+activity, and $n-s(t)$ vote for silence. The probability calculus tells us
+that expectation equalities such as~\eqref{eq:poll_equal_expectations} hold
+for the total activity $S(t)$ of the full population of $N$ neurons that
+could have been recorded.
+
+Expectation properties are especially important for the maximum-entropy
+method, which is used for various purposes in the neuroscientific
+literature. This method is typically applied constraining the equality of
+expectations and empirical averages. Expectation equalities such
+as~\eqref{eq:poll_equal_expectations} then have a far-reaching consequence:
+\emph{from empirical statistics of the activity of a \textbf{sample} of
+  neurons we can construct maximum-entropy distributions for the total
+  activity of the \textbf{full population} of neurons that could have been
+  sampled by the recording device}.
+
+% In the present work we show how to use the maximum-entropy method to assign
+% a probability distribution for the total activity at a new time bin of a
+% population of neurons, \emph{most of which can be unrecorded}, given
+% specific statistics about the activity of a recorded \emph{sample} of the
+% full population. This kind of probability assignment is extremely important
+% in neuroscience: we try to understand the workings of brain areas that may
+% include tens of thousands of neurons, but we can record only few tens or
+% hundreds of them.
+
+% Our approach also corrects a slight misuse of the maximum-entropy method
+% as sometimes appears in the literature. When this method is applied to a
+% set of neurons, it implicitly assumes that the set is completely isolated
+% from other neurons. We'll reveal this implicit assumption in several ways.
+% The probability distribution thus constructed is therefore affected by this
+% hidden assumption. We'll study in which circumstances the application of
+% the method at the sample level and at the full-population level give
+% similar or different probabilities.
+
 %%% Local Variables: 

### 2019-05-12 maxNt_v5.tex
update

--- 2019-04-23 maxNt_v5.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-04-23T13:50:31+0200
+%% Last-Updated: 2019-05-12T18:49:45+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Polling neurons}
-\label{sec:polls}
-
-When political elections are approaching, mass media are constantly
-bubbling with results and discussions of pre-election polls. Why do we
-conduct pre-election polls and other kinds of survey sampling? The
-simplest, intuitive answer is \enquote{because they give us an idea of what
-  the final election outcomes will be}. Can this answer be made more
-quantitative?
-
-Yes. In the simplest cases, the probability calculus give us simple and
-precise quantitative relations between our expectations about a population
--- of voters, for instance -- and about a sample thereof. Suppose there are
-$N$ voters, $S$ of which prefer one political candidate and $N-S$ of which
-prefer another. We don't know $S$ or $N-S$. Now take a sample of $n$
-voters, in such a way that we can assume that every set of $n$ voters is
-equally likely to be sampled. In this sample, $s$ voters prefer the first
-candidate, and $n-s$ the second. The probability calculus tells us that our
-expectations of $s/n$ and $S/N$ are equal:
-\begin{equation}\label{eq:poll_equal_expectations}
-  \E(s/n \| \yH) = \E(S/N \| \yH),
-\end{equation}
-where $\yH$ represents our state of knowledge in this example. A more
-precise answer to our original question then is: \enquote{because we
-  \emph{expect} -- in the full probabilistic sense -- to see the same
-  proportion of political preferences in a sample as in the full
-  population}.
-
-The exact equality above extends to other quantities beyond the means $S/N$
-and $s/n$: it holds for all normalized \emph{factorial} moments
-\citep{potts1953,broca2005}[\sect~A]{portamanaetal2015}. These expectation
-equalities follow from a precise relation between our \dobs\ about the
-polled voters and about all voters. The relations above are powerful
-because they hold no matter what the probability distribution $\pf(S\|\yH)$
-may be.
-
-\bigskip
-
-These probabilistic properties are of outstanding importance in
-neuroscience, where we try to understand the workings of brain areas that
-may include tens of thousands of neurons, but we can record only few tens
-or hundreds of them. Whenever we record the activity of $n$ neurons we're
-effectively polling them on the activity of all $N$ neurons under the
-recording device -- those that could equally likely have been sampled
-instead. Suppose the recorded activities are time-binned and binarized. At
-every time bin $t$, $s(t)$ out of the $n$ neurons \enquote{vote} for
-activity, and $n-s(t)$ vote for silence. The probability calculus tells us
-that expectation equalities such as~\eqref{eq:poll_equal_expectations} hold
-for the total activity $S(t)$ of the full population of $N$ neurons that
-could have been recorded.
-
-Expectation properties are especially important for the maximum-entropy
-method, which is used for various purposes in the neuroscientific
-literature. This method is typically applied constraining the equality of
-expectations and empirical averages. Expectation equalities such
-as~\eqref{eq:poll_equal_expectations} then have a far-reaching consequence:
-\emph{from empirical statistics of the activity of a \textbf{sample} of
-  neurons we can construct maximum-entropy distributions for the total
-  activity of the \textbf{full population} of neurons that could have been
-  sampled by the recording device}.
-
-% In the present work we show how to use the maximum-entropy method to assign
-% a probability distribution for the total activity at a new time bin of a
-% population of neurons, \emph{most of which can be unrecorded}, given
-% specific statistics about the activity of a recorded \emph{sample} of the
-% full population. This kind of probability assignment is extremely important
-% in neuroscience: we try to understand the workings of brain areas that may
-% include tens of thousands of neurons, but we can record only few tens or
-% hundreds of them.
-
-% Our approach also corrects a slight misuse of the maximum-entropy method
-% as sometimes appears in the literature. When this method is applied to a
-% set of neurons, it implicitly assumes that the set is completely isolated
-% from other neurons. We'll reveal this implicit assumption in several ways.
-% The probability distribution thus constructed is therefore affected by this
-% hidden assumption. We'll study in which circumstances the application of
-% the method at the sample level and at the full-population level give
-% similar or different probabilities.
-
-\mynote{integrate above with Yasser's below}
+\mynote{comment about the possibility of drawing conclusions about a brain
+  area using different sets of neurons (eg because of recording across many
+  sessions)}
+
+\section{Introduction}
+\label{sec:intro}
+
+What correlations are important for the description of the multi-neuronal
+activity in a specific brain area? How does such activity change when
+external stimuli or experimental conditions change? Does such activity
+range over all its theoretically possible values, or only over a
+submanifold thereof?
+
+Answering this kind of questions always engages an element of uncertainty.
+Our answers therefore involve experimental data, such as neuronal
+recordings from specific brain areas, and also probabilistic assumptions --
+probabilistic \enquote{models} -- about biological conditions and
+mechanisms that cannot be experimentally ascertained.
+
+Despite remarkable advances in recording technologies, our best experimental
+measurements of neuronal activity can still only sample a very small number
+of neurons compared to the number that constitutes a brain area. Many
+probabilistic models focus on such samples only, somehow neglecting, in
+their assumptions, that the recorded neurons are a sample from a larger
+network. This kind of isolation assumptions sometimes escape attention
+because it's subtly hidden in the mathematics. Some probabilistic models
+try to take also unrecorded neurons into account and in doing so become
+very complex.
+
+In the present work
 

### 2019-05-13 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-05-12T18:49:45+0200
+%% Last-Updated: 2019-05-13T10:06:49+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Despite remarkable advances in recording technologies, our best experimental
-measurements of neuronal activity can still only sample a very small number
-of neurons compared to the number that constitutes a brain area. Many
-probabilistic models focus on such samples only, somehow neglecting, in
-their assumptions, that the recorded neurons are a sample from a larger
-network. This kind of isolation assumptions sometimes escape attention
-because it's subtly hidden in the mathematics. Some probabilistic models
-try to take also unrecorded neurons into account and in doing so become
-very complex.
-
-In the present work
+Despite remarkable advances in recording technologies, the best
+experimental measurements of neuronal activity can still only record a very
+small sample of neurons compared to the number that constitutes a brain
+area. Many probabilistic models focus on such samples only, somehow
+neglecting, in their assumptions, that the recorded neurons are a sample
+from a larger network. This kind of isolation assumptions sometimes escape
+attention, being subtly hidden in the mathematics. Some probabilistic
+models try to take also  unrecorded neurons into account, but
+become very complex in doing so.
+
+In the present work we give an answer to this question: What kind of
+inferences about the total activity of a large neuronal network can we make
+by recording a very small sample thereof? Our answer comes from the
+simple combination of basic sampling probability relations and the maximum-entropy
+method.\mynote{shall we give a two-sentence summary of the main idea here?}
+
+***
 

### 2019-05-14 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-05-13T10:06:49+0200
+%% Last-Updated: 2019-05-13T11:41:45+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 small sample of neurons compared to the number that constitutes a brain
-area. Many probabilistic models focus on such samples only, somehow
+region. Many probabilistic models focus on such samples only, somehow
 neglecting, in their assumptions, that the recorded neurons are a sample
@@
 attention, being subtly hidden in the mathematics. Some probabilistic
-models try to take also  unrecorded neurons into account, but
-become very complex in doing so.
+models try to take also unrecorded neurons into account, but become very
+complex in doing so.
 
@@
 method.\mynote{shall we give a two-sentence summary of the main idea here?}
+
+The maximum-entropy method\mynote{refs} has been used for a variety of
+inferences about neuronal activity from various brain areas and about other
+phenomena of importance to the neurosciences, for example gene regulation
+and protein sequencing\mynote{what's this exactly? refs}. This method is
+often used to test whether some statistics of the data, for example
+second-order time correlatins, is sufficient for making other inferences
+about the data. It can be considered as an approximation of probabilistic
+models based on various assumptions of inferential
+sufficiency\mynote{refs} -- for example of empirical moments or of long-run
+frequencies.
+
+\mynote{Say something more about advantages of such question/answer: for
+  example we can make statements about total activity of brain region even
+  across recordings when sampled neurons aren't the same.}
 

### 2019-05-14 scripts/ME_1fmoms.m

@@ -1,11 +1,11 @@
-(* Homog maxent with 4 moments - corresponds to constraining variances *)
+(* Homog maxent with any number of moments *)
 
 Clear[ME1fmoms,ME1fmoms0]; 
 
 ME1fmoms[eemoms_, nn_, wp_:MachinePrecision,pg_:4] := 
   Block[{ll=Length[eemoms],js=Table[Unique["js"],Length[eemoms]], freeenm, soluzm, prob, bvalues,fmoms,emoms},
-	emoms=1-eemoms;
-	fmoms[s_]=Table[1-Binomial[s,r]/Binomial[nn,r], {r,ll}];
+	emoms=eemoms;
+	fmoms[s_]=Table[Binomial[s,r]/Binomial[nn,r], {r,ll}];
 	freeenm = Log@Sum[(1-ss+nn)/((nn+1)*(nn+2))*
 			 Exp[js.(fmoms[ss]-emoms)],{ss, 0, nn}];
 

### 2019-05-17 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-05-13T11:41:45+0200
+%% Last-Updated: 2019-05-17T16:54:33+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\newcommand*{\yXv}{\varSigma}
-\newcommand*{\yRv}{S}
+\newcommand*{\yAv}{A}
 %\newcommand*{\yxv}{s}
-\newcommand*{\yrv}{s}
+\newcommand*{\yav}{a}
 \newcommand*{\yNv}{N}
 \newcommand*{\yNN}{\varNu}
+\newcommand*{\yFF}{F}
+\newcommand*{\yff}{f}
+\newcommand*{\yF}{\bm{\yFF}}
+\newcommand*{\yf}{\bm{\yff}}
 
@@
 %\newcommand*{\yXf}{\av{\yX}}%pop av state
-\newcommand*{\yrrs}{\sav{\yr\yr}}%subpop av state
-%\newcommand*{\yXXf}{\av{\yR\yR}}%subpop av state
-\newcommand*{\yr}{\bm{\yrv}}%subpop value
-\newcommand*{\yrs}{\yrv}%subpop av value
-\newcommand*{\yR}{\bm{\yRv}}%pop value
-\newcommand*{\yRf}{\yRv}%pop av value
+\newcommand*{\yaa}{\bm{\yav}}%subpop value
+\newcommand*{\yaas}{\sav{\yaa\yaa}}%subpop av state
+%\newcommand*{\yXXf}{\av{\yAA\yAA}}%subpop av state
+\newcommand*{\ya}{\yav}%subpop av value
+\newcommand*{\yAA}{\bm{\yAv}}%pop value
+\newcommand*{\yA}{\yAv}%pop av value
 %conditional assumptions
@@
 
-\newcommand*{\ynu}{\bm{\nu}}
+\newcommand*{\yR}{\bm{R}}
+
+\newcommand*{\ynuu}{\nu}
+\newcommand*{\ynu}{\bm{\ynuu}}
+\newcommand*{\yjj}{\phi}
+\newcommand*{\yj}{\bm{\yjj}}
 
 \newcommand*{\prop}[1]{`#1'}
+
+\newcommand{\appropto}{\mathrel{\vcenter{
+  \offinterlineskip\halign{\hfil$##$\cr
+    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
+% https://tex.stackexchange.com/a/33547/97039
 
@@
 
-  Let the variable $\yrv_i(t)\in\set{0,1}$ denote the activity of the $i$th
+  Let the variable $\yav_i(t)\in\set{0,1}$ denote the activity of the $i$th
   sampled neuron at time bin $t$. Collectively denote the $n$ activities
-  with $\yr(t) \defd \bigl(\yrv_1(t),\dotsc,\yrv_n(t)\bigr)$. The
+  with $\yaa(t) \defd \bigl(\yav_1(t),\dotsc,\yav_n(t)\bigr)$. The
   population-averaged activity at that bin is
-  $\yrs(t) \defd \sum_i \yrv_i(t)/n$. If we count the number of distinct
+  $\ya(t) \defd \sum_i \yav_i(t)/n$. If we count the number of distinct
   pairs of active neurons at that bin we combinatorially find
-  $\tbinom{n\yrs(t)}{2}\equiv n\yrs(t)\,[n\yrs(t)-1]/2$. There can be at
+  $\tbinom{n\ya(t)}{2}\equiv n\ya(t)\,[n\ya(t)-1]/2$. There can be at
   most $\tbinom{n}{2}$ simultaneously active pairs, so the
   population-averaged pair activity is
-  $\av{\yr \yr}(t) \defd \tbinom{n}{2}^{-1}\tbinom{n\yrs(t)}{2}$. With some
+  $\av{\yaa \yaa}(t) \defd \tbinom{n}{2}^{-1}\tbinom{n\ya(t)}{2}$. With some
   combinatorics we see that the population-averaged activity of $m$-tuples
@@
     \label{eq:products_intermsof_average}
-    \av{\underbrace{\yr\dotsm\yr}_{\text{$m$ terms}}}(t)
-    = \binom{n}{m}^{-1}\binom{n\yrs(t)}{m}.
+    \av{\underbrace{\yaa\dotsm\yaa}_{\text{$m$ terms}}}(t)
+    = \binom{n}{m}^{-1}\binom{n\ya(t)}{m}.
   \end{equation}
@@
   For brevity let us agree to simply call \enquote{activity} the average
-  $\yrs$, \enquote{pair-activity} the average $\av{\yr \yr}$, and so on.
+  $\ya$, \enquote{pair-activity} the average $\av{\yaa \yaa}$, and so on.
 
 \item Construct a sequence of relative-\me\ distributions for the activity
-  $\yrs$, using this sequence of constraints:
+  $\ya$, using this sequence of constraints:
   \begin{itemize}
-  \item the time average of the activity: $\tav{\yrs} \defd \sum_t\yrs(t)/T$;
+  \item the time average of the activity: $\tav{\ya} \defd \sum_t\ya(t)/T$;
   \item the time averages of the activity and of the pair-activity
-    $\tav{\av{\yr \yr}} \defd \sum_t\av{\yr \yr}(t)/T$;
+    $\tav{\av{\yaa \yaa}} \defd \sum_t\av{\yaa \yaa}(t)/T$;
   \item \ldots
@@
   \end{itemize}
-  Call the resulting distributions $p_1(\yrs), p_2(\yrs),\dotsc,p_k(\yrs)$.
+  Call the resulting distributions $p_1(\ya), p_2(\ya),\dotsc,p_k(\ya)$.
   The time-bin dependence is now absent because these distributions can be
@@
   We also have the empirical frequency distribution of the total activity,
-  $f(\yrs)$, counted from the time bins.
+  $f(\ya)$, counted from the time bins.
 
@@
 larger population, but using the corresponding capital letters:
-$\yRv_{i}(t)$ is the activity of the $i$th neuron at time bin $t$,
-$\yRf(t) \defd \sum_{i} \yRv_{i}(t)/\yNv$ is the activity at that
+$\yAv_{i}(t)$ is the activity of the $i$th neuron at time bin $t$,
+$\yA(t) \defd \sum_{i} \yAv_{i}(t)/\yNv$ is the activity at that
 bin averaged over the larger population, and so on.
@@
 %   \begin{aligned}
-%     &\text{\enquote{$\yXv_{\iota} = \yRv_{\iota}$} 
-%       stands for \enquote{the activity of the $\iota$th neuron is $\yRv_{\iota}$}},
+%     &\text{\enquote{$\yXv_{\iota} = \yAv_{\iota}$} 
+%       stands for \enquote{the activity of the $\iota$th neuron is $\yAv_{\iota}$}},
 %     \\
-%     &\text{\enquote{$\yXf = \yRf$} 
-%       stands for \enquote{the (population-averaged) activity of the neurons is $\yRf$}},
+%     &\text{\enquote{$\yXf = \yA$} 
+%       stands for \enquote{the (population-averaged) activity of the neurons is $\yA$}},
 %     \\
-%     &\text{\enquote{$\yxv_i = \yrv_i$} 
-%       stands for \enquote{the activity of the $i$th sample neuron is $\yrv_i$}},
+%     &\text{\enquote{$\yxv_i = \yav_i$} 
+%       stands for \enquote{the activity of the $i$th sample neuron is $\yav_i$}},
 %   \end{aligned}
@@
   \label{eq:joint_plaus}
-  \pf(\yRv_1, \yRv_2, \dotsc, \yRv_\yNv \| \yHc)
+  \pf(\yAv_1, \yAv_2, \dotsc, \yAv_\yNv \| \yHc)
   \quad\text{or}\quad
-\pf(\yR \| \yHc), \quad \yR \in \set{0,1}^\yNv.
+\pf(\yAA \| \yHc), \quad \yAA \in \set{0,1}^\yNv.
 \end{equation}
@@
   \label{eq:sample_plaus}
-  \pf(\yrv_1, \yrv_2, \dotsc, \yrv_n \| \yHc) \quad\text{or}\quad
-\pf(\yr \| \yHc), \quad \yr \in \set{0,1}^n.
+  \pf(\yav_1, \yav_2, \dotsc, \yav_n \| \yHc) \quad\text{or}\quad
+\pf(\yaa \| \yHc), \quad \yaa \in \set{0,1}^n.
 \end{equation}
@@
   \label{eq:joint_plaus_N_homog}
-  \pf(\yR \|  \yH) \equiv
-  \sum_{\yRf}\pf(\yR \| \yRf, \yH)\,
-  \pf(\yRf \| \yH) =
-  \binom{\yNv}{\yNv\yRf}^{-1} \pf(\yRf \| \yH).
+  \pf(\yAA \|  \yH) \equiv
+  \sum_{\yA}\pf(\yAA \| \yA, \yH)\,
+  \pf(\yA \| \yH) =
+  \binom{\yNv}{\yNv\yA}^{-1} \pf(\yA \| \yH).
 \end{equation}
@@
 This result is intuitive: owing to symmetry, we must assign equal
-probabilities to all $\tbinom{\yNv}{\yNv\yRf}$ activity vectors with
-$\yNv\yRf$ active neurons; the probability of each activity vector is
+probabilities to all $\tbinom{\yNv}{\yNv\yA}$ activity vectors with
+$\yNv\yA$ active neurons; the probability of each activity vector is
 therefore given by that of the average activity divided by the number of
@@
   \label{eq:conditional_hypergeometric}
-  \pf(\yrs \|\yRf, \yH)=
-  \binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}
-  \defs \ypp_{\yrs\yRf},
+  \pf(\ya \|\yA, \yH)=
+  \binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}
+  \defs \ypp_{\ya\yA},
 \end{equation}
@@
   \label{eq:subpop_average}
-  \pf(\yrs \| \yH) = \sum_{\yRf}
-  \pf(\yrs \|\yRf, \yH)\,
-  \pf(\yRf \| \yH)
+  \pf(\ya \| \yH) = \sum_{\yA}
+  \pf(\ya \|\yA, \yH)\,
+  \pf(\yA \| \yH)
   = 
-  \sum_{\yRf}
-  \ypp_{\yrs\yRf}\,
-  \pf(\yRf \| \yH).
+  \sum_{\yA}
+  \ypp_{\ya\yA}\,
+  \pf(\yA \| \yH).
 \end{gather}
@@
   \label{eq:marginal}
-  \pf(\yr \| \yH) = \binom{n}{n\yrs}^{-1} \pf(\yrs \| \yH).
+  \pf(\yaa \| \yH) = \binom{n}{n\ya}^{-1} \pf(\ya \| \yH).
 \end{equation}
@@
 The conditional probability
-$\pf(\yrs \|\yRf, \yH) \equiv \ypp_{\yrs \yRf}$, besides
+$\pf(\ya \|\yA, \yH) \equiv \ypp_{\ya \yA}$, besides
 relating the distributions for the population and sample activities via
 marginalization, also allows us to express the expectation value of any
-function of the sample activity, $\yg_{\yrs}$, in terms of the distribution
+function of the sample activity, $\yg_{\ya}$, in terms of the distribution
 for the full population, as follows:
@@
   \equiv
-  \sum_{\yrs} \yg_{\yrs}\,\pf(\yrs \| \yH)
+  \sum_{\ya} \yg_{\ya}\,\pf(\ya \| \yH)
   =
-  \sum_{\yrs} \yg_{\yrs} \sum_{\yRf} \ypp_{\yrs\yRf}\,\pf(\yRf \| \yH)
+  \sum_{\ya} \yg_{\ya} \sum_{\yA} \ypp_{\ya\yA}\,\pf(\yA \| \yH)
   =%{}\\
-  \sum_{\yRf} \bigl( \tsum_{\yrs} \yg_{\yrs}  \ypp_{\yrs\yRf} \bigr)\,
-  \pf(\yRf \| \yH),
+  \sum_{\yA} \bigl( \tsum_{\ya} \yg_{\ya}  \ypp_{\ya\yA} \bigr)\,
+  \pf(\yA \| \yH),
 %  =
-%  \sum_{N\yRf=0}^N \sum_{n\yrs=0}^n f(\yrs)\ypp_{\yrs\yRf}\,\p(\yXf=\yRf \| \yH)
+%  \sum_{N\yA=0}^N \sum_{n\ya=0}^n f(\ya)\ypp_{\ya\yA}\,\p(\yXf=\yA \| \yH)
 %  \equiv  \expeb{\yg^*}.
@@
 where the second step uses \eqn~\eqref{eq:subpop_average}. The last
-expression shows that the expectation of the function $\yg_{\yrs}$ is equal to
+expression shows that the expectation of the function $\yg_{\ya}$ is equal to
 the expectation of the function
-$\yg^*(\yRf) \defd \sum_{\yrs} \yg_{\yrs}\,\ypp_{\yrs\yRf}$.
+$\yg^*(\yA) \defd \sum_{\ya} \yg_{\ya}\,\ypp_{\ya\yA}$.
 
@@
   \label{eq:constraint_extended}
-  \yc = \E(\yg \| \yH) \equiv \sum_{\yRf} \bigl( \tsum_{\yrs} \yg_{\yrs}  \ypp_{\yrs\yRf} \bigr)\;
-  \pf(\yRf \| \yH).
+  \yc = \E(\yg \| \yH) \equiv \sum_{\yA} \bigl( \tsum_{\ya} \yg_{\ya}  \ypp_{\ya\yA} \bigr)\;
+  \pf(\yA \| \yH).
 \end{equation}
@@
 In particular, when the function $\yg$ is the $m$-activity of the sample,
-$\yg_{\yrs} = \sav{\yr\dotso\yr} \equiv \binom{n\yrs}{m}/\binom{n}{m}$, we
+$\yg_{\ya} = \sav{\yaa\dotso\yaa} \equiv \binom{n\ya}{m}/\binom{n}{m}$, we
 find
@@
     % \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
-  \E(\sav{\underbrace{\yr \dotsm \yr}_{\text{$m$ factors}}} \| \yH)
+  \E(\sav{\underbrace{\yaa \dotsm \yaa}_{\text{$m$ factors}}} \| \yH)
 \equiv
@@
     % \frac{(n-m)!}{n!}
-    \sum_{\yrs} %m!\,
+    \sum_{\ya} %m!\,
     \binom{n}{m}^{-1}
-    \binom{n \yrs}{m}\, \pf(\yrs \| \yH)
-    % \sum_{N \yRf=0}^N %m!\,
-    % \binom{N-m}{N \yRf-m} \binom{N}{N \yRf}^{-1} \p(\yXf=\yRf \| \yH)
+    \binom{n \ya}{m}\, \pf(\ya \| \yH)
+    % \sum_{N \yA=0}^N %m!\,
+    % \binom{N-m}{N \yA-m} \binom{N}{N \yA}^{-1} \p(\yXf=\yA \| \yH)
     ={}\\
@@
     % \frac{(N-m)!}{N!}
-    \sum_{\yRf} %m!\,
-    \binom{\yNv \yRf}{m}\, \pf(\yRf \| \yH)
-\equiv    \E(\av{\underbrace{\yR \dotsm \yR}_{\text{$m$ factors}}} \| \yH),
+    \sum_{\yA} %m!\,
+    \binom{\yNv \yA}{m}\, \pf(\yA \| \yH)
+\equiv    \E(\av{\underbrace{\yAA \dotsm \yAA}_{\text{$m$ factors}}} \| \yH),
 \end{multline}
@@
 distribution and can be found in \textcite{potts1953}. Similar relations
-can be found for the raw moments $\E(\yrs^m)$ and $\E(\yRf^m)$, which
+can be found for the raw moments $\E(\ya^m)$ and $\E(\yA^m)$, which
 can be written in terms of the product expectations using
@@
 \caption{Log-density plot of the hypergeometric distribution
-  $%\pf(\yrs \|\yRf, \yH)=
-\ypp_{\yrs\yRf} \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\yrs}\binom{\yNv-n}{\yNv \yRf-n\yrs}\binom{\yNv}{\yNv \yRf}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
+  $%\pf(\ya \|\yA, \yH)=
+\ypp_{\ya\yA} \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
   colourbar depending on your \textsc{pdf} viewer.)}
@@
 formalize and mathematically express our intuition. The hypergeometric
-distribution $\ypp_{\yrs\yRf}$ plays an important role in this
+distribution $\ypp_{\ya\yA}$ plays an important role in this
 formalization. A look at its plot, \fig~\ref{fig:hypergeom_proj}, reveals
 that it is a sort of \enquote{fuzzy identity transformation}, or fuzzy
-Kronecker delta, between the $\yRf$-space $\set{0,\dotsc,\yNv}$ and
-$\yrs$-space $\set{0,\dotsc,n}$. From \eqn~\eqref{eq:marginal} we thus have
+Kronecker delta, between the $\yA$-space $\set{0,\dotsc,\yNv}$ and
+$\ya$-space $\set{0,\dotsc,n}$. From \eqn~\eqref{eq:marginal} we thus have
 that
@@
   \label{eq:roughly_equal_nN}
-  \pf(\yrs=a \|\yH) \approx \pf(\yRf=a \|\yH),\qquad
-\E[\yg_{\yrs} \|\yH] \approx \E[\yg_{\yRf} \|\yH],
+  \pf(\ya=a \|\yH) \approx \pf(\yA=a \|\yH),\qquad
+\E[\yg_{\ya} \|\yH] \approx \E[\yg_{\yA} \|\yH],
 \end{equation}
@@
   sample is representative of our uncertainty about the population and
-  about other samples}, and vice versa. When $n=\yNv$, $\ypp_{\yrs\yRf}$
+  about other samples}, and vice versa. When $n=\yNv$, $\ypp_{\ya\yA}$
 becomes the identity matrix and the approximate equalities above become
@@
 probability distributions. In the next section we will in fact emphasize
-their differences. If the distribution for the population average $\yRf$ is
+their differences. If the distribution for the population average $\yA$ is
 bimodal, for example, the bimodality can be lost in the distribution for
-the sample average $\yrs$, owing to the coarsening effect of
-$\ypp_{\yrs\yRf}$.
+the sample average $\ya$, owing to the coarsening effect of
+$\ypp_{\ya\yA}$.
 
@@
 % hold even in cases where the conditional probability
-% $\pf(\yxs = \yrs \|\yXf=\yRf)$ is not a fuzzy identity, our uncertainties
+% $\pf(\yxs = \ya \|\yXf=\yA)$ is not a fuzzy identity, our uncertainties
 % about sample and about population can differ wildly, and the approximate
@@
 $\yc \defd (\widehat{c}_1,\dotsc,\widehat{c}_m)$. These functions are typically $k$-activities
-$\sav{\yr\dotso \yr}$, and the values are typically the time averages of
+$\sav{\yaa\dotso \yaa}$, and the values are typically the time averages of
 the observed sample, as discussed in \sect~\ref{sec:intro}:
-$\yc = \sum_t \yg[\yrs(t)]/T$.
+$\yc = \sum_t \yg[\ya(t)]/T$.
 
@@
   \label{eq:app_maxent_sample}
-  \pf(\yrs \|\yHb)
+  \pf(\ya \|\yHb)
   =\frac{1}{\yk(\yl)}\,
-  \binom{n}{n\yrs}
-  \exp[\yl\T \yg_{\yrs}]
+  \binom{n}{n\ya}
+  \exp[\yl\T \yg_{\ya}]
 \end{equation}
@@
 exponential appears because we must account for the multiplicity by which
-the population-average activity $\yrs$ can be realized: $\yrs=0$ can be
-realized in only one way (all neurons inactive), $\yrs=1/n$ can be realized
+the population-average activity $\ya$ can be realized: $\ya=0$ can be
+realized in only one way (all neurons inactive), $\ya=1/n$ can be realized
 in $n$ ways (one active neuron out of $n$), and so on. This term is analogous to
@@
   \yc = \E(\yg \|\yHb) \equiv
-  \frac{1}{\yk(\yl)}\sum_{\yrs}  \yg_{\yrs} \binom{n}{n\yrs}
-  \exp[\yl\T \yg_{\yrs}].%, \qquad k=1,\dotsc,m.
+  \frac{1}{\yk(\yl)}\sum_{\ya}  \yg_{\ya} \binom{n}{n\ya}
+  \exp[\yl\T \yg_{\ya}].%, \qquad k=1,\dotsc,m.
 \end{equation}
@@
   \label{eq:app_maxent_pop}
-  \pf(\yRf \| \yHa)  = \frac{1}{\yK(\yL)}\,
-  \binom{\yNv}{\yNv \yRf}\,\exp\bigl(\yL\T
-  \tsum_{\yrs} \yg_{\yrs}\ypp_{\yrs\yRf}\bigr).
+  \pf(\yA \| \yHa)  = \frac{1}{\yK(\yL)}\,
+  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yL\T
+  \tsum_{\ya} \yg_{\ya}\ypp_{\ya\yA}\bigr).
 \end{equation}
@@
   \yc = \E(\yg \|\yHa)\equiv %{}\\
- \frac{1}{\yK(\yL)}\sum_{\yrs} \sum_{\yRf} \yg_{\yrs} \ypp_{\yrs \yRf}
+ \frac{1}{\yK(\yL)}\sum_{\ya} \sum_{\yA} \yg_{\ya} \ypp_{\ya \yA}
 \,
-  \binom{\yNv}{\yNv \yRf}\,\exp\bigl(\yL\T
-  \tsum_{\yrs} \yg_{\yrs}\ypp_{\yrs\yRf}\bigr).
+  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yL\T
+  \tsum_{\ya} \yg_{\ya}\ypp_{\ya\yA}\bigr).
 %  \\ k=1,\dotsc,m.
@@
   \label{eq:app_maxent_pop_marg}
-  \pf(\yrs \| \yHa)  = \frac{1}{\yK(\yL)}\, 
-  \sum_{\yRf} \ypp_{\yrs \yRf}
+  \pf(\ya \| \yHa)  = \frac{1}{\yK(\yL)}\, 
+  \sum_{\yA} \ypp_{\ya \yA}
 \,
-  \binom{\yNv}{\yNv \yRf}\,\exp\bigl(\yL\T
-  \tsum_{\yrs} \yg_{\yrs}\ypp_{\yrs\yRf}\bigr).
+  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yL\T
+  \tsum_{\ya} \yg_{\ya}\ypp_{\ya\yA}\bigr).
 \end{equation}
@@
 \eqref{eq:app_maxent_sample_constraints}, and in addition also the $n$
-equations $\pf(\yrs \| \yHa) = \pf(\yrs \| \yHb)$,
-$\yrs=1/n,\dotsc,1$ (one equation is taken care of by the normalization of
+equations $\pf(\ya \| \yHa) = \pf(\ya \| \yHb)$,
+$\ya=1/n,\dotsc,1$ (one equation is taken care of by the normalization of
 the distributions). We would have a set of $2m+n$ equations in $2m$
@@
   full-population level are inequivalent}. They lead to numerically
-different distributions for the sample activity $\yr$.
+different distributions for the sample activity $\yaa$.
 
@@
 \includegraphics[width=\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
-\caption{Linear and log-plots of $\pf(\yrs)$ constructed by \me\ at
+\caption{Linear and log-plots of $\pf(\ya)$ constructed by \me\ at
   the population level followed by sample marginalization (blue triangles),
@@
 \includegraphics[width=\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}% 
-\caption{Linear and log-plots of $\pf(\yrs)$ constructed by \me\ at
+\caption{Linear and log-plots of $\pf(\ya)$ constructed by \me\ at
   the population level followed by sample marginalization (blue triangles),
@@
 \includegraphics[width=\linewidth]{comparison3.pdf}%
-\caption{Linear and log-plots of $\pf(\yrs)$ for a sample of $n=200$
+\caption{Linear and log-plots of $\pf(\ya)$ for a sample of $n=200$
   and constraints as in \eqn~\eqref{eq:constraints}, constructed by:
@@
   \label{eq:constraints}
-  \E(\yrs) = 0.0478,\qquad
-  \E(\yrrs) = 0.00257,
+  \E(\ya) = 0.0478,\qquad
+  \E(\yaas) = 0.00257,
   % ,\quad
-  % \expe{\sav{\yr\yr\yr}} = 1.48\times 10^{-4},\quad
-  % \expe{\sav{\yr\yr\yr\yr}} = 8.81 \times 10^{-6}.
+  % \expe{\sav{\yaa\yaa\yaa}} = 1.48\times 10^{-4},\quad
+  % \expe{\sav{\yaa\yaa\yaa\yaa}} = 8.81 \times 10^{-6}.
 \end{equation}
@@
   \label{eq:app_maxent_pop_marg_unknown_N}
-  \pf(\yrs \| \yHd)  =
-  \sum_{\yNv} \pf(\yrs \| \yNv,\yHd) \,
+  \pf(\ya \| \yHd)  =
+  \sum_{\yNv} \pf(\ya \| \yNv,\yHd) \,
   \pf(\yNv \|\yHd)
@@
   \sum_\yNv \biggl\{\frac{1}{\yK(\yL_\yNv)}\, 
-  \sum_{\yRf} \ypp^{(\yNv)}_{\yrs\yRf}
+  \sum_{\yA} \ypp^{(\yNv)}_{\ya\yA}
 \,
-  \tbinom{\yNv}{\yNv \yRf}\,\exp\Bigl[{\yL_\yNv}\T
-  \tsum_{\yrs} \yg_{\yrs}\ypp^{(\yNv)}_{\yrs\yRf}\Bigr]\biggr\}
+  \tbinom{\yNv}{\yNv \yA}\,\exp\Bigl[{\yL_\yNv}\T
+  \tsum_{\ya} \yg_{\ya}\ypp^{(\yNv)}_{\ya\yA}\Bigr]\biggr\}
   \;h(\yNv),
@@
 where the Lagrange multipliers $\yL_{\yNv}$ and the summation range for
-$\yRf$ depend on $\yNv$.
+$\yA$ depend on $\yNv$.
 
@@
   %       \\[3\jot]
-  %    \frac{\int \pf(k \| \yl,\yr,\yS) \sum_{\yf{N}} \delt(\tsum \yo\yf{N}\in \yA)\,
-  %     \binom{N}{N\yf{N}}\,\bigl[ \tprod  \pf(\yk \| \yl,\yr,\yS)^{N\yf{N}} \bigr]
+  %    \frac{\int \pf(k \| \yl,\yaa,\yS) \sum_{\yf{N}} \delt(\tsum \yo\yf{N}\in \yA)\,
+  %     \binom{N}{N\yf{N}}\,\bigl[ \tprod  \pf(\yk \| \yl,\yaa,\yS)^{N\yf{N}} \bigr]
   %      \,
   %     \pf(\yl \|\yS) \,\di\yl}{\int\sum_{\yf{N}} \delt(\tsum \yo\yf{N}\in \yA)\,
-  %     \binom{N}{N\yf{N}}\,\bigl[ \tprod \pf(\yk \| \yl,\yr,\yS)^{N\yf{N}} \bigr]
+  %     \binom{N}{N\yf{N}}\,\bigl[ \tprod \pf(\yk \| \yl,\yaa,\yS)^{N\yf{N}} \bigr]
   %      \,
@@
 \ref{item:suff_route} \emph{only holds if $T$ is much larger than the
-  possible values of the activity $\yRf$}. Finally, we also obtain very
+  possible values of the activity $\yA$}. Finally, we also obtain very
 different expressions depending on whether we're asking about \emph{the
@@
 distributions are often very vague about the latter point.
+
+\bigskip
+
+Here are the distributions for our \dobs\ about three different quantities,
+assuming an entropic pre-data distribution for the long-run frequencies of
+the full population:
+
+\textbf{The frequency distribution $\yF$ of the full-population activity during
+  the recording}
+
+\begin{multline}
+  \label{eq:fullfreq_from_samplefreq}
+  \p(\yF \|\yf,\yH) \propto
+  \int\!\di\ynu\;  \tsum_{\yj}
+  \delt(\tsum_{\yA}\yjj_{\ya\yA} = \yff_{\ya})\;
+  \delt(\tsum_{\ya}\yjj_{\ya\yA} = \yFF_{\yA})\times{}\\
+  \binom{T}{T\yj}\;
+  \biggl[\prod_{\ya\yA}
+(\ypp_{\ya\yA}\ynuu_{\yA})^{T\yjj_{\ya\yA}}
+\biggr]\;
+\exp[-L\;\ysh(\ynu;\yR)]
+\\
+{}\appropto
+  \delt(\tsum_{\yA}\ypp_{\ya\yA}\yFF_{\yA} = \yff_{\ya})\;
+\exp[-L\;\ysh(\yF;\yR)]
+\end{multline}
 

### 2019-05-17 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-05-17T16:54:33+0200
+%% Last-Updated: 2019-05-17T19:34:12+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \SetMathAlphabet{\mathsf}{bold}{T1}{\sfdefault}{b}{sl}
-%\newcommand*{\mathte}[1]{\textbf{\textit{\textsf{#1}}}}
+\newcommand*{\mathte}[1]{\textbf{\textit{\textsf{#1}}}}
 % Upright sans-serif math alphabet
@@
 
-\textbf{The frequency distribution $\yF$ of the full-population activity during
-  the recording}
-
-\begin{multline}
+\textbf{The frequency distribution $\yF$ of the full-population activity
+  \emph{during the recording}}
+
+\begin{equation}
   \label{eq:fullfreq_from_samplefreq}
-  \p(\yF \|\yf,\yH) \propto
+  \begin{split}
+  \p(\yF \|\yf,\yH) &\propto
+  \!\begin{multlined}[t][0.8\linewidth]
   \int\!\di\ynu\;  \tsum_{\yj}
@@
 \exp[-L\;\ysh(\ynu;\yR)]
+\end{multlined}
 \\
-{}\appropto
-  \delt(\tsum_{\yA}\ypp_{\ya\yA}\yFF_{\yA} = \yff_{\ya})\;
+&\appropto \delt(\tsum_{\yA}\ypp_{\ya\yA}\yFF_{\yA} = \yff_{\ya})\;
 \exp[-L\;\ysh(\yF;\yR)]
-\end{multline}
+\end{split}
+\end{equation}
+
+\bigskip
+
+\textbf{The long-run frequency distribution $\ynu$ of the full-population
+  activity}
+
+\begin{equation}
+  \label{eq:longrun_from_samplefreq}
+  \begin{split}
+  \p(\ynu \|\yf,\yH) &\propto
+    \binom{T}{T\yf}\;
+  \biggl[\prod_{\ya}
+(\tsum_{\yA}\ypp_{\ya\yA}\ynuu_{\yA})^{T\yff_{\ya}}
+\biggr]\;
+\exp[-L\;\ysh(\ynu;\yR)]
+\\
+&\appropto 
+\exp[-T\;\ysh(\yf; \mathte{\ypp}\ynu)
+-L\;\ysh(\ynu;\yR)]
+\end{split}
+\end{equation}
+Note that if some $\yff_{\ya}$ are zero, then the first exponential may
+be badly approximated by a delta, because the constraints lie on a facet of the
+simplex of frequencies $\set{\ynu}$.
+
+\bigskip
+
+\textbf{The activity $\yA'$ of the full-population
+  in a \emph{new} time bin}
+
+\begin{equation}
+  \label{eq:newbin_from_samplefreq}
+  \begin{split}
+  \p(\yA' \|\yf,\yH) &\propto
+  \int\!\di\ynu\;\ynuu_{\yA'}\;
+  \binom{T}{T\yf}\;
+  \biggl[\prod_{\ya}
+(\tsum_{\yA}\ypp_{\ya\yA}\ynuu_{\yA})^{T\yff_{\ya}}
+\biggr]\;
+\exp[-L\;\ysh(\ynu;\yR)]
+\\
+&\appropto 
+\int\!\di\ynu\;\ynuu_{\yA'}\;
+\exp[-T\;\ysh(\yf; \mathte{\ypp}\ynu)
+-L\;\ysh(\ynu;\yR)]
+\end{split}
+\end{equation}
+Note that if some $\yff_{\ya}$ are zero, then the first exponential may
+be badly approximated by a delta, because the constraints lie on a facet of the
+simplex of frequencies $\set{\ynu}$.
 

### 2019-06-10 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-05-17T19:34:12+0200
+%% Last-Updated: 2019-06-10T16:14:24+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 external stimuli or experimental conditions change? Does such activity
-range over all its theoretically possible values, or only over a
-submanifold thereof?
+range over all its mathematically possible values, or only over a
+subset thereof?
 
 Answering this kind of questions always engages an element of uncertainty.
-Our answers therefore involve experimental data, such as neuronal
-recordings from specific brain areas, and also probabilistic assumptions --
+Our answers therefore involve both experimental data, such as neuronal
+recordings from specific brain areas, and probabilistic assumptions --
 probabilistic \enquote{models} -- about biological conditions and
@@
 experimental measurements of neuronal activity can still only record a very
-small sample of neurons compared to the number that constitutes a brain
-region. Many probabilistic models focus on such samples only, somehow
-neglecting, in their assumptions, that the recorded neurons are a sample
-from a larger network. This kind of isolation assumptions sometimes escape
-attention, being subtly hidden in the mathematics. Some probabilistic
-models try to take also unrecorded neurons into account, but become very
-complex in doing so.
+small sample of neurons compared to the numbers that constitute one
+functionally distinguished brain region. Many probabilistic models focus on
+such samples only, somehow neglecting, in their assumptions, that the
+recorded neurons are a sample from a larger network. This kind of isolation
+assumptions sometimes escape attention, being subtly hidden in the
+mathematics. Some probabilistic models try to take also unrecorded neurons
+into account, but become very complex in doing so.
 
 In the present work we give an answer to this question: What kind of
-inferences about the total activity of a large neuronal network can we make
-by recording a very small sample thereof? Our answer comes from the
-simple combination of basic sampling probability relations and the maximum-entropy
-method.\mynote{shall we give a two-sentence summary of the main idea here?}
+inferences about the \emph{total activity} of a large neuronal network can
+we make by recording a very small sample thereof? The answer we'll give
+comes from the simple combination of basic sampling probability relations
+and the maximum-entropy method.\mynote{shall we give a two-sentence summary
+  of the main idea here?}
 
 The maximum-entropy method\mynote{refs} has been used for a variety of
-inferences about neuronal activity from various brain areas and about other
-phenomena of importance to the neurosciences, for example gene regulation
-and protein sequencing\mynote{what's this exactly? refs}. This method is
-often used to test whether some statistics of the data, for example
-second-order time correlatins, is sufficient for making other inferences
-about the data. It can be considered as an approximation of probabilistic
-models based on various assumptions of inferential
+inferences about the neuronal activity of various brain areas and about
+other phenomena of importance to the neurosciences, for example gene
+regulation and protein sequencing\mynote{what's this exactly? refs}. This
+method is often used to test whether some statistics of the data, for
+example second-order time correlatins, is sufficient for making other
+inferences about the data. It can be considered as an approximation of
+probabilistic models based on various assumptions of inferential
 sufficiency\mynote{refs} -- for example of empirical moments or of long-run
@@
 
+\bigskip
+
+The maximum-relative-entropy distribution is, in the first two cases, an
+approximation of the most probable frequency distribution $\yF$ or $\ynu$;
+in the third case, an approximation of the probability distribution for
+$\yA'$.
+
+\section{Assumptions and limitations}
+\label{sec:assumptions}
+
+Main assumptions behind this belief distribution:
+
+We are approximating our state of knowledge with a finitely exchangeable
+one. In turn, this is numerically approximated by an infinitely
+exchangeable one for $T$ large. But we don't really have an exchangeable
+belief: our \dob\ that the activity at the next time bin will
+differ from the activity at the present one is roughly proportional to the
+difference in the two subsequent activities.
+
+The formulae say that we have equal beliefs about the underlying network
+states having the same activity. This isn't really our belief, for we
+believe there are interactions between the neurons and subnetworks thereof.
+
 \section{Discussion}

### 2019-06-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-10T16:14:24+0200
+%% Last-Updated: 2019-06-11T11:27:39+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The maximum-entropy method\mynote{refs} has been used for a variety of
-inferences about the neuronal activity of various brain areas and about
-other phenomena of importance to the neurosciences, for example gene
-regulation and protein sequencing\mynote{what's this exactly? refs}. This
-method is often used to test whether some statistics of the data, for
-example second-order time correlatins, is sufficient for making other
+The maximum-entropy method \citep{jaynes1957}[much clearer
+in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b,grandy1980}
+has been used for a variety of inferences about the neuronal activity of
+various brain areas and about other phenomena of importance to the
+neurosciences, for example gene and protein interaction \citep[for
+example][]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,mackeetal2009b,tkaciketal2009,roudietal2009c,barreiroetal2010,gerwinnetal2010,mackeetal2011b,ganmoretal2011,cohenetal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}.
+This method is often used to test whether some statistics of the data, for
+example second-order time correlations, is sufficient for making other
 inferences about the data. It can be considered as an approximation of
 probabilistic models based on various assumptions of inferential
-sufficiency\mynote{refs} -- for example of empirical moments or of long-run
-frequencies.
+sufficiency \citep{jaynes1986d_r1996,portamana2017} -- for example of
+empirical moments or of long-run frequencies.
 

### 2019-06-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-11T11:27:39+0200
+%% Last-Updated: 2019-06-11T11:49:47+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\propertitle}{Maximum-entropy distributions for a 
-  neuronal population\\ from subpopulation constraints%\\{\large ***}%
+  neuronal population\\ from subpopulation data%\\{\large ***}%
 }
 \newcommand*{\pdftitle}{\propertitle}
-\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
+\newcommand*{\headtitle}{Maximum-entropy from sample data}
 \newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
@@
 {\protect\centering P.G.L.  Porta Mana\\%
-\footnotesize\epost{\email{piero.mana}{ntnu.no}}}%
+\footnotesize\epost{\email{pgl}{portamana.org}}}%
 \hspace*{\stretch{1}}%
@@
   model for the total activity of a population of neurons, given only some
-  activity statistics -- for example, empirical moments -- of a
+  activity data or statistics -- for example, empirical moments -- of a
   \emph{subpopulation} thereof. This kind of model is useful because
@@
   The model is applied to two sets of neuronal data available in the
-  literature. In some cases it predicts the larger population to have
-  interesting features -- for example, two modes in the probability for the
-  total activity at low regimes -- that are not visible in the sample or in
-  a maximum-entropy model built for the sample alone.
+  literature. In some cases it makes interesting forecasts about the larger
+  population -- for example, two low-regime modes in the frequency
+  distribution for the total activity -- that are not visible in the sample
+  data or in maximum-entropy models applied only to the sample.
 %

### 2019-06-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-11T11:49:47+0200
+%% Last-Updated: 2019-06-11T14:46:37+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In the present work we give an answer to this question: What kind of
-inferences about the \emph{total activity} of a large neuronal network can
-we make by recording a very small sample thereof? The answer we'll give
-comes from the simple combination of basic sampling probability relations
-and the maximum-entropy method.\mynote{shall we give a two-sentence summary
-  of the main idea here?}
+In the present work we give an answer to this question: What inferences
+about the \emph{total activity} of a large neuronal network can we make by
+recording a very small sample thereof? We'll first give an answer by
+combining, in a straightforward way, the maximum-entropy method and basic
+sampling relations of the probability calculus. Later on we'll show that
+this answer can be obtained by exclusively using the probability calculus;
+this point of view will show that the maximum-entropy answer is a first
+approximation of a more complete one, and the assumptions lying behind it.
+
+We apply this method to two concrete data sets:
+
+\mynote{shall we give a
+  two-sentence summary of the main idea here?}
 

### 2019-06-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-11T14:46:37+0200
+%% Last-Updated: 2019-06-11T15:16:29+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 this answer can be obtained by exclusively using the probability calculus;
-this point of view will show that the maximum-entropy answer is a first
-approximation of a more complete one, and the assumptions lying behind it.
-
-We apply this method to two concrete data sets:
+this point of view shows that the maximum-entropy answer is a first
+approximation to the answers of \emph{three} different questions, and also
+shows the assumptions lying behind it.
+
+We apply our approach to two concrete data sets: the activity of 65 neurons
+recorded from a rat's Medial Entorhinal Cortex \mynote{ref}, and the
+activity of 159 neurons recorded from a macaque's Motor Cortex
+\mynote{ref}\mynote{add recording length}. In the first data set, the most
+interesting inference is that the most probable frequency distribution for
+the total activity of the full population, if the latter is assumed to be
+larger than 1\,000 neurons, has two very distinct modes, both at low
+activities ($0.7\%$ and $2.6\%$ of simultaneously active neurons within
+$3\;\mathrm{ms}$, see \fig***). The analogous distribution for second data
+set doesn't have two modes but still presents one prominent shoulder in the
+low-activity mode. Note that these inferred features for the full
+population aren't observed in the sample, nor are they inferred by the
+application of maximum-entropy to the sample alone.
 

### 2019-06-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-11T15:16:29+0200
+%% Last-Updated: 2019-06-11T15:22:46+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \mynote{ref}\mynote{add recording length}. In the first data set, the most
-interesting inference is that the most probable frequency distribution for
-the total activity of the full population, if the latter is assumed to be
-larger than 1\,000 neurons, has two very distinct modes, both at low
+interesting inference is that the most probable time-frequency distribution
+for the total activity of the full population, if the latter is assumed to
+be larger than 1\,000 neurons, has two very distinct modes, both at low
 activities ($0.7\%$ and $2.6\%$ of simultaneously active neurons within
 $3\;\mathrm{ms}$, see \fig***). The analogous distribution for second data
-set doesn't have two modes but still presents one prominent shoulder in the
-low-activity mode. Note that these inferred features for the full
+set doesn't have two modes but still presents one prominent shoulder in its
+one low-activity mode. Note that these inferred features for the full
 population aren't observed in the sample, nor are they inferred by the
-application of maximum-entropy to the sample alone.
+application of maximum-entropy to the sample alone. \mynote{if we manage
+  the Monte Carlo sampling or other approx: say something about the
+  deviation from the most probable frequency distribution}
 

### 2019-06-12 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-11T15:22:46+0200
+%% Last-Updated: 2019-06-12T11:01:28+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-***
-
-Experimental technologies to record the activity of many neurons at the
-same time in different species and brain areas are rapidly advancing. These
-experimental advancements are paralleled by advances in theoretical and
-computational methods for analyzing the data accumulated using the
-recording technologies. Such theoretical methods usually take the form of
-probabilistic models that try to describe the multi-neuronal activity of
-the recorded neurons. With such probabilistic models one aims to address
-numerous issues: What correlations are important in describing the
-multi-neuronal pattern? How does the pattern of activity covary with
-external stimuli or experimental conditions? What dimensionality does the
-neural data live in and how is this related to the underlying network
-interactions? The probabilistic models can also be used to make predictions
-about the structure of the neural code, by studying the properties of the
-fitted model, or by generating synthetic data from it.
-
-In general, despite the rapid advances in recording technology, the best
-experimental measurements of neuronal activity still only provide data from
-a small subset of neurons that comprise a neuronal network. A wealth of
-studies on building probabilistic models of neural data focuses on
-describing such subsets, ignoring the fact that the observed neurons is a
-small group in a much bigger set of hidden neurons. Some other studies do
-include hidden variables which, amongst other things, aim to model the
-global features of the unrecorded neurons, but we still lack a through
-understanding of how to include the role of hidden neurons in probabilistic
-models, how our inferences about the recorded activity would be affected by
-them, and what we can we say about the rest of the network by studying the
-heavily subsampled recordings. In this paper, we aim to address these
-questions in the case of a simple maximum entropy model, namely the
-homogeneous maximum entropy model.
-
-The maximum entropy approach has been used in a variety of setting for
-building statistical models of complex systems and datasets, ranging from
-neuronal activity in the retina, in the cortex, protein sequences, gene
-regulatory networks and natural images. The general idea is, for a dataset,
-to write down the distribution that maximizes the entropy of state
-variables, given some low order statistics. Now given the fact that the
-recorded neurons are a fraction of the neurons in the network, several
-quantitative questions arise that we will address in this paper: given the
-data from the sampled neurons, can we build a maximum entropy model over
-the whole network? Once we build such a network level maximum entropy
-model, can we see features in the neural activity which cannot be directly
-seen from a model build from the sampled neurons? Since we can marginalize
-down the network level maximum entropy model to the sampled network, how
-does this margianzlied maximum entropy model match the sample level model?
-
-All these questions can be answered in the case of the homogeneous maximum
-entropy model. First we show how to go from the sample level maximum
-entropy level to the network level, by assuming different sizes of the
-network and also by assuming an uninformative prior over the size of the
-network. This is done by inferring the statistics of correlation functions
-at the network level from those of the sample level by using simple
-counting arguments. We then find that, when applied to experimental
-recordings from the Medial Entorhinal Cortex of rats and the monkey visual
-cortex, this network level maximum entropy model may exhibit features that
-the sample level model does not predict. Specifically, we observed modes in
-the distribution of the activity in the network level model that do not
-show up in the sample level. We study how the assumed size of the full
-network affects the appearance of these modes and find that there is a
-minimum size of the full network for which such modes can be observed. We
-then compared the distribution found by marginalizing the full network
-maximum entropy model down to the sample level, and the distribution fit
-directly to the sample level. For the two datasets that we tested, we found
-that the two distributions match each other to a large degree but that
-there are also differences between them. We quantify how these differences
-also depend on the assumed size of the network and find that …(WE SHOULD
-TEST SHI) This predicts that for a large enough population (DO WE PREDICT
-THAT IF THE FULL NETWORK GETS BIGGER THE DIFFERENCE ALSO GET BIGGER)?…
-
-The rest of the paper is organized as follows. We first describe how to go
-from the sample level maximum entropy model to the full network maximum
-entropy model. In section 2, we apply this to the two experimental datasets
-and study the effect of the assumed size of the network as well as the
-moments that we use for building the maximum entropy model. In section 3 we
-compare the distributions found from marginalizing the maximum entropy
-model down to the sample level and the original sample level model.
-
-\mynote{Luca: add this: from sampling theory we know that important
+\mynote{Add this?: from sampling theory we know that important
   features of the full network may not be visible in a sample because
@@
 
-\hrule
-\bigskip
+\mynote{To be continued after structure of the rest of the article is
+  clear. Orig intro is on p.~\pageref{orig_intro}}
 
@@
 
+\clearpage
+
+\hrule
+\newcounter{dummy}\refstepcounter{dummy}\label{orig_intro}
+\mynote{orig intro}
+Experimental technologies to record the activity of many neurons at the
+same time in different species and brain areas are rapidly advancing. These
+experimental advancements are paralleled by advances in theoretical and
+computational methods for analyzing the data accumulated using the
+recording technologies. Such theoretical methods usually take the form of
+probabilistic models that try to describe the multi-neuronal activity of
+the recorded neurons. With such probabilistic models one aims to address
+numerous issues: What correlations are important in describing the
+multi-neuronal pattern? How does the pattern of activity covary with
+external stimuli or experimental conditions? What dimensionality does the
+neural data live in and how is this related to the underlying network
+interactions? The probabilistic models can also be used to make predictions
+about the structure of the neural code, by studying the properties of the
+fitted model, or by generating synthetic data from it.
+
+In general, despite the rapid advances in recording technology, the best
+experimental measurements of neuronal activity still only provide data from
+a small subset of neurons that comprise a neuronal network. A wealth of
+studies on building probabilistic models of neural data focuses on
+describing such subsets, ignoring the fact that the observed neurons is a
+small group in a much bigger set of hidden neurons. Some other studies do
+include hidden variables which, amongst other things, aim to model the
+global features of the unrecorded neurons, but we still lack a through
+understanding of how to include the role of hidden neurons in probabilistic
+models, how our inferences about the recorded activity would be affected by
+them, and what we can we say about the rest of the network by studying the
+heavily subsampled recordings. In this paper, we aim to address these
+questions in the case of a simple maximum entropy model, namely the
+homogeneous maximum entropy model.
+
+The maximum entropy approach has been used in a variety of setting for
+building statistical models of complex systems and datasets, ranging from
+neuronal activity in the retina, in the cortex, protein sequences, gene
+regulatory networks and natural images. The general idea is, for a dataset,
+to write down the distribution that maximizes the entropy of state
+variables, given some low order statistics. Now given the fact that the
+recorded neurons are a fraction of the neurons in the network, several
+quantitative questions arise that we will address in this paper: given the
+data from the sampled neurons, can we build a maximum entropy model over
+the whole network? Once we build such a network level maximum entropy
+model, can we see features in the neural activity which cannot be directly
+seen from a model build from the sampled neurons? Since we can marginalize
+down the network level maximum entropy model to the sampled network, how
+does this margianzlied maximum entropy model match the sample level model?
+
+All these questions can be answered in the case of the homogeneous maximum
+entropy model. First we show how to go from the sample level maximum
+entropy level to the network level, by assuming different sizes of the
+network and also by assuming an uninformative prior over the size of the
+network. This is done by inferring the statistics of correlation functions
+at the network level from those of the sample level by using simple
+counting arguments. We then find that, when applied to experimental
+recordings from the Medial Entorhinal Cortex of rats and the monkey visual
+cortex, this network level maximum entropy model may exhibit features that
+the sample level model does not predict. Specifically, we observed modes in
+the distribution of the activity in the network level model that do not
+show up in the sample level. We study how the assumed size of the full
+network affects the appearance of these modes and find that there is a
+minimum size of the full network for which such modes can be observed. We
+then compared the distribution found by marginalizing the full network
+maximum entropy model down to the sample level, and the distribution fit
+directly to the sample level. For the two datasets that we tested, we found
+that the two distributions match each other to a large degree but that
+there are also differences between them. We quantify how these differences
+also depend on the assumed size of the network and find that …(WE SHOULD
+TEST SHI) This predicts that for a large enough population (DO WE PREDICT
+THAT IF THE FULL NETWORK GETS BIGGER THE DIFFERENCE ALSO GET BIGGER)?…
+
+The rest of the paper is organized as follows. We first describe how to go
+from the sample level maximum entropy model to the full network maximum
+entropy model. In section 2, we apply this to the two experimental datasets
+and study the effect of the assumed size of the network as well as the
+moments that we use for building the maximum entropy model. In section 3 we
+compare the distributions found from marginalizing the maximum entropy
+model down to the sample level and the original sample level model.
+
+\hrule
+\bigskip
+
 %%\setlength{\intextsep}{0.5ex}% with wrapfigure

### 2019-06-12 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-12T11:01:28+0200
+%% Last-Updated: 2019-06-12T16:15:50+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 this point of view shows that the maximum-entropy answer is a first
-approximation to the answers of \emph{three} different questions, and also
+approximation to the answers of \emph{four} different questions, and also
 shows the assumptions lying behind it.
@@
   clear. Orig intro is on p.~\pageref{orig_intro}}
+
+\section{Model: maximum-entropy and sampling}
+\label{sec:model}
+
+The idea behind the approach is easily summarized:
+\begin{enumerate*}[label=(\alph*)]
+  \item\label{item:maxent_step} we build a distribution for the total
+  activity of the full population using the maximum-entropy method;
+  \item\label{item:sample_step} the expectation constraints used in the
+  method for the full population are, in turn, determined via sampling
+  theory from the expectation constraints for the sample.
+\end{enumerate*}
+
+Regarding point~\ref{item:maxent_step}, we assume the reader to be familiar
+with the maximum-entropy method. We add two remarks.
+First, the distribution given by the method is the zeroth-order
+approximation of \emph{four} distinct distributions:
+\begin{itemize}
+\item the probability distribution for the value of the activity at any
+  time bin among those recorded,
+\item the probability distribution for the value of the activity at a \emph{new}
+  time bin,
+\item the most probable time-frequency distribution of the activity during the
+  recording,
+\item the most probable time-frequency distribution of the activity in a very
+  long run of new time bins.
+\end{itemize}
+So the generic term \enquote{distribution} means one of these four; note
+that although these distributions numerically coincide in this basic
+approximation, they numerically differ in higher-order approximations.
+Second, the maximum-entropy method based on the Shannon entropy implicitly
+makes specific assumptions about the probability of the long-run frequency
+distributions \citep{jaynes1986d_r1996,portamana2009,portamana2017}.
+
+In our case, to apply the method to the total activity of the full
+population we need to set some of its expectation values, for some of its
+moments for example. Here enters point~\ref{item:sample_step}:
 

### 2019-06-12 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-12T16:15:50+0200
+%% Last-Updated: 2019-06-12T20:30:19+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Regarding point~\ref{item:maxent_step}, we assume the reader to be familiar
-with the maximum-entropy method. We add two remarks.
-First, the distribution given by the method is the zeroth-order
-approximation of \emph{four} distinct distributions:
+with the maximum-entropy method. In short, it amounts to choosing the
+distribution having maximum Shannon entropy among those satisfying specific
+convex constraints, such as expectation values. We add two remarks. First,
+the \enquote{distribution} given by this method is the zeroth-order
+approximation of four different distributions:
 \begin{itemize}
-\item the probability distribution for the value of the activity at any
-  time bin among those recorded,
-\item the probability distribution for the value of the activity at a \emph{new}
-  time bin,
-\item the most probable time-frequency distribution of the activity during the
-  recording,
-\item the most probable time-frequency distribution of the activity in a very
-  long run of new time bins.
+\item the \emph{probability} distribution for the value of the activity at any
+  time bin among those \emph{recorded},
+\item the \emph{probability} distribution for the value of the activity at
+  a \emph{new} time bin,
+\item the most probable \emph{time-frequency} distribution of the activity
+  during the \emph{recording},
+\item the most probable \emph{time-frequency} distribution of the activity
+  in a very long run of \emph{new} time bins.
 \end{itemize}
-So the generic term \enquote{distribution} means one of these four; note
-that although these distributions numerically coincide in this basic
-approximation, they numerically differ in higher-order approximations.
-Second, the maximum-entropy method based on the Shannon entropy implicitly
-makes specific assumptions about the probability of the long-run frequency
-distributions \citep{jaynes1986d_r1996,portamana2009,portamana2017}.
+So the maximum-entropy \enquote{distribution} is an approximate answer to
+four conceptually different inferences. Note that although these four
+distributions numerically coincide in this approximation, they numerically
+differ in higher-order approximations. Second, the maximum-entropy method
+based on the Shannon entropy implicitly makes specific assumptions about
+the probability of the long-run frequency distributions
+\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
+assumptions in \sect***
 
 In our case, to apply the method to the total activity of the full
-population we need to set some of its expectation values, for some of its
-moments for example. Here enters point~\ref{item:sample_step}:
+population we need to set some of its expectation values, for example its
+moments. Here enters point~\ref{item:sample_step}:
 

### 2019-06-13 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-12T20:30:19+0200
+%% Last-Updated: 2019-06-13T11:51:25+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Regarding point~\ref{item:maxent_step}, we assume the reader to be familiar
-with the maximum-entropy method. In short, it amounts to choosing the
+Regarding point~\ref{item:maxent_step}, we assume you to be familiar with
+the maximum-entropy method. In short, it amounts to choosing the
 distribution having maximum Shannon entropy among those satisfying specific
-convex constraints, such as expectation values. We add two remarks. First,
-the \enquote{distribution} given by this method is the zeroth-order
-approximation of four different distributions:
+convex constraints, such as expectation values; these expectations are
+typically chosen to be equal to measured averages. We add two remarks that
+you may not be familiar with. First, the \enquote{distribution} given by
+this method is the zeroth-order approximation of four different
+distributions:
 \begin{itemize}
@@
 \end{itemize}
-So the maximum-entropy \enquote{distribution} is an approximate answer to
-four conceptually different inferences. Note that although these four
-distributions numerically coincide in this approximation, they numerically
-differ in higher-order approximations. Second, the maximum-entropy method
-based on the Shannon entropy implicitly makes specific assumptions about
-the probability of the long-run frequency distributions
-\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
-assumptions in \sect***
+The maximum-entropy distribution is thus an approximate answer to four
+conceptually different inferences. Note that although the four
+distributions above numerically coincide in this approximation, they
+numerically differ in higher-order approximations. Second, the
+maximum-entropy method based on the Shannon entropy implicitly makes
+specific assumptions about the probability of the long-run frequency
+distributions \citep{jaynes1986d_r1996,portamana2009,portamana2017}. We
+discuss these assumptions in \sect***
 
 In our case, to apply the method to the total activity of the full
-population we need to set some of its expectation values, for example its
-moments. Here enters point~\ref{item:sample_step}:
+population we need to fix some of its expectation values, for example its
+moments. But we don't have any measured averages for the full population to
+equate our expectations to. Here enters point~\ref{item:sample_step}:
+classical sampling theory
+\citep[for example][\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
+also][]{whitworth1897}.
 

### 2019-06-13 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-13T11:51:25+0200
+%% Last-Updated: 2019-06-13T14:42:35+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Regarding point~\ref{item:maxent_step}, we assume you to be familiar with
-the maximum-entropy method. In short, it amounts to choosing the
-distribution having maximum Shannon entropy among those satisfying specific
-convex constraints, such as expectation values; these expectations are
-typically chosen to be equal to measured averages. We add two remarks that
-you may not be familiar with. First, the \enquote{distribution} given by
-this method is the zeroth-order approximation of four different
-distributions:
+the maximum-entropy method. In short, it amounts to a pair of
+prescriptions: choose the distribution having maximum Shannon entropy among
+those satisfying specific convex constraints, such as expectation values;
+and judge these expectations to be equal to measured averages. We add two
+remarks that you may not be familiar with. First, the
+\enquote{distribution} given by this method is the zeroth-order
+approximation of four different distributions:
 \begin{itemize}
@@
 moments. But we don't have any measured averages for the full population to
-equate our expectations to. Here enters point~\ref{item:sample_step}:
-classical sampling theory
-\citep[for example][\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
-also][]{whitworth1897}.
+equate our expectations to. Here enters point~\ref{item:sample_step}: the
+probability calculus gives an exact relation between the first $m$ moments
+for the full population and the first $m$ for the sample, so that the ones
+determine the others and vice versa at every time bin; this is a classical
+result of sampling theory \citep[for
+example][\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
+also][]{whitworth1897}. Combining this result with the maximum-entropy
+prescription \enquote{expectations${}={}$measured averages} for the sample,
+we have that the latter averages determine the moments for the full
+population:
+\begin{equation}
+  \label{eq:chain_prescription_moments}
+  \text{measured averages} \longrightarrow
+  \text{sample moments} \longrightarrow
+  \text{full-population moments}
+\end{equation}
+
+This step is even more straightforward if instead of power moments we use
+\emph{normalized factorial moments}. The $m$th normalized factorial moment
+is defined as the expectation
+\begin{equation}
+  \E\condt[\Bigg]{
+    \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
+  \equiv
+  \sum_{\ya} %m!\,
+  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pf(\ya \| \yH),
+\end{equation}
+where $\ya\in\set{0, 1/n, 2/n, \dotsc, (n-1)/n, 1}$ is the normalized total
+activity of a population of $n$ neurons. The $m$th factorial moment can be
+interpreted as the expectation of the number of distinct $m$-tuples of
+simultaneously active neurons, normalized by the maximum possible number of
+distinct $m$-tuples.
 

### 2019-06-13 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-13T14:42:35+0200
+%% Last-Updated: 2019-06-13T15:02:32+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 probability calculus gives an exact relation between the first $m$ moments
-for the full population and the first $m$ for the sample, so that the ones
-determine the others and vice versa at every time bin; this is a classical
+for the full population and the first $m$ for the sample; the ones
+determine the others and vice versa at every time bin. This is a classical
 result of sampling theory \citep[for
@@
 prescription \enquote{expectations${}={}$measured averages} for the sample,
-we have that the latter averages determine the moments for the full
-population:
-\begin{equation}
+we have that the measured averages for the sample determine the moments for
+the full population:
+\begin{equation*}
   \label{eq:chain_prescription_moments}
-  \text{measured averages} \longrightarrow
-  \text{sample moments} \longrightarrow
-  \text{full-population moments}
-\end{equation}
+  \mathrlap{\overbracket{\phantom{\text{measured averages} \rightarrow
+        \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
+  \text{measured averages} \rightarrow
+        \underbracket{\text{sample moments}
+  \rightarrow
+    \text{full-population moments}}_{\textit{sampling theory}}
+\end{equation*}
 
@@
 where $\ya\in\set{0, 1/n, 2/n, \dotsc, (n-1)/n, 1}$ is the normalized total
-activity of a population of $n$ neurons. The $m$th factorial moment can be
-interpreted as the expectation of the number of distinct $m$-tuples of
-simultaneously active neurons, normalized by the maximum possible number of
-distinct $m$-tuples.
+activity of a population of $n$ neurons. This moment can be interpreted as
+the expectation of the number of distinct $m$-tuples of simultaneously
+active neurons, normalized by the maximum possible number of distinct
+$m$-tuples.
 

### 2019-06-13 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-13T15:02:32+0200
+%% Last-Updated: 2019-06-13T16:07:01+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 where $\ya\in\set{0, 1/n, 2/n, \dotsc, (n-1)/n, 1}$ is the normalized total
-activity of a population of $n$ neurons. This moment can be interpreted as
-the expectation of the number of distinct $m$-tuples of simultaneously
-active neurons, normalized by the maximum possible number of distinct
-$m$-tuples.
+activity of a population of $n$ neurons, and $\pf(\ya \| \yH)$ is the
+probability for activity $\ya$ given assumptions $\yH$. This moment can be
+interpreted as the expectation of the number of distinct $m$-tuples of
+simultaneously active neurons, normalized by the maximum possible number of
+distinct $m$-tuples. Note that the first $m$ factorial moments determine
+the first $m$ moments and vice versa, so fixing the ones is equivalent to
+fixing the others. Now assume that these $n$ neurons are a sample from a
+full population of $N$ neurons, with possible normalized activity
+$\yA \in\set{0,1/N,\dotsc,1}$. We have this extremely convenient property:
+\emph{The first $n-1$ normalized factorial moments for the sample and for
+  the full population are numerically identical}:
+\begin{equation}
+  \label{eq:equal_m_factorialmoments}
+  \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
+  \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
+  \qquad m < n.
+\end{equation}
+
+Hence we can apply the maximum-entropy method to obtain a distribution for
+the full population by constraining, for any $m$s we please, its $m$th
+factorial moment to be equal to the sample's recorded average of
+$\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$. If $\ya_{t}$ is the normalized
+total activity of the sample during the time bin $t$, during a recording
+with $T$ time bins, the constraint on the distribution $\pf(\yA \| \yH)$
+for the full population is
+\begin{equation}
+  \label{eq:final_constraint}
+  \frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
+  =
+   \sum_{\yA}
+  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \pf(\yA \| \yH)
+\end{equation}
+for any $m$s we wish.
 

### 2019-06-14 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-13T16:07:01+0200
+%% Last-Updated: 2019-06-14T14:06:05+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The idea behind the approach is easily summarized:
-\begin{enumerate*}[label=(\alph*)]
+The idea behind the approach presented here is easily summarized:
+\begin{enumerate}[label=(\alph*)]
   \item\label{item:maxent_step} we build a distribution for the total
@@
   theory from the expectation constraints for the sample.
-\end{enumerate*}
+\end{enumerate}
+Let's discuss these points in more detail.
 
@@
 those satisfying specific convex constraints, such as expectation values;
-and judge these expectations to be equal to measured averages. We add two
-remarks that you may not be familiar with. First, the
+and judge these expectation values to be equal to measured averages. We add
+two remarks that are seldom made in the literature. First, the
 \enquote{distribution} given by this method is the zeroth-order
@@
 equate our expectations to. Here enters point~\ref{item:sample_step}: the
-probability calculus gives an exact relation between the first $m$ moments
-for the full population and the first $m$ for the sample; the ones
-determine the others and vice versa at every time bin. This is a classical
-result of sampling theory \citep[for
-example][\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
-also][]{whitworth1897}. Combining this result with the maximum-entropy
-prescription \enquote{expectations${}={}$measured averages} for the sample,
-we have that the measured averages for the sample determine the moments for
-the full population:
+probability calculus gives an exact, linear relation between the first $m$
+moments for the full population and the first $m$ for the sample; the ones
+determine the others and vice versa at every time bin \citep[for
+example][\eqns~(16)]{portamanaetal2015}. This is a classical result of
+sampling theory \citep[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
+also][]{whitworth1897}.
+
+Combining this result with the maximum-entropy prescription
+\enquote{expectations${}={}$measured averages} for the sample, we have that
+the measured averages for the sample determine the moments for the full
+population:
 \begin{equation*}
@@
 
-This step is even more straightforward if instead of power moments we use
-\emph{normalized factorial moments}. The $m$th normalized factorial moment
-is defined as the expectation
+These two steps are even more straightforward if instead of power moments
+we use \emph{normalized factorial moments}. The $m$th normalized factorial
+moment is defined as the expectation
 \begin{equation}
@@
 simultaneously active neurons, normalized by the maximum possible number of
-distinct $m$-tuples. Note that the first $m$ factorial moments determine
-the first $m$ moments and vice versa, so fixing the ones is equivalent to
-fixing the others. Now assume that these $n$ neurons are a sample from a
-full population of $N$ neurons, with possible normalized activity
-$\yA \in\set{0,1/N,\dotsc,1}$. We have this extremely convenient property:
-\emph{The first $n-1$ normalized factorial moments for the sample and for
-  the full population are numerically identical}:
+distinct $m$-tuples. For example, with $m=2$, if $na=4$ neurons are active
+in a population of $n=5$, we have $\tbinom{4}{2}=6$ distinct pairs of
+simultaneously active neurons, and if all five neurons were active we would
+have a maximum of $\tbinom{5}{2}=10$ distinct pairs; the normalized number
+of pairs is therefore $6/10$. Note that the first $m$ factorial moments
+linearly determine the first $m$ moments and vice versa, so fixing the ones is
+equivalent to fixing the others. Now assume that these $n$ neurons are a
+sample from a full population of $N$ neurons, with possible normalized
+activity $\yA \in\set{0,1/N,\dotsc,1}$. We have this extremely convenient
+property: \emph{The first $n-1$ normalized factorial moments for the sample
+  and for the full population are numerically identical}:
 \begin{equation}
@@
   \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
-  \qquad m < n.
-\end{equation}
-
-Hence we can apply the maximum-entropy method to obtain a distribution for
-the full population by constraining, for any $m$s we please, its $m$th
-factorial moment to be equal to the sample's recorded average of
-$\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$. If $\ya_{t}$ is the normalized
-total activity of the sample during the time bin $t$, during a recording
-with $T$ time bins, the constraint on the distribution $\pf(\yA \| \yH)$
-for the full population is
+  \qquad m \le n.
+\end{equation}
+
+We can hence apply the maximum-entropy method to obtain a distribution for
+the full population by constraining its $m$th factorial moment to be equal
+to the sample's recorded average of $\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$,
+for as many $m$ as we please with $1\le m \le n$. If $\ya_{t}$ is the
+normalized total activity of the sample during the time bin $t$, during a
+recording with $T$ time bins, the constraint on the distribution
+$\pf(\yA \| \yH)$ for the full population is
 \begin{equation}

### 2019-06-14 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-14T14:06:05+0200
+%% Last-Updated: 2019-06-14T15:23:13+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+Let's introduce the context and notation for our problem.
+
+We have recorded the activities of $n$ neurons for a given amount of time.
+These activities are binned into $T$ time bins and binarized to $\set{0,1}$
+values. Denote by $\ya_{t}$ the number of active neurons during time bin
+$t$, divided by the total number of neurons $n$; we call this the
+\emph{normalized total activity} or just \enquote{total activity} for
+short. Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$: if $\ya_{t}=0$,
+all $n$ neurons were silent during time bin $t$; if $\ya_{t}=1$, all were
+active.
+
+\mynote{maybe move this paragraph to the intro?}Present-day technologies
+allow the recording of neuronal activity from small brain regions
+\mynote{be more specific}. The specific neurons that happen to be recorded
+from a region are not and cannot yet be specifically chosen, however, owing
+to several factors -- for example, limitations in the precision with which
+electrodes are inserted, or in the targeting of neurons by viruses, and so
+on. In fact, the set of recorded neurons may vary during very long
+recordings or across experimental trials. We assume that there's a small
+region, comprising a population of $N$ neurons, for which other sets of
+neurons had equal chances of being \enquote{selected} as the sample of $n$
+neurons that have actually been recorded. During the recording, the
+normalized total activity of these $N$ neurons at time bin $t$ was
+$\yA_{t}$; note that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
+$N\yA_{t} \ge n\ya_{t}$ for obvious reasons.
+
+Our questions concern general features of the total activity $\yA$ of the
+full population; for example: what was its time-frequency distribution? How
+much activity should we expect in a time bin during the recording, or in a
+new time bin? The answers to these questions are probabilistic and come in
+the form of distributions.
+
 The idea behind the approach presented here is easily summarized:
@@
 activity $\yA \in\set{0,1/N,\dotsc,1}$. We have this extremely convenient
-property: \emph{The first $n-1$ normalized factorial moments for the sample
+property: \emph{The first $n$ normalized factorial moments for the sample
   and for the full population are numerically identical}:
@@
 for as many $m$ as we please with $1\le m \le n$. If $\ya_{t}$ is the
-normalized total activity of the sample during the time bin $t$, during a
+normalized total activity of the sample during the time bin $t$, and the
 recording with $T$ time bins, the constraint on the distribution

### 2019-06-16 maxNt_v5.tex
update

--- 2019-06-14 maxNt_v5.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-14T14:06:05+0200
+%% Last-Updated: 2019-06-16T12:46:55+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+Let's introduce some context and notation for our inference problem.
+
+The context we consider is as follows. We have recorded the spiking
+activities of $n$ neurons for a given amount of time. These spikes are
+binned into $T$ time bins and binarized to $\set{0,1}$ values in each bin.
+Denote by $\ya_{t}$ the number of neurons that fire during time bin $t$,
+divided by the total number of neurons $n$; we call this the
+\emph{normalized total activity} or just \enquote{total activity} for
+short. Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$,
+no neuron spiked during time bin $t$; if $\ya_{t}=1$, all spiked at some
+point during time bin $t$. For brevity, let's say \enquote{at $t$} for
+\enquote{during time bin $t$}.
+
+\mynote{maybe move this paragraph to the intro?}Present-day technologies
+enable the recording of neuronal activity from small brain regions
+\mynote{be more specific}. The neurons that are recorded within the region
+are not, and cannot yet be, specifically chosen from among the rest, owing
+to several limiting factors -- for example, limitations in the precision
+with which electrodes are inserted, or with which neurons are targeted by
+viruses, and so on. In fact, the set of recorded neurons may even vary
+during very long recordings or across experimental trials. We assume that
+there's a small region, comprising a population of $N$ neurons, from which
+other sets of neurons had the same probability of being recorded as the set
+of $n$ neurons that have actually been recorded. During the recording, the
+normalized total activity of these $N$ neurons at $t$ was $\yA_{t}$. We
+don't know the values $\yA_{t}$; we only know for certain that
+$\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
+for obvious reasons. For the moment we assume that we know $N$; we'll
+discuss later the consequences of our lack of precise knowledge about $N$.
+
+Our questions concern general features of the total activity $\yA$ of the
+full population; for example: what was its time-frequency distribution? How
+much activity should we expect in a time bin during the recording, or in a
+new time bin? The answers to these questions are probabilistic and come in
+the form of distributions.
+
 The idea behind the approach presented here is easily summarized:

### 2019-06-16 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-16T12:46:55+0200
+%% Last-Updated: 2019-06-16T21:39:12+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 divided by the total number of neurons $n$; we call this the
-\emph{normalized total activity} or just \enquote{total activity} for
-short. Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$,
-no neuron spiked during time bin $t$; if $\ya_{t}=1$, all spiked at some
-point during time bin $t$. For brevity, let's say \enquote{at $t$} for
-\enquote{during time bin $t$}.
+\emph{normalized total activity} or just \enquote{total activity} or
+\enquote{activity} for short. Obviously
+$\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no neuron spiked
+during time bin $t$; if $\ya_{t}=1$, all spiked at some point during time
+bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during time
+  bin $t$}.
 
@@
 with which electrodes are inserted, or with which neurons are targeted by
-viruses, and so on. In fact, the set of recorded neurons may even vary
+viruses, and so on. In fact, the set of recorded neurons may even change
 during very long recordings or across experimental trials. We assume that
-there's a small region, comprising a population of $N$ neurons, from which
-other sets of neurons had the same probability of being recorded as the set
-of $n$ neurons that have actually been recorded. During the recording, the
-normalized total activity of these $N$ neurons at $t$ was $\yA_{t}$. We
-don't know the values $\yA_{t}$; we only know for certain that
+there's a region, comprising a population of $N$ neurons, from which other
+sets of neurons had the same probability of being recorded as the set of
+$n$ neurons that have actually been recorded. We call this the
+\enquote{full population}. During the recording, the normalized total
+activity of these $N$ neurons at $t$ was $\yA_{t}$. We don't know the
+values $\yA_{t}$ at each $t$; we only know for certain that
 $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
-for obvious reasons. For the moment we assume that we know $N$; we'll
-discuss later the consequences of our lack of precise knowledge about $N$.
+for obvious reasons. For the moment we assume that we know $N$; in \sect***
+we discuss the consequences of our lack of precise knowledge about this
+number.
 
 Our questions concern general features of the total activity $\yA$ of the
-full population; for example: what was its time-frequency distribution? How
-much activity should we expect in a time bin during the recording, or in a
-new time bin? The answers to these questions are probabilistic and come in
-the form of distributions.
-
-The idea behind the approach presented here is easily summarized:
+full population during and after the recording. For example: what was its
+frequency distribution in time? How much total activity should we expect at
+a time bin during the recording, or at a new time bin? We cannot answer
+these questions with certainty; we can only give a distribution of
+probability over the possible answers. Our approach gives such a
+probability distribution.
+
+\bigskip
+
+The idea behind the approach is easily summarized:
 \begin{enumerate}[label=(\alph*)]
@@
   \item\label{item:sample_step} the expectation constraints used in the
-  method for the full population are, in turn, determined via sampling
-  theory from the expectation constraints for the sample.
+  maximum-entropy method for the full population are, in turn, determined
+  via sampling theory from the expectation constraints for the sample.
 \end{enumerate}
-Let's discuss these points in more detail.
-
-Regarding point~\ref{item:maxent_step}, we assume you to be familiar with
-the maximum-entropy method. In short, it amounts to a pair of
-prescriptions: choose the distribution having maximum Shannon entropy among
-those satisfying specific convex constraints, such as expectation values;
-and judge these expectation values to be equal to measured averages. We add
-two remarks that are seldom made in the literature. First, the
-\enquote{distribution} given by this method is the zeroth-order
+Let's discuss these points in detail.
+
+Regarding~\ref{item:maxent_step}, we assume you to be familiar with the
+maximum-entropy method. In short, it amounts to a pair of prescriptions:
+choose the distribution having maximum Shannon entropy among those
+satisfying specific convex constraints, such as fixed expectation values; and
+judge these expectation values to be equal to measured averages. We add two
+remarks about this method that are seldom made in the literature. First,
+the \enquote{distribution} given by this method is the zeroth-order
 approximation of four different distributions:
 \begin{itemize}
-\item the \emph{probability} distribution for the value of the activity at any
-  time bin among those \emph{recorded},
-\item the \emph{probability} distribution for the value of the activity at
-  a \emph{new} time bin,
-\item the most probable \emph{time-frequency} distribution of the activity
-  during the \emph{recording},
-\item the most probable \emph{time-frequency} distribution of the activity
-  in a very long run of \emph{new} time bins.
+\item the \emph{probability} distribution for the value of the total
+  activity at any time bin among those \emph{recorded},
+\item the \emph{probability} distribution for the value of the total
+  activity at a \emph{new} time bin,
+\item the most probable \emph{time-frequency} distribution of the total
+  activity across the \emph{recorded} bins,
+\item the most probable \emph{time-frequency} distribution of the total
+  activity in a very long run of \emph{new} time bins.
 \end{itemize}
@@
 conceptually different inferences. Note that although the four
-distributions above numerically coincide in this approximation, they
-numerically differ in higher-order approximations. Second, the
-maximum-entropy method based on the Shannon entropy implicitly makes
-specific assumptions about the probability of the long-run frequency
-distributions \citep{jaynes1986d_r1996,portamana2009,portamana2017}. We
-discuss these assumptions in \sect***
-
-In our case, to apply the method to the total activity of the full
-population we need to fix some of its expectation values, for example its
-moments. But we don't have any measured averages for the full population to
-equate our expectations to. Here enters point~\ref{item:sample_step}: the
-probability calculus gives an exact, linear relation between the first $m$
-moments for the full population and the first $m$ for the sample; the ones
-determine the others and vice versa at every time bin \citep[for
-example][\eqns~(16)]{portamanaetal2015}. This is a classical result of
-sampling theory \citep[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
+distributions above numerically coincide in this approximation, they differ
+in higher-order approximations. Second, the maximum-entropy method based on
+the Shannon entropy implicitly makes specific assumptions about the
+probability of the long-run frequency distributions
+\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
+assumptions in \sect***
+
+In our case, to apply the method for the total activity of the full
+population we need to fix some of the latter's expectation values, for
+example its moments. But we don't have any measured averages for the full
+population to equate our expectations to. Here enters
+point~\ref{item:sample_step}: the probability calculus gives an exact,
+linear relation between the first $m$ moments for the full population and
+the first $m$ for the sample; the ones determine the others and vice versa
+at every time bin \citep[for example][\eqns~(16)]{portamanaetal2015}. This
+relation is a classical result of sampling theory
+\citep[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
 also][]{whitworth1897}.
@@
 
-These two steps are even more straightforward if instead of power moments
-we use \emph{normalized factorial moments}. The $m$th normalized factorial
-moment is defined as the expectation
+These two steps are more straightforward if instead of power moments we use
+\emph{normalized factorial moments}. The $m$th normalized factorial moment
+of the probability distribution for the activity of the recorded neurons is
+defined as the expectation
 \begin{equation}
@@
   \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pf(\ya \| \yH),
-\end{equation}
-where $\ya\in\set{0, 1/n, 2/n, \dotsc, (n-1)/n, 1}$ is the normalized total
-activity of a population of $n$ neurons, and $\pf(\ya \| \yH)$ is the
-probability for activity $\ya$ given assumptions $\yH$. This moment can be
-interpreted as the expectation of the number of distinct $m$-tuples of
-simultaneously active neurons, normalized by the maximum possible number of
-distinct $m$-tuples. For example, with $m=2$, if $na=4$ neurons are active
-in a population of $n=5$, we have $\tbinom{4}{2}=6$ distinct pairs of
-simultaneously active neurons, and if all five neurons were active we would
-have a maximum of $\tbinom{5}{2}=10$ distinct pairs; the normalized number
+  \qquad 1\le m \le n
+\end{equation}
+where $\pf(\ya \| \yH)$ is the probability for activity $\ya$ given
+assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
+new bin. This moment can be interpreted as the expectation of the number of
+distinct $m$-tuples of simultaneously spiking neurons (within a bin's time
+width), normalized by the number of distinct $m$-tuples. For example, with
+$m=2$, if $na=4$ neurons spike in a population of $n=5$, we have
+$\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons, and the
+total number of distinct pairs is $\tbinom{5}{2}=10$. The normalized number
 of pairs is therefore $6/10$. Note that the first $m$ factorial moments
-linearly determine the first $m$ moments and vice versa, so fixing the ones is
-equivalent to fixing the others. Now assume that these $n$ neurons are a
-sample from a full population of $N$ neurons, with possible normalized
-activity $\yA \in\set{0,1/N,\dotsc,1}$. We have this extremely convenient
-property: \emph{The first $n$ normalized factorial moments for the sample
-  and for the full population are numerically identical}:
+linearly determine the first $m$ moments and vice versa, so fixing the ones
+is equivalent to fixing the others. Now, since the recorded neurons are a
+sample from the full population of $N$ neurons, we have this extremely
+convenient property: \emph{The first $n$ normalized factorial moments for
+  the sample and for the full population are numerically identical}:
 \begin{equation}
@@
 
-We can hence apply the maximum-entropy method to obtain a distribution for
-the full population by constraining its $m$th factorial moment to be equal
-to the sample's recorded average of $\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$,
-for as many $m$ as we please with $1\le m \le n$. If $\ya_{t}$ is the
-normalized total activity of the sample during the time bin $t$, and the
-recording with $T$ time bins, the constraint on the distribution
-$\pf(\yA \| \yH)$ for the full population is
+We can therefore apply the maximum-entropy method to obtain a distribution
+for the full population, by constraining its $m$th factorial moment to be
+equal to the sample's recorded average of
+$\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$, for as many $m$ as we please with
+$1\le m \le n$. The constraints on the distribution $\pf(\yA \| \yH)$ for
+the full population are
 \begin{equation}
@@
 for any $m$s we wish.
+
+\bigskip
+
+In the next section we apply this method to the data sets from two actual
+recordings, and discuss the properties of the resulting distributions.
+
+\section{Application: two data sets}
+\label{sec:application}
 

### 2019-06-17 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-16T21:39:12+0200
+%% Last-Updated: 2019-06-17T10:37:37+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\propertitle}{Maximum-entropy distributions for a 
-  neuronal population\\ from subpopulation data%\\{\large ***}%
+  neuronal network\\ from subnetwork data%\\{\large ***}%
 }
@@
   \noindent This work shows how to build a maximum-entropy probabilistic
-  model for the total activity of a population of neurons, given only some
+  model for the total activity of a network of neurons, given only some
   activity data or statistics -- for example, empirical moments -- of a
-  \emph{subpopulation} thereof. This kind of model is useful because
+  \emph{subnetwork} thereof. This kind of model is useful because
   neuronal recordings are always limited to a very small sample of a
-  population of neurons.
+  network of neurons.
 %
@@
   literature. In some cases it makes interesting forecasts about the larger
-  population -- for example, two low-regime modes in the frequency
+  network -- for example, two low-regime modes in the frequency
   distribution for the total activity -- that are not visible in the sample
@@
   For the two datasets, the maximum-entropy probability model applied only
-  to the subpopulation is compared with the marginal probability
+  to the subnetwork is compared with the marginal probability
   distribution obtained from the maximum-entropy model applied to the full
-  population. On a linear probability scale no large differences are
+  network. On a linear probability scale no large differences are
   visible, but on a logarithmic scale the two distributions show very
@@
 interesting inference is that the most probable time-frequency distribution
-for the total activity of the full population, if the latter is assumed to
+for the total activity of the full network, if the latter is assumed to
 be larger than 1\,000 neurons, has two very distinct modes, both at low
@@
 one low-activity mode. Note that these inferred features for the full
-population aren't observed in the sample, nor are they inferred by the
+network aren't observed in the sample, nor are they inferred by the
 application of maximum-entropy to the sample alone. \mynote{if we manage
@@
 during very long recordings or across experimental trials. We assume that
-there's a region, comprising a population of $N$ neurons, from which other
+there's a region, comprising a network of $N$ neurons, from which other
 sets of neurons had the same probability of being recorded as the set of
 $n$ neurons that have actually been recorded. We call this the
-\enquote{full population}. During the recording, the normalized total
+\enquote{full network}. During the recording, the normalized total
 activity of these $N$ neurons at $t$ was $\yA_{t}$. We don't know the
@@
 Our questions concern general features of the total activity $\yA$ of the
-full population during and after the recording. For example: what was its
+full network during and after the recording. For example: what was its
 frequency distribution in time? How much total activity should we expect at
@@
   \item\label{item:maxent_step} we build a distribution for the total
-  activity of the full population using the maximum-entropy method;
+  activity of the full network using the maximum-entropy method;
   \item\label{item:sample_step} the expectation constraints used in the
-  maximum-entropy method for the full population are, in turn, determined
+  maximum-entropy method for the full network are, in turn, determined
   via sampling theory from the expectation constraints for the sample.
@@
 In our case, to apply the method for the total activity of the full
-population we need to fix some of the latter's expectation values, for
+network we need to fix some of the latter's expectation values, for
 example its moments. But we don't have any measured averages for the full
-population to equate our expectations to. Here enters
+network to equate our expectations to. Here enters
 point~\ref{item:sample_step}: the probability calculus gives an exact,
-linear relation between the first $m$ moments for the full population and
+linear relation between the first $m$ moments for the full network and
 the first $m$ for the sample; the ones determine the others and vice versa
@@
 the measured averages for the sample determine the moments for the full
-population:
+network:
 \begin{equation*}
@@
   \rightarrow
-    \text{full-population moments}}_{\textit{sampling theory}}
+    \text{full-network moments}}_{\textit{sampling theory}}
 \end{equation*}
@@
 width), normalized by the number of distinct $m$-tuples. For example, with
-$m=2$, if $na=4$ neurons spike in a population of $n=5$, we have
+$m=2$, if $na=4$ neurons spike in a network of $n=5$, we have
 $\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons, and the
@@
 is equivalent to fixing the others. Now, since the recorded neurons are a
-sample from the full population of $N$ neurons, we have this extremely
+sample from the full network of $N$ neurons, we have this extremely
 convenient property: \emph{The first $n$ normalized factorial moments for
-  the sample and for the full population are numerically identical}:
+  the sample and for the full network are numerically identical}:
 \begin{equation}
@@
 We can therefore apply the maximum-entropy method to obtain a distribution
-for the full population, by constraining its $m$th factorial moment to be
+for the full network, by constraining its $m$th factorial moment to be
 equal to the sample's recorded average of
@@
 $1\le m \le n$. The constraints on the distribution $\pf(\yA \| \yH)$ for
-the full population are
+the full network are
 \begin{equation}
@@
   with $\yaa(t) \defd \bigl(\yav_1(t),\dotsc,\yav_n(t)\bigr)$. The
-  population-averaged activity at that bin is
+  network-averaged activity at that bin is
   $\ya(t) \defd \sum_i \yav_i(t)/n$. If we count the number of distinct
@@
   most $\tbinom{n}{2}$ simultaneously active pairs, so the
-  population-averaged pair activity is
+  network-averaged pair activity is
   $\av{\yaa \yaa}(t) \defd \tbinom{n}{2}^{-1}\tbinom{n\ya(t)}{2}$. With some
-  combinatorics we see that the population-averaged activity of $m$-tuples
+  combinatorics we see that the network-averaged activity of $m$-tuples
   of neurons is
@@
 into account the activities or pair activities of specific neurons rather
-than population averages, or even protocols that take into account time
+than network averages, or even protocols that take into account time
 dependence.
@@
 that the recorded neurons are a \emph{sample} from a larger, unrecorded
-population, but the \me\ method as applied above is treating them as
+network, but the \me\ method as applied above is treating them as
 isolated from the rest of the brain. Hence, the results it provides cannot
@@
 some square millimetres \citep[\cf][]{berenyietal2014}. This recording is a
-sample of the activity of the neuronal population under the recording
+sample of the activity of the neuronal network under the recording
 device, which can amount to tens of thousands of neurons
@@
 little meaning if we did not expect its results to generalize to a larger,
-unrecorded population -- at the very least the population under the
+unrecorded network -- at the very least the network under the
 recording device. In other words, we expect that the conclusions drawn with
@@
 neurons are a \emph{representative sample} of some larger neuronal
-population. Probability theory tells us how to make inferences from a
-sample to the larger population from which it is sampled \parentext{see
+network. Probability theory tells us how to make inferences from a
+sample to the larger network from which it is sampled \parentext{see
   references below}.
@@
 sample. But, given that our sample is representative of a larger
-population, we can also apply the \me\ method to the larger (unrecorded)
-population. The constraints are the same: the time averages of the sampled
-data, since they constitute representative data about the larger population
+network, we can also apply the \me\ method to the larger (unrecorded)
+network. The constraints are the same: the time averages of the sampled
+data, since they constitute representative data about the larger network
 as well. The method thus yields a probability distribution for the larger
-population, and the distribution for the sample is obtained by
+network, and the distribution for the sample is obtained by
 marginalization. The problem is that \emph{the distributions obtained from
@@
 In this work we develop the second way of applying the \me\ method, at the
-level of the larger population, and show that its results differ from the
+level of the larger network, and show that its results differ from the
 application at the sample level. We also consider the case where the size
-of the larger population is unknown.
-
-To apply the \me\ method to the larger, unsampled population, it is
+of the larger network is unknown.
+
+To apply the \me\ method to the larger, unsampled network, it is
 necessary to use probability relations relevant to sampling
@@
 
-\section{Probability relations between population and  sample}
+\section{Probability relations between network and  sample}
 \label{sec:prob_samples}
@@
 introduce an analogous notation for the $\yNv$ neurons constituting the
-larger population, but using the corresponding capital letters:
+larger network, but using the corresponding capital letters:
 $\yAv_{i}(t)$ is the activity of the $i$th neuron at time bin $t$,
 $\yA(t) \defd \sum_{i} \yAv_{i}(t)/\yNv$ is the activity at that
-bin averaged over the larger population, and so on.
-
-The probability relations between sample and larger population are valid at
+bin averaged over the larger network, and so on.
+
+The probability relations between sample and larger network are valid at
 every time bin. As we mentioned above, the \me\ distribution refers to any
@@
 %     &\text{\enquote{$\yXf = \yA$} 
-%       stands for \enquote{the (population-averaged) activity of the neurons is $\yA$}},
+%       stands for \enquote{the (network-averaged) activity of the neurons is $\yA$}},
 %     \\
@@
 backing our probability assignments -- our uncertainty about the full
-activity of the larger population is expressed by the joint probability
+activity of the larger network is expressed by the joint probability
 distribution
@@
 We need to make an initial probability assignment for the state of the full
-population before any experimental observations are made. This initial
+network before any experimental observations are made. This initial
 assignment will be modified by our experimental observations, and these can
-involve just a sample of the population. Our state of knowledge and initial
+involve just a sample of the network. Our state of knowledge and initial
 probability assignment should reflect that samples are somehow
-representative of the whole population.
+representative of the whole network.
 
 In this state of knowledge, denoted $\yH$, we know that the neurons in the
-population are biologically or functionally similar, for example in
+network are biologically or functionally similar, for example in
 morphology or the kind of input or output they receive or give. But we are
@@
 probability assignment is just an expression of the symmetry of our
-\emph{ignorance} about the state of the population, not an expression of
+\emph{ignorance} about the state of the network, not an expression of
 some biologic or physical symmetry or identity of the neurons.
@@
 activity is completely determined by the distribution for its
-population-average:
+network-average:
 \begin{equation}
@@
 
-Our uncertainties about the full population and the sample are connected
+Our uncertainties about the full network and the sample are connected
 via the conditional probability
@@
 of the sample is also symmetric and completely determined by the
-distribution of its population-averaged activity:
+distribution of its network-averaged activity:
 \begin{equation}
@@
 theorem~\eqref{eq:joint_plaus_N_homog} as the distribution for the full
-population.
+network.
 
@@
 $\pf(\ya \|\yA, \yH) \equiv \ypp_{\ya \yA}$, besides
-relating the distributions for the population and sample activities via
+relating the distributions for the network and sample activities via
 marginalization, also allows us to express the expectation value of any
 function of the sample activity, $\yg_{\ya}$, in terms of the distribution
-for the full population, as follows:
+for the full network, as follows:
 \begin{equation}
@@
 \emph{translates into a linear constraint for the distribution of the full
-  population}:
+  network}:
 \begin{equation}
@@
 that is, \emph{the expected values of the $m$-activities of the sample and
-  of the full population are equal}. The proof of the middle equality uses
+  of the full network are equal}. The proof of the middle equality uses
 the expression for the $m$th factorial moment of the hypergeometric
@@
 $m$-activity of a sample to have a particular value, we are also
-requiring  the expectation of the $m$-activity of the full population to
+requiring  the expectation of the $m$-activity of the full network to
 have the same value.
@@
 \end{figure}%hypergeometric_identity.nb
-These expectation equalities between sample and full population should not
+These expectation equalities between sample and full network should not
 be surprising: we intuitively \emph{expect} that the proportion of coloured
@@
 equalities express the intuitive fact that \emph{our uncertainty about the
-  sample is representative of our uncertainty about the population and
+  sample is representative of our uncertainty about the network and
   about other samples}, and vice versa. When $n=\yNv$, $\ypp_{\ya\yA}$
 becomes the identity matrix and the approximate equalities above become
-exact -- of course, since we have sampled the full population.
+exact -- of course, since we have sampled the full network.
 
@@
 probability distributions. In the next section we will in fact emphasize
-their differences. If the distribution for the population average $\yA$ is
+their differences. If the distribution for the network average $\yA$ is
 bimodal, for example, the bimodality can be lost in the distribution for
@@
 % $n\to \yNv$, do: these limits are trivially, universally valid because the
-% sample becomes the full population as $n\to \yNv$. In particular, these limits
+% sample becomes the full network as $n\to \yNv$. In particular, these limits
 % hold even in cases where the conditional probability
 % $\pf(\yxs = \ya \|\yXf=\yA)$ is not a fuzzy identity, our uncertainties
-% about sample and about population can differ wildly, and the approximate
+% about sample and about network can differ wildly, and the approximate
 % equalities~\eqref{eq:roughly_equal_nN} do not hold.
 
-\section{Maximum-entropy: sample level \vs\ full-population level}
+\section{Maximum-entropy: sample level \vs\ full-network level}
 \label{sec:specific_initial_probability}
@@
 be used as constraints on the distribution for the activity of the full
-population. Let us use such constraints with the \me\ method. Suppose that
+network. Let us use such constraints with the \me\ method. Suppose that
 we want to constrain $m$ functions of the sample activity, vectorially
@@
 directly to sampled neurons; denote this approach by $\yHb$. Then we apply
-the method to the full population of neurons, most of which are unsampled;
+the method to the full network of neurons, most of which are unsampled;
 denote this approach by $\yHa$.
@@
 exponential appears because we must account for the multiplicity by which
-the population-average activity $\ya$ can be realized: $\ya=0$ can be
+the network-average activity $\ya$ can be realized: $\ya=0$ can be
 realized in only one way (all neurons inactive), $\ya=1/n$ can be realized
@@
 
-Applied to the full population, using the constraint
+Applied to the full network, using the constraint
 expression~\eqref{eq:constraint_extended} derived in the previous section,
-the method yields the distribution for the full-population activity
+the method yields the distribution for the full-network activity
 \begin{equation}
@@
 Hence, \emph{the applications of \me\ at the sample level and at the
-  full-population level are inequivalent}. They lead to numerically
+  full-network level are inequivalent}. They lead to numerically
 different distributions for the sample activity $\yaa$.
@@
 \caption{Linear and log-plots of $\pf(\ya)$ constructed by \me\ at
-  the population level followed by sample marginalization (blue triangles),
+  the network level followed by sample marginalization (blue triangles),
   \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
@@
   \textcolor{mypurpleblue}{\textbf{blue triangles:}} \me\ at the
-  population level, \eqn~\eqref{eq:app_maxent_pop_marg} with
+  network level, \eqn~\eqref{eq:app_maxent_pop_marg} with
   $\yNv=10\,000$, followed by sample marginalization;
-  \textcolor{myyellow}{\textbf{yellow circles:}} \me\ at the population
-  level with unknown population size,
+  \textcolor{myyellow}{\textbf{yellow circles:}} \me\ at the network
+  level with unknown network size,
   \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}, according to the
-  distribution~\eqref{eq:pdf_popsize} for the population.}
+  distribution~\eqref{eq:pdf_popsize} for the network.}
 \label{fig:all_three}
@@
 The distribution obtained at the sample level will show different
-features from the one obtained at the population level, like displaced or
+features from the one obtained at the network level, like displaced or
 additional modes or particular tail behaviour. We show an example of this
@@
 motor cortex \citep{rostamietal2016_r2017}. The distribution obtained at the
-population level (blue triangles) has a higher and displaced mode and a quite
+network level (blue triangles) has a higher and displaced mode and a quite
 different behaviour for activities around $0.5$ than the distribution
@@
 In our discussion we have so far assumed the size $\yNv$ of the larger
-population to be known. This is rarely the case, however. We usually are
+network to be known. This is rarely the case, however. We usually are
 uncertain about $\yNv$ and can only guess its order of magnitude. In such a
@@
 assuming an entropic pre-data distribution for the long-run frequencies of
-the full population:
-
-\textbf{The frequency distribution $\yF$ of the full-population activity
+the full network:
+
+\textbf{The frequency distribution $\yF$ of the full-network activity
   \emph{during the recording}}
@@
 
-\textbf{The long-run frequency distribution $\ynu$ of the full-population
+\textbf{The long-run frequency distribution $\ynu$ of the full-network
   activity}
@@
 
-\textbf{The activity $\yA'$ of the full-population
+\textbf{The activity $\yA'$ of the full-network
   in a \emph{new} time bin}
@@
 set-up, that the \me\ method can be applied to recorded neuronal data in a
-way that accounts for the larger population from which the data are
+way that accounts for the larger network from which the data are
 sampled, \eqns~\eqref{eq:app_maxent_pop}--\eqref{eq:app_maxent_pop_marg}.
@@
 We gave a numerical example of this difference. We have also shown how to
-extend the new application when the size of the larger population is
+extend the new application when the size of the larger network is
 unknown, \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}.
@@
 \me\ 
-implicitly assumes that \emph{no} larger population exists beyond the
+implicitly assumes that \emph{no} larger network exists beyond the
 recorded sample of neurons. One could in fact object to the application at
-the population level, and say that the traditional way of applying \me,
+the network level, and say that the traditional way of applying \me,
 \eqn~\eqref{eq:app_maxent_sample}, yields different results because it does
 not make assumptions about the size $\yNv$ of a possibly existing larger
-population. Such a state of uncertainty, however, is correctly formalized
+network. Such a state of uncertainty, however, is correctly formalized
 according to the laws of probability by introducing a probability
@@
 for $\yNv$ gives unit probability to $\yNv=n$; that is, unless the sample
-\emph{is} the full population, and no larger population exists.
+\emph{is} the full network, and no larger network exists.
 
@@
 recording device. This assumption is unrealistic. The \me\ approach at the
-population level does not make such assumption and is therefore preferable.
+network level does not make such assumption and is therefore preferable.
 It may reveal features in a data set that were unnoticed by the standard
@@
 % very special, isolated set among all those that could have been recorded.
-% It is for this reason that we find the \me\ application at the population
+% It is for this reason that we find the \me\ application at the network
 % level preferable. Physical models of neuronal networks usually include some
 % sort of external input, mimicking an embedding in a larger network. The
-% \me\ distribution obtained at the population level may reveal features in a
+% \me\ distribution obtained at the network level may reveal features in a
 % data set that were unnoticed by the standard \me\ model.
@@
 The difference in the resulting distributions between the applications at
-the sample and at the population levels appears in the use of Boltzmann
+the sample and at the network levels appears in the use of Boltzmann
 machines with hidden units \citep{lerouxetal2008}, although by a different
@@
 neuroscientific applications, the difference in predictions of the sample
-\vs\ full-population applications can instead be very relevant.
+\vs\ full-network applications can instead be very relevant.
 
@@
 Finally, we would like to point out the usefulness and importance of the
-probability formulae that relate our states of knowledge about a population
+probability formulae that relate our states of knowledge about a network
 and its samples, presented in \sect~\ref{sec:prob_samples}. This kind of
@@
 also depend on the assumed size of the network and find that …(WE SHOULD
-TEST SHI) This predicts that for a large enough population (DO WE PREDICT
+TEST SHI) This predicts that for a large enough network (DO WE PREDICT
 THAT IF THE FULL NETWORK GETS BIGGER THE DIFFERENCE ALSO GET BIGGER)?…
@@
 Yes. In the simplest cases, the probability calculus give us simple and
-precise quantitative relations between our expectations about a population
+precise quantitative relations between our expectations about a network
 -- of voters, for instance -- and about a sample thereof. Suppose there are
@@
   proportion of political preferences in a sample as in the full
-  population}.
+  network}.
 
@@
 that expectation equalities such as~\eqref{eq:poll_equal_expectations} hold
-for the total activity $S(t)$ of the full population of $N$ neurons that
+for the total activity $S(t)$ of the full network of $N$ neurons that
 could have been recorded.
@@
   neurons we can construct maximum-entropy distributions for the total
-  activity of the \textbf{full population} of neurons that could have been
+  activity of the \textbf{full network} of neurons that could have been
   sampled by the recording device}.
@@
 % a probability distribution for the total activity at a new time bin of a
-% population of neurons, \emph{most of which can be unrecorded}, given
+% network of neurons, \emph{most of which can be unrecorded}, given
 % specific statistics about the activity of a recorded \emph{sample} of the
-% full population. This kind of probability assignment is extremely important
+% full network. This kind of probability assignment is extremely important
 % in neuroscience: we try to understand the workings of brain areas that may
@@
 % hidden assumption. We'll study in which circumstances the application of
-% the method at the sample level and at the full-population level give
+% the method at the sample level and at the full-network level give
 % similar or different probabilities.

### 2019-06-17 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-17T10:37:37+0200
+%% Last-Updated: 2019-06-17T10:56:29+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The context we consider is as follows. We have recorded the spiking
-activities of $n$ neurons for a given amount of time. These spikes are
-binned into $T$ time bins and binarized to $\set{0,1}$ values in each bin.
-Denote by $\ya_{t}$ the number of neurons that fire during time bin $t$,
-divided by the total number of neurons $n$; we call this the
-\emph{normalized total activity} or just \enquote{total activity} or
-\enquote{activity} for short. Obviously
-$\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no neuron spiked
-during time bin $t$; if $\ya_{t}=1$, all spiked at some point during time
-bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during time
-  bin $t$}.
+activities of $n$ neurons for a given amount of time. Let's call these
+neurons the \enquote{sample}. Their spikes are binned into $T$ time bins and
+binarized to $\set{0,1}$ values in each bin. Denote by $\ya_{t}$ the number
+of neurons that fire during time bin $t$, divided by the total number of
+neurons $n$; we call this the \emph{normalized total activity} of the
+sample, or just its \enquote{total activity} or \enquote{activity} for
+short. Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$,
+no neuron spiked during time bin $t$; if $\ya_{t}=1$, all spiked at some
+point during time bin $t$. For brevity, let's say \enquote{at $t$} for
+\enquote{during time bin $t$}.
 
@@
 sets of neurons had the same probability of being recorded as the set of
-$n$ neurons that have actually been recorded. We call this the
-\enquote{full network}. During the recording, the normalized total
-activity of these $N$ neurons at $t$ was $\yA_{t}$. We don't know the
-values $\yA_{t}$ at each $t$; we only know for certain that
-$\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
-for obvious reasons. For the moment we assume that we know $N$; in \sect***
-we discuss the consequences of our lack of precise knowledge about this
-number.
+$n$ neurons that has actually been recorded. We call this the \enquote{full
+  network}\mynote{how about calling it \enquote{the pool}?}. During the
+recording, the normalized total activity of these $N$ neurons at $t$ was
+$\yA_{t}$. We don't know the values $\yA_{t}$ at each $t$; we only know for
+certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
+$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the moment we assume that
+we know $N$; in \sect*** we discuss the consequences of our lack of precise
+knowledge about this number.
 
@@
 
-Regarding~\ref{item:maxent_step}, we assume you to be familiar with the
+Regarding~\ref{item:maxent_step}, we assume that you're familiar with the
 maximum-entropy method. In short, it amounts to a pair of prescriptions:
 choose the distribution having maximum Shannon entropy among those
-satisfying specific convex constraints, such as fixed expectation values; and
-judge these expectation values to be equal to measured averages. We add two
-remarks about this method that are seldom made in the literature. First,
-the \enquote{distribution} given by this method is the zeroth-order
+satisfying specific convex constraints, such as fixed expectation values;
+and judge these expectation values to be equal to measured averages. We add
+two remarks about this method that are seldom made in the literature.
+First, the \enquote{distribution} given by this method is the zeroth-order
 approximation of four different distributions:
@@
 distributions above numerically coincide in this approximation, they differ
-in higher-order approximations. Second, the maximum-entropy method based on
-the Shannon entropy implicitly makes specific assumptions about the
-probability of the long-run frequency distributions
-\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
-assumptions in \sect***
+in higher-order approximations. We discuss their differences in \sect***.
+Second, the maximum-entropy method based on the Shannon entropy implicitly
+makes specific assumptions about the probability of the long-run frequency
+distributions \citep{jaynes1986d_r1996,portamana2009,portamana2017}. We
+discuss these assumptions in \sect***.
 

### 2019-06-17 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-17T10:56:29+0200
+%% Last-Updated: 2019-06-17T11:00:59+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 activities of $n$ neurons for a given amount of time. Let's call these
-neurons the \enquote{sample}. Their spikes are binned into $T$ time bins and
-binarized to $\set{0,1}$ values in each bin. Denote by $\ya_{t}$ the number
-of neurons that fire during time bin $t$, divided by the total number of
-neurons $n$; we call this the \emph{normalized total activity} of the
-sample, or just its \enquote{total activity} or \enquote{activity} for
-short. Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$,
-no neuron spiked during time bin $t$; if $\ya_{t}=1$, all spiked at some
-point during time bin $t$. For brevity, let's say \enquote{at $t$} for
-\enquote{during time bin $t$}.
+neurons the \enquote{sample} or the \enquote{subnetwork}. Their spikes are
+binned into $T$ time bins and binarized to $\set{0,1}$ values in each bin.
+Denote by $\ya_{t}$ the number of neurons that fire during time bin $t$,
+divided by the total number of neurons $n$; we call this the
+\emph{normalized total activity} of the sample, or just its \enquote{total
+  activity} or \enquote{activity} for short. Obviously
+$\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no neuron spiked
+during time bin $t$; if $\ya_{t}=1$, all spiked at some point during time
+bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during time
+  bin $t$}.
 
@@
 
-In our case, to apply the method for the total activity of the full
-network we need to fix some of the latter's expectation values, for
-example its moments. But we don't have any measured averages for the full
-network to equate our expectations to. Here enters
-point~\ref{item:sample_step}: the probability calculus gives an exact,
-linear relation between the first $m$ moments for the full network and
-the first $m$ for the sample; the ones determine the others and vice versa
-at every time bin \citep[for example][\eqns~(16)]{portamanaetal2015}. This
-relation is a classical result of sampling theory
+In our case, to apply the method for the total activity of the full network
+we need to fix some of the latter's expectation values, for example its
+moments. But we don't have any measured averages for the full network to
+equate our expectations to. Here enters point~\ref{item:sample_step}: the
+probability calculus gives an exact, linear relation between the first $m$
+moments for the full network and the first $m$ for the sample
+\citep[\eqns~(16)]{portamanaetal2015}; the ones determine the others and
+vice versa at every time bin. This relation is a classical result of
+sampling theory
 \citep[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see

### 2019-06-17 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-17T11:00:59+0200
+%% Last-Updated: 2019-06-17T16:26:01+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % https://tex.stackexchange.com/a/33547/97039
+\newcommand*{\fm}{f}
+\newcommand*{\Fm}{F}
 
@@
 \begin{itemize}
+\item the most probable \emph{time-frequency} distribution of the total
+  activity across the \emph{recorded} bins,
+\item the most probable \emph{time-frequency} distribution of the total
+  activity in a very long run of \emph{new} time bins,
 \item the \emph{probability} distribution for the value of the total
@@
 \item the \emph{probability} distribution for the value of the total
-  activity at a \emph{new} time bin,
-\item the most probable \emph{time-frequency} distribution of the total
-  activity across the \emph{recorded} bins,
-\item the most probable \emph{time-frequency} distribution of the total
-  activity in a very long run of \emph{new} time bins.
+  activity at a \emph{new} time bin.
 \end{itemize}
@@
 in higher-order approximations. We discuss their differences in \sect***.
+The maximum-entropy distribution approximates best the first of the four
+distributions, so that's the one we'll mainly have in mind, and we'll
+denote it by $\Fm(\yA)$.***
 Second, the maximum-entropy method based on the Shannon entropy implicitly
@@
 total number of distinct pairs is $\tbinom{5}{2}=10$. The normalized number
-of pairs is therefore $6/10$. Note that the first $m$ factorial moments
-linearly determine the first $m$ moments and vice versa, so fixing the ones
-is equivalent to fixing the others. Now, since the recorded neurons are a
-sample from the full network of $N$ neurons, we have this extremely
-convenient property: \emph{The first $n$ normalized factorial moments for
-  the sample and for the full network are numerically identical}:
+of spiking pairs is therefore $6/10$. Note that the first $m$ factorial
+moments linearly determine the first $m$ power moments and vice versa,
+because $\tbinom{n\ya}{m}$ is a polynomial in $\ya$ of degree $m$; so
+fixing the ones is equivalent to fixing the others. Now we have this
+extremely convenient property: \emph{the first $n$ normalized factorial
+  moments for a sample and for the full network are numerically identical}:
 \begin{equation}
@@
 \label{sec:application}
+
+We apply the approach just described to two data sets publicly available in
+the literature:
+
+The first data set, from \textcite[rat 14147]{stensolaetal2012}, consists
+of $n=65$ neurons (27 of which classified as grid cells) from rat Medial
+Entorhinal Cortex, recorded for about 20 minutes. Their spikes are
+binned ainto $T=417\,641$ bins of $3\;\textrm{ms}$ width.
+
+The second data set, from \textcite{rostamietal2016_r2017}, consists of
+$n=159$ neurons from macaque Motor Cortex, recorded for about 15 minutes.
+Their spikes are binned ainto $T=300\,394$ bins of $3\;\textrm{ms}$ width.
+
+\texttt{data\_RS.npy}
 

### 2019-06-17 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-17T16:26:01+0200
+%% Last-Updated: 2019-06-17T17:13:33+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In the next section we apply this method to the data sets from two actual
-recordings, and discuss the properties of the resulting distributions.
+The number of moments used with this method depends on the specific
+questions and hypotheses that a researcher is exploring; for example,
+hypotheses of sufficient statistics, such as the sufficiency of pairwise
+correlations for inferential purposes. In the present work we only want to
+present the general method without entering into biological questions of
+this kind. The particular case in which $n$ moments are constrained
+corresponds to fully specify the measured time-frequency distribution of
+the sample activity. In this case, our inferences about the full-network
+activity are using all available measured frequency data. Using the full
+frequency distribution of the sample may be a bad idea, however, because
+the maximum-entropy distribution becomes a bad approximation of some of the
+four distributions described above. It is preferable to use a moderately
+high number of moments smaller than $n$. We explain this point in \sect***.
+
+\bigskip
+
+In the next section we apply the method to the data sets from two actual
+recordings, using *** moments, and discuss the properties of the resulting
+distributions.
 
@@
 Entorhinal Cortex, recorded for about 20 minutes. Their spikes are
-binned ainto $T=417\,641$ bins of $3\;\textrm{ms}$ width.
-
-The second data set, from \textcite{rostamietal2016_r2017}, consists of
-$n=159$ neurons from macaque Motor Cortex, recorded for about 15 minutes.
-Their spikes are binned ainto $T=300\,394$ bins of $3\;\textrm{ms}$ width.
-
-\texttt{data\_RS.npy}
+binned into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
+
+The second data set, from
+\textcite{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
+consists of $n=159$ neurons from macaque Motor Cortex, recorded for about
+15 minutes. Their spikes are binned into $T=300\,394$ bins of
+$3\;\textrm{ms}$ width.
 

### 2019-06-18 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-17T17:13:33+0200
+%% Last-Updated: 2019-06-18T15:05:04+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % https://tex.stackexchange.com/a/33547/97039
-\newcommand*{\fm}{f}
-\newcommand*{\Fm}{F}
+%\newcommand*{\fm}{f}
+\newcommand*{\px}{p}
 
@@
 \mynote{ref}\mynote{add recording length}. In the first data set, the most
-interesting inference is that the most probable time-frequency distribution
-for the total activity of the full network, if the latter is assumed to
-be larger than 1\,000 neurons, has two very distinct modes, both at low
+interesting inference is that the most probable frequency distribution for
+the total activity of the full network, if the latter is assumed to be
+larger than 1\,000 neurons, has two very distinct modes, both at low
 activities ($0.7\%$ and $2.6\%$ of simultaneously active neurons within
@@
 full network during and after the recording. For example: what was its
-frequency distribution in time? How much total activity should we expect at
-a time bin during the recording, or at a new time bin? We cannot answer
-these questions with certainty; we can only give a distribution of
-probability over the possible answers. Our approach gives such a
-probability distribution.
+frequency distribution in time during the recording? How much total
+activity should we expect at a time bin during the recording? And at a new
+time bin? We cannot answer these questions with certainty; we can only give
+distributions of probability over their possible answers. Our approach
+gives such probability distributions.
 
@@
   activity of the full network using the maximum-entropy method;
-  \item\label{item:sample_step} the expectation constraints used in the
-  maximum-entropy method for the full network are, in turn, determined
-  via sampling theory from the expectation constraints for the sample.
+  \item\label{item:sample_step} the constrained averages used in the
+  maximum-entropy method for the full network are, in turn, determined via
+  sampling theory from the constrained averages for the sample.
 \end{enumerate}
@@
 choose the distribution having maximum Shannon entropy among those
-satisfying specific convex constraints, such as fixed expectation values;
-and judge these expectation values to be equal to measured averages. We add
-two remarks about this method that are seldom made in the literature.
-First, the \enquote{distribution} given by this method is the zeroth-order
-approximation of four different distributions:
+satisfying specific convex constraints, such as fixed averages; and judge
+these averages to be equal to measured averages. We add two remarks about
+this method that are seldom made in the literature. First, the
+\enquote{distribution} given by this method, which we denote by $\px(\yA)$,
+is the zeroth-order approximation of four different distributions for the
+full network:
 \begin{itemize}
-\item the most probable \emph{time-frequency} distribution of the total
-  activity across the \emph{recorded} bins,
-\item the most probable \emph{time-frequency} distribution of the total
-  activity in a very long run of \emph{new} time bins,
+\item the most probable \emph{frequency} distribution for the total
+  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
+\item the most probable \emph{frequency} distribution for the total
+  activity in a very long run of \emph{new} time bins, $\ynuu(\yA)$,
 \item the \emph{probability} distribution for the value of the total
   activity at any time bin among those \emph{recorded},
+  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
 \item the \emph{probability} distribution for the value of the total
   activity at a \emph{new} time bin.
+  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
 \end{itemize}
 The maximum-entropy distribution is thus an approximate answer to four
-conceptually different inferences. Note that although the four
-distributions above numerically coincide in this approximation, they differ
-in higher-order approximations. We discuss their differences in \sect***.
-The maximum-entropy distribution approximates best the first of the four
-distributions, so that's the one we'll mainly have in mind, and we'll
-denote it by $\Fm(\yA)$.***
+completely different inferences. Note that although the four distributions
+above numerically coincide in this approximation, they differ in
+higher-order approximations. We discuss their differences in \sect***.
 Second, the maximum-entropy method based on the Shannon entropy implicitly
-makes specific assumptions about the probability of the long-run frequency
-distributions \citep{jaynes1986d_r1996,portamana2009,portamana2017}. We
-discuss these assumptions in \sect***.
+makes specific assumptions about the probabilities for the long-run
+frequency distributions
+\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
+assumptions in \sect***.
 
 In our case, to apply the method for the total activity of the full network
-we need to fix some of the latter's expectation values, for example its
-moments. But we don't have any measured averages for the full network to
-equate our expectations to. Here enters point~\ref{item:sample_step}: the
+we need to fix some of the latter's averages, for example its moments. But
+we don't have any measured moments for the full network to equate the
+distribution moments to. Here enters point~\ref{item:sample_step}: the
 probability calculus gives an exact, linear relation between the first $m$
@@
 Combining this result with the maximum-entropy prescription
-\enquote{expectations${}={}$measured averages} for the sample, we have that
-the measured averages for the sample determine the moments for the full
-network:
+\enquote{moments${}={}$measured moments} for the sample, we have that the
+measured moments for the sample determine the moments for the full network:
 \begin{equation*}
   \label{eq:chain_prescription_moments}
-  \mathrlap{\overbracket{\phantom{\text{measured averages} \rightarrow
+  \mathrlap{\overbracket{\phantom{\text{measured moments} \rightarrow
         \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
-  \text{measured averages} \rightarrow
+  \text{measured moments} \rightarrow
         \underbracket{\text{sample moments}
@@
 These two steps are more straightforward if instead of power moments we use
-\emph{normalized factorial moments}. The $m$th normalized factorial moment
-of the probability distribution for the activity of the recorded neurons is
-defined as the expectation
-\begin{equation}
-  \E\condt[\Bigg]{
-    \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
-  \equiv
+\emph{normalized factorial moments} \citep{potts1953}. The $m$th normalized
+factorial moment of a distribution $\px(\ya)$ for the activity of the
+sample neurons is defined as the average
+\begin{equation}
+  % \E\condt[\Bigg]{
+  %   \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
+  % \equiv
   \sum_{\ya} %m!\,
-  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pf(\ya \| \yH),
+  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \px(\ya),
   \qquad 1\le m \le n
 \end{equation}
-where $\pf(\ya \| \yH)$ is the probability for activity $\ya$ given
-assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
-new bin. This moment can be interpreted as the expectation of the number of
-distinct $m$-tuples of simultaneously spiking neurons (within a bin's time
-width), normalized by the number of distinct $m$-tuples. For example, with
-$m=2$, if $na=4$ neurons spike in a network of $n=5$, we have
-$\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons, and the
-total number of distinct pairs is $\tbinom{5}{2}=10$. The normalized number
-of spiking pairs is therefore $6/10$. Note that the first $m$ factorial
-moments linearly determine the first $m$ power moments and vice versa,
-because $\tbinom{n\ya}{m}$ is a polynomial in $\ya$ of degree $m$; so
-fixing the ones is equivalent to fixing the others. Now we have this
-extremely convenient property: \emph{the first $n$ normalized factorial
-  moments for a sample and for the full network are numerically identical}:
+% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
+% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
+% new bin.
+This moment can be interpreted as the expectation of the number of distinct
+$m$-tuples of simultaneously spiking neurons (within a bin's time width),
+normalized by the number of distinct $m$-tuples. For example, with $m=2$,
+if $na=4$ neurons spike in a network of $n=5$, we have $\tbinom{4}{2}=6$
+distinct pairs of simultaneously spiking neurons, and the total number of
+distinct pairs is $\tbinom{5}{2}=10$. The normalized number of spiking
+pairs is therefore $6/10$. Note that the first $m$ factorial moments
+linearly determine the first $m$ power moments and vice versa, because
+$\tbinom{n\ya}{m}$ is a polynomial in $\ya$ of degree $m$; so fixing the
+ones is equivalent to fixing the others. Now we have this extremely
+convenient property: \emph{the first $n$ normalized factorial moments for a
+  sample and for the full network are numerically identical}:
 \begin{equation}
   \label{eq:equal_m_factorialmoments}
-  \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
-  \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
-  \qquad m \le n.
-\end{equation}
+    \sum_{\ya} %m!\,
+  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \px(\ya)
+=    \sum_{\yA} %m!\,
+  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA),
+  % \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
+  % \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
+  \qquad 1\le m \le n,
+\end{equation}
+where $\px(\yA)$ is the distribution for the full-network activity.
 
@@
 $\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$, for as many $m$ as we please with
-$1\le m \le n$. The constraints on the distribution $\pf(\yA \| \yH)$ for
+$1\le m \le n$. The constraints on the distribution $\px(\yA)$ for
 the full network are
@@
    \sum_{\yA}
-  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \pf(\yA \| \yH)
+  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \pf(\yA)
 \end{equation}
@@
 
-The number of moments used with this method depends on the specific
-questions and hypotheses that a researcher is exploring; for example,
-hypotheses of sufficient statistics, such as the sufficiency of pairwise
-correlations for inferential purposes. In the present work we only want to
-present the general method without entering into biological questions of
-this kind. The particular case in which $n$ moments are constrained
-corresponds to fully specify the measured time-frequency distribution of
-the sample activity. In this case, our inferences about the full-network
-activity are using all available measured frequency data. Using the full
-frequency distribution of the sample may be a bad idea, however, because
-the maximum-entropy distribution becomes a bad approximation of some of the
-four distributions described above. It is preferable to use a moderately
-high number of moments smaller than $n$. We explain this point in \sect***.
+The number of moments used with this method depends on the questions and
+hypotheses that a researcher is exploring; for example, hypotheses of
+sufficient statistics, such as the sufficiency of pairwise correlations for
+inferential purposes. In the present work we only want to present the
+general method without entering into biological questions of this kind.
+
+The particular case in which $n$ moments are constrained is especially
+important: it corresponds to fully specify the measured frequency
+distribution for the activity of the sample neurons. In this case, our
+inferences about the full-network activity are using all available measured
+frequency data. Note that application of the maximum-entropy method
+\emph{at the sample level} is in this case trivial and meaningless -- it
+just gives back the measured frequency distribution. But application of the
+method \emph{at the level of the full network}, as done here, is not
+trivial.
+
+Using the full frequency distribution of the sample may be a bad idea,
+however, because the maximum-entropy distribution becomes a bad
+approximation of some of the four distributions described above. It is
+preferable to use a moderately high number of moments smaller than $n$. We
+explain this point in \sect***.
 
@@
 the literature:
-
-The first data set, from \textcite[rat 14147]{stensolaetal2012}, consists
-of $n=65$ neurons (27 of which classified as grid cells) from rat Medial
-Entorhinal Cortex, recorded for about 20 minutes. Their spikes are
-binned into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
-
-The second data set, from
+\begin{itemize}[wide]
+\item The first data set, from \textcite[rat 14147]{stensolaetal2012},
+  consists of $n=65$ neurons (27 of which classified as grid cells) from
+  rat Medial Entorhinal Cortex, recorded for about 20 minutes. Their spikes
+  are binned into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
+\item The second data set, from
   \textcite{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
@@
   $3\;\textrm{ms}$ width.
+\end{itemize}
 

### 2019-06-18 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-18T15:05:04+0200
+%% Last-Updated: 2019-06-18T16:16:26+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\newcommand*{\fm}{f}
-\newcommand*{\px}{p}
+\newcommand*{\px}{p_{\text{me}}}
 
@@
 bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during time
-  bin $t$}.
+  bin $t$}. From the activities $\set{\ya_{t}}$ we can count how many times
+the activity $\ya=0$, $\ya=1/n$, and so on each appeared during the
+recording, obtaining the measured relative-frequency distribution
+$\yff(\ya) \defs \yf$.
 
@@
 recording, the normalized total activity of these $N$ neurons at $t$ was
-$\yA_{t}$. We don't know the values $\yA_{t}$ at each $t$; we only know for
-certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
-$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the moment we assume that
-we know $N$; in \sect*** we discuss the consequences of our lack of precise
-knowledge about this number.
+$\yA_{t}$; the relative frequencies of the various activity levels were
+$\yFF(\yA) \defs \yF$. We don't know the values $\yA_{t}$ at each $t$ or
+the frequency distribution $\yF$. We only know for certain that
+$\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
+for obvious reasons. For the moment we assume that we know $N$; in \sect***
+we discuss the consequences of our lack of precise knowledge about this
+number.
 
@@
 full network during and after the recording. For example: what was its
-frequency distribution in time during the recording? How much total
-activity should we expect at a time bin during the recording? And at a new
-time bin? We cannot answer these questions with certainty; we can only give
+frequency distribution $\yF$ during the recording? How much total activity
+should we expect at a time bin during the recording? And at a new time bin?
+We cannot answer these questions with certainty; we can only give
 distributions of probability over their possible answers. Our approach
@@
 \begin{enumerate}[label=(\alph*)]
-  \item\label{item:maxent_step} we build a distribution for the total
+  \item\label{item:maxent_step} we build a distribution $\px(\yA)$ for the total
   activity of the full network using the maximum-entropy method;
@@
 these averages to be equal to measured averages. We add two remarks about
-this method that are seldom made in the literature. First, the
-\enquote{distribution} given by this method, which we denote by $\px(\yA)$,
-is the zeroth-order approximation of four different distributions for the
-full network:
+this method that are seldom made in the literature. First, the distribution
+$\px(\yA)$ given by this method is the zeroth-order approximation of four
+different distributions for the full network:
 \begin{itemize}
@@
    \sum_{\yA}
-  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \pf(\yA)
+  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA)
 \end{equation}

### 2019-06-18 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-18T16:16:26+0200
+%% Last-Updated: 2019-06-18T19:00:36+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The context we consider is as follows. We have recorded the spiking
-activities of $n$ neurons for a given amount of time. Let's call these
-neurons the \enquote{sample} or the \enquote{subnetwork}. Their spikes are
-binned into $T$ time bins and binarized to $\set{0,1}$ values in each bin.
-Denote by $\ya_{t}$ the number of neurons that fire during time bin $t$,
-divided by the total number of neurons $n$; we call this the
-\emph{normalized total activity} of the sample, or just its \enquote{total
-  activity} or \enquote{activity} for short. Obviously
+activities of $n$ neurons for a given amount of time. These neurons are our
+\enquote{sample} or \enquote{subnetwork}. Their spikes are binned into $T$
+time bins and binarized to $\set{0,1}$ values in each bin. Denote by
+$\ya_{t}$ the number of neurons that fire during time bin $t$, divided by
+the total number of neurons $n$; this is the \emph{normalized total
+  activity} of the sample, or just its \enquote{total activity} or
+\enquote{activity} for short. Obviously
 $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no neuron spiked
-during time bin $t$; if $\ya_{t}=1$, all spiked at some point during time
-bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during time
-  bin $t$}. From the activities $\set{\ya_{t}}$ we can count how many times
-the activity $\ya=0$, $\ya=1/n$, and so on each appeared during the
+during bin $t$; if $\ya_{t}=1$, all spiked at some point during bin $t$.
+For brevity, let's say \enquote{at $t$} for \enquote{during time bin $t$}.
+From the activities $\set{\ya_{t}}$ we can count how many times the
+activity levels $\ya=0$, $\ya=1/n$, and so on appeared during the
 recording, obtaining the measured relative-frequency distribution
-$\yff(\ya) \defs \yf$.
+$(\yff_{\ya}) \defs \yf$.
 
@@
 enable the recording of neuronal activity from small brain regions
-\mynote{be more specific}. The neurons that are recorded within the region
-are not, and cannot yet be, specifically chosen from among the rest, owing
-to several limiting factors -- for example, limitations in the precision
-with which electrodes are inserted, or with which neurons are targeted by
-viruses, and so on. In fact, the set of recorded neurons may even change
-during very long recordings or across experimental trials. We assume that
-there's a region, comprising a network of $N$ neurons, from which other
-sets of neurons had the same probability of being recorded as the set of
-$n$ neurons that has actually been recorded. We call this the \enquote{full
-  network}\mynote{how about calling it \enquote{the pool}?}. During the
-recording, the normalized total activity of these $N$ neurons at $t$ was
-$\yA_{t}$; the relative frequencies of the various activity levels were
-$\yFF(\yA) \defs \yF$. We don't know the values $\yA_{t}$ at each $t$ or
-the frequency distribution $\yF$. We only know for certain that
-$\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
-for obvious reasons. For the moment we assume that we know $N$; in \sect***
-we discuss the consequences of our lack of precise knowledge about this
-number.
+\mynote{be more specific}. For many animal species, the neurons that are
+recorded within the region are not, and at present cannot be, specifically
+chosen from among the rest, owing to several limiting factors -- for
+example, limitations in the precision with which electrodes are inserted,
+or with which neurons are targeted by viruses, and so on. In fact, the set
+of recorded neurons may even change during very long recordings or across
+experimental sessions. We assume that there's a region, comprising a
+network of $N$ neurons, from which other sets of neurons had the same
+probability of being recorded as the set of $n$ neurons that was actually
+recorded. This is our \enquote{full network}\mynote{how about calling it
+  \enquote{the pool}?}. During the recording, the normalized total activity
+of these $N$ neurons at $t$ was $\yA_{t}$; the relative frequencies of the
+various activity levels were $(\yFF_{yA}) \defs \yF$. We don't know the
+values $\yA_{t}$ at each $t$ or the frequency distribution $\yF$. We only
+know for certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
+$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the moment we assume that
+we know $N$; in \sect*** we discuss the consequences of our lack of precise
+knowledge about this number.
 
@@
 frequency distribution $\yF$ during the recording? How much total activity
-should we expect at a time bin during the recording? And at a new time bin?
-We cannot answer these questions with certainty; we can only give
-distributions of probability over their possible answers. Our approach
-gives such probability distributions.
+$\yA_{t}$ should we expect at a time bin $t$ during the recording? And at a
+new time bin? We cannot answer these questions with certainty; we can only
+give distributions of probability over their possible answers. The approach
+presented here gives such probability distributions.
 
@@
 
-The idea behind the approach is easily summarized:
+The idea behind our approach is easily summarized:
 \begin{enumerate}[label=(\alph*)]
@@
 \item the most probable \emph{frequency} distribution for the total
-  activity in a very long run of \emph{new} time bins, $\ynuu(\yA)$,
+  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
 \item the \emph{probability} distribution for the value of the total

### 2019-06-18 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-18T19:00:36+0200
+%% Last-Updated: 2019-06-18T19:09:53+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\px}{p_{\text{me}}}
+\newcommand*{\pxx}{p}
 
@@
 \emph{normalized factorial moments} \citep{potts1953}. The $m$th normalized
-factorial moment of a distribution $\px(\ya)$ for the activity of the
+factorial moment of a distribution $\pxx(\ya)$ for the activity of the
 sample neurons is defined as the average
@@
   \sum_{\ya} %m!\,
-  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \px(\ya),
+  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pxx(\ya),
   \qquad 1\le m \le n
@@
     \sum_{\ya} %m!\,
-  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \px(\ya)
+  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pxx(\ya)
 =    \sum_{\yA} %m!\,
-  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA),
+  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \pxx(\yA),
   % \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
@@
 \end{equation}
-where $\px(\yA)$ is the distribution for the full-network activity.
+where $\pxx(\yA)$ is the distribution for the full-network activity.
 
@@
 $\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$, for as many $m$ as we please with
-$1\le m \le n$. The constraints on the distribution $\px(\yA)$ for
-the full network are
+$1\le m \le n$. The constraints on the maximum-entropy distribution
+$\px(\yA)$ for the full network are
 \begin{equation}
@@
 The particular case in which $n$ moments are constrained is especially
-important: it corresponds to fully specify the measured frequency
+important: it corresponds to fully constraining the marginal frequency
 distribution for the activity of the sample neurons. In this case, our
@@
 frequency data. Note that application of the maximum-entropy method
-\emph{at the sample level} is in this case trivial and meaningless -- it
+\emph{at the sample level} is trivial and meaningless in this case -- it
 just gives back the measured frequency distribution. But application of the
-method \emph{at the level of the full network}, as done here, is not
-trivial.
+method \emph{at the level of the full network} is not trivial.
 
 Using the full frequency distribution of the sample may be a bad idea,
-however, because the maximum-entropy distribution becomes a bad
+however, because the maximum-entropy distribution may become a bad
 approximation of some of the four distributions described above. It is
@@
 
-In the next section we apply the method to the data sets from two actual
-recordings, using *** moments, and discuss the properties of the resulting
-distributions.
+In the next section we apply the method just described to the data sets
+from two actual recordings, using *** moments, and discuss the properties
+of the resulting distributions.
 

### 2019-06-18 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-18T19:09:53+0200
+%% Last-Updated: 2019-06-18T20:14:11+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The context we consider is as follows. We have recorded the spiking
-activities of $n$ neurons for a given amount of time. These neurons are our
-\enquote{sample} or \enquote{subnetwork}. Their spikes are binned into $T$
-time bins and binarized to $\set{0,1}$ values in each bin. Denote by
+activities of $n$ neurons for a certain amount of time. These neurons are
+our \enquote{sample} or \enquote{subnetwork}. Their spikes are binned into
+$T$ time bins and binarized to $\set{0,1}$ values in each bin. Call
 $\ya_{t}$ the number of neurons that fire during time bin $t$, divided by
 the total number of neurons $n$; this is the \emph{normalized total
-  activity} of the sample, or just its \enquote{total activity} or
-\enquote{activity} for short. Obviously
-$\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no neuron spiked
-during bin $t$; if $\ya_{t}=1$, all spiked at some point during bin $t$.
-For brevity, let's say \enquote{at $t$} for \enquote{during time bin $t$}.
-From the activities $\set{\ya_{t}}$ we can count how many times the
-activity levels $\ya=0$, $\ya=1/n$, and so on appeared during the
-recording, obtaining the measured relative-frequency distribution
-$(\yff_{\ya}) \defs \yf$.
+  activity} of the sample, or just \enquote{activity} for short. Obviously
+$\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no neuron spikes
+during bin $t$; if $\ya_{t}=1$, all spike at some point during bin $t$. For
+brevity, let's say \enquote{at $t$} for \enquote{during time bin $t$}. From
+the activities $\set{\ya_{t}}$ we can count how many times the activity
+levels $\ya=0$, $\ya=1/n$, and so on appeared during the recording,
+obtaining the measured relative-frequency distribution
+$(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
+bins \emph{outside} of the recorded range. Such activity is unknown to us,
+of course.
 
@@
 \mynote{be more specific}. For many animal species, the neurons that are
-recorded within the region are not, and at present cannot be, specifically
-chosen from among the rest, owing to several limiting factors -- for
-example, limitations in the precision with which electrodes are inserted,
-or with which neurons are targeted by viruses, and so on. In fact, the set
+recorded within the region are not -- and at present cannot be --
+specifically chosen from among the rest, owing to several limiting factors;
+for example, limitations in the precision with which electrodes are
+inserted, or with which neurons are targeted by viruses. In fact, the set
 of recorded neurons may even change during very long recordings or across
@@
 recorded. This is our \enquote{full network}\mynote{how about calling it
-  \enquote{the pool}?}. During the recording, the normalized total activity
-of these $N$ neurons at $t$ was $\yA_{t}$; the relative frequencies of the
-various activity levels were $(\yFF_{yA}) \defs \yF$. We don't know the
-values $\yA_{t}$ at each $t$ or the frequency distribution $\yF$. We only
+  \enquote{the pool}?}. The normalized total activity of these $N$ neurons
+at $t$ is $\yA_{t}$. The relative frequencies of the various activity
+levels during the recording were $(\yFF_{yA}) \defs \yF$. We don't know the
+values $\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We only
 know for certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
-$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the moment we assume that
-we know $N$; in \sect*** we discuss the consequences of our lack of precise
-knowledge about this number.
+$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the time being we assume
+that we know $N$; in \sect*** we discuss the consequences of our lack of
+precise knowledge about this number.
 
@@
 The maximum-entropy distribution is thus an approximate answer to four
-completely different inferences. Note that although the four distributions
-above numerically coincide in this approximation, they differ in
-higher-order approximations. We discuss their differences in \sect***.
-Second, the maximum-entropy method based on the Shannon entropy implicitly
-makes specific assumptions about the probabilities for the long-run
-frequency distributions
+completely different inferences. Note that the four distributions above
+numerically differ in higher-order approximations. We discuss their
+differences in \sect***. Second, the maximum-entropy method based on the
+Shannon entropy implicitly makes some assumptions about the probabilities
+for the long-run frequency distributions
 \citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
@@
   \frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
+  \equiv
+   \sum_{\ya}
+  \binom{n \ya}{m}\,  \binom{n}{m}^{-1} \yff_{\ya}
   =

### 2019-06-19 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-18T20:14:11+0200
+%% Last-Updated: 2019-06-19T11:23:16+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The context we consider is as follows. We have recorded the spiking
-activities of $n$ neurons for a certain amount of time. These neurons are
-our \enquote{sample} or \enquote{subnetwork}. Their spikes are binned into
-$T$ time bins and binarized to $\set{0,1}$ values in each bin. Call
-$\ya_{t}$ the number of neurons that fire during time bin $t$, divided by
-the total number of neurons $n$; this is the \emph{normalized total
-  activity} of the sample, or just \enquote{activity} for short. Obviously
-$\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no neuron spikes
-during bin $t$; if $\ya_{t}=1$, all spike at some point during bin $t$. For
-brevity, let's say \enquote{at $t$} for \enquote{during time bin $t$}. From
-the activities $\set{\ya_{t}}$ we can count how many times the activity
-levels $\ya=0$, $\ya=1/n$, and so on appeared during the recording,
-obtaining the measured relative-frequency distribution
-$(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
-bins \emph{outside} of the recorded range. Such activity is unknown to us,
-of course.
+The context we consider is as follows. During an experimental session we
+have recorded the spiking activities of $n$ neurons for a certain amount of
+time. These neurons are our \enquote{sample} or \enquote{subnetwork}. Their
+spikes are binned into $T$ time bins and binarized to $\set{0,1}$ values in
+each bin. Call $\ya_{t}$ the number of neurons that fire during time bin
+$t$, divided by the total number of neurons $n$; this is the
+\emph{normalized total activity} of the sample, or just \enquote{activity}
+for short. Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if
+$\ya_{t}=0$, no neuron spikes during bin $t$; if $\ya_{t}=1$, all spike at
+some point during bin $t$. For brevity, let's say \enquote{at $t$} for
+\enquote{during time bin $t$}. From the activities $\set{\ya_{t}}$ we can
+count how many times the activity levels $\ya=0$, $\ya=1/n$, and so on
+appeared during the recording, obtaining the measured relative-frequency
+distribution $(\yff_{\ya}) \defs \yf$. We can also consider the sample
+activity at time bins \emph{outside} of the recorded range. Such activity
+is unknown to us, of course.
 
@@
 experimental sessions. We assume that there's a region, comprising a
-network of $N$ neurons, from which other sets of neurons had the same
-probability of being recorded as the set of $n$ neurons that was actually
-recorded. This is our \enquote{full network}\mynote{how about calling it
+network of $N$ neurons, from which other sets of neurons had or will have
+the same probability of being recorded, even in other experimental
+sessions, as the set of $n$ neurons that was actually recorded in this
+session. This is our \enquote{full network}\mynote{how about calling it
   \enquote{the pool}?}. The normalized total activity of these $N$ neurons
 at $t$ is $\yA_{t}$. The relative frequencies of the various activity
-levels during the recording were $(\yFF_{yA}) \defs \yF$. We don't know the
+levels during the recording were $(\yFF_{\yA}) \defs \yF$. We don't know the
 values $\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We only
@@
 Our questions concern general features of the total activity $\yA$ of the
-full network during and after the recording. For example: what was its
-frequency distribution $\yF$ during the recording? How much total activity
-$\yA_{t}$ should we expect at a time bin $t$ during the recording? And at a
-new time bin? We cannot answer these questions with certainty; we can only
-give distributions of probability over their possible answers. The approach
-presented here gives such probability distributions.
+full network during and after the recording, and across sessions under the
+same study conditions. For example: what was its frequency distribution
+during the recording? How much does this frequency distribution change
+across sessions? How much total activity should we expect at any time bin
+during a recording? We cannot answer these questions with certainty; we can
+only give distributions of probability over their possible answers. The
+approach presented here gives such probability distributions.
+
+\mynote{Move this to intro?}We want to stress the usefulness of making
+inferences about the full-network activity. First of all, this seems to be
+the primary idea behind recording a sample. Second, it allows us to make
+comparisons across experimental sessions; such comparisons would me
+difficult or meaningless made at the level of recorded samples, which
+generally comprise non-overlapping sets of neurons differing also in size.
 
@@
 
+The calculation amounts to a convex optimization
+\citep{meadetal1984}[\chap~10]{pressetal1988_r2007}{fangetal1997,boydetal2004_r2009,portamana2017b}
+and for the numbers $N,n,m$ considered in the next section it can be done
+on a modern computer without approximations of normalization constants or
+potential functions.
+
 \bigskip
@@
 \begin{itemize}[wide]
-\item The first data set, from \textcite[rat 14147]{stensolaetal2012},
+\item The first, from \textcite[rat 14147]{stensolaetal2012},
   consists of $n=65$ neurons (27 of which classified as grid cells) from
@@
   are binned into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
-\item The second data set, from
+\item The second, from
   \textcite{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
@@
 \end{itemize}
+
+We first calculate the maximum-entropy distribution for the full network by
+using six moments. This number already provides as much information as the
+full frequency distribution of the sample, and at the same time illustrates
+the use of the approach in questions of statistic sufficiency (typically
+limited to two or three moments). We show the results for full-network
+sizes $N=1\,000$, $N=5\,000$, $N=10\,000$\mynote{motivate?}. Thereafter we
+show the results for two moments and six moments using $N=10\,000$.
+
+\begin{figure}[!b]
+\centering
+\includegraphics[width=\linewidth]{scripts/paper_stensola_6mom_distributionsN_prior2.pdf}%
+\\%
+\includegraphics[width=\linewidth]{scripts/paper_riehle_6mom_distributionsN_prior2.pdf}%
+\caption{***}
+\label{fig:plots_6mom}
+\end{figure}
+%maxent_solutions_different_moments_prior2_6moments.nb
 

### 2019-06-19 scripts/ME_1fmoms.m

@@ -4,8 +4,8 @@ Clear[ME1fmoms,ME1fmoms0,ME1fmomsc,ME1fmomsn];
 
 ME1fmoms[eemoms_, nn_, wp_:MachinePrecision,pg_:4] := 
   Block[{ll=Length[eemoms],js=Table[Unique["js"],Length[eemoms]], freeenm, soluzm, prob, bvalues,fmoms,emoms},
-	emoms=1-eemoms;
-	fmoms[s_]=Table[1-Binomial[s,r]/Binomial[nn,r], {r,ll}];
+	emoms=1/2-eemoms;
+	fmoms[s_]=Table[1/2-Binomial[s,r]/Binomial[nn,r], {r,ll}];
 	freeenm = Log@Sum[(1-ss+nn)/((nn+1)*(nn+2))*
 			 Exp[js.(fmoms[ss]-emoms)],{ss, 0, nn}];
 

### 2019-06-19 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-19T11:23:16+0200
+%% Last-Updated: 2019-06-19T13:30:42+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 the use of the approach in questions of statistic sufficiency (typically
-limited to two or three moments). We show the results for full-network
-sizes $N=1\,000$, $N=5\,000$, $N=10\,000$\mynote{motivate?}. Thereafter we
-show the results for two moments and six moments using $N=10\,000$.
-
-\begin{figure}[!b]
+limited to two or three moments). Figure~\ref{fig:plots_6mom} shows the
+resulting densities (distribution/activity-step, that is,
+distribution${}\times N$) for full-network sizes $N=n$, $N=1\,000$,
+$N=5\,000$, $N=10\,000$\mynote{motivate?}. The case $N=n$ corresponds to
+applying the maximum-entropy method at the sample level.
+
+Thereafter we show the results for two moments and six moments using
+$N=10\,000$.
+
+\begin{figure}[!p]
 \centering
 \includegraphics[width=\linewidth]{scripts/paper_stensola_6mom_distributionsN_prior2.pdf}%
-\\%
+\\[3em]%
 \includegraphics[width=\linewidth]{scripts/paper_riehle_6mom_distributionsN_prior2.pdf}%

### 2019-06-19 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-19T13:30:42+0200
+%% Last-Updated: 2019-06-19T14:57:21+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\bigskip
+
 We first calculate the maximum-entropy distribution for the full network by
@@
 $N=5\,000$, $N=10\,000$\mynote{motivate?}. The case $N=n$ corresponds to
-applying the maximum-entropy method at the sample level.
-
-Thereafter we show the results for two moments and six moments using
-$N=10\,000$.
-
+applying the maximum-entropy method at the sample level; it can be observed
+that with six moments it reproduces almost exactly the measured frequency
+distribution.
 \begin{figure}[!p]
@@
 \includegraphics[width=\linewidth]{scripts/paper_riehle_6mom_distributionsN_prior2.pdf}%
+\caption{***}
+\label{fig:plots_6mom}
+\end{figure}
+%maxent_solutions_different_moments_prior2_6moments.nb
+
+The distribution for the full-network is sharper than the measured
+frequency distribution for the sample; the sharper the larger $N$ is. Most
+remarkably, it has two distinct low-activity modes for the first data set.
+But also the second data set presents, upon closer inspection, a small
+shoulder on the right of the mode, suggestive of two activity regimes.
+These features are clearly not present in the sample or in the
+maximum-entropy distribution at the sample level. The application of the
+probability calculus to make forecasts about the full network thus reveals
+interesting and useful features.
+
+Thereafter we show the results for two moments and for four moments using
+$N=10\,000$.
+
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\linewidth]{scripts/paper2_stensola_twomom_distributions10000_prior2.pdf}%
+\\[3em]%
+\includegraphics[width=\linewidth]{scripts/paper_riehle_twomom_distributions10000_prior2.pdf}%
 \caption{***}

### 2019-06-19 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-19T14:57:21+0200
+%% Last-Updated: 2019-06-19T15:25:54+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Thereafter we show the results for two moments and for four moments using
-$N=10\,000$.
+\bigskip
+
+As an illustration of our approach in questions of sufficient statistics,
+\fig~\ref{fig:plots_2_4mom} shows the results for two constrained moments
+(equivalent to constraining means and correlations only) and for four
+constrained moments, using $N=10\,000$. In either data set we obtain very
+distinct distributions for the two kinds of constraints. In particular, for
+the first data set with four moments we obtain a bimodal distribution, and
+with two moments a unimodal one, showing that two moments would be clearly
+insufficient. The inset in the plot for the first data set shows the
+distributions obtained applying maximum-entropy at the sample level: their
+differences aren't so glaring as in the full-network application.
 
@@
 \caption{***}
-\label{fig:plots_6mom}
+\label{fig:plots_2_4mom}
 \end{figure}

### 2019-06-19 maxNt_v5.tex
update

--- 2019-06-19 maxNt_v5.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-19T14:57:21+0200
+%% Last-Updated: 2019-06-19T15:34:37+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Thereafter we show the results for two moments and for four moments using
-$N=10\,000$.
+\bigskip
 
@@
 \caption{***}
-\label{fig:plots_6mom}
+\label{fig:plots_2_4mom}
 \end{figure}
 %maxent_solutions_different_moments_prior2_6moments.nb
+As an illustration of our approach in questions of sufficient statistics,
+\fig~\ref{fig:plots_2_4mom} shows the results for two constrained moments
+(equivalent to constraining means and correlations only) and for four
+constrained moments, with $N=10\,000$. In either data set we obtain two
+very distinct distributions. For the first data set in particular, four
+moments lead to a bimodal distribution, whereas two to a unimodal one,
+showing that two moments would not be a sufficient statistics. The inset in
+the plot for the first data set shows, for comparison, the two
+distributions obtained applying maximum-entropy at the sample level: their
+differences aren't so glaring as in the full-network application.
 

### 2019-06-19 maxNt_v5.tex
update

--- 2019-06-19 maxNt_v5.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-19T14:57:21+0200
+%% Last-Updated: 2019-06-19T19:49:04+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 ]{newpxmath}
-\useosf\linespread{1.083}
+\linespread{1.083}%\useosf
 %% smaller operators for old version of newpxmath
@@
 this method that are seldom made in the literature. First, the distribution
-$\px(\yA)$ given by this method is the zeroth-order approximation of four
-different distributions for the full network:
+$\px(\yA)$ given by this method is the zeroth-order approximation \citep[in
+the sense of Laplace's
+method:][\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
+four different distributions for the full network:
 \begin{itemize}
@@
 \end{equation}
-for any $m$s we wish.
+for all $m$ we wish.
 
@@
 
+For concreteness's sake we'll consider the maximum-entropy distribution
+$\px(\yA)$ as \emph{the most probable frequency distribution} for the
+full-network activity during the recording; but remember that it is also
+the approximation of three other distributions, as discussed in the
+previous section.
+
 \bigskip
 
-We first calculate the maximum-entropy distribution for the full network by
-using six moments. This number already provides as much information as the
-full frequency distribution of the sample, and at the same time illustrates
-the use of the approach in questions of statistic sufficiency (typically
-limited to two or three moments). Figure~\ref{fig:plots_6mom} shows the
-resulting densities (distribution/activity-step, that is,
-distribution${}\times N$) for full-network sizes $N=n$, $N=1\,000$,
-$N=5\,000$, $N=10\,000$\mynote{motivate?}. The case $N=n$ corresponds to
-applying the maximum-entropy method at the sample level; it can be observed
-that with six moments it reproduces almost exactly the measured frequency
+We first calculate the distribution by using six moments. This number
+already provides as much information as the full frequency distribution of
+the sample, and at the same time illustrates the use of the approach in
+questions of statistic sufficiency (typically limited to two or three
+moments). Figure~\ref{fig:plots_6mom} shows the resulting densities
+(distribution/activity-step, that is, distribution${}\times N$) for
+full-network sizes $N=n$, $N=1\,000$, $N=5\,000$,
+$N=10\,000$\mynote{motivate?}. The case $N=n$ corresponds to applying the
+maximum-entropy method at the sample level; it can be observed that with
+six moments it reproduces almost exactly the measured frequency
 distribution.
@@
 But also the second data set presents, upon closer inspection, a small
-shoulder on the right of the mode, suggestive of two activity regimes.
-These features are clearly not present in the sample or in the
-maximum-entropy distribution at the sample level. The application of the
-probability calculus to make forecasts about the full network thus reveals
-interesting and useful features.
-
-Thereafter we show the results for two moments and for four moments using
-$N=10\,000$.
-
+shoulder on the right of the mode, suggestive of two activity regimes. We
+frame no hypotheses about the biological cause of these two modes (they
+could stem from the presence of different kinds of cells or modules). These
+features are clearly not present in the sample or in the maximum-entropy
+distribution at the sample level. The application of the probability
+calculus to make forecasts about the full network thus reveals interesting
+and useful features.
+
+\bigskip
+
+To illustrate how our approach can be applied to studies of sufficient
+statistics, \fig~\ref{fig:plots_2_4mom} shows the results for two
+constrained moments (equivalent to constraining means and correlations
+only) and for four constrained moments, with $N=10\,000$. In either data
+set we obtain two very distinct distributions. For the first data set in
+particular, four moments lead to a bimodal distribution, whereas two to a
+unimodal one, showing that two moments would not be a sufficient
+statistics. For comparison, the two distributions obtained applying
+maximum-entropy \emph{at the sample level} are shown as an inset in the
+plot for the first data set: their differences aren't so glaring as in the
+full-network application.
 \begin{figure}[!p]
@@
 \caption{***}
-\label{fig:plots_6mom}
+\label{fig:plots_2_4mom}
 \end{figure}
 %maxent_solutions_different_moments_prior2_6moments.nb
+
+\section{Marginalization to the sample level}
+\label{sec:marginalization}
+
+\section{Assumptions and corrections}
+\label{sec:assumptions_formulae}
+
+in depth discussion about underlying assumptions, derivation from
+probability calculus, higher-order approximations beyond maxent and
+"error bars" around the distributions
+
+\section{Summary and discussion}
+\label{sec:summary_discussion}
+
+\clearpage
+\hrule
+OLD TEXT
+\hrule
 

### 2019-06-22 scripts/mctest1.R
update

@@ -1,0 +1,64 @@
+## this script computes the probability for the conditional frequencies
+## frequency(symptom | snp)
+## where each symptom can assume values listed in 'symptomvariants'
+## and each snp can assume values (alleles) listed in 'snpvariants'
+## such values can be tuples
+library('ggplot2')
+library('RColorBrewer')
+library('cowplot')
+library('png')
+library('plot3D')
+library('doParallel')
+library('LaplacesDemon')
+library('dplyr')
+
+filename <- 'mcsamples.csv'
+
+cores <- 1
+thinning <- 1
+mciterations <- 10000
+mcdiscard <- 10
+
+## setup parallel processing
+`%dox%` <- `%do%`
+if(cores>1){
+`%dox%` <- `%dopar%`
+cl <- makeCluster(cores, outfile="")
+registerDoParallel(cl)
+}
+
+### Here is the set-up for the Markov-chain Monte Carlo sampling of the parameters
+### Uses package LaplacesDemon
+
+        PGF <- function(data){1}
+
+        mydata <- list(y=1, PGF=PGF,
+               parm.names=c('x'),
+               mon.names=c('')
+               )
+
+        logprob <- function(parm,data){
+            parm <- interval(parm,0,1)
+            LP <- 1
+            list(LP=LP, Dev=-2*LP, Monitor=1, yhat=1, parm=parm)
+}
+
+        ##Initial.Values <- c(0,0) #log(mf/sum(mf)) #GIV(prob, mydata, n=1000, PGF=T)
+        
+        Initial.Values <- GIV(logprob, mydata, n=1000, PGF=F)
+
+        usamples <- LaplacesDemon(logprob, mydata, Initial.Values,
+                        Covar=NULL,
+                        Thinning=thinning,
+                        Iterations=mciterations, Status=mciterations,
+                        #Chains=nchains,CPUs=nchains,LogFile=paste0(filename,'_LDlog'), #Packages=c('Matrix'),#Type="MPI",
+                        ##Algorithm="RDMH"#, Specs=list(B=list(1:d,d1:d2,d3:dnp))
+                        ##Algorithm="Slice", Specs=list(B=NULL, Bounds=c(0,1), m=Inf, Type="Continuous", w=0.001)
+                        Algorithm="AFSS", Specs=list(A=mcdiscard, B=NULL, m=100, n=0, w=1)
+                        )
+    samples2 <- usamples$Posterior2
+    samples1 <- usamples$Posterior1
[…]

### 2019-06-22 scripts/mctest2.R

@@ -1,0 +1,64 @@
+## this script computes the probability for the conditional frequencies
+## frequency(symptom | snp)
+## where each symptom can assume values listed in 'symptomvariants'
+## and each snp can assume values (alleles) listed in 'snpvariants'
+## such values can be tuples
+library('ggplot2')
+library('RColorBrewer')
+library('cowplot')
+library('png')
+library('plot3D')
+library('doParallel')
+library('LaplacesDemon')
+library('dplyr')
+
+filename <- 'mcsamples.csv'
+
+cores <- 1
+thinning <- 1
+mciterations <- 10000
+mcdiscard <- 10
+
+## setup parallel processing
+`%dox%` <- `%do%`
+if(cores>1){
+`%dox%` <- `%dopar%`
+cl <- makeCluster(cores, outfile="")
+registerDoParallel(cl)
+}
+
+### Here is the set-up for the Markov-chain Monte Carlo sampling of the parameters
+### Uses package LaplacesDemon
+
+        PGF <- function(data){1}
+
+        mydata <- list(y=1, PGF=PGF,
+               parm.names=c('y'),
+               mon.names=c('')
+               )
+
+        logprob <- function(parm,data){
+            parm <- interval(parm,0,1)
+            LP <- log(2*parm)
+            list(LP=LP, Dev=-2*LP, Monitor=1, yhat=1, parm=parm)
+}
+
+        ##Initial.Values <- c(0,0) #log(mf/sum(mf)) #GIV(prob, mydata, n=1000, PGF=T)
+        
+        Initial.Values <- GIV(logprob, mydata, n=1000, PGF=F)
+
+        usamples <- LaplacesDemon(logprob, mydata, Initial.Values,
+                        Covar=NULL,
+                        Thinning=thinning,
+                        Iterations=mciterations, Status=mciterations,
+                        #Chains=nchains,CPUs=nchains,LogFile=paste0(filename,'_LDlog'), #Packages=c('Matrix'),#Type="MPI",
+                        ##Algorithm="RDMH"#, Specs=list(B=list(1:d,d1:d2,d3:dnp))
+                        ##Algorithm="Slice", Specs=list(B=NULL, Bounds=c(0,1), m=Inf, Type="Continuous", w=0.001)
+                        Algorithm="AFSS", Specs=list(A=mcdiscard, B=NULL, m=100, n=0, w=1)
+                        )
+    samples2 <- usamples$Posterior2
[…]

### 2019-06-22 scripts/mctestf1.R

@@ -1,0 +1,184 @@
+## this script computes the probability for the conditional frequencies
+## frequency(symptom | snp)
+## where each symptom can assume values listed in 'symptomvariants'
+## and each snp can assume values (alleles) listed in 'snpvariants'
+## such values can be tuples
+#library('ggplot2')
+#library('RColorBrewer')
+#library('cowplot')
+#library('png')
+#library('plot3D')
+library('doParallel')
+library('foreach')
+library('LaplacesDemon')
+#library('dplyr')
+
+#pre <- FALSE
+pre <- TRUE
+
+filename <- paste0(if(pre){'pre_'},'x4testmc_p1_')
+
+cores <- 30
+thinning <- 10
+mciterations <- 100
+mcstatus <- 10
+mcdiscard <- 10
+nn <- 10000
+## f <- choose(n,(0:n)+1)
+## f <- f/sum(f)
+f <- unlist(read.csv('stensola_activity_freqs_bins417641.csv',header=FALSE,sep=','))
+ind <- which(f>0)
+n <- length(f)-1
+L <- 10
+T <- 417641
+smooth <- nn*10
+#prior <- (1-(0:nn)+nn)/((nn+1)*(nn+2))
+poste <- unlist(read.csv('stensola_6mom_prior2_N10000.csv',header=FALSE,sep=','))
+#dirich <- nn*(1/(nn+1)+unlist(read.csv('stensola_6mom_prior2_N1000.csv',header=FALSE,sep=',')))
+
+nchains <- cores
+
+#if(cores>1){registerDoMC(cores=cores)}
+
+
+##condp <- function(n,a,nn,aa){exp(lchoose(n,a)+lchoose(nn-n,aa-a)-lchoose(nn,aa))}
+## G <- t(sapply(0:n, function(a){
+##     exp(lchoose(n,a)+lchoose(nn-n,(0:nn)-a)-lchoose(nn,(0:nn)))
+## }))
+
+# this is slower
+#condtt <- sapply(0:nn, function(aa){
+#    exp(lchoose(n,(0:n))+lchoose(nn-n,aa-(0:n))-lchoose(nn,aa))
+#})
+
+if(!pre){
+## setup parallel processing
+## `%dox%` <- `%do%`
+ if(cores>1){
+## `%dox%` <- `%dopar%`
+ cl <- makeCluster(cores, outfile="")
+     registerDoParallel(cl)
+ }
+print('calculating G...')
[…]

### 2019-06-22 scripts/mctestf2.R

@@ -1,0 +1,186 @@
+## this script computes the probability for the conditional frequencies
+## frequency(symptom | snp)
+## where each symptom can assume values listed in 'symptomvariants'
+## and each snp can assume values (alleles) listed in 'snpvariants'
+## such values can be tuples
+#library('ggplot2')
+#library('RColorBrewer')
+#library('cowplot')
+#library('png')
+#library('plot3D')
+library('doParallel')
+library('foreach')
+library('LaplacesDemon')
+#library('dplyr')
+
+pre <- FALSE
+#pre <- TRUE
+
+filename <- paste0(if(pre){'pre_'},'testmc_c6_p1_')
+
+cores <- 30
+thinning <- 10
+mciterations <- 100
+mcstatus <- 10
+mcdiscard <- 10
+nn <- 1000
+n <- 6
+moms <- unlist(read.csv('stensola_norm_allbinomeans.csv',header=FALSE,sep=','))
+moms <- moms[1:n]
+L <- 500
+T <- 417641^2
+smooth <- nn^2
+#prior <- (1-(0:nn)+nn)/((nn+1)*(nn+2))
+poste <- unlist(read.csv('stensola_6mom_prior2_N1000.csv',header=FALSE,sep=','))
+#dirich <- nn*(1/(nn+1)+unlist(read.csv('stensola_6mom_prior2_N1000.csv',header=FALSE,sep=',')))
+
+nchains <- cores
+
+#if(cores>1){registerDoMC(cores=cores)}
+
+
+##condp <- function(n,a,nn,aa){exp(lchoose(n,a)+lchoose(nn-n,aa-a)-lchoose(nn,aa))}
+## G <- t(sapply(0:n, function(a){
+##     exp(lchoose(n,a)+lchoose(nn-n,(0:nn)-a)-lchoose(nn,(0:nn)))
+## }))
+
+# this is slower
+#condtt <- sapply(0:nn, function(aa){
+#    exp(lchoose(n,(0:n))+lchoose(nn-n,aa-(0:n))-lchoose(nn,aa))
+#})
+
+if(!pre){
+## setup parallel processing
+## `%dox%` <- `%do%`
+ if(cores>1){
+## `%dox%` <- `%dopar%`
+ cl <- makeCluster(cores, outfile="")
+     registerDoParallel(cl)
+ }
+print('calculating G...')
+    G <- foreach(m=1:n, .combine='cbind'#, .export=c('moms')
+               ) %:% 
[…]

### 2019-06-22 scripts/parallel.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 #################### Set the name of the job
-#SBATCH --job-name marg2e3
+#SBATCH --job-name testmc2
 # Launch an array of 100 jobs
 ## SBATCH --array 1-10
 # Specify a time limit
@@ -8,8 +8,8 @@
 # Redirect stderr and stdout to the same file:
 # %A will be replaced by the job ID and %a by the array index
 ####################
-#SBATCH -o marg2e3.out
-#SBATCH -e marg2e3.out
+#SBATCH -o testmc2.out
+#SBATCH -e testmc2.out
 # Send email notifications
 ## SBATCH --mail-type=ALL
 # We request an exclusive node for every job in the array
@@ -27,4 +27,4 @@
 # we use slurm's environment variables to create unique output files and echo the name of the executing node in that file
 ##module load anaconda
 ##source activate r
-srun Rscript -e 'source("marginalize.R")'
+srun Rscript -e 'source("mctestf1.R")'

### 2019-06-22 scripts/parallel2.sh

@@ -1,0 +1,30 @@
+#!/bin/bash
+#################### Set the name of the job
+#SBATCH --job-name testmc3
+# Launch an array of 100 jobs
+## SBATCH --array 1-10
+# Specify a time limit
+# SBATCH --time 96:00:00
+# Redirect stderr and stdout to the same file:
+# %A will be replaced by the job ID and %a by the array index
+####################
+#SBATCH -o testmc3.out
+#SBATCH -e testmc3.out
+# Send email notifications
+## SBATCH --mail-type=ALL
+# We request an exclusive node for every job in the array
+## SBATCH --exclusive
+# reserve MB of memory
+#SBATCH --mem=60000
+# Specify the number of tasks (processes)
+#SBATCH --ntasks 1
+# Our job is multithreaded, so we ask 4 CPUs per process
+#SBATCH --cpus-per-task=31
+# So, in total we will have 4x8 CPUs for us
+## module load nest/2.10.0
+##################################################################
+# from here on we can run whatever command we want (e.g., srun python simulate.py $SLURM_ARRAY_JOB_ID)
+# we use slurm's environment variables to create unique output files and echo the name of the executing node in that file
+##module load anaconda
+##source activate r
+srun Rscript -e 'source("mctestf1.R")'

### 2019-06-22 scripts/mctestf2.R
update

@@ -3,6 +3,9 @@
 ## where each symptom can assume values listed in 'symptomvariants'
 ## and each snp can assume values (alleles) listed in 'snpvariants'
 ## such values can be tuples
+
+set.seed(181225)
+
 #library('ggplot2')
 #library('RColorBrewer')
 #library('cowplot')
@@ -13,25 +16,26 @@ library('foreach')
 library('LaplacesDemon')
 #library('dplyr')
 
+
 pre <- FALSE
 #pre <- TRUE
 
-filename <- paste0(if(pre){'pre_'},'testmc_c6_p1_')
+filename <- paste0(if(pre){'pre_'},'mc_6mom_unif_')
 
 cores <- 30
 thinning <- 10
-mciterations <- 100
-mcstatus <- 10
-mcdiscard <- 10
+mciterations <- 1000
+mcstatus <- 50
+mcdiscard <- 100
 nn <- 1000
 n <- 6
 moms <- unlist(read.csv('stensola_norm_allbinomeans.csv',header=FALSE,sep=','))
 moms <- moms[1:n]
-L <- 500
+L <- 1000
 T <- 417641^2
-smooth <- nn^2
+smooth <- round(nn^(3/2))
 #prior <- (1-(0:nn)+nn)/((nn+1)*(nn+2))
-poste <- unlist(read.csv('stensola_6mom_prior2_N1000.csv',header=FALSE,sep=','))
+poste <- unlist(read.csv(paste0('stensola_6mom_prior2_N',nn,'.csv'),header=FALSE,sep=','))
 #dirich <- nn*(1/(nn+1)+unlist(read.csv('stensola_6mom_prior2_N1000.csv',header=FALSE,sep=',')))
 
 nchains <- cores
@@ -76,8 +80,9 @@ if(cores>1){
 if(pre){
 
     PGF <- function(data){nn <- data$nn
-    ff <- 0.9/(nn+1)+0.1*rdirichlet(1,rep(1,nn+1))
-    (1-cumsum(ff[-(nn+1)]))/c(1,1-cumsum(ff[-c(nn,nn+1)]))
+        ff <- 0.9/(nn+1)+0.1*rdirichlet(1,rep(1,nn+1))
+        ff <- ff[-(nn+1)]/sum(ff)
+    (1-cumsum(ff))/c(1,1-cumsum(ff[-nn]))
 }
 
 mydata <- list(y=1, PGF=PGF,
[…]

### 2019-06-24 scripts/ME_1fmoms.m
update

@@ -1,89 +1,0 @@
-(* Homog maxent with any number of moments *)
-
-Clear[ME1fmoms,ME1fmoms0,ME1fmomsc,ME1fmomsn]; 
-
-ME1fmoms[eemoms_, nn_, wp_:MachinePrecision,pg_:4] := 
-  Block[{ll=Length[eemoms],js=Table[Unique["js"],Length[eemoms]], freeenm, soluzm, prob, bvalues,fmoms,emoms},
-	emoms=1/2-eemoms;
-	fmoms[s_]=Table[1/2-Binomial[s,r]/Binomial[nn,r], {r,ll}];
-	freeenm = Log@Sum[(1-ss+nn)/((nn+1)*(nn+2))*
-			 Exp[js.(fmoms[ss]-emoms)],{ss, 0, nn}];
-
-  soluzm = 
-   NMinimize[freeenm, js, MaxIterations -> 10000, 
-    PrecisionGoal -> pg, AccuracyGoal -> Infinity,WorkingPrecision->wp];
-
-  prob = Table[
-    (1-ss+nn)/((nn+1)*(nn+2))*
-    Exp[(js /. soluzm[[2]]).fmoms[ss]] ,{ss, 0, nn}];
-  totalprob = Total[prob];prob=prob/totalprob;
-  
-  bvalues=(Total[Table[fmoms[ss],{ss,0,nn}]*prob]/emoms-1);
-  Print["rel. discrepancies %: ", bvalues*100//N];
-  
- {js /. soluzm[[2]],prob,totalprob}]
-
-
-ME1fmoms0[eemoms_, nn_, wp_:MachinePrecision,pg_:4] := 
-  Block[{ll=Length[eemoms],js=Table[Unique["js"],Length[eemoms]], freeenm, soluzm, prob, bvalues,fmoms,emoms},
-	emoms=1-eemoms;
-	fmoms[s_]=Table[1-Binomial[s,r]/Binomial[nn,r], {r,ll}];
-	freeenm = Log@Sum[Binomial[nn,ss]*
-			 Exp[js.(fmoms[ss]-emoms)],{ss, 0, nn}]/2^nn;
-
-  soluzm = 
-   NMinimize[freeenm, js, MaxIterations -> 10000, 
-    PrecisionGoal -> pg, AccuracyGoal -> Infinity,WorkingPrecision->wp];
-
-  prob = Table[
-    Binomial[nn,ss]*
[…]

### 2019-06-24 scripts/ME_4momsolution.m

@@ -1,120 +1,0 @@
-(* Homog maxent with 4 moments - corresponds to constraining variances *)
-
-Clear[MEhom4moments,MEhom4thmoment,MEhomvmoments,MEhom4momentsprob]; 
-
-MEhom4moments[mr1_, mr2_, mr3_, mr4_, nn_, wp_:MachinePrecision,pg_:6,mi_:10000] := 
-  Block[{totalprob,
-	 m1=Rationalize[mr1,0],m2=Rationalize[mr2,0],m3=Rationalize[mr3,0],m4=Rationalize[mr4,0],
-	js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
-	nn2=Binomial[nn,2], nn3=Binomial[nn,3], nn4=Binomial[nn,4], 
-	bvalues},
-       freeenm = Log@Sum[Binomial[nn, m]*
-			 Exp[js1*(m/nn-m1) + js2*(Binomial[m,2]/nn2-m2) +
-			     js3*(Binomial[m,3]/nn3-m3) +
-			     js4*(Binomial[m,4]/nn4-m4)],
-       {m, 0, nn}];
-  soluzm = 
-   NMinimize[freeenm, {js1, js2, js3, js4}, MaxIterations -> mi, 
-    PrecisionGoal -> pg, AccuracyGoal -> Infinity,WorkingPrecision->wp];
-  prob = Table[
-	  Binomial[nn, m]*Exp[js1*m/nn + js2*Binomial[m,2]/nn2 +
-		      js3*Binomial[m,3]/nn3 +
-		      js4*Binomial[m,4]/nn4]/. soluzm[[2]],
-	  {m, 0, nn}];
-  totalprob = Total[prob];prob=prob/totalprob;
-    bvalues={(ran.prob/nn)/m1-1,
-	     ((Binomial[ran,2]/nn2).prob)/m2-1,
-	     ((Binomial[ran,3]/nn3).prob)/m3-1,
-	     ((Binomial[ran,4]/nn4).prob)/m4-1};
- Print["rel. discrepancies %: ", bvalues*100//N];
- {js1 /. soluzm[[2]], js2 /. soluzm[[2]],
-  js3 /. soluzm[[2]], js4 /. soluzm[[2]],prob,totalprob}]
-
[…]

### 2019-06-24 scripts/ME_4momsolution2.m

@@ -1,93 +1,0 @@
-(* Homog maxent with 4 moments - corresponds to constraining variances *)
-
-Clear[MEhom4moments2,MEhom4thmoment,MEhomvmoments,MEhom4momentsprob]; 
-
-MEhom4moments2[m1_, m2_, m3_, m4_, nn_, wp_:MachinePrecision] := 
- Block[{totalprob,
-	zzzm, js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
-	nn2=Binomial[nn,2], nn3=Binomial[nn,3], nn4=Binomial[nn,4], 
-	bvalues},
-       zzzm = Sum[Binomial[nn, m]*
-		  Exp[js1*m + js2*Binomial[m,2] +
-		      js3*Binomial[m,3] +
-		      js4*Binomial[m,4]],
-       {m, 0, nn}];
-  freeenm = Log[zzzm] - js1*m1*nn - js2*m2*nn2 -js3*m3*nn3 -js4*m4*nn4;
-  soluzm = 
-   NMinimize[freeenm, {js1, js2, js3, js4}, MaxIterations -> 10000, 
-    PrecisionGoal -> 6, AccuracyGoal -> Infinity,WorkingPrecision->wp];
-  prob = Table[
-	  Binomial[nn, m]*
-	  		  Exp[js1*m + js2*Binomial[m,2] +
-		      js3*Binomial[m,3] +
-		      js4*Binomial[m,4]]/. soluzm[[2]],
-	  {m, 0, nn}];
-  totalprob = Total[prob];prob=prob/totalprob;
-    bvalues={(ran.prob/nn)/m1-1,
-	     ((Binomial[ran,2]/nn2).prob)/m2-1,
-	     ((Binomial[ran,3]/nn3).prob)/m3-1,
-	     ((Binomial[ran,4]/nn4).prob)/m4-1};
- Print["rel. discrepancies %: ", bvalues*100//N];
- {js1/nn /. soluzm[[2]], js2/nn2 /. soluzm[[2]],
-  js3/nn3 /. soluzm[[2]], js4/nn4 /. soluzm[[2]],prob,totalprob}]
-
-MEhom4thmoment[m1_, m2_, m4_, nn_, wp_:MachinePrecision] := 
- Block[{totcorr = meancorr,
[…]

### 2019-06-24 scripts/ME_4momsolution3.m

@@ -1,93 +1,0 @@
-(* Homog maxent with 4 moments - corresponds to constraining variances *)
-
-Clear[MEhom4moments3,MEhom4thmoment,MEhomvmoments,MEhom4momentsprob]; 
-
-MEhom4moments3[m1_, m2_, m3_, m4_, nn_, wp_:MachinePrecision] := 
- Block[{totalprob,
-	zzzm, js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
-	nn1=Sqrt[nn], nn2=Sqrt@Binomial[nn,2], nn3=Sqrt@Binomial[nn,3], nn4=Sqrt@Binomial[nn,4], 
-	bvalues},
-       zzzm = Sum[Binomial[nn, m]*
-		  Exp[js1*m/nn1 + js2*Binomial[m,2]/nn2 +
-		      js3*Binomial[m,3]/nn2 +
-		      js4*Binomial[m,4]/nn4],
-       {m, 0, nn}];
-  freeenm = Log[zzzm] - js1*m1*nn1 - js2*m2*nn2 -js3*m3*nn3 -js4*m4*nn4;
-  soluzm = 
-   NMinimize[freeenm, {js1, js2, js3, js4}, MaxIterations -> 10000, 
-    PrecisionGoal -> 6, AccuracyGoal -> Infinity,WorkingPrecision->wp];
-  prob = Table[
-	  Binomial[nn, m]*
-	  		  Exp[js1*m/nn1 + js2*Binomial[m,2]/nn2 +
-		      js3*Binomial[m,3]/nn3 +
-		      js4*Binomial[m,4]/nn4]/. soluzm[[2]],
-	  {m, 0, nn}];
-  totalprob = Total[prob];prob=prob/totalprob;
-    bvalues={(ran.prob/nn)/m1-1,
-	     ((Binomial[ran,2]/nn2^2).prob)/m2-1,
-	     ((Binomial[ran,3]/nn3^2).prob)/m3-1,
-	     ((Binomial[ran,4]/nn4^2).prob)/m4-1};
- Print["rel. discrepancies %: ", bvalues*100//N];
- {js1/nn1 /. soluzm[[2]], js2/nn2 /. soluzm[[2]],
-  js3/nn3 /. soluzm[[2]], js4/nn4 /. soluzm[[2]],prob,totalprob}]
-
[…]

### 2019-06-24 scripts/ME_4momsolution_prior2.m

@@ -1,120 +1,0 @@
-(* Homog maxent with 4 moments - corresponds to constraining variances *)
-
-Clear[MEhom4moments,MEhom4thmoment,MEhomvmoments,MEhom4momentsprob]; 
-
-MEhom4moments[mr1_, mr2_, mr3_, mr4_, nn_, wp_:MachinePrecision,pg_:4] := 
-  Block[{totalprob,
-	 	 m1=Rationalize[mr1,0],m2=Rationalize[mr2,0],m3=Rationalize[mr3,0],m4=Rationalize[mr4,0],
-	js1, js2, js3, js4, freeenm, varm, soluzm, prob,ran=Range[0,nn],
-	nn2=Binomial[nn,2], nn3=Binomial[nn,3], nn4=Binomial[nn,4], 
-	bvalues},
-       freeenm = Log@Sum[(1-m+nn)/((nn+1)*(nn+2))*
-			 Exp[js1*(m/nn-m1) + js2*(Binomial[m,2]/nn2-m2) +
-			     js3*(Binomial[m,3]/nn3-m3) +
-			     js4*(Binomial[m,4]/nn4-m4)],
-       {m, 0, nn}];
-  soluzm = 
-   NMinimize[freeenm, {js1, js2, js3, js4}, MaxIterations -> 10000, 
-    PrecisionGoal -> pg, AccuracyGoal -> Infinity,WorkingPrecision->wp];
-  prob = Table[
-	  (1-m+nn)/((nn+1)*(nn+2))*Exp[js1*m/nn + js2*Binomial[m,2]/nn2 +
-		      js3*Binomial[m,3]/nn3 +
-		      js4*Binomial[m,4]/nn4]/. soluzm[[2]],
-	  {m, 0, nn}];
-  totalprob = Total[prob];prob=prob/totalprob;
-    bvalues={(ran.prob/nn)/m1-1,
-	     ((Binomial[ran,2]/nn2).prob)/m2-1,
-	     ((Binomial[ran,3]/nn3).prob)/m3-1,
-	     ((Binomial[ran,4]/nn4).prob)/m4-1};
- Print["rel. discrepancies %: ", bvalues*100//N];
- {js1 /. soluzm[[2]], js2 /. soluzm[[2]],
-  js3 /. soluzm[[2]], js4 /. soluzm[[2]],prob,totalprob}]
-
[…]

### 2019-06-24 scripts/ME_fmoms.m

@@ -1,23 +1,0 @@
-(* Homog maxent with 4 moments - corresponds to constraining variances *)
-
-Clear[MEfmoms]; 
-
-MEfmoms[emoms_, nn_, wp_:MachinePrecision,pg_:4] := 
-  Block[{ll=Length[emoms],js=Table[Unique["js"],Length[emoms]], freeenm, soluzm, prob, bvalues,fmoms},
-	fmoms[s_]=Table[Binomial[s,r]/Binomial[nn,r], {r,ll}];
-	freeenm = Log@Sum[(1-ss+nn)/((nn+1)*(nn+2))*
-			 Exp[js.(fmoms[ss]-emoms)],{ss, 0, nn}];
-
-  soluzm = 
-   NMinimize[freeenm, js, MaxIterations -> 10000, 
-    PrecisionGoal -> pg, AccuracyGoal -> Infinity,WorkingPrecision->wp];
-
-  prob = Table[
-    (1-ss+nn)/((nn+1)*(nn+2))*
-    Exp[(js /. soluzm[[2]]).fmoms[ss]] ,{ss, 0, nn}];
-  totalprob = Total[prob];prob=prob/totalprob;
-  
-  bvalues=(Total[Table[fmoms[ss],{ss,0,nn}]*prob]/emoms-1);
-  Print["rel. discrepancies %: ", bvalues*100//N];
-  
- {js /. soluzm[[2]],prob,totalprob}]

### 2019-06-24 scripts/ME_freqs.m

@@ -1,25 +1,0 @@
-(* Homog maxent with 4 moments - corresponds to constraining variances *)
-
-Clear[MEfreqs]; 
-
-MEfreqs[freqs_, n_,nn_, wp_:MachinePrecision,pg_:4] := 
-  Block[{r=Length[freqs]-1,js=Table[Unique["js"],Length[freqs]], freeenm, varm, soluzm, prob, bvalues},
-	
-       freeenm = Log@Sum[(1-ss+nn)/((nn+1)*(nn+2))*
-			 Exp[js.(Table[Binomial[n,s]*Binomial[nn-nn,ss-s]/Binomial[nn,ss],{s,0,r}]-freqs)]
-       ,{ss, 0, nn}];
-
-  soluzm = 
-   NMinimize[freeenm, js, MaxIterations -> 10000, 
-    PrecisionGoal -> pg, AccuracyGoal -> Infinity,WorkingPrecision->wp];
-
-  prob = Table[
-    (1-ss+nn)/((nn+1)*(nn+2))*
-    Exp[js.(Table[Binomial[n,s]*Binomial[nn-nn,ss-s]/Binomial[nn,ss],{s,r}]-freqs)]
-	 /. soluzm[[2]],{ss, 0, nn}];
-  totalprob = Total[prob];prob=prob/totalprob;
-  bvalues=(Table[
-    Table[Binomial[n,s]*Binomial[nn-nn,ss-s]/Binomial[nn,ss],{ss,0,nn}].prob
-	   ,{s,0,r}]/freqs-1);
- Print["rel. discrepancies %: ", bvalues*100//N];
- {js /. soluzm[[2]],prob,totalprob}]

### 2019-06-24 scripts/ME_homsolution.m

@@ -1,96 +1,0 @@
-(* Homog maxent *)
-
-Clear[MEhomsolution,MEhomsolutiontruncated,MEhomprob]; 
-
-MEhomsolution::usage = "Input: pop-avgd activity [0,1], pop-avgd couple activity, pop size, machine precision. Output: Lagr multipl, Lagr multipl, distribution, Z"
-(* it uses normalized values in the optimization; usually gives more precise results than MEhomsolution2, although it reports difficulties in finding the optimum *)
-MEhomsolution[meanmm_, meancorr_, nn_, wp_:MachinePrecision, pg_:6] := 
-  Block[{totcorr = Rationalize[meancorr,0],
-	 totmm = Rationalize[meanmm,0],totalprob,
-   zzzm, js2, js1, freeenm, varm, soluzm, prob,ran=Range[0,nn],bvalues},
-  zzzm = Sum[Binomial[nn, m]*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn], {m, 0, nn}]/2^nn;
-  freeenm = Log[zzzm] - js2*totcorr - js1*totmm;
-  soluzm = 
-   NMinimize[freeenm, {js1, js2}, MaxIterations -> 10000, 
-    PrecisionGoal -> pg,Method->Automatic, AccuracyGoal -> Infinity,WorkingPrecision->wp];
-  prob = Table[
-    Binomial[nn, m]*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn]/2^nn /. soluzm[[2]], {m, 
-     0, nn}];
-  totalprob = Total[prob];prob=prob/totalprob;
-  bvalues={(ran.prob/nn)/meanmm-1,((ran^2-ran).prob/(nn^2-nn))/meancorr-1};
- Print["rel. discrepancies %: ", bvalues*100//N];
-  {js1 /. soluzm[[2]], js2/. soluzm[[2]], prob,totalprob}]
-
-MEhomsolutiont[meanmm_, meancorr_, nn_, wp_:MachinePrecision] := 
- Block[{totcorr = meancorr,
-   totmm = meanmm,totalprob,
-   js2, js1, freeenm, varm, soluzm, prob,ran=Range[0,nn],bvalues},
[…]

### 2019-06-24 scripts/ME_homsolution2.m

@@ -1,49 +1,0 @@
-(* Homog maxent *)
-
-Clear[MEhomsolution2,MEhomsolutiontruncated,MEhomprob]; 
-
-MEhomsolution2::usage = "Input: pop-avgd activity [0,1], pop-avgd couple activity, pop size, machine precision. Output: Lagr multipl, Lagr multipl, distribution, Z"
-(* it uses absolute values in the optimization; usually gives much less precise results than MEhomsolutio2, but it doesn't complain about difficulties in finding the optimum *)
-MEhomsolution2[meanmm_, meancorr_, nn_, wp_:MachinePrecision] := 
-  Block[{totcorr = meancorr*(nn^2-nn)/2,(* it uses absolute values in the optimization *)
-   totmm = meanmm*nn,totalprob,
-   zzzm, js2, js1, freeenm, varm, soluzm, prob,ran=Range[0,nn],bvalues},
-  zzzm = Sum[Binomial[nn, m]*Exp[js2*(m^2-m)/2 + js1*m], {m, 0, nn}]/2^nn;
-  freeenm = Log[zzzm] - js2*totcorr - js1*totmm;
-  soluzm = 
-   NMinimize[freeenm, {js1, js2}, MaxIterations -> 10000, 
-    PrecisionGoal -> 6,Method->Automatic, AccuracyGoal -> Infinity,WorkingPrecision->wp];
-  prob = Table[
-    Binomial[nn, m]*Exp[js2*(m^2-m)/2 + js1*m]/2^nn /. soluzm[[2]], {m, 
-     0, nn}];
-  totalprob = Total[prob];prob=prob/totalprob;
-  bvalues={(ran.prob/nn)/meanmm-1,((ran^2-ran).prob/(nn^2-nn))/meancorr-1};
- Print["rel. discrepancies %: ", bvalues*100//N];
-  {js1*nn /. soluzm[[2]], js2*(nn^2-nn)/2 /. soluzm[[2]], prob,totalprob}]
-
- MEhomsolutiontruncated[meanmm_, meancorr_, nn_, trunc_, wp_:MachinePrecision] := 
- Block[{totcorr = meancorr,
-   totmm = meanmm,totalprob,
[…]

### 2019-06-24 scripts/ME_homsolution_prior2.m

@@ -1,96 +1,0 @@
-(* Homog maxent *)
-
-Clear[MEhomsolution,MEhomsolutiontruncated,MEhomprob]; 
-
-MEhomsolution::usage = "Input: pop-avgd activity [0,1], pop-avgd couple activity, pop size, machine precision. Output: Lagr multipl, Lagr multipl, distribution, Z"
-(* it uses normalized values in the optimization; usually gives more precise results than MEhomsolution2, although it reports difficulties in finding the optimum *)
-MEhomsolution[meanmm_, meancorr_, nn_, wp_:MachinePrecision,pg_:6] := 
-  Block[{totcorr = Rationalize[meancorr,0],
-	 totmm = Rationalize[meanmm,0],totalprob,
-   zzzm, js2, js1, freeenm, varm, soluzm, prob,ran=Range[0,nn],bvalues},
-       zzzm = Sum[(1-m+nn)/((nn+1)*(nn+2))*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn], {m, 0, nn}];
-  freeenm = Log[zzzm] - js2*totcorr - js1*totmm;
-  soluzm = 
-   NMinimize[freeenm, {js1, js2}, MaxIterations -> 10000, 
-    PrecisionGoal -> pg,Method->Automatic, AccuracyGoal -> Infinity,WorkingPrecision->wp];
-  prob = Table[
-    (1-m+nn)/((nn+1)*(nn+2))*Exp[js2*(m^2-m)/(nn^2-nn) + js1*m/nn] /. soluzm[[2]], {m, 
-     0, nn}];
-  totalprob = Total[prob];prob=prob/totalprob;
-  bvalues={(ran.prob/nn)/meanmm-1,((ran^2-ran).prob/(nn^2-nn))/meancorr-1};
- Print["rel. discrepancies %: ", bvalues*100//N];
-  {js1 /. soluzm[[2]], js2/. soluzm[[2]], prob,totalprob}]
-
-MEhomsolutiont[meanmm_, meancorr_, nn_, wp_:MachinePrecision] := 
- Block[{totcorr = meancorr,
-   totmm = meanmm,totalprob,
[…]

### 2019-06-24 scripts/marginalize_different_N.R

@@ -1,45 +1,0 @@
-library('doParallel')
-
-cores <- 1
-
-condp <- function(n,a,nn,aa){exp(lchoose(n,a)+lchoose(nn-n,aa-a)-lchoose(nn,aa))}
-
-## setup parallel processing
-`%dox%` <- `%do%`
-if(cores>1){
-`%dox%` <- `%dopar%`
-cl <- makeCluster(cores, outfile="")
-registerDoParallel(cl)
-}
-
-for(nn in c(1000)){message(nn)
-
-    data1a <- unlist(c(read.csv(paste0('stensoladist2mom',nn,'.csv'),header=F)))
-    data1b <- unlist(c(read.csv(paste0('stensoladist4mom',nn,'.csv'),header=F)))
-    n <- 65
-
-    result <- foreach(a=0:n, .combine='cbind') %do% {
-    table <- foreach(aa=0:nn, .combine='c', .export=c('condp','nn','n','a','data1a','data1b')) %dox% {condp(n,a,nn,aa)}
-    
-    c(sum(table*data1a),
-      sum(table*data1b))
-    }
-    write.csv(result[1,],paste0('stensola2mom_marg_from',nn,'to',n,'.csv'),row.names=F)
-    write.csv(result[2,],paste0('stensola4mom_marg_from',nn,'to',n,'.csv'),row.names=F)
-    
-
-    data2a <- unlist(c(read.csv(paste0('riehledist2mom',nn,'.csv'),header=F)))
-    data2b <- unlist(c(read.csv(paste0('riehledist4mom',nn,'.csv'),header=F)))
-    n <- 159
-
-    result <- foreach(a=0:n, .combine='cbind') %do% {
-    table <- foreach(aa=0:nn, .combine='c', .export=c('condp','nn','n','a','data1a','data1b')) %dox% {condp(n,a,nn,aa)}
-    
-    c(sum(table*data2a),
-      sum(table*data2b))
-    }
-    write.csv(result[1,],paste0('riehle2mom_marg_from',nn,'to',n,'.csv'),row.names=F)
-    write.csv(result[2,],paste0('riehle4mom_marg_from',nn,'to',n,'.csv'),row.names=F)
-}
-    
-if(cores>1){stopCluster(cl)}

### 2019-06-24 scripts/mcmc.m

@@ -1,472 +1,0 @@
-(* ::Package:: *)
-
-(* Mathematica Package *)
-
-BeginPackage["MCMC`", {"Units`"}];
-
-MCMC::usage = "MCMC[plogexpr, paramspec, numsteps]
-
-Perform MCMC sampling of the supplied probability distribution.
-
-1. plogexpr should be an expression that gives the unnormalized log
-probability for a particular choice of parameter values.
-
-2. paramspec either gives the results of a previous MCMC run (w/ same
-plogexpr--just to add on more iterations), or lists the model parameters
-like so:
-{{param1, ival1, spread1, domain1}, ...}
-a) Each param should be symbolic.
-b) ival is the initial parameter value.
-c) spread is roughly how far to try to change the parameter each step in
-the Markov chain. In this routine we select new parameters values based
-on an exponential distribution of the form Exp[\[CapitalDelta]param/spread]. My
-Numerical Recipes book advises setting these spreads so that the average
-candidate acceptance is 10-40%.
-d) Each domain is either Reals or a list of all possible values the
-parameter can take on (needs to be a uniform grid).
-
-3. numsteps is the number of Markov chain steps to perform.";
-
-MCMCModelFit::usage = "MCMCModelFit[data, errors, model, paramspec, ivars, numsteps]
-
-Perform MCMC samping of the probability distribution resulting from
-modeling data with model, assuming Gaussian errors. Straightforward
-wrapper around MCMC and GetChisqExpr.
-
-1. data must be given as:
-{{ivar1, dvar1}, {ivar2, dvar2}, ..., {ivarN, dvarN}}
-where ivar is the independent variable, dvar is the dependent variable,
-and N is the number of data points. Either the ivars or dvars can be
-vector valued; if the independent variable is a vector, then we're just
-dealing with a function of multiple variables, and if the dependent variable
-is, then we have a vector field.
-
-2. errors must have the same length as data (= N), with each entry giving
-the errors in the corresponding dependent variable supplied in data. If
-each dvar is just a number, then so too should be each element of errors;
[…]

### 2019-06-24 scripts/mctest1.R

@@ -1,64 +1,0 @@
-## this script computes the probability for the conditional frequencies
-## frequency(symptom | snp)
-## where each symptom can assume values listed in 'symptomvariants'
-## and each snp can assume values (alleles) listed in 'snpvariants'
-## such values can be tuples
-library('ggplot2')
-library('RColorBrewer')
-library('cowplot')
-library('png')
-library('plot3D')
-library('doParallel')
-library('LaplacesDemon')
-library('dplyr')
-
-filename <- 'mcsamples.csv'
-
-cores <- 1
-thinning <- 1
-mciterations <- 10000
-mcdiscard <- 10
-
-## setup parallel processing
-`%dox%` <- `%do%`
-if(cores>1){
-`%dox%` <- `%dopar%`
-cl <- makeCluster(cores, outfile="")
-registerDoParallel(cl)
-}
-
-### Here is the set-up for the Markov-chain Monte Carlo sampling of the parameters
-### Uses package LaplacesDemon
-
-        PGF <- function(data){1}
-
-        mydata <- list(y=1, PGF=PGF,
-               parm.names=c('x'),
-               mon.names=c('')
-               )
-
-        logprob <- function(parm,data){
-            parm <- interval(parm,0,1)
-            LP <- 1
-            list(LP=LP, Dev=-2*LP, Monitor=1, yhat=1, parm=parm)
-}
-
-        ##Initial.Values <- c(0,0) #log(mf/sum(mf)) #GIV(prob, mydata, n=1000, PGF=T)
-        
-        Initial.Values <- GIV(logprob, mydata, n=1000, PGF=F)
-
-        usamples <- LaplacesDemon(logprob, mydata, Initial.Values,
-                        Covar=NULL,
-                        Thinning=thinning,
-                        Iterations=mciterations, Status=mciterations,
-                        #Chains=nchains,CPUs=nchains,LogFile=paste0(filename,'_LDlog'), #Packages=c('Matrix'),#Type="MPI",
-                        ##Algorithm="RDMH"#, Specs=list(B=list(1:d,d1:d2,d3:dnp))
-                        ##Algorithm="Slice", Specs=list(B=NULL, Bounds=c(0,1), m=Inf, Type="Continuous", w=0.001)
-                        Algorithm="AFSS", Specs=list(A=mcdiscard, B=NULL, m=100, n=0, w=1)
-                        )
-    samples2 <- usamples$Posterior2
-    samples1 <- usamples$Posterior1
[…]

### 2019-06-24 scripts/mctest2.R

@@ -1,64 +1,0 @@
-## this script computes the probability for the conditional frequencies
-## frequency(symptom | snp)
-## where each symptom can assume values listed in 'symptomvariants'
-## and each snp can assume values (alleles) listed in 'snpvariants'
-## such values can be tuples
-library('ggplot2')
-library('RColorBrewer')
-library('cowplot')
-library('png')
-library('plot3D')
-library('doParallel')
-library('LaplacesDemon')
-library('dplyr')
-
-filename <- 'mcsamples.csv'
-
-cores <- 1
-thinning <- 1
-mciterations <- 10000
-mcdiscard <- 10
-
-## setup parallel processing
-`%dox%` <- `%do%`
-if(cores>1){
-`%dox%` <- `%dopar%`
-cl <- makeCluster(cores, outfile="")
-registerDoParallel(cl)
-}
-
-### Here is the set-up for the Markov-chain Monte Carlo sampling of the parameters
-### Uses package LaplacesDemon
-
-        PGF <- function(data){1}
-
-        mydata <- list(y=1, PGF=PGF,
-               parm.names=c('y'),
-               mon.names=c('')
-               )
-
-        logprob <- function(parm,data){
-            parm <- interval(parm,0,1)
-            LP <- log(2*parm)
-            list(LP=LP, Dev=-2*LP, Monitor=1, yhat=1, parm=parm)
-}
-
-        ##Initial.Values <- c(0,0) #log(mf/sum(mf)) #GIV(prob, mydata, n=1000, PGF=T)
-        
-        Initial.Values <- GIV(logprob, mydata, n=1000, PGF=F)
-
-        usamples <- LaplacesDemon(logprob, mydata, Initial.Values,
-                        Covar=NULL,
-                        Thinning=thinning,
-                        Iterations=mciterations, Status=mciterations,
-                        #Chains=nchains,CPUs=nchains,LogFile=paste0(filename,'_LDlog'), #Packages=c('Matrix'),#Type="MPI",
-                        ##Algorithm="RDMH"#, Specs=list(B=list(1:d,d1:d2,d3:dnp))
-                        ##Algorithm="Slice", Specs=list(B=NULL, Bounds=c(0,1), m=Inf, Type="Continuous", w=0.001)
-                        Algorithm="AFSS", Specs=list(A=mcdiscard, B=NULL, m=100, n=0, w=1)
-                        )
-    samples2 <- usamples$Posterior2
[…]

### 2019-06-24 scripts/mctestf1.R

@@ -1,184 +1,0 @@
-## this script computes the probability for the conditional frequencies
-## frequency(symptom | snp)
-## where each symptom can assume values listed in 'symptomvariants'
-## and each snp can assume values (alleles) listed in 'snpvariants'
-## such values can be tuples
-#library('ggplot2')
-#library('RColorBrewer')
-#library('cowplot')
-#library('png')
-#library('plot3D')
-library('doParallel')
-library('foreach')
-library('LaplacesDemon')
-#library('dplyr')
-
-#pre <- FALSE
-pre <- TRUE
-
-filename <- paste0(if(pre){'pre_'},'x4testmc_p1_')
-
-cores <- 30
-thinning <- 10
-mciterations <- 100
-mcstatus <- 10
-mcdiscard <- 10
-nn <- 10000
-## f <- choose(n,(0:n)+1)
-## f <- f/sum(f)
-f <- unlist(read.csv('stensola_activity_freqs_bins417641.csv',header=FALSE,sep=','))
-ind <- which(f>0)
-n <- length(f)-1
-L <- 10
-T <- 417641
-smooth <- nn*10
-#prior <- (1-(0:nn)+nn)/((nn+1)*(nn+2))
-poste <- unlist(read.csv('stensola_6mom_prior2_N10000.csv',header=FALSE,sep=','))
-#dirich <- nn*(1/(nn+1)+unlist(read.csv('stensola_6mom_prior2_N1000.csv',header=FALSE,sep=',')))
-
-nchains <- cores
-
-#if(cores>1){registerDoMC(cores=cores)}
-
-
-##condp <- function(n,a,nn,aa){exp(lchoose(n,a)+lchoose(nn-n,aa-a)-lchoose(nn,aa))}
-## G <- t(sapply(0:n, function(a){
-##     exp(lchoose(n,a)+lchoose(nn-n,(0:nn)-a)-lchoose(nn,(0:nn)))
-## }))
-
-# this is slower
-#condtt <- sapply(0:nn, function(aa){
-#    exp(lchoose(n,(0:n))+lchoose(nn-n,aa-(0:n))-lchoose(nn,aa))
-#})
-
-if(!pre){
-## setup parallel processing
-## `%dox%` <- `%do%`
- if(cores>1){
-## `%dox%` <- `%dopar%`
- cl <- makeCluster(cores, outfile="")
-     registerDoParallel(cl)
- }
-print('calculating G...')
[…]

### 2019-06-24 scripts/mctestf2.R

@@ -1,189 +1,0 @@
-## this script computes the probability for the conditional frequencies
-## frequency(symptom | snp)
-## where each symptom can assume values listed in 'symptomvariants'
-## and each snp can assume values (alleles) listed in 'snpvariants'
-## such values can be tuples
-
-set.seed(181225)
-
-#library('ggplot2')
-#library('RColorBrewer')
-#library('cowplot')
-#library('png')
-#library('plot3D')
-library('doParallel')
-library('foreach')
-library('LaplacesDemon')
-#library('dplyr')
-
-
-pre <- FALSE
-#pre <- TRUE
-
-filename <- paste0(if(pre){'pre_'},'mc_6mom_unif_')
-
-cores <- 30
-thinning <- 10
-mciterations <- 1000
-mcstatus <- 50
-mcdiscard <- 100
-nn <- 1000
-n <- 6
-moms <- unlist(read.csv('stensola_norm_allbinomeans.csv',header=FALSE,sep=','))
-moms <- moms[1:n]
-L <- 1000
-T <- 417641^2
-smooth <- round(nn^(3/2))
-#prior <- (1-(0:nn)+nn)/((nn+1)*(nn+2))
-poste <- unlist(read.csv(paste0('stensola_6mom_prior2_N',nn,'.csv'),header=FALSE,sep=','))
-#dirich <- nn*(1/(nn+1)+unlist(read.csv('stensola_6mom_prior2_N1000.csv',header=FALSE,sep=',')))
-
-nchains <- cores
-
-#if(cores>1){registerDoMC(cores=cores)}
-
-
-##condp <- function(n,a,nn,aa){exp(lchoose(n,a)+lchoose(nn-n,aa-a)-lchoose(nn,aa))}
-## G <- t(sapply(0:n, function(a){
-##     exp(lchoose(n,a)+lchoose(nn-n,(0:nn)-a)-lchoose(nn,(0:nn)))
-## }))
-
-# this is slower
-#condtt <- sapply(0:nn, function(aa){
-#    exp(lchoose(n,(0:n))+lchoose(nn-n,aa-(0:n))-lchoose(nn,aa))
-#})
-
-if(!pre){
-## setup parallel processing
-## `%dox%` <- `%do%`
- if(cores>1){
-## `%dox%` <- `%dopar%`
- cl <- makeCluster(cores, outfile="")
-     registerDoParallel(cl)
- }
-print('calculating G...')
[…]

### 2019-06-24 scripts/me_solution.R

@@ -1,94 +1,0 @@
-## Calculation of expected value, std, thetas, spread of conditional frequencies
-## of each symptom given each SNP - extra data set
-
-#### PREAMBLE
-## libraries and colour-blind palette from http://www.sron.nl/~pault/
-##runfunction <- function(aa=1e3L){
-library('ggplot2')
-library('RColorBrewer')
-library('cowplot')
-library('png')
-library('plot3D')
-library('doParallel')
-library('LaplacesDemon')
-library('dplyr')
-library('GA')
-#
-mypurpleblue <- '#4477AA'
-myblue <- '#66CCEE'
-mygreen <- '#228833'
-myyellow <- '#CCBB44'
-myred <- '#EE6677'
-myredpurple <- '#AA3377'
-mygrey <- '#BBBBBB'
-mycolours <- c(myblue, myred, mygreen, myyellow, myredpurple, mypurpleblue, mygrey, 'black')
-palette(mycolours)
-barpalette <- colorRampPalette(c(mypurpleblue,'white',myredpurple),space='Lab')
-barpalettepos <- colorRampPalette(c('white','black'),space='Lab')
-dev.off()
-mmtoin <- 0.0393701
-#### END PREAMBLE
-
-mehomsolution <- function(popsize,popmean,popcouple){
-    popsize2 <- (popsize^2-popsize)/2
-    rpopmean <- popmean*popsize
-    rpopcouple <- popcouple*popsize2
-
-    zterm <- function(js){
-        m <- 0:popsize
-        log(sum(exp(lchoose(popsize,m) + js[1]*m + js[2]*(m^2-m)/2))) - sum(js*c(rpopmean,rpopcouple))
-    }
-    
-    zterm <- function(js){
-        m <- 0:popsize
-        log(sum(choose(popsize,m)*exp( js[1]*m + js[2]*(m^2-m)/2))) - sum(js*c(rpopmean,rpopcouple))
-    }
-    
-    result <- nlm(f=zterm,p=c(0,0),iterlim=1e6)
-
-    result$estimate/c(popsize,popsize2)
-
-
-    # this seems to lead to solutions closer to the maximum
-    mehomsolution2 <- function(popsize,popmean,popcouple){
-    popsize2 <- (popsize^2-popsize)/2
-    rpopmean <- popmean*popsize
-    rpopcouple <- popcouple*popsize2
-
-    zterm <- function(js,cadd){
[…]

### 2019-06-24 scripts/parallel.sh

@@ -1,30 +1,0 @@
-#!/bin/bash
-#################### Set the name of the job
-#SBATCH --job-name testmc2
-# Launch an array of 100 jobs
-## SBATCH --array 1-10
-# Specify a time limit
-# SBATCH --time 96:00:00
-# Redirect stderr and stdout to the same file:
-# %A will be replaced by the job ID and %a by the array index
-####################
-#SBATCH -o testmc2.out
-#SBATCH -e testmc2.out
-# Send email notifications
-## SBATCH --mail-type=ALL
-# We request an exclusive node for every job in the array
-## SBATCH --exclusive
-# reserve MB of memory
-#SBATCH --mem=60000
-# Specify the number of tasks (processes)
-#SBATCH --ntasks 1
-# Our job is multithreaded, so we ask 4 CPUs per process
-#SBATCH --cpus-per-task=31
-# So, in total we will have 4x8 CPUs for us
-## module load nest/2.10.0
-##################################################################
-# from here on we can run whatever command we want (e.g., srun python simulate.py $SLURM_ARRAY_JOB_ID)
-# we use slurm's environment variables to create unique output files and echo the name of the executing node in that file
-##module load anaconda
-##source activate r
-srun Rscript -e 'source("mctestf1.R")'

### 2019-06-24 scripts/parallel2.sh

@@ -1,30 +1,0 @@
-#!/bin/bash
-#################### Set the name of the job
-#SBATCH --job-name testmc3
-# Launch an array of 100 jobs
-## SBATCH --array 1-10
-# Specify a time limit
-# SBATCH --time 96:00:00
-# Redirect stderr and stdout to the same file:
-# %A will be replaced by the job ID and %a by the array index
-####################
-#SBATCH -o testmc3.out
-#SBATCH -e testmc3.out
-# Send email notifications
-## SBATCH --mail-type=ALL
-# We request an exclusive node for every job in the array
-## SBATCH --exclusive
-# reserve MB of memory
-#SBATCH --mem=60000
-# Specify the number of tasks (processes)
-#SBATCH --ntasks 1
-# Our job is multithreaded, so we ask 4 CPUs per process
-#SBATCH --cpus-per-task=31
-# So, in total we will have 4x8 CPUs for us
-## module load nest/2.10.0
-##################################################################
-# from here on we can run whatever command we want (e.g., srun python simulate.py $SLURM_ARRAY_JOB_ID)
-# we use slurm's environment variables to create unique output files and echo the name of the executing node in that file
-##module load anaconda
-##source activate r
-srun Rscript -e 'source("mctestf1.R")'

### 2019-06-24 scripts/riehle2mom_1000.m

@@ -1,1067 +1,0 @@
-solution = {-3276.0216806385019419577`12., 3266.0748522927663949072`12., 
-     {2.3127528918720995309`9.764723258042576*^-19, 
-      8.73717620178402731901`9.756524491798812*^-18, 
-      1.6595416806902263182598`9.748461728571172*^-16, 
-      2.11309365116512701713468`9.740530303148072*^-15, 
-      2.029153598310698793908005`9.732725787047318*^-14, 
-      1.567490737712539481664106`9.725043972762512*^-13, 
-      1.01465232900411006384021357`9.717480859298492*^-12, 
-      5.66090361143300819633717926`9.710032638871304*^-12, 
-      2.77884867328331150289856687`9.702695684661906*^-11, 
-      1.2192524915048345053719693378`9.695466539525075*^-10, 
-      4.8413514904468090294`9.688341905565416*^-10, 
-      1.75730906017421372619`9.681318634501777*^-9, 
-      5.87952084391474800143`9.674393718749632*^-9, 
-      1.825887366037138704639`9.667564283158114*^-8, 
-      5.294455151695888793687`9.660827577344886*^-8, 
-      1.4408037138222573182445`9.654180968577693*^-7, 
-      3.6962273216960887836453`9.647621935156463*^-7, 
-      8.9739195553371978969859`9.641148060254293*^-7, 
-      2.06909400892354236842921`9.634757026179644*^-6, 
-      4.54459163470728717814336`9.628446609025707*^-6, 
-      9.53521187034464663460558`9.622214673675927*^-6, 
-      0.00001915902109658590898548452`9.616059169137664, 
-      0.00003694955651535490978457224`9.60997812417844, 
-      0.00006853870095296498594026526`9.603969643241513, 
-      0.00012251084261844954358506205`9.598031902619619, 
-      0.00021138754174924396247130809`9.592163146867527, 
-      0.00035265124770785696981369504`9.586361685435753, 
-      0.00056965918272499522422609761`9.580625889509239, 
-      0.00089224497893855977224249565`9.574954189036266, 
-      0.00135677020328989527263118942`9.569345069933942, 
-      0.0020053856656025581420068589`9.563797071457907, 
-      0.00288430575747729252885687482`9.558308783724733, 
-      0.00404099359455301637423385675`9.552878845376569, 
-      0.00552029595963518441238999264`9.547505941378272, 
-      0.00735973908639680258915907341`9.542188800938185, 
[…]

### 2019-06-24 scripts/riehle2mom_2000.m

@@ -1,2063 +1,0 @@
-solution = {-6552.0686352146698293455`12., 6542.2365636417679637371`12., 
-     {5.38524704897105175324`9.463655365440044*^-38, 
-      4.0688530048600382175546`9.459536977079718*^-36, 
-      1.541389909090459922776544`9.455453220046103*^-34, 
-      3.9036043436366939931411471`9.45140348707773*^-33, 
-      7.43505622940975934725557117`9.447387186795696*^-32, 
-      1.1360475787914977493138461677`9.443403743154485*^-30, 
-      1.4505451117908695786`9.439452594916308*^-29, 
-      1.59192558584678181112`9.4355331951478*^-28, 
-      1.532942351740735842648`9.431645010737897*^-27, 
-      1.3157692581752792539792`9.42778752193588*^-26, 
-      1.01924695395640463214137`9.423960221908532*^-25, 
-      7.19763102878831789638059`9.420162616315519*^-25, 
-      4.672122512153230732343649`9.416394222902053*^-24, 
-      2.807240978665519143354044`9.412654571108042*^-23, 
-      1.57059174467569079432473861`9.4089432016929*^-22, 
-      8.22406281676435988434793041`9.405259666375292*^-22, 
-      4.048399942284606344766322269`9.401603527487097*^-21, 
-      1.8808491206790785049188561977`9.397974357640914*^-20, 
-      8.275672749830806531`9.3943717394105*^-20, 
-      3.4591853937519078665`9.39079526502352*^-19, 
-      1.37743190502667479591`9.387244536066058*^-18, 
-      5.23816522243507153833`9.383719163198341*^-18, 
-      1.906719484711098445395`9.380218765881185*^-17, 
-      6.657192167998627508461`9.376742972112668*^-17, 
-      2.2336426734580011008929`9.37329141817457*^-16, 
-      7.2145483123309376496096`9.36986374838819*^-16, 
-      2.24684313854176856482868`9.366459614879039*^-15, 
-      6.75689752704315124571138`9.36307867735014*^-15, 
-      1.964847825927778095134832`9.35972060286346*^-14, 
-      5.531866704021767921773559`9.356385065629171*^-14, 
-      1.509705680233468613467409`9.353071746802431*^-13, 
-      3.998281554718270122877712`9.34978033428729*^-13, 
-      1.02864851280323861041384635`9.346510522547513*^-12, 
-      2.57334128346435590718835225`9.34326201242395*^-12, 
-      6.26561059386341308496999143`9.34003451095824*^-12, 
[…]

### 2019-06-24 scripts/riehle2mom_3000.m

@@ -1,3065 +1,0 @@
-solution = {-9828.1074649852852415015`12., 9818.3057750917726106543`12., 
-     {1.253774154106195335659`9.28756326257194*^-56, 
-      1.42094179203363176606013`9.284813339601406*^-54, 
-      8.066894132838952915725453`9.282078904634824*^-53, 
-      3.05877977847421855531034356`9.279359775249402*^-51, 
-      8.714737713449937375498224926`9.27665577223629*^-50, 
-      1.9900013160969420826`9.273966719525514*^-48, 
-      3.7937967286369180378`9.271292444113083*^-47, 
-      6.210846715795174874139`9.26863277599019*^-46, 
-      8.913297814455477699792`9.26598754807447*^-45, 
-      1.139137774645836111305128`9.2633565961432*^-43, 
-      1.3126813120518795883372769`9.260739758768395*^-42, 
-      1.37769258502205490502654829`9.258136877253735*^-41, 
-      1.327881697031929766820863571`9.255547795573275*^-40, 
-      1.1836052967920599931945720099`9.252972360311851*^-39, 
-      9.8145917847483914418`9.250410420607137*^-39, 
-      7.60986265523004166315`9.247861828093308*^-38, 
-      5.541854053636123975753`9.245326436846254*^-37, 
-      3.8054566077058618271154`9.242804103330268*^-36, 
-      2.47250509057339811635688`9.24029468634621*^-35, 
-      1.52471539045468343542266`9.237798046981045*^-34, 
-      8.948828713176137404751634`9.235314048558736*^-34, 
-      5.0113700753799054276054999`9.232842556592486*^-33, 
-      2.6837711980473083713239168`9.230383438738183*^-32, 
-      1.377309093120466716486453498`9.227936564749147*^-31, 
-      6.786345351715694816765509166`9.225501806432002*^-31, 
-      3.2159826980526040212443823649`9.223079037603751*^-30, 
-      1.4681148575232315589`9.220668134049939*^-29, 
-      6.4657338168099933401`9.218268973483907*^-29, 
-      2.75095297452566270488`9.215881435507095*^-28, 
-      1.132166552449313809464`9.213505401570362*^-27, 
-      4.512486791440778776055`9.211140754936297*^-27, 
-      1.7437443157066824987792`9.20878738064246*^-26, 
-      6.539778052021031801602`9.206445165465597*^-26, 
-      2.38276290809146940224515`9.204113997886699*^-25, 
-      8.44180283183676667189035`9.201793768056975*^-25, 
[…]

### 2019-06-24 scripts/riehle4mom_1000.m

@@ -1,1068 +1,0 @@
-solution = {-3496.3910425506497829162`12., 2902.9136828413948159499`12., 
-     42990.9504495166447986678`12., -160178.4992672097666502284`12., 
-     {3.6164084150285518996407`9.72215109447946*^-16, 
-      1.096009036232301311827017`9.714215461992431*^-14, 
-      1.6688229119695246781039476`9.706409396856536*^-13, 
-      1.70261513919057416795869856`9.698727990677027*^-12, 
-      1.309771733965096992362428384`9.691166581207664*^-11, 
-      8.105595211642411673863546513`9.68372073589353*^-11, 
-      4.204554979361288697336154342`9.67638623677403*^-10, 
-      1.88079147186424601706`9.66915906661358*^-9, 
-      7.40807057742313962254`9.662035396142238*^-9, 
-      2.610686338548518162583`9.655011572301609*^-8, 
-      8.336516178116664595692`9.648084107402566*^-8, 
-      2.4370144320156209699046`9.641249669111394*^-7, 
-      6.5777313786097716740728`9.634505071189713*^-7, 
-      1.65105017360269201528532`9.627847264921227*^-6, 
-      3.87776402102852069283822`9.621273331165229*^-6, 
-      8.56747154669625906867221`9.614780472982796*^-6, 
-      0.00001788952182996503989444969`9.608366008787007, 
-      0.00003544930669132836355767767`9.602027365973191, 
-      0.00006690703498480478107907268`9.59576207498961, 
-      0.00012067529636572191408470642`9.589567763812555, 
-      0.00020861021640326886688249203`9.583442152793348, 
-      0.00034657094899195526935429687`9.577383049847738, 
-      0.00055469471667993939369572312`9.571388345960784, 
-      0.00085723075094105359574707932`9.565456010982878, 
-      0.00128180416342483832379923052`9.5595840896946, 
-      0.00185804159054090682060740865`9.553770698120159, 
-      0.00261557816564235273067196968`9.548014020070875, 
-      0.00358156643608597713335779636`9.542312303901742, 
-      0.00477790398237743934839707811`9.536663859465591, 
-      0.00621846874183942338156683312`9.531067055250672, 
-      0.00790668408373837495761258385`9.525520315688517, 
-      0.00983372140201637997482977094`9.520022118620236, 
-      0.01197758721382667736981033763`9.51457099291018, 
-      0.01430324391905929957764162104`9.509165516196836, 
[…]

### 2019-06-24 scripts/riehle4mom_2000.m

@@ -1,2068 +1,0 @@
-solution = {-7648.73959277321941686371928068205`16., 
-     17165.34516709959823988135954603357`16., 
-     11685.2759527350185965197786928489`16., 
-     -147285.03294881805915275675364843213`16., 
-     {4.07376950905684987128136561948343`13.367263352769942*^-26, 
-      1.778786192365427077357744`13.363411407149659*^-24, 
-      3.915012208855655375263651`13.359584790774546*^-23, 
-      5.79117923472777716253373732`13.355783127413934*^-22, 
-      6.477113925006658284598991361`13.352006049026306*^-21, 
-      5.8426375191322893861690539733`13.348253195519927*^-20, 
-      4.42773285061961446766638139266`13.344524214522165*^-19, 
-      2.899590773975145186738910656116`13.340818761157177*^-18, 
-      1.6750697898722872293729247095716`13.33713649783154*^-17, 
-      8.6718658316397607003199558439693`13.33347709402755*^-17, 
-      4.0735560973982500194897278431423`13.329840226103796*^-16, 
-      1.75381628434257509953217`13.32622557710278*^-15, 
-      6.97831710856117019805112`13.322632836565214*^-15, 
-      2.584060654064048202514702`13.319061700350753*^-14, 
-      8.958187996383293232843536`13.315511870464919*^-14, 
-      2.9223155874634473745431525`13.311983054891897*^-13, 
-      9.0107404585217890664661656`13.308474967433035*^-13, 
-      2.63646742469904476703624603`13.304987327550775*^-12, 
-      7.34548565174844561229349791`13.301519860217807*^-12, 
-      1.954787014048603273647948932`13.298072295771231*^-11, 
-      4.982714087921838861314614026`13.294644369771556*^-11, 
-      1.2195767788551284557178781505`13.291235822866298*^-10, 
-      2.8728752486975951649031779405`13.287846400658058*^-10, 
-      6.5266276703596987823441999225`13.284475853576824*^-10, 
-      1.4326786422462595165315224474`13.281123936756432*^-9, 
-      3.04405739247710563487234786642`13.27779040991494*^-9, 
-      6.2704383422142481943756829893`13.274475037238831*^-9, 
-      1.254088821885644253297104705722`13.271177587270824*^-8, 
-      2.438605029388865956368715186276`13.267897832801271*^-8, 
-      4.616288103883337585036432975911`13.264635550762867*^-8, 
[…]

### 2019-06-24 scripts/riehle4mom_margto1000.m

@@ -1,6 +1,0 @@
-qmarg = {0.00016757362195149664, 0.005316169478490606, 0.08488531343917713, 
-     0.9096309434603806}
-qmarg = {0.0001675736219514966, 0.005316169478490606, 0.08488531343917712, 
-     0.9096309434603808}
-qmarg = {0.00016757362195149664, 0.005316169478490606, 0.08488531343917713, 
-     0.9096309434603806}

### 2019-06-24 scripts/riehlesample4mom.m

@@ -1,596 +1,0 @@
-riehlesample4mom = {0.0006971622014384971112112985669456149758175143688870421\
-5826757559322434143915`58.59615272022401, 0.004904841709118510338078546236596\
-82477088814405544397522157420253405322392194`58.545760262969175, 0.0172003059\
-9239863758721185724205118954781423264707044442447460223627996306195`58.500567\
-76690646, 0.04028195228811163518463413381754252491189532791063766777976174821\
-894707209836`58.459536710526585, 0.071205263059448446429451160848901940644106\
-4437042388832120169371347044669567`58.42189843424312, 0.101786073622129899511\
-33667639726992774575383292456733105511370491437814648521`58.38706762004686, 0\
-.1230818266861406744698423468025264148691242195767887801464946764449534826144\
-7`58.354587941677934, 0.13001583817109741968788499272590039812115611112528268\
-159051242239334908468688`58.32409652443162, 0.1229391808052028278744738232148\
-0782718876579398099641990429118792341161431137`58.29529991012458, 0.106087149\
-75571430868220277223170434784645173259976198900343062797558198354256`58.26795\
-733147517, 0.0848717340747395788703622677947149168946147424850666692011090234\
-8424498762655`58.24186878302404, 0.063784845424410570858421007212640025532983\
-20212960485188594216292345067524449`58.216866328750605, 0.0455406075327151983\
-4697422850364481374790263822691543658314749096367326998778`58.1928076475845, 
-     0.0311895563137383326969563060083508368781703852384845244708471165875898\
-8410623`58.1695711596204, 0.0206637383151298449523054284858335571019827037446\
-7310031606152643906311324791`58.14705229009713, 0.013341642554415204163218783\
-26707798930120992150994635815026225051829889005911`58.12516056615153, 0.00844\
-971271862048671390115736731650703605909509913119232983870404030996714217`58.1\
-0381733228606, 0.005279717885371641494635918842865945014756811938761858341974\
-53668318237483143`58.082953931701304, 0.0032713888456208742766953572252413516\
-5167286146673140632745817871530559870427`58.062510242647136, 0.00201914288049\
-982319789690387988408602828102162527619134162584353104117359325`58.0424334882\
-7325, 0.001246365455092730453127311448713382489692753041963196147550165607304\
-2182182`58.02267725926482, 0.000772123400364863957682361327766860768948216540\
[…]

### 2019-06-24 scripts/stensola2mom_1000.m

@@ -1,1028 +1,0 @@
-solution = {-4492.3242675362289306417`12., 4482.4132898545298672193`12., 
-     {7.91049480882382492940182`10.245920169588343*^-6, 
-      0.00008855478022495072293256842`10.212842729433458, 
-      0.00049963532236660560823268416`10.18204799473724, 
-      0.00189437225335087367806430619`10.153237536519883, 
-      0.00543001046596453906883719244`10.126167884545389, 
-      0.01255126353402920705452869399`10.100637809304118, 
-      0.02436991410004108910458430064`10.076479082515338, 
-      0.04088220714041069442770645545`10.0535496334263, 
-      0.06048989901344833504487199744`10.031728392045089, 
-      0.08019338723911787409716781042`10.010911343680236, 
-      0.0964486203856138722022566098`9.99100846861537, 
-      0.10629665761519102946062402021`9.97194133882831, 
-      0.10824628638591237160821731032`9.953641209422251, 
-      0.10256571731826156785665190501`9.936047487382966, 
-      0.09096288670528169348707970491`9.919106491539287, 
-      0.07589620447317053861615569252`9.90277043969945, 
-      0.05984167552199169136054199095`9.886996614781005, 
-      0.04476253086483744240544677639`9.871746673268413, 
-      0.03187554578584053757605284233`9.856986067806504, 
-      0.02167578187149012157329206734`9.842683562045144, 
-      0.01411465642404770967850727911`9.828810820594862, 
-      0.00882329686349376817112523009`9.815342060557198, 
-      0.00530691307202063999265358193`9.802253753856865, 
-      0.0030775232034981838089741028`9.789524371739512, 
-      0.00172397226008121479270307176`9.777134164464652, 
-      0.00093450788430625434518160674`9.765064970531691, 
-      0.00049096936792351994826438634`9.75330005081194, 
-      0.00025037272778846636271894682`9.741823943783958, 
-      0.0001241013578150558368484851`9.730622338729932, 
-      0.00005986552225615228004782495`9.719681964283682, 
-      0.00002813869797707134225200621`9.70899049015261, 
-      0.00001290151042887337239506663`9.698536440187993, 
-      5.77614580985609244796574`9.68830911526665*^-6, 
-      2.52767148123571328409346`9.678298524684296*^-6, 
[…]

### 2019-06-24 scripts/stensola2mom_2000.m

@@ -1,2029 +1,0 @@
-solution = {-8984.458335089092200426`12., 8974.5617346377262884321`12., 
-     {6.254288766969503158307937147`9.944888418162927*^-11, 
-      1.40041768302555586537`9.928035278746478*^-9, 
-      1.574127194066642161159`9.911795866655991*^-8, 
-      1.1843041861433611615281`9.896126514217817*^-7, 
-      6.7093481423872238495371`9.880988055396363*^-7, 
-      3.05294953965692856960397`9.866345227537005*^-6, 
-      0.00001162277466172515831400089`9.852166169305178, 
-      0.00003807897505175728747525119`9.838421996837198, 
-      0.00010959750222130764606758603`9.82508644392277, 
-      0.00028151098857478048824779587`9.812135554950752, 
-      0.00065337762642330043262614681`9.799547421597557, 
-      0.00138411467510685312208798115`9.787301955987006, 
-      0.00269850240672702727640368293`9.775380694422683, 
-      0.00487576553393763342216950065`9.763766626877793, 
-      0.00821314322405807904886564485`9.752444048289629, 
-      0.01296414588223358774336704108`9.741398428395946, 
-      0.01926108474425174904092816907`9.730616297406243, 
-      0.0270408020461045611467811233`9.720085145251064, 
-      0.03599694418108770883512835727`9.70979333251912, 
-      0.04557863520474713725995722896`9.69973001149192, 
-      0.05504415454645637519078988157`9.689885055932672, 
-      0.0635626926139278233153036728`9.680248998489898, 
-      0.0703429149498906334963942346`9.670812974745893, 
-      0.07475901145721372524040370242`9.661568673081046, 
-      0.07644577564356084658778017427`9.652508289643364, 
-      0.0753433130093457957725536779`9.643624487811849, 
-      0.07168570016730366610697886513`9.634910361626023, 
-      0.06594164213169877062408644517`9.626359402724878, 
-      0.05872491957779981910189655543`9.61796547039877, 
-      0.05069612188990133590580330799`9.609722764408916, 
-      0.04247498938167560733758591464`9.601625800273256, 
-      0.03457647841101547515907396116`9.593669386754812, 
-      0.02737595132150500872562935649`9.585848605321157, 
-      0.02110196381113306424275790549`9.578158791371477, 
[…]

### 2019-06-24 scripts/stensola2mom_3000.m

@@ -1,3029 +1,0 @@
-solution = {-13476.5957602289250291442`12., 13466.7017593618299366635`12., 
-     {4.9450536240888997986249`9.768806083397697*^-16, 
-      1.660944811545583822196237`9.757498153465962*^-14, 
-      2.7968212956657890262585908`9.74646994527079*^-13, 
-      3.14802548934408574316757522`9.735707785060455*^-12, 
-      2.664576807970195091446769444`9.725198978002046*^-11, 
-      1.8091034557436583447973008979`9.714931716912574*^-10, 
-      1.02629553092884057782`9.704895001384962*^-9, 
-      5.00368262252023408745`9.695078565919847*^-9, 
-      2.140277813653743357685`9.685472815885978*^-8, 
-      8.159301796362427996728`9.676068770307596*^-8, 
-      2.8069409805726585582392`9.666858010623727*^-7, 
-      8.8018863878737450115031`9.657832634686548*^-7, 
-      2.53678992921510293722873`9.64898521536897*^-6, 
-      6.76684028877159936089221`9.640308763238044*^-6, 
-      0.00001680573058825192194836474`9.631796692824215, 
-      0.00003905892879783672318056651`9.62344279207855, 
-      0.00008533145904396469370207295`9.615241194663081, 
-      0.00017592342126078209916544123`9.607186354764682, 
-      0.00034345441798536218784315929`9.599273024161533, 
-      0.00063692448595711845745944356`9.591496231304763, 
-      0.001125083689380344453468716`9.583851262206348, 
-      0.00189778410232215226662048889`9.576333642949367, 
-      0.00306379523284637776881764341`9.568939123658016, 
-      0.00474375014088645035319450637`9.561663663783685, 
-      0.00705756220898055074905670595`9.554503418579413, 
-      0.01010678558345099902797547508`9.547454726649466, 
-      0.01395379044506264657612005414`9.540514098473047, 
-      0.01860094557527406827377274857`9.533678205812079, 
-      0.02397384143751447765251383255`9.526943871922619, 
-      0.02991260717365706832075161663`9.520308062497756, 
-      0.03617442193274817568587863377`9.513767877277443, 
-      0.04244850443509610357743520725`9.507320542267115, 
-      0.04838255002865275073976560432`9.5009634025129, 
-      0.05361730188864450941723893569`9.494693915386284, 
[…]

### 2019-06-24 scripts/stensola2mom_margto1000.m

@@ -1,6 +1,0 @@
-qmarg = {0.002473331771688187, 0.03056674695587842, 0.18887909233039415, 
-     0.7780808289420393}
-qmarg = {0.002473331752115446, 0.030566746714123173, 0.18887909084597623, 
-     0.7780808306877852}
-qmarg = {0.002473331771688187, 0.03056674695587842, 0.18887909233039415, 
-     0.7780808289420393}

### 2019-06-24 scripts/stensola4mom_1000.m

@@ -1,1032 +1,0 @@
-solution = {-5419.6389678540528791535`12., 28178.0964201717551818647`12., 
-     985508.741383959412433281`12., -2.5031294027926990031736749`12.*^7, 
-     {0.00280393134755988956370734367`10.093572468336756, 
-      0.01241789771715728419457546011`10.065315721404486, 
-      0.02906457364212882915843706918`10.038517883966996, 
-      0.04822020363073988904342681755`10.012998855779081, 
-      0.06413700392239619376221089162`9.988604975431608, 
-      0.07329627556044720922565956407`9.965203881342811, 
-      0.07527753486281266184204331432`9.942680587819266, 
-      0.07171647904685520693115311169`9.920934449167154, 
-      0.06488740062541725490500518443`9.899876783063382, 
-      0.05677138684713285015428326466`9.879428990240589, 
-      0.04871462174857456945778579718`9.859521052535706, 
-      0.04145626415533736004320326483`9.840090322644082, 
-      0.03529710800597576121454664751`9.821080541034187, 
-      0.03027493419707517139662285351`9.80244103133644, 
-      0.02629589142721978087358994873`9.78412603704891, 
-      0.02321668656595953385216496635`9.766094170892318, 
-      0.02088900901212392318438074894`9.748307954473846, 
-      0.01917966920547630432921858814`9.730733430689549, 
-      0.01797684923738874374063320792`9.713339834931295, 
-      0.01718911026368950474545492826`9.696099313966407, 
-      0.01674088769623622631006397007`9.678986683540204, 
-      0.01656634335330100390665545123`9.661979217468655, 
-      0.01660244653512568622765940231`9.645056462353214, 
-      0.01678180956053743699968752207`9.628200073145274, 
-      0.01702596767595115099337711833`9.611393665675077, 
-      0.017240370879043155805426629`9.5946226829852, 
-      0.01731318485996050033744350326`9.577874272905566, 
-      0.01712070675810454727181747077`9.561137174801425, 
-      0.01654208775768085276078179138`9.544401613837437, 
-      0.01548417552144179601194198749`9.527659201444319, 
-      0.01391297939428918190391657242`9.510902840961464, 
-      0.01188215680156781899004872504`9.494126637667689, 
[…]

### 2019-06-24 scripts/stensola4mom_2000.m

@@ -1,2031 +1,0 @@
-solution = {-11073.3425032128154787224`12., 68134.1436286657463508225`12., 
-     1.954354673527362413545413`12.*^6, -5.39774526459431462506552932`12.*^7, 
-     {0.00003229783582380820038945255`9.77720794285961, 
-      0.00025448233769449348741647171`9.76304542440075, 
-      0.00103680531275094557036235516`9.749247098647201, 
-      0.00291654234051088066685441054`9.735789665565916, 
-      0.00638154058408607994113801663`9.722651645807964, 
-      0.01160017579080585081452896874`9.709813187617028, 
-      0.01827030394605966589781165974`9.697255899140325, 
-      0.02567437868567287679027748586`9.684962702256223, 
-      0.032895992984126955949814578`9.672917704709029, 
-      0.03907870627965901514447978546`9.66110608788637, 
-      0.04361913461595346384864161776`9.64951400801606, 
-      0.04624594702842306571718299589`9.638128508919182, 
-      0.04699527544067604694591451485`9.626937444750551, 
-      0.04612430095475955658060910499`9.615929411400286, 
-      0.04400756674504106330620422653`9.605093685430559, 
-      0.04104710481547589896517810671`9.594420169588044, 
-      0.03761046852603559861204519736`9.583899344071408, 
-      0.03399769211549870353109635557`9.573522222849357, 
-      0.03043114902918669135889699434`9.563280314422528, 
-      0.02706020560405539755552154022`9.553165586504871, 
-      0.02397343191220924864178861954`9.54317043417003, 
-      0.02121316058091820729101569106`9.53328765106748, 
-      0.01878926643377339864965283693`9.523510403363828, 
-      0.01669067183234651286348808618`9.513832206107843, 
-      0.01489415544216130848902947755`9.504246901754971, 
-      0.01337063885734328431181964839`9.494748640618973, 
-      0.01208938783931758099575635866`9.485331863045912, 
-      0.01102062503441797051695116451`9.475991283129629, 
-      0.01013700712684124309805747361`9.466721873808424, 
-      0.00941433429398817799383986775`9.457518853200922, 
-      0.0088317686911109207206562393`9.448377672054574, 
-      0.00837175809618270708610054744`9.43929400219434, 
-      0.0080197964408379683308204932`9.43026372587088, 
[…]

### 2019-06-24 scripts/stensola4mom_margto1000.m

@@ -1,6 +1,0 @@
-qmarg = {0.02040222177199268, 0.10875463618073142, 0.29987318185155, 
-     0.570969960195726}
-qmarg = {0.020402221771729828, 0.10875463617933047, 0.2998731818477017, 
-     0.5709699602012381}
-qmarg = {0.02040222177199268, 0.10875463618073142, 0.29987318185155, 
-     0.570969960195726}

### 2019-06-24 scripts/stensolasample2mom.m

@@ -1,132 +1,0 @@
-stensolasample2mom = {0.46614974944969130398905583811`11.433018378318666, 
-     0.33838520817042186883083629227`11.08701691373688, 
-     0.13846905359961422285370115332`10.892299224176979, 
-     0.04257770123428630860488599408`10.754979802703232, 
-     0.01106478339043668026811837751`10.648183041894086, 
-     0.00259149617097024079985667298`10.560390715788968, 
-     0.00056966195024979590846220511`10.485592977305123, 
-     0.00012085277895696574947182265`10.420253813903445, 
-     0.00002525218262854800328860944`10.362115348361781, 
-     5.27781736808848804020823`10.30964728924825*^-6, 
-     1.11682285152229980811979`10.261764215830535*^-6, 
-     2.4161006317981038462811`10.217668113249838*^-7, 
-     5.386505059810145180896`10.176754957043418*^-8, 
-     1.245770221189831591402`10.138556422245188*^-8, 
-     3.00559476476563580276`10.102701970775408*^-9, 
-     7.6005554433274011172662033376`10.068893323612794*^-10, 
-     2.0227870204286554727893699882`10.036886758937671*^-10, 
-     5.685529778143856443869104068`10.006480522567658*^-11, 
-     1.6929031333844892531295078`9.977505675107576*^-11, 
-     5.35411407791586161917308782`9.949819307906617*^-12, 
-     1.8027916057166808205171626`9.923299428081076*^-12, 
-     6.4757136080216956623947639`9.897841042750265*^-13, 
-     2.4859096088209497760612705`9.873353120061156*^-13, 
-     1.0214390869781754557219939`9.849756201397666*^-13, 
-     4.498377084383398815061773`9.826980504124831*^-14, 
-     2.125807129470862194991728`9.804964398641095*^-14, 
-     1.079079574717662677322349`9.783653174427505*^-14, 
-     5.88865486375942049290743`9.762998031642923*^-15, 
-     3.45718220211357030101118`9.742955250498067*^-15, 
-     2.18486905423368245158683`9.723485502044594*^-15, 
-     1.48704768353330757236638`9.70455327241037*^-15, 
-     1.0903559478293190408395`9.686126378761644*^-15, 
-     8.614956795706615651897`9.66817555997635*^-16, 
-     7.335467348592542794168`9.650674128586237*^-16, 
-     6.731197766254461794417`9.633597673286747*^-16, 
[…]

### 2019-06-24 scripts/stensolasample4mom.m

@@ -1,122 +1,0 @@
-stensolasample4mom = {0.49733206399905305545999008716231443036481631530922961\
-687752959237371396073947`59.402474749269096, 0.301125154178916672482877348746\
-74016528756018274738227198926194172701949017105`59.0637592838092, 0.129811708\
-20227603061407899399373288048748377408521918533406610402615371252438`58.86392\
-559728395, 0.0487491161134158391127287860216695185962347153890533013625206274\
-8555257730022`58.71716676176557, 0.016490629006918742927996564631670604517924\
-5216392880268438630131400953077676`58.598238725911216, 0.00493435999275953605\
-67971483004925889140000459940805596141856935584237536576`58.49629912755605, 0\
-.0012544425446128404037479948309201897664731363214017288200870367113705053229\
-5`58.4057484489246, 0.0002573969360828042718732114736102973875141070460766062\
-5456338109298492398437`58.32335574080289, 0.000040254326068588284611328827422\
-60324296633334008412445518590060598146427963`58.247112644595504, 
-     4.5163112658096409630910819675527616363877059834351935444829821354246491\
-8`58.175701379614026*^-6, 
-     3.4156645353725147848601756710293046697401480344649752266317862325606962`\
-58.10822016518226*^-7, 
-     1.634973894471768458778657013244757547448056590521415447110186689004737`\
-58.044030095355396*^-8, 4.649973962095806531696318627724391163481340122388888\
-7721224354773036641082204`57.982664613081994*^-10, 
-     7.3785740412647519385689962064597749092296591759864336242020913747497747\
-6849`57.923773536203186*^-12, 
-     6.1376392634998268214020602818725978668720768548609730502357849199622613\
-51`57.86708721141984*^-14, 
-     2.5165261427327738473653143751009482572778503074029358752057976191256446`\
-57.812392908177884*^-16, 
-     4.7867957569082971163259616318876424332838542634942447207146697285859`57\
-.75951891642099*^-19, 
-     3.9798233872771671755533470624070806321327174371509791703831840322149241\
-1648`57.70832362983569*^-22, 
-     1.3642375498918943529608605512991078496088513836059176819265547088569981\
-3`57.65868792924515*^-25, 
-     1.8209004735326114475083232022707090267019314339926078614030819430108`57\
-.61050979163423*^-29, 
-     8.9489194658058763608103701661461543142614915928185364375055011733248214\
-39`57.56370042373525*^-34, 
[…]

### 2019-06-25 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-19T19:49:04+0200
+%% Last-Updated: 2019-06-25T10:55:38+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In the present work we give an answer to this question: What inferences
-about the \emph{total activity} of a large neuronal network can we make by
-recording a very small sample thereof? We'll first give an answer by
-combining, in a straightforward way, the maximum-entropy method and basic
-sampling relations of the probability calculus. Later on we'll show that
-this answer can be obtained by exclusively using the probability calculus;
-this point of view shows that the maximum-entropy answer is a first
-approximation to the answers of \emph{four} different questions, and also
-shows the assumptions lying behind it.
+In the present work we give an answer to this question: How much is our
+uncertainty about the \emph{total activity} of a large neuronal network, if
+we have observed the activity of a very small sample thereof? We'll first
+give an answer by combining, in a straightforward way, the maximum-entropy
+method and basic sampling relations of the probability calculus. Later on
+we'll show that this answer can be obtained by exclusively using the
+probability calculus; this point of view shows that the maximum-entropy
+answer is a first approximation to the answers of \emph{four} different
+questions, and also shows the assumptions lying behind it.
 
@@
 \mynote{ref}\mynote{add recording length}. In the first data set, the most
-interesting inference is that the most probable frequency distribution for
+interesting finding is that the most probable frequency distribution for
 the total activity of the full network, if the latter is assumed to be
@@
 in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b,grandy1980}
-has been used for a variety of inferences about the neuronal activity of
+has been used for a variety of forecasts about the neuronal activity of
 various brain areas and about other phenomena of importance to the
@@
 This method is often used to test whether some statistics of the data, for
-example second-order time correlations, is sufficient for making other
-inferences about the data. It can be considered as an approximation of
+example second-order time correlations, is sufficient for quantifying our
+uncertainty about the data. It can be considered as an approximation of
 probabilistic models based on various assumptions of inferential
@@
 
-Let's introduce some context and notation for our inference problem.
+Let's introduce some context and notation for our problem.
 
@@
 \mynote{Move this to intro?}We want to stress the usefulness of making
-inferences about the full-network activity. First of all, this seems to be
+quantified guesses about the full-network activity. First of all, this seems to be
 the primary idea behind recording a sample. Second, it allows us to make
@@
 \end{itemize}
-The maximum-entropy distribution is thus an approximate answer to four
-completely different inferences. Note that the four distributions above
-numerically differ in higher-order approximations. We discuss their
-differences in \sect***. Second, the maximum-entropy method based on the
-Shannon entropy implicitly makes some assumptions about the probabilities
-for the long-run frequency distributions
+The maximum-entropy distribution is thus an approximation of our
+uncertainty about four completely different quantities. Note that the four
+distributions above numerically differ in higher-order approximations. We
+discuss their differences in \sect***. Second, the maximum-entropy method
+based on the Shannon entropy implicitly makes some assumptions about the
+probabilities for the long-run frequency distributions
 \citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
@@
 hypotheses that a researcher is exploring; for example, hypotheses of
-sufficient statistics, such as the sufficiency of pairwise correlations for
-inferential purposes. In the present work we only want to present the
+sufficient statistics, such as the sufficiency of pairwise correlations to
+quantify our uncertainty. In the present work we only want to present the
 general method without entering into biological questions of this kind.
@@
 distribution for the activity of the sample neurons. In this case, our
-inferences about the full-network activity are using all available measured
-frequency data. Note that application of the maximum-entropy method
-\emph{at the sample level} is trivial and meaningless in this case -- it
-just gives back the measured frequency distribution. But application of the
-method \emph{at the level of the full network} is not trivial.
+uncertaity about the full-network activity is based on all available
+measured frequency data. Note that application of the maximum-entropy
+method \emph{at the sample level} is trivial and meaningless in this case
+-- it just gives back the measured frequency distribution. But application
+of the method \emph{at the level of the full network} is not trivial.
 
@@
 \centering
-\includegraphics[width=\linewidth]{scripts/paper_stensola_6mom_distributionsN_prior2.pdf}%
+\includegraphics[width=\linewidth]{_scripts/paper_stensola_6mom_distributionsN_prior2.pdf}%
 \\[3em]%
-\includegraphics[width=\linewidth]{scripts/paper_riehle_6mom_distributionsN_prior2.pdf}%
+\includegraphics[width=\linewidth]{_scripts/paper_riehle_6mom_distributionsN_prior2.pdf}%
 \caption{***}
@@
 \centering
-\includegraphics[width=\linewidth]{scripts/paper2_stensola_twomom_distributions10000_prior2.pdf}%
+\includegraphics[width=\linewidth]{_scripts/paper2_stensola_twomom_distributions10000_prior2.pdf}%
 \\[3em]%
-\includegraphics[width=\linewidth]{scripts/paper_riehle_twomom_distributions10000_prior2.pdf}%
+\includegraphics[width=\linewidth]{_scripts/paper_riehle_twomom_distributions10000_prior2.pdf}%
 \caption{***}

### 2019-06-25 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-25T10:55:38+0200
+%% Last-Updated: 2019-06-25T11:58:15+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b,grandy1980}
-has been used for a variety of forecasts about the neuronal activity of
-various brain areas and about other phenomena of importance to the
+has been used for different kinds of estimations of the neuronal activity
+of various brain areas and about other phenomena of importance to the
 neurosciences, for example gene and protein interaction \citep[for
@@
 \emph{normalized total activity} of the sample, or just \enquote{activity}
-for short. Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if
-$\ya_{t}=0$, no neuron spikes during bin $t$; if $\ya_{t}=1$, all spike at
-some point during bin $t$. For brevity, let's say \enquote{at $t$} for
-\enquote{during time bin $t$}. From the activities $\set{\ya_{t}}$ we can
-count how many times the activity levels $\ya=0$, $\ya=1/n$, and so on
-appeared during the recording, obtaining the measured relative-frequency
-distribution $(\yff_{\ya}) \defs \yf$. We can also consider the sample
-activity at time bins \emph{outside} of the recorded range. Such activity
-is unknown to us, of course.
+for short. It is also the network-averaged activity of the neurons.
+Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no
+neuron spikes during bin $t$; if $\ya_{t}=1$, all spike at some point
+during bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during
+  time bin $t$}. From the activities $\set{\ya_{t}}$ we can count how many
+times the activity levels $\ya=0$, $\ya=1/n$, and so on appeared during the
+recording, obtaining the measured relative-frequency distribution
+$(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
+bins \emph{outside} of the recorded range. Such activity is unknown to us,
+of course.
 
@@
 sufficient statistics, such as the sufficiency of pairwise correlations to
-quantify our uncertainty. In the present work we only want to present the
-general method without entering into biological questions of this kind.
+quantify our uncertainty about network activity. In the present work we
+only want to introduce the general method without entering into biological
+questions of this kind.
 
@@
 distribution for the activity of the sample neurons. In this case, our
-uncertaity about the full-network activity is based on all available
+uncertainty about the full-network activity is based on all available
 measured frequency data. Note that application of the maximum-entropy
@@
 We first calculate the distribution by using six moments. This number
-already provides as much information as the full frequency distribution of
-the sample, and at the same time illustrates the use of the approach in
-questions of statistic sufficiency (typically limited to two or three
-moments). Figure~\ref{fig:plots_6mom} shows the resulting densities
-(distribution/activity-step, that is, distribution${}\times N$) for
-full-network sizes $N=n$, $N=1\,000$, $N=5\,000$,
-$N=10\,000$\mynote{motivate?}. The case $N=n$ corresponds to applying the
-maximum-entropy method at the sample level; it can be observed that with
-six moments it reproduces almost exactly the measured frequency
-distribution.
+already provides almost as much information as the full frequency
+distribution of the sample, and at the same time illustrates the use of the
+approach in questions of statistic sufficiency (typically limited to two or
+three moments). Figure~\ref{fig:plots_6mom} shows the resulting densities
+(that is, distribution${}\times N$) for full-network sizes $N=n$,
+$N=1\,000$, $N=5\,000$, $N=10\,000$\mynote{motivate?}. The case $N=n$
+corresponds to applying the maximum-entropy method at the sample level; it
+can be observed that with six moments it reproduces almost exactly the
+measured frequency distribution.
 \begin{figure}[!p]
@@
 distribution at the sample level. The application of the probability
-calculus to make forecasts about the full network thus reveals interesting
-and useful features.
+calculus thus reveals interesting possible features of the full network.
 

### 2019-06-25 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-25T11:58:15+0200
+%% Last-Updated: 2019-06-25T17:30:45+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 particular, four moments lead to a bimodal distribution, whereas two to a
-unimodal one, showing that two moments would not be a sufficient
+unimodal one, showing that two moments would not be sufficient
 statistics. For comparison, the two distributions obtained applying
@@
 
-\section{Marginalization to the sample level}
+\section{Marginalization to the sample level:\\ the issue with pairwise
+  correlations and other statistics}
 \label{sec:marginalization}
+
+The maximum-entropy method has been applied in the neurosciences to
+questions of sufficient statistics -- for example, whether we approximately
+reach the same state of uncertainty about network activity by considering
+only its network-averaged pairwise correlations and by considering also
+third-order correlations. The method is typically used as follows. We
+imagine to be ignorant of the measured frequency distribution for the
+network activity, and to know only the measured statistics of interest; two
+maximum-entropy distributions are then built, constrained under the two
+sets of moments under comparison. Finally, the resulting distributions are
+compared with each other and with the measured frequency distribution.
+
+The method introduced in the present work reveals an issue in the recipe
+above. The issue doesn't come from the method itself but from a simple,
+general fact of probability theory: if a probability distribution has some
+sufficient statistics \citep[\sect~4.5]{bernardoetal1994}{fortinietal2000},
+then its marginals \emph{cannot} have the same sufficient statistics, and
+vice versa (except for trivial cases, such as uniform distributions). This
+impossibility is known in statistical mechanics: marginals of Gibbs states
+aren't Gibbs states \citep{maesetal1999}. Mathematically it corresponds to
+the general impossibility of solving a system of $n$ equations in $m$
+unknowns when $n>m$
+\citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.
+
+In the neuroscientific application above, this means for example that if
+pairwise correlations are sufficient statistics for a sample of a network,
+then they \emph{cannot} be sufficient for the full network; and vice versa,
+if we assume that they are sufficient statistics for the full network, then
+we must expect them \emph{not} to be sufficient for any recorded sample
+thereof. Thus, maximum-entropy methods applied at the sample level
+paradoxically give an answer opposite to the one we might think they're
+giving!
 
@@
 
-in depth discussion about underlying assumptions, derivation from
+in-depth discussion about underlying assumptions, derivation from
 probability calculus, higher-order approximations beyond maxent and

### 2019-06-25 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-25T17:30:45+0200
+%% Last-Updated: 2019-06-25T17:40:21+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 vice versa (except for trivial cases, such as uniform distributions). This
-impossibility is known in statistical mechanics: marginals of Gibbs states
-aren't Gibbs states \citep{maesetal1999}. Mathematically it corresponds to
-the general impossibility of solving a system of $n$ equations in $m$
-unknowns when $n>m$
-\citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.
+also holds among marginals of different order (to $n$ neurons or to $n+1$
+neurons, for example). This impossibility is known in statistical
+mechanics: marginals of Gibbs states aren't Gibbs states
+\citep{maesetal1999}. Mathematically it corresponds to the general
+impossibility of solving a system of $n$ equations in $m$ unknowns when
+$n>m$ \citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.
 
@@
 pairwise correlations are sufficient statistics for a sample of a network,
-then they \emph{cannot} be sufficient for the full network; and vice versa,
-if we assume that they are sufficient statistics for the full network, then
-we must expect them \emph{not} to be sufficient for any recorded sample
-thereof. Thus, maximum-entropy methods applied at the sample level
+then they \emph{cannot} be sufficient for the full network or for a sample
+with more or fewer neurons; and vice versa, if we assume that they are
+sufficient statistics for the full network, then we must expect them
+\emph{not} to be sufficient for any recorded sample thereof.
+
+It seems to us that such questions about sufficient statistics are
+implicitly asked for specific brain areas, not for casually recorded
+samples thereof. Thus, maximum-entropy methods applied at the sample level
 paradoxically give an answer opposite to the one we might think they're

### 2019-06-25 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-25T17:40:21+0200
+%% Last-Updated: 2019-06-25T17:44:44+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 sufficient statistics for the full network, then we must expect them
-\emph{not} to be sufficient for any recorded sample thereof.
-
-It seems to us that such questions about sufficient statistics are
-implicitly asked for specific brain areas, not for casually recorded
-samples thereof. Thus, maximum-entropy methods applied at the sample level
+\emph{not} to be sufficient for any recorded sample thereof. It seems to us
+that such questions about sufficient statistics are implicitly meant to
+apply to a specific brain area, not just to some casually recorded samples
+thereof. Thus, maximum-entropy methods applied at the sample level
 paradoxically give an answer opposite to the one we might think they're

### 2019-07-07 maxNt_v5.tex
update

--- 2019-06-25 maxNt_v5.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-06-25T17:40:21+0200
+%% Last-Updated: 2019-07-07T14:24:05+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The maximum-entropy method has been applied in the neurosciences to
-questions of sufficient statistics -- for example, whether we approximately
-reach the same state of uncertainty about network activity by considering
-only its network-averaged pairwise correlations and by considering also
-third-order correlations. The method is typically used as follows. We
-imagine to be ignorant of the measured frequency distribution for the
-network activity, and to know only the measured statistics of interest; two
-maximum-entropy distributions are then built, constrained under the two
-sets of moments under comparison. Finally, the resulting distributions are
-compared with each other and with the measured frequency distribution.
-
-The method introduced in the present work reveals an issue in the recipe
-above. The issue doesn't come from the method itself but from a simple,
-general fact of probability theory: if a probability distribution has some
+The maximum-entropy method has often been applied in the neurosciences to
+questions of sufficient statistics. For example: does knowledge about
+third-order correlations change our state of uncertainty about network
+activity in an appreciable way, compared to knowledge about second-order
+correlations only? The method is used as follows to answer this question.
+We imagine to be ignorant of the measured frequency distribution for the
+network activity, and to know only up to second-order measured statistics,
+or up to third-order. Two maximum-entropy distributions are then built,
+constrained under the two sets of moments under comparison. Finally, the
+resulting distributions are compared with each other and with the measured
+frequency distribution. This recipe generalizes to any subsets of measured
+statistics, of course.
+
+The method introduced in the present work reveals an issue in this recipe;
+the issue doesn't come from the method itself but from a simple, general
+fact of probability theory. If a probability distribution has some
 sufficient statistics \citep[\sect~4.5]{bernardoetal1994}{fortinietal2000},
 then its marginals \emph{cannot} have the same sufficient statistics, and
-vice versa (except for trivial cases, such as uniform distributions). This
-also holds among marginals of different order (to $n$ neurons or to $n+1$
-neurons, for example). This impossibility is known in statistical
-mechanics: marginals of Gibbs states aren't Gibbs states
+vice versa, except for trivial cases such as uniform distributions. This in
+general also holds between different marginals: if one has some sufficient
+statistics, another won't have it. This impossibility is known in
+statistical mechanics: marginals of Gibbs states aren't Gibbs states
 \citep{maesetal1999}. Mathematically it corresponds to the general
@@
 In the neuroscientific application above, this means for example that if
-pairwise correlations are sufficient statistics for a sample of a network,
-then they \emph{cannot} be sufficient for the full network or for a sample
-with more or fewer neurons; and vice versa, if we assume that they are
+pairwise correlations are sufficient statistics for a particular sample of
+a network, then they \emph{cannot} be sufficient for the full networ, or
+for a different sample; and vice versa, if we assume that they are
 sufficient statistics for the full network, then we must expect them
-\emph{not} to be sufficient for any recorded sample thereof.
-
-It seems to us that such questions about sufficient statistics are
-implicitly asked for specific brain areas, not for casually recorded
-samples thereof. Thus, maximum-entropy methods applied at the sample level
-paradoxically give an answer opposite to the one we might think they're
-giving!
+\emph{not} to be sufficient for any recorded sample thereof. But questions
+about sufficient statistics are implicitly meant to apply to a whole brain
+area, not just to some casually recorded sample. Thus, maximum-entropy
+methods applied at the sample level paradoxically give an answer opposite
+to the one we might think they're giving!
+
+The comparisons of different sufficient statistics are only approximate,
+however: we ask only whether our uncertainties conditional on either are
+\emph{appreciably} different in some specified sense. It's therefore worth
+checking whether the different results we get from a sufficient statistics
+for the sample and for the full-network are still very similar when we
+consider an altogether different statistics.
 
@@
 "error bars" around the distributions
+
+%plot_mc_6mom_unif_afss3tr_N10000_t400_L500_T5e+05_s30000_a0_i3
 

### 2019-07-08 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-07T14:24:05+0200
+%% Last-Updated: 2019-07-08T13:40:41+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \mynote{Say something more about advantages of such question/answer: for
-  example we can make statements about total activity of brain region even
+  example we can make statements about total activity of brain area even
   across recordings when sampled neurons aren't the same.}
@@
 \mynote{maybe move this paragraph to the intro?}Present-day technologies
-enable the recording of neuronal activity from small brain regions
+enable the recording of neuronal activity from small brain areas
 \mynote{be more specific}. For many animal species, the neurons that are
-recorded within the region are not -- and at present cannot be --
+recorded within the area are not -- and at present cannot be --
 specifically chosen from among the rest, owing to several limiting factors;
@@
 of recorded neurons may even change during very long recordings or across
-experimental sessions. We assume that there's a region, comprising a
+experimental sessions. We assume that there's a area, comprising a
 network of $N$ neurons, from which other sets of neurons had or will have
@@
 
-In the neuroscientific application above, this means for example that if
-pairwise correlations are sufficient statistics for a particular sample of
-a network, then they \emph{cannot} be sufficient for the full networ, or
-for a different sample; and vice versa, if we assume that they are
-sufficient statistics for the full network, then we must expect them
-\emph{not} to be sufficient for any recorded sample thereof. But questions
-about sufficient statistics are implicitly meant to apply to a whole brain
-area, not just to some casually recorded sample. Thus, maximum-entropy
-methods applied at the sample level paradoxically give an answer opposite
-to the one we might think they're giving!
-
-The comparisons of different sufficient statistics are only approximate,
-however: we ask only whether our uncertainties conditional on either are
-\emph{appreciably} different in some specified sense. It's therefore worth
-checking whether the different results we get from a sufficient statistics
-for the sample and for the full-network are still very similar when we
-consider an altogether different statistics.
+In the neuroscientific application above, this means that if pairwise
+correlations, for example, are sufficient statistics for a particular
+sample of from a brain area, then they \emph{cannot} be sufficient for the
+full network of neurons in that area, or for a different sample; vice
+versa, if we assume that they are sufficient statistics for the full
+network, then we must expect them \emph{not} to be sufficient for any
+recorded sample thereof. But questions about sufficient statistics are
+meant to be addressed to a whole brain area, not just to some casually
+recorded sample. Thus, maximum-entropy methods applied at the sample level
+paradoxically give an answer opposite to the one we might think they're
+giving.
+
+Questions about sufficient statistics are therefore correctly addressed by
+applying the maximum-entropy method at the full-network level. If a
+comparison with the measured frequencies of the sample activities is
+desired, then the marginal distribution to the sample size should be used.
+
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
+\\[3em]%
+\includegraphics[width=\linewidth]{_scripts/comparison_marginals.pdf}%
+\caption{***}
+\label{fig:comparisons_marginals}
+\end{figure}
+%comparison_ME_sample_full_levels.nb
+Does the application to the full-network lead to very different results
+from the one to the sample? Figure~\ref{fig:comparisons_marginals} gives a
+graphical answer in the case of the first data set. The upper plot shows
+the maximum-entropy distributions for a full network of $10\,000$ neurons,
+constructed from two moments (\textcolor{mygreen}{dashed green curve}) and
+from four moments (\textcolor{myblue}{solid blue curve}). The lower plot
+shows the maximum-entropy distributions for the sample (dashed curves),
+from two moments (\textcolor{mygreen}{empty green triangles}) and from four
+moments (\textcolor{mygreen}{empty blue squares}), and the distributions
+obtained by marginalizing the full-network distribution to the sample size
+(solid curves), from two moments (\textcolor{mygreen}{filled green
+  triangles}) and from four moments (\textcolor{mygreen}{filled blue
+  squares}). The measured frequency distribution (\textcolor{myred}{red
+  circles}) is also shown. We note the following:
+\begin{itemize}
+\item The application to the full-network (upper plot) leads to two
+  completely different distributions (even different number of modes);
+  clearly the two statistics are not even approximately equivalent for
+  inferential purposes.
+\item The application to the sample leads to deceivingly similar
+  distributions (empty triangles and empty squares), which can only be
+  distinguished on a logarithmic scale. This could lead to the erroneous
+  conclusion that the two statistics are approximately equivalent.
+\item The difference between marginals from the full-network distribution
+  and the distributions obtained at the sample level (filled vs empty
+  markers) is larger than the difference between different statistics
+  (triangles vs squares).
+\item The marginals of the application to the full network are closer to
+  the measured frequencies than the distributions obtained at the sample
+  level (although this closeness is not a valid criterion for their
+  goodness).
+\end{itemize}
+Therefore, the two maximum-entropy applications lead to different results,
+and the application at the full-network level is clearly superior, leading
+to a clearer distinction between the effect of considering different
+statistics.
 
@@
 "error bars" around the distributions
-
-%plot_mc_6mom_unif_afss3tr_N10000_t400_L500_T5e+05_s30000_a0_i3
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\linewidth]{_scripts/MC_plot1000.pdf}%
+\\[3em]%
+\includegraphics[width=\linewidth]{_scripts/MC_samples1000.pdf}%
+\caption{***}
+\label{fig:comparisons_marginals}
+\end{figure}
+%mctests3.nb
+%plot_mcfd_6mom_unif_afssd26_cok_tr_N1000_t50_L50_T0.333333333333333_s60000_a0
 
@@
 formulae is essential in neuroscience, where we try to understand
-properties of extended brain regions from partial observations. The
+properties of extended brain areas from partial observations. The
 formulae presented here reflect a simple, symmetric state of ignorance.

### 2019-07-08 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-08T13:40:41+0200
+%% Last-Updated: 2019-07-08T16:40:48+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 from four moments (\textcolor{myblue}{solid blue curve}). The lower plot
-shows the maximum-entropy distributions for the sample (dashed curves),
+shows the maximum-entropy distributions for the sample (empty markers),
 from two moments (\textcolor{mygreen}{empty green triangles}) and from four
-moments (\textcolor{mygreen}{empty blue squares}), and the distributions
+moments (\textcolor{myblue}{empty blue squares}), and the distributions
 obtained by marginalizing the full-network distribution to the sample size
-(solid curves), from two moments (\textcolor{mygreen}{filled green
-  triangles}) and from four moments (\textcolor{mygreen}{filled blue
+(filled markers), from two moments (\textcolor{mygreen}{filled green
+  triangles}) and from four moments (\textcolor{myblue}{filled blue
   squares}). The measured frequency distribution (\textcolor{myred}{red
@@
 
-in-depth discussion about underlying assumptions, derivation from
+***in-depth discussion about underlying assumptions, derivation from
 probability calculus, higher-order approximations beyond maxent and
 "error bars" around the distributions
+
+***We can ask about probability of bimodality etc
+
 \begin{figure}[!p]

### 2019-07-08 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-08T16:40:48+0200
+%% Last-Updated: 2019-07-08T23:05:19+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The maximum-entropy method has often been applied in the neurosciences to
-questions of sufficient statistics. For example: does knowledge about
-third-order correlations change our state of uncertainty about network
-activity in an appreciable way, compared to knowledge about second-order
-correlations only? The method is used as follows to answer this question.
-We imagine to be ignorant of the measured frequency distribution for the
-network activity, and to know only up to second-order measured statistics,
-or up to third-order. Two maximum-entropy distributions are then built,
-constrained under the two sets of moments under comparison. Finally, the
-resulting distributions are compared with each other and with the measured
-frequency distribution. This recipe generalizes to any subsets of measured
-statistics, of course.
-
-The method introduced in the present work reveals an issue in this recipe;
-the issue doesn't come from the method itself but from a simple, general
-fact of probability theory. If a probability distribution has some
-sufficient statistics \citep[\sect~4.5]{bernardoetal1994}{fortinietal2000},
-then its marginals \emph{cannot} have the same sufficient statistics, and
-vice versa, except for trivial cases such as uniform distributions. This in
-general also holds between different marginals: if one has some sufficient
-statistics, another won't have it. This impossibility is known in
-statistical mechanics: marginals of Gibbs states aren't Gibbs states
+In the neurosciences, the maximum-entropy method has often been applied to
+questions of sufficient statistics. In this section we discuss how our
+proposed application bears on this kind of questions. For concreteness
+we use an example with specific statistics: first and second moments
+(equivalent to mean and correlations), versus first-to-fourth moments. But
+our discussion holds for any other sets of statistics.
+
+The question is as follows. We want to assess our degree of belief about
+the frequencies of the activities of the sample, or about the sample
+activity in a new time bin. For this assessment we use any measured data we
+have. But it sometimes happens that dropping part of the data -- for
+example, the measured third- and higher-order moments -- leaves our
+assessment almost unchanged: the dropped data are \emph{informationally
+  irrelevant}, or \emph{almost} so. The data that remain -- for example,
+first and second moments -- are \emph{informationally sufficient}. This
+informational quasi-sufficiency may be connected to biological or physical
+mechanisms; that's why it's interesting.
+
+We can approximately check which data are irrelevant by simply dropping
+them from the conditional of our degree of belief, and see if the latter
+changes appreciably. When our degree of belief is built with the
+maximum-entropy method, we simply drop some of the data used as constraints
+and see how the resulting distribution changes. In the neurosciences this
+has been done with maximum-entropy distributions constructed \emph{for the
+  sample}.
+
+% For example: does knowledge about
+% third-order correlations change our state of uncertainty about network
+% activity in an appreciable way, compared to knowledge about second-order
+% correlations only? The method is used as follows to answer this question.
+% We imagine to be ignorant of the measured frequency distribution for the
+% network activity, and to know only up to second-order measured statistics,
+% or up to third-order. Two maximum-entropy distributions are then built,
+% constrained under the two sets of moments under comparison. Finally, the
+% resulting distributions are compared with each other and with the measured
+% frequency distribution. This recipe generalizes to any subsets of measured
+% statistics, of course.
+
+The method introduced in the present work reveals an issue with this
+application to the sample. The issue doesn't come from the method itself
+but from a simple, general fact of probability theory. If a probability
+distribution has some sufficient statistics
+\citep[\sect~4.5]{bernardoetal1994}{fortinietal2000}, then its marginals
+\emph{cannot} have the same sufficient statistics, and vice versa, except
+for trivial cases such as uniform distributions. This in general also holds
+between different marginals: if a marginal has some sufficient statistics,
+another won't have it. This impossibility is known in statistical
+mechanics: marginals of Gibbs states aren't Gibbs states
 \citep{maesetal1999}. Mathematically it corresponds to the general

### 2019-07-09 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-08T23:05:19+0200
+%% Last-Updated: 2019-07-09T13:33:41+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In the neurosciences, the maximum-entropy method has often been applied to
+In the neurosciences the maximum-entropy method is often applied to
 questions of sufficient statistics. In this section we discuss how our
-proposed application bears on this kind of questions. For concreteness
-we use an example with specific statistics: first and second moments
+proposed application bears on this kind of questions. For concreteness we
+use an example with specific statistics: first and second moments
 (equivalent to mean and correlations), versus first-to-fourth moments. But
-our discussion holds for any other sets of statistics.
+our discussion holds for any  sets of statistics.
 
@@
 the frequencies of the activities of the sample, or about the sample
-activity in a new time bin. For this assessment we use any measured data we
-have. But it sometimes happens that dropping part of the data -- for
-example, the measured third- and higher-order moments -- leaves our
+activity in a new time bin. For this assessment we should use all measured
+data we have. But it sometimes happens that dropping part of the data --
+for example, the measured third- and higher-order moments -- leaves our
 assessment almost unchanged: the dropped data are \emph{informationally
-  irrelevant}, or \emph{almost} so. The data that remain -- for example,
-first and second moments -- are \emph{informationally sufficient}. This
-informational quasi-sufficiency may be connected to biological or physical
-mechanisms; that's why it's interesting.
-
-We can approximately check which data are irrelevant by simply dropping
-them from the conditional of our degree of belief, and see if the latter
-changes appreciably. When our degree of belief is built with the
+  irrelevant}, or almost so. The remaining data -- for example, first and
+second moments -- are \emph{informationally sufficient}. This informational
+quasi-sufficiency is interesting because it may hint at peculiar biological
+or physical mechanisms or dependencies.
+
+We can approximately check how much some data are irrelevant by simply
+dropping them from the conditional of our degree of belief, and see if the
+latter changes appreciably. When our degree of belief is built with the
 maximum-entropy method, we simply drop some of the data used as constraints
-and see how the resulting distribution changes. In the neurosciences this
-has been done with maximum-entropy distributions constructed \emph{for the
-  sample}.
+and see how the resulting distribution changes. For example, we may only
+retain the first two moments (means and correlations) or the first four
+moments. In the neurosciences this check is often done with maximum-entropy
+distributions constructed \emph{for the sample}, ignoring the full network
+it's sampled from.
 
@@
 The method introduced in the present work reveals an issue with this
-application to the sample. The issue doesn't come from the method itself
-but from a simple, general fact of probability theory. If a probability
+application at the sample level. The issue doesn't come from the method
+itself but from a general fact of probability theory. If a probability
 distribution has some sufficient statistics
@@
 \emph{cannot} have the same sufficient statistics, and vice versa, except
-for trivial cases such as uniform distributions. This in general also holds
-between different marginals: if a marginal has some sufficient statistics,
+for trivial cases such as uniform distributions. This generally also holds
+among different marginals: if a marginal has some sufficient statistics,
 another won't have it. This impossibility is known in statistical
@@
 
-In the neuroscientific application above, this means that if pairwise
-correlations, for example, are sufficient statistics for a particular
-sample of from a brain area, then they \emph{cannot} be sufficient for the
-full network of neurons in that area, or for a different sample; vice
-versa, if we assume that they are sufficient statistics for the full
-network, then we must expect them \emph{not} to be sufficient for any
-recorded sample thereof. But questions about sufficient statistics are
-meant to be addressed to a whole brain area, not just to some casually
-recorded sample. Thus, maximum-entropy methods applied at the sample level
-paradoxically give an answer opposite to the one we might think they're
-giving.
-
-Questions about sufficient statistics are therefore correctly addressed by
-applying the maximum-entropy method at the full-network level. If a
+In our example this means that if means and pairwise correlations or the
+first four moments are informationally sufficient for a particular sample
+from a brain area, then they \emph{cannot} be sufficient for the full
+network of neurons in that area, or for a different sample. Vice versa, if
+we assume that they are informationally sufficient for the full network,
+then we must expect them \emph{not} to be sufficient for any recorded
+sample thereof. Now, questions about sufficient statistics are meant to be
+addressed to a whole brain area, not just to some casually recorded sample.
+Thus, paradoxically, maximum-entropy methods applied at the sample level
+give an answer opposite to the one we might think they're giving.
+
+Questions about informational sufficiency are therefore correctly addressed
+by applying the maximum-entropy method at the full-network level. If a
 comparison with the measured frequencies of the sample activities is
@@
 %comparison_ME_sample_full_levels.nb
-Does the application to the full-network lead to very different results
-from the one to the sample? Figure~\ref{fig:comparisons_marginals} gives a
-graphical answer in the case of the first data set. The upper plot shows
-the maximum-entropy distributions for a full network of $10\,000$ neurons,
-constructed from two moments (\textcolor{mygreen}{dashed green curve}) and
-from four moments (\textcolor{myblue}{solid blue curve}). The lower plot
-shows the maximum-entropy distributions for the sample (empty markers),
-from two moments (\textcolor{mygreen}{empty green triangles}) and from four
-moments (\textcolor{myblue}{empty blue squares}), and the distributions
-obtained by marginalizing the full-network distribution to the sample size
-(filled markers), from two moments (\textcolor{mygreen}{filled green
-  triangles}) and from four moments (\textcolor{myblue}{filled blue
-  squares}). The measured frequency distribution (\textcolor{myred}{red
-  circles}) is also shown. We note the following:
+But does the application at the full-network level lead to appreciably
+different results from that at the sample level? after all we're interested
+in an approximate informational sufficiency, not in an analytically exact
+one. Figure~\ref{fig:comparisons_marginals} gives a graphical answer to
+this question in the case of our first data set. The upper plot shows the
+maximum-entropy distributions for a full network of $10\,000$ neurons,
+constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
+from four moments (\textcolor{myblue}{\textbf{------}}). The lower plot
+shows the maximum-entropy distributions for the sample, from two moments
+(\textcolor{mygreen}{$\bm{\triangledown}$}) and from four moments
+(\textcolor{myblue}{$\bm{\square}$}), and the distributions obtained by
+marginalizing the full-network distribution to the sample size, from two moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$})
+and from four moments (\textcolor{myblue}{$\bm{\blacksquare}$}). The
+measured frequency distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also
+shown. We note the following:
 \begin{itemize}
@@
   completely different distributions (even different number of modes);
-  clearly the two statistics are not even approximately equivalent for
-  inferential purposes.
-\item The application to the sample leads to deceivingly similar
-  distributions (empty triangles and empty squares), which can only be
-  distinguished on a logarithmic scale. This could lead to the erroneous
-  conclusion that the two statistics are approximately equivalent.
+  clearly the two sets of statistics are not even approximately equivalent
+  for inferential purposes.
+\item The application at the sample level leads to deceivingly similar
+  distributions (\textcolor{mygreen}{$\bm{\triangledown}$} and
+  \textcolor{myblue}{$\bm{\square}$}), which can only be distinguished on a
+  logarithmic scale. This could lead to the erroneous conclusion that the
+  two statistics are approximately equivalent.
 \item The difference between marginals from the full-network distribution
@@
 \end{itemize}
-Therefore, the two maximum-entropy applications lead to different results,
-and the application at the full-network level is clearly superior, leading
-to a clearer distinction between the effect of considering different
-statistics.
+Therefore, the maximum-entropy application at the sample level and at the
+full-network level lead to different results; the latter is not only more
+meaningful but also superior because it shows clearly the different
+informational sufficiency of the two sets of statistics.
 

### 2019-07-09 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-09T13:33:41+0200
+%% Last-Updated: 2019-07-09T16:57:47+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Assumptions and corrections}
+\section{Assumptions and corrections:\\beyond the maximum-entropy approximation}
 \label{sec:assumptions_formulae}
+
+In \sect~\ref{sec:model} we mentioned that maximum-entropy distributions
+are approximations to the full answers provided by the probability
+calculus. For example we ask: what is the frequency distribution for the
+total activity of the full network across the recorded bins? Several
+frequency distributions are possible, each with a specific probability. The
+maximum-entropy distribution is
+
+***mode of marginal can be very different from the one in full space (minka1998)
 

### 2019-07-10 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-09T16:57:47+0200
+%% Last-Updated: 2019-07-10T10:12:32+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The method introduced in the present work reveals an issue with this
-application at the sample level. The issue doesn't come from the method
+application at the sample level, however. The issue doesn't come from the method
 itself but from a general fact of probability theory. If a probability
@@
 sample thereof. Now, questions about sufficient statistics are meant to be
-addressed to a whole brain area, not just to some casually recorded sample.
-Thus, paradoxically, maximum-entropy methods applied at the sample level
-give an answer opposite to the one we might think they're giving.
+addressed to a whole brain area, not just to some specific, casually
+recorded sample. Thus, paradoxically, maximum-entropy methods applied at
+the sample level give an answer opposite to the one we might think they're
+giving.
 
@@
 in an approximate informational sufficiency, not in an analytically exact
-one. Figure~\ref{fig:comparisons_marginals} gives a graphical answer to
-this question in the case of our first data set. The upper plot shows the
+one. A correct answer can only be given case by case;
+\fig~\ref{fig:comparisons_marginals} gives a graphical answer to this
+question in the case of our first data set. The upper plot shows the
 maximum-entropy distributions for a full network of $10\,000$ neurons,
@@
 (\textcolor{myblue}{$\bm{\square}$}), and the distributions obtained by
-marginalizing the full-network distribution to the sample size, from two moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$})
-and from four moments (\textcolor{myblue}{$\bm{\blacksquare}$}). The
-measured frequency distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also
-shown. We note the following:
+marginalizing the full-network distribution to the sample size, from two
+moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
+moments (\textcolor{myblue}{$\bm{\blacksquare}$}). The measured frequency
+distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also shown. We
+note the following:
 \begin{itemize}
@@
 
-***mode of marginal can be very different from the one in full space (minka1998)
+***mode of marginal can be very different from the one in full space
+(minka1998)
+
+*** diff position of mode in marginals says that full distr is extremely
+skewed – not surprising since the base space is bounded.
 

### 2019-07-10 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-10T10:12:32+0200
+%% Last-Updated: 2019-07-10T11:49:38+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\yH}{\varIota}
-\newcommand*{\yD}{\varDelta}
 \newcommand*{\yHa}{\varIota_\textrm{p}}
@@
 \newcommand*{\yHi}{\varIota''}
+\newcommand*{\yI}{\varIota}
+\newcommand*{\yD}{D}
 %Lagr multipliers
@@
 In \sect~\ref{sec:model} we mentioned that maximum-entropy distributions
-are approximations to the full answers provided by the probability
-calculus. For example we ask: what is the frequency distribution for the
-total activity of the full network across the recorded bins? Several
-frequency distributions are possible, each with a specific probability. The
-maximum-entropy distribution is
+come from a zeroth-order Laplace approximation of the densities obtained
+from the probability calculus. Zeroth-order means that we are simply
+considering the mode of such densities. Let's be more concrete. Consider
+the question: What is the relative-frequency distribution for the different
+activity levels for the full-network, across the recorded bins? Call this
+distribution $(\yFF_{\yA}) \defs \yF$: activity level $\yA$ appeared in
+$T\,\yFF(\yA)$ out of $T$ time bins. We don't know $\yF$, and our
+uncertainty about its possible values is a probability distribution
+$\pf(\yF\|\yD\yI)$ conditional on data $\yD$ and prior knowledge $\yI$,
+which we approximate with a continuous density
+\begin{equation}
+  \label{eq:density_F}
+  \pf(\yF \| \yD \yI)\;\di\yF.
+\end{equation}
+Now consider the related question: What is the activity $\yA_{t}$ of the
+full network at some specific time bin $t$? If we don't know the exact
+sequence of activities during the recording then we don't know $\yA_{t}$;
+if we know the frequency distribution $\yF$ of the activity levels, then
+our probability for $\yA_{t}$ is $\yFF_{\yA_{t}}$, out of combinatorial
+reasons. But if we are also uncertain about $\yF$, with
+probability~\eqref{eq:density_F}, then our probability for $\yA_{t}$ by
+marginalization is
+\begin{equation}
+  \label{eq:probability_AT}
+  \pf(\yA_{t} \| \yD \yI) =
+  \int \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF.
+\end{equation}
 
@@
 \end{figure}
-%mctests3.nb
-%plot_mcfd_6mom_unif_afssd26_cok_tr_N1000_t50_L50_T0.333333333333333_s60000_a0
+% mctests3.nb
+% mcfd_6mom_unif_afssd40st0np_cok_tr_N1000_t50_L50_T0.333333333333333_s60000_a15758.0500981316
+
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\linewidth]{_scripts/MC_plot5000.pdf}%
+\\[3em]%
+\includegraphics[width=\linewidth]{_scripts/MC_samples5000.pdf}%
+\caption{***}
+\label{fig:comparisons_marginals}
+\end{figure}
+% mctests3.nb
+% mcfd_6mom_unif_afssd32s_cok_tr_N5000_t180_L50_T0.333333333333333_s25800_a329212306.071189
 

### 2019-07-10 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-10T11:49:38+0200
+%% Last-Updated: 2019-07-10T14:04:12+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\yf}{\bm{\yff}}
+\newcommand*{\yFFm}{\yFF^{*}}
+\newcommand*{\yFm}{\yF^{*}}
 
@@
 \newcommand*{\yc}{\widehat{\bm{c}}}
-\newcommand*{\yL}{\bm{\lambda}}
+\newcommand*{\yll}{\bm{\lambda}}
 \newcommand*{\yl}{\bm{l}}
@@
 \newcommand*{\pxx}{p}
+\newcommand*{\Nd}{\textrm{N}}
+\newcommand*{\ycm}{\mathte{M}}
+\newcommand*{\ycs}{\sigma}
+\newcommand*{\yL}{L}
+\newcommand*{\yd}{\bm{\varDelta}}
+\newcommand*{\yds}{\delta}
 
@@
 from the probability calculus. Zeroth-order means that we are simply
-considering the mode of such densities. Let's be more concrete. Consider
-the question: What is the relative-frequency distribution for the different
-activity levels for the full-network, across the recorded bins? Call this
-distribution $(\yFF_{\yA}) \defs \yF$: activity level $\yA$ appeared in
-$T\,\yFF(\yA)$ out of $T$ time bins. We don't know $\yF$, and our
-uncertainty about its possible values is a probability distribution
+considering the mode of such densities. Let's be more concrete.
+
+Consider the question: What is the relative-frequency distribution for the
+different activity levels for the full-network, across the recorded bins?
+Call this distribution $(\yFF_{\yA}) \defs \yF$: activity level $\yA$
+appeared in $T\times\yFF(\yA)$ out of $T$ time bins. We don't know $\yF$,
+and our uncertainty about its possible values is a probability distribution
 $\pf(\yF\|\yD\yI)$ conditional on data $\yD$ and prior knowledge $\yI$,
@@
 \end{equation}
+In appendix*** we give a summary of how this density is derived. Note that
+this is a probability distribution over frequency distributions, so we must
+be careful with terminology to avoid confusion.
+
 Now consider the related question: What is the activity $\yA_{t}$ of the
@@
   \pf(\yA_{t} \| \yD \yI) =
-  \int \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF.
+  \int \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF;
+\end{equation}
+our probability distribution for $\yA_{t}$ is thus simply the mean
+$\int \yF\; \pf(\yF \| \yD \yI)\;\di\yF$.
+
+Formulae~\eqref{eq:density_F} and \eqref{eq:probability_AT} show that the
+most probable frequency distribution $\yF$ (if it's unique) and the
+probability distribution for $\yA_{t}$ are generally different: the former
+is the mode of $\pf(\yF \| \yD \yI)$, the latter its mean.
+
+If the probability density $\pf(\yF \| \yD \yI)$ is extremely peaked at some
+frequency distribution $\yFm$, it can be approximately treated as a delta
+density,
+\begin{equation}
+  \label{eq:approx_delta_pF}
+  \pf(\yF \| \yD \yI)\;\di\yF \approx \delta(\yF-\yFm)\;\di\yF,
+\end{equation}
+and consequently the probability for the activity at bin $t$, by
+\eqn~\eqref{eq:probability_AT}, is simply the mode $\yFm$ itself:
+\begin{equation}
+  \label{eq:approx_delta_pA}
+  \pf(\yA_{t} \| \yD \yI) \approx \yFFm_{\yA_{t}}.
+\end{equation}
+The maximum-entropy distribution is exactly this mode $\yFm$, for
+\emph{particular} kinds of probability densities $\pf(\yF \| \yD \yI)$
+which we'll discuss shortly. Formulae~\eqref{eq:approx_delta_pF}
+and~\eqref{eq:approx_delta_pA} show why the maximum-entropy distribution is
+an approximate answer to two different questions.
+
+If the probability density $\pf(\yF \| \yD \yI)$ is not sufficiently
+peaked, however, its mode and mean can be quite different. The
+maximum-entropy approximation can then be inadequate for two reasons.
+First, despite being the mode it can be a poor representative of all the
+possible frequency distributions around the mode. Second, it can be a bad
+approximation of the probability distribution~\eqref{eq:probability_AT}.
+
+\bigskip
+
+We therefore want to assess how well the maximum-entropy distributions of
+\fig~\ref{fig:plots_6mom} answer our two questions above. We do this by
+sampling from the probability density $\pf(\yF \| \yD \yI)$ through
+Markov-chain Monte Carlo. The density is
+\begin{equation}
+  \label{eq:density_proportional}
+  \pf(\yF \| \yD \yI) \propto
+  \Nd(\ycm\yF \| \yc, \ycs)\;
+  \Nd(\yd\yF \| 0, \yds)\;
+  \exp[-\yL\;\ysh(\yF;\yR)]
 \end{equation}
@@
   \label{eq:app_maxent_pop}
-  \pf(\yA \| \yHa)  = \frac{1}{\yK(\yL)}\,
-  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yL\T
+  \pf(\yA \| \yHa)  = \frac{1}{\yK(\yll)}\,
+  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
   \tsum_{\ya} \yg_{\ya}\ypp_{\ya\yA}\bigr).
 \end{equation}
-The $m$ Lagrange multipliers $\yL\defd (\lambda_1,\dotsc,\lambda_m)$ must
+The $m$ Lagrange multipliers $\yll\defd (\lambda_1,\dotsc,\lambda_m)$ must
 satisfy the $m$ constraint equations
@@
   \yc = \E(\yg \|\yHa)\equiv %{}\\
- \frac{1}{\yK(\yL)}\sum_{\ya} \sum_{\yA} \yg_{\ya} \ypp_{\ya \yA}
+ \frac{1}{\yK(\yll)}\sum_{\ya} \sum_{\yA} \yg_{\ya} \ypp_{\ya \yA}
 \,
-  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yL\T
+  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
   \tsum_{\ya} \yg_{\ya}\ypp_{\ya\yA}\bigr).
@@
   \label{eq:app_maxent_pop_marg}
-  \pf(\ya \| \yHa)  = \frac{1}{\yK(\yL)}\, 
+  \pf(\ya \| \yHa)  = \frac{1}{\yK(\yll)}\, 
   \sum_{\yA} \ypp_{\ya \yA}
 \,
-  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yL\T
+  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
   \tsum_{\ya} \yg_{\ya}\ypp_{\ya\yA}\bigr).
@@
 yet they cannot be exactly equal, because their equality would require the
-$2m$ quantities $\yL$ and $\yl$ to satisfy the constraint equations
+$2m$ quantities $\yll$ and $\yl$ to satisfy the constraint equations
 \eqref{eq:app_maxent_pop_constraints} and
@@
   ={}\\[-\jot]
-  \sum_\yNv \biggl\{\frac{1}{\yK(\yL_\yNv)}\, 
+  \sum_\yNv \biggl\{\frac{1}{\yK(\yll_\yNv)}\, 
   \sum_{\yA} \ypp^{(\yNv)}_{\ya\yA}
 \,
-  \tbinom{\yNv}{\yNv \yA}\,\exp\Bigl[{\yL_\yNv}\T
+  \tbinom{\yNv}{\yNv \yA}\,\exp\Bigl[{\yll_\yNv}\T
   \tsum_{\ya} \yg_{\ya}\ypp^{(\yNv)}_{\ya\yA}\Bigr]\biggr\}
@@
 \end{multline}
-where the Lagrange multipliers $\yL_{\yNv}$ and the summation range for
+where the Lagrange multipliers $\yll_{\yNv}$ and the summation range for
 $\yA$ depend on $\yNv$.

### 2019-07-10 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-10T14:04:12+0200
+%% Last-Updated: 2019-07-10T21:31:16+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\newcommand*{\yR}{\bm{R}}
+\newcommand*{\yRR}{R}
+\newcommand*{\yR}{\bm{\yRR}}
 
@@
 
-We therefore want to assess how well the maximum-entropy distributions of
-\fig~\ref{fig:plots_6mom} answer our two questions above. We do this by
-sampling from the probability density $\pf(\yF \| \yD \yI)$ through
-Markov-chain Monte Carlo. The density is
+Let's therefore examine the full density $\pf(\yF \| \yD \yI)$ given by the
+probability calculus. Its derivation is given in appendix***; here we
+briefly discuss its mathematical features and examine the answers it
+provides. Its form is this:
 \begin{equation}
@@
 \end{equation}
+and the three main terms have the following meaning:
+
+$\exp[-\yL\;\ysh(\yF;\yR)]$ expresses our pre-data uncertainty about the
+possible frequency distributions. Here
+$\ysh(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln\frac{\yFF_{\yA}}{\yRR_{\yA}}$
+is the relative entropy or discrimination information
+\citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is a
+reference distribution, and $\yL$ a positive parameter. This density is
+proportional to the number of ways in which the frequency distribution
+$\yF$ can be realized if we assume that the activities appear with
+frequencies $\yR$ in the long run, as can be seen using Stirling's
+approximation. For $\yR$ we use the uniform distribution, but any
+biologically sensible distribution, which gives higher frequencies to low
+activities, leads to the same final results. The parameter $\yL$ roughly
+corresponds to the number of data points necessary to change our initial
+belief, and we set it equal to $50$. \mynote{connection with entropic
+  priors
+  \citep{rodriguez1991,skilling1998,catichaetal2004,neumann2007,portamana2017}
+}
+
+$\Nd(\yd\yF \| 0, \yds)$ expresses our belief, hinging on biological
+arguments, that the frequency distribution should be somewhat smooth: if
+the activity levels $\yA-1$ and $\yA+1$ appear roughly equally often, we
+don't believe that the activity $\yA$ could appear with an extremely
+different frequency than theirs. The matrix $\yd$ is the discrete
+equivalent of a fourth-order derivative \mynote{ref}; the probability of
+its squared integral decreases exponentially with standard deviation
+$\yds$. We set this such that the smoothness of the maximum-entropy
+distribution is within two standard deviations, thus allowing for more
+maxima and minima in $\yF$.
+
+The two densities above together express our pre-data beliefs about the
+possible frequency distributions.
+
+$\Nd(\ycm\yF \| \yc, \ycs)$ ***
 

### 2019-07-11 maxNt_v5.tex
update

--- 2019-07-10 maxNt_v5.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-10T14:04:12+0200
+%% Last-Updated: 2019-07-11T11:28:12+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We therefore want to assess how well the maximum-entropy distributions of
-\fig~\ref{fig:plots_6mom} answer our two questions above. We do this by
-sampling from the probability density $\pf(\yF \| \yD \yI)$ through
-Markov-chain Monte Carlo. The density is
+Let's therefore examine the full density $\pf(\yF \| \yD \yI)$ given by the
+probability calculus. Its derivation is given in appendix***; here we
+ discuss its mathematical features and examine the answers it
+provides. Its expression,
 \begin{equation}
   \label{eq:density_proportional}
-  \pf(\yF \| \yD \yI) \propto
-  \Nd(\ycm\yF \| \yc, \ycs)\;
-  \Nd(\yd\yF \| 0, \yds)\;
-  \exp[-\yL\;\ysh(\yF;\yR)]
-\end{equation}
+  \pf(\yF \| \yD \yI) \propto \exp\biggl[
+-  \frac{\abs{\ycm\yF - \yc}^{2}}{2\ycs^{2}}
+-  \frac{\abs{\yd\yF}^{2}}{2\yds^{2}}
+-\yL\;\ysh(\yF;\yR)\biggr],
+\end{equation}
+consists of three terms:
+
+\textbullet\ The density $\exp[-\yL\;\ysh(\yF;\yR)]$ expresses our pre-data
+uncertainty about the possible frequency distributions. Here
+$\ysh(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln(\yFF_{\yA}/\yRR_{\yA})$ is the
+relative entropy or discrimination information
+\citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is a
+reference distribution, and $\yL$ a positive parameter. This density
+expresses a belief proportional to the number of ways in which the
+frequency distribution $\yF$ can be realized if we assume that the
+activities appear with frequencies $\yR$ in the long run, as can be seen
+using Stirling's approximation. For $\yR$ we choose the uniform
+distribution, but any biologically sensible distribution, such as one
+decreasing exponentially with the activity, leads to the same final
+results. The parameter $\yL$ roughly represents the number of data points
+necessary to change our initial belief, and we set it equal to $50$. This
+density is similar to an \enquote{entropic prior}
+\citep{rodriguez1991,skilling1998,catichaetal2004,portamana2017}[see
+esp.][\sect~3]{neumann2007}.
+
+\textbullet\ The previous density is corrected by the normal density
+$\exp\Bigl(-\frac{\abs{\yd\yF}^{2}}{2\yds^{2}}\Bigr)$, where the matrix
+$\yd$ is the discrete equivalent of a fourth-order derivative
+operator. %ref?
+This density expresses our belief, hinging on biological arguments, that
+the frequency distribution should be somewhat smooth, without huge
+discontinuities between one activity level and the next. The fourth
+derivative is chosen because it still allows for the presence of modulated
+maxima and minima more than derivatives of lower order do. The standard
+deviation $\yds$ is such that the smoothness of the maximum-entropy
+distribution, taken as a touchstone, is within two standard deviations.
+
+\textbullet\ The normal density
+$\exp\Bigl(-\frac{\abs{\ycm\yF - \yc}^{2}}{2\ycs^{2}}\Bigr)$ is the
+likelihood of the frequency $\yF$ in view of the data $\yc$. The matrix
+$\ycm$
+
+The two densities above together express our pre-data beliefs about the
+possible frequency distributions.
+
+$\Nd(\ycm\yF \| \yc, \ycs)$ ***
 

### 2019-07-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-11T11:28:12+0200
+%% Last-Updated: 2019-07-11T11:58:39+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\Nd}{\textrm{N}}
-\newcommand*{\ycm}{\mathte{M}}
+\newcommand*{\ycmm}{C}
+\newcommand*{\ycm}{\mathte{\ycmm}}
 \newcommand*{\ycs}{\sigma}
@@
 $\exp\Bigl(-\frac{\abs{\ycm\yF - \yc}^{2}}{2\ycs^{2}}\Bigr)$ is the
-likelihood of the frequency $\yF$ in view of the data $\yc$. The matrix
-$\ycm$
-
-The two densities above together express our pre-data beliefs about the
-possible frequency distributions.
-
-$\Nd(\ycm\yF \| \yc, \ycs)$ ***
+likelihood of the frequency $\yF$ in view of our data. The matrix $\ycm$
+expresses the linear constraints on $\yF$, and the tuple $\yc$ the measured
+values of such constraints. For example, if we consider the full frequency
+distribution $\yf$ of the sample, then $\ycm$ effects the marginalization
+to a sample of size $n$ through a hypergeometric distribution (sampling
+without replacement):
+\begin{equation}
+  \label{eq:constraint_hypergeometric}
+\ycmm_{\ya,\yA} \defd
+  \binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}
+\end{equation}
+and $\yc\equiv\yf$ are the measured frequencies. If we only consider some
+factorial moments, instead, then
+\begin{equation}
+  \label{eq:constraint_hypergeometric}
+  \ycmm_{m,\yA} \defd
+  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}
+\end{equation}
+as for \eqn~\eqref{eq:equal_m_factorialmoments}, and $\yc$ is the tuple of
+measured factorial moments.
 

### 2019-07-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-11T11:58:39+0200
+%% Last-Updated: 2019-07-11T13:15:49+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\yg}{\bm{c}}
-\newcommand*{\yc}{\widehat{\bm{c}}}
+\newcommand*{\yc}{\Hat{\bm{c}}}
 \newcommand*{\yll}{\bm{\lambda}}
@@
 \newcommand*{\ycm}{\mathte{\ycmm}}
-\newcommand*{\ycs}{\sigma}
+\newcommand*{\ycs}{\bm{\sigma}}
 \newcommand*{\yL}{L}
@@
   \pf(\yA_{t} \| \yD \yI) =
-  \int \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF;
+  \smallint \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF;
 \end{equation}
@@
 \end{equation}
-and $\yc\equiv\yf$ are the measured frequencies. If we only consider some
+and $\yc \defd \yf$ are the measured frequencies. If we only consider some
 factorial moments, instead, then
@@
 as for \eqn~\eqref{eq:equal_m_factorialmoments}, and $\yc$ is the tuple of
-measured factorial moments.
+measured factorial moments. The standard deviations $\ycs$ reflect our
+uncertainty about the measurement results, which can come for example from
+the spike-sorting procedure. Their values are discussed in appendix***.
+
+\bigskip
 

### 2019-07-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-11T13:15:49+0200
+%% Last-Updated: 2019-07-11T13:29:31+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \pf(\yA_{t} \| \yD \yI) =
-  \smallint \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF;
+  \int \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF;
 \end{equation}
@@
   \label{eq:approx_delta_pF}
-  \pf(\yF \| \yD \yI)\;\di\yF \approx \delta(\yF-\yFm)\;\di\yF,
+  \pf(\yF \| \yD \yI)\;\di\yF \approx \delt(\yF-\yFm)\;\di\yF,
 \end{equation}
@@
 
-\textbullet\ The density $\exp[-\yL\;\ysh(\yF;\yR)]$ expresses our pre-data
+\textbullet\ The density
+\begin{equation}
+\exp[-\yL\;\ysh(\yF;\yR)]\label{eq:entropic_prior}
+\end{equation}
+expresses our pre-data
 uncertainty about the possible frequency distributions. Here
@@
 \bigskip
+
+The post-data belief density~\eqref{eq:density_proportional} has a
+maximum-entropy distribution as a mode approximation when $\ycs$ is small,
+$\yL$ is large, and $\ycs L$ is small. Its mode in fact will be the
+constrained $\yF$ that maximizes the relative entropy $\ysh(\yF;\yR)$. It
+should be noted that a pre-data belief different from the entropic
+one~\eqref{eq:entropic_prior} will lead to a different mode approximation.
+For example, a Dirichlet density would lead to a maximum-entropy
+distribution with \emph{Burg's} entropy \citey{burg1975} instead of
+Shannon's \citep{jaynes1986d_r1996}.
 

### 2019-07-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-11T13:29:31+0200
+%% Last-Updated: 2019-07-11T14:53:17+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 our probability distribution for $\yA_{t}$ is thus simply the mean
-$\int \yF\; \pf(\yF \| \yD \yI)\;\di\yF$.
+$\int \yF \pf(\yF \| \yD \yI)\,\di\yF$.
 
@@
 
-***mode of marginal can be very different from the one in full space
-(minka1998)
-
-*** diff position of mode in marginals says that full distr is extremely
-skewed – not surprising since the base space is bounded.
-
-***in-depth discussion about underlying assumptions, derivation from
-probability calculus, higher-order approximations beyond maxent and
-"error bars" around the distributions
-
-***We can ask about probability of bimodality etc
+\bigskip
 
@@
 \\[3em]%
-\includegraphics[width=\linewidth]{_scripts/MC_samples1000.pdf}%
+\includegraphics[width=\linewidth]{_scripts/MC_marginal1000.pdf}%
 \caption{***}
-\label{fig:comparisons_marginals}
+\label{fig:MC1000}
 \end{figure}
@@
 % mcfd_6mom_unif_afssd40st0np_cok_tr_N1000_t50_L50_T0.333333333333333_s60000_a15758.0500981316
-
 \begin{figure}[!p]
@@
 \\[3em]%
-\includegraphics[width=\linewidth]{_scripts/MC_samples5000.pdf}%
+\includegraphics[width=\linewidth]{_scripts/MC_marginal5000.pdf}%
 \caption{***}
-\label{fig:comparisons_marginals}
+\label{fig:MC5000}
 \end{figure}
@@
 % mcfd_6mom_unif_afssd32s_cok_tr_N5000_t180_L50_T0.333333333333333_s25800_a329212306.071189
+The properties of the belief density~\eqref{eq:density_proportional} are
+shown in \fig~\ref{fig:MC1000} for $N=1\,000$ and in \fig~\ref{fig:MC5000}
+for $N=5\,000$, considering the first six factorial moments as our data. It
+turns out that that six moments lead to practically the same result as
+specifying the full frequency distribution for the sample; they are therefore
+informationally sufficient.
+
+The top plots in the figures show the mean
+$\int \yF \pf(\yF \| \yD \yI)\,\di\yF$
+(\textcolor{myblue}{\textbf{------}}), which is also our belief about the
+activity at any single time bin, \eqn~\eqref{eq:probability_AT}, is clearly
+different from the maximum-entropy mode. The latter is therefore not a good
+approximation for the full density~\eqref{eq:density_proportional}, which
+is quite broad, as can be seen from the ten samples
+(\textcolor{mygreen}{------}) drawn from it. The mean and the
+maximum-entropy mode are nevertheless qualitatively similar. In fact, for
+$N=1\,000$ the second peak of the mean is even higher than the
+corresponding maximum-entropy one. The samples also show that generally two
+main frequency peaks of activity are expected, especially for $N=5\,000$.
 

### 2019-07-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-11T14:53:17+0200
+%% Last-Updated: 2019-07-11T15:32:36+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
-from four moments (\textcolor{myblue}{\textbf{------}}). The lower plot
+from four moments (\textcolor{mypurpleblue}{\textbf{------}}). The lower plot
 shows the maximum-entropy distributions for the sample, from two moments
 (\textcolor{mygreen}{$\bm{\triangledown}$}) and from four moments
-(\textcolor{myblue}{$\bm{\square}$}), and the distributions obtained by
+(\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained by
 marginalizing the full-network distribution to the sample size, from two
 moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
-moments (\textcolor{myblue}{$\bm{\blacksquare}$}). The measured frequency
+moments (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The measured frequency
 distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also shown. We
@@
   distributions (\textcolor{mygreen}{$\bm{\triangledown}$} and
-  \textcolor{myblue}{$\bm{\square}$}), which can only be distinguished on a
+  \textcolor{mypurpleblue}{$\bm{\square}$}), which can only be distinguished on a
   logarithmic scale. This could lead to the erroneous conclusion that the
@@
 $\int \yF \pf(\yF \| \yD \yI)\,\di\yF$
-(\textcolor{myblue}{\textbf{------}}), which is also our belief about the
+(\textcolor{mypurpleblue}{\textbf{------}}), which is also our belief about the
 activity at any single time bin, \eqn~\eqref{eq:probability_AT}, is clearly
@@
 is quite broad, as can be seen from the ten samples
-(\textcolor{mygreen}{------}) drawn from it. The mean and the
-maximum-entropy mode are nevertheless qualitatively similar. In fact, for
-$N=1\,000$ the second peak of the mean is even higher than the
-corresponding maximum-entropy one. The samples also show that generally two
-main frequency peaks of activity are expected, especially for $N=5\,000$.
+(\textcolor{mygreen}{------}) drawn from it. This broadness is not
+surprising: we're making guesses about thousands of neurons from the
+observation of fewer than 100. Yet, the mean and the maximum-entropy mode
+are qualitatively similar. In fact, for $N=1\,000$ the second peak of the
+mean is even higher than the corresponding maximum-entropy one. The samples
+also show that generally two high-frequency regions of activity are
+expected, especially for $N=5\,000$.
+
+The bottom plots in the figures show our belief distributions for the
+marginal frequency of each activity level, $\pf(\yFF_{\yA} \| \yD \yI)$,
+with the averages (\textcolor{mypurpleblue}{\textbf{------}}), medians
+(\textcolor{mygreen}{\textbf{--~--~--}}), and 16\% and 84\% quantiles
+(\textcolor{myyellow}{$\bm{\cdot \cdot \cdot \cdot}$} corresponding to one
+standard deviation for a normal distribution).
 

### 2019-07-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-11T15:32:36+0200
+%% Last-Updated: 2019-07-11T15:55:31+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 our probability distribution for $\yA_{t}$ is thus simply the mean
-$\int \yF \pf(\yF \| \yD \yI)\,\di\yF$.
+$\int \yF\, \pf(\yF \| \yD \yI)\,\di\yF$.
 
@@
 For example, a Dirichlet density would lead to a maximum-entropy
-distribution with \emph{Burg's} entropy \citey{burg1975} instead of
+distribution with \emph{Burg's} entropy \citep{burg1975} instead of
 Shannon's \citep{jaynes1986d_r1996}.
@@
 % mcfd_6mom_unif_afssd32s_cok_tr_N5000_t180_L50_T0.333333333333333_s25800_a329212306.071189
-The properties of the belief density~\eqref{eq:density_proportional} are
-shown in \fig~\ref{fig:MC1000} for $N=1\,000$ and in \fig~\ref{fig:MC5000}
-for $N=5\,000$, considering the first six factorial moments as our data. It
-turns out that that six moments lead to practically the same result as
-specifying the full frequency distribution for the sample; they are therefore
-informationally sufficient.
+The properties of the belief density~\eqref{eq:density_proportional} in the
+case of our first data set are shown in \fig~\ref{fig:MC1000} for
+$N=1\,000$ and in \fig~\ref{fig:MC5000} for $N=5\,000$. Only the first six
+factorial moments were considered as our data, but it turns out that they
+are informationally sufficient, leading to practically the same result as
+specifying the full frequency distribution for the sample.
 
 The top plots in the figures show the mean
-$\int \yF \pf(\yF \| \yD \yI)\,\di\yF$
+$\int \yF\, \pf(\yF \| \yD \yI)\,\di\yF$
 (\textcolor{mypurpleblue}{\textbf{------}}), which is also our belief about the
@@
 marginal frequency of each activity level, $\pf(\yFF_{\yA} \| \yD \yI)$,
-with the averages (\textcolor{mypurpleblue}{\textbf{------}}), medians
+with the means (\textcolor{mypurpleblue}{\textbf{------}}), medians
 (\textcolor{mygreen}{\textbf{--~--~--}}), and 16\% and 84\% quantiles
 (\textcolor{myyellow}{$\bm{\cdot \cdot \cdot \cdot}$} corresponding to one
-standard deviation for a normal distribution).
+standard deviation for a normal distribution). The coincidence of means
+and medians indicate that these marginals are very close to normal
+distributions. This happens thanks to the marginalization, which integrates
+out $N-2$ quantities. Note, however, that the full distribution for $\yF$
+is far from normal, as indicated by the pronounced differences between mode
+and mean.
+
+\bigskip
+
+These results caution us against taking maximum-entropy solutions too
+literally. The full probabilistic analysis allows us to see more clearly
+the extent of our uncertainty and to make better-informed guesses.
 

### 2019-07-11 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-11T15:55:31+0200
+%% Last-Updated: 2019-07-11T16:10:41+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 distribution with \emph{Burg's} entropy \citep{burg1975} instead of
-Shannon's \citep{jaynes1986d_r1996}.
+Shannon's \citep{jaynes1986d_r1996,portamana2009}.
 

### 2019-07-12 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-11T16:10:41+0200
+%% Last-Updated: 2019-07-12T11:25:44+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Answering this kind of questions always engages an element of uncertainty.
-Our answers therefore involve both experimental data, such as neuronal
-recordings from specific brain areas, and probabilistic assumptions --
-probabilistic \enquote{models} -- about biological conditions and
-mechanisms that cannot be experimentally ascertained.
+Our answers therefore involve experimental data, such as neuronal
+recordings from specific brain areas, and probabilities or \dobs, based on
+prior knowledge, about biological conditions and mechanisms that cannot be
+experimentally ascertained. Such \dobs\ are often formulated as simplified
+\enquote{models} to be mathematically more tractable.
 
@@
 experimental measurements of neuronal activity can still only record a very
-small sample of neurons compared to the numbers that constitute one
+small sample of neurons compared to the numbers that constitute a
 functionally distinguished brain region. Many probabilistic models focus on
@@
 
-In the present work we give an answer to this question: How much is our
-uncertainty about the \emph{total activity} of a large neuronal network, if
-we have observed the activity of a very small sample thereof? We'll first
-give an answer by combining, in a straightforward way, the maximum-entropy
-method and basic sampling relations of the probability calculus. Later on
-we'll show that this answer can be obtained by exclusively using the
-probability calculus; this point of view shows that the maximum-entropy
-answer is a first approximation to the answers of \emph{four} different
-questions, and also shows the assumptions lying behind it.
+In the present work we give an answer to this question: How much can the
+total activity of a large neuronal network be, if we have observed the
+activity of a very small sample thereof? We'll quantify our \dobs\ about
+the possible answers by combining, in a straightforward way, the
+maximum-entropy method and basic sampling relations of the probability
+calculus. Later on we'll show that our \dobs\ can be quantified by
+exclusively using the probability calculus. This derivation will provide a
+more accurate quantification, revealing that the maximum-entropy answer is
+only a first approximation.
 
@@
 interesting finding is that the most probable frequency distribution for
-the total activity of the full network, if the latter is assumed to be
-larger than 1\,000 neurons, has two very distinct modes, both at low
-activities ($0.7\%$ and $2.6\%$ of simultaneously active neurons within
-$3\;\mathrm{ms}$, see \fig***). The analogous distribution for second data
-set doesn't have two modes but still presents one prominent shoulder in its
-one low-activity mode. Note that these inferred features for the full
-network aren't observed in the sample, nor are they inferred by the
-application of maximum-entropy to the sample alone. \mynote{if we manage
-  the Monte Carlo sampling or other approx: say something about the
-  deviation from the most probable frequency distribution}
+the total activity of the full network has two very distinct modes, both at
+low activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
+% within $3\;\mathrm{ms}$,
+see \fig~\ref{fig:plots_6mom}. The analogous frequency distribution for the
+second data set doesn't have two modes but still presents one prominent
+shoulder in its one low-activity mode. Note that these guessed features of
+the full network aren't observed in the sample, nor can they be inferred by
+the application of maximum-entropy \emph{to the sample alone}.
+\mynote{Monte Carlo sampling: say something about the deviation from the
+  most probable frequency distribution}
 
@@
 
-The maximum-entropy method \citep{jaynes1957}[much clearer
+The maximum-entropy or minimum-relative-entropy method
+\citep{jaynes1957}[much clearer
 in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b,grandy1980}
@@
 example second-order time correlations, is sufficient for quantifying our
-uncertainty about the data. It can be considered as an approximation of
-probabilistic models based on various assumptions of inferential
-sufficiency \citep{jaynes1986d_r1996,portamana2017} -- for example of
-empirical moments or of long-run frequencies.
+\dob\ about some quantities of the system. It can be considered as an
+approximation of probabilistic models based on various assumptions of
+inferential sufficiency \citep{jaynes1986d_r1996,portamana2017}.
 
@@
   clear. Orig intro is on p.~\pageref{orig_intro}}
+
+Our notation and terminology follow \textsc{iso} standards
+\citep{iso1993,iso2006,iso2006b} and Jaynes \citey{jaynes1994_r2003} for
+\dobs. We often simply say \enquote{belief} for \enquote{\dob}.
 
@@
 during bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during
-  time bin $t$}. From the activities $\set{\ya_{t}}$ we can count how many
-times the activity levels $\ya=0$, $\ya=1/n$, and so on appeared during the
-recording, obtaining the measured relative-frequency distribution
+  time bin $t$}. From the activities $\set{\ya_{t}}$ we can count how often
+the activity levels $\ya=0$, $\ya=1/n$, and so on appeared during the
+recording, obtaining the distribution of measured relative frequencies
 $(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
@@
 sessions, as the set of $n$ neurons that was actually recorded in this
-session. This is our \enquote{full network}\mynote{how about calling it
-  \enquote{the pool}?}. The normalized total activity of these $N$ neurons
+session. This is our \enquote{full network}.\mynote{how about calling it
+  \enquote{the pool}?} The normalized total activity of these $N$ neurons
 at $t$ is $\yA_{t}$. The relative frequencies of the various activity
@@
 during a recording? We cannot answer these questions with certainty; we can
-only give distributions of probability over their possible answers. The
+only give distributions of probability or \dobs\ over their possible answers. The
 approach presented here gives such probability distributions.
@@
 \mynote{Move this to intro?}We want to stress the usefulness of making
-quantified guesses about the full-network activity. First of all, this seems to be
-the primary idea behind recording a sample. Second, it allows us to make
-comparisons across experimental sessions; such comparisons would me
-difficult or meaningless made at the level of recorded samples, which
-generally comprise non-overlapping sets of neurons differing also in size.
+quantified guesses about the full-network activity. First of all, this
+seems to be the primary idea behind recording a sample. Second, it allows
+us to make comparisons across experimental sessions; such comparisons would
+me difficult or meaningless if made with the recorded samples, which
+generally comprise non-overlapping sets of neurons and differ in size.
 
@@
 Regarding~\ref{item:maxent_step}, we assume that you're familiar with the
-maximum-entropy method. In short, it amounts to a pair of prescriptions:
-choose the distribution having maximum Shannon entropy among those
-satisfying specific convex constraints, such as fixed averages; and judge
-these averages to be equal to measured averages. We add two remarks about
-this method that are seldom made in the literature. First, the distribution
-$\px(\yA)$ given by this method is the zeroth-order approximation \citep[in
-the sense of Laplace's
+maximum-entropy method. We actually use the \emph{minimum-relative-entropy}
+method \citep{hobsonetal1973}, but call it \enquote{maximum-entropy} for
+brevity. It amounts to a pair of prescriptions: choose the distribution,
+among those satisfying specific convex constraints, such as fixed
+expectations, that has minimum relative entropy with respect to a reference
+distribution, often taken to be the uniform one; and judge those expectations
+to be equal to measured averages. We add two remarks about this method that
+are seldom made in the literature. First, the distribution $\px(\yA)$ given
+by this method is the zeroth-order approximation \citep[in the sense of
+Laplace's
 method:][\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
@@
   activity across the \emph{recorded} bins, %$\yFF(\yA)$,
+\item the \emph{belief} distribution for the value of the total
+  activity at any time bin among those \emph{recorded},
+  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
 \item the most probable \emph{frequency} distribution for the total
   activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
-\item the \emph{probability} distribution for the value of the total
-  activity at any time bin among those \emph{recorded},
-  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
-\item the \emph{probability} distribution for the value of the total
+\item the \emph{belief} distribution for the value of the total
   activity at a \emph{new} time bin.
@@
 \end{itemize}
-The maximum-entropy distribution is thus an approximation of our
-uncertainty about four completely different quantities. Note that the four
+The maximum-entropy distribution is thus an approximation of our belief
+distribution about four completely different quantities. Note that the four
 distributions above numerically differ in higher-order approximations. We
@@
 sufficient statistics, such as the sufficiency of pairwise correlations to
-quantify our uncertainty about network activity. In the present work we
+quantify our \dob\ about network activity. In the present work we
 only want to introduce the general method without entering into biological
@@
 distribution for the activity of the sample neurons. In this case, our
-uncertainty about the full-network activity is based on all available
+belief about the full-network activity is based on all available
 measured frequency data. Note that application of the maximum-entropy
@@
 appeared in $T\times\yFF(\yA)$ out of $T$ time bins. We don't know $\yF$,
-and our uncertainty about its possible values is a probability distribution
+and our beliefs about its possible values have a distribution
 $\pf(\yF\|\yD\yI)$ conditional on data $\yD$ and prior knowledge $\yI$,
-which we approximate with a continuous density
+which we approximate with a continuous distribution
 \begin{equation}
@@
 \end{equation}
-In appendix*** we give a summary of how this density is derived. Note that
+In appendix*** we give a summary of how this distribution is derived. Note that
 this is a probability distribution over frequency distributions, so we must
@@
 if we know the frequency distribution $\yF$ of the activity levels, then
-our probability for $\yA_{t}$ is $\yFF_{\yA_{t}}$, out of combinatorial
+our \dob\ about $\yA_{t}$ is $\yFF_{\yA_{t}}$, out of combinatorial
 reasons. But if we are also uncertain about $\yF$, with
-probability~\eqref{eq:density_F}, then our probability for $\yA_{t}$ by
+belief distribution~\eqref{eq:density_F}, then our \dob\ about $\yA_{t}$ by
 marginalization is
@@
 \end{equation}
-our probability distribution for $\yA_{t}$ is thus simply the mean
+our belief distribution for $\yA_{t}$ is thus simply the mean
 $\int \yF\, \pf(\yF \| \yD \yI)\,\di\yF$.
@@
 expresses our pre-data
-uncertainty about the possible frequency distributions. Here
+beliefs about the possible frequency distributions. Here
 $\ysh(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln(\yFF_{\yA}/\yRR_{\yA})$ is the
@@
 measured factorial moments. The standard deviations $\ycs$ reflect our
-uncertainty about the measurement results, which can come for example from
+beliefs about the measurement results, which can come for example from
 the spike-sorting procedure. Their values are discussed in appendix***.
@@
 maximum-entropy distribution as a mode approximation when $\ycs$ is small,
-$\yL$ is large, and $\ycs L$ is small. Its mode in fact will be the
-constrained $\yF$ that maximizes the relative entropy $\ysh(\yF;\yR)$. It
-should be noted that a pre-data belief different from the entropic
+$\yL$ is large, and $\ycs L$ is small: the mode will be the constrained
+$\yF$ that minimizes the relative entropy $\ysh(\yF;\yR)$. It should be
+noted that a pre-data belief different from the entropic
 one~\eqref{eq:entropic_prior} will lead to a different mode approximation.

### 2019-07-13 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-12T11:25:44+0200
+%% Last-Updated: 2019-07-13T11:35:37+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \mynote{maybe move this paragraph to the intro?}Present-day technologies
-enable the recording of neuronal activity from small brain areas
-\mynote{be more specific}. For many animal species, the neurons that are
-recorded within the area are not -- and at present cannot be --
-specifically chosen from among the rest, owing to several limiting factors;
-for example, limitations in the precision with which electrodes are
-inserted, or with which neurons are targeted by viruses. In fact, the set
-of recorded neurons may even change during very long recordings or across
-experimental sessions. We assume that there's a area, comprising a
-network of $N$ neurons, from which other sets of neurons had or will have
-the same probability of being recorded, even in other experimental
-sessions, as the set of $n$ neurons that was actually recorded in this
-session. This is our \enquote{full network}.\mynote{how about calling it
-  \enquote{the pool}?} The normalized total activity of these $N$ neurons
-at $t$ is $\yA_{t}$. The relative frequencies of the various activity
-levels during the recording were $(\yFF_{\yA}) \defs \yF$. We don't know the
-values $\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We only
-know for certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
-$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the time being we assume
-that we know $N$; in \sect*** we discuss the consequences of our lack of
-precise knowledge about this number.
+enable the recording of neuronal activity from small brain areas \mynote{be
+  more specific}. For many animal species, the neurons that are recorded
+within the area are not -- and at present cannot be -- specifically chosen
+from among the rest, owing to several limiting factors; for example,
+limitations in how precisely electrodes are inserted or neurons are
+targeted by viruses. In fact, the set of recorded neurons may even change
+during very long recordings or across experimental sessions. We assume that
+there's an area, comprising a network of $N$ neurons, from which other sets
+of neurons could have been or will be recorded***have the same probability of being recorded, even in
+other experimental sessions, as the set of $n$ neurons that was actually
+recorded in this session. This is our \enquote{full network}.\mynote{how
+  about calling it \enquote{the pool}?} The normalized total activity of
+these $N$ neurons at $t$ is $\yA_{t}$. The relative frequencies of the
+various activity levels during the recording were $(\yFF_{\yA}) \defs \yF$.
+We don't know the values $\yA_{t}$ at each $t$, or the frequency
+distribution $\yF$. We only know for certain that
+$\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
+for obvious reasons. For the time being we assume that we know $N$; in
+\sect*** we discuss the consequences of our lack of precise knowledge about
+this number.
 
@@
 one~\eqref{eq:entropic_prior} will lead to a different mode approximation.
-For example, a Dirichlet density would lead to a maximum-entropy
-distribution with \emph{Burg's} entropy \citep{burg1975} instead of
-Shannon's \citep{jaynes1986d_r1996,portamana2009}.
+For example, a Dirichlet density (which is equivalent
+to~\eqref{eq:entropic_prior} with $\yF$ and $\yR$ exchanged) would lead to
+a maximum-entropy distribution with \emph{Burg's} entropy \citep{burg1975}
+instead of Shannon's \citep{jaynes1986d_r1996,portamana2009}.
 
@@
 \includegraphics[width=\linewidth]{_scripts/MC_marginal1000.pdf}%
-\caption{***}
+\caption{Entropic prior}
 \label{fig:MC1000}
@@
 % mcfd_6mom_unif_afssd40st0np_cok_tr_N1000_t50_L50_T0.333333333333333_s60000_a15758.0500981316
+%
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\linewidth]{_scripts/MC_plot_unif1000.pdf}%
+\\[3em]%
+\includegraphics[width=\linewidth]{_scripts/MC_marginal_unif1000.pdf}%
+\caption{Prior uniform in $\di\yF$}
+\label{fig:MC1000}
+\end{figure}
+% mctests3.nb
+% mcfd_6mom_uunif_afssd40st0np_cok_tr_N1000_t50_L0_T0.333333333333333_s60000_a15758.0500981316
+%
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\linewidth]{_scripts/MC_plot_dirich1000.pdf}%
+\\[3em]%
+\includegraphics[width=\linewidth]{_scripts/MC_marginal_dirich1000.pdf}%
+\caption{Dirichlet prior (entropic with reversed relative entropy)}
+\label{fig:MC1000}
+\end{figure}
+% mctests3.nb
+% mcfd_6mom_iunif_afssd40st0np_cok_tr_N1000_t50_L1_T0.333333333333333_s60000_a15758.0500981316
+%
 \begin{figure}[!p]
@@
 \includegraphics[width=\linewidth]{_scripts/MC_marginal5000.pdf}%
-\caption{***}
+\caption{Entropic prior}
 \label{fig:MC5000}

### 2019-07-20 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-13T11:35:37+0200
+%% Last-Updated: 2019-07-20T12:48:31+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\mynote{Move this to intro?}We want to stress the usefulness of making
-quantified guesses about the full-network activity. First of all, this
-seems to be the primary idea behind recording a sample. Second, it allows
-us to make comparisons across experimental sessions; such comparisons would
-me difficult or meaningless if made with the recorded samples, which
-generally comprise non-overlapping sets of neurons and differ in size.
+\mynote{Move this to intro?}We want to stress the usefulness of making a
+quantified guess about the full-network activity. First of all, this guess
+about a brain area seems to be the primary idea behind recording a sample
+from that area. Second, it allows us to make comparisons across
+experimental sessions; such comparisons would me difficult or meaningless
+if made with the recorded samples, which generally comprise non-overlapping
+sets of neurons and differ in size. Third, only at the full-network level
+can we meaningfully assess whether correlations of some order are
+informationally sufficient: as explained in
+\sect~\ref{sec:marginalization}, any such assessment at the sample level
+leads to conclusions opposite to the ones we think we're reaching.
 
@@
 calculus thus reveals interesting possible features of the full network.
+\textcolor{white}{If you find this you can claim a postcard from us.}
 
@@
 
-\section{Assumptions and corrections:\\beyond the maximum-entropy approximation}
-\label{sec:assumptions_formulae}
-
-In \sect~\ref{sec:model} we mentioned that maximum-entropy distributions
-come from a zeroth-order Laplace approximation of the densities obtained
-from the probability calculus. Zeroth-order means that we are simply
-considering the mode of such densities. Let's be more concrete.
-
-Consider the question: What is the relative-frequency distribution for the
-different activity levels for the full-network, across the recorded bins?
-Call this distribution $(\yFF_{\yA}) \defs \yF$: activity level $\yA$
-appeared in $T\times\yFF(\yA)$ out of $T$ time bins. We don't know $\yF$,
-and our beliefs about its possible values have a distribution
-$\pf(\yF\|\yD\yI)$ conditional on data $\yD$ and prior knowledge $\yI$,
-which we approximate with a continuous distribution
-\begin{equation}
-  \label{eq:density_F}
-  \pf(\yF \| \yD \yI)\;\di\yF.
-\end{equation}
-In appendix*** we give a summary of how this distribution is derived. Note that
-this is a probability distribution over frequency distributions, so we must
-be careful with terminology to avoid confusion.
-
-Now consider the related question: What is the activity $\yA_{t}$ of the
-full network at some specific time bin $t$? If we don't know the exact
-sequence of activities during the recording then we don't know $\yA_{t}$;
-if we know the frequency distribution $\yF$ of the activity levels, then
-our \dob\ about $\yA_{t}$ is $\yFF_{\yA_{t}}$, out of combinatorial
-reasons. But if we are also uncertain about $\yF$, with
-belief distribution~\eqref{eq:density_F}, then our \dob\ about $\yA_{t}$ by
-marginalization is
-\begin{equation}
-  \label{eq:probability_AT}
-  \pf(\yA_{t} \| \yD \yI) =
-  \int \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF;
-\end{equation}
-our belief distribution for $\yA_{t}$ is thus simply the mean
-$\int \yF\, \pf(\yF \| \yD \yI)\,\di\yF$.
-
-Formulae~\eqref{eq:density_F} and \eqref{eq:probability_AT} show that the
-most probable frequency distribution $\yF$ (if it's unique) and the
-probability distribution for $\yA_{t}$ are generally different: the former
-is the mode of $\pf(\yF \| \yD \yI)$, the latter its mean.
-
-If the probability density $\pf(\yF \| \yD \yI)$ is extremely peaked at some
-frequency distribution $\yFm$, it can be approximately treated as a delta
-density,
-\begin{equation}
-  \label{eq:approx_delta_pF}
-  \pf(\yF \| \yD \yI)\;\di\yF \approx \delt(\yF-\yFm)\;\di\yF,
-\end{equation}
-and consequently the probability for the activity at bin $t$, by
-\eqn~\eqref{eq:probability_AT}, is simply the mode $\yFm$ itself:
-\begin{equation}
-  \label{eq:approx_delta_pA}
-  \pf(\yA_{t} \| \yD \yI) \approx \yFFm_{\yA_{t}}.
-\end{equation}
-The maximum-entropy distribution is exactly this mode $\yFm$, for
-\emph{particular} kinds of probability densities $\pf(\yF \| \yD \yI)$
-which we'll discuss shortly. Formulae~\eqref{eq:approx_delta_pF}
-and~\eqref{eq:approx_delta_pA} show why the maximum-entropy distribution is
-an approximate answer to two different questions.
-
-If the probability density $\pf(\yF \| \yD \yI)$ is not sufficiently
-peaked, however, its mode and mean can be quite different. The
-maximum-entropy approximation can then be inadequate for two reasons.
-First, despite being the mode it can be a poor representative of all the
-possible frequency distributions around the mode. Second, it can be a bad
-approximation of the probability distribution~\eqref{eq:probability_AT}.
-
-\bigskip
-
-Let's therefore examine the full density $\pf(\yF \| \yD \yI)$ given by the
-probability calculus. Its derivation is given in appendix***; here we
- discuss its mathematical features and examine the answers it
-provides. Its expression,
-\begin{equation}
-  \label{eq:density_proportional}
-  \pf(\yF \| \yD \yI) \propto \exp\biggl[
--  \frac{\abs{\ycm\yF - \yc}^{2}}{2\ycs^{2}}
--  \frac{\abs{\yd\yF}^{2}}{2\yds^{2}}
--\yL\;\ysh(\yF;\yR)\biggr],
-\end{equation}
-consists of three terms:
-
-\textbullet\ The density
-\begin{equation}
-\exp[-\yL\;\ysh(\yF;\yR)]\label{eq:entropic_prior}
-\end{equation}
-expresses our pre-data
-beliefs about the possible frequency distributions. Here
-$\ysh(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln(\yFF_{\yA}/\yRR_{\yA})$ is the
-relative entropy or discrimination information
-\citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is a
-reference distribution, and $\yL$ a positive parameter. This density
-expresses a belief proportional to the number of ways in which the
-frequency distribution $\yF$ can be realized if we assume that the
-activities appear with frequencies $\yR$ in the long run, as can be seen
-using Stirling's approximation. For $\yR$ we choose the uniform
-distribution, but any biologically sensible distribution, such as one
-decreasing exponentially with the activity, leads to the same final
-results. The parameter $\yL$ roughly represents the number of data points
-necessary to change our initial belief, and we set it equal to $50$. This
-density is similar to an \enquote{entropic prior}
-\citep{rodriguez1991,skilling1998,catichaetal2004,portamana2017}[see
-esp.][\sect~3]{neumann2007}.
-
-\textbullet\ The previous density is corrected by the normal density
-$\exp\Bigl(-\frac{\abs{\yd\yF}^{2}}{2\yds^{2}}\Bigr)$, where the matrix
-$\yd$ is the discrete equivalent of a fourth-order derivative
-operator. %ref?
-This density expresses our belief, hinging on biological arguments, that
-the frequency distribution should be somewhat smooth, without huge
-discontinuities between one activity level and the next. The fourth
-derivative is chosen because it still allows for the presence of modulated
-maxima and minima more than derivatives of lower order do. The standard
-deviation $\yds$ is such that the smoothness of the maximum-entropy
-distribution, taken as a touchstone, is within two standard deviations.
-
-\textbullet\ The normal density
-$\exp\Bigl(-\frac{\abs{\ycm\yF - \yc}^{2}}{2\ycs^{2}}\Bigr)$ is the
-likelihood of the frequency $\yF$ in view of our data. The matrix $\ycm$
-expresses the linear constraints on $\yF$, and the tuple $\yc$ the measured
-values of such constraints. For example, if we consider the full frequency
-distribution $\yf$ of the sample, then $\ycm$ effects the marginalization
-to a sample of size $n$ through a hypergeometric distribution (sampling
-without replacement):
-\begin{equation}
-  \label{eq:constraint_hypergeometric}
-\ycmm_{\ya,\yA} \defd
-  \binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}
-\end{equation}
-and $\yc \defd \yf$ are the measured frequencies. If we only consider some
-factorial moments, instead, then
-\begin{equation}
-  \label{eq:constraint_hypergeometric}
-  \ycmm_{m,\yA} \defd
-  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}
-\end{equation}
-as for \eqn~\eqref{eq:equal_m_factorialmoments}, and $\yc$ is the tuple of
-measured factorial moments. The standard deviations $\ycs$ reflect our
-beliefs about the measurement results, which can come for example from
-the spike-sorting procedure. Their values are discussed in appendix***.
-
-\bigskip
-
-The post-data belief density~\eqref{eq:density_proportional} has a
-maximum-entropy distribution as a mode approximation when $\ycs$ is small,
-$\yL$ is large, and $\ycs L$ is small: the mode will be the constrained
-$\yF$ that minimizes the relative entropy $\ysh(\yF;\yR)$. It should be
-noted that a pre-data belief different from the entropic
-one~\eqref{eq:entropic_prior} will lead to a different mode approximation.
-For example, a Dirichlet density (which is equivalent
-to~\eqref{eq:entropic_prior} with $\yF$ and $\yR$ exchanged) would lead to
-a maximum-entropy distribution with \emph{Burg's} entropy \citep{burg1975}
-instead of Shannon's \citep{jaynes1986d_r1996,portamana2009}.
-
-\bigskip
-
-\begin{figure}[!p]
-\centering
-\includegraphics[width=\linewidth]{_scripts/MC_plot1000.pdf}%
-\\[3em]%
-\includegraphics[width=\linewidth]{_scripts/MC_marginal1000.pdf}%
-\caption{Entropic prior}
-\label{fig:MC1000}
-\end{figure}
-% mctests3.nb
-% mcfd_6mom_unif_afssd40st0np_cok_tr_N1000_t50_L50_T0.333333333333333_s60000_a15758.0500981316
-%
-\begin{figure}[!p]
-\centering
-\includegraphics[width=\linewidth]{_scripts/MC_plot_unif1000.pdf}%
-\\[3em]%
-\includegraphics[width=\linewidth]{_scripts/MC_marginal_unif1000.pdf}%
-\caption{Prior uniform in $\di\yF$}
-\label{fig:MC1000}
-\end{figure}
-% mctests3.nb
-% mcfd_6mom_uunif_afssd40st0np_cok_tr_N1000_t50_L0_T0.333333333333333_s60000_a15758.0500981316
-%
-\begin{figure}[!p]
-\centering
-\includegraphics[width=\linewidth]{_scripts/MC_plot_dirich1000.pdf}%
-\\[3em]%
-\includegraphics[width=\linewidth]{_scripts/MC_marginal_dirich1000.pdf}%
-\caption{Dirichlet prior (entropic with reversed relative entropy)}
-\label{fig:MC1000}
-\end{figure}
-% mctests3.nb
-% mcfd_6mom_iunif_afssd40st0np_cok_tr_N1000_t50_L1_T0.333333333333333_s60000_a15758.0500981316
-%
-\begin{figure}[!p]
-\centering
-\includegraphics[width=\linewidth]{_scripts/MC_plot5000.pdf}%
-\\[3em]%
-\includegraphics[width=\linewidth]{_scripts/MC_marginal5000.pdf}%
-\caption{Entropic prior}
-\label{fig:MC5000}
-\end{figure}
-% mctests3.nb
-% mcfd_6mom_unif_afssd32s_cok_tr_N5000_t180_L50_T0.333333333333333_s25800_a329212306.071189
-The properties of the belief density~\eqref{eq:density_proportional} in the
-case of our first data set are shown in \fig~\ref{fig:MC1000} for
-$N=1\,000$ and in \fig~\ref{fig:MC5000} for $N=5\,000$. Only the first six
-factorial moments were considered as our data, but it turns out that they
-are informationally sufficient, leading to practically the same result as
-specifying the full frequency distribution for the sample.
-
-The top plots in the figures show the mean
-$\int \yF\, \pf(\yF \| \yD \yI)\,\di\yF$
-(\textcolor{mypurpleblue}{\textbf{------}}), which is also our belief about the
-activity at any single time bin, \eqn~\eqref{eq:probability_AT}, is clearly
-different from the maximum-entropy mode. The latter is therefore not a good
-approximation for the full density~\eqref{eq:density_proportional}, which
-is quite broad, as can be seen from the ten samples
-(\textcolor{mygreen}{------}) drawn from it. This broadness is not
-surprising: we're making guesses about thousands of neurons from the
-observation of fewer than 100. Yet, the mean and the maximum-entropy mode
-are qualitatively similar. In fact, for $N=1\,000$ the second peak of the
-mean is even higher than the corresponding maximum-entropy one. The samples
-also show that generally two high-frequency regions of activity are
-expected, especially for $N=5\,000$.
-
-The bottom plots in the figures show our belief distributions for the
-marginal frequency of each activity level, $\pf(\yFF_{\yA} \| \yD \yI)$,
-with the means (\textcolor{mypurpleblue}{\textbf{------}}), medians
-(\textcolor{mygreen}{\textbf{--~--~--}}), and 16\% and 84\% quantiles
-(\textcolor{myyellow}{$\bm{\cdot \cdot \cdot \cdot}$} corresponding to one
-standard deviation for a normal distribution). The coincidence of means
-and medians indicate that these marginals are very close to normal
-distributions. This happens thanks to the marginalization, which integrates
-out $N-2$ quantities. Note, however, that the full distribution for $\yF$
-is far from normal, as indicated by the pronounced differences between mode
-and mean.
-
-\bigskip
+\section{Summary and discussion}
+\label{sec:summary_discussion}
 
@@
 the extent of our uncertainty and to make better-informed guesses.
-
-\section{Summary and discussion}
-\label{sec:summary_discussion}
 
@@
 dependence.
-\textcolor{white}{If you find this you can claim a postcard from me.}
 

### 2019-07-20 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-20T12:48:31+0200
+%% Last-Updated: 2019-07-20T18:44:47+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 during very long recordings or across experimental sessions. We assume that
-there's an area, comprising a network of $N$ neurons, from which other sets
-of neurons could have been or will be recorded***have the same probability of being recorded, even in
-other experimental sessions, as the set of $n$ neurons that was actually
-recorded in this session. This is our \enquote{full network}.\mynote{how
-  about calling it \enquote{the pool}?} The normalized total activity of
-these $N$ neurons at $t$ is $\yA_{t}$. The relative frequencies of the
-various activity levels during the recording were $(\yFF_{\yA}) \defs \yF$.
-We don't know the values $\yA_{t}$ at each $t$, or the frequency
-distribution $\yF$. We only know for certain that
+there's an area, comprising a network of $N$ neurons, for which we equally
+believe that any other sample of size $n$ could have been recorded instead
+of the sample of $n$ neurons that was actually recorded. This is our
+\enquote{full network}.\mynote{how about calling it \enquote{the pool}?}
+The normalized total activity of these $N$ neurons at $t$ is $\yA_{t}$. The
+relative frequencies of the various activity levels during the recording
+were $(\yFF_{\yA}) \defs \yF$. We don't know the values $\yA_{t}$ at each
+$t$, or the frequency distribution $\yF$. We only know for certain that
 $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
@@
 informational sufficiency of the two sets of statistics.
+
+\section{The full-network size $N$}
+\label{sec:N}
+
+The calculations and conclusions presented in the preceding sections depend
+on the size of the full network, $N$. The full network can't be the whole
+brain, of course; that would be silly. How large can $N$ be?
+
+The crucial point is formula~\eqref{eq:equal_m_factorialmoments}, on which
+our method is based, and which asserts the equality of the factorial
+moments of the full network and its sample, (or, equivalently, asserting
+that the power moments for the one are linear functions of the power
+moments for the other). This formula is only valid if our beliefs about the
+activity of the sample  $\pxx(\ya)$ and that of the full network
+$\pxx(\yA)$ are related by a hypergeometric distribution:
+\begin{equation}
+  \label{eq:hypergeometric_relation}
+    \pxx(\ya)=
+  \binom{n}{n\ya}\;\binom{\yNv-n}{\yNv \yA-n\ya}\;\binom{\yNv}{\yNv \yA}^{-1}
+\;\pxx(\yA).
+\end{equation}
+This is a relation of \enquote{drawing without replacement}
+\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
+In other words, $N$ must be such that we equally believe other samples of
+size $n$ could have been recorded by the electrodes or the probe used. This
+is indeed what we stated in \sect~\ref{sec:model}.
+
+This requirement delimits an area around the probe. Its number of neurons
+depends on the animal species and on the brain region from which the
+recording was made, as the neuron density can be very different. It's
+practically impossible to specify the exact number of neurons, but its
+order of magnitude is enough for making qualitative inferences. As we'll
+discuss in the next section, even if the exact number were known the
+maximum-entropy distribution ought still to be interpreted qualitatively.
+
+It would be possible to derive a distribution $\pxx(A)$ for the full
+network based on an unknown $N$, by  expressing our belief distribution
+$\pxx(N)$ about $N$ and using this to marginalize $N$ out:
+\begin{equation}
+  \label{eq:marginalize_N}
+  \pxx(A \| \text{$N$ \small unknown}) = \sum_{N} \px(\yA \| N)\;\pxx(N).
+\end{equation}
+But this would be overkill, owing to the qualitative character of the
+maximum-entropy distribution constructed with exact $N$.
 

### 2019-07-21 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-20T18:44:47+0200
+%% Last-Updated: 2019-07-21T17:05:52+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Our notation and terminology follow \textsc{iso} standards
-\citep{iso1993,iso2006,iso2006b} and Jaynes \citey{jaynes1994_r2003} for
-\dobs. We often simply say \enquote{belief} for \enquote{\dob}.
+Our notation and terminology follow \textsc{iso}
+\citey{iso1993,iso2006,iso2006b} standards and Jaynes
+\citey{jaynes1994_r2003} for \dobs. We often simply say \enquote{belief}
+for \enquote{\dob}.
 
@@
 during very long recordings or across experimental sessions. We assume that
-there's an area, comprising a network of $N$ neurons, for which we equally
-believe that any other sample of size $n$ could have been recorded instead
-of the sample of $n$ neurons that was actually recorded. This is our
-\enquote{full network}.\mynote{how about calling it \enquote{the pool}?}
-The normalized total activity of these $N$ neurons at $t$ is $\yA_{t}$. The
-relative frequencies of the various activity levels during the recording
-were $(\yFF_{\yA}) \defs \yF$. We don't know the values $\yA_{t}$ at each
-$t$, or the frequency distribution $\yF$. We only know for certain that
-$\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
-for obvious reasons. For the time being we assume that we know $N$; in
-\sect*** we discuss the consequences of our lack of precise knowledge about
-this number.
+there's an area, comprising a network of $N$ neurons, for which we believe
+that any other sample of size $n$ could have equally plausibly been
+recorded instead of the sample of $n$ neurons that was actually recorded.
+This is our \enquote{full network}.\mynote{how about calling it
+  \enquote{the pool}?} The normalized total activity of these $N$ neurons
+at $t$ is $\yA_{t}$. The relative frequencies of the various activity
+levels during the recording were $(\yFF_{\yA}) \defs \yF$. We don't know
+the values $\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We
+only know for certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and
+that $N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the time being we
+assume that we know $N$; in \sect*** we discuss the consequences of our
+lack of precise knowledge about this number.
 
@@
 expectations, that has minimum relative entropy with respect to a reference
-distribution, often taken to be the uniform one; and judge those expectations
-to be equal to measured averages. We add two remarks about this method that
-are seldom made in the literature. First, the distribution $\px(\yA)$ given
-by this method is the zeroth-order approximation \citep[in the sense of
-Laplace's
+distribution, often taken to be the uniform one; and judge those
+expectations to be equal to measured averages. We add two remarks about
+this method that are seldom made in the literature. First remark: the
+distribution $\px(\yA)$ given by this method is the zeroth-order
+approximation \citep[in the sense of Laplace's
 method:][\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
@@
 distribution about four completely different quantities. Note that the four
-distributions above numerically differ in higher-order approximations. We
-discuss their differences in \sect***. Second, the maximum-entropy method
-based on the Shannon entropy implicitly makes some assumptions about the
-probabilities for the long-run frequency distributions
-\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
-assumptions in \sect***.
-
-In our case, to apply the method for the total activity of the full network
-we need to fix some of the latter's averages, for example its moments. But
-we don't have any measured moments for the full network to equate the
-distribution moments to. Here enters point~\ref{item:sample_step}: the
-probability calculus gives an exact, linear relation between the first $m$
-moments for the full network and the first $m$ for the sample
+distributions above numerically differ in higher-order approximations.
+Second remark: the maximum-entropy method based on the Shannon entropy
+implicitly makes some assumptions about the probabilities for the long-run
+frequency distributions
+\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We say more about
+these two points in \sect***.
+
+In our case, to apply the maximum-entropy prescriptions to the total
+activity of the full network we need to fix some averages of its belief
+distribution, for example the distribution's moments. But we don't have any
+measured moments for the full network to equate the distribution moments
+to. Here enters point~\ref{item:sample_step}: the probability calculus
+gives an exact, linear relation between the first $m$ moments for the full
+network and the first $m$ for the sample
 \citep[\eqns~(16)]{portamanaetal2015}; the ones determine the others and
@@
 pairs is therefore $6/10$. Note that the first $m$ factorial moments
-linearly determine the first $m$ power moments and vice versa, because
-$\tbinom{n\ya}{m}$ is a polynomial in $\ya$ of degree $m$; so fixing the
-ones is equivalent to fixing the others. Now we have this extremely
-convenient property: \emph{the first $n$ normalized factorial moments for a
-  sample and for the full network are numerically identical}:
+provide the same information as the first $m$ power moments and vice versa:
+they linearly determine each other because $\tbinom{n\ya}{m}$ is a
+polynomial in $\ya$ of degree $m$. Fixing the ones is therefore equivalent
+to fixing the others. But the normalized factorial moments have this
+extremely convenient property: \emph{the first $n$ normalized factorial
+  moments for a sample and for the full network are numerically identical}:
 \begin{equation}
@@
 \end{equation}
-where $\pxx(\yA)$ is the distribution for the full-network activity.
+where $\pxx(\yA)$ is our distribution of belief about the full-network
+activity.
 
 We can therefore apply the maximum-entropy method to obtain a distribution
-for the full network, by constraining its $m$th factorial moment to be
-equal to the sample's recorded average of
-$\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$, for as many $m$ as we please with
-$1\le m \le n$. The constraints on the maximum-entropy distribution
-$\px(\yA)$ for the full network are
+for the full network by constraining its $m$th factorial moment to be equal
+to the sample's recorded average of $\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$,
+for as many $m$ as we please with $1\le m \le n$. In formulae these
+constraints on $\px(\yA)$ are
 \begin{equation}
   \label{eq:final_constraint}
-  \frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
+\underbracket{\frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
   \equiv
    \sum_{\ya}
-  \binom{n \ya}{m}\,  \binom{n}{m}^{-1} \yff_{\ya}
+  \binom{n \ya}{m}\,  \binom{n}{m}^{-1} \yff_{\ya}}_{\text{measured moments}}
   =
-   \sum_{\yA}
-  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA)
+   \underbracket{\sum_{\yA}
+  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA)}_{\text{distribution moments}}
 \end{equation}
@@
 \citep{meadetal1984}[\chap~10]{pressetal1988_r2007}{fangetal1997,boydetal2004_r2009,portamana2017b}
-and for the numbers $N,n,m$ considered in the next section it can be done
+and for the numbers $N,n,m$ considered in the present work it can be done
 on a modern computer without approximations of normalization constants or
@@
 sufficient statistics, such as the sufficiency of pairwise correlations to
-quantify our \dob\ about network activity. In the present work we
-only want to introduce the general method without entering into biological
-questions of this kind.
+quantify our \dob\ about network activity. We discuss this kind of use in
+\sect~\ref{sec:marginalization}.%, but our main goal is to simply introduce
+% the general method without entering into biological questions of this
+% kind.
 
@@
 important: it corresponds to fully constraining the marginal frequency
-distribution for the activity of the sample neurons. In this case, our
-belief about the full-network activity is based on all available
-measured frequency data. Note that application of the maximum-entropy
+distribution for the activity of the sample neurons, $\yf$. In this case,
+our belief about the full-network activity is based on all available
+measured frequency data. Note that the application of the maximum-entropy
 method \emph{at the sample level} is trivial and meaningless in this case
@@
 In the next section we apply the method just described to the data sets
-from two actual recordings, using *** moments, and discuss the properties
+from two actual recordings, using six moments, and discuss the properties
 of the resulting distributions.

### 2019-07-22 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-21T17:05:52+0200
+%% Last-Updated: 2019-07-22T11:30:14+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In the neurosciences the maximum-entropy method is often applied to
-questions of sufficient statistics. In this section we discuss how our
-proposed application bears on this kind of questions. For concreteness we
-use an example with specific statistics: first and second moments
-(equivalent to mean and correlations), versus first-to-fourth moments. But
-our discussion holds for any  sets of statistics.
-
-The question is as follows. We want to assess our degree of belief about
-the frequencies of the activities of the sample, or about the sample
-activity in a new time bin. For this assessment we should use all measured
-data we have. But it sometimes happens that dropping part of the data --
-for example, the measured third- and higher-order moments -- leaves our
-assessment almost unchanged: the dropped data are \emph{informationally
-  irrelevant}, or almost so. The remaining data -- for example, first and
-second moments -- are \emph{informationally sufficient}. This informational
-quasi-sufficiency is interesting because it may hint at peculiar biological
-or physical mechanisms or dependencies.
-
-We can approximately check how much some data are irrelevant by simply
-dropping them from the conditional of our degree of belief, and see if the
-latter changes appreciably. When our degree of belief is built with the
-maximum-entropy method, we simply drop some of the data used as constraints
-and see how the resulting distribution changes. For example, we may only
-retain the first two moments (means and correlations) or the first four
-moments. In the neurosciences this check is often done with maximum-entropy
-distributions constructed \emph{for the sample}, ignoring the full network
-it's sampled from.
+questions of sufficient statistics; for example: do pairwise correlations
+in the neuronal activity provide the same information as the full frequency
+distribution of activities? In this section we discuss how our proposed
+application bears on this kind of questions. % For concreteness we use an
+% example with specific statistics: first and second moments (equivalent to
+% mean and correlations), versus first-to-fourth moments. But our discussion
+% holds for any sets of statistics.
+
+Let's make the question more precise first. We want to assess our degree of
+belief about the frequencies of the activities of the sample, or about the
+activity of the sample in a new time bin. For this assessment we should use
+all measured data we have. But it sometimes happens that dropping part of
+the data -- for example, the measured third- and higher-order moments --
+leaves our assessment almost unchanged: the dropped data are
+\emph{informationally irrelevant}, or almost so. The remaining data -- for
+example, first and second moments -- are then \emph{informationally
+  sufficient}. This informational quasi-sufficiency is interesting because
+it may hint at peculiar biological  mechanisms or dependencies.
+
+We can approximately check how much a part of the data is irrelevant by
+simply dropping it from the conditional of our degree of belief, and see
+if the latter changes appreciably. When our degree of belief is built with
+the maximum-entropy method, we simply drop part of the data used as
+constraints and see how the resulting distribution changes. For example, we
+may only retain the first two moments (means and correlations) or the first
+four moments.
 
@@
 
-The method introduced in the present work reveals an issue with this
-application at the sample level, however. The issue doesn't come from the method
-itself but from a general fact of probability theory. If a probability
-distribution has some sufficient statistics
+In the neurosciences this check is often done with maximum-entropy
+distributions constructed \emph{for the sample}, ignoring the full network
+it's sampled from. The method introduced in the present work, however,
+reveals an issue with this application at the sample level. The issue
+doesn't come from the method itself but from a general fact of probability
+theory. If a probability distribution has some sufficient statistics
 \citep[\sect~4.5]{bernardoetal1994}{fortinietal2000}, then its marginals
@@
 mechanics: marginals of Gibbs states aren't Gibbs states
-\citep{maesetal1999}. Mathematically it corresponds to the general
-impossibility of solving a system of $n$ equations in $m$ unknowns when
-$n>m$ \citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.
+\citep{maesetal1999}. Mathematically this impossibility translates into a
+system of $n$ independent equations in $m$ unknowns with $n>m$
+\citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.
 
@@
 addressed to a whole brain area, not just to some specific, casually
-recorded sample. Thus, paradoxically, maximum-entropy methods applied at
-the sample level give an answer opposite to the one we might think they're
-giving.
+recorded sample. Thus, paradoxically, maximum-entropy methods applied
+\emph{at the sample level} give an answer opposite to the one we might
+think they're giving.
 
@@
 in an approximate informational sufficiency, not in an analytically exact
-one. A correct answer can only be given case by case;
-\fig~\ref{fig:comparisons_marginals} gives a graphical answer to this
+one. A correct answer can only be given case by case.
+Figure~\ref{fig:comparisons_marginals} gives a graphical answer to this
 question in the case of our first data set. The upper plot shows the
@@
 constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
-from four moments (\textcolor{mypurpleblue}{\textbf{------}}). The lower plot
-shows the maximum-entropy distributions for the sample, from two moments
-(\textcolor{mygreen}{$\bm{\triangledown}$}) and from four moments
-(\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained by
-marginalizing the full-network distribution to the sample size, from two
+from four moments (\textcolor{mypurpleblue}{\textbf{------}}). The lower
+plot shows the maximum-entropy distributions for the sample, from two
+moments (\textcolor{mygreen}{$\bm{\triangledown}$}) and from four moments
+(\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained
+by marginalizing the full-network distribution to the sample size, from two
 moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
-moments (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The measured frequency
-distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also shown. We
-note the following:
+moments (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The measured
+frequency distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also
+shown. We note the following:
 \begin{itemize}
@@
 
+\section{Assumptions and limitations}
+\label{sec:assumptions}
+
+In the study or use of the frequency distribution obtained with the
+procedure here presented we must take into account two important points.
+
+The first point is that there are many possible frequency distributions
+which we believe, to different degrees, could be the true one that happened
+during the recording. The one given by our procedure is simply the one with
+the largest degree of belief, the mode of the belief distribution. The
+space of possible frequency distributions has many dimensions, however --
+thousands or tens of thousands. We must remember that belief distributions
+in high dimensions have counter-intuitive properties. For example, the mode
+or mean can have \emph{atypical} features when compared with the features
+of most other points of the space. The mode and mean can also be very
+different from each other.
+
+The question, then, is \emph{which features of the maximum-entropy
+  frequency distribution are typical of the majority of plausible frequency
+  distributions?} We can only answer for sure by using the full-fledged
+probability calculus. A more complete study (in preparation) with the first
+data set reveals that most of the plausible frequency distributions have
+these important features in common with the maximum-entropy one:
+\begin{itemize}
+\item all activities $\yA \lesssim 0.05\%$ have practically zero frequencies;
+\item there are two regions of activity levels with high frequencies,
+  roughly separated by a trough of lower frequencies. The frequencies of
+  the region on the left ($\yA \lesssim 0.018\%$) are higher than those of
+  the region on the right ($\yA \gtrsim 0.018\%$).
+\end{itemize}
+But there are also differences. For example, plausible frequency
+distributions may have three or four modes instead of just two; and these
+modes are higher than of the the maximum-entropy distribution.
+
+The second point is that our degrees of belief about the frequency
+distribution for the full network depend not only on the measured data in
+the sample, but also on our pre-data beliefs about the distribution. Which
+assumptions lie behind the maximum-entropy approximation?
+
 \section{Summary and discussion}
 \label{sec:summary_discussion}
+
+We have presented a procedure to construct the most plausible frequency
+distribution of population-averaged activities of a network of neurons,
+given the recording about a small sample thereof. This procedure combines
+the maximum-entropy method and basic identities from sampling theory. From
+the application to two real data sets we saw that the frequency
+distributions obtained with our procedure can have features very different
+from the one measured in the sample, such as multiple modes. This procedure
+can also be used with moment constraints of different order -- means,
+population-averaged pairwise correlations, or higher-order correlations --
+thus giving an approximate assessment of the informational sufficiency of
+specific subsets of moments. In fact, we saw that the application of
+maximum-entropy only at the sample level leads to misleading results about
+this kind of sufficiency questions.
+
+Let's
 

### 2019-07-22 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-22T11:30:14+0200
+%% Last-Updated: 2019-07-22T12:48:21+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 each bin. Call $\ya_{t}$ the number of neurons that fire during time bin
-$t$, divided by the total number of neurons $n$; this is the
-\emph{normalized total activity} of the sample, or just \enquote{activity}
-for short. It is also the network-averaged activity of the neurons.
-Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no
-neuron spikes during bin $t$; if $\ya_{t}=1$, all spike at some point
-during bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during
-  time bin $t$}. From the activities $\set{\ya_{t}}$ we can count how often
-the activity levels $\ya=0$, $\ya=1/n$, and so on appeared during the
-recording, obtaining the distribution of measured relative frequencies
-$(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
-bins \emph{outside} of the recorded range. Such activity is unknown to us,
-of course.
+$t$: this is the \emph{total activity} of the sample, or just
+\enquote{activity} for short. Obviously $\ya_{t} \in\set{0,1,\dotsc,n}$; if
+$\ya_{t}=0$, no neuron spikes during bin $t$; if $\ya_{t}=n$, all spike at
+some point during bin $t$. For brevity, let's say \enquote{at $t$} for
+\enquote{during time bin $t$}. If we divide the total activity by the
+population size, $\ya/n \in \set{0,1/n,\dotsc,1}$ we have the normalized
+total activity or population-averaged activity. From the activities
+$\set{\ya_{t}}$ we can count how often the activity levels $\ya=0$,
+$\ya=1$, and so on appeared during the recording, obtaining the
+distribution of measured relative frequencies $(\yff_{\ya}) \defs \yf$. We
+can also consider the sample activity at time bins \emph{outside} of the
+recorded range. Such activity is unknown to us, of course.
 
@@
 This is our \enquote{full network}.\mynote{how about calling it
-  \enquote{the pool}?} The normalized total activity of these $N$ neurons
-at $t$ is $\yA_{t}$. The relative frequencies of the various activity
-levels during the recording were $(\yFF_{\yA}) \defs \yF$. We don't know
-the values $\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We
-only know for certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and
-that $N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the time being we
-assume that we know $N$; in \sect*** we discuss the consequences of our
-lack of precise knowledge about this number.
+  \enquote{the pool}?} The total activity of these $N$ neurons at $t$ is
+$\yA_{t}$. The relative frequencies of the various activity levels during
+the recording were $(\yFF_{\yA}) \defs \yF$. We don't know the values
+$\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We only know
+for certain that $\yA_{t} \in \set{0, 1, \dotsc, N}$ and that
+$\yA_{t} \ge \ya_{t}$ for obvious reasons. For the time being we assume
+that we know $N$; in \sect*** we discuss the consequences of our lack of
+precise knowledge about this number.
 
@@
   % \E\condt[\Bigg]{
-  %   \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
+  %   \binom{\ya}{m}\,\binom{n}{m}^{-1}}{\yH}
   % \equiv
   \sum_{\ya} %m!\,
-  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pxx(\ya),
+  \binom{\ya}{m}\,  \binom{n}{m}^{-1}\, \pxx(\ya),
   \qquad 1\le m \le n
@@
 provide the same information as the first $m$ power moments and vice versa:
-they linearly determine each other because $\tbinom{n\ya}{m}$ is a
+they linearly determine each other because $\tbinom{\ya}{m}$ is a
 polynomial in $\ya$ of degree $m$. Fixing the ones is therefore equivalent
@@
     \sum_{\ya} %m!\,
-  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pxx(\ya)
+  \binom{\ya}{m}\,  \binom{n}{m}^{-1}\, \pxx(\ya)
 =    \sum_{\yA} %m!\,
-  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \pxx(\yA),
-  % \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
-  % \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
+  \binom{\yA}{m}\,  \binom{N}{m}^{-1}\, \pxx(\yA),
+  % \E\condt[\big]{\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
+  % \E\condt[\big]{\tbinom{\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
   \qquad 1\le m \le n,
@@
 for the full network by constraining its $m$th factorial moment to be equal
-to the sample's recorded average of $\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$,
+to the sample's recorded average of $\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}$,
 for as many $m$ as we please with $1\le m \le n$. In formulae these
@@
   \label{eq:final_constraint}
-\underbracket{\frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
+\underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\,\binom{n}{m}^{-1}
   \equiv
    \sum_{\ya}
-  \binom{n \ya}{m}\,  \binom{n}{m}^{-1} \yff_{\ya}}_{\text{measured moments}}
+  \binom{\ya}{m}\,  \binom{n}{m}^{-1} \yff_{\ya}}_{\text{measured moments}}
   =
    \underbracket{\sum_{\yA}
-  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA)}_{\text{distribution moments}}
+  \binom{\yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA)}_{\text{distribution moments}}
 \end{equation}
@@
     \pxx(\ya)=
-  \binom{n}{n\ya}\;\binom{\yNv-n}{\yNv \yA-n\ya}\;\binom{\yNv}{\yNv \yA}^{-1}
+  \binom{n}{\ya}\;\binom{N-n}{\yA-\ya}\;\binom{N}{\yA}^{-1}
 \;\pxx(\yA).
@@
 
-\section{Assumptions and limitations}
+\section{Limitations and assumptions}
 \label{sec:assumptions}
@@
 \begin{itemize}
-\item all activities $\yA \lesssim 0.05\%$ have practically zero frequencies;
+\item all activities $\yA/N \lesssim 5\%$ have practically zero frequencies;
 \item there are two regions of activity levels with high frequencies,
   roughly separated by a trough of lower frequencies. The frequencies of
-  the region on the left ($\yA \lesssim 0.018\%$) are higher than those of
-  the region on the right ($\yA \gtrsim 0.018\%$).
+  the region on the left ($\yA/N \lesssim 1.8\%$) are higher than those of
+  the region on the right ($\yA/N \gtrsim 1.8\%$).
 \end{itemize}
@@
 
+\medskip
+
 The second point is that our degrees of belief about the frequency
 distribution for the full network depend not only on the measured data in
-the sample, but also on our pre-data beliefs about the distribution. Which
-assumptions lie behind the maximum-entropy approximation?
+the sample, but also on our pre-data beliefs $\yI$ about the distribution. Which
+assumptions lead to the maximum-entropy result? This distribution appear
+when our initial belief about the possible frequency distributions $\yF$ is
+quantified by an entropic prior:
+\begin{equation}
+  \label{eq:entropic_prior}
+  \pf(\yF \| \yI) \propto  \exp[-\yL\;\ysh(\yF;\yR)]
+  \approx \binom{L}{L\yFF_{0}, \dotsc, L\yFF_{N}}\;
+  \prod_{\yA}{\yRR_{\yA}}^{L\yFF_{\yA}}
+\end{equation}
+where $\ysh(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln(\yFF_{\yA}/\yRR_{\yA})$
+is the relative entropy or discrimination information
+\citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is a
+reference distribution, and $\yL$ a positive parameter.
 

### 2019-07-22 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-22T12:48:21+0200
+%% Last-Updated: 2019-07-22T13:06:53+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %entropies
-\newcommand*{\ysh}{H}
+\newcommand*{\re}{\textrm{H}}
 \newcommand*{\ybu}{H_{\text{B}}}
@@
 \begin{itemize}
-\item all activities $\yA/N \lesssim 5\%$ have practically zero frequencies;
+\item all activities $\yA/N \gtrsim 5\%$ have practically zero frequencies;
 \item there are two regions of activity levels with high frequencies,
@@
 \end{itemize}
-But there are also differences. For example, plausible frequency
-distributions may have three or four modes instead of just two; and these
-modes are higher than of the the maximum-entropy distribution.
+But there are also differences. For example, many plausible frequency
+distributions have three or four modes instead of just two; and these modes
+are higher than those of the maximum-entropy distribution.
 
@@
 distribution for the full network depend not only on the measured data in
-the sample, but also on our pre-data beliefs $\yI$ about the distribution. Which
-assumptions lead to the maximum-entropy result? This distribution appear
-when our initial belief about the possible frequency distributions $\yF$ is
-quantified by an entropic prior:
+the sample, but also on our pre-data beliefs $\yI$ about the distribution.
+Which assumptions lead to the maximum-entropy result? This distribution
+appear when our initial belief about the possible frequency distributions
+$\yF$ is quantified by an entropic prior
+\citep{neumann2007,rodriguez1991,skilling1998,catichaetal2004,portamana2017}:
 \begin{equation}
   \label{eq:entropic_prior}
-  \pf(\yF \| \yI) \propto  \exp[-\yL\;\ysh(\yF;\yR)]
+  \pf(\yF \| \yI) \propto  \exp[-\yL\;\re(\yF;\yR)]
   \approx \binom{L}{L\yFF_{0}, \dotsc, L\yFF_{N}}\;
@@
 \end{equation}
-where $\ysh(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln(\yFF_{\yA}/\yRR_{\yA})$
+where $\re(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln(\yFF_{\yA}/\yRR_{\yA})$
 is the relative entropy or discrimination information
-\citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is a
-reference distribution, and $\yL$ a positive parameter.
+\citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is the
+reference distribution, and $\yL$ a positive parameter. The approximate
+equality (obtained through Stirling's approximation) shows that this prior
+belief is proportional to the number of ways in which the distribution $\yF$
+can be realized in $L$ time bins. The parameter $L$ roughly quantifies how
+many time bins our data set must have to affect our initial belief.
 
@@
 this kind of sufficiency questions.
-
-Let's
 
@@
 \biggr]\;
-\exp[-L\;\ysh(\ynu;\yR)]
+\exp[-L\;\re(\ynu;\yR)]
 \end{multlined}
@@
 &\appropto \delt(\tsum_{\yA}\ypp_{\ya\yA}\yFF_{\yA} = \yff_{\ya})\;
-\exp[-L\;\ysh(\yF;\yR)]
+\exp[-L\;\re(\yF;\yR)]
 \end{split}
@@
 \biggr]\;
-\exp[-L\;\ysh(\ynu;\yR)]
+\exp[-L\;\re(\ynu;\yR)]
 \\
 &\appropto 
-\exp[-T\;\ysh(\yf; \mathte{\ypp}\ynu)
--L\;\ysh(\ynu;\yR)]
+\exp[-T\;\re(\yf; \mathte{\ypp}\ynu)
+-L\;\re(\ynu;\yR)]
 \end{split}
@@
 \biggr]\;
-\exp[-L\;\ysh(\ynu;\yR)]
+\exp[-L\;\re(\ynu;\yR)]
 \\
@@
 \int\!\di\ynu\;\ynuu_{\yA'}\;
-\exp[-T\;\ysh(\yf; \mathte{\ypp}\ynu)
--L\;\ysh(\ynu;\yR)]
+\exp[-T\;\re(\yf; \mathte{\ypp}\ynu)
+-L\;\re(\ynu;\yR)]
 \end{split}

### 2019-07-22 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-22T13:06:53+0200
+%% Last-Updated: 2019-07-22T13:24:52+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 data set reveals that most of the plausible frequency distributions have
-these important features in common with the maximum-entropy one:
+three important features in common with the maximum-entropy one:
 \begin{itemize}
@@
 \item there are two regions of activity levels with high frequencies,
-  roughly separated by a trough of lower frequencies. The frequencies of
-  the region on the left ($\yA/N \lesssim 1.8\%$) are higher than those of
-  the region on the right ($\yA/N \gtrsim 1.8\%$).
+  roughly separated by a trough of lower frequencies.
+\item The frequencies of the region on the left ($\yA/N \lesssim 1.8\%$)
+  are higher than those of the region on the right ($\yA/N \gtrsim 1.8\%$).
 \end{itemize}
 But there are also differences. For example, many plausible frequency
-distributions have three or four modes instead of just two; and these modes
-are higher than those of the maximum-entropy distribution.
+distributions have three or four modes instead of just two; these modes are
+higher than those of the maximum-entropy distribution; and the bump of high
+frequencies on the right is slightly shifted towards lower activities than
+the corresponding maximum in the maximum-entropy distribution.
 
@@
 Which assumptions lead to the maximum-entropy result? This distribution
-appear when our initial belief about the possible frequency distributions
+appears when our initial belief about the possible frequency distributions
 $\yF$ is quantified by an entropic prior
@@
 \end{equation}
-where $\re(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln(\yFF_{\yA}/\yRR_{\yA})$
-is the relative entropy or discrimination information
+where $\re(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln(\yFF_{\yA}/\yRR_{\yA})$ is
+the relative entropy or discrimination information
 \citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is the
 reference distribution, and $\yL$ a positive parameter. The approximate
-equality (obtained through Stirling's approximation) shows that this prior
-belief is proportional to the number of ways in which the distribution $\yF$
-can be realized in $L$ time bins. The parameter $L$ roughly quantifies how
-many time bins our data set must have to affect our initial belief.
+equality (obtained through Stirling's approximation), where the large
+parentheses denote a multinomial coefficient, shows that this prior belief
+is proportional to the number of ways in which the distribution $\yF$ can
+be realized in $L$ time bins. The parameter $L$ roughly quantifies how many
+time bins our data set must have to affect our initial belief. The
+maximum-entropy approximation is valid when $L$ is large, but small
+compared to the sharpness of the constraints on $\yF$; in our case this
+means $\yL \approx 10$, give or take an order of magnitude.
+
+We could obviously consider pre-data beliefs different
+from~\eqref{eq:entropic_prior}, for example one quantified by a Dirichlet
+distribution (which is equivalent to the above but with $\yF$ and $\yR$
+switched), or a uniform distribution in $\yF$-space. Would these lead to
+markedly different post-data beliefs? A full-fledged probabilistic analysis
+shows that the three typical features listed above still appear with these
+different initial beliefs. They are therefore robust.
 
@@
 
-These results caution us against taking maximum-entropy solutions too
-literally. The full probabilistic analysis allows us to see more clearly
-the extent of our uncertainty and to make better-informed guesses.
-
-\clearpage
+%%\setlength{\intextsep}{0.5ex}% with wrapfigure
+%\begin{figure}[p!]%{r}{0.4\linewidth} % with wrapfigure
+%  \centering\includegraphics[trim={12ex 0 18ex 0},clip,width=\linewidth]{maxent_saddle.png}\\
+%\caption{caption}\label{fig:comparison_a5}
+%\end{figure}% exp_family_maxent.nb
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Acknowledgements
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\begin{acknowledgements}
+  This work is financially supported by the Kavli Foundation and the Centre
+  of Excellence scheme of the  Research Council of Norway (Yasser Roudi group).\\
+  PGLPM thanks Mari, Miri, \amp\ Emma for continuous encouragement and
+  affection; Buster Keaton and Saitama for filling life with awe and
+  inspiration; the developers and maintainers of \LaTeX, Emacs, AUC\TeX,
+  Open Science Framework, Python, Inkscape, Sci-Hub for making a free and
+  unfiltered scientific exchange possible.
+%\rotatebox{15}{P}\rotatebox{5}{I}\rotatebox{-10}{P}\rotatebox{10}{\reflectbox{P}}\rotatebox{-5}{O}.
+%\sourceatright{\autanet}
+\end{acknowledgements}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Appendices
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+%\clearpage
+% %\renewcommand*{\appendixpagename}{Appendix}
+% %\renewcommand*{\appendixname}{Appendix}
+% %\appendixpage
+% \appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Bibliography
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\defbibnote{prenote}{{\footnotesize (\enquote{de $X$} is listed under D,
+    \enquote{van $X$} under V, and so on, regardless of national
+    conventions.)\par}}
+% \defbibnote{postnote}{\par\medskip\noindent{\footnotesize% Note:
+%     \arxivp \mparcp \philscip \biorxivp}}
+
+\printbibliography[prenote=prenote%,postnote=postnote
+]
+
+\end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Cut text (won't be compiled)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+When political elections are approaching, mass media are constantly
+bubbling with results and discussions of pre-election polls. Why do we
+conduct pre-election polls and other kinds of survey sampling? The
+simplest, intuitive answer is \enquote{because they give us an idea of what
+  the final election outcomes will be}. Can this answer be made more
+quantitative?
+
+Yes. In the simplest cases, the probability calculus give us simple and
+precise quantitative relations between our expectations about a network
+-- of voters, for instance -- and about a sample thereof. Suppose there are
+$N$ voters, $S$ of which prefer one political candidate and $N-S$ of which
+prefer another. We don't know $S$ or $N-S$. Now take a sample of $n$
+voters, in such a way that we can assume that every set of $n$ voters is
+equally likely to be sampled. In this sample, $s$ voters prefer the first
+candidate, and $n-s$ the second. The probability calculus tells us that our
+expectations of $s/n$ and $S/N$ are equal:
+\begin{equation}\label{eq:poll_equal_expectations}
+  \E(s/n \| \yH) = \E(S/N \| \yH),
+\end{equation}
+where $\yH$ represents our state of knowledge in this example. A more
+precise answer to our original question then is: \enquote{because we
+  \emph{expect} -- in the full probabilistic sense -- to see the same
+  proportion of political preferences in a sample as in the full
+  network}.
+
+The exact equality above extends to other quantities beyond the means $S/N$
+and $s/n$: it holds for all normalized \emph{factorial} moments
+\citep{potts1953,broca2005}[\sect~A]{portamanaetal2015}. These expectation
+equalities follow from a precise relation between our \dobs\ about the
+polled voters and about all voters. The relations above are powerful
+because they hold no matter what the probability distribution $\pf(S\|\yH)$
+may be.
+
+\bigskip
+
+These probabilistic properties are of outstanding importance in
+neuroscience, where we try to understand the workings of brain areas that
+may include tens of thousands of neurons, but we can record only few tens
+or hundreds of them. Whenever we record the activity of $n$ neurons we're
+effectively polling them on the activity of all $N$ neurons under the
+recording device -- those that could equally likely have been sampled
+instead. Suppose the recorded activities are time-binned and binarized. At
+every time bin $t$, $s(t)$ out of the $n$ neurons \enquote{vote} for
+activity, and $n-s(t)$ vote for silence. The probability calculus tells us
+that expectation equalities such as~\eqref{eq:poll_equal_expectations} hold
+for the total activity $S(t)$ of the full network of $N$ neurons that
+could have been recorded.
+
+Expectation properties are especially important for the maximum-entropy
+method, which is used for various purposes in the neuroscientific
+literature. This method is typically applied constraining the equality of
+expectations and empirical averages. Expectation equalities such
+as~\eqref{eq:poll_equal_expectations} then have a far-reaching consequence:
+\emph{from empirical statistics of the activity of a \textbf{sample} of
+  neurons we can construct maximum-entropy distributions for the total
+  activity of the \textbf{full network} of neurons that could have been
+  sampled by the recording device}.
+
+% In the present work we show how to use the maximum-entropy method to assign
+% a probability distribution for the total activity at a new time bin of a
+% network of neurons, \emph{most of which can be unrecorded}, given
+% specific statistics about the activity of a recorded \emph{sample} of the
+% full network. This kind of probability assignment is extremely important
+% in neuroscience: we try to understand the workings of brain areas that may
+% include tens of thousands of neurons, but we can record only few tens or
+% hundreds of them.
+
+% Our approach also corrects a slight misuse of the maximum-entropy method
+% as sometimes appears in the literature. When this method is applied to a
+% set of neurons, it implicitly assumes that the set is completely isolated
+% from other neurons. We'll reveal this implicit assumption in several ways.
+% The probability distribution thus constructed is therefore affected by this
+% hidden assumption. We'll study in which circumstances the application of
+% the method at the sample level and at the full-network level give
+% similar or different probabilities.
+
 \hrule
@@
 
-%%\setlength{\intextsep}{0.5ex}% with wrapfigure
-%\begin{figure}[p!]%{r}{0.4\linewidth} % with wrapfigure
-%  \centering\includegraphics[trim={12ex 0 18ex 0},clip,width=\linewidth]{maxent_saddle.png}\\
-%\caption{caption}\label{fig:comparison_a5}
-%\end{figure}% exp_family_maxent.nb
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% Acknowledgements
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\iffalse
-\begin{acknowledgements}
-  PGLPM thanks Mari, Miri, \amp\ Emma for continuous encouragement and
-  affection; Buster Keaton and Saitama for filling life with awe and
-  inspiration; the developers and maintainers of \LaTeX, Emacs, AUC\TeX,
-  Open Science Framework, Python, Inkscape, Sci-Hub for making a free and
-  unfiltered scientific exchange possible.
-%\rotatebox{15}{P}\rotatebox{5}{I}\rotatebox{-10}{P}\rotatebox{10}{\reflectbox{P}}\rotatebox{-5}{O}.
-\sourceatright{\autanet}
-\end{acknowledgements}
-\fi
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% Appendices
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\clearpage
-% %\renewcommand*{\appendixpagename}{Appendix}
-% %\renewcommand*{\appendixname}{Appendix}
-% %\appendixpage
-% \appendix
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% Bibliography
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\defbibnote{prenote}{{\footnotesize (\enquote{de $X$} is listed under D,
-    \enquote{van $X$} under V, and so on, regardless of national
-    conventions.)\par}}
-% \defbibnote{postnote}{\par\medskip\noindent{\footnotesize% Note:
-%     \arxivp \mparcp \philscip \biorxivp}}
-
-\printbibliography[prenote=prenote%,postnote=postnote
-]
-
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% Cut text (won't be compiled)
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-When political elections are approaching, mass media are constantly
-bubbling with results and discussions of pre-election polls. Why do we
-conduct pre-election polls and other kinds of survey sampling? The
-simplest, intuitive answer is \enquote{because they give us an idea of what
-  the final election outcomes will be}. Can this answer be made more
-quantitative?
-
-Yes. In the simplest cases, the probability calculus give us simple and
-precise quantitative relations between our expectations about a network
--- of voters, for instance -- and about a sample thereof. Suppose there are
-$N$ voters, $S$ of which prefer one political candidate and $N-S$ of which
-prefer another. We don't know $S$ or $N-S$. Now take a sample of $n$
-voters, in such a way that we can assume that every set of $n$ voters is
-equally likely to be sampled. In this sample, $s$ voters prefer the first
-candidate, and $n-s$ the second. The probability calculus tells us that our
-expectations of $s/n$ and $S/N$ are equal:
-\begin{equation}\label{eq:poll_equal_expectations}
-  \E(s/n \| \yH) = \E(S/N \| \yH),
-\end{equation}
-where $\yH$ represents our state of knowledge in this example. A more
-precise answer to our original question then is: \enquote{because we
-  \emph{expect} -- in the full probabilistic sense -- to see the same
-  proportion of political preferences in a sample as in the full
-  network}.
-
-The exact equality above extends to other quantities beyond the means $S/N$
-and $s/n$: it holds for all normalized \emph{factorial} moments
-\citep{potts1953,broca2005}[\sect~A]{portamanaetal2015}. These expectation
-equalities follow from a precise relation between our \dobs\ about the
-polled voters and about all voters. The relations above are powerful
-because they hold no matter what the probability distribution $\pf(S\|\yH)$
-may be.
-
-\bigskip
-
-These probabilistic properties are of outstanding importance in
-neuroscience, where we try to understand the workings of brain areas that
-may include tens of thousands of neurons, but we can record only few tens
-or hundreds of them. Whenever we record the activity of $n$ neurons we're
-effectively polling them on the activity of all $N$ neurons under the
-recording device -- those that could equally likely have been sampled
-instead. Suppose the recorded activities are time-binned and binarized. At
-every time bin $t$, $s(t)$ out of the $n$ neurons \enquote{vote} for
-activity, and $n-s(t)$ vote for silence. The probability calculus tells us
-that expectation equalities such as~\eqref{eq:poll_equal_expectations} hold
-for the total activity $S(t)$ of the full network of $N$ neurons that
-could have been recorded.
-
-Expectation properties are especially important for the maximum-entropy
-method, which is used for various purposes in the neuroscientific
-literature. This method is typically applied constraining the equality of
-expectations and empirical averages. Expectation equalities such
-as~\eqref{eq:poll_equal_expectations} then have a far-reaching consequence:
-\emph{from empirical statistics of the activity of a \textbf{sample} of
-  neurons we can construct maximum-entropy distributions for the total
-  activity of the \textbf{full network} of neurons that could have been
-  sampled by the recording device}.
-
-% In the present work we show how to use the maximum-entropy method to assign
-% a probability distribution for the total activity at a new time bin of a
-% network of neurons, \emph{most of which can be unrecorded}, given
-% specific statistics about the activity of a recorded \emph{sample} of the
-% full network. This kind of probability assignment is extremely important
-% in neuroscience: we try to understand the workings of brain areas that may
-% include tens of thousands of neurons, but we can record only few tens or
-% hundreds of them.
-
-% Our approach also corrects a slight misuse of the maximum-entropy method
-% as sometimes appears in the literature. When this method is applied to a
-% set of neurons, it implicitly assumes that the set is completely isolated
-% from other neurons. We'll reveal this implicit assumption in several ways.
-% The probability distribution thus constructed is therefore affected by this
-% hidden assumption. We'll study in which circumstances the application of
-% the method at the sample level and at the full-network level give
-% similar or different probabilities.
-
 %%% Local Variables: 

### 2019-07-22 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-22T13:24:52+0200
+%% Last-Updated: 2019-07-22T13:43:19+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\mynote{To be continued after structure of the rest of the article is
-  clear. Orig intro is on p.~\pageref{orig_intro}}
+% \mynote{To be continued after structure of the rest of the article is
+%   clear. Orig intro is on p.~\pageref{orig_intro}}
 
@@
 \begin{itemize}[wide]
-\item The first, from \textcite[rat 14147]{stensolaetal2012},
-  consists of $n=65$ neurons (27 of which classified as grid cells) from
-  rat Medial Entorhinal Cortex, recorded for about 20 minutes. Their spikes
-  are binned into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
-\item The second, from
+\item The \textbf{first data set}, from \textcite[rat
+  14147]{stensolaetal2012}, consists of $n=65$ neurons (27 of which
+  classified as grid cells) from rat Medial Entorhinal Cortex, recorded for
+  about 20 minutes. Their spikes are binned into $T=417\,641$ bins of
+  $3\;\textrm{ms}$ width.
+\item The \textbf{second data set}, from
   \textcite{rostamietal2016_r2017}, %\texttt{data\_RS.npy}

### 2019-07-25 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-22T13:43:19+0200
+%% Last-Updated: 2019-07-25T11:36:58+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Answering this kind of questions always engages an element of uncertainty.
-Our answers therefore involve experimental data, such as neuronal
-recordings from specific brain areas, and probabilities or \dobs, based on
-prior knowledge, about biological conditions and mechanisms that cannot be
-experimentally ascertained. Such \dobs\ are often formulated as simplified
-\enquote{models} to be mathematically more tractable.
+We can't say \enquote{the answer is such and such}, but can at best assign
+a degree of reasonable belief -- that is, probability -- to every possible
+answer. The assessment of this distribution of belief involves experimental
+data, such as recordings of neuronal activity from specific brain areas,
+and pre-data knowledge about biological conditions and mechanisms. Our
+pre-data \dobs\ are often simplified and called \enquote{models} to be
+mathematically more tractable.
 
@@
 mathematics. Some probabilistic models try to take also unrecorded neurons
-into account, but become very complex in doing so.
-
-In the present work we give an answer to this question: How much can the
-total activity of a large neuronal network be, if we have observed the
-activity of a very small sample thereof? We'll quantify our \dobs\ about
-the possible answers by combining, in a straightforward way, the
-maximum-entropy method and basic sampling relations of the probability
-calculus. Later on we'll show that our \dobs\ can be quantified by
-exclusively using the probability calculus. This derivation will provide a
-more accurate quantification, revealing that the maximum-entropy answer is
-only a first approximation.
-
-We apply our approach to two concrete data sets: the activity of 65 neurons
-recorded from a rat's Medial Entorhinal Cortex \mynote{ref}, and the
-activity of 159 neurons recorded from a macaque's Motor Cortex
-\mynote{ref}\mynote{add recording length}. In the first data set, the most
+into account individually, and therefore become very complex\mynote{refs}.
+It would be useful to have models that operate in between: addressing the
+larger brain area where a sample comes from, but without adding too much
+complexity and detail about it. Such intermediate models would be useful,
+for example, for preliminary investigations, to help us to decide which
+hypotheses to discard and which to consider for more complex and costly
+studies.
+
+In the present work we propose such an intermediate model. It answers this
+question: How much can the \emph{total} activity of a large neuronal
+network be, if we have observed the activity of a very small sample
+thereof? This model addresses a larger brain area, avoiding the assumption
+of isolation of the sample. And by focusing on the total activity, rather
+than the activity of individual neurons, it remains simple and numerically
+tractable.
+
+The proposed model is based on the straightforward combination of the
+maximum-entropy method and basic sampling relations from the probability
+calculus. The maximum-entropy or minimum-relative-entropy method
+\citep{jaynes1957}[much clearer
+in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b,grandy1980}
+has been used for different kinds of estimations of the neuronal activity
+of various brain areas and about other phenomena of importance to the
+neurosciences, for example gene and protein interaction \citep[for
+example][]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,mackeetal2009b,tkaciketal2009,roudietal2009c,barreiroetal2010,gerwinnetal2010,mackeetal2011b,ganmoretal2011,cohenetal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}.
+This method has often been used to test whether a smaller part of the data,
+for example time averages and second-order correlations, is sufficient for
+quantifying our \dob\ about some aspects of the neuronal activity. We'll
+consider this particular application also for our proposed model.
+Maximum-entropy models can be considered as approximations of Bayesian
+models based on various assumptions of inferential sufficiency
+\citep{jaynes1986d_r1996,portamana2017}.
+% Later on we'll show that our \dobs\ can be quantified by
+% exclusively using the probability calculus. This derivation will provide a
+% more accurate quantification, revealing that the maximum-entropy answer is
+% only a first approximation.
+
+We apply our proposed model to two concrete data sets:
+\begin{enumerate*}[label=(\alph*)]
+\item the activity of 65 neurons recorded from a rat's Medial Entorhinal
+  Cortex \citep{stensolaetal2012}, \item the activity of 159 neurons
+  recorded from a macaque's Motor Cortex \citep{rostamietal2016_r2017}.
+\end{enumerate*}
+For each data set the model yields***
+
+In the first data set, the most
 interesting finding is that the most probable frequency distribution for
@@
   two-sentence summary of the main idea here?}
-
-The maximum-entropy or minimum-relative-entropy method
-\citep{jaynes1957}[much clearer
-in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b,grandy1980}
-has been used for different kinds of estimations of the neuronal activity
-of various brain areas and about other phenomena of importance to the
-neurosciences, for example gene and protein interaction \citep[for
-example][]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,mackeetal2009b,tkaciketal2009,roudietal2009c,barreiroetal2010,gerwinnetal2010,mackeetal2011b,ganmoretal2011,cohenetal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}.
-This method is often used to test whether some statistics of the data, for
-example second-order time correlations, is sufficient for quantifying our
-\dob\ about some quantities of the system. It can be considered as an
-approximation of probabilistic models based on various assumptions of
-inferential sufficiency \citep{jaynes1986d_r1996,portamana2017}.
 

### 2019-07-25 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-25T11:36:58+0200
+%% Last-Updated: 2019-07-25T13:53:17+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 It would be useful to have models that operate in between: addressing the
-larger brain area where a sample comes from, but without adding too much
+larger brain area whence the sample comes, but without adding too much
 complexity and detail about it. Such intermediate models would be useful,
@@
 question: How much can the \emph{total} activity of a large neuronal
-network be, if we have observed the activity of a very small sample
-thereof? This model addresses a larger brain area, avoiding the assumption
-of isolation of the sample. And by focusing on the total activity, rather
-than the activity of individual neurons, it remains simple and numerically
-tractable.
+network have been, given the observation of the activity of a very small
+sample thereof? This model addresses a larger brain area, avoiding the
+assumption of isolation of the sample; and by focusing on the total
+activity, rather than the activity of individual neurons, it remains simple
+and numerically tractable.
 
@@
 
-We apply our proposed model to two concrete data sets:
+We illustrate possible uses of our proposed model by applying it to two
+concrete data sets:
 \begin{enumerate*}[label=(\alph*)]
-\item the activity of 65 neurons recorded from a rat's Medial Entorhinal
-  Cortex \citep{stensolaetal2012}, \item the activity of 159 neurons
-  recorded from a macaque's Motor Cortex \citep{rostamietal2016_r2017}.
+\item the activity of 65 neurons recorded for 20 min from a rat's Medial
+  Entorhinal Cortex \citep{stensolaetal2012}, \item the activity of 159
+  neurons recorded for 15 min from a macaque's Motor Cortex
+  \citep{rostamietal2016_r2017}.
 \end{enumerate*}
-For each data set the model yields***
-
-In the first data set, the most
-interesting finding is that the most probable frequency distribution for
-the total activity of the full network has two very distinct modes, both at
-low activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
+For each data set the model gives the most plausible frequency distribution
+with which all levels of total activity of a larger network of thousands of
+neurons appeared during the recording. These guessed frequency
+distributions are distinctly different from the recorded ones for the
+sampled neurons. For the first data set, for example, the frequency
+distribution for the full network has two very distinct modes, both at low
+activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
 % within $3\;\mathrm{ms}$,
-see \fig~\ref{fig:plots_6mom}. The analogous frequency distribution for the
-second data set doesn't have two modes but still presents one prominent
-shoulder in its one low-activity mode. Note that these guessed features of
-the full network aren't observed in the sample, nor can they be inferred by
-the application of maximum-entropy \emph{to the sample alone}.
-\mynote{Monte Carlo sampling: say something about the deviation from the
-  most probable frequency distribution}
-
-\mynote{shall we give a
-  two-sentence summary of the main idea here?}
+see \fig~\ref{fig:plots_6mom}, whereas the frequency distribution for the
+sample is monotonically decreasing with its maximum at zero activity. The
+frequency distribution for the second data set doesn't have two modes but
+still presents one prominent shoulder in its low-activity mode. Note that
+these guessed features of the full network cannot be inferred by the
+application of maximum-entropy \emph{to the sample alone}.
 

### 2019-07-25 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-25T13:53:17+0200
+%% Last-Updated: 2019-07-25T14:43:18+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 example][]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,mackeetal2009b,tkaciketal2009,roudietal2009c,barreiroetal2010,gerwinnetal2010,mackeetal2011b,ganmoretal2011,cohenetal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}.
-This method has often been used to test whether a smaller part of the data,
-for example time averages and second-order correlations, is sufficient for
-quantifying our \dob\ about some aspects of the neuronal activity. We'll
-consider this particular application also for our proposed model.
-Maximum-entropy models can be considered as approximations of Bayesian
-models based on various assumptions of inferential sufficiency
-\citep{jaynes1986d_r1996,portamana2017}.
-% Later on we'll show that our \dobs\ can be quantified by
-% exclusively using the probability calculus. This derivation will provide a
-% more accurate quantification, revealing that the maximum-entropy answer is
-% only a first approximation.
 
@@
 For each data set the model gives the most plausible frequency distribution
-with which all levels of total activity of a larger network of thousands of
+with which all levels of total activity of a larger network of 1\,000--10\,000
 neurons appeared during the recording. These guessed frequency
@@
 these guessed features of the full network cannot be inferred by the
-application of maximum-entropy \emph{to the sample alone}.
+application of maximum-entropy \emph{to the sample alone}. Although we don't
+formulate any hypotheses on why the larger network could have two distinct
+low-activity regimes -- the purpose of the present work is only
+methodological -- these results shows that the proposed method can lead to
+the formulation or preliminary assessment of interesting hypotheses.
+
+Maximum-entropy models have often been used to test whether a smaller part
+of the sample data, for example time averages and second-order
+correlations, is sufficient for quantifying our \dob\ about some aspects of
+the neuronal activity of the sample. This use of maximum-entropy at the
+sample level, however, has an important issue. We discuss this issue and
+explain how our approach bypasses it. Moreover, our proposed model gives
+clearer answers to questions of correlation sufficiency: we show this by
+applying it to the two data set above addressing an example question of
+sufficiency: two moments versus four moments.
+
+Maximum-entropy models can be
+considered as approximations of Bayesian models based on various
+assumptions of inferential sufficiency
+\citep{jaynes1986d_r1996,portamana2017}.
+% Later on we'll show that our \dobs\ can be quantified by
+% exclusively using the probability calculus. This derivation will provide a
+% more accurate quantification, revealing that the maximum-entropy answer is
+% only a first approximation.
+
+We also apply the model to the question of informational sufficiency of
+lower-order moments
 

### 2019-07-25 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-25T14:43:18+0200
+%% Last-Updated: 2019-07-25T15:46:11+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %entropies
-\newcommand*{\re}{\textrm{H}}
-\newcommand*{\ybu}{H_{\text{B}}}
+\newcommand*{\re}{\Eta}
+\newcommand*{\ybu}{\re_{\text{B}}}
 
@@
 and pre-data knowledge about biological conditions and mechanisms. Our
-pre-data \dobs\ are often simplified and called \enquote{models} to be
-mathematically more tractable.
+pre-data \dobs\ are often simplified to be mathematically more tractable,
+and are therefore called \enquote{models}.
 
@@
 maximum-entropy method and basic sampling relations from the probability
-calculus. The maximum-entropy or minimum-relative-entropy method
-\citep{jaynes1957}[much clearer
+calculus, discussed in \sect~\ref{sec:model}. The maximum-entropy or
+minimum-relative-entropy method \citep{jaynes1957}[much clearer
 in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b,grandy1980}
@@
 
-We illustrate possible uses of our proposed model by applying it to two
-concrete data sets:
+We illustrate possible uses of our proposed model in
+\sect~\ref{sec:application}, by applying it to two concrete data sets:
 \begin{enumerate*}[label=(\alph*)]
@@
 sample level, however, has an important issue. We discuss this issue and
-explain how our approach bypasses it. Moreover, our proposed model gives
-clearer answers to questions of correlation sufficiency: we show this by
-applying it to the two data set above addressing an example question of
-sufficiency: two moments versus four moments.
-
-Maximum-entropy models can be
-considered as approximations of Bayesian models based on various
-assumptions of inferential sufficiency
-\citep{jaynes1986d_r1996,portamana2017}.
+explain how our approach bypasses it in \sect~\ref{sec:marginalization}.
+Moreover, our proposed model gives clearer answers to questions of
+correlation sufficiency: we show this by applying it to the two data set
+above addressing an example question of sufficiency, two moments versus
+four moments.
+
+We obviously don't know whether the frequency distribution obtained with
+our approach is the \emph{actual} one which the activity levels of the full
+network had during the recording; it's only the most plausible. In high
+dimensions, however, the features of the most plausible distribution may
+\emph{not} be typical of the majority of most plausible distributions; and
+the set of all possible frequency distributions, if the full network for
+example comprises 1\,000 neurons, is a 1\,000-dimensional space. In
+\sect~\ref{sec:assumptions} we therefore try to assess which features of
+the frequency distribution delivered by our method may be typical and
+therefore expected of the actual one. We find that general features such as
+the bimodality of the first data set are indeed typical. Maximum-entropy
+models can be considered as approximations of Bayesian models based on
+various assumptions of inferential sufficiency
+\citep{jaynes1986d_r1996,portamana2017}. How do our guesses change if we
+modify our pre-data assumptions? We show, in the same section, that the
+typical features indicated by our method are robust against such changes.
+
+\medskip
+
+% (for a full network of 1\,000 neurons and
+% a recording of 500\,000 time bins, for example, there are
+% $\binom{500\,000+1\,000}{1\,000} \approx 10^{3\,131}$ possible frequency
+% distributions).
+
 % Later on we'll show that our \dobs\ can be quantified by
@@
 % only a first approximation.
-
-We also apply the model to the question of informational sufficiency of
-lower-order moments
 

### 2019-07-25 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-25T15:46:11+0200
+%% Last-Updated: 2019-07-25T16:20:19+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Introduction}
+\section{Introduction:\\a simple model for questions about large neuronal
+  networks}
 \label{sec:intro}
@@
 
+How large is the full network addressed by the method proposed here? Its
+size must have some limit and can't obviously include the full brain. The
+size is determined by the validity of the formulae from sampling theory and
+discussed in \sect~\ref{sec:N}.
+
 We obviously don't know whether the frequency distribution obtained with
@@
 
+A summary of all points above is given in the final
+\sect~\ref{sec:summary_discussion}.
+
 \medskip
+
+Our notation and terminology follow \textsc{iso}
+\citey{iso1993,iso2006,iso2006b} standards and Jaynes
+\citey{jaynes1994_r2003} for \dobs. We often simply say \enquote{belief}
+for \enquote{\dob}.
 
@@
 
-Our notation and terminology follow \textsc{iso}
-\citey{iso1993,iso2006,iso2006b} standards and Jaynes
-\citey{jaynes1994_r2003} for \dobs. We often simply say \enquote{belief}
-for \enquote{\dob}.
-
-\section{Model: maximum-entropy and sampling}
+\section{The approach: maximum-entropy and sampling}
 \label{sec:model}
@@
 
-\section{Application: two data sets}
+\section{Example application: two data sets}
 \label{sec:application}
@@
 
-\section{Marginalization to the sample level:\\ the issue with pairwise
-  correlations and other statistics}
+\section{Example application for sufficiency of partial statistics:\\the
+  issue with methods at the sample level}
 \label{sec:marginalization}

### 2019-07-25 maxNt_v5.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-25T16:20:19+0200
+%% Last-Updated: 2019-07-25T16:30:40+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}\labelsep 0pt%
-  \noindent This work shows how to build a maximum-entropy probabilistic
-  model for the total activity of a network of neurons, given only some
-  activity data or statistics -- for example, empirical moments -- of a
-  \emph{subnetwork} thereof. This kind of model is useful because
-  neuronal recordings are always limited to a very small sample of a
-  network of neurons.
+  \noindent \mynote{Abstract must be rewritten once paper is ready}This
+  work shows how to build a maximum-entropy probabilistic model for the
+  total activity of a network of neurons, given only some activity data or
+  statistics -- for example, empirical moments -- of a \emph{subnetwork}
+  thereof. This kind of model is useful because neuronal recordings are
+  always limited to a very small sample of a network of neurons.
 %

### 2019-11-13 maxNt_v6.tex
update

--- 2019-07-25 maxNt_v5.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-07-25T16:30:40+0200
+%% Last-Updated: 2019-11-13T16:55:17+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\propertitle}{Maximum-entropy distributions for a 
-  neuronal network\\ from subnetwork data%\\{\large ***}%
+  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
 }
@@
 Despite remarkable advances in recording technologies, the best
-experimental measurements of neuronal activity can still only record a very
-small sample of neurons compared to the numbers that constitute a
-functionally distinguished brain region. Many probabilistic models focus on
-such samples only, somehow neglecting, in their assumptions, that the
-recorded neurons are a sample from a larger network. This kind of isolation
-assumptions sometimes escape attention, being subtly hidden in the
-mathematics. Some probabilistic models try to take also unrecorded neurons
-into account individually, and therefore become very complex\mynote{refs}.
-It would be useful to have models that operate in between: addressing the
-larger brain area whence the sample comes, but without adding too much
-complexity and detail about it. Such intermediate models would be useful,
-for example, for preliminary investigations, to help us to decide which
-hypotheses to discard and which to consider for more complex and costly
-studies.
-
+experimental measurements of instantaneous neuronal activity can still only
+record a very small sample of neurons -- hundreds at most -- compared to
+the numbers that constitute a functionally distinguished brain region. Many
+probabilistic models focus on such samples only, somehow neglecting, in
+their assumptions, that the recorded neurons are a sample from a larger
+network. This kind of isolation assumptions sometimes escape attention,
+being subtly hidden in the mathematics. Some probabilistic models try to
+take also unrecorded neurons into account individually, and therefore
+become very complex\mynote{Yasser \amp\ Vahid: refs for here?}. It would be
+useful to have models that operate in between: addressing the larger brain
+area whence the sample comes, but without adding too much complexity and
+detail about it. Such intermediate models would be useful, for example, for
+preliminary investigations, to help us to decide which hypotheses to
+discard and which to consider for more complex and costly studies.
+
+\mynote{[Yasser:] In the present work we aim to address the issue of building probability distributions over a large network using recording from subnetworks by considering the distribution of total population activity as an example.
+[}
 In the present work we propose such an intermediate model. It answers this
-question: How much can the \emph{total} activity of a large neuronal
+question:\mynote{]} How much can the \emph{total} activity of a large neuronal
 network have been, given the observation of the activity of a very small
@@
 
-The proposed model is based on the straightforward combination of the
+We want to stress the usefulness of making a quantified guess about the
+activity of a larger brain area. Such a guess seems indeed to be the
+primary idea behind recording a sample from that area. %  Second, it allows us
+% to make comparisons across experimental sessions; such comparisons would me
+% difficult or meaningless if made with the recorded samples, which generally
+% comprise non-overlapping sets of neurons and differ in size.
+There's also an important methodological reason: the question whether
+correlations of some order are informationally sufficient cannot be
+properly assessed at the sample level. When the difference in size between
+sample and full network is too large, correlation sufficiency for the
+sample implies the \emph{lack} of correlation sufficiency for the larger
+network, and vice versa. This point is explained in detail in
+\sect~\ref{sec:marginalization}.
+
+The model we propose is based on the straightforward combination of the
 maximum-entropy method and basic sampling relations from the probability
@@
 
-\mynote{Say something more about advantages of such question/answer: for
-  example we can make statements about total activity of brain area even
-  across recordings when sampled neurons aren't the same.}
-
-\mynote{Add this?: from sampling theory we know that important
-  features of the full network may not be visible in a sample because
-  smoothed out. But using sampling theory in the inverse direction we can
-  infer such full-network features from the sample.}
+% \mynote{Say something more about advantages of such question/answer: for
+%   example we can make statements about total activity of brain area even
+%   across recordings when sampled neurons aren't the same.}
+
+% \mynote{Add this?: from sampling theory we know that important
+%   features of the full network may not be visible in a sample because
+%   smoothed out. But using sampling theory in the inverse direction we can
+%   infer such full-network features from the sample.}
 
@@
 \enquote{during time bin $t$}. If we divide the total activity by the
-population size, $\ya/n \in \set{0,1/n,\dotsc,1}$ we have the normalized
-total activity or population-averaged activity. From the activities
-$\set{\ya_{t}}$ we can count how often the activity levels $\ya=0$,
-$\ya=1$, and so on appeared during the recording, obtaining the
-distribution of measured relative frequencies $(\yff_{\ya}) \defs \yf$. We
-can also consider the sample activity at time bins \emph{outside} of the
-recorded range. Such activity is unknown to us, of course.
-
-\mynote{maybe move this paragraph to the intro?}Present-day technologies
-enable the recording of neuronal activity from small brain areas \mynote{be
-  more specific}. For many animal species, the neurons that are recorded
-within the area are not -- and at present cannot be -- specifically chosen
-from among the rest, owing to several limiting factors; for example,
-limitations in how precisely electrodes are inserted or neurons are
-targeted by viruses. In fact, the set of recorded neurons may even change
-during very long recordings or across experimental sessions. We assume that
-there's an area, comprising a network of $N$ neurons, for which we believe
-that any other sample of size $n$ could have equally plausibly been
-recorded instead of the sample of $n$ neurons that was actually recorded.
-This is our \enquote{full network}.\mynote{how about calling it
-  \enquote{the pool}?} The total activity of these $N$ neurons at $t$ is
-$\yA_{t}$. The relative frequencies of the various activity levels during
-the recording were $(\yFF_{\yA}) \defs \yF$. We don't know the values
-$\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We only know
-for certain that $\yA_{t} \in \set{0, 1, \dotsc, N}$ and that
-$\yA_{t} \ge \ya_{t}$ for obvious reasons. For the time being we assume
-that we know $N$; in \sect*** we discuss the consequences of our lack of
-precise knowledge about this number.
+population size we have the normalized total activity or
+population-averaged activity $\ya/n$, ranging from $0$ to $1$ in $1/n$
+steps. From the activities $\set{\ya_{t}}$ we can count how often the
+activity levels $\ya=0$, $\ya=1$, and so on appeared during the recording,
+obtaining the distribution of measured relative frequencies
+$(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
+bins \emph{outside} of the recorded period. Such activity is unknown to us,
+of course.
+
+For many animal species, the neurons that are recorded within a brain area
+are not specifically chosen from among the rest, owing to several limiting
+factors; for example, limitations in how precisely electrodes are inserted
+or neurons are targeted by viruses. The set of recorded neurons may even
+change slightly across experimental sessions that are very far apart in
+time. We assume that there's an area, comprising a network of $N$ neurons,
+for which we believe that any other sample of size $n$ could have equally
+plausibly been recorded instead of the sample of $n$ neurons that was
+actually recorded. We call this larger network the \enquote{full network}.
+The total activity of these $N$ neurons at $t$ is $\yA_{t}$. The relative
+frequencies of the various activity levels during the recording were
+$(\yFF_{\yA}) \defs \yF$. We don't know the values $\yA_{t}$ at each $t$,
+or the frequency distribution $\yF$. We only know for certain that
+$\yA_{t} \in \set{0, 1, \dotsc, N}$ and that $\yA_{t} \ge \ya_{t}$ for
+obvious reasons. For the time being we assume that we know $N$; in
+\sect~\ref{sec:N} we discuss the consequences of our lack of precise
+knowledge about this number.
 
@@
 across sessions? How much total activity should we expect at any time bin
-during a recording? We cannot answer these questions with certainty; we can
-only give distributions of probability or \dobs\ over their possible answers. The
-approach presented here gives such probability distributions.
-
-\mynote{Move this to intro?}We want to stress the usefulness of making a
-quantified guess about the full-network activity. First of all, this guess
-about a brain area seems to be the primary idea behind recording a sample
-from that area. Second, it allows us to make comparisons across
-experimental sessions; such comparisons would me difficult or meaningless
-if made with the recorded samples, which generally comprise non-overlapping
-sets of neurons and differ in size. Third, only at the full-network level
-can we meaningfully assess whether correlations of some order are
-informationally sufficient: as explained in
-\sect~\ref{sec:marginalization}, any such assessment at the sample level
-leads to conclusions opposite to the ones we think we're reaching.
+during a recording? The approach presented here gives probability
+distributions over the possible answers to these questions.
 
@@
 
-Regarding~\ref{item:maxent_step}, we assume that you're familiar with the
+Regarding~\ref{item:maxent_step}, we assume familiarity with the
 maximum-entropy method. We actually use the \emph{minimum-relative-entropy}
@@
 The maximum-entropy distribution is thus an approximation of our belief
-distribution about four completely different quantities. Note that the four
+distribution about four different quantities. Note that the four
 distributions above numerically differ in higher-order approximations.
@@
 frequency distributions
-\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We say more about
-these two points in \sect***.
+\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these two
+points further in \sect~\ref{sec:assumptions}.
 
@@
   % \equiv
-  \sum_{\ya} %m!\,
-  \binom{\ya}{m}\,  \binom{n}{m}^{-1}\, \pxx(\ya),
+  \sum_{\ya=0}^{n} %m!\,
+  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya),
   \qquad 1\le m \le n
@@
 normalized by the number of distinct $m$-tuples. For example, with $m=2$,
-if $na=4$ neurons spike in a network of $n=5$, we have $\tbinom{4}{2}=6$
+if $a=4$ neurons spike in a network of $n=5$, we have $\tbinom{4}{2}=6$
 distinct pairs of simultaneously spiking neurons, and the total number of

### 2019-11-13 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-13T16:55:17+0100
+%% Last-Updated: 2019-11-13T19:14:02+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\providecommand{\mparcnote}{\texttt{mp_arc} eprints available at \url{http://www.ma.utexas.edu/mp_arc/}.}
+}
+\newcommand*{\haleprint}[1]{%\global\def\arxivp{\arxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
+\texttt{\urlalt{https://hal.archives-ouvertes.fr/#1}{HAL:\hspace{0pt}#1}}%
+%\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
+%\renewcommand{\arxivnote}{\texttt{arXiv} eprints available at \url{http://arxiv.org/}.}
 }
@@
 %\newcommand*{\fm}{f}
-\newcommand*{\px}{p_{\text{me}}}
+\newcommand*{\px}{P_{\text{me}}}
 \newcommand*{\pxx}{p}
+\newcommand*{\pxxx}{P}
 \newcommand*{\Nd}{\textrm{N}}
@@
 minimum-relative-entropy method \citep{jaynes1957}[much clearer
-in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b,grandy1980}
+in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b_r1996,grandy1980}
 has been used for different kinds of estimations of the neuronal activity
@@
 four different distributions for the full network:
-\begin{itemize}
+\begin{enumerate}[label=(\roman*)]
 \item the most probable \emph{frequency} distribution for the total
@@
   % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
-\end{itemize}
+\end{enumerate}
 The maximum-entropy distribution is thus an approximation of our belief
@@
 \citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these two
-points further in \sect~\ref{sec:assumptions}.
+remarks further in \sect~\ref{sec:assumptions}.
 
@@
 they linearly determine each other because $\tbinom{\ya}{m}$ is a
-polynomial in $\ya$ of degree $m$. Fixing the ones is therefore equivalent
-to fixing the others. But the normalized factorial moments have this
-extremely convenient property: \emph{the first $n$ normalized factorial
-  moments for a sample and for the full network are numerically identical}:
+polynomial in $\ya$ of degree $m$. Specifying one set is therefore
+equivalent to specifying the other set. But the normalized factorial
+moments have a convenient property for our analysis
+\citep[\sect~2.3]{portamanaetal2015}{potts1953}: \emph{the first $n$
+  normalized factorial moments for the sample and for the full network are
+  numerically identical}:
 \begin{equation}
   \label{eq:equal_m_factorialmoments}
-    \sum_{\ya} %m!\,
-  \binom{\ya}{m}\,  \binom{n}{m}^{-1}\, \pxx(\ya)
-=    \sum_{\yA} %m!\,
-  \binom{\yA}{m}\,  \binom{N}{m}^{-1}\, \pxx(\yA),
+    \sum_{\ya=0}^{n} %m!\,
+  \binom{\ya}{m}\bigg/  \binom{n}{m}\; \pxx(\ya)
+=    \sum_{\yA=0}^{N} %m!\,
+  \binom{\yA}{m}\bigg/  \binom{N}{m}\; \pxxx(\yA),
   % \E\condt[\big]{\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
@@
 \end{equation}
-where $\pxx(\yA)$ is our distribution of belief about the full-network
+where $\pxxx(\yA)$ is our distribution of belief about the full-network
 activity.
@@
 We can therefore apply the maximum-entropy method to obtain a distribution
-for the full network by constraining its $m$th factorial moment to be equal
-to the sample's recorded average of $\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}$,
-for as many $m$ as we please with $1\le m \le n$. In formulae these
-constraints on $\px(\yA)$ are
+$\px(\yA)$ for the full network by constraining its $m$th factorial moment
+to be equal to the sample's recorded average of
+$\tbinom{\ya}{m}/\tbinom{n}{m}$, for as many $m$ as we please with
+$1\le m \le n$. In formulae, the constraints on $\px(\yA)$ are
 \begin{equation}
   \label{eq:final_constraint}
-\underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\,\binom{n}{m}^{-1}
+\underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\bigg/\binom{n}{m}
   \equiv
    \sum_{\ya}
-  \binom{\ya}{m}\,  \binom{n}{m}^{-1} \yff_{\ya}}_{\text{measured moments}}
-  =
+  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured moments}}
+  \mathrel{\;=\;}
    \underbracket{\sum_{\yA}
-  \binom{\yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA)}_{\text{distribution moments}}
-\end{equation}
-for all $m$ we wish.
-
-The calculation amounts to a convex optimization
+  \binom{\yA}{m}\bigg/  \binom{N}{m} \; \px(\yA)}_{\text{distribution moments}}.
+\end{equation}
+
+Finding the constrained maximum-entropy distribution $\px(\yA)$ amounts to
+a convex optimization
 \citep{meadetal1984}[\chap~10]{pressetal1988_r2007}{fangetal1997,boydetal2004_r2009,portamana2017b}
 and for the numbers $N,n,m$ considered in the present work it can be done
-on a modern computer without approximations of normalization constants or
-potential functions.
+on a modern computer without approximations of the partition functions that
+appear in the formalism.
 
@@
 
-For concreteness's sake we'll consider the maximum-entropy distribution
-$\px(\yA)$ as \emph{the most probable frequency distribution} for the
-full-network activity during the recording; but remember that it is also
-the approximation of three other distributions, as discussed in the
-previous section.
+% In the followingFrom a Bayesian point of view, the maximum-entropy distribution $\px(\yA)$
+% can be considered as \emph{the most probable frequency distribution} for
+% the full-network activity during the recording. But it also approximates
+% three other distributions, as discussed in the previous section.
 
@@
 distribution of the sample, and at the same time illustrates the use of the
-approach in questions of statistic sufficiency (typically limited to two or
-three moments). Figure~\ref{fig:plots_6mom} shows the resulting densities
-(that is, distribution${}\times N$) for full-network sizes $N=n$,
-$N=1\,000$, $N=5\,000$, $N=10\,000$\mynote{motivate?}. The case $N=n$
-corresponds to applying the maximum-entropy method at the sample level; it
-can be observed that with six moments it reproduces almost exactly the
-measured frequency distribution.
+approach in questions of statistical sufficiency \citep[cf. for
+example][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
+Figure~\ref{fig:plots_6mom} shows the resulting densities (that is,
+distribution${}\times N$) for full-network sizes $N=n$, $N=1\,000$,
+$N=5\,000$, $N=10\,000$. The case $N=n$ corresponds to applying the
+maximum-entropy method at the sample level; it can be observed that with
+six moments it reproduces almost exactly the measured frequency
+distribution.
 \begin{figure}[!p]

### 2019-11-13 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-13T19:14:02+0100
+%% Last-Updated: 2019-11-13T20:29:21+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We want to stress the usefulness of making a quantified guess about the
-activity of a larger brain area. Such a guess seems indeed to be the
-primary idea behind recording a sample from that area. %  Second, it allows us
-% to make comparisons across experimental sessions; such comparisons would me
-% difficult or meaningless if made with the recorded samples, which generally
-% comprise non-overlapping sets of neurons and differ in size.
-There's also an important methodological reason: the question whether
-correlations of some order are informationally sufficient cannot be
-properly assessed at the sample level. When the difference in size between
-sample and full network is too large, correlation sufficiency for the
-sample implies the \emph{lack} of correlation sufficiency for the larger
-network, and vice versa. This point is explained in detail in
-\sect~\ref{sec:marginalization}.
-
 The model we propose is based on the straightforward combination of the
@@
 
-Maximum-entropy models have often been used to test whether a smaller part
-of the sample data, for example time averages and second-order
-correlations, is sufficient for quantifying our \dob\ about some aspects of
-the neuronal activity of the sample. This use of maximum-entropy at the
-sample level, however, has an important issue. We discuss this issue and
-explain how our approach bypasses it in \sect~\ref{sec:marginalization}.
-Moreover, our proposed model gives clearer answers to questions of
-correlation sufficiency: we show this by applying it to the two data set
-above addressing an example question of sufficiency, two moments versus
-four moments.
+We want to stress the usefulness of making a quantified guess about the
+activity of a larger brain area. Such a guess seems indeed to be the
+primary idea behind recording a sample from that area. %  Second, it allows us
+% to make comparisons across experimental sessions; such comparisons would me
+% difficult or meaningless if made with the recorded samples, which generally
+% comprise non-overlapping sets of neurons and differ in size.
+There's also an important methodological reason. Maximum-entropy methods
+have been used to assess the informational sufficiency of pairwise
+correlations versus correlations of higher order \citep[see for
+example][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
+But when the difference in size between sample and full network is too
+large, correlation sufficiency for the sample implies the \emph{lack} of
+correlation sufficiency for the larger network, and vice versa, as will be
+explained in detail in \sect~\ref{sec:marginalization}. Therefore,
+maximum-entropy applications at the sample level can be deceptive for
+questions of statistical sufficiency, and a full-network application is
+more reliable. We show this point in \sect~\ref{sec:application} for a real
+data set.
+
+% Maximum-entropy models have often been used to test whether a smaller part
+% of the sample data, for example time averages and second-order
+% correlations, is sufficient for quantifying our \dob\ about some aspects of
+% the neuronal activity of the sample. This use of maximum-entropy at the
+% sample level, however, has an important issue. We discuss this issue and
+% explain how our approach bypasses it in \sect~\ref{sec:marginalization}.
+% Moreover, our proposed model gives clearer answers to questions of
+% correlation sufficiency: we show this by applying it to the two data set
+% above addressing an example question of sufficiency, two moments versus
+% four moments.
 
@@
 distribution of the sample, and at the same time illustrates the use of the
-approach in questions of statistical sufficiency \citep[cf. for
-example][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
+approach in questions of statistical sufficiency.
 Figure~\ref{fig:plots_6mom} shows the resulting densities (that is,
-distribution${}\times N$) for full-network sizes $N=n$, $N=1\,000$,
-$N=5\,000$, $N=10\,000$. The case $N=n$ corresponds to applying the
-maximum-entropy method at the sample level; it can be observed that with
-six moments it reproduces almost exactly the measured frequency
-distribution.
+distribution${}\times N$) for four example values of full-network sizes:
+$N=n$, $N=1\,000$, $N=5\,000$, $N=10\,000$. The case $N=n$ corresponds to
+applying the maximum-entropy method at the sample level; it can be observed
+that with six moments it reproduces almost exactly the measured frequency
+distribution. We discuss the case of unknown $N$ in \sect~\ref{sec:N}.
 \begin{figure}[!p]
@@
 
-The distribution for the full-network is sharper than the measured
-frequency distribution for the sample; the sharper the larger $N$ is. Most
-remarkably, it has two distinct low-activity modes for the first data set.
-But also the second data set presents, upon closer inspection, a small
-shoulder on the right of the mode, suggestive of two activity regimes. We
-frame no hypotheses about the biological cause of these two modes (they
-could stem from the presence of different kinds of cells or modules). These
-features are clearly not present in the sample or in the maximum-entropy
-distribution at the sample level. The application of the probability
-calculus thus reveals interesting possible features of the full network.
-\textcolor{white}{If you find this you can claim a postcard from us.}
+The figure shows that the distribution for the full-network is less flat
+than the measured frequency distribution for the sample; their difference
+increases with $N$. Most remarkably, for the first data set the
+distribution has two distinct low-activity modes. For the second data
+the distribution presents a small shoulder on the right of the
+mode.% , suggestive of two activity regimes. We frame no hypotheses about the
+% biological cause of these two modes (they could stem from the presence of
+% different kinds of cells or modules)
+These features are clearly not present in the sample or in the
+maximum-entropy distribution at the sample level. The application of the
+probability calculus thus reveals interesting possible features of the full
+network. \textcolor{white}{If you find this you can claim a postcard from
+  us.}
 
@@
 
-To illustrate how our approach can be applied to studies of sufficient
-statistics, \fig~\ref{fig:plots_2_4mom} shows the results for two
-constrained moments (equivalent to constraining means and correlations
-only) and for four constrained moments, with $N=10\,000$. In either data
-set we obtain two very distinct distributions. For the first data set in
-particular, four moments lead to a bimodal distribution, whereas two to a
-unimodal one, showing that two moments would not be sufficient
-statistics. For comparison, the two distributions obtained applying
-maximum-entropy \emph{at the sample level} are shown as an inset in the
-plot for the first data set: their differences aren't so glaring as in the
-full-network application.
 \begin{figure}[!p]
@@
 %maxent_solutions_different_moments_prior2_6moments.nb
+To illustrate how our approach can be applied to studies of sufficient
+statistics, \fig~\ref{fig:plots_2_4mom} shows the full-network
+distributions, for $N=10\,000$, obtained constraining the first two moments
+(equivalent to constraining means and correlations) and the first four
+moments; the frequency distribution of the sample is also shown for
+comparison. In both data sets, the two-moment constrain leads to a
+distribution quite different from the four-moment constrain. For the first
+data set in particular, one distribution is bimodal, the other unimodal,
+showing that two moments are not informationally sufficient.
+
+It's useful to contrast the distributions for the full network with those
+obtained applying the maximum-entropy method \emph{at the sample level}
+with the same constraints, as traditionally done. The sample-level
+distributions are shown as an inset in the plot for the first data set. We
+see that constraining four moments doesn't lead to a distribution much
+different from constraining only two. This would suggest that two moments
+are informationally almost sufficient. This conclusion would be deceptive:
+it obviously cannot be extended to the full network. We discuss this point
+more in detail in the next section.
 
@@
 
-In the neurosciences the maximum-entropy method is often applied to
-questions of sufficient statistics; for example: do pairwise correlations
-in the neuronal activity provide the same information as the full frequency
+As mentioned in the Introduction (see references there), in the
+neurosciences the maximum-entropy method has also been applied to questions
+of sufficient statistics. For example: do pairwise correlations in the
+neuronal activity provide the same information as the full frequency
 distribution of activities? In this section we discuss how our proposed
@@
 
-Let's make the question more precise first. We want to assess our degree of
-belief about the frequencies of the activities of the sample, or about the
-activity of the sample in a new time bin. For this assessment we should use
-all measured data we have. But it sometimes happens that dropping part of
-the data -- for example, the measured third- and higher-order moments --
+The question must be made more precise first. We want to assess our degree
+of belief about the frequencies of the activities of the sample, or about
+the activity of the sample in a new time bin. For this assessment we should
+use all measured data we have. But it sometimes happens that dropping part
+of the data -- for example, the measured third- and higher-order moments --
 leaves our assessment almost unchanged: the dropped data are
@@
 example, first and second moments -- are then \emph{informationally
-  sufficient}. This informational quasi-sufficiency is interesting because
-it may hint at peculiar biological  mechanisms or dependencies.
+  sufficient}. % This informational quasi-sufficiency is interesting because
+% it may give clues about network organization and dynamics.
 
@@
 distributions constructed \emph{for the sample}, ignoring the full network
-it's sampled from. The method introduced in the present work, however,
-reveals an issue with this application at the sample level. The issue
-doesn't come from the method itself but from a general fact of probability
-theory. If a probability distribution has some sufficient statistics
+it's sampled from \citep[for
+example][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
+The method introduced in the present work, however, reveals an issue with
+this application at the sample level. The issue doesn't come from the
+method itself but from a general fact of probability theory. If a
+probability distribution has some sufficient statistics
 \citep[\sect~4.5]{bernardoetal1994}{fortinietal2000}, then its marginals
@@
 on the size of the full network, $N$. The full network can't be the whole
-brain, of course; that would be silly. How large can $N$ be?
+brain, of course. How large should or can $N$ be in our formulae?
 
@@
 practically impossible to specify the exact number of neurons, but its
-order of magnitude is enough for making qualitative inferences. As we'll
-discuss in the next section, even if the exact number were known the
-maximum-entropy distribution ought still to be interpreted qualitatively.
+order of magnitude is enough for making qualitative inferences. As
+discussed in the next section, even if the exact number were known, the
+maximum-entropy distribution ought to be interpreted qualitatively.
 

### 2019-11-14 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-13T20:29:21+0100
+%% Last-Updated: 2019-11-14T22:48:47+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\subsection*{Use with moment constraints}
+\label{sec:use_moments}
+
 \begin{figure}[!p]
@@
 (equivalent to constraining means and correlations) and the first four
-moments; the frequency distribution of the sample is also shown for
-comparison. In both data sets, the two-moment constrain leads to a
-distribution quite different from the four-moment constrain. For the first
-data set in particular, one distribution is bimodal, the other unimodal,
-showing that two moments are not informationally sufficient.
-
-It's useful to contrast the distributions for the full network with those
-obtained applying the maximum-entropy method \emph{at the sample level}
-with the same constraints, as traditionally done. The sample-level
+moments. The frequency distribution of the sample is also shown for
+comparison. In both data sets the two-moment constraint leads to a
+distribution quite different from the four-moment constraint. For the first
+data set in particular, one distribution is bimodal, the other unimodal:
+this difference shows that the first two moments alone are \emph{not}
+informationally sufficient for this specific data set.
+
+It is instructive to contrast the distributions for the full network with
+those obtained applying the maximum-entropy method \emph{at the sample
+  level} with the same constraints, as traditionally done. The sample-level
 distributions are shown as an inset in the plot for the first data set. We
@@
 \\[3em]%
-\includegraphics[width=\linewidth]{_scripts/comparison_marginals.pdf}%
+\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
 \caption{***}
@@
 \end{figure}
-%comparison_ME_sample_full_levels.nb
+% comparison_ME_sample_full_levels.nb
+%
+% relentropies of sample2mom, marg2mom, sample4mom, marg4mom:
+% 322.72, -85.7627, -42.8005, -4.23746
+% correspond to probability ratios of
+% 1., 8.11718\times 10^{102}, 3.69524\times 10^{121}, 2.06704\times 10^{138}
+% comparison_ME_sample_full_levels.nb
+%
 But does the application at the full-network level lead to appreciably

### 2019-11-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-14T22:48:47+0100
+%% Last-Updated: 2019-11-15T15:15:50+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\newcommand*{\ca}{{c.}}
-%\newcommand*{\eg}{{e.g.}}
+\newcommand*{\eg}{{e.g.}}
 \newcommand*{\foll}{{ff.}}
@@
 have been used to assess the informational sufficiency of pairwise
-correlations versus correlations of higher order \citep[see for
+correlations and correlations of higher order \citep[see for
 example][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
@@
 large, correlation sufficiency for the sample implies the \emph{lack} of
-correlation sufficiency for the larger network, and vice versa, as will be
-explained in detail in \sect~\ref{sec:marginalization}. Therefore,
+correlation sufficiency for the larger network, and vice versa. Therefore,
 maximum-entropy applications at the sample level can be deceptive for
 questions of statistical sufficiency, and a full-network application is
-more reliable. We show this point in \sect~\ref{sec:application} for a real
-data set.
+more reliable. We discuss this point in detail in \sect~\ref{sec:marginalization}.
 
@@
 
-\bigskip
-
 \subsection*{Use with moment constraints}
@@
 distribution quite different from the four-moment constraint. For the first
-data set in particular, one distribution is bimodal, the other unimodal:
-this difference shows that the first two moments alone are \emph{not}
-informationally sufficient for this specific data set.
-
-It is instructive to contrast the distributions for the full network with
-those obtained applying the maximum-entropy method \emph{at the sample
-  level} with the same constraints, as traditionally done. The sample-level
-distributions are shown as an inset in the plot for the first data set. We
-see that constraining four moments doesn't lead to a distribution much
-different from constraining only two. This would suggest that two moments
-are informationally almost sufficient. This conclusion would be deceptive:
-it obviously cannot be extended to the full network. We discuss this point
-more in detail in the next section.
-
-\section{Example application for sufficiency of partial statistics:\\the
-  issue with methods at the sample level}
+data set in particular, one distribution is bimodal, the other unimodal.
+This visual difference shows directly -- without the need to calculate
+relative entropies, for example -- that the first two moments alone are
+\emph{not} informationally sufficient for this specific data set.
+
+It is instructive to contrast the conclusions about correlation sufficiency
+reached through this full-network analysis, with those we would have
+reached by applying the maximum-entropy method \emph{at the sample
+  level}, as traditionally done. Such a comparison involves an important
+methodological caveat, so we discuss it in full in the next section.
+
+\section{Quantifying the importance of higher-order correlations:\\
+  the limitations of methods at the sample level}
 \label{sec:marginalization}
@@
 As mentioned in the Introduction (see references there), in the
-neurosciences the maximum-entropy method has also been applied to questions
-of sufficient statistics. For example: do pairwise correlations in the
-neuronal activity provide the same information as the full frequency
-distribution of activities? In this section we discuss how our proposed
-application bears on this kind of questions. % For concreteness we use an
+neurosciences the maximum-entropy method has also been used as a way of
+quantifying the \enquote{cooperativity} \citep{gersteinetal1985} or
+\enquote{interaction}
+\citep{martignonetal1995,schneidmanetal2006,shlensetal2006} or
+\enquote{synchrony} \citep{bohteetal2000,amarietal2003} of neuronal
+activity. In this section we discuss how our proposed
+application bears on this kind of quantification. % For concreteness we use an
 % example with specific statistics: first and second moments (equivalent to
-% mean and correlations), versus first-to-fourth moments. But our discussion
-% holds for any sets of statistics.
-
-The question must be made more precise first. We want to assess our degree
-of belief about the frequencies of the activities of the sample, or about
-the activity of the sample in a new time bin. For this assessment we should
-use all measured data we have. But it sometimes happens that dropping part
-of the data -- for example, the measured third- and higher-order moments --
-leaves our assessment almost unchanged: the dropped data are
-\emph{informationally irrelevant}, or almost so. The remaining data -- for
-example, first and second moments -- are then \emph{informationally
+% mean and correlations), versus first-to-fourth moments. But our
+% discussion holds for any sets of statistics.
+
+Let us translate this question into more precise terms first, because
+\enquote{cooperativity}, \enquote{interaction}, and similar notions are
+vague. A translation in terms of \emph{informational sufficiency} seems to
+suit all these notions \parentext{for an intuitive understanding of
+  sufficiency, which goes back to \cite{fisher1922}, see \eg\
+  \cite[\sect~4.5]{bernardoetal1994}; \cite[\chap~8,
+  \sect~14.2]{jaynes1994_r2003}; \cite{cifarellietal1982};
+  \cite{kullbacketal1951}. For technical results and connections with the
+  notion of symmetry see \eg\
+  \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,cifarellietal1980,cifarellietal1981,diaconisetal1981,lauritzen1982_r1988,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,lauritzen2007,ayetal2015}}.
+The idea is as follows. We want to assess our degree of belief about the
+frequencies of the activities of the sample, or about the activity of the
+sample in a new time bin. For this assessment we should use all measured
+data we have. But it sometimes happens that dropping part of the data --
+for example, the measured third- and higher-order moments -- leaves our
+assessment almost unchanged: the dropped data are \emph{informationally
+  irrelevant}, or almost so. The remaining data -- for example, first and
+second moments -- are then \emph{informationally
   sufficient}. % This informational quasi-sufficiency is interesting because
@@
 % relentropies of sample2mom, marg2mom, sample4mom, marg4mom:
-% 322.72, -85.7627, -42.8005, -4.23746
+% -322.72, -85.7627, -42.8005, -4.23746 nats =
+% -1401.56, -372.462, -185.88, -18.403 db
 % correspond to probability ratios of
 % 1., 8.11718\times 10^{102}, 3.69524\times 10^{121}, 2.06704\times 10^{138}
+%
+% difference in relentropies between 4 mom and 2 mom:
+% sample level: 279.92 nats = 1215.68 db
+% marg level: 81.5252 nats =  354.059 db
+% correspond to prob ratios of
+% 3.69524\times 10^{121},2.5465\times 10^{35}
+%
 % comparison_ME_sample_full_levels.nb

### 2019-11-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-15T15:15:50+0100
+%% Last-Updated: 2019-11-15T15:18:17+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 assessment almost unchanged: the dropped data are \emph{informationally
-  irrelevant}, or almost so. The remaining data -- for example, first and
+  irrelevant} or almost so. The remaining data -- for example, first and
 second moments -- are then \emph{informationally
@@
 We can approximately check how much a part of the data is irrelevant by
-simply dropping it from the conditional of our degree of belief, and see
-if the latter changes appreciably. When our degree of belief is built with
-the maximum-entropy method, we simply drop part of the data used as
-constraints and see how the resulting distribution changes. For example, we
-may only retain the first two moments (means and correlations) or the first
-four moments.
+simply dropping it from the conditional of our probability, and see if the
+latter changes appreciably. When our degree of belief is built with the
+maximum-entropy method, this is done by dropping part of the constraints
+and see how the resulting distribution changes. For example, we may only
+retain the first two moments (means and correlations) or the first four
+moments.
 

### 2019-11-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-15T15:18:17+0100
+%% Last-Updated: 2019-11-15T15:51:26+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We can approximately check how much a part of the data is irrelevant by
-simply dropping it from the conditional of our probability, and see if the
-latter changes appreciably. When our degree of belief is built with the
-maximum-entropy method, this is done by dropping part of the constraints
-and see how the resulting distribution changes. For example, we may only
-retain the first two moments (means and correlations) or the first four
-moments.
+We can check how much a part of the data is irrelevant by simply dropping
+it from the conditional of our probabilities, and see if the latter change
+appreciably. When our probabilities are built with the maximum-entropy
+method, this is done by dropping part of the constraints and see how the
+resulting distribution changes. For example, we may only retain the first
+two moments (means and correlations) or the first four moments. This
+procedure is equivalent to calculating and comparing the probabilities of
+our hypotheses about sufficiency, conditional the data, assuming their
+pre-data probabilities to be equal. We perform this calculation below.
 
@@
 %
+
+ This procedure is equivalent to calculating and comparing the
+probabilities of our hypotheses about sufficiency, conditional the data, if
+their pre-data probability to be equal; so that for example
+\begin{equation}
+  \label{eq:compare_hypoth_conditional}
+  \pf(\text{\small\enquote{first two moments are sufficient}} \| \text{\small data}) \propto
+  \pf(\text{\small data} \|
+  \text{\small\enquote{first two moments are sufficient}}).
+\end{equation}
+
 But does the application at the full-network level lead to appreciably

### 2019-11-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-15T15:51:26+0100
+%% Last-Updated: 2019-11-15T19:22:27+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 neurosciences the maximum-entropy method has also been used as a way of
-quantifying the \enquote{cooperativity} \citep{gersteinetal1985} or
+quantifying the \enquote{cooperativity} \citep[\eg][]{gersteinetal1985} or
 \enquote{interaction}
-\citep{martignonetal1995,schneidmanetal2006,shlensetal2006} or
-\enquote{synchrony} \citep{bohteetal2000,amarietal2003} of neuronal
+\citep[\eg][]{martignonetal1995,schneidmanetal2006,shlensetal2006} or
+\enquote{synchrony} \citep[\eg][we're only citing the first papers using
+these terms]{bohteetal2000,amarietal2003} of neuronal
 activity. In this section we discuss how our proposed
@@
 vague. A translation in terms of \emph{informational sufficiency} seems to
-suit all these notions \parentext{for an intuitive understanding of
-  sufficiency, which goes back to \cite{fisher1922}, see \eg\
-  \cite[\sect~4.5]{bernardoetal1994}; \cite[\chap~8,
-  \sect~14.2]{jaynes1994_r2003}; \cite{cifarellietal1982};
+suit all these notions \parentext{for the technical definition and
+  intuitive understanding of sufficiency, which goes back to
+  \cite{fisher1922}, see \eg\ \cite[\sect~4.5]{bernardoetal1994};
+  \cite[\chap~8, \sect~14.2]{jaynes1994_r2003}; \cite{cifarellietal1982};
   \cite{kullbacketal1951}. For technical results and connections with the
@@
   \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,cifarellietal1980,cifarellietal1981,diaconisetal1981,lauritzen1982_r1988,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,lauritzen2007,ayetal2015}}.
-The idea is as follows. We want to assess our degree of belief about the
+The idea is as follows. We want to assess our probabilities about the
 frequencies of the activities of the sample, or about the activity of the
 sample in a new time bin. For this assessment we should use all measured
-data we have. But it sometimes happens that dropping part of the data --
-for example, the measured third- and higher-order moments -- leaves our
-assessment almost unchanged: the dropped data are \emph{informationally
-  irrelevant} or almost so. The remaining data -- for example, first and
-second moments -- are then \emph{informationally
+data and statistics we have. But it can be the case that dropping part of
+the data or statistics -- for example, the measured third- and higher-order
+moments -- leaves our probabilities almost unchanged. This means that the dropped
+statistics are \emph{informationally irrelevant} or almost so. The
+remaining statistics -- for example, first and second moments -- are then
+\emph{informationally
   sufficient}. % This informational quasi-sufficiency is interesting because
@@
 
-We can check how much a part of the data is irrelevant by simply dropping
-it from the conditional of our probabilities, and see if the latter change
-appreciably. When our probabilities are built with the maximum-entropy
-method, this is done by dropping part of the constraints and see how the
-resulting distribution changes. For example, we may only retain the first
-two moments (means and correlations) or the first four moments. This
-procedure is equivalent to calculating and comparing the probabilities of
-our hypotheses about sufficiency, conditional the data, assuming their
-pre-data probabilities to be equal. We perform this calculation below.
+We can thus quantify how much a subset of statistics is relevant with
+respect to a larger set by keeping only the subset in the conditional of
+our probabilities, and checking how much the latter change. When our
+probabilities are built with the maximum-entropy method, this means keeping
+only part of the constraints (and re-maximizing the entropy). For example,
+keeping only the first two moments (means and correlations) out of the
+first four moments. This procedure is actually equivalent to comparing the
+probabilities of our hypotheses about sufficiency, conditional the
+\emph{full} data, assuming their pre-data probabilities to be equal. We
+show this equivalence below and perform the calculations for our first data
+set.
 
@@
 
-In the neurosciences this check is often done with maximum-entropy
-distributions constructed \emph{for the sample}, ignoring the full network
-it's sampled from \citep[for
-example][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
-The method introduced in the present work, however, reveals an issue with
-this application at the sample level. The issue doesn't come from the
-method itself but from a general fact of probability theory. If a
-probability distribution has some sufficient statistics
-\citep[\sect~4.5]{bernardoetal1994}{fortinietal2000}, then its marginals
-\emph{cannot} have the same sufficient statistics, and vice versa, except
-for trivial cases such as uniform distributions. This generally also holds
-among different marginals: if a marginal has some sufficient statistics,
-another won't have it. This impossibility is known in statistical
-mechanics: marginals of Gibbs states aren't Gibbs states
-\citep{maesetal1999}. Mathematically this impossibility translates into a
-system of $n$ independent equations in $m$ unknowns with $n>m$
-\citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.
-
-In our example this means that if means and pairwise correlations or the
-first four moments are informationally sufficient for a particular sample
-from a brain area, then they \emph{cannot} be sufficient for the full
-network of neurons in that area, or for a different sample. Vice versa, if
-we assume that they are informationally sufficient for the full network,
-then we must expect them \emph{not} to be sufficient for any recorded
-sample thereof. Now, questions about sufficient statistics are meant to be
-addressed to a whole brain area, not just to some specific, casually
-recorded sample. Thus, paradoxically, maximum-entropy methods applied
-\emph{at the sample level} give an answer opposite to the one we might
-think they're giving.
+Before applying this idea to neuronal activity, however, it is important to
+keep in mind that \emph{informational sufficiency is not preserved under
+  sampling}: if a probability distribution has some sufficient statistics,
+then its marginals \emph{cannot} have the same sufficient statistics, and
+vice versa, except for trivial cases such as uniform probability
+distributions. This generally also holds among different marginals: if a
+statistics is sufficient for some marginal, then it will not be sufficient
+for another marginal. This impossibility is known in statistical mechanics:
+marginals of Gibbs states are not Gibbs states \citep{maesetal1999}.
+Mathematically this impossibility comes from the Pitman-Koopman theorem
+\citep{koopman1936,pitman1936,darmois1935} connecting sufficient statistics
+and the exponential family, and translates into the general impossibility
+of solving a system of $n$ independent equations in $m$ unknowns with more
+equations than unknowns, $n>m$ \citep[\sect~3.1]{portamanaetal2015}.
+
+This fact is important in the analysis of neural activity when it is based
+on small samples from brain regions. If means and pairwise correlations
+seem informationally sufficient for a particular sample from a brain area,
+then they may well be \emph{insufficient} for the full network of neurons
+constituting that area. Vice versa, if they are informationally sufficient
+for the full network, then they may turn out \emph{not} to be sufficient
+(or less so) for any recorded sample thereof. Hence, if we wonder about
+\enquote{cooperativity} or \enquote{interaction} of a brain area, we cannot
+reliably use maximum-entropy distributions \emph{for a sample only}. Yet
+the neuroscientific literature is besprinkled with such sample-level
+analyses
+\citep[\eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
 

### 2019-11-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-15T19:22:27+0100
+%% Last-Updated: 2019-11-15T21:36:39+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \citep[\eg][]{martignonetal1995,schneidmanetal2006,shlensetal2006} or
-\enquote{synchrony} \citep[\eg][we're only citing the first papers using
+\enquote{synchrony} \citep[\eg][-- we're only citing early papers using
 these terms]{bohteetal2000,amarietal2003} of neuronal
@@
 
-Let us translate this question into more precise terms first, because
-\enquote{cooperativity}, \enquote{interaction}, and similar notions are
-vague. A translation in terms of \emph{informational sufficiency} seems to
-suit all these notions \parentext{for the technical definition and
-  intuitive understanding of sufficiency, which goes back to
-  \cite{fisher1922}, see \eg\ \cite[\sect~4.5]{bernardoetal1994};
-  \cite[\chap~8, \sect~14.2]{jaynes1994_r2003}; \cite{cifarellietal1982};
-  \cite{kullbacketal1951}. For technical results and connections with the
-  notion of symmetry see \eg\
-  \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,cifarellietal1980,cifarellietal1981,diaconisetal1981,lauritzen1982_r1988,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,lauritzen2007,ayetal2015}}.
-The idea is as follows. We want to assess our probabilities about the
+\enquote{Cooperativity}, \enquote{interaction}, and similar terms are
+vague, so we need to translate them into a more precise notion first. We
+propose the notion of \emph{informational sufficiency}
+\citep[\sect~4.5]{bernardoetal1994}[\chap~8 \amp\
+\sect~14.2]{jaynes1994_r2003}{cifarellietal1982,kullbacketal1951}[the
+notion goes back to][]{fisher1922} because it relates to those terms, is
+intuitive, and connects directly to probability theory and maximum-entropy
+distributions. Its idea is as follows. Our probabilities about the
 frequencies of the activities of the sample, or about the activity of the
-sample in a new time bin. For this assessment we should use all measured
-data and statistics we have. But it can be the case that dropping part of
+sample in a new time bin, are in principle conditional on all experimental
+data and statistics we have. But it can be the case that discarding part of
 the data or statistics -- for example, the measured third- and higher-order
-moments -- leaves our probabilities almost unchanged. This means that the dropped
-statistics are \emph{informationally irrelevant} or almost so. The
-remaining statistics -- for example, first and second moments -- are then
-\emph{informationally
-  sufficient}. % This informational quasi-sufficiency is interesting because
+moments -- leaves our probabilities almost unchanged. This means that the
+discarded statistics are \emph{informationally irrelevant} or almost so.
+The remaining statistics -- for example, first and second moments -- are
+\emph{informationally sufficient}. \parentext{For more technical results
+  and connections with the notion of symmetry see \eg\
+  \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,cifarellietal1980,cifarellietal1981,diaconisetal1981,lauritzen1982_r1988,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,lauritzen2007,ayetal2015}.}
+% This informational quasi-sufficiency is interesting because
 % it may give clues about network organization and dynamics.
 
-We can thus quantify how much a subset of statistics is relevant with
-respect to a larger set by keeping only the subset in the conditional of
-our probabilities, and checking how much the latter change. When our
-probabilities are built with the maximum-entropy method, this means keeping
-only part of the constraints (and re-maximizing the entropy). For example,
-keeping only the first two moments (means and correlations) out of the
-first four moments. This procedure is actually equivalent to comparing the
-probabilities of our hypotheses about sufficiency, conditional the
-\emph{full} data, assuming their pre-data probabilities to be equal. We
-show this equivalence below and perform the calculations for our first data
-set.
-
-% For example: does knowledge about
-% third-order correlations change our state of uncertainty about network
-% activity in an appreciable way, compared to knowledge about second-order
-% correlations only? The method is used as follows to answer this question.
-% We imagine to be ignorant of the measured frequency distribution for the
-% network activity, and to know only up to second-order measured statistics,
-% or up to third-order. Two maximum-entropy distributions are then built,
-% constrained under the two sets of moments under comparison. Finally, the
-% resulting distributions are compared with each other and with the measured
-% frequency distribution. This recipe generalizes to any subsets of measured
-% statistics, of course.
-
-Before applying this idea to neuronal activity, however, it is important to
-keep in mind that \emph{informational sufficiency is not preserved under
-  sampling}: if a probability distribution has some sufficient statistics,
-then its marginals \emph{cannot} have the same sufficient statistics, and
-vice versa, except for trivial cases such as uniform probability
-distributions. This generally also holds among different marginals: if a
-statistics is sufficient for some marginal, then it will not be sufficient
-for another marginal. This impossibility is known in statistical mechanics:
-marginals of Gibbs states are not Gibbs states \citep{maesetal1999}.
+We can thus quantify the informational relevance of a subset of statistics,
+for example first and second moments (means and correlations), with respect
+to a larger set, for example the first four moments, by comparing the
+probabilities conditional on the subset and on the full set. For
+probabilities built with the maximum-entropy method, this means comparing
+those constrained on the subset and on the full set. This procedure is
+actually equivalent to comparing the probabilities of the two hypotheses
+about sufficiency, conditional the \emph{full} data, assuming their
+pre-data probabilities to be equal. We show this equivalence below and
+perform the calculations for our first data set.
+
+\medskip
+
+Before applying this notion to neuronal activity, however, we must keep in
+mind that \emph{informational sufficiency is not preserved under sampling}:
+if a probability distribution has some sufficient statistics, then its
+marginals, such as the distribution for a sample, \emph{cannot} have the
+same sufficient statistics, and vice versa; except for trivial cases such
+as uniform probability distributions. % This generally holds also among
+% different marginals: if a statistics is sufficient for some marginal, then
+% it will not be sufficient for another marginal.
+This impossibility is known in statistical mechanics: if a system is
+described by a Gibbs state, its subsystems cannot be perfectly described by
+Gibbs states \citep[\eg][and references therein]{maesetal1999}.
 Mathematically this impossibility comes from the Pitman-Koopman theorem
 \citep{koopman1936,pitman1936,darmois1935} connecting sufficient statistics
-and the exponential family, and translates into the general impossibility
-of solving a system of $n$ independent equations in $m$ unknowns with more
-equations than unknowns, $n>m$ \citep[\sect~3.1]{portamanaetal2015}.
-
-This fact is important in the analysis of neural activity when it is based
-on small samples from brain regions. If means and pairwise correlations
-seem informationally sufficient for a particular sample from a brain area,
-then they may well be \emph{insufficient} for the full network of neurons
-constituting that area. Vice versa, if they are informationally sufficient
-for the full network, then they may turn out \emph{not} to be sufficient
-(or less so) for any recorded sample thereof. Hence, if we wonder about
+and the exponential family of distributions, and translates into the
+general impossibility of solving a system of $n$ independent equations in
+$m$ unknowns with more equations than unknowns, $n>m$
+\citep[\sect~3.1]{portamanaetal2015}.
+
+This fact is important for our analysis. If, say, means and pairwise
+correlations seem informationally sufficient for a particular sample from a
+brain area, then they may well be \emph{insufficient} for the full network
+of neurons constituting that area, and vice versa. If what interests us is
 \enquote{cooperativity} or \enquote{interaction} of a brain area, we cannot
-reliably use maximum-entropy distributions \emph{for a sample only}. Yet
-the neuroscientific literature is besprinkled with such sample-level
-analyses
+therefore reliably use maximum-entropy distributions \emph{for a sample
+  only} thereof. Yet the neuroscientific literature is besprinkled with
+such sample-level analyses
 \citep[\eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
-
-Questions about informational sufficiency are therefore correctly addressed
-by applying the maximum-entropy method at the full-network level. If a
-comparison with the measured frequencies of the sample activities is
-desired, then the marginal distribution to the sample size should be used.
+The approach presented here bypasses this problem instead, because the
+maximum-entropy method is applied to obtain the distribution of the full
+network, not of the sample alone.
 

### 2019-11-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-15T21:36:39+0100
+%% Last-Updated: 2019-11-16T00:42:52+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 correlations seem informationally sufficient for a particular sample from a
-brain area, then they may well be \emph{insufficient} for the full network
-of neurons constituting that area, and vice versa. If what interests us is
-\enquote{cooperativity} or \enquote{interaction} of a brain area, we cannot
-therefore reliably use maximum-entropy distributions \emph{for a sample
-  only} thereof. Yet the neuroscientific literature is besprinkled with
-such sample-level analyses
-\citep[\eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
-The approach presented here bypasses this problem instead, because the
+brain area, then they may well not be sufficient for the full network of
+neurons constituting that area, and vice versa. So if what interests us is
+\enquote{cooperativity} or \enquote{interaction} of a brain area, it is
+unreliable to use a maximum-entropy distribution constructed only for the
+sample. % \citep[contrast this with some literature, \eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}
+The approach presented here avoids this problem, because the
 maximum-entropy method is applied to obtain the distribution of the full
@@
 %
-% relentropies of sample2mom, marg2mom, sample4mom, marg4mom:
-% -322.72, -85.7627, -42.8005, -4.23746 nats =
-% -1401.56, -372.462, -185.88, -18.403 db
+% relentropies of sample2mom, marg2mom, sample4mom, marg4mom, marg6mom
+% -322.72, -85.7627, -42.8005, -4.23746, -4.26796 nats =
+% -1401.56, -372.462, -185.88, -18.403, -18.5355 dB
 % correspond to probability ratios of
-% 1., 8.11718\times 10^{102}, 3.69524\times 10^{121}, 2.06704\times 10^{138}
+% 1., 8.11718\times 10^{102}, 3.69524\times 10^{121}, 2.06704\times 10^{138},
+% 2.00495\times 10^{138}
 %
 % difference in relentropies between 4 mom and 2 mom:
-% sample level: 279.92 nats = 1215.68 db
-% marg level: 81.5252 nats =  354.059 db
+% sample level: 279.92 nats = 1215.68 dB
+% marg level: 81.5252 nats =  354.059 dB
+% difference in relentropies between 6 mom and 4 mom:
+% marg level: -0.0305024 nats = -0.13247 dB
 % correspond to prob ratios of
-% 3.69524\times 10^{121},2.5465\times 10^{35}
+% 3.69524\times 10^{121},2.5465\times 10^{35}, 0.969958
 %

### 2019-11-16 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-16T00:42:52+0100
+%% Last-Updated: 2019-11-16T18:08:44+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\newcommand*{\ypp}{G}
+\newcommand*{\yG}{G}
 
@@
 \newcommand*{\me}{maximum-entropy}
+
+\newcommand*{\sh}{\mathrm{\varEta}}
 %%% Custom macros end @@@
@@
 The number of moments used with this method depends on the questions and
-hypotheses that a researcher is exploring; for example, hypotheses of
-sufficient statistics, such as the sufficiency of pairwise correlations to
-quantify our \dob\ about network activity. We discuss this kind of use in
+hypotheses that a researcher is exploring; for example, hypotheses about
+the \enquote{cooperativity} or \enquote{interaction} -- for example,
+pairwise or higher-order -- of the network activity. We discuss this kind
+of use in
 \sect~\ref{sec:marginalization}.%, but our main goal is to simply introduce
@@
 \end{figure}
-%maxent_solutions_different_moments_prior2_6moments.nb
+% maxent_solutions_different_moments_prior2_6moments.nb
 To illustrate how our approach can be applied to studies of sufficient
@@
 \enquote{Cooperativity}, \enquote{interaction}, and similar terms are
-vague, so we need to translate them into a more precise notion first. We
-propose the notion of \emph{informational sufficiency}
+vague, so we need to translate them into a more precise notion first. Here
+we use the notion of \emph{informational sufficiency}
 \citep[\sect~4.5]{bernardoetal1994}[\chap~8 \amp\
@@
 notion goes back to][]{fisher1922} because it relates to those terms, is
-intuitive, and connects directly to probability theory and maximum-entropy
-distributions. Its idea is as follows. Our probabilities about the
-frequencies of the activities of the sample, or about the activity of the
-sample in a new time bin, are in principle conditional on all experimental
-data and statistics we have. But it can be the case that discarding part of
-the data or statistics -- for example, the measured third- and higher-order
-moments -- leaves our probabilities almost unchanged. This means that the
-discarded statistics are \emph{informationally irrelevant} or almost so.
-The remaining statistics -- for example, first and second moments -- are
+intuitive, and is connected with maximum-entropy distributions. Its idea is
+as follows. Our probabilities about the frequencies of the activities of
+the sample, or about the activity of the sample in a new time bin, are in
+principle conditional on all experimental data and statistics we have. But
+it can be the case that discarding part of the data or statistics -- for
+example, the measured third- and higher-order moments -- leaves our
+probabilities almost unchanged. This means that the discarded statistics
+are \emph{informationally irrelevant} or almost so. The remaining
+statistics -- for example, first and second moments -- are
 \emph{informationally sufficient}. \parentext{For more technical results
@@
 % it may give clues about network organization and dynamics.
+
+There's a tight connection between informational sufficiency and
+maximum-entropy distributions
+\citep{jaynes1982b}[\sect~4.5.4]{bernardoetal1994}: if a probability
+distribution for repetitive phenomena has a sufficient statistics, then by
+the Pitman-Koopman theorem \citep{koopman1936,pitman1936,darmois1935}[for
+later analyses and the discrete case
+see][]{hipp1974,andersen1970,denny1967,denny1972,fraser1963,barankinetal1963,barndorffnielsen1978_r2014}
+it is a mixture of exponential distributions of maximum-entropy form.
 
@@
 actually equivalent to comparing the probabilities of the two hypotheses
-about sufficiency, conditional the \emph{full} data, assuming their
+about sufficiency, conditional on the \emph{full} data, assuming their
 pre-data probabilities to be equal. We show this equivalence below and
 perform the calculations for our first data set.
-
 \medskip
@@
 
+\bigskip
+
+Let us illustrate the remarks above with our first data set.
+
+A natural procedure to compare the informational sufficiency of two
+different sets of moments $M'\defd\set{1,\dotsc,m'}$ and
+$M''\defd \set{1,\dotsc,m''}$ is as follows:
+\begin{enumerate*}[label=(\roman*)]
+\item calculate the relative entropies of the measured frequencies $\yf$
+  and the sample marginals of the maximum-entropy distributions built with
+  the two corresponding sets of constraints;
+\item multiply these relative entropies by the number of time bins of the
+  recording;
+\item take their difference.
+\end{enumerate*}
+In formulae,
+\begin{multline}
+  \label{eq:measure_suffic}
+T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{
+  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M')}
+-
+T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{
+  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')}\\
+{}\equiv
+T  \sum_{\ya} \yff_{\ya} \log\frac{
+  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')}{
+  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M')}.
+\end{multline}
+
+This is a natural measure because
+
 \begin{figure}[!p]
@@
 %
-% relentropies of sample2mom, marg2mom, sample4mom, marg4mom, marg6mom
-% -322.72, -85.7627, -42.8005, -4.23746, -4.26796 nats =
-% -1401.56, -372.462, -185.88, -18.403, -18.5355 dB
-% correspond to probability ratios of
-% 1., 8.11718\times 10^{102}, 3.69524\times 10^{121}, 2.06704\times 10^{138},
-% 2.00495\times 10^{138}
+% relentropies of sample2mom, sample4mom, sample5mom, marg2mom, marg4mom, marg5mom
+% -322.72,-42.8005,-39.4917,-85.7626,-4.23725,-4.20059 nats =
+% -1401.56,-185.88,-171.51,-372.462,-18.4021,-18.2429 dB
 %
@@
 % sample level: 279.92 nats = 1215.68 dB
-% marg level: 81.5252 nats =  354.059 dB
-% difference in relentropies between 6 mom and 4 mom:
-% marg level: -0.0305024 nats = -0.13247 dB
+% marg level: 81.525 nats =  354.06 dB
+% difference in relentropies between 5 mom and 4 mom:
+% sample level: 3.30882 nats = 14.37 dB
+% marg level: 0.0366623 nats = 0.159223 dB
 % correspond to prob ratios of
-% 3.69524\times 10^{121},2.5465\times 10^{35}, 0.969958
+% sample: 3.69524\times 10^{121}, 27.3529
+% marg: 2.5469\times 10^{35}, 1.03734
 % 
@@
   \binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}
-  \defs \ypp_{\ya\yA},
+  \defs \yG_{\ya\yA},
 \end{equation}
@@
   \sum_{\yA}
-  \ypp_{\ya\yA}\,
+  \yG_{\ya\yA}\,
   \pf(\yA \| \yH).
@@
 The conditional probability
-$\pf(\ya \|\yA, \yH) \equiv \ypp_{\ya \yA}$, besides
+$\pf(\ya \|\yA, \yH) \equiv \yG_{\ya \yA}$, besides
 relating the distributions for the network and sample activities via
@@
   =
-  \sum_{\ya} \yg_{\ya} \sum_{\yA} \ypp_{\ya\yA}\,\pf(\yA \| \yH)
+  \sum_{\ya} \yg_{\ya} \sum_{\yA} \yG_{\ya\yA}\,\pf(\yA \| \yH)
   =%{}\\
-  \sum_{\yA} \bigl( \tsum_{\ya} \yg_{\ya}  \ypp_{\ya\yA} \bigr)\,
+  \sum_{\yA} \bigl( \tsum_{\ya} \yg_{\ya}  \yG_{\ya\yA} \bigr)\,
   \pf(\yA \| \yH),
 %  =
-%  \sum_{N\yA=0}^N \sum_{n\ya=0}^n f(\ya)\ypp_{\ya\yA}\,\p(\yXf=\yA \| \yH)
+%  \sum_{N\yA=0}^N \sum_{n\ya=0}^n f(\ya)\yG_{\ya\yA}\,\p(\yXf=\yA \| \yH)
 %  \equiv  \expeb{\yg^*}.
@@
 the expectation of the function
-$\yg^*(\yA) \defd \sum_{\ya} \yg_{\ya}\,\ypp_{\ya\yA}$.
+$\yg^*(\yA) \defd \sum_{\ya} \yg_{\ya}\,\yG_{\ya\yA}$.
 
@@
   \label{eq:constraint_extended}
-  \yc = \E(\yg \| \yH) \equiv \sum_{\yA} \bigl( \tsum_{\ya} \yg_{\ya}  \ypp_{\ya\yA} \bigr)\;
+  \yc = \E(\yg \| \yH) \equiv \sum_{\yA} \bigl( \tsum_{\ya} \yg_{\ya}  \yG_{\ya\yA} \bigr)\;
   \pf(\yA \| \yH).
@@
 %  \label{eq:pullback_m_expectations}
-%  \ypp^*\colon \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}
+%  \yG^*\colon \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}
 %  \mapsto
@@
   $%\pf(\ya \|\yA, \yH)=
-\ypp_{\ya\yA} \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
+\yG_{\ya\yA} \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
   colourbar depending on your \textsc{pdf} viewer.)}
@@
 formalize and mathematically express our intuition. The hypergeometric
-distribution $\ypp_{\ya\yA}$ plays an important role in this
+distribution $\yG_{\ya\yA}$ plays an important role in this
 formalization. A look at its plot, \fig~\ref{fig:hypergeom_proj}, reveals
@@
   sample is representative of our uncertainty about the network and
-  about other samples}, and vice versa. When $n=\yNv$, $\ypp_{\ya\yA}$
+  about other samples}, and vice versa. When $n=\yNv$, $\yG_{\ya\yA}$
 becomes the identity matrix and the approximate equalities above become
@@
 the sample average $\ya$, owing to the coarsening effect of
-$\ypp_{\ya\yA}$.
+$\yG_{\ya\yA}$.
 
@@
   \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
-  \tsum_{\ya} \yg_{\ya}\ypp_{\ya\yA}\bigr).
+  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
 \end{equation}
@@
   \yc = \E(\yg \|\yHa)\equiv %{}\\
- \frac{1}{\yK(\yll)}\sum_{\ya} \sum_{\yA} \yg_{\ya} \ypp_{\ya \yA}
+ \frac{1}{\yK(\yll)}\sum_{\ya} \sum_{\yA} \yg_{\ya} \yG_{\ya \yA}
 \,
   \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
-  \tsum_{\ya} \yg_{\ya}\ypp_{\ya\yA}\bigr).
+  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
 %  \\ k=1,\dotsc,m.
@@
   \pf(\ya \| \yHa)  = \frac{1}{\yK(\yll)}\, 
-  \sum_{\yA} \ypp_{\ya \yA}
+  \sum_{\yA} \yG_{\ya \yA}
 \,
   \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
-  \tsum_{\ya} \yg_{\ya}\ypp_{\ya\yA}\bigr).
+  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
 \end{equation}
@@
   \sum_\yNv \biggl\{\frac{1}{\yK(\yll_\yNv)}\, 
-  \sum_{\yA} \ypp^{(\yNv)}_{\ya\yA}
+  \sum_{\yA} \yG^{(\yNv)}_{\ya\yA}
 \,
   \tbinom{\yNv}{\yNv \yA}\,\exp\Bigl[{\yll_\yNv}\T
-  \tsum_{\ya} \yg_{\ya}\ypp^{(\yNv)}_{\ya\yA}\Bigr]\biggr\}
+  \tsum_{\ya} \yg_{\ya}\yG^{(\yNv)}_{\ya\yA}\Bigr]\biggr\}
   \;h(\yNv),
@@
   \biggl[\prod_{\ya\yA}
-(\ypp_{\ya\yA}\ynuu_{\yA})^{T\yjj_{\ya\yA}}
+(\yG_{\ya\yA}\ynuu_{\yA})^{T\yjj_{\ya\yA}}
 \biggr]\;
@@
 \\
-&\appropto \delt(\tsum_{\yA}\ypp_{\ya\yA}\yFF_{\yA} = \yff_{\ya})\;
+&\appropto \delt(\tsum_{\yA}\yG_{\ya\yA}\yFF_{\yA} = \yff_{\ya})\;
 \exp[-L\;\re(\yF;\yR)]
@@
   \biggl[\prod_{\ya}
-(\tsum_{\yA}\ypp_{\ya\yA}\ynuu_{\yA})^{T\yff_{\ya}}
+(\tsum_{\yA}\yG_{\ya\yA}\ynuu_{\yA})^{T\yff_{\ya}}
 \biggr]\;
@@
 &\appropto 
-\exp[-T\;\re(\yf; \mathte{\ypp}\ynu)
+\exp[-T\;\re(\yf; \mathte{\yG}\ynu)
 -L\;\re(\ynu;\yR)]
@@
 \int\!\di\ynu\;\ynuu_{\yA'}\;
-\exp[-T\;\re(\yf; \mathte{\ypp}\ynu)
+\exp[-T\;\re(\yf; \mathte{\yG}\ynu)
 -L\;\re(\ynu;\yR)]

### 2019-11-16 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-16T18:08:44+0100
+%% Last-Updated: 2019-11-16T19:24:32+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \label{eq:measure_suffic}
-T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{
+  \text{difference in informational sufficiency between $M'$ and $M''$} \defd{}\\
+T  \sum_{\ya} \yff_{\ya} \ln\frac{\yff_{\ya}}{
   \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M')}
 -
-T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{
-  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')}\\
-{}\equiv
-T  \sum_{\ya} \yff_{\ya} \log\frac{
+T  \sum_{\ya} \yff_{\ya} \ln\frac{\yff_{\ya}}{
+  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')} \equiv{}\\
+T  \sum_{\ya} \yff_{\ya} \ln\frac{
   \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')}{
@@
 \end{multline}
-
-This is a natural measure because
+Why is this a natural measure? Let's ask how much more probable is the
+sufficiency of one set with respect to the other, conditional on our data
+$\yf$:
+\begin{equation}
+  \label{eq:ratio_prob_suffic}
+  \pf(M'' \| \yf)/
+  \pf(M' \| \yf).
+\end{equation}
+Now, the probability of observing activity $\ya$ in the sample at any time
+bin is the sample marginal of the maximum-entropy distribution for the full
+network, owing to the excheangeability assumption implicit in the
+maximum-entropy method:
+\begin{equation}
+  \label{eq:sample_marginal}
+  \pxx(\ya_{t} \| M) = \sum_{\yA}\yG_{\ya_{t}\,\yA}\px(\yA \| M).
+\end{equation}
+The probability of observing one sequence $(\ya_{t})$ with frequencies
+$\yf$ is therefore
+\begin{equation}
+  \label{eq:prob_frequencies}
+  \prod_{t=1}^{T} \pxx(\ya_{t} \| M) \equiv
+  \prod_{\ya=0}^{n} \pxx(\ya \| M)^{T\,\yff_{\ya}} \equiv
+  \prod_{\ya=0}^{n}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{T\,\yff_{\ya}}.
+\end{equation}
+The probability of observing the frequencies $\yf$ is obtained multiplying this
+by their multiplicity factor, the multinomial coefficient
+\begin{equation}
+  \label{eq:multiplicity_factor}
+  \binom{T}{T\yf} \defd \frac{T!}{\prod_{\ya}(T\,\yff_{\ya})!}
+  \approx \prod_{\ya}{\yff_{\ya}}^{-T\,\yff_{\ya}},
+\end{equation}
+the last expression coming from Stirling's approximation
+\citep[Lemma~2.2]{csiszaretal2004b}.
+If we assign equal pre-data probabilities to the two hypotheses $M'$ and $M''$,
+each probability in the ratio~\eqref{eq:ratio_prob_suffic} then becomes, by
+Bayes's theorem,
+\begin{multline}
+  \label{eq:prob_proportional_relentropy}
+  \pf(M \| \yf) \propto
+  \pf(\yf \| M) \times\text{\small const} \propto
+  \binom{T}{T\yf}\,  \prod_{\ya}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{
+    T\,\yff_{\ya}}
+   \approx{}\\
+  \prod_{\ya}{\yff_{\ya}}^{-T\,\yff_{\ya}}
+  \times
+  \prod_{\ya}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{
+    T\,\yff_{\ya}}.
+\end{multline}
+The logarithm of the probability above is easily seen to be the number of
+bins $T$ multiplied by relative entropy between the frequency distribution
+$\yf$ and the sample marginal of the maximum-entropy distribution.
+
+Thus, the difference~\eqref{eq:measure_suffic} is the logarithm of the
+probability ratio~\eqref{eq:ratio_prob_suffic}. The exponential of the
+difference~\eqref{eq:measure_suffic} tells us how much more probable is the
+set $M''$ to be sufficient than the set $M'$.
 
@@
 %
-
- This procedure is equivalent to calculating and comparing the
-probabilities of our hypotheses about sufficiency, conditional the data, if
-their pre-data probability to be equal; so that for example
-\begin{equation}
-  \label{eq:compare_hypoth_conditional}
-  \pf(\text{\small\enquote{first two moments are sufficient}} \| \text{\small data}) \propto
-  \pf(\text{\small data} \|
-  \text{\small\enquote{first two moments are sufficient}}).
-\end{equation}
 

### 2019-11-16 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-16T19:24:32+0100
+%% Last-Updated: 2019-11-17T00:27:31+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\yL}{L}
-\newcommand*{\yd}{\bm{\varDelta}}
-\newcommand*{\yds}{\delta}
+%\newcommand*{\yd}{\bm{\varDelta}}
+\newcommand*{\ydi}{\varDelta}
 
@@
 
-A natural procedure to compare the informational sufficiency of two
-different sets of moments $M'\defd\set{1,\dotsc,m'}$ and
-$M''\defd \set{1,\dotsc,m''}$ is as follows:
-\begin{enumerate*}[label=(\roman*)]
-\item calculate the relative entropies of the measured frequencies $\yf$
-  and the sample marginals of the maximum-entropy distributions built with
-  the two corresponding sets of constraints;
-\item multiply these relative entropies by the number of time bins of the
-  recording;
-\item take their difference.
-\end{enumerate*}
-In formulae,
+We measure the difference $\ydi(M'',M')$ in informational sufficiency
+between a set of moments, say $M''\defd\set{1,\dotsc,m''}$, and another, say
+$M'\defd \set{1,\dotsc,m'}$, as follows:
+\begin{enumerate}[label=(\roman*)]
+\item from each maximum-entropy distributions $\px(\yA \| M)$ for the full
+  network, built from each set of constraints $M=M',M''$, calculate the
+  marginal distribution for the sample:
+\begin{equation}
+  \label{eq:sample_marginal}
+  \pxx(\ya \| M) = \sum_{\yA}\yG_{\ya \yA}\px(\yA \| M),
+  \qquad M=M',M'';
+\end{equation}
+\item calculate the relative entropies of the measured frequency
+  distribution $\yf$ with respect to each sample marginal, and multiply
+  them by the number of time bins $T$:
+  \begin{equation}
+    \label{eq:relentropy_T}
+    T\,\sh[\yf; \pxx(\ya \| M)] \defd  T  \sum_{\ya} \yff_{\ya} \ln\frac{\yff_{\ya}}{
+      \pxx(\ya \| M)},
+    \qquad M=M',M'';    
+  \end{equation}
+\item take the difference: 
 \begin{multline}
   \label{eq:measure_suffic}
-  \text{difference in informational sufficiency between $M'$ and $M''$} \defd{}\\
-T  \sum_{\ya} \yff_{\ya} \ln\frac{\yff_{\ya}}{
-  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M')}
+  \ydi(M'',M') \defd
+  T \,\sh[\yf; \pxx(\ya \| M')]
   -
-T  \sum_{\ya} \yff_{\ya} \ln\frac{\yff_{\ya}}{
-  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')} \equiv{}\\
+  T \,\sh[\yf; \pxx(\ya \| M'')]
+  \equiv{}\\
 T  \sum_{\ya} \yff_{\ya} \ln\frac{
@@
 \end{multline}
-Why is this a natural measure? Let's ask how much more probable is the
-sufficiency of one set with respect to the other, conditional on our data
-$\yf$:
-\begin{equation}
-  \label{eq:ratio_prob_suffic}
-  \pf(M'' \| \yf)/
-  \pf(M' \| \yf).
-\end{equation}
-Now, the probability of observing activity $\ya$ in the sample at any time
-bin is the sample marginal of the maximum-entropy distribution for the full
-network, owing to the excheangeability assumption implicit in the
-maximum-entropy method:
-\begin{equation}
-  \label{eq:sample_marginal}
-  \pxx(\ya_{t} \| M) = \sum_{\yA}\yG_{\ya_{t}\,\yA}\px(\yA \| M).
-\end{equation}
-The probability of observing one sequence $(\ya_{t})$ with frequencies
-$\yf$ is therefore
-\begin{equation}
-  \label{eq:prob_frequencies}
-  \prod_{t=1}^{T} \pxx(\ya_{t} \| M) \equiv
-  \prod_{\ya=0}^{n} \pxx(\ya \| M)^{T\,\yff_{\ya}} \equiv
-  \prod_{\ya=0}^{n}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{T\,\yff_{\ya}}.
-\end{equation}
-The probability of observing the frequencies $\yf$ is obtained multiplying this
-by their multiplicity factor, the multinomial coefficient
-\begin{equation}
-  \label{eq:multiplicity_factor}
-  \binom{T}{T\yf} \defd \frac{T!}{\prod_{\ya}(T\,\yff_{\ya})!}
-  \approx \prod_{\ya}{\yff_{\ya}}^{-T\,\yff_{\ya}},
-\end{equation}
-the last expression coming from Stirling's approximation
-\citep[Lemma~2.2]{csiszaretal2004b}.
-If we assign equal pre-data probabilities to the two hypotheses $M'$ and $M''$,
-each probability in the ratio~\eqref{eq:ratio_prob_suffic} then becomes, by
-Bayes's theorem,
-\begin{multline}
-  \label{eq:prob_proportional_relentropy}
-  \pf(M \| \yf) \propto
-  \pf(\yf \| M) \times\text{\small const} \propto
-  \binom{T}{T\yf}\,  \prod_{\ya}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{
-    T\,\yff_{\ya}}
-   \approx{}\\
-  \prod_{\ya}{\yff_{\ya}}^{-T\,\yff_{\ya}}
-  \times
-  \prod_{\ya}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{
-    T\,\yff_{\ya}}.
-\end{multline}
-The logarithm of the probability above is easily seen to be the number of
-bins $T$ multiplied by relative entropy between the frequency distribution
-$\yf$ and the sample marginal of the maximum-entropy distribution.
-
-Thus, the difference~\eqref{eq:measure_suffic} is the logarithm of the
-probability ratio~\eqref{eq:ratio_prob_suffic}. The exponential of the
-difference~\eqref{eq:measure_suffic} tells us how much more probable is the
-set $M''$ to be sufficient than the set $M'$.
+\end{enumerate}
+The measure $\ydi(M'',M')$ so defined is positive if $M''$ is \enquote{more
+  informationally sufficient} than $M'$.
+
+Why is this a natural measure? Because, if we assign equal pre-data
+probabilities to the two hypotheses $M''$ and $M'$, then the exponential of
+$\ydi(M'',M')$ is approximately equal to the ratio of their probabilities
+conditional on the data $\yf$ -- their Bayes factor:
+\begin{equation}
+  \label{eq:bayes_factor}
+  \ydi(M'',M') \approx
+  \ln\bigl[\pf(M'' \| \yf)/
+  \pf(M' \| \yf)\bigr].
+\end{equation}
+So the exponential of $\ydi(M'',M')$ tells us how much more probable $M''$
+is than $M'$, in view of the data $\yf$. We prove this in appendix**
 
@@
 % \appendix
+
+Let's ask how much more probable is the
+sufficiency of one set with respect to the other, conditional on our data
+$\yf$:
+\begin{equation}
+  \label{eq:ratio_prob_suffic}
+  \pf(M'' \| \yf)/
+  \pf(M' \| \yf).
+\end{equation}
+Now, the probability of observing activity $\ya$ in the sample at any time
+bin is the sample marginal of the maximum-entropy distribution for the full
+network, owing to the excheangeability assumption implicit in the
+maximum-entropy method:
+\begin{equation}
+  \label{eq:sample_marginal}
+  \pxx(\ya_{t} \| M) = \sum_{\yA}\yG_{\ya_{t}\,\yA}\px(\yA \| M).
+\end{equation}
+The probability of observing one sequence $(\ya_{t})$ with frequencies
+$\yf$ is therefore
+\begin{equation}
+  \label{eq:prob_frequencies}
+  \prod_{t=1}^{T} \pxx(\ya_{t} \| M) \equiv
+  \prod_{\ya=0}^{n} \pxx(\ya \| M)^{T\,\yff_{\ya}} \equiv
+  \prod_{\ya=0}^{n}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{T\,\yff_{\ya}}.
+\end{equation}
+The probability of observing the frequencies $\yf$ is obtained multiplying this
+by their multiplicity factor, the multinomial coefficient
+\begin{equation}
+  \label{eq:multiplicity_factor}
+  \binom{T}{T\yf} \defd \frac{T!}{\prod_{\ya}(T\,\yff_{\ya})!}
+  \approx \prod_{\ya}{\yff_{\ya}}^{-T\,\yff_{\ya}},
+\end{equation}
+the last expression coming from Stirling's approximation
+\citep[Lemma~2.2]{csiszaretal2004b}.
+If we assign equal pre-data probabilities to the two hypotheses $M'$ and $M''$,
+each probability in the ratio~\eqref{eq:ratio_prob_suffic} then becomes, by
+Bayes's theorem,
+\begin{multline}
+  \label{eq:prob_proportional_relentropy}
+  \pf(M \| \yf) \propto
+  \pf(\yf \| M) \times\text{\small const} \propto
+  \binom{T}{T\yf}\,  \prod_{\ya}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{
+    T\,\yff_{\ya}}
+   \approx{}\\
+  \prod_{\ya}{\yff_{\ya}}^{-T\,\yff_{\ya}}
+  \times
+  \prod_{\ya}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{
+    T\,\yff_{\ya}}.
+\end{multline}
+The logarithm of the probability above is easily seen to be the number of
+bins $T$ multiplied by relative entropy between the frequency distribution
+$\yf$ and the sample marginal of the maximum-entropy distribution.
+
+Thus, the difference~\eqref{eq:measure_suffic} is the logarithm of the
+probability ratio~\eqref{eq:ratio_prob_suffic}. The exponential of the
+difference~\eqref{eq:measure_suffic} tells us how much more probable is the
+set $M''$ to be sufficient than the set $M'$.
 

### 2019-11-17 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-17T00:27:31+0100
+%% Last-Updated: 2019-11-17T15:08:12+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\sh}{\mathrm{\varEta}}
+\newcommand*{\nat}{\textrm{nat}}
+\newcommand*{\hart}{\textrm{Hart}}
 %%% Custom macros end @@@
@@
     \label{eq:relentropy_T}
-    T\,\sh[\yf; \pxx(\ya \| M)] \defd  T  \sum_{\ya} \yff_{\ya} \ln\frac{\yff_{\ya}}{
+    T\,\sh[\yf; \pxx(\ya \| M)] \defd  T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{
       \pxx(\ya \| M)},
@@
   \equiv{}\\
-T  \sum_{\ya} \yff_{\ya} \ln\frac{
+T  \sum_{\ya} \yff_{\ya} \log\frac{
   \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')}{
@@
 
-Why is this a natural measure? Because, if we assign equal pre-data
-probabilities to the two hypotheses $M''$ and $M'$, then the exponential of
-$\ydi(M'',M')$ is approximately equal to the ratio of their probabilities
-conditional on the data $\yf$ -- their Bayes factor:
+Why is this a natural measure? Because $\ydi(M'',M')$ is equal to the
+log-ratio of the likelihoods of the hypotheses $M''$ and $M'$ in view of
+the data $\yf$:
 \begin{equation}
   \label{eq:bayes_factor}
-  \ydi(M'',M') \approx
-  \ln\bigl[\pf(M'' \| \yf)/
-  \pf(M' \| \yf)\bigr].
-\end{equation}
-So the exponential of $\ydi(M'',M')$ tells us how much more probable $M''$
-is than $M'$, in view of the data $\yf$. We prove this in appendix**
+  \ydi(M'',M') =
+  \log\bigl[\pf(\yf \| M'')/
+  \pf(\yf \| M')\bigr].
+\end{equation}
+This is called their \emph{relative weight of evidence}, the logarithm of
+their \emph{relative Bayes factor}
+\citep[\chap~6]{good1950}{good1975,good1981,good1985,good1983}[\sect~1.4]{osteyeeetal1974}{mackay1992,kassetal1995}[see
+also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
+this in appendix**. So the exponential (same base) of $\ydi(M'',M')$ tells
+us how much more probable the data $\yf$ are, in view of $M''$ vs $M'$. If
+we also introduce the pre-data probabilities of the two hypotheses we
+obtain their probabilities conditional on the data.
 
@@
 % 
+% comparison N vs n:
+% 2 mom: 236.958 nats = 1029.09 dB, 8.11758\times 10^{102}
+% 4 mom: 38.5632 nats = 167.478 dB, 5.59496*10^16
+% 5 mom: 35.2911 nats = 153.267 dB, 2.12185*10^15
+%
 % comparison_ME_sample_full_levels.nb
 %
+
+\bigskip
+
+Consider the three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$ consisting
+in the first two, four, five moments. For a full network of size
+$N=10\,000$, we obtain the following differences:
+\begin{equation}
+  \label{eq:diff_suff_N}
+  \begin{split}
+  \ydi(M_{4},M_{2}) &= 81\;\nat = 35\;\hart,
+\\    \ydi(M_{5},M_{4}) &= 0.037\;\nat = 0.016\;\hart,
+  \end{split}
+\end{equation}
+where the Hartley ($\hart$) denotes base-10 logarithms
+\citep[\sect~C.4]{iso2009}. In words, the measured frequencies of the
+sample activity are 35 orders of magnitude more probable under sufficiency
+of the first four moments are sufficient, than under sufficiency of the
+first two only; but they are about equally probable ($10^{0.016} = 1.04$)
+whether we consider the first five moments to be sufficient or just the
+first four.
+
+Compare the results above with those obtained by applying the method at the
+sample level only:
+\begin{equation}
+  \label{eq:diff_suff_n}
+  \begin{split}
+  \ydi_{\text{sample level}}(M_{4},M_{2}) &= 280\;\nat = 1\,220\;\hart,
+\\    \ydi_{\text{sample level}}(M_{5},M_{4}) &= 3.3\;\nat = 1.4\;\hart,
+\end{split}
+\end{equation}
+that is, the data are 1\,220 orders of magnitude more probable conditional
+on four moments as opposed to two, and $10^{1.4}=25$ times more probable
+conditional on five moments as opposed to five.
+
+The application at the full-network level therefore leads to different
+conclusions about the relative importance of the moments from the
+application at the sample level. It is also possible to use the
+measure~\eqref{eq:measure_suffic} and its probabilistic
+meaninig~\eqref{eq:bayes_factor} to compare the probability of the data
+$\yf$ conditional on the hypothesis $M_{5, N}$ of 5-moment sufficiency at
+the full-network level, $N=10\,000$, versus the hypothesis $M_{5, n}$ of
+5-moment sufficiency at the sample level, $N=n=65$. We obtain
+\begin{equation}
+  \label{eq:diff_suff_n}
+  \ydi(M_{5, N}, M_{5, n}) = 35\;\nat = 15\;\hart,
+\end{equation}
+that is, the data are 15 orders of magnitude more likely under the first
+hypothesis than under the second.
 
@@
 \end{equation}
-where $\re(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\ln(\yFF_{\yA}/\yRR_{\yA})$ is
+where $\re(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\log(\yFF_{\yA}/\yRR_{\yA})$ is
 the relative entropy or discrimination information

### 2019-11-17 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-17T15:08:12+0100
+%% Last-Updated: 2019-11-17T16:12:37+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{figure}[!p]
-\centering
-\includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
-\\[3em]%
-\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
-\caption{***}
-\label{fig:comparisons_marginals}
-\end{figure}
-% comparison_ME_sample_full_levels.nb
+% NUMERICAL RESULTS
 %
@@
 
-But does the application at the full-network level lead to appreciably
-different results from that at the sample level? after all we're interested
-in an approximate informational sufficiency, not in an analytically exact
-one. A correct answer can only be given case by case.
-Figure~\ref{fig:comparisons_marginals} gives a graphical answer to this
-question in the case of our first data set. The upper plot shows the
-maximum-entropy distributions for a full network of $10\,000$ neurons,
-constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
-from four moments (\textcolor{mypurpleblue}{\textbf{------}}). The lower
-plot shows the maximum-entropy distributions for the sample, from two
-moments (\textcolor{mygreen}{$\bm{\triangledown}$}) and from four moments
-(\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained
-by marginalizing the full-network distribution to the sample size, from two
-moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
-moments (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The measured
-frequency distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also
-shown. We note the following:
-\begin{itemize}
-\item The application to the full-network (upper plot) leads to two
-  completely different distributions (even different number of modes);
-  clearly the two sets of statistics are not even approximately equivalent
-  for inferential purposes.
-\item The application at the sample level leads to deceivingly similar
-  distributions (\textcolor{mygreen}{$\bm{\triangledown}$} and
-  \textcolor{mypurpleblue}{$\bm{\square}$}), which can only be distinguished on a
-  logarithmic scale. This could lead to the erroneous conclusion that the
-  two statistics are approximately equivalent.
-\item The difference between marginals from the full-network distribution
-  and the distributions obtained at the sample level (filled vs empty
-  markers) is larger than the difference between different statistics
-  (triangles vs squares).
-\item The marginals of the application to the full network are closer to
-  the measured frequencies than the distributions obtained at the sample
-  level (although this closeness is not a valid criterion for their
-  goodness).
-\end{itemize}
-Therefore, the maximum-entropy application at the sample level and at the
-full-network level lead to different results; the latter is not only more
-meaningful but also superior because it shows clearly the different
-informational sufficiency of the two sets of statistics.
+\bigskip
+
+The conclusions as regards \enquote{cooperativity} or \enquote{interaction}
+that we reach by applying the maximum-entropy method to the full network,
+as proposed here, are not only different from the application at the sample
+level, but also visually clearer. The plot of \fig~\ref{fig:plots_2_4mom},
+showing the two clearly different distributions constrained by two and four
+moments, should be compared with the plot for the distributions obtained at
+the sample level under the same constraints, shown in
+\fig~\ref{fig:comparisons_marginals} as dashed lines with empty red
+circles (\textcolor{myred}{\LARGE$\bm{\mathord{\circ}}$}) and empty blue
+squares (\textcolor{mypurpleblue}{$\bm{\square}$}). The measured
+frequencies are shown as filled black triangles
+(\textcolor{black}{$\bm{\triangledown}$}). It is necessary to use a
+logarithmic scale to see the differences, which appear mainly in the tail.
+\begin{figure}[!b]
+\centering
+% \includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
+% \\[3em]%
+\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
+\caption{***}
+\label{fig:comparisons_marginals}
+\end{figure}
+% comparison_ME_sample_full_levels.nb
+
+The same plot also shows the marginals obtained from applying the method to
+a full network with $N=10\,000$ and the same constraints, that is, the
+sample marginals of \fig~\ref{fig:plots_2_4mom}. The two-moment
+distribution is represented by filled red circles
+(\textcolor{myred}{\LARGE$\bm{\bullet}$}), and the four-moment one by
+filled blue squares (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The
+full-network distribution fits the measured frequencies very well if we
+only keep four moments.
+
+% But does the application at the full-network level lead to appreciably
+% different results from that at the sample level? after all we're interested
+% in an approximate informational sufficiency, not in an analytically exact
+% one. A correct answer can only be given case by case.
+% Figure~\ref{fig:comparisons_marginals} gives a graphical answer to this
+% question in the case of our first data set. The upper plot shows the
+% maximum-entropy distributions for a full network of $10\,000$ neurons,
+% constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
+% from four moments (\textcolor{mypurpleblue}{\textbf{------}}). The lower
+% plot shows the maximum-entropy distributions for the sample, from two
+% moments (\textcolor{mygreen}{$\bm{\triangledown}$}) and from four moments
+% (\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained
+% by marginalizing the full-network distribution to the sample size, from two
+% moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
+% moments (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The measured
+% frequency distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also
+% shown. We note the following:
+% \begin{itemize}
+% \item The application to the full-network (upper plot) leads to two
+%   completely different distributions (even different number of modes);
+%   clearly the two sets of statistics are not even approximately equivalent
+%   for inferential purposes.
+% \item The application at the sample level leads to deceivingly similar
+%   distributions (\textcolor{mygreen}{$\bm{\triangledown}$} and
+%   \textcolor{mypurpleblue}{$\bm{\square}$}), which can only be distinguished on a
+%   logarithmic scale. This could lead to the erroneous conclusion that the
+%   two statistics are approximately equivalent.
+% \item The difference between marginals from the full-network distribution
+%   and the distributions obtained at the sample level (filled vs empty
+%   markers) is larger than the difference between different statistics
+%   (triangles vs squares).
+% \item The marginals of the application to the full network are closer to
+%   the measured frequencies than the distributions obtained at the sample
+%   level (although this closeness is not a valid criterion for their
+%   goodness).
+% \end{itemize}
+% Therefore, the maximum-entropy application at the sample level and at the
+% full-network level lead to different results; the latter is not only more
+% meaningful but also superior because it shows clearly the different
+% informational sufficiency of the two sets of statistics.
 

### 2019-11-17 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-17T16:12:37+0100
+%% Last-Updated: 2019-11-17T17:07:11+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \centering
-\includegraphics[width=\linewidth]{_scripts/paper2_stensola_twomom_distributions10000_prior2.pdf}%
+\includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_twomom_distributions10000_priorc.pdf}%
 \\[3em]%
-\includegraphics[width=\linewidth]{_scripts/paper_riehle_twomom_distributions10000_prior2.pdf}%
+\includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_twomom_distributions10000_prior2.pdf}%
 \caption{***}
@@
 filled blue squares (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The
-full-network distribution fits the measured frequencies very well if we
-only keep four moments.
+full-network distribution fits the measured frequencies
+(\textcolor{black}{$\bm{\triangledown}$}) very well if we constrain only four
+moments.
 

### 2019-11-18 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-17T17:07:11+0100
+%% Last-Updated: 2019-11-18T10:30:55+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Mathematically this impossibility comes from the Pitman-Koopman theorem
-\citep{koopman1936,pitman1936,darmois1935} connecting sufficient statistics
-and the exponential family of distributions, and translates into the
-general impossibility of solving a system of $n$ independent equations in
-$m$ unknowns with more equations than unknowns, $n>m$
-\citep[\sect~3.1]{portamanaetal2015}.
+mentioned above, and translates into the general impossibility of solving a
+system of $n$ independent equations in $m$ unknowns with more equations
+than unknowns, $n>m$ \citep[\sect~3.1]{portamanaetal2015}.
 

### 2019-11-18 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-18T10:30:55+0100
+%% Last-Updated: 2019-11-18T10:37:30+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
-this in appendix**. So the exponential (same base) of $\ydi(M'',M')$ tells
-us how much more probable the data $\yf$ are, in view of $M''$ vs $M'$. If
-we also introduce the pre-data probabilities of the two hypotheses we
-obtain their probabilities conditional on the data.
+this equality in appendix**. So the exponential of $\ydi(M'',M')$ tells
+us how much more probable the data $\yf$ are, conditional on $M''$ versus $M'$. If
+we assign the pre-data probabilities of the two hypotheses, this measure
+also gives us the ratio of their probabilities conditional on the data.
 
@@
 sample activity are 35 orders of magnitude more probable under sufficiency
-of the first four moments are sufficient, than under sufficiency of the
-first two only; but they are about equally probable ($10^{0.016} = 1.04$)
-whether we consider the first five moments to be sufficient or just the
-first four.
+of the first four moments than under sufficiency of the first two only; but
+they are about equally probable ($10^{0.016} = 1.04$) whether we consider
+the first five moments to be sufficient or just the first four.
 
@@
 that is, the data are 1\,220 orders of magnitude more probable conditional
-on four moments as opposed to two, and $10^{1.4}=25$ times more probable
-conditional on five moments as opposed to five.
+on four moments than conditional on two, and $10^{1.4}=25$ times more
+probable conditional on five moments than on four moments.
 
@@
 measure~\eqref{eq:measure_suffic} and its probabilistic
-meaninig~\eqref{eq:bayes_factor} to compare the probability of the data
+meaning~\eqref{eq:bayes_factor} to compare the probability of the data
 $\yf$ conditional on the hypothesis $M_{5, N}$ of 5-moment sufficiency at

### 2019-11-18 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-18T10:37:30+0100
+%% Last-Updated: 2019-11-18T13:35:04+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % within $3\;\mathrm{ms}$,
-see \fig~\ref{fig:plots_6mom}, whereas the frequency distribution for the
+see \fig~\ref{fig:plots_5mom}, whereas the frequency distribution for the
 sample is monotonically decreasing with its maximum at zero activity. The
@@
 
-We first calculate the distribution by using six moments. This number
+We first calculate the distribution by using five moments. This number
 already provides almost as much information as the full frequency
-distribution of the sample, and at the same time illustrates the use of the
-approach in questions of statistical sufficiency.
-Figure~\ref{fig:plots_6mom} shows the resulting densities (that is,
+distribution of the sample (see \sect~\ref{sec:marginalization}).
+Figure~\ref{fig:plots_5mom} shows the resulting densities (that is,
 distribution${}\times N$) for four example values of full-network sizes:
@@
 \centering
-\includegraphics[width=\linewidth]{_scripts/paper_stensola_6mom_distributionsN_prior2.pdf}%
+\includegraphics[width=\linewidth]{_scripts/newpaper_stensola_5mom_distributionsN_priorc.pdf}%
 \\[3em]%
-\includegraphics[width=\linewidth]{_scripts/paper_riehle_6mom_distributionsN_prior2.pdf}%
+\includegraphics[width=\linewidth]{_scripts/newpaper_riehle_5mom_distributionsN_priorc.pdf}%
 \caption{***}
-\label{fig:plots_6mom}
+\label{fig:plots_5mom}
 \end{figure}
@@
 Consider the three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$ consisting
-in the first two, four, five moments. For a full network of size
+of the first two, four, five moments. For a full network of size
 $N=10\,000$, we obtain the following differences:
@@
 
-The conclusions as regards \enquote{cooperativity} or \enquote{interaction}
+The conclusions regarding \enquote{cooperativity} or \enquote{interaction}
 that we reach by applying the maximum-entropy method to the full network,

### 2019-11-18 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-18T13:35:04+0100
+%% Last-Updated: 2019-11-18T20:56:20+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 }
-\newcommand*{\citep}{\parencites}
+\newcommand*{\citep}{\footcites}
 \newcommand*{\citey}{\parencites*}
@@
 statistics -- for example, first and second moments -- are
-\emph{informationally sufficient}. \parentext{For more technical results
+\emph{informationally sufficient}.\footnote{For more technical results
   and connections with the notion of symmetry see \eg\
-  \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,cifarellietal1980,cifarellietal1981,diaconisetal1981,lauritzen1982_r1988,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,lauritzen2007,ayetal2015}.}
+  \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,lauritzen1982_r1988,lauritzen2007,cifarellietal1980,cifarellietal1981,diaconisetal1981,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,ayetal2015}.}
 % This informational quasi-sufficiency is interesting because
@@
 The measure $\ydi(M'',M')$ so defined is positive if $M''$ is \enquote{more
-  informationally sufficient} than $M'$.
+  informationally sufficient} than $M'$, and negative otherwise.
 
@@
 also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
-this equality in appendix**. So the exponential of $\ydi(M'',M')$ tells
-us how much more probable the data $\yf$ are, conditional on $M''$ versus $M'$. If
-we assign the pre-data probabilities of the two hypotheses, this measure
-also gives us the ratio of their probabilities conditional on the data.
+this equality in appendix**. So the exponential of $\ydi(M'',M')$ tells us
+how much more probable the data $\yf$ are conditional on $M''$, than
+conditional on $M'$. We can also combine this measure with pre-data
+probabilities for the two hypotheses to obtain the ratio of their
+probabilities conditional on the data.
 
@@
 
-Consider the three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$ consisting
-of the first two, four, five moments. For a full network of size
-$N=10\,000$, we obtain the following differences:
+In our case consider for example three sets of constraints $M_{2}$,
+$M_{4}$, $M_{5}$, consisting of the first two, four, five moments. For a
+full network of size $N=10\,000$, we obtain the following differences:
 \begin{equation}
@@
 where the Hartley ($\hart$) denotes base-10 logarithms
-\citep[\sect~C.4]{iso2009}. In words, the measured frequencies of the
-sample activity are 35 orders of magnitude more probable under sufficiency
-of the first four moments than under sufficiency of the first two only; but
-they are about equally probable ($10^{0.016} = 1.04$) whether we consider
-the first five moments to be sufficient or just the first four.
+\citep[\sect~C.4]{iso2009}[it was called \enquote{ban} and used by Turing
+and Good in their code-breaking work at Bletchley
+Park:][]{good1985,good1950,good1969}[\sect~4.2]{jaynes1994_r2003}. In
+words, the measured frequencies of the sample activity are 35 orders of
+magnitude more probable under sufficiency of the first four moments than
+under sufficiency of the first two only; but they are about equally
+probable ($10^{0.016} = 1.04$) whether we consider the first five moments
+to be sufficient or just the first four.
 
@@
 
+% results for post-data probabilities of N given data and 5 moments:
+% N=65,1000,2000,5000,10000
+% 1.40169\times 10^{-16}, 0.0439602, 0.139657, 0.251064, 0.297726, 0.267594
+%
+% using 1/N pre-data prob:
+% 1.04099\times 10^{-14}, 0.212211, 0.337085, 0.242394, 0.143722, 0.0645882
+% 
+% maxent_solutions_different_moments_prior2_6moments.nb
+
 The calculations and conclusions presented in the preceding sections depend
@@
 our method is based, and which asserts the equality of the factorial
-moments of the full network and its sample, (or, equivalently, asserting
-that the power moments for the one are linear functions of the power
-moments for the other). This formula is only valid if our beliefs about the
-activity of the sample  $\pxx(\ya)$ and that of the full network
-$\pxx(\yA)$ are related by a hypergeometric distribution:
+moments of the full network and those of the sample. This formula is only
+valid if our beliefs about the activity of the sample $\pxx(\ya)$ and that
+of the full network $\pxx(\yA)$ are related by a hypergeometric
+distribution:
 \begin{equation}
@@
   \binom{n}{\ya}\;\binom{N-n}{\yA-\ya}\;\binom{N}{\yA}^{-1}
-\;\pxx(\yA).
-\end{equation}
-This is a relation of \enquote{drawing without replacement}
+\;\pxx(\yA),
+\end{equation}
+characteristic of \enquote{drawing without replacement}
 \citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.

### 2019-11-18 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-18T20:56:20+0100
+%% Last-Updated: 2019-11-18T22:24:37+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The conclusions regarding \enquote{cooperativity} or \enquote{interaction}
+The conclusions about \enquote{cooperativity} or \enquote{interaction}
 that we reach by applying the maximum-entropy method to the full network,
@@
   \binom{n}{\ya}\;\binom{N-n}{\yA-\ya}\;\binom{N}{\yA}^{-1}
-\;\pxx(\yA),
+\;\pxxx(\yA),
 \end{equation}
@@
 \citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-In other words, $N$ must be such that we equally believe other samples of
+In other words, $N$ is the size of the pool of all neurons that could just
+as well have been recorded by our measuring procedure. \enquote{Just as
+  well} indicates equal degrees of belief. Whether the equality of such
+beliefs is justified, or for which sampling procedures it could be
+justified, is the fundamental, deep question of sampling theory, which we
+cannot discuss here of course \citep[see \eg\ the discussions and
+references in][]{ericson1969,smith1976}.
+
+must be such that we equally believe other samples of
 size $n$ could have been recorded by the electrodes or the probe used. This

### 2019-11-18 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-18T22:24:37+0100
+%% Last-Updated: 2019-11-18T22:37:13+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 cannot discuss here of course \citep[see \eg\ the discussions and
-references in][]{ericson1969,smith1976}.
-
-must be such that we equally believe other samples of
-size $n$ could have been recorded by the electrodes or the probe used. This
-is indeed what we stated in \sect~\ref{sec:model}.
-
-This requirement delimits an area around the probe. Its number of neurons
-depends on the animal species and on the brain region from which the
-recording was made, as the neuron density can be very different. It's
-practically impossible to specify the exact number of neurons, but its
-order of magnitude is enough for making qualitative inferences. As
+references in][]{ericson1969,smith1976}. $N$ therefore depends on factors
+such as: the shape and dimensions of the recording probe; the inaccuracy in
+its insertion, leading for example to slightly different insertion points
+or angles; the density of neurons around the point of insertion of the
+probe.
+
+Its order of magnitude is enough for making qualitative inferences. As
 discussed in the next section, even if the exact number were known, the

### 2019-11-19 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-18T22:37:13+0100
+%% Last-Updated: 2019-11-19T08:53:02+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Why is this a natural measure? Because $\ydi(M'',M')$ is equal to the
-log-ratio of the likelihoods of the hypotheses $M''$ and $M'$ in view of
-the data $\yf$:
+log-ratio of the probabilities of the data $\yf$ conditional on the
+hypotheses $M''$ and $M'$:
 \begin{equation}
@@
 \end{equation}
-that is, the data are 15 orders of magnitude more likely under the first
+that is, the data are 15 orders of magnitude more probable under the first
 hypothesis than under the second.
@@
 justified, is the fundamental, deep question of sampling theory, which we
-cannot discuss here of course \citep[see \eg\ the discussions and
+cannot discuss here of course \citep[see \eg\ the discussions, reviews, and
 references in][]{ericson1969,smith1976}. $N$ therefore depends on factors
-such as: the shape and dimensions of the recording probe; the inaccuracy in
-its insertion, leading for example to slightly different insertion points
-or angles; the density of neurons around the point of insertion of the
-probe.
-
-Its order of magnitude is enough for making qualitative inferences. As
-discussed in the next section, even if the exact number were known, the
-maximum-entropy distribution ought to be interpreted qualitatively.
-
-It would be possible to derive a distribution $\pxx(A)$ for the full
-network based on an unknown $N$, by  expressing our belief distribution
-$\pxx(N)$ about $N$ and using this to marginalize $N$ out:
-\begin{equation}
-  \label{eq:marginalize_N}
-  \pxx(A \| \text{$N$ \small unknown}) = \sum_{N} \px(\yA \| N)\;\pxx(N).
-\end{equation}
-But this would be overkill, owing to the qualitative character of the
-maximum-entropy distribution constructed with exact $N$.
+such as: the shape, dimensions, and technical specifications of the
+recording probe; the inaccuracy in the insertion of the probe, leading for
+example to slightly different insertion points or angles; the density of
+neurons around the probe; and similar. $N$ can only be assessed case by
+case. The plots of \fig~\ref{fig:plots_5mom} suggest that its order of
+might be enough for qualitative inferences. As discussed in the next
+section, even if the exact number $N$ were known, the maximum-entropy
+distribution ought to be interpreted qualitatively or semi-quantitatively.
+
+If our uncertainty about $N$ spans several orders of magnitude, we can of
+course assign a probability to the possible values of $N$ based on our
+background information and the frequency data $\yf$. The procedure is
+similar to that of \sect~\ref{sec:marginalization} and appendix***. The
+probability for the data $\yf$ conditional on a set of constraints $M$ and
+$N$ is the exponential of the relative entropy~\eqref{eq:relentropy_T}:
+\begin{equation}
+  \label{eq:prob_data_from_N}
+  \pf(\yf \| M, N) \approx \exp\bigl\{
+  T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
+\end{equation}
+If our pre-data probability for $N$ is $p(N)$, then by Bayes's theorem
+\begin{equation}
+  \label{eq:prob_N_from_data}
+  \pf(N \| M, \yf) \propto p(N)\;\exp\bigl\{
+  T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
+\end{equation}
+
+Here is a simple illustrative example with our first data set. Suppose our
+uncertainty spans slightly more than an order of magnitude, from $N=1\,000$
+to $N=20\,000$. Divide this range roughly into thirds of order of
+magnitude, considering the values
+$N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming five
+moments $M$ are sufficient,
+formula~\eqref{eq:prob_data_from_N} gives
+\begin{equation}
+  \label{eq:data1_f_from_N}
+  \begin{gathered}
+  \pf(\yf \| N=1\,000, M) =  0.00222,\qquad
+  \pf(\yf \| N=2\,000, M) =  0.00704,\\
+  \pf(\yf \| N=5\,000, M) =  0.0127,\qquad
+  \pf(\yf \| N=10\,000, M) =  0.0150,\\
+  \pf(\yf \| N=20\,000, M) =  0.0135.
+\end{gathered}
+\end{equation}
+Our uncertainty is about a scale factor, so it can be represented by equal
+pre-data probabilities $1/5$ about these partial orders of magnitude. Thus
+from~\eqref{eq:data1_f_from_N} we find 
+\begin{equation}
+  \label{eq:data1_N_from_f}
+  \begin{gathered}
+  \pf(N=1\,000 \| \yf, M) =  0.044,\qquad
+  \pf(N=2\,000 \| \yf, M) =  0.140,\\
+  \pf(N=5\,000 \| \yf, M) =  0.251,\qquad
+  \pf(N=10\,000 \| \yf, M) =  0.298,\\
+  \pf(N=20\,000 \| \yf, M) =  0.267,
+\end{gathered}
+\end{equation}
+which gives a slightly higher probability to $N=10\,000$.
 

### 2019-11-19 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-19T08:53:02+0100
+%% Last-Updated: 2019-11-19T14:47:34+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 case. The plots of \fig~\ref{fig:plots_5mom} suggest that its order of
-might be enough for qualitative inferences. As discussed in the next
-section, even if the exact number $N$ were known, the maximum-entropy
+magnitude might be enough for qualitative inferences. As discussed in the
+next section, even if the exact number $N$ were known, the maximum-entropy
 distribution ought to be interpreted qualitatively or semi-quantitatively.
@@
 probability for the data $\yf$ conditional on a set of constraints $M$ and
-$N$ is the exponential of the relative entropy~\eqref{eq:relentropy_T}:
+$N$ is the exponential of minus the relative entropy~\eqref{eq:relentropy_T}:
 \begin{equation}
   \label{eq:prob_data_from_N}
-  \pf(\yf \| M, N) \approx \exp\bigl\{
-  T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
+  \pf(\yf \| M, N) = \binom{T}{T\yf}\;\prod_{a}\pxx(\ya \| M, N)^{T\,\yff_{a}}
+  \approx \exp\bigl\{  -T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
 \end{equation}
@@
   \label{eq:prob_N_from_data}
-  \pf(N \| M, \yf) \propto p(N)\;\exp\bigl\{
-  T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
+  \pf(N \| M, \yf) \propto
+  p(N)\;\exp\bigl\{-T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
 \end{equation}
@@
 magnitude, considering the values
-$N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming five
-moments $M$ are sufficient,
-formula~\eqref{eq:prob_data_from_N} gives
+$N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming the set $M$
+of first five moments is sufficient, formula~\eqref{eq:prob_data_from_N}
+gives
 \begin{equation}
@@
 \end{equation}
-Our uncertainty is about a scale factor, so it can be represented by equal
-pre-data probabilities $1/5$ about these partial orders of magnitude. Thus
-from~\eqref{eq:data1_f_from_N} we find 
+Since our uncertainty regards a scale factor, it can be represented by
+equal pre-data probabilities $1/5$ about these partial orders of magnitude.
+Thus from~\eqref{eq:data1_f_from_N} we find
 \begin{equation}
@@
 which gives a slightly higher probability to $N=10\,000$.
+
+\bigskip
+
+In this way we can make inferences -- for example about the sufficiency of
+a set of moments, or about the marginal sample distribution -- that take
+into account our uncertainty about $N$. We must make sure to avoid
+circularities, though: for example, we can't assume a set of moments $M$ to
+be sufficient and assess our uncertainty about $N$ conditional on it, and
+then use this uncertainty to assess the sufficiency of $M$. But we can
+approximately assess the sufficiency of a subset of moments $M' \subset M$
+taking into account the uncertainty of $N$ conditional on $M$.
+
+A rigorous assessment would involve a more expensive, full Bayesian
+calculation; but such calculation would make the whole maximum-entropy
+approach superfluous. We discuss this final point in the next section.
 

### 2019-11-19 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-19T14:47:34+0100
+%% Last-Updated: 2019-11-19T16:21:29+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{equation}
+  \label{eq:def_factorial_moment_a}
   % \E\condt[\Bigg]{
@@
 % %\appendixpage
-% \appendix
+\appendix
+
+\section{Derivation of the maximum-entropy distribution}
+\label{sec:derivation_maxent_fullnet}
+
+Here is a summary derivation of the maximum-entropy distribution for the
+full network constrained by a set of factorial moments. For further details
+see Porta Mana \etal\ \citey{portamanaetal2015}.
+
+\bigskip
+
+First of all the sampling relation. We have a set of $N$ units, $\yA$ of
+which have some specific property, and we sample in an unknown way $n$ of
+the $N$ units. The probability that $\ya$ of the $n$ sampled units have that
+property is then given by the hypergeometric distribution
+\begin{equation}
+  \label{eq:conditional_hypergeometric}
+  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
+  \binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}.
+\end{equation}
+typical of \enquote{drawing without replacement}
+\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
+In the following leave $n$, $N$ implicit in the conditional. If we are
+uncertain about the number $\yA$, with belief $P(\yA)$, then by the theorem
+of total probability our belief about $\ya$ is
+\begin{equation}
+  \label{eq:belief_sample_unknown_A}
+  p(\ya)= \sum_{\yA} \yG_{\ya\yA} \;P(\A).
+\end{equation}
+In our case the units are neurons, the set is the full network, and the
+property is their being active in a specific time bin.
+
+From the definition of normalized factorial
+moment~\eqref{eq:def_factorial_moment_a}, the expression for the hypergeometric
+distribution~\eqref{eq:conditional_hypergeometric}, and the
+relation~\eqref{eq:belief_sample_unknown_A} between our beliefs about $\ya$
+and $\yA$, it is possible to prove using some combinatorial juggling the
+equality~\eqref{eq:equal_m_factorialmoments} between the factorial moments
+of sample and full network.
+
+\bigskip
+
+For the construction of a maximum-entropy distribution from generic
+constraints see Jaynes \citey{jaynes1963}[\chap~11]{jaynes1994_r2003}. More
+precisely we use the minimum relative-entropy principle
+\citep{hobsonetal1973,csiszar1985}[\sect~5.2.2]{sivia1996_r2006} with
+respect to a reference distribution. For the solution of the extremization
+problem see Mead \etal\ \citey{meadetal1984} and the extensive text by Fang
+\etal\ \citey{fangetal1997}.
+
+Let us apply the relative-\me\ method \citep{sivia1996_r2006,meadetal1984}
+directly to sampled neurons; denote this approach by $\yHb$. Then we apply
+the method to the full network of neurons, most of which are unsampled;
+denote this approach by $\yHa$.
+
+Applied directly to the sampled neurons, the method yields the distribution
+\begin{equation}
+  \label{eq:app_maxent_sample}
+  \pf(\ya \|\yHb)
+  =\frac{1}{\yk(\yl)}\,
+  \binom{n}{n\ya}
+  \exp[\yl\T \yg_{\ya}]
+\end{equation}
+where $\yk(\yl)$ is a normalization constant. The binomial in front of the
+exponential appears because we must account for the multiplicity by which
+the network-average activity $\ya$ can be realized: $\ya=0$ can be
+realized in only one way (all neurons inactive), $\ya=1/n$ can be realized
+in $n$ ways (one active neuron out of $n$), and so on. This term is analogous to
+the \enquote{density of states} in front of the Boltzmann factor in
+statistical mechanics \citep[\chap~16]{callen1960_r1985}. The $m$ Lagrange
+multipliers $\yl\defd (l_1,\dotsc,l_m)$ must satisfy the $m$ constraint
+equations
+\begin{equation}
+  \label{eq:app_maxent_sample_constraints}
+  \yc = \E(\yg \|\yHb) \equiv
+  \frac{1}{\yk(\yl)}\sum_{\ya}  \yg_{\ya} \binom{n}{n\ya}
+  \exp[\yl\T \yg_{\ya}].%, \qquad k=1,\dotsc,m.
+\end{equation}
+
+Applied to the full network, using the constraint
+expression~\eqref{eq:constraint_extended} derived in the previous section,
+the method yields the distribution for the full-network activity
+\begin{equation}
+  \label{eq:app_maxent_pop}
+  \pf(\yA \| \yHa)  = \frac{1}{\yK(\yll)}\,
+  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
+  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
+\end{equation}
+The $m$ Lagrange multipliers $\yll\defd (\lambda_1,\dotsc,\lambda_m)$ must
+satisfy the $m$ constraint equations
+\begin{equation}
+  \label{eq:app_maxent_pop_constraints}
+  \yc = \E(\yg \|\yHa)\equiv %{}\\
+ \frac{1}{\yK(\yll)}\sum_{\ya} \sum_{\yA} \yg_{\ya} \yG_{\ya \yA}
+\,
+  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
+  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
+%  \\ k=1,\dotsc,m.
+\end{equation}
+
+We obtain the distribution for the sample activity by marginalization, using
+\eqn~\eqref{eq:marginal}:
+\begin{equation}
+  \label{eq:app_maxent_pop_marg}
+  \pf(\ya \| \yHa)  = \frac{1}{\yK(\yll)}\, 
+  \sum_{\yA} \yG_{\ya \yA}
+\,
+  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
+  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
+\end{equation}
+
+The distributions for the sample activity,
+\eqns~\eqref{eq:app_maxent_pop_marg} and \eqref{eq:app_maxent_sample},
+obtained with the two approaches $\yHb$ and $\yHa$, are different. From the
+discussion in the previous section we expect them to be vaguely similar;
+yet they cannot be exactly equal, because their equality would require the
+$2m$ quantities $\yll$ and $\yl$ to satisfy the constraint equations
+\eqref{eq:app_maxent_pop_constraints} and
+\eqref{eq:app_maxent_sample_constraints}, and in addition also the $n$
+equations $\pf(\ya \| \yHa) = \pf(\ya \| \yHb)$,
+$\ya=1/n,\dotsc,1$ (one equation is taken care of by the normalization of
+the distributions). We would have a set of $2m+n$ equations in $2m$
+unknowns.
+
+\section{Measure of informational sufficiency}
+\label{sec:measure_suff}
 

### 2019-11-19 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-19T16:21:29+0100
+%% Last-Updated: 2019-11-19T17:09:08+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\yll}{\bm{\lambda}}
-\newcommand*{\yl}{\bm{l}}
+\newcommand*{\yl}{\lambda}
 \newcommand*{\yk}{z}
@@
 \citey{iso1993,iso2006,iso2006b} standards and Jaynes
-\citey{jaynes1994_r2003} for \dobs. We often simply say \enquote{belief}
-for \enquote{\dob}.
+\citey{jaynes1994_r2003} for \dobs. We use \enquote{\dob},
+\enquote{belief}, and \enquote{probability} interchangeably.
 
@@
   \label{eq:belief_sample_unknown_A}
-  p(\ya)= \sum_{\yA} \yG_{\ya\yA} \;P(\A).
+  p(\ya)= \sum_{\yA} \yG_{\ya\yA} \;P(\yA).
 \end{equation}
@@
 From the definition of normalized factorial
-moment~\eqref{eq:def_factorial_moment_a}, the expression for the hypergeometric
-distribution~\eqref{eq:conditional_hypergeometric}, and the
+moment~\eqref{eq:def_factorial_moment_a}, the expression for the
+hypergeometric distribution~\eqref{eq:conditional_hypergeometric}, and the
 relation~\eqref{eq:belief_sample_unknown_A} between our beliefs about $\ya$
-and $\yA$, it is possible to prove using some combinatorial juggling the
+and $\yA$, and using some combinatorial juggling, one can prove the
 equality~\eqref{eq:equal_m_factorialmoments} between the factorial moments
@@
 For the construction of a maximum-entropy distribution from generic
-constraints see Jaynes \citey{jaynes1963}[\chap~11]{jaynes1994_r2003}. More
-precisely we use the minimum relative-entropy principle
+expectation constraints see Jaynes
+\citey{jaynes1963}[\chap~11]{jaynes1994_r2003}. More precisely we use the
+minimum relative-entropy method
 \citep{hobsonetal1973,csiszar1985}[\sect~5.2.2]{sivia1996_r2006} with
 respect to a reference distribution. For the solution of the extremization
-problem see Mead \etal\ \citey{meadetal1984} and the extensive text by Fang
-\etal\ \citey{fangetal1997}.
-
-Let us apply the relative-\me\ method \citep{sivia1996_r2006,meadetal1984}
-directly to sampled neurons; denote this approach by $\yHb$. Then we apply
-the method to the full network of neurons, most of which are unsampled;
-denote this approach by $\yHa$.
-
-Applied directly to the sampled neurons, the method yields the distribution
-\begin{equation}
-  \label{eq:app_maxent_sample}
-  \pf(\ya \|\yHb)
-  =\frac{1}{\yk(\yl)}\,
-  \binom{n}{n\ya}
-  \exp[\yl\T \yg_{\ya}]
-\end{equation}
-where $\yk(\yl)$ is a normalization constant. The binomial in front of the
-exponential appears because we must account for the multiplicity by which
-the network-average activity $\ya$ can be realized: $\ya=0$ can be
-realized in only one way (all neurons inactive), $\ya=1/n$ can be realized
-in $n$ ways (one active neuron out of $n$), and so on. This term is analogous to
-the \enquote{density of states} in front of the Boltzmann factor in
-statistical mechanics \citep[\chap~16]{callen1960_r1985}. The $m$ Lagrange
-multipliers $\yl\defd (l_1,\dotsc,l_m)$ must satisfy the $m$ constraint
-equations
-\begin{equation}
-  \label{eq:app_maxent_sample_constraints}
-  \yc = \E(\yg \|\yHb) \equiv
-  \frac{1}{\yk(\yl)}\sum_{\ya}  \yg_{\ya} \binom{n}{n\ya}
-  \exp[\yl\T \yg_{\ya}].%, \qquad k=1,\dotsc,m.
-\end{equation}
-
-Applied to the full network, using the constraint
-expression~\eqref{eq:constraint_extended} derived in the previous section,
-the method yields the distribution for the full-network activity
+problem using Lagrangians and Lagrange multipliers see Mead \etal\
+\citey{meadetal1984} and the extensive texts by Fang \etal\
+\citey{fangetal1997} and Boyd \etal\ \citey{boydetal2004_r2009}. For a
+geometric understanding of the extremization and of the relation between
+expectations and multipliers see Porta Mana \citey{portamana2017b}.
+
+The result has the standard exponential-family form
 \begin{equation}
   \label{eq:app_maxent_pop}
-  \pf(\yA \| \yHa)  = \frac{1}{\yK(\yll)}\,
-  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
-  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
-\end{equation}
-The $m$ Lagrange multipliers $\yll\defd (\lambda_1,\dotsc,\lambda_m)$ must
-satisfy the $m$ constraint equations
-\begin{equation}
-  \label{eq:app_maxent_pop_constraints}
-  \yc = \E(\yg \|\yHa)\equiv %{}\\
- \frac{1}{\yK(\yll)}\sum_{\ya} \sum_{\yA} \yg_{\ya} \yG_{\ya \yA}
-\,
-  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
-  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
-%  \\ k=1,\dotsc,m.
-\end{equation}
-
-We obtain the distribution for the sample activity by marginalization, using
-\eqn~\eqref{eq:marginal}:
-\begin{equation}
-  \label{eq:app_maxent_pop_marg}
-  \pf(\ya \| \yHa)  = \frac{1}{\yK(\yll)}\, 
-  \sum_{\yA} \yG_{\ya \yA}
-\,
-  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
-  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
-\end{equation}
-
-The distributions for the sample activity,
-\eqns~\eqref{eq:app_maxent_pop_marg} and \eqref{eq:app_maxent_sample},
-obtained with the two approaches $\yHb$ and $\yHa$, are different. From the
-discussion in the previous section we expect them to be vaguely similar;
-yet they cannot be exactly equal, because their equality would require the
-$2m$ quantities $\yll$ and $\yl$ to satisfy the constraint equations
-\eqref{eq:app_maxent_pop_constraints} and
-\eqref{eq:app_maxent_sample_constraints}, and in addition also the $n$
-equations $\pf(\ya \| \yHa) = \pf(\ya \| \yHb)$,
-$\ya=1/n,\dotsc,1$ (one equation is taken care of by the normalization of
-the distributions). We would have a set of $2m+n$ equations in $2m$
-unknowns.
+  \begin{gathered}
+  \px(\yA)  = \frac{1}{Z(\yll)}\;
+  g(\yA)\;
+  \exp\Biggl[\sum_{m}\yl_{m}\;
+  \binom{\yA}{m}\binom{N}{m}^{-1}
+  \Biggr],
+  \\
+  Z(\yll)\defd \sum_{\yA}g(\yA)\;
+  \exp\Biggl[\sum_{m}\yl_{m}\;
+  \binom{\yA}{m}\binom{N}{m}^{-1}
+  \Biggr],
+\end{gathered}
+\end{equation}
+where $g(\yA)$ is the reference distribution and $\yll\defd (\yl_{m})$ are
+the Lagrange multipliers, satisfying the implicit
+constraint equations~\eqref{eq:final_constraint}:
+\begin{equation}
+  \label{eq:constraint_eq_lambda}
+  \sum_{\yA}
+  \tbinom{\yA}{m}\tbinom{N}{m}^{-1}\;
+  \tfrac{1}{Z(\yll)}\;
+  g(\yA)\;
+  \exp\biggl[\sum_{m}\yl_{m}\;
+  \tbinom{\yA}{m}\tbinom{N}{m}^{-1}
+  \biggr]
+=
+  \sum_{\ya}
+  \tbinom{\ya}{m} \tbinom{n}{m}^{-1} \; \yff_{\ya},\quad
+  m\in M.
+\end{equation}
+
+The reference distribution $g(\yA)$ represents our pre-data beliefs about
+the activity levels $\yA$. We know that the majority of neurons in a brain
+area are rarely simultaneously active within a window of some milliseconds,
+so we could choose a distribution with slightly higher weights on low
+values of $\yA$. On the other hand, considering the number of ways in which
+$\yA$ out of $N$ neurons can be simultaneously active would suggest the
+multiplicity distribution proportional to $\binom{N}{\yA}$. It turns out
+that our results of \sects~\ref{sec:application}--\ref{sec:marginalization}
+are actually quite insensitive to the choice between these two possible
+reference distributions, or even a uniform reference distribution.
 

### 2019-11-20 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-19T17:09:08+0100
+%% Last-Updated: 2019-11-20T11:41:20+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 across sessions? How much total activity should we expect at any time bin
-during a recording? The approach presented here gives probability
-distributions over the possible answers to these questions.
+during a recording? The approach presented here gives a probability
+distribution that approximately answers all these questions.
 
@@
 \begin{enumerate}[label=(\alph*)]
-  \item\label{item:maxent_step} we build a distribution $\px(\yA)$ for the total
-  activity of the full network using the maximum-entropy method;
-  \item\label{item:sample_step} the constrained averages used in the
-  maximum-entropy method for the full network are, in turn, determined via
-  sampling theory from the constrained averages for the sample.
+\item \label{item:sample_step} Using sampling theory we determine the
+  relation between some expected values -- specifically, moments -- for the
+  total activity $\ya$ of the sample and corresponding expected values $M$
+  for the total activity $\yA$ of the full network.
+\item \label{item:maxent_step} Using the maximum-entropy method we build a
+  distribution $\px(\yA \| M, N)$ for the total activity of the full
+  network of $N$ neurons, using as constraints the expected values $M$
+  found in the previous step.
 \end{enumerate}
-Let's discuss these points in detail.
-
-Regarding~\ref{item:maxent_step}, we assume familiarity with the
-maximum-entropy method. We actually use the \emph{minimum-relative-entropy}
-method \citep{hobsonetal1973}, but call it \enquote{maximum-entropy} for
-brevity. It amounts to a pair of prescriptions: choose the distribution,
-among those satisfying specific convex constraints, such as fixed
-expectations, that has minimum relative entropy with respect to a reference
-distribution, often taken to be the uniform one; and judge those
-expectations to be equal to measured averages. We add two remarks about
-this method that are seldom made in the literature. First remark: the
-distribution $\px(\yA)$ given by this method is the zeroth-order
-approximation \citep[in the sense of Laplace's
-method:][\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
-four different distributions for the full network:
-\begin{enumerate}[label=(\roman*)]
-\item the most probable \emph{frequency} distribution for the total
-  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
-\item the \emph{belief} distribution for the value of the total
-  activity at any time bin among those \emph{recorded},
-  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
-\item the most probable \emph{frequency} distribution for the total
-  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
-\item the \emph{belief} distribution for the value of the total
-  activity at a \emph{new} time bin.
-  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
-\end{enumerate}
-The maximum-entropy distribution is thus an approximation of our belief
-distribution about four different quantities. Note that the four
-distributions above numerically differ in higher-order approximations.
-Second remark: the maximum-entropy method based on the Shannon entropy
-implicitly makes some assumptions about the probabilities for the long-run
-frequency distributions
-\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these two
-remarks further in \sect~\ref{sec:assumptions}.
-
-In our case, to apply the maximum-entropy prescriptions to the total
-activity of the full network we need to fix some averages of its belief
-distribution, for example the distribution's moments. But we don't have any
-measured moments for the full network to equate the distribution moments
-to. Here enters point~\ref{item:sample_step}: the probability calculus
+We now discuss the ideas behind these steps more in detail, leaving their
+precise mathematical implementation to
+appendix~\ref{sec:derivation_maxent_fullnet}.
+
+Step~\ref{item:sample_step} is possible because the probability calculus
 gives an exact, linear relation between the first $m$ moments for the full
@@
 \citep[\eqns~(16)]{portamanaetal2015}; the ones determine the others and
-vice versa at every time bin. This relation is a classical result of
-sampling theory
+vice versa at every time bin. This relation holds for any belief
+distribution $\pxxx(\yA)$ for the full-network activity and its marginal
+$\pxx(\ya)$ for the sample activity. It is a classical result of sampling
+theory.
+
+***
+
 \citep[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
@@
 activity.
+
+Regarding~\ref{item:maxent_step}, we assume familiarity with the
+maximum-entropy method. We actually use the \emph{minimum-relative-entropy}
+method \citep{hobsonetal1973}, but call it \enquote{maximum-entropy} for
+brevity. It amounts to a pair of prescriptions: choose the distribution,
+among those satisfying specific convex constraints, such as fixed
+expectations, that has minimum relative entropy with respect to a reference
+distribution, often taken to be the uniform one; and judge those
+expectations to be equal to measured averages. We add two remarks about
+this method that are seldom made in the literature. First remark: the
+distribution $\px(\yA \| M, N)$ given by this method is the zeroth-order
+approximation \citep[\mynote{move to sect}in the sense of Laplace's
+method:][\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
+four different distributions for the full network:
+\begin{enumerate}[label=(\roman*)]
+\item the most probable \emph{frequency} distribution for the total
+  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
+\item the \emph{belief} distribution for the value of the total
+  activity at any time bin among those \emph{recorded},
+  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
+\item the most probable \emph{frequency} distribution for the total
+  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
+\item the \emph{belief} distribution for the value of the total
+  activity at a \emph{new} time bin.
+  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
+\end{enumerate}
+Note that the four distributions above numerically differ in higher-order
+approximations. Second remark: the maximum-entropy method based on the
+Shannon entropy implicitly makes some assumptions about the probabilities
+for the long-run frequency distributions \citep[\mynote{move to
+  sect}]{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
+two remarks further in \sect~\ref{sec:assumptions}.
+
+In our case, to apply the maximum-entropy prescription to the total
+activity of the full network we need to fix some averages of its belief
+distribution, for example the distribution's moments. But we don't have any
+measured moments for the full network to equate the distribution moments
+to.
 

### 2019-11-20 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-20T11:41:20+0100
+%% Last-Updated: 2019-11-20T16:55:01+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{enumerate}[label=(\alph*)]
-\item \label{item:sample_step} Using sampling theory we determine the
+\item\label{item:sample_step} Using sampling theory we determine the
   relation between some expected values -- specifically, moments -- for the
@@
   for the total activity $\yA$ of the full network.
-\item \label{item:maxent_step} Using the maximum-entropy method we build a
+\item\label{item:maxent_step} Using the maximum-entropy method we build a
   distribution $\px(\yA \| M, N)$ for the total activity of the full
@@
 We now discuss the ideas behind these steps more in detail, leaving their
-precise mathematical implementation to
-appendix~\ref{sec:derivation_maxent_fullnet}.
-
-Step~\ref{item:sample_step} is possible because the probability calculus
-gives an exact, linear relation between the first $m$ moments for the full
-network and the first $m$ for the sample
-\citep[\eqns~(16)]{portamanaetal2015}; the ones determine the others and
-vice versa at every time bin. This relation holds for any belief
-distribution $\pxxx(\yA)$ for the full-network activity and its marginal
-$\pxx(\ya)$ for the sample activity. It is a classical result of sampling
-theory.
-
-***
-
-\citep[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
-also][]{whitworth1897}.
-
-Combining this result with the maximum-entropy prescription
-\enquote{moments${}={}$measured moments} for the sample, we have that the
-measured moments for the sample determine the moments for the full network:
+precise mathematical implementation and a more detailed list of references
+to appendix~\ref{sec:derivation_maxent_fullnet}.
+
+\medskip
+
+Step~\ref{item:sample_step}, a result of sampling theory, is just an
+application of the probability calculus, which gives an exact linear
+relation between the first $m$ moments for the full network and the first
+$m$ for the sample \citep[\eqns~(16)]{portamanaetal2015}. The ones
+determine the others and vice versa at every time bin. This relation holds
+for any belief distribution $\pxxx(\yA_{t})$ for the full-network activity
+and its marginal $\pxx(\ya_{t})$ for the sample activity at that bin.
+
+This sampling relation is even more straightforward if instead of power
+moments we use \emph{normalized factorial moments} \citep{potts1953}. The
+$m$th normalized factorial moment of a distribution $\pxx(\ya)$ is
+\begin{equation}
+  \label{eq:def_factorial_moment_a}
+  \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
+  \equiv
+  \sum_{\ya=0}^{n} %m!\,
+  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya),
+  \qquad 1\le m \le n,
+\end{equation}
+% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
+% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
+% new bin.
+that is, the expected number of distinct $m$-tuples of simultaneously
+spiking neurons (within a time-bin's width), normalized by the maximum
+possible number of distinct $m$-tuples. % For
+% example, with $m=2$, if $a=4$ neurons spike in a network of $n=5$, we
+% have $\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons,
+% and the total number of distinct pairs is $\tbinom{5}{2}=10$. The
+% normalized number of spiking pairs is therefore $6/10$.
+Note that the first $m$ factorial moments together provide the same
+information as the first $m$ power moments together, and vice versa: they
+are linearly related because $\tbinom{\ya}{m}$ is a polynomial in $\ya$ of
+degree $m$. % Specifying one set is therefore
+% equivalent to specifying the other set.
+But the normalized factorial moments have a convenient property for our
+analysis: \emph{the first $n$ normalized factorial moments for the sample
+  and for the full network are numerically identical}:
+\begin{equation}
+  \label{eq:equal_m_factorialmoments}
+  \begin{split}
+  \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
+  &=
+  \E\Biggl[\binom{\yA}{m} \Biggr]\bigg/\binom{N}{m}
+  \qquad\text{or}\\
+    \sum_{\ya=0}^{n} %m!\,
+  \binom{\ya}{m}\bigg/  \binom{n}{m}\; \pxx(\ya)
+&=    \sum_{\yA=0}^{N} %m!\,
+  \binom{\yA}{m}\bigg/  \binom{N}{m}\; \pxxx(\yA),\qquad
+  % \E\condt[\big]{\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
+  % \E\condt[\big]{\tbinom{\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
+  1\le m \le n.
+\end{split}
+\end{equation}
+% where $\pxxx(\yA)$ is our distribution of belief about the full-network
+% activity, and $\pxx$
+
+\medskip
+
+In step~\ref{item:maxent_step} we actually use the
+\emph{minimum-relative-entropy} method
+\citep{hobsonetal1973,csiszar1985}[\sect~5.2.2]{sivia1996_r2006} with
+respect to a uniform reference distribution. We'll still call it
+\enquote{maximum-entropy} for brevity. It amounts to two prescriptions:
+choose the distribution, among those satisfying specific convex constraints
+such as fixed expectations, that has minimum relative entropy with respect
+to a reference distribution; and judge those expectations to be equal to
+some measured averages, typically time averages.
+
+In our case we don't have measured time averages of the quantity
+$\binom{\yA}{m}$ to equate to the factorial moment
+$\E\bigl[\binom{\yA}{m}\bigr]$ of the full-network distribution
+$\pxxx(\yA)$. But \eqn~\eqref{eq:equal_m_factorialmoments} of
+step~\ref{item:sample_step} comes to our rescue, because it says that the
+expectations for the full network are determined by those for the sample
+$\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time averages of the
+corresponding sample quantity $\binom{\ya}{m}$. So we can combine the two
+steps:
 \begin{equation*}
@@
 \end{equation*}
-
-These two steps are more straightforward if instead of power moments we use
-\emph{normalized factorial moments} \citep{potts1953}. The $m$th normalized
-factorial moment of a distribution $\pxx(\ya)$ for the activity of the
-sample neurons is defined as the average
-\begin{equation}
-  \label{eq:def_factorial_moment_a}
-  % \E\condt[\Bigg]{
-  %   \binom{\ya}{m}\,\binom{n}{m}^{-1}}{\yH}
-  % \equiv
-  \sum_{\ya=0}^{n} %m!\,
-  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya),
-  \qquad 1\le m \le n
-\end{equation}
-% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
-% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
-% new bin.
-This moment can be interpreted as the expectation of the number of distinct
-$m$-tuples of simultaneously spiking neurons (within a bin's time width),
-normalized by the number of distinct $m$-tuples. For example, with $m=2$,
-if $a=4$ neurons spike in a network of $n=5$, we have $\tbinom{4}{2}=6$
-distinct pairs of simultaneously spiking neurons, and the total number of
-distinct pairs is $\tbinom{5}{2}=10$. The normalized number of spiking
-pairs is therefore $6/10$. Note that the first $m$ factorial moments
-provide the same information as the first $m$ power moments and vice versa:
-they linearly determine each other because $\tbinom{\ya}{m}$ is a
-polynomial in $\ya$ of degree $m$. Specifying one set is therefore
-equivalent to specifying the other set. But the normalized factorial
-moments have a convenient property for our analysis
-\citep[\sect~2.3]{portamanaetal2015}{potts1953}: \emph{the first $n$
-  normalized factorial moments for the sample and for the full network are
-  numerically identical}:
-\begin{equation}
-  \label{eq:equal_m_factorialmoments}
-    \sum_{\ya=0}^{n} %m!\,
-  \binom{\ya}{m}\bigg/  \binom{n}{m}\; \pxx(\ya)
-=    \sum_{\yA=0}^{N} %m!\,
-  \binom{\yA}{m}\bigg/  \binom{N}{m}\; \pxxx(\yA),
-  % \E\condt[\big]{\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
-  % \E\condt[\big]{\tbinom{\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
-  \qquad 1\le m \le n,
-\end{equation}
-where $\pxxx(\yA)$ is our distribution of belief about the full-network
-activity.
-
-Regarding~\ref{item:maxent_step}, we assume familiarity with the
-maximum-entropy method. We actually use the \emph{minimum-relative-entropy}
-method \citep{hobsonetal1973}, but call it \enquote{maximum-entropy} for
-brevity. It amounts to a pair of prescriptions: choose the distribution,
-among those satisfying specific convex constraints, such as fixed
-expectations, that has minimum relative entropy with respect to a reference
-distribution, often taken to be the uniform one; and judge those
-expectations to be equal to measured averages. We add two remarks about
-this method that are seldom made in the literature. First remark: the
-distribution $\px(\yA \| M, N)$ given by this method is the zeroth-order
-approximation \citep[\mynote{move to sect}in the sense of Laplace's
+and apply the maximum-entropy method to obtain a distribution
+$\px(\yA \| M, N)$ for the full network of $N$ neurons by constraining some
+of its factorial moments, for example $M=\set{1,\dotsc,m'}$ with $m'\le n$,
+to be equal to the sample's recorded averages. In formulae, the constraints
+on $\pxxx(\yA)$ are
+\begin{equation}
+  \label{eq:final_constraint}
+\underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\bigg/\binom{n}{m}
+  \equiv
+   \sum_{\ya}
+  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured moments}}
+  \mathrel{\;=\;}
+   \underbracket{\sum_{\yA}
+     \binom{\yA}{m}\bigg/  \binom{N}{m} \; \pxxx(\yA)}_{\text{distribution moments}},
+   \quad m\in M.
+\end{equation}
+
+The result is the
+distribution of the form
+\begin{equation}
+  \label{eq:app_maxent_pop_first}
+  \begin{gathered}
+  \px(\yA \| M, N)  = \frac{1}{Z(\yll)}\;
+%  g(\yA)\;
+  \exp\Biggl[\sum_{m}\yl_{m}\;
+  \binom{\yA}{m}\bigg/\binom{N}{m}
+  \Biggr]
+  \\
+\text{with}\qquad  Z(\yll)\defd \sum_{\yA}
+  % g(\yA)\;
+  \exp\Biggl[\sum_{m}\yl_{m}\;
+  \binom{\yA}{m}\bigg/\binom{N}{m}
+  \Biggr],
+\end{gathered}
+\end{equation}
+where the parameters $\yll$ are determined by the
+constraints~\eqref{eq:final_constraint}. It is further discussed in
+appendix~\ref{sec:derivation_maxent_fullnet}.
+
+\bigskip
+
+Before applying the formula above to two concrete data sets we want to add
+two remarks about the maximum-entropy method that are seldom made in the
+neuroscientific literature.
+
+First remark: the distribution $\px(\yA \| M, N)$ given by this method is
+the zeroth-order approximation \citep[\mynote{move to sect}in the sense of
+Laplace's
 method:][\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
@@
 
-In our case, to apply the maximum-entropy prescription to the total
-activity of the full network we need to fix some averages of its belief
-distribution, for example the distribution's moments. But we don't have any
-measured moments for the full network to equate the distribution moments
-to.
-
-We can therefore apply the maximum-entropy method to obtain a distribution
-$\px(\yA)$ for the full network by constraining its $m$th factorial moment
-to be equal to the sample's recorded average of
-$\tbinom{\ya}{m}/\tbinom{n}{m}$, for as many $m$ as we please with
-$1\le m \le n$. In formulae, the constraints on $\px(\yA)$ are
-\begin{equation}
-  \label{eq:final_constraint}
-\underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\bigg/\binom{n}{m}
-  \equiv
-   \sum_{\ya}
-  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured moments}}
-  \mathrel{\;=\;}
-   \underbracket{\sum_{\yA}
-  \binom{\yA}{m}\bigg/  \binom{N}{m} \; \px(\yA)}_{\text{distribution moments}}.
-\end{equation}
-
 Finding the constrained maximum-entropy distribution $\px(\yA)$ amounts to
@@
 appear in the formalism.
+
+***
 
@@
 relation~\eqref{eq:belief_sample_unknown_A} between our beliefs about $\ya$
-and $\yA$, and using some combinatorial juggling, one can prove the
-equality~\eqref{eq:equal_m_factorialmoments} between the factorial moments
-of sample and full network.
+and $\yA$, and using some combinatorial juggling
+\citep[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[appendix~A]{portamanaetal2015}{potts1953}, one
+can prove the equality~\eqref{eq:equal_m_factorialmoments} between the
+factorial moments of sample and full network.
 

### 2019-11-20 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-20T16:55:01+0100
+%% Last-Updated: 2019-11-20T17:26:55+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 two remarks about the maximum-entropy method that are seldom made in the
-neuroscientific literature.
-
-First remark: the distribution $\px(\yA \| M, N)$ given by this method is
-the zeroth-order approximation \citep[\mynote{move to sect}in the sense of
-Laplace's
-method:][\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
-four different distributions for the full network:
+neuroscientific literature. They are important for the interpretation of
+the results and are further discussed in \sect~\ref{sec:assumptions}.
+
+First, the distribution $\px(\yA \| M, N)$ given by this method is the
+zeroth-order approximation (in the sense of Laplace's
+method)\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}
+of four different distributions for the full network, which differ
+numerically from one another in higher-order approximations:
 \begin{enumerate}[label=(\roman*)]
@@
 \end{enumerate}
-Note that the four distributions above numerically differ in higher-order
-approximations. Second remark: the maximum-entropy method based on the
-Shannon entropy implicitly makes some assumptions about the probabilities
-for the long-run frequency distributions \citep[\mynote{move to
-  sect}]{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
-two remarks further in \sect~\ref{sec:assumptions}.
-
-Finding the constrained maximum-entropy distribution $\px(\yA)$ amounts to
-a convex optimization
-\citep{meadetal1984}[\chap~10]{pressetal1988_r2007}{fangetal1997,boydetal2004_r2009,portamana2017b}
-and for the numbers $N,n,m$ considered in the present work it can be done
-on a modern computer without approximations of the partition functions that
-appear in the formalism.
-
-***
+
+Second, the maximum-entropy method % (based on the \emph{Shannon} relative
+% entropy, as opposed to other entropy functions such as Burg's
+% \citeyear{burg1975})
+rests on some implicit assumptions about the probabilities for the long-run
+frequency distribution of activities. The often heard statement that it's
+\enquote{the maximally unbiased (or non-committal) distribution} must be
+taken with a grain of salt. \citep[\mynote{move to
+  sect}]{jaynes1986d_r1996,portamana2009,portamana2017}.
 

### 2019-11-20 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-20T17:26:55+0100
+%% Last-Updated: 2019-11-20T20:23:20+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+The number and degrees of the constraining moments depend on the questions
+and hypotheses that a researcher is exploring; for example, hypotheses
+about the \enquote{cooperativity} or \enquote{interaction} in the network
+activity. We discuss this kind of use in \sect~\ref{sec:marginalization}.
+Note that the constraint equation~\eqref{eq:final_constraint} for the
+distribution~\eqref{eq:app_maxent_pop_first} \emph{may not have exact
+  solutions} if $N$ is strictly larger than $n$; and the possible
+discrepancy increases with the number of constraints. The essential reason
+for this discrepancy is the approximate character of two assumptions on
+which the maximum-entropy method is based and of one assumption specific to
+our application: measured moments equal expectations (or equivalently that
+the number of time bins $T$ is infinite), the order of the time bins is
+irrelevant, and that all full-network subsets with $\yA$ neurons are
+equally likely to be generating a total activity $\yA$. We discuss these
+assumptions in \sect~\ref{sec:assumptions}.
+
+%, but our main goal is to simply introduce
+% the general method without entering into biological questions of this
+% kind.
+
 \bigskip
@@
 Before applying the formula above to two concrete data sets we want to add
-two remarks about the maximum-entropy method that are seldom made in the
+three remarks on the maximum-entropy method that are seldom made in the
 neuroscientific literature. They are important for the interpretation of
@@
 
-First, the distribution $\px(\yA \| M, N)$ given by this method is the
+First, a maximum-entropy distribution, like $\px(\yA \| M, N)$, is the
 zeroth-order approximation (in the sense of Laplace's
-method)\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}
+approximation)\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}
 of four different distributions for the full network, which differ

### 2019-11-21 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-20T20:23:20+0100
+%% Last-Updated: 2019-11-21T11:47:50+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Note that the constraint equation~\eqref{eq:final_constraint} for the
-distribution~\eqref{eq:app_maxent_pop_first} \emph{may not have exact
-  solutions} if $N$ is strictly larger than $n$; and the possible
-discrepancy increases with the number of constraints. The essential reason
-for this discrepancy is the approximate character of two assumptions on
-which the maximum-entropy method is based and of one assumption specific to
-our application: measured moments equal expectations (or equivalently that
-the number of time bins $T$ is infinite), the order of the time bins is
-irrelevant, and that all full-network subsets with $\yA$ neurons are
-equally likely to be generating a total activity $\yA$. We discuss these
-assumptions in \sect~\ref{sec:assumptions}.
+distribution~\eqref{eq:app_maxent_pop_first} may not have exact solutions
+if $N$ is strictly larger than $n$; and the possible discrepancy typically
+increases with the number of constraints. The essential reason for this
+discrepancy is the approximate character of two assumptions behind the
+maximum-entropy and of one assumption specific to our application: measured
+moments equal expectations (or equivalently, the number of time bins $T$ is
+infinite), the order of the time bins is irrelevant, and an activity level
+$\yA$ can be equally likely be generated by any set of $\yA$ neurons in the
+full network. The magnitude of this discrepancy can be a signature of the
+(at least temporary) presence of a \enquote{neuronal assembly}
+\citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss this matter
+in \sect~\ref{sec:assumptions}.
 
@@
 Before applying the formula above to two concrete data sets we want to add
-three remarks on the maximum-entropy method that are seldom made in the
+two remarks on the maximum-entropy method that are seldom made in the
 neuroscientific literature. They are important for the interpretation of
 the results and are further discussed in \sect~\ref{sec:assumptions}.
-
 First, a maximum-entropy distribution, like $\px(\yA \| M, N)$, is the
@@
 \end{enumerate}
-
-Second, the maximum-entropy method % (based on the \emph{Shannon} relative
+The validity of the maximum-entropy approximation decreases as the ratio
+$N/T$ increases. Second, the maximum-entropy
+method % (based on the \emph{Shannon} relative
 % entropy, as opposed to other entropy functions such as Burg's
@@
 rests on some implicit assumptions about the probabilities for the long-run
-frequency distribution of activities. The often heard statement that it's
-\enquote{the maximally unbiased (or non-committal) distribution} must be
-taken with a grain of salt. \citep[\mynote{move to
+frequency distribution of activities, besides the assumptions mentioned at
+the end of the previous paragraph. So the often heard statement that it
+gives \enquote{the maximally unbiased (or non-committal) distribution} must
+be taken with a grain of salt. \citep[\mynote{move to
   sect}]{jaynes1986d_r1996,portamana2009,portamana2017}.
-
-\bigskip
-
-The number of moments used with this method depends on the questions and
-hypotheses that a researcher is exploring; for example, hypotheses about
-the \enquote{cooperativity} or \enquote{interaction} -- for example,
-pairwise or higher-order -- of the network activity. We discuss this kind
-of use in
-\sect~\ref{sec:marginalization}.%, but our main goal is to simply introduce
-% the general method without entering into biological questions of this
-% kind.
-
-The particular case in which $n$ moments are constrained is especially
-important: it corresponds to fully constraining the marginal frequency
-distribution for the activity of the sample neurons, $\yf$. In this case,
-our belief about the full-network activity is based on all available
-measured frequency data. Note that the application of the maximum-entropy
-method \emph{at the sample level} is trivial and meaningless in this case
--- it just gives back the measured frequency distribution. But application
-of the method \emph{at the level of the full network} is not trivial.
-
-Using the full frequency distribution of the sample may be a bad idea,
-however, because the maximum-entropy distribution may become a bad
-approximation of some of the four distributions described above. It is
-preferable to use a moderately high number of moments smaller than $n$. We
-explain this point in \sect***.
 
@@
 also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
-this equality in appendix**. So the exponential of $\ydi(M'',M')$ tells us
-how much more probable the data $\yf$ are conditional on $M''$, than
-conditional on $M'$. We can also combine this measure with pre-data
-probabilities for the two hypotheses to obtain the ratio of their
-probabilities conditional on the data.
+this equality in appendix~\ref{sec:measure_suff}. So the exponential of
+$\ydi(M'',M')$ tells us how much more probable the data $\yf$ are
+conditional on $M''$, than conditional on $M'$. We can also combine this
+measure with pre-data probabilities for the two hypotheses to obtain the
+ratio of their probabilities conditional on the data
+\citep[\cf][]{bretthorst2013}.
 

### 2019-11-21 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-21T11:47:50+0100
+%% Last-Updated: 2019-11-21T16:46:09+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
-  \equiv
+  \defd
   \sum_{\ya=0}^{n} %m!\,
@@
 Before applying the formula above to two concrete data sets we want to add
-two remarks on the maximum-entropy method that are seldom made in the
+two remarks about the maximum-entropy method that are seldom made in the
 neuroscientific literature. They are important for the interpretation of
 the results and are further discussed in \sect~\ref{sec:assumptions}.
-First, a maximum-entropy distribution, like $\px(\yA \| M, N)$, is the
+First, a maximum-entropy distribution like $\px(\yA \| M, N)$ is the
 zeroth-order approximation (in the sense of Laplace's
@@
 In the next section we apply the method just described to the data sets
-from two actual recordings, using six moments, and discuss the properties
-of the resulting distributions.
+from two actual recordings and discuss the properties of the resulting
+distributions.
 
@@
 
-We apply the approach just described to two data sets publicly available in
+We apply the method just described to two data sets publicly available in
 the literature:
@@
 already provides almost as much information as the full frequency
-distribution of the sample (see \sect~\ref{sec:marginalization}).
-Figure~\ref{fig:plots_5mom} shows the resulting densities (that is,
-distribution${}\times N$) for four example values of full-network sizes:
-$N=n$, $N=1\,000$, $N=5\,000$, $N=10\,000$. The case $N=n$ corresponds to
-applying the maximum-entropy method at the sample level; it can be observed
-that with six moments it reproduces almost exactly the measured frequency
-distribution. We discuss the case of unknown $N$ in \sect~\ref{sec:N}.
+distribution of the sample (see next section). Figure~\ref{fig:plots_5mom}
+shows the resulting densities (distribution${}\times N$) for three example
+values of full-network sizes: $N=1\,000$, $N=5\,000$, $N=10\,000$. The
+frequency distribution of the sample activity is also shown, and in the
+plot it is indistinguishable from the maximum-entropy distribution for
+$N=n$, that is, applied at the sample level. We discuss the case of unknown
+$N$ in \sect~\ref{sec:N}.
 \begin{figure}[!p]
@@
 \end{figure}
-%maxent_solutions_different_moments_prior2_6moments.nb
+% maxent_solutions_different_moments_prior2_6moments.nb
 
@@
 increases with $N$. Most remarkably, for the first data set the
-distribution has two distinct low-activity modes. For the second data
+distribution has two distinct low-activity modes. For the second data set
 the distribution presents a small shoulder on the right of the
@@
 % different kinds of cells or modules)
-These features are clearly not present in the sample or in the
+These features are clearly not present in the sample frequencies or in the
 maximum-entropy distribution at the sample level. The application of the
@@
 
-\subsection*{Use with moment constraints}
-\label{sec:use_moments}
+\bigskip
 
@@
 % maxent_solutions_different_moments_prior2_6moments.nb
-To illustrate how our approach can be applied to studies of sufficient
-statistics, \fig~\ref{fig:plots_2_4mom} shows the full-network
-distributions, for $N=10\,000$, obtained constraining the first two moments
-(equivalent to constraining means and correlations) and the first four
-moments. The frequency distribution of the sample is also shown for
-comparison. In both data sets the two-moment constraint leads to a
-distribution quite different from the four-moment constraint. For the first
-data set in particular, one distribution is bimodal, the other unimodal.
-This visual difference shows directly -- without the need to calculate
-relative entropies, for example -- that the first two moments alone are
-\emph{not} informationally sufficient for this specific data set.
-
-It is instructive to contrast the conclusions about correlation sufficiency
-reached through this full-network analysis, with those we would have
-reached by applying the maximum-entropy method \emph{at the sample
-  level}, as traditionally done. Such a comparison involves an important
+To illustrate how our approach can be applied to studies of
+\enquote{cooperativity} or \enquote{interaction},
+\fig~\ref{fig:plots_2_4mom} shows the full-network distributions, for
+$N=10\,000$, obtained constraining the first two moments (in
+\textcolor{red}{red}, equivalent to constraining means and correlations)
+and the first four moments (in \textcolor{blue}{blue}). The frequency
+distribution for the sample is also shown for comparison. In both data sets
+the two-moment constraint leads to a distribution quite different from the
+that of the four-moment constraint. For the first data set in particular,
+one distribution is bimodal, the other unimodal. Interpreted in terms of
+This visual difference shows directly that the two sets of moments are not
+even approximately equivalent.
+
+It is instructive to contrast this conclusion with that we would have
+reached by applying the maximum-entropy method \emph{at the sample level},
+as traditionally done. Such a comparison involves an important
 methodological caveat, so we discuss it in full in the next section.

### 2019-11-22 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-21T16:46:09+0100
+%% Last-Updated: 2019-11-22T15:47:19+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 are linearly related because $\tbinom{\ya}{m}$ is a polynomial in $\ya$ of
-degree $m$. % Specifying one set is therefore
+degree $m$. So we'll just say \enquote{first $m$ moments} from now
+on. % Specifying one set is therefore
 % equivalent to specifying the other set.
@@
   $3\;\textrm{ms}$ width.
-\item The \textbf{second data set}, from
-  \textcite{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
+\item The \textbf{second data set}, from \textcite[data courtesy by A.
+  Riehle and T. Brochier]{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
   consists of $n=159$ neurons from macaque Motor Cortex, recorded for about
@@
 
-We first calculate the distribution by using five moments. This number
-already provides almost as much information as the full frequency
-distribution of the sample (see next section). Figure~\ref{fig:plots_5mom}
-shows the resulting densities (distribution${}\times N$) for three example
-values of full-network sizes: $N=1\,000$, $N=5\,000$, $N=10\,000$. The
-frequency distribution of the sample activity is also shown, and in the
-plot it is indistinguishable from the maximum-entropy distribution for
-$N=n$, that is, applied at the sample level. We discuss the case of unknown
-$N$ in \sect~\ref{sec:N}.
+The full-network maximum-entropy distribution is calculated using five
+moments. This number already provides almost as much information as the
+full frequency distribution of the sample (see next section).
+Figure~\ref{fig:plots_5mom} shows the resulting densities
+(distribution${}\times N$) for three example values of full-network sizes:
+$N=1\,000$, $N=5\,000$, $N=10\,000$. The frequency distribution of the
+sample activity is also shown, and in the plot it would be indistinguishable from
+the maximum-entropy distribution for $N=n$, that is, applied at the sample
+level. We discuss the case of unknown $N$ in \sect~\ref{sec:N}.
 \begin{figure}[!p]
@@
 % maxent_solutions_different_moments_prior2_6moments.nb
+
+The most expensive calculation, for $N=10\,000$, takes less than 15 minutes
+on a laptop with two $2\;\textrm{GHz}$ cores. In all cases the moments were
+recovered with relative errors smaller than $10^{-12}$.
+% 1000:  3*10^-13 %
+% 5000: 7*10^-13 %, 581 s
+% 10000: 6*10^-11 %, 737 s
 
@@
   us.}
+
+These results show the potential usefulness of this maximum-entropy
+approach. For example, some researchers could be interested in the
+hypothesis that two distinct cell types or assemblies be present in the
+region where the recording was made. Finding a full-network distribution
+such as that in the upper panel of \fig~\ref{fig:plots_5mom} would provide
+some evidence for this hypothesis. If the researchers also have reasons to
+suspect that a specific set of the sampled neurons are of one type, and the
+remaining of the other type, they could further apply the method to the two
+sets separately. Finding two distributions with a single peak each would
+then provide some evidence that their separation was correct. In the case
+of data set 1, 27--29 sampled neurons were identified as grid cells
+belonging to 3--4 functional modules, see \cite{dunnetal2015}. The two
+peaks in the distribution obtained with our approach could reflect the grid
+vs non-grid distinction, or two groups of modules, or a combination of
+these cases.
+
+The important point is that ***
 

### 2019-11-25 maxNt_v6.tex
update

--- 2019-11-21 maxNt_v6.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-21T16:46:09+0100
+%% Last-Updated: 2019-11-25T17:35:14+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\propertitle}{Maximum-entropy distributions for a 
-  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
+  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
 }
@@
   work shows how to build a maximum-entropy probabilistic model for the
-  total activity of a network of neurons, given only some activity data or
-  statistics -- for example, empirical moments -- of a \emph{subnetwork}
+  total activity of a population of neurons, given only some activity data or
+  statistics -- for example, empirical moments -- of a \emph{subpopulation}
   thereof. This kind of model is useful because neuronal recordings are
-  always limited to a very small sample of a network of neurons.
+  always limited to a very small sample of a population of neurons.
 %
@@
   literature. In some cases it makes interesting forecasts about the larger
-  network -- for example, two low-regime modes in the frequency
+  population -- for example, two low-regime modes in the frequency
   distribution for the total activity -- that are not visible in the sample
@@
   For the two datasets, the maximum-entropy probability model applied only
-  to the subnetwork is compared with the marginal probability
+  to the subpopulation is compared with the marginal probability
   distribution obtained from the maximum-entropy model applied to the full
-  network. On a linear probability scale no large differences are
+  population. On a linear probability scale no large differences are
   visible, but on a logarithmic scale the two distributions show very
@@
 \section{Introduction:\\a simple model for questions about large neuronal
-  networks}
+  populations}
 \label{sec:intro}
@@
 their assumptions, that the recorded neurons are a sample from a larger
-network. This kind of isolation assumptions sometimes escape attention,
+population. This kind of isolation assumptions sometimes escape attention,
 being subtly hidden in the mathematics. Some probabilistic models try to
@@
 
-\mynote{[Yasser:] In the present work we aim to address the issue of building probability distributions over a large network using recording from subnetworks by considering the distribution of total population activity as an example.
+\mynote{[Yasser:] In the present work we aim to address the issue of building probability distributions over a large population using recording from subpopulations by considering the distribution of total population activity as an example.
 [}
@@
 question:\mynote{]} How much can the \emph{total} activity of a large neuronal
-network have been, given the observation of the activity of a very small
+population have been, given the observation of the activity of a very small
 sample thereof? This model addresses a larger brain area, avoiding the
@@
 For each data set the model gives the most plausible frequency distribution
-with which all levels of total activity of a larger network of 1\,000--10\,000
+with which all levels of total activity of a larger population of 1\,000--10\,000
 neurons appeared during the recording. These guessed frequency
@@
 sampled neurons. For the first data set, for example, the frequency
-distribution for the full network has two very distinct modes, both at low
+distribution for the full population has two very distinct modes, both at low
 activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
@@
 still presents one prominent shoulder in its low-activity mode. Note that
-these guessed features of the full network cannot be inferred by the
+these guessed features of the full population cannot be inferred by the
 application of maximum-entropy \emph{to the sample alone}. Although we don't
-formulate any hypotheses on why the larger network could have two distinct
+formulate any hypotheses on why the larger population could have two distinct
 low-activity regimes -- the purpose of the present work is only
@@
 example][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
-But when the difference in size between sample and full network is too
+But when the difference in size between sample and full population is too
 large, correlation sufficiency for the sample implies the \emph{lack} of
-correlation sufficiency for the larger network, and vice versa. Therefore,
+correlation sufficiency for the larger population, and vice versa. Therefore,
 maximum-entropy applications at the sample level can be deceptive for
-questions of statistical sufficiency, and a full-network application is
+questions of statistical sufficiency, and a full-population application is
 more reliable. We discuss this point in detail in \sect~\ref{sec:marginalization}.
@@
 
-How large is the full network addressed by the method proposed here? Its
+How large is the full population addressed by the method proposed here? Its
 size must have some limit and can't obviously include the full brain. The
@@
 our approach is the \emph{actual} one which the activity levels of the full
-network had during the recording; it's only the most plausible. In high
+population had during the recording; it's only the most plausible. In high
 dimensions, however, the features of the most plausible distribution may
 \emph{not} be typical of the majority of most plausible distributions; and
-the set of all possible frequency distributions, if the full network for
+the set of all possible frequency distributions, if the full population for
 example comprises 1\,000 neurons, is a 1\,000-dimensional space. In
@@
 
-% (for a full network of 1\,000 neurons and
+% (for a full population of 1\,000 neurons and
 % a recording of 500\,000 time bins, for example, there are
@@
 % \mynote{Add this?: from sampling theory we know that important
-%   features of the full network may not be visible in a sample because
+%   features of the full population may not be visible in a sample because
 %   smoothed out. But using sampling theory in the inverse direction we can
-%   infer such full-network features from the sample.}
+%   infer such full-population features from the sample.}
 
@@
 have recorded the spiking activities of $n$ neurons for a certain amount of
-time. These neurons are our \enquote{sample} or \enquote{subnetwork}. Their
+time. These neurons are our \enquote{sample} or \enquote{subpopulation}. Their
 spikes are binned into $T$ time bins and binarized to $\set{0,1}$ values in
@@
 change slightly across experimental sessions that are very far apart in
-time. We assume that there's an area, comprising a network of $N$ neurons,
+time. We assume that there's an area, comprising a population of $N$ neurons,
 for which we believe that any other sample of size $n$ could have equally
 plausibly been recorded instead of the sample of $n$ neurons that was
-actually recorded. We call this larger network the \enquote{full network}.
+actually recorded. We call this larger population the \enquote{full population}.
 The total activity of these $N$ neurons at $t$ is $\yA_{t}$. The relative
@@
 Our questions concern general features of the total activity $\yA$ of the
-full network during and after the recording, and across sessions under the
+full population during and after the recording, and across sessions under the
 same study conditions. For example: what was its frequency distribution
@@
   total activity $\ya$ of the sample and corresponding expected values $M$
-  for the total activity $\yA$ of the full network.
+  for the total activity $\yA$ of the full population.
 \item\label{item:maxent_step} Using the maximum-entropy method we build a
   distribution $\px(\yA \| M, N)$ for the total activity of the full
-  network of $N$ neurons, using as constraints the expected values $M$
+  population of $N$ neurons, using as constraints the expected values $M$
   found in the previous step.
@@
 application of the probability calculus, which gives an exact linear
-relation between the first $m$ moments for the full network and the first
+relation between the first $m$ moments for the full population and the first
 $m$ for the sample \citep[\eqns~(16)]{portamanaetal2015}. The ones
 determine the others and vice versa at every time bin. This relation holds
-for any belief distribution $\pxxx(\yA_{t})$ for the full-network activity
+for any belief distribution $\pxxx(\yA_{t})$ for the full-population activity
 and its marginal $\pxx(\ya_{t})$ for the sample activity at that bin.
@@
 possible number of distinct $m$-tuples. % For
-% example, with $m=2$, if $a=4$ neurons spike in a network of $n=5$, we
+% example, with $m=2$, if $a=4$ neurons spike in a population of $n=5$, we
 % have $\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons,
@@
 analysis: \emph{the first $n$ normalized factorial moments for the sample
-  and for the full network are numerically identical}:
+  and for the full population are numerically identical}:
 \begin{equation}
@@
 \end{equation}
-% where $\pxxx(\yA)$ is our distribution of belief about the full-network
+% where $\pxxx(\yA)$ is our distribution of belief about the full-population
 % activity, and $\pxx$
@@
 $\binom{\yA}{m}$ to equate to the factorial moment
-$\E\bigl[\binom{\yA}{m}\bigr]$ of the full-network distribution
+$\E\bigl[\binom{\yA}{m}\bigr]$ of the full-population distribution
 $\pxxx(\yA)$. But \eqn~\eqref{eq:equal_m_factorialmoments} of
 step~\ref{item:sample_step} comes to our rescue, because it says that the
-expectations for the full network are determined by those for the sample
+expectations for the full population are determined by those for the sample
 $\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time averages of the
@@
   \rightarrow
-    \text{full-network moments}}_{\textit{sampling theory}}
+    \text{full-population moments}}_{\textit{sampling theory}}
 \end{equation*}
 and apply the maximum-entropy method to obtain a distribution
-$\px(\yA \| M, N)$ for the full network of $N$ neurons by constraining some
+$\px(\yA \| M, N)$ for the full population of $N$ neurons by constraining some
 of its factorial moments, for example $M=\set{1,\dotsc,m'}$ with $m'\le n$,
@@
 and hypotheses that a researcher is exploring; for example, hypotheses
-about the \enquote{cooperativity} or \enquote{interaction} in the network
+about the \enquote{cooperativity} or \enquote{interaction} in the population
 activity. We discuss this kind of use in \sect~\ref{sec:marginalization}.
@@
 $\yA$ can be equally likely be generated by any set of $\yA$ neurons in the
-full network. The magnitude of this discrepancy can be a signature of the
+full population. The magnitude of this discrepancy can be a signature of the
 (at least temporary) presence of a \enquote{neuronal assembly}
@@
 approximation)\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}
-of four different distributions for the full network, which differ
+of four different distributions for the full population, which differ
 numerically from one another in higher-order approximations:
@@
 % can be considered as \emph{the most probable frequency distribution} for
-% the full-network activity during the recording. But it also approximates
+% the full-population activity during the recording. But it also approximates
 % three other distributions, as discussed in the previous section.
@@
 
-We first calculate the distribution by using five moments. This number
-already provides almost as much information as the full frequency
-distribution of the sample (see next section). Figure~\ref{fig:plots_5mom}
-shows the resulting densities (distribution${}\times N$) for three example
-values of full-network sizes: $N=1\,000$, $N=5\,000$, $N=10\,000$. The
-frequency distribution of the sample activity is also shown, and in the
-plot it is indistinguishable from the maximum-entropy distribution for
-$N=n$, that is, applied at the sample level. We discuss the case of unknown
-$N$ in \sect~\ref{sec:N}.
+The full-population maximum-entropy distribution is calculated using five
+moments. This number seems to provide almost as much information as the
+full frequency distribution of the sample (see next section).
+Figure~\ref{fig:plots_5mom} shows the resulting densities
+(distribution${}\times N$) for three example values of full-population
+sizes: $N=1\,000$ (\textcolor{mygreen}{green diamonds}), $N=5\,000$
+(\textcolor{myred}{red circles}), $N=10\,000$ (\textcolor{myblue}{blue
+  curve}). The frequency distribution of the sample activity is also shown
+(\textcolor{black}{black triangles}), and in the plot it would be
+indistinguishable from the maximum-entropy distribution for $N=n$, that is,
+applied at the sample level. We discuss the case of unknown $N$ in
+\sect~\ref{sec:N}.
 \begin{figure}[!p]
@@
 
-The figure shows that the distribution for the full-network is less flat
-than the measured frequency distribution for the sample; their difference
-increases with $N$. Most remarkably, for the first data set the
-distribution has two distinct low-activity modes. For the second data set
-the distribution presents a small shoulder on the right of the
-mode.% , suggestive of two activity regimes. We frame no hypotheses about the
+The most expensive calculation, for $N=10\,000$, takes less than 15 minutes
+on a laptop with two $2\;\textrm{GHz}$ cores. In all cases the moments were
+recovered with relative errors smaller than $10^{-12}$.
+% 1000:  3*10^-13 %
+% 5000: 7*10^-13 %, 581 s
+% 10000: 6*10^-11 %, 737 s
+
+The figure shows that the distribution for the full-population is more
+peaked than the measured frequency distribution for the sample; their
+difference increases with $N$. Most remarkably, for the first data set
+(upper panel) the distribution for the full population has two distinct
+low-activity modes. For the second data set (lower panel) the distribution
+presents a small shoulder on the right of the mode.
+% , suggestive of two activity regimes. We frame no hypotheses about the
 % biological cause of these two modes (they could stem from the presence of
 % different kinds of cells or modules)
-These features are clearly not present in the sample frequencies or in the
+Such features are clearly not present in the sample frequencies or in the
 maximum-entropy distribution at the sample level. The application of the
 probability calculus thus reveals interesting possible features of the full
-network. \textcolor{white}{If you find this you can claim a postcard from
+population. \textcolor{white}{If you find this you can claim a postcard from
   us.}
 
-\bigskip
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\linewidth]{_scripts/newpaper3_gridvsnogrid.pdf}%
+\\[3em]%
+\includegraphics[width=\linewidth]{_scripts/newpaper3_gridconv.pdf}%
+\caption{***}
+\label{fig:plots_cellsubsets}
+\end{figure}
+% maxent_solutions_different_moments_partial.nb
+These results show the potential usefulness of this maximum-entropy
+approach. Here is a \emph{toy} example of possible uses. We could be
+interested in the hypothesis that two distinct cell types or assemblies be
+present in the region where the recording was made. Finding a
+full-population distribution such as that in the upper panel of
+\fig~\ref{fig:plots_5mom} would provide some evidence for this hypothesis.
+Let's further imagine that we have reasons for suspecting that a specific
+set of the sampled neurons are of one type, and the remaining of the other
+type. In the case of data set 1, 27 of the 65 sampled neurons were
+identified as grid cells belonging to 3--4 functional modules, see
+\cite{dunnetal2015}. Could the two peaks in the distribution of
+\fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
+cells? We apply the method to these two sets of neurons individually, using
+$N=4\,150$ for the grid set and $N=5\,850$ for the non-grid set, to reflect
+their proportions in the recorded sample. The results are shown in the
+upper panel of \fig~\ref{fig:plots_cellsubsets}. The distribution for the
+larger population of grid cells (\textcolor{mygreen}{green triangles})
+seems to have one, broad peak, close to the first peak of the
+full-population distribution. The distribution for the larger population of
+non-grid cells (\textcolor{myred}{red circles}) has two peaks instead,
+roughly at the same positions as the full-population distribution
+(\textcolor{myblue}{blue curve}) but closer in height. So it would seem
+that the population of grid cells is contributing to the first peak of the
+full population, but it isn't the sole contributor. We can also assess how
+much the distributions for the two sets are independent. If they were
+completely independent, the full-population distribution would be given by
+their convolution:
+\begin{equation}
+  \label{eq:convolution}
+  P_{\text{full}}(\yA) =  \smashoperator{\sum_{\yA'=0}^{\yA}} P_{\text{grid}}(\yA')\;
+  P_{\text{non-grid}}(\yA-\yA'). %, \qquad \yA \in \set{1,\dotsc,10\,000}
+\end{equation}
+However, the lower panel of
+figure~\ref{fig:plots_cellsubsets} shows that such convolution
+(\textcolor{black}{black diamonds}) is different from the full-population
+distribution (\textcolor{myblue}{blue curve}): the two peaks of the former
+are closer in position and height. This means that the population
+distributions of grid and non-grid cells are \emph{not} independent:
+knowledge of the activity of either set gives us some information about the
+activity of the other.
+
+\medskip
+
+The toy analysis just given should  be taken not literally, but just
+as an illustration of the method's possible applications. The important point
+is that this method is computationally very cheap but can provide useful
+insights, even if just qualitative ones.
+
+\section{Quantifying the importance of higher-order correlations:\\
+  the limitations of methods at the sample level}
+\label{sec:marginalization}
 
@@
 \enquote{cooperativity} or \enquote{interaction},
-\fig~\ref{fig:plots_2_4mom} shows the full-network distributions, for
+\fig~\ref{fig:plots_2_4mom} shows the full-population distributions, for
 $N=10\,000$, obtained constraining the first two moments (in
@@
 reached by applying the maximum-entropy method \emph{at the sample level},
-as traditionally done. Such a comparison involves an important
-methodological caveat, so we discuss it in full in the next section.
-
-\section{Quantifying the importance of higher-order correlations:\\
-  the limitations of methods at the sample level}
-\label{sec:marginalization}
+as traditionally done. Such a comparison involves some important
+methodological caveat which we wish to discuss first.
 
@@
 % This informational quasi-sufficiency is interesting because
-% it may give clues about network organization and dynamics.
+% it may give clues about population organization and dynamics.
 
@@
 correlations seem informationally sufficient for a particular sample from a
-brain area, then they may well not be sufficient for the full network of
+brain area, then they may well not be sufficient for the full population of
 neurons constituting that area, and vice versa. So if what interests us is
@@
 maximum-entropy method is applied to obtain the distribution of the full
-network, not of the sample alone.
+population, not of the sample alone.
 
@@
 \item from each maximum-entropy distributions $\px(\yA \| M)$ for the full
-  network, built from each set of constraints $M=M',M''$, calculate the
+  population, built from each set of constraints $M=M',M''$, calculate the
   marginal distribution for the sample:
@@
 $M_{4}$, $M_{5}$, consisting of the first two, four, five moments. For a
-full network of size $N=10\,000$, we obtain the following differences:
+full population of size $N=10\,000$, we obtain the following differences:
 \begin{equation}
@@
 
-The application at the full-network level therefore leads to different
+The application at the full-population level therefore leads to different
 conclusions about the relative importance of the moments from the
@@
 $\yf$ conditional on the hypothesis $M_{5, N}$ of 5-moment sufficiency at
-the full-network level, $N=10\,000$, versus the hypothesis $M_{5, n}$ of
+the full-population level, $N=10\,000$, versus the hypothesis $M_{5, n}$ of
 5-moment sufficiency at the sample level, $N=n=65$. We obtain
@@
 The conclusions about \enquote{cooperativity} or \enquote{interaction}
-that we reach by applying the maximum-entropy method to the full network,
+that we reach by applying the maximum-entropy method to the full population,
 as proposed here, are not only different from the application at the sample
@@
 The same plot also shows the marginals obtained from applying the method to
-a full network with $N=10\,000$ and the same constraints, that is, the
+a full population with $N=10\,000$ and the same constraints, that is, the
 sample marginals of \fig~\ref{fig:plots_2_4mom}. The two-moment
@@
 filled blue squares (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The
-full-network distribution fits the measured frequencies
+full-population distribution fits the measured frequencies
 (\textcolor{black}{$\bm{\triangledown}$}) very well if we constrain only four
@@
 
-% But does the application at the full-network level lead to appreciably
+% But does the application at the full-population level lead to appreciably
 % different results from that at the sample level? after all we're interested
@@
 % question in the case of our first data set. The upper plot shows the
-% maximum-entropy distributions for a full network of $10\,000$ neurons,
+% maximum-entropy distributions for a full population of $10\,000$ neurons,
 % constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
@@
 % (\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained
-% by marginalizing the full-network distribution to the sample size, from two
+% by marginalizing the full-population distribution to the sample size, from two
 % moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
@@
 % \begin{itemize}
-% \item The application to the full-network (upper plot) leads to two
+% \item The application to the full-population (upper plot) leads to two
 %   completely different distributions (even different number of modes);
@@
 %   two statistics are approximately equivalent.
-% \item The difference between marginals from the full-network distribution
+% \item The difference between marginals from the full-population distribution
 %   and the distributions obtained at the sample level (filled vs empty
@@
 %   (triangles vs squares).
-% \item The marginals of the application to the full network are closer to
+% \item The marginals of the application to the full population are closer to
 %   the measured frequencies than the distributions obtained at the sample
@@
 % Therefore, the maximum-entropy application at the sample level and at the
-% full-network level lead to different results; the latter is not only more
+% full-population level lead to different results; the latter is not only more
 % meaningful but also superior because it shows clearly the different
@@
 
-\section{The full-network size $N$}
+\section{The full-population size $N$}
 \label{sec:N}
@@
 The calculations and conclusions presented in the preceding sections depend
-on the size of the full network, $N$. The full network can't be the whole
+on the size of the full population, $N$. The full population can't be the whole
 brain, of course. How large should or can $N$ be in our formulae?
@@
 our method is based, and which asserts the equality of the factorial
-moments of the full network and those of the sample. This formula is only
+moments of the full population and those of the sample. This formula is only
 valid if our beliefs about the activity of the sample $\pxx(\ya)$ and that
-of the full network $\pxx(\yA)$ are related by a hypergeometric
+of the full population $\pxx(\yA)$ are related by a hypergeometric
 distribution:
@@
 The second point is that our degrees of belief about the frequency
-distribution for the full network depend not only on the measured data in
+distribution for the full population depend not only on the measured data in
 the sample, but also on our pre-data beliefs $\yI$ about the distribution.
@@
 We have presented a procedure to construct the most plausible frequency
-distribution of population-averaged activities of a network of neurons,
+distribution of population-averaged activities of a population of neurons,
 given the recording about a small sample thereof. This procedure combines
@@
 Here is a summary derivation of the maximum-entropy distribution for the
-full network constrained by a set of factorial moments. For further details
+full population constrained by a set of factorial moments. For further details
 see Porta Mana \etal\ \citey{portamanaetal2015}.
@@
 \end{equation}
-In our case the units are neurons, the set is the full network, and the
+In our case the units are neurons, the set is the full population, and the
 property is their being active in a specific time bin.
@@
 can prove the equality~\eqref{eq:equal_m_factorialmoments} between the
-factorial moments of sample and full network.
+factorial moments of sample and full population.
 
@@
 bin is the sample marginal of the maximum-entropy distribution for the full
-network, owing to the excheangeability assumption implicit in the
+population, owing to the excheangeability assumption implicit in the
 maximum-entropy method:
@@
 Yes. In the simplest cases, the probability calculus give us simple and
-precise quantitative relations between our expectations about a network
+precise quantitative relations between our expectations about a population
 -- of voters, for instance -- and about a sample thereof. Suppose there are
@@
   proportion of political preferences in a sample as in the full
-  network}.
+  population}.
 
@@
 that expectation equalities such as~\eqref{eq:poll_equal_expectations} hold
-for the total activity $S(t)$ of the full network of $N$ neurons that
+for the total activity $S(t)$ of the full population of $N$ neurons that
 could have been recorded.
@@
   neurons we can construct maximum-entropy distributions for the total
-  activity of the \textbf{full network} of neurons that could have been
+  activity of the \textbf{full population} of neurons that could have been
   sampled by the recording device}.
@@
 % a probability distribution for the total activity at a new time bin of a
-% network of neurons, \emph{most of which can be unrecorded}, given
+% population of neurons, \emph{most of which can be unrecorded}, given
 % specific statistics about the activity of a recorded \emph{sample} of the
-% full network. This kind of probability assignment is extremely important
+% full population. This kind of probability assignment is extremely important
 % in neuroscience: we try to understand the workings of brain areas that may
@@
 % hidden assumption. We'll study in which circumstances the application of
-% the method at the sample level and at the full-network level give
+% the method at the sample level and at the full-population level give
 % similar or different probabilities.
@@
   with $\yaa(t) \defd \bigl(\yav_1(t),\dotsc,\yav_n(t)\bigr)$. The
-  network-averaged activity at that bin is
+  population-averaged activity at that bin is
   $\ya(t) \defd \sum_i \yav_i(t)/n$. If we count the number of distinct
@@
   most $\tbinom{n}{2}$ simultaneously active pairs, so the
-  network-averaged pair activity is
+  population-averaged pair activity is
   $\av{\yaa \yaa}(t) \defd \tbinom{n}{2}^{-1}\tbinom{n\ya(t)}{2}$. With some
-  combinatorics we see that the network-averaged activity of $m$-tuples
+  combinatorics we see that the population-averaged activity of $m$-tuples
   of neurons is
@@
 into account the activities or pair activities of specific neurons rather
-than network averages, or even protocols that take into account time
+than population averages, or even protocols that take into account time
 dependence.
@@
 that the recorded neurons are a \emph{sample} from a larger, unrecorded
-network, but the \me\ method as applied above is treating them as
+population, but the \me\ method as applied above is treating them as
 isolated from the rest of the brain. Hence, the results it provides cannot
@@
 some square millimetres \citep[\cf][]{berenyietal2014}. This recording is a
-sample of the activity of the neuronal network under the recording
+sample of the activity of the neuronal population under the recording
 device, which can amount to tens of thousands of neurons
@@
 little meaning if we did not expect its results to generalize to a larger,
-unrecorded network -- at the very least the network under the
+unrecorded population -- at the very least the population under the
 recording device. In other words, we expect that the conclusions drawn with
@@
 neurons are a \emph{representative sample} of some larger neuronal
-network. Probability theory tells us how to make inferences from a
-sample to the larger network from which it is sampled \parentext{see
+population. Probability theory tells us how to make inferences from a
+sample to the larger population from which it is sampled \parentext{see
   references below}.
@@
 sample. But, given that our sample is representative of a larger
-network, we can also apply the \me\ method to the larger (unrecorded)
-network. The constraints are the same: the time averages of the sampled
-data, since they constitute representative data about the larger network
+population, we can also apply the \me\ method to the larger (unrecorded)
+population. The constraints are the same: the time averages of the sampled
+data, since they constitute representative data about the larger population
 as well. The method thus yields a probability distribution for the larger
-network, and the distribution for the sample is obtained by
+population, and the distribution for the sample is obtained by
 marginalization. The problem is that \emph{the distributions obtained from
@@
 In this work we develop the second way of applying the \me\ method, at the
-level of the larger network, and show that its results differ from the
+level of the larger population, and show that its results differ from the
 application at the sample level. We also consider the case where the size
-of the larger network is unknown.
-
-To apply the \me\ method to the larger, unsampled network, it is
+of the larger population is unknown.
+
+To apply the \me\ method to the larger, unsampled population, it is
 necessary to use probability relations relevant to sampling
@@
 
-\section{Probability relations between network and  sample}
+\section{Probability relations between population and  sample}
 \label{sec:prob_samples}
@@
 introduce an analogous notation for the $\yNv$ neurons constituting the
-larger network, but using the corresponding capital letters:
+larger population, but using the corresponding capital letters:
 $\yAv_{i}(t)$ is the activity of the $i$th neuron at time bin $t$,
 $\yA(t) \defd \sum_{i} \yAv_{i}(t)/\yNv$ is the activity at that
-bin averaged over the larger network, and so on.
-
-The probability relations between sample and larger network are valid at
+bin averaged over the larger population, and so on.
+
+The probability relations between sample and larger population are valid at
 every time bin. As we mentioned above, the \me\ distribution refers to any
@@
 %     &\text{\enquote{$\yXf = \yA$} 
-%       stands for \enquote{the (network-averaged) activity of the neurons is $\yA$}},
+%       stands for \enquote{the (population-averaged) activity of the neurons is $\yA$}},
 %     \\
@@
 backing our probability assignments -- our uncertainty about the full
-activity of the larger network is expressed by the joint probability
+activity of the larger population is expressed by the joint probability
 distribution
@@
 We need to make an initial probability assignment for the state of the full
-network before any experimental observations are made. This initial
+population before any experimental observations are made. This initial
 assignment will be modified by our experimental observations, and these can
-involve just a sample of the network. Our state of knowledge and initial
+involve just a sample of the population. Our state of knowledge and initial
 probability assignment should reflect that samples are somehow
-representative of the whole network.
+representative of the whole population.
 
 In this state of knowledge, denoted $\yH$, we know that the neurons in the
-network are biologically or functionally similar, for example in
+population are biologically or functionally similar, for example in
 morphology or the kind of input or output they receive or give. But we are
@@
 probability assignment is just an expression of the symmetry of our
-\emph{ignorance} about the state of the network, not an expression of
+\emph{ignorance} about the state of the population, not an expression of
 some biologic or physical symmetry or identity of the neurons.
@@
 activity is completely determined by the distribution for its
-network-average:
+population-average:
 \begin{equation}
@@
 
-Our uncertainties about the full network and the sample are connected
+Our uncertainties about the full population and the sample are connected
 via the conditional probability
@@
 of the sample is also symmetric and completely determined by the
-distribution of its network-averaged activity:
+distribution of its population-averaged activity:
 \begin{equation}
@@
 theorem~\eqref{eq:joint_plaus_N_homog} as the distribution for the full
-network.
+population.
 
@@
 $\pf(\ya \|\yA, \yH) \equiv \yG_{\ya \yA}$, besides
-relating the distributions for the network and sample activities via
+relating the distributions for the population and sample activities via
 marginalization, also allows us to express the expectation value of any
 function of the sample activity, $\yg_{\ya}$, in terms of the distribution
-for the full network, as follows:
+for the full population, as follows:
 \begin{equation}
@@
 \emph{translates into a linear constraint for the distribution of the full
-  network}:
+  population}:
 \begin{equation}
@@
 that is, \emph{the expected values of the $m$-activities of the sample and
-  of the full network are equal}. The proof of the middle equality uses
+  of the full population are equal}. The proof of the middle equality uses
 the expression for the $m$th factorial moment of the hypergeometric
@@
 $m$-activity of a sample to have a particular value, we are also
-requiring  the expectation of the $m$-activity of the full network to
+requiring  the expectation of the $m$-activity of the full population to
 have the same value.
@@
 \end{figure}%hypergeometric_identity.nb
-These expectation equalities between sample and full network should not
+These expectation equalities between sample and full population should not
 be surprising: we intuitively \emph{expect} that the proportion of coloured
@@
 equalities express the intuitive fact that \emph{our uncertainty about the
-  sample is representative of our uncertainty about the network and
+  sample is representative of our uncertainty about the population and
   about other samples}, and vice versa. When $n=\yNv$, $\yG_{\ya\yA}$
 becomes the identity matrix and the approximate equalities above become
-exact -- of course, since we have sampled the full network.
+exact -- of course, since we have sampled the full population.
 
@@
 probability distributions. In the next section we will in fact emphasize
-their differences. If the distribution for the network average $\yA$ is
+their differences. If the distribution for the population average $\yA$ is
 bimodal, for example, the bimodality can be lost in the distribution for
@@
 % $n\to \yNv$, do: these limits are trivially, universally valid because the
-% sample becomes the full network as $n\to \yNv$. In particular, these limits
+% sample becomes the full population as $n\to \yNv$. In particular, these limits
 % hold even in cases where the conditional probability
 % $\pf(\yxs = \ya \|\yXf=\yA)$ is not a fuzzy identity, our uncertainties
-% about sample and about network can differ wildly, and the approximate
+% about sample and about population can differ wildly, and the approximate
 % equalities~\eqref{eq:roughly_equal_nN} do not hold.
 
-\section{Maximum-entropy: sample level \vs\ full-network level}
+\section{Maximum-entropy: sample level \vs\ full-population level}
 \label{sec:specific_initial_probability}
@@
 be used as constraints on the distribution for the activity of the full
-network. Let us use such constraints with the \me\ method. Suppose that
+population. Let us use such constraints with the \me\ method. Suppose that
 we want to constrain $m$ functions of the sample activity, vectorially
@@
 directly to sampled neurons; denote this approach by $\yHb$. Then we apply
-the method to the full network of neurons, most of which are unsampled;
+the method to the full population of neurons, most of which are unsampled;
 denote this approach by $\yHa$.
@@
 exponential appears because we must account for the multiplicity by which
-the network-average activity $\ya$ can be realized: $\ya=0$ can be
+the population-average activity $\ya$ can be realized: $\ya=0$ can be
 realized in only one way (all neurons inactive), $\ya=1/n$ can be realized
@@
 
-Applied to the full network, using the constraint
+Applied to the full population, using the constraint
 expression~\eqref{eq:constraint_extended} derived in the previous section,
-the method yields the distribution for the full-network activity
+the method yields the distribution for the full-population activity
 \begin{equation}
@@
 Hence, \emph{the applications of \me\ at the sample level and at the
-  full-network level are inequivalent}. They lead to numerically
+  full-population level are inequivalent}. They lead to numerically
 different distributions for the sample activity $\yaa$.
@@
 \caption{Linear and log-plots of $\pf(\ya)$ constructed by \me\ at
-  the network level followed by sample marginalization (blue triangles),
+  the population level followed by sample marginalization (blue triangles),
   \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
@@
   \textcolor{mypurpleblue}{\textbf{blue triangles:}} \me\ at the
-  network level, \eqn~\eqref{eq:app_maxent_pop_marg} with
+  population level, \eqn~\eqref{eq:app_maxent_pop_marg} with
   $\yNv=10\,000$, followed by sample marginalization;
-  \textcolor{myyellow}{\textbf{yellow circles:}} \me\ at the network
-  level with unknown network size,
+  \textcolor{myyellow}{\textbf{yellow circles:}} \me\ at the population
+  level with unknown population size,
   \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}, according to the
-  distribution~\eqref{eq:pdf_popsize} for the network.}
+  distribution~\eqref{eq:pdf_popsize} for the population.}
 \label{fig:all_three}
@@
 The distribution obtained at the sample level will show different
-features from the one obtained at the network level, like displaced or
+features from the one obtained at the population level, like displaced or
 additional modes or particular tail behaviour. We show an example of this
@@
 motor cortex \citep{rostamietal2016_r2017}. The distribution obtained at the
-network level (blue triangles) has a higher and displaced mode and a quite
+population level (blue triangles) has a higher and displaced mode and a quite
 different behaviour for activities around $0.5$ than the distribution
@@
 In our discussion we have so far assumed the size $\yNv$ of the larger
-network to be known. This is rarely the case, however. We usually are
+population to be known. This is rarely the case, however. We usually are
 uncertain about $\yNv$ and can only guess its order of magnitude. In such a
@@
 assuming an entropic pre-data distribution for the long-run frequencies of
-the full network:
-
-\textbf{The frequency distribution $\yF$ of the full-network activity
+the full population:
+
+\textbf{The frequency distribution $\yF$ of the full-population activity
   \emph{during the recording}}
@@
 
-\textbf{The long-run frequency distribution $\ynu$ of the full-network
+\textbf{The long-run frequency distribution $\ynu$ of the full-population
   activity}
@@
 
-\textbf{The activity $\yA'$ of the full-network
+\textbf{The activity $\yA'$ of the full-population
   in a \emph{new} time bin}
@@
 
-The formulae say that we have equal beliefs about the underlying network
+The formulae say that we have equal beliefs about the underlying population
 states having the same activity. This isn't really our belief, for we
-believe there are interactions between the neurons and subnetworks thereof.
+believe there are interactions between the neurons and subpopulations thereof.
 
@@
 set-up, that the \me\ method can be applied to recorded neuronal data in a
-way that accounts for the larger network from which the data are
+way that accounts for the larger population from which the data are
 sampled, \eqns~\eqref{eq:app_maxent_pop}--\eqref{eq:app_maxent_pop_marg}.
@@
 We gave a numerical example of this difference. We have also shown how to
-extend the new application when the size of the larger network is
+extend the new application when the size of the larger population is
 unknown, \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}.
@@
 \me\ 
-implicitly assumes that \emph{no} larger network exists beyond the
+implicitly assumes that \emph{no} larger population exists beyond the
 recorded sample of neurons. One could in fact object to the application at
-the network level, and say that the traditional way of applying \me,
+the population level, and say that the traditional way of applying \me,
 \eqn~\eqref{eq:app_maxent_sample}, yields different results because it does
 not make assumptions about the size $\yNv$ of a possibly existing larger
-network. Such a state of uncertainty, however, is correctly formalized
+population. Such a state of uncertainty, however, is correctly formalized
 according to the laws of probability by introducing a probability
@@
 for $\yNv$ gives unit probability to $\yNv=n$; that is, unless the sample
-\emph{is} the full network, and no larger network exists.
+\emph{is} the full population, and no larger population exists.
 
 The standard \me\ approach therefore assumes that the recorded neurons
-constitute a special subnetwork, isolated from the larger network of
+constitute a special subpopulation, isolated from the larger population of
 neurons in which it is embedded, and which was also present under the
 recording device. This assumption is unrealistic. The \me\ approach at the
-network level does not make such assumption and is therefore preferable.
+population level does not make such assumption and is therefore preferable.
 It may reveal features in a data set that were unnoticed by the standard
@@
 % very special, isolated set among all those that could have been recorded.
-% It is for this reason that we find the \me\ application at the network
-% level preferable. Physical models of neuronal networks usually include some
-% sort of external input, mimicking an embedding in a larger network. The
-% \me\ distribution obtained at the network level may reveal features in a
+% It is for this reason that we find the \me\ application at the population
+% level preferable. Physical models of neuronal populations usually include some
+% sort of external input, mimicking an embedding in a larger population. The
+% \me\ distribution obtained at the population level may reveal features in a
 % data set that were unnoticed by the standard \me\ model.
@@
 The difference in the resulting distributions between the applications at
-the sample and at the network levels appears in the use of Boltzmann
+the sample and at the population levels appears in the use of Boltzmann
 machines with hidden units \citep{lerouxetal2008}, although by a different
@@
 neuroscientific applications, the difference in predictions of the sample
-\vs\ full-network applications can instead be very relevant.
+\vs\ full-population applications can instead be very relevant.
 
@@
 Finally, we would like to point out the usefulness and importance of the
-probability formulae that relate our states of knowledge about a network
+probability formulae that relate our states of knowledge about a population
 and its samples, presented in \sect~\ref{sec:prob_samples}. This kind of
@@
 More work is needed \citep[\cf][]{levinaetal2017} to extend these formulae
-to account for finer knowledge of the cerebral cortex and its network
+to account for finer knowledge of the cerebral cortex and its population
 properties.
@@
 external stimuli or experimental conditions? What dimensionality does the
-neural data live in and how is this related to the underlying network
+neural data live in and how is this related to the underlying population
 interactions? The probabilistic models can also be used to make predictions
@@
 experimental measurements of neuronal activity still only provide data from
-a small subset of neurons that comprise a neuronal network. A wealth of
+a small subset of neurons that comprise a neuronal population. A wealth of
 studies on building probabilistic models of neural data focuses on
@@
 models, how our inferences about the recorded activity would be affected by
-them, and what we can we say about the rest of the network by studying the
+them, and what we can we say about the rest of the population by studying the
 heavily subsampled recordings. In this paper, we aim to address these
@@
 neuronal activity in the retina, in the cortex, protein sequences, gene
-regulatory networks and natural images. The general idea is, for a dataset,
+regulatory populations and natural images. The general idea is, for a dataset,
 to write down the distribution that maximizes the entropy of state
 variables, given some low order statistics. Now given the fact that the
-recorded neurons are a fraction of the neurons in the network, several
+recorded neurons are a fraction of the neurons in the population, several
 quantitative questions arise that we will address in this paper: given the
 data from the sampled neurons, can we build a maximum entropy model over
-the whole network? Once we build such a network level maximum entropy
+the whole population? Once we build such a population level maximum entropy
 model, can we see features in the neural activity which cannot be directly
 seen from a model build from the sampled neurons? Since we can marginalize
-down the network level maximum entropy model to the sampled network, how
+down the population level maximum entropy model to the sampled population, how
 does this margianzlied maximum entropy model match the sample level model?
@@
 entropy model. First we show how to go from the sample level maximum
-entropy level to the network level, by assuming different sizes of the
-network and also by assuming an uninformative prior over the size of the
-network. This is done by inferring the statistics of correlation functions
-at the network level from those of the sample level by using simple
+entropy level to the population level, by assuming different sizes of the
+population and also by assuming an uninformative prior over the size of the
+population. This is done by inferring the statistics of correlation functions
+at the population level from those of the sample level by using simple
 counting arguments. We then find that, when applied to experimental
 recordings from the Medial Entorhinal Cortex of rats and the monkey visual
-cortex, this network level maximum entropy model may exhibit features that
+cortex, this population level maximum entropy model may exhibit features that
 the sample level model does not predict. Specifically, we observed modes in
-the distribution of the activity in the network level model that do not
+the distribution of the activity in the population level model that do not
 show up in the sample level. We study how the assumed size of the full
-network affects the appearance of these modes and find that there is a
-minimum size of the full network for which such modes can be observed. We
-then compared the distribution found by marginalizing the full network
+population affects the appearance of these modes and find that there is a
+minimum size of the full population for which such modes can be observed. We
+then compared the distribution found by marginalizing the full population
 maximum entropy model down to the sample level, and the distribution fit
@@
 there are also differences between them. We quantify how these differences
-also depend on the assumed size of the network and find that …(WE SHOULD
-TEST SHI) This predicts that for a large enough network (DO WE PREDICT
-THAT IF THE FULL NETWORK GETS BIGGER THE DIFFERENCE ALSO GET BIGGER)?…
+also depend on the assumed size of the population and find that …(WE SHOULD
+TEST SHI) This predicts that for a large enough population (DO WE PREDICT
+THAT IF THE FULL POPULATION GETS BIGGER THE DIFFERENCE ALSO GET BIGGER)?…
 
 The rest of the paper is organized as follows. We first describe how to go
-from the sample level maximum entropy model to the full network maximum
+from the sample level maximum entropy model to the full population maximum
 entropy model. In section 2, we apply this to the two experimental datasets
-and study the effect of the assumed size of the network as well as the
+and study the effect of the assumed size of the population as well as the
 moments that we use for building the maximum entropy model. In section 3 we

### 2019-11-25 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-25T17:35:14+0100
+%% Last-Updated: 2019-11-25T17:52:38+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % maxent_solutions_different_moments_partial.nb
-These results show the potential usefulness of this maximum-entropy
-approach. Here is a \emph{toy} example of possible uses. We could be
-interested in the hypothesis that two distinct cell types or assemblies be
-present in the region where the recording was made. Finding a
-full-population distribution such as that in the upper panel of
-\fig~\ref{fig:plots_5mom} would provide some evidence for this hypothesis.
-Let's further imagine that we have reasons for suspecting that a specific
-set of the sampled neurons are of one type, and the remaining of the other
-type. In the case of data set 1, 27 of the 65 sampled neurons were
-identified as grid cells belonging to 3--4 functional modules, see
-\cite{dunnetal2015}. Could the two peaks in the distribution of
-\fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
-cells? We apply the method to these two sets of neurons individually, using
-$N=4\,150$ for the grid set and $N=5\,850$ for the non-grid set, to reflect
-their proportions in the recorded sample. The results are shown in the
-upper panel of \fig~\ref{fig:plots_cellsubsets}. The distribution for the
-larger population of grid cells (\textcolor{mygreen}{green triangles})
-seems to have one, broad peak, close to the first peak of the
-full-population distribution. The distribution for the larger population of
-non-grid cells (\textcolor{myred}{red circles}) has two peaks instead,
-roughly at the same positions as the full-population distribution
-(\textcolor{myblue}{blue curve}) but closer in height. So it would seem
-that the population of grid cells is contributing to the first peak of the
-full population, but it isn't the sole contributor. We can also assess how
-much the distributions for the two sets are independent. If they were
-completely independent, the full-population distribution would be given by
-their convolution:
+These results suggest that this maximum-entropy approach could be very
+useful. Here is a toy example of possible uses. We could be interested in
+the hypothesis that two distinct cell types or assemblies be present in the
+region where the recording was made. Finding a full-population distribution
+with two peaks, as in the upper panel of \fig~\ref{fig:plots_5mom}, would
+provide some evidence for this hypothesis. Let's further imagine that we
+have reasons for suspecting that a specific set of the sampled neurons are
+of one type, and the remaining of the other type. In the case of the first
+data set, 27 of the 65 sampled neurons were identified as grid cells
+belonging to 3--4 functional modules, see \cite{dunnetal2015}. Could the
+two peaks in the distribution of \fig~\ref{fig:plots_5mom} reflect the
+activities of grid versus non-grid cells? We apply the method to these two
+sets of neurons individually, using $N=4\,150$ for the grid set and
+$N=5\,850$ for the non-grid set, to reflect their proportions in the
+recorded sample. The results are shown in the upper panel of
+\fig~\ref{fig:plots_cellsubsets}. The distribution for the larger
+population of grid cells (\textcolor{mygreen}{green triangles}) seems to
+have one, broad peak, close to the first peak of the full-population
+distribution. The distribution for the larger population of non-grid cells
+(\textcolor{myred}{red circles}) has two peaks instead, roughly at the same
+positions as the full-population distribution (\textcolor{myblue}{blue
+  curve}) but closer in height. So it would seem that the population of
+grid cells is contributing to the first peak of the full population, but it
+isn't the sole contributor. We can also assess how much the distributions
+for the two sets are independent. If they were completely independent, the
+full-population distribution would be given by their convolution:
 \begin{equation}

### 2019-11-25 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-25T17:52:38+0100
+%% Last-Updated: 2019-11-25T18:32:19+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \label{eq:convolution}
-  P_{\text{full}}(\yA) =  \smashoperator{\sum_{\yA'=0}^{\yA}} P_{\text{grid}}(\yA')\;
+  P_{\text{full}}(\yA) =  \smashoperator{\sum_{\yA'=0}^{4150}} P_{\text{grid}}(\yA')\;
   P_{\text{non-grid}}(\yA-\yA'). %, \qquad \yA \in \set{1,\dotsc,10\,000}
@@
 
-The toy analysis just given should  be taken not literally, but just
-as an illustration of the method's possible applications. The important point
-is that this method is computationally very cheap but can provide useful
-insights, even if just qualitative ones.
+The toy analysis just given should be taken not literally, but just as an
+illustration of the method's possible applications. The important point is
+that this method is computationally very cheap but can provide useful
+insights, even if just qualitatively.
 
@@
 methodological caveat which we wish to discuss first.
+
+\bigskip
 

### 2019-11-25 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-25T18:32:19+0100
+%% Last-Updated: 2019-11-25T19:19:26+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 grid cells is contributing to the first peak of the full population, but it
-isn't the sole contributor. We can also assess how much the distributions
-for the two sets are independent. If they were completely independent, the
-full-population distribution would be given by their convolution:
+isn't the sole contributor. We can also assess if the distributions for the
+two sets are independent. If they were independent, the full-population
+distribution would be given by their convolution:
 \begin{equation}
   \label{eq:convolution}
-  P_{\text{full}}(\yA) =  \smashoperator{\sum_{\yA'=0}^{4150}} P_{\text{grid}}(\yA')\;
+  P_{\text{full}}(\yA) =  \smashoperator{\sum_{\yA'=\max(0,\yA-5850)}^{\min(\yA,4150)}} P_{\text{grid}}(\yA')\;
   P_{\text{non-grid}}(\yA-\yA'). %, \qquad \yA \in \set{1,\dotsc,10\,000}

### 2019-11-26 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-25T19:19:26+0100
+%% Last-Updated: 2019-11-26T10:47:14+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\yav}{a}
-\newcommand*{\yNv}{N}
+\newcommand*{\yNg}{N_{\text{g}}}
+\newcommand*{\yNng}{N_{\text{ng}}}
 \newcommand*{\yNN}{\varNu}
@@
 
-What correlations are important for the description of the multi-neuronal
-activity in a specific brain area? How does such activity change when
-external stimuli or experimental conditions change? Does such activity
-range over all its mathematically possible values, or only over a
-subset thereof?
+What correlations are dominant in the neuronal activity of a specific brain
+area? How does such activity change when external stimuli or the activity
+of other areas change? Does such activity range over all its mathematically
+possible values, or only within restricted bounds?
 
 Answering this kind of questions always engages an element of uncertainty.
-We can't say \enquote{the answer is such and such}, but can at best assign
-a degree of reasonable belief -- that is, probability -- to every possible
-answer. The assessment of this distribution of belief involves experimental
-data, such as recordings of neuronal activity from specific brain areas,
-and pre-data knowledge about biological conditions and mechanisms. Our
-pre-data \dobs\ are often simplified to be mathematically more tractable,
-and are therefore called \enquote{models}.
-
-Despite remarkable advances in recording technologies, the best
-experimental measurements of instantaneous neuronal activity can still only
-record a very small sample of neurons -- hundreds at most -- compared to
-the numbers that constitute a functionally distinguished brain region. Many
-probabilistic models focus on such samples only, somehow neglecting, in
-their assumptions, that the recorded neurons are a sample from a larger
-population. This kind of isolation assumptions sometimes escape attention,
-being subtly hidden in the mathematics. Some probabilistic models try to
-take also unrecorded neurons into account individually, and therefore
-become very complex\mynote{Yasser \amp\ Vahid: refs for here?}. It would be
-useful to have models that operate in between: addressing the larger brain
-area whence the sample comes, but without adding too much complexity and
-detail about it. Such intermediate models would be useful, for example, for
-preliminary investigations, to help us to decide which hypotheses to
-discard and which to consider for more complex and costly studies.
-
-\mynote{[Yasser:] In the present work we aim to address the issue of building probability distributions over a large population using recording from subpopulations by considering the distribution of total population activity as an example.
-[}
-In the present work we propose such an intermediate model. It answers this
-question:\mynote{]} How much can the \emph{total} activity of a large neuronal
-population have been, given the observation of the activity of a very small
-sample thereof? This model addresses a larger brain area, avoiding the
-assumption of isolation of the sample; and by focusing on the total
-activity, rather than the activity of individual neurons, it remains simple
-and numerically tractable.
+We cannot say \enquote{the answer is such and such}; at best we can assign
+degrees of reasonable belief -- that is, probabilities -- to the possible
+answers. The assessment of this distribution of belief involves
+experimental data, such as recordings of neuronal activity from specific
+brain areas, and pre-data knowledge about biological conditions and
+mechanisms. Our pre-data \dobs\ are often simplified, to be mathematically
+tractable, and are therefore called \enquote{models}.
+
+The best experimental measurements of instantaneous neuronal activity use
+remarkable technologies, but can still only record a very small sample of
+neurons -- hundreds at most -- compared to the numbers that constitute a
+functionally distinguished brain region. Many probabilistic models focus on
+such samples only: they somehow neglect, in their mathematical assumptions,
+that the recorded neurons are a sample from a larger population. Such
+isolating assumptions sometimes escape attention, being subtly hidden in
+the mathematics. Some probabilistic models try to take unrecorded neurons
+into account -- but by describing each neuron individually, thus becoming
+very complex \citep{huang2015,battistinetal2017}. It would be useful to
+explore models that operate in between, addressing the larger brain area
+whence the sample comes, but collectively, without asking about individual
+neuronal details. Such intermediate models would be useful for preliminary
+investigations, to help us to decide which hypotheses to discard and which
+to consider for more complex and costly studies, or to suggest new
+hypotheses.
+
+\mynote{[Yasser:] In the present work we aim to address the issue of
+  building probability distributions over a large population using
+  recording from subpopulations by considering the distribution of total
+  population activity as an example. [} In the present work we propose such
+an intermediate probability model. It answers this question: How much was
+the \emph{total} activity of a large neuronal population, given the
+observation of the activity of a very small sample thereof? This model
+addresses a larger brain area, avoiding the assumption of isolation of the
+sample; and by focusing on the total activity, rather than the activity of
+individual neurons, it remains simple and numerically tractable.
 
@@
 in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b_r1996,grandy1980}
-has been used for different kinds of estimations of the neuronal activity
-of various brain areas and about other phenomena of importance to the
-neurosciences, for example gene and protein interaction \citep[for
+has been used for different kinds of investigations about the neuronal
+activity of various brain areas and about other phenomena of importance to
+the neurosciences, for example gene and protein interaction \citep[for
 example][]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,mackeetal2009b,tkaciketal2009,roudietal2009c,barreiroetal2010,gerwinnetal2010,mackeetal2011b,ganmoretal2011,cohenetal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}.
@@
 \end{enumerate*}
-For each data set the model gives the most plausible frequency distribution
-with which all levels of total activity of a larger population of 1\,000--10\,000
-neurons appeared during the recording. These guessed frequency
-distributions are distinctly different from the recorded ones for the
-sampled neurons. For the first data set, for example, the frequency
-distribution for the full population has two very distinct modes, both at low
-activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
+For each data set the model gives us the most plausible frequencies with
+which all levels of total activity of a larger population of
+1\,000--10\,000 neurons appeared during the recording. These guessed
+frequency distributions are distinctly different from the recorded ones for
+the sampled neurons. For the first data set, for example, the frequency
+distribution for the larger population has two very distinct modes, both at
+low activities %($0.7\%$ and $2.6\%$ of simultaneously active neurons
 % within $3\;\mathrm{ms}$,
-see \fig~\ref{fig:plots_5mom}, whereas the frequency distribution for the
+(see \fig~\ref{fig:plots_5mom}), whereas the frequency distribution for the
 sample is monotonically decreasing with its maximum at zero activity. The
-frequency distribution for the second data set doesn't have two modes but
-still presents one prominent shoulder in its low-activity mode. Note that
-these guessed features of the full population cannot be inferred by the
-application of maximum-entropy \emph{to the sample alone}. Although we don't
-formulate any hypotheses on why the larger population could have two distinct
-low-activity regimes -- the purpose of the present work is only
-methodological -- these results shows that the proposed method can lead to
-the formulation or preliminary assessment of interesting hypotheses.
-
-We want to stress the usefulness of making a quantified guess about the
-activity of a larger brain area. Such a guess seems indeed to be the
-primary idea behind recording a sample from that area. %  Second, it allows us
-% to make comparisons across experimental sessions; such comparisons would me
-% difficult or meaningless if made with the recorded samples, which generally
-% comprise non-overlapping sets of neurons and differ in size.
-There's also an important methodological reason. Maximum-entropy methods
-have been used to assess the informational sufficiency of pairwise
-correlations and correlations of higher order \citep[see for
+frequency distribution for the second data set presents one prominent
+shoulder in its low-activity mode. These results show that the proposed
+method can lead to the formulation or preliminary assessment of interesting
+hypotheses, as we will illustrate with a toy example. Note that these
+guessed features of the full population could not be inferred by the
+application of maximum-entropy \emph{to the sample alone}.
+
+% We want to stress the usefulness of making a quantified guess about the
+% activity of a larger brain area. Such a guess seems indeed to be the
+% primary idea behind recording a sample from that area. %  Second, it allows us
+% % to make comparisons across experimental sessions; such comparisons would me
+% % difficult or meaningless if made with the recorded samples, which generally
+% % comprise non-overlapping sets of neurons and differ in size.
+% There's also an important methodological reason.
+
+Our approach also solves a methodological problem in the use of
+maximum-entropy methods to assess the \enquote{cooperativity},
+\enquote{interaction}, or \enquote{synchrony} in neuronal activity, for
+example studying its pairwise correlations and correlations of higher order
+\citep[see for
 example][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
-But when the difference in size between sample and full population is too
-large, correlation sufficiency for the sample implies the \emph{lack} of
-correlation sufficiency for the larger population, and vice versa. Therefore,
-maximum-entropy applications at the sample level can be deceptive for
-questions of statistical sufficiency, and a full-population application is
-more reliable. We discuss this point in detail in \sect~\ref{sec:marginalization}.
+When the difference in size between a large population and a sample thereof
+is too large, the presence of some inferential properties of correlations
+for the sample implies the \emph{lack} of such properties for the larger
+population, and vice versa. Maximum-entropy applications at the sample
+level can therefore deceive us in questions regarding the cooperativity of
+the larger population. The application to the larger population is more
+reliable. We discuss this point in detail in
+\sect~\ref{sec:marginalization}.
 
@@
 
-We obviously don't know whether the frequency distribution obtained with
-our approach is the \emph{actual} one which the activity levels of the full
-population had during the recording; it's only the most plausible. In high
+\mynote{review this}We obviously don't know whether the activity levels of
+the larger population really had the frequencies given by the model, during
+the recording. The model gives our most plausible guess. In high
 dimensions, however, the features of the most plausible distribution may
@@
 activities of grid versus non-grid cells? We apply the method to these two
-sets of neurons individually, using $N=4\,150$ for the grid set and
-$N=5\,850$ for the non-grid set, to reflect their proportions in the
-recorded sample. The results are shown in the upper panel of
-\fig~\ref{fig:plots_cellsubsets}. The distribution for the larger
+sets of neurons individually, using $\yNg=4\,150$ for the grid set and
+$\yNng=5\,850$ for the non-grid set, to reflect their proportions (27/65
+and 38/65) in the recorded sample. The results are shown in the upper panel
+of \fig~\ref{fig:plots_cellsubsets}. The distribution for the larger
 population of grid cells (\textcolor{mygreen}{green triangles}) seems to
@@
 (\textcolor{myred}{red circles}) has two peaks instead, roughly at the same
-positions as the full-population distribution (\textcolor{myblue}{blue
-  curve}) but closer in height. So it would seem that the population of
-grid cells is contributing to the first peak of the full population, but it
-isn't the sole contributor. We can also assess if the distributions for the
-two sets are independent. If they were independent, the full-population
-distribution would be given by their convolution:
+normalized activities as the peaks of the full-population distribution
+(\textcolor{myblue}{blue curve}) but closer in height. It would thus seem
+that the population of grid cells is contributing to the first peak of the
+full population, but it is not the sole contributor. We can also assess if
+the distributions for the two sets of neurons are independent. If they were
+independent, the full-population distribution would be given by their
+convolution:
 \begin{equation}
   \label{eq:convolution}
-  P_{\text{full}}(\yA) =  \smashoperator{\sum_{\yA'=\max(0,\yA-5850)}^{\min(\yA,4150)}} P_{\text{grid}}(\yA')\;
-  P_{\text{non-grid}}(\yA-\yA'). %, \qquad \yA \in \set{1,\dotsc,10\,000}
-\end{equation}
-However, the lower panel of
+  P_{\text{full}}(\yA) = \sum_{\yA'} P_{\text{grid}}(\yA')\;
+  P_{\text{non-grid}}(\yA-\yA') %, \qquad \yA \in \set{1,\dotsc,10\,000}
+\end{equation}
+where the index $\yA'$ runs from $\max(0,\yA-\yNng)$ to $\min(\yA,\yNg)$.
+But this is not the case: the lower panel of
 figure~\ref{fig:plots_cellsubsets} shows that such convolution
-(\textcolor{black}{black diamonds}) is different from the full-population
-distribution (\textcolor{myblue}{blue curve}): the two peaks of the former
-are closer in position and height. This means that the population
-distributions of grid and non-grid cells are \emph{not} independent:
-knowledge of the activity of either set gives us some information about the
-activity of the other.
+(\textcolor{black}{black diamonds}) is quite different from the
+full-population distribution (\textcolor{myblue}{blue curve}): the two
+peaks of the former are closer in position and height than those of the
+latter. The population distributions of grid and non-grid cells are
+therefore \emph{not} independent: knowledge of the activity of either set
+gives us some information about the activity of the other.
 
@@
 illustration of the method's possible applications. The important point is
-that this method is computationally very cheap but can provide useful
-insights, even if just qualitatively.
+that this method is computationally very cheap and yet can provide useful
+insights, even if just qualitative ones.
 
@@
   \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
-  \binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}.
+  \binom{n}{n\ya}\binom{N-n}{N \yA-n\ya}\binom{N}{N \yA}^{-1}.
 \end{equation}
@@
 We have already introduced the notation for the sample neurons. We
-introduce an analogous notation for the $\yNv$ neurons constituting the
+introduce an analogous notation for the $N$ neurons constituting the
 larger population, but using the corresponding capital letters:
 $\yAv_{i}(t)$ is the activity of the $i$th neuron at time bin $t$,
-$\yA(t) \defd \sum_{i} \yAv_{i}(t)/\yNv$ is the activity at that
+$\yA(t) \defd \sum_{i} \yAv_{i}(t)/N$ is the activity at that
 bin averaged over the larger population, and so on.
@@
   \label{eq:joint_plaus}
-  \pf(\yAv_1, \yAv_2, \dotsc, \yAv_\yNv \| \yHc)
+  \pf(\yAv_1, \yAv_2, \dotsc, \yAv_N \| \yHc)
   \quad\text{or}\quad
-\pf(\yAA \| \yHc), \quad \yAA \in \set{0,1}^\yNv.
+\pf(\yAA \| \yHc), \quad \yAA \in \set{0,1}^N.
 \end{equation}
@@
   \pf(\yA \| \yH) =
-  \binom{\yNv}{\yNv\yA}^{-1} \pf(\yA \| \yH).
+  \binom{N}{N\yA}^{-1} \pf(\yA \| \yH).
 \end{equation}
@@
 This result is intuitive: owing to symmetry, we must assign equal
-probabilities to all $\tbinom{\yNv}{\yNv\yA}$ activity vectors with
-$\yNv\yA$ active neurons; the probability of each activity vector is
+probabilities to all $\tbinom{N}{N\yA}$ activity vectors with
+$N\yA$ active neurons; the probability of each activity vector is
 therefore given by that of the average activity divided by the number of
@@
   \pf(\ya \|\yA, \yH)=
-  \binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}
+  \binom{n}{n\ya}\binom{N-n}{N \yA-n\ya}\binom{N}{N \yA}^{-1}
   \defs \yG_{\ya\yA},
@@
     ={}\\
-    \binom{\yNv}{m}^{-1}
+    \binom{N}{m}^{-1}
     % \frac{(N-m)!}{N!}
     \sum_{\yA} %m!\,
-    \binom{\yNv \yA}{m}\, \pf(\yA \| \yH)
+    \binom{N \yA}{m}\, \pf(\yA \| \yH)
 \equiv    \E(\av{\underbrace{\yAA \dotsm \yAA}_{\text{$m$ factors}}} \| \yH),
@@
   $%\pf(\ya \|\yA, \yH)=
-\yG_{\ya\yA} \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}$ for $\yNv=5000$, $n=200$. (Band artifacts may appear in the
+\yG_{\ya\yA} \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\ya}\binom{N-n}{N \yA-n\ya}\binom{N}{N \yA}^{-1}$ for $N=5000$, $n=200$. (Band artifacts may appear in the
   colourbar depending on your \textsc{pdf} viewer.)}
@@
 that it is a sort of \enquote{fuzzy identity transformation}, or fuzzy
-Kronecker delta, between the $\yA$-space $\set{0,\dotsc,\yNv}$ and
+Kronecker delta, between the $\yA$-space $\set{0,\dotsc,N}$ and
 $\ya$-space $\set{0,\dotsc,n}$. From \eqn~\eqref{eq:marginal} we thus have
@@
   sample is representative of our uncertainty about the population and
-  about other samples}, and vice versa. When $n=\yNv$, $\yG_{\ya\yA}$
+  about other samples}, and vice versa. When $n=N$, $\yG_{\ya\yA}$
 becomes the identity matrix and the approximate equalities above become
@@
 % $\pf(\yxs=a) \to \pf(\yXf=a)$ and $\expeb{\yg_{\yxs}} \to \expeb{\yg_{\yXf}}$, as
-% $n\to \yNv$, do: these limits are trivially, universally valid because the
-% sample becomes the full population as $n\to \yNv$. In particular, these limits
+% $n\to N$, do: these limits are trivially, universally valid because the
+% sample becomes the full population as $n\to N$. In particular, these limits
 % hold even in cases where the conditional probability
@@
   \pf(\yA \| \yHa)  = \frac{1}{\yK(\yll)}\,
-  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
+  \binom{N}{N \yA}\,\exp\bigl(\yll\T
   \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
@@
 \,
-  \binom{\yNv}{\yNv \yA}\,\exp\bigl(\yll\T
+  \binom{N}{N \yA}\,\exp\bigl(\yll\T
   \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
@@
   \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:app_maxent_sample}, with $\yNv=5000$,
+  circles), \eqn~\eqref{eq:app_maxent_sample}, with $N=5000$,
   $n=200$, and constraints as in \eqn~\eqref{eq:constraints}.}
@@
   population level, \eqn~\eqref{eq:app_maxent_pop_marg} with
-  $\yNv=10\,000$, followed by sample marginalization;
+  $N=10\,000$, followed by sample marginalization;
   \textcolor{myyellow}{\textbf{yellow circles:}} \me\ at the population
@@
 discrepancy in \fig~\ref{fig:all_three}, for
-$\yNv=10\,000$, $n=200$, and the two constraints
+$N=10\,000$, $n=200$, and the two constraints
 \begin{equation}
@@
 
-In our discussion we have so far assumed the size $\yNv$ of the larger
+In our discussion we have so far assumed the size $N$ of the larger
 population to be known. This is rarely the case, however. We usually are
-uncertain about $\yNv$ and can only guess its order of magnitude. In such a
-state of knowledge $\yHd$ our ignorance about the possible value of $\yNv$
-is expressed by a probability distribution $\pf(\yNN=\yNv \|\yHd)=h(\yNv)$,
+uncertain about $N$ and can only guess its order of magnitude. In such a
+state of knowledge $\yHd$ our ignorance about the possible value of $N$
+is expressed by a probability distribution $\pf(\yNN=N \|\yHd)=h(N)$,
 and the marginal distribution for the sample
@@
   \pf(\ya \| \yHd)  =
-  \sum_{\yNv} \pf(\ya \| \yNv,\yHd) \,
-  \pf(\yNv \|\yHd)
+  \sum_{N} \pf(\ya \| N,\yHd) \,
+  \pf(N \|\yHd)
   ={}\\[-\jot]
-  \sum_\yNv \biggl\{\frac{1}{\yK(\yll_\yNv)}\, 
-  \sum_{\yA} \yG^{(\yNv)}_{\ya\yA}
+  \sum_N \biggl\{\frac{1}{\yK(\yll_N)}\, 
+  \sum_{\yA} \yG^{(N)}_{\ya\yA}
 \,
-  \tbinom{\yNv}{\yNv \yA}\,\exp\Bigl[{\yll_\yNv}\T
-  \tsum_{\ya} \yg_{\ya}\yG^{(\yNv)}_{\ya\yA}\Bigr]\biggr\}
-  \;h(\yNv),
+  \tbinom{N}{N \yA}\,\exp\Bigl[{\yll_N}\T
+  \tsum_{\ya} \yg_{\ya}\yG^{(N)}_{\ya\yA}\Bigr]\biggr\}
+  \;h(N),
 \end{multline}
-where the Lagrange multipliers $\yll_{\yNv}$ and the summation range for
-$\yA$ depend on $\yNv$.
+where the Lagrange multipliers $\yll_{N}$ and the summation range for
+$\yA$ depend on $N$.
 
@@
 (yellow circles) for the same constraints as above, and a probability
-distribution for $\yNv$ inspired by Jeffreys
+distribution for $N$ inspired by Jeffreys
 \citep[\sect~4.8]{jeffreys1939_r1983}:
@@
   \label{eq:pdf_popsize}
-  h(\yNv) \propto 1/\yNv, \qquad
-  \yNv \in\set{1\,000,\; 2\,000,\; \dotsc,\;10\,000}.
+  h(N) \propto 1/N, \qquad
+  N \in\set{1\,000,\; 2\,000,\; \dotsc,\;10\,000}.
 \end{equation}
@@
 \eqn~\eqref{eq:app_maxent_sample}, yields different results because it does
-not make assumptions about the size $\yNv$ of a possibly existing larger
+not make assumptions about the size $N$ of a possibly existing larger
 population. Such a state of uncertainty, however, is correctly formalized
 according to the laws of probability by introducing a probability
-distribution for $\yNv$, and is expressed by
+distribution for $N$, and is expressed by
 \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}. This expression cannot
 generally be equal to~\eqref{eq:app_maxent_sample} unless the distribution
-for $\yNv$ gives unit probability to $\yNv=n$; that is, unless the sample
+for $N$ gives unit probability to $N=n$; that is, unless the sample
 \emph{is} the full population, and no larger population exists.

### 2019-11-26 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-26T10:47:14+0100
+%% Last-Updated: 2019-11-26T12:30:07+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 For each data set the model gives us the most plausible frequencies with
-which all levels of total activity of a larger population of
-1\,000--10\,000 neurons appeared during the recording. These guessed
-frequency distributions are distinctly different from the recorded ones for
-the sampled neurons. For the first data set, for example, the frequency
-distribution for the larger population has two very distinct modes, both at
-low activities %($0.7\%$ and $2.6\%$ of simultaneously active neurons
+which all levels of total activity of a much larger population,
+1\,000--10\,000 neurons, appeared during the recording. For example it can
+tell us that 60 out of 10\,000 neurons were likely active during 1\% of the
+recording time (though not necessarily always the same 60), 250 neurons out
+of 10\,000 were active during 0.4\% of the time, and so on. The precise
+meaning of this frequency distribution is explained in
+\sect~\ref{sec:model}. For the two example data sets, the guessed frequency
+distributions in the larger population are distinctly different from those
+in the sample. For the first data set the frequency distribution for the
+larger population has two very distinct modes, both at low
+activities %($0.7\%$ and $2.6\%$ of simultaneously active neurons
 % within $3\;\mathrm{ms}$,
@@
 
-\section{The approach: maximum-entropy and sampling}
+\section{Method: maximum-entropy and sampling}
 \label{sec:model}
 
-Let's introduce some context and notation for our problem.
+Let's introduce some context and mathematical notation.
 
@@
 have recorded the spiking activities of $n$ neurons for a certain amount of
-time. These neurons are our \enquote{sample} or \enquote{subpopulation}. Their
-spikes are binned into $T$ time bins and binarized to $\set{0,1}$ values in
-each bin. Call $\ya_{t}$ the number of neurons that fire during time bin
-$t$: this is the \emph{total activity} of the sample, or just
-\enquote{activity} for short. Obviously $\ya_{t} \in\set{0,1,\dotsc,n}$; if
-$\ya_{t}=0$, no neuron spikes during bin $t$; if $\ya_{t}=n$, all spike at
-some point during bin $t$. For brevity, let's say \enquote{at $t$} for
-\enquote{during time bin $t$}. If we divide the total activity by the
-population size we have the normalized total activity or
-population-averaged activity $\ya/n$, ranging from $0$ to $1$ in $1/n$
-steps. From the activities $\set{\ya_{t}}$ we can count how often the
-activity levels $\ya=0$, $\ya=1$, and so on appeared during the recording,
-obtaining the distribution of measured relative frequencies
-$(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
-bins \emph{outside} of the recorded period. Such activity is unknown to us,
-of course.
+time. These neurons are our \enquote{sample}. Their spikes are binned into
+$T$ time bins and binarized to $\set{0,1}$ values in each bin. Call
+$\ya_{t}$ the number of neurons that fire during time bin $t$: this is the
+\emph{total activity} of the sample, or just \enquote{activity} for short.
+Obviously $\ya_{t} \in\set{0,1,\dotsc,n}$; if $\ya_{t}=0$, no neuron spikes
+during bin $t$; if $\ya_{t}=n$, all spike at some point during bin $t$, and
+so on. For brevity, let's say \enquote{at $t$} for \enquote{during time bin
+  $t$}. If we divide the total activity by the population size we have the
+\emph{normalized total activity} or population-averaged activity $\ya/n$,
+ranging from $0$ to $1$ in $1/n$ steps. From the activities $\set{\ya_{t}}$
+we can count how often the activity levels $\ya=0$, $\ya=1$, and so on
+appeared during the recording, obtaining the distribution of measured
+relative frequencies $(\yff_{\ya}) \defs \yf$. We can also consider the
+(unknown) sample activity at time bins \emph{outside} of the recorded
+period.
 
@@
 are not specifically chosen from among the rest, owing to several limiting
-factors; for example, limitations in how precisely electrodes are inserted
-or neurons are targeted by viruses. The set of recorded neurons may even
-change slightly across experimental sessions that are very far apart in
-time. We assume that there's an area, comprising a population of $N$ neurons,
-for which we believe that any other sample of size $n$ could have equally
-plausibly been recorded instead of the sample of $n$ neurons that was
-actually recorded. We call this larger population the \enquote{full population}.
-The total activity of these $N$ neurons at $t$ is $\yA_{t}$. The relative
+factors; for example, limitations in how precisely electrodes are inserted.
+The sample of recorded neurons may even change slightly across experimental
+sessions that are very far apart in time. We assume that there's an area,
+comprising a population of $N$ neurons, for which we believe that any other
+sample of size $n$ could have equally plausibly been recorded instead of
+the sample of $n$ neurons that was actually recorded. This is what we will
+mean with \enquote{larger population}. This population need not be a whole
+functionally or anatomically distinct region. Loosely speaking it is the
+area of which we believe our sample to be \enquote{representative}
+\citep[with the warnings that accompany that term:][]{kruskaletal1979c}.
+
+The total activity of the $N$ neurons at $t$ is $\yA_{t}$. The relative
 frequencies of the various activity levels during the recording were
@@
 or the frequency distribution $\yF$. We only know for certain that
-$\yA_{t} \in \set{0, 1, \dotsc, N}$ and that $\yA_{t} \ge \ya_{t}$ for
-obvious reasons. For the time being we assume that we know $N$; in
-\sect~\ref{sec:N} we discuss the consequences of our lack of precise
-knowledge about this number.
+$\yA_{t} \in \set{0, 1, \dotsc, N}$, that $\yA_{t} \ge \ya_{t}$, and that
+$N-\yA_{t} \ge n-\ya_{t}$ for obvious reasons. For the time being we assume
+that we know $N$; in \sect~\ref{sec:N} we discuss the consequences of our
+lack of precise knowledge about this number.
 
 Our questions concern general features of the total activity $\yA$ of the
-full population during and after the recording, and across sessions under the
-same study conditions. For example: what was its frequency distribution
-during the recording? How much does this frequency distribution change
-across sessions? How much total activity should we expect at any time bin
-during a recording? The approach presented here gives a probability
-distribution that approximately answers all these questions.
+larger population during and after the recording, and across sessions under
+the same study conditions. For example: what was its frequency distribution
+$\yF$ during the recording? How much does this frequency distribution
+change across sessions? How much total activity should we expect at any
+time bin during a recording? The approach presented here gives a
+probability distribution that approximately answers all these questions.
 
@@
   total activity $\ya$ of the sample and corresponding expected values $M$
-  for the total activity $\yA$ of the full population.
+  for the total activity $\yA$ of the larger population.
 \item\label{item:maxent_step} Using the maximum-entropy method we build a
-  distribution $\px(\yA \| M, N)$ for the total activity of the full
+  distribution $\px(\yA \| M, N)$ for the total activity of the larger
   population of $N$ neurons, using as constraints the expected values $M$
@@
 application of the probability calculus, which gives an exact linear
-relation between the first $m$ moments for the full population and the first
+relation between the first $m$ moments for the larger population and the first
 $m$ for the sample \citep[\eqns~(16)]{portamanaetal2015}. The ones
 determine the others and vice versa at every time bin. This relation holds
-for any belief distribution $\pxxx(\yA_{t})$ for the full-population activity
+for any belief distribution $\pxxx(\yA_{t})$ for the larger-population activity
 and its marginal $\pxx(\ya_{t})$ for the sample activity at that bin.
@@
 analysis: \emph{the first $n$ normalized factorial moments for the sample
-  and for the full population are numerically identical}:
+  and for the larger population are numerically identical}:
 \begin{equation}
@@
 \end{equation}
-% where $\pxxx(\yA)$ is our distribution of belief about the full-population
+% where $\pxxx(\yA)$ is our distribution of belief about the larger-population
 % activity, and $\pxx$
@@
 $\binom{\yA}{m}$ to equate to the factorial moment
-$\E\bigl[\binom{\yA}{m}\bigr]$ of the full-population distribution
+$\E\bigl[\binom{\yA}{m}\bigr]$ of the larger-population distribution
 $\pxxx(\yA)$. But \eqn~\eqref{eq:equal_m_factorialmoments} of
 step~\ref{item:sample_step} comes to our rescue, because it says that the
-expectations for the full population are determined by those for the sample
+expectations for the larger population are determined by those for the sample
 $\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time averages of the
@@
   \rightarrow
-    \text{full-population moments}}_{\textit{sampling theory}}
+    \text{larger-population moments}}_{\textit{sampling theory}}
 \end{equation*}
 and apply the maximum-entropy method to obtain a distribution
-$\px(\yA \| M, N)$ for the full population of $N$ neurons by constraining some
+$\px(\yA \| M, N)$ for the larger population of $N$ neurons by constraining some
 of its factorial moments, for example $M=\set{1,\dotsc,m'}$ with $m'\le n$,
@@
 $\yA$ can be equally likely be generated by any set of $\yA$ neurons in the
-full population. The magnitude of this discrepancy can be a signature of the
+larger population. The magnitude of this discrepancy can be a signature of the
 (at least temporary) presence of a \enquote{neuronal assembly}
@@
 approximation)\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}
-of four different distributions for the full population, which differ
+of four different distributions for the larger population, which differ
 numerically from one another in higher-order approximations:
@@
 % can be considered as \emph{the most probable frequency distribution} for
-% the full-population activity during the recording. But it also approximates
+% the larger-population activity during the recording. But it also approximates
 % three other distributions, as discussed in the previous section.
@@
 
-The full-population maximum-entropy distribution is calculated using five
+The larger-population maximum-entropy distribution is calculated using five
 moments. This number seems to provide almost as much information as the
@@
 Figure~\ref{fig:plots_5mom} shows the resulting densities
-(distribution${}\times N$) for three example values of full-population
+(distribution${}\times N$) for three example values of larger-population
 sizes: $N=1\,000$ (\textcolor{mygreen}{green diamonds}), $N=5\,000$
@@
 
-The figure shows that the distribution for the full-population is more
+The figure shows that the distribution for the larger-population is more
 peaked than the measured frequency distribution for the sample; their
 difference increases with $N$. Most remarkably, for the first data set
-(upper panel) the distribution for the full population has two distinct
+(upper panel) the distribution for the larger population has two distinct
 low-activity modes. For the second data set (lower panel) the distribution
@@
 maximum-entropy distribution at the sample level. The application of the
-probability calculus thus reveals interesting possible features of the full
+probability calculus thus reveals interesting possible features of the larger
 population. \textcolor{white}{If you find this you can claim a postcard from
@@
 the hypothesis that two distinct cell types or assemblies be present in the
-region where the recording was made. Finding a full-population distribution
+region where the recording was made. Finding a larger-population distribution
 with two peaks, as in the upper panel of \fig~\ref{fig:plots_5mom}, would
@@
 population of grid cells (\textcolor{mygreen}{green triangles}) seems to
-have one, broad peak, close to the first peak of the full-population
+have one, broad peak, close to the first peak of the larger-population
 distribution. The distribution for the larger population of non-grid cells
 (\textcolor{myred}{red circles}) has two peaks instead, roughly at the same
-normalized activities as the peaks of the full-population distribution
+normalized activities as the peaks of the larger-population distribution
 (\textcolor{myblue}{blue curve}) but closer in height. It would thus seem
 that the population of grid cells is contributing to the first peak of the
-full population, but it is not the sole contributor. We can also assess if
+larger population, but it is not the sole contributor. We can also assess if
 the distributions for the two sets of neurons are independent. If they were
-independent, the full-population distribution would be given by their
+independent, the larger-population distribution would be given by their
 convolution:
@@
 (\textcolor{black}{black diamonds}) is quite different from the
-full-population distribution (\textcolor{myblue}{blue curve}): the two
+larger-population distribution (\textcolor{myblue}{blue curve}): the two
 peaks of the former are closer in position and height than those of the
@@
 \enquote{cooperativity} or \enquote{interaction},
-\fig~\ref{fig:plots_2_4mom} shows the full-population distributions, for
+\fig~\ref{fig:plots_2_4mom} shows the larger-population distributions, for
 $N=10\,000$, obtained constraining the first two moments (in
@@
 correlations seem informationally sufficient for a particular sample from a
-brain area, then they may well not be sufficient for the full population of
+brain area, then they may well not be sufficient for the larger population of
 neurons constituting that area, and vice versa. So if what interests us is
@@
 The approach presented here avoids this problem, because the
-maximum-entropy method is applied to obtain the distribution of the full
+maximum-entropy method is applied to obtain the distribution of the larger
 population, not of the sample alone.
@@
 \begin{enumerate}[label=(\roman*)]
-\item from each maximum-entropy distributions $\px(\yA \| M)$ for the full
+\item from each maximum-entropy distributions $\px(\yA \| M)$ for the larger
   population, built from each set of constraints $M=M',M''$, calculate the
@@
 $M_{4}$, $M_{5}$, consisting of the first two, four, five moments. For a
-full population of size $N=10\,000$, we obtain the following differences:
+larger population of size $N=10\,000$, we obtain the following differences:
 \begin{equation}
@@
 
-The application at the full-population level therefore leads to different
+The application at the larger-population level therefore leads to different
 conclusions about the relative importance of the moments from the
@@
 $\yf$ conditional on the hypothesis $M_{5, N}$ of 5-moment sufficiency at
-the full-population level, $N=10\,000$, versus the hypothesis $M_{5, n}$ of
+the larger-population level, $N=10\,000$, versus the hypothesis $M_{5, n}$ of
 5-moment sufficiency at the sample level, $N=n=65$. We obtain
@@
 The conclusions about \enquote{cooperativity} or \enquote{interaction}
-that we reach by applying the maximum-entropy method to the full population,
+that we reach by applying the maximum-entropy method to the larger population,
 as proposed here, are not only different from the application at the sample
@@
 The same plot also shows the marginals obtained from applying the method to
-a full population with $N=10\,000$ and the same constraints, that is, the
+a larger population with $N=10\,000$ and the same constraints, that is, the
 sample marginals of \fig~\ref{fig:plots_2_4mom}. The two-moment
@@
 filled blue squares (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The
-full-population distribution fits the measured frequencies
+larger-population distribution fits the measured frequencies
 (\textcolor{black}{$\bm{\triangledown}$}) very well if we constrain only four
@@
 
-% But does the application at the full-population level lead to appreciably
+% But does the application at the larger-population level lead to appreciably
 % different results from that at the sample level? after all we're interested
@@
 % question in the case of our first data set. The upper plot shows the
-% maximum-entropy distributions for a full population of $10\,000$ neurons,
+% maximum-entropy distributions for a larger population of $10\,000$ neurons,
 % constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
@@
 % (\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained
-% by marginalizing the full-population distribution to the sample size, from two
+% by marginalizing the larger-population distribution to the sample size, from two
 % moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
@@
 % \begin{itemize}
-% \item The application to the full-population (upper plot) leads to two
+% \item The application to the larger-population (upper plot) leads to two
 %   completely different distributions (even different number of modes);
@@
 %   two statistics are approximately equivalent.
-% \item The difference between marginals from the full-population distribution
+% \item The difference between marginals from the larger-population distribution
 %   and the distributions obtained at the sample level (filled vs empty
@@
 %   (triangles vs squares).
-% \item The marginals of the application to the full population are closer to
+% \item The marginals of the application to the larger population are closer to
 %   the measured frequencies than the distributions obtained at the sample
@@
 % Therefore, the maximum-entropy application at the sample level and at the
-% full-population level lead to different results; the latter is not only more
+% larger-population level lead to different results; the latter is not only more
 % meaningful but also superior because it shows clearly the different
@@
 
-\section{The full-population size $N$}
+\section{The larger-population size $N$}
 \label{sec:N}
@@
 The calculations and conclusions presented in the preceding sections depend
-on the size of the full population, $N$. The full population can't be the whole
+on the size of the larger population, $N$. The larger population can't be the whole
 brain, of course. How large should or can $N$ be in our formulae?
@@
 our method is based, and which asserts the equality of the factorial
-moments of the full population and those of the sample. This formula is only
+moments of the larger population and those of the sample. This formula is only
 valid if our beliefs about the activity of the sample $\pxx(\ya)$ and that
-of the full population $\pxx(\yA)$ are related by a hypergeometric
+of the larger population $\pxx(\yA)$ are related by a hypergeometric
 distribution:
@@
 The second point is that our degrees of belief about the frequency
-distribution for the full population depend not only on the measured data in
+distribution for the larger population depend not only on the measured data in
 the sample, but also on our pre-data beliefs $\yI$ about the distribution.
@@
 Here is a summary derivation of the maximum-entropy distribution for the
-full population constrained by a set of factorial moments. For further details
+larger population constrained by a set of factorial moments. For further details
 see Porta Mana \etal\ \citey{portamanaetal2015}.
@@
 \end{equation}
-In our case the units are neurons, the set is the full population, and the
+In our case the units are neurons, the set is the larger population, and the
 property is their being active in a specific time bin.
@@
 can prove the equality~\eqref{eq:equal_m_factorialmoments} between the
-factorial moments of sample and full population.
+factorial moments of sample and larger population.
 
@@
 Now, the probability of observing activity $\ya$ in the sample at any time
-bin is the sample marginal of the maximum-entropy distribution for the full
+bin is the sample marginal of the maximum-entropy distribution for the larger
 population, owing to the excheangeability assumption implicit in the
@@
   \emph{expect} -- in the full probabilistic sense -- to see the same
-  proportion of political preferences in a sample as in the full
+  proportion of political preferences in a sample as in the larger
   population}.
@@
 that expectation equalities such as~\eqref{eq:poll_equal_expectations} hold
-for the total activity $S(t)$ of the full population of $N$ neurons that
+for the total activity $S(t)$ of the larger population of $N$ neurons that
 could have been recorded.
@@
   neurons we can construct maximum-entropy distributions for the total
-  activity of the \textbf{full population} of neurons that could have been
+  activity of the \textbf{larger population} of neurons that could have been
   sampled by the recording device}.
@@
 % specific statistics about the activity of a recorded \emph{sample} of the
-% full population. This kind of probability assignment is extremely important
+% larger population. This kind of probability assignment is extremely important
 % in neuroscience: we try to understand the workings of brain areas that may
@@
 % hidden assumption. We'll study in which circumstances the application of
-% the method at the sample level and at the full-population level give
+% the method at the sample level and at the larger-population level give
 % similar or different probabilities.
@@
 
-We need to make an initial probability assignment for the state of the full
+We need to make an initial probability assignment for the state of the larger
 population before any experimental observations are made. This initial
@@
 
-Our uncertainties about the full population and the sample are connected
+Our uncertainties about the larger population and the sample are connected
 via the conditional probability
@@
 sample~\eqref{eq:subpop_average} indeed satisfies the same representation
-theorem~\eqref{eq:joint_plaus_N_homog} as the distribution for the full
+theorem~\eqref{eq:joint_plaus_N_homog} as the distribution for the larger
 population.
@@
 function of the sample activity, $\yg_{\ya}$, in terms of the distribution
-for the full population, as follows:
+for the larger population, as follows:
 \begin{equation}
@@
 sample, have a  value $\yc$ obtained from observed data,
-\emph{translates into a linear constraint for the distribution of the full
+\emph{translates into a linear constraint for the distribution of the larger
   population}:
@@
 that is, \emph{the expected values of the $m$-activities of the sample and
-  of the full population are equal}. The proof of the middle equality uses
+  of the larger population are equal}. The proof of the middle equality uses
 the expression for the $m$th factorial moment of the hypergeometric
@@
 $m$-activity of a sample to have a particular value, we are also
-requiring  the expectation of the $m$-activity of the full population to
+requiring  the expectation of the $m$-activity of the larger population to
 have the same value.
@@
 \end{figure}%hypergeometric_identity.nb
-These expectation equalities between sample and full population should not
+These expectation equalities between sample and larger population should not
 be surprising: we intuitively \emph{expect} that the proportion of coloured
@@
 becomes the identity matrix and the approximate equalities above become
-exact -- of course, since we have sampled the full population.
+exact -- of course, since we have sampled the larger population.
 
@@
 % $n\to N$, do: these limits are trivially, universally valid because the
-% sample becomes the full population as $n\to N$. In particular, these limits
+% sample becomes the larger population as $n\to N$. In particular, these limits
 % hold even in cases where the conditional probability
@@
 
-\section{Maximum-entropy: sample level \vs\ full-population level}
+\section{Maximum-entropy: sample level \vs\ larger-population level}
 \label{sec:specific_initial_probability}
@@
 In the previous section we have seen that observations about a sample can
-be used as constraints on the distribution for the activity of the full
+be used as constraints on the distribution for the activity of the larger
 population. Let us use such constraints with the \me\ method. Suppose that
@@
 directly to sampled neurons; denote this approach by $\yHb$. Then we apply
-the method to the full population of neurons, most of which are unsampled;
+the method to the larger population of neurons, most of which are unsampled;
 denote this approach by $\yHa$.
@@
 
-Applied to the full population, using the constraint
+Applied to the larger population, using the constraint
 expression~\eqref{eq:constraint_extended} derived in the previous section,
-the method yields the distribution for the full-population activity
+the method yields the distribution for the larger-population activity
 \begin{equation}
@@
 Hence, \emph{the applications of \me\ at the sample level and at the
-  full-population level are inequivalent}. They lead to numerically
+  larger-population level are inequivalent}. They lead to numerically
 different distributions for the sample activity $\yaa$.
@@
 assuming an entropic pre-data distribution for the long-run frequencies of
-the full population:
-
-\textbf{The frequency distribution $\yF$ of the full-population activity
+the larger population:
+
+\textbf{The frequency distribution $\yF$ of the larger-population activity
   \emph{during the recording}}
@@
 
-\textbf{The long-run frequency distribution $\ynu$ of the full-population
+\textbf{The long-run frequency distribution $\ynu$ of the larger-population
   activity}
@@
 
-\textbf{The activity $\yA'$ of the full-population
+\textbf{The activity $\yA'$ of the larger-population
   in a \emph{new} time bin}
@@
 for $N$ gives unit probability to $N=n$; that is, unless the sample
-\emph{is} the full population, and no larger population exists.
+\emph{is} the larger population, and no larger population exists.
 
@@
 neuroscientific applications, the difference in predictions of the sample
-\vs\ full-population applications can instead be very relevant.
+\vs\ larger-population applications can instead be very relevant.
 
@@
 the distribution of the activity in the population level model that do not
-show up in the sample level. We study how the assumed size of the full
+show up in the sample level. We study how the assumed size of the larger
 population affects the appearance of these modes and find that there is a
-minimum size of the full population for which such modes can be observed. We
-then compared the distribution found by marginalizing the full population
+minimum size of the larger population for which such modes can be observed. We
+then compared the distribution found by marginalizing the larger population
 maximum entropy model down to the sample level, and the distribution fit
@@
 TEST SHI) This predicts that for a large enough population (DO WE PREDICT
-THAT IF THE FULL POPULATION GETS BIGGER THE DIFFERENCE ALSO GET BIGGER)?…
+THAT IF THE LARGER POPULATION GETS BIGGER THE DIFFERENCE ALSO GET BIGGER)?…
 
 The rest of the paper is organized as follows. We first describe how to go
-from the sample level maximum entropy model to the full population maximum
+from the sample level maximum entropy model to the larger population maximum
 entropy model. In section 2, we apply this to the two experimental datasets

### 2019-11-26 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-26T12:30:07+0100
+%% Last-Updated: 2019-11-26T13:28:00+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 area of which we believe our sample to be \enquote{representative}
-\citep[with the warnings that accompany that term:][]{kruskaletal1979c}.
+\citep[with the warnings that accompany that
+term:][]{kruskaletal1979b,kruskaletal1979c}.
 
@@
   relation between some expected values -- specifically, moments -- for the
-  total activity $\ya$ of the sample and corresponding expected values $M$
+  total activity $\ya$ of the sample and corresponding expected values %$M$
   for the total activity $\yA$ of the larger population.
@@
 \end{enumerate}
-We now discuss the ideas behind these steps more in detail, leaving their
+We now discuss the ideas behind these steps more in detail, but leave their
 precise mathematical implementation and a more detailed list of references
@@
 
-Step~\ref{item:sample_step}, a result of sampling theory, is just an
-application of the probability calculus, which gives an exact linear
-relation between the first $m$ moments for the larger population and the first
-$m$ for the sample \citep[\eqns~(16)]{portamanaetal2015}. The ones
-determine the others and vice versa at every time bin. This relation holds
-for any belief distribution $\pxxx(\yA_{t})$ for the larger-population activity
-and its marginal $\pxx(\ya_{t})$ for the sample activity at that bin.
+Step~\ref{item:sample_step} is just an application of the probability
+calculus, which gives an exact linear relation between the first $m$
+moments for the larger population and the first $m$ for the sample
+\citep[\eqns~(16)]{portamanaetal2015}. The ones determine the others and
+vice versa at every time bin. This relation holds for any belief
+distribution $\pxxx(\yA_{t})$ for the larger-population activity and its
+marginal $\pxx(\ya_{t})$ for the sample activity at that bin.
 
@@
 that is, the expected number of distinct $m$-tuples of simultaneously
-spiking neurons (within a time-bin's width), normalized by the maximum
+active neurons (within a time-bin's width), normalized by the maximum
 possible number of distinct $m$-tuples. % For
@@
 % equivalent to specifying the other set.
-But the normalized factorial moments have a convenient property for our
-analysis: \emph{the first $n$ normalized factorial moments for the sample
-  and for the larger population are numerically identical}:
+But the normalized factorial moments have a mathematically convenient
+property for our analysis: \emph{the first $n$ normalized factorial moments
+  for the sample and for the larger population are numerically identical}:
 \begin{equation}
@@
 \enquote{maximum-entropy} for brevity. It amounts to two prescriptions:
-choose the distribution, among those satisfying specific convex constraints
-such as fixed expectations, that has minimum relative entropy with respect
-to a reference distribution; and judge those expectations to be equal to
-some measured averages, typically time averages.
-
-In our case we don't have measured time averages of the quantity
-$\binom{\yA}{m}$ to equate to the factorial moment
+first, take the distributions satisfying specific convex constraints --
+such as fixed expectations -- and among them select the one having minimum
+relative entropy with respect to a reference distribution; second, judge
+those expectations to be equal to some measured averages, typically time
+averages.
+
+In our case we don't know the time averages of the quantity
+$\binom{\yA}{m}$, so we cannot directly equate them to the factorial moment
 $\E\bigl[\binom{\yA}{m}\bigr]$ of the larger-population distribution
@@
 step~\ref{item:sample_step} comes to our rescue, because it says that the
-expectations for the larger population are determined by those for the sample
-$\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time averages of the
+expectation for the larger population are determined by that for the sample
+$\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time average of its
 corresponding sample quantity $\binom{\ya}{m}$. So we can combine the two
@@
 \end{equation*}
-and apply the maximum-entropy method to obtain a distribution
-$\px(\yA \| M, N)$ for the larger population of $N$ neurons by constraining some
-of its factorial moments, for example $M=\set{1,\dotsc,m'}$ with $m'\le n$,
-to be equal to the sample's recorded averages. In formulae, the constraints
-on $\pxxx(\yA)$ are
+We obtain a distribution $\px(\yA \| M, N)$ for the larger population of
+$N$ neurons by constraining some of its factorial moments, for example
+$M=\set{1,\dotsc,m'}$ with $m'\le n$, to be equal to the sample's recorded
+averages. In formulae, the constraints on $\pxxx(\yA)$ are
 \begin{equation}
@@
 
-The result is the
-distribution of the form
+The result is the distribution of the form
 \begin{equation}
@@
 where the parameters $\yll$ are determined by the
-constraints~\eqref{eq:final_constraint}. It is further discussed in
-appendix~\ref{sec:derivation_maxent_fullnet}.
-
-The number and degrees of the constraining moments depend on the questions
-and hypotheses that a researcher is exploring; for example, hypotheses
-about the \enquote{cooperativity} or \enquote{interaction} in the population
-activity. We discuss this kind of use in \sect~\ref{sec:marginalization}.
-Note that the constraint equation~\eqref{eq:final_constraint} for the
+constraints~\eqref{eq:final_constraint}. This formula is further discussed
+in appendix~\ref{sec:derivation_maxent_fullnet}.
+
+The amount and degrees of the constraining moments depend on the questions
+and hypotheses that a researcher is exploring. We give some examples in the
+next two sections. Note that the constraint
+equation~\eqref{eq:final_constraint} for the
 distribution~\eqref{eq:app_maxent_pop_first} may not have exact solutions
 if $N$ is strictly larger than $n$; and the possible discrepancy typically
-increases with the number of constraints. The essential reason for this
-discrepancy is the approximate character of two assumptions behind the
-maximum-entropy and of one assumption specific to our application: measured
-moments equal expectations (or equivalently, the number of time bins $T$ is
-infinite), the order of the time bins is irrelevant, and an activity level
-$\yA$ can be equally likely be generated by any set of $\yA$ neurons in the
-larger population. The magnitude of this discrepancy can be a signature of the
-(at least temporary) presence of a \enquote{neuronal assembly}
-\citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss this matter
-in \sect~\ref{sec:assumptions}.
+increases with the number of constraints. This discrepancy comes from the
+approximate character of two assumptions behind the maximum-entropy method
+and of one assumption specific to our application: that the order of the
+time bins is irrelevant, that measured moments equal expectations (or
+equivalently, the number of time bins $T$ is infinite), and that an
+activity level $\yA$ can equally likely be generated by any set of $\yA$
+neurons in the larger population. The magnitude of this discrepancy can be
+a signature of the (at least temporary) presence of a \enquote{neuronal
+  assembly} \citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss
+this matter in \sect~\ref{sec:assumptions}.
 
@@
 First, a maximum-entropy distribution like $\px(\yA \| M, N)$ is the
-zeroth-order approximation (in the sense of Laplace's
-approximation)\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}
-of four different distributions for the larger population, which differ
+zeroth-order approximation (in the sense of Laplace's approximation
+\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}) of
+four distinct distributions for the larger population, which differ
 numerically from one another in higher-order approximations:

### 2019-11-26 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-26T13:28:00+0100
+%% Last-Updated: 2019-11-26T13:40:59+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 the results and are further discussed in \sect~\ref{sec:assumptions}.
-First, a maximum-entropy distribution like $\px(\yA \| M, N)$ is the
-zeroth-order approximation (in the sense of Laplace's approximation
+First, the maximum-entropy
+method % (based on the \emph{Shannon} relative
+% entropy, as opposed to other entropy functions such as Burg's
+% \citeyear{burg1975})
+rests on some implicit assumptions about the probabilities for the long-run
+frequency distribution of activities, besides the assumptions just
+mentioned at the end of the previous paragraph. So the often-heard
+statement that it gives \enquote{the maximally unbiased (or non-committal)
+  distribution} must be taken with a grain of salt. \citep[\mynote{move to
+  sect}]{jaynes1986d_r1996,portamana2009,portamana2017}. Second, a
+maximum-entropy distribution like $\px(\yA \| M, N)$ is the zeroth-order
+approximation (in the sense of Laplace's approximation
 \citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}) of
@@
 \item the \emph{belief} distribution for the value of the total
-  activity at any time bin among those \emph{recorded},
+  activity at any time bin among the \emph{recorded} ones,
   %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
@@
 The validity of the maximum-entropy approximation decreases as the ratio
-$N/T$ increases. Second, the maximum-entropy
-method % (based on the \emph{Shannon} relative
-% entropy, as opposed to other entropy functions such as Burg's
-% \citeyear{burg1975})
-rests on some implicit assumptions about the probabilities for the long-run
-frequency distribution of activities, besides the assumptions mentioned at
-the end of the previous paragraph. So the often heard statement that it
-gives \enquote{the maximally unbiased (or non-committal) distribution} must
-be taken with a grain of salt. \citep[\mynote{move to
-  sect}]{jaynes1986d_r1996,portamana2009,portamana2017}.
-
-\bigskip
-
-In the next section we apply the method just described to the data sets
-from two actual recordings and discuss the properties of the resulting
-distributions.
+$N/T$ increases.
 
@@
 \item The \textbf{first data set}, from \textcite[rat
-  14147]{stensolaetal2012}, consists of $n=65$ neurons (27 of which
-  classified as grid cells) from rat Medial Entorhinal Cortex, recorded for
-  about 20 minutes. Their spikes are binned into $T=417\,641$ bins of
-  $3\;\textrm{ms}$ width.
+  14147]{stensolaetal2012}, consists of $n=65$ neurons from rat Medial
+  Entorhinal Cortex, recorded for about 20 minutes. Their spikes are binned
+  into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
 \item The \textbf{second data set}, from \textcite[data courtesy by A.
@@
 
-The larger-population maximum-entropy distribution is calculated using five
-moments. This number seems to provide almost as much information as the
-full frequency distribution of the sample (see next section).
+The maximum-entropy distribution for the larger population is calculated
+using five moments. This number seems to provide almost as much information
+as the full frequency distribution of the sample (see next section).
 Figure~\ref{fig:plots_5mom} shows the resulting densities
@@
 (\textcolor{myred}{red circles}), $N=10\,000$ (\textcolor{myblue}{blue
-  curve}). The frequency distribution of the sample activity is also shown
+  curve}). The frequency density of the sample activity is also shown
 (\textcolor{black}{black triangles}), and in the plot it would be
-indistinguishable from the maximum-entropy distribution for $N=n$, that is,
+indistinguishable from the maximum-entropy density for $N=n$, that is,
 applied at the sample level. We discuss the case of unknown $N$ in
@@
 recovered with relative errors smaller than $10^{-12}$.
-% 1000:  3*10^-13 %
-% 5000: 7*10^-13 %, 581 s
-% 10000: 6*10^-11 %, 737 s
+% 1000:  3*10^-13 \%
+% 5000: 7*10^-13 \%, 581 s
+% 10000: 6*10^-11 \%, 737 s
 
@@
 low-activity modes. For the second data set (lower panel) the distribution
-presents a small shoulder on the right of the mode.
+presents a small shoulder on the right of its mode.
 % , suggestive of two activity regimes. We frame no hypotheses about the
@@
 maximum-entropy distribution at the sample level. The application of the
-probability calculus thus reveals interesting possible features of the larger
-population. \textcolor{white}{If you find this you can claim a postcard from
-  us.}
+probability calculus thus reveals interesting features of the larger
+population. \textcolor{white}{If you find this you can claim a postcard
+  from us.}
 
@@
 % maxent_solutions_different_moments_partial.nb
-These results suggest that this maximum-entropy approach could be very
-useful. Here is a toy example of possible uses. We could be interested in
-the hypothesis that two distinct cell types or assemblies be present in the
-region where the recording was made. Finding a larger-population distribution
-with two peaks, as in the upper panel of \fig~\ref{fig:plots_5mom}, would
-provide some evidence for this hypothesis. Let's further imagine that we
-have reasons for suspecting that a specific set of the sampled neurons are
-of one type, and the remaining of the other type. In the case of the first
+Here is a toy example of possible uses of this maximum-entropy approach,
+based on the first data set. We could be interested in the hypothesis that
+two distinct cell types or assemblies be present in the region where the
+recording was made. Finding a larger-population distribution with two
+peaks, as in the upper panel of \fig~\ref{fig:plots_5mom}, would provide
+some evidence for this hypothesis. Let's further imagine that we have
+reasons for suspecting that a specific set of the sampled neurons is of the
+first type, and the remaining of the second type. In the case of the first
 data set, 27 of the 65 sampled neurons were identified as grid cells
-belonging to 3--4 functional modules, see \cite{dunnetal2015}. Could the
+belonging to 3--4 functional modules \citep{dunnetal2015}. Could the
 two peaks in the distribution of \fig~\ref{fig:plots_5mom} reflect the
@@
 that the population of grid cells is contributing to the first peak of the
-larger population, but it is not the sole contributor. We can also assess if
-the distributions for the two sets of neurons are independent. If they were
-independent, the larger-population distribution would be given by their
-convolution:
+larger population, but it is not the sole contributor. We can also assess
+if the distributions for the two sets of neurons are independent. If they
+were independent, the larger-population distribution would be given by
+their convolution:
 \begin{equation}

### 2019-11-26 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-26T13:40:59+0100
+%% Last-Updated: 2019-11-26T14:07:21+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 population of grid cells (\textcolor{mygreen}{green triangles}) seems to
-have one, broad peak, close to the first peak of the larger-population
+have one broad peak, close to the first peak of the larger-population
 distribution. The distribution for the larger population of non-grid cells
@@
 (\textcolor{myblue}{blue curve}) but closer in height. It would thus seem
-that the population of grid cells is contributing to the first peak of the
-larger population, but it is not the sole contributor. We can also assess
+that the population of grid cells is contributing to the first mode of the
+larger population, but it is not its sole contributor. We can also assess
 if the distributions for the two sets of neurons are independent. If they
@@
 
-The toy analysis just given should be taken not literally, but just as an
+The toy analysis above should not be taken literally, but just as an
 illustration of the method's possible applications. The important point is
-that this method is computationally very cheap and yet can provide useful
-insights, even if just qualitative ones.
+that this method is computationally very cheap and yet it can provide
+useful insights, even if just qualitative ones, and even suggest new
+hypotheses.
 
@@
 \label{sec:marginalization}
-
-\begin{figure}[!p]
-\centering
-\includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_twomom_distributions10000_priorc.pdf}%
-\\[3em]%
-\includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_twomom_distributions10000_prior2.pdf}%
-\caption{***}
-\label{fig:plots_2_4mom}
-\end{figure}
-% maxent_solutions_different_moments_prior2_6moments.nb
-To illustrate how our approach can be applied to studies of
-\enquote{cooperativity} or \enquote{interaction},
-\fig~\ref{fig:plots_2_4mom} shows the larger-population distributions, for
-$N=10\,000$, obtained constraining the first two moments (in
-\textcolor{red}{red}, equivalent to constraining means and correlations)
-and the first four moments (in \textcolor{blue}{blue}). The frequency
-distribution for the sample is also shown for comparison. In both data sets
-the two-moment constraint leads to a distribution quite different from the
-that of the four-moment constraint. For the first data set in particular,
-one distribution is bimodal, the other unimodal. Interpreted in terms of
-This visual difference shows directly that the two sets of moments are not
-even approximately equivalent.
-
-It is instructive to contrast this conclusion with that we would have
-reached by applying the maximum-entropy method \emph{at the sample level},
-as traditionally done. Such a comparison involves some important
-methodological caveat which we wish to discuss first.
-
-\bigskip
 
@@
 \enquote{synchrony} \citep[\eg][-- we're only citing early papers using
-these terms]{bohteetal2000,amarietal2003} of neuronal
-activity. In this section we discuss how our proposed
-application bears on this kind of quantification. % For concreteness we use an
+these terms]{bohteetal2000,amarietal2003} of neuronal activity. Most, if
+not all, such applications apply the method \emph{at the sample level}. In
+this section we discuss how our proposed application bears on this kind of
+quantification and compare it with the traditional application at the
+sample level. But such a comparison involves some important methodological
+caveats which we wish to discuss first.% For concreteness we use an
 % example with specific statistics: first and second moments (equivalent to
@@
 % discussion holds for any sets of statistics.
+
+\bigskip
 
@@
 \bigskip
+
+\begin{figure}[!p]
+\centering
+\includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_twomom_distributions10000_priorc.pdf}%
+\\[3em]%
+\includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_twomom_distributions10000_prior2.pdf}%
+\caption{***}
+\label{fig:plots_2_4mom}
+\end{figure}
+% maxent_solutions_different_moments_prior2_6moments.nb
+Figure~\ref{fig:plots_2_4mom} shows the larger-population distributions, for
+$N=10\,000$, obtained constraining the first two moments (in
+\textcolor{red}{red}, equivalent to constraining means and correlations)
+and the first four moments (in \textcolor{blue}{blue}). The frequency
+distribution for the sample is also shown for comparison. In both data sets
+the two-moment constraint leads to a distribution quite different from the
+that of the four-moment constraint. For the first data set in particular,
+one distribution is bimodal, the other unimodal. This visual difference
+shows directly that the two sets of moments are not even approximately
+equivalent; a quantitative comparison is given below.
 

### 2019-11-26 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-26T14:07:21+0100
+%% Last-Updated: 2019-11-26T14:21:20+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \label{eq:chain_prescription_moments}
-  \mathrlap{\overbracket{\phantom{\text{measured moments} \rightarrow
+  \mathrlap{\overbracket{\phantom{\text{measured averages} \rightarrow
         \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
@@
    \sum_{\ya}
-  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured moments}}
+  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured averages}}
   \mathrel{\;=\;}
@@
 and of one assumption specific to our application: that the order of the
-time bins is irrelevant, that measured moments equal expectations (or
+time bins is irrelevant, that measured averages equal expected values (or
 equivalently, the number of time bins $T$ is infinite), and that an
@@
 it can be the case that discarding part of the data or statistics -- for
-example, the measured third- and higher-order moments -- leaves our
+example, the third- and higher-order empirical moments -- leaves our
 probabilities almost unchanged. This means that the discarded statistics
 are \emph{informationally irrelevant} or almost so. The remaining
-statistics -- for example, first and second moments -- are
-\emph{informationally sufficient}.\footnote{For more technical results
-  and connections with the notion of symmetry see \eg\
+statistics -- for example, first and second empirical moments -- are
+\emph{informationally sufficient}.\footnote{For more technical results and
+  connections with the notion of symmetry see \eg\
   \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,lauritzen1982_r1988,lauritzen2007,cifarellietal1980,cifarellietal1981,diaconisetal1981,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,ayetal2015}.}
@@
 Before applying this notion to neuronal activity, however, we must keep in
-mind that \emph{informational sufficiency is not preserved under sampling}:
-if a probability distribution has some sufficient statistics, then its
+mind that \emph{informational sufficiency is not preserved under sampling}.
+If a probability distribution has some sufficient statistics, then its
 marginals, such as the distribution for a sample, \emph{cannot} have the
@@
 mentioned above, and translates into the general impossibility of solving a
-system of $n$ independent equations in $m$ unknowns with more equations
-than unknowns, $n>m$ \citep[\sect~3.1]{portamanaetal2015}.
+system of independent equations with more equations than unknowns
+\citep[\sect~3.1]{portamanaetal2015}.
 
@@
 correlations seem informationally sufficient for a particular sample from a
-brain area, then they may well not be sufficient for the larger population of
-neurons constituting that area, and vice versa. So if what interests us is
-\enquote{cooperativity} or \enquote{interaction} of a brain area, it is
-unreliable to use a maximum-entropy distribution constructed only for the
-sample. % \citep[contrast this with some literature, \eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}
-The approach presented here avoids this problem, because the
-maximum-entropy method is applied to obtain the distribution of the larger
-population, not of the sample alone.
+brain area, then they may well not be sufficient for the larger population
+of neurons constituting that area, and vice versa. So if we are interested
+in the \enquote{cooperativity} or \enquote{interaction} of a brain area, it
+is unreliable to use a maximum-entropy distribution constructed only for a
+sample
+thereof. % \citep[contrast this with some literature, \eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}
+The approach presented here avoids this problem because the maximum-entropy
+method is applied to obtain the distribution of the larger population, not
+of the sample alone.
 
@@
 \begin{enumerate}[label=(\roman*)]
-\item from each maximum-entropy distributions $\px(\yA \| M)$ for the larger
+\item from each maximum-entropy distributions $\px(\yA \| M,N)$ for the larger
   population, built from each set of constraints $M=M',M''$, calculate the
@@
   \label{eq:sample_marginal}
-  \pxx(\ya \| M) = \sum_{\yA}\yG_{\ya \yA}\px(\yA \| M),
+  \pxx(\ya \| M, N) = \sum_{\yA}\yG_{\ya \yA}\px(\yA \| M,N),
   \qquad M=M',M'';
@@
     \label{eq:relentropy_T}
-    T\,\sh[\yf; \pxx(\ya \| M)] \defd  T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{
-      \pxx(\ya \| M)},
+    T\,\sh[\yf; \pxx(\ya \| M,N)] \defd
+    T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{\pxx(\ya \| M,N)},
     \qquad M=M',M'';    
@@
 \item take the difference: 
-\begin{multline}
+\begin{equation}
   \label{eq:measure_suffic}
-  \ydi(M'',M') \defd
-  T \,\sh[\yf; \pxx(\ya \| M')]
+  \begin{split}
+  \ydi_{N}(M'',M') &\defd
+  T \,\sh[\yf; \pxx(\ya \| M',N)]
   -
-  T \,\sh[\yf; \pxx(\ya \| M'')]
-  \equiv{}\\
+  T \,\sh[\yf; \pxx(\ya \| M'',N)]
+  \\ &\equiv
 T  \sum_{\ya} \yff_{\ya} \log\frac{
-  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'')}{
-  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M')}.
-\end{multline}
+  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'',N)}{
+  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M',N)}.
+\end{split}
+\end{equation}
 \end{enumerate}
-The measure $\ydi(M'',M')$ so defined is positive if $M''$ is \enquote{more
+The measure $\ydi_{N}(M'',M')$ so defined is positive if $M''$ is \enquote{more
   informationally sufficient} than $M'$, and negative otherwise.
 
-Why is this a natural measure? Because $\ydi(M'',M')$ is equal to the
+Why is this a natural measure? Because $\ydi_{N}(M'',M')$ is equal to the
 log-ratio of the probabilities of the data $\yf$ conditional on the

### 2019-11-26 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-26T14:21:20+0100
+%% Last-Updated: 2019-11-26T14:27:05+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \label{eq:bayes_factor}
-  \ydi(M'',M') =
-  \log\bigl[\pf(\yf \| M'')/
-  \pf(\yf \| M')\bigr].
+  \ydi_{N}(M'',M') =
+  \log\bigl[\pf(\yf \| M'',N)/
+  \pf(\yf \| M',N)\bigr].
 \end{equation}
@@
 also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
-this equality in appendix~\ref{sec:measure_suff}. So the exponential of
-$\ydi(M'',M')$ tells us how much more probable the data $\yf$ are
+this equality in appendix~\ref{sec:measure_suff}. The exponential of
+$\ydi_{N}(M'',M')$ tells us how much more probable the data $\yf$ are
 conditional on $M''$, than conditional on $M'$. We can also combine this
@@
   \begin{split}
-  \ydi(M_{4},M_{2}) &= 81\;\nat = 35\;\hart,
-\\    \ydi(M_{5},M_{4}) &= 0.037\;\nat = 0.016\;\hart,
+  \ydi_{N}(M_{4},M_{2}) &= 81\;\nat = 35\;\hart,
+\\    \ydi_{N}(M_{5},M_{4}) &= 0.037\;\nat = 0.016\;\hart,
   \end{split}
@@
 magnitude more probable under sufficiency of the first four moments than
-under sufficiency of the first two only; but they are about equally
-probable ($10^{0.016} = 1.04$) whether we consider the first five moments
-to be sufficient or just the first four.
+under sufficiency of the first two only. But they are about as probable
+($10^{0.016} = 1.04$) considering the first five moments to be sufficient
+as considering the first four moments only.
 
 Compare the results above with those obtained by applying the method at the
-sample level only:
+sample level, that is, with $N=n$:
 \begin{equation}
@@
   \begin{split}
-  \ydi_{\text{sample level}}(M_{4},M_{2}) &= 280\;\nat = 1\,220\;\hart,
-\\    \ydi_{\text{sample level}}(M_{5},M_{4}) &= 3.3\;\nat = 1.4\;\hart,
+  \ydi_{n}(M_{4},M_{2}) &= 280\;\nat = 1\,220\;\hart,
+\\    \ydi_{n}(M_{5},M_{4}) &= 3.3\;\nat = 1.4\;\hart.
 \end{split}
 \end{equation}
-that is, the data are 1\,220 orders of magnitude more probable conditional
-on four moments than conditional on two, and $10^{1.4}=25$ times more
-probable conditional on five moments than on four moments.
+The data are 1\,220 orders of magnitude more probable conditional on four
+moments than conditional on two, and $10^{1.4}=25$ times more probable
+conditional on five moments than on four moments.
+
+***
 

### 2019-11-26 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-26T14:27:05+0100
+%% Last-Updated: 2019-11-26T16:21:29+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 words, the measured frequencies of the sample activity are 35 orders of
-magnitude more probable under sufficiency of the first four moments than
-under sufficiency of the first two only. But they are about as probable
-($10^{0.016} = 1.04$) considering the first five moments to be sufficient
-as considering the first four moments only.
-
-Compare the results above with those obtained by applying the method at the
-sample level, that is, with $N=n$:
-\begin{equation}
-  \label{eq:diff_suff_n}
-  \begin{split}
-  \ydi_{n}(M_{4},M_{2}) &= 280\;\nat = 1\,220\;\hart,
-\\    \ydi_{n}(M_{5},M_{4}) &= 3.3\;\nat = 1.4\;\hart.
-\end{split}
-\end{equation}
-The data are 1\,220 orders of magnitude more probable conditional on four
-moments than conditional on two, and $10^{1.4}=25$ times more probable
-conditional on five moments than on four moments.
-
-***
-
-The application at the larger-population level therefore leads to different
-conclusions about the relative importance of the moments from the
-application at the sample level. It is also possible to use the
-measure~\eqref{eq:measure_suffic} and its probabilistic
-meaning~\eqref{eq:bayes_factor} to compare the probability of the data
-$\yf$ conditional on the hypothesis $M_{5, N}$ of 5-moment sufficiency at
-the larger-population level, $N=10\,000$, versus the hypothesis $M_{5, n}$ of
-5-moment sufficiency at the sample level, $N=n=65$. We obtain
-\begin{equation}
-  \label{eq:diff_suff_n}
-  \ydi(M_{5, N}, M_{5, n}) = 35\;\nat = 15\;\hart,
-\end{equation}
-that is, the data are 15 orders of magnitude more probable under the first
-hypothesis than under the second.
-
-\bigskip
+magnitude more probable assuming sufficiency of the first four moments than
+assuming sufficiency of the first two only. But they are about as probable
+($10^{0.016} = 1.04$) assuming sufficiency of the first five moments as of
+the first four moments only.
 
@@
 % maxent_solutions_different_moments_prior2_6moments.nb
-Figure~\ref{fig:plots_2_4mom} shows the larger-population distributions, for
-$N=10\,000$, obtained constraining the first two moments (in
-\textcolor{red}{red}, equivalent to constraining means and correlations)
-and the first four moments (in \textcolor{blue}{blue}). The frequency
-distribution for the sample is also shown for comparison. In both data sets
-the two-moment constraint leads to a distribution quite different from the
-that of the four-moment constraint. For the first data set in particular,
-one distribution is bimodal, the other unimodal. This visual difference
-shows directly that the two sets of moments are not even approximately
-equivalent; a quantitative comparison is given below.
-
-The conclusions about \enquote{cooperativity} or \enquote{interaction}
-that we reach by applying the maximum-entropy method to the larger population,
-as proposed here, are not only different from the application at the sample
-level, but also visually clearer. The plot of \fig~\ref{fig:plots_2_4mom},
-showing the two clearly different distributions constrained by two and four
-moments, should be compared with the plot for the distributions obtained at
-the sample level under the same constraints, shown in
-\fig~\ref{fig:comparisons_marginals} as dashed lines with empty red
-circles (\textcolor{myred}{\LARGE$\bm{\mathord{\circ}}$}) and empty blue
-squares (\textcolor{mypurpleblue}{$\bm{\square}$}). The measured
-frequencies are shown as filled black triangles
-(\textcolor{black}{$\bm{\triangledown}$}). It is necessary to use a
-logarithmic scale to see the differences, which appear mainly in the tail.
-\begin{figure}[!b]
+These informational differences shows clearly in the shapes of the
+distribution themselves, plotted in the upper panel of
+\fig~\ref{fig:plots_2_4mom}. The two-moment distribution
+(\textcolor{myred}{dashed red}) is unimodal, the four-moment distribution
+(\textcolor{myblue}{solid blue}) is bimodal. The five-moment distribution
+would not be distinguishable from the four-moment one. The lower panel in
+the figure shows the corresponding distributions for the second data set;
+also in this case they are visually very distinct.
+
+\medskip
+
+Compare the results above with those obtained by applying the method at the
+sample level, that is, with $N=n$:
+\begin{equation}
+  \label{eq:diff_suff_n}
+  \begin{split}
+  \ydi_{n}(M_{4},M_{2}) &= 280\;\nat = 1\,220\;\hart,
+\\    \ydi_{n}(M_{5},M_{4}) &= 3.3\;\nat = 1.4\;\hart.
+\end{split}
+\end{equation}
+The data are 1\,220 orders of magnitude more probable conditional on four
+moments than conditional on two, and $10^{1.4}=25$ times more probable
+conditional on five moments than on four moments.
+
+The conclusions about \enquote{cooperativity} or \enquote{interaction} that
+we reach by applying the maximum-entropy method to the larger population,
+as proposed here, are therefore different from those applying it at the
+sample level -- and also visually clearer. The plot in the upper panel of
+\fig~\ref{fig:plots_2_4mom}, showing the two clearly different
+distributions constrained by two and four moments, should be compared with
+the plot for the distributions obtained at the sample level under the same
+constraints, shown in \fig~\ref{fig:comparisons_marginals} as
+\textcolor{myred}{red circles} and \textcolor{myblue}{blue squares}. The
+measured frequencies (black triangles) are also shown. It is necessary to
+use a logarithmic scale to see the differences, which appear mainly in the
+tail.
+\begin{figure}[!p]
 \centering
@@
 
-The same plot also shows the marginals obtained from applying the method to
-a larger population with $N=10\,000$ and the same constraints, that is, the
-sample marginals of \fig~\ref{fig:plots_2_4mom}. The two-moment
-distribution is represented by filled red circles
-(\textcolor{myred}{\LARGE$\bm{\bullet}$}), and the four-moment one by
-filled blue squares (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The
-larger-population distribution fits the measured frequencies
-(\textcolor{black}{$\bm{\triangledown}$}) very well if we constrain only four
-moments.
+\medskip
+
+% 4 mom: 38.5632 nats = 167.478 dB, 5.59496*10^16
+% 5 mom: 35.2911 nats = 153.267 dB, 2.12185*10^15
+
+It is also possible to use the measure~\eqref{eq:measure_suffic} and its
+probabilistic meaning~\eqref{eq:bayes_factor} to compare the probability of
+the data $\yf$ conditional on the hypothesis $(M_{4},N)$ of four-moment
+sufficiency at the larger-population level, $N=10\,000$, versus the
+hypothesis $(M_{4}, n)$ of four-moment sufficiency at the sample level,
+$N=n=65$. We obtain
+\begin{equation}
+  \label{eq:diff_suff_n}
+  \ydi[(M_{4}, N), (M_{4}, n)] = 39\;\nat = 17\;\hart,
+\end{equation}
+that is, the data are 17 orders of magnitude more probable under the first
+hypothesis than under the second. This result is also shown in
+\fig~\ref{fig:comparisons_marginals}, where we see that the four-moment
+distribution constructed at the larger-population level
+(\textcolor{myblue}{blue filled squares}) fits the measured frequency
+distribution (black triangles) more closely than the corresponding
+sample-level distribution (\textcolor{myblue}{blue empty squares}).
 
@@
 % plot shows the maximum-entropy distributions for the sample, from two
-% moments (\textcolor{mygreen}{$\bm{\triangledown}$}) and from four moments
+% moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four moments
 % (\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained
@@
 % \item The application at the sample level leads to deceivingly similar
-%   distributions (\textcolor{mygreen}{$\bm{\triangledown}$} and
+%   distributions (\textcolor{mygreen}{$\bm{\blacktriangledown}$} and
 %   \textcolor{mypurpleblue}{$\bm{\square}$}), which can only be distinguished on a

### 2019-11-26 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-26T16:21:29+0100
+%% Last-Updated: 2019-11-26T22:02:42+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The calculations and conclusions presented in the preceding sections depend
-on the size of the larger population, $N$. The larger population can't be the whole
-brain, of course. How large should or can $N$ be in our formulae?
-
-The crucial point is formula~\eqref{eq:equal_m_factorialmoments}, on which
-our method is based, and which asserts the equality of the factorial
-moments of the larger population and those of the sample. This formula is only
-valid if our beliefs about the activity of the sample $\pxx(\ya)$ and that
-of the larger population $\pxx(\yA)$ are related by a hypergeometric
-distribution:
-\begin{equation}
-  \label{eq:hypergeometric_relation}
-    \pxx(\ya)=
-  \binom{n}{\ya}\;\binom{N-n}{\yA-\ya}\;\binom{N}{\yA}^{-1}
-\;\pxxx(\yA),
-\end{equation}
-characteristic of \enquote{drawing without replacement}
+on the size $N$ of the larger population. The larger population can't be
+the whole brain, of course. How large should or can $N$ be in our formulae?
+
+The crucial point is our belief distribution for the activity of the
+sample conditional on the activity of the larger population, given by
+formula~\eqref{eq:conditional_hypergeometric} in
+appendix~\ref{sec:derivation_maxent_fullnet}, reproduced here:
+\begin{equation}
+  \labelbis{eq:conditional_hypergeometric}
+  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
+  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
+\end{equation}
+which leads to the equality of factorial
+moments~\eqref{eq:equal_m_factorialmoments}, on which our approach rests.
+The conditional probability above is characteristic of \enquote{drawing
+  without replacement}
 \citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-In other words, $N$ is the size of the pool of all neurons that could just
-as well have been recorded by our measuring procedure. \enquote{Just as
-  well} indicates equal degrees of belief. Whether the equality of such
-beliefs is justified, or for which sampling procedures it could be
-justified, is the fundamental, deep question of sampling theory, which we
+In other words, consider the pool of neurons which we equally believe could
+have been recorded. $N$ is the size of that pool. Whether the equality of
+such beliefs is justified, or for which sampling procedures it could be
+justified, is a fundamental, deep question of sampling theory, which we
 cannot discuss here of course \citep[see \eg\ the discussions, reviews, and
@@
 
+\mynote{cite  \cite[p.~440]{maxwell1873c}}
+
 We have presented a procedure to construct the most plausible frequency
@@
   \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
-  \binom{n}{n\ya}\binom{N-n}{N \yA-n\ya}\binom{N}{N \yA}^{-1}.
+  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
 \end{equation}

### 2019-11-27 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-26T22:02:42+0100
+%% Last-Updated: 2019-11-27T11:44:52+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-In other words, consider the pool of neurons which we equally believe could
-have been recorded. $N$ is the size of that pool. Whether the equality of
-such beliefs is justified, or for which sampling procedures it could be
-justified, is a fundamental, deep question of sampling theory, which we
-cannot discuss here of course \citep[see \eg\ the discussions, reviews, and
-references in][]{ericson1969,smith1976}. $N$ therefore depends on factors
-such as: the shape, dimensions, and technical specifications of the
-recording probe; the inaccuracy in the insertion of the probe, leading for
-example to slightly different insertion points or angles; the density of
-neurons around the probe; and similar. $N$ can only be assessed case by
-case. The plots of \fig~\ref{fig:plots_5mom} suggest that its order of
-magnitude might be enough for qualitative inferences. As discussed in the
-next section, even if the exact number $N$ were known, the maximum-entropy
-distribution ought to be interpreted qualitatively or semi-quantitatively.
+In other words, consider the pool of all neurons which we believe -- with
+equal degree for each neuron -- could have been recorded. $N$ is the size
+of that pool. $N$ therefore depends on factors such as: the shape,
+dimensions, and technical specifications of the recording probe; the
+inaccuracy in the insertion of the probe, leading for example to slightly
+different insertion points or angles; the density of neurons around the
+probe. But it also depends on the homogeneity of the brain region where the
+recording was made: if we believe that it doesn't matter whether the probe
+had been inserted in some other point of the same brain region, then
+formula~\eqref{eq:conditional_hypergeometric} is appropriate. Thus $N$ can
+only be assessed case by case.
+
+Whether such beliefs are justified, or for which sampling procedures they
+can be justified, is a fundamental, deep question of sampling theory, which
+we cannot discuss here \citep[see \eg\ the discussions, reviews, and
+references in][]{ericson1969,smith1976,kruskaletal1979c,kruskaletal1980}.
+It is of course possible to probabilistically assess, case by case, whether
+this assumption is appropriate. It good to keep in mind that such
+assessment rests, in turn, on similar assumptions at a higher level.
+
+Luckily the plots of \fig~\ref{fig:plots_5mom} suggest that the order of
+magnitude of $N$ ought to be enough for qualitative inferences. As
+discussed in the next section, even if the exact number $N$ were known, the
+maximum-entropy distribution ought to be interpreted qualitatively or
+semi-quantitatively.
 

### 2019-11-27 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-27T11:44:52+0100
+%% Last-Updated: 2019-11-27T14:35:27+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{equation}
-which leads to the equality of factorial
+leading to the equality of factorial
 moments~\eqref{eq:equal_m_factorialmoments}, on which our approach rests.
-The conditional probability above is characteristic of \enquote{drawing
-  without replacement}
-\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-In other words, consider the pool of all neurons which we believe -- with
-equal degree for each neuron -- could have been recorded. $N$ is the size
-of that pool. $N$ therefore depends on factors such as: the shape,
-dimensions, and technical specifications of the recording probe; the
-inaccuracy in the insertion of the probe, leading for example to slightly
-different insertion points or angles; the density of neurons around the
-probe. But it also depends on the homogeneity of the brain region where the
-recording was made: if we believe that it doesn't matter whether the probe
-had been inserted in some other point of the same brain region, then
-formula~\eqref{eq:conditional_hypergeometric} is appropriate. Thus $N$ can
-only be assessed case by case.
-
-Whether such beliefs are justified, or for which sampling procedures they
-can be justified, is a fundamental, deep question of sampling theory, which
-we cannot discuss here \citep[see \eg\ the discussions, reviews, and
-references in][]{ericson1969,smith1976,kruskaletal1979c,kruskaletal1980}.
-It is of course possible to probabilistically assess, case by case, whether
-this assumption is appropriate. It good to keep in mind that such
-assessment rests, in turn, on similar assumptions at a higher level.
+This conditional probability, characteristic of \enquote{drawing without
+  replacement}
+\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
+ensues when our degree of belief that the $i$th unit (ball or neuron) will
+be drawn has the same value for all $i \in \set{1,\dotsc,N}$. Thus,
+consider the pool of all neurons which we believe -- with equal degree for
+each neuron -- could have been recorded. $N$ is the size of that pool. $N$
+therefore depends on factors such as: the shape, dimensions, and technical
+specifications of the recording probe; the inaccuracy in the insertion of
+the probe, leading for example to slightly different insertion points or
+angles; the density of neurons around the probe. But it also depends on the
+homogeneity of the brain region where the recording was made: if we believe
+that it doesn't matter whether the probe had been inserted in some other
+point of the same brain region, then
+formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
+therefore only be assessed case by case.
+
+Whether such equal beliefs are justified, or for which sampling procedures
+they can be justified, is the fundamental, deep question of sampling
+theory, which we cannot discuss here \citep[see \eg\ the discussions,
+reviews, and references
+in][]{ericson1969,smith1976,kruskaletal1979c,kruskaletal1980}. It is of
+course possible to probabilistically assess, case by case, whether this
+equality assumption is appropriate. But we must also be aware that such
+assessment in turn rests on analogous assumptions at a higher level, and so
+on. The probability calculus, just like the logical calculus, cannot yield
+conclusions without premisses \citep[p.~182]{johnson1924}.
 
@@
 
-If our uncertainty about $N$ spans several orders of magnitude, we can of
-course assign a probability to the possible values of $N$ based on our
-background information and the frequency data $\yf$. The procedure is
-similar to that of \sect~\ref{sec:marginalization} and appendix***. The
+If our uncertainty about $N$ spans several orders of magnitude we can
+assign a probability to the possible values of $N$ based on our background
+information and the frequency data $\yf$, similarly to the procedure in
+\sect~\ref{sec:marginalization} and appendix~\ref{sec:measure_suff}. The
 probability for the data $\yf$ conditional on a set of constraints $M$ and
-$N$ is the exponential of minus the relative entropy~\eqref{eq:relentropy_T}:
+size $N$ is the exponential of the negative relative
+entropy~\eqref{eq:relentropy_T}:
 \begin{equation}
@@
 
-Here is a simple illustrative example with our first data set. Suppose our
+Here is an illustrative example with our first data set. Suppose our
 uncertainty spans slightly more than an order of magnitude, from $N=1\,000$
@@
 $N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming the set $M$
-of first five moments is sufficient, formula~\eqref{eq:prob_data_from_N}
+of first five moments to be sufficient, formula~\eqref{eq:prob_data_from_N}
 gives
@@
 Since our uncertainty regards a scale factor, it can be represented by
-equal pre-data probabilities $1/5$ about these partial orders of magnitude.
-Thus from~\eqref{eq:data1_f_from_N} we find
+equal pre-data probabilities of $1/5$ about these partial orders of
+magnitude. Thus from~\eqref{eq:data1_f_from_N} we find
 \begin{equation}
@@
 typical of \enquote{drawing without replacement}
-\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
+\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}[\sects~2.1,
+3.2]{jeffreys1939_r1983}.
 In the following leave $n$, $N$ implicit in the conditional. If we are

### 2019-11-29 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-27T14:35:27+0100
+%% Last-Updated: 2019-11-29T17:03:17+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{enumerate}
-The validity of the maximum-entropy approximation decreases as the ratio
-$N/T$ increases.
+In the present case the validity of the maximum-entropy approximation
+decreases as the ratio $nN/T$ increases (see \sect~\ref{sec:assumptions}).
 
@@
 
-\section{Limitations and assumptions}
+\section{Assumptions and limitations}
 \label{sec:assumptions}
+
+The limitations of the proposed approach can be appraised by examining the
+assumptions on which it rests. And its assumptions become clear within a
+full-fledged probabilistic approach to our inference problem, at least
+qualitatively sketched.
 

### 2019-12-01 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-11-29T17:03:17+0100
+%% Last-Updated: 2019-12-01T15:24:42+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Assumptions and limitations}
+\section{Assumptions, approximations, limitations}
 \label{sec:assumptions}
@@
 The limitations of the proposed approach can be appraised by examining the
-assumptions on which it rests. And its assumptions become clear within a
-full-fledged probabilistic approach to our inference problem, at least
-qualitatively sketched.
+assumptions and approximations on which it rests. Its assumptions and
+approximations become clear within a full-fledged probabilistic approach to
+our inference problem, at least qualitatively sketched.
+
+The first and most important assumption of this kind of maximum-entropy
+application in neuroscience is that the order of the time bins has no
+relevance; or equivalently that the frequencies of the activity during a
+recording are informationally sufficient. Note that the maximum-entropy
+method in statistical physics is meant to be used not with this assumption,
+but with averages across \emph{repetitions} of the same experiment: its
+constraints represent macroscopically \emph{reproducible} quantities and
+kinematics, even in non-equilibrium
+\citep{jaynes1957,guntonetal1983,tikochinskyetal1984,grandy1988,berryetal1988,deroecketal2005_r2006,touchette2009}.
+
+An assumption of our specific application is that all subsets of $\yA$
+neurons in the larger population are equally likely to give rise to total
+activity $\yA$, at every time bin. A mathematically equivalent assumption
+is that $n$ neurons are sampled anew at each time bin.
+
+Both assumptions above make sense as hypotheses for judging the
+informational relevance of correlations of some order, as in
+\sect~\ref{sec:marginalization}, to be compared with hypotheses about more
+realistic time dependences and clustering.
+
+In an exact calculation based on the two assumptions above, the probability
+of the sequences of sample and larger-population activities,
+$(\ya_{t}, \yA_{t})$, would involve the joint frequencies of total
+activities of sample and larger population $F_{\ya,\yA}$, multiplied by the
+number of time bins $T$:
+\begin{equation}
+  \label{eq:prob_joint_frequencies}
+  \underbracket{\prod_{t} \pxx(\ya_{t}, \yA_{t})}_{\mathclap{\textit{prob. of joint sequence}}} \equiv
+  \prod_{\ya,\yA} \pxx(\ya, \yA)^{T\,F_{\ya,\yA}} =
+  \prod_{\ya,\yA}\bigl[\underbracket{\vphantom{\underline{g}}\yG_{\ya\yA}\;\pxx(\yA)}_{\mathclap{\textit{sampling relation}}}\bigr]^{T\,F_{\ya,\yA}}.
+\end{equation}
+The reasoning would be similar to that leading to
+\eqn~\eqref{eq:prob_frequencies} of appendix~\ref{sec:measure_suff}.
+Conditioning on the frequencies $\yf$ of the sample would correspond to
+constraining the marginal $\sum_{\yA}F_{\ya,\yA}=\yff_{\ya}$. The
+maximum-entropy method approximates $T$ with infinity, thus treating the
+probability above as a delta. The result is that
+\begin{enumerate*}[label=(\roman*)]
+\item the probability for the total activity $\yA$ becomes equal to its
+  observed frequency;
+\item the frequency of sample activity $\ya$ conditional on
+  larger-population activity $\yA$ is constrained to equal the
+  hypergeometric distribution $\yG_{\ya\yA}$,
+  \eqn~\eqref{eq:conditional_hypergeometric}.
+\end{enumerate*}
+Without this approximation the latter conditional frequency can differ from
+$\yG_{\ya\yA}$, and this is expected if the number of bins $T$ is not much
+larger than the number of possible joint possibilities, roughly
+$n\times N$. Discrepancies of order $nN/T$ are therefore expected.
 
@@
   \label{eq:sample_marginal}
-  \pxx(\ya_{t} \| M) = \sum_{\yA}\yG_{\ya_{t}\,\yA}\px(\yA \| M).
+  \pxx(\ya_{t} \| M) = \sum_{\yA}\yG_{\ya_{t}\,\yA}\;\px(\yA \| M).
 \end{equation}
@@
   \prod_{\ya=0}^{n} \pxx(\ya \| M)^{T\,\yff_{\ya}} \equiv
-  \prod_{\ya=0}^{n}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{T\,\yff_{\ya}}.
+  \prod_{\ya=0}^{n}\biggl[\sum_{\yA}\yG_{\ya\yA}\;\px(\yA \| M)\biggr]^{T\,\yff_{\ya}}.
 \end{equation}

### 2019-12-01 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-01T15:24:42+0100
+%% Last-Updated: 2019-12-01T17:40:51+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 of the sequences of sample and larger-population activities,
-$(\ya_{t}, \yA_{t})$, would involve the joint frequencies of total
-activities of sample and larger population $F_{\ya,\yA}$, multiplied by the
-number of time bins $T$:
+$(\ya_{t}, \yA_{t})$, involves the joint frequencies of total activities of
+sample and larger population $F_{\ya,\yA}$, multiplied by the number of
+time bins $T$:
 \begin{equation}
@@
   \prod_{\ya,\yA} \pxx(\ya, \yA)^{T\,F_{\ya,\yA}} =
-  \prod_{\ya,\yA}\bigl[\underbracket{\vphantom{\underline{g}}\yG_{\ya\yA}\;\pxx(\yA)}_{\mathclap{\textit{sampling relation}}}\bigr]^{T\,F_{\ya,\yA}}.
+  \prod_{\ya,\yA}\bigl[\underbracket{\vphantom{\underline{g}}\yG_{\ya\yA}\;\pxxx(\yA)}_{\mathclap{\textit{sampling relation}}}\bigr]^{T\,F_{\ya,\yA}}.
 \end{equation}
@@
 \begin{enumerate*}[label=(\roman*)]
-\item the probability for the total activity $\yA$ becomes equal to its
-  observed frequency;
+\item the probability and observed frequency for the total activity $\yA$
+  are constrained to be equal,
 \item the frequency of sample activity $\ya$ conditional on
@@
 larger than the number of possible joint possibilities, roughly
-$n\times N$. Discrepancies of order $nN/T$ are therefore expected.
+$n\times N$.
+
+The final assumption in the maximum-entropy method regards the form of the
+probability $\pxxx(\yA)$ in the expression above.
 

### 2019-12-01 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-01T17:40:51+0100
+%% Last-Updated: 2019-12-01T23:27:33+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 relevance; or equivalently that the frequencies of the activity during a
-recording are informationally sufficient. Note that the maximum-entropy
-method in statistical physics is meant to be used not with this assumption,
-but with averages across \emph{repetitions} of the same experiment: its
-constraints represent macroscopically \emph{reproducible} quantities and
-kinematics, even in non-equilibrium
+recording are informationally sufficient. *** markov here \citep{diaconisetal1980c,geetal2012} *** Note that the
+maximum-entropy method in statistical physics is meant to be used not with
+this assumption, but with averages across \emph{repetitions} of the same
+experiment: its constraints represent macroscopically \emph{reproducible}
+quantities and kinematics, even in non-equilibrium
 \citep{jaynes1957,guntonetal1983,tikochinskyetal1984,grandy1988,berryetal1988,deroecketal2005_r2006,touchette2009}.

### 2019-12-02 maxNt_v6.tex
update

--- 2019-12-01 maxNt_v6.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-01T17:40:51+0100
+%% Last-Updated: 2019-12-02T10:04:27+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The limitations of the proposed approach can be appraised by examining the
+The limitations of approach here proposed can be appraised by examining the
 assumptions and approximations on which it rests. Its assumptions and
-approximations become clear within a full-fledged probabilistic approach to
-our inference problem, at least qualitatively sketched.
+approximations become clear within a full-fledged probabilistic inference
+problem, which we sketch qualitatively.
 
 The first and most important assumption of this kind of maximum-entropy
-application in neuroscience is that the order of the time bins has no
-relevance; or equivalently that the frequencies of the activity during a
-recording are informationally sufficient. Note that the maximum-entropy
-method in statistical physics is meant to be used not with this assumption,
-but with averages across \emph{repetitions} of the same experiment: its
-constraints represent macroscopically \emph{reproducible} quantities and
-kinematics, even in non-equilibrium
+application in neuroscience can be expressed in three equivalent ways:
+\begin{enumerate*}[(\roman*)]
+\item the order of the time bins has no relevance; \item the frequencies of
+  the activity during a recording are informationally sufficient for
+  inferences about unrecorded times; \item our degree of belief about the
+  sequence of activities is invariant under permutations of the time bins,
+  that is, exchangeable \citep[\chap~4]{bernardoetal1994_r2000}{dawid2013}.
+\end{enumerate*}
+Equating time averages and expectations, \eqn~\eqref{eq:final_constraint},
+would not make sense without this assumption. Its mathematical consequence
+\citep{definetti1930,hewittetal1955} is that the probability of observing a
+frequency distribution $(\yFF_{\yA})$ in $T$ time bins is given by a
+mixture of terms
+\begin{equation}
+  \label{eq:mixture_from_definetti}
+  \exp[-T\;\sh(\yF;\ynu)] \quad
+  \text{with weights }q(\ynu),
+\end{equation}
+where $\ynu\defd (\nu_{\yA})$ represents the long-run frequency
+distribution for the activity. % Other maximum-entropy models are also used
+% in neuroscience, such as kinetic Ising models
+% \cite{kawasaki1972,mezardetal2011,battistinetal2015}, based on the slightly
+% less drastic assumption of invariance under permutations of sequences
+% having the same starting state and the same transition counts
+% \citep{diaconisetal1980c,geetal2012}.
+Note that the maximum-entropy method in statistical physics is meant to be
+used not with this assumption, but with averages across \emph{repetitions}
+of the same experiment: its constraints represent macroscopically
+\emph{reproducible} quantities and kinematics, even in non-equilibrium
 \citep{jaynes1957,guntonetal1983,tikochinskyetal1984,grandy1988,berryetal1988,deroecketal2005_r2006,touchette2009}.
@@
 
-Both assumptions above make sense as hypotheses for judging the
+Both assumptions above make sense as toy hypotheses for judging the
 informational relevance of correlations of some order, as in

### 2019-12-02 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-02T10:04:27+0100
+%% Last-Updated: 2019-12-02T14:16:29+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-An assumption of our specific application is that all subsets of $\yA$
+The second assumption is specific to our application: all subsets of $\yA$
 neurons in the larger population are equally likely to give rise to total
@@
 is that $n$ neurons are sampled anew at each time bin.
+\begin{equation}
+  \label{eq:assume_sampled_anew}
+  \pxx[(\ya_{1},\dotsc,\ya_{T}), (\yA_{1}, \dotsc, \yA_{T})]
+  = \prod_{t=1}^{T} \pxx(\ya_{t}, \yA_{t})
+  = \prod_{t=1}^{T} \yG_{\ya_{t}\yA_{t}}\;\nu_{\yA_{t}}
+\end{equation}
 
@@
 $(\ya_{t}, \yA_{t})$, involves the joint frequencies of total activities of
-sample and larger population $F_{\ya,\yA}$, multiplied by the number of
+sample and larger population $J_{\ya,\yA}$, multiplied by the number of
 time bins $T$:
@@
   \underbracket{\prod_{t} \pxx(\ya_{t}, \yA_{t})}_{\mathclap{\textit{prob. of joint sequence}}} \equiv
-  \prod_{\ya,\yA} \pxx(\ya, \yA)^{T\,F_{\ya,\yA}} =
-  \prod_{\ya,\yA}\bigl[\underbracket{\vphantom{\underline{g}}\yG_{\ya\yA}\;\pxxx(\yA)}_{\mathclap{\textit{sampling relation}}}\bigr]^{T\,F_{\ya,\yA}}.
+  \prod_{\ya,\yA} \pxx(\ya, \yA)^{T\,J_{\ya,\yA}} =
+  \prod_{\ya,\yA}\bigl[\underbracket{\vphantom{\underline{g}}\yG_{\ya\yA}\;\nu_{\yA}}_{\mathclap{\textit{sampling relation}}}\bigr]^{T\,J_{\ya,\yA}}.
 \end{equation}

### 2019-12-02 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-02T14:16:29+0100
+%% Last-Updated: 2019-12-02T15:04:31+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \citep{definetti1930,hewittetal1955} is that the probability of observing a
-frequency distribution $(\yFF_{\yA})$ in $T$ time bins is given by a
-mixture of terms
+sequence of activities $(\yA_{1},\dotsc,\yA_{T})$ with frequencies
+$(\yFF_{\yA})$ is a mixture of terms
 \begin{equation}
   \label{eq:mixture_from_definetti}
-  \exp[-T\;\sh(\yF;\ynu)] \quad
+%  \exp[-T\;\sh(\yF;\ynu)] \quad
+%  \pxxx[(\yA_{1},\dotsc,\yA_{T})] \propto
+  \prod_{\yA=1}^{N}{\nu_{\yA}}^{T\,\yFF_{\yA}} \quad
   \text{with weights }q(\ynu),
@@
 activity $\yA$, at every time bin. A mathematically equivalent assumption
-is that $n$ neurons are sampled anew at each time bin.
+is that $n$ neurons are sampled anew at each time bin. This assumption
+allows us to factorize the joint probability of the two sequences of
+activities and apply the sampling
+formula~\eqref{eq:conditional_hypergeometric} at each bin:
 \begin{equation}
   \label{eq:assume_sampled_anew}
-  \pxx[(\ya_{1},\dotsc,\ya_{T}), (\yA_{1}, \dotsc, \yA_{T})]
-  = \prod_{t=1}^{T} \pxx(\ya_{t}, \yA_{t})
-  = \prod_{t=1}^{T} \yG_{\ya_{t}\yA_{t}}\;\nu_{\yA_{t}}
-\end{equation}
-
-Both assumptions above make sense as toy hypotheses for judging the
+  \pxx[(\ya_{1},\dotsc,\ya_{T}) \| (\yA_{1}, \dotsc, \yA_{T}) ]
+  = \prod_{t=1}^{T} \pxx(\ya_{t} \| \yA_{t})
+  = \prod_{t=1}^{T} \yG_{\ya_{t}\yA_{t}}.
+\end{equation}
+
+Both assumptions above make sense as reference hypotheses for judging the
 informational relevance of correlations of some order, as in
@@
 
-In an exact calculation based on the two assumptions above, the probability
-of the sequences of sample and larger-population activities,
-$(\ya_{t}, \yA_{t})$, involves the joint frequencies of total activities of
-sample and larger population $J_{\ya,\yA}$, multiplied by the number of
-time bins $T$:
+From these two assumptions the probability calculus allows us to calculate
+the probability for the frequency distribution $\yF$ given the observed
+frequency distribution for the sample $\yf$. The result is a mixture of
+terms involving the joint frequencies of total activities of sample and
+larger population $J_{\ya,\yA}$, multiplied by the number of time bins $T$:
 \begin{equation}
   \label{eq:prob_joint_frequencies}
-  \underbracket{\prod_{t} \pxx(\ya_{t}, \yA_{t})}_{\mathclap{\textit{prob. of joint sequence}}} \equiv
-  \prod_{\ya,\yA} \pxx(\ya, \yA)^{T\,J_{\ya,\yA}} =
-  \prod_{\ya,\yA}\bigl[\underbracket{\vphantom{\underline{g}}\yG_{\ya\yA}\;\nu_{\yA}}_{\mathclap{\textit{sampling relation}}}\bigr]^{T\,J_{\ya,\yA}}.
-\end{equation}
-The reasoning would be similar to that leading to
-\eqn~\eqref{eq:prob_frequencies} of appendix~\ref{sec:measure_suff}.
-Conditioning on the frequencies $\yf$ of the sample would correspond to
-constraining the marginal $\sum_{\yA}F_{\ya,\yA}=\yff_{\ya}$. The
-maximum-entropy method approximates $T$ with infinity, thus treating the
-probability above as a delta. The result is that
+%\pxx[\yF \| \yf ] \propto
+\prod_{\ya,\yA}\bigl(\yG_{\ya\yA}\;\nu_{\yA}\bigr)^{T\,J_{\ya,\yA}},
+\end{equation}
+with weights $q(\ynu)$, and with the marginal of the joint frequency
+constrained Conditioning on the frequencies $\yf$ of the sample
+would correspond to constraining the marginal
+$\sum_{\yA}F_{\ya,\yA}=\yff_{\ya}$. The maximum-entropy method approximates
+$T$ with infinity, thus treating the probability above as a delta. The
+result is that
 \begin{enumerate*}[label=(\roman*)]

### 2019-12-02 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-02T15:04:31+0100
+%% Last-Updated: 2019-12-02T17:12:16+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 From these two assumptions the probability calculus allows us to calculate
-the probability for the frequency distribution $\yF$ given the observed
-frequency distribution for the sample $\yf$. The result is a mixture of
-terms involving the joint frequencies of total activities of sample and
-larger population $J_{\ya,\yA}$, multiplied by the number of time bins $T$:
+the probability $\pxx(\yF \| \yf )$ for the frequency distribution $\yF$
+given the observed frequency distribution for the sample $\yf$. The result
+is a mixture of terms involving the joint frequencies of total activities
+of sample and larger population $J_{\ya,\yA}$, multiplied by the number of
+time bins $T$:
 \begin{equation}
@@
 %\pxx[\yF \| \yf ] \propto
-\prod_{\ya,\yA}\bigl(\yG_{\ya\yA}\;\nu_{\yA}\bigr)^{T\,J_{\ya,\yA}},
-\end{equation}
-with weights $q(\ynu)$, and with the marginal of the joint frequency
-constrained Conditioning on the frequencies $\yf$ of the sample
-would correspond to constraining the marginal
-$\sum_{\yA}F_{\ya,\yA}=\yff_{\ya}$. The maximum-entropy method approximates
-$T$ with infinity, thus treating the probability above as a delta. The
-result is that
+ \sum_{\bm{J}} \prod_{\ya,\yA}\bigl(\yG_{\ya\yA}\;\nu_{\yA}\bigr)^{T\,J_{\ya,\yA}}
+  \qquad
+  \text{with}\quad \tsum_{\yA}J_{\ya,\yA}=\yff_{\ya},
+  \quad \tsum_{\ya}J_{\ya,\yA}=\yFF_{\yA},
+\end{equation}
+with weights $q(\ynu)$. The maximum-entropy method approximates $T$ with
+infinity, thus treating the probability above as a delta. The result is
+that
 \begin{enumerate*}[label=(\roman*)]
-\item the probability and observed frequency for the total activity $\yA$
-  are constrained to be equal,
-\item the frequency of sample activity $\ya$ conditional on
+\item the frequency distribution $\yF$ for the experiment and the long-run
+  one $\ynu$ are constrained to be equal,
+\item the frequency of the sample activity $\ya$ conditional on the
   larger-population activity $\yA$ is constrained to equal the
-  hypergeometric distribution $\yG_{\ya\yA}$,
-  \eqn~\eqref{eq:conditional_hypergeometric}.
+  hypergeometric distribution $\yG_{\ya\yA}$.
 \end{enumerate*}
@@
 The final assumption in the maximum-entropy method regards the form of the
-probability $\pxxx(\yA)$ in the expression above.
+probability $\pxx(\ynu)$ in the expressions above.
 

### 2019-12-02 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-02T17:12:16+0100
+%% Last-Updated: 2019-12-02T22:37:28+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The first and most important assumption of this kind of maximum-entropy
-application in neuroscience can be expressed in three equivalent ways:
+applications in neuroscience can be expressed in two equivalent ways:
 \begin{enumerate*}[(\roman*)]
-\item the order of the time bins has no relevance; \item the frequencies of
-  the activity during a recording are informationally sufficient for
-  inferences about unrecorded times; \item our degree of belief about the
-  sequence of activities is invariant under permutations of the time bins,
-  that is, exchangeable \citep[\chap~4]{bernardoetal1994_r2000}{dawid2013}.
+  % \item the order of the time bins has no relevance;
+\item the frequencies of the activity during a recording are
+  informationally sufficient for inferences about unrecorded times; \item
+  our degree of belief about the sequence of activities is exchangeable
+  \citep[\chap~4]{bernardoetal1994_r2000}{dawid2013}, that is, invariant
+  under permutations of the time bins.
 \end{enumerate*}
@@
 used not with this assumption, but with averages across \emph{repetitions}
-of the same experiment: its constraints represent macroscopically
+of the same experiment, because its constraints represent macroscopically
 \emph{reproducible} quantities and kinematics, even in non-equilibrium
@@
 activities and apply the sampling
-formula~\eqref{eq:conditional_hypergeometric} at each bin:
+relation~\eqref{eq:conditional_hypergeometric} of
+appendix~\ref{sec:derivation_maxent_fullnet} at each bin:
 \begin{equation}
@@
 
-From these two assumptions the probability calculus allows us to calculate
-the probability $\pxx(\yF \| \yf )$ for the frequency distribution $\yF$
-given the observed frequency distribution for the sample $\yf$. The result
-is a mixture of terms involving the joint frequencies of total activities
-of sample and larger population $J_{\ya,\yA}$, multiplied by the number of
-time bins $T$:
+From these two assumptions the probability calculus allows us to derive not
+only the probability for the frequency distribution $\yF$, but also that
+for the joint frequencies of total activities of sample and larger
+population, $J_{\ya,\yA}$, conditional on the observed frequency
+distribution for the sample $\yf$. The result is again a mixture of terms
 \begin{equation}
@@
 %\pxx[\yF \| \yf ] \propto
- \sum_{\bm{J}} \prod_{\ya,\yA}\bigl(\yG_{\ya\yA}\;\nu_{\yA}\bigr)^{T\,J_{\ya,\yA}}
+  \prod_{\ya,\yA}\bigl(\yG_{\ya\yA}\;\nu_{\yA}\bigr)^{T\,J_{\ya,\yA}}
   \qquad
-  \text{with}\quad \tsum_{\yA}J_{\ya,\yA}=\yff_{\ya},
-  \quad \tsum_{\ya}J_{\ya,\yA}=\yFF_{\yA},
-\end{equation}
-with weights $q(\ynu)$. The maximum-entropy method approximates $T$ with
-infinity, thus treating the probability above as a delta. The result is
-that
+  \text{with } \tsum_{\yA}J_{\ya,\yA}=\yff_{\ya}
+  %\quad \tsum_{\ya}J_{\ya,\yA}=\yFF_{\yA},
+\end{equation}
+with weights $q(\ynu)$.
+
+The maximum-entropy method approximates $T$ in the probability above with
+infinity, thus treating it as a delta. This leads to two approximate equalities:
 \begin{enumerate*}[label=(\roman*)]
-\item the frequency distribution $\yF$ for the experiment and the long-run
-  one $\ynu$ are constrained to be equal,
-\item the frequency of the sample activity $\ya$ conditional on the
-  larger-population activity $\yA$ is constrained to equal the
-  hypergeometric distribution $\yG_{\ya\yA}$.
+\item of the recorded and long-run frequency distributions $\yF$ and
+  $\ynu$, and therefore also of their probability distributions,
+\item of the hypergeometric distribution $\yG_{\ya\yA}$ and the frequency
+  of the sample activity $\ya$ conditional on the larger-population
+  activity $\yA$.
 \end{enumerate*}
 Without this approximation the latter conditional frequency can differ from
-$\yG_{\ya\yA}$, and this is expected if the number of bins $T$ is not much
+$\yG_{\ya\yA}$ -- and this is expected if the number of bins $T$ is not much
 larger than the number of possible joint possibilities, roughly
@@
 
-The final assumption in the maximum-entropy method regards the form of the
-probability $\pxx(\ynu)$ in the expressions above.
+The final assumption of the maximum-entropy method concerns the form of the
+probability density $\pxx(\ynu)$ in the expressions above. It is taken to
+be either an \enquote{entropic prior}
+\citep{skillingetal1984,rodriguez1991,neumann2007}, which partially takes
+into account the multiplicity of long-run frequencies, or a prior for a
+model with sufficient statistics, which expresses the sufficiency of the
+model under study \citep{portamana2017}. This assumption is crucial for
+obtaining the maximum-entropy method with the Shannon entropy. A different
+density would lead to alternative maximum-entropy methods, for example
+based on the Burg entropy \citep{burg1975} if a Dirichlet density is used
+\citep{jaynes1986d_r1996}.
 

### 2019-12-03 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-02T22:37:28+0100
+%% Last-Updated: 2019-12-03T09:50:35+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The limitations of approach here proposed can be appraised by examining the
-assumptions and approximations on which it rests. Its assumptions and
-approximations become clear within a full-fledged probabilistic inference
-problem, which we sketch qualitatively.
+The limitations of the approach here proposed can be appraised by examining
+the assumptions and approximations on which it rests: three assumptions and
+two approximation all in all. These become in turn clear within a
+full-fledged probabilistic approach. We summarize them here and give a more
+mathematical sketch in appendix~\ref{sec:inference_prob}.
 
@@
   informationally sufficient for inferences about unrecorded times; \item
-  our degree of belief about the sequence of activities is exchangeable
-  \citep[\chap~4]{bernardoetal1994_r2000}{dawid2013}, that is, invariant
-  under permutations of the time bins.
+  our degree of belief about the sequence of activities is invariant under
+  permutations of the time bins.
 \end{enumerate*}
@@
 would not make sense without this assumption. Its mathematical consequence
-\citep{definetti1930,hewittetal1955} is that the probability of observing a
-sequence of activities $(\yA_{1},\dotsc,\yA_{T})$ with frequencies
-$(\yFF_{\yA})$ is a mixture of terms
-\begin{equation}
-  \label{eq:mixture_from_definetti}
-%  \exp[-T\;\sh(\yF;\ynu)] \quad
-%  \pxxx[(\yA_{1},\dotsc,\yA_{T})] \propto
-  \prod_{\yA=1}^{N}{\nu_{\yA}}^{T\,\yFF_{\yA}} \quad
-  \text{with weights }q(\ynu),
-\end{equation}
-where $\ynu\defd (\nu_{\yA})$ represents the long-run frequency
-distribution for the activity. % Other maximum-entropy models are also used
-% in neuroscience, such as kinetic Ising models
-% \cite{kawasaki1972,mezardetal2011,battistinetal2015}, based on the slightly
-% less drastic assumption of invariance under permutations of sequences
-% having the same starting state and the same transition counts
-% \citep{diaconisetal1980c,geetal2012}.
-Note that the maximum-entropy method in statistical physics is meant to be
-used not with this assumption, but with averages across \emph{repetitions}
-of the same experiment, because its constraints represent macroscopically
-\emph{reproducible} quantities and kinematics, even in non-equilibrium
+is to relate the probability for the frequency distribution $\yF$ during
+the recording to that of the \emph{long-run} frequency distribution for the
+activities of the larger population. Note that the maximum-entropy method
+in statistical physics is meant to be used not with this assumption, but
+with averages across \emph{repetitions} of the same experiment, because its
+constraints represent macroscopically \emph{reproducible} quantities and
+kinematics, even in non-equilibrium
 \citep{jaynes1957,guntonetal1983,tikochinskyetal1984,grandy1988,berryetal1988,deroecketal2005_r2006,touchette2009}.
 
-The second assumption is specific to our application: all subsets of $\yA$
+The second assumption of the maximum-entropy method concerns the form of
+the probability distribution for the long-run frequencies ensuing from the
+first assumption. It is taken to be either an \enquote{entropic prior}
+\citep{skillingetal1984,rodriguez1991,neumann2007}, which takes into
+account the multiplicity of long-run frequencies through a coefficient $L$,
+or a prior for a model with sufficient statistics, which expresses the
+sufficiency of the moments under study \citep{portamana2017}. This
+assumption is crucial for obtaining the maximum-entropy method with the
+Shannon entropy. A different density leads to alternative maximum-entropy
+methods, for example based on the Burg entropy \citep{burg1975} if a
+Dirichlet density is used \citep{jaynes1986d_r1996}.
+
+The third assumption is specific to our application: all subsets of $\yA$
 neurons in the larger population are equally likely to give rise to total
@@
 relation~\eqref{eq:conditional_hypergeometric} of
-appendix~\ref{sec:derivation_maxent_fullnet} at each bin:
-\begin{equation}
-  \label{eq:assume_sampled_anew}
-  \pxx[(\ya_{1},\dotsc,\ya_{T}) \| (\yA_{1}, \dotsc, \yA_{T}) ]
-  = \prod_{t=1}^{T} \pxx(\ya_{t} \| \yA_{t})
-  = \prod_{t=1}^{T} \yG_{\ya_{t}\yA_{t}}.
-\end{equation}
-
-Both assumptions above make sense as reference hypotheses for judging the
-informational relevance of correlations of some order, as in
-\sect~\ref{sec:marginalization}, to be compared with hypotheses about more
-realistic time dependences and clustering.
-
-From these two assumptions the probability calculus allows us to derive not
-only the probability for the frequency distribution $\yF$, but also that
-for the joint frequencies of total activities of sample and larger
-population, $J_{\ya,\yA}$, conditional on the observed frequency
-distribution for the sample $\yf$. The result is again a mixture of terms
-\begin{equation}
-  \label{eq:prob_joint_frequencies}
-%\pxx[\yF \| \yf ] \propto
-  \prod_{\ya,\yA}\bigl(\yG_{\ya\yA}\;\nu_{\yA}\bigr)^{T\,J_{\ya,\yA}}
-  \qquad
-  \text{with } \tsum_{\yA}J_{\ya,\yA}=\yff_{\ya}
-  %\quad \tsum_{\ya}J_{\ya,\yA}=\yFF_{\yA},
-\end{equation}
-with weights $q(\ynu)$.
-
-The maximum-entropy method approximates $T$ in the probability above with
-infinity, thus treating it as a delta. This leads to two approximate equalities:
-\begin{enumerate*}[label=(\roman*)]
-\item of the recorded and long-run frequency distributions $\yF$ and
-  $\ynu$, and therefore also of their probability distributions,
-\item of the hypergeometric distribution $\yG_{\ya\yA}$ and the frequency
-  of the sample activity $\ya$ conditional on the larger-population
-  activity $\yA$.
-\end{enumerate*}
-Without this approximation the latter conditional frequency can differ from
-$\yG_{\ya\yA}$ -- and this is expected if the number of bins $T$ is not much
-larger than the number of possible joint possibilities, roughly
-$n\times N$.
-
-The final assumption of the maximum-entropy method concerns the form of the
-probability density $\pxx(\ynu)$ in the expressions above. It is taken to
-be either an \enquote{entropic prior}
-\citep{skillingetal1984,rodriguez1991,neumann2007}, which partially takes
-into account the multiplicity of long-run frequencies, or a prior for a
-model with sufficient statistics, which expresses the sufficiency of the
-model under study \citep{portamana2017}. This assumption is crucial for
-obtaining the maximum-entropy method with the Shannon entropy. A different
-density would lead to alternative maximum-entropy methods, for example
-based on the Burg entropy \citep{burg1975} if a Dirichlet density is used
-\citep{jaynes1986d_r1996}.
-
-In the study or use of the frequency distribution obtained with the
-procedure here presented we must take into account two important points.
-
-The first point is that there are many possible frequency distributions
-which we believe, to different degrees, could be the true one that happened
-during the recording. The one given by our procedure is simply the one with
-the largest degree of belief, the mode of the belief distribution. The
-space of possible frequency distributions has many dimensions, however --
-thousands or tens of thousands. We must remember that belief distributions
-in high dimensions have counter-intuitive properties. For example, the mode
-or mean can have \emph{atypical} features when compared with the features
-of most other points of the space. The mode and mean can also be very
-different from each other.
-
-The question, then, is \emph{which features of the maximum-entropy
-  frequency distribution are typical of the majority of plausible frequency
-  distributions?} We can only answer for sure by using the full-fledged
-probability calculus. A more complete study (in preparation) with the first
-data set reveals that most of the plausible frequency distributions have
-three important features in common with the maximum-entropy one:
+appendix~\ref{sec:derivation_maxent_fullnet} at each bin.
+
+Finally, the maximum-entropy method typically approximates the number $T$
+of time bins, the prior coefficient $L$, and the ratio $T/L$ to infinity.
+These approximations lead to the identification of the four distributions
+listed at the end of \sect~\ref{sec:model}: the recorded frequency
+distribution becomes equal to the long-run one, and also to the probability
+distribution for the activity at recorded and new time bins. In our
+application it also leads to the equality of the hypergeometric
+distribution~\eqref{eq:conditional_hypergeometric} and the conditional
+frequencies of the sample activity $\ya$ given $\yA$. The goodness of this
+approximation decreases as the ratio between the number of possible joint
+states and the number of bins -- roughly $nN/T$ -- increases, and also as
+observed quantities approach their mathematical bounds, such as some
+frequencies $\yff_{\ya}=0$. Without these approximations it is still
+possible to interpret the maximum-entropy
+distribution~\eqref{eq:app_maxent_pop_first} as the mode of the probability
+distribution for the recorded frequencies $\yF$, that is, the most probable
+recorded frequency.
+
+\medskip
+
+The first and third assumptions above are the least realistic, but they
+make sense as reference hypotheses for judging the informational relevance
+of correlations of some order, as in \sect~\ref{sec:marginalization}, to be
+compared with hypotheses about more realistic time dependences and
+clustering.
+
+The two infinity approximations are at the limit of their validity in our
+specific application, so the distributions discussed in our example are
+best interpreted as \emph{most probable frequency distributions}.
+
+The space of possible frequency distributions for the larger population has
+$N$ dimensions, however, and probability distributions in high dimensions
+have counter-intuitive properties. For example, the mode or mean can have
+\emph{atypical} features when compared with the features of most other
+points of the space. The question, then, is \emph{which features of the
+  maximum-entropy frequency distribution are typical of the majority of
+  plausible frequency distributions?} We can only answer for sure by using
+the full-fledged probability calculus. A more complete study (in
+preparation) with the first data set reveals that most of the plausible
+frequency distributions have three important features in common with the
+maximum-entropy one:
 \begin{itemize}
@@
 set $M''$ to be sufficient than the set $M'$.
+
+\section{Sketch of inference with the probability calculus}
+\label{sec:inference_prob}
+
+The first and most important assumption of this kind of maximum-entropy
+applications in neuroscience can be expressed in two equivalent ways:
+\begin{enumerate*}[(\roman*)]
+  % \item the order of the time bins has no relevance;
+\item the frequencies of the activity during a recording are
+  informationally sufficient for inferences about unrecorded times; \item
+  our degree of belief about the sequence of activities is exchangeable
+  \citep[\chap~4]{bernardoetal1994_r2000}{dawid2013}, that is, invariant
+  under permutations of the time bins.
+\end{enumerate*}
+Equating time averages and expectations, \eqn~\eqref{eq:final_constraint},
+would not make sense without this assumption. Its mathematical consequence
+\citep{definetti1930,hewittetal1955} is that the probability of observing a
+sequence of activities $(\yA_{1},\dotsc,\yA_{T})$ with frequencies
+$(\yFF_{\yA})$ is a mixture of terms
+\begin{equation}
+  \label{eq:mixture_from_definetti}
+%  \exp[-T\;\sh(\yF;\ynu)] \quad
+%  \pxxx[(\yA_{1},\dotsc,\yA_{T})] \propto
+  \prod_{\yA=1}^{N}{\nu_{\yA}}^{T\,\yFF_{\yA}} \quad
+  \text{with weights }q(\ynu),
+\end{equation}
+where $\ynu\defd (\nu_{\yA})$ represents the long-run frequency
+distribution for the activity. % Other maximum-entropy models are also used
+% in neuroscience, such as kinetic Ising models
+% \cite{kawasaki1972,mezardetal2011,battistinetal2015}, based on the slightly
+% less drastic assumption of invariance under permutations of sequences
+% having the same starting state and the same transition counts
+% \citep{diaconisetal1980c,geetal2012}.
+Note that the maximum-entropy method in statistical physics is meant to be
+used not with this assumption, but with averages across \emph{repetitions}
+of the same experiment, because its constraints represent macroscopically
+\emph{reproducible} quantities and kinematics, even in non-equilibrium
+\citep{jaynes1957,guntonetal1983,tikochinskyetal1984,grandy1988,berryetal1988,deroecketal2005_r2006,touchette2009}.
+
+The second assumption is specific to our application: all subsets of $\yA$
+neurons in the larger population are equally likely to give rise to total
+activity $\yA$, at every time bin. A mathematically equivalent assumption
+is that $n$ neurons are sampled anew at each time bin. This assumption
+allows us to factorize the joint probability of the two sequences of
+activities and apply the sampling
+relation~\eqref{eq:conditional_hypergeometric} of
+appendix~\ref{sec:derivation_maxent_fullnet} at each bin:
+\begin{equation}
+  \label{eq:assume_sampled_anew}
+  \pxx[(\ya_{1},\dotsc,\ya_{T}) \| (\yA_{1}, \dotsc, \yA_{T}) ]
+  = \prod_{t=1}^{T} \pxx(\ya_{t} \| \yA_{t})
+  = \prod_{t=1}^{T} \yG_{\ya_{t}\yA_{t}}.
+\end{equation}
+
+Both assumptions above make sense as reference hypotheses for judging the
+informational relevance of correlations of some order, as in
+\sect~\ref{sec:marginalization}, to be compared with hypotheses about more
+realistic time dependences and clustering.
+
+From these two assumptions the probability calculus allows us to derive not
+only the probability for the frequency distribution $\yF$, but also that
+for the joint frequencies of total activities of sample and larger
+population, $J_{\ya,\yA}$, conditional on the observed frequency
+distribution for the sample $\yf$. The result is again a mixture of terms
+\begin{equation}
+  \label{eq:prob_joint_frequencies}
+%\pxx[\yF \| \yf ] \propto
+  \prod_{\ya,\yA}\bigl(\yG_{\ya\yA}\;\nu_{\yA}\bigr)^{T\,J_{\ya,\yA}}
+  \qquad
+  \text{with } \tsum_{\yA}J_{\ya,\yA}=\yff_{\ya}
+  %\quad \tsum_{\ya}J_{\ya,\yA}=\yFF_{\yA},
+\end{equation}
+with weights $q(\ynu)$.
+
+The maximum-entropy method approximates $T$ in the probability above with
+infinity, thus treating it as a delta. This leads to two approximate equalities:
+\begin{enumerate*}[label=(\roman*)]
+\item of the recorded and long-run frequency distributions $\yF$ and
+  $\ynu$, and therefore also of their probability distributions,
+\item of the hypergeometric distribution $\yG_{\ya\yA}$ and the frequency
+  of the sample activity $\ya$ conditional on the larger-population
+  activity $\yA$.
+\end{enumerate*}
+Without this approximation the latter conditional frequency can differ from
+$\yG_{\ya\yA}$ -- and this is expected if the number of bins $T$ is not much
+larger than the number of possible joint possibilities, roughly
+$n\times N$.
 

### 2019-12-03 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-03T09:50:35+0100
+%% Last-Updated: 2019-12-03T09:54:02+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\medskip
-
-The second point is that our degrees of belief about the frequency
-distribution for the larger population depend not only on the measured data in
-the sample, but also on our pre-data beliefs $\yI$ about the distribution.
-Which assumptions lead to the maximum-entropy result? This distribution
-appears when our initial belief about the possible frequency distributions
-$\yF$ is quantified by an entropic prior
-\citep{neumann2007,rodriguez1991,skilling1998,catichaetal2004,portamana2017}:
-\begin{equation}
-  \label{eq:entropic_prior}
-  \pf(\yF \| \yI) \propto  \exp[-\yL\;\re(\yF;\yR)]
-  \approx \binom{L}{L\yFF_{0}, \dotsc, L\yFF_{N}}\;
-  \prod_{\yA}{\yRR_{\yA}}^{L\yFF_{\yA}}
-\end{equation}
-where $\re(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\log(\yFF_{\yA}/\yRR_{\yA})$ is
-the relative entropy or discrimination information
-\citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is the
-reference distribution, and $\yL$ a positive parameter. The approximate
-equality (obtained through Stirling's approximation), where the large
-parentheses denote a multinomial coefficient, shows that this prior belief
-is proportional to the number of ways in which the distribution $\yF$ can
-be realized in $L$ time bins. The parameter $L$ roughly quantifies how many
-time bins our data set must have to affect our initial belief. The
-maximum-entropy approximation is valid when $L$ is large, but small
-compared to the sharpness of the constraints on $\yF$; in our case this
-means $\yL \approx 10$, give or take an order of magnitude.
-
-We could obviously consider pre-data beliefs different
+% \medskip
+
+% The second point is that our degrees of belief about the frequency
+% distribution for the larger population depend not only on the measured data in
+% the sample, but also on our pre-data beliefs $\yI$ about the distribution.
+% Which assumptions lead to the maximum-entropy result? This distribution
+% appears when our initial belief about the possible frequency distributions
+% $\yF$ is quantified by an entropic prior
+% \citep{neumann2007,rodriguez1991,skilling1998,catichaetal2004,portamana2017}:
+% \begin{equation}
+%   \label{eq:entropic_prior}
+%   \pf(\yF \| \yI) \propto  \exp[-\yL\;\re(\yF;\yR)]
+%   \approx \binom{L}{L\yFF_{0}, \dotsc, L\yFF_{N}}\;
+%   \prod_{\yA}{\yRR_{\yA}}^{L\yFF_{\yA}}
+% \end{equation}
+% where $\re(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\log(\yFF_{\yA}/\yRR_{\yA})$ is
+% the relative entropy or discrimination information
+% \citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is the
+% reference distribution, and $\yL$ a positive parameter. The approximate
+% equality (obtained through Stirling's approximation), where the large
+% parentheses denote a multinomial coefficient, shows that this prior belief
+% is proportional to the number of ways in which the distribution $\yF$ can
+% be realized in $L$ time bins. The parameter $L$ roughly quantifies how many
+% time bins our data set must have to affect our initial belief. The
+% maximum-entropy approximation is valid when $L$ is large, but small
+% compared to the sharpness of the constraints on $\yF$; in our case this
+% means $\yL \approx 10$, give or take an order of magnitude.
+
+\mynote{add this above}We could obviously consider pre-data beliefs different
 from~\eqref{eq:entropic_prior}, for example one quantified by a Dirichlet

### 2019-12-03 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-03T09:54:02+0100
+%% Last-Updated: 2019-12-03T19:25:44+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
         \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
-  \text{measured moments} \rightarrow
+  \text{measured averages} \rightarrow
         \underbracket{\text{sample moments}
@@
 compared with hypotheses about more realistic time dependences and
-clustering.
+clustering. They may turn out to approximate well more realistic assumptions.
 
@@
 have counter-intuitive properties. For example, the mode or mean can have
-\emph{atypical} features when compared with the features of most other
-points of the space. The question, then, is \emph{which features of the
-  maximum-entropy frequency distribution are typical of the majority of
-  plausible frequency distributions?} We can only answer for sure by using
-the full-fledged probability calculus. A more complete study (in
-preparation) with the first data set reveals that most of the plausible
-frequency distributions have three important features in common with the
-maximum-entropy one:
-\begin{itemize}
-\item all activities $\yA/N \gtrsim 5\%$ have practically zero frequencies;
-\item there are two regions of activity levels with high frequencies,
-  roughly separated by a trough of lower frequencies.
-\item The frequencies of the region on the left ($\yA/N \lesssim 1.8\%$)
-  are higher than those of the region on the right ($\yA/N \gtrsim 1.8\%$).
-\end{itemize}
-But there are also differences. For example, many plausible frequency
-distributions have three or four modes instead of just two; these modes are
-higher than those of the maximum-entropy distribution; and the bump of high
-frequencies on the right is slightly shifted towards lower activities than
-the corresponding maximum in the maximum-entropy distribution.
+\emph{atypical} features when compared with the majority of other probable
+points of the probability space. The question then is \emph{which features
+  of the maximum-entropy frequency distribution are typical of the majority
+  of plausible frequency distributions?} Preliminary studies using a full
+probability calculation (to be published separately) show that several
+semi-quantitative features of the maximum-entropy distribution are indeed
+typical. For the first data set, for example, most frequency distributions
+have two high-frequency regimes of activity, as in the upper panel of
+\fig~\ref{fig:plots_5mom}, with roughly the same heights and almost the
+same locations, within a few percentages of normalized total activity. Each
+of high-frequency regimes, though, may consist of two or three very close
+modes rather than one.
+
+This means that the maximum-entropy approximation presented here can be
+favourably used to get an idea of what a fully probabilistic calculation
+could yield. -- With a far lower computational cost: for example, for
+$N=1\,000$ the maximum-entropy distribution can be calculated in ten
+minutes on a laptop, whereas the full probability calculation takes several
+days on a high-performance computing cluster.
 

### 2019-12-04 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-03T19:25:44+0100
+%% Last-Updated: 2019-12-04T17:01:22+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
-  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
+    \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
+\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
 \end{equation}
@@
 
-\section{Assumptions, approximations, limitations}
+\section{Assumptions and approximations}
 \label{sec:assumptions}
 
-The limitations of the approach here proposed can be appraised by examining
-the assumptions and approximations on which it rests: three assumptions and
-two approximation all in all. These become in turn clear within a
-full-fledged probabilistic approach. We summarize them here and give a more
-mathematical sketch in appendix~\ref{sec:inference_prob}.
-
-The first and most important assumption of this kind of maximum-entropy
-applications in neuroscience can be expressed in two equivalent ways:
+\mynote{Rewrite!}
+
+The possible uses of the method here presented depend on the assumptions
+and approximations on which it rests: three assumptions and two
+approximation all in all. These become clear within a full-fledged
+probabilistic approach. We summarize them here and give a more mathematical
+sketch in appendix~\ref{sec:inference_prob}.
+
+The first and most important assumption, typical of this kind of
+maximum-entropy application in neuroscience, can be expressed in two
+equivalent ways:
 \begin{enumerate*}[(\roman*)]
@@
 
-The second assumption of the maximum-entropy method concerns the form of
-the probability distribution for the long-run frequencies ensuing from the
-first assumption. It is taken to be either an \enquote{entropic prior}
-\citep{skillingetal1984,rodriguez1991,neumann2007}, which takes into
-account the multiplicity of long-run frequencies through a coefficient $L$,
-or a prior for a model with sufficient statistics, which expresses the
-sufficiency of the moments under study \citep{portamana2017}. This
-assumption is crucial for obtaining the maximum-entropy method with the
-Shannon entropy. A different density leads to alternative maximum-entropy
-methods, for example based on the Burg entropy \citep{burg1975} if a
-Dirichlet density is used \citep{jaynes1986d_r1996}.
+The second assumption, typical of maximum-entropy in general, concerns the
+form of the probability distribution for the long-run frequencies ensuing
+from the first assumption. It is taken to be either an \enquote{entropic
+  prior} \citep{skillingetal1984,rodriguez1991,neumann2007}, which takes
+into account the multiplicity of long-run frequencies through a coefficient
+$L$, and a reference distribution, or a prior for a model with sufficient
+statistics, which expresses the sufficiency of the moments under study
+\citep{portamana2017}. This assumption is crucial for obtaining the
+maximum-entropy method with the Shannon entropy. A different density leads
+to alternative maximum-entropy methods, for example based on the Burg
+entropy \citep{burg1975} if a Dirichlet density is used
+\citep{jaynes1986d_r1996}.
 
@@
 compared with hypotheses about more realistic time dependences and
-clustering. They may turn out to approximate well more realistic assumptions.
+clustering. %They may turn out to approximate well more realistic assumptions.
 
@@
   \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
-  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
+  \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
+\equiv  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
 \end{equation}

### 2019-12-04 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-04T17:01:22+0100
+%% Last-Updated: 2019-12-04T20:38:52+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\mynote{Rewrite!}
-
 The possible uses of the method here presented depend on the assumptions
@@
   prior} \citep{skillingetal1984,rodriguez1991,neumann2007}, which takes
-into account the multiplicity of long-run frequencies through a coefficient
-$L$, and a reference distribution, or a prior for a model with sufficient
-statistics, which expresses the sufficiency of the moments under study
-\citep{portamana2017}. This assumption is crucial for obtaining the
-maximum-entropy method with the Shannon entropy. A different density leads
-to alternative maximum-entropy methods, for example based on the Burg
-entropy \citep{burg1975} if a Dirichlet density is used
-\citep{jaynes1986d_r1996}.
+into account the multiplicity of long-run frequencies, or a prior for a
+model with sufficient statistics, which expresses the sufficiency of the
+moments under study \citep{portamana2017}. This assumption is crucial for
+obtaining the maximum-entropy method with the Shannon entropy. A different
+density leads to alternative maximum-entropy methods, for example based on
+the Burg entropy \citep{burg1975} if a Dirichlet density is used
+\citep{jaynes1986d_r1996}. The entropic prior depends on a reference
+distribution and on a coefficient $L$ which determines the sharpness of the
+prior's peak. The computations of the preceding sections were insensitive
+to the reference distribution in their significant digits. We used three
+biologically reasonable alternatives: a flat distribution, one linearly
+decreasing with $\yA$, and a normal one with a broad peak at low
+activities.
 
@@
 of high-frequency regimes, though, may consist of two or three very close
-modes rather than one.
+modes rather than one. The same preliminary studies also indicate that
+these typical features remain with prior choices in the second assumption
+above, for example using a Dirichlet prior instead of an entropic one.
 
@@
 
-\mynote{add this above}We could obviously consider pre-data beliefs different
-from~\eqref{eq:entropic_prior}, for example one quantified by a Dirichlet
-distribution (which is equivalent to the above but with $\yF$ and $\yR$
-switched), or a uniform distribution in $\yF$-space. Would these lead to
-markedly different post-data beliefs? A full-fledged probabilistic analysis
-shows that the three typical features listed above still appear with these
-different initial beliefs. They are therefore robust.
-
 \section{Summary and discussion}
@@
 
-\mynote{cite  \cite[p.~440]{maxwell1873c}}
-
-We have presented a procedure to construct the most plausible frequency
-distribution of population-averaged activities of a population of neurons,
-given the recording about a small sample thereof. This procedure combines
-the maximum-entropy method and basic identities from sampling theory. From
-the application to two real data sets we saw that the frequency
-distributions obtained with our procedure can have features very different
-from the one measured in the sample, such as multiple modes. This procedure
-can also be used with moment constraints of different order -- means,
-population-averaged pairwise correlations, or higher-order correlations --
-thus giving an approximate assessment of the informational sufficiency of
-specific subsets of moments. In fact, we saw that the application of
-maximum-entropy only at the sample level leads to misleading results about
-this kind of sufficiency questions.
+The reason why we record neurons from a brain region is to have an idea of
+the activity of the other neurons in that region, similarly to survey
+sampling. If we thought that \emph{precisely} the recorded ones were
+exceptions or completely unrelated to the others surrounding the recording
+probe, our recording would serve little purpose. Probabilistic methods that
+allow us to make direct inferences about the larger population are
+therefore useful. But they become complex and sensitive if we ask for very
+detailed inferences.
+
+We have presented a method that allows us to make direct inferences about a
+larger population from very small samples. Its inferences regard only the
+frequencies of the total activity of the larger population, but such
+limited scope makes the method computationally very cheap.
+
+Yet, and most important, the method leads to inferences that are not at all
+trivial and not easily discernible by looking at the sample. We showed this
+by applying the method to two real data sets.
+
+By \enquote{inference} we mean the quantification of a degree of belief
+about possible observations, based on specific assumptions; not the
+response of an oracle. The method gives us a \emph{forecast} based on those
+assumptions, and therefore has two main uses, just like any other
+probabilistic method. If our assumptions are only hypotheses under testing,
+then the subsequent confirmation or confutation of the forecast will
+increase or decrease our confidence in those hypotheses. If we strongly
+trust our assumptions, on the other hand, then we rely on the forecast and
+even use it in place of actual observations, especially if the latter are
+difficult to make. (In fact, we typically shift from the first to the
+second use.) In the preceding sections we have given examples of hypotheses
+that could be explored with the method. The method itself rests on
+assumptions, discussed in \sect~\ref{sec:assumptions}, that can either be
+under test or relied upon, depending on the case.
+
+Some approximations implicit in the method make its numerical results
+semi-quantitative with respect to an analysis based on the full probability
+calculus. The results should not be off by more than a few percentages for
+the normalized total activity and for the frequency density. This
+imprecision, however, is compensated by the extreme computational cheapness
+of the method, far less costly than a full probabilistic analysis. The
+method is therefore very convenient at least for essaying hypotheses. We
+hope to present an exact quantitative comparison with the full
+probabilistic approach in a future publication.
+
+% \mynote{cite  \cite[p.~440]{maxwell1873c}}
+
+% We have presented a procedure to construct the most plausible frequency
+% distribution of population-averaged activities of a population of neurons,
+% given the recording about a small sample thereof. This procedure combines
+% the maximum-entropy method and basic identities from sampling theory. From
+% the application to two real data sets we saw that the frequency
+% distributions obtained with our procedure can have features very different
+% from the one measured in the sample, such as multiple modes. This procedure
+% can also be used with moment constraints of different order -- means,
+% population-averaged pairwise correlations, or higher-order correlations --
+% thus giving an approximate assessment of the informational sufficiency of
+% specific subsets of moments. In fact, we saw that the application of
+% maximum-entropy only at the sample level leads to misleading results about
+% this kind of sufficiency questions.
 

### 2019-12-04 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-04T20:38:52+0100
+%% Last-Updated: 2019-12-04T20:53:39+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\mynote{comment about the possibility of drawing conclusions about a brain
-  area using different sets of neurons (eg because of recording across many
-  sessions)}
+% \mynote{comment about the possibility of drawing conclusions about a brain
+%   area using different sets of neurons (eg because of recording across many
+%   sessions)}
 
@@
 
-\mynote{review this}We obviously don't know whether the activity levels of
-the larger population really had the frequencies given by the model, during
-the recording. The model gives our most plausible guess. In high
-dimensions, however, the features of the most plausible distribution may
-\emph{not} be typical of the majority of most plausible distributions; and
-the set of all possible frequency distributions, if the full population for
-example comprises 1\,000 neurons, is a 1\,000-dimensional space. In
-\sect~\ref{sec:assumptions} we therefore try to assess which features of
-the frequency distribution delivered by our method may be typical and
-therefore expected of the actual one. We find that general features such as
-the bimodality of the first data set are indeed typical. Maximum-entropy
-models can be considered as approximations of Bayesian models based on
-various assumptions of inferential sufficiency
-\citep{jaynes1986d_r1996,portamana2017}. How do our guesses change if we
-modify our pre-data assumptions? We show, in the same section, that the
-typical features indicated by our method are robust against such changes.
-
-A summary of all points above is given in the final
-\sect~\ref{sec:summary_discussion}.
+In \sect~\ref{sec:assumptions} we discuss in detail the assumptions and
+approximations on which the method is based, from the point of view of the
+probability calculus.
+
+% \mynote{review this}We obviously don't know whether the activity levels of
+% the larger population really had the frequencies given by the model, during
+% the recording. The model gives our most plausible guess. In high
+% dimensions, however, the features of the most plausible distribution may
+% \emph{not} be typical of the majority of most plausible distributions; and
+% the set of all possible frequency distributions, if the full population for
+% example comprises 1\,000 neurons, is a 1\,000-dimensional space. In
+% \sect~\ref{sec:assumptions} we therefore try to assess which features of
+% the frequency distribution delivered by our method may be typical and
+% therefore expected of the actual one. We find that general features such as
+% the bimodality of the first data set are indeed typical. Maximum-entropy
+% models can be considered as approximations of Bayesian models based on
+% various assumptions of inferential sufficiency
+% \citep{jaynes1986d_r1996,portamana2017}. How do our guesses change if we
+% modify our pre-data assumptions? We show, in the same section, that the
+% typical features indicated by our method are robust against such changes.
+
+A summary of all points above and a discussion of the usefulness of the
+method is given in the final \sect~\ref{sec:summary_discussion}.
 

### 2019-12-05 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-04T20:53:39+0100
+%% Last-Updated: 2019-12-05T10:25:02+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\newcommand*{\yRR}{R}
-\newcommand*{\yR}{\bm{\yRR}}
+\newcommand*{\yr}{\bm{r}}
 
@@
 \newcommand*{\yL}{L}
+
+\newcommand*{\yla}{\bm{\lambda}}
+\newcommand*{\yJ}{\bm{J}}
 %\newcommand*{\yd}{\bm{\varDelta}}
@@
 
-\section{Sketch of inference with the probability calculus}
+\section{Assumptions and approximations: mathematical sketch}
 \label{sec:inference_prob}
 
-The first and most important assumption of this kind of maximum-entropy
-applications in neuroscience can be expressed in two equivalent ways:
-\begin{enumerate*}[(\roman*)]
-  % \item the order of the time bins has no relevance;
-\item the frequencies of the activity during a recording are
-  informationally sufficient for inferences about unrecorded times; \item
-  our degree of belief about the sequence of activities is exchangeable
-  \citep[\chap~4]{bernardoetal1994_r2000}{dawid2013}, that is, invariant
-  under permutations of the time bins.
-\end{enumerate*}
-Equating time averages and expectations, \eqn~\eqref{eq:final_constraint},
-would not make sense without this assumption. Its mathematical consequence
-\citep{definetti1930,hewittetal1955} is that the probability of observing a
-sequence of activities $(\yA_{1},\dotsc,\yA_{T})$ with frequencies
-$(\yFF_{\yA})$ is a mixture of terms
+We give here a sketch of the mathematical formulae behind the assumptions
+and approximations discussed in \sect~\ref{sec:assumptions}. We indicate
+them all collectively with $\yI$.
+
+The first assumption, of infinite exchangeability
+\citep[\chap~4]{bernardoetal1994_r2000}{dawid2013} of the probability
+distribution $\pf[(\yA_{1},\dotsc, \yA_{T}) \| \yI]$ for the sequence of
+activities, implies the following mixture form by de~Finetti's theorem
+\citep{definetti1930,hewittetal1955}:
 \begin{equation}
@@
 %  \pxxx[(\yA_{1},\dotsc,\yA_{T})] \propto
-  \prod_{\yA=1}^{N}{\nu_{\yA}}^{T\,\yFF_{\yA}} \quad
-  \text{with weights }q(\ynu),
-\end{equation}
-where $\ynu\defd (\nu_{\yA})$ represents the long-run frequency
-distribution for the activity. % Other maximum-entropy models are also used
-% in neuroscience, such as kinetic Ising models
-% \cite{kawasaki1972,mezardetal2011,battistinetal2015}, based on the slightly
-% less drastic assumption of invariance under permutations of sequences
-% having the same starting state and the same transition counts
-% \citep{diaconisetal1980c,geetal2012}.
-Note that the maximum-entropy method in statistical physics is meant to be
-used not with this assumption, but with averages across \emph{repetitions}
-of the same experiment, because its constraints represent macroscopically
-\emph{reproducible} quantities and kinematics, even in non-equilibrium
-\citep{jaynes1957,guntonetal1983,tikochinskyetal1984,grandy1988,berryetal1988,deroecketal2005_r2006,touchette2009}.
-
-The second assumption is specific to our application: all subsets of $\yA$
-neurons in the larger population are equally likely to give rise to total
-activity $\yA$, at every time bin. A mathematically equivalent assumption
-is that $n$ neurons are sampled anew at each time bin. This assumption
-allows us to factorize the joint probability of the two sequences of
-activities and apply the sampling
-relation~\eqref{eq:conditional_hypergeometric} of
-appendix~\ref{sec:derivation_maxent_fullnet} at each bin:
-\begin{equation}
+  \pf[(\yA_{1},\dotsc, \yA_{T}) \| \yI]
+  = \int\!\!\!\di\ynu\;\pf(\ynu \|\yI)\,\prod_{\yA=1}^{N}{\nu_{\yA}}^{T\,\yFF_{\yA}},
+\end{equation}
+where $\ynu\defd(\nu_{\yA})$ is the long-run frequency distribution of
+activities in an imaginary continuation or repetition of the recording, and
+$\pf(\ynu \|\yI)\;\di\ynu$ our belief distribution about it.
+
+The second assumption concerns the latter degree distribution. It has one
+of these expressions:
+\begin{subequations}\label{eq:maxent_prior}
+  \begin{equation}
+    \label{eq:entropic_prior}
+    \pf(\ynu \| \yI) \propto  \exp[-\yL\;\sh(\yF;\yr)]
+    \approx \binom{L}{L\nu_{0}, \dotsc, L\nu_{N}}\;
+    \prod_{\yA}{r_{\yA}}^{L\nu_{\yA}}  
+  \end{equation}
+  where $\yr\defd(r_{\yA})$ is a reference distribution and the expression
+  in larger parentheses is a multinomial coefficient, or
+  \begin{equation}
+    \label{eq:suff_prior}
+    \pf(\ynu \| \yI) \propto
+    \int\!\!\!\di\yla\;\pf(\yla \| \yI)\;
+    \prod_{\yA}\delt\Biggl\{\ynu_{\yA}-
+    \frac{r_{\yA}}{Z(\yla)}\,\exp\Bigl[
+    \tsum_{m}\lambda_{m}\tbinom{\yA}{m}/\tbinom{N}{m}
+    \Bigr]\Biggr\},
+  \end{equation}
+\end{subequations}
+where $\yla\defd(\lambda_{m})$ are Lagrange multipliers and $Z(\yla)$ a
+normalization factor for the exponential distribution within the delta.
+This is a formal expression saying that our belief about $\ynu$ is an
+exponential parametric model for which the normalized factorial moments are
+informationally sufficient.
+
+The third assumption corresponds to the formula
+\begin{multline}
   \label{eq:assume_sampled_anew}
-  \pxx[(\ya_{1},\dotsc,\ya_{T}) \| (\yA_{1}, \dotsc, \yA_{T}) ]
-  = \prod_{t=1}^{T} \pxx(\ya_{t} \| \yA_{t})
-  = \prod_{t=1}^{T} \yG_{\ya_{t}\yA_{t}}.
-\end{equation}
-
-Both assumptions above make sense as reference hypotheses for judging the
-informational relevance of correlations of some order, as in
-\sect~\ref{sec:marginalization}, to be compared with hypotheses about more
-realistic time dependences and clustering.
-
-From these two assumptions the probability calculus allows us to derive not
-only the probability for the frequency distribution $\yF$, but also that
-for the joint frequencies of total activities of sample and larger
-population, $J_{\ya,\yA}$, conditional on the observed frequency
-distribution for the sample $\yf$. The result is again a mixture of terms
-\begin{equation}
+  \pf[(\ya_{1},\dotsc,\ya_{T}) \| (\yA_{1}, \dotsc, \yA_{T}), \yI ]
+  ={}\\ \prod_{t=1}^{T} \pf(\ya_{t} \| \yA_{t}, \yI)
+  = \prod_{t=1}^{T} \yG_{\ya_{t}\yA_{t}}
+  \equiv \prod_{\ya,\yA}{\yG_{\ya_{t}\yA_{t}}}^{T\,J_{\ya\yA}},
+\end{multline}
+where $J_{\ya\yA}$ is the joint frequency distribution of the activity
+pairs $(\ya,\yA)$, so that its marginals are the frequency distributions
+$\yf$ and $\yF$. As discussed in \sect~\ref{sec:assumptions} this formula
+follows from either of two beliefs: that all subsets of $\yA$ neurons in
+the larger population are always equally likely to give rise to total
+activity $\yA$, or that a new sampling is done at every time bin.
+
+It can be proved that the formulae above lead to the following belief
+distribution for the joint frequency distribution $\yJ \defd (J_{\ya\yA})$
+conditional on $\yf$:
+\begin{multline}
   \label{eq:prob_joint_frequencies}
-%\pxx[\yF \| \yf ] \propto
-  \prod_{\ya,\yA}\bigl(\yG_{\ya\yA}\;\nu_{\yA}\bigr)^{T\,J_{\ya,\yA}}
-  \qquad
-  \text{with } \tsum_{\yA}J_{\ya,\yA}=\yff_{\ya}
-  %\quad \tsum_{\ya}J_{\ya,\yA}=\yFF_{\yA},
-\end{equation}
-with weights $q(\ynu)$.
-
-The maximum-entropy method approximates $T$ in the probability above with
-infinity, thus treating it as a delta. This leads to two approximate equalities:
-\begin{enumerate*}[label=(\roman*)]
-\item of the recorded and long-run frequency distributions $\yF$ and
-  $\ynu$, and therefore also of their probability distributions,
-\item of the hypergeometric distribution $\yG_{\ya\yA}$ and the frequency
-  of the sample activity $\ya$ conditional on the larger-population
-  activity $\yA$.
-\end{enumerate*}
-Without this approximation the latter conditional frequency can differ from
-$\yG_{\ya\yA}$ -- and this is expected if the number of bins $T$ is not much
-larger than the number of possible joint possibilities, roughly
-$n\times N$.
+  \pf(\yJ \| \yf, \yI) \propto
+                         \int\!\!\!\di\ynu\;\pf(\ynu\|\yI)\;
+                         \binom{T}{TJ_{0 0},\dotsc,TJ_{n N}}\,
+                         \prod_{\ya,\yA}\bigl(\yG_{\ya\yA}\;\nu_{\yA}\bigr)^{T\,J_{\ya\yA}}
+  \approx{}  \\
+\shoveright{\int\!\!\!\di\ynu\;\pf(\ynu\|\yI)\;
+                         \exp\biggl(
+-T \tsum_{\ya,\yA}J_{\ya\yA}\ln\frac{J_{\ya\yA}}{\yG_{\ya\yA}\;\nu_{\yA}}
+                         \biggr)}
+  \\
+  \text{with}\quad \tsum_{\yA}J_{\ya\yA}=\yff_{\ya},
+\end{multline}
+where $\pf(\ynu\|\yI)$ is either of the
+distributions~\eqref{eq:maxent_prior}. From this belief distribution we can
+find the one for $\yFF_{\yA}\equiv \sum_{\ya}J_{\ya\yA}$ by
+marginalization. Within the exponential we can recognize the relative
+entropy of $\yJ$ with respect to $(\yG_{\ya\yA}\;\nu_{\yA})$, which is the
+reason why these two joint distributions must be equal if $T$ tends to
+infinity.
 

### 2019-12-05 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-05T10:25:02+0100
+%% Last-Updated: 2019-12-05T10:55:47+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\hart}{\textrm{Hart}}
+
+\newsubfloat{figure}
+\subcaptionsize{\normalfont}
+\subcaptionlabelfont{\normalfont}
+\subcaptionfont{\normalfont}
 %%% Custom macros end @@@
@@
   \centering
+\subcaption[]{\label{fig:plots_5mom_a}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper_stensola_5mom_distributionsN_priorc.pdf}%
 \\[3em]%
+\subcaption[]{\label{fig:plots_5mom_b}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper_riehle_5mom_distributionsN_priorc.pdf}%
@@
 peaked than the measured frequency distribution for the sample; their
-difference increases with $N$. Most remarkably, for the first data set
-(upper panel) the distribution for the larger population has two distinct
-low-activity modes. For the second data set (lower panel) the distribution
-presents a small shoulder on the right of its mode.
+difference increases with $N$. Most remarkably, for the first data set,
+\fig~\ref{fig:plots_5mom_a}, the distribution for the larger population has
+two distinct low-activity modes. For the second data set,
+\fig~\ref{fig:plots_5mom_b}, the distribution presents a small shoulder on
+the right of its mode.
 % , suggestive of two activity regimes. We frame no hypotheses about the
@@
 \centering
+\subcaption[]{\label{fig:plots_cellsubsets_a}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper3_gridvsnogrid.pdf}%
 \\[3em]%
+\subcaption[]{\label{fig:plots_cellsubsets_b}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper3_gridconv.pdf}%
@@
 recording was made. Finding a larger-population distribution with two
-peaks, as in the upper panel of \fig~\ref{fig:plots_5mom}, would provide
+peaks, as in \fig~\ref{fig:plots_5mom_a}, would provide
 some evidence for this hypothesis. Let's further imagine that we have
@@
 $\yNng=5\,850$ for the non-grid set, to reflect their proportions (27/65
-and 38/65) in the recorded sample. The results are shown in the upper panel
-of \fig~\ref{fig:plots_cellsubsets}. The distribution for the larger
+and 38/65) in the recorded sample. The results are shown in 
+ \fig~\ref{fig:plots_cellsubsets_a}. The distribution for the larger
 population of grid cells (\textcolor{mygreen}{green triangles}) seems to
@@
 where the index $\yA'$ runs from $\max(0,\yA-\yNng)$ to $\min(\yA,\yNg)$.
-But this is not the case: the lower panel of
-figure~\ref{fig:plots_cellsubsets} shows that such convolution
+But this is not the case: 
+figure~\ref{fig:plots_cellsubsets_b} shows that such convolution
 (\textcolor{black}{black diamonds}) is quite different from the
@@
 \centering
+\subcaption[]{\label{fig:plots_2_4mom_a}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_twomom_distributions10000_priorc.pdf}%
 \\[3em]%
+\subcaption[]{\label{fig:plots_2_4mom_b}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_twomom_distributions10000_prior2.pdf}%
@@
 These informational differences shows clearly in the shapes of the
-distribution themselves, plotted in the upper panel of
-\fig~\ref{fig:plots_2_4mom}. The two-moment distribution
+distribution themselves, plotted in 
+\fig~\ref{fig:plots_2_4mom_a}. The two-moment distribution
 (\textcolor{myred}{dashed red}) is unimodal, the four-moment distribution
 (\textcolor{myblue}{solid blue}) is bimodal. The five-moment distribution
-would not be distinguishable from the four-moment one. The lower panel in
-the figure shows the corresponding distributions for the second data set;
+would not be distinguishable from the four-moment one.  Figure~\ref{fig:plots_2_4mom_b} shows the corresponding distributions for the second data set;
 also in this case they are visually very distinct.
@@
 as proposed here, are therefore different from those applying it at the
-sample level -- and also visually clearer. The plot in the upper panel of
-\fig~\ref{fig:plots_2_4mom}, showing the two clearly different
+sample level -- and also visually clearer. The plot in 
+\fig~\ref{fig:plots_2_4mom_a}, showing the two clearly different
 distributions constrained by two and four moments, should be compared with
@@
 typical. For the first data set, for example, most frequency distributions
-have two high-frequency regimes of activity, as in the upper panel of
-\fig~\ref{fig:plots_5mom}, with roughly the same heights and almost the
+have two high-frequency regimes of activity, as in 
+\fig~\ref{fig:plots_5mom_a}, with roughly the same heights and almost the
 same locations, within a few percentages of normalized total activity. Each

### 2019-12-05 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-05T10:55:47+0100
+%% Last-Updated: 2019-12-05T11:13:56+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \includegraphics[width=\linewidth]{_scripts/newpaper_stensola_5mom_distributionsN_priorc.pdf}%
-\\[3em]%
+\\[2em]%
 \subcaption[]{\label{fig:plots_5mom_b}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper_riehle_5mom_distributionsN_priorc.pdf}%
-\caption{***}
+\\[2em]%
+\caption{Maximum-entropy distributions for the normalized total activity of
+  the larger population, assuming several population sizes $N$. Five
+  moments are constrained.}
 \label{fig:plots_5mom}
@@
 \includegraphics[width=\linewidth]{_scripts/newpaper3_gridvsnogrid.pdf}%
-\\[3em]%
+\\[2em]%
 \subcaption[]{\label{fig:plots_cellsubsets_b}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper3_gridconv.pdf}%
-\caption{***}
+\\[2em]%
+\caption{\subcaptionref{fig:plots_cellsubsets_a} Maximum-entropy
+  distributions for larger subpopulations or grid and non-grid cells.
+  \subcaptionref{fig:plots_cellsubsets_b} Convolution of the two
+  subpopulation distributions; it would be equal to the total distribution
+  if the two distributions were independent.}
 \label{fig:plots_cellsubsets}
@@
 \includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_twomom_distributions10000_priorc.pdf}%
-\\[3em]%
+\\[2em]%
 \subcaption[]{\label{fig:plots_2_4mom_b}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_twomom_distributions10000_prior2.pdf}%
-\caption{***}
+\\[2em]%
+\caption{Maximum-entropy distributions for the normalized total activity of
+  the larger population, assuming two different sets of constraining moments
+  and a larger population size $N=10\,000$.}
 \label{fig:plots_2_4mom}
@@
 \includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
-\caption{***}
+\caption{Marginals for the normalized total activity of the sample,
+  obtained from the maximum-entropy distributions with $N=10\,000$ from
+  \fig~\ref{fig:plots_2_4mom_a}, and from maximum-entropy distributions
+  calculated at the sample level, that is with $N=n=65$.}
 \label{fig:comparisons_marginals}

### 2019-12-05 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-05T11:13:56+0100
+%% Last-Updated: 2019-12-05T11:17:52+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \\[3em]%
-\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
+\includegraphics[width=0.95\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
 \caption{Marginals for the normalized total activity of the sample,
@@
   \fig~\ref{fig:plots_2_4mom_a}, and from maximum-entropy distributions
-  calculated at the sample level, that is with $N=n=65$.}
+  calculated at the sample level, that is with $N=n=65$, with two and four
+  moments in either case.}
 \label{fig:comparisons_marginals}

### 2019-12-05 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-05T11:17:52+0100
+%% Last-Updated: 2019-12-05T11:25:20+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 statement that it gives \enquote{the maximally unbiased (or non-committal)
-  distribution} must be taken with a grain of salt. \citep[\mynote{move to
-  sect}]{jaynes1986d_r1996,portamana2009,portamana2017}. Second, a
+  distribution} must be taken with a grain of salt.
+\citep{jaynes1986d_r1996,portamana2009,portamana2017}. Second, a
 maximum-entropy distribution like $\px(\yA \| M, N)$ is the zeroth-order
@@
   of Excellence scheme of the  Research Council of Norway (Yasser Roudi group).\\
-  PGLPM thanks Mari, Miri, \amp\ Emma for continuous encouragement and
+  PGLPM thanks the staff of the NTNU library for their always prompt
+  support; Mari, Miri, \amp\ Emma for continuous encouragement and
   affection; Buster Keaton and Saitama for filling life with awe and

### 2019-12-05 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-05T11:25:20+0100
+%% Last-Updated: 2019-12-05T11:35:44+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}\labelsep 0pt%
-  \noindent \mynote{Abstract must be rewritten once paper is ready}This
-  work shows how to build a maximum-entropy probabilistic model for the
-  total activity of a population of neurons, given only some activity data or
-  statistics -- for example, empirical moments -- of a \emph{subpopulation}
-  thereof. This kind of model is useful because neuronal recordings are
-  always limited to a very small sample of a population of neurons.
-%
-  The model is applied to two sets of neuronal data available in the
-  literature. In some cases it makes interesting forecasts about the larger
-  population -- for example, two low-regime modes in the frequency
-  distribution for the total activity -- that are not visible in the sample
-  data or in maximum-entropy models applied only to the sample.
-%
-  For the two datasets, the maximum-entropy probability model applied only
-  to the subpopulation is compared with the marginal probability
-  distribution obtained from the maximum-entropy model applied to the full
-  population. On a linear probability scale no large differences are
-  visible, but on a logarithmic scale the two distributions show very
-  different behaviours, especially in the tails.
+
+% \noindent \mynote{Abstract must be rewritten once paper is ready}This
+%   work shows how to build a maximum-entropy probabilistic model for the
+%   total activity of a population of neurons, given only some activity data or
+%   statistics -- for example, empirical moments -- of a \emph{subpopulation}
+%   thereof. This kind of model is useful because neuronal recordings are
+%   always limited to a very small sample of a population of neurons.
+% %
+%   The model is applied to two sets of neuronal data available in the
+%   literature. In some cases it makes interesting forecasts about the larger
+%   population -- for example, two low-regime modes in the frequency
+%   distribution for the total activity -- that are not visible in the sample
+%   data or in maximum-entropy models applied only to the sample.
+% %
+%   For the two datasets, the maximum-entropy probability model applied only
+%   to the subpopulation is compared with the marginal probability
+%   distribution obtained from the maximum-entropy model applied to the full
+%   population. On a linear probability scale no large differences are
+%   visible, but on a logarithmic scale the two distributions show very
+%   different behaviours, especially in the tails.
   \\\noindent\emph{\footnotesize Note: Dear Reader \amp\ Peer, this
@@
 
-When political elections are approaching, mass media are constantly
-bubbling with results and discussions of pre-election polls. Why do we
-conduct pre-election polls and other kinds of survey sampling? The
-simplest, intuitive answer is \enquote{because they give us an idea of what
-  the final election outcomes will be}. Can this answer be made more
-quantitative?
-
-Yes. In the simplest cases, the probability calculus give us simple and
-precise quantitative relations between our expectations about a population
--- of voters, for instance -- and about a sample thereof. Suppose there are
-$N$ voters, $S$ of which prefer one political candidate and $N-S$ of which
-prefer another. We don't know $S$ or $N-S$. Now take a sample of $n$
-voters, in such a way that we can assume that every set of $n$ voters is
-equally likely to be sampled. In this sample, $s$ voters prefer the first
-candidate, and $n-s$ the second. The probability calculus tells us that our
-expectations of $s/n$ and $S/N$ are equal:
-\begin{equation}\label{eq:poll_equal_expectations}
-  \E(s/n \| \yH) = \E(S/N \| \yH),
-\end{equation}
-where $\yH$ represents our state of knowledge in this example. A more
-precise answer to our original question then is: \enquote{because we
-  \emph{expect} -- in the full probabilistic sense -- to see the same
-  proportion of political preferences in a sample as in the larger
-  population}.
-
-The exact equality above extends to other quantities beyond the means $S/N$
-and $s/n$: it holds for all normalized \emph{factorial} moments
-\citep{potts1953,broca2005}[\sect~A]{portamanaetal2015}. These expectation
-equalities follow from a precise relation between our \dobs\ about the
-polled voters and about all voters. The relations above are powerful
-because they hold no matter what the probability distribution $\pf(S\|\yH)$
-may be.
-
-\bigskip
-
-These probabilistic properties are of outstanding importance in
-neuroscience, where we try to understand the workings of brain areas that
-may include tens of thousands of neurons, but we can record only few tens
-or hundreds of them. Whenever we record the activity of $n$ neurons we're
-effectively polling them on the activity of all $N$ neurons under the
-recording device -- those that could equally likely have been sampled
-instead. Suppose the recorded activities are time-binned and binarized. At
-every time bin $t$, $s(t)$ out of the $n$ neurons \enquote{vote} for
-activity, and $n-s(t)$ vote for silence. The probability calculus tells us
-that expectation equalities such as~\eqref{eq:poll_equal_expectations} hold
-for the total activity $S(t)$ of the larger population of $N$ neurons that
-could have been recorded.
-
-Expectation properties are especially important for the maximum-entropy
-method, which is used for various purposes in the neuroscientific
-literature. This method is typically applied constraining the equality of
-expectations and empirical averages. Expectation equalities such
-as~\eqref{eq:poll_equal_expectations} then have a far-reaching consequence:
-\emph{from empirical statistics of the activity of a \textbf{sample} of
-  neurons we can construct maximum-entropy distributions for the total
-  activity of the \textbf{larger population} of neurons that could have been
-  sampled by the recording device}.
-
-% In the present work we show how to use the maximum-entropy method to assign
-% a probability distribution for the total activity at a new time bin of a
-% population of neurons, \emph{most of which can be unrecorded}, given
-% specific statistics about the activity of a recorded \emph{sample} of the
-% larger population. This kind of probability assignment is extremely important
-% in neuroscience: we try to understand the workings of brain areas that may
-% include tens of thousands of neurons, but we can record only few tens or
-% hundreds of them.
-
-% Our approach also corrects a slight misuse of the maximum-entropy method
-% as sometimes appears in the literature. When this method is applied to a
-% set of neurons, it implicitly assumes that the set is completely isolated
-% from other neurons. We'll reveal this implicit assumption in several ways.
-% The probability distribution thus constructed is therefore affected by this
-% hidden assumption. We'll study in which circumstances the application of
-% the method at the sample level and at the larger-population level give
-% similar or different probabilities.
-
-\hrule
-OLD TEXT
-\hrule
-
-\section{Mathematical development}
-\label{sec:maths_development}
-
-\begin{subequations}
-  \label{eq:all_equal_expectations_poll}
-\begin{equation}
-  \label{eq:equal_factorialmoments_polls}
-  \begin{split}
-  \E\condt[\big]{\tfrac{s\;(s-1)}{n\;(n-1)}}{\yH} &= 
-  \E\condt[\big]{\tfrac{S\;(S-1)}{N\;(N-1)}}{\yH},\\
-  \E\condt[\big]{\tfrac{s\;(s-1)\;(s-2)}{n\;(n-1)\;(n-2)}}{\yH} &= 
-  \E\condt[\big]{\tfrac{S\;(S-1)\;(S-2)}{N\;(N-1)\;(N-2)}}{\yH},
-  \\
-  \text{and in general}\quad
-  \E\condt[\big]{\tbinom{s}{r}\;\tbinom{n}{r}^{-1}}{\yH} &=
-  \E\condt[\big]{\tbinom{S}{r}\;\tbinom{N}{r}^{-1}}{\yH},
-  \quad r<N.
-\end{split}
-\end{equation}
-
-Note that from the first $r$ factorial moments we can calculate the first
-$r$ power moments and vice versa; but the simple equalities above don't
-hold for the power moments \citep[\sect~A]{portamanaetal2015}.
-
-\begin{equation}
-  \label{eq:relation_ps_pS}
-  \pf(s \| \yH) = \sum_{S} \binom{n}{s}\;\binom{N-n}{S-s}\;\binom{N}{S}^{-1}\;
-  \pf(S \| \yH).
-\end{equation}
-\end{subequations}
-
-Suppose we have recorded the firing activity of a hundred neurons,
-sampled from a particular brain area. What are we to do with such data?
-Gerstein \etal\ \citey{gersteinetal1985} posed this question very
-tersely (our emphasis):
-\begin{quote}\small
-  The principal conceptual problems are (1) \emph{defining cooperativity or
-    functional grouping} among neurons and (2) \emph{formulating
-    quantitative criteria} for recognizing and characterizing such
-  cooperativity.
-\end{quote}
-These questions have a long history, of course; see for instance the 1966
-review by Moore \etal\
-\citey{mooreetal1966}. % This scenario is a concrete possibility
-% thanks to recent electrophysiological techniques \citep{berenyietal2014}.
-The neuroscientific literature has offered several mathematical definitions
-of \enquote{cooperativity} or \enquote{functional grouping} and criteria to
-quantify it.
-
-One such quantitative criterion relies on the \me\ or relative-\me\ method
-\citep{jaynes1957,jaynes1963,hobsonetal1973,sivia1996_r2006,meadetal1984}.
-This criterion has been used in neuroscience at least since the 1990s,
-applied to data recorded from brain areas as diverse as retina and motor
-cortex
-\citep{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015},
-and it has been subjected to mathematical and conceptual scrutiny
-\citep{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.
-
-\enquote{Cooperativity} can be quantified and characterized with \me\
-methods in several ways. The simplest way roughly proceeds along the
-following steps. Consider the recorded activity of a sample of $n$ neurons.
-\begin{enumerate}
-\item The activity of each neuron, a continuous signal, is divided into $T$
-  time bins and binarized in intensity, and thus transformed into a
-  sequence of digits \enquote{$0$}s (inactive) and \enquote{$1$}s (active)
-  \citep[\cf][]{caianiello1961,caianiello1986}.
-
-  Let the variable $\yav_i(t)\in\set{0,1}$ denote the activity of the $i$th
-  sampled neuron at time bin $t$. Collectively denote the $n$ activities
-  with $\yaa(t) \defd \bigl(\yav_1(t),\dotsc,\yav_n(t)\bigr)$. The
-  population-averaged activity at that bin is
-  $\ya(t) \defd \sum_i \yav_i(t)/n$. If we count the number of distinct
-  pairs of active neurons at that bin we combinatorially find
-  $\tbinom{n\ya(t)}{2}\equiv n\ya(t)\,[n\ya(t)-1]/2$. There can be at
-  most $\tbinom{n}{2}$ simultaneously active pairs, so the
-  population-averaged pair activity is
-  $\av{\yaa \yaa}(t) \defd \tbinom{n}{2}^{-1}\tbinom{n\ya(t)}{2}$. With some
-  combinatorics we see that the population-averaged activity of $m$-tuples
-  of neurons is
-  \begin{equation}
-    \label{eq:products_intermsof_average}
-    \av{\underbrace{\yaa\dotsm\yaa}_{\text{$m$ terms}}}(t)
-    = \binom{n}{m}^{-1}\binom{n\ya(t)}{m}.
-  \end{equation}
-
-  For brevity let us agree to simply call \enquote{activity} the average
-  $\ya$, \enquote{pair-activity} the average $\av{\yaa \yaa}$, and so on.
-
-\item Construct a sequence of relative-\me\ distributions for the activity
-  $\ya$, using this sequence of constraints:
-  \begin{itemize}
-  \item the time average of the activity: $\tav{\ya} \defd \sum_t\ya(t)/T$;
-  \item the time averages of the activity and of the pair-activity
-    $\tav{\av{\yaa \yaa}} \defd \sum_t\av{\yaa \yaa}(t)/T$;
-  \item \ldots
-  \item the time averages of the activity, of the pair-activity, and so on, up
-    to the $k$-activity.
-  \end{itemize}
-  Call the resulting distributions $p_1(\ya), p_2(\ya),\dotsc,p_k(\ya)$.
-  The time-bin dependence is now absent because these distributions can be
-  interpreted as referring to any one of the time bins $t$, or to a new
-  time bin (in the future or in the past) containing new data.
-
-  We also have the empirical frequency distribution of the total activity,
-  $f(\ya)$, counted from the time bins.
-
-\item Now compare the distributions above with one another and with the
-  frequency distribution, using some probability-space distance like the
-  relative entropy or discrimination information
-  \citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}. If we find,
-  say, that such distance is very high between $p_1$ and $f$, very low
-  between $p_2$ and $f$, and is more or less the same between all $p_m$ and
-  $f$ for $m \ge 2$, then we can say that there is a \enquote{pairwise
-    cooperativity}, and that any higher-order cooperativity is just a
-  reflection or consequence of the pairwise one. The reason is that the
-  information from higher-order simultaneous activities did not lead to
-  appreciable changes in the distribution obtained from pair activities.
-\end{enumerate}
-The protocol above needs to be made precise by specifying various
-parameters, such as the width of the time bins or the probability
-distance used.
-
-We hurry to say that the description just given is just \emph{one} way to
-quantify and characterize cooperativity and functional grouping, not
-\emph{the only} way. It can surely be criticized from many points of view.
-Yet, it is quantitative and bears a more precise meaning than an undefined,
-vague notion of \enquote{cooperativity}. Two persons who apply this
-procedure to the same data will obtain the same numbers. Different
-protocols can be based on the \me\ method, for instance protocols that take
-into account the activities or pair activities of specific neurons rather
-than population averages, or even protocols that take into account time
-dependence.
-
-The purpose of the present work is not to assess the merits of \me\ methods
-with respect to other methods. Its main purpose is to show that there is a
-problem in the way the \me\ method itself, as sketched above, is applied to
-the activity of the recorded neurons. We believe that this problem is at
-the root of some quirks about this method that were pointed out in the
-literature \citep{roudietal2009b}. This problems extends also to more
-complex versions of the method, possibly except versions that use
-\enquote{hidden} neurons
-\citep{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}. The problem is
-that the recorded neurons are a \emph{sample} from a larger, unrecorded
-population, but the \me\ method as applied above is treating them as
-isolated from the rest of the brain. Hence, the results it provides cannot
-be rightfully extrapolated. We will give a mathematical proof of this. Let
-us first analyse this issue in more detail.
-
-Suppose that the neurons were recorded with electrodes covering an area of
-some square millimetres \citep[\cf][]{berenyietal2014}. This recording is a
-sample of the activity of the neuronal population under the recording
-device, which can amount to tens of thousands of neurons
-\citep{abeles1991}. We could even consider the recorded neurons as a sample
-of a brain area more extended than the recording device.
-
-The characterization of the cooperativity of the recorded sample would have
-little meaning if we did not expect its results to generalize to a larger,
-unrecorded population -- at the very least the population under the
-recording device. In other words, we expect that the conclusions drawn with
-the \me\ methods about the sampled neurons should somehow extrapolate to
-unrecorded neurons in some larger area, from which the recorded neurons
-were sampled. In statistical terms we are assuming that the recorded
-neurons are a \emph{representative sample} of some larger neuronal
-population. Probability theory tells us how to make inferences from a
-sample to the larger population from which it is sampled \parentext{see
-  references below}.
-
-We can apply the \me\ method to the sample, as described in the above
-protocol, to generate probability distributions for the activity of the
-sample. But, given that our sample is representative of a larger
-population, we can also apply the \me\ method to the larger (unrecorded)
-population. The constraints are the same: the time averages of the sampled
-data, since they constitute representative data about the larger population
-as well. The method thus yields a probability distribution for the larger
-population, and the distribution for the sample is obtained by
-marginalization. The problem is that \emph{the distributions obtained from
-  these two applications differ}. Which choice is most meaningful?
-
-In this work we develop the second way of applying the \me\ method, at the
-level of the larger population, and show that its results differ from the
-application at the sample level. We also consider the case where the size
-of the larger population is unknown.
-
-To apply the \me\ method to the larger, unsampled population, it is
-necessary to use probability relations relevant to sampling
-\citep{ghoshetal1997}[parts~I,
-VI]{freedmanetal1978_r2007}[\chap~8]{gelmanetal1995_r2014}[\chap~3]{jaynes1994_r2003}.
-The relations we present are well-known in survey sampling and in the
-pedagogic problem of drawing from an urn without replacement, yet they are
-somewhat hard to find explicitly written in the neuroscientific literature.
-We present and discuss them in the next section. A minor purpose of this
-paper is to make these relations more widely known, because they can be
-useful independently of \me\ methods.
-
-The notation and terminology in the present work follow ISO and ANSI standards
-\citep{iso1993,ieee1993,nist1995,iso2006,iso2006b} but for the use of the
-comma \enquote{,} to denote logical conjunction. Probability notation
-follows Jaynes \citep{jaynes1994_r2003}. By \enquote{probability} we mean a
-degree of belief which \enquote{would be agreed by all rational men if
-  there were any rational men} \citep{good1966}.
-
-\section{Probability relations between population and  sample}
-\label{sec:prob_samples}
-
-We have already introduced the notation for the sample neurons. We
-introduce an analogous notation for the $N$ neurons constituting the
-larger population, but using the corresponding capital letters:
-$\yAv_{i}(t)$ is the activity of the $i$th neuron at time bin $t$,
-$\yA(t) \defd \sum_{i} \yAv_{i}(t)/N$ is the activity at that
-bin averaged over the larger population, and so on.
-
-The probability relations between sample and larger population are valid at
-every time bin. As we mentioned above, the \me\ distribution refers to any
-time bin or to a new bin. For these reasons we will now omit the time-bin
-argument \enquote{$(t)$} from our expressions.
-
-% Probabilities refer to statements about the quantities we observe. We use
-% the standard notation:
-% \begin{equation}
-%   \label{eq:notation_statements}
-%   \begin{aligned}
-%     &\text{\enquote{$\yXv_{\iota} = \yAv_{\iota}$} 
-%       stands for \enquote{the activity of the $\iota$th neuron is $\yAv_{\iota}$}},
-%     \\
-%     &\text{\enquote{$\yXf = \yA$} 
-%       stands for \enquote{the (population-averaged) activity of the neurons is $\yA$}},
-%     \\
-%     &\text{\enquote{$\yxv_i = \yav_i$} 
-%       stands for \enquote{the activity of the $i$th sample neuron is $\yav_i$}},
-%   \end{aligned}
-% \end{equation}
-% and similarly for other quantities.
-
-If $\yHc$ denotes our state of knowledge -- the evidence and assumptions
-backing our probability assignments -- our uncertainty about the full
-activity of the larger population is expressed by the joint probability
-distribution
-\begin{equation}
-  \label{eq:joint_plaus}
-  \pf(\yAv_1, \yAv_2, \dotsc, \yAv_N \| \yHc)
-  \quad\text{or}\quad
-\pf(\yAA \| \yHc), \quad \yAA \in \set{0,1}^N.
-\end{equation}
-Our uncertainty about the state of the sample is likewise expressed by
-\begin{equation}
-  \label{eq:sample_plaus}
-  \pf(\yav_1, \yav_2, \dotsc, \yav_n \| \yHc) \quad\text{or}\quad
-\pf(\yaa \| \yHc), \quad \yaa \in \set{0,1}^n.
-\end{equation}
-
-\bigskip
-
-The theory of statistical sampling is covered in many excellent texts, for
-example Ghosh \amp\ Meeden \citey{ghoshetal1997} or Freedman, Pisani, \amp\
-Purves \citey[parts~I, VI]{freedmanetal1978_r2007}; summaries can be found
-in Gelman \etal\ \citey[\chap~8]{gelmanetal1995_r2014} and Jaynes
-\citey[\chap~3]{jaynes1994_r2003}.
-
-We need to make an initial probability assignment for the state of the larger
-population before any experimental observations are made. This initial
-assignment will be modified by our experimental observations, and these can
-involve just a sample of the population. Our state of knowledge and initial
-probability assignment should reflect that samples are somehow
-representative of the whole population.
-
-In this state of knowledge, denoted $\yH$, we know that the neurons in the
-population are biologically or functionally similar, for example in
-morphology or the kind of input or output they receive or give. But we are
-completely ignorant about the physical details of the individual neurons.
-Our ignorance is therefore symmetric under permutations of neuron
-identities. This ignorance is represented by a probability distribution
-that is symmetric under permutations of neuron identities; such a
-distribution is usually called \emph{finitely exchangeable}
-\citep{ericson1969}[\chap~1]{ghoshetal1997}. We stress that this
-probability assignment is just an expression of the symmetry of our
-\emph{ignorance} about the state of the population, not an expression of
-some biologic or physical symmetry or identity of the neurons.
-
-The \emph{representation theorem for finite exchangeability} states that,
-in the state of knowledge $\yH$, the symmetric distribution for the full
-activity is completely determined by the distribution for its
-population-average:
-\begin{equation}
-  \label{eq:joint_plaus_N_homog}
-  \pf(\yAA \|  \yH) \equiv
-  \sum_{\yA}\pf(\yAA \| \yA, \yH)\,
-  \pf(\yA \| \yH) =
-  \binom{N}{N\yA}^{-1} \pf(\yA \| \yH).
-\end{equation}
-The equivalence on the left is just an application of the law of total
-probability; the equality on the right is the statement of the theorem.
-This result is intuitive: owing to symmetry, we must assign equal
-probabilities to all $\tbinom{N}{N\yA}$ activity vectors with
-$N\yA$ active neurons; the probability of each activity vector is
-therefore given by that of the average activity divided by the number of
-possible vector values. Proof of this theorem and generalizations to
-non-binary and continuum cases are given by de~Finetti
-\citey{definetti1959b}, Kendall \citey{kendall1967}, Ericson
-\citey{ericson1976}, Diaconis \amp\ Freedman
-\citey{diaconis1977,diaconisetal1980}, Heath \amp\ Sudderth
-\citey{heathetal1976}.
-
-Our uncertainties about the larger population and the sample are connected
-via the conditional probability
-\begin{equation}
-  \label{eq:conditional_hypergeometric}
-  \pf(\ya \|\yA, \yH)=
-  \binom{n}{n\ya}\binom{N-n}{N \yA-n\ya}\binom{N}{N \yA}^{-1}
-  \defs \yG_{\ya\yA},
-\end{equation}
-which is a hypergeometric distribution, typical of \enquote{drawing without
-  replacement} problems. The combinatorial proof of this expression is in
-fact the same as for this class of problems
-\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-
-Using the conditional probability above we obtain the probability for the
-activity of the sample:
-\begin{gather}
-  \label{eq:subpop_average}
-  \pf(\ya \| \yH) = \sum_{\yA}
-  \pf(\ya \|\yA, \yH)\,
-  \pf(\yA \| \yH)
-  = 
-  \sum_{\yA}
-  \yG_{\ya\yA}\,
-  \pf(\yA \| \yH).
-\end{gather}
-It should be proved that the probability distribution for the full activity
-of the sample is also symmetric and completely determined by the
-distribution of its population-averaged activity:
-\begin{equation}
-  \label{eq:marginal}
-  \pf(\yaa \| \yH) = \binom{n}{n\ya}^{-1} \pf(\ya \| \yH).
-\end{equation}
-This is intuitively clear: our initial symmetric ignorance should also
-apply to the sample. The distribution for the
-sample~\eqref{eq:subpop_average} indeed satisfies the same representation
-theorem~\eqref{eq:joint_plaus_N_homog} as the distribution for the larger
-population.
-
-The conditional probability
-$\pf(\ya \|\yA, \yH) \equiv \yG_{\ya \yA}$, besides
-relating the distributions for the population and sample activities via
-marginalization, also allows us to express the expectation value of any
-function of the sample activity, $\yg_{\ya}$, in terms of the distribution
-for the larger population, as follows:
-\begin{equation}
-  \label{eq:pullback_P}
-  \E(\yg\|I)
-  \equiv
-  \sum_{\ya} \yg_{\ya}\,\pf(\ya \| \yH)
-  =
-  \sum_{\ya} \yg_{\ya} \sum_{\yA} \yG_{\ya\yA}\,\pf(\yA \| \yH)
-  =%{}\\
-  \sum_{\yA} \bigl( \tsum_{\ya} \yg_{\ya}  \yG_{\ya\yA} \bigr)\,
-  \pf(\yA \| \yH),
-%  =
-%  \sum_{N\yA=0}^N \sum_{n\ya=0}^n f(\ya)\yG_{\ya\yA}\,\p(\yXf=\yA \| \yH)
-%  \equiv  \expeb{\yg^*}.
-\end{equation}
-where the second step uses \eqn~\eqref{eq:subpop_average}. The last
-expression shows that the expectation of the function $\yg_{\ya}$ is equal to
-the expectation of the function
-$\yg^*(\yA) \defd \sum_{\ya} \yg_{\ya}\,\yG_{\ya\yA}$.
-
-\bigskip
-
-The final expression in \eqn~\eqref{eq:pullback_P} is important for our
-\me\ application: the requirement that the function $\yg$, defined for the
-sample, have a  value $\yc$ obtained from observed data,
-\emph{translates into a linear constraint for the distribution of the larger
-  population}:
-\begin{equation}
-  \label{eq:constraint_extended}
-  \yc = \E(\yg \| \yH) \equiv \sum_{\yA} \bigl( \tsum_{\ya} \yg_{\ya}  \yG_{\ya\yA} \bigr)\;
-  \pf(\yA \| \yH).
-\end{equation}
-
-In particular, when the function $\yg$ is the $m$-activity of the sample,
-$\yg_{\ya} = \sav{\yaa\dotso\yaa} \equiv \binom{n\ya}{m}/\binom{n}{m}$, we
-find
-\begin{multline}
-  \label{eq:expe_products}
-%  \label{eq:pullback_m_expectations}
-%  \yG^*\colon \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}
-%  \mapsto
-    % \bigl( \sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}}\bigr)^* =
-    % \av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}},
-    % \\[2\jot]
-    % \expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH} =
-    % \expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH}
-  \E(\sav{\underbrace{\yaa \dotsm \yaa}_{\text{$m$ factors}}} \| \yH)
-\equiv
-    % \\[2\jot]
-    % {}
-    % \frac{(n-m)!}{n!}
-    \sum_{\ya} %m!\,
-    \binom{n}{m}^{-1}
-    \binom{n \ya}{m}\, \pf(\ya \| \yH)
-    % \sum_{N \yA=0}^N %m!\,
-    % \binom{N-m}{N \yA-m} \binom{N}{N \yA}^{-1} \p(\yXf=\yA \| \yH)
-    ={}\\
-    \binom{N}{m}^{-1}
-    % \frac{(N-m)!}{N!}
-    \sum_{\yA} %m!\,
-    \binom{N \yA}{m}\, \pf(\yA \| \yH)
-\equiv    \E(\av{\underbrace{\yAA \dotsm \yAA}_{\text{$m$ factors}}} \| \yH),
-\end{multline}
-that is, \emph{the expected values of the $m$-activities of the sample and
-  of the larger population are equal}. The proof of the middle equality uses
-the expression for the $m$th factorial moment of the hypergeometric
-distribution and can be found in \textcite{potts1953}. Similar relations
-can be found for the raw moments $\E(\ya^m)$ and $\E(\yA^m)$, which
-can be written in terms of the product expectations using
-\eqn~\eqref{eq:products_intermsof_average}.
-
-Thus, in a \me\ application, when we require the expectation of the
-$m$-activity of a sample to have a particular value, we are also
-requiring  the expectation of the $m$-activity of the larger population to
-have the same value.
-
-\bigskip
-
-\begin{figure}[!b]
-\centering
-\includegraphics[width=\linewidth]{pop_sample_projection3.pdf}%
-\caption{Log-density plot of the hypergeometric distribution
-  $%\pf(\ya \|\yA, \yH)=
-\yG_{\ya\yA} \defd  \raisebox{0pt}[0pt][1ex]{}\binom{n}{n\ya}\binom{N-n}{N \yA-n\ya}\binom{N}{N \yA}^{-1}$ for $N=5000$, $n=200$. (Band artifacts may appear in the
-  colourbar depending on your \textsc{pdf} viewer.)}
-\label{fig:hypergeom_proj}
-\end{figure}%hypergeometric_identity.nb
-These expectation equalities between sample and larger population should not
-be surprising: we intuitively \emph{expect} that the proportion of coloured
-balls sampled from an urn should be roughly equal to the proportion of
-coloured ball contained in the urn. The formulae in the present section
-formalize and mathematically express our intuition. The hypergeometric
-distribution $\yG_{\ya\yA}$ plays an important role in this
-formalization. A look at its plot, \fig~\ref{fig:hypergeom_proj}, reveals
-that it is a sort of \enquote{fuzzy identity transformation}, or fuzzy
-Kronecker delta, between the $\yA$-space $\set{0,\dotsc,N}$ and
-$\ya$-space $\set{0,\dotsc,n}$. From \eqn~\eqref{eq:marginal} we thus have
-that
-\begin{equation}
-  \label{eq:roughly_equal_nN}
-  \pf(\ya=a \|\yH) \approx \pf(\yA=a \|\yH),\qquad
-\E[\yg_{\ya} \|\yH] \approx \E[\yg_{\yA} \|\yH],
-\end{equation}
-where $\yg$ is any smooth function defined on $\clcl{0,1}$. These approximate
-equalities express the intuitive fact that \emph{our uncertainty about the
-  sample is representative of our uncertainty about the population and
-  about other samples}, and vice versa. When $n=N$, $\yG_{\ya\yA}$
-becomes the identity matrix and the approximate equalities above become
-exact -- of course, since we have sampled the larger population.
-
-But the approximate equalities above may miss important features of the two
-probability distributions. In the next section we will in fact emphasize
-their differences. If the distribution for the population average $\yA$ is
-bimodal, for example, the bimodality can be lost in the distribution for
-the sample average $\ya$, owing to the coarsening effect of
-$\yG_{\ya\yA}$.
-
-% They should be contrasted with the limits
-% $\pf(\yxs=a) \to \pf(\yXf=a)$ and $\expeb{\yg_{\yxs}} \to \expeb{\yg_{\yXf}}$, as
-% $n\to N$, do: these limits are trivially, universally valid because the
-% sample becomes the larger population as $n\to N$. In particular, these limits
-% hold even in cases where the conditional probability
-% $\pf(\yxs = \ya \|\yXf=\yA)$ is not a fuzzy identity, our uncertainties
-% about sample and about population can differ wildly, and the approximate
-% equalities~\eqref{eq:roughly_equal_nN} do not hold.
-
-\section{Maximum-entropy: sample level \vs\ larger-population level}
-\label{sec:specific_initial_probability}
-
-In the previous section we have seen that observations about a sample can
-be used as constraints on the distribution for the activity of the larger
-population. Let us use such constraints with the \me\ method. Suppose that
-we want to constrain $m$ functions of the sample activity, vectorially
-written $\yg \defd (c_1,\dotsc,c_m)$, to $m$ values
-$\yc \defd (\widehat{c}_1,\dotsc,\widehat{c}_m)$. These functions are typically $k$-activities
-$\sav{\yaa\dotso \yaa}$, and the values are typically the time averages of
-the observed sample, as discussed in \sect~\ref{sec:intro}:
-$\yc = \sum_t \yg[\ya(t)]/T$.
-
-Let us apply the relative-\me\ method \citep{sivia1996_r2006,meadetal1984}
-directly to sampled neurons; denote this approach by $\yHb$. Then we apply
-the method to the larger population of neurons, most of which are unsampled;
-denote this approach by $\yHa$.
-
-Applied directly to the sampled neurons, the method yields the distribution
-\begin{equation}
-  \label{eq:app_maxent_sample}
-  \pf(\ya \|\yHb)
-  =\frac{1}{\yk(\yl)}\,
-  \binom{n}{n\ya}
-  \exp[\yl\T \yg_{\ya}]
-\end{equation}
-where $\yk(\yl)$ is a normalization constant. The binomial in front of the
-exponential appears because we must account for the multiplicity by which
-the population-average activity $\ya$ can be realized: $\ya=0$ can be
-realized in only one way (all neurons inactive), $\ya=1/n$ can be realized
-in $n$ ways (one active neuron out of $n$), and so on. This term is analogous to
-the \enquote{density of states} in front of the Boltzmann factor in
-statistical mechanics \citep[\chap~16]{callen1960_r1985}. The $m$ Lagrange
-multipliers $\yl\defd (l_1,\dotsc,l_m)$ must satisfy the $m$ constraint
-equations
-\begin{equation}
-  \label{eq:app_maxent_sample_constraints}
-  \yc = \E(\yg \|\yHb) \equiv
-  \frac{1}{\yk(\yl)}\sum_{\ya}  \yg_{\ya} \binom{n}{n\ya}
-  \exp[\yl\T \yg_{\ya}].%, \qquad k=1,\dotsc,m.
-\end{equation}
-
-Applied to the larger population, using the constraint
-expression~\eqref{eq:constraint_extended} derived in the previous section,
-the method yields the distribution for the larger-population activity
-\begin{equation}
-  \label{eq:app_maxent_pop}
-  \pf(\yA \| \yHa)  = \frac{1}{\yK(\yll)}\,
-  \binom{N}{N \yA}\,\exp\bigl(\yll\T
-  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
-\end{equation}
-The $m$ Lagrange multipliers $\yll\defd (\lambda_1,\dotsc,\lambda_m)$ must
-satisfy the $m$ constraint equations
-\begin{equation}
-  \label{eq:app_maxent_pop_constraints}
-  \yc = \E(\yg \|\yHa)\equiv %{}\\
- \frac{1}{\yK(\yll)}\sum_{\ya} \sum_{\yA} \yg_{\ya} \yG_{\ya \yA}
-\,
-  \binom{N}{N \yA}\,\exp\bigl(\yll\T
-  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
-%  \\ k=1,\dotsc,m.
-\end{equation}
-
-We obtain the distribution for the sample activity by marginalization, using
-\eqn~\eqref{eq:marginal}:
-\begin{equation}
-  \label{eq:app_maxent_pop_marg}
-  \pf(\ya \| \yHa)  = \frac{1}{\yK(\yll)}\, 
-  \sum_{\yA} \yG_{\ya \yA}
-\,
-  \binom{N}{N \yA}\,\exp\bigl(\yll\T
-  \tsum_{\ya} \yg_{\ya}\yG_{\ya\yA}\bigr).
-\end{equation}
-
-The distributions for the sample activity,
-\eqns~\eqref{eq:app_maxent_pop_marg} and \eqref{eq:app_maxent_sample},
-obtained with the two approaches $\yHb$ and $\yHa$, are different. From the
-discussion in the previous section we expect them to be vaguely similar;
-yet they cannot be exactly equal, because their equality would require the
-$2m$ quantities $\yll$ and $\yl$ to satisfy the constraint equations
-\eqref{eq:app_maxent_pop_constraints} and
-\eqref{eq:app_maxent_sample_constraints}, and in addition also the $n$
-equations $\pf(\ya \| \yHa) = \pf(\ya \| \yHb)$,
-$\ya=1/n,\dotsc,1$ (one equation is taken care of by the normalization of
-the distributions). We would have a set of $2m+n$ equations in $2m$
-unknowns.
-
-Hence, \emph{the applications of \me\ at the sample level and at the
-  larger-population level are inequivalent}. They lead to numerically
-different distributions for the sample activity $\yaa$.
-
-%%% FIGURES %%%
-\iffalse
-\begin{figure}[!t]
-\centering
-\includegraphics[width=\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
-\caption{Linear and log-plots of $\pf(\ya)$ constructed by \me\ at
-  the population level followed by sample marginalization (blue triangles),
-  \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:app_maxent_sample}, with $N=5000$,
-  $n=200$, and constraints as in \eqn~\eqref{eq:constraints}.}
-\label{fig:diff_maxent_pop_sample}
-\end{figure}%maxent_pop_or_sample.nb
-%
-\begin{figure}[!t]
-\centering
-\includegraphics[width=\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}% 
-\caption{Linear and log-plots of $\pf(\ya)$ constructed by \me\ at
-  the population level followed by sample marginalization (blue triangles),
-  \eqn~\eqref{eq:app_maxent_pop_marg}, and at the sample level (red
-  circles), \eqn~\eqref{eq:app_maxent_sample}, with $N=5000$,
-  $n=200$, and constraints as in \eqn~\eqref{eq:constraints}.}
-\label{fig:diff_maxent_pop_sample_realdata}
-\end{figure}%maxent_pop_or_sample.nb
-\fi%
-\begin{figure}[!t]
-\centering
-\includegraphics[width=\linewidth]{comparison3.pdf}%
-\caption{Linear and log-plots of $\pf(\ya)$ for a sample of $n=200$
-  and constraints as in \eqn~\eqref{eq:constraints}, constructed by:
-  \textcolor{myred}{\textbf{red squares:}} \me\ at the sample level,
-  \eqn~\eqref{eq:app_maxent_sample};
-  \textcolor{mypurpleblue}{\textbf{blue triangles:}} \me\ at the
-  population level, \eqn~\eqref{eq:app_maxent_pop_marg} with
-  $N=10\,000$, followed by sample marginalization;
-  \textcolor{myyellow}{\textbf{yellow circles:}} \me\ at the population
-  level with unknown population size,
-  \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}, according to the
-  distribution~\eqref{eq:pdf_popsize} for the population.}
-\label{fig:all_three}
-\end{figure}%maxent_pop_or_sample.nb\fi
-The distribution obtained at the sample level will show different
-features from the one obtained at the population level, like displaced or
-additional modes or particular tail behaviour. We show an example of this
-discrepancy in \fig~\ref{fig:all_three}, for
-$N=10\,000$, $n=200$, and the two constraints
-\begin{equation}
-  \label{eq:constraints}
-  \E(\ya) = 0.0478,\qquad
-  \E(\yaas) = 0.00257,
-  % ,\quad
-  % \expe{\sav{\yaa\yaa\yaa}} = 1.48\times 10^{-4},\quad
-  % \expe{\sav{\yaa\yaa\yaa\yaa}} = 8.81 \times 10^{-6}.
-\end{equation}
-which come from the actual recording of circa 200 neurons from macaque
-motor cortex \citep{rostamietal2016_r2017}. The distribution obtained at the
-population level (blue triangles) has a higher and displaced mode and a quite
-different behaviour for activities around $0.5$ than the distribution
-obtained at the sample level (red squares).
-
-\bigskip
-
-In our discussion we have so far assumed the size $N$ of the larger
-population to be known. This is rarely the case, however. We usually are
-uncertain about $N$ and can only guess its order of magnitude. In such a
-state of knowledge $\yHd$ our ignorance about the possible value of $N$
-is expressed by a probability distribution $\pf(\yNN=N \|\yHd)=h(N)$,
-and the marginal distribution for the sample
-activity~\eqref{eq:app_maxent_pop_marg} is modified, by the law of total
-probability, to
-\begin{multline}
-  \label{eq:app_maxent_pop_marg_unknown_N}
-  \pf(\ya \| \yHd)  =
-  \sum_{N} \pf(\ya \| N,\yHd) \,
-  \pf(N \|\yHd)
-  ={}\\[-\jot]
-  \sum_N \biggl\{\frac{1}{\yK(\yll_N)}\, 
-  \sum_{\yA} \yG^{(N)}_{\ya\yA}
-\,
-  \tbinom{N}{N \yA}\,\exp\Bigl[{\yll_N}\T
-  \tsum_{\ya} \yg_{\ya}\yG^{(N)}_{\ya\yA}\Bigr]\biggr\}
-  \;h(N),
-\end{multline}
-where the Lagrange multipliers $\yll_{N}$ and the summation range for
-$\yA$ depend on $N$.
-
-As a proof of concept, \fig~\ref{fig:all_three} also shows such a distribution
-(yellow circles) for the same constraints as above, and a probability
-distribution for $N$ inspired by Jeffreys
-\citep[\sect~4.8]{jeffreys1939_r1983}:
-\begin{equation}
-  \label{eq:pdf_popsize}
-  h(N) \propto 1/N, \qquad
-  N \in\set{1\,000,\; 2\,000,\; \dotsc,\;10\,000}.
-\end{equation}
-
-\section{Derivation from the probability calculus}
-\label{sec:derivation_prob_calculus}
-
-There are three inequivalent main routes that lead to a probability
-distribution of the maximum-entropy form~\eqref{eq:app_maxent_sample}
-or~\eqref{eq:app_maxent_pop}. The distribution carries a different
-interpretation under each route \citep[pp.~52--55,
-72--77]{jaynes1979b}[pp.~25--28]{jaynes1982}[\sect~I]{jaynes1982b}{jaynes1986d_r1996}[\sect~11.1]{jaynes1994_r2003}.
-
-\begin{enumerate}[wide,label=(\alph*)]
-\item \label{item:entropy_route}One route is the choice of the distribution having the highest
-  Shannon entropy, given only a quantitative assessment some of its
-  properties, such as expectations. The numerical choice of the value of
-  such properties is a (subjective) assumption.
-
-  \medskip
-
-  In the two other routes the maximum-entropy distribution is obtained as
-  an \emph{approximation} of a distribution obtained via the probability
-  calculus, using data coming from a set of $T$ measurements -- as in our
-  present case.\mynote{refs here} Also in this case some (subjective)
-  assumptions are necessary: they concern our beliefs about the long-run
-  relative frequencies of the measurement outcomes:\footnote{Such
-    assumptions are always necessary at the beginning of an inference:
-    \enquote{Now the axioms of probability enable us to infer any
-      probability-conclusion \emph{only} from probability-premisses. In
-      other words, the calculus of probability does not enable us to infer
-      any probability-value unless we have some probabilities or
-      probability relations \emph{given}. Such data cannot be supplied by
-      the mathematician. E.g. the rules of arithmetic and the axioms of the
-      probability-calculus are utterly impotent to determine, on the
-      supposed knowledge that the throw of a coin must yield either head or
-      tail and cannot yield both, the probability that it will yield head
-      or that it will yield tail. We must assume that the two co-exclusive
-      and co-exhaustive possibilities are \emph{equally probable}, before
-      we can estimate the probability of either as being a half of
-      certitude} \citep[\emph{Appendix on eduction}, \sect~5,
-    p.~182]{johnson1924}.}
-
-\item \label{item:entr_prior_route}In one case we consider all possible
-  sets of measurement outcomes to be \emph{roughly} equally likely; this
-  leads to a probability for the frequencies $\ynu$ proportional to a
-  multinomial coefficient $\binom{L}{L\ynu}$, with $L$ large but smaller
-  than $T$. (We cannot assume the sets of measurement outcomes to be
-  exactly equally likely, because this is equivalent to their independence:
-  \cf\
-  \cites[\sect~6.7]{jaynes1994_r2003}[\sect~B]{portamana2009}[\sect~2]{portamana2017}.)
-  The exact expression is\mynote{equation here}
-  %   \begin{multline}
-  %   \label{eq:prob_conditional_average_exch}
-  %   \p\bigl[\yE{N+1}{k} \bigcond \tsum\yo\yf{N}\in\yA, \yX\bigr] ={}\\
-  %   \frac{\int\yqq_k\sum_{\yf{N}}  \delt(\tsum \yo\yf{N}\in \yA)\,
-  %     \binom{N}{N\yf{N}}\,\bigl( \tprod \yq^{N\yf{N}} \bigr)  \,
-  %     \pf(\yq \|\yX)  \,\di\yq}{ \int\sum_{\yf{N}} \delt(\tsum \yo\yf{N}\in \yA)\,
-  %     \binom{N}{N\yf{N}}\,\bigl( \tprod \yq^{N\yf{N}} \bigr) \,
-  %     \pf(\yq \|\yX) \,\di\yq },
-  % \end{multline}
-
-\item \label{item:suff_route}In the other case we assume that the
-  measurements have a \emph{sufficient statistics}: the same as appears in
-  the exponential of the maximum-entropy distribution. The exact expression is\mynote{equation here}
-  % \begin{multline}
-  %   \label{eq:prob_conditional_average_suff}
-  %   \p\bigl[\yE{N+1}{k} \bigcond \tsum\yo\yf{N}\in\yA, \yS\bigr] ={}
-  %       \\[3\jot]
-  %    \frac{\int \pf(k \| \yl,\yaa,\yS) \sum_{\yf{N}} \delt(\tsum \yo\yf{N}\in \yA)\,
-  %     \binom{N}{N\yf{N}}\,\bigl[ \tprod  \pf(\yk \| \yl,\yaa,\yS)^{N\yf{N}} \bigr]
-  %      \,
-  %     \pf(\yl \|\yS) \,\di\yl}{\int\sum_{\yf{N}} \delt(\tsum \yo\yf{N}\in \yA)\,
-  %     \binom{N}{N\yf{N}}\,\bigl[ \tprod \pf(\yk \| \yl,\yaa,\yS)^{N\yf{N}} \bigr]
-  %      \,
-  %     \pf(\yl \|\yS) \,\di\yl}.
-  % \end{multline}
-\end{enumerate}
-
-It's important to keep in mind that the approximate equivalence of these
-three routes only holds under very specific assumptions -- which \emph{have
-  physical and biological meanings and consequences}. In particular,
-route~\ref{item:suff_route} implies that we can discard other empirical
-statistics of the data, if they are known; whereas
-route~\ref{item:entr_prior_route} requires us to specify all known
-empirical statistics, because using only a subset of them may lead to
-different results. Route~\ref{item:entropy_route} is also supposed to be
-used with all known data. Moreover, the approximate equivalence of
-route~\ref{item:entropy_route} with routes~\ref{item:entr_prior_route} and
-\ref{item:suff_route} \emph{only holds if $T$ is much larger than the
-  possible values of the activity $\yA$}. Finally, we also obtain very
-different expressions depending on whether we're asking about \emph{the
-  activity in one of the \textbf{recorded} time bins} or about \emph{the
-  activity in a \textbf{new} time bin}. Works that use maximum-entropy
-distributions are often very vague about the latter point.
-
-\bigskip
-
-Here are the distributions for our \dobs\ about three different quantities,
-assuming an entropic pre-data distribution for the long-run frequencies of
-the larger population:
-
-\textbf{The frequency distribution $\yF$ of the larger-population activity
-  \emph{during the recording}}
-
-\begin{equation}
-  \label{eq:fullfreq_from_samplefreq}
-  \begin{split}
-  \p(\yF \|\yf,\yH) &\propto
-  \!\begin{multlined}[t][0.8\linewidth]
-  \int\!\di\ynu\;  \tsum_{\yj}
-  \delt(\tsum_{\yA}\yjj_{\ya\yA} = \yff_{\ya})\;
-  \delt(\tsum_{\ya}\yjj_{\ya\yA} = \yFF_{\yA})\times{}\\
-  \binom{T}{T\yj}\;
-  \biggl[\prod_{\ya\yA}
-(\yG_{\ya\yA}\ynuu_{\yA})^{T\yjj_{\ya\yA}}
-\biggr]\;
-\exp[-L\;\re(\ynu;\yR)]
-\end{multlined}
-\\
-&\appropto \delt(\tsum_{\yA}\yG_{\ya\yA}\yFF_{\yA} = \yff_{\ya})\;
-\exp[-L\;\re(\yF;\yR)]
-\end{split}
-\end{equation}
-
-\bigskip
-
-\textbf{The long-run frequency distribution $\ynu$ of the larger-population
-  activity}
-
-\begin{equation}
-  \label{eq:longrun_from_samplefreq}
-  \begin{split}
-  \p(\ynu \|\yf,\yH) &\propto
-    \binom{T}{T\yf}\;
-  \biggl[\prod_{\ya}
-(\tsum_{\yA}\yG_{\ya\yA}\ynuu_{\yA})^{T\yff_{\ya}}
-\biggr]\;
-\exp[-L\;\re(\ynu;\yR)]
-\\
-&\appropto 
-\exp[-T\;\re(\yf; \mathte{\yG}\ynu)
--L\;\re(\ynu;\yR)]
-\end{split}
-\end{equation}
-Note that if some $\yff_{\ya}$ are zero, then the first exponential may
-be badly approximated by a delta, because the constraints lie on a facet of the
-simplex of frequencies $\set{\ynu}$.
-
-\bigskip
-
-\textbf{The activity $\yA'$ of the larger-population
-  in a \emph{new} time bin}
-
-\begin{equation}
-  \label{eq:newbin_from_samplefreq}
-  \begin{split}
-  \p(\yA' \|\yf,\yH) &\propto
-  \int\!\di\ynu\;\ynuu_{\yA'}\;
-  \binom{T}{T\yf}\;
-  \biggl[\prod_{\ya}
-(\tsum_{\yA}\yG_{\ya\yA}\ynuu_{\yA})^{T\yff_{\ya}}
-\biggr]\;
-\exp[-L\;\re(\ynu;\yR)]
-\\
-&\appropto 
-\int\!\di\ynu\;\ynuu_{\yA'}\;
-\exp[-T\;\re(\yf; \mathte{\yG}\ynu)
--L\;\re(\ynu;\yR)]
-\end{split}
-\end{equation}
-Note that if some $\yff_{\ya}$ are zero, then the first exponential may
-be badly approximated by a delta, because the constraints lie on a facet of the
-simplex of frequencies $\set{\ynu}$.
-
-\bigskip
-
-The maximum-relative-entropy distribution is, in the first two cases, an
-approximation of the most probable frequency distribution $\yF$ or $\ynu$;
-in the third case, an approximation of the probability distribution for
-$\yA'$.
-
-\section{Assumptions and limitations}
-\label{sec:assumptions}
-
-Main assumptions behind this belief distribution:
-
-We are approximating our state of knowledge with a finitely exchangeable
-one. In turn, this is numerically approximated by an infinitely
-exchangeable one for $T$ large. But we don't really have an exchangeable
-belief: our \dob\ that the activity at the next time bin will
-differ from the activity at the present one is roughly proportional to the
-difference in the two subsequent activities.
-
-The formulae say that we have equal beliefs about the underlying population
-states having the same activity. This isn't really our belief, for we
-believe there are interactions between the neurons and subpopulations thereof.
-
-\section{Discussion}
-\label{sec:discussion}
-
-The purpose of the present work was to point out and show, in a simple
-set-up, that the \me\ method can be applied to recorded neuronal data in a
-way that accounts for the larger population from which the data are
-sampled, \eqns~\eqref{eq:app_maxent_pop}--\eqref{eq:app_maxent_pop_marg}.
-This application leads to results that differ from the standard application
-which only considers the sample in isolation,
-\eqns~\eqref{eq:app_maxent_sample}--\eqref{eq:app_maxent_sample_constraints}.
-We gave a numerical example of this difference. We have also shown how to
-extend the new application when the size of the larger population is
-unknown, \eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}.
-
-The latter formula, in particular, shows that the standard way of applying
-\me\ 
-implicitly assumes that \emph{no} larger population exists beyond the
-recorded sample of neurons. One could in fact object to the application at
-the population level, and say that the traditional way of applying \me,
-\eqn~\eqref{eq:app_maxent_sample}, yields different results because it does
-not make assumptions about the size $N$ of a possibly existing larger
-population. Such a state of uncertainty, however, is correctly formalized
-according to the laws of probability by introducing a probability
-distribution for $N$, and is expressed by
-\eqn~\eqref{eq:app_maxent_pop_marg_unknown_N}. This expression cannot
-generally be equal to~\eqref{eq:app_maxent_sample} unless the distribution
-for $N$ gives unit probability to $N=n$; that is, unless the sample
-\emph{is} the larger population, and no larger population exists.
-
-The standard \me\ approach therefore assumes that the recorded neurons
-constitute a special subpopulation, isolated from the larger population of
-neurons in which it is embedded, and which was also present under the
-recording device. This assumption is unrealistic. The \me\ approach at the
-population level does not make such assumption and is therefore preferable.
-It may reveal features in a data set that were unnoticed by the standard
-\me\ approach.
-
-% This hidden assumption of isolation is unreasonable. It amounts to say that
-% the neurons we distinguished and tracked with our recording device are a
-% very special, isolated set among all those that could have been recorded.
-% It is for this reason that we find the \me\ application at the population
-% level preferable. Physical models of neuronal populations usually include some
-% sort of external input, mimicking an embedding in a larger population. The
-% \me\ distribution obtained at the population level may reveal features in a
-% data set that were unnoticed by the standard \me\ model.
-
-The difference in the resulting distributions between the applications at
-the sample and at the population levels appears in the use of Boltzmann
-machines with hidden units \citep{lerouxetal2008}, although by a different
-conceptual route. It also appears in statistical mechanics: if a system is
-statistically described by a \me\ Gibbs state, its subsystems cannot be
-described by a Gibbs state \citep{maesetal1999}. A somewhat similar
-situation also appears in the statistical description of the final state of
-a non-equilibrium process starting and ending in two equilibrium states: we
-can describe our knowledge about the final state either by (1) a Gibbs
-distribution, calculated from the final equilibrium macrovariables, or (2)
-by the distribution obtained from the Liouville evolution of the Gibbs
-distribution assigned to the initial state. The two distributions differ
-(even though the final \emph{physical} state is obviously exactly the same
-\citep[\sect~4]{jaynes1985d_r1993}), and the second allows us to make
-sharper predictions about the final physical state thanks to our knowledge
-of its preceding dynamics. In this example, though, both distributions are
-usually extremely sharp and practically lead to the same predictions. In
-neuroscientific applications, the difference in predictions of the sample
-\vs\ larger-population applications can instead be very relevant.
-
-The idea of the new application leads in fact to more questions. For
-instance:
-\begin{itemize}
-\item Do the standard and new applications lead to different or contrasting
-  conclusions about \enquote{cooperativity}, when applied to real data
-  sets?
-\item How to extend the new application to the \enquote{inhomogeneous} case
-  \citep{schneidmanetal2006,shlensetal2006,roudietal2009b}, in which
-  expectations for individual neurons or groups of neurons are constrained?
-\item What is the mathematical relation between the new application and
-  \me\ models with hidden neurons
-  \citep{smolensky1986,kulkarnietal2007,huang2015,dunnetal2017}?
-\end{itemize}
-Owing to space limitations we must leave a thorough investigation of these
-questions to future work.
-
-Finally, we would like to point out the usefulness and importance of the
-probability formulae that relate our states of knowledge about a population
-and its samples, presented in \sect~\ref{sec:prob_samples}. This kind of
-formulae is essential in neuroscience, where we try to understand
-properties of extended brain areas from partial observations. The
-formulae presented here reflect a simple, symmetric state of ignorance.
-More work is needed \citep[\cf][]{levinaetal2017} to extend these formulae
-to account for finer knowledge of the cerebral cortex and its population
-properties.
-
-\clearpage
-
-\hrule
-\newcounter{dummy}\refstepcounter{dummy}\label{orig_intro}
-\mynote{orig intro}
-Experimental technologies to record the activity of many neurons at the
-same time in different species and brain areas are rapidly advancing. These
-experimental advancements are paralleled by advances in theoretical and
-computational methods for analyzing the data accumulated using the
-recording technologies. Such theoretical methods usually take the form of
-probabilistic models that try to describe the multi-neuronal activity of
-the recorded neurons. With such probabilistic models one aims to address
-numerous issues: What correlations are important in describing the
-multi-neuronal pattern? How does the pattern of activity covary with
-external stimuli or experimental conditions? What dimensionality does the
-neural data live in and how is this related to the underlying population
-interactions? The probabilistic models can also be used to make predictions
-about the structure of the neural code, by studying the properties of the
-fitted model, or by generating synthetic data from it.
-
-In general, despite the rapid advances in recording technology, the best
-experimental measurements of neuronal activity still only provide data from
-a small subset of neurons that comprise a neuronal population. A wealth of
-studies on building probabilistic models of neural data focuses on
-describing such subsets, ignoring the fact that the observed neurons is a
-small group in a much bigger set of hidden neurons. Some other studies do
-include hidden variables which, amongst other things, aim to model the
-global features of the unrecorded neurons, but we still lack a through
-understanding of how to include the role of hidden neurons in probabilistic
-models, how our inferences about the recorded activity would be affected by
-them, and what we can we say about the rest of the population by studying the
-heavily subsampled recordings. In this paper, we aim to address these
-questions in the case of a simple maximum entropy model, namely the
-homogeneous maximum entropy model.
-
-The maximum entropy approach has been used in a variety of setting for
-building statistical models of complex systems and datasets, ranging from
-neuronal activity in the retina, in the cortex, protein sequences, gene
-regulatory populations and natural images. The general idea is, for a dataset,
-to write down the distribution that maximizes the entropy of state
-variables, given some low order statistics. Now given the fact that the
-recorded neurons are a fraction of the neurons in the population, several
-quantitative questions arise that we will address in this paper: given the
-data from the sampled neurons, can we build a maximum entropy model over
-the whole population? Once we build such a population level maximum entropy
-model, can we see features in the neural activity which cannot be directly
-seen from a model build from the sampled neurons? Since we can marginalize
-down the population level maximum entropy model to the sampled population, how
-does this margianzlied maximum entropy model match the sample level model?
-
-All these questions can be answered in the case of the homogeneous maximum
-entropy model. First we show how to go from the sample level maximum
-entropy level to the population level, by assuming different sizes of the
-population and also by assuming an uninformative prior over the size of the
-population. This is done by inferring the statistics of correlation functions
-at the population level from those of the sample level by using simple
-counting arguments. We then find that, when applied to experimental
-recordings from the Medial Entorhinal Cortex of rats and the monkey visual
-cortex, this population level maximum entropy model may exhibit features that
-the sample level model does not predict. Specifically, we observed modes in
-the distribution of the activity in the population level model that do not
-show up in the sample level. We study how the assumed size of the larger
-population affects the appearance of these modes and find that there is a
-minimum size of the larger population for which such modes can be observed. We
-then compared the distribution found by marginalizing the larger population
-maximum entropy model down to the sample level, and the distribution fit
-directly to the sample level. For the two datasets that we tested, we found
-that the two distributions match each other to a large degree but that
-there are also differences between them. We quantify how these differences
-also depend on the assumed size of the population and find that …(WE SHOULD
-TEST SHI) This predicts that for a large enough population (DO WE PREDICT
-THAT IF THE LARGER POPULATION GETS BIGGER THE DIFFERENCE ALSO GET BIGGER)?…
-
-The rest of the paper is organized as follows. We first describe how to go
-from the sample level maximum entropy model to the larger population maximum
-entropy model. In section 2, we apply this to the two experimental datasets
-and study the effect of the assumed size of the population as well as the
-moments that we use for building the maximum entropy model. In section 3 we
-compare the distributions found from marginalizing the maximum entropy
-model down to the sample level and the original sample level model.
-
-\hrule
-\bigskip
-
 %%% Local Variables: 

### 2019-12-05 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-05T11:35:44+0100
+%% Last-Updated: 2019-12-05T23:42:10+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\firstpublished}{***}
-\newcommand*{\propertitle}{Maximum-entropy distributions for a 
-  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
-}
+\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
 \newcommand*{\pdftitle}{\propertitle}
@@
 \begin{abstract}\labelsep 0pt%
-
+***
 % \noindent \mynote{Abstract must be rewritten once paper is ready}This
@@
 
+The $T= \infty$ approximation can break the method for specific
+distributions of sample frequencies, if too many moments are used. For
+example take $n=2$ with frequency distribution $\yf=(0.5,0,0.5)$ for the
+activities $\ya=0,1,2$; that is, the sample is silent half of the time and
+completely active half of the time. The first two normalized factorial
+moments have both value $1/2$, and together they uniquely determine $\yf$.
+If we now take $N=3$, it is possible to show that
+\eqn~\eqref{eq:app_maxent_pop_first}, constrained on those two moments, has
+no solution. The simple reason is clear from a sampling perspective: if
+there is zero probability of sampling one active neuron, then there must be
+zero probability of sampling two active ones. Without the $T=\infty$
+approximation this doesn't happen, because the conditional frequency of the
+activity levels and the sampling conditional probability are kept distinct.
+
 The two infinity approximations are at the limit of their validity in our
 specific application, so the distributions discussed in our example are
-best interpreted as \emph{most probable frequency distributions}.
-
-The space of possible frequency distributions for the larger population has
-$N$ dimensions, however, and probability distributions in high dimensions
-have counter-intuitive properties. For example, the mode or mean can have
+best interpreted as \emph{most probable frequency distributions}. The space
+of possible frequency distributions for the larger population has $N$
+dimensions, however, and probability distributions in high dimensions have
+counter-intuitive properties. For example, the mode or mean can have
 \emph{atypical} features when compared with the majority of other probable

### 2019-12-06 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-05T23:42:10+0100
+%% Last-Updated: 2019-12-06T11:13:57+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 completely active half of the time. The first two normalized factorial
-moments have both value $1/2$, and together they uniquely determine $\yf$.
-If we now take $N=3$, it is possible to show that
+moments have both value $0.5$, and together they uniquely determine $\yf$.
+If we now take $N=3$, it can be proved that
 \eqn~\eqref{eq:app_maxent_pop_first}, constrained on those two moments, has
-no solution. The simple reason is clear from a sampling perspective: if
-there is zero probability of sampling one active neuron, then there must be
-zero probability of sampling two active ones. Without the $T=\infty$
-approximation this doesn't happen, because the conditional frequency of the
-activity levels and the sampling conditional probability are kept distinct.
+no solution. The reason is clear from a sampling perspective: if there is
+zero probability of sampling one active neuron, then the larger population
+cannot have any active neurons, and there must be zero probability of
+sampling two active ones. But this reasoning holds only for each single
+time bin, not for the averages over time bins. The $T=\infty$ approximation
+identifies the two instead. Outside this approximation there are no
+contradictions because the conditional frequencies and conditional
+probabilities are kept distinct.
 

### 2019-12-06 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-06T11:13:57+0100
+%% Last-Updated: 2019-12-06T11:26:19+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Introduction:\\a simple model for questions about large neuronal
+\section{Introduction:\\a model for questions about large neuronal
   populations}
@@
 degrees of reasonable belief -- that is, probabilities -- to the possible
-answers. The assessment of this distribution of belief involves
+answers. The assessment of such distributions of belief involves
 experimental data, such as recordings of neuronal activity from specific
-brain areas, and pre-data knowledge about biological conditions and
-mechanisms. Our pre-data \dobs\ are often simplified, to be mathematically
-tractable, and are therefore called \enquote{models}.
+brain areas, and pre-data knowledge and assumptions about biological
+conditions and mechanisms. Our pre-data \dobs\ are often simplified, to be
+mathematically tractable, and are therefore called \enquote{models}.
 
@@
 
-\mynote{[Yasser:] In the present work we aim to address the issue of
-  building probability distributions over a large population using
-  recording from subpopulations by considering the distribution of total
-  population activity as an example. [} In the present work we propose such
+% \mynote{[Yasser:] In the present work we aim to address the issue of
+%   building probability distributions over a large population using
+%   recording from subpopulations by considering the distribution of total
+%   population activity as an example. [} 
+In the present work we propose such
 an intermediate probability model. It answers this question: How much was

### 2019-12-06 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-06T11:26:19+0100
+%% Last-Updated: 2019-12-06T12:02:48+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 shoulder in its low-activity mode. These results show that the proposed
-method can lead to the formulation or preliminary assessment of interesting
+model can lead to the formulation or preliminary assessment of interesting
 hypotheses, as we will illustrate with a toy example. Note that these
@@
 
-Our approach also solves a methodological problem in the use of
+Our model also solves a methodological problem in the use of
 maximum-entropy methods to assess the \enquote{cooperativity},
@@
 
-How large is the full population addressed by the method proposed here? Its
+How large is the full population addressed by the model proposed here? Its
 size must have some limit and can't obviously include the full brain. The
@@
 In \sect~\ref{sec:assumptions} we discuss in detail the assumptions and
-approximations on which the method is based, from the point of view of the
+approximations on which the model is based, from the point of view of the
 probability calculus.
@@
 A summary of all points above and a discussion of the usefulness of the
-method is given in the final \sect~\ref{sec:summary_discussion}.
+model is given in the final \sect~\ref{sec:summary_discussion}.
 
@@
 
-\section{Method: maximum-entropy and sampling}
+\section{Model: maximum-entropy and sampling}
 \label{sec:model}
@@
 change across sessions? How much total activity should we expect at any
-time bin during a recording? The approach presented here gives a
+time bin during a recording? The model presented here gives a
 probability distribution that approximately answers all these questions.
@@
 
-We apply the method just described to two data sets publicly available in
+We apply the model just described to two data sets publicly available in
 the literature:
@@
 Such features are clearly not present in the sample frequencies or in the
-maximum-entropy distribution at the sample level. The application of the
-probability calculus thus reveals interesting features of the larger
-population. \textcolor{white}{If you find this you can claim a postcard
-  from us.}
+maximum-entropy distribution at the sample level. The model thus suggests
+interesting features of the larger population. \textcolor{white}{If you
+  find this you can claim a postcard from us.}
 
@@
 % maxent_solutions_different_moments_partial.nb
-Here is a toy example of possible uses of this maximum-entropy approach,
+Here is a toy example of possible uses of this model,
 based on the first data set. We could be interested in the hypothesis that
@@
 two peaks in the distribution of \fig~\ref{fig:plots_5mom} reflect the
-activities of grid versus non-grid cells? We apply the method to these two
+activities of grid versus non-grid cells? We apply the model to these two
 sets of neurons individually, using $\yNg=4\,150$ for the grid set and
@@
 The toy analysis above should not be taken literally, but just as an
-illustration of the method's possible applications. The important point is
-that this method is computationally very cheap and yet it can provide
+illustration of the model's possible applications. The important point is
+that this model is computationally very cheap and yet it can provide
 useful insights, even if just qualitative ones, and even suggest new

### 2019-12-06 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-06T12:02:48+0100
+%% Last-Updated: 2019-12-06T17:01:42+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 experimental data, such as recordings of neuronal activity from specific
-brain areas, and pre-data knowledge and assumptions about biological
-conditions and mechanisms. Our pre-data \dobs\ are often simplified, to be
-mathematically tractable, and are therefore called \enquote{models}.
+brain areas, and pre-data knowledge and hypotheses about biological
+conditions and mechanisms. When we translate such pre-data beliefs into an
+initial probability distribution we often simplify them in more or less
+realistic ways, to make the analysis mathematically tractable. This is why
+such initial probability distributions are called \enquote{models}.
 
@@
 functionally distinguished brain region. Many probabilistic models focus on
-such samples only: they somehow neglect, in their mathematical assumptions,
-that the recorded neurons are a sample from a larger population. Such
-isolating assumptions sometimes escape attention, being subtly hidden in
-the mathematics. Some probabilistic models try to take unrecorded neurons
-into account -- but by describing each neuron individually, thus becoming
-very complex \citep{huang2015,battistinetal2017}. It would be useful to
-explore models that operate in between, addressing the larger brain area
-whence the sample comes, but collectively, without asking about individual
-neuronal details. Such intermediate models would be useful for preliminary
-investigations, to help us to decide which hypotheses to discard and which
-to consider for more complex and costly studies, or to suggest new
-hypotheses.
+such samples only: they neglect that the recorded neurons are a sample from
+a larger population. Such isolating assumptions sometimes escape attention,
+being subtly hidden in the mathematics. Some probabilistic models try to
+take unrecorded neurons into account, but they do so by describing each
+neuron individually, thus becoming very complex
+\citep{huang2015,battistinetal2017}. It would be useful to explore models
+that operate in between, addressing the activity of the larger brain area
+whence the sample comes, but \emph{collectively}, that is, without asking
+about individual neuronal details. Such intermediate models would be useful
+at least for preliminary investigations, to help us to decide which
+hypotheses to discard and which to consider for more complex and costly
+models, or to suggest new hypotheses.
 
@@
 %   population activity as an example. [} 
-In the present work we propose such
-an intermediate probability model. It answers this question: How much was
-the \emph{total} activity of a large neuronal population, given the
-observation of the activity of a very small sample thereof? This model
-addresses a larger brain area, avoiding the assumption of isolation of the
-sample; and by focusing on the total activity, rather than the activity of
-individual neurons, it remains simple and numerically tractable.
-
-The model we propose is based on the straightforward combination of the
+In the present work we propose such an intermediate probabilistic model. It
+answers this question: How much was the \emph{total} activity of a large
+neuronal population, given the observation of the activity of a very small
+sample thereof? This model addresses a larger brain area, avoiding the
+assumption of isolation of the sample; and by focusing on the total
+activity, rather than the activity of individual neurons, it remains simple
+and numerically tractable.
+
+The model can be viewed as a straightforward combination of the
 maximum-entropy method and basic sampling relations from the probability
@@
 in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b_r1996,grandy1980}
-has been used for different kinds of investigations about the neuronal
-activity of various brain areas and about other phenomena of importance to
-the neurosciences, for example gene and protein interaction \citep[for
+to build initial probability distributions has been used for different
+kinds of investigations about the neuronal activity of various brain areas
+and about other phenomena of importance to the neurosciences, for example
+gene and protein interaction \citep[for
 example][]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,mackeetal2009b,tkaciketal2009,roudietal2009c,barreiroetal2010,gerwinnetal2010,mackeetal2011b,ganmoretal2011,cohenetal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}.
-
-We illustrate possible uses of our proposed model in
-\sect~\ref{sec:application}, by applying it to two concrete data sets:
+However, our model can also be viewed as an approximation of a more
+detailed model within a fully Bayesian approach; we shall constantly keep
+this point of view in mind in our discussion.
+
+We illustrate possible uses of our model in \sect~\ref{sec:application}, by
+applying it to two concrete data sets:
 \begin{enumerate*}[label=(\alph*)]
 \item the activity of 65 neurons recorded for 20 min from a rat's Medial
-  Entorhinal Cortex \citep{stensolaetal2012}, \item the activity of 159
-  neurons recorded for 15 min from a macaque's Motor Cortex
-  \citep{rostamietal2016_r2017}.
+  Entorhinal Cortex \citep{stensolaetal2012},
+\item the activity of 159 neurons recorded for 15 min from a macaque's
+  Motor Cortex \citep{rostamietal2016_r2017}.
 \end{enumerate*}
-For each data set the model gives us the most plausible frequencies with
-which all levels of total activity of a much larger population,
-1\,000--10\,000 neurons, appeared during the recording. For example it can
-tell us that 60 out of 10\,000 neurons were likely active during 1\% of the
-recording time (though not necessarily always the same 60), 250 neurons out
-of 10\,000 were active during 0.4\% of the time, and so on. The precise
-meaning of this frequency distribution is explained in
+For each data set the model gives us the most plausible frequencies of all
+levels of total activity of a much larger population, 1\,000--10\,000
+neurons in size, during the recording. For example it can tell us that 60
+out of 10\,000 neurons were most likely active during 1\% of the recording
+time (though not necessarily always the same 60), 250 neurons out of
+10\,000 were active during 0.4\% of the recording time, and so on. The
+precise meaning of this frequency distribution is explained in
 \sect~\ref{sec:model}. For the two example data sets, the guessed frequency
-distributions in the larger population are distinctly different from those
-in the sample. For the first data set the frequency distribution for the
+distributions for the larger population are distinctly different from those
+for the sample. For the first data set the frequency distribution for the
 larger population has two very distinct modes, both at low
@@
 model can lead to the formulation or preliminary assessment of interesting
-hypotheses, as we will illustrate with a toy example. Note that these
-guessed features of the full population could not be inferred by the
-application of maximum-entropy \emph{to the sample alone}.
+hypotheses, as we will illustrate with toy examples in
+\sects~\ref{sec:application} and~\ref{sec:marginalization}. Note that it is
+not possible to guess these features of the larger population by applying
+the maximum-entropy method \emph{at the level of the sample alone}.
 
@@
 When the difference in size between a large population and a sample thereof
-is too large, the presence of some inferential properties of correlations
-for the sample implies the \emph{lack} of such properties for the larger
-population, and vice versa. Maximum-entropy applications at the sample
-level can therefore deceive us in questions regarding the cooperativity of
-the larger population. The application to the larger population is more
-reliable. We discuss this point in detail in
+is too large, the sufficiency of correlations of some order for the sample
+implies the \emph{lack} of sufficiency of correlations of the same order
+for the larger population, and vice versa. Maximum-entropy applications at
+the sample level can therefore deceive us in questions regarding the
+cooperativity of the larger population. Our model addresses directly the
+latter instead. We explain this point more precisely in
 \sect~\ref{sec:marginalization}.
@@
 \section{Quantifying the importance of higher-order correlations:\\
-  the limitations of methods at the sample level}
+  the limitations of models at the sample level}
 \label{sec:marginalization}
@@
 these terms]{bohteetal2000,amarietal2003} of neuronal activity. Most, if
-not all, such applications apply the method \emph{at the sample level}. In
+not all, such applications use the method \emph{at the sample level}. In
 this section we discuss how our proposed application bears on this kind of
 quantification and compare it with the traditional application at the
-sample level. But such a comparison involves some important methodological
+sample level. But such a comparison involves some methodological
 caveats which we wish to discuss first.% For concreteness we use an
@@
 thereof. % \citep[contrast this with some literature, \eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}
-The approach presented here avoids this problem because the maximum-entropy
+The model presented here avoids this problem because the maximum-entropy
 method is applied to obtain the distribution of the larger population, not
@@
 
-Compare the results above with those obtained by applying the method at the
+Compare the results above with those obtained by applying the model at the
 sample level, that is, with $N=n$:
@@
 leading to the equality of factorial
-moments~\eqref{eq:equal_m_factorialmoments}, on which our approach rests.
+moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests.
 This conditional probability, characteristic of \enquote{drawing without
@@
 A rigorous assessment would involve a more expensive, full Bayesian
-calculation; but such calculation would make the whole maximum-entropy
-approach superfluous. We discuss this final point in the next section.
+calculation; but such calculation would make the maximum-entropy
+method superfluous. We discuss this final point in the next section.
 
@@
 
-The possible uses of the method here presented depend on the assumptions
-and approximations on which it rests: three assumptions and two
+The possible uses of the model here presented depend on the assumptions and
+approximations which define it: three main assumptions and two main
 approximation all in all. These become clear within a full-fledged
@@
 moments under study \citep{portamana2017}. This assumption is crucial for
-obtaining the maximum-entropy method with the Shannon entropy. A different
+deriving the maximum-entropy based on with the Shannon entropy. A different
 density leads to alternative maximum-entropy methods, for example based on
@@
 
-Finally, the maximum-entropy method typically approximates the number $T$
-of time bins, the prior coefficient $L$, and the ratio $T/L$ to infinity.
-These approximations lead to the identification of the four distributions
-listed at the end of \sect~\ref{sec:model}: the recorded frequency
-distribution becomes equal to the long-run one, and also to the probability
-distribution for the activity at recorded and new time bins. In our
-application it also leads to the equality of the hypergeometric
+Finally, the maximum-entropy method is typically equivalent to
+approximating the number $T$ of time bins, the prior coefficient $L$, and
+the ratio $T/L$ to infinity. These approximations lead to the
+identification of the four distributions listed at the end of
+\sect~\ref{sec:model}: the recorded frequency distribution becomes equal to
+the long-run one, and also to the probability distribution for the activity
+at recorded and new time bins. In our application it also leads to the
+equality of the hypergeometric
 distribution~\eqref{eq:conditional_hypergeometric} and the conditional
@@
 
-The $T= \infty$ approximation can break the method for specific
+The $T= \infty$ approximation can break our model for specific
 distributions of sample frequencies, if too many moments are used. For
@@
 exceptions or completely unrelated to the others surrounding the recording
-probe, our recording would serve little purpose. Probabilistic methods that
+probe, our recording would serve little purpose. Probabilistic models that
 allow us to make direct inferences about the larger population are
@@
 
-We have presented a method that allows us to make direct inferences about a
+We have presented a model that allows us to make direct inferences about a
 larger population from very small samples. Its inferences regard only the
 frequencies of the total activity of the larger population, but such
-limited scope makes the method computationally very cheap.
-
-Yet, and most important, the method leads to inferences that are not at all
+limited inferential scope makes the model computationally very cheap.
+
+Yet, and most important, this model leads to inferences that are not at all
 trivial and not easily discernible by looking at the sample. We showed this
-by applying the method to two real data sets.
+by applying the model to two real data sets.
 
@@
 about possible observations, based on specific assumptions; not the
-response of an oracle. The method gives us a \emph{forecast} based on those
-assumptions, and therefore has two main uses, just like any other
-probabilistic method. If our assumptions are only hypotheses under testing,
-then the subsequent confirmation or confutation of the forecast will
-increase or decrease our confidence in those hypotheses. If we strongly
-trust our assumptions, on the other hand, then we rely on the forecast and
-even use it in place of actual observations, especially if the latter are
-difficult to make. (In fact, we typically shift from the first to the
-second use.) In the preceding sections we have given examples of hypotheses
-that could be explored with the method. The method itself rests on
-assumptions, discussed in \sect~\ref{sec:assumptions}, that can either be
-under test or relied upon, depending on the case.
-
-Some approximations implicit in the method make its numerical results
-semi-quantitative with respect to an analysis based on the full probability
-calculus. The results should not be off by more than a few percentages for
-the normalized total activity and for the frequency density. This
-imprecision, however, is compensated by the extreme computational cheapness
-of the method, far less costly than a full probabilistic analysis. The
-method is therefore very convenient at least for essaying hypotheses. We
-hope to present an exact quantitative comparison with the full
-probabilistic approach in a future publication.
+response of an oracle. The model here presented gives us a \emph{forecast}
+based on specific assumptions discussed in \sect~\ref{sec:assumptions}, and
+therefore has two main uses, just like any other probabilistic model. If
+our assumptions are only hypotheses under testing, then the subsequent
+confirmation or confutation of the forecast will increase or decrease our
+confidence in those hypotheses. If we strongly trust our assumptions, on
+the other hand, then we rely on the forecast and even use it in place of
+actual observations, especially if the latter are difficult to make. (In
+fact, we typically shift from the first to the second use.) In the
+preceding sections we have given examples of additional hypotheses that
+could be explored with the present model. The assumptions of the model
+itself (\sect~\ref{sec:assumptions}) can either be under test or relied
+upon, depending on the case.
+
+Some approximations implicit in the model make its numerical results
+semi-quantitative with respect to an exact analysis. The results should not
+be off by more than a few percentages for the normalized total activity and
+for the frequency density. This imprecision, however, is compensated by the
+extreme computational cheapness of the model, far less costly than an exact
+analysis. The model is therefore very convenient at least for essaying
+hypotheses. We hope to present an exact quantitative comparison with the
+full probabilistic approach in a future publication.
 

### 2019-12-06 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-06T17:01:42+0100
+%% Last-Updated: 2019-12-06T17:16:23+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 neurons in the larger population. The magnitude of this discrepancy can be
-a signature of the (at least temporary) presence of a \enquote{neuronal
-  assembly} \citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss
-this matter in \sect~\ref{sec:assumptions}.
+a signature of the (at least temporary) presence of neuronal
+\enquote{assemblies} or \enquote{clusters}
+\citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss this matter
+in \sect~\ref{sec:assumptions}.
 
@@
 
-\section{Assumptions and approximations}
+\section{The assumptions and approximations that define this model}
 \label{sec:assumptions}

### 2019-12-07 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-06T17:16:23+0100
+%% Last-Updated: 2019-12-07T14:46:25+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 realistic ways, to make the analysis mathematically tractable. This is why
-such initial probability distributions are called \enquote{models}.
+such initial probability distributions are called \enquote{models}; and
+that is how we intend this word in the present work.
 
@@
 at least for preliminary investigations, to help us to decide which
-hypotheses to discard and which to consider for more complex and costly
-models, or to suggest new hypotheses.
+hypotheses to consider or discard for more complex and costly models, or to
+suggest new hypotheses.
 
@@
 In the present work we propose such an intermediate probabilistic model. It
-answers this question: How much was the \emph{total} activity of a large
+answers this question: how much was the \emph{total} activity of a large
 neuronal population, given the observation of the activity of a very small
@@
 distributions for the larger population are distinctly different from those
-for the sample. For the first data set the frequency distribution for the
-larger population has two very distinct modes, both at low
-activities %($0.7\%$ and $2.6\%$ of simultaneously active neurons
+for the sample. For example, for the first data set the frequency
+distribution for the larger population has two very distinct modes, both at
+low activities %($0.7\%$ and $2.6\%$ of simultaneously active neurons
 % within $3\;\mathrm{ms}$,
 (see \fig~\ref{fig:plots_5mom}), whereas the frequency distribution for the
-sample is monotonically decreasing with its maximum at zero activity. The
-frequency distribution for the second data set presents one prominent
-shoulder in its low-activity mode. These results show that the proposed
-model can lead to the formulation or preliminary assessment of interesting
-hypotheses, as we will illustrate with toy examples in
-\sects~\ref{sec:application} and~\ref{sec:marginalization}. Note that it is
-not possible to guess these features of the larger population by applying
-the maximum-entropy method \emph{at the level of the sample alone}.
+sample is monotonically decreasing with its maximum at zero activity. These
+results show that the proposed model can be used for the formulation or
+preliminary assessment of interesting hypotheses. We illustrate this use
+with toy examples in \sects~\ref{sec:application}
+and~\ref{sec:marginalization}. Note that it is not possible to guess these
+features of the larger population by applying the maximum-entropy method
+\emph{at the level of the sample alone}.
 
@@
 
-How large is the full population addressed by the model proposed here? Its
-size must have some limit and can't obviously include the full brain. The
-size is determined by the validity of the formulae from sampling theory and
-discussed in \sect~\ref{sec:N}.
+How large is the \enquote{larger population} considered by the model
+proposed here? It obviously cannot include the full brain. The size of the
+larger population is determined by the validity of the formulae from
+sampling theory, and can range from a set of neurons around the recording
+probe to part of a brain area, depending on some assumptions about the
+recording procedure and the brain region under consideration. This matter
+is discussed in detail in \sect~\ref{sec:N}.
 
 In \sect~\ref{sec:assumptions} we discuss in detail the assumptions and
-approximations on which the model is based, from the point of view of the
+approximations which define our model, from the point of view of the
 probability calculus.
@@
 \citey{iso1993,iso2006,iso2006b} standards and Jaynes
-\citey{jaynes1994_r2003} for \dobs. We use \enquote{\dob},
+\citey{jaynes1994_r2003} for probability. We use \enquote{\dob},
 \enquote{belief}, and \enquote{probability} interchangeably.
@@
 change across sessions? How much total activity should we expect at any
-time bin during a recording? The model presented here gives a
-probability distribution that approximately answers all these questions.
+time bin during a recording? The model presented here gives a probability
+distribution over the answers to these questions.
 
@@
 
-The idea behind our approach is easily summarized:
+The idea behind our approach can be easily summarized in terms of the
+maximum-entropy method:
 \begin{enumerate}[label=(\alph*)]
@@
 \end{enumerate}
-We now discuss the ideas behind these steps more in detail, but leave their
-precise mathematical implementation and a more detailed list of references
-to appendix~\ref{sec:derivation_maxent_fullnet}.
+We now discuss these steps more in detail, but leave their precise
+mathematical implementation and a more detailed list of references to
+appendix~\ref{sec:derivation_maxent_fullnet}.
 
@@
 distribution $\pxxx(\yA_{t})$ for the larger-population activity and its
-marginal $\pxx(\ya_{t})$ for the sample activity at that bin.
+marginal $\pxx(\ya_{t})$ for the sample activity at that bin. These two
+distributions are related by
+\begin{equation}
+  \label{eq:conditional_hypergeometric_t}
+  \pxx(\ya_{t})= \sum_{\yA_{t}=0}^{N} \yG_{\ya_{t}\,\yA_{t}} \;\pxxx(\yA_{t})
+  \qquad \ya_{t} \in \set{0,\dotsc,n},
+\end{equation}
+where $\yG_{\ya\yA}$ is the hypergeometric distribution defined in
+\eqn~\eqref{eq:conditional_hypergeometric}, characteristic of
+\enquote{drawing without replacement}
+\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
 
@@
   \label{eq:def_factorial_moment_a}
-  \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
-  \defd
+  % \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
+  % \defd
   \sum_{\ya=0}^{n} %m!\,
-  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya),
-  \qquad 1\le m \le n,
+  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya)
+  \qquad m \in \set{1,\dotsc,n},
 \end{equation}
@@
 But the normalized factorial moments have a mathematically convenient
-property for our analysis: \emph{the first $n$ normalized factorial moments
-  for the sample and for the larger population are numerically identical}:
+property: \emph{the first $n$ normalized factorial moments for the sample
+  and for the larger population are numerically identical}:
 \begin{equation}
   \label{eq:equal_m_factorialmoments}
-  \begin{split}
-  \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
-  &=
-  \E\Biggl[\binom{\yA}{m} \Biggr]\bigg/\binom{N}{m}
-  \qquad\text{or}\\
+  % \begin{split}
+  % \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
+  % &=
+  % \E\Biggl[\binom{\yA}{m} \Biggr]\bigg/\binom{N}{m}
+  % \qquad\text{or}\\
     \sum_{\ya=0}^{n} %m!\,
   \binom{\ya}{m}\bigg/  \binom{n}{m}\; \pxx(\ya)
-&=    \sum_{\yA=0}^{N} %m!\,
+=    \sum_{\yA=0}^{N} %m!\,
   \binom{\yA}{m}\bigg/  \binom{N}{m}\; \pxxx(\yA),\qquad
@@
   % \E\condt[\big]{\tbinom{\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
-  1\le m \le n.
-\end{split}
+  m \in \set{1,\dotsc,n}.
+%\end{split}
 \end{equation}
@@
   \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured averages}}
-  \mathrel{\;=\;}
+  \mathrel{\;\;=\;\;}
    \underbracket{\sum_{\yA}
-     \binom{\yA}{m}\bigg/  \binom{N}{m} \; \pxxx(\yA)}_{\text{distribution moments}},
-   \quad m\in M.
-\end{equation}
+     \binom{\yA}{m}\bigg/  \binom{N}{m} \; \pxxx(\yA)}_{\text{distribution moments}}
+   \qquad m\in M.
+\end{equation}
+\smallskip
 
@@
 \end{equation}
-where the parameters $\yll$ are determined by the
+where the $m'$ parameters $\yll \defd (\yl_{m})$ are determined by the
 constraints~\eqref{eq:final_constraint}. This formula is further discussed
@@
 distribution~\eqref{eq:app_maxent_pop_first} may not have exact solutions
-if $N$ is strictly larger than $n$; and the possible discrepancy typically
-increases with the number of constraints. This discrepancy comes from the
-approximate character of two assumptions behind the maximum-entropy method
-and of one assumption specific to our application: that the order of the
-time bins is irrelevant, that measured averages equal expected values (or
-equivalently, the number of time bins $T$ is infinite), and that an
-activity level $\yA$ can equally likely be generated by any set of $\yA$
-neurons in the larger population. The magnitude of this discrepancy can be
-a signature of the (at least temporary) presence of neuronal
-\enquote{assemblies} or \enquote{clusters}
+if $N$ is strictly larger than $n$. The possible discrepancy between its
+two sides typically increases with the number of constraints. This
+discrepancy comes from an approximation implicit in the maximum-entropy
+method combined with one assumption specific to our application: that
+measured averages equal expected values (or equivalently, the number of
+time bins $T$ is infinite), and that an activity level $\yA$ can equally
+likely be generated by any set of $\yA$ neurons in the larger population.
+The magnitude of this discrepancy can be a signature of the presence of
+neuronal \enquote{assemblies} or \enquote{clusters}
 \citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss this matter
@@
 
-%, but our main goal is to simply introduce
-% the general method without entering into biological questions of this
-% kind.
-
 \bigskip
 
-Before applying the formula above to two concrete data sets we want to add
-two remarks about the maximum-entropy method that are seldom made in the
-neuroscientific literature. They are important for the interpretation of
-the results and are further discussed in \sect~\ref{sec:assumptions}.
-First, the maximum-entropy
-method % (based on the \emph{Shannon} relative
-% entropy, as opposed to other entropy functions such as Burg's
-% \citeyear{burg1975})
-rests on some implicit assumptions about the probabilities for the long-run
-frequency distribution of activities, besides the assumptions just
-mentioned at the end of the previous paragraph. So the often-heard
-statement that it gives \enquote{the maximally unbiased (or non-committal)
-  distribution} must be taken with a grain of salt.
-\citep{jaynes1986d_r1996,portamana2009,portamana2017}. Second, a
-maximum-entropy distribution like $\px(\yA \| M, N)$ is the zeroth-order
-approximation (in the sense of Laplace's approximation
+One important question remains about the distribution $\px$ of
+\eqn~\eqref{eq:app_maxent_pop_first}: \emph{what} is it a distribution of?
+-- Of probability? Of relative frequency? The answer is related to an
+important fact about the maximum-entropy method, and also to our
+alternative point of view of the present model.
+
+% Before applying the formula above to two concrete data sets we want to add
+% two remarks about the maximum-entropy method that are seldom made in the
+% neuroscientific literature. They are important for the interpretation of
+% the results and are further discussed in \sect~\ref{sec:assumptions}.
+% First, the maximum-entropy
+% method % (based on the \emph{Shannon} relative
+% % entropy, as opposed to other entropy functions such as Burg's
+% % \citeyear{burg1975})
+% rests on some implicit assumptions about the probabilities for the long-run
+% frequency distribution of activities, besides the assumptions just
+% mentioned at the end of the previous paragraph. So the often-heard
+% statement that it gives \enquote{the maximally unbiased (or non-committal)
+%   distribution} must be taken with a grain of salt.
+% \citep{jaynes1986d_r1996,portamana2009,portamana2017}. Second,
+A maximum-entropy distribution like $\px(\yA \| M, N)$ is equivalent to the
+zeroth-order approximation (in the sense of Laplace's approximation
 \citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}) of
@@
 \begin{enumerate}[label=(\roman*)]
-\item the most probable \emph{frequency} distribution for the total
-  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
+\item\label{item:prob_freq} the most probable \emph{frequency} distribution
+  for the total activity across the \emph{recorded} bins, %$\yFF(\yA)$,
 \item the \emph{belief} distribution for the value of the total
@@
 \end{enumerate}
-In the present case the validity of the maximum-entropy approximation
-decreases as the ratio $nN/T$ increases (see \sect~\ref{sec:assumptions}).
+Their mathematical relationship is presented more explicitly in
+appendix~\ref{sec:inference_prob}. The
+distribution~\eqref{eq:app_maxent_pop_first} can therefore be interpreted
+in each of the four ways above, in this approximate sense. In the present
+case the validity of the approximation decreases as the ratio $nN/T$
+increases. The most robust interpretation is~\ref{item:prob_freq}; and this
+is how we prefer to interpret the distribution $\px$: as \emph{the most
+  probable relative-frequency distribution} of the activity $\yA$. See
+\sect~\ref{sec:assumptions} for further discussion about assumptions and
+approximations.
 
@@
 
-The crucial point is our belief distribution for the activity of the
-sample conditional on the activity of the larger population, given by
-formula~\eqref{eq:conditional_hypergeometric} in
-appendix~\ref{sec:derivation_maxent_fullnet}, reproduced here:
+The crucial point is our belief distribution for the activity of the sample
+\emph{conditional} on the activity of the larger population. It is given by
+the hypergeometric distribution~\eqref{eq:conditional_hypergeometric},
+reprinted here:
 \begin{equation}
@@
     \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
-\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
-\end{equation}
-leading to the equality of factorial
-moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests.
-This conditional probability, characteristic of \enquote{drawing without
+\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}.
+\end{equation}
+This distribution leads to the equality of factorial
+moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests. This
+conditional probability, characteristic of \enquote{drawing without
   replacement}
@@
 The two infinity approximations are at the limit of their validity in our
-specific application, so the distributions discussed in our example are
-best interpreted as \emph{most probable frequency distributions}. The space
-of possible frequency distributions for the larger population has $N$
-dimensions, however, and probability distributions in high dimensions have
-counter-intuitive properties. For example, the mode or mean can have
-\emph{atypical} features when compared with the majority of other probable
-points of the probability space. The question then is \emph{which features
-  of the maximum-entropy frequency distribution are typical of the majority
-  of plausible frequency distributions?} Preliminary studies using a full
-probability calculation (to be published separately) show that several
-semi-quantitative features of the maximum-entropy distribution are indeed
-typical. For the first data set, for example, most frequency distributions
-have two high-frequency regimes of activity, as in
-\fig~\ref{fig:plots_5mom_a}, with roughly the same heights and almost the
-same locations, within a few percentages of normalized total activity. Each
-of high-frequency regimes, though, may consist of two or three very close
-modes rather than one. The same preliminary studies also indicate that
-these typical features remain with prior choices in the second assumption
-above, for example using a Dirichlet prior instead of an entropic one.
+examples. For this reason we interpret the distributions as \emph{most
+  probable frequency distributions}, as mentioned at the end of
+\sect~\ref{sec:model}. The space of possible frequency distributions for
+the larger population has $N$ dimensions, however, and probability
+distributions in high dimensions have counter-intuitive properties. For
+example, the mode or mean can have \emph{atypical} features when compared
+with the majority of other probable points of the probability space. The
+question then is \emph{which features of the maximum-entropy frequency
+  distribution are typical of the majority of plausible frequency
+  distributions?} Preliminary studies using a full probability calculation
+(to be published separately) show that several semi-quantitative features
+of the maximum-entropy distribution are indeed typical. For the first data
+set, for example, most frequency distributions have two high-frequency
+regimes of activity, as in \fig~\ref{fig:plots_5mom_a}, with roughly the
+same heights and almost the same locations, within a few percentages of
+normalized total activity. Each of high-frequency regimes, though, may
+consist of two or three very close modes rather than one. The same
+preliminary studies also indicate that these typical features remain with
+prior choices in the second assumption above, for example using a Dirichlet
+prior instead of an entropic one.
 

### 2019-12-07 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-07T14:46:25+0100
+%% Last-Updated: 2019-12-07T15:36:17+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Here is a summary derivation of the maximum-entropy distribution for the
-larger population constrained by a set of factorial moments. For further details
-see Porta Mana \etal\ \citey{portamanaetal2015}.
+larger population constrained by a set of factorial moments.%  For further details
+% see Porta Mana \etal\ \citey{portamanaetal2015}.
 
@@
 
-First of all the sampling relation. We have a set of $N$ units, $\yA$ of
-which have some specific property, and we sample in an unknown way $n$ of
-the $N$ units. The probability that $\ya$ of the $n$ sampled units have that
-property is then given by the hypergeometric distribution
+First of all the sampling relation. We have a population of $N$ units --
+neurons -- $\yA$ of which have some specific property -- being active in a
+specific time bin -- and we sample in an unknown way $n$ of the $N$ units.
+The probability that $\ya$ of the $n$ sampled neurons are active is then
+given by the hypergeometric distribution
 \begin{equation}
@@
 \citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}[\sects~2.1,
-3.2]{jeffreys1939_r1983}.
-In the following leave $n$, $N$ implicit in the conditional. If we are
-uncertain about the number $\yA$, with belief $P(\yA)$, then by the theorem
-of total probability our belief about $\ya$ is
+3.2]{jeffreys1939_r1983}. In the following we leave $n$, $N$ implicit in
+the conditional. If we are uncertain about the number $\yA$, with belief
+$P(\yA)$, then by the theorem of total probability our belief about $\ya$
+is
 \begin{equation}
@@
 \end{equation}
-In our case the units are neurons, the set is the larger population, and the
-property is their being active in a specific time bin.
 
@@
   \px(\yA)  = \frac{1}{Z(\yll)}\;
-  g(\yA)\;
+  r_{\yA}\;
   \exp\Biggl[\sum_{m}\yl_{m}\;
@@
   \\
-  Z(\yll)\defd \sum_{\yA}g(\yA)\;
+  Z(\yll)\defd \sum_{\yA} r_{\yA}\;
   \exp\Biggl[\sum_{m}\yl_{m}\;
@@
 \end{equation}
-where $g(\yA)$ is the reference distribution and $\yll\defd (\yl_{m})$ are
+where $(r_{\yA})$ is the reference distribution and $\yll\defd (\yl_{m})$ are
 the Lagrange multipliers, satisfying the implicit
@@
   \tfrac{1}{Z(\yll)}\;
-  g(\yA)\;
+  r_{\yA}\;
   \exp\biggl[\sum_{m}\yl_{m}\;
@@
 
-The reference distribution $g(\yA)$ represents our pre-data beliefs about
-the activity levels $\yA$. We know that the majority of neurons in a brain
-area are rarely simultaneously active within a window of some milliseconds,
-so we could choose a distribution with slightly higher weights on low
-values of $\yA$. On the other hand, considering the number of ways in which
-$\yA$ out of $N$ neurons can be simultaneously active would suggest the
-multiplicity distribution proportional to $\binom{N}{\yA}$. It turns out
-that our results of \sects~\ref{sec:application}--\ref{sec:marginalization}
-are actually quite insensitive to the choice between these two possible
-reference distributions, or even a uniform reference distribution.
+The reference distribution $(r_{\yA})$ represents our pre-data beliefs
+about the activity levels $\yA$. We know that the majority of neurons in a
+brain area are rarely simultaneously active within a window of some
+milliseconds, so we could choose a distribution with slightly higher
+weights on low values of $\yA$. On the other hand, considering the number
+of ways in which $\yA$ out of $N$ neurons can be simultaneously active
+would suggest the multiplicity distribution proportional to
+$\binom{N}{\yA}$. It turns out that our results of
+\sects~\ref{sec:application}--\ref{sec:marginalization} are insensitive to
+the choice between these two possible reference distributions, or even a
+uniform reference distribution.
+
+For a derivation from a Bayesian perspective see
+appendix~\ref{sec:inference_prob}.
 
@@
   \label{eq:ratio_prob_suffic}
-  \pf(M'' \| \yf)/
-  \pf(M' \| \yf).
-\end{equation}
-Now, the probability of observing activity $\ya$ in the sample at any time
-bin is the sample marginal of the maximum-entropy distribution for the larger
+  \frac{p(M'' \| \yf)}{ p(M' \| \yf)}.
+\end{equation}
+The probability of observing activity $\ya$ in the sample at any time bin
+is the sample marginal of the maximum-entropy distribution for the larger
 population, owing to the excheangeability assumption implicit in the
@@
 probability ratio~\eqref{eq:ratio_prob_suffic}. The exponential of the
-difference~\eqref{eq:measure_suffic} tells us how much more probable is the
-set $M''$ to be sufficient than the set $M'$.
+difference~\eqref{eq:measure_suffic} tells us how much more probable the
+hypothesis that the set $M''$ is sufficient than the hypothesis that the
+set $M'$ is.
 

### 2019-12-07 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-07T15:36:17+0100
+%% Last-Updated: 2019-12-07T17:52:23+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   for the total activity across the \emph{recorded} bins, %$\yFF(\yA)$,
-\item the \emph{belief} distribution for the value of the total
-  activity at any time bin among the \emph{recorded} ones,
+\item\label{item:prob_binin} the \emph{belief} distribution for the value
+  of the total activity at any time bin among the \emph{recorded} ones,
   %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
-\item the most probable \emph{frequency} distribution for the total
-  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
-\item the \emph{belief} distribution for the value of the total
-  activity at a \emph{new} time bin.
+\item\label{item:prob_longrun} the most probable \emph{frequency}
+  distribution for the total activity in a very long run of \emph{new} time
+  bins, %$\ynuu(\yA)$,
+\item\label{item:prob_binout} the \emph{belief} distribution for the value
+  of the total activity at a \emph{new} time bin.
   % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
@@
 
+From the formulae above we can also find the four distributions mentioned
+at the end of \sect~\ref{sec:model}. The frequency
+distribution~\ref{item:prob_freq} is the mode of $\pf(\yF \| \yf,\yI)$. The
+belief distribution~\ref{item:prob_binin} is
+\begin{equation}
+  \label{eq:prob_binin}
+  \pf(\yA_{t} \| \yf, \yI) =
+  \sum_{\yF} \yFF_{\yA_{t}}\; \pf(\yF \| \yf, \yI),
+  \qquad t \in \set{1,\dotsc,T}.
+\end{equation}
+The frequency distribution~\ref{item:prob_longrun} is the mode of $\pf(\ynu
+\| \yf,\yI)$. Finally, the belief distribution~\ref{item:prob_binout} is
+\begin{equation}
+  \label{eq:prob_binout}
+  \pf(\yA_{t} \| \yf, \yI) =
+  \int\!\!\!\di\ynu\;\nu_{\yA_{t}}\;\pf(\ynu \| \yf, \yI),
+  \qquad t \notin \set{1,\dotsc,T}.
+\end{equation}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-12-07 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-07T17:52:23+0100
+%% Last-Updated: 2019-12-07T18:10:15+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\pdftitle}{\propertitle}
-\newcommand*{\headtitle}{Maximum-entropy from sample data}
+\newcommand*{\headtitle}{Inference about large neuronal populations}
 \newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
@@
 
-\section{Introduction:\\a model for questions about large neuronal
+\section{Introduction:\\a model for collective inferences about large neuronal
   populations}
@@
 
-\section{Model: maximum-entropy and sampling}
+\section{Model: maximum-entropy and sampling,\\ and a Bayesian perspective}
 \label{sec:model}
@@
 
-\section{The larger-population size $N$}
+\section{The size $N$ of the larger population}
 \label{sec:N}
@@
 
-\section{The assumptions and approximations that define this model}
+\section{Assumptions and approximations that define the model}
 \label{sec:assumptions}

### 2019-12-07 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-07T18:10:15+0100
+%% Last-Updated: 2019-12-07T21:28:11+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\pdftitle}{\propertitle}
-\newcommand*{\headtitle}{Inference about large neuronal populations}
+\newcommand*{\headtitle}{Inferences about large neuronal populations}
 \newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
@@
 -- Of probability? Of relative frequency? The answer is related to an
-important fact about the maximum-entropy method, and also to our
-alternative point of view of the present model.
+important relation between maximum-entropy distributions and the
+probability calculus, and also to our alternative point of view of the
+present model.
 
@@
 A maximum-entropy distribution like $\px(\yA \| M, N)$ is equivalent to the
-zeroth-order approximation (in the sense of Laplace's approximation
-\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}) of
-four distinct distributions for the larger population, which differ
-numerically from one another in higher-order approximations:
+zeroth-order (Laplace
+\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000})
+approximation of four distinct distributions for the larger population,
+which differ numerically from one another in higher-order approximations:
 \begin{enumerate}[label=(\roman*)]

### 2019-12-08 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-07T21:28:11+0100
+%% Last-Updated: 2019-12-08T20:13:23+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 about individual neuronal details. Such intermediate models would be useful
-at least for preliminary investigations, to help us to decide which
+at least for preliminary investigations, to help us  decide which
 hypotheses to consider or discard for more complex and costly models, or to
@@
 However, our model can also be viewed as an approximation of a more
-detailed model within a fully Bayesian approach; we shall constantly keep
-this point of view in mind in our discussion.
+detailed model within Bayesian theory; we shall constantly keep this point
+of view in mind in our discussion.
 
@@
 case the validity of the approximation decreases as the ratio $nN/T$
-increases. The most robust interpretation is~\ref{item:prob_freq}; and this
+increases. The most robust interpretation is~\ref{item:prob_freq}, and that
 is how we prefer to interpret the distribution $\px$: as \emph{the most
@@
 this section we discuss how our proposed application bears on this kind of
-quantification and compare it with the traditional application at the
-sample level. But such a comparison involves some methodological
-caveats which we wish to discuss first.% For concreteness we use an
+quantification and compares with the traditional application at the sample
+level. But such a comparison involves some methodological caveats which we
+wish to discuss first.% For concreteness we use an
 % example with specific statistics: first and second moments (equivalent to
@@
 % maxent_solutions_different_moments_prior2_6moments.nb
-These informational differences shows clearly in the shapes of the
+These informational differences show clearly in the shapes of the
 distribution themselves, plotted in 
@@
 It is also possible to use the measure~\eqref{eq:measure_suffic} and its
-probabilistic meaning~\eqref{eq:bayes_factor} to compare the probability of
+probabilistic interpretation~\eqref{eq:bayes_factor} to compare the probability of
 the data $\yf$ conditional on the hypothesis $(M_{4},N)$ of four-moment
@@
   \label{eq:diff_suff_n}
-  \ydi[(M_{4}, N), (M_{4}, n)] = 39\;\nat = 17\;\hart,
-\end{equation}
-that is, the data are 17 orders of magnitude more probable under the first
-hypothesis than under the second. This result is also shown in
+  \ydi[(M_{4}, N), (M_{4}, n)] = 39\;\nat = 17\;\hart.
+\end{equation}
+% that is, the data are 17 orders of magnitude more probable under the first
+% hypothesis than under the second. 
+This result is also shown in
 \fig~\ref{fig:comparisons_marginals}, where we see that the four-moment
@@
 moments under study \citep{portamana2017}. This assumption is crucial for
-deriving the maximum-entropy based on with the Shannon entropy. A different
+deriving the maximum-entropy based on the Shannon entropy. A different
 density leads to alternative maximum-entropy methods, for example based on
@@
 favourably used to get an idea of what a fully probabilistic calculation
-could yield. -- With a far lower computational cost: for example, for
+could yield, but with a far lower computational cost. For example, for
 $N=1\,000$ the maximum-entropy distribution can be calculated in ten

### 2019-12-09 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-08T20:13:23+0100
+%% Last-Updated: 2019-12-09T10:56:16+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The model can be viewed as a straightforward combination of the
-maximum-entropy method and basic sampling relations from the probability
-calculus, discussed in \sect~\ref{sec:model}. The maximum-entropy or
+maximum-entropy method and basic sampling relations from the probability-calculus, discussed in \sect~\ref{sec:model}. The maximum-entropy or
 minimum-relative-entropy method \citep{jaynes1957}[much clearer
@@
 approximations which define our model, from the point of view of the
-probability calculus.
+probability-calculus.
 
@@
 % Later on we'll show that our \dobs\ can be quantified by
-% exclusively using the probability calculus. This derivation will provide a
+% exclusively using the probability-calculus. This derivation will provide a
 % more accurate quantification, revealing that the maximum-entropy answer is
@@
 
-Step~\ref{item:sample_step} is just an application of the probability
-calculus, which gives an exact linear relation between the first $m$
+Step~\ref{item:sample_step} is just an application of the probability-calculus, which gives an exact linear relation between the first $m$
 moments for the larger population and the first $m$ for the sample
@@
 important relation between maximum-entropy distributions and the
-probability calculus, and also to our alternative point of view of the
+probability-calculus, and also to our alternative point of view of the
 present model.
@@
 assessment in turn rests on analogous assumptions at a higher level, and so
-on. The probability calculus, just like the logical calculus, cannot yield
+on. The probability-calculus, just like the truth-calculus, cannot yield
 conclusions without premisses \citep[p.~182]{johnson1924}.

### 2019-12-14 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-09T10:56:16+0100
+%% Last-Updated: 2019-12-14T18:18:53+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \raggedbottom
+
+\paragraphfootnotes
+% \threecolumnfootnotes
+\setlength{\footmarkwidth}{0em}
+%\setlength{\footmarksep}{0em}
+\footmarkstyle{\textsuperscript{\color{myred}\bfseries#1}~}
+%\footmarkstyle{\textsuperscript{[#1]}~}
+
 \selectlanguage{british}\frenchspacing
@@
 The best experimental measurements of instantaneous neuronal activity use
-remarkable technologies, but can still only record a very small sample of
-neurons -- hundreds at most -- compared to the numbers that constitute a
-functionally distinguished brain region. Many probabilistic models focus on
-such samples only: they neglect that the recorded neurons are a sample from
-a larger population. Such isolating assumptions sometimes escape attention,
-being subtly hidden in the mathematics. Some probabilistic models try to
-take unrecorded neurons into account, but they do so by describing each
-neuron individually, thus becoming very complex
-\citep{huang2015,battistinetal2017}. It would be useful to explore models
+remarkable technologies \citep{hongetal2019}, but can still only record a
+very small sample of neurons -- hundreds at most -- compared to the numbers
+that constitute a functionally distinguished brain region. Many
+probabilistic models focus on such samples only: they neglect that the
+recorded neurons are a sample from a larger population. Some probabilistic
+models try to take unrecorded neurons into account, but they do so by
+describing each neuron individually, thus becoming very complex
+\citep{dunnetal2013,huang2015,battistinetal2017,pillowetal2008}. It would be useful to explore models
 that operate in between, addressing the activity of the larger brain area
@@
 about individual neuronal details. Such intermediate models would be useful
-at least for preliminary investigations, to help us  decide which
-hypotheses to consider or discard for more complex and costly models, or to
-suggest new hypotheses.
+at least for preliminary investigations, to help us decide which hypotheses
+to consider or discard for more complex and costly models, or to suggest
+new hypotheses.
 
@@
 The model can be viewed as a straightforward combination of the
-maximum-entropy method and basic sampling relations from the probability-calculus, discussed in \sect~\ref{sec:model}. The maximum-entropy or
-minimum-relative-entropy method \citep{jaynes1957}[much clearer
-in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b_r1996,grandy1980}
-to build initial probability distributions has been used for different
-kinds of investigations about the neuronal activity of various brain areas
-and about other phenomena of importance to the neurosciences, for example
-gene and protein interaction \citep[for
-example][]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,mackeetal2009b,tkaciketal2009,roudietal2009c,barreiroetal2010,gerwinnetal2010,mackeetal2011b,ganmoretal2011,cohenetal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}.
-However, our model can also be viewed as an approximation of a more
-detailed model within Bayesian theory; we shall constantly keep this point
-of view in mind in our discussion.
+maximum-entropy method and basic sampling relations from the
+probability-calculus, discussed in \sect~\ref{sec:model}. The
+maximum-entropy or minimum-relative-entropy method
+\citep{jaynes1957}[\chap~11]{jaynes1994_r2003}[for the minimum-relative
+entropy method see:][\sect~4.b]{jaynes1963}[for a more rigorous
+derivation]{hobsonetal1973}[for applications]{sivia1996_r2006}
+%[see also][for applications to irreversible statistical mechanics]{grandy1980}
+to build initial probability distributions has been used in the
+neurosciences for different kinds of investigations about the neuronal
+activity of various brain areas and about other phenomena of importance,
+for example gene and protein interaction
+\citep[]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,roudietal2009c,barreiroetal2010,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}[for
+further references see][]{lathametal2013}. However, our model can also be
+viewed as an approximation of a more detailed model within Bayesian theory;
+we shall constantly keep this point of view in mind in our discussion.
 
@@
 
-Let's introduce some context and mathematical notation.
+Some context and mathematical notation first.
 
@@
 during bin $t$; if $\ya_{t}=n$, all spike at some point during bin $t$, and
-so on. For brevity, let's say \enquote{at $t$} for \enquote{during time bin
-  $t$}. If we divide the total activity by the population size we have the
-\emph{normalized total activity} or population-averaged activity $\ya/n$,
-ranging from $0$ to $1$ in $1/n$ steps. From the activities $\set{\ya_{t}}$
+so on. For brevity we say \enquote{at $t$} for \enquote{during time bin
+  $t$}. % If we divide the total activity by the population size we have the
+% \emph{normalized total activity} or population-averaged activity $\ya/n$,
+% ranging from $0$ to $1$ in $1/n$ steps.
+From the activities $\set{\ya_{t}}$
 we can count how often the activity levels $\ya=0$, $\ya=1$, and so on
@@
 area of which we believe our sample to be \enquote{representative}
-\citep[with the warnings that accompany that
-term:][]{kruskaletal1979b,kruskaletal1979c}.
+\citep[we intend this term in the Meaning~2 of][]{kruskaletal1979b}[or
+Meanings~8 and~9 of][]{kruskaletal1979c}.
 
@@
 frequencies of the various activity levels during the recording were
-$(\yFF_{\yA}) \defs \yF$. We don't know the values $\yA_{t}$ at each $t$,
-or the frequency distribution $\yF$. We only know for certain that
+$(\yFF_{\yA}) \defs \yF$. The values $\yA_{t}$ at each $t$ and
+the frequency distribution $\yF$ are unknown to us. We only know for certain that
 $\yA_{t} \in \set{0, 1, \dotsc, N}$, that $\yA_{t} \ge \ya_{t}$, and that
@@
 moments for the larger population and the first $m$ for the sample
-\citep[\eqns~(16)]{portamanaetal2015}. The ones determine the others and
+\citep[Appendix]{portamanaetal2015}. The ones determine the others and
 vice versa at every time bin. This relation holds for any belief
@@
 \enquote{drawing without replacement}
-\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
+\citep[\chap~3]{jaynes1994_r2003}.%[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
 
@@
 property: \emph{the first $n$ normalized factorial moments for the sample
-  and for the larger population are numerically identical}:
+  and for the larger population are numerically identical}:\citep[Appendix]{portamanaetal2015}
 \begin{equation}
@@
 \citep{hobsonetal1973,csiszar1985}[\sect~5.2.2]{sivia1996_r2006} with
-respect to a uniform reference distribution. We'll still call it
+respect to a uniform reference distribution. We shall still call it
 \enquote{maximum-entropy} for brevity. It amounts to two prescriptions:
@@
 such as fixed expectations -- and among them select the one having minimum
-relative entropy with respect to a reference distribution; second, judge
-those expectations to be equal to some measured averages, typically time
-averages.
+relative entropy with respect to a reference distribution; second, judge --
+and therefore set -- those expectations to be equal to some measured
+averages, typically time averages. (A judgement is required because
+probability expectations and empirical averages are numerically
+different in general; a misjudgement of their approximate equality can lead to
+\enquote{silly answers}, to quote Mackay \cite*[p.~308]{mackay1995_r2003}.)
 
@@
 (\textcolor{myred}{red circles}) has two peaks instead, roughly at the same
-normalized activities as the peaks of the larger-population distribution
+population-averaged activities as the peaks of the larger-population distribution
 (\textcolor{myblue}{blue curve}) but closer in height. It would thus seem
@@
 same heights and almost the same locations, within a few percentages of
-normalized total activity. Each of high-frequency regimes, though, may
+population-averaged activity. Each of high-frequency regimes, though, may
 consist of two or three very close modes rather than one. The same
@@
 semi-quantitative with respect to an exact analysis. The results should not
-be off by more than a few percentages for the normalized total activity and
+be off by more than a few percentages for the population-averaged activity and
 for the frequency density. This imprecision, however, is compensated by the

### 2019-12-14 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-14T18:18:53+0100
+%% Last-Updated: 2019-12-14T18:54:42+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 describing each neuron individually, thus becoming very complex
-\citep{dunnetal2013,huang2015,battistinetal2017,pillowetal2008}. It would be useful to explore models
-that operate in between, addressing the activity of the larger brain area
-whence the sample comes, but \emph{collectively}, that is, without asking
-about individual neuronal details. Such intermediate models would be useful
-at least for preliminary investigations, to help us decide which hypotheses
-to consider or discard for more complex and costly models, or to suggest
-new hypotheses.
+\citep{pillowetal2008,dunnetal2013,tyrchaetal2014,huang2015,battistinetal2017}.
+It would be useful to explore models that operate in between, addressing
+the activity of the larger brain area whence the sample comes, but
+\emph{collectively}, that is, without asking about individual neuronal
+details. Such intermediate models would be useful at least for preliminary
+investigations, to help us decide which hypotheses to consider or discard
+for more complex and costly models, or to suggest new hypotheses.
 
@@
 different in general; a misjudgement of their approximate equality can lead to
-\enquote{silly answers}, to quote Mackay \cite*[p.~308]{mackay1995_r2003}.)
+\enquote{silly answers}, to quote MacKay \citep[p.~308]{mackay1995_r2003}.)
 

### 2019-12-14 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-14T18:54:42+0100
+%% Last-Updated: 2019-12-15T00:52:50+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\paragraphfootnotes
-% \threecolumnfootnotes
-\setlength{\footmarkwidth}{0em}
+%\paragraphfootnotes\setlength{\footmarkwidth}{0em}
+%\twocolumnfootnotes
 %\setlength{\footmarksep}{0em}
-\footmarkstyle{\textsuperscript{\color{myred}\bfseries#1}~}
+\footmarkstyle{\textsuperscript{\color{myred}\scriptsize\bfseries#1}~}
 %\footmarkstyle{\textsuperscript{[#1]}~}
@@
 %   population activity as an example. [} 
-In the present work we propose such an intermediate probabilistic model. It
-answers this question: how much was the \emph{total} activity of a large
-neuronal population, given the observation of the activity of a very small
-sample thereof? This model addresses a larger brain area, avoiding the
-assumption of isolation of the sample; and by focusing on the total
-activity, rather than the activity of individual neurons, it remains simple
-and numerically tractable.
+In the present work we propose such an intermediate probabilistic model
+\citep[\cf][]{portamanaetal2015}. It answers this question: how much was
+the \emph{total} activity of a large neuronal population, given the
+observation of the activity of a very small sample thereof? This model
+addresses a larger brain area, avoiding the assumption of isolation of the
+sample; and by focusing on the total activity, rather than the activity of
+individual neurons, it remains simple and numerically tractable.
 
@@
 A maximum-entropy distribution like $\px(\yA \| M, N)$ is equivalent to the
-zeroth-order (Laplace
-\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000})
-approximation of four distinct distributions for the larger population,
-which differ numerically from one another in higher-order approximations:
+zeroth-order Laplace approximation
+\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
+four distinct distributions for the larger population, which differ
+numerically from one another in higher-order approximations:
 \begin{enumerate}[label=(\roman*)]

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T00:52:50+0100
+%% Last-Updated: 2019-12-15T16:26:55+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Step~\ref{item:sample_step} is just an application of the probability-calculus, which gives an exact linear relation between the first $m$
-moments for the larger population and the first $m$ for the sample
-\citep[Appendix]{portamanaetal2015}. The ones determine the others and
-vice versa at every time bin. This relation holds for any belief
-distribution $\pxxx(\yA_{t})$ for the larger-population activity and its
-marginal $\pxx(\ya_{t})$ for the sample activity at that bin. These two
-distributions are related by
+Step~\ref{item:sample_step} is just an application of the
+probability-calculus, which gives an exact linear relation between the
+first $m$ moments for the larger population and the first $m$ for the
+sample \citep[see][for a complete discussion and proofs. Some proofs are
+summarized again here in
+appendix~\ref{sec:derivation_maxent_fullnet}]{portamanaetal2015}. The ones
+determine the others and vice versa at every time bin. This relation holds
+for any belief distribution $\pxxx(\yA_{t})$ for the larger-population
+activity and its marginal $\pxx(\ya_{t})$ for the sample activity at that
+bin. These two distributions are related by
 \begin{equation}
@@
 averages, typically time averages. (A judgement is required because
-probability expectations and empirical averages are numerically
-different in general; a misjudgement of their approximate equality can lead to
+probability expectations and empirical averages differ numerically in
+general; a misjudgement of their approximate equality can lead to
 \enquote{silly answers}, to quote MacKay \citep[p.~308]{mackay1995_r2003}.)
@@
 The amount and degrees of the constraining moments depend on the questions
-and hypotheses that a researcher is exploring. We give some examples in the
-next two sections. Note that the constraint
-equation~\eqref{eq:final_constraint} for the
-distribution~\eqref{eq:app_maxent_pop_first} may not have exact solutions
-if $N$ is strictly larger than $n$. The possible discrepancy between its
+and hypotheses under study. We give some examples in the
+next two sections.
+
+Note that the constraint equation~\eqref{eq:final_constraint} for the
+distribution~\eqref{eq:app_maxent_pop_first} may not have solutions
+if $N$ is strictly larger than $n$, but it's still possible to find The possible discrepancy between its
 two sides typically increases with the number of constraints. This
@@
   \pf[(\yA_{1},\dotsc, \yA_{T}) \| \yI]
-  = \int\!\!\!\di\ynu\;\pf(\ynu \|\yI)\,\prod_{\yA=1}^{N}{\nu_{\yA}}^{T\,\yFF_{\yA}},
+  = \int\!\!\!\di\ynu\;\pf(\ynu \|\yI)\,\prod_{\yA=0}^{N}{\nu_{\yA}}^{T\,\yFF_{\yA}},
 \end{equation}
@@
 activities in an imaginary continuation or repetition of the recording, and
-$\pf(\ynu \|\yI)\;\di\ynu$ our belief distribution about it.
-
-The second assumption concerns the latter degree distribution. It has one
-of these expressions:
+$\pf(\ynu \|\yI)\;\di\ynu$ our belief distribution about it. The
+integration is over an $N$-dimensional simplex.
+
+The second assumption concerns the latter distribution. Its density
+function has one of the following expressions:
+\begin{enumerate}[label=(\textit{\alph*}),wide]
+\item an entropic prior \citep{skillingetal1984,rodriguez1991,neumann2007}
   \begin{subequations}\label{eq:maxent_prior}
@@
     \end{equation}
-  where $\yr\defd(r_{\yA})$ is a reference distribution and the expression
-  in larger parentheses is a multinomial coefficient, or
+    where $\yr\defd(r_{\yA})$ is a reference distribution and the
+    expression in larger parentheses is a multinomial coefficient;
+  \item a \enquote{sufficiency prior}, from the exponential family, which
+    follows from the Pitman-Koopman theorem if some moments $\set{m}$ are
+    sufficient statistics for our belief about $\ynu$:
     \begin{equation}
@@
   normalization factor for the exponential distribution within the delta.
-This is a formal expression saying that our belief about $\ynu$ is an
-exponential parametric model for which the normalized factorial moments are
-informationally sufficient.
+  This is a formal expression, the delta symbolizing that the possible
+  domain of $\ynu$ is restricted from an $N$-dimensional simplex to a
+  lower-dimensional subset thereof, parametrized by $\yla$.
+\end{enumerate}
 

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T16:26:55+0100
+%% Last-Updated: 2019-12-15T17:33:33+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 where the $m'$ parameters $\yll \defd (\yl_{m})$ are determined by the
-constraints~\eqref{eq:final_constraint}. This formula is further discussed
-in appendix~\ref{sec:derivation_maxent_fullnet}.
-
-The amount and degrees of the constraining moments depend on the questions
-and hypotheses under study. We give some examples in the
-next two sections.
-
-Note that the constraint equation~\eqref{eq:final_constraint} for the
-distribution~\eqref{eq:app_maxent_pop_first} may not have solutions
-if $N$ is strictly larger than $n$, but it's still possible to find The possible discrepancy between its
-two sides typically increases with the number of constraints. This
-discrepancy comes from an approximation implicit in the maximum-entropy
-method combined with one assumption specific to our application: that
-measured averages equal expected values (or equivalently, the number of
-time bins $T$ is infinite), and that an activity level $\yA$ can equally
+constraints~\eqref{eq:final_constraint} and can be found by convex
+extremization, discussed further in
+appendix~\ref{sec:derivation_maxent_fullnet}. Note that for $N>n$ the
+constraint equation~\eqref{eq:final_constraint} may never be exactly
+satisfied by the distribution~\eqref{eq:app_maxent_pop_first} for any
+$\yla$. In this case the extremization minimizes the discrepancy between
+the left and right sides of~\eqref{eq:final_constraint}. This discrepancy
+comes from an approximation implicit in the maximum-entropy method (namely
+that the number of time bins $T$ is infinite) combined with one assumption
+specific to our application: that an activity level $\yA$ can equally
 likely be generated by any set of $\yA$ neurons in the larger population.
@@
   14147]{stensolaetal2012}, consists of $n=65$ neurons from rat Medial
-  Entorhinal Cortex, recorded for about 20 minutes. Their spikes are binned
-  into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
+  Entorhinal Cortex, simultaneously recorded for about 20 minutes. Their
+  spikes are binned into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
 \item The \textbf{second data set}, from \textcite[data courtesy by A.
   Riehle and T. Brochier]{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
-  consists of $n=159$ neurons from macaque Motor Cortex, recorded for about
-  15 minutes. Their spikes are binned into $T=300\,394$ bins of
-  $3\;\textrm{ms}$ width.
+  consists of $n=159$ neurons from macaque Motor Cortex, simultaneously
+  recorded for about 15 minutes. Their spikes are binned into $T=300\,394$
+  bins of $3\;\textrm{ms}$ width.
 \end{itemize}
@@
   domain of $\ynu$ is restricted from an $N$-dimensional simplex to a
-  lower-dimensional subset thereof, parametrized by $\yla$.
+  lower-dimensional subset thereof, parametrized by $\yla$. When this prior
+  is used, all integrations over $\ynu$ are replaced by integrations over
+  $\yla$.
 \end{enumerate}
@@
 distribution for the joint frequency distribution $\yJ \defd (J_{\ya\yA})$
-conditional on $\yf$:
+and the long-run frequency distribution $\ynu$, conditional on $\yf$:
 \begin{multline}
   \label{eq:prob_joint_frequencies}
-  \pf(\yJ \| \yf, \yI) \propto
-                         \int\!\!\!\di\ynu\;\pf(\ynu\|\yI)\;
+  \pf(\yJ, \ynu \| \yf, \yI) \propto
+  %\int\!\!\!\di\ynu\;
+  \pf(\ynu\|\yI)\;
                          \binom{T}{TJ_{0 0},\dotsc,TJ_{n N}}\,
@@
   \approx{}  \\
-\shoveright{\int\!\!\!\di\ynu\;\pf(\ynu\|\yI)\;
+  \shoveright{%\int\!\!\!\di\ynu\;
+    \pf(\ynu\|\yI)\;
                          \exp\biggl(
@@
 where $\pf(\ynu\|\yI)$ is either of the
-distributions~\eqref{eq:maxent_prior}. From this belief distribution we can
-find the one for $\yFF_{\yA}\equiv \sum_{\ya}J_{\ya\yA}$ by
+distributions~\eqref{eq:maxent_prior}. Note the relative entropy in the
+exponential. If we want to conditionalize not on $\yf$ but on a set of
+measured averages $C(m) = \sum_{\ya}c_{\ya}(m) \yff_{\ya}$, $m\in M$, such
+as~\eqref{eq:chain_prescription_moments}, we have to replace the
+integration constraint $\sum_{\yA}J_{\ya\yA}=\yff_{\ya}$ with
+$\sum_{\ya\yA}c_{\ya}(m)\,J_{\ya\yA}=C(m)$.
+
+From this belief distribution we can find the one for
+$\yFF_{\yA}\equiv \sum_{\ya}J_{\ya\yA}$ and any others of interest by
 marginalization. Within the exponential we can recognize the relative

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T17:33:33+0100
+%% Last-Updated: 2019-12-15T17:46:35+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   spikes are binned into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
-\item The \textbf{second data set}, from \textcite[data courtesy by A.
-  Riehle and T. Brochier]{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
+\item The \textbf{second data set}, from \textcite[supplementary
+  material]{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
   consists of $n=159$ neurons from macaque Motor Cortex, simultaneously

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T17:46:35+0100
+%% Last-Updated: 2019-12-15T18:01:21+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The maximum-entropy distribution for the larger population is calculated
-using five moments. This number seems to provide almost as much information
-as the full frequency distribution of the sample (see next section).
-Figure~\ref{fig:plots_5mom} shows the resulting densities
-(distribution${}\times N$) for three example values of larger-population
-sizes: $N=1\,000$ (\textcolor{mygreen}{green diamonds}), $N=5\,000$
-(\textcolor{myred}{red circles}), $N=10\,000$ (\textcolor{myblue}{blue
-  curve}). The frequency density of the sample activity is also shown
-(\textcolor{black}{black triangles}), and in the plot it would be
-indistinguishable from the maximum-entropy density for $N=n$, that is,
-applied at the sample level. We discuss the case of unknown $N$ in
-\sect~\ref{sec:N}.
+using five moments, because this number provides almost as much information
+as the full frequency distribution of the sample, as discussed in
+\sect~\ref{sec:marginalization}. Figure~\ref{fig:plots_5mom} shows the
+resulting distributions for three example values of
+larger-population sizes: $N=1\,000$ (\textcolor{mygreen}{green diamonds}),
+$N=5\,000$ (\textcolor{myred}{red circles}), $N=10\,000$
+(\textcolor{myblue}{blue curve}). The frequency density of the sample
+activity is also shown (\textcolor{black}{black triangles}), and in the
+plot it would be indistinguishable from the maximum-entropy density for
+$N=n$, that is, applied at the sample level. We discuss the case of unknown
+$N$ in \sect~\ref{sec:N}.
 \begin{figure}[!p]
@@
 \includegraphics[width=\linewidth]{_scripts/newpaper_stensola_5mom_distributionsN_priorc.pdf}%
-\\[2em]%
+\\[1em]%\mbox{}\vfill\mbox{}%
 \subcaption[]{\label{fig:plots_5mom_b}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper_riehle_5mom_distributionsN_priorc.pdf}%
-\\[2em]%
-\caption{Maximum-entropy distributions for the normalized total activity of
-  the larger population, assuming several population sizes $N$. Five
-  moments are constrained.}
+\\[1em]%\mbox{}\vfill\mbox{}%
+\caption{Maximum-entropy distributions for the total activity of the larger
+  population, using five moments and assuming several population sizes $N$.
+  In order to compare different $N$, the distributions are multiplied by
+  $N$ (obtaining densities) and the activity divided by $N$ (obtaining a
+  population average). Note how the distributions for large $N$ have very
+  different features from the measured frequency distribution, and how
+  their peaks become more pronounced as $N$ increases.}
 \label{fig:plots_5mom}
@@
 % maxent_solutions_different_moments_partial.nb
-Here is a toy example of possible uses of this model,
-based on the first data set. We could be interested in the hypothesis that
-two distinct cell types or assemblies be present in the region where the
-recording was made. Finding a larger-population distribution with two
-peaks, as in \fig~\ref{fig:plots_5mom_a}, would provide
-some evidence for this hypothesis. Let's further imagine that we have
-reasons for suspecting that a specific set of the sampled neurons is of the
-first type, and the remaining of the second type. In the case of the first
-data set, 27 of the 65 sampled neurons were identified as grid cells
-belonging to 3--4 functional modules \citep{dunnetal2015}. Could the
-two peaks in the distribution of \fig~\ref{fig:plots_5mom} reflect the
-activities of grid versus non-grid cells? We apply the model to these two
-sets of neurons individually, using $\yNg=4\,150$ for the grid set and
-$\yNng=5\,850$ for the non-grid set, to reflect their proportions (27/65
-and 38/65) in the recorded sample. The results are shown in 
- \fig~\ref{fig:plots_cellsubsets_a}. The distribution for the larger
-population of grid cells (\textcolor{mygreen}{green triangles}) seems to
-have one broad peak, close to the first peak of the larger-population
-distribution. The distribution for the larger population of non-grid cells
-(\textcolor{myred}{red circles}) has two peaks instead, roughly at the same
-population-averaged activities as the peaks of the larger-population distribution
-(\textcolor{myblue}{blue curve}) but closer in height. It would thus seem
-that the population of grid cells is contributing to the first mode of the
-larger population, but it is not its sole contributor. We can also assess
-if the distributions for the two sets of neurons are independent. If they
-were independent, the larger-population distribution would be given by
-their convolution:
+Here is a toy example of possible uses of this model, based on the first
+data set. We could be interested in the hypothesis that two distinct cell
+types or assemblies be present in the region where the recording was made.
+Finding a larger-population distribution with two peaks, as in
+\fig~\ref{fig:plots_5mom_a}, would provide some evidence for this
+hypothesis. We might have reasons for suspecting that a specific set of the
+sampled neurons is of the first type, and the remaining of the second type.
+In the case of the first data set, 27 of the 65 sampled neurons were
+identified as grid cells belonging to 3--4 functional modules
+\citep{stensolaetal2012}. Could the two peaks in the distribution of
+\fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
+cells? We apply the model to these two sets of neurons individually, using
+$\yNg=4\,150$ for the grid set and $\yNng=5\,850$ for the non-grid set, to
+reflect their proportions (27/65 and 38/65) in the recorded sample. The
+results are shown in \fig~\ref{fig:plots_cellsubsets_a}. The distribution
+for the larger population of grid cells (\textcolor{mygreen}{green
+  triangles}) seems to have one broad peak, close to the first peak of the
+larger-population distribution. The distribution for the larger population
+of non-grid cells (\textcolor{myred}{red circles}) has two peaks instead,
+roughly at the same population-averaged activities as the peaks of the
+larger-population distribution (\textcolor{myblue}{blue curve}) but closer
+in height. It would thus seem that the population of grid cells is
+contributing to the first mode of the larger population, but it is not its
+sole contributor. We can also assess if the distributions for the two sets
+of neurons are independent. If they were independent, the larger-population
+distribution would be given by their convolution:
 \begin{equation}

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T18:01:21+0100
+%% Last-Updated: 2019-12-15T18:25:32+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+% maxent_solutions_different_moments_partial.nb
+Here is a toy example of possible uses of this model, based on the first
+data set. We could be interested in the hypothesis that two distinct cell
+types or assemblies be present in the region where the recording was made.
+Finding a larger-population distribution with two peaks, as in
+\fig~\ref{fig:plots_5mom_a}, would provide some evidence for this
+hypothesis. We might have reasons for suspecting that a specific set of the
+sampled neurons is of the first type, and the remaining of the second type.
+In the case of the first data set, 27 of the 65 sampled neurons were
+identified as grid cells belonging to 3--4 functional modules
+\citep{stensolaetal2012}. Could the two peaks in the distribution of
+\fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
+cells? We apply the model to these two sets of neurons individually, using
+$\yNg=4\,150$ for the grid set and $\yNng=5\,850$ for the non-grid set, to
+reflect their proportions (27/65 and 38/65) in the recorded sample. The
+results are shown in \fig~\ref{fig:plots_cellsubsets_a}.
 \begin{figure}[!p]
@@
 \end{figure}
-% maxent_solutions_different_moments_partial.nb
-Here is a toy example of possible uses of this model, based on the first
-data set. We could be interested in the hypothesis that two distinct cell
-types or assemblies be present in the region where the recording was made.
-Finding a larger-population distribution with two peaks, as in
-\fig~\ref{fig:plots_5mom_a}, would provide some evidence for this
-hypothesis. We might have reasons for suspecting that a specific set of the
-sampled neurons is of the first type, and the remaining of the second type.
-In the case of the first data set, 27 of the 65 sampled neurons were
-identified as grid cells belonging to 3--4 functional modules
-\citep{stensolaetal2012}. Could the two peaks in the distribution of
-\fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
-cells? We apply the model to these two sets of neurons individually, using
-$\yNg=4\,150$ for the grid set and $\yNng=5\,850$ for the non-grid set, to
-reflect their proportions (27/65 and 38/65) in the recorded sample. The
-results are shown in \fig~\ref{fig:plots_cellsubsets_a}. The distribution
-for the larger population of grid cells (\textcolor{mygreen}{green
-  triangles}) seems to have one broad peak, close to the first peak of the
-larger-population distribution. The distribution for the larger population
-of non-grid cells (\textcolor{myred}{red circles}) has two peaks instead,
-roughly at the same population-averaged activities as the peaks of the
-larger-population distribution (\textcolor{myblue}{blue curve}) but closer
-in height. It would thus seem that the population of grid cells is
-contributing to the first mode of the larger population, but it is not its
-sole contributor. We can also assess if the distributions for the two sets
-of neurons are independent. If they were independent, the larger-population
-distribution would be given by their convolution:
+The distribution for the larger population of grid cells
+(\textcolor{mygreen}{green triangles}) seems to have one broad peak, close
+to the first peak of the larger-population distribution. The distribution
+for the larger population of non-grid cells (\textcolor{myred}{red
+  circles}) has two peaks instead, roughly at the same population-averaged
+activities as the peaks of the larger-population distribution
+(\textcolor{myblue}{blue curve}) but closer in height. It would thus seem
+that the population of grid cells is contributing to the first mode of the
+larger population, but it is not its sole contributor. We can also assess
+if the distributions for the two sets of neurons are independent. If they
+were independent, the larger-population distribution would be given by
+their convolution:
 \begin{equation}

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T18:25:32+0100
+%% Last-Updated: 2019-12-15T19:16:46+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 sampled neurons is of the first type, and the remaining of the second type.
-In the case of the first data set, 27 of the 65 sampled neurons were
-identified as grid cells belonging to 3--4 functional modules
-\citep{stensolaetal2012}. Could the two peaks in the distribution of
+In the case of the first data set, 27 of the 65 sampled cells were
+identified as grid cells belonging to 3--4 functional modules, the
+remaining 38 \enquote{non-grid} cells being interneurons and unidentified
+cells \citep{stensolaetal2012}. Could the two peaks in the distribution of
 \fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
@@
 \citep[\eg][]{martignonetal1995,schneidmanetal2006,shlensetal2006} or
-\enquote{synchrony} \citep[\eg][-- we're only citing early papers using
-these terms]{bohteetal2000,amarietal2003} of neuronal activity. Most, if
-not all, such applications use the method \emph{at the sample level}. In
-this section we discuss how our proposed application bears on this kind of
-quantification and compares with the traditional application at the sample
-level. But such a comparison involves some methodological caveats which we
-wish to discuss first.% For concreteness we use an
+\enquote{synchrony} \citep[\eg]{bohteetal2000,amarietal2003} of neuronal
+activity. Most, if not all, such applications use the method \emph{at the
+  sample level}. In this section we discuss how our proposed application
+bears on this kind of quantification and compares with the traditional
+application at the sample level. But such a comparison involves some
+methodological caveats which we wish to discuss
+first.% For concreteness we use an
 % example with specific statistics: first and second moments (equivalent to
@@
 \sect~14.2]{jaynes1994_r2003}{cifarellietal1982,kullbacketal1951}[the
-notion goes back to][]{fisher1922} because it relates to those terms, is
-intuitive, and is connected with maximum-entropy distributions. Its idea is
-as follows. Our probabilities about the frequencies of the activities of
-the sample, or about the activity of the sample in a new time bin, are in
-principle conditional on all experimental data and statistics we have. But
-it can be the case that discarding part of the data or statistics -- for
-example, the third- and higher-order empirical moments -- leaves our
-probabilities almost unchanged. This means that the discarded statistics
-are \emph{informationally irrelevant} or almost so. The remaining
-statistics -- for example, first and second empirical moments -- are
-\emph{informationally sufficient}.\footnote{For more technical results and
-  connections with the notion of symmetry see \eg\
+notion goes back to][]{fisher1922} because it relates to those terms and
+especially because is tightly connected with maximum-entropy distributions.
+Its idea is as follows. Our probabilities about the frequencies of the
+activities of the sample, or about the activity of the sample in a new time
+bin, are in principle conditional on all experimental data and statistics
+we have. But it can be the case that discarding part of the data or
+statistics -- for example, the third- and higher-order empirical moments --
+leaves our probabilities almost unchanged. This means that the discarded
+statistics are \emph{informationally irrelevant} or almost so. The
+remaining statistics -- for example, first and second empirical moments --
+are \emph{informationally sufficient}.\footnote{For more technical results
+  and connections with the notion of symmetry see \eg\
   \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,lauritzen1982_r1988,lauritzen2007,cifarellietal1980,cifarellietal1981,diaconisetal1981,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,ayetal2015}.}
@@
 
-There's a tight connection between informational sufficiency and
-maximum-entropy distributions
-\citep{jaynes1982b}[\sect~4.5.4]{bernardoetal1994}: if a probability
-distribution for repetitive phenomena has a sufficient statistics, then by
-the Pitman-Koopman theorem \citep{koopman1936,pitman1936,darmois1935}[for
-later analyses and the discrete case
+An important feature of this notion is its tight connection with
+maximum-entropy distributions: \emph{if a probability distribution for
+  repetitive phenomena has a sufficient statistics, then it is a mixture of
+  distributions of maximum-entropy form}. \mynote{move the following
+  somewhere else? }This mathematical result comes from the Pitman-Koopman
+theorem\citep{jaynes1982b}[\sect~4.5.4]{bernardoetal1994}
+\citep{koopman1936,pitman1936,darmois1935}[for later analyses and the
+discrete case
 see][]{hipp1974,andersen1970,denny1967,denny1972,fraser1963,barankinetal1963,barndorffnielsen1978_r2014}
-it is a mixture of exponential distributions of maximum-entropy form.
-
-We can thus quantify the informational relevance of a subset of statistics,
-for example first and second moments (means and correlations), with respect
-to a larger set, for example the first four moments, by comparing the
-probabilities conditional on the subset and on the full set. For
-probabilities built with the maximum-entropy method, this means comparing
-those constrained on the subset and on the full set. This procedure is
-actually equivalent to comparing the probabilities of the two hypotheses
-about sufficiency, conditional on the \emph{full} data, assuming their
-pre-data probabilities to be equal. We show this equivalence below and
-perform the calculations for our first data set.
-\medskip
+
+\mynote{modify this paragraph }We can thus quantify the informational
+relevance of a subset of statistics, for example first and second moments
+(means and correlations), with respect to a larger set, for example the
+first four moments, by comparing the probabilities conditional on the
+subset and on the full set. For probabilities built with the
+maximum-entropy method, this means comparing those constrained on the
+subset and on the full set. This procedure is actually equivalent to
+comparing the probabilities of the two hypotheses about sufficiency,
+conditional on the \emph{full} data, assuming their pre-data probabilities
+to be equal. We show this equivalence below and perform the calculations
+for our first data set. \medskip
 

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T19:16:46+0100
+%% Last-Updated: 2019-12-15T19:19:48+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-An important feature of this notion is its tight connection with
-maximum-entropy distributions: \emph{if a probability distribution for
+This notion is useful in our present problem, owing to its tight connection
+with maximum-entropy distributions: \emph{if a probability distribution for
   repetitive phenomena has a sufficient statistics, then it is a mixture of

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T19:19:48+0100
+%% Last-Updated: 2019-12-15T19:33:22+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \citep[\sect~4.5]{bernardoetal1994}[\chap~8 \amp\
-\sect~14.2]{jaynes1994_r2003}{cifarellietal1982,kullbacketal1951}[the
-notion goes back to][]{fisher1922} because it relates to those terms and
-especially because is tightly connected with maximum-entropy distributions.
-Its idea is as follows. Our probabilities about the frequencies of the
-activities of the sample, or about the activity of the sample in a new time
-bin, are in principle conditional on all experimental data and statistics
-we have. But it can be the case that discarding part of the data or
-statistics -- for example, the third- and higher-order empirical moments --
-leaves our probabilities almost unchanged. This means that the discarded
-statistics are \emph{informationally irrelevant} or almost so. The
-remaining statistics -- for example, first and second empirical moments --
-are \emph{informationally sufficient}.\footnote{For more technical results
-  and connections with the notion of symmetry see \eg\
+\sect~14.2]{jaynes1994_r2003} because it relates to those terms and
+especially because it is tightly connected with maximum-entropy
+distributions. We now shortly explain its meaning and use; please see the
+end of this section for a bibliography with in-depth discussions.
+
+Our probabilities about the frequencies of the activities of the sample, or
+about the activity of the sample in a new time bin, are in principle
+conditional on all experimental data and statistics we have. But it can be
+the case that discarding part of the data or statistics -- for example, the
+third- and higher-order empirical moments -- leaves our probabilities
+almost unchanged. This means that the discarded statistics are
+\emph{informationally irrelevant} or almost so. The remaining statistics --
+for example, first and second empirical moments -- are
+\emph{informationally sufficient}.\footnote{For more technical results and
+  connections with the notion of symmetry see \eg\
   \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,lauritzen1982_r1988,lauritzen2007,cifarellietal1980,cifarellietal1981,diaconisetal1981,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,ayetal2015}.}
@@
   somewhere else? }This mathematical result comes from the Pitman-Koopman
-theorem\citep{jaynes1982b}[\sect~4.5.4]{bernardoetal1994}
+theorem\citep{cifarellietal1982,kullbacketal1951}[the
+notion goes back to][]{fisher1922} \citep{jaynes1982b}[\sect~4.5.4]{bernardoetal1994}
 \citep{koopman1936,pitman1936,darmois1935}[for later analyses and the

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T19:33:22+0100
+%% Last-Updated: 2019-12-15T19:43:48+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 we use the notion of \emph{informational sufficiency}
-\citep[\sect~4.5]{bernardoetal1994}[\chap~8 \amp\
+\citep[\sect~4.5]{bernardoetal1994}[\chap~8 and
 \sect~14.2]{jaynes1994_r2003} because it relates to those terms and
@@
 about the activity of the sample in a new time bin, are in principle
-conditional on all experimental data and statistics we have. But it can be
-the case that discarding part of the data or statistics -- for example, the
-third- and higher-order empirical moments -- leaves our probabilities
-almost unchanged. This means that the discarded statistics are
+conditional on all experimental data and statistics we have. But our belief
+distribution can be such that discarding part of the data or statistics --
+for example, the third- and higher-order empirical moments -- leaves it
+unchanged, or almost so. This means that the discarded statistics are
 \emph{informationally irrelevant} or almost so. The remaining statistics --

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T19:43:48+0100
+%% Last-Updated: 2019-12-15T19:49:18+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 we use the notion of \emph{informational sufficiency}
-\citep[\sect~4.5]{bernardoetal1994}[\chap~8 and
+\citep{kullbacketal1951}[\sect~4.5]{bernardoetal1994}[\chap~8 and
 \sect~14.2]{jaynes1994_r2003} because it relates to those terms and
@@
 for example, first and second empirical moments -- are
-\emph{informationally sufficient}.\footnote{For more technical results and
-  connections with the notion of symmetry see \eg\
-  \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,lauritzen1982_r1988,lauritzen2007,cifarellietal1980,cifarellietal1981,diaconisetal1981,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,ayetal2015}.}
+\emph{informationally sufficient}.
 % This informational quasi-sufficiency is interesting because
@@
 
-This notion is useful in our present problem, owing to its tight connection
+This notion is useful in our present problem owing to its tight connection
 with maximum-entropy distributions: \emph{if a probability distribution for
   repetitive phenomena has a sufficient statistics, then it is a mixture of
-  distributions of maximum-entropy form}. \mynote{move the following
-  somewhere else? }This mathematical result comes from the Pitman-Koopman
-theorem\citep{cifarellietal1982,kullbacketal1951}[the
-notion goes back to][]{fisher1922} \citep{jaynes1982b}[\sect~4.5.4]{bernardoetal1994}
-\citep{koopman1936,pitman1936,darmois1935}[for later analyses and the
-discrete case
-see][]{hipp1974,andersen1970,denny1967,denny1972,fraser1963,barankinetal1963,barndorffnielsen1978_r2014}
+  distributions of maximum-entropy form}.
 
@@
 sample-level distribution (\textcolor{myblue}{blue empty squares}).
+
+\bigskip
+\mynote{Short reasoned bibliography about sufficient statistics and its
+  relation to maximum-entropy?} blah blah \footnote{For more technical results and
+  connections with the notion of symmetry see \eg\
+  \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,lauritzen1982_r1988,lauritzen2007,cifarellietal1980,cifarellietal1981,diaconisetal1981,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,ayetal2015}.}
+This mathematical result comes from the Pitman-Koopman
+theorem\citep{cifarellietal1982,kullbacketal1951}[the
+notion goes back to][]{fisher1922} \citep{jaynes1982b}[\sect~4.5.4]{bernardoetal1994}
+\citep{koopman1936,pitman1936,darmois1935}[for later analyses and the
+discrete case
+see][]{hipp1974,andersen1970,denny1967,denny1972,fraser1963,barankinetal1963,barndorffnielsen1978_r2014}
 

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T19:49:18+0100
+%% Last-Updated: 2019-12-15T20:01:34+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 distributions. We now shortly explain its meaning and use; please see the
-end of this section for a bibliography with in-depth discussions.
+end of this section for references to in-depth discussions and technical
+results.
 
@@
 
-This notion is useful in our present problem owing to its tight connection
-with maximum-entropy distributions: \emph{if a probability distribution for
-  repetitive phenomena has a sufficient statistics, then it is a mixture of
-  distributions of maximum-entropy form}.
+This notion is useful in our present problem owing to its tight
+mathematical connection with maximum-entropy distributions, in the form of
+a theorem: \emph{if a probability distribution for repetitive phenomena has
+  a sufficient statistics, then it is a mixture of distributions of
+  maximum-entropy form, and vice versa}. Thus, maximum-entropy
+distributions automatically appear (as approximations) from the
+probability-calculus when we test hypotheses about sufficiency.
 

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T20:01:34+0100
+%% Last-Updated: 2019-12-15T20:13:45+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\enquote{Cooperativity}, \enquote{interaction}, and similar terms are
-vague, so we need to translate them into a more precise notion first. Here
-we use the notion of \emph{informational sufficiency}
+\enquote{Cooperativity}, \enquote{interaction}, and similar terms are vague
+and need to be translated into a more precise notion first. We propose here
+the notion of \emph{informational sufficiency}
 \citep{kullbacketal1951}[\sect~4.5]{bernardoetal1994}[\chap~8 and
-\sect~14.2]{jaynes1994_r2003} because it relates to those terms and
-especially because it is tightly connected with maximum-entropy
-distributions. We now shortly explain its meaning and use; please see the
-end of this section for references to in-depth discussions and technical
-results.
+\sect~14.2]{jaynes1994_r2003}. For three reasons:
+\begin{enumerate*}[label=(\roman*)]
+\item it seems closely related to those terms, the way they are used in the
+  literature, \item it is tightly connected with maximum-entropy
+  distributions, \item it reveals important consequences of sampling.
+\end{enumerate*}
+We now shortly illustrate its meaning and use; please see the end of this
+section for references to in-depth discussions and technical results.
 

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T20:13:45+0100
+%% Last-Updated: 2019-12-15T20:16:51+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \item it seems closely related to those terms, the way they are used in the
-  literature, \item it is tightly connected with maximum-entropy
-  distributions, \item it reveals important consequences of sampling.
+  literature; \item it is tightly connected with maximum-entropy
+  distributions; \item it reveals important consequences of sampling.
 \end{enumerate*}
-We now shortly illustrate its meaning and use; please see the end of this
-section for references to in-depth discussions and technical results.
+We shortly illustrate its meaning and use; please see the end of this
+section for references to in-depth discussions and proofs of technical
+results.
 
@@
 
-This notion is useful in our present problem owing to its tight
-mathematical connection with maximum-entropy distributions, in the form of
-a theorem: \emph{if a probability distribution for repetitive phenomena has
-  a sufficient statistics, then it is a mixture of distributions of
-  maximum-entropy form, and vice versa}. Thus, maximum-entropy
-distributions automatically appear (as approximations) from the
-probability-calculus when we test hypotheses about sufficiency.
+This notion is useful for us owing to a mathematical theorem that tightly
+connects it with maximum-entropy distributions: \emph{if a probability
+  distribution for a repetitive phenomenon has a sufficient statistics,
+  then it is a mixture of distributions of maximum-entropy form, and vice
+  versa}. Thus, maximum-entropy distributions automatically appear (as
+approximations) from the probability-calculus when we test hypotheses about
+sufficiency.
 

### 2019-12-15 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T20:16:51+0100
+%% Last-Updated: 2019-12-15T21:12:20+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\setlength{\footmarksep}{0em}
-\footmarkstyle{\textsuperscript{\color{myred}\scriptsize\bfseries#1}~}
+\footmarkstyle{\textsuperscript{%\color{myred}
+\scriptsize\bfseries#1}~}
 %\footmarkstyle{\textsuperscript{[#1]}~}
@@
 \begin{enumerate*}[label=(\roman*)]
-\item it seems closely related to those terms, the way they are used in the
-  literature; \item it is tightly connected with maximum-entropy
+\item it seems closely related to those terms as used in the
+  literature; \item it is mathematically connected with maximum-entropy
   distributions; \item it reveals important consequences of sampling.

### 2019-12-16 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-15T21:12:20+0100
+%% Last-Updated: 2019-12-16T13:30:23+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-As mentioned in the Introduction (see references there), in the
-neurosciences the maximum-entropy method has also been used as a way of
-quantifying the \enquote{cooperativity} \citep[\eg][]{gersteinetal1985} or
+As mentioned in the Introduction, in the neurosciences the maximum-entropy
+method has also been used as a way of quantifying the
+\enquote{cooperativity} \citep[\eg][]{gersteinetal1985} or
 \enquote{interaction}
@@
 activity. Most, if not all, such applications use the method \emph{at the
-  sample level}. In this section we discuss how our proposed application
-bears on this kind of quantification and compares with the traditional
-application at the sample level. But such a comparison involves some
-methodological caveats which we wish to discuss
-first.% For concreteness we use an
+  sample level}, aiming to find a distribution for the sample without
+considering that it comes from a larger population. In this section we
+discuss how our proposed application bears on this kind of quantification
+and compares with the application at the sample level. We also wish to
+highlight some methodological caveats.% For concreteness we use an
 % example with specific statistics: first and second moments (equivalent to
@@
 
-\enquote{Cooperativity}, \enquote{interaction}, and similar terms are vague
-and need to be translated into a more precise notion first. We propose here
-the notion of \emph{informational sufficiency}
+As a first step we need to translate vague terms like
+\enquote{cooperativity}, \enquote{interaction}, and similar into a more
+precise, operational \citep{bridgman1927_r1958} notion. Here we propose a
+notion that
+\begin{enumerate*}[label=(\roman*)]
+\item seems closely related to those terms as used in the
+  literature; \item is mathematically connected with maximum-entropy
+  distributions; \item reveals important consequences of sampling.
+\end{enumerate*}
+This notion is \emph{informational sufficiency}
 \citep{kullbacketal1951}[\sect~4.5]{bernardoetal1994}[\chap~8 and
-\sect~14.2]{jaynes1994_r2003}. For three reasons:
-\begin{enumerate*}[label=(\roman*)]
-\item it seems closely related to those terms as used in the
-  literature; \item it is mathematically connected with maximum-entropy
-  distributions; \item it reveals important consequences of sampling.
-\end{enumerate*}
-We shortly illustrate its meaning and use; please see the end of this
-section for references to in-depth discussions and proofs of technical
-results.
-
-Our probabilities about the frequencies of the activities of the sample, or
-about the activity of the sample in a new time bin, are in principle
-conditional on all experimental data and statistics we have. But our belief
-distribution can be such that discarding part of the data or statistics --
-for example, the third- and higher-order empirical moments -- leaves it
-unchanged, or almost so. This means that the discarded statistics are
-\emph{informationally irrelevant} or almost so. The remaining statistics --
-for example, first and second empirical moments -- are
-\emph{informationally sufficient}.
+\sect~14.2]{jaynes1994_r2003}. Its meaning and the three points above will
+now be illustrated; please see the end of this section for references to
+in-depth discussions and proofs of technical results.
+
+Our belief distribution about the frequencies of the neuronal activities,
+or about the activity in a new time bin, is in principle conditional on all
+experimental data and statistics we have. This distribution, however, can
+be such that it remains unchanged if we discard part of the data or
+statistics -- for example, the third- and higher-order empirical moments --
+but it does change if we discard more. This means that the discarded
+statistics are \emph{informationally irrelevant}: their knowledge does not
+change our beliefs. The remaining statistics -- in the example, first and
+second empirical moments -- are \emph{minimally informationally
+  sufficient}. The qualification \enquote{minimally} will be omitted for
+brevity.
 % This informational quasi-sufficiency is interesting because
@@
 
-This notion is useful for us owing to a mathematical theorem that tightly
-connects it with maximum-entropy distributions: \emph{if a probability
-  distribution for a repetitive phenomenon has a sufficient statistics,
-  then it is a mixture of distributions of maximum-entropy form, and vice
-  versa}. Thus, maximum-entropy distributions automatically appear (as
-approximations) from the probability-calculus when we test hypotheses about
-sufficiency.
-
-\mynote{modify this paragraph }We can thus quantify the informational
-relevance of a subset of statistics, for example first and second moments
-(means and correlations), with respect to a larger set, for example the
-first four moments, by comparing the probabilities conditional on the
-subset and on the full set. For probabilities built with the
-maximum-entropy method, this means comparing those constrained on the
-subset and on the full set. This procedure is actually equivalent to
-comparing the probabilities of the two hypotheses about sufficiency,
-conditional on the \emph{full} data, assuming their pre-data probabilities
-to be equal. We show this equivalence below and perform the calculations
-for our first data set. \medskip
-
-Before applying this notion to neuronal activity, however, we must keep in
-mind that \emph{informational sufficiency is not preserved under sampling}.
-If a probability distribution has some sufficient statistics, then its
-marginals, such as the distribution for a sample, \emph{cannot} have the
-same sufficient statistics, and vice versa; except for trivial cases such
-as uniform probability distributions. % This generally holds also among
+Now let us turn the informational sufficiency of a subset of statistics and
+the informational sufficiency of another subset into two hypotheses. We can
+then ask: which hypothesis is more probable, given the data? equivalently,
+what is the ratio of their probabilities? Translating our question about
+\enquote{cooperativity} into a hypothesis-testing problem about sufficiency
+gives it a well-defined answer determined by the
+probability-calculus. % The ratio of
+% the two probabilities gives us a natural measure of the
+% For example, do
+% the data say that the first two empirical moments are probably enough to
+% determine our beliefs? or is it more probable that the first four are
+% needed instead?
+
+But it does more: through a mathematical theorem, it leads to the automatic
+appearance of maximum-entropy distributions: \emph{if a probability
+  distribution for a repetitive phenomenon has some sufficient statistics,
+  then it is a mixture of distributions of maximum-entropy form having
+  those statistics as constraints, and vice versa}. We give detailed
+references about this theorem at the end of this section.
+% Thus, maximum-entropy distributions automatically appear (as
+% approximations) from the probability-calculus when we test hypotheses
+% about informational sufficiency.
+Its most important consequence is that the probabilities for the two
+hypotheses are approximately proportional, by Bayes's theorem, to the
+probabilities for all the data according to the two maximum-entropy
+distributions constrained on the subsets of statistics (provided we have
+equal pre-data beliefs about the two hypotheses). We shall shortly explain
+and perform this calculation for the first data set.
+
+% For probabilities built with the maximum-entropy method, this means
+% comparing those constrained on the subset and on the full set. This
+% procedure is actually equivalent to comparing the probabilities of the two
+% hypotheses about sufficiency, conditional on the \emph{full} data, assuming
+% their pre-data probabilities to be equal. We show this equivalence below
+% and perform the calculations for our first data set. \medskip
+
+\medskip
+
+An important fact must be taken into account, however: \emph{informational
+  sufficiency is not preserved under sampling}. If a probability
+distribution has some sufficient statistics, then its sample marginals
+\emph{cannot} have the same sufficient statistics, and vice versa
+\citep[\sect~3.1]{portamanaetal2015}; except for trivial cases such as
+uniform probability distributions. % This generally holds also among
 % different marginals: if a statistics is sufficient for some marginal, then
 % it will not be sufficient for another marginal.
-This impossibility is known in statistical mechanics: if a system is
-described by a Gibbs state, its subsystems cannot be perfectly described by
-Gibbs states \citep[\eg][and references therein]{maesetal1999}.
-Mathematically this impossibility comes from the Pitman-Koopman theorem
-mentioned above, and translates into the general impossibility of solving a
-system of independent equations with more equations than unknowns
-\citep[\sect~3.1]{portamanaetal2015}.
-
-This fact is important for our analysis. If, say, means and pairwise
-correlations seem informationally sufficient for a particular sample from a
-brain area, then they may well not be sufficient for the larger population
-of neurons constituting that area, and vice versa. So if we are interested
-in the \enquote{cooperativity} or \enquote{interaction} of a brain area, it
-is unreliable to use a maximum-entropy distribution constructed only for a
-sample
+This impossibility is known in another guise in statistical mechanics: if a
+system is described by a Gibbs state, its subsystems cannot be perfectly
+described by Gibbs states \citep[\eg][and references
+therein]{maesetal1999}. % Mathematically this impossibility comes from the
+% Pitman-Koopman theorem mentioned above, and translates into the general
+% impossibility of solving a system of independent equations with more
+% equations than unknowns .
+
+This fact is important for interpreting the results of our application. For
+example, if means and pairwise correlations seem informationally sufficient
+for a particular sample from a brain area, then they may well not be
+sufficient for the larger population of neurons constituting that area, and
+vice versa. So if we are interested in the \enquote{cooperativity} or
+\enquote{interaction} of the neurons in a brain area, it is unreliable to
+use a maximum-entropy distribution constructed only for a sample
 thereof. % \citep[contrast this with some literature, \eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}
 The model presented here avoids this problem because the maximum-entropy
-method is applied to obtain the distribution of the larger population, not
-of the sample alone.
+method is applied directly to the larger population, not to the sample
+alone.
 
@@
 
-Let us illustrate the remarks above with our first data set.
+Let us finally illustrate the idea and remarks above with our first data set.
 
@@
 \begin{enumerate}[label=(\roman*)]
-\item from each maximum-entropy distributions $\px(\yA \| M,N)$ for the larger
-  population, built from each set of constraints $M=M',M''$, calculate the
-  marginal distribution for the sample:
+\item From each maximum-entropy distribution $\px(\yA \| M,N)$ for the
+  larger population, built from each set of constraints $M=M',M''$,
+  calculate the marginal distribution for the sample by
+  \eqn~\eqref{eq:conditional_hypergeometric_t}:
 \begin{equation}
@@
   \pxx(\ya \| M, N) = \sum_{\yA}\yG_{\ya \yA}\px(\yA \| M,N),
-  \qquad M=M',M'';
-\end{equation}
-\item calculate the relative entropies of the measured frequency
-  distribution $\yf$ with respect to each sample marginal, and multiply
-  them by the number of time bins $T$:
+  \qquad M=M',M''.
+\end{equation}
+\item Calculate the relative entropies of the measured frequency
+  distribution $\yf$ with respect to each sample marginal from the previous
+  step, and multiply them by the number of time bins $T$:
   \begin{equation}
@@
     T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{\pxx(\ya \| M,N)},
-    \qquad M=M',M'';    
-  \end{equation}
-\item take the difference: 
+    \qquad M=M',M''.
+  \end{equation}
+\item Take the difference: 
 \begin{equation}
@@
 \end{enumerate}
-The measure $\ydi_{N}(M'',M')$ so defined is positive if $M''$ is \enquote{more
-  informationally sufficient} than $M'$, and negative otherwise.
-
-Why is this a natural measure? Because $\ydi_{N}(M'',M')$ is equal to the
-log-ratio of the probabilities of the data $\yf$ conditional on the
+The measure $\ydi_{N}(M'',M')$ so defined is positive if $M''$ is more
+probable to be informationally sufficient than $M'$, and negative
+otherwise.
+
+We show in appendix~\ref{sec:measure_suff} that $\ydi_{N}(M'',M')$ is equal
+to the log-ratio of the probabilities of the data $\yf$ conditional on the
 hypotheses $M''$ and $M'$:
@@
   \ydi_{N}(M'',M') =
-  \log\bigl[\pf(\yf \| M'',N)/
-  \pf(\yf \| M',N)\bigr].
-\end{equation}
-This is called their \emph{relative weight of evidence}, the logarithm of
-their \emph{relative Bayes factor}
-\citep[\chap~6]{good1950}{good1975,good1981,good1985,good1983}[\sect~1.4]{osteyeeetal1974}{mackay1992,kassetal1995}[see
-also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
-this equality in appendix~\ref{sec:measure_suff}. The exponential of
-$\ydi_{N}(M'',M')$ tells us how much more probable the data $\yf$ are
-conditional on $M''$, than conditional on $M'$. We can also combine this
-measure with pre-data probabilities for the two hypotheses to obtain the
-ratio of their probabilities conditional on the data
-\citep[\cf][]{bretthorst2013}.
+  \log\frac{\pf(\yf \| M'',N)}{\pf(\yf \| M',N)}.
+\end{equation}
+This is their \emph{relative weight of evidence}
+\citep[\chap~6]{good1950}{good1985}[\sect~1.4]{osteyeeetal1974}, the
+logarithm of their \emph{relative Bayes factor}
+\citep[p.~421]{jeffreys1936}[and references
+therein]{kassetal1995}. %[\chaps~V, VI, A]{jeffreys1939_r1983}.
+The exponential of $\ydi_{N}(M'',M')$ tells us how much more probable the
+data $\yf$ are, conditional on the moments $M''$, than conditional on the
+moments $M'$. We can also combine this measure with pre-data probabilities
+for the two hypotheses to obtain the ratio of their probabilities
+conditional on the data \citep[\cf][]{bretthorst2013}.
 
@@
 In our case consider for example three sets of constraints $M_{2}$,
-$M_{4}$, $M_{5}$, consisting of the first two, four, five moments. For a
+$M_{4}$, $M_{5}$, consisting of the first two, four, five moments; and a
+larger population of size $N=10\,000$. Figure~\ref{fig:plots_2_4mom_a}
+shows the resulting distributions for the two data sets.
+
+We note***
+
+These informational differences show clearly in the shapes of the
+distribution themselves, plotted in 
+\fig~\ref{fig:plots_2_4mom_a}. The two-moment distribution
+(\textcolor{myred}{dashed red}) is unimodal, the four-moment distribution
+(\textcolor{myblue}{solid blue}) is bimodal. The five-moment distribution
+would not be distinguishable from the four-moment one.  Figure~\ref{fig:plots_2_4mom_b} shows the corresponding distributions for the second data set;
+also in this case they are visually very distinct.
+For a
 larger population of size $N=10\,000$, we obtain the following differences:
@@
 and Good in their code-breaking work at Bletchley
-Park:][]{good1985,good1950,good1969}[\sect~4.2]{jaynes1994_r2003}. In
+Park: see][]{good1985%,good1950,good1969
+}[\sect~4.2]{jaynes1994_r2003}. In
 words, the measured frequencies of the sample activity are 35 orders of
@@
 \subcaption[]{\label{fig:plots_2_4mom_a}}%  
-\includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_twomom_distributions10000_priorc.pdf}%
+\includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_threemom_distributions10000_priorc.pdf}%
 \\[2em]%
 \subcaption[]{\label{fig:plots_2_4mom_b}}%  
-\includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_twomom_distributions10000_prior2.pdf}%
+\includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_threemom_distributions10000_prior2.pdf}%
 \\[2em]%

### 2019-12-16 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-16T13:30:23+0100
+%% Last-Updated: 2019-12-16T16:01:08+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In our case consider for example three sets of constraints $M_{2}$,
-$M_{4}$, $M_{5}$, consisting of the first two, four, five moments; and a
-larger population of size $N=10\,000$. Figure~\ref{fig:plots_2_4mom_a}
-shows the resulting distributions for the two data sets.
-
-We note***
-
-These informational differences show clearly in the shapes of the
-distribution themselves, plotted in 
-\fig~\ref{fig:plots_2_4mom_a}. The two-moment distribution
-(\textcolor{myred}{dashed red}) is unimodal, the four-moment distribution
-(\textcolor{myblue}{solid blue}) is bimodal. The five-moment distribution
-would not be distinguishable from the four-moment one.  Figure~\ref{fig:plots_2_4mom_b} shows the corresponding distributions for the second data set;
-also in this case they are visually very distinct.
-For a
-larger population of size $N=10\,000$, we obtain the following differences:
-\begin{equation}
-  \label{eq:diff_suff_N}
-  \begin{split}
-  \ydi_{N}(M_{4},M_{2}) &= 81\;\nat = 35\;\hart,
-\\    \ydi_{N}(M_{5},M_{4}) &= 0.037\;\nat = 0.016\;\hart,
-  \end{split}
-\end{equation}
-where the Hartley ($\hart$) denotes base-10 logarithms
-\citep[\sect~C.4]{iso2009}[it was called \enquote{ban} and used by Turing
-and Good in their code-breaking work at Bletchley
-Park: see][]{good1985%,good1950,good1969
-}[\sect~4.2]{jaynes1994_r2003}. In
-words, the measured frequencies of the sample activity are 35 orders of
-magnitude more probable assuming sufficiency of the first four moments than
-assuming sufficiency of the first two only. But they are about as probable
-($10^{0.016} = 1.04$) assuming sufficiency of the first five moments as of
-the first four moments only.
-
+Let's show a concrete application. Consider for example three sets of
+constraints $M_{2}$, $M_{4}$, $M_{5}$, consisting of the first two, four,
+five moments; and a larger population of size $N=10\,000$.
+Figure~\ref{fig:plots_2_4mom} shows the resulting distributions for the two
+data sets.
 \begin{figure}[!p]
@@
 % maxent_solutions_different_moments_prior2_6moments.nb
-These informational differences show clearly in the shapes of the
-distribution themselves, plotted in 
-\fig~\ref{fig:plots_2_4mom_a}. The two-moment distribution
-(\textcolor{myred}{dashed red}) is unimodal, the four-moment distribution
-(\textcolor{myblue}{solid blue}) is bimodal. The five-moment distribution
-would not be distinguishable from the four-moment one.  Figure~\ref{fig:plots_2_4mom_b} shows the corresponding distributions for the second data set;
-also in this case they are visually very distinct.
+There is a clear visual difference between the distributions based on two
+(\textcolor{myred}{dashed red}) and on four (\textcolor{myblue}{solid
+  blue}) moments; for the first data set, panel
+\subcaptionref{fig:plots_2_4mom_a}, they even differ in the number of
+modes. On the other hand there is barely any visual difference between the
+distributions based on four and on five (\textcolor{myyellow}{dotted
+  yellow}) moments. From these visual differences we immediately conclude
+that two moments are probably not informationally sufficient, whereas four
+and five should both be. We shall see that the
+measure~\eqref{eq:measure_suffic} reflects what we immediately see.
+
+Compare these plots with those obtained applying the model at the sample
+level (equivalent to $N=n$). With the $M_{2}$ and $M_{4}$ constraints we
+obtain the distributions partly shown in
+\fig~\ref{fig:comparisons_marginals} as \textcolor{myred}{red empty
+  circles} for two moments, and \textcolor{myblue}{blue empty squares} for
+four moments. The measured frequencies (black triangles) are also shown.
+The two distributions can only be distinguished on a logarithmic scale, and
+only in their tails. It would be difficult to visually assess which is more
+probable to be informationally sufficient.
 
@@
 
-Compare the results above with those obtained by applying the model at the
-sample level, that is, with $N=n$:
+Calculation of the weight of evidence~\eqref{eq:measure_suffic} for the
+first data set and $N=10\,000$, corresponding to
+\fig~\ref{fig:plots_2_4mom_a}, leads to the following results:
+\begin{equation}
+  \label{eq:diff_suff_N}
+  \begin{split}
+  \ydi_{N}(M_{4},M_{2}) &= 81\;\nat = 35\;\hart,
+\\    \ydi_{N}(M_{5},M_{4}) &= 0.037\;\nat = 0.016\;\hart,
+  \end{split}
+\end{equation}
+where the Hartley ($\hart$) denotes base-10 logarithms
+\citep[\sect~C.4]{iso2009}[called \enquote{ban} and used by Turing and Good
+in their code-breaking work at Bletchley Park:
+see][]{good1985%,good1950,good1969
+}[\sect~4.2]{jaynes1994_r2003}. In words, our data are 35 orders of
+magnitude more probable by assuming sufficiency of $M_{4}$ than by assuming
+sufficiency of $M_{2}$. But the data are about as probable
+($10^{0.016} = 1.04$) by assuming sufficiency of $M_{4}$ as of $M_{5}$. The
+qualitative visual information of \fig~\ref{fig:plots_2_4mom_a} was
+therefore reliable.
+
+For the model at the sample level, that is, with $N=n$, the weights of
+evidence are
 \begin{equation}

### 2019-12-19 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-16T16:01:08+0100
+%% Last-Updated: 2019-12-19T20:55:39+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \enquote{cooperativity}, \enquote{interaction}, and similar into a more
-precise, operational \citep{bridgman1927_r1958} notion. Here we propose a
-notion that
+precise, operational \citep[\chap~I]{bridgman1927_r1958} notion. Here we
+propose the notion of \emph{informational sufficiency} (also called
+\enquote{predictive sufficiency})
+\citep{kullbacketal1951}[\sect~4.5]{bernardoetal1994}[\chap~8 and
+\sect~14.2]{jaynes1994_r2003}, because
 \begin{enumerate*}[label=(\roman*)]
-\item seems closely related to those terms as used in the
-  literature; \item is mathematically connected with maximum-entropy
-  distributions; \item reveals important consequences of sampling.
+\item it seems closely related to those terms as used in the
+  literature; \item it is mathematically connected with maximum-entropy
+  distributions; \item it reveals important consequences of sampling.
 \end{enumerate*}
-This notion is \emph{informational sufficiency}
-\citep{kullbacketal1951}[\sect~4.5]{bernardoetal1994}[\chap~8 and
-\sect~14.2]{jaynes1994_r2003}. Its meaning and the three points above will
-now be illustrated; please see the end of this section for references to
-in-depth discussions and proofs of technical results.
+Its meaning and the three points above will now be illustrated; please see
+the end of this section for references to in-depth discussions and proofs
+of technical results.
 
@@
 be such that it remains unchanged if we discard part of the data or
-statistics -- for example, the third- and higher-order empirical moments --
-but it does change if we discard more. This means that the discarded
-statistics are \emph{informationally irrelevant}: their knowledge does not
-change our beliefs. The remaining statistics -- in the example, first and
-second empirical moments -- are \emph{minimally informationally
-  sufficient}. The qualification \enquote{minimally} will be omitted for
-brevity.
+statistics -- for example, the third- and higher-order empirical moments. % --
+% but it does change if we discard more.
+This means that the discarded statistics are \emph{informationally
+  irrelevant}: knowing them does not change our beliefs. The remaining
+statistics -- in the example, first and second empirical moments -- are
+\emph{informationally sufficient}.%  The qualification \enquote{minimally}
+% will be omitted for brevity.
 % This informational quasi-sufficiency is interesting because
@@
 then ask: which hypothesis is more probable, given the data? equivalently,
-what is the ratio of their probabilities? Translating our question about
-\enquote{cooperativity} into a hypothesis-testing problem about sufficiency
-gives it a well-defined answer determined by the
+what is the ratio of their probabilities? By translating our question about
+\enquote{cooperativity} into a hypothesis-testing problem about sufficiency,
+we can find a well-defined answer determined by the
 probability-calculus. % The ratio of
@@
 
-But it does more: through a mathematical theorem, it leads to the automatic
-appearance of maximum-entropy distributions: \emph{if a probability
-  distribution for a repetitive phenomenon has some sufficient statistics,
-  then it is a mixture of distributions of maximum-entropy form having
-  those statistics as constraints, and vice versa}. We give detailed
-references about this theorem at the end of this section.
+But such a translation does more: through a mathematical theorem, it leads
+to the automatic appearance of maximum-entropy distributions: \emph{if a
+  probability distribution for a repetitive phenomenon has some sufficient
+  statistics, then it is a mixture of distributions of maximum-entropy form
+  having those statistics as constraints, and vice versa}. See the end of
+this section for detailed references about this theorem.
 % Thus, maximum-entropy distributions automatically appear (as
@@
 % about informational sufficiency.
-Its most important consequence is that the probabilities for the two
-hypotheses are approximately proportional, by Bayes's theorem, to the
-probabilities for all the data according to the two maximum-entropy
-distributions constrained on the subsets of statistics (provided we have
-equal pre-data beliefs about the two hypotheses). We shall shortly explain
-and perform this calculation for the first data set.
+Its most important consequence is that, by Bayes's formula, the probability
+that some statistics are informationally sufficient can be approximated by
+finding the maximum-entropy distribution constrained by those statistics
+and then calculating the probability for the whole data under that
+distribution. We shall shortly illustrate this kind of calculation for the
+first data set.
 
@@
 
-An important fact must be taken into account, however: \emph{informational
-  sufficiency is not preserved under sampling}. If a probability
-distribution has some sufficient statistics, then its sample marginals
-\emph{cannot} have the same sufficient statistics, and vice versa
-\citep[\sect~3.1]{portamanaetal2015}; except for trivial cases such as
-uniform probability distributions. % This generally holds also among
+The theorem above makes it very natural to reinterpret studies about
+cooperativity, interaction, and similar, based on maximum-entropy, as
+studies about informational sufficiency. An important fact must then be
+taken into account: \emph{informational sufficiency is not preserved under
+  sampling}. If a probability distribution has some sufficient statistics,
+then its sample marginals \emph{cannot} have the same sufficient
+statistics, and vice versa \citep[\sect~3.1]{portamanaetal2015}; except for
+trivial cases such as uniform probability
+distributions. % This generally holds also among
 % different marginals: if a statistics is sufficient for some marginal, then
@@
 
-This fact is important for interpreting the results of our application. For
-example, if means and pairwise correlations seem informationally sufficient
-for a particular sample from a brain area, then they may well not be
-sufficient for the larger population of neurons constituting that area, and
-vice versa. So if we are interested in the \enquote{cooperativity} or
-\enquote{interaction} of the neurons in a brain area, it is unreliable to
+This fact affects the interpretation of the results obtained from
+maximum-entropy methods. For example, if means and pairwise correlations
+seem informationally sufficient for a particular sample from a brain area,
+then they are probably not sufficient for the larger population of neurons
+constituting that area, and vice versa. If we are interested in the
+cooperativity of the neurons in a brain area, it is therefore unreliable to
 use a maximum-entropy distribution constructed only for a sample
@@
 \end{equation}
-The data are 1\,220 orders of magnitude more probable conditional on four
-moments than conditional on two, and $10^{1.4}=25$ times more probable
-conditional on five moments than on four moments.
-
-The conclusions about \enquote{cooperativity} or \enquote{interaction} that
-we reach by applying the maximum-entropy method to the larger population,
-as proposed here, are therefore different from those applying it at the
-sample level -- and also visually clearer. The plot in 
-\fig~\ref{fig:plots_2_4mom_a}, showing the two clearly different
-distributions constrained by two and four moments, should be compared with
-the plot for the distributions obtained at the sample level under the same
-constraints, shown in \fig~\ref{fig:comparisons_marginals} as
-\textcolor{myred}{red circles} and \textcolor{myblue}{blue squares}. The
-measured frequencies (black triangles) are also shown. It is necessary to
-use a logarithmic scale to see the differences, which appear mainly in the
-tail.
+So the application at the sample level says that the data are 1\,220 orders
+of magnitude more probable conditional on four moments than conditional on
+two, and $10^{1.4}=25$ times more probable conditional on five moments than
+on four moments. These results are different from the ones obtained from
+the larger population, and they would be hard to deduce visually from a
+look at the distributions of \fig~\ref{fig:comparisons_marginals}.
+
+% The conclusions about \enquote{cooperativity} or \enquote{interaction} that
+% we reach by applying the maximum-entropy method to the larger population,
+% as proposed here, are therefore different from those applying it at the
+% sample level -- and also visually clearer. The plot in 
+% \fig~\ref{fig:plots_2_4mom_a}, showing the two clearly different
+% distributions constrained by two and four moments, should be compared with
+% the plot for the distributions obtained at the sample level under the same
+% constraints, shown in \fig~\ref{fig:comparisons_marginals} as
+% \textcolor{myred}{red circles} and \textcolor{myblue}{blue squares}. The
+% measured frequencies (black triangles) are also shown. It is necessary to
+% use a logarithmic scale to see the differences, which appear mainly in the
+% tail.
 \begin{figure}[!p]
@@
 
-It is also possible to use the measure~\eqref{eq:measure_suffic} and its
-probabilistic interpretation~\eqref{eq:bayes_factor} to compare the probability of
-the data $\yf$ conditional on the hypothesis $(M_{4},N)$ of four-moment
-sufficiency at the larger-population level, $N=10\,000$, versus the
-hypothesis $(M_{4}, n)$ of four-moment sufficiency at the sample level,
+It is in fact possible to use the measure~\eqref{eq:measure_suffic} and its
+probabilistic interpretation~\eqref{eq:bayes_factor} to compare the
+probability of the data $\yf$ conditional on the hypothesis $(M_{4},N)$ of
+four-moment sufficiency at the larger-population level, $N=10\,000$, versus
+the hypothesis $(M_{4}, n)$ of four-moment sufficiency at the sample level,
 $N=n=65$. We obtain

### 2019-12-20 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-19T20:55:39+0100
+%% Last-Updated: 2019-12-20T11:10:29+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 This fact affects the interpretation of the results obtained from
-maximum-entropy methods. For example, if means and pairwise correlations
+maximum-entropy distributions. For example, if means and pairwise correlations
 seem informationally sufficient for a particular sample from a brain area,
@@
 
-Let us finally illustrate the idea and remarks above with our first data set.
-
-We measure the difference $\ydi(M'',M')$ in informational sufficiency
-between a set of moments, say $M''\defd\set{1,\dotsc,m''}$, and another, say
-$M'\defd \set{1,\dotsc,m'}$, as follows:
+We compare the informational sufficiency between of a set of moments, say
+$M''\defd\set{1,\dotsc,m''}$, with respect to another, say
+$M'\defd \set{1,\dotsc,m'}$, by introducing the quantity $\ydi(M'',M')$
+defined as follows:
 \begin{enumerate}[label=(\roman*)]
@@
 \end{enumerate}
-The measure $\ydi_{N}(M'',M')$ so defined is positive if $M''$ is more
+The quantity $\ydi_{N}(M'',M')$ so defined is positive if $M''$ is more
 probable to be informationally sufficient than $M'$, and negative
@@
 
-We show in appendix~\ref{sec:measure_suff} that $\ydi_{N}(M'',M')$ is equal
-to the log-ratio of the probabilities of the data $\yf$ conditional on the
-hypotheses $M''$ and $M'$:
+We show in appendix~\ref{sec:measure_suff} that $\ydi_{N}(M'',M')$ is
+approximately equal to the log-ratio of the probabilities of the data $\yf$
+conditional on the hypotheses $M''$ and $M'$:
 \begin{equation}
   \label{eq:bayes_factor}
-  \ydi_{N}(M'',M') =
+  \ydi_{N}(M'',M') \approx
   \log\frac{\pf(\yf \| M'',N)}{\pf(\yf \| M',N)}.
@@
 The exponential of $\ydi_{N}(M'',M')$ tells us how much more probable the
-data $\yf$ are, conditional on the moments $M''$, than conditional on the
-moments $M'$. We can also combine this measure with pre-data probabilities
-for the two hypotheses to obtain the ratio of their probabilities
-conditional on the data \citep[\cf][]{bretthorst2013}.
+data $\yf$ are, conditional on the sufficiency of $M''$, than conditional
+on the sufficiency of $M'$. We can also combine this measure with pre-data
+probabilities for the two hypotheses to obtain the ratio of their
+probabilities conditional on the data \citep[\cf][]{bretthorst2013}.
 
@@
 
-Let's show a concrete application. Consider for example three sets of
-constraints $M_{2}$, $M_{4}$, $M_{5}$, consisting of the first two, four,
-five moments; and a larger population of size $N=10\,000$.
-Figure~\ref{fig:plots_2_4mom} shows the resulting distributions for the two
-data sets.
+Let us illustrate the ideas, the remarks, and the quantity defined so far
+with our first data set. Consider for example three sets of constraints
+$M_{2}$, $M_{4}$, $M_{5}$, consisting of the first two, four, five moments;
+and a larger population of size $N=10\,000$. Figure~\ref{fig:plots_2_4mom}
+shows the resulting distributions for the two data sets.
 \begin{figure}[!p]
@@
   yellow}) moments. From these visual differences we immediately conclude
-that two moments are probably not informationally sufficient, whereas four
-and five should both be. We shall see that the
+that two moments are less likely to be informationally sufficient than
+four; but four should almost as likely as five. We shall see that the
 measure~\eqref{eq:measure_suffic} reflects what we immediately see.
@@
 only in their tails. It would be difficult to visually assess which is more
-probable to be informationally sufficient.
+likely to be informationally sufficient.
 
@@
 This result is also shown in
-\fig~\ref{fig:comparisons_marginals}, where we see that the four-moment
+\fig~\ref{fig:comparisons_marginals}. We see that the four-moment
 distribution constructed at the larger-population level
@@
 \bigskip
+
+We close this section with some references. The notion of informational or
+predictive sufficiency, probably first presented by Kullback \amp\ Leibler
+\parencite*{kullbacketal1951}, is discussed in Bernardo \etal\
+\parencite*[\sect~4.5]{bernardoetal1994}, who explain its connections with
+Fisher's\parencite*{fisher1922} parametric sufficiency. Jaynes
+\parencite*[\chap~8 and \sect~14.2] shows the connections with information
+theory. The mathematical relation between sufficiency and the exponential
+family comes from the Pitman-Koopman
+theorem \citep{koopman1936,pitman1936,darmois1935} for continuous variables.
+
 \mynote{Short reasoned bibliography about sufficient statistics and its

### 2019-12-20 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-20T11:10:29+0100
+%% Last-Updated: 2019-12-20T20:13:38+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \enquote{cooperativity}, \enquote{interaction}, and similar into a more
-precise, operational \citep[\chap~I]{bridgman1927_r1958} notion. Here we
-propose the notion of \emph{informational sufficiency} (also called
-\enquote{predictive sufficiency})
-\citep{kullbacketal1951}[\sect~4.5]{bernardoetal1994}[\chap~8 and
-\sect~14.2]{jaynes1994_r2003}, because
+precise, operational \citep[\cf][\chap~I]{bridgman1927_r1958} notion. Here
+we propose the notion of \emph{informational sufficiency}, because
 \begin{enumerate*}[label=(\roman*)]
@@
 \end{enumerate*}
-Its meaning and the three points above will now be illustrated; please see
-the end of this section for references to in-depth discussions and proofs
-of technical results.
+Its meaning and the three points above will now be illustrated. For an
+extensive discussion of this notion, probably first suggested by Kullback
+\amp\ Leibler \parencite*{kullbacketal1951}, see Bernardo \etal\
+\parencite*[\sect~4.5]{bernardoetal1994} and Jaynes \parencite*[\chap~8 and
+\sect~14.2]{jaynes1994_r2003}.
 
@@
 
-But such a translation does more: through a mathematical theorem, it leads
-to the automatic appearance of maximum-entropy distributions: \emph{if a
-  probability distribution for a repetitive phenomenon has some sufficient
-  statistics, then it is a mixture of distributions of maximum-entropy form
-  having those statistics as constraints, and vice versa}. See the end of
-this section for detailed references about this theorem.
+But such a translation does more: through a mathematical theorem \citep[and
+references therein]{andersen1970}[this theorem was first proved for
+continuous probability distributions
+by][]{koopman1936,pitman1936,darmois1935}[see][\sect~4.5.3 for further
+references]{bernardoetal1994}, it leads to the automatic appearance of
+maximum-entropy distributions: \emph{if a probability distribution for a
+  repetitive phenomenon has some sufficient statistics, then it is a
+  mixture of distributions of maximum-entropy form having those statistics
+  as constraints, and vice versa}. 
 % Thus, maximum-entropy distributions automatically appear (as
@@
 
-We close this section with some references. The notion of informational or
-predictive sufficiency, probably first presented by Kullback \amp\ Leibler
-\parencite*{kullbacketal1951}, is discussed in Bernardo \etal\
-\parencite*[\sect~4.5]{bernardoetal1994}, who explain its connections with
-Fisher's\parencite*{fisher1922} parametric sufficiency. Jaynes
-\parencite*[\chap~8 and \sect~14.2] shows the connections with information
-theory. The mathematical relation between sufficiency and the exponential
-family comes from the Pitman-Koopman
-theorem \citep{koopman1936,pitman1936,darmois1935} for continuous variables.
-
-\mynote{Short reasoned bibliography about sufficient statistics and its
-  relation to maximum-entropy?} blah blah \footnote{For more technical results and
-  connections with the notion of symmetry see \eg\
-  \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,lauritzen1982_r1988,lauritzen2007,cifarellietal1980,cifarellietal1981,diaconisetal1981,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,ayetal2015}.}
-This mathematical result comes from the Pitman-Koopman
-theorem\citep{cifarellietal1982,kullbacketal1951}[the
-notion goes back to][]{fisher1922} \citep{jaynes1982b}[\sect~4.5.4]{bernardoetal1994}
-\citep{koopman1936,pitman1936,darmois1935}[for later analyses and the
-discrete case
-see][]{hipp1974,andersen1970,denny1967,denny1972,fraser1963,barankinetal1963,barndorffnielsen1978_r2014}
+% We close this section with some references. The notion of informational or
+% predictive sufficiency, probably first presented by Kullback \amp\ Leibler
+% \parencite*{kullbacketal1951}, is discussed in Bernardo \etal\
+% \parencite*[\sect~4.5]{bernardoetal1994}, who explain its connections with
+% Fisher's\parencite*{fisher1922} parametric sufficiency. Jaynes
+% \parencite*[\chap~8 and \sect~14.2] shows the connections with information
+% theory. The mathematical relation between sufficiency and the exponential
+% family comes from the Pitman-Koopman
+% theorem \citep{koopman1936,pitman1936,darmois1935} for continuous variables.
+
+% \mynote{Short reasoned bibliography about sufficient statistics and its
+%   relation to maximum-entropy?} blah blah \footnote{For more technical results and
+%   connections with the notion of symmetry see \eg\
+%   \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,lauritzen1982_r1988,lauritzen2007,cifarellietal1980,cifarellietal1981,diaconisetal1981,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,ayetal2015}.}
+% This mathematical result comes from the Pitman-Koopman
+% theorem\citep{kullbacketal1951}[the
+% notion goes back to][]{fisher1922} \citep{jaynes1982b}[\sect~4.5.4]{bernardoetal1994}
+% \citep{koopman1936,pitman1936,darmois1935}[for later analyses and the
+% discrete case
+% see][]{andersen1970,%discrete
+% denny1967,denny1972}
 
@@
   replacement}
-\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
+\citep[\eg][\chap~3]{jaynes1994_r2003}, % [\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}
 ensues when our degree of belief that the $i$th unit (ball or neuron) will
@@
 they can be justified, is the fundamental, deep question of sampling
-theory, which we cannot discuss here \citep[see \eg\ the discussions,
-reviews, and references
-in][]{ericson1969,smith1976,kruskaletal1979c,kruskaletal1980}. It is of
-course possible to probabilistically assess, case by case, whether this
-equality assumption is appropriate. But we must also be aware that such
-assessment in turn rests on analogous assumptions at a higher level, and so
-on. The probability-calculus, just like the truth-calculus, cannot yield
-conclusions without premisses \citep[p.~182]{johnson1924}.
+theory, which we cannot discuss here \citep[see \eg\ the discussions and
+references
+in][]{ericson1969,smith1976}. % {kruskaletal1979c,kruskaletal1980}
+It is of course possible to probabilistically assess, case by case, whether
+this equality assumption is appropriate. But we must also be aware that
+such assessment in turn rests on analogous assumptions at a higher level,
+and so on. The probability-calculus, just like the truth-calculus, cannot
+yield conclusions without premisses \citep{hailperin2011}[p.~182]{johnson1924}.
 
@@
 kinematics, even in non-equilibrium
-\citep{jaynes1957,guntonetal1983,tikochinskyetal1984,grandy1988,berryetal1988,deroecketal2005_r2006,touchette2009}.
+\citep{jaynes1985b_r1996,%jaynes1957,%tikochinskyetal1984,
+grandy1988,%berryetal1988,
+deroecketal2005_r2006%,touchette2009
+}[for experiments \eg][]{guntonetal1983}.
 
@@
 typical of \enquote{drawing without replacement}
-\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}[\sects~2.1,
-3.2]{jeffreys1939_r1983}. In the following we leave $n$, $N$ implicit in
-the conditional. If we are uncertain about the number $\yA$, with belief
-$P(\yA)$, then by the theorem of total probability our belief about $\ya$
-is
+\citep[\eg][\chap~3]{jaynes1994_r2003}. %[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}[\sects~2.1, 3.2]{jeffreys1939_r1983}.
+In the following we leave $n$, $N$ implicit in the conditional. If we are
+uncertain about the number $\yA$, with belief $P(\yA)$, then by the theorem
+of total probability our belief about $\ya$ is
 \begin{equation}
@@
 and $\yA$, and using some combinatorial juggling
-\citep[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[appendix~A]{portamanaetal2015}{potts1953}, one
-can prove the equality~\eqref{eq:equal_m_factorialmoments} between the
-factorial moments of sample and larger population.
+\citep%[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}
+[appendix~A and references therein]{portamanaetal2015}, one can prove the
+equality~\eqref{eq:equal_m_factorialmoments} between the factorial moments
+of sample and larger population.
 
@@
 expectation constraints see Jaynes
-\citey{jaynes1963}[\chap~11]{jaynes1994_r2003}. More precisely we use the
-minimum relative-entropy method
-\citep{hobsonetal1973,csiszar1985}[\sect~5.2.2]{sivia1996_r2006} with
-respect to a reference distribution. For the solution of the extremization
-problem using Lagrangians and Lagrange multipliers see Mead \etal\
-\citey{meadetal1984} and the extensive texts by Fang \etal\
+\citey{jaynes1963}[\chap~11]{jaynes1994_r2003}, and Hobson \etal\
+\citey{hobsonetal1973} and Sivia \citey[\sect~5.2.2]{sivia1996_r2006} for
+the minimum-relative-entropy construction. For the solution of the
+extremization problem using Lagrangians and Lagrange multipliers see Mead
+\etal\ \citey{meadetal1984} and the extensive texts by Fang \etal\
 \citey{fangetal1997} and Boyd \etal\ \citey{boydetal2004_r2009}. For a
@@
 The first assumption, of infinite exchangeability
-\citep[\chap~4]{bernardoetal1994_r2000}{dawid2013} of the probability
-distribution $\pf[(\yA_{1},\dotsc, \yA_{T}) \| \yI]$ for the sequence of
-activities, implies the following mixture form by de~Finetti's theorem
-\citep{definetti1930,hewittetal1955}:
+\citep[\chap~4]{bernardoetal1994_r2000}{dawid2013%cifarellietal1982
+} of the probability distribution $\pf[(\yA_{1},\dotsc, \yA_{T}) \| \yI]$
+for the sequence of activities, implies the following mixture form by
+de~Finetti's theorem \citep{definetti1930,hewittetal1955}:
 \begin{equation}

### 2019-12-20 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-20T20:13:38+0100
+%% Last-Updated: 2019-12-21T00:40:05+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\nat}{\textrm{nat}}
+\newcommand*{\bit}{\textrm{bit}}
 \newcommand*{\hart}{\textrm{Hart}}
@@
 As mentioned in the Introduction, in the neurosciences the maximum-entropy
-method has also been used as a way of quantifying the
-\enquote{cooperativity} \citep[\eg][]{gersteinetal1985} or
-\enquote{interaction}
+method has also been used to quantify the \enquote{cooperativity}
+\citep[\eg][]{gersteinetal1985} or \enquote{interaction}
 \citep[\eg][]{martignonetal1995,schneidmanetal2006,shlensetal2006} or
@@
 activity. Most, if not all, such applications use the method \emph{at the
-  sample level}, aiming to find a distribution for the sample without
+  sample level}, that is, they find a distribution for the sample without
 considering that it comes from a larger population. In this section we
-discuss how our proposed application bears on this kind of quantification
-and compares with the application at the sample level. We also wish to
-highlight some methodological caveats.% For concreteness we use an
+discuss how our proposed application to a larger population bears on this
+kind of quantification and compares with the application at the sample
+level. %We also highlight some methodological caveats.% For concreteness we use an
 % example with specific statistics: first and second moments (equivalent to
@@
 
-As a first step we need to translate vague terms like
-\enquote{cooperativity}, \enquote{interaction}, and similar into a more
-precise, operational \citep[\cf][\chap~I]{bridgman1927_r1958} notion. Here
-we propose the notion of \emph{informational sufficiency}, because
+As a first step, terms like \enquote{cooperativity}, \enquote{interaction},
+and similar need to be translated into a more precise, operational
+\citep[\cf][\chap~I]{bridgman1927_r1958} notion. Here we propose the notion
+of \emph{informational sufficiency}, because
 \begin{enumerate*}[label=(\roman*)]
@@
 \end{enumerate*}
-Its meaning and the three points above will now be illustrated. For an
-extensive discussion of this notion, probably first suggested by Kullback
-\amp\ Leibler \parencite*{kullbacketal1951}, see Bernardo \etal\
+Its meaning will now be briefly explained. For an extensive discussion of
+this notion, which was probably first suggested by Kullback \amp\ Leibler
+\parencite*{kullbacketal1951}, see Bernardo \etal\
 \parencite*[\sect~4.5]{bernardoetal1994} and Jaynes \parencite*[\chap~8 and
@@
 Our belief distribution about the frequencies of the neuronal activities,
-or about the activity in a new time bin, is in principle conditional on all
-experimental data and statistics we have. This distribution, however, can
-be such that it remains unchanged if we discard part of the data or
-statistics -- for example, the third- and higher-order empirical moments. % --
+or about the activity in a new time bin, is conditional on all experimental
+data and statistics we have collected. This distribution, however, can be such that
+it remains unchanged if we discard part of the data or statistics -- for
+example, the third- and higher-order empirical moments. % --
 % but it does change if we discard more.
@@
 
-Now let us turn the informational sufficiency of a subset of statistics and
-the informational sufficiency of another subset into two hypotheses. We can
-then ask: which hypothesis is more probable, given the data? equivalently,
-what is the ratio of their probabilities? By translating our question about
-\enquote{cooperativity} into a hypothesis-testing problem about sufficiency,
-we can find a well-defined answer determined by the
-probability-calculus. % The ratio of
+Now let us turn the informational sufficiencies of two different subsets of
+statistics into two hypotheses. We can then ask: which hypothesis is more
+probable, given the data? equivalently, what is the ratio of their
+probabilities? By translating our question about \enquote{cooperativity}
+into a hypothesis-testing problem about sufficiency, we can find a
+well-defined answer determined by the probability-calculus. % The ratio of
 % the two probabilities gives us a natural measure of the
@@
 % about informational sufficiency.
-Its most important consequence is that, by Bayes's formula, the probability
-that some statistics are informationally sufficient can be approximated by
-finding the maximum-entropy distribution constrained by those statistics
-and then calculating the probability for the whole data under that
-distribution. We shall shortly illustrate this kind of calculation for the
-first data set.
+The most important consequence of this theorem is that, by Bayes's formula,
+the probability that some statistics are informationally sufficient can
+approximately be found by first building the maximum-entropy distribution
+constrained by those statistics, and then calculating the probability for
+the whole data under that distribution. We shall shortly illustrate this
+procedure with the data sets of \sect~\ref{sec:application}.
 
@@
 
-The theorem above makes it very natural to reinterpret studies about
-cooperativity, interaction, and similar, based on maximum-entropy, as
-studies about informational sufficiency. An important fact must then be
-taken into account: \emph{informational sufficiency is not preserved under
-  sampling}. If a probability distribution has some sufficient statistics,
-then its sample marginals \emph{cannot} have the same sufficient
-statistics, and vice versa \citep[\sect~3.1]{portamanaetal2015}; except for
-trivial cases such as uniform probability
-distributions. % This generally holds also among
+Owing to the theorem above, studies about cooperativity (and similar ideas)
+that use maximum-entropy can be naturally reinterpreted as studies about
+informational sufficiency. An important fact must then be taken into
+account: \emph{informational sufficiency is not preserved under sampling}.
+If a probability distribution has some sufficient statistics, then its
+sample marginals \emph{cannot} have the same sufficient statistics, and
+vice versa \citep[\sect~3.1]{portamanaetal2015}; except for trivial cases
+(\eg\ uniform probability distributions). % This generally holds also among
 % different marginals: if a statistics is sufficient for some marginal, then
@@
 This fact affects the interpretation of the results obtained from
-maximum-entropy distributions. For example, if means and pairwise correlations
-seem informationally sufficient for a particular sample from a brain area,
-then they are probably not sufficient for the larger population of neurons
-constituting that area, and vice versa. If we are interested in the
+maximum-entropy distributions. For example, if means and pairwise
+correlations seem informationally sufficient for a sample from a brain
+area, then they are probably not sufficient for the larger population of
+neurons constituting that area, and vice versa. If we are interested in the
 cooperativity of the neurons in a brain area, it is therefore unreliable to
@@
 thereof. % \citep[contrast this with some literature, \eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}
-The model presented here avoids this problem because the maximum-entropy
-method is applied directly to the larger population, not to the sample
-alone.
+The model presented here avoids this problem because it targets the larger
+population, not the sample alone.
 
@@
 
-We compare the informational sufficiency between of a set of moments, say
+We can compare the informational sufficiency between of a set of moments, say
 $M''\defd\set{1,\dotsc,m''}$, with respect to another, say
@@
 \item From each maximum-entropy distribution $\px(\yA \| M,N)$ for the
-  larger population, built from each set of constraints $M=M',M''$,
-  calculate the marginal distribution for the sample by
-  \eqn~\eqref{eq:conditional_hypergeometric_t}:
+  larger population, \eqn~\eqref{eq:app_maxent_pop_first}, built from each
+  set of constraints $M=M',M''$, calculate the marginal distribution for
+  the sample by \eqn~\eqref{eq:conditional_hypergeometric_t}:
 \begin{equation}
@@
 
-% NUMERICAL RESULTS
-%
-% relentropies of sample2mom, sample4mom, sample5mom, marg2mom, marg4mom, marg5mom
-% -322.72,-42.8005,-39.4917,-85.7626,-4.23725,-4.20059 nats =
-% -1401.56,-185.88,-171.51,-372.462,-18.4021,-18.2429 dB
-%
-% difference in relentropies between 4 mom and 2 mom:
-% sample level: 279.92 nats = 1215.68 dB
-% marg level: 81.525 nats =  354.06 dB
-% difference in relentropies between 5 mom and 4 mom:
-% sample level: 3.30882 nats = 14.37 dB
-% marg level: 0.0366623 nats = 0.159223 dB
-% correspond to prob ratios of
-% sample: 3.69524\times 10^{121}, 27.3529
-% marg: 2.5469\times 10^{35}, 1.03734
-% 
-% comparison N vs n:
-% 2 mom: 236.958 nats = 1029.09 dB, 8.11758\times 10^{102}
-% 4 mom: 38.5632 nats = 167.478 dB, 5.59496*10^16
-% 5 mom: 35.2911 nats = 153.267 dB, 2.12185*10^15
-%
-% comparison_ME_sample_full_levels.nb
-%
-
 \bigskip
 
-Let us illustrate the ideas, the remarks, and the quantity defined so far
-with our first data set. Consider for example three sets of constraints
-$M_{2}$, $M_{4}$, $M_{5}$, consisting of the first two, four, five moments;
-and a larger population of size $N=10\,000$. Figure~\ref{fig:plots_2_4mom}
-shows the resulting distributions for the two data sets.
+Let us illustrate the quantity just defined and the previous discussion
+about sufficiency and sampling by using our data sets. Consider for example
+three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$, consisting of the
+first two, four, five moments; and a larger population of size $N=10\,000$.
+Figure~\ref{fig:plots_2_4mom} shows the resulting distributions for the two
+data sets.
 \begin{figure}[!p]
@@
 that two moments are less likely to be informationally sufficient than
-four; but four should almost as likely as five. We shall see that the
-measure~\eqref{eq:measure_suffic} reflects what we immediately see.
+four; but four should be almost as likely as five. We shall see that the
+measure~\eqref{eq:measure_suffic} reflects this swift visual judgement.
 
 Compare these plots with those obtained applying the model at the sample
-level (equivalent to $N=n$). With the $M_{2}$ and $M_{4}$ constraints we
-obtain the distributions partly shown in
-\fig~\ref{fig:comparisons_marginals} as \textcolor{myred}{red empty
-  circles} for two moments, and \textcolor{myblue}{blue empty squares} for
-four moments. The measured frequencies (black triangles) are also shown.
-The two distributions can only be distinguished on a logarithmic scale, and
-only in their tails. It would be difficult to visually assess which is more
-likely to be informationally sufficient.
+level (equivalent to $N=n$) to the first data set. With the $M_{2}$ and
+$M_{4}$ constraints we obtain the distributions partly shown in
+\fig~\ref{fig:comparisons_marginals} for two (\textcolor{myred}{red empty
+  circles}) and four (\textcolor{myblue}{blue empty squares}) moments. The
+measured frequencies (black triangles) are also shown. The two
+distributions can only be distinguished on a logarithmic scale, and only in
+their tails. It would be difficult to visually assess which is more likely
+to be informationally sufficient.
 
@@
 
+% NUMERICAL RESULTS
+%
+% relentropies of sample2mom, sample4mom, sample5mom, marg2mom, marg4mom, marg5mom
+% -322.72,-42.8005,-39.4917,-85.7626,-4.23725,-4.20059 nats =
+% -1401.56,-185.88,-171.51,-372.462,-18.4021,-18.2429 dB
+%
+% difference in relentropies between 4 mom and 2 mom:
+%** sample level: 279.92 nats = 403.839 bit = 1215.68 dB
+%** marg level: 81.525 nats =  117.616 bit = 354.06 dB
+% difference in relentropies between 5 mom and 4 mom:
+%** sample level: 3.30882 nats = 4.77362 bit = 14.37 dB
+%** marg level: 0.0366623 nats = 0.0543836 bit = 0.159223 dB
+% correspond to prob ratios of
+% sample: 3.69524\times 10^{121}, 27.3529
+% marg: 2.5469\times 10^{35}, 1.03734
+% 
+% comparison N vs n:
+% 2 mom: 236.958 nats = 341.858 bit = 1029.09 dB, 8.11758\times 10^{102}
+% 4 mom: 38.5632 nats = 55.635 bit = 167.478 dB, 5.59496*10^16
+% 5 mom: 35.2911 nats = 50.9157 bit = 153.267 dB, 2.12185*10^15
+%
+% comparison_ME_sample_full_levels.nb
+%
 Calculation of the weight of evidence~\eqref{eq:measure_suffic} for the
@@
   \begin{split}
-  \ydi_{N}(M_{4},M_{2}) &= 81\;\nat = 35\;\hart,
-\\    \ydi_{N}(M_{5},M_{4}) &= 0.037\;\nat = 0.016\;\hart,
+  \ydi_{N}(M_{4},M_{2}) &= 118\;\bit \equiv 35\;\hart,
+\\    \ydi_{N}(M_{5},M_{4}) &= 0.054\;\bit \equiv 0.016\;\hart,
   \end{split}
@@
   \begin{split}
-  \ydi_{n}(M_{4},M_{2}) &= 280\;\nat = 1\,220\;\hart,
-\\    \ydi_{n}(M_{5},M_{4}) &= 3.3\;\nat = 1.4\;\hart.
+  \ydi_{n}(M_{4},M_{2}) &= 400\;\bit = 120\;\hart,
+\\    \ydi_{n}(M_{5},M_{4}) &= 4.8\;\bit = 1.4\;\hart.
 \end{split}
 \end{equation}
-So the application at the sample level says that the data are 1\,220 orders
-of magnitude more probable conditional on four moments than conditional on
-two, and $10^{1.4}=25$ times more probable conditional on five moments than
-on four moments. These results are different from the ones obtained from
-the larger population, and they would be hard to deduce visually from a
-look at the distributions of \fig~\ref{fig:comparisons_marginals}.
+So the application at the sample level says that the data are 120 orders of
+magnitude more probable conditional on $M_{4}$ than on $M_{2}$, and
+$10^{1.4}=25$ times more probable conditional on $M_{5}$ than on $M_{4}$.
+These results are different from those of the application at the larger
+population, and they would be hard to deduce visually from the
+distributions of \fig~\ref{fig:comparisons_marginals}.
 
@@
 
-It is in fact possible to use the measure~\eqref{eq:measure_suffic} and its
+It is possible to use the measure~\eqref{eq:measure_suffic} and its
 probabilistic interpretation~\eqref{eq:bayes_factor} to compare the
@@
   \label{eq:diff_suff_n}
-  \ydi[(M_{4}, N), (M_{4}, n)] = 39\;\nat = 17\;\hart.
+  \ydi[(M_{4}, N), (M_{4}, n)] = 56\;\bit = 17\;\hart.
 \end{equation}
@@
 The calculations and conclusions presented in the preceding sections depend
-on the size $N$ of the larger population. The larger population can't be
+on the size $N$ of the larger population. The larger population cannot be
 the whole brain, of course. How large should or can $N$ be in our formulae?
@@
 \citep[\eg][\chap~3]{jaynes1994_r2003}, % [\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}
-ensues when our degree of belief that the $i$th unit (ball or neuron) will
-be drawn has the same value for all $i \in \set{1,\dotsc,N}$. Thus,
-consider the pool of all neurons which we believe -- with equal degree for
-each neuron -- could have been recorded. $N$ is the size of that pool. $N$
-therefore depends on factors such as: the shape, dimensions, and technical
-specifications of the recording probe; the inaccuracy in the insertion of
-the probe, leading for example to slightly different insertion points or
-angles; the density of neurons around the probe. But it also depends on the
-homogeneity of the brain region where the recording was made: if we believe
-that it doesn't matter whether the probe had been inserted in some other
-point of the same brain region, then
-formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
-therefore only be assessed case by case.
+ensues when we believe, with the same degree, that any of the $N$ units
+(balls or neurons) could be drawn or sampled. Thus, consider the pool of
+all neurons which we believe -- with equal degree for each neuron -- could
+have been recorded. Then $N$ is the size of that pool. $N$ therefore depends on
+factors such as: the shape, dimensions, and technical specifications of the
+recording probe; the inaccuracy in the insertion of the probe, leading for
+example to slightly different insertion points or angles; the density of
+neurons around the probe. But it also depends on the homogeneity of the
+brain region where the recording was made: if we believe that it doesn't
+matter whether the probe had been inserted in some other point of the same
+brain region, then formula~\eqref{eq:conditional_hypergeometric} is
+appropriate. $N$ can therefore only be assessed case by case.
 

### 2019-12-20 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-21T00:40:05+0100
+%% Last-Updated: 2019-12-21T00:45:03+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}\labelsep 0pt%
-***
+\mynote{to be written}
 % \noindent \mynote{Abstract must be rewritten once paper is ready}This
@@
 
+\medskip
+
 If our uncertainty about $N$ spans several orders of magnitude we can
@@
 
-In this way we can make inferences -- for example about the sufficiency of
-a set of moments, or about the marginal sample distribution -- that take
-into account our uncertainty about $N$. We must make sure to avoid
-circularities, though: for example, we can't assume a set of moments $M$ to
-be sufficient and assess our uncertainty about $N$ conditional on it, and
-then use this uncertainty to assess the sufficiency of $M$. But we can
-approximately assess the sufficiency of a subset of moments $M' \subset M$
-taking into account the uncertainty of $N$ conditional on $M$.
+In this way we can make inferences % -- for example about the sufficiency of
+% a set of moments, or about the marginal sample distribution --
+that take into account our uncertainty about $N$. We must make sure to
+avoid circularities, though: for example, we cannot assume a set of moments
+$M$ to be sufficient and assess our uncertainty about $N$ conditional on
+it, and then use this uncertainty to assess the sufficiency of $M$. But we
+can approximately assess the sufficiency of a subset of moments
+$M' \subset M$ taking into account the uncertainty of $N$ conditional on
+$M$.
 

### 2019-12-21 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-21T00:45:03+0100
+%% Last-Updated: 2019-12-21T11:22:03+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%\paragraphfootnotes\setlength{\footmarkwidth}{0em}
+\paragraphfootnotes\setlength{\footmarkwidth}{1em}
 %\twocolumnfootnotes
@@
   \begin{split}
-  \ydi_{n}(M_{4},M_{2}) &= 400\;\bit = 120\;\hart,
-\\    \ydi_{n}(M_{5},M_{4}) &= 4.8\;\bit = 1.4\;\hart.
+  \ydi_{n}(M_{4},M_{2}) &= 400\;\bit \equiv 120\;\hart,
+\\    \ydi_{n}(M_{5},M_{4}) &= 4.8\;\bit \equiv 1.4\;\hart.
 \end{split}
@@
   \label{eq:diff_suff_n}
-  \ydi[(M_{4}, N), (M_{4}, n)] = 56\;\bit = 17\;\hart.
+  \ydi[(M_{4}, N), (M_{4}, n)] = 56\;\bit \equiv 17\;\hart.
 \end{equation}
@@
 approximation all in all. These become clear within a full-fledged
-probabilistic approach. We summarize them here and give a more mathematical
-sketch in appendix~\ref{sec:inference_prob}.
+probabilistic approach. We summarize them here, leaving a more mathematical
+sketch to appendix~\ref{sec:inference_prob}.
 
 The first and most important assumption, typical of this kind of
-maximum-entropy application in neuroscience, can be expressed in two
+maximum-entropy application in the neurosciences, can be expressed in two
 equivalent ways:

### 2019-12-21 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-21T11:22:03+0100
+%% Last-Updated: 2019-12-21T21:36:19+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Compare these plots with those obtained applying the model at the sample
-level (equivalent to $N=n$) to the first data set. With the $M_{2}$ and
-$M_{4}$ constraints we obtain the distributions partly shown in
-\fig~\ref{fig:comparisons_marginals} for two (\textcolor{myred}{red empty
-  circles}) and four (\textcolor{myblue}{blue empty squares}) moments. The
-measured frequencies (black triangles) are also shown. The two
-distributions can only be distinguished on a logarithmic scale, and only in
-their tails. It would be difficult to visually assess which is more likely
-to be informationally sufficient.
+level (equivalent to $N=n$) to the first data set, shown in
+\fig~\ref{fig:comparisons_marginals}. The distributions based on two
+(\textcolor{myred}{red circles}), four (\textcolor{myblue}{blue empty
+  squares}), and five (\textcolor{myyellow}{yellow lozenges}) moments can
+only be distinguished on a logarithmic scale, and only in their tails. The
+measured frequencies (black triangles) are also shown. Two moments seem
+slightly less likely to be sufficient than four; and four and five moment
+seem about equally likely. But it is difficult to visually make these
+comparisons.
 
@@
 % \\[3em]%
-\includegraphics[width=0.95\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
+\includegraphics[width=0.95\linewidth]{_scripts/trunct2_comparison_marginals.pdf}%
 \caption{Marginals for the normalized total activity of the sample,
@@
 probabilistic interpretation~\eqref{eq:bayes_factor} to compare the
-probability of the data $\yf$ conditional on the hypothesis $(M_{4},N)$ of
+probability of the data $\yf$ conditional on the hypothesis $(M_{5},N)$ of
 four-moment sufficiency at the larger-population level, $N=10\,000$, versus
-the hypothesis $(M_{4}, n)$ of four-moment sufficiency at the sample level,
+the hypothesis $(M_{5}, n)$ of four-moment sufficiency at the sample level,
 $N=n=65$. We obtain
@@
   \label{eq:diff_suff_n}
-  \ydi[(M_{4}, N), (M_{4}, n)] = 56\;\bit \equiv 17\;\hart.
-\end{equation}
-% that is, the data are 17 orders of magnitude more probable under the first
+  \ydi[(M_{5}, N), (M_{5}, n)] = 51\;\bit \equiv 15\;\hart.
+\end{equation}
+% that is, the data are 15 orders of magnitude more probable under the first
 % hypothesis than under the second. 

### 2019-12-21 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-21T21:36:19+0100
+%% Last-Updated: 2019-12-21T23:54:35+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   $N$ (obtaining densities) and the activity divided by $N$ (obtaining a
-  population average). Note how the distributions for large $N$ have very
-  different features from the measured frequency distribution, and how
-  their peaks become more pronounced as $N$ increases.}
+  population average). The distributions for large $N$ have very different
+  features from the measured frequency distribution, and how their peaks
+  become more pronounced as $N$ increases.}
 \label{fig:plots_5mom}
@@
 \includegraphics[width=\linewidth]{_scripts/newpaper3_gridvsnogrid.pdf}%
-\\[2em]%
+\\[1em]%
 \subcaption[]{\label{fig:plots_cellsubsets_b}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper3_gridconv.pdf}%
-\\[2em]%
+\\[1em]%
 \caption{\subcaptionref{fig:plots_cellsubsets_a} Maximum-entropy
   distributions for larger subpopulations or grid and non-grid cells.
-  \subcaptionref{fig:plots_cellsubsets_b} Convolution of the two
-  subpopulation distributions; it would be equal to the total distribution
-  if the two distributions were independent.}
+  \subcaptionref{fig:plots_cellsubsets_b} The convolution of the two
+  subpopulation distributions is very different from the distribution for
+  all cells; this means that the two subpopulation distributions are not
+  independent.}
 \label{fig:plots_cellsubsets}
@@
 \includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_threemom_distributions10000_priorc.pdf}%
-\\[2em]%
+\\[1em]%
 \subcaption[]{\label{fig:plots_2_4mom_b}}%  
 \includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_threemom_distributions10000_prior2.pdf}%
-\\[2em]%
-\caption{Maximum-entropy distributions for the normalized total activity of
-  the larger population, assuming two different sets of constraining moments
-  and a larger population size $N=10\,000$.}
+\\[1em]%
+\caption{Maximum-entropy distributions for the total activity of the larger
+  population, based on the first two, four, and five moments, and a
+  larger-population size $N=10\,000$. The two- and four-moment
+  distributions are very different, suggesting that two moments are not
+  informationally sufficient. The four- and five-moment distributions are
+  almost indistinguishable, suggesting that four and five moments are
+  equally likely to be sufficient.}
 \label{fig:plots_2_4mom}
@@
   \begin{split}
-  \ydi_{N}(M_{4},M_{2}) &= 118\;\bit \equiv 35\;\hart,
-\\    \ydi_{N}(M_{5},M_{4}) &= 0.054\;\bit \equiv 0.016\;\hart,
+  \ydi_{N}(M_{4},M_{2}) &= 118\;\bit \equiv 35.4\;\hart,
+\\    \ydi_{N}(M_{5},M_{4}) &= 0.0544\;\bit \equiv 0.0159\;\hart,
   \end{split}
@@
   \begin{split}
-  \ydi_{n}(M_{4},M_{2}) &= 400\;\bit \equiv 120\;\hart,
-\\    \ydi_{n}(M_{5},M_{4}) &= 4.8\;\bit \equiv 1.4\;\hart.
+  \ydi_{n}(M_{4},M_{2}) &= 404\;\bit \equiv 122\;\hart,
+\\    \ydi_{n}(M_{5},M_{4}) &= 4.77\;\bit \equiv 1.44\;\hart.
 \end{split}
@@
 \includegraphics[width=0.95\linewidth]{_scripts/trunct2_comparison_marginals.pdf}%
-\caption{Marginals for the normalized total activity of the sample,
-  obtained from the maximum-entropy distributions with $N=10\,000$ from
-  \fig~\ref{fig:plots_2_4mom_a}, and from maximum-entropy distributions
-  calculated at the sample level, that is with $N=n=65$, with two and four
-  moments in either case.}
+\caption{Maximum-entropy distributions constructed at the sample level
+  ($N=n$), based on the first two, four, and five moments; and sample
+  marginal of the distribution constructed with five moments for a larger
+  population with $N=10\,000$ (dotted yellow in
+  \subcaptionref{fig:plots_2_4mom_a}). The distributions at the sample
+  level are all very similar, only distinguishable on a logarithmic scale.
+  The five-moment distribution constructed at the larger-population level
+  fits the frequency data better than the one at the sample level.}
 \label{fig:comparisons_marginals}
@@
 probability of the data $\yf$ conditional on the hypothesis $(M_{5},N)$ of
-four-moment sufficiency at the larger-population level, $N=10\,000$, versus
-the hypothesis $(M_{5}, n)$ of four-moment sufficiency at the sample level,
+five-moment sufficiency at the larger-population level, $N=10\,000$, versus
+the hypothesis $(M_{5}, n)$ of five-moment sufficiency at the sample level,
 $N=n=65$. We obtain
@@
   \label{eq:diff_suff_n}
-  \ydi[(M_{5}, N), (M_{5}, n)] = 51\;\bit \equiv 15\;\hart.
+  \ydi[(M_{5}, N), (M_{5}, n)] = 50.9\;\bit \equiv 15.3\;\hart.
 \end{equation}
@@
 This result is also shown in
-\fig~\ref{fig:comparisons_marginals}. We see that the four-moment
+\fig~\ref{fig:comparisons_marginals}. We see that the five-moment
 distribution constructed at the larger-population level
-(\textcolor{myblue}{blue filled squares}) fits the measured frequency
+(\textcolor{myyellow}{yellow stars}) fits the measured frequency
 distribution (black triangles) more closely than the corresponding
-sample-level distribution (\textcolor{myblue}{blue empty squares}).
+sample-level distribution (\textcolor{myyellow}{yellow lozenges}).
 

### 2019-12-21 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-21T23:54:35+0100
+%% Last-Updated: 2019-12-22T00:23:26+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % \\[3em]%
-\includegraphics[width=0.95\linewidth]{_scripts/trunct2_comparison_marginals.pdf}%
-\caption{Maximum-entropy distributions constructed at the sample level
-  ($N=n$), based on the first two, four, and five moments; and sample
-  marginal of the distribution constructed with five moments for a larger
-  population with $N=10\,000$ (dotted yellow in
-  \subcaptionref{fig:plots_2_4mom_a}). The distributions at the sample
-  level are all very similar, only distinguishable on a logarithmic scale.
-  The five-moment distribution constructed at the larger-population level
-  fits the frequency data better than the one at the sample level.}
+\includegraphics[width=0.92\linewidth]{_scripts/trunct2_comparison_marginals.pdf}%
+\caption{Maximum-entropy distributions constructed at the sample level,
+  using the first two, four, and five moments; and sample marginal of the
+  distribution constructed with five moments for a larger population with
+  $N=10\,000$ (dotted yellow in \ref{fig:plots_2_4mom_a}). The
+  distributions are multiplied by $T$, obtaining absolute frequencies; note
+  that activities with frequency lower than 1 cannot be observed in $T$
+  bins. The distributions at the sample level are only distinguishable on a
+  logarithmic scale. The five-moment distribution constructed at the
+  larger-population level fits the frequency data better than the one at
+  the sample level.}
 \label{fig:comparisons_marginals}

### 2019-12-22 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-22T00:23:26+0100
+%% Last-Updated: 2019-12-22T15:04:32+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The calculations and conclusions presented in the preceding sections depend
-on the size $N$ of the larger population. The larger population cannot be
-the whole brain, of course. How large should or can $N$ be in our formulae?
+The calculations presented in the preceding sections depend on the size $N$
+of the larger population, which of course cannot be the whole brain. How
+large should or can $N$ be in our formulae?
 
@@
 (balls or neurons) could be drawn or sampled. Thus, consider the pool of
-all neurons which we believe -- with equal degree for each neuron -- could
-have been recorded. Then $N$ is the size of that pool. $N$ therefore depends on
-factors such as: the shape, dimensions, and technical specifications of the
-recording probe; the inaccuracy in the insertion of the probe, leading for
-example to slightly different insertion points or angles; the density of
-neurons around the probe. But it also depends on the homogeneity of the
-brain region where the recording was made: if we believe that it doesn't
-matter whether the probe had been inserted in some other point of the same
-brain region, then formula~\eqref{eq:conditional_hypergeometric} is
-appropriate. $N$ can therefore only be assessed case by case.
-
-Whether such equal beliefs are justified, or for which sampling procedures
+all neurons which we believe -- with equal degree of belief for each neuron
+-- could have been recorded. Then $N$ is the size of that pool. $N$
+therefore depends on factors such as: the shape, dimensions, and technical
+specifications of the recording probe; the inaccuracy in the insertion of
+the probe, leading for example to slightly different insertion points or
+angles; the density of neurons around the probe. But it also depends on the
+homogeneity of the brain region where the recording is made: if we believe
+that it does not matter whether the probe had been inserted in some other
+point of the same brain region, then
+formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
+therefore only be assessed case by case.
+
+Whether such symmetric beliefs are justified, or for which sampling procedures
 they can be justified, is the fundamental, deep question of sampling
@@
 It is of course possible to probabilistically assess, case by case, whether
-this equality assumption is appropriate. But we must also be aware that
-such assessment in turn rests on analogous assumptions at a higher level,
+this symmetry assumption is appropriate. But we must also be aware that
+such assessment in turn rests on analogous symmetry assumptions at a higher level,
 and so on. The probability-calculus, just like the truth-calculus, cannot
@@
 
-Luckily the plots of \fig~\ref{fig:plots_5mom} suggest that the order of
-magnitude of $N$ ought to be enough for qualitative inferences. As
+The plots of \fig~\ref{fig:plots_5mom} nevertheless suggest that the order
+of magnitude of $N$ ought to be enough for qualitative inferences. As
 discussed in the next section, even if the exact number $N$ were known, the
@@
   \label{eq:prob_data_from_N}
-  \pf(\yf \| M, N) = \binom{T}{T\yf}\;\prod_{a}\pxx(\ya \| M, N)^{T\,\yff_{a}}
+  \pf(\yf \| M, N) \approx \binom{T}{T\yf}\;\prod_{a}\pxx(\ya \| M, N)^{T\,\yff_{a}}
   \approx \exp\bigl\{  -T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
@@
 Here is an illustrative example with our first data set. Suppose our
-uncertainty spans slightly more than an order of magnitude, from $N=1\,000$
+uncertainty spans slightly more than one order of magnitude, from $N=1\,000$
 to $N=20\,000$. Divide this range roughly into thirds of order of
@@
 that take into account our uncertainty about $N$. We must make sure to
-avoid circularities, though: for example, we cannot assume a set of moments
+avoid circularities: for example, we cannot assume a set of moments
 $M$ to be sufficient and assess our uncertainty about $N$ conditional on
@@
 approximations which define it: three main assumptions and two main
-approximation all in all. These become clear within a full-fledged
+approximations all in all. They become clear within a full-fledged
 probabilistic approach. We summarize them here, leaving a more mathematical
@@
 would not make sense without this assumption. Its mathematical consequence
-is to relate the probability for the frequency distribution $\yF$ during
-the recording to that of the \emph{long-run} frequency distribution for the
-activities of the larger population. Note that the maximum-entropy method
-in statistical physics is meant to be used not with this assumption, but
-with averages across \emph{repetitions} of the same experiment, because its
-constraints represent macroscopically \emph{reproducible} quantities and
-kinematics, even in non-equilibrium
+is to relate the probability for the frequency distribution $\yF$ for the
+activity of the larger population during the recording, to the probability
+for the \emph{long-run} frequency distribution. In statistical physics, as
+a comparison, the maximum-entropy method is instead used with the
+assumption of permutation-invariance across \emph{repetitions} of the same
+experiment, because its constraints represent macroscopically
+\emph{reproducible} quantities and kinematics, even in non-equilibrium
 \citep{jaynes1985b_r1996,%jaynes1957,%tikochinskyetal1984,
@@
   deroecketal2005_r2006%,touchette2009
-}[for experiments \eg][]{guntonetal1983}.
+}[for experiments \eg][]{guntonetal1983}. The kinetic-Ising models
+mentioned in \sect~\ref{sec:intro} are closer to this latter approach.
 
 The second assumption, typical of maximum-entropy in general, concerns the
-form of the probability distribution for the long-run frequencies ensuing
-from the first assumption. It is taken to be either an \enquote{entropic
-  prior} \citep{skillingetal1984,rodriguez1991,neumann2007}, which takes
-into account the multiplicity of long-run frequencies, or a prior for a
-model with sufficient statistics, which expresses the sufficiency of the
-moments under study \citep{portamana2017}. This assumption is crucial for
-deriving the maximum-entropy based on the Shannon entropy. A different
-density leads to alternative maximum-entropy methods, for example based on
-the Burg entropy \citep{burg1975} if a Dirichlet density is used
+probability distribution for the long-run frequencies ensuing from the
+first assumption \citep[and references therein]{portamana2017}. It is taken
+to be either an \enquote{entropic prior}
+\citep{skillingetal1984,rodriguez1991,neumann2007}, which takes into
+account the multiplicity of long-run frequencies, or a prior for a model
+with sufficient statistics, which expresses the sufficiency of the moments
+under study, according to the theorem discussed in
+\sect~\ref{sec:marginalization}. This assumption is crucial for the
+maximum-entropy method based on the Shannon entropy. Different prior
+densities lead to alternative maximum-entropy methods; for example one
+based on the Burg entropy \citep{burg1975} if a Dirichlet density is used
 \citep{jaynes1986d_r1996}. The entropic prior depends on a reference
 distribution and on a coefficient $L$ which determines the sharpness of the
-prior's peak. The computations of the preceding sections were insensitive
-to the reference distribution in their significant digits. We used three
+prior's peak. The computations of \sects~\ref{sec:application}
+and~\ref{sec:marginalization} were insensitive to the choice of the
+reference distribution in their significant digits. We used three
 biologically reasonable alternatives: a flat distribution, one linearly
@@
 is that $n$ neurons are sampled anew at each time bin. This assumption
-allows us to factorize the joint probability of the two sequences of
-activities and apply the sampling
-relation~\eqref{eq:conditional_hypergeometric} of
+allows us to factorize the joint probability of the sequence of activities
+and apply the sampling relation~\eqref{eq:conditional_hypergeometric} of
 appendix~\ref{sec:derivation_maxent_fullnet} at each bin.
@@
 frequencies of the sample activity $\ya$ given $\yA$. The goodness of this
-approximation decreases as the ratio between the number of possible joint
-states and the number of bins -- roughly $nN/T$ -- increases, and also as
-observed quantities approach their mathematical bounds, such as some
+approximation decreases as the ratio -- roughly $nN/T$ -- between the
+number of possible joint states and the number of bins increases, and also
+as observed quantities approach their mathematical bounds, for example some
 frequencies $\yff_{\ya}=0$. Without these approximations it is still
@@
 make sense as reference hypotheses for judging the informational relevance
-of correlations of some order, as in \sect~\ref{sec:marginalization}, to be
-compared with hypotheses about more realistic time dependences and
+of correlations of some order, as discussed in
+\sect~\ref{sec:marginalization}, to be compared with hypotheses about more
+realistic time dependences and
 clustering. %They may turn out to approximate well more realistic assumptions.
@@
 distributions of sample frequencies, if too many moments are used. For
-example take $n=2$ with frequency distribution $\yf=(0.5,0,0.5)$ for the
+example, take $n=2$ with frequency distribution $\yf=(0.5,0,0.5)$ for the
 activities $\ya=0,1,2$; that is, the sample is silent half of the time and
@@
 zero probability of sampling one active neuron, then the larger population
-cannot have any active neurons, and there must be zero probability of
-sampling two active ones. But this reasoning holds only for each single
-time bin, not for the averages over time bins. The $T=\infty$ approximation
-identifies the two instead. Outside this approximation there are no
-contradictions because the conditional frequencies and conditional
-probabilities are kept distinct.
-
-The two infinity approximations are at the limit of their validity in our
-examples. For this reason we interpret the distributions as \emph{most
-  probable frequency distributions}, as mentioned at the end of
-\sect~\ref{sec:model}. The space of possible frequency distributions for
-the larger population has $N$ dimensions, however, and probability
-distributions in high dimensions have counter-intuitive properties. For
-example, the mode or mean can have \emph{atypical} features when compared
-with the majority of other probable points of the probability space. The
-question then is \emph{which features of the maximum-entropy frequency
-  distribution are typical of the majority of plausible frequency
-  distributions?} Preliminary studies using a full probability calculation
-(to be published separately) show that several semi-quantitative features
-of the maximum-entropy distribution are indeed typical. For the first data
-set, for example, most frequency distributions have two high-frequency
-regimes of activity, as in \fig~\ref{fig:plots_5mom_a}, with roughly the
-same heights and almost the same locations, within a few percentages of
-population-averaged activity. Each of high-frequency regimes, though, may
-consist of two or three very close modes rather than one. The same
-preliminary studies also indicate that these typical features remain with
-prior choices in the second assumption above, for example using a Dirichlet
-prior instead of an entropic one.
+cannot have any active neurons at all, and a fortiori there must be zero
+probability of sampling two active ones. But this reasoning holds only for
+each single time bin, not for the averages over time bins. The $T=\infty$
+approximation identifies the two instead. Outside of this approximation
+there are no contradictions because the conditional frequencies and
+conditional probabilities are kept distinct.
+
+The two infinity approximations are at the boundary of their validity in
+our examples. For this reason we interpret the maximum-entropy distribution
+as the \emph{most probable frequency distributions}, as mentioned at the
+end of \sect~\ref{sec:model}. An important question, then, is \emph{which
+  features of the maximum-entropy frequency distribution are typical of the
+  majority of plausible frequency distributions?} Especially for features
+such as number, position, height of modes, or tail behaviour. Preliminary
+studies using a full probability calculation (to be published separately)
+show that several semi-quantitative features of the maximum-entropy
+distribution are indeed typical. For the first data set, for example, most
+frequency distributions have two high-frequency regimes of activity, as in
+\fig~\ref{fig:plots_5mom_a}, with roughly the same heights and almost the
+same locations, within a few percentages of population-averaged activity.
+Each of the two high-frequency regimes, though, may consist of two or three
+very close modes rather than one. The same preliminary studies also
+indicate that these typical features are robust for different prior choices
+in the second assumption above, for example using a Dirichlet prior instead
+of an entropic one.
 
@@
 favourably used to get an idea of what a fully probabilistic calculation
-could yield, but with a far lower computational cost. For example, for
-$N=1\,000$ the maximum-entropy distribution can be calculated in ten
-minutes on a laptop, whereas the full probability calculation takes several
-days on a high-performance computing cluster.
+could yield -- but with a far lower computational cost than the latter. For
+example, for $N=1\,000$ the maximum-entropy distribution can be calculated
+in ten minutes on a laptop, whereas the full probability calculation takes
+several days on a high-performance computing cluster.
 

### 2019-12-22 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-22T15:04:32+0100
+%% Last-Updated: 2019-12-22T15:09:24+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 frequencies of the total activity of the larger population, but such
-limited inferential scope makes the model computationally very cheap.
-
-Yet, and most important, this model leads to inferences that are not at all
+limited inferential scope makes the model computationally very cheap and,
+and most important, it still leads to inferences that are not at all
 trivial and not easily discernible by looking at the sample. We showed this

### 2019-12-26 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-22T15:09:24+0100
+%% Last-Updated: 2019-12-26T15:35:20+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\arxiveprint}[1]{%\global\def\arxivp{\arxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
-\texttt{\urlalt{https://arxiv.org/abs/#1}{arXiv:\hspace{0pt}#1}}%
+\texttt{arXiv:\urlalt{https://arxiv.org/abs/#1}{#1}}%
 %\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
@@
 \newcommand*{\mparceprint}[1]{%\global\def\mparcp{\mparcsi}%\citeauthor{0mparccite}\addtocategory{ifarchcit}{0mparccite}%eprint
-\texttt{\urlalt{http://www.ma.utexas.edu/mp_arc-bin/mpa?yn=#1}{mp\_arc:\hspace{0pt}#1}}%
+\texttt{mp\_arc:\urlalt{http://www.ma.utexas.edu/mp_arc-bin/mpa?yn=#1}{#1}}%
 %\texttt{\href{http://www.ma.utexas.edu/mp_arc-bin/mpa?yn=#1}{\protect\url{mp_arc:#1}}}%
@@
 \newcommand*{\haleprint}[1]{%\global\def\arxivp{\arxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
-\texttt{\urlalt{https://hal.archives-ouvertes.fr/#1}{HAL:\hspace{0pt}#1}}%
+\texttt{HAL:\urlalt{https://hal.archives-ouvertes.fr/#1}{#1}}%
 %\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
@@
 \newcommand*{\philscieprint}[1]{%\global\def\philscip{\philscisi}%\citeauthor{0philscicite}\addtocategory{ifarchcit}{0philscicite}%eprint
-\texttt{\urlalt{http://philsci-archive.pitt.edu/archive/#1}{PhilSci:\hspace{0pt}#1}}%
+\texttt{PhilSci:\urlalt{http://philsci-archive.pitt.edu/archive/#1}{#1}}%
 %\texttt{\href{http://philsci-archive.pitt.edu/archive/#1}{\protect\url{PhilSci:#1}}}%
@@
 \newcommand*{\biorxiveprint}[1]{%\global\def\biorxivp{\biorxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
-\texttt{\urlalt{https://doi.org/10.1101/#1}{bioRxiv doi:\hspace{0pt}10.1101/#1}}%
+\texttt{bioRxiv doi:\urlalt{https://doi.org/10.1101/#1}{10.1101/#1}}%
 %\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
@@
 \newcommand*{\osfeprint}[1]{%
-\texttt{\urlalt{https://doi.org/10.17605/osf.io/#1}{Open Science Framework doi:10.17605/osf.io/#1}}%
+\texttt{Open Science Framework doi:\urlalt{https://doi.org/10.17605/osf.io/#1}{10.17605/osf.io/#1}}%
 }
@@
 \begin{abstract}\labelsep 0pt%
-\mynote{to be written}
+\mynote{Abstract to be written}
 % \noindent \mynote{Abstract must be rewritten once paper is ready}This
@@
 could be explored with the present model. The assumptions of the model
-itself (\sect~\ref{sec:assumptions}) can either be under test or relied
-upon, depending on the case.
+itself, discussed in \sect~\ref{sec:assumptions}, can either be under test
+or relied upon, depending on the case.
 

### 2019-12-27 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-26T15:35:20+0100
+%% Last-Updated: 2019-12-27T12:18:07+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Our notation and terminology follow \textsc{iso}
-\citey{iso1993,iso2006,iso2006b} standards and Jaynes
+\citey{iso2006,iso2006b,jcgm1993_r2008} standards and Jaynes
 \citey{jaynes1994_r2003} for probability. We use \enquote{\dob},
@@
 \citep[we intend this term in the Meaning~2 of][]{kruskaletal1979b}[or
-Meanings~8 and~9 of][]{kruskaletal1979c}.
+Meanings~8 and~9 of][]{kruskaletal1979c}[note that the standard][\sect~3.1.14 states \enquote{The notion of representative sample is fraught with controversy, with some survey practitioners rejecting the term altogether}]{iso2014}.
 

### 2020-01-13 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2019-12-27T12:18:07+0100
+%% Last-Updated: 2020-01-13T15:39:05+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 realistic ways, to make the analysis mathematically tractable. This is why
-such initial probability distributions are called \enquote{models}; and
-that is how we intend this word in the present work.
+such initial probability distributions are called \enquote{(probabilistic)
+  models}; and that is how we intend this word in the present work.
 
@@
 very small sample of neurons -- hundreds at most -- compared to the numbers
-that constitute a functionally distinguished brain region. Many
-probabilistic models focus on such samples only: they neglect that the
-recorded neurons are a sample from a larger population. Some probabilistic
-models try to take unrecorded neurons into account, but they do so by
-describing each neuron individually, thus becoming very complex
+that constitute a functionally distinct brain region. Many probabilistic
+models focus on such samples only: they neglect that the recorded neurons
+are a sample from a larger population. Some probabilistic models try to
+take unrecorded neurons into account, but they do so by describing each
+neuron individually, thus becoming very complex
 \citep{pillowetal2008,dunnetal2013,tyrchaetal2014,huang2015,battistinetal2017}.
@@
 details. Such intermediate models would be useful at least for preliminary
-investigations, to help us decide which hypotheses to consider or discard
-for more complex and costly models, or to suggest new hypotheses.
+investigations, to help us decide which hypotheses to discard, which to
+consider for a deeper analysis with more complex and costly models, or to
+suggest new hypotheses.
 
@@
 In the present work we propose such an intermediate probabilistic model
-\citep[\cf][]{portamanaetal2015}. It answers this question: how much was
+\citep[\cf][\sect~3]{portamanaetal2015}. It answers this question: how much was
 the \emph{total} activity of a large neuronal population, given the
@@
 neurosciences for different kinds of investigations about the neuronal
-activity of various brain areas and about other phenomena of importance,
-for example gene and protein interaction
-\citep[]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,roudietal2009c,barreiroetal2010,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}[for
+activity of various brain areas and about other phenomena of importance % ,
+% for example gene and protein interaction
+\citep[]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,roudietal2009c,barreiroetal2010,shimazakietal2015,moraetal2015%,lezonetal2006,weigtetal2009
+}[for
 further references see][]{lathametal2013}. However, our model can also be
@@
 time. These neurons are our \enquote{sample}. Their spikes are binned into
-$T$ time bins and binarized to $\set{0,1}$ values in each bin. Call
-$\ya_{t}$ the number of neurons that fire during time bin $t$: this is the
-\emph{total activity} of the sample, or just \enquote{activity} for short.
-Obviously $\ya_{t} \in\set{0,1,\dotsc,n}$; if $\ya_{t}=0$, no neuron spikes
-during bin $t$; if $\ya_{t}=n$, all spike at some point during bin $t$, and
-so on. For brevity we say \enquote{at $t$} for \enquote{during time bin
+$T$ time bins, and a binary value 1 is assigned to each bin containing at
+least one spike, 0 to the bins without spikes. Call $\ya_{t}$ the number of
+neurons that fire during time bin $t$: this is the \emph{total activity} of
+the sample, or just \enquote{activity} for short. Obviously
+$\ya_{t} \in\set{0,1,\dotsc,n}$; if $\ya_{t}=0$, no neuron spikes during
+bin $t$; if $\ya_{t}=n$, all spike (at least once) at some point during bin
+$t$, and so on. For brevity we say \enquote{at $t$} for \enquote{during
+  time bin
   $t$}. % If we divide the total activity by the population size we have the
@@
 % ranging from $0$ to $1$ in $1/n$ steps.
-From the activities $\set{\ya_{t}}$
-we can count how often the activity levels $\ya=0$, $\ya=1$, and so on
-appeared during the recording, obtaining the distribution of measured
-relative frequencies $(\yff_{\ya}) \defs \yf$. We can also consider the
-(unknown) sample activity at time bins \emph{outside} of the recorded
-period.
+From the sequence of activities
+$(\ya_{t})\equiv (\ya_{1}, \dotso, \ya_{T})$ we can count how often the
+activity levels $\ya=0$, $\ya=1$, and so on appeared during the recording,
+obtaining the distribution of measured relative frequencies
+$(\yff_{\ya}) \equiv (\yff_{0},\dotso, \yff_{n}) \defs \yf$. We can also
+consider the (unknown) sample activity at time bins \emph{outside} of the
+recorded period.
 
@@
 The sample of recorded neurons may even change slightly across experimental
-sessions that are very far apart in time. We assume that there's an area,
+sessions that are very far apart in time. We assume that there's a volume,
 comprising a population of $N$ neurons, for which we believe that any other
@@
 \citep[we intend this term in the Meaning~2 of][]{kruskaletal1979b}[or
-Meanings~8 and~9 of][]{kruskaletal1979c}[note that the standard][\sect~3.1.14 states \enquote{The notion of representative sample is fraught with controversy, with some survey practitioners rejecting the term altogether}]{iso2014}.
-
-The total activity of the $N$ neurons at $t$ is $\yA_{t}$. The relative
-frequencies of the various activity levels during the recording were
-$(\yFF_{\yA}) \defs \yF$. The values $\yA_{t}$ at each $t$ and
-the frequency distribution $\yF$ are unknown to us. We only know for certain that
-$\yA_{t} \in \set{0, 1, \dotsc, N}$, that $\yA_{t} \ge \ya_{t}$, and that
-$N-\yA_{t} \ge n-\ya_{t}$ for obvious reasons. For the time being we assume
-that we know $N$; in \sect~\ref{sec:N} we discuss the consequences of our
-lack of precise knowledge about this number.
+Meanings~8 and~9 of][]{kruskaletal1979c}[note that the
+standard][\sect~3.1.14 states \enquote{The notion of representative sample
+  is fraught with controversy, with some survey practitioners rejecting the
+  term altogether}]{iso2014}.
+
+The total activity of the larger population of $N$ neurons at $t$ is
+$\yA_{t}$. The relative frequencies of the various activity levels during
+the recording were
+$(\yFF_{\yA}) \equiv (\yFF_{0},\dotso,\yFF_{N}) \defs \yF$. The values
+$\yA_{t}$ at each $t$ and the frequency distribution $\yF$ are unknown to
+us. We only know for certain that $\yA_{t} \in \set{0, 1, \dotsc, N}$, that
+$\yA_{t} \ge \ya_{t}$, and that $N-\yA_{t} \ge n-\ya_{t}$ for obvious
+reasons. For the time being we assume that we know $N$; in
+\sect~\ref{sec:N} we discuss the consequences of our lack of precise
+knowledge about this number.
 
@@
   \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured averages}}
-  \mathrel{\;\;=\;\;}
+  \mathrel{\;\;\underset{\mathclap{\overset{\Bigg\uparrow}{\text{our judgement}}}}{=}\;\;}
    \underbracket{\sum_{\yA}
@@
 \end{equation}
-where the $m'$ parameters $\yll \defd (\yl_{m})$ are determined by the
-constraints~\eqref{eq:final_constraint} and can be found by convex
+where the $m'$ parameters
+$\yll \defd (\yl_{m}) \equiv(\yl_{1},\dotso,\yl_{m'})$ are determined by
+the constraints~\eqref{eq:final_constraint} and can be found by convex
 extremization, discussed further in
@@
 $N=5\,000$ (\textcolor{myred}{red circles}), $N=10\,000$
-(\textcolor{myblue}{blue curve}). The frequency density of the sample
+(small \textcolor{myblue}{blue squares}). The frequency density of the sample
 activity is also shown (\textcolor{black}{black triangles}), and in the
@@
 \caption{Maximum-entropy distributions for the total activity of the larger
-  population, using five moments and assuming several population sizes $N$.
+  population, using $M={}$five moments and assuming several population sizes $N$.
   In order to compare different $N$, the distributions are multiplied by
@@
 two distinct low-activity modes. For the second data set,
-\fig~\ref{fig:plots_5mom_b}, the distribution presents a small shoulder on
-the right of its mode.
+\fig~\ref{fig:plots_5mom_b}, the distribution presents a small shoulder
+(barely visible) on the right of its mode.
 % , suggestive of two activity regimes. We frame no hypotheses about the
@@
 sampled neurons is of the first type, and the remaining of the second type.
-In the case of the first data set, 27 of the 65 sampled cells were
+In the case of the first data set, indeed, 27 of the 65 sampled cells were
 identified as grid cells belonging to 3--4 functional modules, the
@@
 activities as the peaks of the larger-population distribution
-(\textcolor{myblue}{blue curve}) but closer in height. It would thus seem
+(small \textcolor{myblue}{blue squares}) but closer in height. It would thus seem
 that the population of grid cells is contributing to the first mode of the
@@
 (\textcolor{black}{black diamonds}) is quite different from the
-larger-population distribution (\textcolor{myblue}{blue curve}): the two
+larger-population distribution (small \textcolor{myblue}{blue squares}): the two
 peaks of the former are closer in position and height than those of the
@@
 The toy analysis above should not be taken literally, but just as an
-illustration of the model's possible applications. The important point is
-that this model is computationally very cheap and yet it can provide
-useful insights, even if just qualitative ones, and even suggest new
-hypotheses.
+illustration of the model's possible applications. Instead of a grid cell
+\vs\ non-grid cell divisions we could for example have examined the modules
+of the cells. In the absence of such auxiliary information about the cells,
+one could in principle try all possible partitions of the recorded cells
+into two groups, and check whether some partition leads to independent
+distributions (computationally very expensive in practice); and so on.
+
+The important point is that this model is computationally very cheap and
+yet it can provide useful insights, even if just qualitative ones, and even
+suggest new hypotheses.
 
@@
 and similar need to be translated into a more precise, operational
-\citep[\cf][\chap~I]{bridgman1927_r1958} notion. Here we propose the notion
+\citep[\cf][\chap~I]{bridgman1927_r1958} notion. Here we propose to base
+this kind of analysis on the notion
 of \emph{informational sufficiency}, because
@@
 or about the activity in a new time bin, is conditional on all experimental
-data and statistics we have collected. This distribution, however, can be such that
-it remains unchanged if we discard part of the data or statistics -- for
-example, the third- and higher-order empirical moments. % --
+data and statistics we have collected. This belief distribution, however,
+can be such that it remains unchanged if we discard part of the data or
+statistics -- for example, the third- and higher-order empirical
+moments. % --
 % but it does change if we discard more.
@@
 
-Now let us turn the informational sufficiencies of two different subsets of
-statistics into two hypotheses. We can then ask: which hypothesis is more
-probable, given the data? equivalently, what is the ratio of their
+It seems to us that the informational sufficiency of a set of $m$ moments
+captures the idea of the \enquote{importance} of the almost-simultaneous
+activity of $m$ or fewer neurons: such simultaneous activity is
+\enquote{unimportant} if its knowledge does not alter our forecasts. We are
+thus speaking of informational importance, which may or may not imply or be
+implied by biophysical importance -- \enquote{correlation is no proof of
+  causation} \citep[\eg][]{simon1954,freedman2006}. But it lends itself to a
+precise probabilistic analysis, with consequences determined by
+mathematical theorems, as we shall now see.
+
+Our question is whether some amount of \enquote{cooperation} is more
+\enquote{important} than another amount or not. In terms of informational
+sufficiency, its definition might suggest to compare the belief
+distributions conditional on two different subsets of statistics. But which
+measure should be used for such a comparison? and how to motivate it? We
+prefer to take a different approach, leading to an automatic answer: let us
+thus turn the informational sufficiencies of two different subsets of
+statistics into two \emph{hypotheses}. We can then ask: which hypothesis is
+more probable, given the data? Equivalently, what is the ratio of their
 probabilities? By translating our question about \enquote{cooperativity}
@@
 % needed instead?
+We shall see that this approach is indeed more powerful than a simple
+comparison of distributions.
 
@@
 
-We can compare the informational sufficiency between of a set of moments, say
+We can compare the informational sufficiency of a set of moments, say
 $M''\defd\set{1,\dotsc,m''}$, with respect to another, say
@@
 Let us illustrate the quantity just defined and the previous discussion
-about sufficiency and sampling by using our data sets. Consider for example
-three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$, consisting of the
-first two, four, five moments; and a larger population of size $N=10\,000$.
-Figure~\ref{fig:plots_2_4mom} shows the resulting distributions for the two
-data sets.
+about sufficiency and sampling by using our data sets. We want to compare,
+for example, three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$,
+consisting of the first two, four, five moments, for a larger population of
+size $N=10\,000$. Figure~\ref{fig:plots_2_4mom} shows the resulting
+distributions for the two data sets.
 \begin{figure}[!p]
@@
 
-The crucial point is our belief distribution for the activity of the sample
-\emph{conditional} on the activity of the larger population. It is given by
-the hypergeometric distribution~\eqref{eq:conditional_hypergeometric},
-reprinted here:
+The answer comes by examining the origin of our belief distribution for the
+activity of the sample \emph{conditional} on the activity of the larger
+population. It is given by the hypergeometric
+distribution~\eqref{eq:conditional_hypergeometric}, reprinted here:
 \begin{equation}
@@
 
-Whether such symmetric beliefs are justified, or for which sampling procedures
-they can be justified, is the fundamental, deep question of sampling
-theory, which we cannot discuss here \citep[see \eg\ the discussions and
-references
+Whether this symmetry in our beliefs is justified, or for which sampling
+procedures it can be justified, is the fundamental, deep question of
+sampling theory, which we cannot discuss here \citep[see \eg\ the
+discussions and references
 in][]{ericson1969,smith1976}. % {kruskaletal1979c,kruskaletal1980}
@@
 this symmetry assumption is appropriate. But we must also be aware that
-such assessment in turn rests on analogous symmetry assumptions at a higher level,
-and so on. The probability-calculus, just like the truth-calculus, cannot
-yield conclusions without premisses \citep{hailperin2011}[p.~182]{johnson1924}.
+such assessment in turn rests on analogous symmetry assumptions at a higher
+level, and so on. The probability-calculus, just like the truth-calculus,
+cannot yield conclusions without premisses
+\citep{hailperin2011}[p.~182]{johnson1924}.
 
@@
 magnitude, considering the values
-$N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming the set $M$
+$N\in \set{10^{3}=1\,000, 10^{3+1/3}\approx 2\,000, 10^{3+2/3}\approx
+  5\,000, 10^{4}=10\,000, 10^{4+1/3}\approx 20\,000}$. Assuming the set $M$
 of first five moments to be sufficient, formula~\eqref{eq:prob_data_from_N}
@@
 approximating the number $T$ of time bins, the prior coefficient $L$, and
-the ratio $T/L$ to infinity. These approximations lead to the
-identification of the four distributions listed at the end of
+the ratio $T/L$ to infinity (that is, $T\gg L \gg 1$). These approximations
+lead to the identification of the four distributions listed at the end of
 \sect~\ref{sec:model}: the recorded frequency distribution becomes equal to
@@
 equality of the hypergeometric
-distribution~\eqref{eq:conditional_hypergeometric} and the conditional
+distribution~\eqref{eq:conditional_hypergeometric} and the \emph{conditional}
 frequencies of the sample activity $\ya$ given $\yA$. The goodness of this
@@
 distribution~\eqref{eq:app_maxent_pop_first} as the mode of the probability
-distribution for the recorded frequencies $\yF$, that is, the most probable
+distribution for the frequencies $\yF$, that is, the most probable
 recorded frequency.
@@
 
-The $T= \infty$ approximation can break our model for specific
-distributions of sample frequencies, if too many moments are used. For
-example, take $n=2$ with frequency distribution $\yf=(0.5,0,0.5)$ for the
-activities $\ya=0,1,2$; that is, the sample is silent half of the time and
-completely active half of the time. The first two normalized factorial
-moments have both value $0.5$, and together they uniquely determine $\yf$.
-If we now take $N=3$, it can be proved that
-\eqn~\eqref{eq:app_maxent_pop_first}, constrained on those two moments, has
-no solution. The reason is clear from a sampling perspective: if there is
-zero probability of sampling one active neuron, then the larger population
-cannot have any active neurons at all, and a fortiori there must be zero
-probability of sampling two active ones. But this reasoning holds only for
-each single time bin, not for the averages over time bins. The $T=\infty$
-approximation identifies the two instead. Outside of this approximation
-there are no contradictions because the conditional frequencies and
-conditional probabilities are kept distinct.
+The $T= \infty$ approximation (together with the third assumption above)
+can break our model for specific distributions of sample frequencies, if
+too many moments are used. For example, take $n=2$ with frequency
+distribution $\yf=(0.5,0,0.5)$ for the activities $\ya=0,1,2$; that is, the
+sample is silent half of the time and completely active half of the time.
+The first two normalized factorial moments have both value $0.5$, and
+together they uniquely determine $\yf$. If we now take $N=3$, it can be
+proved that \eqn~\eqref{eq:app_maxent_pop_first}, constrained on those two
+moments, has no solution. The reason is clear from a sampling perspective:
+if there is zero probability of sampling one active neuron, then the larger
+population cannot have any active neurons at all, and a fortiori there must
+be zero probability of sampling two active ones. But this reasoning holds
+only for each single time bin, not for the averages over time bins. The
+$T=\infty$ approximation identifies the two instead. Outside of this
+approximation there are no contradictions because the conditional
+frequencies and conditional probabilities are kept distinct.
 
@@
 
-This means that the maximum-entropy approximation presented here can be
-favourably used to get an idea of what a fully probabilistic calculation
-could yield -- but with a far lower computational cost than the latter. For
-example, for $N=1\,000$ the maximum-entropy distribution can be calculated
-in ten minutes on a laptop, whereas the full probability calculation takes
-several days on a high-performance computing cluster.
+This suggests that the maximum-entropy approximation presented here can be
+favourably used to get a qualitative or semi-quantitative idea of what a
+fully probabilistic calculation could yield -- but with a far lower
+computational cost than the latter. For example, for $N=1\,000$ the
+maximum-entropy distribution can be calculated in ten minutes on a laptop,
+whereas the full probability calculation takes several days on a
+high-performance computing cluster.
 
@@
 limited inferential scope makes the model computationally very cheap and,
-and most important, it still leads to inferences that are not at all
-trivial and not easily discernible by looking at the sample. We showed this
-by applying the model to two real data sets.
+most important, it still leads to inferences that are not at all trivial
+and could not be made just by looking at the sample and its distributions.
+We showed this by applying the model to two real data sets.
 
@@
 
-First of all the sampling relation. We have a population of $N$ units --
-neurons -- $\yA$ of which have some specific property -- being active in a
-specific time bin -- and we sample in an unknown way $n$ of the $N$ units.
+First of all the sampling relation. We have a population of $N$ units
+(neurons), $\yA$ of which have some specific property (being active in a
+specific time bin), and we sample $n$ of the $N$ units in such a way
+(partly uncontrollable experimental procedure) that we are led to believe,
+with the same degree, that any of the $N$ units could have been sampled.
 The probability that $\ya$ of the $n$ sampled neurons are active is then
@@
   \label{eq:sample_marginal}
-  \pxx(\ya_{t} \| M) = \sum_{\yA}\yG_{\ya_{t}\,\yA}\;\px(\yA \| M).
-\end{equation}
-The probability of observing one sequence $(\ya_{t})$ with frequencies
-$\yf$ is therefore
+  \pxx(\ya \| M) = \sum_{\yA}\yG_{\ya\,\yA}\;\px(\yA \| M).
+\end{equation}
+The probability of observing a specific sequence $(\ya_{t})$ with
+frequencies $\yf$ is therefore
 \begin{equation}
@@
   \pf(M \| \yf) \propto
-  \pf(\yf \| M) \times\text{\small const} \propto
+  \pf(\yf \| M) % \times\text{\small const}
+  \propto
   \binom{T}{T\yf}\,  \prod_{\ya}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{
@@
 distributions~\eqref{eq:maxent_prior}. Note the relative entropy in the
-exponential. If we want to conditionalize not on $\yf$ but on a set of
-measured averages $C(m) = \sum_{\ya}c_{\ya}(m) \yff_{\ya}$, $m\in M$, such
+exponential. If we want to condition not on $\yf$ but on a set of measured
+averages $C(m) = \sum_{\ya}c_{\ya}(m) \yff_{\ya}$, $m\in M$, such
 as~\eqref{eq:chain_prescription_moments}, we have to replace the

### 2020-01-13 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-01-13T15:39:05+0100
+%% Last-Updated: 2020-01-13T20:37:41+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 From the sequence of activities
-$(\ya_{t})\equiv (\ya_{1}, \dotso, \ya_{T})$ we can count how often the
-activity levels $\ya=0$, $\ya=1$, and so on appeared during the recording,
-obtaining the distribution of measured relative frequencies
-$(\yff_{\ya}) \equiv (\yff_{0},\dotso, \yff_{n}) \defs \yf$. We can also
+$(\ya_{t})\equiv (\ya_{1}, \dotso, \ya_{T})$ we can count in how many bins
+each activity level $\ya=0$, $\ya=1$, \ldots, $\ya=n$ appeared during the
+recording. Dividing each number of bins by the total number of bins $T$ we
+obtain the distribution of measured relative frequencies
+$(\yff_{0},\dotso, \yff_{n}) \equiv (\yff_{\ya}) \defs \yf$. We can also
 consider the (unknown) sample activity at time bins \emph{outside} of the
@@
 the recording were
-$(\yFF_{\yA}) \equiv (\yFF_{0},\dotso,\yFF_{N}) \defs \yF$. The values
+$(\yFF_{0},\dotso,\yFF_{N}) \equiv (\yFF_{\yA}) \defs \yF$. The values
 $\yA_{t}$ at each $t$ and the frequency distribution $\yF$ are unknown to
@@
 
-Our question is whether some amount of \enquote{cooperation} is more
-\enquote{important} than another amount or not. In terms of informational
-sufficiency, its definition might suggest to compare the belief
-distributions conditional on two different subsets of statistics. But which
-measure should be used for such a comparison? and how to motivate it? We
-prefer to take a different approach, leading to an automatic answer: let us
-thus turn the informational sufficiencies of two different subsets of
-statistics into two \emph{hypotheses}. We can then ask: which hypothesis is
-more probable, given the data? Equivalently, what is the ratio of their
+Our question is whether some degree of \enquote{cooperation} is more
+\enquote{important} than some other degree, or not. In terms of
+informational sufficiency, its definition might suggest to compare the
+belief distributions conditional on two different subsets of statistics.
+But which measure should be used for such a comparison? and how to motivate
+it? We prefer to take a different approach, leading to an automatic answer:
+let us thus turn the informational sufficiencies of two different subsets
+of statistics into two \emph{hypotheses}. We can then ask: which hypothesis
+is more probable, given the data? Equivalently, what is the ratio of their
 probabilities? By translating our question about \enquote{cooperativity}
@@
 (balls or neurons) could be drawn or sampled. Thus, consider the pool of
-all neurons which we believe -- with equal degree of belief for each neuron
--- could have been recorded. Then $N$ is the size of that pool. $N$
-therefore depends on factors such as: the shape, dimensions, and technical
-specifications of the recording probe; the inaccuracy in the insertion of
-the probe, leading for example to slightly different insertion points or
-angles; the density of neurons around the probe. But it also depends on the
-homogeneity of the brain region where the recording is made: if we believe
-that it does not matter whether the probe had been inserted in some other
-point of the same brain region, then
+all neurons which we believe -- with an \emph{equal} degree of belief for
+each neuron -- could have been recorded. Then $N$ is the size of that pool.
+$N$ therefore depends on factors such as: the shape, dimensions, and
+technical specifications of the recording probe; the inaccuracy in the
+insertion of the probe, leading for example to slightly different insertion
+points or angles; the density of neurons around the probe. But it also
+depends on the homogeneity of the brain region where the recording is made:
+if we believe that it does not matter whether the probe had been inserted
+in some other point of the same brain region, then
 formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
@@
 
-Whether this symmetry in our beliefs is justified, or for which sampling
-procedures it can be justified, is the fundamental, deep question of
-sampling theory, which we cannot discuss here \citep[see \eg\ the
+Whether this equality in our degrees of belief is justified, or for which
+sampling procedures it can be justified, is the fundamental, deep question
+of sampling theory, which we cannot discuss here \citep[see \eg\ the
 discussions and references
@@
   of Excellence scheme of the  Research Council of Norway (Yasser Roudi group).\\
-  PGLPM thanks the staff of the NTNU library for their always prompt
-  support; Mari, Miri, \amp\ Emma for continuous encouragement and
-  affection; Buster Keaton and Saitama for filling life with awe and
-  inspiration; the developers and maintainers of \LaTeX, Emacs, AUC\TeX,
-  Open Science Framework, Python, Inkscape, Sci-Hub for making a free and
-  unfiltered scientific exchange possible.
+  We thank Iv\'an Davidovich for very insightful and helpful comments on a
+  previous draft of this manuscript. PGLPM thanks the staff of the NTNU
+  library for their always prompt support; Mari, Miri, \amp\ Emma for
+  continuous encouragement and affection; Buster Keaton and Saitama for
+  filling life with awe and inspiration; the developers and maintainers of
+  \LaTeX, Emacs, AUC\TeX, Open Science Framework, Python, Inkscape, Sci-Hub
+  for making a free and unfiltered scientific exchange possible.
 %\rotatebox{15}{P}\rotatebox{5}{I}\rotatebox{-10}{P}\rotatebox{10}{\reflectbox{P}}\rotatebox{-5}{O}.

### 2020-01-14 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-01-13T20:37:41+0100
+%% Last-Updated: 2020-01-14T13:49:20+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\firstdraft}{4 November 2015}
-\newcommand*{\firstpublished}{***}
+\newcommand*{\firstpublished}{4 November 2015}
 \newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
@@
 %% this fixes missing white spaces
-\pdfmapline{+dummy-space <dummy-space.pfb}\pdfinterwordspaceon%
+%\pdfmapline{+dummy-space <dummy-space.pfb}
+%%\pdfglyphtounicode{space}{0020}
+%\pdfinterwordspaceon%
 
@@
 \mynote{Abstract to be written}
-% \noindent \mynote{Abstract must be rewritten once paper is ready}This
-%   work shows how to build a maximum-entropy probabilistic model for the
-%   total activity of a population of neurons, given only some activity data or
-%   statistics -- for example, empirical moments -- of a \emph{subpopulation}
-%   thereof. This kind of model is useful because neuronal recordings are
-%   always limited to a very small sample of a population of neurons.
-% %
-%   The model is applied to two sets of neuronal data available in the
-%   literature. In some cases it makes interesting forecasts about the larger
-%   population -- for example, two low-regime modes in the frequency
-%   distribution for the total activity -- that are not visible in the sample
-%   data or in maximum-entropy models applied only to the sample.
-% %
-%   For the two datasets, the maximum-entropy probability model applied only
-%   to the subpopulation is compared with the marginal probability
-%   distribution obtained from the maximum-entropy model applied to the full
-%   population. On a linear probability scale no large differences are
-%   visible, but on a logarithmic scale the two distributions show very
-%   different behaviours, especially in the tails.
-  \\\noindent\emph{\footnotesize Note: Dear Reader \amp\ Peer, this
-    manuscript is being peer-reviewed by you. Thank you.}
+\noindent This work shows how to build a maximum-entropy probabilistic
+model for the total activity of a large population of unrecorded neurons,
+given only some activity data or statistics recorded from a small sample
+thereof. The model addresses only the collective behaviour of the larger
+population, thus gaining in simplicity and computational effectiveness at
+the expense of more detailed inferential power.
+
+The model is applied to two sets of neuronal data from the literature. It
+makes interesting forecasts about the larger population, such as the
+probable presence of two low-regime modes in the total activity, which are
+not present in the sample data and could not be found with maximum-entropy
+models applied only to the sample.
+
+Using the same data sets it is shown how to use the model for questions
+about the \enquote{cooperativity}, \enquote{interaction}, or
+\enquote{synchrony} of neuronal activity. These questions are translated
+into a hypothesis-testing problem about informational sufficiency, and
+effectively answered with Bayes's theorem. The answers given by the present
+model can be at variance from those given by a maximum-entropy model at the
+sample level only. The reason is discussed: informational sufficiency is not
+preserved under sampling.
+
+The question of how large can the larger population addressed by
+model be, and the assumptions on which the model is based, from a Bayesian
+point of view, are also discussed.
+\\\noindent\emph{\footnotesize Note: Dear Reader
+  \amp\ Peer, this manuscript is being peer-reviewed by you. Thank you.}
 % \par%\\[\jot]

### 2020-01-14 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-01-14T13:49:20+0100
+%% Last-Updated: 2020-01-14T18:37:20+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}\labelsep 0pt%
-\mynote{Abstract to be written}
 \noindent This work shows how to build a maximum-entropy probabilistic
@@
 
-The question of how large can the larger population addressed by
-model be, and the assumptions on which the model is based, from a Bayesian
-point of view, are also discussed.
-\\\noindent\emph{\footnotesize Note: Dear Reader
-  \amp\ Peer, this manuscript is being peer-reviewed by you. Thank you.}
+The question of how large the larger population addressed by model should
+or can be, and the assumptions on which the model is based, from a Bayesian
+point of view, are also discussed. \\\noindent\emph{\footnotesize Note:
+  Dear Reader \amp\ Peer, this manuscript is being peer-reviewed by you.
+  Thank you.}
 % \par%\\[\jot]

### 2020-01-14 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-01-14T18:37:20+0100
+%% Last-Updated: 2020-01-14T19:53:07+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   of Excellence scheme of the  Research Council of Norway (Yasser Roudi group).\\
-  We thank Iv\'an Davidovich for very insightful and helpful comments on a
+  We thank Iv\'an Davidovich for very insightful and valuable comments on a
   previous draft of this manuscript. PGLPM thanks the staff of the NTNU

### 2020-01-16 maxNt_v6.tex
update

--- 2020-01-14 maxNt_v6.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-01-14T18:37:20+0100
+%% Last-Updated: 2020-01-16T10:11:59+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   This work is financially supported by the Kavli Foundation and the Centre
-  of Excellence scheme of the  Research Council of Norway (Yasser Roudi group).\\
-  We thank Iv\'an Davidovich for very insightful and helpful comments on a
+  of Excellence scheme of the  Research Council of Norway (Yasser Roudi
+  group).
+
+  We thank Iv\'an Davidovich for very insightful and valuable comments on a
   previous draft of this manuscript. PGLPM thanks the staff of the NTNU
-  library for their always prompt support; Mari, Miri, \amp\ Emma for
+  library for their always prompt support; Mari, Miri, Emma for
   continuous encouragement and affection; Buster Keaton and Saitama for

### 2020-01-16 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-01-16T10:11:59+0100
+%% Last-Updated: 2020-01-16T12:46:05+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 to build initial probability distributions has been used in the
-neurosciences for different kinds of investigations about the neuronal
+neurosciences for several kinds of investigations about the neuronal
 activity of various brain areas and about other phenomena of importance % ,
@@
 case the validity of the approximation decreases as the ratio $nN/T$
-increases. The most robust interpretation is~\ref{item:prob_freq}, and that
+increases. The most accurate interpretation is~\ref{item:prob_freq}, and that
 is how we prefer to interpret the distribution $\px$: as \emph{the most
@@
 it? We prefer to take a different approach, leading to an automatic answer:
-let us thus turn the informational sufficiencies of two different subsets
-of statistics into two \emph{hypotheses}. We can then ask: which hypothesis
-is more probable, given the data? Equivalently, what is the ratio of their
+let us turn the informational sufficiencies of two different subsets of
+statistics into two \emph{hypotheses}. We can then ask: which hypothesis is
+more probable, given the data? Equivalently, what is the ratio of their
 probabilities? By translating our question about \enquote{cooperativity}
@@
 \sect~\ref{sec:marginalization}. This assumption is crucial for the
-maximum-entropy method based on the Shannon entropy. Different prior
+maximum-entropy method based on the Shannon entropy. Alternative prior
 densities lead to alternative maximum-entropy methods; for example one
@@
 very close modes rather than one. The same preliminary studies also
-indicate that these typical features are robust for different prior choices
+indicate that these typical features are insensitive to alternative prior choices
 in the second assumption above, for example using a Dirichlet prior instead

### 2020-01-24 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-01-16T12:46:05+0100
+%% Last-Updated: 2020-01-24T12:54:04+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\headtitle}{Inferences about large neuronal populations}
-\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
+\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, E. Torre, V. Rostami}
 \newcommand*{\headauthor}{Porta Mana \etal}
@@
 \parbox{0.5\linewidth}%\makebox[0pt][c]%
-{\protect\centering Y. Roudi\\%
-\footnotesize\epost{\email{yasser.roudi}{ntnu.no}}}%
+{\protect\centering E. Torre\\%
+\footnotesize\epost{\email{torre}{ibk.baug.ethz.ch}}}%
 \hspace*{\stretch{1}}\\%
@@
 \parbox{0.5\linewidth}%\makebox[0pt][c]%
-{\protect\centering E. Torre\\%
-\footnotesize\epost{\email{torre}{ibk.baug.ethz.ch}}}%
+{\protect\centering Y. Roudi\\%
+\footnotesize\epost{\email{yasser.roudi}{ntnu.no}}}%
 \hspace*{\stretch{1}}%
@@
   This work is financially supported by the Kavli Foundation and the Centre
-  of Excellence scheme of the  Research Council of Norway (Yasser Roudi
-  group).
+  of Excellence scheme of the  Research Council of Norway.
 

### 2020-02-03 maxNt_v6.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-01-24T12:54:04+0100
+%% Last-Updated: 2020-02-03T11:24:40+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 {\protect\centering P.G.L.  Porta Mana\\%
-\footnotesize\epost{\email{pgl}{portamana.org}}}%
+\footnotesize\epost{\email{piero.mana}{ntnu.no}}}%
 \hspace*{\stretch{1}}%
@@
 {\protect\centering E. Torre\\%
-\footnotesize\epost{\email{torre}{ibk.baug.ethz.ch}}}%
+\footnotesize\epost{\email{emiliano.torre}{yahoo.com}}}%
 \hspace*{\stretch{1}}\\%
@@
 or can be, and the assumptions on which the model is based, from a Bayesian
-point of view, are also discussed. \\\noindent\emph{\footnotesize Note:
-  Dear Reader \amp\ Peer, this manuscript is being peer-reviewed by you.
-  Thank you.}
+point of view, are also discussed.
+% \\\noindent\emph{\footnotesize Note:
+%   Dear Reader \amp\ Peer, this manuscript is being peer-reviewed by you.
+%   Thank you.}
 % \par%\\[\jot]
@@
 
-The two infinity approximations are at the boundary of their validity in
+The two infinity approximations are at the edge of their validity range, in
 our examples. For this reason we interpret the maximum-entropy distribution
-as the \emph{most probable frequency distributions}, as mentioned at the
-end of \sect~\ref{sec:model}. An important question, then, is \emph{which
+as \emph{the most probable frequency distributions}, as mentioned at the
+end of \sect~\ref{sec:model}. In high dimensions, however, the features of
+the most plausible distribution may not be \emph{typical} of the majority
+of plausible distributions. An important question, then, is \emph{which
   features of the maximum-entropy frequency distribution are typical of the
-  majority of plausible frequency distributions?} Especially for features
-such as number, position, height of modes, or tail behaviour. Preliminary
-studies using a full probability calculation (to be published separately)
-show that several semi-quantitative features of the maximum-entropy
-distribution are indeed typical. For the first data set, for example, most
-frequency distributions have two high-frequency regimes of activity, as in
-\fig~\ref{fig:plots_5mom_a}, with roughly the same heights and almost the
-same locations, within a few percentages of population-averaged activity.
-Each of the two high-frequency regimes, though, may consist of two or three
-very close modes rather than one. The same preliminary studies also
-indicate that these typical features are insensitive to alternative prior choices
-in the second assumption above, for example using a Dirichlet prior instead
-of an entropic one.
+  majority of plausible frequency distributions?} -- especially for
+features such as the number, position, and heights of modes; or tail
+behaviour. Preliminary studies using a full probability calculation (in
+preparation, to be published separately) indicate that several
+semi-quantitative features of the maximum-entropy distribution are indeed
+typical. For the first data set, for example, most frequency distributions
+have a high-frequency regime at a population-averaged activity of around
+0.075, as in \fig~\ref{fig:plots_5mom_a}, but their second high-frequency
+regime is typically lower and appears for $N>5\,000$. Most frequency
+distributions have two or three very close modes within each high-frequency
+regime, rather than one. The same preliminary studies also indicate that
+these typical features are insensitive to alternative prior choices in the
+second assumption above, for example using a Dirichlet prior instead of an
+entropic one.
 
@@
 Some approximations implicit in the model make its numerical results
-semi-quantitative with respect to an exact analysis. The results should not
-be off by more than a few percentages for the population-averaged activity and
-for the frequency density. This imprecision, however, is compensated by the
+semi-quantitative with respect to an exact analysis.%  The results should not
+% be off by more than a few percentages for the population-averaged
+% activity and for the frequency density.
+This imprecision, however, is compensated by the
 extreme computational cheapness of the model, far less costly than an exact
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\clearpage
 \defbibnote{prenote}{{\footnotesize (\enquote{de $X$} is listed under D,

### 2020-03-16 replies1_ncomp.tex
update

--- 2020-01-24 maxNt_v6.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-01-24T12:54:04+0100
+%% Last-Updated: 2020-03-16T14:36:02+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newif\iftypodisclaim % typographical disclaim on the side
-\typodisclaimtrue
+\typodisclaimfalse
 \newcommand*{\memfontfamily}{zplx}
@@
 ]{memoir}
-\newcommand*{\updated}{\today}
-\newcommand*{\firstdraft}{4 November 2015}
-\newcommand*{\firstpublished}{4 November 2015}
-\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
+\newcommand*{\firstdraft}{16 March 2020}
+\newcommand*{\firstpublished}{16 March 2020}
+\newcommand*{\updated}{\ifarxiv***\else\today\fi}
+\newcommand*{\propertitle}{Sketch of replies to N. Comp. reviews%\\{\large ***}%
+}% title uses LARGE; set Large for smaller
 \newcommand*{\pdftitle}{\propertitle}
-\newcommand*{\headtitle}{Inferences about large neuronal populations}
-\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, E. Torre, V. Rostami}
-\newcommand*{\headauthor}{Porta Mana \etal}
-\newcommand*{\reporthead}{}% Report number
+\newcommand*{\headtitle}{}
+\newcommand*{\pdfauthor}{}
+\newcommand*{\headauthor}{}
+\newcommand*{\reporthead}{}%\ifarxiv\else Open Science Framework \href{https://doi.org/10.31219/osf.io/***}{\textsc{doi}:10.31219/osf.io/***}\fi}% Report number
 
@@
 
+%%%% empheq above seems more versatile than these:
 %\usepackage{fancybox}
-
 %\usepackage{framed}
@@
 % \begin{inparaenum} = \begin{enumerate*}
+\setlist{itemsep=0pt,topsep=\parsep}
 \setlist[enumerate,2]{label=\alph*.}
@@
 \SetMathAlphabet{\mathsf}{bold}{T1}{\sfdefault}{b}{sl}
-\newcommand*{\mathte}[1]{\textbf{\textit{\textsf{#1}}}}
+%\newcommand*{\mathte}[1]{\textbf{\textit{\textsf{#1}}}}
 % Upright sans-serif math alphabet
@@
 \usepackage[backend=biber,mcite,%subentry,
-citestyle=authoryear-comp,bibstyle=pglpm-authoryear,autopunct=false,sorting=ny,sortcites=false,natbib=false,maxcitenames=1,maxbibnames=8,minbibnames=8,giveninits=true,uniquename=false,uniquelist=false,maxalphanames=1,block=space,hyperref=true,defernumbers=false,useprefix=true,sortupper=false,language=british,parentracker=false]{biblatex}
+citestyle=authoryear-comp,bibstyle=pglpm-authoryear,autopunct=false,sorting=ny,sortcites=false,natbib=false,maxcitenames=2,maxbibnames=8,minbibnames=8,giveninits=true,uniquename=false,uniquelist=false,maxalphanames=1,block=space,hyperref=true,defernumbers=false,useprefix=true,sortupper=false,language=british,parentracker=false]{biblatex}
 \DeclareSortingScheme{ny}{\sort{\field{sortname}\field{author}\field{editor}}\sort{\field{year}}}
@@
 \DefineBibliographyExtras{british}{\def\finalandcomma{\addcomma}}
-\renewcommand*{\finalnamedelim}{\addcomma\space}
+\renewcommand*{\finalnamedelim}{\addspace\amp\space}
+%\renewcommand*{\finalnamedelim}{\addcomma\space}
 \setcounter{biburlnumpenalty}{1}
@@
 \setcounter{biburllcpenalty}{1}
-\DeclareDelimFormat{multicitedelim}{\addsemicolon\space}
-\DeclareDelimFormat{compcitedelim}{\addsemicolon\space}
-\DeclareDelimFormat{postnotedelim}{\space}
+\DeclareDelimFormat{multicitedelim}{\addsemicolon\addspace\space}
+\DeclareDelimFormat{compcitedelim}{\addsemicolon\addspace\space}
+\DeclareDelimFormat{postnotedelim}{\addspace}
 \ifarxiv\else\addbibresource{portamanabib.bib}\fi
@@
 }
-\newcommand*{\citep}{\footcites}
+\newcommand*{\citep}{\parencites}
 \newcommand*{\citey}{\parencites*}
+\newcommand*{\ibid}{\unspace\addtocounter{footnote}{-1}\footnotemark{}}
 %\renewcommand*{\cite}{\parencite}
-%\newcommand*{\citeps}{\parencites}
+%\renewcommand*{\cites}{\parencites}
 \providecommand{\href}[2]{#2}
@@
 \newcommand*{\biorxiveprint}[1]{%\global\def\biorxivp{\biorxivsi}%\citeauthor{0arxivcite}\addtocategory{ifarchcit}{0arxivcite}%eprint
-\texttt{bioRxiv doi:\urlalt{https://doi.org/10.1101/#1}{10.1101/#1}}%
+bioRxiv \texttt{doi:\urlalt{https://doi.org/10.1101/#1}{10.1101/#1}}%
 %\texttt{\href{http://arxiv.org/abs/#1}{\protect\url{arXiv:#1}}}%
@@
 \newcommand*{\osfeprint}[1]{%
-\texttt{Open Science Framework doi:\urlalt{https://doi.org/10.17605/osf.io/#1}{10.17605/osf.io/#1}}%
+Open Science Framework \texttt{doi:\urlalt{https://doi.org/10.17605/osf.io/#1}{10.17605/osf.io/#1}}%
 }
@@
 %% this fixes missing white spaces
-%\pdfmapline{+dummy-space <dummy-space.pfb}
-%%\pdfglyphtounicode{space}{0020}
-%\pdfinterwordspaceon%
+\pdfmapline{+dummy-space <dummy-space.pfb}\pdfinterwordspaceon%
 
@@
 \setsecnumformat{\upshape\csname the#1\endcsname\quad}
-\setsecheadstyle{\large\bfseries\sffamily%
-\centering}
+\setsecheadstyle{\bfseries\sffamily%
+\raggedright}
 \setsubsecheadstyle{\bfseries\sffamily%
@@
 \fi%typodisclaimtrue
-\makeoddhead{plain}{}{}{\footnotesize\reporthead}
+\makeoddhead{plain}{\scriptsize\reporthead}{}{}
 % \copypagestyle{manainitial}{plain}
@@
 
-\pagestyle{manaart}
+\pagestyle{plain}
 
 \setlength{\droptitle}{-3.9\onelineskip}
-\pretitle{\begin{center}\Large\sffamily%
+\pretitle{\noindent\large\sffamily%
 \bfseries}
-\posttitle{\bigskip\end{center}}
+\posttitle{\bigskip\mbox{}}
 
@@
 \postauthor{\par\end{center}}
-\predate{\DTMsetdatestyle{mydate}\vspace{0.5\baselineskip}%
-\begin{center}\footnotesize}
+\predate{\DTMsetdatestyle{mydate}\begin{center}\footnotesize}
 \postdate{\end{center}\vspace{-\medskipamount}}
@@
 \captiontitlefont{\footnotesize}
-\firmlists*
+%\firmlists*
 \midsloppy
@@
 
-\paragraphfootnotes\setlength{\footmarkwidth}{1em}
-%\twocolumnfootnotes
+\paragraphfootnotes
+%\setlength{\footmarkwidth}{0em}
+% \threecolumnfootnotes
 %\setlength{\footmarksep}{0em}
@@
 \scriptsize\bfseries#1}~}
+%\footmarkstyle{\textsuperscript{\color{myred}\scriptsize\bfseries#1}~}
 %\footmarkstyle{\textsuperscript{[#1]}~}
@@
 \author{%
+\iffalse\hspace*{\stretch{1}}%
+%% uncomment if additional authors present
+% \parbox{0.5\linewidth}%\makebox[0pt][c]%
+% {\protect\centering ***\\%
+% \footnotesize\epost{\email{***}{***}}}%
+% \hspace*{\stretch{1}}%
+\parbox{0.75\linewidth}%\makebox[0pt][c]%
+{\protect\centering P.G.L.  Porta Mana  \href{https://orcid.org/0000-0002-6070-0784}{\protect\includegraphics[scale=0.16]{orcid_32x32.png}}\\%
+\footnotesize Kavli Institute, Trondheim\quad\epost{\email{piero.mana}{ntnu.no}}}%
+%% uncomment if additional authors present
+% \hspace*{\stretch{1}}%
+% \parbox{0.5\linewidth}%\makebox[0pt][c]%
+% {\protect\centering ***\\%
+% \footnotesize\epost{\email{***}{***}}}%
 \hspace*{\stretch{1}}%
-\parbox{0.5\linewidth}%\makebox[0pt][c]%
-{\protect\centering P.G.L.  Porta Mana\\%
-\footnotesize\epost{\email{pgl}{portamana.org}}}%
-\hspace*{\stretch{1}}%
-%% uncomment if additional authors present
-\parbox{0.5\linewidth}%\makebox[0pt][c]%
-{\protect\centering E. Torre\\%
-\footnotesize\epost{\email{torre}{ibk.baug.ethz.ch}}}%
-\hspace*{\stretch{1}}\\%
-\parbox{0.5\linewidth}%\makebox[0pt][c]%
-{\protect\centering V. Rostami\\%
-\footnotesize\epost{\email{vrostami}{uni-koeln.de}}}%
-\hspace*{\stretch{1}}%
-\parbox{0.5\linewidth}%\makebox[0pt][c]%
-{\protect\centering Y. Roudi\\%
-\footnotesize\epost{\email{yasser.roudi}{ntnu.no}}}%
-\hspace*{\stretch{1}}%
-%\quad\href{https://orcid.org/0000-0002-6070-0784}{\protect\includegraphics[scale=0.16]{orcid_32x32.png}\textsc{orcid}:0000-0002-6070-0784}%
-}
-
-\date{Draft of \today\ (first drafted \firstdraft)}
-%\date{\firstpublished; updated \updated}
+\fi}
+
+%\date{Draft of \today\ (first drafted \firstdraft)}
+\date{\firstpublished; updated \updated}
 
@@
 %\DeclareMathOperator{\ent}{ent}%integer less or equal to
-%\DeclareMathOperator{\Ord}{O}%same order as
+\DeclareMathOperator{\Ord}{O}%same order as
 %\DeclareMathOperator{\ord}{o}%lower order than
@@
 \renewcommand*{\ge}{\geqslant}%greater or equal
-\DeclarePairedDelimiter\clcl{[}{]}
+%\DeclarePairedDelimiter\clcl{[}{]}
 %\DeclarePairedDelimiter\clop{[}{[}
@@
 \newcommand*{\p}{\mathrm{P}}%probability
-\newcommand*{\E}{\mathrm{E}}
-\renewcommand*{\|}{\nonscript\,\vert\nonscript\;\mathopen{}}
+%\newcommand*{\E}{\mathrm{E}}
+%\renewcommand*{\|}{\nonscript\,\vert\nonscript\;\mathopen{}}
+\renewcommand*{\|}[1][]{\nonscript\,#1\vert\nonscript\;\mathopen{}}
 %\DeclarePairedDelimiterX{\cond}[2]{(}{)}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
-\DeclarePairedDelimiterX{\condt}[2]{[}{]}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
+%\DeclarePairedDelimiterX{\condt}[2]{[}{]}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
 %\DeclarePairedDelimiterX{\conds}[2]{\{}{\}}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
@@
 \newcommand*{\vs}{{vs}}
-%\newcommand*{\etc}{{etc.}}
-%\newcommand*{\ie}{{i.e.}}
+\newcommand*{\eg}{{e.g.}}
+\newcommand*{\etc}{{etc.}}
+\newcommand*{\ie}{{i.e.}}
 %\newcommand*{\ca}{{c.}}
-\newcommand*{\eg}{{e.g.}}
 \newcommand*{\foll}{{ff.}}
@@
 %\def\hy{-\penalty0\hskip0pt\relax}
-\newcommand*{\labelbis}[1]{\tag*{(\ref{#1})$_\text{r}$}}
+%\newcommand*{\labelbis}[1]{\tag*{(\ref{#1})$_\text{r}$}}
 %\newcommand*{\mathbox}[2][.8]{\parbox[t]{#1\columnwidth}{#2}}
@@
 %\newcommand*{\slor}{\mathop{\textstyle\bigvee}}
-\newcommand*{\T}{^\intercal}%transpose
+%\newcommand*{\T}{^\intercal}%transpose
 %%\newcommand*{\QEM}%{\textnormal{$\Box$}}%{\ding{167}}
@@
 
-\newcommand*{\tav}{\widehat} %time average
-\newcommand*{\av}{\widebar} %pop average
-\newcommand*{\sav}{\widebar} %subpop average
-
-\newcommand*{\yG}{G}
-
-%\newcommand*{\yXv}{\varSigma}
-\newcommand*{\yAv}{A}
-%\newcommand*{\yxv}{s}
-\newcommand*{\yav}{a}
-\newcommand*{\yNg}{N_{\text{g}}}
-\newcommand*{\yNng}{N_{\text{ng}}}
-\newcommand*{\yNN}{\varNu}
-\newcommand*{\yFF}{F}
-\newcommand*{\yff}{f}
-\newcommand*{\yF}{\bm{\yFF}}
-\newcommand*{\yf}{\bm{\yff}}
-\newcommand*{\yFFm}{\yFF^{*}}
-\newcommand*{\yFm}{\yF^{*}}
-
-%\newcommand*{\yx}{\bm{\yxv}}%subpop state
-%\newcommand*{\yxs}{\sav{\yx}}%subpop av state
-%\newcommand*{\yX}{\bm{\yXv}}%pop state
-%\newcommand*{\yXf}{\av{\yX}}%pop av state
-\newcommand*{\yaa}{\bm{\yav}}%subpop value
-\newcommand*{\yaas}{\sav{\yaa\yaa}}%subpop av state
-%\newcommand*{\yXXf}{\av{\yAA\yAA}}%subpop av state
-\newcommand*{\ya}{\yav}%subpop av value
-\newcommand*{\yAA}{\bm{\yAv}}%pop value
-\newcommand*{\yA}{\yAv}%pop av value
-%conditional assumptions
-\newcommand*{\yH}{\varIota}
-\newcommand*{\yHa}{\varIota_\textrm{p}}
-\newcommand*{\yHb}{\varIota_\textrm{s}}
-\newcommand*{\yHc}{\varKappa}
-\newcommand*{\yHd}{\varIota_\textrm{u}}
-\newcommand*{\yHp}{\varIota'}
-\newcommand*{\yHi}{\varIota''}
-\newcommand*{\yI}{\varIota}
-\newcommand*{\yD}{D}
-%Lagr multipliers
-\newcommand*{\yg}{\bm{c}}
-\newcommand*{\yc}{\Hat{\bm{c}}}
-\newcommand*{\yll}{\bm{\lambda}}
-\newcommand*{\yl}{\lambda}
-\newcommand*{\yk}{z}
-\newcommand*{\yK}{\zeta}
-%entropies
-\newcommand*{\re}{\Eta}
-\newcommand*{\ybu}{\re_{\text{B}}}
-
-\newcommand*{\yr}{\bm{r}}
-
-\newcommand*{\ynuu}{\nu}
-\newcommand*{\ynu}{\bm{\ynuu}}
-\newcommand*{\yjj}{\phi}
-\newcommand*{\yj}{\bm{\yjj}}
-
-\newcommand*{\prop}[1]{`#1'}
-
-\newcommand{\appropto}{\mathrel{\vcenter{
-  \offinterlineskip\halign{\hfil$##$\cr
-    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
-% https://tex.stackexchange.com/a/33547/97039
-%\newcommand*{\fm}{f}
-\newcommand*{\px}{P_{\text{me}}}
-\newcommand*{\pxx}{p}
-\newcommand*{\pxxx}{P}
-\newcommand*{\Nd}{\textrm{N}}
-\newcommand*{\ycmm}{C}
-\newcommand*{\ycm}{\mathte{\ycmm}}
-\newcommand*{\ycs}{\bm{\sigma}}
-\newcommand*{\yL}{L}
-
-\newcommand*{\yla}{\bm{\lambda}}
-\newcommand*{\yJ}{\bm{J}}
-%\newcommand*{\yd}{\bm{\varDelta}}
-\newcommand*{\ydi}{\varDelta}
-
-\newcommand*{\mee}{maximum entropy}
-\newcommand*{\me}{maximum-entropy}
-
-\newcommand*{\sh}{\mathrm{\varEta}}
-\newcommand*{\nat}{\textrm{nat}}
-\newcommand*{\bit}{\textrm{bit}}
-\newcommand*{\hart}{\textrm{Hart}}
-
-\newsubfloat{figure}
-\subcaptionsize{\normalfont}
-\subcaptionlabelfont{\normalfont}
-\subcaptionfont{\normalfont}
+\newcounter{rpoint}
+\newcommand{\point}{\bigskip\refstepcounter{rpoint}\therpoint}
 %%% Custom macros end @@@
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\firmlists
-\raggedbottom
+%\firmlists
 \begin{document}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\abstractrunin
+\iffalse\abstractrunin
 \abslabeldelim{}
@@
 \begin{abstract}\labelsep 0pt%
-\noindent This work shows how to build a maximum-entropy probabilistic
-model for the total activity of a large population of unrecorded neurons,
-given only some activity data or statistics recorded from a small sample
-thereof. The model addresses only the collective behaviour of the larger
-population, thus gaining in simplicity and computational effectiveness at
-the expense of more detailed inferential power.
-
-The model is applied to two sets of neuronal data from the literature. It
-makes interesting forecasts about the larger population, such as the
-probable presence of two low-regime modes in the total activity, which are
-not present in the sample data and could not be found with maximum-entropy
-models applied only to the sample.
-
-Using the same data sets it is shown how to use the model for questions
-about the \enquote{cooperativity}, \enquote{interaction}, or
-\enquote{synchrony} of neuronal activity. These questions are translated
-into a hypothesis-testing problem about informational sufficiency, and
-effectively answered with Bayes's theorem. The answers given by the present
-model can be at variance from those given by a maximum-entropy model at the
-sample level only. The reason is discussed: informational sufficiency is not
-preserved under sampling.
-
-The question of how large the larger population addressed by model should
-or can be, and the assumptions on which the model is based, from a Bayesian
-point of view, are also discussed. \\\noindent\emph{\footnotesize Note:
-  Dear Reader \amp\ Peer, this manuscript is being peer-reviewed by you.
-  Thank you.}
+  \noindent ***
+% \\\noindent\emph{\footnotesize Note: Dear Reader
+%     \amp\ Peer, this manuscript is being peer-reviewed by you. Thank you.}
 % \par%\\[\jot]
@@
 %\qquad{\footnotesize Keywords: ***}
-\end{abstract}
+\end{abstract}\fi
 \selectlanguage{british}\frenchspacing
@@
 
-% \mynote{comment about the possibility of drawing conclusions about a brain
-%   area using different sets of neurons (eg because of recording across many
-%   sessions)}
-
-\section{Introduction:\\a model for collective inferences about large neuronal
-  populations}
-\label{sec:intro}
-
-What correlations are dominant in the neuronal activity of a specific brain
-area? How does such activity change when external stimuli or the activity
-of other areas change? Does such activity range over all its mathematically
-possible values, or only within restricted bounds?
-
-Answering this kind of questions always engages an element of uncertainty.
-We cannot say \enquote{the answer is such and such}; at best we can assign
-degrees of reasonable belief -- that is, probabilities -- to the possible
-answers. The assessment of such distributions of belief involves
-experimental data, such as recordings of neuronal activity from specific
-brain areas, and pre-data knowledge and hypotheses about biological
-conditions and mechanisms. When we translate such pre-data beliefs into an
-initial probability distribution we often simplify them in more or less
-realistic ways, to make the analysis mathematically tractable. This is why
-such initial probability distributions are called \enquote{(probabilistic)
-  models}; and that is how we intend this word in the present work.
-
-The best experimental measurements of instantaneous neuronal activity use
-remarkable technologies \citep{hongetal2019}, but can still only record a
-very small sample of neurons -- hundreds at most -- compared to the numbers
-that constitute a functionally distinct brain region. Many probabilistic
-models focus on such samples only: they neglect that the recorded neurons
-are a sample from a larger population. Some probabilistic models try to
-take unrecorded neurons into account, but they do so by describing each
-neuron individually, thus becoming very complex
-\citep{pillowetal2008,dunnetal2013,tyrchaetal2014,huang2015,battistinetal2017}.
-It would be useful to explore models that operate in between, addressing
-the activity of the larger brain area whence the sample comes, but
-\emph{collectively}, that is, without asking about individual neuronal
-details. Such intermediate models would be useful at least for preliminary
-investigations, to help us decide which hypotheses to discard, which to
-consider for a deeper analysis with more complex and costly models, or to
-suggest new hypotheses.
-
-% \mynote{[Yasser:] In the present work we aim to address the issue of
-%   building probability distributions over a large population using
-%   recording from subpopulations by considering the distribution of total
-%   population activity as an example. [} 
-In the present work we propose such an intermediate probabilistic model
-\citep[\cf][\sect~3]{portamanaetal2015}. It answers this question: how much was
-the \emph{total} activity of a large neuronal population, given the
-observation of the activity of a very small sample thereof? This model
-addresses a larger brain area, avoiding the assumption of isolation of the
-sample; and by focusing on the total activity, rather than the activity of
-individual neurons, it remains simple and numerically tractable.
-
-The model can be viewed as a straightforward combination of the
-maximum-entropy method and basic sampling relations from the
-probability-calculus, discussed in \sect~\ref{sec:model}. The
-maximum-entropy or minimum-relative-entropy method
-\citep{jaynes1957}[\chap~11]{jaynes1994_r2003}[for the minimum-relative
-entropy method see:][\sect~4.b]{jaynes1963}[for a more rigorous
-derivation]{hobsonetal1973}[for applications]{sivia1996_r2006}
-%[see also][for applications to irreversible statistical mechanics]{grandy1980}
-to build initial probability distributions has been used in the
-neurosciences for several kinds of investigations about the neuronal
-activity of various brain areas and about other phenomena of importance % ,
-% for example gene and protein interaction
-\citep[]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,roudietal2009c,barreiroetal2010,shimazakietal2015,moraetal2015%,lezonetal2006,weigtetal2009
-}[for
-further references see][]{lathametal2013}. However, our model can also be
-viewed as an approximation of a more detailed model within Bayesian theory;
-we shall constantly keep this point of view in mind in our discussion.
-
-We illustrate possible uses of our model in \sect~\ref{sec:application}, by
-applying it to two concrete data sets:
-\begin{enumerate*}[label=(\alph*)]
-\item the activity of 65 neurons recorded for 20 min from a rat's Medial
-  Entorhinal Cortex \citep{stensolaetal2012},
-\item the activity of 159 neurons recorded for 15 min from a macaque's
-  Motor Cortex \citep{rostamietal2016_r2017}.
-\end{enumerate*}
-For each data set the model gives us the most plausible frequencies of all
-levels of total activity of a much larger population, 1\,000--10\,000
-neurons in size, during the recording. For example it can tell us that 60
-out of 10\,000 neurons were most likely active during 1\% of the recording
-time (though not necessarily always the same 60), 250 neurons out of
-10\,000 were active during 0.4\% of the recording time, and so on. The
-precise meaning of this frequency distribution is explained in
-\sect~\ref{sec:model}. For the two example data sets, the guessed frequency
-distributions for the larger population are distinctly different from those
-for the sample. For example, for the first data set the frequency
-distribution for the larger population has two very distinct modes, both at
-low activities %($0.7\%$ and $2.6\%$ of simultaneously active neurons
-% within $3\;\mathrm{ms}$,
-(see \fig~\ref{fig:plots_5mom}), whereas the frequency distribution for the
-sample is monotonically decreasing with its maximum at zero activity. These
-results show that the proposed model can be used for the formulation or
-preliminary assessment of interesting hypotheses. We illustrate this use
-with toy examples in \sects~\ref{sec:application}
-and~\ref{sec:marginalization}. Note that it is not possible to guess these
-features of the larger population by applying the maximum-entropy method
-\emph{at the level of the sample alone}.
-
-% We want to stress the usefulness of making a quantified guess about the
-% activity of a larger brain area. Such a guess seems indeed to be the
-% primary idea behind recording a sample from that area. %  Second, it allows us
-% % to make comparisons across experimental sessions; such comparisons would me
-% % difficult or meaningless if made with the recorded samples, which generally
-% % comprise non-overlapping sets of neurons and differ in size.
-% There's also an important methodological reason.
-
-Our model also solves a methodological problem in the use of
-maximum-entropy methods to assess the \enquote{cooperativity},
-\enquote{interaction}, or \enquote{synchrony} in neuronal activity, for
-example studying its pairwise correlations and correlations of higher order
-\citep[see for
-example][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
-When the difference in size between a large population and a sample thereof
-is too large, the sufficiency of correlations of some order for the sample
-implies the \emph{lack} of sufficiency of correlations of the same order
-for the larger population, and vice versa. Maximum-entropy applications at
-the sample level can therefore deceive us in questions regarding the
-cooperativity of the larger population. Our model addresses directly the
-latter instead. We explain this point more precisely in
-\sect~\ref{sec:marginalization}.
-
-% Maximum-entropy models have often been used to test whether a smaller part
-% of the sample data, for example time averages and second-order
-% correlations, is sufficient for quantifying our \dob\ about some aspects of
-% the neuronal activity of the sample. This use of maximum-entropy at the
-% sample level, however, has an important issue. We discuss this issue and
-% explain how our approach bypasses it in \sect~\ref{sec:marginalization}.
-% Moreover, our proposed model gives clearer answers to questions of
-% correlation sufficiency: we show this by applying it to the two data set
-% above addressing an example question of sufficiency, two moments versus
-% four moments.
-
-How large is the \enquote{larger population} considered by the model
-proposed here? It obviously cannot include the full brain. The size of the
-larger population is determined by the validity of the formulae from
-sampling theory, and can range from a set of neurons around the recording
-probe to part of a brain area, depending on some assumptions about the
-recording procedure and the brain region under consideration. This matter
-is discussed in detail in \sect~\ref{sec:N}.
-
-In \sect~\ref{sec:assumptions} we discuss in detail the assumptions and
-approximations which define our model, from the point of view of the
-probability-calculus.
-
-% \mynote{review this}We obviously don't know whether the activity levels of
-% the larger population really had the frequencies given by the model, during
-% the recording. The model gives our most plausible guess. In high
-% dimensions, however, the features of the most plausible distribution may
-% \emph{not} be typical of the majority of most plausible distributions; and
-% the set of all possible frequency distributions, if the full population for
-% example comprises 1\,000 neurons, is a 1\,000-dimensional space. In
-% \sect~\ref{sec:assumptions} we therefore try to assess which features of
-% the frequency distribution delivered by our method may be typical and
-% therefore expected of the actual one. We find that general features such as
-% the bimodality of the first data set are indeed typical. Maximum-entropy
-% models can be considered as approximations of Bayesian models based on
-% various assumptions of inferential sufficiency
-% \citep{jaynes1986d_r1996,portamana2017}. How do our guesses change if we
-% modify our pre-data assumptions? We show, in the same section, that the
-% typical features indicated by our method are robust against such changes.
-
-A summary of all points above and a discussion of the usefulness of the
-model is given in the final \sect~\ref{sec:summary_discussion}.
+%\vspace{-3\bigskipamount}
+
+\addsec{Reviewer 1}
+\label{sec:rev1}
+
+(1).
+% \textsf{\footnotesize In Fig.1(a), the authors demonstrated that the ME model exhibits the bimodal structure if they extended the (imaginary) number of neurons, even if the subsamples show a monotonically decreasing distribution.
+The authors then speculate that this bimodal distribution might result from two different cell types.}
+{\footnotesize\textsf{First, I would like to direct attention to the paper
+  by Amari, Nakahara, Wu and Sakai in Neural Computation 2003, which the
+  authors also cited in this paper. In this paper, Amari et al showed that
+  "weak higher-order interactions of almost all orders are required for
+  realizing a widespread activity distribution." (Theorem 3). As shown in
+  the case of pairwise model (Eq.3.19 in their paper), if the fixed moments
+  are limited and do not have higher-order, a distribution will be
+  concentrated when we increase the number of neurons.}
+
+\textsf{The suggested ME model by this paper constrains a few
+  moments as opposed to larger number of neurons, which can not result in
+  the wide-spread distributions according to the above theory. In other
+  words, I speculate that the bimodal structure is an artifact of
+  constraining fewer number of moments for a model with such large number
+  of neurons. This could happen even if we use up to all 65th moments. I
+  would like to know how the authors think of Amari's paper in the light of
+  the current manuscript.}
+
+}
 
@@
 
-Our notation and terminology follow \textsc{iso}
-\citey{iso2006,iso2006b,jcgm1993_r2008} standards and Jaynes
-\citey{jaynes1994_r2003} for probability. We use \enquote{\dob},
-\enquote{belief}, and \enquote{probability} interchangeably.
-
-% (for a full population of 1\,000 neurons and
-% a recording of 500\,000 time bins, for example, there are
-% $\binom{500\,000+1\,000}{1\,000} \approx 10^{3\,131}$ possible frequency
-% distributions).
-
-% Later on we'll show that our \dobs\ can be quantified by
-% exclusively using the probability-calculus. This derivation will provide a
-% more accurate quantification, revealing that the maximum-entropy answer is
-% only a first approximation.
-
-% \mynote{Say something more about advantages of such question/answer: for
-%   example we can make statements about total activity of brain area even
-%   across recordings when sampled neurons aren't the same.}
-
-% \mynote{Add this?: from sampling theory we know that important
-%   features of the full population may not be visible in a sample because
-%   smoothed out. But using sampling theory in the inverse direction we can
-%   infer such full-population features from the sample.}
-
-% \mynote{To be continued after structure of the rest of the article is
-%   clear. Orig intro is on p.~\pageref{orig_intro}}
-
-\section{Model: maximum-entropy and sampling,\\ and a Bayesian perspective}
-\label{sec:model}
-
-Some context and mathematical notation first.
-
-The context we consider is as follows. During an experimental session we
-have recorded the spiking activities of $n$ neurons for a certain amount of
-time. These neurons are our \enquote{sample}. Their spikes are binned into
-$T$ time bins, and a binary value 1 is assigned to each bin containing at
-least one spike, 0 to the bins without spikes. Call $\ya_{t}$ the number of
-neurons that fire during time bin $t$: this is the \emph{total activity} of
-the sample, or just \enquote{activity} for short. Obviously
-$\ya_{t} \in\set{0,1,\dotsc,n}$; if $\ya_{t}=0$, no neuron spikes during
-bin $t$; if $\ya_{t}=n$, all spike (at least once) at some point during bin
-$t$, and so on. For brevity we say \enquote{at $t$} for \enquote{during
-  time bin
-  $t$}. % If we divide the total activity by the population size we have the
-% \emph{normalized total activity} or population-averaged activity $\ya/n$,
-% ranging from $0$ to $1$ in $1/n$ steps.
-From the sequence of activities
-$(\ya_{t})\equiv (\ya_{1}, \dotso, \ya_{T})$ we can count in how many bins
-each activity level $\ya=0$, $\ya=1$, \ldots, $\ya=n$ appeared during the
-recording. Dividing each number of bins by the total number of bins $T$ we
-obtain the distribution of measured relative frequencies
-$(\yff_{0},\dotso, \yff_{n}) \equiv (\yff_{\ya}) \defs \yf$. We can also
-consider the (unknown) sample activity at time bins \emph{outside} of the
-recorded period.
-
-For many animal species, the neurons that are recorded within a brain area
-are not specifically chosen from among the rest, owing to several limiting
-factors; for example, limitations in how precisely electrodes are inserted.
-The sample of recorded neurons may even change slightly across experimental
-sessions that are very far apart in time. We assume that there's a volume,
-comprising a population of $N$ neurons, for which we believe that any other
-sample of size $n$ could have equally plausibly been recorded instead of
-the sample of $n$ neurons that was actually recorded. This is what we will
-mean with \enquote{larger population}. This population need not be a whole
-functionally or anatomically distinct region. Loosely speaking it is the
-area of which we believe our sample to be \enquote{representative}
-\citep[we intend this term in the Meaning~2 of][]{kruskaletal1979b}[or
-Meanings~8 and~9 of][]{kruskaletal1979c}[note that the
-standard][\sect~3.1.14 states \enquote{The notion of representative sample
-  is fraught with controversy, with some survey practitioners rejecting the
-  term altogether}]{iso2014}.
-
-The total activity of the larger population of $N$ neurons at $t$ is
-$\yA_{t}$. The relative frequencies of the various activity levels during
-the recording were
-$(\yFF_{0},\dotso,\yFF_{N}) \equiv (\yFF_{\yA}) \defs \yF$. The values
-$\yA_{t}$ at each $t$ and the frequency distribution $\yF$ are unknown to
-us. We only know for certain that $\yA_{t} \in \set{0, 1, \dotsc, N}$, that
-$\yA_{t} \ge \ya_{t}$, and that $N-\yA_{t} \ge n-\ya_{t}$ for obvious
-reasons. For the time being we assume that we know $N$; in
-\sect~\ref{sec:N} we discuss the consequences of our lack of precise
-knowledge about this number.
-
-Our questions concern general features of the total activity $\yA$ of the
-larger population during and after the recording, and across sessions under
-the same study conditions. For example: what was its frequency distribution
-$\yF$ during the recording? How much does this frequency distribution
-change across sessions? How much total activity should we expect at any
-time bin during a recording? The model presented here gives a probability
-distribution over the answers to these questions.
+We think that the arguments of Amari \etal\ \citey{amarietal2003} do not
+apply to the inference in our manuscript, for several reasons:
+\begin{enumerate}[label=\roman*.]
+\item\label{item:maxmomentcheck} We did perform the calculations with an
+  increasing number of moments as sufficient statistics, eventually using
+  the full sample distribution. As mentioned on p.~9 and shown in
+  \sect~4, from about 5 moments upwards the final result doesn't change
+  appreciably.
+
+  Most important, as we discuss in \sects~6 and~7, our results are the same
+  \emph{without} using any sufficient statistics -- that is, \emph{no
+    maximum-entropy} model at all -- under a fully Bayesian nonparametric
+  approach. In fact, the fully nonparametric approach leads to slightly
+  wider distributions (we hope to publish these results soon).
+
+\item As $N$ increases, the two peaks we infer do become more pronounced,
+  as suggested by our Fig.~1. So our inference is not really in
+  contradiction with Amari \etal. The results of that paper hold for the
+  $N\to\infty$, and no explicit form for the limit-remainder is given. It
+  is therefore unclear whether at $N=10\,000$ we should already observe a
+  delta. Our findings~\ref{item:maxmomentcheck}, above, suggest that we
+  should not. Limit-reminders such as \enquote{$\Ord(1/N^{k})$} as given in
+  Amari \etal\ do not help unless we know their precise form. As an extreme
+  example, $\frac{10^{10}}{N}$ is $\Ord(1/N)$, and in this case the $1/N$
+  term cannot really be neglected unless $N \gg 10^{10}$.
+
+\item The results of Amari \etal\ assume that the probability of the
+  neuronal activity factorizes over time. In our manuscript we assume, as
+  explained in \sect~6, that our degree of belief is \emph{exchangeable}
+  under time-bin permutations, but this exchangeability \emph{does not
+    imply factorizability} over time, as is clear from de~Finetti's
+  theorem. So we are not sure whether Amari \etal's results apply at all to
+  our inference.
+\end{enumerate}
 
@@
 
-The idea behind our approach can be easily summarized in terms of the
-maximum-entropy method:
-\begin{enumerate}[label=(\alph*)]
-\item\label{item:sample_step} Using sampling theory we determine the
-  relation between some expected values -- specifically, moments -- for the
-  total activity $\ya$ of the sample and corresponding expected values %$M$
-  for the total activity $\yA$ of the larger population.
-\item\label{item:maxent_step} Using the maximum-entropy method we build a
-  distribution $\px(\yA \| M, N)$ for the total activity of the larger
-  population of $N$ neurons, using as constraints the expected values $M$
-  found in the previous step.
-\end{enumerate}
-We now discuss these steps more in detail, but leave their precise
-mathematical implementation and a more detailed list of references to
-appendix~\ref{sec:derivation_maxent_fullnet}.
+(2) [we can reply by applying to real data.]
+
+\bigskip
+
+(3) {\footnotesize\textsf{I was not satisfied with the section of the Bayes
+  factor. The Bayes factor (Eq.10) was calculated for models with different
+  dimensionality. Then, a larger model should better account for the data.
+  How do you determine whether you include the higher-order or not, in
+  other words how do you determine the significance of the delta? If the
+  models have the same dimension, one may use heuristics, for example given
+  by Kass and Raftery (JASA 1995). If you have nesting models, one may
+  resort to chi-squared test? Just computing the BF without such a
+  verification method, providing the quantity BF adds only marginally to
+  the visual inspection of the distributions.}\par}
 
@@
 
-Step~\ref{item:sample_step} is just an application of the
-probability-calculus, which gives an exact linear relation between the
-first $m$ moments for the larger population and the first $m$ for the
-sample \citep[see][for a complete discussion and proofs. Some proofs are
-summarized again here in
-appendix~\ref{sec:derivation_maxent_fullnet}]{portamanaetal2015}. The ones
-determine the others and vice versa at every time bin. This relation holds
-for any belief distribution $\pxxx(\yA_{t})$ for the larger-population
-activity and its marginal $\pxx(\ya_{t})$ for the sample activity at that
-bin. These two distributions are related by
-\begin{equation}
-  \label{eq:conditional_hypergeometric_t}
-  \pxx(\ya_{t})= \sum_{\yA_{t}=0}^{N} \yG_{\ya_{t}\,\yA_{t}} \;\pxxx(\yA_{t})
-  \qquad \ya_{t} \in \set{0,\dotsc,n},
-\end{equation}
-where $\yG_{\ya\yA}$ is the hypergeometric distribution defined in
-\eqn~\eqref{eq:conditional_hypergeometric}, characteristic of
-\enquote{drawing without replacement}
-\citep[\chap~3]{jaynes1994_r2003}.%[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
-
-This sampling relation is even more straightforward if instead of power
-moments we use \emph{normalized factorial moments} \citep{potts1953}. The
-$m$th normalized factorial moment of a distribution $\pxx(\ya)$ is
-\begin{equation}
-  \label{eq:def_factorial_moment_a}
-  % \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
-  % \defd
-  \sum_{\ya=0}^{n} %m!\,
-  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya)
-  \qquad m \in \set{1,\dotsc,n},
-\end{equation}
-% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
-% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
-% new bin.
-that is, the expected number of distinct $m$-tuples of simultaneously
-active neurons (within a time-bin's width), normalized by the maximum
-possible number of distinct $m$-tuples. % For
-% example, with $m=2$, if $a=4$ neurons spike in a population of $n=5$, we
-% have $\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons,
-% and the total number of distinct pairs is $\tbinom{5}{2}=10$. The
-% normalized number of spiking pairs is therefore $6/10$.
-Note that the first $m$ factorial moments together provide the same
-information as the first $m$ power moments together, and vice versa: they
-are linearly related because $\tbinom{\ya}{m}$ is a polynomial in $\ya$ of
-degree $m$. So we'll just say \enquote{first $m$ moments} from now
-on. % Specifying one set is therefore
-% equivalent to specifying the other set.
-But the normalized factorial moments have a mathematically convenient
-property: \emph{the first $n$ normalized factorial moments for the sample
-  and for the larger population are numerically identical}:\citep[Appendix]{portamanaetal2015}
-\begin{equation}
-  \label{eq:equal_m_factorialmoments}
-  % \begin{split}
-  % \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
-  % &=
-  % \E\Biggl[\binom{\yA}{m} \Biggr]\bigg/\binom{N}{m}
-  % \qquad\text{or}\\
-    \sum_{\ya=0}^{n} %m!\,
-  \binom{\ya}{m}\bigg/  \binom{n}{m}\; \pxx(\ya)
-=    \sum_{\yA=0}^{N} %m!\,
-  \binom{\yA}{m}\bigg/  \binom{N}{m}\; \pxxx(\yA),\qquad
-  % \E\condt[\big]{\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
-  % \E\condt[\big]{\tbinom{\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
-  m \in \set{1,\dotsc,n}.
-%\end{split}
-\end{equation}
-% where $\pxxx(\yA)$ is our distribution of belief about the larger-population
-% activity, and $\pxx$
+Let us emphasize, first of all, that the purpose of our paper is to make
+inferences about the total activity of the larger population, not to decide
+among different sufficient statistics or correlation orders. Section~4 is
+meant to show the implications of our method for that kind of discussions.
+In particular we numerically show that conclusions based on a sample and
+conclusions based on the larger population can be very different. Our
+method can bypass that difficulty, if one wants to use it for such
+comparisons.
+
+There are many differing views on the questions of comparison, complexity,
+nesting, and \enquote{penalization} of models. Our view is close to that
+expressed in several works, studies, and commentaries
+\citep[\eg][]{good1985,jonesetal1986,chatfield1995,copasetal1995,draper1995,spiegelhalteretal1995,raftery1995,gelmanetal1995,draperetal1996,hoetingetal1999,clydeetal1999,browneetal2006},
+and well-explained in MacKay \citey{mackay1992}. We summarize this view
+below. According to it, the analysis in our \sect~4, based on the weight of
+evidence, is consistent and does not need any corrections.
+
+Note in particular that we absolutely avoid speaking about
+\enquote{significance}, also as recommended by the American Statistical
+Association in their recent official statements \citep{asa2016}[especially
+\sect~2]{asa2019}. We are simply calculating the evidence \citep{good1950}
+$\p(D \|M_{i})$ for several hypotheses $M_{i}$ based on data $D$. Evidence
+is a continuous quantity and we are not setting any arbitrary thresholds.
+Multiplied by a pre-data probability, this evidence gives the post-data
+probability for each hypothesis. The pre-data probability will depend on
+the specific case, animal, brain region, and additional knowledge. If the
+evidence is $10^{35}:1$ for the first hypothesis against the second, but
+our pre-data probability for the second is $10^{36}$ times that for the
+first, then the second hypothesis will have higher post-data probability.
+But there is still a continuum of cases, no dichotomy. No such dichotomy is
+necessary.
+
+If a \emph{choice} among two or more hypotheses has to be made for some
+purpose (say, hardware algorithm implementation in some automated device),
+such choice requires not only the probabilities $p(H_{i}\|D)$, but also the
+gain/loss matrix for the various choices conditional on the various
+hypotheses. The choice is then made by maximization of the expected gain.
+Some gain matrices can even lead to choosing the hypothesis having
+\emph{lower} posterior probability \citep[this frequently happens in
+medical decision problems, see \eg][]{soxetal1988_r2013}.
 
@@
 
-In step~\ref{item:maxent_step} we actually use the
-\emph{minimum-relative-entropy} method
-\citep{hobsonetal1973,csiszar1985}[\sect~5.2.2]{sivia1996_r2006} with
-respect to a uniform reference distribution. We shall still call it
-\enquote{maximum-entropy} for brevity. It amounts to two prescriptions:
-first, take the distributions satisfying specific convex constraints --
-such as fixed expectations -- and among them select the one having minimum
-relative entropy with respect to a reference distribution; second, judge --
-and therefore set -- those expectations to be equal to some measured
-averages, typically time averages. (A judgement is required because
-probability expectations and empirical averages differ numerically in
-general; a misjudgement of their approximate equality can lead to
-\enquote{silly answers}, to quote MacKay \citep[p.~308]{mackay1995_r2003}.)
-
-In our case we don't know the time averages of the quantity
-$\binom{\yA}{m}$, so we cannot directly equate them to the factorial moment
-$\E\bigl[\binom{\yA}{m}\bigr]$ of the larger-population distribution
-$\pxxx(\yA)$. But \eqn~\eqref{eq:equal_m_factorialmoments} of
-step~\ref{item:sample_step} comes to our rescue, because it says that the
-expectation for the larger population are determined by that for the sample
-$\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time average of its
-corresponding sample quantity $\binom{\ya}{m}$. So we can combine the two
-steps:
-\begin{equation*}
-  \label{eq:chain_prescription_moments}
-  \mathrlap{\overbracket{\phantom{\text{measured averages} \rightarrow
-        \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
-  \text{measured averages} \rightarrow
-        \underbracket{\text{sample moments}
-  \rightarrow
-    \text{larger-population moments}}_{\textit{sampling theory}}
-\end{equation*}
-We obtain a distribution $\px(\yA \| M, N)$ for the larger population of
-$N$ neurons by constraining some of its factorial moments, for example
-$M=\set{1,\dotsc,m'}$ with $m'\le n$, to be equal to the sample's recorded
-averages. In formulae, the constraints on $\pxxx(\yA)$ are
-\begin{equation}
-  \label{eq:final_constraint}
-  \underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\bigg/\binom{n}{m}
-  \equiv
-   \sum_{\ya}
-  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured averages}}
-  \mathrel{\;\;\underset{\mathclap{\overset{\Bigg\uparrow}{\text{our judgement}}}}{=}\;\;}
-   \underbracket{\sum_{\yA}
-     \binom{\yA}{m}\bigg/  \binom{N}{m} \; \pxxx(\yA)}_{\text{distribution moments}}
-   \qquad m\in M.
-\end{equation}
-\smallskip
-
-The result is the distribution of the form
-\begin{equation}
-  \label{eq:app_maxent_pop_first}
-  \begin{gathered}
-  \px(\yA \| M, N)  = \frac{1}{Z(\yll)}\;
-%  g(\yA)\;
-  \exp\Biggl[\sum_{m}\yl_{m}\;
-  \binom{\yA}{m}\bigg/\binom{N}{m}
-  \Biggr]
-  \\
-\text{with}\qquad  Z(\yll)\defd \sum_{\yA}
-  % g(\yA)\;
-  \exp\Biggl[\sum_{m}\yl_{m}\;
-  \binom{\yA}{m}\bigg/\binom{N}{m}
-  \Biggr],
-\end{gathered}
-\end{equation}
-where the $m'$ parameters
-$\yll \defd (\yl_{m}) \equiv(\yl_{1},\dotso,\yl_{m'})$ are determined by
-the constraints~\eqref{eq:final_constraint} and can be found by convex
-extremization, discussed further in
-appendix~\ref{sec:derivation_maxent_fullnet}. Note that for $N>n$ the
-constraint equation~\eqref{eq:final_constraint} may never be exactly
-satisfied by the distribution~\eqref{eq:app_maxent_pop_first} for any
-$\yla$. In this case the extremization minimizes the discrepancy between
-the left and right sides of~\eqref{eq:final_constraint}. This discrepancy
-comes from an approximation implicit in the maximum-entropy method (namely
-that the number of time bins $T$ is infinite) combined with one assumption
-specific to our application: that an activity level $\yA$ can equally
-likely be generated by any set of $\yA$ neurons in the larger population.
-The magnitude of this discrepancy can be a signature of the presence of
-neuronal \enquote{assemblies} or \enquote{clusters}
-\citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss this matter
-in \sect~\ref{sec:assumptions}.
-
-\bigskip
-
-One important question remains about the distribution $\px$ of
-\eqn~\eqref{eq:app_maxent_pop_first}: \emph{what} is it a distribution of?
--- Of probability? Of relative frequency? The answer is related to an
-important relation between maximum-entropy distributions and the
-probability-calculus, and also to our alternative point of view of the
-present model.
-
-% Before applying the formula above to two concrete data sets we want to add
-% two remarks about the maximum-entropy method that are seldom made in the
-% neuroscientific literature. They are important for the interpretation of
-% the results and are further discussed in \sect~\ref{sec:assumptions}.
-% First, the maximum-entropy
-% method % (based on the \emph{Shannon} relative
-% % entropy, as opposed to other entropy functions such as Burg's
-% % \citeyear{burg1975})
-% rests on some implicit assumptions about the probabilities for the long-run
-% frequency distribution of activities, besides the assumptions just
-% mentioned at the end of the previous paragraph. So the often-heard
-% statement that it gives \enquote{the maximally unbiased (or non-committal)
-%   distribution} must be taken with a grain of salt.
-% \citep{jaynes1986d_r1996,portamana2009,portamana2017}. Second,
-A maximum-entropy distribution like $\px(\yA \| M, N)$ is equivalent to the
-zeroth-order Laplace approximation
-\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
-four distinct distributions for the larger population, which differ
-numerically from one another in higher-order approximations:
-\begin{enumerate}[label=(\roman*)]
-\item\label{item:prob_freq} the most probable \emph{frequency} distribution
-  for the total activity across the \emph{recorded} bins, %$\yFF(\yA)$,
-\item\label{item:prob_binin} the \emph{belief} distribution for the value
-  of the total activity at any time bin among the \emph{recorded} ones,
-  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
-\item\label{item:prob_longrun} the most probable \emph{frequency}
-  distribution for the total activity in a very long run of \emph{new} time
-  bins, %$\ynuu(\yA)$,
-\item\label{item:prob_binout} the \emph{belief} distribution for the value
-  of the total activity at a \emph{new} time bin.
-  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
-\end{enumerate}
-Their mathematical relationship is presented more explicitly in
-appendix~\ref{sec:inference_prob}. The
-distribution~\eqref{eq:app_maxent_pop_first} can therefore be interpreted
-in each of the four ways above, in this approximate sense. In the present
-case the validity of the approximation decreases as the ratio $nN/T$
-increases. The most accurate interpretation is~\ref{item:prob_freq}, and that
-is how we prefer to interpret the distribution $\px$: as \emph{the most
-  probable relative-frequency distribution} of the activity $\yA$. See
-\sect~\ref{sec:assumptions} for further discussion about assumptions and
-approximations.
-
-\section{Example application: two data sets}
-\label{sec:application}
-
-We apply the model just described to two data sets publicly available in
-the literature:
-\begin{itemize}[wide]
-\item The \textbf{first data set}, from \textcite[rat
-  14147]{stensolaetal2012}, consists of $n=65$ neurons from rat Medial
-  Entorhinal Cortex, simultaneously recorded for about 20 minutes. Their
-  spikes are binned into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
-\item The \textbf{second data set}, from \textcite[supplementary
-  material]{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
-  consists of $n=159$ neurons from macaque Motor Cortex, simultaneously
-  recorded for about 15 minutes. Their spikes are binned into $T=300\,394$
-  bins of $3\;\textrm{ms}$ width.
-\end{itemize}
-
-% In the followingFrom a Bayesian point of view, the maximum-entropy distribution $\px(\yA)$
-% can be considered as \emph{the most probable frequency distribution} for
-% the larger-population activity during the recording. But it also approximates
-% three other distributions, as discussed in the previous section.
-
-\bigskip
-
-The maximum-entropy distribution for the larger population is calculated
-using five moments, because this number provides almost as much information
-as the full frequency distribution of the sample, as discussed in
-\sect~\ref{sec:marginalization}. Figure~\ref{fig:plots_5mom} shows the
-resulting distributions for three example values of
-larger-population sizes: $N=1\,000$ (\textcolor{mygreen}{green diamonds}),
-$N=5\,000$ (\textcolor{myred}{red circles}), $N=10\,000$
-(small \textcolor{myblue}{blue squares}). The frequency density of the sample
-activity is also shown (\textcolor{black}{black triangles}), and in the
-plot it would be indistinguishable from the maximum-entropy density for
-$N=n$, that is, applied at the sample level. We discuss the case of unknown
-$N$ in \sect~\ref{sec:N}.
-\begin{figure}[!p]
-  \centering
-\subcaption[]{\label{fig:plots_5mom_a}}%  
-\includegraphics[width=\linewidth]{_scripts/newpaper_stensola_5mom_distributionsN_priorc.pdf}%
-\\[1em]%\mbox{}\vfill\mbox{}%
-\subcaption[]{\label{fig:plots_5mom_b}}%  
-\includegraphics[width=\linewidth]{_scripts/newpaper_riehle_5mom_distributionsN_priorc.pdf}%
-\\[1em]%\mbox{}\vfill\mbox{}%
-\caption{Maximum-entropy distributions for the total activity of the larger
-  population, using $M={}$five moments and assuming several population sizes $N$.
-  In order to compare different $N$, the distributions are multiplied by
-  $N$ (obtaining densities) and the activity divided by $N$ (obtaining a
-  population average). The distributions for large $N$ have very different
-  features from the measured frequency distribution, and how their peaks
-  become more pronounced as $N$ increases.}
-\label{fig:plots_5mom}
-\end{figure}
-% maxent_solutions_different_moments_prior2_6moments.nb
-
-The most expensive calculation, for $N=10\,000$, takes less than 15 minutes
-on a laptop with two $2\;\textrm{GHz}$ cores. In all cases the moments were
-recovered with relative errors smaller than $10^{-12}$.
-% 1000:  3*10^-13 \%
-% 5000: 7*10^-13 \%, 581 s
-% 10000: 6*10^-11 \%, 737 s
-
-The figure shows that the distribution for the larger-population is more
-peaked than the measured frequency distribution for the sample; their
-difference increases with $N$. Most remarkably, for the first data set,
-\fig~\ref{fig:plots_5mom_a}, the distribution for the larger population has
-two distinct low-activity modes. For the second data set,
-\fig~\ref{fig:plots_5mom_b}, the distribution presents a small shoulder
-(barely visible) on the right of its mode.
-% , suggestive of two activity regimes. We frame no hypotheses about the
-% biological cause of these two modes (they could stem from the presence of
-% different kinds of cells or modules)
-Such features are clearly not present in the sample frequencies or in the
-maximum-entropy distribution at the sample level. The model thus suggests
-interesting features of the larger population. \textcolor{white}{If you
-  find this you can claim a postcard from us.}
-
-% maxent_solutions_different_moments_partial.nb
-Here is a toy example of possible uses of this model, based on the first
-data set. We could be interested in the hypothesis that two distinct cell
-types or assemblies be present in the region where the recording was made.
-Finding a larger-population distribution with two peaks, as in
-\fig~\ref{fig:plots_5mom_a}, would provide some evidence for this
-hypothesis. We might have reasons for suspecting that a specific set of the
-sampled neurons is of the first type, and the remaining of the second type.
-In the case of the first data set, indeed, 27 of the 65 sampled cells were
-identified as grid cells belonging to 3--4 functional modules, the
-remaining 38 \enquote{non-grid} cells being interneurons and unidentified
-cells \citep{stensolaetal2012}. Could the two peaks in the distribution of
-\fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
-cells? We apply the model to these two sets of neurons individually, using
-$\yNg=4\,150$ for the grid set and $\yNng=5\,850$ for the non-grid set, to
-reflect their proportions (27/65 and 38/65) in the recorded sample. The
-results are shown in \fig~\ref{fig:plots_cellsubsets_a}.
-\begin{figure}[!p]
-\centering
-\subcaption[]{\label{fig:plots_cellsubsets_a}}%  
-\includegraphics[width=\linewidth]{_scripts/newpaper3_gridvsnogrid.pdf}%
-\\[1em]%
-\subcaption[]{\label{fig:plots_cellsubsets_b}}%  
-\includegraphics[width=\linewidth]{_scripts/newpaper3_gridconv.pdf}%
-\\[1em]%
-\caption{\subcaptionref{fig:plots_cellsubsets_a} Maximum-entropy
-  distributions for larger subpopulations or grid and non-grid cells.
-  \subcaptionref{fig:plots_cellsubsets_b} The convolution of the two
-  subpopulation distributions is very different from the distribution for
-  all cells; this means that the two subpopulation distributions are not
-  independent.}
-\label{fig:plots_cellsubsets}
-\end{figure}
-The distribution for the larger population of grid cells
-(\textcolor{mygreen}{green triangles}) seems to have one broad peak, close
-to the first peak of the larger-population distribution. The distribution
-for the larger population of non-grid cells (\textcolor{myred}{red
-  circles}) has two peaks instead, roughly at the same population-averaged
-activities as the peaks of the larger-population distribution
-(small \textcolor{myblue}{blue squares}) but closer in height. It would thus seem
-that the population of grid cells is contributing to the first mode of the
-larger population, but it is not its sole contributor. We can also assess
-if the distributions for the two sets of neurons are independent. If they
-were independent, the larger-population distribution would be given by
-their convolution:
-\begin{equation}
-  \label{eq:convolution}
-  P_{\text{full}}(\yA) = \sum_{\yA'} P_{\text{grid}}(\yA')\;
-  P_{\text{non-grid}}(\yA-\yA') %, \qquad \yA \in \set{1,\dotsc,10\,000}
-\end{equation}
-where the index $\yA'$ runs from $\max(0,\yA-\yNng)$ to $\min(\yA,\yNg)$.
-But this is not the case: 
-figure~\ref{fig:plots_cellsubsets_b} shows that such convolution
-(\textcolor{black}{black diamonds}) is quite different from the
-larger-population distribution (small \textcolor{myblue}{blue squares}): the two
-peaks of the former are closer in position and height than those of the
-latter. The population distributions of grid and non-grid cells are
-therefore \emph{not} independent: knowledge of the activity of either set
-gives us some information about the activity of the other.
-
-\medskip
-
-The toy analysis above should not be taken literally, but just as an
-illustration of the model's possible applications. Instead of a grid cell
-\vs\ non-grid cell divisions we could for example have examined the modules
-of the cells. In the absence of such auxiliary information about the cells,
-one could in principle try all possible partitions of the recorded cells
-into two groups, and check whether some partition leads to independent
-distributions (computationally very expensive in practice); and so on.
-
-The important point is that this model is computationally very cheap and
-yet it can provide useful insights, even if just qualitative ones, and even
-suggest new hypotheses.
-
-\section{Quantifying the importance of higher-order correlations:\\
-  the limitations of models at the sample level}
-\label{sec:marginalization}
-
-As mentioned in the Introduction, in the neurosciences the maximum-entropy
-method has also been used to quantify the \enquote{cooperativity}
-\citep[\eg][]{gersteinetal1985} or \enquote{interaction}
-\citep[\eg][]{martignonetal1995,schneidmanetal2006,shlensetal2006} or
-\enquote{synchrony} \citep[\eg]{bohteetal2000,amarietal2003} of neuronal
-activity. Most, if not all, such applications use the method \emph{at the
-  sample level}, that is, they find a distribution for the sample without
-considering that it comes from a larger population. In this section we
-discuss how our proposed application to a larger population bears on this
-kind of quantification and compares with the application at the sample
-level. %We also highlight some methodological caveats.% For concreteness we use an
-% example with specific statistics: first and second moments (equivalent to
-% mean and correlations), versus first-to-fourth moments. But our
-% discussion holds for any sets of statistics.
-
-\bigskip
-
-As a first step, terms like \enquote{cooperativity}, \enquote{interaction},
-and similar need to be translated into a more precise, operational
-\citep[\cf][\chap~I]{bridgman1927_r1958} notion. Here we propose to base
-this kind of analysis on the notion
-of \emph{informational sufficiency}, because
-\begin{enumerate*}[label=(\roman*)]
-\item it seems closely related to those terms as used in the
-  literature; \item it is mathematically connected with maximum-entropy
-  distributions; \item it reveals important consequences of sampling.
-\end{enumerate*}
-Its meaning will now be briefly explained. For an extensive discussion of
-this notion, which was probably first suggested by Kullback \amp\ Leibler
-\parencite*{kullbacketal1951}, see Bernardo \etal\
-\parencite*[\sect~4.5]{bernardoetal1994} and Jaynes \parencite*[\chap~8 and
-\sect~14.2]{jaynes1994_r2003}.
-
-Our belief distribution about the frequencies of the neuronal activities,
-or about the activity in a new time bin, is conditional on all experimental
-data and statistics we have collected. This belief distribution, however,
-can be such that it remains unchanged if we discard part of the data or
-statistics -- for example, the third- and higher-order empirical
-moments. % --
-% but it does change if we discard more.
-This means that the discarded statistics are \emph{informationally
-  irrelevant}: knowing them does not change our beliefs. The remaining
-statistics -- in the example, first and second empirical moments -- are
-\emph{informationally sufficient}.%  The qualification \enquote{minimally}
-% will be omitted for brevity.
-% This informational quasi-sufficiency is interesting because
-% it may give clues about population organization and dynamics.
-
-It seems to us that the informational sufficiency of a set of $m$ moments
-captures the idea of the \enquote{importance} of the almost-simultaneous
-activity of $m$ or fewer neurons: such simultaneous activity is
-\enquote{unimportant} if its knowledge does not alter our forecasts. We are
-thus speaking of informational importance, which may or may not imply or be
-implied by biophysical importance -- \enquote{correlation is no proof of
-  causation} \citep[\eg][]{simon1954,freedman2006}. But it lends itself to a
-precise probabilistic analysis, with consequences determined by
-mathematical theorems, as we shall now see.
-
-Our question is whether some degree of \enquote{cooperation} is more
-\enquote{important} than some other degree, or not. In terms of
-informational sufficiency, its definition might suggest to compare the
-belief distributions conditional on two different subsets of statistics.
-But which measure should be used for such a comparison? and how to motivate
-it? We prefer to take a different approach, leading to an automatic answer:
-let us turn the informational sufficiencies of two different subsets of
-statistics into two \emph{hypotheses}. We can then ask: which hypothesis is
-more probable, given the data? Equivalently, what is the ratio of their
-probabilities? By translating our question about \enquote{cooperativity}
-into a hypothesis-testing problem about sufficiency, we can find a
-well-defined answer determined by the probability-calculus. % The ratio of
-% the two probabilities gives us a natural measure of the
-% For example, do
-% the data say that the first two empirical moments are probably enough to
-% determine our beliefs? or is it more probable that the first four are
-% needed instead?
-We shall see that this approach is indeed more powerful than a simple
-comparison of distributions.
-
-But such a translation does more: through a mathematical theorem \citep[and
-references therein]{andersen1970}[this theorem was first proved for
-continuous probability distributions
-by][]{koopman1936,pitman1936,darmois1935}[see][\sect~4.5.3 for further
-references]{bernardoetal1994}, it leads to the automatic appearance of
-maximum-entropy distributions: \emph{if a probability distribution for a
-  repetitive phenomenon has some sufficient statistics, then it is a
-  mixture of distributions of maximum-entropy form having those statistics
-  as constraints, and vice versa}. 
-% Thus, maximum-entropy distributions automatically appear (as
-% approximations) from the probability-calculus when we test hypotheses
-% about informational sufficiency.
-The most important consequence of this theorem is that, by Bayes's formula,
-the probability that some statistics are informationally sufficient can
-approximately be found by first building the maximum-entropy distribution
-constrained by those statistics, and then calculating the probability for
-the whole data under that distribution. We shall shortly illustrate this
-procedure with the data sets of \sect~\ref{sec:application}.
-
-% For probabilities built with the maximum-entropy method, this means
-% comparing those constrained on the subset and on the full set. This
-% procedure is actually equivalent to comparing the probabilities of the two
-% hypotheses about sufficiency, conditional on the \emph{full} data, assuming
-% their pre-data probabilities to be equal. We show this equivalence below
-% and perform the calculations for our first data set. \medskip
-
-\medskip
-
-Owing to the theorem above, studies about cooperativity (and similar ideas)
-that use maximum-entropy can be naturally reinterpreted as studies about
-informational sufficiency. An important fact must then be taken into
-account: \emph{informational sufficiency is not preserved under sampling}.
-If a probability distribution has some sufficient statistics, then its
-sample marginals \emph{cannot} have the same sufficient statistics, and
-vice versa \citep[\sect~3.1]{portamanaetal2015}; except for trivial cases
-(\eg\ uniform probability distributions). % This generally holds also among
-% different marginals: if a statistics is sufficient for some marginal, then
-% it will not be sufficient for another marginal.
-This impossibility is known in another guise in statistical mechanics: if a
-system is described by a Gibbs state, its subsystems cannot be perfectly
-described by Gibbs states \citep[\eg][and references
-therein]{maesetal1999}. % Mathematically this impossibility comes from the
-% Pitman-Koopman theorem mentioned above, and translates into the general
-% impossibility of solving a system of independent equations with more
-% equations than unknowns .
-
-This fact affects the interpretation of the results obtained from
-maximum-entropy distributions. For example, if means and pairwise
-correlations seem informationally sufficient for a sample from a brain
-area, then they are probably not sufficient for the larger population of
-neurons constituting that area, and vice versa. If we are interested in the
-cooperativity of the neurons in a brain area, it is therefore unreliable to
-use a maximum-entropy distribution constructed only for a sample
-thereof. % \citep[contrast this with some literature, \eg][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}
-The model presented here avoids this problem because it targets the larger
-population, not the sample alone.
-
-\bigskip
-
-We can compare the informational sufficiency of a set of moments, say
-$M''\defd\set{1,\dotsc,m''}$, with respect to another, say
-$M'\defd \set{1,\dotsc,m'}$, by introducing the quantity $\ydi(M'',M')$
-defined as follows:
-\begin{enumerate}[label=(\roman*)]
-\item From each maximum-entropy distribution $\px(\yA \| M,N)$ for the
-  larger population, \eqn~\eqref{eq:app_maxent_pop_first}, built from each
-  set of constraints $M=M',M''$, calculate the marginal distribution for
-  the sample by \eqn~\eqref{eq:conditional_hypergeometric_t}:
-\begin{equation}
-  \label{eq:sample_marginal}
-  \pxx(\ya \| M, N) = \sum_{\yA}\yG_{\ya \yA}\px(\yA \| M,N),
-  \qquad M=M',M''.
-\end{equation}
-\item Calculate the relative entropies of the measured frequency
-  distribution $\yf$ with respect to each sample marginal from the previous
-  step, and multiply them by the number of time bins $T$:
-  \begin{equation}
-    \label{eq:relentropy_T}
-    T\,\sh[\yf; \pxx(\ya \| M,N)] \defd
-    T  \sum_{\ya} \yff_{\ya} \log\frac{\yff_{\ya}}{\pxx(\ya \| M,N)},
-    \qquad M=M',M''.
-  \end{equation}
-\item Take the difference: 
-\begin{equation}
-  \label{eq:measure_suffic}
-  \begin{split}
-  \ydi_{N}(M'',M') &\defd
-  T \,\sh[\yf; \pxx(\ya \| M',N)]
-  -
-  T \,\sh[\yf; \pxx(\ya \| M'',N)]
-  \\ &\equiv
-T  \sum_{\ya} \yff_{\ya} \log\frac{
-  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M'',N)}{
-  \sum_{\yA}\yG_{\ya\yA}\px(\yA \| M',N)}.
-\end{split}
-\end{equation}
-\end{enumerate}
-The quantity $\ydi_{N}(M'',M')$ so defined is positive if $M''$ is more
-probable to be informationally sufficient than $M'$, and negative
-otherwise.
-
-We show in appendix~\ref{sec:measure_suff} that $\ydi_{N}(M'',M')$ is
-approximately equal to the log-ratio of the probabilities of the data $\yf$
-conditional on the hypotheses $M''$ and $M'$:
-\begin{equation}
-  \label{eq:bayes_factor}
-  \ydi_{N}(M'',M') \approx
-  \log\frac{\pf(\yf \| M'',N)}{\pf(\yf \| M',N)}.
-\end{equation}
-This is their \emph{relative weight of evidence}
-\citep[\chap~6]{good1950}{good1985}[\sect~1.4]{osteyeeetal1974}, the
-logarithm of their \emph{relative Bayes factor}
-\citep[p.~421]{jeffreys1936}[and references
-therein]{kassetal1995}. %[\chaps~V, VI, A]{jeffreys1939_r1983}.
-The exponential of $\ydi_{N}(M'',M')$ tells us how much more probable the
-data $\yf$ are, conditional on the sufficiency of $M''$, than conditional
-on the sufficiency of $M'$. We can also combine this measure with pre-data
-probabilities for the two hypotheses to obtain the ratio of their
-probabilities conditional on the data \citep[\cf][]{bretthorst2013}.
-
-\bigskip
-
-Let us illustrate the quantity just defined and the previous discussion
-about sufficiency and sampling by using our data sets. We want to compare,
-for example, three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$,
-consisting of the first two, four, five moments, for a larger population of
-size $N=10\,000$. Figure~\ref{fig:plots_2_4mom} shows the resulting
-distributions for the two data sets.
-\begin{figure}[!p]
-\centering
-\subcaption[]{\label{fig:plots_2_4mom_a}}%  
-\includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_threemom_distributions10000_priorc.pdf}%
-\\[1em]%
-\subcaption[]{\label{fig:plots_2_4mom_b}}%  
-\includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_threemom_distributions10000_prior2.pdf}%
-\\[1em]%
-\caption{Maximum-entropy distributions for the total activity of the larger
-  population, based on the first two, four, and five moments, and a
-  larger-population size $N=10\,000$. The two- and four-moment
-  distributions are very different, suggesting that two moments are not
-  informationally sufficient. The four- and five-moment distributions are
-  almost indistinguishable, suggesting that four and five moments are
-  equally likely to be sufficient.}
-\label{fig:plots_2_4mom}
-\end{figure}
-% maxent_solutions_different_moments_prior2_6moments.nb
-There is a clear visual difference between the distributions based on two
-(\textcolor{myred}{dashed red}) and on four (\textcolor{myblue}{solid
-  blue}) moments; for the first data set, panel
-\subcaptionref{fig:plots_2_4mom_a}, they even differ in the number of
-modes. On the other hand there is barely any visual difference between the
-distributions based on four and on five (\textcolor{myyellow}{dotted
-  yellow}) moments. From these visual differences we immediately conclude
-that two moments are less likely to be informationally sufficient than
-four; but four should be almost as likely as five. We shall see that the
-measure~\eqref{eq:measure_suffic} reflects this swift visual judgement.
-
-Compare these plots with those obtained applying the model at the sample
-level (equivalent to $N=n$) to the first data set, shown in
-\fig~\ref{fig:comparisons_marginals}. The distributions based on two
-(\textcolor{myred}{red circles}), four (\textcolor{myblue}{blue empty
-  squares}), and five (\textcolor{myyellow}{yellow lozenges}) moments can
-only be distinguished on a logarithmic scale, and only in their tails. The
-measured frequencies (black triangles) are also shown. Two moments seem
-slightly less likely to be sufficient than four; and four and five moment
-seem about equally likely. But it is difficult to visually make these
-comparisons.
-
-\medskip
-
-% NUMERICAL RESULTS
-%
-% relentropies of sample2mom, sample4mom, sample5mom, marg2mom, marg4mom, marg5mom
-% -322.72,-42.8005,-39.4917,-85.7626,-4.23725,-4.20059 nats =
-% -1401.56,-185.88,-171.51,-372.462,-18.4021,-18.2429 dB
-%
-% difference in relentropies between 4 mom and 2 mom:
-%** sample level: 279.92 nats = 403.839 bit = 1215.68 dB
-%** marg level: 81.525 nats =  117.616 bit = 354.06 dB
-% difference in relentropies between 5 mom and 4 mom:
-%** sample level: 3.30882 nats = 4.77362 bit = 14.37 dB
-%** marg level: 0.0366623 nats = 0.0543836 bit = 0.159223 dB
-% correspond to prob ratios of
-% sample: 3.69524\times 10^{121}, 27.3529
-% marg: 2.5469\times 10^{35}, 1.03734
-% 
-% comparison N vs n:
-% 2 mom: 236.958 nats = 341.858 bit = 1029.09 dB, 8.11758\times 10^{102}
-% 4 mom: 38.5632 nats = 55.635 bit = 167.478 dB, 5.59496*10^16
-% 5 mom: 35.2911 nats = 50.9157 bit = 153.267 dB, 2.12185*10^15
-%
-% comparison_ME_sample_full_levels.nb
-%
-Calculation of the weight of evidence~\eqref{eq:measure_suffic} for the
-first data set and $N=10\,000$, corresponding to
-\fig~\ref{fig:plots_2_4mom_a}, leads to the following results:
-\begin{equation}
-  \label{eq:diff_suff_N}
-  \begin{split}
-  \ydi_{N}(M_{4},M_{2}) &= 118\;\bit \equiv 35.4\;\hart,
-\\    \ydi_{N}(M_{5},M_{4}) &= 0.0544\;\bit \equiv 0.0159\;\hart,
-  \end{split}
-\end{equation}
-where the Hartley ($\hart$) denotes base-10 logarithms
-\citep[\sect~C.4]{iso2009}[called \enquote{ban} and used by Turing and Good
-in their code-breaking work at Bletchley Park:
-see][]{good1985%,good1950,good1969
-}[\sect~4.2]{jaynes1994_r2003}. In words, our data are 35 orders of
-magnitude more probable by assuming sufficiency of $M_{4}$ than by assuming
-sufficiency of $M_{2}$. But the data are about as probable
-($10^{0.016} = 1.04$) by assuming sufficiency of $M_{4}$ as of $M_{5}$. The
-qualitative visual information of \fig~\ref{fig:plots_2_4mom_a} was
-therefore reliable.
-
-For the model at the sample level, that is, with $N=n$, the weights of
-evidence are
-\begin{equation}
-  \label{eq:diff_suff_n}
-  \begin{split}
-  \ydi_{n}(M_{4},M_{2}) &= 404\;\bit \equiv 122\;\hart,
-\\    \ydi_{n}(M_{5},M_{4}) &= 4.77\;\bit \equiv 1.44\;\hart.
-\end{split}
-\end{equation}
-So the application at the sample level says that the data are 120 orders of
-magnitude more probable conditional on $M_{4}$ than on $M_{2}$, and
-$10^{1.4}=25$ times more probable conditional on $M_{5}$ than on $M_{4}$.
-These results are different from those of the application at the larger
-population, and they would be hard to deduce visually from the
-distributions of \fig~\ref{fig:comparisons_marginals}.
-
-% The conclusions about \enquote{cooperativity} or \enquote{interaction} that
-% we reach by applying the maximum-entropy method to the larger population,
-% as proposed here, are therefore different from those applying it at the
-% sample level -- and also visually clearer. The plot in 
-% \fig~\ref{fig:plots_2_4mom_a}, showing the two clearly different
-% distributions constrained by two and four moments, should be compared with
-% the plot for the distributions obtained at the sample level under the same
-% constraints, shown in \fig~\ref{fig:comparisons_marginals} as
-% \textcolor{myred}{red circles} and \textcolor{myblue}{blue squares}. The
-% measured frequencies (black triangles) are also shown. It is necessary to
-% use a logarithmic scale to see the differences, which appear mainly in the
-% tail.
-\begin{figure}[!p]
-\centering
-% \includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
-% \\[3em]%
-\includegraphics[width=0.92\linewidth]{_scripts/trunct2_comparison_marginals.pdf}%
-\caption{Maximum-entropy distributions constructed at the sample level,
-  using the first two, four, and five moments; and sample marginal of the
-  distribution constructed with five moments for a larger population with
-  $N=10\,000$ (dotted yellow in \ref{fig:plots_2_4mom_a}). The
-  distributions are multiplied by $T$, obtaining absolute frequencies; note
-  that activities with frequency lower than 1 cannot be observed in $T$
-  bins. The distributions at the sample level are only distinguishable on a
-  logarithmic scale. The five-moment distribution constructed at the
-  larger-population level fits the frequency data better than the one at
-  the sample level.}
-\label{fig:comparisons_marginals}
-\end{figure}
-% comparison_ME_sample_full_levels.nb
-
-\medskip
-
-% 4 mom: 38.5632 nats = 167.478 dB, 5.59496*10^16
-% 5 mom: 35.2911 nats = 153.267 dB, 2.12185*10^15
-
-It is possible to use the measure~\eqref{eq:measure_suffic} and its
-probabilistic interpretation~\eqref{eq:bayes_factor} to compare the
-probability of the data $\yf$ conditional on the hypothesis $(M_{5},N)$ of
-five-moment sufficiency at the larger-population level, $N=10\,000$, versus
-the hypothesis $(M_{5}, n)$ of five-moment sufficiency at the sample level,
-$N=n=65$. We obtain
-\begin{equation}
-  \label{eq:diff_suff_n}
-  \ydi[(M_{5}, N), (M_{5}, n)] = 50.9\;\bit \equiv 15.3\;\hart.
-\end{equation}
-% that is, the data are 15 orders of magnitude more probable under the first
-% hypothesis than under the second. 
-This result is also shown in
-\fig~\ref{fig:comparisons_marginals}. We see that the five-moment
-distribution constructed at the larger-population level
-(\textcolor{myyellow}{yellow stars}) fits the measured frequency
-distribution (black triangles) more closely than the corresponding
-sample-level distribution (\textcolor{myyellow}{yellow lozenges}).
-
-\bigskip
-
-% We close this section with some references. The notion of informational or
-% predictive sufficiency, probably first presented by Kullback \amp\ Leibler
-% \parencite*{kullbacketal1951}, is discussed in Bernardo \etal\
-% \parencite*[\sect~4.5]{bernardoetal1994}, who explain its connections with
-% Fisher's\parencite*{fisher1922} parametric sufficiency. Jaynes
-% \parencite*[\chap~8 and \sect~14.2] shows the connections with information
-% theory. The mathematical relation between sufficiency and the exponential
-% family comes from the Pitman-Koopman
-% theorem \citep{koopman1936,pitman1936,darmois1935} for continuous variables.
-
-% \mynote{Short reasoned bibliography about sufficient statistics and its
-%   relation to maximum-entropy?} blah blah \footnote{For more technical results and
-%   connections with the notion of symmetry see \eg\
-%   \cite{darmois1935,neyman1935,koopman1936,pitman1936,halmosetal1949,bahadur1954,berk1972,lauritzen1974,lauritzen1982_r1988,lauritzen2007,cifarellietal1980,cifarellietal1981,diaconisetal1981,diaconis1992,furmanczyketal1998,fortinietal2000,nogalesetal2000,kallenberg2005,ayetal2015}.}
-% This mathematical result comes from the Pitman-Koopman
-% theorem\citep{kullbacketal1951}[the
-% notion goes back to][]{fisher1922} \citep{jaynes1982b}[\sect~4.5.4]{bernardoetal1994}
-% \citep{koopman1936,pitman1936,darmois1935}[for later analyses and the
-% discrete case
-% see][]{andersen1970,%discrete
-% denny1967,denny1972}
-
-% But does the application at the larger-population level lead to appreciably
-% different results from that at the sample level? after all we're interested
-% in an approximate informational sufficiency, not in an analytically exact
-% one. A correct answer can only be given case by case.
-% Figure~\ref{fig:comparisons_marginals} gives a graphical answer to this
-% question in the case of our first data set. The upper plot shows the
-% maximum-entropy distributions for a larger population of $10\,000$ neurons,
-% constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
-% from four moments (\textcolor{mypurpleblue}{\textbf{------}}). The lower
-% plot shows the maximum-entropy distributions for the sample, from two
-% moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four moments
-% (\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained
-% by marginalizing the larger-population distribution to the sample size, from two
-% moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
-% moments (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The measured
-% frequency distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also
-% shown. We note the following:
-% \begin{itemize}
-% \item The application to the larger-population (upper plot) leads to two
-%   completely different distributions (even different number of modes);
-%   clearly the two sets of statistics are not even approximately equivalent
-%   for inferential purposes.
-% \item The application at the sample level leads to deceivingly similar
-%   distributions (\textcolor{mygreen}{$\bm{\blacktriangledown}$} and
-%   \textcolor{mypurpleblue}{$\bm{\square}$}), which can only be distinguished on a
-%   logarithmic scale. This could lead to the erroneous conclusion that the
-%   two statistics are approximately equivalent.
-% \item The difference between marginals from the larger-population distribution
-%   and the distributions obtained at the sample level (filled vs empty
-%   markers) is larger than the difference between different statistics
-%   (triangles vs squares).
-% \item The marginals of the application to the larger population are closer to
-%   the measured frequencies than the distributions obtained at the sample
-%   level (although this closeness is not a valid criterion for their
-%   goodness).
-% \end{itemize}
-% Therefore, the maximum-entropy application at the sample level and at the
-% larger-population level lead to different results; the latter is not only more
-% meaningful but also superior because it shows clearly the different
-% informational sufficiency of the two sets of statistics.
-
-\section{The size $N$ of the larger population}
-\label{sec:N}
-
-% results for post-data probabilities of N given data and 5 moments:
-% N=65,1000,2000,5000,10000
-% 1.40169\times 10^{-16}, 0.0439602, 0.139657, 0.251064, 0.297726, 0.267594
-%
-% using 1/N pre-data prob:
-% 1.04099\times 10^{-14}, 0.212211, 0.337085, 0.242394, 0.143722, 0.0645882
-% 
-% maxent_solutions_different_moments_prior2_6moments.nb
-
-The calculations presented in the preceding sections depend on the size $N$
-of the larger population, which of course cannot be the whole brain. How
-large should or can $N$ be in our formulae?
-
-The answer comes by examining the origin of our belief distribution for the
-activity of the sample \emph{conditional} on the activity of the larger
-population. It is given by the hypergeometric
-distribution~\eqref{eq:conditional_hypergeometric}, reprinted here:
-\begin{equation}
-  \labelbis{eq:conditional_hypergeometric}
-  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
-    \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
-\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}.
-\end{equation}
-This distribution leads to the equality of factorial
-moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests. This
-conditional probability, characteristic of \enquote{drawing without
-  replacement}
-\citep[\eg][\chap~3]{jaynes1994_r2003}, % [\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}
-ensues when we believe, with the same degree, that any of the $N$ units
-(balls or neurons) could be drawn or sampled. Thus, consider the pool of
-all neurons which we believe -- with an \emph{equal} degree of belief for
-each neuron -- could have been recorded. Then $N$ is the size of that pool.
-$N$ therefore depends on factors such as: the shape, dimensions, and
-technical specifications of the recording probe; the inaccuracy in the
-insertion of the probe, leading for example to slightly different insertion
-points or angles; the density of neurons around the probe. But it also
-depends on the homogeneity of the brain region where the recording is made:
-if we believe that it does not matter whether the probe had been inserted
-in some other point of the same brain region, then
-formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
-therefore only be assessed case by case.
-
-Whether this equality in our degrees of belief is justified, or for which
-sampling procedures it can be justified, is the fundamental, deep question
-of sampling theory, which we cannot discuss here \citep[see \eg\ the
-discussions and references
-in][]{ericson1969,smith1976}. % {kruskaletal1979c,kruskaletal1980}
-It is of course possible to probabilistically assess, case by case, whether
-this symmetry assumption is appropriate. But we must also be aware that
-such assessment in turn rests on analogous symmetry assumptions at a higher
-level, and so on. The probability-calculus, just like the truth-calculus,
-cannot yield conclusions without premisses
-\citep{hailperin2011}[p.~182]{johnson1924}.
-
-The plots of \fig~\ref{fig:plots_5mom} nevertheless suggest that the order
-of magnitude of $N$ ought to be enough for qualitative inferences. As
-discussed in the next section, even if the exact number $N$ were known, the
-maximum-entropy distribution ought to be interpreted qualitatively or
-semi-quantitatively.
-
-\medskip
-
-If our uncertainty about $N$ spans several orders of magnitude we can
-assign a probability to the possible values of $N$ based on our background
-information and the frequency data $\yf$, similarly to the procedure in
-\sect~\ref{sec:marginalization} and appendix~\ref{sec:measure_suff}. The
-probability for the data $\yf$ conditional on a set of constraints $M$ and
-size $N$ is the exponential of the negative relative
-entropy~\eqref{eq:relentropy_T}:
-\begin{equation}
-  \label{eq:prob_data_from_N}
-  \pf(\yf \| M, N) \approx \binom{T}{T\yf}\;\prod_{a}\pxx(\ya \| M, N)^{T\,\yff_{a}}
-  \approx \exp\bigl\{  -T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
-\end{equation}
-If our pre-data probability for $N$ is $p(N)$, then by Bayes's theorem
-\begin{equation}
-  \label{eq:prob_N_from_data}
-  \pf(N \| M, \yf) \propto
-  p(N)\;\exp\bigl\{-T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
-\end{equation}
-
-Here is an illustrative example with our first data set. Suppose our
-uncertainty spans slightly more than one order of magnitude, from $N=1\,000$
-to $N=20\,000$. Divide this range roughly into thirds of order of
-magnitude, considering the values
-$N\in \set{10^{3}=1\,000, 10^{3+1/3}\approx 2\,000, 10^{3+2/3}\approx
-  5\,000, 10^{4}=10\,000, 10^{4+1/3}\approx 20\,000}$. Assuming the set $M$
-of first five moments to be sufficient, formula~\eqref{eq:prob_data_from_N}
-gives
-\begin{equation}
-  \label{eq:data1_f_from_N}
-  \begin{gathered}
-  \pf(\yf \| N=1\,000, M) =  0.00222,\qquad
-  \pf(\yf \| N=2\,000, M) =  0.00704,\\
-  \pf(\yf \| N=5\,000, M) =  0.0127,\qquad
-  \pf(\yf \| N=10\,000, M) =  0.0150,\\
-  \pf(\yf \| N=20\,000, M) =  0.0135.
-\end{gathered}
-\end{equation}
-Since our uncertainty regards a scale factor, it can be represented by
-equal pre-data probabilities of $1/5$ about these partial orders of
-magnitude. Thus from~\eqref{eq:data1_f_from_N} we find
-\begin{equation}
-  \label{eq:data1_N_from_f}
-  \begin{gathered}
-  \pf(N=1\,000 \| \yf, M) =  0.044,\qquad
-  \pf(N=2\,000 \| \yf, M) =  0.140,\\
-  \pf(N=5\,000 \| \yf, M) =  0.251,\qquad
-  \pf(N=10\,000 \| \yf, M) =  0.298,\\
-  \pf(N=20\,000 \| \yf, M) =  0.267,
-\end{gathered}
-\end{equation}
-which gives a slightly higher probability to $N=10\,000$.
-
-\bigskip
-
-In this way we can make inferences % -- for example about the sufficiency of
-% a set of moments, or about the marginal sample distribution --
-that take into account our uncertainty about $N$. We must make sure to
-avoid circularities: for example, we cannot assume a set of moments
-$M$ to be sufficient and assess our uncertainty about $N$ conditional on
-it, and then use this uncertainty to assess the sufficiency of $M$. But we
-can approximately assess the sufficiency of a subset of moments
-$M' \subset M$ taking into account the uncertainty of $N$ conditional on
-$M$.
-
-A rigorous assessment would involve a more expensive, full Bayesian
-calculation; but such calculation would make the maximum-entropy
-method superfluous. We discuss this final point in the next section.
-
-\section{Assumptions and approximations that define the model}
-\label{sec:assumptions}
-
-The possible uses of the model here presented depend on the assumptions and
-approximations which define it: three main assumptions and two main
-approximations all in all. They become clear within a full-fledged
-probabilistic approach. We summarize them here, leaving a more mathematical
-sketch to appendix~\ref{sec:inference_prob}.
-
-The first and most important assumption, typical of this kind of
-maximum-entropy application in the neurosciences, can be expressed in two
-equivalent ways:
-\begin{enumerate*}[(\roman*)]
-  % \item the order of the time bins has no relevance;
-\item the frequencies of the activity during a recording are
-  informationally sufficient for inferences about unrecorded times; \item
-  our degree of belief about the sequence of activities is invariant under
-  permutations of the time bins.
-\end{enumerate*}
-Equating time averages and expectations, \eqn~\eqref{eq:final_constraint},
-would not make sense without this assumption. Its mathematical consequence
-is to relate the probability for the frequency distribution $\yF$ for the
-activity of the larger population during the recording, to the probability
-for the \emph{long-run} frequency distribution. In statistical physics, as
-a comparison, the maximum-entropy method is instead used with the
-assumption of permutation-invariance across \emph{repetitions} of the same
-experiment, because its constraints represent macroscopically
-\emph{reproducible} quantities and kinematics, even in non-equilibrium
-\citep{jaynes1985b_r1996,%jaynes1957,%tikochinskyetal1984,
-  grandy1988,%berryetal1988,
-  deroecketal2005_r2006%,touchette2009
-}[for experiments \eg][]{guntonetal1983}. The kinetic-Ising models
-mentioned in \sect~\ref{sec:intro} are closer to this latter approach.
-
-The second assumption, typical of maximum-entropy in general, concerns the
-probability distribution for the long-run frequencies ensuing from the
-first assumption \citep[and references therein]{portamana2017}. It is taken
-to be either an \enquote{entropic prior}
-\citep{skillingetal1984,rodriguez1991,neumann2007}, which takes into
-account the multiplicity of long-run frequencies, or a prior for a model
-with sufficient statistics, which expresses the sufficiency of the moments
-under study, according to the theorem discussed in
-\sect~\ref{sec:marginalization}. This assumption is crucial for the
-maximum-entropy method based on the Shannon entropy. Alternative prior
-densities lead to alternative maximum-entropy methods; for example one
-based on the Burg entropy \citep{burg1975} if a Dirichlet density is used
-\citep{jaynes1986d_r1996}. The entropic prior depends on a reference
-distribution and on a coefficient $L$ which determines the sharpness of the
-prior's peak. The computations of \sects~\ref{sec:application}
-and~\ref{sec:marginalization} were insensitive to the choice of the
-reference distribution in their significant digits. We used three
-biologically reasonable alternatives: a flat distribution, one linearly
-decreasing with $\yA$, and a normal one with a broad peak at low
-activities.
-
-The third assumption is specific to our application: all subsets of $\yA$
-neurons in the larger population are equally likely to give rise to total
-activity $\yA$, at every time bin. A mathematically equivalent assumption
-is that $n$ neurons are sampled anew at each time bin. This assumption
-allows us to factorize the joint probability of the sequence of activities
-and apply the sampling relation~\eqref{eq:conditional_hypergeometric} of
-appendix~\ref{sec:derivation_maxent_fullnet} at each bin.
-
-Finally, the maximum-entropy method is typically equivalent to
-approximating the number $T$ of time bins, the prior coefficient $L$, and
-the ratio $T/L$ to infinity (that is, $T\gg L \gg 1$). These approximations
-lead to the identification of the four distributions listed at the end of
-\sect~\ref{sec:model}: the recorded frequency distribution becomes equal to
-the long-run one, and also to the probability distribution for the activity
-at recorded and new time bins. In our application it also leads to the
-equality of the hypergeometric
-distribution~\eqref{eq:conditional_hypergeometric} and the \emph{conditional}
-frequencies of the sample activity $\ya$ given $\yA$. The goodness of this
-approximation decreases as the ratio -- roughly $nN/T$ -- between the
-number of possible joint states and the number of bins increases, and also
-as observed quantities approach their mathematical bounds, for example some
-frequencies $\yff_{\ya}=0$. Without these approximations it is still
-possible to interpret the maximum-entropy
-distribution~\eqref{eq:app_maxent_pop_first} as the mode of the probability
-distribution for the frequencies $\yF$, that is, the most probable
-recorded frequency.
-
-\medskip
-
-The first and third assumptions above are the least realistic, but they
-make sense as reference hypotheses for judging the informational relevance
-of correlations of some order, as discussed in
-\sect~\ref{sec:marginalization}, to be compared with hypotheses about more
-realistic time dependences and
-clustering. %They may turn out to approximate well more realistic assumptions.
-
-The $T= \infty$ approximation (together with the third assumption above)
-can break our model for specific distributions of sample frequencies, if
-too many moments are used. For example, take $n=2$ with frequency
-distribution $\yf=(0.5,0,0.5)$ for the activities $\ya=0,1,2$; that is, the
-sample is silent half of the time and completely active half of the time.
-The first two normalized factorial moments have both value $0.5$, and
-together they uniquely determine $\yf$. If we now take $N=3$, it can be
-proved that \eqn~\eqref{eq:app_maxent_pop_first}, constrained on those two
-moments, has no solution. The reason is clear from a sampling perspective:
-if there is zero probability of sampling one active neuron, then the larger
-population cannot have any active neurons at all, and a fortiori there must
-be zero probability of sampling two active ones. But this reasoning holds
-only for each single time bin, not for the averages over time bins. The
-$T=\infty$ approximation identifies the two instead. Outside of this
-approximation there are no contradictions because the conditional
-frequencies and conditional probabilities are kept distinct.
-
-The two infinity approximations are at the boundary of their validity in
-our examples. For this reason we interpret the maximum-entropy distribution
-as the \emph{most probable frequency distributions}, as mentioned at the
-end of \sect~\ref{sec:model}. An important question, then, is \emph{which
-  features of the maximum-entropy frequency distribution are typical of the
-  majority of plausible frequency distributions?} Especially for features
-such as number, position, height of modes, or tail behaviour. Preliminary
-studies using a full probability calculation (to be published separately)
-show that several semi-quantitative features of the maximum-entropy
-distribution are indeed typical. For the first data set, for example, most
-frequency distributions have two high-frequency regimes of activity, as in
-\fig~\ref{fig:plots_5mom_a}, with roughly the same heights and almost the
-same locations, within a few percentages of population-averaged activity.
-Each of the two high-frequency regimes, though, may consist of two or three
-very close modes rather than one. The same preliminary studies also
-indicate that these typical features are insensitive to alternative prior choices
-in the second assumption above, for example using a Dirichlet prior instead
-of an entropic one.
-
-This suggests that the maximum-entropy approximation presented here can be
-favourably used to get a qualitative or semi-quantitative idea of what a
-fully probabilistic calculation could yield -- but with a far lower
-computational cost than the latter. For example, for $N=1\,000$ the
-maximum-entropy distribution can be calculated in ten minutes on a laptop,
-whereas the full probability calculation takes several days on a
-high-performance computing cluster.
-
-% \medskip
-
-% The second point is that our degrees of belief about the frequency
-% distribution for the larger population depend not only on the measured data in
-% the sample, but also on our pre-data beliefs $\yI$ about the distribution.
-% Which assumptions lead to the maximum-entropy result? This distribution
-% appears when our initial belief about the possible frequency distributions
-% $\yF$ is quantified by an entropic prior
-% \citep{neumann2007,rodriguez1991,skilling1998,catichaetal2004,portamana2017}:
-% \begin{equation}
-%   \label{eq:entropic_prior}
-%   \pf(\yF \| \yI) \propto  \exp[-\yL\;\re(\yF;\yR)]
-%   \approx \binom{L}{L\yFF_{0}, \dotsc, L\yFF_{N}}\;
-%   \prod_{\yA}{\yRR_{\yA}}^{L\yFF_{\yA}}
-% \end{equation}
-% where $\re(\yF;\yR)\defd \sum_{\yA}\yFF_{\yA}\log(\yFF_{\yA}/\yRR_{\yA})$ is
-% the relative entropy or discrimination information
-% \citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}, $\yR$ is the
-% reference distribution, and $\yL$ a positive parameter. The approximate
-% equality (obtained through Stirling's approximation), where the large
-% parentheses denote a multinomial coefficient, shows that this prior belief
-% is proportional to the number of ways in which the distribution $\yF$ can
-% be realized in $L$ time bins. The parameter $L$ roughly quantifies how many
-% time bins our data set must have to affect our initial belief. The
-% maximum-entropy approximation is valid when $L$ is large, but small
-% compared to the sharpness of the constraints on $\yF$; in our case this
-% means $\yL \approx 10$, give or take an order of magnitude.
-
-\section{Summary and discussion}
-\label{sec:summary_discussion}
-
-The reason why we record neurons from a brain region is to have an idea of
-the activity of the other neurons in that region, similarly to survey
-sampling. If we thought that \emph{precisely} the recorded ones were
-exceptions or completely unrelated to the others surrounding the recording
-probe, our recording would serve little purpose. Probabilistic models that
-allow us to make direct inferences about the larger population are
-therefore useful. But they become complex and sensitive if we ask for very
-detailed inferences.
-
-We have presented a model that allows us to make direct inferences about a
-larger population from very small samples. Its inferences regard only the
-frequencies of the total activity of the larger population, but such
-limited inferential scope makes the model computationally very cheap and,
-most important, it still leads to inferences that are not at all trivial
-and could not be made just by looking at the sample and its distributions.
-We showed this by applying the model to two real data sets.
-
-By \enquote{inference} we mean the quantification of a degree of belief
-about possible observations, based on specific assumptions; not the
-response of an oracle. The model here presented gives us a \emph{forecast}
-based on specific assumptions discussed in \sect~\ref{sec:assumptions}, and
-therefore has two main uses, just like any other probabilistic model. If
-our assumptions are only hypotheses under testing, then the subsequent
-confirmation or confutation of the forecast will increase or decrease our
-confidence in those hypotheses. If we strongly trust our assumptions, on
-the other hand, then we rely on the forecast and even use it in place of
-actual observations, especially if the latter are difficult to make. (In
-fact, we typically shift from the first to the second use.) In the
-preceding sections we have given examples of additional hypotheses that
-could be explored with the present model. The assumptions of the model
-itself, discussed in \sect~\ref{sec:assumptions}, can either be under test
-or relied upon, depending on the case.
-
-Some approximations implicit in the model make its numerical results
-semi-quantitative with respect to an exact analysis. The results should not
-be off by more than a few percentages for the population-averaged activity and
-for the frequency density. This imprecision, however, is compensated by the
-extreme computational cheapness of the model, far less costly than an exact
-analysis. The model is therefore very convenient at least for essaying
-hypotheses. We hope to present an exact quantitative comparison with the
-full probabilistic approach in a future publication.
-
-% \mynote{cite  \cite[p.~440]{maxwell1873c}}
-
-% We have presented a procedure to construct the most plausible frequency
-% distribution of population-averaged activities of a population of neurons,
-% given the recording about a small sample thereof. This procedure combines
-% the maximum-entropy method and basic identities from sampling theory. From
-% the application to two real data sets we saw that the frequency
-% distributions obtained with our procedure can have features very different
-% from the one measured in the sample, such as multiple modes. This procedure
-% can also be used with moment constraints of different order -- means,
-% population-averaged pairwise correlations, or higher-order correlations --
-% thus giving an approximate assessment of the informational sufficiency of
-% specific subsets of moments. In fact, we saw that the application of
-% maximum-entropy only at the sample level leads to misleading results about
-% this kind of sufficiency questions.
+(4) {\footnotesize\textsf{\footnotesize I enjoyed somewhat philosophical arguments regarding the interpretation of the ME distribution (P8 and Appendix).
+However, I was not sure how these are relevant to the paper if it is just an interpretation.
+Can we consider different outcomes arising the different interpretations?
+If the authors are preparing another manuscript in this direction, these arguments may better be used for that paper?
+(This is just a weak suggestion.)}\par}
+
+\addsec{Reviewer 2}
+\label{sec:rev2}
+
+2.
+{\footnotesize\textsf{\footnotesize The method requires that the measurements on n neurons are 'representative' of the full population of N neurons.
+The authors show one example, where they divide a dataset into two subsets of neurons with different statistics.
+They show that the extrapolations from these two subsets are significantly different.
+But this particular analysis is somewhat ad hoc.
+Are there any more systematic tests of procedures for understanding how sensitive this analysis is to choosing different subsets of neurons?
+Or different recoding periods?}\par}
+
+3.
+{\footnotesize\textsf{\footnotesize In particular, correlations tend to be stronger between nearby neurons than far-away neurons.
+This should be a consistent pattern that shows up in most experimental datasets and which should lead to systematic biases in inferring the properties of populations, N, which are large enough to result in lower average correlation (see Nonnenmacher et al., PLoS CB 2017).
+Is there any way to take this dependence into account?}\par}
+
+%%%% examples use empheq
+%   \begin{empheq}[left={\mathllap{\begin{aligned}    \de\yF_{\yc}/\de\yp&=0\text{:} \\
+%         \de\yF_{\yc}/\de\ym&=0\text{:}\\ \de\yF_{\yc}/\de\yl&=0\text{:}\end{aligned}}\qquad}\empheqlbrace]{align}
+%     \label{eq:con_p}
+% %    \de\yF_{\yc}/\de\yp &\equiv
+%     -\ln\yp + \ln\yq + \yl\yM + \ym\yu &=0,\\
+%     \label{eq:con_u}
+% %    \de\yF_{\yc}/\de\ym &\equiv
+%     \yu\yp-1 &=0,\\
+%     \label{eq:con_l}
+%     %\de\yF_{\yc}/\de\yl &\equiv
+%     \yM\yp-\yc &=0.
+%   \end{empheq}
+%%%%
+% \begin{empheq}[box=\widefbox]{equation}
+%   \label{eq:maxent_question}
+%   \p\bigl[\yE{N+1}{k} \bigcond \tsum\yo\yf{N}\in\yA, \yM\bigr] = \mathord{?}
+% \end{empheq}
+
+% \[
+%   \begin{tikzcd}
+%       M_{n,n}(\CC) \arrow{r}{R'_{a}(\Hat{U})} & M_{n,n}(\CC)
+%     \\
+%     L(\mathcal{H}) \arrow{r}{\Hat{U}} \arrow[swap]{d}{R_*}\arrow[swap]{u}{R'_*} & L(\mathcal{H}) \arrow{d}{R_*}\arrow{u}{R'_*} \\
+%       M_{n,n}(\CC) \arrow{r}{R_{a}(\Hat{U})} & M_{n,n}(\CC)
+%   \end{tikzcd}
+% \]
+
+% \[
+%   \begin{tikzcd}
+%       \CC^n \arrow{r}{R'_*(A)} & \CC^n
+%     \\
+%     \mathcal{H} \arrow{r}{A} \arrow[swap]{d}{R}\arrow[swap]{u}{R'} & \mathcal{H} \arrow{d}{R}\arrow{u}{R'} \\
+%       \CC^n \arrow{r}{R_*(A)} & \CC^n
+%   \end{tikzcd}
+% \]
+
+% \[
+%   \begin{tikzcd}
+%     \mathcal{H} \arrow{r}{A} \arrow[swap]{d}{R} & \mathcal{H} \arrow{d}{R} \\
+%       \CC^n \arrow{r}{R_*(A)} & \CC^n
+%   \end{tikzcd}
+% \]
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\iffalse
 \begin{acknowledgements}
-  This work is financially supported by the Kavli Foundation and the Centre
-  of Excellence scheme of the  Research Council of Norway.
-
-  We thank Iv\'an Davidovich for very insightful and valuable comments on a
-  previous draft of this manuscript. PGLPM thanks the staff of the NTNU
-  library for their always prompt support; Mari, Miri, Emma for
-  continuous encouragement and affection; Buster Keaton and Saitama for
-  filling life with awe and inspiration; the developers and maintainers of
-  \LaTeX, Emacs, AUC\TeX, Open Science Framework, Python, Inkscape, Sci-Hub
-  for making a free and unfiltered scientific exchange possible.
+  \ldots to Mari \amp\ Miri for continuous encouragement and affection, and
+  to Buster Keaton and Saitama for filling life with awe and inspiration.
+  To the developers and maintainers of \LaTeX, Emacs, AUC\TeX, Open Science
+  Framework, R, Python, Inkscape, Sci-Hub for making a free and impartial
+  scientific exchange possible.
 %\rotatebox{15}{P}\rotatebox{5}{I}\rotatebox{-10}{P}\rotatebox{10}{\reflectbox{P}}\rotatebox{-5}{O}.
 %\sourceatright{\autanet}
+\mbox{}\hfill\autanet
 \end{acknowledgements}
+\fi
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-%\clearpage
+\bigskip
 % %\renewcommand*{\appendixpagename}{Appendix}
@@
 % %\appendixpage
-\appendix
-
-\section{Derivation of the maximum-entropy distribution}
-\label{sec:derivation_maxent_fullnet}
-
-Here is a summary derivation of the maximum-entropy distribution for the
-larger population constrained by a set of factorial moments.%  For further details
-% see Porta Mana \etal\ \citey{portamanaetal2015}.
-
-\bigskip
-
-First of all the sampling relation. We have a population of $N$ units
-(neurons), $\yA$ of which have some specific property (being active in a
-specific time bin), and we sample $n$ of the $N$ units in such a way
-(partly uncontrollable experimental procedure) that we are led to believe,
-with the same degree, that any of the $N$ units could have been sampled.
-The probability that $\ya$ of the $n$ sampled neurons are active is then
-given by the hypergeometric distribution
-\begin{equation}
-  \label{eq:conditional_hypergeometric}
-  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
-  \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
-\equiv  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
-\end{equation}
-typical of \enquote{drawing without replacement}
-\citep[\eg][\chap~3]{jaynes1994_r2003}. %[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}[\sects~2.1, 3.2]{jeffreys1939_r1983}.
-In the following we leave $n$, $N$ implicit in the conditional. If we are
-uncertain about the number $\yA$, with belief $P(\yA)$, then by the theorem
-of total probability our belief about $\ya$ is
-\begin{equation}
-  \label{eq:belief_sample_unknown_A}
-  p(\ya)= \sum_{\yA} \yG_{\ya\yA} \;P(\yA).
-\end{equation}
-
-From the definition of normalized factorial
-moment~\eqref{eq:def_factorial_moment_a}, the expression for the
-hypergeometric distribution~\eqref{eq:conditional_hypergeometric}, and the
-relation~\eqref{eq:belief_sample_unknown_A} between our beliefs about $\ya$
-and $\yA$, and using some combinatorial juggling
-\citep%[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}
-[appendix~A and references therein]{portamanaetal2015}, one can prove the
-equality~\eqref{eq:equal_m_factorialmoments} between the factorial moments
-of sample and larger population.
-
-\bigskip
-
-For the construction of a maximum-entropy distribution from generic
-expectation constraints see Jaynes
-\citey{jaynes1963}[\chap~11]{jaynes1994_r2003}, and Hobson \etal\
-\citey{hobsonetal1973} and Sivia \citey[\sect~5.2.2]{sivia1996_r2006} for
-the minimum-relative-entropy construction. For the solution of the
-extremization problem using Lagrangians and Lagrange multipliers see Mead
-\etal\ \citey{meadetal1984} and the extensive texts by Fang \etal\
-\citey{fangetal1997} and Boyd \etal\ \citey{boydetal2004_r2009}. For a
-geometric understanding of the extremization and of the relation between
-expectations and multipliers see Porta Mana \citey{portamana2017b}.
-
-The result has the standard exponential-family form
-\begin{equation}
-  \label{eq:app_maxent_pop}
-  \begin{gathered}
-  \px(\yA)  = \frac{1}{Z(\yll)}\;
-  r_{\yA}\;
-  \exp\Biggl[\sum_{m}\yl_{m}\;
-  \binom{\yA}{m}\binom{N}{m}^{-1}
-  \Biggr],
-  \\
-  Z(\yll)\defd \sum_{\yA} r_{\yA}\;
-  \exp\Biggl[\sum_{m}\yl_{m}\;
-  \binom{\yA}{m}\binom{N}{m}^{-1}
-  \Biggr],
-\end{gathered}
-\end{equation}
-where $(r_{\yA})$ is the reference distribution and $\yll\defd (\yl_{m})$ are
-the Lagrange multipliers, satisfying the implicit
-constraint equations~\eqref{eq:final_constraint}:
-\begin{equation}
-  \label{eq:constraint_eq_lambda}
-  \sum_{\yA}
-  \tbinom{\yA}{m}\tbinom{N}{m}^{-1}\;
-  \tfrac{1}{Z(\yll)}\;
-  r_{\yA}\;
-  \exp\biggl[\sum_{m}\yl_{m}\;
-  \tbinom{\yA}{m}\tbinom{N}{m}^{-1}
-  \biggr]
-=
-  \sum_{\ya}
-  \tbinom{\ya}{m} \tbinom{n}{m}^{-1} \; \yff_{\ya},\quad
-  m\in M.
-\end{equation}
-
-The reference distribution $(r_{\yA})$ represents our pre-data beliefs
-about the activity levels $\yA$. We know that the majority of neurons in a
-brain area are rarely simultaneously active within a window of some
-milliseconds, so we could choose a distribution with slightly higher
-weights on low values of $\yA$. On the other hand, considering the number
-of ways in which $\yA$ out of $N$ neurons can be simultaneously active
-would suggest the multiplicity distribution proportional to
-$\binom{N}{\yA}$. It turns out that our results of
-\sects~\ref{sec:application}--\ref{sec:marginalization} are insensitive to
-the choice between these two possible reference distributions, or even a
-uniform reference distribution.
-
-For a derivation from a Bayesian perspective see
-appendix~\ref{sec:inference_prob}.
-
-\section{Measure of informational sufficiency}
-\label{sec:measure_suff}
-
-Let's ask how much more probable is the
-sufficiency of one set with respect to the other, conditional on our data
-$\yf$:
-\begin{equation}
-  \label{eq:ratio_prob_suffic}
-  \frac{p(M'' \| \yf)}{ p(M' \| \yf)}.
-\end{equation}
-The probability of observing activity $\ya$ in the sample at any time bin
-is the sample marginal of the maximum-entropy distribution for the larger
-population, owing to the excheangeability assumption implicit in the
-maximum-entropy method:
-\begin{equation}
-  \label{eq:sample_marginal}
-  \pxx(\ya \| M) = \sum_{\yA}\yG_{\ya\,\yA}\;\px(\yA \| M).
-\end{equation}
-The probability of observing a specific sequence $(\ya_{t})$ with
-frequencies $\yf$ is therefore
-\begin{equation}
-  \label{eq:prob_frequencies}
-  \prod_{t=1}^{T} \pxx(\ya_{t} \| M) \equiv
-  \prod_{\ya=0}^{n} \pxx(\ya \| M)^{T\,\yff_{\ya}} \equiv
-  \prod_{\ya=0}^{n}\biggl[\sum_{\yA}\yG_{\ya\yA}\;\px(\yA \| M)\biggr]^{T\,\yff_{\ya}}.
-\end{equation}
-The probability of observing the frequencies $\yf$ is obtained multiplying this
-by their multiplicity factor, the multinomial coefficient
-\begin{equation}
-  \label{eq:multiplicity_factor}
-  \binom{T}{T\yf} \defd \frac{T!}{\prod_{\ya}(T\,\yff_{\ya})!}
-  \approx \prod_{\ya}{\yff_{\ya}}^{-T\,\yff_{\ya}},
-\end{equation}
-the last expression coming from Stirling's approximation
-\citep[Lemma~2.2]{csiszaretal2004b}.
-If we assign equal pre-data probabilities to the two hypotheses $M'$ and $M''$,
-each probability in the ratio~\eqref{eq:ratio_prob_suffic} then becomes, by
-Bayes's theorem,
-\begin{multline}
-  \label{eq:prob_proportional_relentropy}
-  \pf(M \| \yf) \propto
-  \pf(\yf \| M) % \times\text{\small const}
-  \propto
-  \binom{T}{T\yf}\,  \prod_{\ya}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{
-    T\,\yff_{\ya}}
-   \approx{}\\
-  \prod_{\ya}{\yff_{\ya}}^{-T\,\yff_{\ya}}
-  \times
-  \prod_{\ya}\biggl[\sum_{\yA}\yG_{\ya\yA}\px(\yA \| M)\biggr]^{
-    T\,\yff_{\ya}}.
-\end{multline}
-The logarithm of the probability above is easily seen to be the number of
-bins $T$ multiplied by relative entropy between the frequency distribution
-$\yf$ and the sample marginal of the maximum-entropy distribution.
-
-Thus, the difference~\eqref{eq:measure_suffic} is the logarithm of the
-probability ratio~\eqref{eq:ratio_prob_suffic}. The exponential of the
-difference~\eqref{eq:measure_suffic} tells us how much more probable the
-hypothesis that the set $M''$ is sufficient than the hypothesis that the
-set $M'$ is.
-
-\section{Assumptions and approximations: mathematical sketch}
-\label{sec:inference_prob}
-
-We give here a sketch of the mathematical formulae behind the assumptions
-and approximations discussed in \sect~\ref{sec:assumptions}. We indicate
-them all collectively with $\yI$.
-
-The first assumption, of infinite exchangeability
-\citep[\chap~4]{bernardoetal1994_r2000}{dawid2013%cifarellietal1982
-} of the probability distribution $\pf[(\yA_{1},\dotsc, \yA_{T}) \| \yI]$
-for the sequence of activities, implies the following mixture form by
-de~Finetti's theorem \citep{definetti1930,hewittetal1955}:
-\begin{equation}
-  \label{eq:mixture_from_definetti}
-%  \exp[-T\;\sh(\yF;\ynu)] \quad
-%  \pxxx[(\yA_{1},\dotsc,\yA_{T})] \propto
-  \pf[(\yA_{1},\dotsc, \yA_{T}) \| \yI]
-  = \int\!\!\!\di\ynu\;\pf(\ynu \|\yI)\,\prod_{\yA=0}^{N}{\nu_{\yA}}^{T\,\yFF_{\yA}},
-\end{equation}
-where $\ynu\defd(\nu_{\yA})$ is the long-run frequency distribution of
-activities in an imaginary continuation or repetition of the recording, and
-$\pf(\ynu \|\yI)\;\di\ynu$ our belief distribution about it. The
-integration is over an $N$-dimensional simplex.
-
-The second assumption concerns the latter distribution. Its density
-function has one of the following expressions:
-\begin{enumerate}[label=(\textit{\alph*}),wide]
-\item an entropic prior \citep{skillingetal1984,rodriguez1991,neumann2007}
-  \begin{subequations}\label{eq:maxent_prior}
-    \begin{equation}
-      \label{eq:entropic_prior}
-      \pf(\ynu \| \yI) \propto  \exp[-\yL\;\sh(\yF;\yr)]
-      \approx \binom{L}{L\nu_{0}, \dotsc, L\nu_{N}}\;
-      \prod_{\yA}{r_{\yA}}^{L\nu_{\yA}}  
-    \end{equation}
-    where $\yr\defd(r_{\yA})$ is a reference distribution and the
-    expression in larger parentheses is a multinomial coefficient;
-  \item a \enquote{sufficiency prior}, from the exponential family, which
-    follows from the Pitman-Koopman theorem if some moments $\set{m}$ are
-    sufficient statistics for our belief about $\ynu$:
-    \begin{equation}
-      \label{eq:suff_prior}
-      \pf(\ynu \| \yI) \propto
-      \int\!\!\!\di\yla\;\pf(\yla \| \yI)\;
-      \prod_{\yA}\delt\Biggl\{\ynu_{\yA}-
-      \frac{r_{\yA}}{Z(\yla)}\,\exp\Bigl[
-      \tsum_{m}\lambda_{m}\tbinom{\yA}{m}/\tbinom{N}{m}
-      \Bigr]\Biggr\},
-    \end{equation}
-  \end{subequations}
-  where $\yla\defd(\lambda_{m})$ are Lagrange multipliers and $Z(\yla)$ a
-  normalization factor for the exponential distribution within the delta.
-  This is a formal expression, the delta symbolizing that the possible
-  domain of $\ynu$ is restricted from an $N$-dimensional simplex to a
-  lower-dimensional subset thereof, parametrized by $\yla$. When this prior
-  is used, all integrations over $\ynu$ are replaced by integrations over
-  $\yla$.
-\end{enumerate}
-
-The third assumption corresponds to the formula
-\begin{multline}
-  \label{eq:assume_sampled_anew}
-  \pf[(\ya_{1},\dotsc,\ya_{T}) \| (\yA_{1}, \dotsc, \yA_{T}), \yI ]
-  ={}\\ \prod_{t=1}^{T} \pf(\ya_{t} \| \yA_{t}, \yI)
-  = \prod_{t=1}^{T} \yG_{\ya_{t}\yA_{t}}
-  \equiv \prod_{\ya,\yA}{\yG_{\ya_{t}\yA_{t}}}^{T\,J_{\ya\yA}},
-\end{multline}
-where $J_{\ya\yA}$ is the joint frequency distribution of the activity
-pairs $(\ya,\yA)$, so that its marginals are the frequency distributions
-$\yf$ and $\yF$. As discussed in \sect~\ref{sec:assumptions} this formula
-follows from either of two beliefs: that all subsets of $\yA$ neurons in
-the larger population are always equally likely to give rise to total
-activity $\yA$, or that a new sampling is done at every time bin.
-
-It can be proved that the formulae above lead to the following belief
-distribution for the joint frequency distribution $\yJ \defd (J_{\ya\yA})$
-and the long-run frequency distribution $\ynu$, conditional on $\yf$:
-\begin{multline}
-  \label{eq:prob_joint_frequencies}
-  \pf(\yJ, \ynu \| \yf, \yI) \propto
-  %\int\!\!\!\di\ynu\;
-  \pf(\ynu\|\yI)\;
-                         \binom{T}{TJ_{0 0},\dotsc,TJ_{n N}}\,
-                         \prod_{\ya,\yA}\bigl(\yG_{\ya\yA}\;\nu_{\yA}\bigr)^{T\,J_{\ya\yA}}
-  \approx{}  \\
-  \shoveright{%\int\!\!\!\di\ynu\;
-    \pf(\ynu\|\yI)\;
-                         \exp\biggl(
--T \tsum_{\ya,\yA}J_{\ya\yA}\ln\frac{J_{\ya\yA}}{\yG_{\ya\yA}\;\nu_{\yA}}
-                         \biggr)}
-  \\
-  \text{with}\quad \tsum_{\yA}J_{\ya\yA}=\yff_{\ya},
-\end{multline}
-where $\pf(\ynu\|\yI)$ is either of the
-distributions~\eqref{eq:maxent_prior}. Note the relative entropy in the
-exponential. If we want to condition not on $\yf$ but on a set of measured
-averages $C(m) = \sum_{\ya}c_{\ya}(m) \yff_{\ya}$, $m\in M$, such
-as~\eqref{eq:chain_prescription_moments}, we have to replace the
-integration constraint $\sum_{\yA}J_{\ya\yA}=\yff_{\ya}$ with
-$\sum_{\ya\yA}c_{\ya}(m)\,J_{\ya\yA}=C(m)$.
-
-From this belief distribution we can find the one for
-$\yFF_{\yA}\equiv \sum_{\ya}J_{\ya\yA}$ and any others of interest by
-marginalization. Within the exponential we can recognize the relative
-entropy of $\yJ$ with respect to $(\yG_{\ya\yA}\;\nu_{\yA})$, which is the
-reason why these two joint distributions must be equal if $T$ tends to
-infinity.
-
-From the formulae above we can also find the four distributions mentioned
-at the end of \sect~\ref{sec:model}. The frequency
-distribution~\ref{item:prob_freq} is the mode of $\pf(\yF \| \yf,\yI)$. The
-belief distribution~\ref{item:prob_binin} is
-\begin{equation}
-  \label{eq:prob_binin}
-  \pf(\yA_{t} \| \yf, \yI) =
-  \sum_{\yF} \yFF_{\yA_{t}}\; \pf(\yF \| \yf, \yI),
-  \qquad t \in \set{1,\dotsc,T}.
-\end{equation}
-The frequency distribution~\ref{item:prob_longrun} is the mode of $\pf(\ynu
-\| \yf,\yI)$. Finally, the belief distribution~\ref{item:prob_binout} is
-\begin{equation}
-  \label{eq:prob_binout}
-  \pf(\yA_{t} \| \yf, \yI) =
-  \int\!\!\!\di\ynu\;\nu_{\yA_{t}}\;\pf(\ynu \| \yf, \yI),
-  \qquad t \notin \set{1,\dotsc,T}.
-\end{equation}
+% \appendix
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\renewcommand*{\finalnamedelim}{\addcomma\space}
 \defbibnote{prenote}{{\footnotesize (\enquote{de $X$} is listed under D,
@@
 
-\printbibliography[prenote=prenote%,postnote=postnote
+\printbibliography[%prenote=prenote%,postnote=postnote
 ]

### 2020-03-16 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-16T14:36:02+0100
+%% Last-Updated: 2020-03-16T16:48:19+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\newcommand*{\limplies}{\mathbin{\Rightarrow}}%implies
-%\newcommand*{\suchthat}{\mid}%{\mathpunct{|}}%such that (eg in sets)
+\newcommand*{\suchthat}{\mid}%{\mathpunct{|}}%such that (eg in sets)
 %\newcommand*{\with}{\colon}%with (list of indices)
@@
 %\DeclarePairedDelimiter\opcl{]}{]}
-%\DeclarePairedDelimiter\opop{]}{[}
+\DeclarePairedDelimiter\opop{]}{[}
 \DeclarePairedDelimiter\abs{\lvert}{\rvert}
@@
 %\DeclarePairedDelimiterX{\condt}[2]{[}{]}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
-%\DeclarePairedDelimiterX{\conds}[2]{\{}{\}}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
+\DeclarePairedDelimiterX{\conds}[2]{\{}{\}}{#1\nonscript\,\delimsize\vert\nonscript\;\mathopen{}#2}
 %\newcommand*{\+}{\lor}
@@
 
-There are many differing views on the questions of comparison, complexity,
-nesting, and \enquote{penalization} of models. Our view is close to that
-expressed in several works, studies, and commentaries
-\citep[\eg][]{good1985,jonesetal1986,chatfield1995,copasetal1995,draper1995,spiegelhalteretal1995,raftery1995,gelmanetal1995,draperetal1996,hoetingetal1999,clydeetal1999,browneetal2006},
+There are many differing, fragmented views on the questions of comparison,
+complexity, nesting, and \enquote{penalization} of models. Our view is
+close to that expressed in several works, studies, and commentaries
+\citep[\eg][]{good1950,good1985,kadaneetal1980b,jonesetal1986,chatfield1995,copasetal1995,draper1995,spiegelhalteretal1995,raftery1995,gelmanetal1995,mackay1995_r2003,draperetal1996,hoetingetal1999,clydeetal1999,browneetal2006},
 and well-explained in MacKay \citey{mackay1992}. We summarize this view
@@
 purpose (say, hardware algorithm implementation in some automated device),
-such choice requires not only the probabilities $p(H_{i}\|D)$, but also the
-gain/loss matrix for the various choices conditional on the various
-hypotheses. The choice is then made by maximization of the expected gain.
-Some gain matrices can even lead to choosing the hypothesis having
+such choice requires not only the probabilities $\p(H_{i}\|D)$, but also
+the utility/loss matrix for the various choices conditional on the various
+hypotheses \citep{kadaneetal1980b}. The choice is then made by maximization
+of the expected utility, a procedure dictated by some simple rational
+desiderata
+\citep{fishburn1981,bernardoetal1994_r2000}[\chap~13]{jaynes1994_r2003}.
+Some utility matrices can even lead to choosing the hypothesis having
 \emph{lower} posterior probability \citep[this frequently happens in
@@
 \medskip
+
+From our Bayesian point of view, the question of comparing (as opposed to
+choosing, see above) two models or hypotheses $M_{i}$ is nothing else than
+the calculation of their post-data probabilities $\p(M_{i} \| D, I)$, where
+$I$ denotes the context and prior information. Such calculation depends on
+what we mean by \enquote{model}. There are two main possibilities: model as
+a family, or model as a specific distribution.
+
+\smallskip
+
+\textbullet\ If by \enquote{model} we mean only a family of probability
+distribution, \eg\
+\[\conds*{\exp\biggl[-\frac{(x-\mu)^{2}}{2\sigma^{2}}\biggr]\,\di x}{
+    \mu\in\RR, \sigma \in\opop{0,\infty}}\;,\]
+then model comparison is undefined, because we cannot calculate
+$\p(D \| M_{i})$. Each distribution in the family assigns a different
+probability, and it is not clear which should be chosen or how they should
+be combined and why. A hypothesis is ill-defined, or at least useless, if
+it does not allow us to assign a probability to the data. We need a
+pre-data probability distribution for the model parameters, \eg\
+$\pf(\mu,\sigma \| I)\,\di\mu\,\di\sigma$. \enquote{Model} must therefore
+be redefined as a family of distributions \emph{and} a pre-data
+distribution over such family. This redefinition has several consequences:
+
+First, two models are different if they have the same family but different
+distributions over it. In fact they lead to different probabilities on the
+data space. Two such models are also mutually exclusive,
+$\p(M_{1} \land M_{2} \|I)=0$, because we cannot use the one family
+distribution \emph{and} the other family distribution at the same time.
+It's either-or. This indeed makes sense from a nonparametric perspective,
+where a parametric model (in the redefined sense) is just a singular
+nonparametric distribution, concentrated on a lower-dimensional manifold
+(the family above) of the full space of distributions
+\citep[p.~761]{draperetal1996}.
+
+Second, the \enquote{complexity} of a model cannot be simply judged from
+the dimensionality of its family. Model $M_{1}$ can have a family
+distribution concentrated on a very small region, with respect to model
+$M_{2}$ with the same family. $M_{1}$ is thus effectively using
+\enquote{fewer} parameters than $M_{2}$. Correspondingly the
+\enquote{typical set} of data to which $M_{1}$ assigns appreciable
+probability is smaller than the typical set of $M_{2}$: model $M_{1}$ is
+\enquote{less powerful}. The effective dimensionality or complexity of a
+model is thus best defined by the volume of data that is typical under that
+model. It may thus happen that a model is more powerful than another even
+if it is defined on a lower-dimensional family.
+
+Third, \emph{the evidence $\p(D \| M_{i}, I)$ of a model automatically
+  takes care of its complexity}. This happens because the probability over
+the data space is normalized. If $M_{1}$ has a typical data volume larger
+than $M_{2}$, then it must also assign lower probability to each data point
+in that volume, to ensure normalization. As a result, if the observed data
+$D^{*}$ lie in the typical volumes of both models, then
+$\pf(D^{*} \| M_{1},I)< \pf(D^{*} \| M_{2}, I)$: \emph{the evidence is
+  automatically penalizing the larger model}. There is a continuous
+trade-off between how typical the data is for each model and each model's
+power. This is the so-called automatic Ockham razor of Bayes's theorem,
+extensively explained by MacKay \citey{mackay1992}. And this is the reason
+why we are personally satisfied with simply comparing the evidence for any
+two models.
+
+\smallskip
+
+\textbullet\
 

### 2020-03-16 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-16T16:48:19+0100
+%% Last-Updated: 2020-03-16T17:07:32+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\textbullet\ If by \enquote{model} we mean only a family of probability
-distribution, \eg\
+\textbf{(A)} If by \enquote{model} we mean only a family of probability
+distributions, \eg\ (apart from normalization)
 \[\conds*{\exp\biggl[-\frac{(x-\mu)^{2}}{2\sigma^{2}}\biggr]\,\di x}{
@@
   takes care of its complexity}. This happens because the probability over
-the data space is normalized. If $M_{1}$ has a typical data volume larger
-than $M_{2}$, then it must also assign lower probability to each data point
-in that volume, to ensure normalization. As a result, if the observed data
-$D^{*}$ lie in the typical volumes of both models, then
+the data space is normalized. If $M_{1}$ has a typical-data volume larger
+than $M_{2}$'s, then it must also assign lower probability to each data
+point in that volume, to ensure normalization. As a result, if the observed
+data $D^{*}$ lie in the typical volumes of both models, then
 $\pf(D^{*} \| M_{1},I)< \pf(D^{*} \| M_{2}, I)$: \emph{the evidence is
@@
 power. This is the so-called automatic Ockham razor of Bayes's theorem,
-extensively explained by MacKay \citey{mackay1992}. And this is the reason
-why we are personally satisfied with simply comparing the evidence for any
-two models.
+extensively explained by MacKay
+\citey{mackay1992}[also][\sect~28.1]{mackay1995_r2003}. And this is the
+reason why we are personally satisfied with simply comparing the evidence
+for any two models.
 
@@
 
-\textbullet\
+\textbf{(B)} If by \enquote{model} we mean a specific distribution, \eg\
+\[ \exp\biggl[-\frac{(x-5)^{2}}{18}\biggr]\,\di x \;,\] then two models can
+of course be directly compared; that is $\p(D \|M_{i}, I)$ is well-defined.
+In this case the \enquote{dimensionality} of the model is undefined,
+however. The distribution above could be considered to belong to the family
+$\set[\Big]{\exp\Bigl[-\frac{(x-\mu)^{2}}{18}\Bigr]}$, or
+$\set[\Big]{\exp\Bigl[-\frac{(x-5)^{2}}{2\sigma^{2}}\Bigr]}$, or
+$\set[\big]{\exp[f(x-5)]}$, and so on. A point in a manifold belongs to an
+infinity of submanifolds.
+
+Note that the property of having a minimal sufficient statistic is still
+well-defined for a model in this sense, in an exchangeable context. The
+model above has $\set{n, \widebar{x}, \widebar{x^{2}}}$ as minimal
+sufficient statistics, where the bar denotes sample average. This does not
+authorize us to see it as belonging to a sufficient-statistics family,
+however.
+
+\bigskip
 

### 2020-03-16 replies1_ncomp.tex
update

--- 2020-03-16 replies1_ncomp.tex
@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-16T14:36:02+0100
+%% Last-Updated: 2020-03-16T19:55:02+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newif\ifafour
-\afourfalse% true = A4, false = A5
+\afourtrue% true = A4, false = A5
 \newif\iftypodisclaim % typographical disclaim on the side
@@
 meant to show the implications of our method for that kind of discussions.
-In particular we numerically show that conclusions based on a sample and
-conclusions based on the larger population can be very different. Our
+In particular we want to show that \emph{conclusions based on a sample and
+  conclusions based on the larger population can be very different}. Our
 method can bypass that difficulty, if one wants to use it for such
@@
 
-There are many differing views on the questions of comparison, complexity,
-nesting, and \enquote{penalization} of models. Our view is close to that
-expressed in several works, studies, and commentaries
-\citep[\eg][]{good1985,jonesetal1986,chatfield1995,copasetal1995,draper1995,spiegelhalteretal1995,raftery1995,gelmanetal1995,draperetal1996,hoetingetal1999,clydeetal1999,browneetal2006},
-and well-explained in MacKay \citey{mackay1992}. We summarize this view
-below. According to it, the analysis in our \sect~4, based on the weight of
-evidence, is consistent and does not need any corrections.
-
-Note in particular that we absolutely avoid speaking about
+We partly agree with the reviewer's concern, but for different reasons. The
+simple comparison of the log-evidences is not a problem in itself, as
+explained below. The problem is what is meant by \enquote{model} and what
+is being compared. Under one interpretation, our calculation of the
+log-evidence is missing a factor, which comes from the full Bayesian
+analysis. This factor, however, should be roughly the same in the
+calculations from the sample and from the larger population, so our point
+that the two calculations lead to different conclusions still holds. Under
+another interpretation our calculation is correct as it is. We explain this
+below.
+
+In any case, please note that we avoid speaking about
 \enquote{significance}, also as recommended by the American Statistical
-Association in their recent official statements \citep{asa2016}[especially
-\sect~2]{asa2019}. We are simply calculating the evidence \citep{good1950}
-$\p(D \|M_{i})$ for several hypotheses $M_{i}$ based on data $D$. Evidence
-is a continuous quantity and we are not setting any arbitrary thresholds.
-Multiplied by a pre-data probability, this evidence gives the post-data
-probability for each hypothesis. The pre-data probability will depend on
-the specific case, animal, brain region, and additional knowledge. If the
-evidence is $10^{35}:1$ for the first hypothesis against the second, but
-our pre-data probability for the second is $10^{36}$ times that for the
-first, then the second hypothesis will have higher post-data probability.
-But there is still a continuum of cases, no dichotomy. No such dichotomy is
-necessary.
-
-If a \emph{choice} among two or more hypotheses has to be made for some
-purpose (say, hardware algorithm implementation in some automated device),
-such choice requires not only the probabilities $p(H_{i}\|D)$, but also the
-gain/loss matrix for the various choices conditional on the various
-hypotheses. The choice is then made by maximization of the expected gain.
-Some gain matrices can even lead to choosing the hypothesis having
-\emph{lower} posterior probability \citep[this frequently happens in
-medical decision problems, see \eg][]{soxetal1988_r2013}.
+Association in their recent official statements \citep{asa2016}[see
+especially \sect~2]{asa2019}. When comparing hypotheses $M_{i}$ we simply
+want their post-data probabilities $\p(M_{i} \| D)$, with their continuum
+of cases. Other researchers are of course free to make dichotomies if they
+like; but we personally agree with the American Statistical Association
+that any threshold is unnecessary and arbitrary. If a \emph{choice} among
+the hypotheses has to be made for some purpose (say, hardware algorithm
+implementation in some automated device), such choice requires not only the
+probabilities $\p(M_{i}\|D)$, but also the utility/loss matrix for the
+various choices conditional on the various hypotheses
+\citep{kadaneetal1980b}. The choice is then made by maximization of the
+expected utility, a procedure dictated by some simple rational desiderata
+\citep{fishburn1981,bernardoetal1994_r2000}[\chap~13]{jaynes1994_r2003}.
+Some utility matrices can even lead to choosing the hypothesis having
+\emph{lower} post-data probability; this frequently happens in
+medical decision problems, see \eg\ \textcite{soxetal1988_r2013}.
+
+% We want a continuous quantity, without any arbitrary thresholds. The
+% evidence is such a continuous quantity. Multiplied by a pre-data
+% probability, the evidence gives the post-data probability for each
+% hypothesis. The pre-data probability will depend on the specific case,
+% animal, brain region, and additional knowledge. If the evidence is
+% $10^{35}:1$ for the first hypothesis against the second, but our pre-data
+% probability for the second is $10^{36}$ times that for the first, then the
+% second hypothesis will have higher post-data probability. But there is
+% still a continuum of cases, no dichotomy. No such dichotomy is necessary.
+% simply calculating the evidence \citep{good1950} $\p(D \|M_{i})$ for
+% several hypotheses $M_{i}$ based on data $D$. Evidence is a continuous
+% quantity and we are not setting any arbitrary thresholds. Multiplied by a
+% pre-data probability, this evidence gives the post-data probability for
+% each hypothesis. The pre-data probability will depend on the specific case,
+% animal, brain region, and additional knowledge. If the evidence is
+% $10^{35}:1$ for the first hypothesis against the second, but our pre-data
+% probability for the second is $10^{36}$ times that for the first, then the
+% second hypothesis will have higher post-data probability. But there is
+% still a continuum of cases, no dichotomy. No such dichotomy is necessary.
+
+\smallskip
+
+There are many differing, fragmented views on the questions of comparison,
+complexity, nesting, and \enquote{penalization} of models. Our view is
+close to that expressed in several works, studies, and commentaries
+\footcites[\eg][]{good1950,good1985,kadaneetal1980b,jonesetal1986,chatfield1995,copasetal1995,draper1995,spiegelhalteretal1995,raftery1995,gelmanetal1995,mackay1995_r2003,draperetal1996,hoetingetal1999,clydeetal1999,browneetal2006},
+and well-explained in MacKay \citey{mackay1992}. According to this view,
+which we summarize below, the log-evidence automatically takes care of
+model complexity.
+
+The main question is what we mean by \enquote{model}. There are two main
+possibilities: model as a family, or model as a specific distribution.
+
+\smallskip
+
+\textbf{(A)} If by \enquote{model} we mean a family of probability
+distributions, \eg\ (apart from normalization)
+\[\conds*{\exp\biggl[-\frac{(x-\mu)^{2}}{2\sigma^{2}}\biggr]\,\di x}{
+    \mu\in\RR, \sigma \in\opop{0,\infty}}\;,\]
+then model comparison is ill-defined. % Each distribution in the family
+% assigns a different probability, and it is not clear which should be
+% chosen or how they should be combined and why.
+It may happen that the optimal distribution in family $\mathcal{F}_{1}$
+assigns more probability to the observed data than the optimal distribution
+in family $\mathcal{F}_{2}$, and yet $\mathcal{F}_{1}$ assigns \enquote{on
+  average} lower probabilities to the data than $\mathcal{F}_{2}$. It is
+not clear how the two families should be compared in this case, or what
+$\p(D \| \mathcal{F}_{i})$ is. A hypothesis is ill-defined, or at least
+useless, if it does not allow us to assign a probability to the data. Such
+a probability can be obtained if we specify a pre-data probability
+distribution over the distributions in the family, \eg\
+$\pf(\mu,\sigma \| I)\,\di\mu\,\di\sigma$ in the example above.
+\enquote{Model} must therefore be redefined as a family of distributions
+\emph{and} a pre-data distribution over such family. This redefinition has
+several consequences:
+
+First, two models are different if they have the same family but different
+distributions over it. In fact they lead to different probabilities on the
+data space. Two such models are also mutually exclusive,
+$\p(M_{1} \land M_{2} \|I)=0$, because we cannot use the one family
+distribution \emph{and} the other family distribution at the same time.
+It's either-or. This consequence makes sense from a nonparametric
+perspective, where a parametric model (in the redefined sense) is just a
+delta-like nonparametric distribution, concentrated on a lower-dimensional
+manifold (the family above) of the full space of distributions
+\citep[p.~761]{draperetal1996}.
+
+Second, the \enquote{complexity} of a model cannot be judged from the
+dimensionality of its family. Model $M_{1}$ can have a family distribution
+concentrated on a very small region, with respect to model $M_{2}$ having
+the same family. $M_{1}$ is thus effectively using fewer parameters than
+$M_{2}$. Correspondingly the \enquote{typical set} of data to which $M_{1}$
+assigns appreciable probability is smaller than the typical set of $M_{2}$:
+model $M_{1}$ is less \enquote{powerful}. The effective dimensionality or
+complexity of a model is thus best defined by the volume of data that is
+typical under that model. Most important, it may then happen that a model
+is more powerful than another even if it is defined on a lower-dimensional
+family.
+
+Third, \emph{the evidence $\p(D \| M_{i})$ of a model automatically takes
+  care of its complexity}. This happens because the probability over the
+data space is normalized. If $M_{1}$ has a typical-data volume larger than
+$M_{2}$'s, then it must also assign lower probability to each data point in
+that volume, to ensure normalization. As a result, if the observed data
+$D^{*}$ lie in the typical volumes of both models, then
+$\pf(D^{*} \| M_{1})< \pf(D^{*} \| M_{2})$: \emph{the evidence is
+  automatically penalizing the larger model}. There is a continuous
+trade-off between how typical the data is for each model and each model's
+power. This is the so-called automatic Ockham razor of Bayes's theorem,
+extensively explained by MacKay
+\citey{mackay1992}[also][\sect~28.1]{mackay1995_r2003}. And this is the
+reason why we are personally satisfied with simply comparing the evidence
+for any two models.
+
+In our calculations, however, we are using a zeroth-order Laplace
+approximation for calculating the evidence in \sect~4. A first-order
+approximation (containing the width of the distribution, influenced by the
+typical-set volume above) would correct for the model's power. We believe
+that the correction would nevertheless be the same in the calculation based
+on the larger population, \eqns~(11), and on the sample, \eqns~(12),
+because the data space is the same. Hence the difference in the two
+approaches -- our main message -- should persist. But it would be better to
+check this.
+
+\smallskip
+
+\textbf{(B)} If by \enquote{model} we mean a specific distribution, \eg\
+\[ \exp\biggl[-\frac{(x-5)^{2}}{18}\biggr]\,\di x \;,\] then two models can
+of course be directly compared: $\p(D \|M_{i}, I)$ is well-defined. In this
+case the \enquote{dimensionality} of the model is undefined, however. The
+distribution above could be considered as belonging to the family
+$\set[\Big]{\exp\Bigl[-\frac{(x-\mu)^{2}}{18}\Bigr]}$, or
+$\set[\Big]{\exp\Bigl[-\frac{(x-5)^{2}}{2\sigma^{2}}\Bigr]}$, or
+$\set[\big]{\exp[f(x-5)]}$, and so on. A point in a manifold belongs to an
+infinity of submanifolds.
+
+But the property of having a minimal sufficient statistic is still
+well-defined for a model in this sense, in an exchangeable context. The
+model above has $\set[\big]{n, \widebar{x}, \widebar{x^{2}}}$ (or any
+equivalent set) as minimal sufficient statistics, where the bar denotes the
+sample average. Thus two models having different sufficient statistics can
+be compared without reference to any family. Interpreted in this sense our
+calculations of \sect~4 are correct.
 
@@
 
-(4) {\footnotesize\textsf{\footnotesize I enjoyed somewhat philosophical arguments regarding the interpretation of the ME distribution (P8 and Appendix).
-However, I was not sure how these are relevant to the paper if it is just an interpretation.
-Can we consider different outcomes arising the different interpretations?
-If the authors are preparing another manuscript in this direction, these arguments may better be used for that paper?
-(This is just a weak suggestion.)}\par}
+The question, in summary, is what exactly is being compared, whether
+families (with accompanying distribution) or single distributions. The
+literature is not at all clear. But, in either case, our calculations show
+that the approaches based on a population and a sample thereof can lead to
+different conclusions -- in accordance with the non-preservation of
+sufficiency under sampling -- and they provide a way to bypass this
+problem.
+
+\bigskip
+
+(4) {\footnotesize\textsf{\footnotesize I enjoyed somewhat philosophical
+    arguments regarding the interpretation of the ME distribution (P8 and
+    Appendix). However, I was not sure how these are relevant to the paper
+    if it is just an interpretation. Can we consider different outcomes
+    arising the different interpretations? If the authors are preparing
+    another manuscript in this direction, these arguments may better be
+    used for that paper? (This is just a weak suggestion.)}\par}
+
+\medskip
+
+As stated on p.~8, the interpretation is important because it leads to
+numerically different results in first- and higher-order approximations;
+Appendix~C explains where the differences appear. Also, it is important to
+know whether we are calculating the frequency of past results, or the
+probability for a new result, or the most probable frequency of future
+results. These three quantities are very different when the ratio
+$\frac{\text{\# samples}}{\text{data-space dimension}}$ is high, and they are
+to be used in different ways.
+
+So these are not philosophical matters, it's a matter of understanding what
+we are doing and stating it clearly in our work. % Unfortunately the
+% neuroscience literature is very unclear on such matters.
+There's also another important reason. We are generally concerned for
+master's and PhD students, who often learn concepts and methods through
+actual research, perusing works like the present one. Young researchers too
+often end up learning incorrect or confused ideas, gathered from unclear
+literature. They trust it because it's peer-reviewed. Sometimes the result,
+quoting Feynman \citey{feynman1985_r1989}, is that \enquote{they don't
+  learn by understanding; they learn by some other way -- by rote, or
+  something. Their knowledge is so fragile!}. We owe them clarity,
+precision, honesty, and at least a glimpse of the bigger picture.
 

### 2020-03-16 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-16T19:55:02+0100
+%% Last-Updated: 2020-03-16T20:21:43+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   contradiction with Amari \etal. The results of that paper hold for the
-  $N\to\infty$, and no explicit form for the limit-remainder is given. It
-  is therefore unclear whether at $N=10\,000$ we should already observe a
-  delta. Our findings~\ref{item:maxmomentcheck}, above, suggest that we
-  should not. Limit-reminders such as \enquote{$\Ord(1/N^{k})$} as given in
+  $N\to\infty$ limit, and no explicit form for the limit-remainder is given
+  there. Limit-reminders such as \enquote{$\Ord(1/N^{k})$} as given in
   Amari \etal\ do not help unless we know their precise form. As an extreme
   example, $\frac{10^{10}}{N}$ is $\Ord(1/N)$, and in this case the $1/N$
-  term cannot really be neglected unless $N \gg 10^{10}$.
+  term cannot really be neglected unless $N \gg 10^{10}$. It is therefore
+  unclear whether at $N=10\,000$ we should already observe a delta. Our
+  findings~\ref{item:maxmomentcheck}, above, suggest that we should not.
 
@@
 Association in their recent official statements \citep{asa2016}[see
-especially \sect~2]{asa2019}. When comparing hypotheses $M_{i}$ we simply
-want their post-data probabilities $\p(M_{i} \| D)$, with their continuum
+especially \sect~2]{asa2019}. When comparing hypotheses $M_{i}$ we want
+their full post-data probabilities $\p(M_{i} \| D)$, with their continuum
 of cases. Other researchers are of course free to make dichotomies if they
@@
 Some utility matrices can even lead to choosing the hypothesis having
-\emph{lower} post-data probability; this frequently happens in
-medical decision problems, see \eg\ \textcite{soxetal1988_r2013}.
+\emph{lower} post-data probability; this frequently happens in medical
+decision problems, see \eg\ \textcite{soxetal1988_r2013}.
 
@@
 
-The main question is what we mean by \enquote{model}. There are two main
-possibilities: model as a family, or model as a specific distribution.
+The main question, on the other hand, is what we mean by \enquote{model}.
+There are two main possibilities\ibid: model as a family, or model as a specific
+distribution.
 
@@
 \[\conds*{\exp\biggl[-\frac{(x-\mu)^{2}}{2\sigma^{2}}\biggr]\,\di x}{
-    \mu\in\RR, \sigma \in\opop{0,\infty}}\;,\]
+    \mu\in\RR, \sigma \in \RR_{>0}}\;,\]
 then model comparison is ill-defined. % Each distribution in the family
@@
 data space. Two such models are also mutually exclusive,
-$\p(M_{1} \land M_{2} \|I)=0$, because we cannot use the one family
-distribution \emph{and} the other family distribution at the same time.
+$\p(M_{1} \land M_{2} \|I)=0$, because we cannot use $M_{1}$'s family
+distribution \emph{and} $M_{2}$'s family distribution at the same time.
 It's either-or. This consequence makes sense from a nonparametric
-perspective, where a parametric model (in the redefined sense) is just a
-delta-like nonparametric distribution, concentrated on a lower-dimensional
-manifold (the family above) of the full space of distributions
-\citep[p.~761]{draperetal1996}.
+perspective, where a parametric model is just a delta-like nonparametric
+distribution, concentrated on a lower-dimensional manifold (the family
+above) of the full space of distributions \citep[p.~761]{draperetal1996}.
+The notion of nested models becomes unimportant, from this perspective.
 
 Second, the \enquote{complexity} of a model cannot be judged from the
-dimensionality of its family. Model $M_{1}$ can have a family distribution
-concentrated on a very small region, with respect to model $M_{2}$ having
-the same family. $M_{1}$ is thus effectively using fewer parameters than
-$M_{2}$. Correspondingly the \enquote{typical set} of data to which $M_{1}$
-assigns appreciable probability is smaller than the typical set of $M_{2}$:
-model $M_{1}$ is less \enquote{powerful}. The effective dimensionality or
-complexity of a model is thus best defined by the volume of data that is
+dimensionality of its family. Models $M_{1}$ and $M_{2}$ can have the same
+family, but $M_{1}$ can have a family distribution concentrated on a very
+small region, with respect to that of $M_{2}$. Model $M_{1}$ is thus
+effectively using fewer parameters than $M_{2}$. Correspondingly the
+\enquote{typical set} of data to which $M_{1}$ assigns appreciable
+probability is smaller than the typical set of $M_{2}$: model $M_{1}$ is
+less \enquote{powerful}. The effective dimensionality, complexity, or power
+of a model is thus best defined by the volume of \emph{data space} that is
 typical under that model. Most important, it may then happen that a model
-is more powerful than another even if it is defined on a lower-dimensional
+is more powerful than another even if it is based on a lower-dimensional
 family.
@@
   care of its complexity}. This happens because the probability over the
-data space is normalized. If $M_{1}$ has a typical-data volume larger than
-$M_{2}$'s, then it must also assign lower probability to each data point in
-that volume, to ensure normalization. As a result, if the observed data
-$D^{*}$ lie in the typical volumes of both models, then
-$\pf(D^{*} \| M_{1})< \pf(D^{*} \| M_{2})$: \emph{the evidence is
+data space is normalized. If $M_{2}$ has a typical-data volume larger than
+$M_{1}$'s, then it must also assign lower probability to each volume
+element in that volume, to ensure normalization. As a result, if the
+observed data $D^{*}$ lie in the typical volumes of both models, then
+$\pf(D^{*} \| M_{1}) > \pf(D^{*} \| M_{2})$: \emph{the evidence is
   automatically penalizing the larger model}. There is a continuous
-trade-off between how typical the data is for each model and each model's
-power. This is the so-called automatic Ockham razor of Bayes's theorem,
-extensively explained by MacKay
+trade-off between how typical the observed data are for each model and each
+model's power. This is the so-called automatic Ockham razor of Bayes's
+theorem, extensively explained by MacKay
 \citey{mackay1992}[also][\sect~28.1]{mackay1995_r2003}. And this is the
@@
 on the larger population, \eqns~(11), and on the sample, \eqns~(12),
-because the data space is the same. Hence the difference in the two
+because the data space is the same in both. Hence the difference in the two
 approaches -- our main message -- should persist. But it would be better to
@@
 
-But the property of having a minimal sufficient statistic is still
+Yet, the property of having a minimal sufficient statistic is still
 well-defined for a model in this sense, in an exchangeable context. The
@@
 The question, in summary, is what exactly is being compared, whether
-families (with accompanying distribution) or single distributions. The
-literature is not at all clear. But, in either case, our calculations show
-that the approaches based on a population and a sample thereof can lead to
-different conclusions -- in accordance with the non-preservation of
-sufficiency under sampling -- and they provide a way to bypass this
-problem.
+families (and what their accompanying distributions are) or single
+distributions. The literature is not at all clear. But, in either case, our
+calculations show that the approaches based on a population and a sample
+thereof can lead to different conclusions -- in accordance with the
+non-preservation of sufficiency under sampling -- and they provide a way to
+bypass this problem.
 

### 2020-03-16 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-16T20:21:43+0100
+%% Last-Updated: 2020-03-16T21:23:04+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The notion of nested models becomes unimportant, from this perspective.
+Some authors try to define some \enquote{standard} distribution associated
+to a family, but we agree with those
+\citep[\eg][]{copasetal1979,lindleyetal1982} who find such definitions
+ultimately arbitrary. Many such distributions lead, moreover, to
+mathematical singularities.
 
@@
 
-Third, \emph{the evidence $\p(D \| M_{i})$ of a model automatically takes
+Third, \emph{the evidence $\p(D \| M_{i})$ for a model automatically takes
   care of its complexity}. This happens because the probability over the
@@
 $M_{1}$'s, then it must also assign lower probability to each volume
-element in that volume, to ensure normalization. As a result, if the
-observed data $D^{*}$ lie in the typical volumes of both models, then
+element therein, to ensure normalization. As a result, if the observed data
+$D^{*}$ lie in the typical volumes of both models, then
 $\pf(D^{*} \| M_{1}) > \pf(D^{*} \| M_{2})$: \emph{the evidence is
@@
 typical-set volume above) would correct for the model's power. We believe
-that the correction would nevertheless be the same in the calculation based
-on the larger population, \eqns~(11), and on the sample, \eqns~(12),
-because the data space is the same in both. Hence the difference in the two
-approaches -- our main message -- should persist. But it would be better to
-check this.
-
-\smallskip
+that the correction would nevertheless be the same in the calculations
+based on the larger population, \eqns~(11), and on the sample, \eqns~(12).
+Hence the difference in the two approaches -- our main message -- should
+persist. But it would be better to check this.
+
+\medskip
 
@@
 families (and what their accompanying distributions are) or single
-distributions. The literature is not at all clear. But, in either case, our
-calculations show that the approaches based on a population and a sample
-thereof can lead to different conclusions -- in accordance with the
-non-preservation of sufficiency under sampling -- and they provide a way to
-bypass this problem.
+distributions. The literature is unclear and contradictory on this. But, in
+either case, our calculations show that the approaches based on a
+population and a sample thereof can lead to different conclusions -- in
+accordance with the non-preservation of sufficiency under sampling -- and
+they provide a way to bypass this problem.
 
@@
 As stated on p.~8, the interpretation is important because it leads to
-numerically different results in first- and higher-order approximations;
-Appendix~C explains where the differences appear. Also, it is important to
-know whether we are calculating the frequency of past results, or the
-probability for a new result, or the most probable frequency of future
-results. These three quantities are very different when the ratio
-$\frac{\text{\# samples}}{\text{data-space dimension}}$ is high, and they are
-to be used in different ways.
-
-So these are not philosophical matters, it's a matter of understanding what
-we are doing and stating it clearly in our work. % Unfortunately the
+numerically different results in first- and higher-order approximations.
+And such approximations may be necessary in the precise calculation of the
+evidence, as explained in the reply to the previous point. Appendix~C
+explains where and why the numerical differences appear. Also, it is
+important to know whether we are calculating the frequency of past results,
+or the probability for a new result, or the most probable frequency of
+future results. These three quantities can be numerically very different
+when the ratio $\frac{\text{\# samples}}{\text{data-space dimension}}$ is
+high, and they are to be used in different ways.
+
+These are therefore not philosophical matters. It's a matter of
+understanding what we are doing and stating it clearly in our
+work. % Unfortunately the
 % neuroscience literature is very unclear on such matters.
@@
 often end up learning incorrect or confused ideas, gathered from unclear
-literature. They trust it because it's peer-reviewed. Sometimes the result,
-quoting Feynman \citey{feynman1985_r1989}, is that \enquote{they don't
-  learn by understanding; they learn by some other way -- by rote, or
-  something. Their knowledge is so fragile!}. We owe them clarity,
-precision, honesty, and at least a glimpse of the bigger picture.
+literature. % They trust it because it's peer-reviewed.
+Sometimes the result, quoting Feynman \citey{feynman1985_r1989}, is that
+\enquote{they don't learn by understanding; they learn by some other way --
+  by rote, or something. Their knowledge is so fragile!}. We owe them
+clarity, precision, honesty, and a glimpse of the bigger picture.
 

### 2020-03-16 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-16T21:23:04+0100
+%% Last-Updated: 2020-03-16T22:02:36+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \enquote{Model} must therefore be redefined as a family of distributions
-\emph{and} a pre-data distribution over such family. This redefinition has
-several consequences:
+\emph{and} a pre-data distribution over such family
+\citep{dawid1982,dawid1993}; under exchangeability, this is mathematically
+equivalent to the definition (B) below
+\citep[\chap~4]{bernardoetal1994_r2000}. This redefinition has several
+consequences:
 
 First, two models are different if they have the same family but different
-distributions over it. In fact they lead to different probabilities on the
-data space. Two such models are also mutually exclusive,
+distributions over it. In fact they clearly lead to different probabilities
+on the data space. Two such models are also mutually exclusive,
 $\p(M_{1} \land M_{2} \|I)=0$, because we cannot use $M_{1}$'s family
@@
 Some authors try to define some \enquote{standard} distribution associated
-to a family, but we agree with those
-\citep[\eg][]{copasetal1979,lindleyetal1982} who find such definitions
-ultimately arbitrary. Many such distributions lead, moreover, to
-mathematical singularities.
+to a family, but we agree with those \citep[\eg][]{copasetal1979}[DeGroot
+in][]{lindleyetal1982} who find such definitions ultimately arbitrary. Many
+such distributions lead, moreover, to mathematical singularities.
 

### 2020-03-16 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-16T22:02:36+0100
+%% Last-Updated: 2020-03-16T22:12:34+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 above) of the full space of distributions \citep[p.~761]{draperetal1996}.
-The notion of nested models becomes unimportant, from this perspective.
+The notion of nested models becomes unimportant from this perspective.
 Some authors try to define some \enquote{standard} distribution associated
@@
   care of its complexity}. This happens because the probability over the
-data space is normalized. If $M_{2}$ has a typical-data volume larger than
-$M_{1}$'s, then it must also assign lower probability to each volume
-element therein, to ensure normalization. As a result, if the observed data
-$D^{*}$ lie in the typical volumes of both models, then
-$\pf(D^{*} \| M_{1}) > \pf(D^{*} \| M_{2})$: \emph{the evidence is
+data space is normalized. If the more powerful $M_{2}$ has a typical-data
+volume larger than $M_{1}$'s, then it must also assign lower probability to
+each volume element therein, to ensure normalization. As a result, if the
+observed data $D^{*}$ lie in the typical volumes of both models, then
+$\pf(D^{*} \| M_{2}) < \pf(D^{*} \| M_{1})$: \emph{the evidence is
   automatically penalizing the larger model}. There is a continuous
@@
 approximation (containing the width of the distribution, influenced by the
-typical-set volume above) would correct for the model's power. We believe
-that the correction would nevertheless be the same in the calculations
-based on the larger population, \eqns~(11), and on the sample, \eqns~(12).
-Hence the difference in the two approaches -- our main message -- should
-persist. But it would be better to check this.
+typical-set volume above) would correct for the model's power
+\citep[\sect~2.2]{mackay1992}. We believe that the correction would
+nevertheless be the same in the calculations based on the larger
+population, \eqns~(11), and on the sample, \eqns~(12). Hence the difference
+in the two approaches -- our main message -- should persist. But it would
+be better to check this.
 

### 2020-03-16 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-16T22:12:34+0100
+%% Last-Updated: 2020-03-17T00:35:35+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 when the ratio $\frac{\text{\# samples}}{\text{data-space dimension}}$ is
-high, and they are to be used in different ways.
+high, and they are to be used in different ways. Preliminary calculations
+from the full Bayesian analysis show that%
+\setlength{\intextsep}{-1ex}% with wrapfigure
+\setlength{\columnsep}{0ex}% with wrapfigure
+\begin{wrapfigure}{r}{0pt} % with wrapfigure
+\centering\includegraphics[width=0.5\linewidth]{example_fullbayesian.pdf}
+%\caption{caption}\label{fig:comparison_a5}
+\end{wrapfigure}% exp_family_maxent.nb
+the most probable frequency distribution across recorded bins (item i on
+p.~8) and the belief distribution over the total activity at any recorded
+bin (item ii on p.~8) are visibly different. See figure on the right: the
+first is the red dashed red curve, the second the solid green line.
 

### 2020-03-17 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-17T00:35:35+0100
+%% Last-Updated: 2020-03-17T01:02:47+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 their full post-data probabilities $\p(M_{i} \| D)$, with their continuum
-of cases. Other researchers are of course free to make dichotomies if they
-like; but we personally agree with the American Statistical Association
-that any threshold is unnecessary and arbitrary. If a \emph{choice} among
-the hypotheses has to be made for some purpose (say, hardware algorithm
-implementation in some automated device), such choice requires not only the
-probabilities $\p(M_{i}\|D)$, but also the utility/loss matrix for the
-various choices conditional on the various hypotheses
-\citep{kadaneetal1980b}. The choice is then made by maximization of the
-expected utility, a procedure dictated by some simple rational desiderata
+of cases. We respect the freedom of other researchers to make dichotomies,
+if they so prefer. But our personal point of view agrees with the American
+Statistical Association: any threshold is unnecessary and arbitrary. If a
+\emph{choice} among the hypotheses has to be made for some purpose (say,
+hardware algorithm implementation in some automated device), such choice
+requires not only the probabilities $\p(M_{i}\|D)$, but also the
+utility/loss matrix for the various choices conditional on the various
+hypotheses \citep{kadaneetal1980b}. The choice is then made by maximization
+of the expected utility, a procedure dictated by some simple rational
+desiderata
 \citep{fishburn1981,bernardoetal1994_r2000}[\chap~13]{jaynes1994_r2003}.
@@
 
-2.
-{\footnotesize\textsf{\footnotesize The method requires that the measurements on n neurons are 'representative' of the full population of N neurons.
-The authors show one example, where they divide a dataset into two subsets of neurons with different statistics.
-They show that the extrapolations from these two subsets are significantly different.
-But this particular analysis is somewhat ad hoc.
-Are there any more systematic tests of procedures for understanding how sensitive this analysis is to choosing different subsets of neurons?
-Or different recoding periods?}\par}
+2. {\footnotesize\textsf{\footnotesize The method requires that the
+    measurements on n neurons are 'representative' of the full population
+    of N neurons. The authors show one example, where they divide a dataset
+    into two subsets of neurons with different statistics.  They show that
+    the extrapolations from these two subsets are significantly different.
+    But this particular analysis is somewhat ad hoc. Are there any more
+    systematic tests of procedures for understanding how sensitive this
+    analysis is to choosing different subsets of neurons? Or different
+    recoding periods?}\par}
+
+*** equiv to unsystematic sampling for extended recording arrays
 
@@
 Is there any way to take this dependence into account?}\par}
+
+*** analogous to \enquote{design} in survey sampling
 

### 2020-03-17 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-17T01:02:47+0100
+%% Last-Updated: 2020-03-17T02:31:16+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 simple comparison of the log-evidences is not a problem in itself, as
-explained below. The problem is what is meant by \enquote{model} and what
-is being compared. Under one interpretation, our calculation of the
-log-evidence is missing a factor, which comes from the full Bayesian
-analysis. This factor, however, should be roughly the same in the
-calculations from the sample and from the larger population, so our point
-that the two calculations lead to different conclusions still holds. Under
-another interpretation our calculation is correct as it is. We explain this
-below.
+explained below. % The problem is what is meant by \enquote{model} and what
+% is being compared. Under one interpretation,
+Our calculation of the log-evidence is missing a factor, which comes from
+the full Bayesian analysis. This factor, however, should be roughly the
+same in the calculations from the sample and from the larger population, so
+our point that the two calculations lead to different conclusions still
+holds. Under another interpretation our calculation is correct as it is. We
+explain this below.
 
@@
 which we summarize below, the log-evidence automatically takes care of
-model complexity.
-
-The main question, on the other hand, is what we mean by \enquote{model}.
-There are two main possibilities\ibid: model as a family, or model as a specific
-distribution.
-
-\smallskip
-
-\textbf{(A)} If by \enquote{model} we mean a family of probability
-distributions, \eg\ (apart from normalization)
+model complexity. This view hinges on a clear definition of what a
+\enquote{model} is.
+
+If by \enquote{model} we mean a family of probability distributions, \eg\
+(apart from normalization)
 \[\conds*{\exp\biggl[-\frac{(x-\mu)^{2}}{2\sigma^{2}}\biggr]\,\di x}{
@@
 $\pf(\mu,\sigma \| I)\,\di\mu\,\di\sigma$ in the example above.
-\enquote{Model} must therefore be redefined as a family of distributions
+\enquote{Model} must be therefore redefined as a family of distributions
 \emph{and} a pre-data distribution over such family
-\citep{dawid1982,dawid1993}; under exchangeability, this is mathematically
-equivalent to the definition (B) below
-\citep[\chap~4]{bernardoetal1994_r2000}. This redefinition has several
-consequences:
+\citep{dawid1982,dawid1993}. This redefinition also turns out to be
+equivalent, under exchangeability, to the specification of a single
+predictive distribution, that is, a joint probability for any number of
+samples
+\citep[\chap~4]{bernardoetal1994_r2000}{fortinietal2000,ressel1983_r1985,lauritzen1984}.
+This redefinition has several consequences:
 
 First, two models are different if they have the same family but different
-distributions over it. In fact they clearly lead to different probabilities
-on the data space. Two such models are also mutually exclusive,
-$\p(M_{1} \land M_{2} \|I)=0$, because we cannot use $M_{1}$'s family
-distribution \emph{and} $M_{2}$'s family distribution at the same time.
-It's either-or. This consequence makes sense from a nonparametric
+distributions over it. In fact they clearly lead to different predictive
+probabilities on the data space. Two such models are also mutually
+exclusive, $\p(M_{1} \land M_{2} \|I)=0$, because we cannot use $M_{1}$'s
+family distribution \emph{and} $M_{2}$'s family distribution at the same
+time. It's either-or. This consequence makes sense from a nonparametric
 perspective, where a parametric model is just a delta-like nonparametric
@@
 above) of the full space of distributions \citep[p.~761]{draperetal1996}.
-The notion of nested models becomes unimportant from this perspective.
-Some authors try to define some \enquote{standard} distribution associated
-to a family, but we agree with those \citep[\eg][]{copasetal1979}[DeGroot
+The notion of nested models becomes unimportant from this perspective. Some
+authors try to define some \enquote{standard} distribution associated to a
+family, but we agree with those \citep[\eg][]{copasetal1979}[DeGroot
 in][]{lindleyetal1982} who find such definitions ultimately arbitrary. Many
@@
 
-\medskip
-
-\textbf{(B)} If by \enquote{model} we mean a specific distribution, \eg\
-\[ \exp\biggl[-\frac{(x-5)^{2}}{18}\biggr]\,\di x \;,\] then two models can
-of course be directly compared: $\p(D \|M_{i}, I)$ is well-defined. In this
-case the \enquote{dimensionality} of the model is undefined, however. The
-distribution above could be considered as belonging to the family
-$\set[\Big]{\exp\Bigl[-\frac{(x-\mu)^{2}}{18}\Bigr]}$, or
-$\set[\Big]{\exp\Bigl[-\frac{(x-5)^{2}}{2\sigma^{2}}\Bigr]}$, or
-$\set[\big]{\exp[f(x-5)]}$, and so on. A point in a manifold belongs to an
-infinity of submanifolds.
-
-Yet, the property of having a minimal sufficient statistic is still
-well-defined for a model in this sense, in an exchangeable context. The
-model above has $\set[\big]{n, \widebar{x}, \widebar{x^{2}}}$ (or any
-equivalent set) as minimal sufficient statistics, where the bar denotes the
-sample average. Thus two models having different sufficient statistics can
-be compared without reference to any family. Interpreted in this sense our
-calculations of \sect~4 are correct.
-
-\medskip
-
-The question, in summary, is what exactly is being compared, whether
-families (and what their accompanying distributions are) or single
-distributions. The literature is unclear and contradictory on this. But, in
-either case, our calculations show that the approaches based on a
-population and a sample thereof can lead to different conclusions -- in
-accordance with the non-preservation of sufficiency under sampling -- and
-they provide a way to bypass this problem.
+% \medskip
+
+% \textbf{(B)} If by \enquote{model} we mean a specific distribution, \eg\
+% \[ \exp\biggl[-\frac{(x-5)^{2}}{18}\biggr]\,\di x \;,\] then two models can
+% of course be directly compared: $\p(D \|M_{i}, I)$ is well-defined. In this
+% case the \enquote{dimensionality} of the model is undefined, however. The
+% distribution above could be considered as belonging to the family
+% $\set[\Big]{\exp\Bigl[-\frac{(x-\mu)^{2}}{18}\Bigr]}$, or
+% $\set[\Big]{\exp\Bigl[-\frac{(x-5)^{2}}{2\sigma^{2}}\Bigr]}$, or
+% $\set[\big]{\exp[f(x-5)]}$, and so on. A point in a manifold belongs to an
+% infinity of submanifolds.
+
+% Yet, the property of having a minimal sufficient statistic is still
+% well-defined for a model in this sense, in an exchangeable context. The
+% model above has $\set[\big]{n, \widebar{x}, \widebar{x^{2}}}$ (or any
+% equivalent set) as minimal sufficient statistics, where the bar denotes the
+% sample average. Thus two models having different sufficient statistics can
+% be compared without reference to any family. Interpreted in this sense our
+% calculations of \sect~4 are correct.
+
+% \medskip
+
+% The question, in summary, is what exactly is being compared, whether
+% families (and what their accompanying distributions are) or single
+% distributions. The literature is unclear and contradictory on this. But, in
+% either case, our calculations show that the approaches based on a
+% population and a sample thereof can lead to different conclusions -- in
+% accordance with the non-preservation of sufficiency under sampling -- and
+% they provide a way to bypass this problem.
 

### 2020-03-17 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-17T02:31:16+0100
+%% Last-Updated: 2020-03-17T02:34:29+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In our calculations, however, we are using a zeroth-order Laplace
-approximation for calculating the evidence in \sect~4. A first-order
-approximation (containing the width of the distribution, influenced by the
-typical-set volume above) would correct for the model's power
-\citep[\sect~2.2]{mackay1992}. We believe that the correction would
-nevertheless be the same in the calculations based on the larger
-population, \eqns~(11), and on the sample, \eqns~(12). Hence the difference
-in the two approaches -- our main message -- should persist. But it would
-be better to check this.
+In our calculations of the evidence in \sect~4, however, we are using a
+zeroth-order Laplace approximation. A first-order approximation (containing
+the width of the distribution, influenced by the typical-set volume above)
+would correct for the model's power \citep[\sect~2.2]{mackay1992}. We
+believe that the correction would nevertheless be the same in the
+calculations based on the larger population, \eqns~(11), and on the sample,
+\eqns~(12). Hence the difference in the two approaches -- our main message
+-- should persist. But it would be better to check this.
 

### 2020-03-17 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-17T02:34:29+0100
+%% Last-Updated: 2020-03-17T12:01:54+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 close to that expressed in several works, studies, and commentaries
-\footcites[\eg][]{good1950,good1985,kadaneetal1980b,jonesetal1986,chatfield1995,copasetal1995,draper1995,spiegelhalteretal1995,raftery1995,gelmanetal1995,mackay1995_r2003,draperetal1996,hoetingetal1999,clydeetal1999,browneetal2006},
+\footcites[\eg][]{good1950,good1985,kadaneetal1980b,dawid1982,dawid1993,jonesetal1986,chatfield1995,copasetal1995,draper1995,spiegelhalteretal1995,raftery1995,gelmanetal1995,mackay1995_r2003,draperetal1996,hoetingetal1999,clydeetal1999,mccullagh2002,browneetal2006},
 and well-explained in MacKay \citey{mackay1992}. According to this view,
@@
 
-If by \enquote{model} we mean a family of probability distributions, \eg\
-(apart from normalization)
+If by \enquote{model} we mean a family of probability distributions, for
+example (apart from normalization)
 \[\conds*{\exp\biggl[-\frac{(x-\mu)^{2}}{2\sigma^{2}}\biggr]\,\di x}{
@@
 predictive distribution, that is, a joint probability for any number of
-samples
+data samples. Such a predictive distribution automatically determines a
+family of parametric distributions and a distribution over such family
 \citep[\chap~4]{bernardoetal1994_r2000}{fortinietal2000,ressel1983_r1985,lauritzen1984}.
@@
 p.~8) and the belief distribution over the total activity at any recorded
-bin (item ii on p.~8) are visibly different. See figure on the right: the
-first is the red dashed red curve, the second the solid green line.
+bin (item ii on p.~8) are visibly different. See figure on the right: item
+(i) is the red dashed red curve, item (ii) the solid green line.
 

### 2020-03-17 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-17T12:01:54+0100
+%% Last-Updated: 2020-03-17T12:05:53+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 family, but we agree with those \citep[\eg][]{copasetal1979}[DeGroot
-in][]{lindleyetal1982} who find such definitions ultimately arbitrary. Many
-such distributions lead, moreover, to mathematical singularities.
+in][]{lindleyetal1982} who find such definitions ultimately arbitrary. The
+pre-data distribution over the family expresses our prior information and
+is therefore highly dependent on the specific problem. Many such
+\enquote{standard} distributions lead, moreover, to mathematical
+singularities.
 
@@
 small region, with respect to that of $M_{2}$. Model $M_{1}$ is thus
-effectively using fewer parameters than $M_{2}$. Correspondingly the
-\enquote{typical set} of data to which $M_{1}$ assigns appreciable
+effectively using a smaller parameter range than $M_{2}$. Correspondingly
+the \enquote{typical set} of data to which $M_{1}$ assigns appreciable
 probability is smaller than the typical set of $M_{2}$: model $M_{1}$ is

### 2020-03-17 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-17T12:05:53+0100
+%% Last-Updated: 2020-03-17T12:14:41+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 of a model is thus best defined by the volume of \emph{data space} that is
-typical under that model. Most important, it may then happen that a model
-is more powerful than another even if it is based on a lower-dimensional
-family.
+typical under that model \citep[an \enquote{effective dimensionality} of
+the family can be calculated from this volume:][\sect~4]{mackay1992}. Most
+important, it may then happen that a model is more powerful than another
+even if it is based on a lower-dimensional family
+\citep[\sect~6]{mackay1992}[esp.\ Fig.~4]{mackay1992b}.
 

### 2020-03-17 replies1_ncomp.tex
update

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-03-17T12:14:41+0100
+%% Last-Updated: 2020-03-17T16:56:56+0100
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-As stated on p.~8, the interpretation is important because it leads to
-numerically different results in first- and higher-order approximations.
-And such approximations may be necessary in the precise calculation of the
-evidence, as explained in the reply to the previous point. Appendix~C
-explains where and why the numerical differences appear. Also, it is
-important to know whether we are calculating the frequency of past results,
-or the probability for a new result, or the most probable frequency of
-future results. These three quantities can be numerically very different
-when the ratio $\frac{\text{\# samples}}{\text{data-space dimension}}$ is
-high, and they are to be used in different ways. Preliminary calculations
-from the full Bayesian analysis show that%
+Perhaps we chose the wrong term, speaking of \enquote{interpretation}. What
+we meant is that the formula we find is valid, as a first approximation,
+for four different quantities of interest. It is obviously important to
+specify for which quantities the formula can be used and, among those, for
+which it provides a better or worse approximation.
+
+On p.~8 we also state that first- and higher-order approximations will be
+different for each of those four quantities, so their numerical equality is
+only valid at zeroth-order. Preliminary calculations from the
+full Bayesian%
 \setlength{\intextsep}{-1ex}% with wrapfigure
@@
 \end{wrapfigure}% exp_family_maxent.nb
-the most probable frequency distribution across recorded bins (item i on
-p.~8) and the belief distribution over the total activity at any recorded
-bin (item ii on p.~8) are visibly different. See figure on the right: item
-(i) is the red dashed red curve, item (ii) the solid green line.
+ analysis show that the most probable frequency distribution across recorded
+bins (item i on p.~8) and the belief distribution over the total activity
+at any recorded bin (item ii on p.~8) are visibly different. See figure on
+the right: item (i) is the red dashed red curve, item (ii) the solid green
+line.
+
+Regarding the quantities mentioned on p.~8, it is important to know whether
+we are calculating the frequency of past results, or the probability for a
+new result, or the most probable frequency of future results. These
+quantities can be numerically very different when the ratio
+$\frac{\text{\# samples}}{\text{data-space dimension}}$ is high, and they
+are to be used in different ways.
 
 These are therefore not philosophical matters. It's a matter of
-understanding what we are doing and stating it clearly in our
-work. % Unfortunately the
+understanding and clearly stating what we are
+calculating. % Unfortunately the
 % neuroscience literature is very unclear on such matters.

### 2020-04-26 maxNt_v6.tex
Add files via upload

@@
 %% Created: 2015-05-01T20:53:34+0200
-%% Last-Updated: 2020-02-03T11:24:40+0100
+%% Last-Updated: 2020-04-26T13:23:00+0200
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand*{\headtitle}{Inferences about large neuronal populations}
-\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, E. Torre, V. Rostami}
+\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, V. Rostami, E. Torre}
 \newcommand*{\headauthor}{Porta Mana \etal}
@@
 \hspace*{\stretch{1}}%
-\parbox{0.5\linewidth}%\makebox[0pt][c]%
+\parbox{0.3\linewidth}%\makebox[0pt][c]%
 {\protect\centering P.G.L.  Porta Mana\\%
@@
 \hspace*{\stretch{1}}%
+\parbox{0.3\linewidth}%\makebox[0pt][c]%
+{\protect\centering V. Rostami\\%
+\footnotesize\epost{\email{vrostami}{uni-koeln.de}}}%
+\hspace*{\stretch{1}}%
 %% uncomment if additional authors present
-\parbox{0.5\linewidth}%\makebox[0pt][c]%
+\parbox{0.4\linewidth}%\makebox[0pt][c]%
 {\protect\centering E. Torre\\%
@@
 \hspace*{\stretch{1}}\\%
-\parbox{0.5\linewidth}%\makebox[0pt][c]%
-{\protect\centering V. Rostami\\%
-\footnotesize\epost{\email{vrostami}{uni-koeln.de}}}%
-\hspace*{\stretch{1}}%
-\parbox{0.5\linewidth}%\makebox[0pt][c]%
-{\protect\centering Y. Roudi\\%
-\footnotesize\epost{\email{yasser.roudi}{ntnu.no}}}%
-\hspace*{\stretch{1}}%
 %\quad\href{https://orcid.org/0000-0002-6070-0784}{\protect\includegraphics[scale=0.16]{orcid_32x32.png}\textsc{orcid}:0000-0002-6070-0784}%
@@
   This work is financially supported by the Kavli Foundation and the Centre
-  of Excellence scheme of the  Research Council of Norway.
-
-  We thank Iv\'an Davidovich for very insightful and valuable comments on a
-  previous draft of this manuscript. PGLPM thanks the staff of the NTNU
+  of Excellence scheme, Yasser Roudi group, of the  Research Council of Norway.
+
+  We thank Yasser Roudi and Iv\'an Davidovich for very insightful and valuable comments on 
+  previous drafts of this manuscript. PGLPM thanks the staff of the NTNU
   library for their always prompt support; Mari, Miri, Emma for
@@
   filling life with awe and inspiration; the developers and maintainers of
-  \LaTeX, Emacs, AUC\TeX, Open Science Framework, Python, Inkscape, Sci-Hub
+  \LaTeX, Emacs, AUC\TeX, Open Science Framework, R, Python, Inkscape, Sci-Hub
   for making a free and unfiltered scientific exchange possible.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\clearpage
+%\clearpage
 \defbibnote{prenote}{{\footnotesize (\enquote{de $X$} is listed under D,
